### 2012-06-29 app-chemistry.tex
first commit of paper.

\section{Primordial chemistry and cooling}\label{app:chemistry}

\red{This currently just describes the rates solver without any radiation
backgrounds.  Improve this to include terms for radiation backgrounds (including
photoionizing and lyman-werner) and, if necessary, explain how these
terms fit in there.}

The primordial chemistry network implemented in Enzo is discussed in 
Section~\ref{sec.ov.chem} and also in
much more detail by~\citet{abel97} and~\citet{anninos97}.  
These papers describe the chemistry and cooling behavior of low-density primordial gas
($n \simeq 10^4$ and below), as well as
the steps that are necessary to obtain fast and accurate numerical solutions of the 
nonequilibrium chemical reaction rate network.  This network is extended by
~\citet{ABN02} to include the 3-body molecular hydrogen creation process, 
which becomes important at higher densities, and extends the validity of the reaction
network several more orders of magnitude in density, essentially until the gas becomes
optically thick to cooling by $H_2$ line emission. 
For the sake of completeness, and because the properties the primordial gas are so 
crucial to the results that are discussed in this work, we describe the chemistry 
network in this appendix.

Tables~\ref{table.collisional} and~\ref{table.radiative} summarize the collisional
and radiative processes solved in the Enzo nonequilibrium chemistry routines.  
\citet{abel97} show that accurate results can be obtained if several unnecessary
reactions are eliminated and the reaction network is reduced to the following:

\begin{eqnarray}
\frac{d \nd{H} }{dt}    &=& \k{2} \nd{\mHp} \nd{e} - \k{1} \nd{H} \nd{e} + 2 \k{31} \nd{\mHH}  \\
\frac{d\nd{\mHp}}{dt}  &=& \k{1} \nd{H}   \nd{e} - \k{2} \nd{\mHp}  \nd{e}      \\
\frac{d\nd{He} }{dt}  &=& \k{4} \nd{\mHep}\nd{e} - \k{3} \nd{He}   \nd{e}       \\
\frac{d\nd{\mHep}}{dt} &=& \k{3} \nd{He} \nd{e} + \k{6} \nd{\mHepp}\nd{e} -
                       \k{4} \nd{\mHep}\nd{e}                           \\
\frac{d\nd{\mHepp}}{dt}&=& \k{5} \nd{\mHep}\nd{e} - \k{6} \nd{\mHepp} \nd{e}    \\
\frac{d\nd{\mHH} }{dt}  &=& \k{8} \nd{\mHm} \nd{H} + \k{22} \nd{H}^3 - \nd{\mHH}   \left( 
        \k{31} + \k{11} \nd{\mHp} + \k{12} \nd{e} \right), 
\end{eqnarray}

where the number density of \Hm is given by the equilibrium condition

\begin{eqnarray}
\nd{\mHm} = {\k{7} \nd{H} \nd{e}  \over \k{8} \nd{H}  + 
 \k{16} \nd{\mHp} + \k{14} \nd{e}}.
\end{eqnarray}

The \Hm number density can be calculated in equilibrium because the timescale at which the
reactions controlling its number density occur are much shorter than the rest of the 
reactions in this system.  The rate coefficients used in the equations above are defined as follows:

\begin{eqnarray}       
 k_1 & = & \exp[-32.71396786 + 13.536556  \ln(T) - 5.73932875   \ln(T)^2 \nonumber \\ 
     &   & + 1.56315498  \ln(T)^3  - 0.2877056 \ln(T)^4 + 3.48255977\times 10^{-2} \times\ln(T)^5  \nonumber \\
     &   & - 2.63197617\times 10^{-3}\times\ln(T)^6 +  1.11954395\times 10^{-4}   \ln(T)^7 \nonumber \\
     &   & -  2.03914985 \times 10^{-6}   \ln(T)^8]~{\rm cm^3~s^{-1}}
\end{eqnarray}

\begin{eqnarray}
 k_2 &=& \exp[-28.6130338 - 0.72411256  \ln(T) - 2.02604473\times10^{-2}  \ln(T)^2  \nonumber \\
     & &   - 2.38086188\times10^{-3}  \ln(T)^3 - 3.21260521\times10^{-4}  \ln(T)^4  \nonumber \\
     & &   - 1.42150291\times10^{-5}  \ln(T)^5 + 4.98910892  \times10^{-6} \ln(T)^6  \nonumber \\
     & &   + 5.75561414 \times10^{-7}   \ln(T)^7 - 1.85676704 \times10^{-8}   \ln(T)^8 \nonumber \\
     & &   - 3.07113524 \times10^{-9}   \ln(T)^9]~{\rm cm^3~s^{-1}}
\end{eqnarray}

\begin{eqnarray}
 k_3 &=& \exp[ (-44.09864886 + 23.91596563   \ln(T) -  10.7532302   \ln(T)^2 \nonumber \\
     & &   + 3.05803875   \ln(T)^3 - 0.56851189   \ln(T)^4 + 6.79539123 \times 10^{-2}  \ln(T)^5  \nonumber \\
     & &   - 5.00905610 \times 10^{-3}   \ln(T)^6 + 2.06723616 \times 10^{-4} \ln(T)^7 \nonumber \\
     & &   - 3.64916141 \times 10^{-6}   \ln(T)^8)~{\rm cm^3~s^{-1}}
\end{eqnarray}

\begin{eqnarray}
k_{4r} & = & 3.925 \times 10^{-13}T^{-0.6353}~{\rm cm^3 s^{-1} } \\
k_{4d} &=& 1.544 \times 10^{-9} T^{-\frac{3}{2}} 
     \exp\left( - \frac{48.596~eV}{T} \right) \times \nonumber \\
   & &     \left[0.3 + \exp\left(\frac{8.10~eV}{T}\right)\right]~{\rm cm^3~s^{-1}}
\end{eqnarray}

\begin{eqnarray}
k_5 & = & \exp[-68.71040990 + 43.93347633   \ln(T) - 18.4806699   \ln(T)^2 \nonumber \\
    & &   + 4.70162649   \ln(T)^3 - 0.76924663   \ln(T)^4 + 8.113042 \times 10^{-2} \ln(T)^5 \nonumber \\
    & &   - 5.32402063 \times 10^{-3}  \ln(T)^6 +  1.97570531\times 10^{-4}  \ln(T)^7 \nonumber \\
    & &   - 3.16558106 \times 10^{-6}  \ln(T)^8]~{\rm cm^3~s^{-1}}
\end{eqnarray}

\begin{eqnarray}
k_6 = 3.36 \times 10^{-10} T^{-\frac{1}{2}}
\left(\frac{T}{1000\K}\right)^{-0.2}
\left(1+\left(\frac{T}{10^6\K}\right)^{0.7}\right)^{-1}~{\rm cm^3~s^{-1}}
\end{eqnarray}

$k_7$ for T $\leq 6000$~K:

\begin{eqnarray} 
k_7 & = & 1.429 \times 10^{-18} T^{0.7620} T^{0.1523 \log_{10}(T)}  T^{-3.274 \times 10^{-2} \log_{10}^2(T)}~{\rm cm^3~s^{-1}} \nonumber \\
\end{eqnarray}

$k_7$ for T $ > 6000$~K:

\begin{eqnarray}  
k_7 & = & 3.802 \times 10^{-17} T^{0.1998 \log_{10}(T)}  \nonumber \\
    &   & {\rm dex} \left( { 4.0415 \times 10^{-5} \log_{10}^6(T)  - 5.447 \times 10^{-3} \log_{10}^4(T)}~{\rm cm^3~s^{-1}}\right)
\end{eqnarray}

\begin{eqnarray}
T&>&0.1~{\rm eV}:\ k_8 = \exp[-20.06913897 +0.22898   \ln(T) +  3.5998377  \nonumber \\
 &&       \times 10^{-2}   \ln(T)^2 - 4.55512 \times 10^{-3}  \ln(T)^3- 3.10511544\times 10^{-4}   \ln(T)^4   \nonumber \\
 &&       + 1.0732940 \times 10^{-4}  \ln(T)^5 -  8.36671960 \times 10^{-6}   \ln(T)^6 + 2.23830623 \nonumber \\
  &&      \times 10^{-7}   \ln(T)^7]~{\rm cm^3~s^{-1}} . \\
T &<& 0.1~{\rm eV}:  k_8  =  1.428 \times 10^{-9}~{\rm cm^3~s^{-1}}
\end{eqnarray}

\begin{eqnarray}
\ln(k_{11}) & = &   -24.24914687 + 3.40082444  \ln(T) - 3.89800396  \ln(T)^2  \nonumber \\
             & &       +  2.04558782  \ln(T)^3 -  0.541618285  \ln(T)^4   + 8.41077503 \times 10^{-2}  \ln(T)^5 \nonumber \\
             & &       - 7.87902615 \times 10^{-3} \ln(T)^6 \nonumber + 4.13839842 \times 10^{-4} \ln(T)^7 \nonumber \\
             & &       - 9.36345888\times 10^{-6}  \ln(T)^8 {\rm cm^3 s^{-1}}
\end{eqnarray}

\begin{eqnarray}
\ \ k_{12} = 5.6 \times 10^{-11} T^\frac{1}{2} \exp(-\frac{102,124 K}{T}) {\rm cm^3 s^{-1}}
\end{eqnarray}

\begin{eqnarray}
k_{14}&=& \exp[-18.01849334 + 2.3608522   \ln(T) - 0.28274430   \ln(T)^2   \nonumber \\
      & &    + 1.62331664 \times 10^{-2}  \ln(T)^3 - 3.36501203 \times 10^{-2} \ln(T)^4   \nonumber \\
      & &    + 1.17832978 \times 10^{-2}   \ln(T)^5 - 1.65619470 \times 10^{-3} \ln(T)^6   \nonumber \\
      & &    + 1.06827520 \times 10^{-4}   \ln(T)^7 - 2.63128581 \times 10^{-6} \ln(T)^8) {\rm cm^3 s^{-1}}
\end{eqnarray}

\begin{eqnarray}\ \ k_{16} = 7 \times 10^{-8} \left( \frac{T}{100{\rm
        K}} \right)^{-\frac{1}{2}} {\rm cm^3 s^{-1}} 
\end{eqnarray}

$k_{22}$ For T $< 300$~K:

\begin{eqnarray}
k_{22} & = & 1.3 \times 10^{-32} (T / 300 {\rm K} )^{-0.38} {\rm cm^6 s^{-1}}
\end{eqnarray}

$k_{22}$ for T $\geq 300$~K:

\begin{eqnarray}
k_{22} & = & 1.3 \times 10^{-32} (T / 300 {\rm K} )^{-1} {\rm cm^6 s^{-1}}
\end{eqnarray}

The photodissociation of molecular hydrogen by Lyman-Werner radiation is controlled by the $k_{31}$ parameter:

\begin{eqnarray}
k_{31} & = & 1.13 \times 10^8~{\rm F_{LW} }~|t|
\end{eqnarray}

where $|t| \equiv ( 4 \pi G \Omega_m \rho_c (1+z_i)^3)^{-1/2}$
and has units of seconds, and F$_{LW}$ is the Lyman-Werner background flux in 
units of~erg~s$^{-1}$~cm$^{-2}$~Hz$^{-1}$.

% this is the end!

### 2012-06-29 app-datastruct.tex
first commit of paper.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\section{Data structures}\label{app.datastructure}

The primary data structure in ENZO\ is the {\tt grid}.  
\gtt{It has yet to be determined how much of this
appendix will be in the Method Paper, and how much will be in the
User/Developer Guide.  Green type denotes text that will definitely be
in the User/Developer guide.} \red{Red denotes meta notes-- things to
write, fix, find.}

\subsection{Outline and List of Routines.}
\begin{enumerate}
 \item Creation and Destruction
   \begin{enumerate}
     \item Flow of Rebuild hierarchy: 
       \begin{enumerate}
	 \item delete hierarchy
	 \item build new hierarchy
	 \item scavange from old hierarchy
       \end{enumerate}
     \item Cell Flagging Methods
     \item Grid Fracturing Methods
   \end{enumerate}    
 \item Parallelism
  \begin{enumerate}
    \item Top Grid (block domain decomposition)
    \item Subgrids (memory balance, level by level)(Actually, I think
          its Parent Grid by Parent Grid.)
  \end{enumerate}
 \item Relation to other objects: 
   \begin{enumerate}
     \item Parents (No refinement jump requirements.  Talk about flux
        correction improvement to this effect here?)
     \item Siblings
     \item Children
     \end{enumerate}
 \item Traversing the hierarchy (this might be rolled into the
   previous bullet)
  \begin{enumerate}
    \item Level iterator 
    \item Hierarchy iterator
    \item Grid Array
    \item Chaining Mesh
  \end{enumerate}
  \item Particle treatement
   \begin{enumerate}
     \item Which grid gets the particles
     \item How particles are refined: criteria
     \item How particles are refined: how it fits in with the Grid refinement
   \end{enumerate}
  \item Ghost zones
    \begin{enumerate}
      \item how, when filled
    \end{enumerate}
  \item Timestepping
    \begin{enumerate}
      \item How timestep is determined
      \item Order of level operations: W cycle
    \end{enumerate}
\end{enumerate}
A list of routines to mention: \red{Zeus does this in a big table that
  references code name, what it does, and what equations it
  solves. Might be a useful idea to steal.}
\begin{description}
 \item[\tt{EvolveHierarchy}] Primary timestep control routine. 
 \item[\tt{EvolveLevel}] Primary evolution routine.
 \item[\tt{RebuildHierarchy}] Rebuilds the grid hierarchy
\end{description}

\subsection{Grid Intro}
The life of an \enzo\ simulation centers around the {\tt grid}
object and relationships between individual {\tt grid} instances.  
Grids are independant cartesian patches in space, which contain meta data about
that space (for instance its size, its position in the volume, 
size of the cells in that patch) as well as the data that belongs in
that space (fluid quntities like density and energy, as well as
particle quantites like position and velosity.)

Grids are treated, as individual instances of fluid
dynamics problems, with Dirichlet boundary conditions.  Boundary
conditions, stored in ghost zones, are determined by
one of 3 methods; from user defined exterior conditions like
outflow or reflecting, copied from active zones of 
neighbors, or interpolated from parents.  Grids are group in two
logical structures: a tree like hierarchy that relates parent grids to
children grids; and a level structure that relates grids of the same
refinement.

\subsection{Creation and Destruction: \tt{RebuildHierarchy}}
Creation and destruction of grids happens in the routine
{\tt RebuildHierarchy}.  
At the end of every timestep, the Hierarchy is rebuilt.
Rebuilding starts at the coarsest level at the point in time (note
that coarser levels may be further ahead in their integration, due to
their larger time step and the W-cycle)  Beginning with this coarsest level, each
grid on each level is examined for refinement.  Cells are first flagged
based on a variety of criterion, described in section
\ref{sec.flagging}.  The set of flagged cells, stored in the \grid\ memeber
array {\tt FlaggingField}, is then devided into grids as described
in section \ref{app.subgrid_creation}.  Data in these subgrids
is then created by interpolation from the parent using \red{what kind
  of interpolation?}  After this New Subgrid is filled with
interpolated data, the set of Old Subgrids is examined for overlapping
data.  All Old Cells that correspond to New Cells are coppied to the
New Grid.

\subsection{Cell Flagging Methods}\label{sec.flagging}
\enzo\ selects cells to refine based on the physics of the simulation.
Essentially any criterion that can be derived locally from the fluid
data and its derivatives can, in principal, be used to flag the cells
for refinement. Currently, \enzo\ has 8 methods of refinement.
installed, which we will discuss here.  Any or all of these methods may be employed in
any given \enzo\ simulation.  \gtt{Information on how to write your own
refinement criterion can be found Somewhere, to be written}
\begin{description}
  \item[Field Slope] \gtt{grid::FlagCellsToBeRefinedBySlope()} For each
  fluid quantity $u$, a cell $i$ is flagged if 
$$ \frac{u_{i+1}-u_{i-1}}{u_i} > slope $$
  where $slope$ \gtt{\tt MinimumSlopeForRefinement} is
  a  user-defined
  parameter. This is repeated for all
  3 directions. 
  \item[Baryon mass] \gtt{
  grid::FlagCellsToBeRefinedByMass()}  Cells
  are flagged if $$ \rho  V > m_{f} r^{\alpha L} $$ 
  where $V$ is the
  volume of the cell, $r$ is the refinement ratio, $L$ is the grid
  level, $m_{f}$ and ${\alpha}$ are user parameters. \gtt{MinimumMassForRefinement} and 
  \gtt{ MinimumMassForRefinementLevelExponent\[\]}
  \item[Baryon Overdensity] As the name suggests, refines if the
  density is greater than the mean density in the simulation.  In
  practice, this refinement method uses the same machenery as flagging
  by Baryon Mass, with the minimum
  mass for flagging, $m_f$ \gtt{ = MinimumMassForRefinement}, set to the
  input parameter \gtt{MinimumOverDensityForRefinement} times the
  mean density.  \gtt{This is calculted differently for each problem
  generator: consult the source for further documentation}
  \red{This could probably use a word on the effects of
  parameter choices.}
  \item[Shocks] \gtt{
  grid::FlagCellsToBeRefinedByShocks()} \red{This text should be
  updated to reflect Alexei's improvements} A cell $i$ is flagged if it is
  determined to be taking part in a shock.  This is determined if it
  has each of the following:
  \begin{enumerate}
    \item Large relative pressure jump: 
      $$\frac{|p_{i+1} - p_{i-1}| }{min( p_{i+1}, p_{i-1} ) }
      > \gtt{MinimumPressureJumpForRefinement} $$
     \item compression: $$ u_{i+1} < u_{i-1} $$
    \item More thermal energy than Kinetic: $$ \frac{
    \frac{p_i}{\gamma -1} }{ \frac{1}{2}\rho v^2 } > \gtt{
    MinimumEnergyRatioForRefinement } $$ \red{why?}
  \end{enumerate}  
  \item[Particle Mass] Using the Cloud In Cell interpolation method,
  described  \red{elsewhere?}, particle data is interpolated to the
  Eulerian grid cell centers.  A cell is flagged if the resulting mass field passes
  the ``RefineByMass'' test, described above.
  \item[Jeans Length] If a gravitationally collapsing simulation is
  under-resolved, artificial fragmentation can occur.  It was
  descoverd experimentally by \red{Truelove; get citation}
  that sufficient resolution was achieved if the Jeans Length is
  resolved by at least 4 cells.  This is determined if 
  $$ dx^2 >
  \frac{\alpha}{N_{cells} \gtt{= RefineByJeansLengthSafetyFactor}}\frac{T_{i,j,k}}{\rho_{i,j,k}} $$

  and $\alpha = \frac{\pi k_b }{G m_{proton} }*
  \frac{LengthUnits^2}{DensityUnits}$

  \red{dude, double check this.}
  \item[Cooling Time]  \red{Finish this.}

  \item[Shear] For highly turbulent flows, it may be useful to resolve
  shear layers.  This is done whenever
  $$\frac{\partial v_x}{\partial y}^2 + \frac{\partial v_x}{\partial z}^2 
  > \eta \gtt{= MinimumShearForRefinement}
  $$
  In practice, $$\frac{\partial v_x}{\partial y} = 
  u_{j+1}-u_{j-1}$$
  The same is done for $v_y$ and $v_z$

\end{description}

\subsection{Grid Fracturing}\label{app.subgrid_creation}
Once the cells in a given grid are flagged for refinement, they're
broken into a set of subgrids by an iterative process.  This process
first draws a minimally enclosing box around the set of flagging
cells, and generates a {\tt ProtoSubgrid}.  A {\tt ProtoSubgrid} is
essentially just the spatial extent and flagging field that will
define the spatial extent of the future \grid.  Each {\tt
  ProtoSubgrid} is then examined for efficiency (the ratio of flagged cells to
total), total size (for cache optimization) and minimum side
length (to avoid a flock of one zone grids).  {\tt ProtoSubgrid} that are too
large or not efficient enough are broken into smaller grids, by a
process to be described shortly, new minimally enclosing boxes are
drawn around each of the two remaining flagging fields, and new {\tt
  ProtoSubgrid}s are created.  This process is then repeated until
all {\tt ProtoSubgrid}s are both small enough and efficient enough.
It should be noted that \enzo\ does not glue grids back together-- the
Minimum Side Length is enforced by not examining the other two
conditions if the {\tt ProtoSubgrid} is too small.

Maximum size, minimum side length,  and minimum efficiency are
controlled by the parameters {\tt MaximumSubgridSize}, {\tt MinimumSubgridEdge} and {\tt
  MinimumEfficiency}, respectivly. Typical maximum sizes are 10000 zones, minimum
side lengths are typically 5 zones, and minimum efficiencies are  0.2.
Choice of these parameters, especially maximum size, are
based on the architecture of the computer in question.  
Machines with larger cache sizes will benifit from larger values for
{\tt MaximumSubgridSize}, as communication and memory overhead from
ghost zones will be reduced; machines with smaller cache sizes benifit
from lower values, in that the computational time is reduced for a
grid that can fit entirely into cache.

A {\tt ProtoSubgrid} that fails either the efficency or size
requirement is then broken into smaller {\tt ProtoSubgrid}s by a
method that tries to preserve physically continuous structures onto
individual grids.  The process examines the 1d Signature of the Flagging
field, which is the projection of the Flagging Field along two of the
coordinate axes.  First, each of the 3 signature fields is examined
for zeros.  Zeros in the signature field indicates disjoint
structures, so the {\tt ProtoSubgrid} is first cut into two smaller
ones along that line.  If no such cuts can be made, zero crossings of
the second derivative in the signature are then looked for.  Zero
crossings in the second derivative indicates edges of structures.  If
more than one exists, the 'strongest crossing,' that is the largest
jump, is selected for the cut.  If no such zero crossings can be
found, the {\tt ProtoSubgrid} is then cut evenly along the longest
axis.  See figure \ref{fig.signature} for a schematic of this process.

\begin{figure}
\begin{center}
\includegraphics[width=0.4\textwidth]{figures/SignatureSchematic.eps}
\end{center}
\caption{
Schematic of Flagging Field and Signature.  Cuts will be made along
some line I'll draw.  The figure will be improved in the future.
}
\label{fig.signature}
\end{figure}

### 2012-06-29 app-gravity.tex
first commit of paper.

\section{Gravity solver}

\red{This needs to be fleshed out a lot - basically we want to emphasize how we solve the
root grid (mention parallelization strategy for FFT), how the potentials calculated are 
interpolated to subgrids for the multigrid solution on the subgrids, and how particles are 
actually moved.
}

\red{**********  Which version of the gravity solver will we be using for the public release?  
The new Dan Reynolds mulgrid-everywhere solver, or the one that's currently in the public
release?  Alter paper and other documentation accordingly.}

With and without cosmology - how does this affect what is solved/how it's solved?
Why do we do $\rho-\bar{\rho}$ instead of just $\rho$?  What is the finite difference
equation that you actually solve in the end?

How are particles interpolated to grids, and accelerations from grid
to particles?

\subsection{multigrid}
This is for James.

\subsection{leapfrog integrator}
Talk about how particles' accelerations are calculated, etc.

### 2012-06-29 app-ppmhydro.tex
first commit of paper.

\section{The PPM hydro method}\label{app.ppm}
\red{
\begin{enumerate}
  \item Cite \citet{1995CoPhC..89..149B} for Lagrangian Remap version
  \item State that Lagrangian Remap it doesn't really work and nobody uses it (why is that?)
  \item Brief overview of PPM: 
  \begin{enumerate}
    \item Parabolic reconstruction
    \item Order Lowering near shocks
    \item Characteristic tracing for Left and Right states
    \item I don't remember if we use a contact steepener:  If yes,
    describe.  If no, say so.
  \end{enumerate}
  \item Repeat bits of \citet{1984JCoPh..54..174C} that have been altered
        for cosmology 
  \item I think C\&W 1984 discuss gravity: double check we do the
  same.  If not, comment.
  \item Discus TwoShock solver, 
    \begin{enumerate} 
       \item Short method review, I think Woodward 1986.
       \item Discuss Rarefaction updates 
       \item Discuss alterations for Expansion
       \item Discuss alterations for Gravity (if not in Woodward 1986)
    \end{enumerate}
\end{enumerate}
}%end red

### 2012-06-29 code-tests.tex
first commit of paper.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Code tests}\label{sec.tests}

\red{
\begin{enumerate}
\item Things that need to be addressed:  Can we refine by non-powers of two?  What test problems to do?  Order of test problems?
\item Some of the test problems are motivated by the FLASH code method paper \citep{flash_method}, and are referred to there.  Others are motivated by the GADGET-2 method paper \citep{gadget_2_method}, and are referred to as such.
\item In this paper we want to make sure to refer to the Santa Barbara Cluster comparison project \citep{SantaBarbara} and to our ENZO/GADGET-2 comparison project~\citep{2005ApJS..160....1O}.
\end{enumerate}
}

\red{
General structure for each test problem:  Outline how the test problem is constructed (initial and boundary conditions), 
the analytical solution, why we have in the paper (what does it break, or what flaw does it expose (not in enzo of course)),
a plot showing how well enzo solves said test problem, and a brief description of the plot and how awesome enzo is.
}

\red{*** WE NEED SUGGESTIONS FOR CODE TESTS***}

\subsection{Advection problem}\label{sec.tests.advect}
Problem 7.1 in the flash method paper~\citep{flash_method}.

\subsection{Sod Shock Tube}\label{sec.tests.sodshock}
Problem type 1.  Both unigrid and AMR.  Tests the hydro.
This is also problem 7.2 in the FLASH method paper.

\subsection{Wave pool}\label{sec.tests.wavepool}
Problem type 2, unigrid and fixed AMR 1D mostly.  Tests reflections of waves at boundaries.

\subsection{Shock pool}\label{sec.tests.shockpool}
Problem type 3, unigrid and AMR 2D.  Tests passage of shock through a refinement boundary.

\subsection{Interacting Blast-Wave problem}\label{sec.tests.interblast}
Problem 7.5 in the FLASH method paper.

\subsection{Double mach reflection}\label{sec.tests.doublemach}
Problem type 4.  This is one of Alexei's test problems: 
http://cosmos.ucsd.edu/~akritsuk/renzo/mach/mach.html

This tests the hydro.

\subsection{Sedov Explosion}\label{sec.tests.sedov}
Problem type 7.  One of Alexei's test problems:
http://cosmos.ucsd.edu/~akritsuk/renzo/sedov/sedov.html
Also problem 7.4 in the FLASH method paper.
This tests the hydro.

\subsection{Noh problem}\label{sec.tests.noh}
Problem type 9.  One of Alexei's test problems:
http://cosmos.ucsd.edu/~akritsuk/renzo/noh/noh.html
This tests the hydro.

\subsection{Radiating shock problem}\label{sec.test.radshock}
In~\citet{anninos97} -- tests the energy equation.  Followon
to Noh problem.  Original reference is cited in this paper.

\subsection{Point source gravity test}\label{sec.test.gravitypointsource}
This is the TestGravity (problem 23) test problem.  It tests gravity around a point source, using fixed AMR.

\subsection{Orbit Test}\label{sec.test.testorbit}
This is the TestOrbit problem (problem 29 for GB).  This tests gravity and particle integration (no hydro).

\subsection{Pressureless Sphere Collapse}\label{sec.test.pressurelesscollapse}
This is in the zeus paper and in the kronos test paper -- it tests mostly gravity and advection.  Test problem 21.

\subsection{Collapse of an adiabatic gas sphere}\label{sec.tests.adiasphere}
I think this is problem type 27, and also is problem 6.2 in the
GADGET-2 method paper~\citep{gadget_2_method}.

\subsection{Self-Similar infall test}\label{sec.tests.infall}
Problem type 24.  This is a test based on Bertschinger's 1985 3D self-similar infall
solution, and tests gravity + hydro.

\subsection{Zel`Dovich Pancake}\label{sec.tests.pancake}
Problem type 20, unigrid and AMR both.  Tests gravity + hydro.

\subsection{Santa Barbara Cluster comparison}\label{sec.tests.santabarbara}
This is the classic test -- we can mostly just point to the enzo results in that paper.  Maybe update them?

\subsection{Dark matter halo mass function and clustering}\label{sec.tests.dm}
Standard problem, also problem 6.4 in GADGET-2 method paper.  Should this be here, or just a reference to another paper?

### 2012-06-29 discussion.tex
first commit of paper.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Discussion}

\red{
\begin{enumerate}
\item What physics will be put into the code in the next couple of years? (ideal mhd, FLD, ray tracing, more...)
\item What will never be in the code (non-cartesian coordinates, SPH)
\item What are clear areas of improvement?  (gravity solver, maybe more advanced/different hydro methods, parallelization, load balancing)
\item Discuss relative ease of adding new particle types (dust, ``galaxy particles,'' etc.) and some other new packages (cooling).  Other things are harder (MHD, radiation transport).  Outline what's easy and what's hard, in very general terms.
\end{enumerate}
} %end red

### 2012-06-29 introduction.tex
first commit of paper.

\section{Introduction}\label{sec.intro}

The numerical simulation of astronomical phenomena now plays a key role, along with observations and analytic theory, in pushing forward our understanding of the cosmos 
%\citep[e.g.,][]{DecadalSurvey01}.  
But we believe that with this role comes responsibility.  Those developing simulation tools must fulfill two key obligations: the first is to make those tools available to the community as a whole, much in the way that astronomical data is usually made publicly available.  The second is document, test and refinement those methods so that they can be critically evaluated and expanded by others.  This paper represents our attempt to meet our obligations with respect to the \enzo\ code.

\enzo\ is a structured adaptive mesh refinement (SAMR) code that was originally developed for cosmological hydrodynamics and has been used on a wide variety of problems.  It has been expanded as a general tool for astrophysical fluid dynamics and is intended to be efficient, accurate and easily expanded.  Although many of the components of the code have been described in previous publications \citep{1995CoPhC..89..149B, BryanThesis96, Bryan97a, Bryan97b, Norman99, BryanCompSci99, Bryan01, Oshea04}, there has been no systematic and complete description of the code.  In this paper we provide that description, filling in many gaps and showing the result of the code on a wide variety of test problems.

There are many diverse phenomena in astrophysics and cosmology which require a high spatial and temporal dynamical range. There are a number of numerical techniques which have been used to
model such problems.  The most common is a gridless, fully Lagrangian method call Smoothed Particle Hydrodynamcs \citep[SPH;][]{Lucy77, SPH} which has been very successful, particularly for those situations dominated by gravity.  Its development, however, is still at a relatively early stage when
compared to the effort put into Eulerian, grid-based hydrodynamic schemes \citep{Woodward84}.

Therefore, we would like to make use of this expertise; unfortunately, none of these schemes provide an easy way to adaptively and dynamically increase the resolution in small volumes, as required by, for example, gravitational instability.  We have instead, adapted an idea first suggested by \citet{Berger89} in the Computational Fluid Dynamics (CFD) community, which has become known as Structured Adaptive Mesh Refinement (SAMR).  The principle is to adaptively add (and modify) additional, finer meshes, or patches, over regions which require higher resolution.  Since each patch has a simple geometry (here, cartesian) the hyperbolic fluid-dynamical equations can be solved on that patch with a well-developed `off-the-shelf' numerical method.  In addition, it will be relatively straightforward to add other advanced physics (e.g. MHD, radiation transport beyond flux limited diffusion, etc.). This paper is devoted to describing how we have implemented this idea for astrophysics, including the addition of self-gravity and collisionless fluids.

There have been a few other numerical methods described in the astronomical literature which contain elements of this idea, or have a similar aim.  For example, an N-body solver \cite{Villumsen89} used
non-adaptive meshing to increase the resolution in pre-selected.  This static approach, applied to hydrodynamics, has been used extensively \citep[e.g.,]{Ruffert94, Anninos94}.  Adding adaptivity is a more recent enhancement, and there are now a number of N-body codes which possess this feature, both with and without  hydrodynamics  \cite{Couchman91, Jessop94, Suisalu95, Splinter96, Gelato97, ART97, Truelove98, flash_method, MLAPM01,  Yahagi01, RAMSES, Quilis04, Ziegler05, Zhang06}. It is also possible to deform the grid to obtain high resolution \citep[e.g.,][]{Gnedin95, Xu97, Pen98}.

The \enzo\ code has been widely used over the lasts decade in a wide variety of problems in topics ranging from star formation to turbulence to large-scale structure.  More than 50 peer-reviewed papers have been published based on results obtained with the code.

The structure of the code is as follows.  In section~\ref{sec.overview}, we first provide a top-level overview of the code method and structure.  This is designed to give a first broad-brush picture of the equations we aim to solve and the methods used to solve them.  Next, in section~\ref{sec.methods}, we describe the methods we use in detail, reserving some of the longer descriptions of particular components for the appendix in order not to interrupt the flow of the paper.  The code tests are described in section~\ref{sec.tests} and the parallelism strategy and scaling results in section~\ref{sec.performance}.

%\red{
%\begin{enumerate}
%  \item Motivate why we're writing the paper now: (because there isn't one, lots of people use it for all sorts of things, given that it's publicly available we should really have a method paper and some general proof that the code works, and a list of what it does at this point in time)
%  \item Why have the \enzo\ code at all?  Many diverse phenomena in astrophysics and cosmology where you need a high spatial and temporal dynamical range.  Regular (i.e. monolithic) grids don't have the required range.  While SPH is the most commonly used tool, and has been quite successful, the lack of a grid structure complicates the usage of advanced hydrodynamical techniques and other advanced physics (e.g. full solution of the MHD equations, radiation transport beyond flux limited diffusion, etc.).  There has been 50(ish) years of advancement in the fields of grid-based solutions to these methods, and the use of adaptive mesh on structured grids allows us to take advantage of this for a wide range of astrophysical and cosmological applications.  
%  \item List the subjects that the code has been used for, citing appropriate papers.  Emphasize heavily landmark contributions (Greg's 1998 galaxy cluster paper and various Motl/Burns/Hallman cluster things, various lyman alpha forest stuff, Pop III papers, alexei's turbulence stuff, etc.).  This code has been used for approximately 10 years, and is publicly available.  List papers by non-collaborators that have already used the code.
%  \item Explain structure of the paper.  Emphasize that we pass through the enzo code in multiple levels of (increasing) detail: The ``physical equations'' section provides bare-bones ``what we're solving'' information, the ``overview of numerical methods'' section briefly explains (i.e. no pseudocode or finite difference stencils) how the various important subsystems of the code work, and the appendices of the paper have the real nitty-gritty details that most people don't want to know, but a select few will find highly useful.
%\end{enumerate}
%} %end red

### 2012-06-29 ms.tex
first commit of paper.

%\documentclass[10pt,preprint]{aastex}  % for e-submission to ApJ
%\documentclass[12pt,preprint2]{aastex}  % for e-submission to ApJ - two column
\documentclass[onecolumn]{emulateapj}  % this makes everything look like ApJ

\usepackage{graphicx, natbib, color}

%%%% PUT NEW COMMANDS AND DEFINITIONS HERE %%%%
\input{local-commands}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

% Paper Outline  (last updated June 15, 2009)
%
% 1.  Introduction [GB,MN]
% 2.  Overview [GB]
% - including both equations and a few paragraph overview of how each method works
% - this might have section 2.1, 2.2, etc. corresponding to very brief overviews of section 3.1, 3.2, etc.
% 3.  Numerical methods
%  3.1 AMR [GB]
%  3.2. Hydro: ZEUS [GB]
%  3.3. Hydro: PPM [GB,AK]
%  3.4. Gravity [GB]
%  3.5. Particle Integration [GB]
%  3.6. Chemistry [BWO]
%  3.7. Star formation & Feedback [BWO]
%  3.8. Timestep constraints [BWO,GB]
% 4. Software Implementation
% - maybe this could be split up by into sections 4.1, 4.2, 4.3, etc. using the same scheme as section 3 [same responsibilities as 3]
% 5. Parallelism 
% 5.1. Parallel strategy [GB, RH]
% 5.2. Parallel performance [JB, GB, RH]
% 6. Code Tests [All]
% 7. Summary and Conclusions
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\citestyle{aa}  % correct formatting for ApJ style files

\begin{document}

\title{Enzo: An Adaptive Mesh Refinement Code for Astrophysics}
\author{Greg Bryan\altaffilmark{1}, Michael L. Norman\altaffilmark{2}, 
Brian W. O'Shea\altaffilmark{3}, Robert Harkness\altaffilmark{4}, Tom Abel\altaffilmark{5},
David C. Collins,\altaffilmark{2} 
Alexei Kritsuk\altaffilmark{2} and James Bordner\altaffilmark{2} }

\altaffiltext{1}{ Department of Astronomy, Columbia University,
New York, NY, 10027, U.S.A.  Email: gbryan@astro.columbia.edu }
\altaffiltext{2}{Center for Astrophysics and Space Sciences,
University of California at San Diego.
Email: mnorman,dcollins,akritsuk,jbordner@cosmos.ucsd.edu}
\altaffiltext{3}{Theoretical Astrophysics (T-6), Los Alamos National
Laboratories.  Email:  bwoshea@lanl.gov}
\altaffiltext{4}{San Diego Supercomputing Center, University of California
at San Diego.  Email: harkness@sdsc.edu}
\altaffiltext{5}{Kavli Institute for Particle Astrophysics and Cosmology,
 Stanford University.  Email:  tabel@stanford.edu}

\begin{abstract}
We describe the \enzo\ code, which uses a structured adaptive mesh refinement technique to provide high spatial and temporal resolution for modeling astrophysical flows.  The code includes support for the following physics: multi-dimensional hydrodynamics, self-gravity, N-body dynamics, primordial gas chemistry, optically-thin radiative cooling, and cosmological expansion.  Two Eulerian hydrodynamic solvers are provided.  This paper describes the method as well as the code's performance on a wide range of test problems.
\end{abstract}

\keywords{methods: numerical --- hydrodynamics}

\maketitle

%% Section: introduction
\input{introduction}

%% Section:  Overview
\input{overview}

%% Section: numerical methods
\section{Numerical Methods}
\label{sec.methods}

We now describe each of the numerical methods in detail.  The structure of this section mirrors the previous one, in which each method was briefly described; here we present the details.

\input{numerical-amr}
\input{numerical-ppm}
\input{numerical-zeus}
\input{numerical-gravity}
\input{numerical-particle}
\input{numerical-chemistry}
\input{numerical-starformation}
\input{numerical-timestep}
%\input{numerical-units}

%% Section: software implementation
% \input{software}

%% Section: code tests
%%  this should probably be done with one file per test
\input{code-tests}

%% Section: parallelism and performance
\input{parallel}

%% Section: discussion
\input{discussion}

%% Section: summary
\input{summary}

\appendix

%% Appendix: interpolation
%\input{app-interpolation}

%% Appendix:  Data Structures
\input{app-datastruct}

%% Appendix:  I/O
\input{app-IO}

%% Appendix:  Zeus hydro
%\input{app-zeushydro}

%% Appendix:  PPM hydro
%\input{app-ppmhydro}

%% Appendix:  gravity solver
%\input{app-gravity}

%% Appendix:  
\input{app-chemistry}

\bibliographystyle{apj}
\bibliography{apj-jour,ms}  % looks in ms.bib for bibliography info

\end{document}

### 2012-06-29 numerical-amr.tex
first commit of paper.

\subsection{The Structured Adaptive Mesh Refinement (SAMR) Method}
\label{sec.amr}

%\subsubsection{SAMR Overview}

Unlike moving mesh methods \citep{1995ApJS..100..269P,1995ApJS...97..231G} or  methods that subdivide individual cells \citep{Adjerid}, Berger \& Collela's AMR (also referred to as \emph{structured} AMR) utilizes an adaptive hierarchy of grid patches at varying levels of resolution.  The SAMR method described here is similar to \citet{Berger89}.  A more complete discussion of the history and development of this type of algorithm is available in \citet{Neeman96}.

The back-bone of the SAMR idea is the patch, or grid, which we take to be rectilinear to simplify booking (in practice, the lower level routines all assume Cartesian coordinates).  The single root grid covers the entire computational domain and plays the role of the root node in the grid hierarchy.  Subgrids are placed such that they cover (hopefully small) sub-volumes of the root grid at higher resolution.  This structure forms a tree may be extended to arbitrary depths by adding additional grids to refine regions in the subgrids.  We use the following notation to describe grid relationships.  A grid's parent is that coarse grid which completely contains it.  The root grid has no parent.  A child grid is the opposite (note that a grid may have many children but only one parent), while sibling grids are those which have the same resolution, but not necessarily the same parent.  We also refer to levels of the hierarchy, where the root grid is labeled level zero and all grids with the same resolution are on the
same level.

In order to simplify bookeeping, a number of restrictions are imposed on the subgrids:
\begin{itemize}
 \item The refinement factor $r$, or the ratio of the coarse cell width to the fine cell width, must be an integer.  In practice, we also always use $r=2$, since tests have shown this provides the best performance.
 \item Each subgrid must begin and end on the boundary of a coarse cell.  This means that the number of cells in a patch will be a multiple of the refinement factor.
 \item A subgrid most be completely enclosed by its parent.
\end{itemize}

The AMR grid patches are the primary data structure in \enzo.  Each patch is treated as an individual object which can contain both field variables and particle data.  Individual grids are organized into a  dynamic, distributed hierarchy of mesh patches.  
%See Figure~\ref{fig.2.amrhierdens} for an example of an \enzo\ simulation performed using this AMR scheme where both the baryon density and AMR grid hierarchy are shown together. 
% GB: I think this figure is too application-specific

Each grid patch in \enzo\ contains arrays of values for baryon and particle quantities.   Grids are partitioned into a core of \emph{real zones} and a surrounding layer of \emph{ghost zones}, as shown in Figure~\ref{fig:amr_hierarchy}.  The real zones store field values and ghost zones are used to temporarily store values which have been obtained directly from neighboring grids or interpolated from a parent grid.  These zones are necessary to accommodate the computational stencil of the hydrodynamics solvers (sections~\ref{sec.hydro.ppm} and~\ref{sec.hydro.zeus}) and the gravity solver (section~\ref{sec.gravity}).  The PPM and Zeus hydro solvers require 3 layers of ghost zones,  and the gravity solver requires 6 (although only for the density and potential fields).  This can lead to significant memory and computational overhead, particularly for smaller grid patches at high levels of refinement.

From the point of view of the hydro solver, each grid is solved as an independent CFD problem, with Dirichlet boundary conditions stored in the ghost zones.  At the end of each time step, cells on a given level fill ghost zones by interpolating from the parent and copying from neighbors on the same level.  Grids then correct the flux at the interface boundary between fine and coarse zones, and finally project their real zone data to the parent grid.

Timesteps are determined on each level as described in \ref{sec.timestepping}, and the hierarchy is advanced on
a level-by-level basis in a W-Cycle.    Beginning with the coarsest level, $l$, all grids on that level are advanced one
timestep.  Then one timestep is taken on all grids at the next level of refinement, $l+1$, and so on until the finest level is resolved. The finest level is then advanced, using as many steps as it takes to reach the level immediately above.  The finest level is then synced to the level above it, which then proceeds forewards one more step.  This is shown graphically in Figure~\ref{fig:wcycle}

\begin{figure}
\begin{center}
\includegraphics[width=0.6\textwidth]{figures/wcycle.eps}
\end{center}
\caption{\emph{Left:}  Example of a simple, distributed AMR hierarchy showing real and ghost grids.
\emph{Right:}  Example 2D \enzo\ grid showing real and ghost zones, as 
needed for the PPM hydro stencil. }
\label{fig:wcycle}
\end{figure}

At the end of every timestep on every level, each grid updates its ghost zones by exchanging information with its neighboring grid patches (if any exist) and/or by interpolating from a parent grid.  In addition, cells are examined to see where refinement is required and the entire grid hierarchy is rebuilt at that level (including all more highly refined levels).  The timestepping and hierarchy advancement/rebuilding process described here is repeated recursively on every level to the specified maximum level of refinement in the simulation.

The basic control algorithm looks very much like any used in a single
grid scheme, and can be written schematically as:
\begin{verbatim}
InitializeHierarchy
While (Time < StopTime)
begin
  dt = ComputeTimeStep(0)
  EvolveLevel(0, dt)
  Time = Time + dt
  CheckForOutput(Time)
end
\end{verbatim}
The AMR control algorithm is contained within the recursive
EvolveLevel algorithm, which takes the level on which to operate as an
argument.  This looks something like the following pseudo-code:
\begin{verbatim}
EvolveLevel(level)
begin
  SetBoundaryValues
  while (Time < ParentTime)
  begin
    dt = ComputeTimeStep(level)
    SolveHydroEquations(dt)
    Time = Time + dt
    SetBoundaryValues
    EvolveLevel(level+1, dt)
    FluxCorrection
    Projection
    RebuildHierarchy(level+1)
  end
end
\end{verbatim}

Each function operates on all the grids on the given level.  In the following sections, we will address each of the major elements in turn.

\subsubsection{Setting the boundary values}
\label{sec:interpolation}

In order to solve the hydrodynamic equations on a patch, it is necessary to specify the boundary conditions on that patch.  There are two cases.  The first is that the boundary is external to the computational domain.  This occurs for all the boundaries of the root grid.  The second possibility is that the boundary is within the computational domain, which is true for most, but not all, subgrids.

In the first case, we allow for four types of boundary conditions:
\begin{enumerate}
  \item{\em reflecting.} In this case, the boundary is like a mirror, with the solution on the boundary reflecting the solution in the computational domain ($q(-x) = q(+x)$) but with the velocity normal to the boundary reversed: $v_x(-x) = -v_x(+x)$.
  \item{\em outflow.} True outflow conditions are extremely difficult (REF?).  We approximate them by duplicating the solution at the edge of the computational domain: $q(-x) = q(0)$.
  \item{\em inflow.} In this case, the boundary values are fixed by a pre-determined function ($q(-x) = q0(x,t)$).  The code provides a simple way to set inflow values which are constant in both time and space over an entire face, or hooks are provided for special cases where the boundary must be time-varying.
  \item{\em periodic.} The boundary solution is obtained from the other side of the grid: $q(-x) = q(x_{\rm{max}}-x)$.
\end{enumerate}

In the second case, the value must be either (1) interpolated from a parent, or (2) copied from a sibling.  In practice, we solve this problem by first interpolating all boundaries from the grid's parent and then performing the sibling copy wherever possible, since these values will always be as good as, or superior to, interpolation.

We focus first on conserved, cell-centered quantities.  The interpolation function should be conservative in the sense that the mean over the interpolated quantity, in a given cell, must return the cell-centered value.  It should also be monotonic, so that new extrema are not introduced into a smoothly varying field.

In one dimension, such functions are well known; examples include
donor cell, Van-Leer and piecewise parabolic interpolation, which are
first, second and third-order accurate respectively \citep[e.g.,][]{Stone92a}.
However, we require fully multi-dimensional interpolation functions, and so
focus on second-order schemes since they are tractable and
relatively accurate. 
\footnote{We could, in principle, build a multi-dimensional interpolator out
of a one-dimensional interpolation function; however, at higher order, this would not necessarily
satisfy our constraints on not introducing new extrema.}
In Appendix~\ref{app:interpolation} we explicitely write down
functions in one, two and three dimensions, but here we discuss the
general design considerations.

In two dimensions, it is relatively easy to build a linear, monotonic,
conservative function by requiring, first, that the function at the
center of the cell return the cell-centered quantitiy (and since it is
linear, this implies that an average over the cell will also be equal
to this value), and second, by treating the cell-diagonals as
one-dimensional Van-Leer type functions.  The reason to focus on the
cell diagonals is that for a linear function, the extrema will occur at the cell 
corners, which are in turn linked by the diagonals.  This process can be done in
two dimensions since there are three constraints (one at the center and one along
each diagonal), and three parameters in a two-dimensional linear
function.

In three dimensions this process breaks down since the number of
constraints (center + four diagonals = five) exceeds the number of
parameters (four).  One way to solve this problem would be to perform
a least-squares fit under the given constraints, however, this is
computationally too expensive.  Therefore, in the appendix we describe
a heuristic approach which satisfies four constaints analytically and
then adjusts the interpolation parameters in such a way as to satisfy the fifth.
While the result is not necessarily optimal, it does seem to perform
adequately in our tests.

%The same procedure can be used for face-centered quantities; however, for the velocity variable in zeus, we drop the extrema constraint.

\subsubsection{Solving the hydrodynamic equations on a grid}
\label{sec:solve_hydro}

Given the grid and it's boundary conditions, the process of solving equations~(\ref{eq:mass_prime})--(\ref{eq:total_energy_prime}) can be detached from the rest of the SAMR technique, so that it is possible to experiment with a number of different methods.  The restrictions are that the method should be flux-conservative, or at least produce an accurate estimate of the flux of conserved quantities at cell interfaces.  This is required in the flux-correction step (see \ref{sec:flux_correction}).  There is also the issue of centering, which refers to the position of physical quantities within a cell (the primary types are cell-centered, face-centered, and node-centered). We will employ both the direct Eulerian version of the piecewise parabolic method \citep{1984JCoPh..54..174C,1995CoPhC..89..149B} which assumes cell-centering for all quantities (actually they are cell-averages).  However, it is also possible to use other schemes, which employ different centering for different variables \citep[e.g.,][]{Stone92a}.   The interpolation step, described in section~\ref{sec:interpolation}, must correctly interpolate with the selected centering.

\subsubsection{Projection}
\label{sec:projection}

Since a refined region is simulated with (at least) two different resolutions (one fine and one coarse), it has two or
more ``solutions'' (i.e. values of density, velocity, etc.) which represent the same volume of space.  This is one important aspect of the SAMR method -- some physical regions are represented by grids at various resolutions.  To
maintain consistency, the coarse grid quantities are updated with the finer values, using a volume-weighted average of the conserved quantities:
\begin{equation}
q^{\rm coarse} = r^{-d} \sum q^{\rm fine}_{i, j,k}
\end{equation}
where $r$ is the refinement factor, $d$ the dimensionality, and the sum is over all fine cells which ``cover" the coarse cell.

\subsubsection{Flux correction}
\label{sec:flux_correction}

One of the big advantages a flux-conservative methods is, of course, their ability to maintain conserved-quantities to machine precision.  Unfortunately, the projection step upsets this flux conservation since, at the boundary of a refined region, the coarse and fine cells on either side of the border are updated with different estimates of the flux across the boundary.  In one case this comes from the coarse grid computation, and in the other from the fine grid solution.  Conservation can be restored by correcting all coarse cells which lie at the boundary of a fine region
with the difference between the coarse and fine estimates of the flux across the boundary:
\begin{equation}
q^{\rm coarse} = \tilde{q}^{\rm coarse} - \Delta t \left( F^{\rm coarse} - \sum_{j,k} F^{fine}_{j,k} \right)
\end{equation}
where $F$ is the flux of quantity $q_i$ in the $i$-direction through the cell interface.   We use $\tilde{q}$ to indicate the uncorrected quantity on the coarse grid and the sum is over the $r^{d-1}$ fine cell interfaces in the perpendicular dimensions which share a face with the coarse cell being corrected.  This corrected is carried out for all coarse cells which share a face with a refined cell (and are not themselves covered with more refined cells).

\subsubsection{Hierarchy reconstruction}
\label{sec:hierarchy_reconstruction}

Since the addition of more highly refined grids is adaptive, the conditions for refinement must be specified.  Up to five different criteria can be specified, and these refinement criteria are discussed in more detail in the following section (section~\ref{sec:refinement_criteria}).   Once all cells in a given grid that require additional refinement have been flagged, rectangular boundaries are determined which minimally encompass the flagged cells.  This is done using an iterative machine-vesion technique as described in (REF).

We first find the smallest region (or ``proto-subgrid") which covers all of the refined cells on a grid.  We compute an efficiency $\epsilon$ for this grid, defined as the ratio of flagged cells to total cells.  If this efficiency is sufficiently larger (typically larger than 30\%), or if the grid is small enough (typically less $4^d$ coarse cells) then we adopt the region as a new grid and move on to the next region (if any).  If the region is not acceptable, then we try to split it into two using the following algorithm.  First, we determine the longest dimension and compute the signature along that dimension.  The signature is found by summing the number of refined cells with the same index along the longest dimension.  Edges are typically indicates by inflection points in this signature, so we calculate the second derivative using a simple three-point finite difference estimator and look for the maximum in the second derivative, with ties going to the index closest to the center.  The original region is then split into two new regions, or proto-subgrids.  This process is then re-applied to each of the new regions.

This algorithm is not necessarily optimal, but experience shows it does a reasonable job of generating an efficient covering of the refined cells.  Note that some cells are unnecessarily refined with the SAMR approach, resulting in wasted resources (unlike in cell-splitting AMR schemes).  However, it has a number of advantages.  First, traditional grid-based hydro method can be simply added with only minor modifications.  Second, it is easier to accommodate larger stencils (as in PPM, which requires three boundary zones), and finally it can be more efficient for some machines since there are more predictable and linear memory accesses.

\subsubsection{Refinement criteria}
\label{sec:refinement_criteria}

Finally, we turn to the refinement criteria themselves.  Although generalized refinement criteria are available based on estimators for the error, such as the Richardson method (REF), we have focused generally on methods which use physical estimators on the assumption that the simulator has a specific quantity to the indicates refinement.  The flagging methods currently built into \enzo\ are:

\begin{enumerate}

\item{\em Slope.}  If the normalized slope $(q_{i+1} - q_{i-1})/ (2 q_i)$ is larger than a specified value (typically 0.3), then the cell is flagged.

\item{\em Baryon mass.}  This refinement criteria is identified to mimic a Lagrangian method in that it tries to keep a fixed mass resolution.  In this method, a cell is refined if the baryonic mass in the cell ($M_g = \rho \Delta_x^d$) is larger than a specified value:
\begin{equation}
M_g > \rho_{\rm flag} {\Delta x^d_{\rm root}} r^{\epsilon_l l},
\end{equation}
where $\rho_{\rm flag}$ is the equivalent density on the root grid required for refinement, $\Delta x_{\rm root}$ is the root grid cell spacing, $r$ is the refinement factor, $l$ is the level and $\epsilon_l$ is a parameter which can be used to make the refinement faster ($\epsilon_l < 0$) or slower than pure Lagrangian refinement ($\epsilon_l > 0$).

\item{\em Shocks.}  In this method, we identify and refine shocks with the following criteria:
\begin{eqnarray}
(p_{i+1} - p_{i-1})/\min(p_{i+1}, p_{i-1}) & > & 0.33  \nonumber \\
u_{i-1} - u_{i+1} & > & 0 \\
e_i / E_i & > & 0.1  \nonumber
\end{eqnarray}

\item{\em Particle (dark matter) mass.}  This refinement criteria is precisely the same as the baryon overdensity criteria except it uses the grided (CIC) particle density.

\item{\em Jeans Length.}  Following \cite{Truelove98}, the code can be forced to refine the Jeans length by a fixed number of cells, or, in more specifically, whenever the following criteria is not met:
\begin{equation}
\Delta x < \left( \frac{\pi k_B T}{J^2 G \rho m_H} \right)^{1/2}
\end{equation}
where $J$ is the required number of cells per Jeans length (4 by default).

\item{\em Cooling time.}  Often in cosmological simulations, cooling is a rapid process and we do not necessarily want to be limited by the cooling time.  However, in some cases it is desirable to resolve the cooling process, in which case we flag on cells which have a cooling time shorter than their sound crossing time.  The cooling time is given by $t_{\rm cool} = 1.5 kT / (n \Lambda(T)$ and the sound crossing time is $t_{\rm cross} = \Delta x / c_s$, where $c_s$ is the sound speed.

\item{\em Must-refine particles}  In order to allow for refinement in moving regions, it is possible to generate ``must-refine" particles which for local refinement over the 8 nearest cells to a specific level.  The particles feel on gravitational forces and their trajectories are followed as for any other collisionless particle.  This feature can be used to ensure that a specific Lagrangian region stays refined to a specified level of resolution.

\item{\em Shear} As described in \citet{KritsukXX}, turbulence calculations can be successfully modeled with AMR if the shear is used as a refinement parameter; in particular, refinement occurs when the finite difference version of the following inequality holds:
\begin{equation}
%\left(\frac{du}{dy}\right)^2 + \left(\frac{du}{dz}\right)^2 + 
%\left(\frac{dv}{dx}\right)^2 + \left(\frac{dv}{dz}\right)^2 + 
%\left(\frac{dw}{dx}\right)^2 + \left(\frac{dw}{dy}\right)^2 
\sum_i \sum_j \left( \frac{\partial v_i}{\partial x_j} \right)^2 -  \sum_i \left( \frac{\partial v_i}{\partial x_i} \right)^2
> \epsilon_s \frac{c_s^2}{\Delta x^2},
\end{equation}
where $\epsilon_s$ is a dimensionless parameter.

\end{enumerate}

Finally, it is also possible to specify a specific rectangular region which is statically refined at a certain level.  These static refined regions are in place throughout the simulation regardless of the grid quantities and can be used for testing purposes, or for multiply-zoomed initial conditions (for example in cosmology simulations) that would not otherwise be refined.

%GB: The following section has now been worked into the main text above and, I think, can be removed.
%
%\subsection{Adaptive mesh refinement}\label{sec.ov.amr}
%\dcc{removed most of the notes.  Added intro blurb.  moved parallelism
%  stuff. Re-worded the bit on ghost zones.  Summary of 'grid as
%  independent update + syncing with neighbors and parents,' W-cycle. Added
%  sentence on subgrid creation.  Made timestepping its own subsection.}
%\red{Talk about reasonable refinement factors - which 
%ones are usable, etc?}

%The primary purpose of the \enzo\ code is its Adaptive Mesh Refinement
%capability, which allows it to reach extremely large dynamical ranges
%with limited computational resources, opening doors previously closed
%by finite memory and computational time. Unlike moving mesh methods
%\citep{1995ApJS..100..269P,1995ApJS...97..231G} or  
%methods that subdivide 
%individual cells \citep{Adjerid}, Berger \& Collela's AMR (also referred 
%to as \emph{structured} AMR) utilizes an adaptive hierarchy of grid 
%patches at varying levels of resolution.  Each rectangular grid patch 
%(referred to as a ``grid'') covers some region of space in its 
%\emph{parent grid} which requires higher resolution, and can itself 
%become the parent grid to an even more highly resolved \emph{child grid}. 
%ENZO's implementation of structured AMR places no restrictions on 
%the number of grids at a given level of refinement, or on the number of 
%levels of refinement.  However, owing to limited computational resources 
%it is practical to institute a maximum level of refinement $\Lmax$.

%The \enzo\ implementation of AMR allows arbitrary integer ratios of parent and
%child grid resolution.  However, in practice 
%refinement factors (defined as the ratio of parent grid resolution to child grid resolution) 
%of two or four are typically used, with a refinement factor
%of two being most commonly used for cosmological simulations for reasons of
%efficiency. The ratio of
%box size to the maximum grid resolution of a given simulation is
%therefore $L/e = \Nroot \times 2^{\Lmax}$, where $\Nroot$ is the
%number of cells along one edge of the root grid, $\Lmax$ is the
%maximum level of refinement allowed, and L and e are the box size
%and grid resolution of the most highly refined region, respectively.

%The AMR grid patches are the primary data structure in \enzo.  Each
%patch is treated as an individual object which can contain both field
%variables and particle data.  Individual grids are organized into a 
%dynamic, distributed hierarchy of mesh patches.  
%See 
%Figure~\ref{fig.2.amrhierdens} for an example of an \enzo\
%simulation performed using this AMR scheme where both the baryon density
%and AMR grid hierarchy are shown together.

%\begin{figure}
%\begin{center}
%\includegraphics[width=0.4\textwidth]{figures/amr-hier-dens.eps}
%\caption{Example of an \enzo\ simulation showing the AMR grid hierarchy.  
%This is a simulation of the collapse of a single massive halo with a 
%$32^3$ root grid and two $32^3$ static nested grids.  AMR is only allowed
%within the most highly refined static nested grid.  Log baryon density is shown
%by colored isocontours with values corresponding to the legend at center left.  
%The rectangular solid wire frames correspond to individual 
%AMR grids, with different colors corresponding to level as indicated by the
%legend at top left.}
%\label{fig.2.amrhierdens}
%\end{center}
%\end{figure}

%Each grid patch in \enzo\ contains arrays of values for baryon and 
%particle quantities.   Grids are 
%partitioned into a core of \emph{real zones} and a surrounding layer 
%of \emph{ghost zones}.  The real zones store field values and ghost 
%zones are used to temporarily store values which have been obtained 
%directly from neighboring grids or interpolated from a parent grid.  
%These zones are necessary to accommodate the computational stencil of 
%the hydrodynamics solvers (Sections~\ref{sec.ov.hydro.ppm} and~\ref{sec.ov.hydro.zeus}) 
%and the gravity solver (Section~\ref{sec.ov.nbody}).  
%The PPM and Zeus hydro solvers require 3 layers of ghost zones,  and the gravity
%solver requires 6.  This can lead to significant memory 
%and computational overhead, particularly for smaller grid patches
%at high levels of refinement.  More information on filling of ghost
%zones can be found in the appendix.

%From the point of view of the hydro solver, each grid is solved as an
%independent CFD problem, with Dirichlet boundary conditions stored in
%the ghost zones.  At the end of each time step, cells on a given level
%fill ghost zones by interpolating from the parent and copying from
%neighbors on the same level.  Grids then correct the flux at the interface boundary between
%fine and coarse zones, and finally project their real zone data to the
%parent grid.

%Timesteps are determined on each level as
%described in \ref{sed.timestepping}, and the hierarchy is advanced on
%a level by level basis in what's known as a W-Cycle.    Beginning with
%the coarsest level, $L$, all grids on that level are advanced one
%timestep.  Then one timestep is taken on all grids at the next level
%of refinement, $L+1$, and so on until the finest level is resolved.
%The finest level is then advanced, using as many steps as it takes to
%reach the level immediately above.  The finest level is then synced
%to the level above it, which then proceeds forewards one more step.
%\red{Needs a picture of the W cycle.}

%At the end of every timestep on every level each grid updates 
%its ghost zones by exchanging information with its neighboring grid patches 
%(if any exist) and/or by interpolating from a parent grid.  
%In addition, cells are examined to see if they should be refined or 
%de-refined, and the entire grid hierarchy is rebuilt at that 
%level (including all more highly refined levels).  The timestepping and 
%hierarchy advancement/rebuilding process described here is repeated 
%recursively on every level to the specified maximum level of 
%refinement in the simulation.

%Since the addition of more highly refined grids is adaptive,  
%the conditions for refinement must be specified.  The criteria of 
%refinement can be set by the threshold value of the overdensity of 
%baryon gas or dark matter in a cell (which is really a refinement on
%the mass of gas or DM in a cell), the local Jeans length, the 
%local density gradient, or local pressure and energy gradients.  
%A cell reaching any or all of these criteria 
%will then be flagged for refinement.  
%Once all cells at a given level have been examined, rectangular boundaries 
%are determined which minimally encompass the flagged cells. The set of
%flagging cells is then further fractured into smaller bounding
%rectangles to ensure computational 
%efficiency.  More details on subgrid creation can be found in Appendix
%\ref{app.subgrid_creation}.  A refined grid patch is 
%introduced within each such bounding rectangle. Thus the cells needing 
%refinement, as well as adjacent cells within the patch which do not need 
%refinement, are refined. While this approach is not as memory efficient as 
%cell-splitting AMR schemes, it offers more freedom with 
%finite difference stencils. For example, PPM requires a stencil of seven 
%cells per dimension. This cannot easily be accommodated in cell-splitting 
%AMR schemes.

### 2012-06-29 numerical-chemistry.tex
first commit of paper.

\subsection{Chemistry, radiation and cooling}
\label{sec.ov.chem}

\red{
As this stands it is the methods that I (brian) use, so I don't really emphasize
either the tabular cooling functions or the various radiation background options.
Add more discussion of this, point towards appendix.
}

Though the equations of hydrodynamics described above are a closed system,
they are still missing a crucial piece of physics: radiative heating and cooling.
Radiative cooling is extremely important in many astrophysical situations, as is 
the heating of gas from some sort of radiation background.  \enzo\ has a very simple
Sutherland and Dopita equilibrium cooling function \citep{1993ApJS...88..253S} 
implemented, which uses a
cooling table assuming a fixed metallicity of $Z = 0.3 Z_\odot$, and also a
nonequilibrium heating/cooling model that assumes gas of primordial composition exposed to a
uniform metagalactic ultraviolet (UV) background that varies with time \citep{1996ApJS..105...19K}.

The simulations discussed in this thesis almost exclusively use the nonequilibrium
routines, described in great detail by Abel et al. and Anninos et al.  
\citep{abel97,anninos97} and summarized in Appendix~\ref{app:chemistry}. These 
routines follow the non-equilibrium chemistry of a 
gas of primordial composition with 9 total species:  
$H$, $H^+$, $He$, $He^+$, $He^{++}$, $H^-$, $H_2^+$, $H_2$, and $e^-$.  The code also calculates 
radiative heating and cooling, following atomic line excitation, recombination,
collisional excitation, free-free transitions, molecular line excitations, and Compton
scattering of the cosmic microwave background, as well as any of
approximately a dozen different models for a metagalactic UV background that heat
the gas via photoionization and photodissociation. 
The multispecies rate equations are solved out of
equilibrium to properly model situations where, e.g., the cooling rate of the gas
is much shorter than the hydrogen recombination time.  The effect of this nonequilibrium
cooling is to leave behind a much larger fraction of residual free electrons than one would
expect if the assumption of equilibrium were being made.  The practical effect of this
is that more $H^-$ is formed, which subsequently produces hydrogen molecules.  If large
amounts of $H_2$ is formed it can greatly increase the cooling rate of primordial gas
at relatively low temperatures ($T \leq 10^4$ K).  This can efficiently cool the gas 
to approximately $200$ K, which significantly reduces the Jeans mass of the gas.  Correct
modeling of the formation of molecular hydrogen is crucial to the study of star formation
in a primordial gas.

A total of 9 kinetic equations are solved, including 29 kinetic and radiative 
processes, for the 9 species mentioned above.  See Table~\ref{table.collisional} 
for the collisional processes and Table~\ref{table.radiative} for the 
radiative processes solved.

%---------------- table of collisional processes

\begin{table}
\begin{center}
{\bfseries Collisional Processes}\\[1ex]
\begin{tabular}{llllllll}
(1) & H & + & e$^-$ & $\rightarrow$ & H$^+$ &+& 2e$^-$ \\
(2) & H$^+$ &+ &e$^-$ & $\rightarrow$ & H &+ &$\gamma$ \\
(3) & He &+& e$^-$ & $\rightarrow$ & He$^+$ &+& 2e$^-$  \\
(4) & He$^+$ &+& e$^-$ & $\rightarrow$ & He &+ &$\gamma$  \\
(5) & He$^{+}$ &+& e$^-$ & $\rightarrow$ & He$^{++}$ &+& 2$e^-$  \\
(6) & He$^{++}$ &+& e$^-$ & $\rightarrow$ & He$^+$ &+& $\gamma$ \\
(7) & H &+& e$^-$ &$\rightarrow$& H$^-$ &+& $\gamma$  \\
(8) & H$^-$ &+& H &$\rightarrow$ & H$_2$ & +& e$^-$ \\
(9) & H &+ &H$^+$ &$\rightarrow$ &H$_2^+$ &+ &$\gamma$ \\
(10) & H$_2^+$ &+ &H &$\rightarrow$ &$H_2$ &+ &$H^+$ \\
(11) & H$_2$ &+ &H$^+$ &$\rightarrow$ &H$_2^+$ & +& H \\
(12) & H$_2$ &+ &e$^-$ & $\rightarrow$ & 2H & + & e$^-$  \\
(13) & H$_2$ & + & H & $\rightarrow$ & 3H &   &      \\
(14) & H$^-$ & + & e$^-$ & $\rightarrow$ & H & + & 2e$^-$ \\
(15) & H$^-$ & + & H & $\rightarrow$ & 2H & + & e$^-$ \\ 
(16) & H$^-$ & + & H$^+$ & $\rightarrow$ & 2H & & \\
(17) & H$^-$ & + & H$^+$ & $\rightarrow$ & H$_2^+$ & + & e$^-$ \\
(18) & H$_2^+$ & + & e$^-$ & $\rightarrow$ & 2H & & \\
(19) & H$_2^+$ & + & H$^-$ & $\rightarrow$ & H$_2$ & + & H  \\
(20) & 3H & & & $\rightarrow$ & H$_2$ & + & H
\end{tabular}
\caption[]{Collisional processes solved in the Enzo nonequilibrium
primordial chemistry routines.}
\label{table.collisional}
\end{center}
\end{table}

\begin{table}
\begin{center}
{\bfseries Radiative Processes}\\[1ex]
\begin{tabular}{llllllll}
(21) & H & + & $\gamma$ & $\rightarrow$ & H$^+$ & + & e$^-$ \\
(22) & He & + & $\gamma$ & $\rightarrow$ & He$^+$ & + & e$^-$ \\
(23) & He$^+$ & + & $\gamma$ & $\rightarrow$ & He$^{++}$ & + & e$^-$ \\
(24) & H$^-$ & + & $\gamma$ & $\rightarrow$ & H & + & e$^-$ \\
(25) & H$_2$ & + & $\gamma$ & $\rightarrow$ & H$_2^+$ & + & e$^-$ \\
(26) & H$_2^+$ & + & $\gamma$ & $\rightarrow$ & H & + & H$^+$ \\
(27) & H$_2^+$ & + & $\gamma$ & $\rightarrow$ & 2H$^+$ & + & e$^-$ \\
(28) & H$_2$ & + & $\gamma$ & $\rightarrow$ & H$_2^*$ & $\rightarrow$ & 2H \\
(29) & H$_2$ & + & $\gamma$ & $\rightarrow$ & 2H &  & 
\end{tabular}
\caption[]{Radiative processes solved in the Enzo nonequilibrium
primordial chemistry routines.}
\label{table.radiative}
\end{center}
\end{table}

The chemical reaction equation network is technically challenging to solve due to 
the huge range of reaction timescales involved -- the characteristic creation
and destruction timescales of the various species and reactions can differ by 
many orders of magnitude.  As a result, the set of rate equations is extremely 
stiff, and an explicit scheme for integration of the rate equations can be 
exceptionally costly if small enough timesteps are taken to keep the network
stable.  This makes an implicit scheme much more preferable for such a set of 
equations.  However, an implicit scheme typically require an iterative 
procedure to converge, and for large networks (such as this one) an implicit
method can be very time-consuming, making it undesirable for a large, three-dimensional
simulation.

\enzo\ solves the rate equations using a method based on a backwards differencing 
formula (BDF) in order to provide a stable and accurate solution.  This technique is
optimized by taking the chemical intermediaries $H^-$ and $H_2^+$, which have 
large rate coefficients and low concentrations, and grouping them into a separate
category of equations.  Due to their fast reactions these species are very sensitive
to small changes in the more abundant species.  However, due to their low overall
concentrations, they do not significantly influence the abundance of species with
higher concentrations.  Therefore, reactions involving these two species can be
decoupled from the rest of the network and treated independently.  In fact, $H^-$ 
and $H_2^+$ are treated as being in equilibrium at all times, independent of 
the other species and the hydrodynamic variables.  This allows a large speedup
in solution as the BDF scheme is then applied only to the slower 7-species network
on timescales closer to those required by the hydrodynamics of the simulation.
Even so, the accuracy and stability of the scheme is maintained by subcycling the 
rate solver over a single hydrodynamic timestep.  These subcycle timesteps are 
determined so that the maximum fractional change in the electron concentration is
limited to no more than $10\%$ per timestep.

It is important to note the regime in which this model is valid.  According to Abel et al. and
Anninos et al. \citep{abel97,anninos97}, the reaction network is valid for temperatures
between $10^0 - 10^8$ K.  The original model discussed in these two references is only
valid up to $n_H \sim 10^4$~cm$^{-3}$.  However, addition of the 3-body $H_2$ formation
process (equation 20 in Table~\ref{table.collisional}) allows 
correct modeling of the chemistry of the gas up
until the point where collisionally-induced emission from molecular hydrogen becomes an important
cooling process, which occurs at $n_H \sim 10^{14}$~cm$^{-3}$.  A further concern is that
the optically thin approximation for radiative cooling breaks down, which
occurs before $n_H \sim 10^{16} - 10^{17}$~cm$^{-3}$.  Beyond this point, 
modifications the cooling function that take into account the non-negligible
opacity in the gas must be made, as discussed by \citet{2004MNRAS.348.1019R}. 
Even with these modifications, a completely correct description of the cooling of
this gas will require some form of radiation transport, which will greatly 
increase the cost of the simulations.

Several processes are neglected.  The deuterium atom and its processes are 
completely ignored, which may have some effect.  Recent work shows that HD is
a more effective coolant than previously thought~\citep{2005MNRAS.361..850L}.  However,
the fractional abundance of HD is so low that under circumstances relevant to
the formation of a Population III star in an un-ionized region it should be
sub-dominant.  However, the enhanced electron fraction in fossil 
HII regions (as discussed later in this thesis) could result in the HD molecule
becoming a dominant cooling mechanism at relatively low ($\sim$ few hundred K)
temperatures, and could potentially cool the gas down to below $100$~K, which can
enhance fragmentation and could have important consequences for the IMF of 
primordial stars forming in a relic HII region~\citep{2005MNRAS.364.1378N}.

Aside from deuterium, the chemical reactions involving lithium 
are also neglected.  According to \citet{1998A&A...335..403G}, these are 
not important for the density and temperature regimes explored 
by the simulations discussed in this thesis.  However, at higher densities it 
is possible that there are regimes where lithium can be an important coolant.

### 2012-06-29 numerical-gravity.tex
first commit of paper.

\subsection{Gravity: Solving the Poisson Equations}
\label{sec.gravity}

There are multiple methods to compute the gravitational potential (which is an elliptic equation in the Newtonian limit) in a structured AMR framework.  One way would be to model the dark matter (or other collisionless particle-like objects, such as stars) as a second fluid in addition to the baryonic fluid and solve the collisionless Boltzmann equation, which follows the evolution of the fluid density in six-dimensional phase space.  However, this is computationally prohibitive owing to the large dimensionality of the problem, making this approach unappealing for the cosmological AMR code.

The dark matter particles are distributed onto the grids using the cloud-in-cell (CIC) interpolation technique to form a spatially discretized density field (analogous to the baryon densities used to calculate the equations of hydrodynamics).  After sampling dark matter density onto the grid and adding baryon density if it exists (to get the total matter density in each cell), the gravitational potential is calculated on the periodic root grid using a fast Fourier transform.  For periodic boundary conditions, we can use either a simple Greens function kernel of $-k^{-2}$, or the finite-difference equivalent \citep{HockneyEastwoord1980}:
\begin{equation}
G(\vec{k}) = - \frac{\Delta x}{2 \left( sin(k_x \Delta x/2)^2 + sin(k_y \Delta y/2)^2 + sin(k_z \Delta z/2)^2 \right) }
\end{equation}
where $k^2 = k_x^2 + k_y^2 + k_z^2$ is the wavenumber in Fourier space and the potential is calculated in k-space as usual with $\tilde{\phi}(k) = G(k) \tilde{\rho}(k)$.

For isolated boundary conditions, we use James method (REF).  In this case, the Greens function is generated in real-space so as to have the correct zero-padding properties and then transformed into the Fourier domain.

In order to calculate more accurate potentials on subgrids, \enzo\ re-samples the dark matter density onto individual subgrids using the same CIC method as on the root grid, but at higher spatial resolution (and again adding baryon densities if applicable). Boundary conditions are then interpolated from the potential values on the parent grid.  We use either tri-linear interpolation or a natural second-order spline: both methods seem to give similar results. The potential equation on each subgrid is then solved with the given Dirichlet boundary conditions.  This solution is done with a multigrid relaxation technique (but only applied to each subgrid).

The region immediately next to the boundary can contain unwanted oscillations (Anninos \etal REF), and so we use an expanded buffer zone around the grid, of size three root grid boundary zones (so typically six refined zones).  The density is computed in this region and the potential solved, but only the region which overlaps with the grid itself is used to calculate accelerations.

It is known (REF) that simply interpolating the potential without feeding it back to higher levels leads to errors in the potential at more refined levels because of the build-up of errors during the interpolation of coarse boundary values.  
In addition, neighboring subgrids are not guaranteed to generate the same potential values because of the lack of a coherent potential solve across the whole hierarchy. In an attempt to partially alleviate this problem, we allow for an iterative procedure across sibling grids, in which the potential values on the boundary of grids can be updated with the potential in `active' regions of neighboring subgrids.  To prevent overshoot, we average the potential on the boundary and allow for (typically) 4 iterations.  This procedure can help in many cases, but does not produce a coherent solution across all grids and so does not solve the problem; we are working on a slower but more accurate method which does a multi-grid solve across the whole grid (Reynolds \etal, in preparation).

Forces are computed on the mesh by finite-differencing potential values and are then interpolated to the particle positions, where they are used to update the particle's position and velocity information.  Potentials on child grids are computed recursively and particle positions are updated using the same timestep as in the hydrodynamic equations.

At this point it is useful to emphasize that the effective force resolution of an adaptive particle-mesh calculation is approximately twice as coarse as the grid spacing at a given level of resolution.  
%The potential is solved in each grid cell; however, the quantity of interest, namely the acceleration, is the gradient of the potential, and hence two potential values are required to calculate this.  In addition, it should be noted that the adaptive particle-mesh technique described here is very memory intensive: in order to ensure reasonably accurate force resolution at grid edges the multigrid relaxation method used in the code requires a layer of ghost zones which is very deep -- typically 6 cells in every direction around the edge of a grid patch.  This greatly adds to the memory requirements of the simulation, particularly because subgrids are typically small (on the order of $12^3 - 16^3$ real cells for a standard cosmological calculation) and ghost zones can dominate the memory and computational 
requirements of the code.

### 2012-06-29 numerical-overview.tex
first commit of paper.

\section{Overview of Numerical Methods}\label{sec.overview}

%---------------------------------------------------------------------------%
\subsection{N-body dynamics}\label{sec.ov.nbody}
fft + multigrid

There are multiple methods to compute the gravitational potential 
(which is an elliptic equation in the Newtonian limit) in a structured 
AMR framework.  One way would be to model the dark matter (or other
collisionless particle-like objects, such as stars) as a second fluid 
in addition to the baryonic fluid and solve the collisionless
Boltzmann equation, which follows the evolution of the fluid density in
six-dimensional phase space.  However, this is computationally prohibitive 
owing to the large dimensionality of the problem, making this approach 
unappealing for the cosmological AMR code.

Instead, \enzo\ uses a particle-mesh N-body method to calculate 
the dynamics of collisionless systems \citep{Hockney88}.  This method 
follows trajectories 
of a representative sample of individual particles and is much more 
efficient than a direct solution of the Boltzmann equation in most 
astrophysical situations. 
The gravitational potential is computed by solving the elliptic 
Poisson's equation (Eqtn.~\ref{enzopoisson}).

These equations are finite-differenced and for simplicity are solved with the
same timestep as the hydrodynamic equations, as discussed in 
Section~\ref{sec.ov.amr}.
The dark matter particles are distributed onto the grids
using the cloud-in-cell (CIC) interpolation technique to form a spatially
discretized density field (analogous to the baryon densities used to calculate
the equations of hydrodynamics).  After sampling dark matter density onto the
grid and adding baryon density if it exists (to get the total matter density
in each cell), the gravitational potential is calculated on the periodic root
grid using a fast Fourier transform.  In order to calculate more accurate
potentials on subgrids, \enzo\ re-samples the dark matter density onto
individual subgrids using the same CIC method as on the root grid, but at
higher spatial resolution (and again adding baryon densities if applicable).
Boundary conditions are then interpolated from the potential values on the
parent grid (with adjacent grid patches on a given level communicating to
ensure that their boundary values are the same), and then a multigrid
relaxation technique is used to calculate the gravitational potential at every
point within a subgrid.  Forces are computed on the mesh by
finite-differencing potential values and are then interpolated to the particle
positions, where they are used to update the particle's position and velocity
information.  Potentials on child grids are computed recursively and particle
positions are updated using the same timestep as in the hydrodynamic
equations.  Particles are stored in the most highly refined grid patch at the
point in space where they exist, and particles which move out of a subgrid
patch are sent to the grid patch covering the adjacent volume with the finest
spatial resolution, which may be of the same spatial resolution, coarser, or finer
than the grid patch that the particles are moved from.  This takes place in a
communication process at the end of each timestep on a level.

At this point it is useful to emphasize that the effective force
resolution of an adaptive particle-mesh calculation is approximately twice
as coarse as the grid spacing at a given level of resolution.  The 
potential is solved in each grid cell;
however, the quantity of interest, namely the acceleration, is the gradient
of the potential, and hence two potential values are required to calculate
this.  In addition, it should be noted that the adaptive particle-mesh
technique described here is very memory intensive: in order to ensure reasonably 
accurate force resolution at grid edges the multigrid relaxation
method used in the code requires a layer of ghost zones which is very deep --
typically 6 cells in every direction around the edge of a grid patch.  This greatly
adds to the memory requirements of the simulation, particularly because subgrids
are typically small (on the order of $12^3 - 16^3$ real cells for a standard 
cosmological calculation) and ghost zones can dominate the memory and computational 
requirements of the code.

%---------------------------------------------------------------------------%
\subsection{Hydrodynamics}\label{sec.ov.hydro}

\subsubsection{The PPM hydro method}\label{sec.ov.hydro.ppm}

\dcc{Removed ``things to do'' list.  Mention ppm-lr and ppm-de, and
  that ppm-de is preferred.  Briefly describe ppm-de itself.}

The primary hydrodynamic method used in \enzo\ is based on the
piecewise parabolic method (PPM) of
\citet{1984JCoPh..54..174C}, which has been
modified for the study of cosmological fluid flows.  
\citet{1984JCoPh..54..174C} describe two variants on this method: a
Lagrangian Remap variant and a Direct Eulerian variant.  Both have
been adapted for \enzo, and the Lagrangian Remap 
variant is described in \citet{1995CoPhC..89..149B}.  Because of its
conservative formulation, however, the Direct Eulerian variant is
preferred for \enzo\ AMR simulations.  We will describe the Direct
Eulerian (PPM-DE) variant and its additions briefly here, and in
further depth in Appendix \ref{app.ppm}

PPM is an explicit, higher order-accurate version of
Godunov's method for ideal gas dynamics with third order-accurate piecewise parabolic
monotonic interpolation and a nonlinear Riemann solver for shock
capturing.  It advances the hydrodynamic equations in the following
steps:
\begin{enumerate}
 \item Construct monotonic parabolic interpolation of cell average
 data, for each fluid quantity.
 \item Compute interface states by averaging the parabola for over the domain of dependence for
 each interface
 \item Use interface data to solve the Riemann problem.
 \item Difference the interface fluxes to update the cell average quantities.
\end{enumerate}
It does an excellent job of capturing strong shocks in at
most two cells.  Multidimensional schemes are built up by directional
splitting and produce a method that is formally second order-accurate
in space and time which explicitly conserves mass, linear momentum, and energy.

To use the above algorithm for cosmological simulations, the
conservation laws for fluid mass, momentum and energy 
density are written in comoving coordinates for a
Friedman-Robertson-Walker space-time, as described previously 
in Equations~\ref{enzoconserve} 
through~\ref{enzoenergy}.  Both the conservation laws and
the Riemann solver are modified to include gravity, which is solved
using an adaptive particle-mesh (PM) technique (see Section~\ref{sec.ov.nbody}).  
The terms due to cosmological expansion, as well as primordial chemistry and
radiative heating and cooling, are solved in a separate step because they have
different numerical behavior, and therefore must be treated differently to ensure
stability.  Note that unlike the \zeus\ hydro scheme, PPM does not need to use 
artificial viscosity to resolve shocks.

The system of equations described above works well for systems with
relatively low Mach numbers, as long as these systems are well resolved.  
However, cosmology is replete with situations where there are bulk hypersonic
flows.  In these situations, the ratio of kinetic to thermal energy can be
very high -- in some situations up to $10^6 - 10^8$.  This implies that the
thermal energy is an extremely tiny fraction of the kinetic energy, which can 
cause numerical problems when one is interested in just the thermal energy of the gas, since
 Equation~\ref{enzoenergy} solves for the total energy.  In this system of
equations, the
 thermal energy $E_{therm}$ is calculated as $E - E_{kin}$, 
where E is the total specific energy as
calculated in equation~\ref{enzoenergy} and $E_{kin}$ is the specific kinetic energy,
$0.5 \vec{v}_b^2$.  
In hypersonic flows $E$ and $E_{kin}$ are nearly the same, and any 
number calculated as 
the difference of these is going to be strongly affected by numerical error.  To 
avoid this problem, \enzo\ also solves the internal energy equation in
comoving coordinates:

\begin{equation}
\frac{\partial e}{\partial t} + \frac{1}{a} \vec{v}_b \cdot \nabla e = 
- \frac{3(\gamma-1)\dot{a}}{a}e - \frac{p}{a \rho} \nabla \cdot \vec{v}_b
\label{enzointenergy}
\end{equation}

In this equation $e$ is the internal energy (defined as $P/\rho (\gamma-1)$ and the other terms are as described
previously.  The code still conserves total energy ($E$) as well.  In order
to maintain consistency, both equations are solved at all times in all cells,
with the equation for the total energy (eqtn.~\ref{enzoenergy}) being used
for hydrodynamics routines and the internal energy $e$ being used when 
temperature is required.  When pressure is required for dynamic purposes,
the total energy is used if the ratio of thermal energy to total energy
is less than some threshold value $\eta$, and the internal energy is used
for values of the ratio larger than $\eta$.  A typical value of this parameter
is $10^{-3}$.  This \emph{dual energy formulation} ensures that the
method produces the correct entropy jump at strong shocks and also
yields accurate pressures and temperatures in cosmological hypersonic
flows.

\subsubsection{The Zeus hydro method}\label{sec.ov.hydro.zeus}

As a check on PPM, \enzo\ also includes an implementation of the
finite-difference hydrodynamic algorithm employed in the compressible 
magnetohydrodynamics code `\zeus'
\citep{Stone92a, Stone92b}.  Fluid transport is solved on a Cartesian
grid using the upwind, monotonic advection scheme of \citet{1977JCoPh..23..276V} 
within a multistep 
(operator split) solution procedure which is fully explicit in time.  
This method is formally second order-accurate in space but 
first order-accurate in time.  It is important to note that 
this method conserves internal energy
rather than total energy, so the ``dual energy
formulation'' discussed in Section~\ref{sec.ov.hydro.ppm} is unnecessary.

Operator split methods break
the solution of the hydrodynamic equations into parts, with each part
representing a single term in the equations.  Each part is evaluated
successively using the results preceding it.  The individual parts of
the solution are grouped into two steps, called the \emph{source} and
\emph{transport} steps.

The \zeus\ method uses a von Neumann-Richtmeyer artificial viscosity 
to smooth shock discontinuities that may
appear in fluid flows and can cause a break-down of finite-difference
equations.  The artificial viscosity term is added in the source terms
as
\begin{eqnarray}
\rho \frac{\partial\textbf{v}}{\partial t} &=& - \nabla p - \rho \nabla \phi 
- \nabla \cdot \textbf{Q} \\
\frac{\partial e}{\partial t} &=& -p \nabla \cdot \textbf{v} - \textbf{Q} : \nabla \textbf{v}, 
\end{eqnarray}
where \textbf{v} is the baryon velocity, $\rho$ is the mass density,
$p$ is the pressure, $e$ is the internal energy density of the gas and \textbf{Q}
is the artificial viscosity stress tensor, such that:

\begin{eqnarray}
Q_{ii} = \left\{ \begin{array}{ll}
Q_{\rm AV} \rho_b  (a \Delta v_{i} + \dot{a} \Delta x_i)^2, 
& \textrm{for $a \Delta v_{i} + \dot{a} \Delta x_i < 0$}  \\
0 & \textrm{otherwise}\\
\end{array} \right. 
\end{eqnarray}
and
\begin{equation}
Q_{ij} = 0  \;\;\;{\rm for}\;\; i \ne j .
\end{equation}

$\Delta x_i$ and $\Delta v_{i}$ refer to the comoving 
width of the grid cell
along the $i$-th axis and the corresponding difference in gas
peculiar velocities across the grid cell, respectively, and $a$ is the
cosmological scale factor.  $Q_{\rm AV}$ is a constant with a typical
value of 2. We refer the interested reader to \citet{1994ApJ...429..434A} for more details.

The limitation of a technique that uses an artificial viscosity is that,
while the correct Rankine-Hugoniot jump conditions are achieved,
shocks are broadened over 6-8 mesh cells and are thus not treated as true
discontinuities. This may cause unphysical pre-heating of gas upstream
of the shock wave, as discussed in \citet{1994ApJ...429..434A}.

%---------------------------------------------------------------------------%
\subsection{Chemistry, radiation and cooling}\label{sec.ov.chem}

\red{
As this stands it is the methods that I (brian) use, so I don't really emphasize
either the tabular cooling functions or the various radiation background options.
Add more discussion of this, point towards appendix.
}

Though the equations of hydrodynamics described above are a closed system,
they are still missing a crucial piece of physics: radiative heating and cooling.
Radiative cooling is extremely important in many astrophysical situations, as is 
the heating of gas from some sort of radiation background.  \enzo\ has a very simple
Sutherland and Dopita equilibrium cooling function \citep{1993ApJS...88..253S} 
implemented, which uses a
cooling table assuming a fixed metallicity of $Z = 0.3 Z_\odot$, and also a
nonequilibrium heating/cooling model that assumes gas of primordial composition exposed to a
uniform metagalactic ultraviolet (UV) background that varies with time \citep{1996ApJS..105...19K}.

The simulations discussed in this thesis almost exclusively use the nonequilibrium
routines, described in great detail by Abel et al. and Anninos et al.  
\citep{abel97,anninos97} and summarized in Appendix~\ref{app:chemistry}. These 
routines follow the non-equilibrium chemistry of a 
gas of primordial composition with 9 total species:  
$H$, $H^+$, $He$, $He^+$, $He^{++}$, $H^-$, $H_2^+$, $H_2$, and $e^-$.  The code also calculates 
radiative heating and cooling, following atomic line excitation, recombination,
collisional excitation, free-free transitions, molecular line excitations, and Compton
scattering of the cosmic microwave background, as well as any of
approximately a dozen different models for a metagalactic UV background that heat
the gas via photoionization and photodissociation. 
The multispecies rate equations are solved out of
equilibrium to properly model situations where, e.g., the cooling rate of the gas
is much shorter than the hydrogen recombination time.  The effect of this nonequilibrium
cooling is to leave behind a much larger fraction of residual free electrons than one would
expect if the assumption of equilibrium were being made.  The practical effect of this
is that more $H^-$ is formed, which subsequently produces hydrogen molecules.  If large
amounts of $H_2$ is formed it can greatly increase the cooling rate of primordial gas
at relatively low temperatures ($T \leq 10^4$ K).  This can efficiently cool the gas 
to approximately $200$ K, which significantly reduces the Jeans mass of the gas.  Correct
modeling of the formation of molecular hydrogen is crucial to the study of star formation
in a primordial gas.

A total of 9 kinetic equations are solved, including 29 kinetic and radiative 
processes, for the 9 species mentioned above.  See Table~\ref{table.collisional} 
for the collisional processes and Table~\ref{table.radiative} for the 
radiative processes solved.

%---------------- table of collisional processes

\begin{table}
\begin{center}
{\bfseries Collisional Processes}\\[1ex]
\begin{tabular}{llllllll}
(1) & H & + & e$^-$ & $\rightarrow$ & H$^+$ &+& 2e$^-$ \\
(2) & H$^+$ &+ &e$^-$ & $\rightarrow$ & H &+ &$\gamma$ \\
(3) & He &+& e$^-$ & $\rightarrow$ & He$^+$ &+& 2e$^-$  \\
(4) & He$^+$ &+& e$^-$ & $\rightarrow$ & He &+ &$\gamma$  \\
(5) & He$^{+}$ &+& e$^-$ & $\rightarrow$ & He$^{++}$ &+& 2$e^-$  \\
(6) & He$^{++}$ &+& e$^-$ & $\rightarrow$ & He$^+$ &+& $\gamma$ \\
(7) & H &+& e$^-$ &$\rightarrow$& H$^-$ &+& $\gamma$  \\
(8) & H$^-$ &+& H &$\rightarrow$ & H$_2$ & +& e$^-$ \\
(9) & H &+ &H$^+$ &$\rightarrow$ &H$_2^+$ &+ &$\gamma$ \\
(10) & H$_2^+$ &+ &H &$\rightarrow$ &$H_2$ &+ &$H^+$ \\
(11) & H$_2$ &+ &H$^+$ &$\rightarrow$ &H$_2^+$ & +& H \\
(12) & H$_2$ &+ &e$^-$ & $\rightarrow$ & 2H & + & e$^-$  \\
(13) & H$_2$ & + & H & $\rightarrow$ & 3H &   &      \\
(14) & H$^-$ & + & e$^-$ & $\rightarrow$ & H & + & 2e$^-$ \\
(15) & H$^-$ & + & H & $\rightarrow$ & 2H & + & e$^-$ \\ 
(16) & H$^-$ & + & H$^+$ & $\rightarrow$ & 2H & & \\
(17) & H$^-$ & + & H$^+$ & $\rightarrow$ & H$_2^+$ & + & e$^-$ \\
(18) & H$_2^+$ & + & e$^-$ & $\rightarrow$ & 2H & & \\
(19) & H$_2^+$ & + & H$^-$ & $\rightarrow$ & H$_2$ & + & H  \\
(20) & 3H & & & $\rightarrow$ & H$_2$ & + & H
\end{tabular}
\caption[]{Collisional processes solved in the Enzo nonequilibrium
primordial chemistry routines.}
\label{table.collisional}
\end{center}
\end{table}

\begin{table}
\begin{center}
{\bfseries Radiative Processes}\\[1ex]
\begin{tabular}{llllllll}
(21) & H & + & $\gamma$ & $\rightarrow$ & H$^+$ & + & e$^-$ \\
(22) & He & + & $\gamma$ & $\rightarrow$ & He$^+$ & + & e$^-$ \\
(23) & He$^+$ & + & $\gamma$ & $\rightarrow$ & He$^{++}$ & + & e$^-$ \\
(24) & H$^-$ & + & $\gamma$ & $\rightarrow$ & H & + & e$^-$ \\
(25) & H$_2$ & + & $\gamma$ & $\rightarrow$ & H$_2^+$ & + & e$^-$ \\
(26) & H$_2^+$ & + & $\gamma$ & $\rightarrow$ & H & + & H$^+$ \\
(27) & H$_2^+$ & + & $\gamma$ & $\rightarrow$ & 2H$^+$ & + & e$^-$ \\
(28) & H$_2$ & + & $\gamma$ & $\rightarrow$ & H$_2^*$ & $\rightarrow$ & 2H \\
(29) & H$_2$ & + & $\gamma$ & $\rightarrow$ & 2H &  & 
\end{tabular}
\caption[]{Radiative processes solved in the Enzo nonequilibrium
primordial chemistry routines.}
\label{table.radiative}
\end{center}
\end{table}

The chemical reaction equation network is technically challenging to solve due to 
the huge range of reaction timescales involved -- the characteristic creation
and destruction timescales of the various species and reactions can differ by 
many orders of magnitude.  As a result, the set of rate equations is extremely 
stiff, and an explicit scheme for integration of the rate equations can be 
exceptionally costly if small enough timesteps are taken to keep the network
stable.  This makes an implicit scheme much more preferable for such a set of 
equations.  However, an implicit scheme typically require an iterative 
procedure to converge, and for large networks (such as this one) an implicit
method can be very time-consuming, making it undesirable for a large, three-dimensional
simulation.

\enzo\ solves the rate equations using a method based on a backwards differencing 
formula (BDF) in order to provide a stable and accurate solution.  This technique is
optimized by taking the chemical intermediaries $H^-$ and $H_2^+$, which have 
large rate coefficients and low concentrations, and grouping them into a separate
category of equations.  Due to their fast reactions these species are very sensitive
to small changes in the more abundant species.  However, due to their low overall
concentrations, they do not significantly influence the abundance of species with
higher concentrations.  Therefore, reactions involving these two species can be
decoupled from the rest of the network and treated independently.  In fact, $H^-$ 
and $H_2^+$ are treated as being in equilibrium at all times, independent of 
the other species and the hydrodynamic variables.  This allows a large speedup
in solution as the BDF scheme is then applied only to the slower 7-species network
on timescales closer to those required by the hydrodynamics of the simulation.
Even so, the accuracy and stability of the scheme is maintained by subcycling the 
rate solver over a single hydrodynamic timestep.  These subcycle timesteps are 
determined so that the maximum fractional change in the electron concentration is
limited to no more than $10\%$ per timestep.

It is important to note the regime in which this model is valid.  According to Abel et al. and
Anninos et al. \citep{abel97,anninos97}, the reaction network is valid for temperatures
between $10^0 - 10^8$ K.  The original model discussed in these two references is only
valid up to $n_H \sim 10^4$~cm$^{-3}$.  However, addition of the 3-body $H_2$ formation
process (equation 20 in Table~\ref{table.collisional}) allows 
correct modeling of the chemistry of the gas up
until the point where collisionally-induced emission from molecular hydrogen becomes an important
cooling process, which occurs at $n_H \sim 10^{14}$~cm$^{-3}$.  A further concern is that
the optically thin approximation for radiative cooling breaks down, which
occurs before $n_H \sim 10^{16} - 10^{17}$~cm$^{-3}$.  Beyond this point, 
modifications the cooling function that take into account the non-negligible
opacity in the gas must be made, as discussed by \citet{2004MNRAS.348.1019R}. 
Even with these modifications, a completely correct description of the cooling of
this gas will require some form of radiation transport, which will greatly 
increase the cost of the simulations.

Several processes are neglected.  The deuterium atom and its processes are 
completely ignored, which may have some effect.  Recent work shows that HD is
a more effective coolant than previously thought~\citep{2005MNRAS.361..850L}.  However,
the fractional abundance of HD is so low that under circumstances relevant to
the formation of a Population III star in an un-ionized region it should be
sub-dominant.  However, the enhanced electron fraction in fossil 
HII regions (as discussed later in this thesis) could result in the HD molecule
becoming a dominant cooling mechanism at relatively low ($\sim$ few hundred K)
temperatures, and could potentially cool the gas down to below $100$~K, which can
enhance fragmentation and could have important consequences for the IMF of 
primordial stars forming in a relic HII region~\citep{2005MNRAS.364.1378N}.

Aside from deuterium, the chemical reactions involving lithium 
are also neglected.  According to \citet{1998A&A...335..403G}, these are 
not important for the density and temperature regimes explored 
by the simulations discussed in this thesis.  However, at higher densities it 
is possible that there are regimes where lithium can be an important coolant.

%---------------------------------------------------------------------------%
\subsection{Star formation and feedback}\label{sec.ov.star}

\red{By the time this paper is submitted my (BWO's) paper on the star formation
and feedback algorithms, and the various parameters contained therein, should be
at least written and submitted as well.  Point to this, and maybe include a couple
interesting plots (perhaps in an appendix?).}

While the physics discussed previously is all crucial to the study of
cosmological structure formation, most cosmological observations are of
stars and related phenomena.  Additionally, stars eject energy and metal-
enriched gas throughout their lives, drastically modifying their own 
environment.  The formation of galaxies and clusters cannot be completely
modeled without including the feedback of energy and metals.  In particular,
it is thought that feedback is crucial for suppressing the large numbers of
dwarf-like galaxies that CDM theories predict 
\citep{1991ApJ...381...14L,1991ApJ...379...52W}.  An early
burst of star formation could remove a large fraction of cold gas from such systems
\citep{1978MNRAS.183..341W,1991ApJ...367...45C}.  Also, the unexpectedly 
low luminosities of small clusters and
groups (relative to the richest clusters) is commonly explained through feedback
\citep{1991ApJ...383..104K}.  Energy ejected during the formation of the cluster ellipticals 
lowers the central density and hence the X-ray luminosity of such clusters 
 \citep{1997ApJ...484L..21C}.  Therefore, the inclusion of star formation and the 
feedback of energy and metals into the intergalactic medium in a cosmological
code is crucial for many reasons.

We have extended the \enzo\ code to include multiple models for star formation
and feedback.  It is difficult to directly simulate the formation of individual
stars in the context of simulations of galaxy formation and evolution due
to the immense computational cost.  Therefore, 
we adopt parametric methods which attempt
to model star forming regions.  One model is based on Kravtsov's method
 \citep{2003ApJ...590L...1K}, and the other is based on the method of 
\citet{1992ApJ...399L.113C}, which has been modified for use in an AMR code.  The basic
ideas behind the methods are straightforward and observationally motivated.
Similar (though somewhat more advanced) methods have been employed for the 
smoothed particle hydrodynamics
(SPH) method, most recently by Springel \& Hernquist 
\citep{sh03a,sh03b,hs03}. This 
method assumes a
multiphase IGM and has been shown to accurately reproduce the cosmic star
formation rate, and will be implemented into \enzo\ in the near future.

In the following sections we will detail both the Kravtsov and Cen \& Ostriker
star formation methods separately.  Tests of the star formation and
feedback algorithms will not be shown, since this work is currently in
progress (and an extension of the AMR/SPH code comparison described in
\citet{2005ApJS..160....1O}).

\subsubsection{The Kravtsov star formation and feedback algorithm}

Kravtsov's method of star formation is designed to reproduce
the global Schmidt law of star 
formation~\citep{2003ApJ...590L...1K,1959ApJ...129..243S}.
This algorithm is deliberately minimal, and is explicitly geared
towards modeling star formation in a phenomenological way on kiloparsec
scales.  Stars are assume to form with a characteristic gas timescale 
$\tau_*$ such that $\dot{\rho}_* = \rho_{gas}/\tau_*$ where $\rho_{gas}$ 
and $\rho_*$ are the baryon gas and stellar densities, respectively.  
This ``constant efficiency'' model on the scale of star formation 
regions is well motivated observationally \citep{1996AJ....112.1903Y,2002ApJ...569..157W}.  
Star formation is only allowed to take place in very dense regions
with $\rho_{gas} \geq \rho_{SF}$, where $\rho_{SF}$ is a constant
proper density threshold above which star formation is allowed to
occur.  No other criteria are imposed.  Kravtsov's typical choices
for $\tau_*$ and $\rho_{SF}$ are $\tau_* = 4$ Gyr and $\rho_{SF} =
1.64~$M$_\odot$~pc$^{-3}$~$(n_H \sim 50$~cm$^{-3})$.  The adopted timescale
is derived from the observationally-determined normalization of the
Schmidt law, and the density threshold is determined by observations
of star forming regions on $\sim 100$ pc scales.  Note that the density
threshold is in proper, not comoving, units.

Algorithmically, the star formation events in Kravtsov's code
are assumed to occur once every global time step 
$\Delta t_0 \leq 10^7$.  In cells where star formation is
determined to occur (i.e. $\rho_{gas} \geq \rho_{SF}$) a
collisionless ``star particle'' is assumed to form, with a 
mass $m_* = \dot{\rho}_* V_{cell} \Delta t_0$, where
$\dot{\rho}_*$ is described above and $V_{cell}$ is the 
volume of the mesh cell.  These star particles are
formed instantaneously, with all $m_*$ of gas being
taken out of the cell and immediately deposited into the
star particle.  This particle is then given the velocity
of the gas in the cell, and thereafter treated as a 
collisionless particle.  The \enzo\ implementation
of this algorithm is similar, except that instead of 
forming stars only at the root grid time step, we allow 
stars to form at the time step of the highest level of
resolution at any particular point in space.  As can be seen
from the equation for $m_*$ above, this can result in very 
small stellar masses.  To avoid memory and processor time
issues related to having very large numbers of star particles
we impose a threshold mass $M_{*,min}$ such that a star
particle only forms if $m_* \geq M_{*,min}$.  An
appropriate choice of $M_{*,min}$ does not significantly 
change the star overall star formation history of a 
simulation, though it may delay the onset of star formation 
in a given cell relative to a simulation without a particle mass
threshold.

Each ``star particle'' is assumed to represent an ensemble
of stars and is treated as a single-age stellar population.
Kravtsov assumes that the IMF is described by a
Miller \& Scalo functional form with stellar masses between
$0.1$ and $100~M_\odot$ \citep{1979ApJS...41..513M}.  All stars in this
IMF with $M_* > 8~M_\odot$ deposit $2 \times 10^{51}$ ergs of
thermal energy and a mass $f_z M_*$ of metals into the
cell in which they form without delay, with
$f_z \equiv min(0.2, 0.01~M_*-0.06)$ (i.e. instantaneous
deposition of metals).  The definition of $f_z$ is a rough
approximation of the results of 
\citet{1995ApJS..101..181W}.

Kravtsov reports that simulations with this algorithm
reliably reproduce the star formation rate-gas surface
density relation of the Schmidt law, and are not particularly
sensitive to numerical parameters \citep{2003ApJ...590L...1K}.
He also notes that this is surprisingly insensitive to
the presence or absence of feedback and details of the cooling
and heating properties of the gas, which suggests that the
global star formation rate is determined by gravitationally-
driven supersonic turbulence (on large scales)
rather than stellar feedback or thermal instabilities on small 
scales.

\subsubsection{The Cen \& Ostriker star formation algorithm}

The Cen \& Ostriker method is a heuristic model of star formation
on galactic scales.  This method, first described by
\citet{1992ApJ...399L.113C}, is similar in some ways to the Kravtsov algorithm
but has more complex criteria for determining where stars should
be formed.  In this method, cells that form stars must have a 
baryon overdensity higher than some threshold 
$\rho_b/\bar{\rho}_b \geq \eta$ 
where $\rho_b$ is the baryon density in that cell, 
$\bar{\rho_b}$ is the mean baryon density in the simulation,
and $\eta$ is the user-defined overdensity threshold.
Additionally, the gas in the cells must be contracting, 
cooling rapidly, and gravitationally unstable, e.g.:

\begin{equation}
\nabla \cdot \vec{v}_b < 0,
\label{cencont}
\end{equation}

\begin{equation}
t_{cool} < t_{dyn} \equiv \sqrt{3 \pi / 32G \rho_{tot}},
\end{equation}

\begin{equation}
m_{b} > m_{J} \equiv G^{-3/2} \rho_{b}^{-1/2}C^{3}
\left[ 1 + \frac{\delta\rho_{d}}{\delta\rho_{b}} \right]^{-3/2}
\end{equation}

where $\vec{v}$ is the velocity of the gas in the cell, $\rho_{b}$ and 
$\rho_{d}$ are the cell's baryon and dark matter density, respectively,
$\rho_{total} = \rho_{b} + \rho_{d}$, $m_{b}$ and $m_{j}$ are the 
baryonic mass in the cell and jeans mass of the cell, and C is the 
isothermal sound speed in the cell.  
If all of these criteria are met, 
the mass of a star particle is calculated as \(m_{*} = m_{b} 
\frac{ \Delta t}{ t_{dyn} } f_{*eff} \), 
where $f_{*eff}$ is the star formation efficiency parameter.

If $m_{*}$ is greater than a minimum star mass $m_{*min}$, a particle 
is created and given several attributes:  Mass, a unique index number, 
the time of formation $t_{form}$, the local dynamical free-fall time 
$t_{dyn}$ and the metallicity fraction of the baryon gas in the cell 
$f_{Zb}$.  There is a user-defined minimum dynamical time
$T_{dyn,min}$ which is observationally motivated and affects the
feedback rates (see below).
The particle is placed in the center of the cell and given 
the same peculiar velocity as the gas in the cell, and is then treated 
in the same manner as the dark matter particles.  An 
amount of baryon gas corresponding to the new particle's mass is 
removed from the cell.

In addition, we have added a stochastic star formation algorithm that keeps 
track of all of the sub-mass threshold stars that should have been created 
and when the total mass of uncreated stars is greater than the minimum mass, 
a star particle is created.

The feedback of energy and metals into the baryon gas is similar to
the Kravtsov feedback, with some important differences.  The star formation 
algorithm creates each star particle instantaneously.  However, feedback
should take place over a significant timescale, as all of the stars contained
within the ``star particle'' would form over a long period of time.  Therefore,
we assume that for the purposes of feedback that the mass of stars formed
at a time $t$ with finite timestep $\Delta t$ is:

\begin{eqnarray}  
\Delta m_{sf} =  m_{*} \left[ \right. \left(1 + \frac{t-t_{form}}{t_{dyn}} 
                    \right) \nonumber \\ 
		   \times exp \left( \frac{-(t-t_{form})}{t_{dyn}} \right) 
                    \nonumber \\ 
	       - \left( 1 + \frac{t+ \Delta t-t_{form}}{t_{dyn}} \right) \nonumber \\ 
                    exp \left( \frac{-(t+\Delta t-t_{form})}{t_{dyn}} \right) 
                   \left. \right]
\end{eqnarray}

which can be represented more clearly in integral form:

\begin{eqnarray}
\int_{t}^{t+Dt} \frac{dM}{dt} dt = \int_{t}^{t+Dt} m_{*} 
\frac{dt}{t_{tyn}} \left( \frac{t-t_{form}}{t_{dyn}} \right) \nonumber \\ 
\times exp \left( - ~ \frac{t-t_{form}}{t_{dyn}} \right) 
\end{eqnarray}

During this timestep, we assume that the star particle feeds back
metal-enriched gas and thermal energy from supernovae and from stellar winds.
Since massive stars have very short lifetimes, we assume that there is an
immediate feedback of some fraction $f_{SN}$ of the rest energy from the
stars created that timestep into the baryon gas, such that 
$E_{add} = f_{SN} \Delta m_{sf} c^2$, where c is the speed of light.
In addition, a fraction $f_{Z*}$ of the metal from the star particle
is fed back into the baryon gas, which takes into account the effects
of metal recycling.  Finally, a fraction of the mass $f_{m*}$ from all
stars (rather than just supernovae) is fed back into the gas along with
momentum in order to simulate the mass ejection from non-exploding stars
via stellar winds.

There are six user-defined parameters in this algorithm:  three deal with the
star formation ($\eta$, $m_{*min}$ and $T_{dyn,min}$), and three 
deal with feedback ($f_{SN}$, $f_{Z*}$ and $f_{m*}$).  Some of these
parameters are completely free, while others can be guided by observation
or theory.  For example, the supernova feedback parameter, $f_{SN}$, can be 
constrained assuming that, for every $200 M_\odot$ of stars created, one
supernova occurs, and this event feeds back approximately $10^{51}$ ergs of
thermal energy, giving:

\begin{equation}
f_{SN} = \frac{10^{51} ergs}{200~M_\odot c^2} \simeq 3 \times 10^{-6}
\end{equation}

The metal yield $f_{Z*}$, defined as the mass in metals produced per unit
mass of stars created, can be constrained by a theoretical model of
 \citet{1995ApJS..101..181W}.  This model suggests that $f_{Z*} = 0.02$
is an appropriate number.  The minimum dynamical time is set to be $T_{dyn,min} = 10^7$
~years to agree with timescales seen in nearby OB associations.

The other parameters, such as the overdensity threshold $\eta$, minimum star
 mass $m_{*min}$, and mass ejection fraction $f_{m*}$ are not well constrained 
either theoretically or observationally.  Indeed, $m_{*min}$ is a purely
numerical parameter designed to keep the code from producing too many star 
particles, and thus has no observational or theoretical counterpart.  These
parameters have to be set by performing parameter studies and comparing to 
observations.  Unfortunately, the range of parameter space is large, and
the results may be degenerate for some combinations of these parameters.

%---------------------------------------------------------------------------%
\subsection{Adaptive mesh refinement}\label{sec.ov.amr}
\dcc{removed most of the notes.  Added intro blurb.  moved parallelism
  stuff. Re-worded the bit on ghost zones.  Summary of 'grid as
  independent update + syncing with neighbors and parents,' W-cycle. Added
  sentence on subgrid creation.  Made timestepping its own subsection.}
\red{Talk about reasonable refinement factors - which 
ones are usable, etc?}

The primary purpose of the \enzo\ code is its Adaptive Mesh Refinement
capability, which allows it to reach extremely large dynamical ranges
with limited computational resources, opening doors previously closed
by finite memory and computational time. Unlike moving mesh methods
\citep{1995ApJS..100..269P,1995ApJS...97..231G} or  
methods that subdivide 
individual cells \citep{Adjerid}, Berger \& Collela's AMR (also referred 
to as \emph{structured} AMR) utilizes an adaptive hierarchy of grid 
patches at varying levels of resolution.  Each rectangular grid patch 
(referred to as a ``grid'') covers some region of space in its 
\emph{parent grid} which requires higher resolution, and can itself 
become the parent grid to an even more highly resolved \emph{child grid}. 
ENZO's implementation of structured AMR places no restrictions on 
the number of grids at a given level of refinement, or on the number of 
levels of refinement.  However, owing to limited computational resources 
it is practical to institute a maximum level of refinement $\Lmax$.

The \enzo\ implementation of AMR allows arbitrary integer ratios of parent and
child grid resolution.  However, in practice 
refinement factors (defined as the ratio of parent grid resolution to child grid resolution) 
of two or four are typically used, with a refinement factor
of two being most commonly used for cosmological simulations for reasons of
efficiency. The ratio of
box size to the maximum grid resolution of a given simulation is
therefore $L/e = \Nroot \times 2^{\Lmax}$, where $\Nroot$ is the
number of cells along one edge of the root grid, $\Lmax$ is the
maximum level of refinement allowed, and L and e are the box size
and grid resolution of the most highly refined region, respectively.

The AMR grid patches are the primary data structure in \enzo.  Each
patch is treated as an individual object which can contain both field
variables and particle data.  Individual grids are organized into a 
dynamic, distributed hierarchy of mesh patches.  
See 
Figure~\ref{fig.2.amrhierdens} for an example of an \enzo\
simulation performed using this AMR scheme where both the baryon density
and AMR grid hierarchy are shown together.

\begin{figure}
\begin{center}
\includegraphics[width=0.4\textwidth]{figures/amr_hierarchy.eps}
\end{center}
\caption{\emph{Left:}  Example of a simple, 
distributed AMR hierarchy showing real and 
ghost grids.
\emph{Right:}  Example 2D \enzo\ grid showing real and ghost zones, as 
needed for the PPM hydro stencil.  Image courtesy of James Bordner (Cent.
Astrophysics and Space Sciences, UCSD).
}
\label{fig.2.amrhierarchy}
\end{figure}

\begin{figure}
\begin{center}
\includegraphics[width=0.4\textwidth]{figures/amr-hier-dens.eps}
\caption{Example of an \enzo\ simulation showing the AMR grid hierarchy.  
This is a simulation of the collapse of a single massive halo with a 
$32^3$ root grid and two $32^3$ static nested grids.  AMR is only allowed
within the most highly refined static nested grid.  Log baryon density is shown
by colored isocontours with values corresponding to the legend at center left.  
The rectangular solid wire frames correspond to individual 
AMR grids, with different colors corresponding to level as indicated by the
legend at top left.}
\label{fig.2.amrhierdens}
\end{center}
\end{figure}

Each grid patch in \enzo\ contains arrays of values for baryon and 
particle quantities.   Grids are 
partitioned into a core of \emph{real zones} and a surrounding layer 
of \emph{ghost zones}.  The real zones store field values and ghost 
zones are used to temporarily store values which have been obtained 
directly from neighboring grids or interpolated from a parent grid.  
These zones are necessary to accommodate the computational stencil of 
the hydrodynamics solvers (Sections~\ref{sec.ov.hydro.ppm} and~\ref{sec.ov.hydro.zeus}) 
and the gravity solver (Section~\ref{sec.ov.nbody}).  
The PPM and Zeus hydro solvers require 3 layers of ghost zones,  and the gravity
solver requires 6.  This can lead to significant memory 
and computational overhead, particularly for smaller grid patches
at high levels of refinement.  More information on filling of ghost
zones can be found in the appendix.

From the point of view of the hydro solver, each grid is solved as an
independent CFD problem, with Dirichlet boundary conditions stored in
the ghost zones.  At the end of each time step, cells on a given level
fill ghost zones by interpolating from the parent and copying from
neighbors on the same level.  Grids then correct the flux at the interface boundary between
fine and coarse zones, and finally project their real zone data to the
parent grid.

Timesteps are determined on each level as
described in \ref{sed.timestepping}, and the hierarchy is advanced on
a level by level basis in what's known as a W-Cycle.    Beginning with
the coarsest level, $L$, all grids on that level are advanced one
timestep.  Then one timestep is taken on all grids at the next level
of refinement, $L+1$, and so on until the finest level is resolved.
The finest level is then advanced, using as many steps as it takes to
reach the level immediately above.  The finest level is then synced
to the level above it, which then proceeds forewards one more step.
\red{Needs a picture of the W cycle.}

At the end of every timestep on every level each grid updates 
its ghost zones by exchanging information with its neighboring grid patches 
(if any exist) and/or by interpolating from a parent grid.  
In addition, cells are examined to see if they should be refined or 
de-refined, and the entire grid hierarchy is rebuilt at that 
level (including all more highly refined levels).  The timestepping and 
hierarchy advancement/rebuilding process described here is repeated 
recursively on every level to the specified maximum level of 
refinement in the simulation.

Since the addition of more highly refined grids is adaptive,  
the conditions for refinement must be specified.  The criteria of 
refinement can be set by the threshold value of the overdensity of 
baryon gas or dark matter in a cell (which is really a refinement on
the mass of gas or DM in a cell), the local Jeans length, the 
local density gradient, or local pressure and energy gradients.  
A cell reaching any or all of these criteria 
will then be flagged for refinement.  
Once all cells at a given level have been examined, rectangular boundaries 
are determined which minimally encompass the flagged cells. The set of
flagging cells is then further fractured into smaller bounding
rectangles to ensure computational 
efficiency.  More details on subgrid creation can be found in Appendix
\ref{app.subgrid_creation}.  A refined grid patch is 
introduced within each such bounding rectangle. Thus the cells needing 
refinement, as well as adjacent cells within the patch which do not need 
refinement, are refined. While this approach is not as memory efficient as 
cell-splitting AMR schemes, it offers more freedom with 
finite difference stencils. For example, PPM requires a stencil of seven 
cells per dimension. This cannot easily be accommodated in cell-splitting 
AMR schemes.

\blue{ \emph{I think this  should be
  re-written to reflect data presented or referenced in this paper,
  whatever that winds up being.}
In the simulations discussed in this thesis we typically use baryon and
dark matter overdensities as our refinement criteria, though
for some higher-resolution simulations we also refine on other
criteria as needed.}

\subsection{Time Stepping}\label{sec.timestepping}
\dcc{Made this its own section (not really part of the AMR, it's more
  general than that.)  Added stuff on harmonic average for 3d. See
  note below on acceleration timestep criterion being crap. Moved
  w-cycle to the previous section}
In \enzo, resolution of the equations being solved is adaptive in time 
as well as in space.  The timestep in \enzo\ is satisfied on a level-by-level
 basis by finding the largest timestep such that multiple criteria are
satisfied on each level.  The timestep criteria used by Enzo are 
(showing the one-dimensional case for clarity):

\begin{equation}
\Delta t_{hydro} = min \left( \kappa_{hydro} \frac{a \Delta x}{c_{s} + |v_x|} \right)_L ,
\label{eqn:dthydro}
\end{equation}

\begin{equation}
\Delta t_{dm} = min \left(\kappa_{dm} \frac{a \Delta x}{v_{dm,x}} \right)_L ,
\label{eqn:dtdarkmatter}
\end{equation}

\begin{equation}
\Delta t_{exp} = f_{exp} \left( \frac{a}{\dot{a}} \right) ,
\label{eqn:dtexpand}
\end{equation}

\begin{equation}
\Delta t_{accel} = min \left( \sqrt{\frac{\Delta x}{\vec{g}}} \right)_L 
\label{eqn:dtaccel}
\end{equation}

 In equations~\ref{eqn:dthydro}-\ref{eqn:dtaccel}, the $min ( \ldots )_L$
formalism means that this value is calculated for all cells on a given level
L and the minimum overall value is taken as the timestep.  
Equation~\ref{eqn:dthydro} ensures that all cells satisfy the 
Courant-Freidrichs-Levy (CFL) condition for accuracy and stability of an explicit
finite difference discretization of the Euler equations.  Effectively this condition
forces the timestep to be small enough such that any changes in the fluid propagate
less than a single grid spacing, $\Delta x$.  In this equation, $\kappa_{hydro}$ is 
a numerical constant with a value of $0 < \kappa_{hydro} \leq 1$ (with a typical
value of $\kappa_{hydro} \sim 0.3-0.5$) that ensures that the CFL condition is always
met, and $c_s$ and $v_x$ are the sound speed and peculiar baryon
velocity in a given cell.

Equation~\ref{eqn:dtdarkmatter} is analogous to Equation~\ref{eqn:dthydro} and ensures
accuracy in the N-body solver by requiring that no dark matter particle travels more than
one cell width.  The parameter $\kappa_{dm}$ is analogous to $\kappa_{dthydro}$, with
an identical range of values.  Equation~\ref{eqn:dtexpand} limits the timestep such 
that the expansion parameter a can only change by a fractional amount of $f_{exp} = \Delta a/a$, 
where $f_{exp}$ is a user-defined parameter and has typical values of $f_{exp} = 0.01-0.02$.
This is required for the stability of the PPM algorithm in comoving coordinates, and
typically limits the timestep only at high redshifts when densities are relatively
homogeneous.

Equation~\ref{eqn:dtaccel} is supplementary to equation~\ref{eqn:dthydro} in that it
takes into account the possibility of large accelerations causing numerical 
instabilities by violating the Courant condition.  In this equation, $\vec{g}$ is the
gravitational acceleration in each cell on level L.  \blue{ \emph{It should be
  noted that the timestep is computed before the acceleration field is computed,
  and the acceleration field is deleted at the end of each time step.
  So this condition isn't ever actually enforced.  We should either
  fix this problem, or not discuss it.
 -dcollins.} }

Equation ~\ref{eqn:dthydro} is valid for one dimension.  For 2 or 3
dimensions, it was shown by \cite{Godunov1959}  that using the
harmonic average of the timestep found along each of the coordinate
axes yields a maximum $\kappa_{hydro} = 0.8$.  So letting $\Delta
t_x$, $\Delta t_y$, and $\Delta t_z$ be the analogues of
eqn. \ref{eqn:dthydro} along the $x,y$ and $z$ axes, 
\begin{equation}
\Delta t_{hydro} = min ( \kappa_{hydro} /( \frac{1}{\Delta t_x}
+\frac{1}{\Delta t_y} + \frac{1}{\Delta t_z} ) )_L
\end{equation}

For all other criteria, multiple dimensions are accounted for by
repeating the one dimensional criterion along each axis, and taking
the minimum.

%---------------------------------------------------------------------------%
\subsection{Parallelism}\label{sec.ov.parallel}
\dcc{moved hierarchy description, parallism notes from AMR section
  here. Mentioned performance/scaling study.  Still short.}

When \enzo\ is run on a distrubuted platform, some of the data is
replicated to all processors, while some remains on a single processor.
A description of the entire hierarchy (the position and size of each grid, as well
some other meta data) is stored on each processor.  This enables each
grid to know what grids are near it in the hierarchy, regardless of
where the data is.  The actual Baryon data (density, velocity, energy,
and any chemistry data) and particle data are stored on only one
processor.  Storing the entire hierarchy on all processor does create
some overhead, but in practice this is not an issue until one reaches
the extremely large scale.
  The code handles load balancing 
on a level-by-level basis such that the workload on each level is 
distributed as uniformly as possible across all processors.  Spatial locality of 
grids is not forced during message passing, for maximum flexibility (though not
necessarily maximum efficiency).  
The MPI message passing library\footnote{http://www-unix.mcs.anl.gov/mpi/}
 is used to transfer data between processors.

The performance of \enzo\ has been extensively studied using the
software package LCAPerf.  More on LCAPerf can be found in section
\ref{sec.performance}

%old text
%Every processor 
%keeps a description of the entire grid hierarchy at all times, so that 
%each processor knows where all grids are.  However, baryon and particle 
%data for a given grid only exists on a single processor.  See 
%Figure~\ref{fig.2.amrhierarchy} for a schematic example of this.

### 2012-06-29 numerical-particle.tex
first commit of paper.

--- 2012-06-29 numerical-overview.tex
@@
-\section{Overview of Numerical Methods}\label{sec.overview}
+\subsection{N-body Dynamics}
+\label{sec.ov.nbody}
 
-%---------------------------------------------------------------------------%
-\subsection{N-body dynamics}\label{sec.ov.nbody}
-fft + multigrid
+The codes needs to be able to model collisionless 
+One way would be to model the dark matter (or other collisionless particle-like objects, such as stars) as a second fluid in addition to the baryonic fluid and solve the collisionless Boltzmann equation, which follows the evolution of the fluid density in six-dimensional phase space.
+However, this is computationally prohibitive owing to the large dimensionality of the problem, making this approach  unappealing for the cosmological AMR code.
+%GB: removed the above (good discussion but not really appropriate in this paper I think, or maybe at the beginning)
 
-There are multiple methods to compute the gravitational potential 
-(which is an elliptic equation in the Newtonian limit) in a structured 
-AMR framework.  One way would be to model the dark matter (or other
-collisionless particle-like objects, such as stars) as a second fluid 
-in addition to the baryonic fluid and solve the collisionless
-Boltzmann equation, which follows the evolution of the fluid density in
-six-dimensional phase space.  However, this is computationally prohibitive 
-owing to the large dimensionality of the problem, making this approach 
-unappealing for the cosmological AMR code.
-
-Instead, \enzo\ uses a particle-mesh N-body method to calculate 
+To solve the, \enzo\ uses a particle-mesh N-body method to calculate 
 the dynamics of collisionless systems \citep{Hockney88}.  This method 
@@
 
-These equations are finite-differenced and for simplicity are solved with the
-same timestep as the hydrodynamic equations, as discussed in 
-Section~\ref{sec.ov.amr}.
-The dark matter particles are distributed onto the grids
-using the cloud-in-cell (CIC) interpolation technique to form a spatially
-discretized density field (analogous to the baryon densities used to calculate
-the equations of hydrodynamics).  After sampling dark matter density onto the
-grid and adding baryon density if it exists (to get the total matter density
-in each cell), the gravitational potential is calculated on the periodic root
-grid using a fast Fourier transform.  In order to calculate more accurate
-potentials on subgrids, \enzo\ re-samples the dark matter density onto
-individual subgrids using the same CIC method as on the root grid, but at
-higher spatial resolution (and again adding baryon densities if applicable).
-Boundary conditions are then interpolated from the potential values on the
-parent grid (with adjacent grid patches on a given level communicating to
-ensure that their boundary values are the same), and then a multigrid
-relaxation technique is used to calculate the gravitational potential at every
-point within a subgrid.  Forces are computed on the mesh by
-finite-differencing potential values and are then interpolated to the particle
-positions, where they are used to update the particle's position and velocity
-information.  Potentials on child grids are computed recursively and particle
-positions are updated using the same timestep as in the hydrodynamic
-equations.  Particles are stored in the most highly refined grid patch at the
-point in space where they exist, and particles which move out of a subgrid
-patch are sent to the grid patch covering the adjacent volume with the finest
-spatial resolution, which may be of the same spatial resolution, coarser, or finer
-than the grid patch that the particles are moved from.  This takes place in a
-communication process at the end of each timestep on a level.
-
-At this point it is useful to emphasize that the effective force
-resolution of an adaptive particle-mesh calculation is approximately twice
-as coarse as the grid spacing at a given level of resolution.  The 
-potential is solved in each grid cell;
-however, the quantity of interest, namely the acceleration, is the gradient
-of the potential, and hence two potential values are required to calculate
-this.  In addition, it should be noted that the adaptive particle-mesh
-technique described here is very memory intensive: in order to ensure reasonably 
-accurate force resolution at grid edges the multigrid relaxation
-method used in the code requires a layer of ghost zones which is very deep --
-typically 6 cells in every direction around the edge of a grid patch.  This greatly
-adds to the memory requirements of the simulation, particularly because subgrids
-are typically small (on the order of $12^3 - 16^3$ real cells for a standard 
-cosmological calculation) and ghost zones can dominate the memory and computational 
-requirements of the code.
-
-%---------------------------------------------------------------------------%
-\subsection{Hydrodynamics}\label{sec.ov.hydro}
-
-\subsubsection{The PPM hydro method}\label{sec.ov.hydro.ppm}
-
-\dcc{Removed ``things to do'' list.  Mention ppm-lr and ppm-de, and
-  that ppm-de is preferred.  Briefly describe ppm-de itself.}
-
-The primary hydrodynamic method used in \enzo\ is based on the
-piecewise parabolic method (PPM) of
-\citet{1984JCoPh..54..174C}, which has been
-modified for the study of cosmological fluid flows.  
-\citet{1984JCoPh..54..174C} describe two variants on this method: a
-Lagrangian Remap variant and a Direct Eulerian variant.  Both have
-been adapted for \enzo, and the Lagrangian Remap 
-variant is described in \citet{1995CoPhC..89..149B}.  Because of its
-conservative formulation, however, the Direct Eulerian variant is
-preferred for \enzo\ AMR simulations.  We will describe the Direct
-Eulerian (PPM-DE) variant and its additions briefly here, and in
-further depth in Appendix \ref{app.ppm}
-
-PPM is an explicit, higher order-accurate version of
-Godunov's method for ideal gas dynamics with third order-accurate piecewise parabolic
-monotonic interpolation and a nonlinear Riemann solver for shock
-capturing.  It advances the hydrodynamic equations in the following
-steps:
-\begin{enumerate}
- \item Construct monotonic parabolic interpolation of cell average
- data, for each fluid quantity.
- \item Compute interface states by averaging the parabola for over the domain of dependence for
- each interface
- \item Use interface data to solve the Riemann problem.
- \item Difference the interface fluxes to update the cell average quantities.
-\end{enumerate}
-It does an excellent job of capturing strong shocks in at
-most two cells.  Multidimensional schemes are built up by directional
-splitting and produce a method that is formally second order-accurate
-in space and time which explicitly conserves mass, linear momentum, and energy.
-
-To use the above algorithm for cosmological simulations, the
-conservation laws for fluid mass, momentum and energy 
-density are written in comoving coordinates for a
-Friedman-Robertson-Walker space-time, as described previously 
-in Equations~\ref{enzoconserve} 
-through~\ref{enzoenergy}.  Both the conservation laws and
-the Riemann solver are modified to include gravity, which is solved
-using an adaptive particle-mesh (PM) technique (see Section~\ref{sec.ov.nbody}).  
-The terms due to cosmological expansion, as well as primordial chemistry and
-radiative heating and cooling, are solved in a separate step because they have
-different numerical behavior, and therefore must be treated differently to ensure
-stability.  Note that unlike the \zeus\ hydro scheme, PPM does not need to use 
-artificial viscosity to resolve shocks.
-
-The system of equations described above works well for systems with
-relatively low Mach numbers, as long as these systems are well resolved.  
-However, cosmology is replete with situations where there are bulk hypersonic
-flows.  In these situations, the ratio of kinetic to thermal energy can be
-very high -- in some situations up to $10^6 - 10^8$.  This implies that the
-thermal energy is an extremely tiny fraction of the kinetic energy, which can 
-cause numerical problems when one is interested in just the thermal energy of the gas, since
- Equation~\ref{enzoenergy} solves for the total energy.  In this system of
-equations, the
- thermal energy $E_{therm}$ is calculated as $E - E_{kin}$, 
-where E is the total specific energy as
-calculated in equation~\ref{enzoenergy} and $E_{kin}$ is the specific kinetic energy,
-$0.5 \vec{v}_b^2$.  
-In hypersonic flows $E$ and $E_{kin}$ are nearly the same, and any 
-number calculated as 
-the difference of these is going to be strongly affected by numerical error.  To 
-avoid this problem, \enzo\ also solves the internal energy equation in
-comoving coordinates:
-
-\begin{equation}
-\frac{\partial e}{\partial t} + \frac{1}{a} \vec{v}_b \cdot \nabla e = 
-- \frac{3(\gamma-1)\dot{a}}{a}e - \frac{p}{a \rho} \nabla \cdot \vec{v}_b
-\label{enzointenergy}
-\end{equation}
-
-In this equation $e$ is the internal energy (defined as $P/\rho (\gamma-1)$ and the other terms are as described
-previously.  The code still conserves total energy ($E$) as well.  In order
-to maintain consistency, both equations are solved at all times in all cells,
-with the equation for the total energy (eqtn.~\ref{enzoenergy}) being used
-for hydrodynamics routines and the internal energy $e$ being used when 
-temperature is required.  When pressure is required for dynamic purposes,
-the total energy is used if the ratio of thermal energy to total energy
-is less than some threshold value $\eta$, and the internal energy is used
-for values of the ratio larger than $\eta$.  A typical value of this parameter
-is $10^{-3}$.  This \emph{dual energy formulation} ensures that the
-method produces the correct entropy jump at strong shocks and also
-yields accurate pressures and temperatures in cosmological hypersonic
-flows.
-
-\subsubsection{The Zeus hydro method}\label{sec.ov.hydro.zeus}
-
-As a check on PPM, \enzo\ also includes an implementation of the
-finite-difference hydrodynamic algorithm employed in the compressible 
-magnetohydrodynamics code `\zeus'
-\citep{Stone92a, Stone92b}.  Fluid transport is solved on a Cartesian
-grid using the upwind, monotonic advection scheme of \citet{1977JCoPh..23..276V} 
-within a multistep 
-(operator split) solution procedure which is fully explicit in time.  
-This method is formally second order-accurate in space but 
-first order-accurate in time.  It is important to note that 
-this method conserves internal energy
-rather than total energy, so the ``dual energy
-formulation'' discussed in Section~\ref{sec.ov.hydro.ppm} is unnecessary.
-
-Operator split methods break
-the solution of the hydrodynamic equations into parts, with each part
-representing a single term in the equations.  Each part is evaluated
-successively using the results preceding it.  The individual parts of
-the solution are grouped into two steps, called the \emph{source} and
-\emph{transport} steps.
-
-The \zeus\ method uses a von Neumann-Richtmeyer artificial viscosity 
-to smooth shock discontinuities that may
-appear in fluid flows and can cause a break-down of finite-difference
-equations.  The artificial viscosity term is added in the source terms
-as
 \begin{eqnarray}
-\rho \frac{\partial\textbf{v}}{\partial t} &=& - \nabla p - \rho \nabla \phi 
-- \nabla \cdot \textbf{Q} \\
-\frac{\partial e}{\partial t} &=& -p \nabla \cdot \textbf{v} - \textbf{Q} : \nabla \textbf{v}, 
-\end{eqnarray}
-where \textbf{v} is the baryon velocity, $\rho$ is the mass density,
-$p$ is the pressure, $e$ is the internal energy density of the gas and \textbf{Q}
-is the artificial viscosity stress tensor, such that:
-
-\begin{eqnarray}
-Q_{ii} = \left\{ \begin{array}{ll}
-Q_{\rm AV} \rho_b  (a \Delta v_{i} + \dot{a} \Delta x_i)^2, 
-& \textrm{for $a \Delta v_{i} + \dot{a} \Delta x_i < 0$}  \\
-0 & \textrm{otherwise}\\
-\end{array} \right. 
-\end{eqnarray}
-and
-\begin{equation}
-Q_{ij} = 0  \;\;\;{\rm for}\;\; i \ne j .
-\end{equation}
-
-$\Delta x_i$ and $\Delta v_{i}$ refer to the comoving 
-width of the grid cell
-along the $i$-th axis and the corresponding difference in gas
-peculiar velocities across the grid cell, respectively, and $a$ is the
-cosmological scale factor.  $Q_{\rm AV}$ is a constant with a typical
-value of 2. We refer the interested reader to \citet{1994ApJ...429..434A} for more details.
-
-The limitation of a technique that uses an artificial viscosity is that,
-while the correct Rankine-Hugoniot jump conditions are achieved,
-shocks are broadened over 6-8 mesh cells and are thus not treated as true
-discontinuities. This may cause unphysical pre-heating of gas upstream
-of the shock wave, as discussed in \citet{1994ApJ...429..434A}.
-
-%---------------------------------------------------------------------------%
-\subsection{Chemistry, radiation and cooling}\label{sec.ov.chem}
-
-\red{
-As this stands it is the methods that I (brian) use, so I don't really emphasize
-either the tabular cooling functions or the various radiation background options.
-Add more discussion of this, point towards appendix.
-}
-
-Though the equations of hydrodynamics described above are a closed system,
-they are still missing a crucial piece of physics: radiative heating and cooling.
-Radiative cooling is extremely important in many astrophysical situations, as is 
-the heating of gas from some sort of radiation background.  \enzo\ has a very simple
-Sutherland and Dopita equilibrium cooling function \citep{1993ApJS...88..253S} 
-implemented, which uses a
-cooling table assuming a fixed metallicity of $Z = 0.3 Z_\odot$, and also a
-nonequilibrium heating/cooling model that assumes gas of primordial composition exposed to a
-uniform metagalactic ultraviolet (UV) background that varies with time \citep{1996ApJS..105...19K}.
-
-The simulations discussed in this thesis almost exclusively use the nonequilibrium
-routines, described in great detail by Abel et al. and Anninos et al.  
-\citep{abel97,anninos97} and summarized in Appendix~\ref{app:chemistry}. These 
-routines follow the non-equilibrium chemistry of a 
-gas of primordial composition with 9 total species:  
-$H$, $H^+$, $He$, $He^+$, $He^{++}$, $H^-$, $H_2^+$, $H_2$, and $e^-$.  The code also calculates 
-radiative heating and cooling, following atomic line excitation, recombination,
-collisional excitation, free-free transitions, molecular line excitations, and Compton
-scattering of the cosmic microwave background, as well as any of
-approximately a dozen different models for a metagalactic UV background that heat
-the gas via photoionization and photodissociation. 
-The multispecies rate equations are solved out of
-equilibrium to properly model situations where, e.g., the cooling rate of the gas
-is much shorter than the hydrogen recombination time.  The effect of this nonequilibrium
-cooling is to leave behind a much larger fraction of residual free electrons than one would
-expect if the assumption of equilibrium were being made.  The practical effect of this
-is that more $H^-$ is formed, which subsequently produces hydrogen molecules.  If large
-amounts of $H_2$ is formed it can greatly increase the cooling rate of primordial gas
-at relatively low temperatures ($T \leq 10^4$ K).  This can efficiently cool the gas 
-to approximately $200$ K, which significantly reduces the Jeans mass of the gas.  Correct
-modeling of the formation of molecular hydrogen is crucial to the study of star formation
-in a primordial gas.
-
-A total of 9 kinetic equations are solved, including 29 kinetic and radiative 
-processes, for the 9 species mentioned above.  See Table~\ref{table.collisional} 
-for the collisional processes and Table~\ref{table.radiative} for the 
-radiative processes solved.
-
-%---------------- table of collisional processes
-
-\begin{table}
-\begin{center}
-{\bfseries Collisional Processes}\\[1ex]
-\begin{tabular}{llllllll}
-(1) & H & + & e$^-$ & $\rightarrow$ & H$^+$ &+& 2e$^-$ \\
-(2) & H$^+$ &+ &e$^-$ & $\rightarrow$ & H &+ &$\gamma$ \\
-(3) & He &+& e$^-$ & $\rightarrow$ & He$^+$ &+& 2e$^-$  \\
-(4) & He$^+$ &+& e$^-$ & $\rightarrow$ & He &+ &$\gamma$  \\
-(5) & He$^{+}$ &+& e$^-$ & $\rightarrow$ & He$^{++}$ &+& 2$e^-$  \\
-(6) & He$^{++}$ &+& e$^-$ & $\rightarrow$ & He$^+$ &+& $\gamma$ \\
-(7) & H &+& e$^-$ &$\rightarrow$& H$^-$ &+& $\gamma$  \\
-(8) & H$^-$ &+& H &$\rightarrow$ & H$_2$ & +& e$^-$ \\
-(9) & H &+ &H$^+$ &$\rightarrow$ &H$_2^+$ &+ &$\gamma$ \\
-(10) & H$_2^+$ &+ &H &$\rightarrow$ &$H_2$ &+ &$H^+$ \\
-(11) & H$_2$ &+ &H$^+$ &$\rightarrow$ &H$_2^+$ & +& H \\
-(12) & H$_2$ &+ &e$^-$ & $\rightarrow$ & 2H & + & e$^-$  \\
-(13) & H$_2$ & + & H & $\rightarrow$ & 3H &   &      \\
-(14) & H$^-$ & + & e$^-$ & $\rightarrow$ & H & + & 2e$^-$ \\
-(15) & H$^-$ & + & H & $\rightarrow$ & 2H & + & e$^-$ \\ 
-(16) & H$^-$ & + & H$^+$ & $\rightarrow$ & 2H & & \\
-(17) & H$^-$ & + & H$^+$ & $\rightarrow$ & H$_2^+$ & + & e$^-$ \\
-(18) & H$_2^+$ & + & e$^-$ & $\rightarrow$ & 2H & & \\
-(19) & H$_2^+$ & + & H$^-$ & $\rightarrow$ & H$_2$ & + & H  \\
-(20) & 3H & & & $\rightarrow$ & H$_2$ & + & H
-\end{tabular}
-\caption[]{Collisional processes solved in the Enzo nonequilibrium
-primordial chemistry routines.}
-\label{table.collisional}
-\end{center}
-\end{table}
-
-\begin{table}
-\begin{center}
-{\bfseries Radiative Processes}\\[1ex]
-\begin{tabular}{llllllll}
-(21) & H & + & $\gamma$ & $\rightarrow$ & H$^+$ & + & e$^-$ \\
-(22) & He & + & $\gamma$ & $\rightarrow$ & He$^+$ & + & e$^-$ \\
-(23) & He$^+$ & + & $\gamma$ & $\rightarrow$ & He$^{++}$ & + & e$^-$ \\
-(24) & H$^-$ & + & $\gamma$ & $\rightarrow$ & H & + & e$^-$ \\
-(25) & H$_2$ & + & $\gamma$ & $\rightarrow$ & H$_2^+$ & + & e$^-$ \\
-(26) & H$_2^+$ & + & $\gamma$ & $\rightarrow$ & H & + & H$^+$ \\
-(27) & H$_2^+$ & + & $\gamma$ & $\rightarrow$ & 2H$^+$ & + & e$^-$ \\
-(28) & H$_2$ & + & $\gamma$ & $\rightarrow$ & H$_2^*$ & $\rightarrow$ & 2H \\
-(29) & H$_2$ & + & $\gamma$ & $\rightarrow$ & 2H &  & 
-\end{tabular}
-\caption[]{Radiative processes solved in the Enzo nonequilibrium
-primordial chemistry routines.}
-\label{table.radiative}
-\end{center}
-\end{table}
-
-The chemical reaction equation network is technically challenging to solve due to 
-the huge range of reaction timescales involved -- the characteristic creation
-and destruction timescales of the various species and reactions can differ by 
-many orders of magnitude.  As a result, the set of rate equations is extremely 
-stiff, and an explicit scheme for integration of the rate equations can be 
-exceptionally costly if small enough timesteps are taken to keep the network
-stable.  This makes an implicit scheme much more preferable for such a set of 
-equations.  However, an implicit scheme typically require an iterative 
-procedure to converge, and for large networks (such as this one) an implicit
-method can be very time-consuming, making it undesirable for a large, three-dimensional
-simulation.
-
-\enzo\ solves the rate equations using a method based on a backwards differencing 
-formula (BDF) in order to provide a stable and accurate solution.  This technique is
-optimized by taking the chemical intermediaries $H^-$ and $H_2^+$, which have 
-large rate coefficients and low concentrations, and grouping them into a separate
-category of equations.  Due to their fast reactions these species are very sensitive
-to small changes in the more abundant species.  However, due to their low overall
-concentrations, they do not significantly influence the abundance of species with
-higher concentrations.  Therefore, reactions involving these two species can be
-decoupled from the rest of the network and treated independently.  In fact, $H^-$ 
-and $H_2^+$ are treated as being in equilibrium at all times, independent of 
-the other species and the hydrodynamic variables.  This allows a large speedup
-in solution as the BDF scheme is then applied only to the slower 7-species network
-on timescales closer to those required by the hydrodynamics of the simulation.
-Even so, the accuracy and stability of the scheme is maintained by subcycling the 
-rate solver over a single hydrodynamic timestep.  These subcycle timesteps are 
-determined so that the maximum fractional change in the electron concentration is
-limited to no more than $10\%$ per timestep.
-
-It is important to note the regime in which this model is valid.  According to Abel et al. and
-Anninos et al. \citep{abel97,anninos97}, the reaction network is valid for temperatures
-between $10^0 - 10^8$ K.  The original model discussed in these two references is only
-valid up to $n_H \sim 10^4$~cm$^{-3}$.  However, addition of the 3-body $H_2$ formation
-process (equation 20 in Table~\ref{table.collisional}) allows 
-correct modeling of the chemistry of the gas up
-until the point where collisionally-induced emission from molecular hydrogen becomes an important
-cooling process, which occurs at $n_H \sim 10^{14}$~cm$^{-3}$.  A further concern is that
-the optically thin approximation for radiative cooling breaks down, which
-occurs before $n_H \sim 10^{16} - 10^{17}$~cm$^{-3}$.  Beyond this point, 
-modifications the cooling function that take into account the non-negligible
-opacity in the gas must be made, as discussed by \citet{2004MNRAS.348.1019R}. 
-Even with these modifications, a completely correct description of the cooling of
-this gas will require some form of radiation transport, which will greatly 
-increase the cost of the simulations.
-
-Several processes are neglected.  The deuterium atom and its processes are 
-completely ignored, which may have some effect.  Recent work shows that HD is
-a more effective coolant than previously thought~\citep{2005MNRAS.361..850L}.  However,
-the fractional abundance of HD is so low that under circumstances relevant to
-the formation of a Population III star in an un-ionized region it should be
-sub-dominant.  However, the enhanced electron fraction in fossil 
-HII regions (as discussed later in this thesis) could result in the HD molecule
-becoming a dominant cooling mechanism at relatively low ($\sim$ few hundred K)
-temperatures, and could potentially cool the gas down to below $100$~K, which can
-enhance fragmentation and could have important consequences for the IMF of 
-primordial stars forming in a relic HII region~\citep{2005MNRAS.364.1378N}.
-
-Aside from deuterium, the chemical reactions involving lithium 
-are also neglected.  According to \citet{1998A&A...335..403G}, these are 
-not important for the density and temperature regimes explored 
-by the simulations discussed in this thesis.  However, at higher densities it 
-is possible that there are regimes where lithium can be an important coolant.
-
-%---------------------------------------------------------------------------%
-\subsection{Star formation and feedback}\label{sec.ov.star}
-
-\red{By the time this paper is submitted my (BWO's) paper on the star formation
-and feedback algorithms, and the various parameters contained therein, should be
-at least written and submitted as well.  Point to this, and maybe include a couple
-interesting plots (perhaps in an appendix?).}
-
-While the physics discussed previously is all crucial to the study of
-cosmological structure formation, most cosmological observations are of
-stars and related phenomena.  Additionally, stars eject energy and metal-
-enriched gas throughout their lives, drastically modifying their own 
-environment.  The formation of galaxies and clusters cannot be completely
-modeled without including the feedback of energy and metals.  In particular,
-it is thought that feedback is crucial for suppressing the large numbers of
-dwarf-like galaxies that CDM theories predict 
-\citep{1991ApJ...381...14L,1991ApJ...379...52W}.  An early
-burst of star formation could remove a large fraction of cold gas from such systems
-\citep{1978MNRAS.183..341W,1991ApJ...367...45C}.  Also, the unexpectedly 
-low luminosities of small clusters and
-groups (relative to the richest clusters) is commonly explained through feedback
-\citep{1991ApJ...383..104K}.  Energy ejected during the formation of the cluster ellipticals 
-lowers the central density and hence the X-ray luminosity of such clusters 
- \citep{1997ApJ...484L..21C}.  Therefore, the inclusion of star formation and the 
-feedback of energy and metals into the intergalactic medium in a cosmological
-code is crucial for many reasons.
-
-We have extended the \enzo\ code to include multiple models for star formation
-and feedback.  It is difficult to directly simulate the formation of individual
-stars in the context of simulations of galaxy formation and evolution due
-to the immense computational cost.  Therefore, 
-we adopt parametric methods which attempt
-to model star forming regions.  One model is based on Kravtsov's method
- \citep{2003ApJ...590L...1K}, and the other is based on the method of 
-\citet{1992ApJ...399L.113C}, which has been modified for use in an AMR code.  The basic
-ideas behind the methods are straightforward and observationally motivated.
-Similar (though somewhat more advanced) methods have been employed for the 
-smoothed particle hydrodynamics
-(SPH) method, most recently by Springel \& Hernquist 
-\citep{sh03a,sh03b,hs03}. This 
-method assumes a
-multiphase IGM and has been shown to accurately reproduce the cosmic star
-formation rate, and will be implemented into \enzo\ in the near future.
-
-In the following sections we will detail both the Kravtsov and Cen \& Ostriker
-star formation methods separately.  Tests of the star formation and
-feedback algorithms will not be shown, since this work is currently in
-progress (and an extension of the AMR/SPH code comparison described in
-\citet{2005ApJS..160....1O}).
-
-\subsubsection{The Kravtsov star formation and feedback algorithm}
-
-Kravtsov's method of star formation is designed to reproduce
-the global Schmidt law of star 
-formation~\citep{2003ApJ...590L...1K,1959ApJ...129..243S}.
-This algorithm is deliberately minimal, and is explicitly geared
-towards modeling star formation in a phenomenological way on kiloparsec
-scales.  Stars are assume to form with a characteristic gas timescale 
-$\tau_*$ such that $\dot{\rho}_* = \rho_{gas}/\tau_*$ where $\rho_{gas}$ 
-and $\rho_*$ are the baryon gas and stellar densities, respectively.  
-This ``constant efficiency'' model on the scale of star formation 
-regions is well motivated observationally \citep{1996AJ....112.1903Y,2002ApJ...569..157W}.  
-Star formation is only allowed to take place in very dense regions
-with $\rho_{gas} \geq \rho_{SF}$, where $\rho_{SF}$ is a constant
-proper density threshold above which star formation is allowed to
-occur.  No other criteria are imposed.  Kravtsov's typical choices
-for $\tau_*$ and $\rho_{SF}$ are $\tau_* = 4$ Gyr and $\rho_{SF} =
-1.64~$M$_\odot$~pc$^{-3}$~$(n_H \sim 50$~cm$^{-3})$.  The adopted timescale
-is derived from the observationally-determined normalization of the
-Schmidt law, and the density threshold is determined by observations
-of star forming regions on $\sim 100$ pc scales.  Note that the density
-threshold is in proper, not comoving, units.
-
-Algorithmically, the star formation events in Kravtsov's code
-are assumed to occur once every global time step 
-$\Delta t_0 \leq 10^7$.  In cells where star formation is
-determined to occur (i.e. $\rho_{gas} \geq \rho_{SF}$) a
-collisionless ``star particle'' is assumed to form, with a 
-mass $m_* = \dot{\rho}_* V_{cell} \Delta t_0$, where
-$\dot{\rho}_*$ is described above and $V_{cell}$ is the 
-volume of the mesh cell.  These star particles are
-formed instantaneously, with all $m_*$ of gas being
-taken out of the cell and immediately deposited into the
-star particle.  This particle is then given the velocity
-of the gas in the cell, and thereafter treated as a 
-collisionless particle.  The \enzo\ implementation
-of this algorithm is similar, except that instead of 
-forming stars only at the root grid time step, we allow 
-stars to form at the time step of the highest level of
-resolution at any particular point in space.  As can be seen
-from the equation for $m_*$ above, this can result in very 
-small stellar masses.  To avoid memory and processor time
-issues related to having very large numbers of star particles
-we impose a threshold mass $M_{*,min}$ such that a star
-particle only forms if $m_* \geq M_{*,min}$.  An
-appropriate choice of $M_{*,min}$ does not significantly 
-change the star overall star formation history of a 
-simulation, though it may delay the onset of star formation 
-in a given cell relative to a simulation without a particle mass
-threshold.
-
-Each ``star particle'' is assumed to represent an ensemble
-of stars and is treated as a single-age stellar population.
-Kravtsov assumes that the IMF is described by a
-Miller \& Scalo functional form with stellar masses between
-$0.1$ and $100~M_\odot$ \citep{1979ApJS...41..513M}.  All stars in this
-IMF with $M_* > 8~M_\odot$ deposit $2 \times 10^{51}$ ergs of
-thermal energy and a mass $f_z M_*$ of metals into the
-cell in which they form without delay, with
-$f_z \equiv min(0.2, 0.01~M_*-0.06)$ (i.e. instantaneous
-deposition of metals).  The definition of $f_z$ is a rough
-approximation of the results of 
-\citet{1995ApJS..101..181W}.
-
-Kravtsov reports that simulations with this algorithm
-reliably reproduce the star formation rate-gas surface
-density relation of the Schmidt law, and are not particularly
-sensitive to numerical parameters \citep{2003ApJ...590L...1K}.
-He also notes that this is surprisingly insensitive to
-the presence or absence of feedback and details of the cooling
-and heating properties of the gas, which suggests that the
-global star formation rate is determined by gravitationally-
-driven supersonic turbulence (on large scales)
-rather than stellar feedback or thermal instabilities on small 
-scales.
-
-\subsubsection{The Cen \& Ostriker star formation algorithm}
-
-The Cen \& Ostriker method is a heuristic model of star formation
-on galactic scales.  This method, first described by
-\citet{1992ApJ...399L.113C}, is similar in some ways to the Kravtsov algorithm
-but has more complex criteria for determining where stars should
-be formed.  In this method, cells that form stars must have a 
-baryon overdensity higher than some threshold 
-$\rho_b/\bar{\rho}_b \geq \eta$ 
-where $\rho_b$ is the baryon density in that cell, 
-$\bar{\rho_b}$ is the mean baryon density in the simulation,
-and $\eta$ is the user-defined overdensity threshold.
-Additionally, the gas in the cells must be contracting, 
-cooling rapidly, and gravitationally unstable, e.g.:
-
-\begin{equation}
-\nabla \cdot \vec{v}_b < 0,
-\label{cencont}
-\end{equation}
-
-\begin{equation}
-t_{cool} < t_{dyn} \equiv \sqrt{3 \pi / 32G \rho_{tot}},
-\end{equation}
-
-\begin{equation}
-m_{b} > m_{J} \equiv G^{-3/2} \rho_{b}^{-1/2}C^{3}
-\left[ 1 + \frac{\delta\rho_{d}}{\delta\rho_{b}} \right]^{-3/2}
-\end{equation}
-
-where $\vec{v}$ is the velocity of the gas in the cell, $\rho_{b}$ and 
-$\rho_{d}$ are the cell's baryon and dark matter density, respectively,
-$\rho_{total} = \rho_{b} + \rho_{d}$, $m_{b}$ and $m_{j}$ are the 
-baryonic mass in the cell and jeans mass of the cell, and C is the 
-isothermal sound speed in the cell.  
-If all of these criteria are met, 
-the mass of a star particle is calculated as \(m_{*} = m_{b} 
-\frac{ \Delta t}{ t_{dyn} } f_{*eff} \), 
-where $f_{*eff}$ is the star formation efficiency parameter.
-
-If $m_{*}$ is greater than a minimum star mass $m_{*min}$, a particle 
-is created and given several attributes:  Mass, a unique index number, 
-the time of formation $t_{form}$, the local dynamical free-fall time 
-$t_{dyn}$ and the metallicity fraction of the baryon gas in the cell 
-$f_{Zb}$.  There is a user-defined minimum dynamical time
-$T_{dyn,min}$ which is observationally motivated and affects the
-feedback rates (see below).
-The particle is placed in the center of the cell and given 
-the same peculiar velocity as the gas in the cell, and is then treated 
-in the same manner as the dark matter particles.  An 
-amount of baryon gas corresponding to the new particle's mass is 
-removed from the cell.
-
-In addition, we have added a stochastic star formation algorithm that keeps 
-track of all of the sub-mass threshold stars that should have been created 
-and when the total mass of uncreated stars is greater than the minimum mass, 
-a star particle is created.
-
-The feedback of energy and metals into the baryon gas is similar to
-the Kravtsov feedback, with some important differences.  The star formation 
-algorithm creates each star particle instantaneously.  However, feedback
-should take place over a significant timescale, as all of the stars contained
-within the ``star particle'' would form over a long period of time.  Therefore,
-we assume that for the purposes of feedback that the mass of stars formed
-at a time $t$ with finite timestep $\Delta t$ is:
-
-\begin{eqnarray}  
-\Delta m_{sf} =  m_{*} \left[ \right. \left(1 + \frac{t-t_{form}}{t_{dyn}} 
-                    \right) \nonumber \\ 
-		   \times exp \left( \frac{-(t-t_{form})}{t_{dyn}} \right) 
-                    \nonumber \\ 
-	       - \left( 1 + \frac{t+ \Delta t-t_{form}}{t_{dyn}} \right) \nonumber \\ 
-                    exp \left( \frac{-(t+\Delta t-t_{form})}{t_{dyn}} \right) 
-                   \left. \right]
+x^{n+1/2} & = & x^n + (\Delta t/2) v^{n} \nonumber \\
+v^{n+1} & = & v^n + \Delta t a^{n+1/2} \\
+x^{n+1} & = & x^{n+1/2} + (\Delta t/2) v^{n+1} \nonumber
 \end{eqnarray}
 
-which can be represented more clearly in integral form:
-
-\begin{eqnarray}
-\int_{t}^{t+Dt} \frac{dM}{dt} dt = \int_{t}^{t+Dt} m_{*} 
-\frac{dt}{t_{tyn}} \left( \frac{t-t_{form}}{t_{dyn}} \right) \nonumber \\ 
-\times exp \left( - ~ \frac{t-t_{form}}{t_{dyn}} \right) 
-\end{eqnarray}
-
-During this timestep, we assume that the star particle feeds back
-metal-enriched gas and thermal energy from supernovae and from stellar winds.
-Since massive stars have very short lifetimes, we assume that there is an
-immediate feedback of some fraction $f_{SN}$ of the rest energy from the
-stars created that timestep into the baryon gas, such that 
-$E_{add} = f_{SN} \Delta m_{sf} c^2$, where c is the speed of light.
-In addition, a fraction $f_{Z*}$ of the metal from the star particle
-is fed back into the baryon gas, which takes into account the effects
-of metal recycling.  Finally, a fraction of the mass $f_{m*}$ from all
-stars (rather than just supernovae) is fed back into the gas along with
-momentum in order to simulate the mass ejection from non-exploding stars
-via stellar winds.
-
-There are six user-defined parameters in this algorithm:  three deal with the
-star formation ($\eta$, $m_{*min}$ and $T_{dyn,min}$), and three 
-deal with feedback ($f_{SN}$, $f_{Z*}$ and $f_{m*}$).  Some of these
-parameters are completely free, while others can be guided by observation
-or theory.  For example, the supernova feedback parameter, $f_{SN}$, can be 
-constrained assuming that, for every $200 M_\odot$ of stars created, one
-supernova occurs, and this event feeds back approximately $10^{51}$ ergs of
-thermal energy, giving:
-
-\begin{equation}
-f_{SN} = \frac{10^{51} ergs}{200~M_\odot c^2} \simeq 3 \times 10^{-6}
-\end{equation}
-
-The metal yield $f_{Z*}$, defined as the mass in metals produced per unit
-mass of stars created, can be constrained by a theoretical model of
- \citet{1995ApJS..101..181W}.  This model suggests that $f_{Z*} = 0.02$
-is an appropriate number.  The minimum dynamical time is set to be $T_{dyn,min} = 10^7$
-~years to agree with timescales seen in nearby OB associations.
-
-The other parameters, such as the overdensity threshold $\eta$, minimum star
- mass $m_{*min}$, and mass ejection fraction $f_{m*}$ are not well constrained 
-either theoretically or observationally.  Indeed, $m_{*min}$ is a purely
-numerical parameter designed to keep the code from producing too many star 
-particles, and thus has no observational or theoretical counterpart.  These
-parameters have to be set by performing parameter studies and comparing to 
-observations.  Unfortunately, the range of parameter space is large, and
-the results may be degenerate for some combinations of these parameters.
-
-%---------------------------------------------------------------------------%
-\subsection{Adaptive mesh refinement}\label{sec.ov.amr}
-\dcc{removed most of the notes.  Added intro blurb.  moved parallelism
-  stuff. Re-worded the bit on ghost zones.  Summary of 'grid as
-  independent update + syncing with neighbors and parents,' W-cycle. Added
-  sentence on subgrid creation.  Made timestepping its own subsection.}
-\red{Talk about reasonable refinement factors - which 
-ones are usable, etc?}
-
-The primary purpose of the \enzo\ code is its Adaptive Mesh Refinement
-capability, which allows it to reach extremely large dynamical ranges
-with limited computational resources, opening doors previously closed
-by finite memory and computational time. Unlike moving mesh methods
-\citep{1995ApJS..100..269P,1995ApJS...97..231G} or  
-methods that subdivide 
-individual cells \citep{Adjerid}, Berger \& Collela's AMR (also referred 
-to as \emph{structured} AMR) utilizes an adaptive hierarchy of grid 
-patches at varying levels of resolution.  Each rectangular grid patch 
-(referred to as a ``grid'') covers some region of space in its 
-\emph{parent grid} which requires higher resolution, and can itself 
-become the parent grid to an even more highly resolved \emph{child grid}. 
-ENZO's implementation of structured AMR places no restrictions on 
-the number of grids at a given level of refinement, or on the number of 
-levels of refinement.  However, owing to limited computational resources 
-it is practical to institute a maximum level of refinement $\Lmax$.
-
-The \enzo\ implementation of AMR allows arbitrary integer ratios of parent and
-child grid resolution.  However, in practice 
-refinement factors (defined as the ratio of parent grid resolution to child grid resolution) 
-of two or four are typically used, with a refinement factor
-of two being most commonly used for cosmological simulations for reasons of
-efficiency. The ratio of
-box size to the maximum grid resolution of a given simulation is
-therefore $L/e = \Nroot \times 2^{\Lmax}$, where $\Nroot$ is the
-number of cells along one edge of the root grid, $\Lmax$ is the
-maximum level of refinement allowed, and L and e are the box size
-and grid resolution of the most highly refined region, respectively.
-
-The AMR grid patches are the primary data structure in \enzo.  Each
-patch is treated as an individual object which can contain both field
-variables and particle data.  Individual grids are organized into a 
-dynamic, distributed hierarchy of mesh patches.  
-See 
-Figure~\ref{fig.2.amrhierdens} for an example of an \enzo\
-simulation performed using this AMR scheme where both the baryon density
-and AMR grid hierarchy are shown together.
-
-\begin{figure}
-\begin{center}
-\includegraphics[width=0.4\textwidth]{figures/amr_hierarchy.eps}
-\end{center}
-\caption{\emph{Left:}  Example of a simple, 
-distributed AMR hierarchy showing real and 
-ghost grids.
-\emph{Right:}  Example 2D \enzo\ grid showing real and ghost zones, as 
-needed for the PPM hydro stencil.  Image courtesy of James Bordner (Cent.
-Astrophysics and Space Sciences, UCSD).
-}
-\label{fig.2.amrhierarchy}
-\end{figure}
-
-\begin{figure}
-\begin{center}
-\includegraphics[width=0.4\textwidth]{figures/amr-hier-dens.eps}
-\caption{Example of an \enzo\ simulation showing the AMR grid hierarchy.  
-This is a simulation of the collapse of a single massive halo with a 
-$32^3$ root grid and two $32^3$ static nested grids.  AMR is only allowed
-within the most highly refined static nested grid.  Log baryon density is shown
-by colored isocontours with values corresponding to the legend at center left.  
-The rectangular solid wire frames correspond to individual 
-AMR grids, with different colors corresponding to level as indicated by the
-legend at top left.}
-\label{fig.2.amrhierdens}
-\end{center}
-\end{figure}
-
-Each grid patch in \enzo\ contains arrays of values for baryon and 
-particle quantities.   Grids are 
-partitioned into a core of \emph{real zones} and a surrounding layer 
-of \emph{ghost zones}.  The real zones store field values and ghost 
-zones are used to temporarily store values which have been obtained 
-directly from neighboring grids or interpolated from a parent grid.  
-These zones are necessary to accommodate the computational stencil of 
-the hydrodynamics solvers (Sections~\ref{sec.ov.hydro.ppm} and~\ref{sec.ov.hydro.zeus}) 
-and the gravity solver (Section~\ref{sec.ov.nbody}).  
-The PPM and Zeus hydro solvers require 3 layers of ghost zones,  and the gravity
-solver requires 6.  This can lead to significant memory 
-and computational overhead, particularly for smaller grid patches
-at high levels of refinement.  More information on filling of ghost
-zones can be found in the appendix.
-
-From the point of view of the hydro solver, each grid is solved as an
-independent CFD problem, with Dirichlet boundary conditions stored in
-the ghost zones.  At the end of each time step, cells on a given level
-fill ghost zones by interpolating from the parent and copying from
-neighbors on the same level.  Grids then correct the flux at the interface boundary between
-fine and coarse zones, and finally project their real zone data to the
-parent grid.
-
-Timesteps are determined on each level as
-described in \ref{sed.timestepping}, and the hierarchy is advanced on
-a level by level basis in what's known as a W-Cycle.    Beginning with
-the coarsest level, $L$, all grids on that level are advanced one
-timestep.  Then one timestep is taken on all grids at the next level
-of refinement, $L+1$, and so on until the finest level is resolved.
-The finest level is then advanced, using as many steps as it takes to
-reach the level immediately above.  The finest level is then synced
-to the level above it, which then proceeds forewards one more step.
-\red{Needs a picture of the W cycle.}
-
-At the end of every timestep on every level each grid updates 
-its ghost zones by exchanging information with its neighboring grid patches 
-(if any exist) and/or by interpolating from a parent grid.  
-In addition, cells are examined to see if they should be refined or 
-de-refined, and the entire grid hierarchy is rebuilt at that 
-level (including all more highly refined levels).  The timestepping and 
-hierarchy advancement/rebuilding process described here is repeated 
-recursively on every level to the specified maximum level of 
-refinement in the simulation.
-
-Since the addition of more highly refined grids is adaptive,  
-the conditions for refinement must be specified.  The criteria of 
-refinement can be set by the threshold value of the overdensity of 
-baryon gas or dark matter in a cell (which is really a refinement on
-the mass of gas or DM in a cell), the local Jeans length, the 
-local density gradient, or local pressure and energy gradients.  
-A cell reaching any or all of these criteria 
-will then be flagged for refinement.  
-Once all cells at a given level have been examined, rectangular boundaries 
-are determined which minimally encompass the flagged cells. The set of
-flagging cells is then further fractured into smaller bounding
-rectangles to ensure computational 
-efficiency.  More details on subgrid creation can be found in Appendix
-\ref{app.subgrid_creation}.  A refined grid patch is 
-introduced within each such bounding rectangle. Thus the cells needing 
-refinement, as well as adjacent cells within the patch which do not need 
-refinement, are refined. While this approach is not as memory efficient as 
-cell-splitting AMR schemes, it offers more freedom with 
-finite difference stencils. For example, PPM requires a stencil of seven 
-cells per dimension. This cannot easily be accommodated in cell-splitting 
-AMR schemes.
-
-\blue{ \emph{I think this  should be
-  re-written to reflect data presented or referenced in this paper,
-  whatever that winds up being.}
-In the simulations discussed in this thesis we typically use baryon and
-dark matter overdensities as our refinement criteria, though
-for some higher-resolution simulations we also refine on other
-criteria as needed.}
-
-\subsection{Time Stepping}\label{sec.timestepping}
-\dcc{Made this its own section (not really part of the AMR, it's more
-  general than that.)  Added stuff on harmonic average for 3d. See
-  note below on acceleration timestep criterion being crap. Moved
-  w-cycle to the previous section}
-In \enzo, resolution of the equations being solved is adaptive in time 
-as well as in space.  The timestep in \enzo\ is satisfied on a level-by-level
- basis by finding the largest timestep such that multiple criteria are
-satisfied on each level.  The timestep criteria used by Enzo are 
-(showing the one-dimensional case for clarity):
-
-\begin{equation}
-\Delta t_{hydro} = min \left( \kappa_{hydro} \frac{a \Delta x}{c_{s} + |v_x|} \right)_L ,
-\label{eqn:dthydro}
-\end{equation}
-
-\begin{equation}
-\Delta t_{dm} = min \left(\kappa_{dm} \frac{a \Delta x}{v_{dm,x}} \right)_L ,
-\label{eqn:dtdarkmatter}
-\end{equation}
-
-\begin{equation}
-\Delta t_{exp} = f_{exp} \left( \frac{a}{\dot{a}} \right) ,
-\label{eqn:dtexpand}
-\end{equation}
-
-\begin{equation}
-\Delta t_{accel} = min \left( \sqrt{\frac{\Delta x}{\vec{g}}} \right)_L 
-\label{eqn:dtaccel}
-\end{equation}
-
- In equations~\ref{eqn:dthydro}-\ref{eqn:dtaccel}, the $min ( \ldots )_L$
-formalism means that this value is calculated for all cells on a given level
-L and the minimum overall value is taken as the timestep.  
-Equation~\ref{eqn:dthydro} ensures that all cells satisfy the 
-Courant-Freidrichs-Levy (CFL) condition for accuracy and stability of an explicit
-finite difference discretization of the Euler equations.  Effectively this condition
-forces the timestep to be small enough such that any changes in the fluid propagate
-less than a single grid spacing, $\Delta x$.  In this equation, $\kappa_{hydro}$ is 
-a numerical constant with a value of $0 < \kappa_{hydro} \leq 1$ (with a typical
-value of $\kappa_{hydro} \sim 0.3-0.5$) that ensures that the CFL condition is always
-met, and $c_s$ and $v_x$ are the sound speed and peculiar baryon
-velocity in a given cell.
-
-Equation~\ref{eqn:dtdarkmatter} is analogous to Equation~\ref{eqn:dthydro} and ensures
-accuracy in the N-body solver by requiring that no dark matter particle travels more than
-one cell width.  The parameter $\kappa_{dm}$ is analogous to $\kappa_{dthydro}$, with
-an identical range of values.  Equation~\ref{eqn:dtexpand} limits the timestep such 
-that the expansion parameter a can only change by a fractional amount of $f_{exp} = \Delta a/a$, 
-where $f_{exp}$ is a user-defined parameter and has typical values of $f_{exp} = 0.01-0.02$.
-This is required for the stability of the PPM algorithm in comoving coordinates, and
-typically limits the timestep only at high redshifts when densities are relatively
-homogeneous.
-
-Equation~\ref{eqn:dtaccel} is supplementary to equation~\ref{eqn:dthydro} in that it
-takes into account the possibility of large accelerations causing numerical 
-instabilities by violating the Courant condition.  In this equation, $\vec{g}$ is the
-gravitational acceleration in each cell on level L.  \blue{ \emph{It should be
-  noted that the timestep is computed before the acceleration field is computed,
-  and the acceleration field is deleted at the end of each time step.
-  So this condition isn't ever actually enforced.  We should either
-  fix this problem, or not discuss it.
- -dcollins.} }
-
-Equation ~\ref{eqn:dthydro} is valid for one dimension.  For 2 or 3
-dimensions, it was shown by \cite{Godunov1959}  that using the
-harmonic average of the timestep found along each of the coordinate
-axes yields a maximum $\kappa_{hydro} = 0.8$.  So letting $\Delta
-t_x$, $\Delta t_y$, and $\Delta t_z$ be the analogues of
-eqn. \ref{eqn:dthydro} along the $x,y$ and $z$ axes, 
-\begin{equation}
-\Delta t_{hydro} = min ( \kappa_{hydro} /( \frac{1}{\Delta t_x}
-+\frac{1}{\Delta t_y} + \frac{1}{\Delta t_z} ) )_L
-\end{equation}
-
-For all other criteria, multiple dimensions are accounted for by
-repeating the one dimensional criterion along each axis, and taking
-the minimum.
-
-%---------------------------------------------------------------------------%
-\subsection{Parallelism}\label{sec.ov.parallel}
-\dcc{moved hierarchy description, parallism notes from AMR section
-  here. Mentioned performance/scaling study.  Still short.}
-
-When \enzo\ is run on a distrubuted platform, some of the data is
-replicated to all processors, while some remains on a single processor.
-A description of the entire hierarchy (the position and size of each grid, as well
-some other meta data) is stored on each processor.  This enables each
-grid to know what grids are near it in the hierarchy, regardless of
-where the data is.  The actual Baryon data (density, velocity, energy,
-and any chemistry data) and particle data are stored on only one
-processor.  Storing the entire hierarchy on all processor does create
-some overhead, but in practice this is not an issue until one reaches
-the extremely large scale.
-  The code handles load balancing 
-on a level-by-level basis such that the workload on each level is 
-distributed as uniformly as possible across all processors.  Spatial locality of 
-grids is not forced during message passing, for maximum flexibility (though not
-necessarily maximum efficiency).  
-The MPI message passing library\footnote{http://www-unix.mcs.anl.gov/mpi/}
- is used to transfer data between processors.
-
-The performance of \enzo\ has been extensively studied using the
-software package LCAPerf.  More on LCAPerf can be found in section
-\ref{sec.performance}
-
-%old text
-%Every processor 
-%keeps a description of the entire grid hierarchy at all times, so that 
-%each processor knows where all grids are.  However, baryon and particle 
-%data for a given grid only exists on a single processor.  See 
-%Figure~\ref{fig.2.amrhierarchy} for a schematic example of this.
+Particles are stored in the most highly refined grid patch at the point in space where they exist, and particles which move out of a subgrid patch are sent to the grid patch covering the adjacent volume with the finest spatial resolution, which may be of the same spatial resolution, coarser, or finer than the grid patch that the particles are moved from.
+This takes place in a communication process at the end of each timestep on a level.

### 2012-06-29 numerical-ppm.tex
first commit of paper.

\subsection{Hydrodynamics: The PPM hydro method}
\label{sec.hydro.ppm}

The primary hydrodynamic method used in \enzo\ is based on the
piecewise parabolic method (PPM) of
\citet{1984JCoPh..54..174C}, which has been
modified for the study of cosmological fluid flows.  
\citet{1984JCoPh..54..174C} describe two variants on this method: a
Lagrangian Remap variant and a Direct Eulerian variant.  Both have
been adapted for \enzo, and the Lagrangian Remap 
variant is described in \citet{1995CoPhC..89..149B}.  Because of its
conservative formulation, however, the Direct Eulerian variant is
preferred for \enzo\ AMR simulations.  We will describe the Direct
Eulerian (PPM-DE) variant here.

We can examine the structure of eqs.~(\ref{eq:mass})-(\ref{eq:total_energy})
to obtain a better understanding of the effect of each term.
The derivatives are first order in time
and provide us with the dependent variables that will be evaluated each
time step: $\rhob$, $\vecvb$, and $E$.

The second term on the left hand side of each relation
is due to our choice of Eulerian coordinates and corresponds to
advection, the flow of conserved quantities 
($\rhob$, $\rhob \vecvb$, and $\rhob E$) 
across the mesh.  These will be termed the transport terms.

The first term on the right hand side of eqs. (\ref{eq:momentum}) and
(\ref{eq:total_energy}) comes from our choice of comoving coordinates
(a similar term also appears in eq.~(\ref{eq:dm_velocity}), the velocity 
relation for the dark matter particles).  It does not
appear in the mass conservation equation (\ref{eq:mass}) because
of the comoving density definition.   
We note that these expansion terms could be
eliminated entirely by the proper choice of variables (including time),
although we have not done so here, as they do not constitute a major
source of error.

The remainder of the terms (due to pressure and gravity) are denoted as
source terms.  They drive instabilities and discontinuities in the flow.
We make these distinctions because each of the three
groups of terms have different numerical behaviour and are therefore
treated differently.  The source and transport steps will be solved in
a self-consistent fashion by the PPM scheme while the expansion terms,
which are spatially localized, are split off and solved in a separate step
(as are radiative cooling and heating, if present).

The remainder of the terms (transport and source) in the fluid equations
involve derivatives.  
Because we are interested in phenomena with no special geometry, we will
restrict discussion to Cartesian coordinates.
Once again we use the concept of operator splitting,
but now apply it spatially.   We can rewrite the one-dimensional (Eulerian) 
versions of eqs.~(\ref{eq:mass})--(\ref{eq:total_energy}) without expansion terms
in conservative form as,
\begin{eqnarray}
\frac{\partial \rho}{\partial t}  + \frac{1}{a} \frac{\partial \rho v }{\partial x}    & = &  
     0 \label{eq:mass1d} \\
\frac{\partial \rho v}{\partial t}  + \frac{1}{a} \frac{\partial \rho v^2}{\partial x}   + 
      \frac{1}{a} \frac{\partial p}{ \partial x} & = & 
      \rho \frac{1}{a} \frac{\partial \phi}{ \partial x}  \label{eq:momentum1d}  \\
\frac{\partial \rho E}{\partial t}  + \frac{1}{a} \frac{\partial \rho v E}{\partial x}  & =  &
      \rho v \frac{1}{a} \frac{\partial \phi}{\partial x}. \label{eq:total_energy1d}
\end{eqnarray}

%
Note that $x$ and $v$ refer to the one dimensional comoving position and 
peculiar velocity of the baryonic gas, and $g$ is the acceleration at the cell center.
We write this including the cosmological scale-factors but it reduces to the non-cosmological case with $a = 1$.
These equations are now in a form that can be solved by the split PPM scheme.

% ------------------------------------------------------------

\subsubsection{The piecewise parabolic method in one dimension} % 2.2

We now restrict ourselves to the solution of eqs. (\ref{eq:mass1d})--(\ref{eq:total_energy1d}), plus transport, in one dimension.  In the Direct Eulerian version of PPM, this is accomplished by a three-step process.  First, we compute effective left and right states at each grid boundary by constructing a piecewise parabolic description of the primary variables ($\rho$, $u$ and $E$) and then averaging over the regions corresponding to the distance each characteristic wave can travel ($u$, $u-c$ and $u+c$).  Second, a Riemann problem is solved using these effective left and right states, and finally the fluxes are computed based on the solution to this Riemann problem.  This is described in detail in \citet{1984JCoPh..54..174C}, but we will briefly outline the procedure here both for completeness and to put the changes we will make in context.

The Eulerian difference equations are:
%
\newcommand{\avp}[1]{\overline{#1}_{j+1/2}}
\newcommand{\avm}[1]{\overline{#1}_{j-1/2}}
\begin{equation}
\rho_j^{n+1} = \rho_j^{n} + \Delta t \left( 
                      \frac{ \avp{\rho}\avp{v} -  \avm{\rho}\avm{v} } {\Delta x_j}
            \right)
     \label{eq:mass_diff}
\end{equation}
\begin{equation}
\rho_j^{n+1} v_{j}^{n+1} =
       \rho_j^{n+1} v_j^n  + \Delta t \left(
          \frac{ \avp{\rho} \avp{v}^2 - \avm{\rho} \avm{v}^2 + \avp{p} - \avm{p}} {\Delta x_j}
             \right)
     + \frac{ \Delta }{2} g_j^{n+1/2} (\rho_j^n + \rho_j^{n+1}), 
     \label{eq:momentum_diff}
\end{equation}
\begin{eqnarray}
\rho_j^{n+1} E_j^{n+1}  = 
       \rho_j^n E_j^n  & + & \Delta t  \left(
            \frac{  \avp{\rho} \avp{v} \avp{E}  - \avm{\rho} \avm{v} \avm{E}  +
                       \avp{v} \avp{p}    - \avm{v} \avm{p} } {\Delta x_j}
             \right) \nonumber \\
         & & + \frac{ \Delta t }{2} g_j^{n+1/2} (\rho_j^n v_j^n + \rho_j^{n+1} v_j^{n+1} ).
     \label{eq:energy_diff}
\end{eqnarray}

We have used subscripts to indicate zone-centered ($j$)
and face-centered ($j+1/2$) quantities, while superscripts refer to
position in time.  The cell width is $\Delta x_j$.
Although they have been discretized
in space, the accuracy of the updates depend on how well we can compute
the fluxes into and out of the cell during $\Delta t$.  
This in turn depends on our ability to compute the time-averaged 
values of $p$, $\rho$ and $v$ at the cell interfaces, denoted here
by $\overline{p}_{j\pm 1/2}$, $\overline{\rho}_{j\pm 1/2}$, and $\overline{v}_{j\pm 1/2}$.  We now
describe the steps required to compute these quantities.
%Note that these difference equations are conservative.

We first construct monotonic piecewise 
parabolic (third order) interpolations in one dimension
for each of $p$, $\rho$, and $v$  The pressure is determined from
eq.~(\ref{eq:state}), the equation of state.
%We have dropped the subscript b in this and the following two sections because we will be referring solely to baryonic quantities.  
The interpolation formula for some quantity $q$ is given by:
%
\begin{eqnarray}
q_j(x) & = &  q_{L,j} + \tilde{x}(\Delta q_j + q_{6,j}(1-\tilde{x})), \\
\tilde{x}      & \equiv & {x - x_{j-1/2} \over \Delta x_j}, \qquad
             x_{j-1/2} \leq x \leq x_{j+1/2}. \nonumber
\end{eqnarray}
%
This is eq.~(1.4) of CW84 (in the radial rather than mass coordinate as is appropriate for the direct rulerian approach). 
%The zones edges in the mass coordinates are $m_{j-1/2}$ and $m_{j+1/2}$ for the left and right edges, respectively.  
The quantities $q_{L,j}$, $\Delta q_j$,
and $q_{6,j}$ can be viewed simply as interpolation constants; however,
they also have more intuitive meanings.  For example, $q_{L,j}$ is the
value of $q$ at the left edge of zone j, while $\Delta q_j$ and $q_{6,j}$ are analogous to the slope and first order correction to the slope of $q$ (see \citet{1984JCoPh..54..174C} for a complete discussion):
\begin{equation}
\Delta q_j \equiv q_{R,j} - q_{L,j} \qquad 
q_{6,j}    \equiv 6\left[q_j - 1/2\left(q_{L,j} + q_{R,j}\right)\right].
\end{equation}

We have reduced the problem to finding $q_{L,j}$ and $q_{R,j}$.  While this
is simple in principle, it is complicated somewhat by the requirement that
these values be of sufficient accuracy and that the resulting distribution
be monotonic.  That is, no new maxima or minima are introduced.
The resulting formulae are straightforward but complicated and are not
reproduced here, but see eqs. 1.7 to 1.10 of \citet{1984JCoPh..54..174C}.
We also optionally allow steepening as described in this reference.

In the Lagrangian case, once we have the interpolation constants we compute the characteristic domain for each zone edge.  This is simply the farthest a sound wave could travel in order to reach the interface by the end of a timestep. The characteristic domain encompasses all of the information that can reach the (Lagrangian) zone edge during the current timestep.  This information would then be fed into the Riemann solver to find the flux at each Lagrangian cell boundary.

In the Eulerian approach, this is more complex because the Eulerian zone-boundary does not naturally correspond to one the three characteristic waves and so a more complicated approach is required.  First, the primary quantities ($p$, $\rho$, $v$ and $E$) are averaged over the domains corresponding to the three characteristics $C = u-c$, $u$ or $u+c$ (where $c$ is the sound speed in a cell):
\begin{eqnarray}
      q^{C+}_{j+1/2} & = & f_{j+1/2,L}(\Delta t C^n_j), \\
      q^{C-}_{j+1/2} & = & f_{j+1/2,R}(-\Delta t C^n_{j+1}), \nonumber
\end{eqnarray}
where $q^{C+}$ ($q^{C-}$) refers to the left (right) characteristic $C$.  The functions $f$ are just integrals
over the parabolic interpolation formula:
%
\begin{eqnarray}
f_{j+1/2,L}(y) & = &
       {1 \over y} \int^{x_{j+1/2}}_{x_{j+1/2}-y} q_j(x) dx, \\
f_{j+1/2,R}(y) & = &
       {1 \over y} \int^{x_{j+1/2}+y}_{x_{j+1/2}} q_{j+1}(x) dx. \nonumber
\end{eqnarray}
The linearized gas dynamics equations are then used to compute second-order correct left and right states which take into account the multiple wave families.  This process is described in \citet{1984JCoPh..54..174C} and we use their equations (3.6) and (3.7).

With these effective states, an approximation to the Riemann problem is found
with an iterative approach \citep[see][]{Woodward86}, producing estimates for 
$\overline{p}_{j\pm 1/2}$, $\overline{\rho}_{j\pm1/2}$, and $\overline{v}_{j\pm 1/2}$ that are third order
accurate in space and second order accurate in time.  These are then used to solve the difference eqs. (\ref{eq:mass_diff})--(\ref{eq:energy_diff}) for $\rho^{n+1}$, $v^{n+1}$, and $E^{n+1}$.

We include an optional diffusive flux (and flattening for the parabolic curves) which can improve the solution in some cases.  Our implementation follows that in the appendix of \citet{1984JCoPh..54..174C}.

In addition, as discussed earlier, the three dimensional scheme is achieved by operator splitting and repeating the above procedure in the other two orthogonal directions.  The transverse velocities and any additional passive quantities are naturally and easily added to this system (see eq. 3.6 of \citet{1984JCoPh..54..174C}).

We note that the acceleration required in eq.~(\ref{eq:momentum_diff}) is actually the
acceleration felt by the entire zone and not just at the zone center.
Therefore, it is possible to find the mass weighted average acceleration over the zone
by expanding the density and acceleration distributions and
retaining all terms up to second order in $\Delta x$:
%
\begin{equation}
g_j^{n+1/2} = 
       \frac{1}{2 a^{n+1/2} \delta x_j} \left[ 
             \phi_{j+1}^{n+1/2} 
           - \phi_{j-1}^{n+1/2} 
           + \frac{1}{12} \left(    \phi^{n+1/2}_{j+1} 
                                 - 2\phi^{n+1/2}_j 
                                 + \phi^{n+1/2}_{j-1} \right) 
                                   {\delta d_j \over d_j}
       \right].
       \label{eq:accel}
\end{equation}
Generally we find that the potential is so slowly varying that this change is unnecessary.

% ------------------------------------------------------------

\subsubsection{The dual energy formulation for  very high Mach flows} % 2.4

The system described so far works well for gravitating systems with
reasonable Mach numbers ($<100$) as long as the structures are well resolved.
This section and the next detail changes that are required
to correctly account for situations in which one or both of these
requirements are not met.

Large, hypersonic bulk flows appear to be very common in cosmological
simulations and they present a problem because of the high ratio of
kinetic energy $E_k$ to gas internal energy $e$, which can reach as
high as $10^8$.  Inverted, we see that the internal energy consists of
an extremely small portion of the total energy.  The pressure then,
proportional to $E - E_{k}$, is the small difference between two large
numbers: a disastrous numerical situation.  This is not as large a
problem as it may at first appear because it only occurs when the
pressure is negligibly small.  Therefore, even if we suffer large
errors in the pressure distribution in these regions, the dynamics and
total energy budget of the flow will remain unaffected.  Nevertheless,
if the temperature distribution is required for other reasons
(e.g. radiative processes), a remedy is required.

To overcome this, we also solve the internal energy equation:
\begin{equation}
 \frac{\partial e}{\partial t} 
           + \frac{1}{a} \vecv \cdot \nabla e
%         = - \frac{3(\gamma -1)\dot{a}}{a} e
         = - 3 \frac{\dot{a}}{a} \frac{p}{\rho}
           - \frac{p}{a\rho} \nabla \cdot \vecv
\end{equation}
in comoving co-ordinates.  The structure is similar to the total energy
equation; the second term on the left hand side represents transport, while
the first term on the right is due to expansion of the coordinate
system.  
It is differenced (again, in Eulerian form without the expansion term) as,
%
\begin{equation}
\rho_j^{n+1} e_j^{n+1}  = 
       \rho_j^n e_j^n   +  \Delta t  \left(
            \frac{  \avp{\rho} \avp{v} \avp{e}  - \avm{\rho} \avm{v} \avm{e}} {\Delta x_j} \right)
            - \Delta t \ p_j^{n} \left( \frac{ \avp{v} - \avm{v}} {\Delta x_j} 
                      \right)
                  \label{eq:gasenergy_diff}
\end{equation}
%
Note that because of the structure of this equation, is not in flux-conservative form.  In particular, the pressure is evaluated at the cell-center.  Unfortunately, time-centering of this pressure has proved difficult to do without generating large errors in the internal energy and so we conservatively leave the pressure at the old time in this difference equation.  This leads to spreading of shocks; however, we note that this equation is only used in hypersonic flows.

It is necessary, however, to conserve the total energy so that the conversion
of kinetic to thermal energy is performed properly.   We must therefore,
combine the two formulations without allowing the separately advected
internal energy $e$ to play a role in the gas dynamics.  This is done by
carrying both terms through the simulation and using
the total energy $E$ for hydrodynamic routines and the internal energy
$e$ when the temperature profile is required.  One way to view this procedure
is to treat $e$ as enhanced precision (extra digits) for $E$ that automatically
`floats' to where it is needed.
We only require that they be kept synchronized when the two levels of precision
overlap.

When the pressure is required solely for dynamic purposes, the 
selection criterion 
operates on a cell by cell basis using,
\begin{equation}
p = \cases{ \rho(\gamma - 1)(E - \vecv^2/2),& 
                  $  (E - \vecv^2/2)/E > {\eta}_1 $; \cr
            \rho(\gamma -1)e,&
                  $  (E - \vecv^2/2)/E < {\eta}_1 $. \cr}
\end{equation}

It should be stressed that as long as the parameter ${\eta}_1$ is small enough
the dual energy method {\it will have no dynamical effect}.  We use
${\eta}_1 = 10^{-3}$ which is consistent with the truncation error of the
scheme.  We are now free to select the method by which the internal energy
field variable $e$ is updated so that it will not become contaminated with
errors advected by the total energy formulation but still give the
correct distribution in shocked regions. 
Since we are concerned with the advection of errors, the selection
criterion must look at each cell's local neighbourhood.
In one dimension, this is done with,
\begin{equation}
e = \cases{ (E - \vecv^2/2),& $\rho(E - \vecv^2/2)/
    \max ({\rho}_{j-1}E_{j-1},{\rho}_j E_j,{\rho}_{j+1}E_{j+1}) > {\eta}_2$,\cr
            e,& $\rho(E - \vecv^2/2)/
    \max ({\rho}_{j-1}E_{j-1},{\rho}_j E_j,{\rho}_{j+1}E_{j+1}) < {\eta}_2$.\cr}
    \label{eq:dualselect}
\end{equation}
Thus, ${\eta}_2$ determines when the synchronization (of $e$ with $E$) occurs.
%is a measure of how much error we will allow to be passed
%from neighboring cells.  
Too high a value may mask relatively weak shocks, 
while spurious heating (via contamination) may occur if it is set too low.
After some experimentation,
we have chosen ${\eta}_2 = 0.08$, a somewhat conservative value.
This scheme is optional and is generally only required in large-scale 
cosmological simulations where the gas cools to the expansion of the universe
but large bulk flows develop due to the formation of structure.

We note that others have independently developed a similar but
distinct scheme for dealing with this problem, which is endemic to
methods adopting the total energy equation.  In \citet{TVD93} the 
two variables adopted are total energy and
entropy (rather than total energy and thermal energy), with an
analogous scheme for choosing which variable to employ.

% ------------------------------------------------------------

% GB: I commented the following out because it has been incorporated into the above
%
%PPM is an explicit, higher order-accurate version of
%Godunov's method for ideal gas dynamics with third order-accurate piecewise parabolic
%monotonic interpolation and a nonlinear Riemann solver for shock
%capturing.  It advances the hydrodynamic equations in the following
%steps:
%\begin{enumerate}
% \item Construct monotonic parabolic interpolation of cell average
% data, for each fluid quantity.
% \item Compute interface states by averaging the parabola for over the domain of dependence for
% each interface
% \item Use interface data to solve the Riemann problem.
% \item Difference the interface fluxes to update the cell average quantities.
%\end{enumerate}
%It does an excellent job of capturing strong shocks in at
%most two cells.  Multidimensional schemes are built up by directional
%splitting and produce a method that is formally second order-accurate
%in space and time which explicitly conserves mass, linear momentum, and energy.

%To use the above algorithm for cosmological simulations, the
%conservation laws for fluid mass, momentum and energy 
%density are written in comoving coordinates for a
%Friedman-Robertson-Walker space-time, as described previously 
%in Equations~\ref{enzoconserve} 
%through~\ref{enzoenergy}.  Both the conservation laws and
%the Riemann solver are modified to include gravity, which is solved
%using an adaptive particle-mesh (PM) technique (see Section~\ref{sec.ov.nbody}).  
%The terms due to cosmological expansion, as well as primordial chemistry and
%radiative heating and cooling, are solved in a separate step because they have
%different numerical behavior, and therefore must be treated differently to ensure
%stability.  Note that unlike the \zeus\ hydro scheme, PPM does not need to use 
%artificial viscosity to resolve shocks.

%The system of equations described above works well for systems with
%relatively low Mach numbers, as long as these systems are well resolved.  
%However, cosmology is replete with situations where there are bulk hypersonic
%flows.  In these situations, the ratio of kinetic to thermal energy can be
%very high -- in some situations up to $10^6 - 10^8$.  This implies that the
%thermal energy is an extremely tiny fraction of the kinetic energy, which can 
%cause numerical problems when one is interested in just the thermal energy of the gas, since
% Equation~\ref{enzoenergy} solves for the total energy.  In this system of
%equations, the
% thermal energy $E_{therm}$ is calculated as $E - E_{kin}$, 
%where E is the total specific energy as
%calculated in equation~\ref{enzoenergy} and $E_{kin}$ is the specific kinetic energy,
%$0.5 \vecv_b^2$.  
%In hypersonic flows $E$ and $E_{kin}$ are nearly the same, and any 
%number calculated as 
%the difference of these is going to be strongly affected by numerical error.  To 
%avoid this problem, \enzo\ also solves the internal energy equation in
%comoving coordinates:

%\begin{equation}
%\frac{\partial e}{\partial t} + \frac{1}{a} \vecv_b \cdot \nabla e = 
%- \frac{3(\gamma-1)\dot{a}}{a}e - \frac{p}{a \rho} \nabla \cdot \vecv_b
%\label{enzointenergy}
%\end{equation}

%In this equation $e$ is the internal energy (defined as $P/\rho (\gamma-1)$ and the other terms are as described
%previously.  The code still conserves total energy ($E$) as well.  In order
%to maintain consistency, both equations are solved at all times in all cells,
%with the equation for the total energy (eqtn.~\ref{enzoenergy}) being used
%for hydrodynamics routines and the internal energy $e$ being used when 
%temperature is required.  When pressure is required for dynamic purposes,
%the total energy is used if the ratio of thermal energy to total energy
%is less than some threshold value $\eta$, and the internal energy is used
%for values of the ratio larger than $\eta$.  A typical value of this parameter
%is $10^{-3}$.  This \emph{dual energy formulation} ensures that the
%method produces the correct entropy jump at strong shocks and also
%yields accurate pressures and temperatures in cosmological hypersonic
%flows.

### 2012-06-29 numerical-starformation.tex
first commit of paper.

--- 2012-06-29 numerical-overview.tex
@@
-\section{Overview of Numerical Methods}\label{sec.overview}
+\subsection{Star formation and feedback}
+\label{sec.ov.star}
 
-%---------------------------------------------------------------------------%
-\subsection{N-body dynamics}\label{sec.ov.nbody}
-fft + multigrid
-
-There are multiple methods to compute the gravitational potential 
-(which is an elliptic equation in the Newtonian limit) in a structured 
-AMR framework.  One way would be to model the dark matter (or other
-collisionless particle-like objects, such as stars) as a second fluid 
-in addition to the baryonic fluid and solve the collisionless
-Boltzmann equation, which follows the evolution of the fluid density in
-six-dimensional phase space.  However, this is computationally prohibitive 
-owing to the large dimensionality of the problem, making this approach 
-unappealing for the cosmological AMR code.
-
-Instead, \enzo\ uses a particle-mesh N-body method to calculate 
-the dynamics of collisionless systems \citep{Hockney88}.  This method 
-follows trajectories 
-of a representative sample of individual particles and is much more 
-efficient than a direct solution of the Boltzmann equation in most 
-astrophysical situations. 
-The gravitational potential is computed by solving the elliptic 
-Poisson's equation (Eqtn.~\ref{enzopoisson}).
-
-These equations are finite-differenced and for simplicity are solved with the
-same timestep as the hydrodynamic equations, as discussed in 
-Section~\ref{sec.ov.amr}.
-The dark matter particles are distributed onto the grids
-using the cloud-in-cell (CIC) interpolation technique to form a spatially
-discretized density field (analogous to the baryon densities used to calculate
-the equations of hydrodynamics).  After sampling dark matter density onto the
-grid and adding baryon density if it exists (to get the total matter density
-in each cell), the gravitational potential is calculated on the periodic root
-grid using a fast Fourier transform.  In order to calculate more accurate
-potentials on subgrids, \enzo\ re-samples the dark matter density onto
-individual subgrids using the same CIC method as on the root grid, but at
-higher spatial resolution (and again adding baryon densities if applicable).
-Boundary conditions are then interpolated from the potential values on the
-parent grid (with adjacent grid patches on a given level communicating to
-ensure that their boundary values are the same), and then a multigrid
-relaxation technique is used to calculate the gravitational potential at every
-point within a subgrid.  Forces are computed on the mesh by
-finite-differencing potential values and are then interpolated to the particle
-positions, where they are used to update the particle's position and velocity
-information.  Potentials on child grids are computed recursively and particle
-positions are updated using the same timestep as in the hydrodynamic
-equations.  Particles are stored in the most highly refined grid patch at the
-point in space where they exist, and particles which move out of a subgrid
-patch are sent to the grid patch covering the adjacent volume with the finest
-spatial resolution, which may be of the same spatial resolution, coarser, or finer
-than the grid patch that the particles are moved from.  This takes place in a
-communication process at the end of each timestep on a level.
-
-At this point it is useful to emphasize that the effective force
-resolution of an adaptive particle-mesh calculation is approximately twice
-as coarse as the grid spacing at a given level of resolution.  The 
-potential is solved in each grid cell;
-however, the quantity of interest, namely the acceleration, is the gradient
-of the potential, and hence two potential values are required to calculate
-this.  In addition, it should be noted that the adaptive particle-mesh
-technique described here is very memory intensive: in order to ensure reasonably 
-accurate force resolution at grid edges the multigrid relaxation
-method used in the code requires a layer of ghost zones which is very deep --
-typically 6 cells in every direction around the edge of a grid patch.  This greatly
-adds to the memory requirements of the simulation, particularly because subgrids
-are typically small (on the order of $12^3 - 16^3$ real cells for a standard 
-cosmological calculation) and ghost zones can dominate the memory and computational 
-requirements of the code.
-
-%---------------------------------------------------------------------------%
-\subsection{Hydrodynamics}\label{sec.ov.hydro}
-
-\subsubsection{The PPM hydro method}\label{sec.ov.hydro.ppm}
-
-\dcc{Removed ``things to do'' list.  Mention ppm-lr and ppm-de, and
-  that ppm-de is preferred.  Briefly describe ppm-de itself.}
-
-The primary hydrodynamic method used in \enzo\ is based on the
-piecewise parabolic method (PPM) of
-\citet{1984JCoPh..54..174C}, which has been
-modified for the study of cosmological fluid flows.  
-\citet{1984JCoPh..54..174C} describe two variants on this method: a
-Lagrangian Remap variant and a Direct Eulerian variant.  Both have
-been adapted for \enzo, and the Lagrangian Remap 
-variant is described in \citet{1995CoPhC..89..149B}.  Because of its
-conservative formulation, however, the Direct Eulerian variant is
-preferred for \enzo\ AMR simulations.  We will describe the Direct
-Eulerian (PPM-DE) variant and its additions briefly here, and in
-further depth in Appendix \ref{app.ppm}
-
-PPM is an explicit, higher order-accurate version of
-Godunov's method for ideal gas dynamics with third order-accurate piecewise parabolic
-monotonic interpolation and a nonlinear Riemann solver for shock
-capturing.  It advances the hydrodynamic equations in the following
-steps:
-\begin{enumerate}
- \item Construct monotonic parabolic interpolation of cell average
- data, for each fluid quantity.
- \item Compute interface states by averaging the parabola for over the domain of dependence for
- each interface
- \item Use interface data to solve the Riemann problem.
- \item Difference the interface fluxes to update the cell average quantities.
-\end{enumerate}
-It does an excellent job of capturing strong shocks in at
-most two cells.  Multidimensional schemes are built up by directional
-splitting and produce a method that is formally second order-accurate
-in space and time which explicitly conserves mass, linear momentum, and energy.
-
-To use the above algorithm for cosmological simulations, the
-conservation laws for fluid mass, momentum and energy 
-density are written in comoving coordinates for a
-Friedman-Robertson-Walker space-time, as described previously 
-in Equations~\ref{enzoconserve} 
-through~\ref{enzoenergy}.  Both the conservation laws and
-the Riemann solver are modified to include gravity, which is solved
-using an adaptive particle-mesh (PM) technique (see Section~\ref{sec.ov.nbody}).  
-The terms due to cosmological expansion, as well as primordial chemistry and
-radiative heating and cooling, are solved in a separate step because they have
-different numerical behavior, and therefore must be treated differently to ensure
-stability.  Note that unlike the \zeus\ hydro scheme, PPM does not need to use 
-artificial viscosity to resolve shocks.
-
-The system of equations described above works well for systems with
-relatively low Mach numbers, as long as these systems are well resolved.  
-However, cosmology is replete with situations where there are bulk hypersonic
-flows.  In these situations, the ratio of kinetic to thermal energy can be
-very high -- in some situations up to $10^6 - 10^8$.  This implies that the
-thermal energy is an extremely tiny fraction of the kinetic energy, which can 
-cause numerical problems when one is interested in just the thermal energy of the gas, since
- Equation~\ref{enzoenergy} solves for the total energy.  In this system of
-equations, the
- thermal energy $E_{therm}$ is calculated as $E - E_{kin}$, 
-where E is the total specific energy as
-calculated in equation~\ref{enzoenergy} and $E_{kin}$ is the specific kinetic energy,
-$0.5 \vec{v}_b^2$.  
-In hypersonic flows $E$ and $E_{kin}$ are nearly the same, and any 
-number calculated as 
-the difference of these is going to be strongly affected by numerical error.  To 
-avoid this problem, \enzo\ also solves the internal energy equation in
-comoving coordinates:
-
-\begin{equation}
-\frac{\partial e}{\partial t} + \frac{1}{a} \vec{v}_b \cdot \nabla e = 
-- \frac{3(\gamma-1)\dot{a}}{a}e - \frac{p}{a \rho} \nabla \cdot \vec{v}_b
-\label{enzointenergy}
-\end{equation}
-
-In this equation $e$ is the internal energy (defined as $P/\rho (\gamma-1)$ and the other terms are as described
-previously.  The code still conserves total energy ($E$) as well.  In order
-to maintain consistency, both equations are solved at all times in all cells,
-with the equation for the total energy (eqtn.~\ref{enzoenergy}) being used
-for hydrodynamics routines and the internal energy $e$ being used when 
-temperature is required.  When pressure is required for dynamic purposes,
-the total energy is used if the ratio of thermal energy to total energy
-is less than some threshold value $\eta$, and the internal energy is used
-for values of the ratio larger than $\eta$.  A typical value of this parameter
-is $10^{-3}$.  This \emph{dual energy formulation} ensures that the
-method produces the correct entropy jump at strong shocks and also
-yields accurate pressures and temperatures in cosmological hypersonic
-flows.
-
-\subsubsection{The Zeus hydro method}\label{sec.ov.hydro.zeus}
-
-As a check on PPM, \enzo\ also includes an implementation of the
-finite-difference hydrodynamic algorithm employed in the compressible 
-magnetohydrodynamics code `\zeus'
-\citep{Stone92a, Stone92b}.  Fluid transport is solved on a Cartesian
-grid using the upwind, monotonic advection scheme of \citet{1977JCoPh..23..276V} 
-within a multistep 
-(operator split) solution procedure which is fully explicit in time.  
-This method is formally second order-accurate in space but 
-first order-accurate in time.  It is important to note that 
-this method conserves internal energy
-rather than total energy, so the ``dual energy
-formulation'' discussed in Section~\ref{sec.ov.hydro.ppm} is unnecessary.
-
-Operator split methods break
-the solution of the hydrodynamic equations into parts, with each part
-representing a single term in the equations.  Each part is evaluated
-successively using the results preceding it.  The individual parts of
-the solution are grouped into two steps, called the \emph{source} and
-\emph{transport} steps.
-
-The \zeus\ method uses a von Neumann-Richtmeyer artificial viscosity 
-to smooth shock discontinuities that may
-appear in fluid flows and can cause a break-down of finite-difference
-equations.  The artificial viscosity term is added in the source terms
-as
-\begin{eqnarray}
-\rho \frac{\partial\textbf{v}}{\partial t} &=& - \nabla p - \rho \nabla \phi 
-- \nabla \cdot \textbf{Q} \\
-\frac{\partial e}{\partial t} &=& -p \nabla \cdot \textbf{v} - \textbf{Q} : \nabla \textbf{v}, 
-\end{eqnarray}
-where \textbf{v} is the baryon velocity, $\rho$ is the mass density,
-$p$ is the pressure, $e$ is the internal energy density of the gas and \textbf{Q}
-is the artificial viscosity stress tensor, such that:
-
-\begin{eqnarray}
-Q_{ii} = \left\{ \begin{array}{ll}
-Q_{\rm AV} \rho_b  (a \Delta v_{i} + \dot{a} \Delta x_i)^2, 
-& \textrm{for $a \Delta v_{i} + \dot{a} \Delta x_i < 0$}  \\
-0 & \textrm{otherwise}\\
-\end{array} \right. 
-\end{eqnarray}
-and
-\begin{equation}
-Q_{ij} = 0  \;\;\;{\rm for}\;\; i \ne j .
-\end{equation}
-
-$\Delta x_i$ and $\Delta v_{i}$ refer to the comoving 
-width of the grid cell
-along the $i$-th axis and the corresponding difference in gas
-peculiar velocities across the grid cell, respectively, and $a$ is the
-cosmological scale factor.  $Q_{\rm AV}$ is a constant with a typical
-value of 2. We refer the interested reader to \citet{1994ApJ...429..434A} for more details.
-
-The limitation of a technique that uses an artificial viscosity is that,
-while the correct Rankine-Hugoniot jump conditions are achieved,
-shocks are broadened over 6-8 mesh cells and are thus not treated as true
-discontinuities. This may cause unphysical pre-heating of gas upstream
-of the shock wave, as discussed in \citet{1994ApJ...429..434A}.
-
-%---------------------------------------------------------------------------%
-\subsection{Chemistry, radiation and cooling}\label{sec.ov.chem}
-
-\red{
-As this stands it is the methods that I (brian) use, so I don't really emphasize
-either the tabular cooling functions or the various radiation background options.
-Add more discussion of this, point towards appendix.
-}
-
-Though the equations of hydrodynamics described above are a closed system,
-they are still missing a crucial piece of physics: radiative heating and cooling.
-Radiative cooling is extremely important in many astrophysical situations, as is 
-the heating of gas from some sort of radiation background.  \enzo\ has a very simple
-Sutherland and Dopita equilibrium cooling function \citep{1993ApJS...88..253S} 
-implemented, which uses a
-cooling table assuming a fixed metallicity of $Z = 0.3 Z_\odot$, and also a
-nonequilibrium heating/cooling model that assumes gas of primordial composition exposed to a
-uniform metagalactic ultraviolet (UV) background that varies with time \citep{1996ApJS..105...19K}.
-
-The simulations discussed in this thesis almost exclusively use the nonequilibrium
-routines, described in great detail by Abel et al. and Anninos et al.  
-\citep{abel97,anninos97} and summarized in Appendix~\ref{app:chemistry}. These 
-routines follow the non-equilibrium chemistry of a 
-gas of primordial composition with 9 total species:  
-$H$, $H^+$, $He$, $He^+$, $He^{++}$, $H^-$, $H_2^+$, $H_2$, and $e^-$.  The code also calculates 
-radiative heating and cooling, following atomic line excitation, recombination,
-collisional excitation, free-free transitions, molecular line excitations, and Compton
-scattering of the cosmic microwave background, as well as any of
-approximately a dozen different models for a metagalactic UV background that heat
-the gas via photoionization and photodissociation. 
-The multispecies rate equations are solved out of
-equilibrium to properly model situations where, e.g., the cooling rate of the gas
-is much shorter than the hydrogen recombination time.  The effect of this nonequilibrium
-cooling is to leave behind a much larger fraction of residual free electrons than one would
-expect if the assumption of equilibrium were being made.  The practical effect of this
-is that more $H^-$ is formed, which subsequently produces hydrogen molecules.  If large
-amounts of $H_2$ is formed it can greatly increase the cooling rate of primordial gas
-at relatively low temperatures ($T \leq 10^4$ K).  This can efficiently cool the gas 
-to approximately $200$ K, which significantly reduces the Jeans mass of the gas.  Correct
-modeling of the formation of molecular hydrogen is crucial to the study of star formation
-in a primordial gas.
-
-A total of 9 kinetic equations are solved, including 29 kinetic and radiative 
-processes, for the 9 species mentioned above.  See Table~\ref{table.collisional} 
-for the collisional processes and Table~\ref{table.radiative} for the 
-radiative processes solved.
-
-%---------------- table of collisional processes
-
-\begin{table}
-\begin{center}
-{\bfseries Collisional Processes}\\[1ex]
-\begin{tabular}{llllllll}
-(1) & H & + & e$^-$ & $\rightarrow$ & H$^+$ &+& 2e$^-$ \\
-(2) & H$^+$ &+ &e$^-$ & $\rightarrow$ & H &+ &$\gamma$ \\
-(3) & He &+& e$^-$ & $\rightarrow$ & He$^+$ &+& 2e$^-$  \\
-(4) & He$^+$ &+& e$^-$ & $\rightarrow$ & He &+ &$\gamma$  \\
-(5) & He$^{+}$ &+& e$^-$ & $\rightarrow$ & He$^{++}$ &+& 2$e^-$  \\
-(6) & He$^{++}$ &+& e$^-$ & $\rightarrow$ & He$^+$ &+& $\gamma$ \\
-(7) & H &+& e$^-$ &$\rightarrow$& H$^-$ &+& $\gamma$  \\
-(8) & H$^-$ &+& H &$\rightarrow$ & H$_2$ & +& e$^-$ \\
-(9) & H &+ &H$^+$ &$\rightarrow$ &H$_2^+$ &+ &$\gamma$ \\
-(10) & H$_2^+$ &+ &H &$\rightarrow$ &$H_2$ &+ &$H^+$ \\
-(11) & H$_2$ &+ &H$^+$ &$\rightarrow$ &H$_2^+$ & +& H \\
-(12) & H$_2$ &+ &e$^-$ & $\rightarrow$ & 2H & + & e$^-$  \\
-(13) & H$_2$ & + & H & $\rightarrow$ & 3H &   &      \\
-(14) & H$^-$ & + & e$^-$ & $\rightarrow$ & H & + & 2e$^-$ \\
-(15) & H$^-$ & + & H & $\rightarrow$ & 2H & + & e$^-$ \\ 
-(16) & H$^-$ & + & H$^+$ & $\rightarrow$ & 2H & & \\
-(17) & H$^-$ & + & H$^+$ & $\rightarrow$ & H$_2^+$ & + & e$^-$ \\
-(18) & H$_2^+$ & + & e$^-$ & $\rightarrow$ & 2H & & \\
-(19) & H$_2^+$ & + & H$^-$ & $\rightarrow$ & H$_2$ & + & H  \\
-(20) & 3H & & & $\rightarrow$ & H$_2$ & + & H
-\end{tabular}
-\caption[]{Collisional processes solved in the Enzo nonequilibrium
-primordial chemistry routines.}
-\label{table.collisional}
-\end{center}
-\end{table}
-
-\begin{table}
-\begin{center}
-{\bfseries Radiative Processes}\\[1ex]
-\begin{tabular}{llllllll}
-(21) & H & + & $\gamma$ & $\rightarrow$ & H$^+$ & + & e$^-$ \\
-(22) & He & + & $\gamma$ & $\rightarrow$ & He$^+$ & + & e$^-$ \\
-(23) & He$^+$ & + & $\gamma$ & $\rightarrow$ & He$^{++}$ & + & e$^-$ \\
-(24) & H$^-$ & + & $\gamma$ & $\rightarrow$ & H & + & e$^-$ \\
-(25) & H$_2$ & + & $\gamma$ & $\rightarrow$ & H$_2^+$ & + & e$^-$ \\
-(26) & H$_2^+$ & + & $\gamma$ & $\rightarrow$ & H & + & H$^+$ \\
-(27) & H$_2^+$ & + & $\gamma$ & $\rightarrow$ & 2H$^+$ & + & e$^-$ \\
-(28) & H$_2$ & + & $\gamma$ & $\rightarrow$ & H$_2^*$ & $\rightarrow$ & 2H \\
-(29) & H$_2$ & + & $\gamma$ & $\rightarrow$ & 2H &  & 
-\end{tabular}
-\caption[]{Radiative processes solved in the Enzo nonequilibrium
-primordial chemistry routines.}
-\label{table.radiative}
-\end{center}
-\end{table}
-
-The chemical reaction equation network is technically challenging to solve due to 
-the huge range of reaction timescales involved -- the characteristic creation
-and destruction timescales of the various species and reactions can differ by 
-many orders of magnitude.  As a result, the set of rate equations is extremely 
-stiff, and an explicit scheme for integration of the rate equations can be 
-exceptionally costly if small enough timesteps are taken to keep the network
-stable.  This makes an implicit scheme much more preferable for such a set of 
-equations.  However, an implicit scheme typically require an iterative 
-procedure to converge, and for large networks (such as this one) an implicit
-method can be very time-consuming, making it undesirable for a large, three-dimensional
-simulation.
-
-\enzo\ solves the rate equations using a method based on a backwards differencing 
-formula (BDF) in order to provide a stable and accurate solution.  This technique is
-optimized by taking the chemical intermediaries $H^-$ and $H_2^+$, which have 
-large rate coefficients and low concentrations, and grouping them into a separate
-category of equations.  Due to their fast reactions these species are very sensitive
-to small changes in the more abundant species.  However, due to their low overall
-concentrations, they do not significantly influence the abundance of species with
-higher concentrations.  Therefore, reactions involving these two species can be
-decoupled from the rest of the network and treated independently.  In fact, $H^-$ 
-and $H_2^+$ are treated as being in equilibrium at all times, independent of 
-the other species and the hydrodynamic variables.  This allows a large speedup
-in solution as the BDF scheme is then applied only to the slower 7-species network
-on timescales closer to those required by the hydrodynamics of the simulation.
-Even so, the accuracy and stability of the scheme is maintained by subcycling the 
-rate solver over a single hydrodynamic timestep.  These subcycle timesteps are 
-determined so that the maximum fractional change in the electron concentration is
-limited to no more than $10\%$ per timestep.
-
-It is important to note the regime in which this model is valid.  According to Abel et al. and
-Anninos et al. \citep{abel97,anninos97}, the reaction network is valid for temperatures
-between $10^0 - 10^8$ K.  The original model discussed in these two references is only
-valid up to $n_H \sim 10^4$~cm$^{-3}$.  However, addition of the 3-body $H_2$ formation
-process (equation 20 in Table~\ref{table.collisional}) allows 
-correct modeling of the chemistry of the gas up
-until the point where collisionally-induced emission from molecular hydrogen becomes an important
-cooling process, which occurs at $n_H \sim 10^{14}$~cm$^{-3}$.  A further concern is that
-the optically thin approximation for radiative cooling breaks down, which
-occurs before $n_H \sim 10^{16} - 10^{17}$~cm$^{-3}$.  Beyond this point, 
-modifications the cooling function that take into account the non-negligible
-opacity in the gas must be made, as discussed by \citet{2004MNRAS.348.1019R}. 
-Even with these modifications, a completely correct description of the cooling of
-this gas will require some form of radiation transport, which will greatly 
-increase the cost of the simulations.
-
-Several processes are neglected.  The deuterium atom and its processes are 
-completely ignored, which may have some effect.  Recent work shows that HD is
-a more effective coolant than previously thought~\citep{2005MNRAS.361..850L}.  However,
-the fractional abundance of HD is so low that under circumstances relevant to
-the formation of a Population III star in an un-ionized region it should be
-sub-dominant.  However, the enhanced electron fraction in fossil 
-HII regions (as discussed later in this thesis) could result in the HD molecule
-becoming a dominant cooling mechanism at relatively low ($\sim$ few hundred K)
-temperatures, and could potentially cool the gas down to below $100$~K, which can
-enhance fragmentation and could have important consequences for the IMF of 
-primordial stars forming in a relic HII region~\citep{2005MNRAS.364.1378N}.
-
-Aside from deuterium, the chemical reactions involving lithium 
-are also neglected.  According to \citet{1998A&A...335..403G}, these are 
-not important for the density and temperature regimes explored 
-by the simulations discussed in this thesis.  However, at higher densities it 
-is possible that there are regimes where lithium can be an important coolant.
-
-%---------------------------------------------------------------------------%
-\subsection{Star formation and feedback}\label{sec.ov.star}
-
-\red{By the time this paper is submitted my (BWO's) paper on the star formation
-and feedback algorithms, and the various parameters contained therein, should be
-at least written and submitted as well.  Point to this, and maybe include a couple
-interesting plots (perhaps in an appendix?).}
+%\red{By the time this paper is submitted my (BWO's) paper on the star formation
+%and feedback algorithms, and the various parameters contained therein, should be
+%at least written and submitted as well.  Point to this, and maybe include a couple
+%interesting plots (perhaps in an appendix?).}
+%GB: can we really point to that paper?
 
@@
 stars in the context of simulations of galaxy formation and evolution due
-to the immense computational cost.  Therefore, 
-we adopt parametric methods which attempt
-to model star forming regions.  One model is based on Kravtsov's method
- \citep{2003ApJ...590L...1K}, and the other is based on the method of 
-\citet{1992ApJ...399L.113C}, which has been modified for use in an AMR code.  The basic
+to the immense computational cost.  Therefore, we adopt a parametric method which attempts
+to model star forming regions in a plausible but heuristic way.  
+%One model is based on Kravtsov's method  \citep{2003ApJ...590L...1K}, and the other is 
+Our model is based on the method of  \citet{CO1992}, which 
+has been modified for use in an AMR code.  The basic
 ideas behind the methods are straightforward and observationally motivated.
@@
 \citep{sh03a,sh03b,hs03}. This 
-method assumes a
-multiphase IGM and has been shown to accurately reproduce the cosmic star
+method assumes a multiphase IGM and has been shown to accurately reproduce the cosmic star
 formation rate, and will be implemented into \enzo\ in the near future.
 
-In the following sections we will detail both the Kravtsov and Cen \& Ostriker
-star formation methods separately.  Tests of the star formation and
+%In the following sections we will detail both the Kravtsov and Cen \& Ostriker star formation methods separately.  
+Tests of the star formation and
 feedback algorithms will not be shown, since this work is currently in
@@
 \citet{2005ApJS..160....1O}).
+%
+%\subsubsection{The Kravtsov star formation and feedback algorithm}
 
-\subsubsection{The Kravtsov star formation and feedback algorithm}
+% GB: Is this going to be in the code????
+% GB: I assume not, so I am commenting it out.  Text to be used in another paper?
 
-Kravtsov's method of star formation is designed to reproduce
-the global Schmidt law of star 
-formation~\citep{2003ApJ...590L...1K,1959ApJ...129..243S}.
-This algorithm is deliberately minimal, and is explicitly geared
-towards modeling star formation in a phenomenological way on kiloparsec
-scales.  Stars are assume to form with a characteristic gas timescale 
-$\tau_*$ such that $\dot{\rho}_* = \rho_{gas}/\tau_*$ where $\rho_{gas}$ 
-and $\rho_*$ are the baryon gas and stellar densities, respectively.  
-This ``constant efficiency'' model on the scale of star formation 
-regions is well motivated observationally \citep{1996AJ....112.1903Y,2002ApJ...569..157W}.  
-Star formation is only allowed to take place in very dense regions
-with $\rho_{gas} \geq \rho_{SF}$, where $\rho_{SF}$ is a constant
-proper density threshold above which star formation is allowed to
-occur.  No other criteria are imposed.  Kravtsov's typical choices
-for $\tau_*$ and $\rho_{SF}$ are $\tau_* = 4$ Gyr and $\rho_{SF} =
-1.64~$M$_\odot$~pc$^{-3}$~$(n_H \sim 50$~cm$^{-3})$.  The adopted timescale
-is derived from the observationally-determined normalization of the
-Schmidt law, and the density threshold is determined by observations
-of star forming regions on $\sim 100$ pc scales.  Note that the density
-threshold is in proper, not comoving, units.
+%Kravtsov's method of star formation is designed to reproduce
+%the global Schmidt law of star 
+%formation~\citep{2003ApJ...590L...1K,1959ApJ...129..243S}.
+%This algorithm is deliberately minimal, and is explicitly geared
+%towards modeling star formation in a phenomenological way on kiloparsec
+%scales.  Stars are assume to form with a characteristic gas timescale 
+%$\tau_*$ such that $\dot{\rho}_* = \rho_{gas}/\tau_*$ where $\rho_{gas}$ 
+%and $\rho_*$ are the baryon gas and stellar densities, respectively.  
+%This ``constant efficiency'' model on the scale of star formation 
+%regions is well motivated observationally \citep{1996AJ....112.1903Y,2002ApJ...569..157W}.  
+%Star formation is only allowed to take place in very dense regions
+%with $\rho_{gas} \geq \rho_{SF}$, where $\rho_{SF}$ is a constant
+%proper density threshold above which star formation is allowed to
+%occur.  No other criteria are imposed.  Kravtsov's typical choices
+%for $\tau_*$ and $\rho_{SF}$ are $\tau_* = 4$ Gyr and $\rho_{SF} =
+%1.64~$M$_\odot$~pc$^{-3}$~$(n_H \sim 50$~cm$^{-3})$.  The adopted timescale
+%is derived from the observationally-determined normalization of the
+%Schmidt law, and the density threshold is determined by observations
+%of star forming regions on $\sim 100$ pc scales.  Note that the density
+%threshold is in proper, not comoving, units.
 
-Algorithmically, the star formation events in Kravtsov's code
-are assumed to occur once every global time step 
-$\Delta t_0 \leq 10^7$.  In cells where star formation is
-determined to occur (i.e. $\rho_{gas} \geq \rho_{SF}$) a
-collisionless ``star particle'' is assumed to form, with a 
-mass $m_* = \dot{\rho}_* V_{cell} \Delta t_0$, where
-$\dot{\rho}_*$ is described above and $V_{cell}$ is the 
-volume of the mesh cell.  These star particles are
-formed instantaneously, with all $m_*$ of gas being
-taken out of the cell and immediately deposited into the
-star particle.  This particle is then given the velocity
-of the gas in the cell, and thereafter treated as a 
-collisionless particle.  The \enzo\ implementation
-of this algorithm is similar, except that instead of 
-forming stars only at the root grid time step, we allow 
-stars to form at the time step of the highest level of
-resolution at any particular point in space.  As can be seen
-from the equation for $m_*$ above, this can result in very 
-small stellar masses.  To avoid memory and processor time
-issues related to having very large numbers of star particles
-we impose a threshold mass $M_{*,min}$ such that a star
-particle only forms if $m_* \geq M_{*,min}$.  An
-appropriate choice of $M_{*,min}$ does not significantly 
-change the star overall star formation history of a 
-simulation, though it may delay the onset of star formation 
-in a given cell relative to a simulation without a particle mass
-threshold.
+%Algorithmically, the star formation events in Kravtsov's code
+%are assumed to occur once every global time step 
+%$\Delta t_0 \leq 10^7$.  In cells where star formation is
+%determined to occur (i.e. $\rho_{gas} \geq \rho_{SF}$) a
+%collisionless ``star particle'' is assumed to form, with a 
+%mass $m_* = \dot{\rho}_* V_{cell} \Delta t_0$, where
+%$\dot{\rho}_*$ is described above and $V_{cell}$ is the 
+%volume of the mesh cell.  These star particles are
+%formed instantaneously, with all $m_*$ of gas being
+%taken out of the cell and immediately deposited into the
+%star particle.  This particle is then given the velocity
+%of the gas in the cell, and thereafter treated as a 
+%collisionless particle.  The \enzo\ implementation
+%of this algorithm is similar, except that instead of 
+%forming stars only at the root grid time step, we allow 
+%stars to form at the time step of the highest level of
+%resolution at any particular point in space.  As can be seen
+%from the equation for $m_*$ above, this can result in very 
+%small stellar masses.  To avoid memory and processor time
+%issues related to having very large numbers of star particles
+%we impose a threshold mass $M_{*,min}$ such that a star
+%particle only forms if $m_* \geq M_{*,min}$.  An
+%appropriate choice of $M_{*,min}$ does not significantly 
+%change the star overall star formation history of a 
+%simulation, though it may delay the onset of star formation 
+%in a given cell relative to a simulation without a particle mass
+%threshold.
 
-Each ``star particle'' is assumed to represent an ensemble
-of stars and is treated as a single-age stellar population.
-Kravtsov assumes that the IMF is described by a
-Miller \& Scalo functional form with stellar masses between
-$0.1$ and $100~M_\odot$ \citep{1979ApJS...41..513M}.  All stars in this
-IMF with $M_* > 8~M_\odot$ deposit $2 \times 10^{51}$ ergs of
-thermal energy and a mass $f_z M_*$ of metals into the
-cell in which they form without delay, with
-$f_z \equiv min(0.2, 0.01~M_*-0.06)$ (i.e. instantaneous
-deposition of metals).  The definition of $f_z$ is a rough
-approximation of the results of 
-\citet{1995ApJS..101..181W}.
+%Each ``star particle'' is assumed to represent an ensemble
+%of stars and is treated as a single-age stellar population.
+%Kravtsov assumes that the IMF is described by a
+%Miller \& Scalo functional form with stellar masses between
+%$0.1$ and $100~M_\odot$ \citep{1979ApJS...41..513M}.  All stars in this
+%IMF with $M_* > 8~M_\odot$ deposit $2 \times 10^{51}$ ergs of
+%thermal energy and a mass $f_z M_*$ of metals into the
+%cell in which they form without delay, with
+%$f_z \equiv min(0.2, 0.01~M_*-0.06)$ (i.e. instantaneous
+%deposition of metals).  The definition of $f_z$ is a rough
+%approximation of the results of 
+%\citet{1995ApJS..101..181W}.
 
-Kravtsov reports that simulations with this algorithm
-reliably reproduce the star formation rate-gas surface
-density relation of the Schmidt law, and are not particularly
-sensitive to numerical parameters \citep{2003ApJ...590L...1K}.
-He also notes that this is surprisingly insensitive to
-the presence or absence of feedback and details of the cooling
-and heating properties of the gas, which suggests that the
-global star formation rate is determined by gravitationally-
-driven supersonic turbulence (on large scales)
-rather than stellar feedback or thermal instabilities on small 
-scales.
+%Kravtsov reports that simulations with this algorithm
+%reliably reproduce the star formation rate-gas surface
+%density relation of the Schmidt law, and are not particularly
+%sensitive to numerical parameters \citep{2003ApJ...590L...1K}.
+%He also notes that this is surprisingly insensitive to
+%the presence or absence of feedback and details of the cooling
+%and heating properties of the gas, which suggests that the
+%global star formation rate is determined by gravitationally-
+%driven supersonic turbulence (on large scales)
+%rather than stellar feedback or thermal instabilities on small 
+%scales.
 
-\subsubsection{The Cen \& Ostriker star formation algorithm}
+%\subsubsection{The Cen \& Ostriker star formation algorithm}
 
@@
 \begin{eqnarray}  
-\Delta m_{sf} =  m_{*} \left[ \right. \left(1 + \frac{t-t_{form}}{t_{dyn}} 
-                    \right) \nonumber \\ 
-		   \times exp \left( \frac{-(t-t_{form})}{t_{dyn}} \right) 
-                    \nonumber \\ 
-	       - \left( 1 + \frac{t+ \Delta t-t_{form}}{t_{dyn}} \right) \nonumber \\ 
-                    exp \left( \frac{-(t+\Delta t-t_{form})}{t_{dyn}} \right) 
-                   \left. \right]
+\Delta m_{sf} =  m_{*} \left[ \left(1 + \frac{t-t_{form}}{t_{dyn}}  \right)
+%\nonumber \times \\ % uncomment for 2-column
+		   \exp{\left( \frac{-(t-t_{form})}{t_{dyn}} \right) }
+%                    \nonumber \\ 
+	       - \left( 1 + \frac{t+ \Delta t-t_{form}}{t_{dyn}} \right) 
+%	       \nonumber \\ 
+                    \exp{ \left( \frac{-(t+\Delta t-t_{form})}{t_{dyn}} \right)  }
+                   \right]
 \end{eqnarray}
@@
 \int_{t}^{t+Dt} \frac{dM}{dt} dt = \int_{t}^{t+Dt} m_{*} 
-\frac{dt}{t_{tyn}} \left( \frac{t-t_{form}}{t_{dyn}} \right) \nonumber \\ 
-\times exp \left( - ~ \frac{t-t_{form}}{t_{dyn}} \right) 
+\frac{dt}{t_{tyn}} \left( \frac{t-t_{form}}{t_{dyn}} \right) 
+%\nonumber \times \\ 
+\exp{ \left( - ~ \frac{t-t_{form}}{t_{dyn}} \right) }
 \end{eqnarray}
@@
 the results may be degenerate for some combinations of these parameters.
-
-%---------------------------------------------------------------------------%
-\subsection{Adaptive mesh refinement}\label{sec.ov.amr}
-\dcc{removed most of the notes.  Added intro blurb.  moved parallelism
-  stuff. Re-worded the bit on ghost zones.  Summary of 'grid as
-  independent update + syncing with neighbors and parents,' W-cycle. Added
-  sentence on subgrid creation.  Made timestepping its own subsection.}
-\red{Talk about reasonable refinement factors - which 
-ones are usable, etc?}
-
-The primary purpose of the \enzo\ code is its Adaptive Mesh Refinement
-capability, which allows it to reach extremely large dynamical ranges
-with limited computational resources, opening doors previously closed
-by finite memory and computational time. Unlike moving mesh methods
-\citep{1995ApJS..100..269P,1995ApJS...97..231G} or  
-methods that subdivide 
-individual cells \citep{Adjerid}, Berger \& Collela's AMR (also referred 
-to as \emph{structured} AMR) utilizes an adaptive hierarchy of grid 
-patches at varying levels of resolution.  Each rectangular grid patch 
-(referred to as a ``grid'') covers some region of space in its 
-\emph{parent grid} which requires higher resolution, and can itself 
-become the parent grid to an even more highly resolved \emph{child grid}. 
-ENZO's implementation of structured AMR places no restrictions on 
-the number of grids at a given level of refinement, or on the number of 
-levels of refinement.  However, owing to limited computational resources 
-it is practical to institute a maximum level of refinement $\Lmax$.
-
-The \enzo\ implementation of AMR allows arbitrary integer ratios of parent and
-child grid resolution.  However, in practice 
-refinement factors (defined as the ratio of parent grid resolution to child grid resolution) 
-of two or four are typically used, with a refinement factor
-of two being most commonly used for cosmological simulations for reasons of
-efficiency. The ratio of
-box size to the maximum grid resolution of a given simulation is
-therefore $L/e = \Nroot \times 2^{\Lmax}$, where $\Nroot$ is the
-number of cells along one edge of the root grid, $\Lmax$ is the
-maximum level of refinement allowed, and L and e are the box size
-and grid resolution of the most highly refined region, respectively.
-
-The AMR grid patches are the primary data structure in \enzo.  Each
-patch is treated as an individual object which can contain both field
-variables and particle data.  Individual grids are organized into a 
-dynamic, distributed hierarchy of mesh patches.  
-See 
-Figure~\ref{fig.2.amrhierdens} for an example of an \enzo\
-simulation performed using this AMR scheme where both the baryon density
-and AMR grid hierarchy are shown together.
-
-\begin{figure}
-\begin{center}
-\includegraphics[width=0.4\textwidth]{figures/amr_hierarchy.eps}
-\end{center}
-\caption{\emph{Left:}  Example of a simple, 
-distributed AMR hierarchy showing real and 
-ghost grids.
-\emph{Right:}  Example 2D \enzo\ grid showing real and ghost zones, as 
-needed for the PPM hydro stencil.  Image courtesy of James Bordner (Cent.
-Astrophysics and Space Sciences, UCSD).
-}
-\label{fig.2.amrhierarchy}
-\end{figure}
-
-\begin{figure}
-\begin{center}
-\includegraphics[width=0.4\textwidth]{figures/amr-hier-dens.eps}
-\caption{Example of an \enzo\ simulation showing the AMR grid hierarchy.  
-This is a simulation of the collapse of a single massive halo with a 
-$32^3$ root grid and two $32^3$ static nested grids.  AMR is only allowed
-within the most highly refined static nested grid.  Log baryon density is shown
-by colored isocontours with values corresponding to the legend at center left.  
-The rectangular solid wire frames correspond to individual 
-AMR grids, with different colors corresponding to level as indicated by the
-legend at top left.}
-\label{fig.2.amrhierdens}
-\end{center}
-\end{figure}
-
-Each grid patch in \enzo\ contains arrays of values for baryon and 
-particle quantities.   Grids are 
-partitioned into a core of \emph{real zones} and a surrounding layer 
-of \emph{ghost zones}.  The real zones store field values and ghost 
-zones are used to temporarily store values which have been obtained 
-directly from neighboring grids or interpolated from a parent grid.  
-These zones are necessary to accommodate the computational stencil of 
-the hydrodynamics solvers (Sections~\ref{sec.ov.hydro.ppm} and~\ref{sec.ov.hydro.zeus}) 
-and the gravity solver (Section~\ref{sec.ov.nbody}).  
-The PPM and Zeus hydro solvers require 3 layers of ghost zones,  and the gravity
-solver requires 6.  This can lead to significant memory 
-and computational overhead, particularly for smaller grid patches
-at high levels of refinement.  More information on filling of ghost
-zones can be found in the appendix.
-
-From the point of view of the hydro solver, each grid is solved as an
-independent CFD problem, with Dirichlet boundary conditions stored in
-the ghost zones.  At the end of each time step, cells on a given level
-fill ghost zones by interpolating from the parent and copying from
-neighbors on the same level.  Grids then correct the flux at the interface boundary between
-fine and coarse zones, and finally project their real zone data to the
-parent grid.
-
-Timesteps are determined on each level as
-described in \ref{sed.timestepping}, and the hierarchy is advanced on
-a level by level basis in what's known as a W-Cycle.    Beginning with
-the coarsest level, $L$, all grids on that level are advanced one
-timestep.  Then one timestep is taken on all grids at the next level
-of refinement, $L+1$, and so on until the finest level is resolved.
-The finest level is then advanced, using as many steps as it takes to
-reach the level immediately above.  The finest level is then synced
-to the level above it, which then proceeds forewards one more step.
-\red{Needs a picture of the W cycle.}
-
-At the end of every timestep on every level each grid updates 
-its ghost zones by exchanging information with its neighboring grid patches 
-(if any exist) and/or by interpolating from a parent grid.  
-In addition, cells are examined to see if they should be refined or 
-de-refined, and the entire grid hierarchy is rebuilt at that 
-level (including all more highly refined levels).  The timestepping and 
-hierarchy advancement/rebuilding process described here is repeated 
-recursively on every level to the specified maximum level of 
-refinement in the simulation.
-
-Since the addition of more highly refined grids is adaptive,  
-the conditions for refinement must be specified.  The criteria of 
-refinement can be set by the threshold value of the overdensity of 
-baryon gas or dark matter in a cell (which is really a refinement on
-the mass of gas or DM in a cell), the local Jeans length, the 
-local density gradient, or local pressure and energy gradients.  
-A cell reaching any or all of these criteria 
-will then be flagged for refinement.  
-Once all cells at a given level have been examined, rectangular boundaries 
-are determined which minimally encompass the flagged cells. The set of
-flagging cells is then further fractured into smaller bounding
-rectangles to ensure computational 
-efficiency.  More details on subgrid creation can be found in Appendix
-\ref{app.subgrid_creation}.  A refined grid patch is 
-introduced within each such bounding rectangle. Thus the cells needing 
-refinement, as well as adjacent cells within the patch which do not need 
-refinement, are refined. While this approach is not as memory efficient as 
-cell-splitting AMR schemes, it offers more freedom with 
-finite difference stencils. For example, PPM requires a stencil of seven 
-cells per dimension. This cannot easily be accommodated in cell-splitting 
-AMR schemes.
-
-\blue{ \emph{I think this  should be
-  re-written to reflect data presented or referenced in this paper,
-  whatever that winds up being.}
-In the simulations discussed in this thesis we typically use baryon and
-dark matter overdensities as our refinement criteria, though
-for some higher-resolution simulations we also refine on other
-criteria as needed.}
-
-\subsection{Time Stepping}\label{sec.timestepping}
-\dcc{Made this its own section (not really part of the AMR, it's more
-  general than that.)  Added stuff on harmonic average for 3d. See
-  note below on acceleration timestep criterion being crap. Moved
-  w-cycle to the previous section}
-In \enzo, resolution of the equations being solved is adaptive in time 
-as well as in space.  The timestep in \enzo\ is satisfied on a level-by-level
- basis by finding the largest timestep such that multiple criteria are
-satisfied on each level.  The timestep criteria used by Enzo are 
-(showing the one-dimensional case for clarity):
-
-\begin{equation}
-\Delta t_{hydro} = min \left( \kappa_{hydro} \frac{a \Delta x}{c_{s} + |v_x|} \right)_L ,
-\label{eqn:dthydro}
-\end{equation}
-
-\begin{equation}
-\Delta t_{dm} = min \left(\kappa_{dm} \frac{a \Delta x}{v_{dm,x}} \right)_L ,
-\label{eqn:dtdarkmatter}
-\end{equation}
-
-\begin{equation}
-\Delta t_{exp} = f_{exp} \left( \frac{a}{\dot{a}} \right) ,
-\label{eqn:dtexpand}
-\end{equation}
-
-\begin{equation}
-\Delta t_{accel} = min \left( \sqrt{\frac{\Delta x}{\vec{g}}} \right)_L 
-\label{eqn:dtaccel}
-\end{equation}
-
- In equations~\ref{eqn:dthydro}-\ref{eqn:dtaccel}, the $min ( \ldots )_L$
-formalism means that this value is calculated for all cells on a given level
-L and the minimum overall value is taken as the timestep.  
-Equation~\ref{eqn:dthydro} ensures that all cells satisfy the 
-Courant-Freidrichs-Levy (CFL) condition for accuracy and stability of an explicit
-finite difference discretization of the Euler equations.  Effectively this condition
-forces the timestep to be small enough such that any changes in the fluid propagate
-less than a single grid spacing, $\Delta x$.  In this equation, $\kappa_{hydro}$ is 
-a numerical constant with a value of $0 < \kappa_{hydro} \leq 1$ (with a typical
-value of $\kappa_{hydro} \sim 0.3-0.5$) that ensures that the CFL condition is always
-met, and $c_s$ and $v_x$ are the sound speed and peculiar baryon
-velocity in a given cell.
-
-Equation~\ref{eqn:dtdarkmatter} is analogous to Equation~\ref{eqn:dthydro} and ensures
-accuracy in the N-body solver by requiring that no dark matter particle travels more than
-one cell width.  The parameter $\kappa_{dm}$ is analogous to $\kappa_{dthydro}$, with
-an identical range of values.  Equation~\ref{eqn:dtexpand} limits the timestep such 
-that the expansion parameter a can only change by a fractional amount of $f_{exp} = \Delta a/a$, 
-where $f_{exp}$ is a user-defined parameter and has typical values of $f_{exp} = 0.01-0.02$.
-This is required for the stability of the PPM algorithm in comoving coordinates, and
-typically limits the timestep only at high redshifts when densities are relatively
-homogeneous.
-
-Equation~\ref{eqn:dtaccel} is supplementary to equation~\ref{eqn:dthydro} in that it
-takes into account the possibility of large accelerations causing numerical 
-instabilities by violating the Courant condition.  In this equation, $\vec{g}$ is the
-gravitational acceleration in each cell on level L.  \blue{ \emph{It should be
-  noted that the timestep is computed before the acceleration field is computed,
-  and the acceleration field is deleted at the end of each time step.
-  So this condition isn't ever actually enforced.  We should either
-  fix this problem, or not discuss it.
- -dcollins.} }
-
-Equation ~\ref{eqn:dthydro} is valid for one dimension.  For 2 or 3
-dimensions, it was shown by \cite{Godunov1959}  that using the
-harmonic average of the timestep found along each of the coordinate
-axes yields a maximum $\kappa_{hydro} = 0.8$.  So letting $\Delta
-t_x$, $\Delta t_y$, and $\Delta t_z$ be the analogues of
-eqn. \ref{eqn:dthydro} along the $x,y$ and $z$ axes, 
-\begin{equation}
-\Delta t_{hydro} = min ( \kappa_{hydro} /( \frac{1}{\Delta t_x}
-+\frac{1}{\Delta t_y} + \frac{1}{\Delta t_z} ) )_L
-\end{equation}
-
-For all other criteria, multiple dimensions are accounted for by
-repeating the one dimensional criterion along each axis, and taking
-the minimum.
-
-%---------------------------------------------------------------------------%
-\subsection{Parallelism}\label{sec.ov.parallel}
-\dcc{moved hierarchy description, parallism notes from AMR section
-  here. Mentioned performance/scaling study.  Still short.}
-
-When \enzo\ is run on a distrubuted platform, some of the data is
-replicated to all processors, while some remains on a single processor.
-A description of the entire hierarchy (the position and size of each grid, as well
-some other meta data) is stored on each processor.  This enables each
-grid to know what grids are near it in the hierarchy, regardless of
-where the data is.  The actual Baryon data (density, velocity, energy,
-and any chemistry data) and particle data are stored on only one
-processor.  Storing the entire hierarchy on all processor does create
-some overhead, but in practice this is not an issue until one reaches
-the extremely large scale.
-  The code handles load balancing 
-on a level-by-level basis such that the workload on each level is 
-distributed as uniformly as possible across all processors.  Spatial locality of 
-grids is not forced during message passing, for maximum flexibility (though not
-necessarily maximum efficiency).  
-The MPI message passing library\footnote{http://www-unix.mcs.anl.gov/mpi/}
- is used to transfer data between processors.
-
-The performance of \enzo\ has been extensively studied using the
-software package LCAPerf.  More on LCAPerf can be found in section
-\ref{sec.performance}
-
-%old text
-%Every processor 
-%keeps a description of the entire grid hierarchy at all times, so that 
-%each processor knows where all grids are.  However, baryon and particle 
-%data for a given grid only exists on a single processor.  See 
-%Figure~\ref{fig.2.amrhierarchy} for a schematic example of this.

### 2012-06-29 numerical-timestep.tex
first commit of paper.

--- 2012-06-29 numerical-overview.tex
@@
-\section{Overview of Numerical Methods}\label{sec.overview}
+\subsection{Time Stepping}
+\label{sec.timestepping}
 
-%---------------------------------------------------------------------------%
-\subsection{N-body dynamics}\label{sec.ov.nbody}
-fft + multigrid
+%\dcc{Made this its own section (not really part of the AMR, it's more
+%  general than that.)  Added stuff on harmonic average for 3d. See
+%  note below on acceleration timestep criterion being crap. Moved
+%  w-cycle to the previous section}
 
-There are multiple methods to compute the gravitational potential 
-(which is an elliptic equation in the Newtonian limit) in a structured 
-AMR framework.  One way would be to model the dark matter (or other
-collisionless particle-like objects, such as stars) as a second fluid 
-in addition to the baryonic fluid and solve the collisionless
-Boltzmann equation, which follows the evolution of the fluid density in
-six-dimensional phase space.  However, this is computationally prohibitive 
-owing to the large dimensionality of the problem, making this approach 
-unappealing for the cosmological AMR code.
-
-Instead, \enzo\ uses a particle-mesh N-body method to calculate 
-the dynamics of collisionless systems \citep{Hockney88}.  This method 
-follows trajectories 
-of a representative sample of individual particles and is much more 
-efficient than a direct solution of the Boltzmann equation in most 
-astrophysical situations. 
-The gravitational potential is computed by solving the elliptic 
-Poisson's equation (Eqtn.~\ref{enzopoisson}).
-
-These equations are finite-differenced and for simplicity are solved with the
-same timestep as the hydrodynamic equations, as discussed in 
-Section~\ref{sec.ov.amr}.
-The dark matter particles are distributed onto the grids
-using the cloud-in-cell (CIC) interpolation technique to form a spatially
-discretized density field (analogous to the baryon densities used to calculate
-the equations of hydrodynamics).  After sampling dark matter density onto the
-grid and adding baryon density if it exists (to get the total matter density
-in each cell), the gravitational potential is calculated on the periodic root
-grid using a fast Fourier transform.  In order to calculate more accurate
-potentials on subgrids, \enzo\ re-samples the dark matter density onto
-individual subgrids using the same CIC method as on the root grid, but at
-higher spatial resolution (and again adding baryon densities if applicable).
-Boundary conditions are then interpolated from the potential values on the
-parent grid (with adjacent grid patches on a given level communicating to
-ensure that their boundary values are the same), and then a multigrid
-relaxation technique is used to calculate the gravitational potential at every
-point within a subgrid.  Forces are computed on the mesh by
-finite-differencing potential values and are then interpolated to the particle
-positions, where they are used to update the particle's position and velocity
-information.  Potentials on child grids are computed recursively and particle
-positions are updated using the same timestep as in the hydrodynamic
-equations.  Particles are stored in the most highly refined grid patch at the
-point in space where they exist, and particles which move out of a subgrid
-patch are sent to the grid patch covering the adjacent volume with the finest
-spatial resolution, which may be of the same spatial resolution, coarser, or finer
-than the grid patch that the particles are moved from.  This takes place in a
-communication process at the end of each timestep on a level.
-
-At this point it is useful to emphasize that the effective force
-resolution of an adaptive particle-mesh calculation is approximately twice
-as coarse as the grid spacing at a given level of resolution.  The 
-potential is solved in each grid cell;
-however, the quantity of interest, namely the acceleration, is the gradient
-of the potential, and hence two potential values are required to calculate
-this.  In addition, it should be noted that the adaptive particle-mesh
-technique described here is very memory intensive: in order to ensure reasonably 
-accurate force resolution at grid edges the multigrid relaxation
-method used in the code requires a layer of ghost zones which is very deep --
-typically 6 cells in every direction around the edge of a grid patch.  This greatly
-adds to the memory requirements of the simulation, particularly because subgrids
-are typically small (on the order of $12^3 - 16^3$ real cells for a standard 
-cosmological calculation) and ghost zones can dominate the memory and computational 
-requirements of the code.
-
-%---------------------------------------------------------------------------%
-\subsection{Hydrodynamics}\label{sec.ov.hydro}
-
-\subsubsection{The PPM hydro method}\label{sec.ov.hydro.ppm}
-
-\dcc{Removed ``things to do'' list.  Mention ppm-lr and ppm-de, and
-  that ppm-de is preferred.  Briefly describe ppm-de itself.}
-
-The primary hydrodynamic method used in \enzo\ is based on the
-piecewise parabolic method (PPM) of
-\citet{1984JCoPh..54..174C}, which has been
-modified for the study of cosmological fluid flows.  
-\citet{1984JCoPh..54..174C} describe two variants on this method: a
-Lagrangian Remap variant and a Direct Eulerian variant.  Both have
-been adapted for \enzo, and the Lagrangian Remap 
-variant is described in \citet{1995CoPhC..89..149B}.  Because of its
-conservative formulation, however, the Direct Eulerian variant is
-preferred for \enzo\ AMR simulations.  We will describe the Direct
-Eulerian (PPM-DE) variant and its additions briefly here, and in
-further depth in Appendix \ref{app.ppm}
-
-PPM is an explicit, higher order-accurate version of
-Godunov's method for ideal gas dynamics with third order-accurate piecewise parabolic
-monotonic interpolation and a nonlinear Riemann solver for shock
-capturing.  It advances the hydrodynamic equations in the following
-steps:
-\begin{enumerate}
- \item Construct monotonic parabolic interpolation of cell average
- data, for each fluid quantity.
- \item Compute interface states by averaging the parabola for over the domain of dependence for
- each interface
- \item Use interface data to solve the Riemann problem.
- \item Difference the interface fluxes to update the cell average quantities.
-\end{enumerate}
-It does an excellent job of capturing strong shocks in at
-most two cells.  Multidimensional schemes are built up by directional
-splitting and produce a method that is formally second order-accurate
-in space and time which explicitly conserves mass, linear momentum, and energy.
-
-To use the above algorithm for cosmological simulations, the
-conservation laws for fluid mass, momentum and energy 
-density are written in comoving coordinates for a
-Friedman-Robertson-Walker space-time, as described previously 
-in Equations~\ref{enzoconserve} 
-through~\ref{enzoenergy}.  Both the conservation laws and
-the Riemann solver are modified to include gravity, which is solved
-using an adaptive particle-mesh (PM) technique (see Section~\ref{sec.ov.nbody}).  
-The terms due to cosmological expansion, as well as primordial chemistry and
-radiative heating and cooling, are solved in a separate step because they have
-different numerical behavior, and therefore must be treated differently to ensure
-stability.  Note that unlike the \zeus\ hydro scheme, PPM does not need to use 
-artificial viscosity to resolve shocks.
-
-The system of equations described above works well for systems with
-relatively low Mach numbers, as long as these systems are well resolved.  
-However, cosmology is replete with situations where there are bulk hypersonic
-flows.  In these situations, the ratio of kinetic to thermal energy can be
-very high -- in some situations up to $10^6 - 10^8$.  This implies that the
-thermal energy is an extremely tiny fraction of the kinetic energy, which can 
-cause numerical problems when one is interested in just the thermal energy of the gas, since
- Equation~\ref{enzoenergy} solves for the total energy.  In this system of
-equations, the
- thermal energy $E_{therm}$ is calculated as $E - E_{kin}$, 
-where E is the total specific energy as
-calculated in equation~\ref{enzoenergy} and $E_{kin}$ is the specific kinetic energy,
-$0.5 \vec{v}_b^2$.  
-In hypersonic flows $E$ and $E_{kin}$ are nearly the same, and any 
-number calculated as 
-the difference of these is going to be strongly affected by numerical error.  To 
-avoid this problem, \enzo\ also solves the internal energy equation in
-comoving coordinates:
-
-\begin{equation}
-\frac{\partial e}{\partial t} + \frac{1}{a} \vec{v}_b \cdot \nabla e = 
-- \frac{3(\gamma-1)\dot{a}}{a}e - \frac{p}{a \rho} \nabla \cdot \vec{v}_b
-\label{enzointenergy}
-\end{equation}
-
-In this equation $e$ is the internal energy (defined as $P/\rho (\gamma-1)$ and the other terms are as described
-previously.  The code still conserves total energy ($E$) as well.  In order
-to maintain consistency, both equations are solved at all times in all cells,
-with the equation for the total energy (eqtn.~\ref{enzoenergy}) being used
-for hydrodynamics routines and the internal energy $e$ being used when 
-temperature is required.  When pressure is required for dynamic purposes,
-the total energy is used if the ratio of thermal energy to total energy
-is less than some threshold value $\eta$, and the internal energy is used
-for values of the ratio larger than $\eta$.  A typical value of this parameter
-is $10^{-3}$.  This \emph{dual energy formulation} ensures that the
-method produces the correct entropy jump at strong shocks and also
-yields accurate pressures and temperatures in cosmological hypersonic
-flows.
-
-\subsubsection{The Zeus hydro method}\label{sec.ov.hydro.zeus}
-
-As a check on PPM, \enzo\ also includes an implementation of the
-finite-difference hydrodynamic algorithm employed in the compressible 
-magnetohydrodynamics code `\zeus'
-\citep{Stone92a, Stone92b}.  Fluid transport is solved on a Cartesian
-grid using the upwind, monotonic advection scheme of \citet{1977JCoPh..23..276V} 
-within a multistep 
-(operator split) solution procedure which is fully explicit in time.  
-This method is formally second order-accurate in space but 
-first order-accurate in time.  It is important to note that 
-this method conserves internal energy
-rather than total energy, so the ``dual energy
-formulation'' discussed in Section~\ref{sec.ov.hydro.ppm} is unnecessary.
-
-Operator split methods break
-the solution of the hydrodynamic equations into parts, with each part
-representing a single term in the equations.  Each part is evaluated
-successively using the results preceding it.  The individual parts of
-the solution are grouped into two steps, called the \emph{source} and
-\emph{transport} steps.
-
-The \zeus\ method uses a von Neumann-Richtmeyer artificial viscosity 
-to smooth shock discontinuities that may
-appear in fluid flows and can cause a break-down of finite-difference
-equations.  The artificial viscosity term is added in the source terms
-as
-\begin{eqnarray}
-\rho \frac{\partial\textbf{v}}{\partial t} &=& - \nabla p - \rho \nabla \phi 
-- \nabla \cdot \textbf{Q} \\
-\frac{\partial e}{\partial t} &=& -p \nabla \cdot \textbf{v} - \textbf{Q} : \nabla \textbf{v}, 
-\end{eqnarray}
-where \textbf{v} is the baryon velocity, $\rho$ is the mass density,
-$p$ is the pressure, $e$ is the internal energy density of the gas and \textbf{Q}
-is the artificial viscosity stress tensor, such that:
-
-\begin{eqnarray}
-Q_{ii} = \left\{ \begin{array}{ll}
-Q_{\rm AV} \rho_b  (a \Delta v_{i} + \dot{a} \Delta x_i)^2, 
-& \textrm{for $a \Delta v_{i} + \dot{a} \Delta x_i < 0$}  \\
-0 & \textrm{otherwise}\\
-\end{array} \right. 
-\end{eqnarray}
-and
-\begin{equation}
-Q_{ij} = 0  \;\;\;{\rm for}\;\; i \ne j .
-\end{equation}
-
-$\Delta x_i$ and $\Delta v_{i}$ refer to the comoving 
-width of the grid cell
-along the $i$-th axis and the corresponding difference in gas
-peculiar velocities across the grid cell, respectively, and $a$ is the
-cosmological scale factor.  $Q_{\rm AV}$ is a constant with a typical
-value of 2. We refer the interested reader to \citet{1994ApJ...429..434A} for more details.
-
-The limitation of a technique that uses an artificial viscosity is that,
-while the correct Rankine-Hugoniot jump conditions are achieved,
-shocks are broadened over 6-8 mesh cells and are thus not treated as true
-discontinuities. This may cause unphysical pre-heating of gas upstream
-of the shock wave, as discussed in \citet{1994ApJ...429..434A}.
-
-%---------------------------------------------------------------------------%
-\subsection{Chemistry, radiation and cooling}\label{sec.ov.chem}
-
-\red{
-As this stands it is the methods that I (brian) use, so I don't really emphasize
-either the tabular cooling functions or the various radiation background options.
-Add more discussion of this, point towards appendix.
-}
-
-Though the equations of hydrodynamics described above are a closed system,
-they are still missing a crucial piece of physics: radiative heating and cooling.
-Radiative cooling is extremely important in many astrophysical situations, as is 
-the heating of gas from some sort of radiation background.  \enzo\ has a very simple
-Sutherland and Dopita equilibrium cooling function \citep{1993ApJS...88..253S} 
-implemented, which uses a
-cooling table assuming a fixed metallicity of $Z = 0.3 Z_\odot$, and also a
-nonequilibrium heating/cooling model that assumes gas of primordial composition exposed to a
-uniform metagalactic ultraviolet (UV) background that varies with time \citep{1996ApJS..105...19K}.
-
-The simulations discussed in this thesis almost exclusively use the nonequilibrium
-routines, described in great detail by Abel et al. and Anninos et al.  
-\citep{abel97,anninos97} and summarized in Appendix~\ref{app:chemistry}. These 
-routines follow the non-equilibrium chemistry of a 
-gas of primordial composition with 9 total species:  
-$H$, $H^+$, $He$, $He^+$, $He^{++}$, $H^-$, $H_2^+$, $H_2$, and $e^-$.  The code also calculates 
-radiative heating and cooling, following atomic line excitation, recombination,
-collisional excitation, free-free transitions, molecular line excitations, and Compton
-scattering of the cosmic microwave background, as well as any of
-approximately a dozen different models for a metagalactic UV background that heat
-the gas via photoionization and photodissociation. 
-The multispecies rate equations are solved out of
-equilibrium to properly model situations where, e.g., the cooling rate of the gas
-is much shorter than the hydrogen recombination time.  The effect of this nonequilibrium
-cooling is to leave behind a much larger fraction of residual free electrons than one would
-expect if the assumption of equilibrium were being made.  The practical effect of this
-is that more $H^-$ is formed, which subsequently produces hydrogen molecules.  If large
-amounts of $H_2$ is formed it can greatly increase the cooling rate of primordial gas
-at relatively low temperatures ($T \leq 10^4$ K).  This can efficiently cool the gas 
-to approximately $200$ K, which significantly reduces the Jeans mass of the gas.  Correct
-modeling of the formation of molecular hydrogen is crucial to the study of star formation
-in a primordial gas.
-
-A total of 9 kinetic equations are solved, including 29 kinetic and radiative 
-processes, for the 9 species mentioned above.  See Table~\ref{table.collisional} 
-for the collisional processes and Table~\ref{table.radiative} for the 
-radiative processes solved.
-
-%---------------- table of collisional processes
-
-\begin{table}
-\begin{center}
-{\bfseries Collisional Processes}\\[1ex]
-\begin{tabular}{llllllll}
-(1) & H & + & e$^-$ & $\rightarrow$ & H$^+$ &+& 2e$^-$ \\
-(2) & H$^+$ &+ &e$^-$ & $\rightarrow$ & H &+ &$\gamma$ \\
-(3) & He &+& e$^-$ & $\rightarrow$ & He$^+$ &+& 2e$^-$  \\
-(4) & He$^+$ &+& e$^-$ & $\rightarrow$ & He &+ &$\gamma$  \\
-(5) & He$^{+}$ &+& e$^-$ & $\rightarrow$ & He$^{++}$ &+& 2$e^-$  \\
-(6) & He$^{++}$ &+& e$^-$ & $\rightarrow$ & He$^+$ &+& $\gamma$ \\
-(7) & H &+& e$^-$ &$\rightarrow$& H$^-$ &+& $\gamma$  \\
-(8) & H$^-$ &+& H &$\rightarrow$ & H$_2$ & +& e$^-$ \\
-(9) & H &+ &H$^+$ &$\rightarrow$ &H$_2^+$ &+ &$\gamma$ \\
-(10) & H$_2^+$ &+ &H &$\rightarrow$ &$H_2$ &+ &$H^+$ \\
-(11) & H$_2$ &+ &H$^+$ &$\rightarrow$ &H$_2^+$ & +& H \\
-(12) & H$_2$ &+ &e$^-$ & $\rightarrow$ & 2H & + & e$^-$  \\
-(13) & H$_2$ & + & H & $\rightarrow$ & 3H &   &      \\
-(14) & H$^-$ & + & e$^-$ & $\rightarrow$ & H & + & 2e$^-$ \\
-(15) & H$^-$ & + & H & $\rightarrow$ & 2H & + & e$^-$ \\ 
-(16) & H$^-$ & + & H$^+$ & $\rightarrow$ & 2H & & \\
-(17) & H$^-$ & + & H$^+$ & $\rightarrow$ & H$_2^+$ & + & e$^-$ \\
-(18) & H$_2^+$ & + & e$^-$ & $\rightarrow$ & 2H & & \\
-(19) & H$_2^+$ & + & H$^-$ & $\rightarrow$ & H$_2$ & + & H  \\
-(20) & 3H & & & $\rightarrow$ & H$_2$ & + & H
-\end{tabular}
-\caption[]{Collisional processes solved in the Enzo nonequilibrium
-primordial chemistry routines.}
-\label{table.collisional}
-\end{center}
-\end{table}
-
-\begin{table}
-\begin{center}
-{\bfseries Radiative Processes}\\[1ex]
-\begin{tabular}{llllllll}
-(21) & H & + & $\gamma$ & $\rightarrow$ & H$^+$ & + & e$^-$ \\
-(22) & He & + & $\gamma$ & $\rightarrow$ & He$^+$ & + & e$^-$ \\
-(23) & He$^+$ & + & $\gamma$ & $\rightarrow$ & He$^{++}$ & + & e$^-$ \\
-(24) & H$^-$ & + & $\gamma$ & $\rightarrow$ & H & + & e$^-$ \\
-(25) & H$_2$ & + & $\gamma$ & $\rightarrow$ & H$_2^+$ & + & e$^-$ \\
-(26) & H$_2^+$ & + & $\gamma$ & $\rightarrow$ & H & + & H$^+$ \\
-(27) & H$_2^+$ & + & $\gamma$ & $\rightarrow$ & 2H$^+$ & + & e$^-$ \\
-(28) & H$_2$ & + & $\gamma$ & $\rightarrow$ & H$_2^*$ & $\rightarrow$ & 2H \\
-(29) & H$_2$ & + & $\gamma$ & $\rightarrow$ & 2H &  & 
-\end{tabular}
-\caption[]{Radiative processes solved in the Enzo nonequilibrium
-primordial chemistry routines.}
-\label{table.radiative}
-\end{center}
-\end{table}
-
-The chemical reaction equation network is technically challenging to solve due to 
-the huge range of reaction timescales involved -- the characteristic creation
-and destruction timescales of the various species and reactions can differ by 
-many orders of magnitude.  As a result, the set of rate equations is extremely 
-stiff, and an explicit scheme for integration of the rate equations can be 
-exceptionally costly if small enough timesteps are taken to keep the network
-stable.  This makes an implicit scheme much more preferable for such a set of 
-equations.  However, an implicit scheme typically require an iterative 
-procedure to converge, and for large networks (such as this one) an implicit
-method can be very time-consuming, making it undesirable for a large, three-dimensional
-simulation.
-
-\enzo\ solves the rate equations using a method based on a backwards differencing 
-formula (BDF) in order to provide a stable and accurate solution.  This technique is
-optimized by taking the chemical intermediaries $H^-$ and $H_2^+$, which have 
-large rate coefficients and low concentrations, and grouping them into a separate
-category of equations.  Due to their fast reactions these species are very sensitive
-to small changes in the more abundant species.  However, due to their low overall
-concentrations, they do not significantly influence the abundance of species with
-higher concentrations.  Therefore, reactions involving these two species can be
-decoupled from the rest of the network and treated independently.  In fact, $H^-$ 
-and $H_2^+$ are treated as being in equilibrium at all times, independent of 
-the other species and the hydrodynamic variables.  This allows a large speedup
-in solution as the BDF scheme is then applied only to the slower 7-species network
-on timescales closer to those required by the hydrodynamics of the simulation.
-Even so, the accuracy and stability of the scheme is maintained by subcycling the 
-rate solver over a single hydrodynamic timestep.  These subcycle timesteps are 
-determined so that the maximum fractional change in the electron concentration is
-limited to no more than $10\%$ per timestep.
-
-It is important to note the regime in which this model is valid.  According to Abel et al. and
-Anninos et al. \citep{abel97,anninos97}, the reaction network is valid for temperatures
-between $10^0 - 10^8$ K.  The original model discussed in these two references is only
-valid up to $n_H \sim 10^4$~cm$^{-3}$.  However, addition of the 3-body $H_2$ formation
-process (equation 20 in Table~\ref{table.collisional}) allows 
-correct modeling of the chemistry of the gas up
-until the point where collisionally-induced emission from molecular hydrogen becomes an important
-cooling process, which occurs at $n_H \sim 10^{14}$~cm$^{-3}$.  A further concern is that
-the optically thin approximation for radiative cooling breaks down, which
-occurs before $n_H \sim 10^{16} - 10^{17}$~cm$^{-3}$.  Beyond this point, 
-modifications the cooling function that take into account the non-negligible
-opacity in the gas must be made, as discussed by \citet{2004MNRAS.348.1019R}. 
-Even with these modifications, a completely correct description of the cooling of
-this gas will require some form of radiation transport, which will greatly 
-increase the cost of the simulations.
-
-Several processes are neglected.  The deuterium atom and its processes are 
-completely ignored, which may have some effect.  Recent work shows that HD is
-a more effective coolant than previously thought~\citep{2005MNRAS.361..850L}.  However,
-the fractional abundance of HD is so low that under circumstances relevant to
-the formation of a Population III star in an un-ionized region it should be
-sub-dominant.  However, the enhanced electron fraction in fossil 
-HII regions (as discussed later in this thesis) could result in the HD molecule
-becoming a dominant cooling mechanism at relatively low ($\sim$ few hundred K)
-temperatures, and could potentially cool the gas down to below $100$~K, which can
-enhance fragmentation and could have important consequences for the IMF of 
-primordial stars forming in a relic HII region~\citep{2005MNRAS.364.1378N}.
-
-Aside from deuterium, the chemical reactions involving lithium 
-are also neglected.  According to \citet{1998A&A...335..403G}, these are 
-not important for the density and temperature regimes explored 
-by the simulations discussed in this thesis.  However, at higher densities it 
-is possible that there are regimes where lithium can be an important coolant.
-
-%---------------------------------------------------------------------------%
-\subsection{Star formation and feedback}\label{sec.ov.star}
-
-\red{By the time this paper is submitted my (BWO's) paper on the star formation
-and feedback algorithms, and the various parameters contained therein, should be
-at least written and submitted as well.  Point to this, and maybe include a couple
-interesting plots (perhaps in an appendix?).}
-
-While the physics discussed previously is all crucial to the study of
-cosmological structure formation, most cosmological observations are of
-stars and related phenomena.  Additionally, stars eject energy and metal-
-enriched gas throughout their lives, drastically modifying their own 
-environment.  The formation of galaxies and clusters cannot be completely
-modeled without including the feedback of energy and metals.  In particular,
-it is thought that feedback is crucial for suppressing the large numbers of
-dwarf-like galaxies that CDM theories predict 
-\citep{1991ApJ...381...14L,1991ApJ...379...52W}.  An early
-burst of star formation could remove a large fraction of cold gas from such systems
-\citep{1978MNRAS.183..341W,1991ApJ...367...45C}.  Also, the unexpectedly 
-low luminosities of small clusters and
-groups (relative to the richest clusters) is commonly explained through feedback
-\citep{1991ApJ...383..104K}.  Energy ejected during the formation of the cluster ellipticals 
-lowers the central density and hence the X-ray luminosity of such clusters 
- \citep{1997ApJ...484L..21C}.  Therefore, the inclusion of star formation and the 
-feedback of energy and metals into the intergalactic medium in a cosmological
-code is crucial for many reasons.
-
-We have extended the \enzo\ code to include multiple models for star formation
-and feedback.  It is difficult to directly simulate the formation of individual
-stars in the context of simulations of galaxy formation and evolution due
-to the immense computational cost.  Therefore, 
-we adopt parametric methods which attempt
-to model star forming regions.  One model is based on Kravtsov's method
- \citep{2003ApJ...590L...1K}, and the other is based on the method of 
-\citet{1992ApJ...399L.113C}, which has been modified for use in an AMR code.  The basic
-ideas behind the methods are straightforward and observationally motivated.
-Similar (though somewhat more advanced) methods have been employed for the 
-smoothed particle hydrodynamics
-(SPH) method, most recently by Springel \& Hernquist 
-\citep{sh03a,sh03b,hs03}. This 
-method assumes a
-multiphase IGM and has been shown to accurately reproduce the cosmic star
-formation rate, and will be implemented into \enzo\ in the near future.
-
-In the following sections we will detail both the Kravtsov and Cen \& Ostriker
-star formation methods separately.  Tests of the star formation and
-feedback algorithms will not be shown, since this work is currently in
-progress (and an extension of the AMR/SPH code comparison described in
-\citet{2005ApJS..160....1O}).
-
-\subsubsection{The Kravtsov star formation and feedback algorithm}
-
-Kravtsov's method of star formation is designed to reproduce
-the global Schmidt law of star 
-formation~\citep{2003ApJ...590L...1K,1959ApJ...129..243S}.
-This algorithm is deliberately minimal, and is explicitly geared
-towards modeling star formation in a phenomenological way on kiloparsec
-scales.  Stars are assume to form with a characteristic gas timescale 
-$\tau_*$ such that $\dot{\rho}_* = \rho_{gas}/\tau_*$ where $\rho_{gas}$ 
-and $\rho_*$ are the baryon gas and stellar densities, respectively.  
-This ``constant efficiency'' model on the scale of star formation 
-regions is well motivated observationally \citep{1996AJ....112.1903Y,2002ApJ...569..157W}.  
-Star formation is only allowed to take place in very dense regions
-with $\rho_{gas} \geq \rho_{SF}$, where $\rho_{SF}$ is a constant
-proper density threshold above which star formation is allowed to
-occur.  No other criteria are imposed.  Kravtsov's typical choices
-for $\tau_*$ and $\rho_{SF}$ are $\tau_* = 4$ Gyr and $\rho_{SF} =
-1.64~$M$_\odot$~pc$^{-3}$~$(n_H \sim 50$~cm$^{-3})$.  The adopted timescale
-is derived from the observationally-determined normalization of the
-Schmidt law, and the density threshold is determined by observations
-of star forming regions on $\sim 100$ pc scales.  Note that the density
-threshold is in proper, not comoving, units.
-
-Algorithmically, the star formation events in Kravtsov's code
-are assumed to occur once every global time step 
-$\Delta t_0 \leq 10^7$.  In cells where star formation is
-determined to occur (i.e. $\rho_{gas} \geq \rho_{SF}$) a
-collisionless ``star particle'' is assumed to form, with a 
-mass $m_* = \dot{\rho}_* V_{cell} \Delta t_0$, where
-$\dot{\rho}_*$ is described above and $V_{cell}$ is the 
-volume of the mesh cell.  These star particles are
-formed instantaneously, with all $m_*$ of gas being
-taken out of the cell and immediately deposited into the
-star particle.  This particle is then given the velocity
-of the gas in the cell, and thereafter treated as a 
-collisionless particle.  The \enzo\ implementation
-of this algorithm is similar, except that instead of 
-forming stars only at the root grid time step, we allow 
-stars to form at the time step of the highest level of
-resolution at any particular point in space.  As can be seen
-from the equation for $m_*$ above, this can result in very 
-small stellar masses.  To avoid memory and processor time
-issues related to having very large numbers of star particles
-we impose a threshold mass $M_{*,min}$ such that a star
-particle only forms if $m_* \geq M_{*,min}$.  An
-appropriate choice of $M_{*,min}$ does not significantly 
-change the star overall star formation history of a 
-simulation, though it may delay the onset of star formation 
-in a given cell relative to a simulation without a particle mass
-threshold.
-
-Each ``star particle'' is assumed to represent an ensemble
-of stars and is treated as a single-age stellar population.
-Kravtsov assumes that the IMF is described by a
-Miller \& Scalo functional form with stellar masses between
-$0.1$ and $100~M_\odot$ \citep{1979ApJS...41..513M}.  All stars in this
-IMF with $M_* > 8~M_\odot$ deposit $2 \times 10^{51}$ ergs of
-thermal energy and a mass $f_z M_*$ of metals into the
-cell in which they form without delay, with
-$f_z \equiv min(0.2, 0.01~M_*-0.06)$ (i.e. instantaneous
-deposition of metals).  The definition of $f_z$ is a rough
-approximation of the results of 
-\citet{1995ApJS..101..181W}.
-
-Kravtsov reports that simulations with this algorithm
-reliably reproduce the star formation rate-gas surface
-density relation of the Schmidt law, and are not particularly
-sensitive to numerical parameters \citep{2003ApJ...590L...1K}.
-He also notes that this is surprisingly insensitive to
-the presence or absence of feedback and details of the cooling
-and heating properties of the gas, which suggests that the
-global star formation rate is determined by gravitationally-
-driven supersonic turbulence (on large scales)
-rather than stellar feedback or thermal instabilities on small 
-scales.
-
-\subsubsection{The Cen \& Ostriker star formation algorithm}
-
-The Cen \& Ostriker method is a heuristic model of star formation
-on galactic scales.  This method, first described by
-\citet{1992ApJ...399L.113C}, is similar in some ways to the Kravtsov algorithm
-but has more complex criteria for determining where stars should
-be formed.  In this method, cells that form stars must have a 
-baryon overdensity higher than some threshold 
-$\rho_b/\bar{\rho}_b \geq \eta$ 
-where $\rho_b$ is the baryon density in that cell, 
-$\bar{\rho_b}$ is the mean baryon density in the simulation,
-and $\eta$ is the user-defined overdensity threshold.
-Additionally, the gas in the cells must be contracting, 
-cooling rapidly, and gravitationally unstable, e.g.:
-
-\begin{equation}
-\nabla \cdot \vec{v}_b < 0,
-\label{cencont}
-\end{equation}
-
-\begin{equation}
-t_{cool} < t_{dyn} \equiv \sqrt{3 \pi / 32G \rho_{tot}},
-\end{equation}
-
-\begin{equation}
-m_{b} > m_{J} \equiv G^{-3/2} \rho_{b}^{-1/2}C^{3}
-\left[ 1 + \frac{\delta\rho_{d}}{\delta\rho_{b}} \right]^{-3/2}
-\end{equation}
-
-where $\vec{v}$ is the velocity of the gas in the cell, $\rho_{b}$ and 
-$\rho_{d}$ are the cell's baryon and dark matter density, respectively,
-$\rho_{total} = \rho_{b} + \rho_{d}$, $m_{b}$ and $m_{j}$ are the 
-baryonic mass in the cell and jeans mass of the cell, and C is the 
-isothermal sound speed in the cell.  
-If all of these criteria are met, 
-the mass of a star particle is calculated as \(m_{*} = m_{b} 
-\frac{ \Delta t}{ t_{dyn} } f_{*eff} \), 
-where $f_{*eff}$ is the star formation efficiency parameter.
-
-If $m_{*}$ is greater than a minimum star mass $m_{*min}$, a particle 
-is created and given several attributes:  Mass, a unique index number, 
-the time of formation $t_{form}$, the local dynamical free-fall time 
-$t_{dyn}$ and the metallicity fraction of the baryon gas in the cell 
-$f_{Zb}$.  There is a user-defined minimum dynamical time
-$T_{dyn,min}$ which is observationally motivated and affects the
-feedback rates (see below).
-The particle is placed in the center of the cell and given 
-the same peculiar velocity as the gas in the cell, and is then treated 
-in the same manner as the dark matter particles.  An 
-amount of baryon gas corresponding to the new particle's mass is 
-removed from the cell.
-
-In addition, we have added a stochastic star formation algorithm that keeps 
-track of all of the sub-mass threshold stars that should have been created 
-and when the total mass of uncreated stars is greater than the minimum mass, 
-a star particle is created.
-
-The feedback of energy and metals into the baryon gas is similar to
-the Kravtsov feedback, with some important differences.  The star formation 
-algorithm creates each star particle instantaneously.  However, feedback
-should take place over a significant timescale, as all of the stars contained
-within the ``star particle'' would form over a long period of time.  Therefore,
-we assume that for the purposes of feedback that the mass of stars formed
-at a time $t$ with finite timestep $\Delta t$ is:
-
-\begin{eqnarray}  
-\Delta m_{sf} =  m_{*} \left[ \right. \left(1 + \frac{t-t_{form}}{t_{dyn}} 
-                    \right) \nonumber \\ 
-		   \times exp \left( \frac{-(t-t_{form})}{t_{dyn}} \right) 
-                    \nonumber \\ 
-	       - \left( 1 + \frac{t+ \Delta t-t_{form}}{t_{dyn}} \right) \nonumber \\ 
-                    exp \left( \frac{-(t+\Delta t-t_{form})}{t_{dyn}} \right) 
-                   \left. \right]
-\end{eqnarray}
-
-which can be represented more clearly in integral form:
-
-\begin{eqnarray}
-\int_{t}^{t+Dt} \frac{dM}{dt} dt = \int_{t}^{t+Dt} m_{*} 
-\frac{dt}{t_{tyn}} \left( \frac{t-t_{form}}{t_{dyn}} \right) \nonumber \\ 
-\times exp \left( - ~ \frac{t-t_{form}}{t_{dyn}} \right) 
-\end{eqnarray}
-
-During this timestep, we assume that the star particle feeds back
-metal-enriched gas and thermal energy from supernovae and from stellar winds.
-Since massive stars have very short lifetimes, we assume that there is an
-immediate feedback of some fraction $f_{SN}$ of the rest energy from the
-stars created that timestep into the baryon gas, such that 
-$E_{add} = f_{SN} \Delta m_{sf} c^2$, where c is the speed of light.
-In addition, a fraction $f_{Z*}$ of the metal from the star particle
-is fed back into the baryon gas, which takes into account the effects
-of metal recycling.  Finally, a fraction of the mass $f_{m*}$ from all
-stars (rather than just supernovae) is fed back into the gas along with
-momentum in order to simulate the mass ejection from non-exploding stars
-via stellar winds.
-
-There are six user-defined parameters in this algorithm:  three deal with the
-star formation ($\eta$, $m_{*min}$ and $T_{dyn,min}$), and three 
-deal with feedback ($f_{SN}$, $f_{Z*}$ and $f_{m*}$).  Some of these
-parameters are completely free, while others can be guided by observation
-or theory.  For example, the supernova feedback parameter, $f_{SN}$, can be 
-constrained assuming that, for every $200 M_\odot$ of stars created, one
-supernova occurs, and this event feeds back approximately $10^{51}$ ergs of
-thermal energy, giving:
-
-\begin{equation}
-f_{SN} = \frac{10^{51} ergs}{200~M_\odot c^2} \simeq 3 \times 10^{-6}
-\end{equation}
-
-The metal yield $f_{Z*}$, defined as the mass in metals produced per unit
-mass of stars created, can be constrained by a theoretical model of
- \citet{1995ApJS..101..181W}.  This model suggests that $f_{Z*} = 0.02$
-is an appropriate number.  The minimum dynamical time is set to be $T_{dyn,min} = 10^7$
-~years to agree with timescales seen in nearby OB associations.
-
-The other parameters, such as the overdensity threshold $\eta$, minimum star
- mass $m_{*min}$, and mass ejection fraction $f_{m*}$ are not well constrained 
-either theoretically or observationally.  Indeed, $m_{*min}$ is a purely
-numerical parameter designed to keep the code from producing too many star 
-particles, and thus has no observational or theoretical counterpart.  These
-parameters have to be set by performing parameter studies and comparing to 
-observations.  Unfortunately, the range of parameter space is large, and
-the results may be degenerate for some combinations of these parameters.
-
-%---------------------------------------------------------------------------%
-\subsection{Adaptive mesh refinement}\label{sec.ov.amr}
-\dcc{removed most of the notes.  Added intro blurb.  moved parallelism
-  stuff. Re-worded the bit on ghost zones.  Summary of 'grid as
-  independent update + syncing with neighbors and parents,' W-cycle. Added
-  sentence on subgrid creation.  Made timestepping its own subsection.}
-\red{Talk about reasonable refinement factors - which 
-ones are usable, etc?}
-
-The primary purpose of the \enzo\ code is its Adaptive Mesh Refinement
-capability, which allows it to reach extremely large dynamical ranges
-with limited computational resources, opening doors previously closed
-by finite memory and computational time. Unlike moving mesh methods
-\citep{1995ApJS..100..269P,1995ApJS...97..231G} or  
-methods that subdivide 
-individual cells \citep{Adjerid}, Berger \& Collela's AMR (also referred 
-to as \emph{structured} AMR) utilizes an adaptive hierarchy of grid 
-patches at varying levels of resolution.  Each rectangular grid patch 
-(referred to as a ``grid'') covers some region of space in its 
-\emph{parent grid} which requires higher resolution, and can itself 
-become the parent grid to an even more highly resolved \emph{child grid}. 
-ENZO's implementation of structured AMR places no restrictions on 
-the number of grids at a given level of refinement, or on the number of 
-levels of refinement.  However, owing to limited computational resources 
-it is practical to institute a maximum level of refinement $\Lmax$.
-
-The \enzo\ implementation of AMR allows arbitrary integer ratios of parent and
-child grid resolution.  However, in practice 
-refinement factors (defined as the ratio of parent grid resolution to child grid resolution) 
-of two or four are typically used, with a refinement factor
-of two being most commonly used for cosmological simulations for reasons of
-efficiency. The ratio of
-box size to the maximum grid resolution of a given simulation is
-therefore $L/e = \Nroot \times 2^{\Lmax}$, where $\Nroot$ is the
-number of cells along one edge of the root grid, $\Lmax$ is the
-maximum level of refinement allowed, and L and e are the box size
-and grid resolution of the most highly refined region, respectively.
-
-The AMR grid patches are the primary data structure in \enzo.  Each
-patch is treated as an individual object which can contain both field
-variables and particle data.  Individual grids are organized into a 
-dynamic, distributed hierarchy of mesh patches.  
-See 
-Figure~\ref{fig.2.amrhierdens} for an example of an \enzo\
-simulation performed using this AMR scheme where both the baryon density
-and AMR grid hierarchy are shown together.
-
-\begin{figure}
-\begin{center}
-\includegraphics[width=0.4\textwidth]{figures/amr_hierarchy.eps}
-\end{center}
-\caption{\emph{Left:}  Example of a simple, 
-distributed AMR hierarchy showing real and 
-ghost grids.
-\emph{Right:}  Example 2D \enzo\ grid showing real and ghost zones, as 
-needed for the PPM hydro stencil.  Image courtesy of James Bordner (Cent.
-Astrophysics and Space Sciences, UCSD).
-}
-\label{fig.2.amrhierarchy}
-\end{figure}
-
-\begin{figure}
-\begin{center}
-\includegraphics[width=0.4\textwidth]{figures/amr-hier-dens.eps}
-\caption{Example of an \enzo\ simulation showing the AMR grid hierarchy.  
-This is a simulation of the collapse of a single massive halo with a 
-$32^3$ root grid and two $32^3$ static nested grids.  AMR is only allowed
-within the most highly refined static nested grid.  Log baryon density is shown
-by colored isocontours with values corresponding to the legend at center left.  
-The rectangular solid wire frames correspond to individual 
-AMR grids, with different colors corresponding to level as indicated by the
-legend at top left.}
-\label{fig.2.amrhierdens}
-\end{center}
-\end{figure}
-
-Each grid patch in \enzo\ contains arrays of values for baryon and 
-particle quantities.   Grids are 
-partitioned into a core of \emph{real zones} and a surrounding layer 
-of \emph{ghost zones}.  The real zones store field values and ghost 
-zones are used to temporarily store values which have been obtained 
-directly from neighboring grids or interpolated from a parent grid.  
-These zones are necessary to accommodate the computational stencil of 
-the hydrodynamics solvers (Sections~\ref{sec.ov.hydro.ppm} and~\ref{sec.ov.hydro.zeus}) 
-and the gravity solver (Section~\ref{sec.ov.nbody}).  
-The PPM and Zeus hydro solvers require 3 layers of ghost zones,  and the gravity
-solver requires 6.  This can lead to significant memory 
-and computational overhead, particularly for smaller grid patches
-at high levels of refinement.  More information on filling of ghost
-zones can be found in the appendix.
-
-From the point of view of the hydro solver, each grid is solved as an
-independent CFD problem, with Dirichlet boundary conditions stored in
-the ghost zones.  At the end of each time step, cells on a given level
-fill ghost zones by interpolating from the parent and copying from
-neighbors on the same level.  Grids then correct the flux at the interface boundary between
-fine and coarse zones, and finally project their real zone data to the
-parent grid.
-
-Timesteps are determined on each level as
-described in \ref{sed.timestepping}, and the hierarchy is advanced on
-a level by level basis in what's known as a W-Cycle.    Beginning with
-the coarsest level, $L$, all grids on that level are advanced one
-timestep.  Then one timestep is taken on all grids at the next level
-of refinement, $L+1$, and so on until the finest level is resolved.
-The finest level is then advanced, using as many steps as it takes to
-reach the level immediately above.  The finest level is then synced
-to the level above it, which then proceeds forewards one more step.
-\red{Needs a picture of the W cycle.}
-
-At the end of every timestep on every level each grid updates 
-its ghost zones by exchanging information with its neighboring grid patches 
-(if any exist) and/or by interpolating from a parent grid.  
-In addition, cells are examined to see if they should be refined or 
-de-refined, and the entire grid hierarchy is rebuilt at that 
-level (including all more highly refined levels).  The timestepping and 
-hierarchy advancement/rebuilding process described here is repeated 
-recursively on every level to the specified maximum level of 
-refinement in the simulation.
-
-Since the addition of more highly refined grids is adaptive,  
-the conditions for refinement must be specified.  The criteria of 
-refinement can be set by the threshold value of the overdensity of 
-baryon gas or dark matter in a cell (which is really a refinement on
-the mass of gas or DM in a cell), the local Jeans length, the 
-local density gradient, or local pressure and energy gradients.  
-A cell reaching any or all of these criteria 
-will then be flagged for refinement.  
-Once all cells at a given level have been examined, rectangular boundaries 
-are determined which minimally encompass the flagged cells. The set of
-flagging cells is then further fractured into smaller bounding
-rectangles to ensure computational 
-efficiency.  More details on subgrid creation can be found in Appendix
-\ref{app.subgrid_creation}.  A refined grid patch is 
-introduced within each such bounding rectangle. Thus the cells needing 
-refinement, as well as adjacent cells within the patch which do not need 
-refinement, are refined. While this approach is not as memory efficient as 
-cell-splitting AMR schemes, it offers more freedom with 
-finite difference stencils. For example, PPM requires a stencil of seven 
-cells per dimension. This cannot easily be accommodated in cell-splitting 
-AMR schemes.
-
-\blue{ \emph{I think this  should be
-  re-written to reflect data presented or referenced in this paper,
-  whatever that winds up being.}
-In the simulations discussed in this thesis we typically use baryon and
-dark matter overdensities as our refinement criteria, though
-for some higher-resolution simulations we also refine on other
-criteria as needed.}
-
-\subsection{Time Stepping}\label{sec.timestepping}
-\dcc{Made this its own section (not really part of the AMR, it's more
-  general than that.)  Added stuff on harmonic average for 3d. See
-  note below on acceleration timestep criterion being crap. Moved
-  w-cycle to the previous section}
 In \enzo, resolution of the equations being solved is adaptive in time 
@@
 the minimum.
-
-%---------------------------------------------------------------------------%
-\subsection{Parallelism}\label{sec.ov.parallel}
-\dcc{moved hierarchy description, parallism notes from AMR section
-  here. Mentioned performance/scaling study.  Still short.}
-
-When \enzo\ is run on a distrubuted platform, some of the data is
-replicated to all processors, while some remains on a single processor.
-A description of the entire hierarchy (the position and size of each grid, as well
-some other meta data) is stored on each processor.  This enables each
-grid to know what grids are near it in the hierarchy, regardless of
-where the data is.  The actual Baryon data (density, velocity, energy,
-and any chemistry data) and particle data are stored on only one
-processor.  Storing the entire hierarchy on all processor does create
-some overhead, but in practice this is not an issue until one reaches
-the extremely large scale.
-  The code handles load balancing 
-on a level-by-level basis such that the workload on each level is 
-distributed as uniformly as possible across all processors.  Spatial locality of 
-grids is not forced during message passing, for maximum flexibility (though not
-necessarily maximum efficiency).  
-The MPI message passing library\footnote{http://www-unix.mcs.anl.gov/mpi/}
- is used to transfer data between processors.
-
-The performance of \enzo\ has been extensively studied using the
-software package LCAPerf.  More on LCAPerf can be found in section
-\ref{sec.performance}
-
-%old text
-%Every processor 
-%keeps a description of the entire grid hierarchy at all times, so that 
-%each processor knows where all grids are.  However, baryon and particle 
-%data for a given grid only exists on a single processor.  See 
-%Figure~\ref{fig.2.amrhierarchy} for a schematic example of this.

### 2012-06-29 numerical-zeus.tex
first commit of paper.

\subsection{The Zeus hydro method}
\label{sec.hydro.zeus}

As an alternative to PPM, \enzo\ also includes an implementation of the finite-difference hydrodynamic algorithm employed in the compressible magnetohydrodynamics code \zeus \citep{Stone92a, Stone92b}.  Fluid transport is solved on a Cartesian grid using the upwind, monotonic advection scheme of \citet{1977JCoPh..23..276V} within a multistep (operator split) solution procedure which is fully explicit in time.  This method is formally second order-accurate in space but first order-accurate in time.

As discussed in the section describing PPM, operator split methods break the solution of the hydrodynamic equations into parts, with each part representing a single term in the equations.  Each part is evaluated successively using the results preceding it.  In this method, in addition to operator splitting the expansion terms (i.e. those terms in eqs.~(\ref{eq:mass}) -- (\ref{eq:total_energy}) which depend on $\dot{a}$), we divide the remaining terms into \emph{source} and \emph{transport} steps.  The terms to be solved in the source step are those on the right-hand side of eqs.~(\ref{eq:mass}) -- (\ref{eq:total_energy}), while the transport terms are on the left-hand side of these equations and are responsible for the advection of mass, momentum and energy across the grid.

The \zeus\ method uses a von Neumann-Richtmeyer artificial viscosity to smooth shock discontinuities that may appear in fluid flows and can cause a break-down of finite-difference equations.  The artificial viscosity term is added in the source terms as:
\begin{eqnarray}
\rho \frac{\partial\vecv}{\partial t} &=& - \nabla p - \rho \nabla \phi 
- \nabla \cdot \textbf{Q} \\
\frac{\partial e}{\partial t} &=& -p \nabla \cdot \vecv - \textbf{Q} : \nabla \vecv, 
\end{eqnarray}
%where \textbf{v} is the baryon velocity, $\rho$ is the mass density, $p$ is the pressure, $e$ is the internal energy density of the gas and 
Here \textbf{Q} is the artificial viscosity stress tensor, which we take to be diagonal with on-axis terms given by $l^2 \rho (\partial v / \partial x)^2$ as proposed by von Neumann \& Richtmyer.  The length scale $l$ determines the width of shocks and is typically a few times the cell spacing.

In our Cartesian coordinate system, the finite difference version of these equations is particularly simple, although there is one important complication.  In the \zeus\ formalism, the velocity is a face-centered quantity -- that is the velocity is recorded on a grid which is staggered as compared to the density, pressure and energy, which are at the cell center.  Therefore we must remember that $v_j$ is at position $x_{j-1/2}$ (we use notation, rather than writing $v_{j+1/2}$ both to match the original \zeus\ paper and also to make it easier to compare these equations to what is actually in the code).

As in the original \zeus\ paper, the source terms are added in three steps. First we add the pressure and gravity forces:
\begin{equation}
v_j^{n+a}  =  v_j^n - \frac{\Delta t}{\Delta x_j} \frac{p^n_j - p^n_{j-1}} {(\rho^n_j + \rho^n_{j-1})/2}.
\end{equation}
Partial updates are denoted by the $n+a$, $n+b$ notation.  We show only one-dimension as the extension to the multi-dimensional case is straightforward (note that all dimensions are carried out for each substep before progressing to the next substep). Then we add the artificial viscosity:
\begin{eqnarray}
v_j^{n+b} & = & v_j^{n+a} - \frac{\Delta t}{\Delta x_j} 
                             \frac{q^{n+a}_j - q^{n+a}_{j-1}} {(\rho^n_j + \rho^n_{j-1})/2} \\
e_j^{n+b} & = & e_j^n - \frac{\Delta t}{\Delta x_j} q^{n+a}_j (v^{n+a}_{j+1} - v^{n+a}_{j}).
\end{eqnarray}
The artificial viscosity coefficient $q_j$ is given by:
\begin{equation}
q_j = \left\{ \begin{array}{ll}
              Q_{\rm AV} \rho_j (v_{j+1} - v_j)^2 \quad & \rm{if} (v_{j+1} - v_j) < 0 \\
               0 & \rm{otherwise}
               \end{array} \right.
\end{equation}
where $Q_{\rm AV}$ is a constant with a typical value of 2. We refer the interested reader to \citet{Stone92a} and \citet{1994ApJ...429..434A} for more details.  We also include the option (turned off by default) of adding a linear artificial viscosity as suggested in the \zeus\ paper for stagnant flow regions.  This is given by
\begin{equation}
q_{{\rm lin},j} = Q_{\rm LIN} \rho c_j (v_{j+1} - v_{j})
\end{equation}
where $c_j^2 = \gamma p/\rho$ is the adiabatic sounds speed.

Finally, the third source step is the compression term and is given by
\begin{equation}
e^{n+c}_j = e^{n+b}_j \left( \frac{1 - (\Delta t/2) (\gamma - 1) (\nabla \cdot \vecv)_j }
                           {1 + (\Delta t/2) (\gamma - 1) (\nabla \cdot \vecv)_j } \right)
\end{equation}
We have used the notation $(\nabla \cdot \vecv)_j$ to indicate the (potentially) multi-dimensional velocity divergence evaluated at the cell center position $x_j$.  This equation differs from the previous ones in that in the multi-dimensional case, still only one finite difference equation is evaluated, but the divergence becomes multi-dimensional.

Next, we examine the transport step, which is conservative.  Once again, we dimensionally split the equations and so present only the one-dimensional version.  The finite-difference equations actually solved are:
\begin{equation}
\rho_j^{n+d} = \rho_j^{n} - \frac{\Delta t}{\Delta x} (v^{n+c}_{j+1/2} \rho^{*}_{j+1/2} - v^{n+c}_{j-1/2} \rho^{*}_{j-1/2} )
\end{equation}
Here $\rho^*_j$ is the correctly upwinded value of $\rho$ evaluated at the cell-face corresponding to $v_j$, making $\rho^*_j v_j$ the mass flux at the cell boundary and guaranteeing mass conservation.   This requires interpolating each cell-centered quantity to the cell edge.  As recomended in \citet{Stone92a}, we use the second-order van Leer scheme, which uses piecewise linear functions.  These are given by equations (48) and (49) of \citet{Stone92a}.  The transport steps for the other variables are similar.  Note that we advect the specific energy and specific momenta using the mass flux, as dictated by the principle of consistent transport.  This requires appropriate averaging for the momenta in the perpendicular directions as outline in equations (57)-(72) of \citet{Stone92a}.

%GB: the following is not correct (it is based on the Anninos & Norman paper which uses a
%GB: different set of definitions for the comoving coordinates).
%\begin{eqnarray}
%Q_{ii} & = & \left\{ 
%   \begin{array}{ll}
%      Q_{\rm AV} \rho_b  (a \Delta v_{i} + \dot{a} \Delta x_i)^2 \quad
%       & \textrm{for $a \Delta v_{i} + \dot{a} \Delta x_i < 0$}  \\
%      0 & \textrm{otherwise}\\
%\end{array} \right. \\
%%\end{eqnarray}
%%and
%%\begin{equation}
%Q_{ij} & = & 0  \;\;\;{\rm for}\;\; i \ne j . 
%%\end{equation}
%\end{eqnarray}
%$\Delta x_i$ and $\Delta v_{i}$ refer to the comoving width of the grid cell
%along the $i$-th axis and the corresponding difference in gas
%peculiar velocities across the grid cell, respectively, and $a$ is the
%cosmological scale factor.

A limitation of a technique that uses an artificial viscosity is that, while the correct Rankine-Hugoniot jump conditions are achieved, shocks are broadened over 3-4 mesh cells and are thus not treated as true discontinuities. This may cause unphysical pre-heating of gas upstream of the shock wave, as discussed in \citet{1994ApJ...429..434A}.  On the other hand, it is much more robust than PPM and is easy to add additional physics.  We also note that this method solves only the internal energy equation rather than total energy, so the ``dual energy formulation'' discussed in Section~\ref{sec.hydro.ppm} is unnecessary.

### 2012-06-29 overview.tex
first commit of paper.

\section{Physical Equations and Overview of Numerical Methodology}
\label{sec.overview}

We begin by laying out the equations to be solved and then discuss briefly the techniques used to solve each part.

% ====================================================

\subsection{Physical Equations}

%GB: The division of this subsection into subsubsections is somewhat arbitrary.
%GB: I wanted to do it at first in the same divisions as section 3, but this is not really
%GB: possible because there is no "AMR" equation.  Maybe somebody can think 
%GB: of a clever way to fix this.

% ----------------------------------------------------------------------------------------

\subsubsection{Hydrodynamics and Gravity}

The usual Eulerian equations of gas dynamics with gravity are given by

\begin{eqnarray} 
\left( \frac{\partial \rhop}{\partial t} \right)_r 
          + \vecvp \cdot \nabla_r \rhop
    & = & - \rhop \nabla_r \cdot \vecvp,
        \label{eq:mass_prime} \\
%
\left( \frac{\partial \vecvp }{ \partial t} \right)_r 
          + ( \vecvp \cdot \nabla_r ) \vecvp 
    & = & - \frac{1}{\rhop} \nabla_r \pp 
          - \nabla_r \phip,
        \label{eq:momentum_prime} \\
%
\left( \frac{\partial \Ep}{\partial t} \right)_r + \vecvp \cdot \nabla_r \Ep
    & = & - \frac{1}{\rhop} \nabla_r \cdot (\pp \vecvp) 
          - \vecvp \cdot \nabla_r \phip
%\nonumber          \\  & &     % uncomment for 2-column
     + \Gamma - \Lambda
        \label{eq:total_energy_prime}
\end{eqnarray}
%
Here $E$, $e$, $\rho$, $\vecv$ and $p$ are the baryonic total specific fluid energy, the baryonic specific thermal energy, the baryonic density, velocity and pressure, respectively.  Here we use primes to indicate variables defined in the absolute frame of reference (to distinguish them from a comoving frame).  The first equation represents conservation of mass, the second, conservation of momentum, and the third, conservation of total fluid (kinetic plus thermal) energy.  They are respectively, the first, second and third moments of the Boltzman equation. 
Terms representing radiative cooling ($\Lambda$) and heating ($\Gamma$) enter on the right-hand side of the energy equation (\ref{eq:total_energy_prime}).

The total fluid specific energy ($\Ep$) is given by:
\begin{equation}
\Ep =  e^\prime + \frac{1}{2} \vecvp \cdot \vecvp.
        \label{eq:total_energy_def_prime}
\end{equation}
It is useful to solve the equation of total energy because doing so enforces local energy conservation.

The equations are closed with the equation of state and Poisson's equation (we work in the Newtonian limit):
%
\begin{eqnarray}
%
e^\prime    & = & \pp / [(\gamma - 1) \rhop ],
        \label{eq:eq_of_state_prime} \\
%
\nabla_r^2 \phip & = & 4 \pi G \rho_{total}^\prime 
%                   + 3 p_{total}^\prime/c^2) - \Lambda. 
        \label{eq:potential_prime}
\end{eqnarray}
%
For simplicity, the equation of state is shown here for an ideal gas with a ratio of specific heats $\gamma$.
Equation (\ref{eq:potential_prime}) includes contributions to the gravitational potential $\phi^\prime$ from all  populations ($\rho_{total}^\prime$).
% as well as any radiation or massless neutrino background ($p_{total}^\prime$) or cosmological constant ($\Lambda_c$), if we are interested in the cosmological case.

% ----------------------------------------------------------------------------------------

\subsubsection{Comoving Coordinates}

This set of coupled hydrodynamic plus gravity equations is the standard one that \enzo solves.  However, in the cosmological case it is useful to transform these to a coordinate system that is comoving with the expanding universe \citep{Peebles93} (but note that \enzo\ can also operate in a fixed coordinate system). Specifically, we define:
\begin{eqnarray}
\vecx & = & \vecr/a, \\
\vecv & = & a \hspace{0.5mm} d\vecx/dt = 
              \vecvp - \dot{a}\vecx, \\
\rho    & = & a^3 \rhop,   \\
p       & = & a^3 \pp, \\
E       & = & \Ep - 
              \dot{a} \vecx \cdot \vecv - 
              \frac{1}{2} \dot{a}^2 \vecx^2, \\
\phi    & = & \phip + \frac{1}{2} a \ddot{a} \vecx^2.
\end{eqnarray}

These definitions result in the following comoving equivalents to eqs. 
(\ref{eq:mass_prime})-(\ref{eq:potential_prime}).
%

\begin{eqnarray} 
\frac{\partial \rhob}{\partial t} 
          + \frac{1}{a} \vecvb \cdot \nabla \rhob 
    & = & - \frac{1}{a} \rhob \nabla \cdot \vecvb, 
        \label{eq:mass} \\
%
\frac{\partial \vecvb}{\partial t}  
          + \frac{1}{a} ( \vecvb \cdot \nabla ) \vecvb 
    & = & - \frac{\dot{a}}{a} \vecvb 
          - \frac{1}{a \rhob} \nabla p 
          - \frac{1}{a} \nabla \phi, 
        \label{eq:momentum} \\
%
\frac{\partial E}{\partial t} 
         + \frac{1}{a} \vecvb \cdot \nabla E 
   & = & - \frac{\dot{a}}{a} (3 \frac{p}{\rhob} + \vecvb^2 )
         - \frac{1}{a \rhob } \nabla \cdot (p \vecvb) 
%\nonumber   & &   %uncomment for 2-column
         - \frac{1}{a} \vecvb \cdot \nabla \phi 
                + \Gamma - \Lambda.
       \label{eq:total_energy} \\
%
E  & = & e + \frac{1}{2} {\vecvb}^2,
        \label{eq:total_energy_def} \\
%
e                  & = & 
        p/\left[ \left( \gamma - 1 \right) \rhob \right],
      \label{eq:state} \\ 
{\nabla}^2 \phi    & = & 
        \frac{4\pi G}{a} ( \rho_{total} - \rho_0 ).
      \label{eq:poisson}
\end{eqnarray}
%
%We use the proper peculiar baryonic velocity $\vecv_b \equiv a(t) d\vecx / dt$, proper pressure $p$, and modified gravitational potential $ \phi $ which is related to the potential in proper coordinates $\Phi$ by 
%$$\phi \equiv \Phi + \frac{1}{2} a \ddot{a} \vecx^2. $$
%The density, however, is comoving:
%$$\rho_b \equiv \rho_{b,proper}a(t)^3, $$
The expansion parameter $a \equiv 1/(1 + z)$ follows the expansion of a smooth, homogeneous background, where $z$, the redshift, is a function only of $t$.  All derivatives are determined with respect to the comoving position $\vecx$, which is defined simply to remove the universal expansion from the coordinate system.

The evolution of $a(t)$ is governed by the formula for the expansion of an isotropic, homogeneous universe:
%
\begin{eqnarray}
\frac{\ddot{a}}{a} & = & 
      - \frac{4 \pi G }{3 a^3 } (\rho_0 
      + 3p_0/c^2) 
      + \Lambda_c /3 .
      \label{eq:expansion} 
\end{eqnarray}
%
Here, $\rho_0$ is the mean (baryonic and dark) comoving mass density and $p_0$ is the comoving background pressure contribution, and $\Lambda_c$ is the cosmological constant.
%$\gamma$ is the ratio of specific heats, and $\Lambda$ is the cosmological constant.
%We note that if $\gamma$ is not $5/3$ then $E$ in the first term on
%the right side of eq.~(ref{eq:total_energy}) must be replaced with 
%eq.~(\ref{eq:state}).
This system of equations is limited to the non-relativistic regime and assumes that curvature effects are not important --- both assumptions are reasonable as long as the  size of the simulated region is small compared to the radius of curvature and the Hubble length $c/H$ ($c$ is the speed of light and $H$ is the Hubble constant).

It is clear that these equation are equivalent to the absolute coordinate version in the limit in which $a = 1$ and $\dot{a} = 0$.  In the rest of this paper we will employ only the unprimed variables with the understanding that we are not restricting ourselves to cosmological applications.

\subsubsection{Particles and Chemistry}

Any collisionless component (e.g. dark matter or stars) are modeled by particles which are governed by Newton's equations in comoving coordinates:
%
\begin{eqnarray}
\frac{d \vecx}{dt} 
    & = & \frac{1}{a} \vecv, 
          \label{eq:dm_position} \\
\frac{d \vecv}{dt} 
    & = & - \frac{\dot{a}}{a} \vecv
          - \frac{1}{a} \nabla \phi, 
          \label{eq:dm_velocity} 
\end{eqnarray}
%
The particles also make a contribution to the potential through their contribution to Poisson's equations.

In addition, we solve mass conservation equations for any chemical species.  For any species $i$, these equations have the form:

\begin{eqnarray}
\frac{\partial \rho_i}{\partial t} 
          + \frac{1}{a} \vecvb \cdot \nabla \rho_i 
     & = &  - \frac{1}{a} \rho_i \nabla \cdot \vecvb 
        + \sum_j \sum_l k_{jl}(T) \rho_j \rho_l 
% \nonumber \\    & &    %uncomment for 2-column
      + \sum_j I_j \rho_j 
        \label{eq:mass_species}
\end{eqnarray}
where $k_{ij}$ are the rate coefficients for the two-body reactions and are usually functions of only the gas species (we will specifically note the cases where we either include three-body reactions or have density-dependent rates).  The $I_j$ are destruction/creation rates due to photoionizations and/or photodissociations.

%\red{
%There is some debate over whether to put in the hydro equations without cosmology, in addition
%to the equations that are already here.  Would this
%be more or less clear?  We're leaning towards saying ``when cosmology is turned off,
% $\dot{a}$ goes to zero and $a$ goes to 1, and the poisson equation changes to blah blah blah'' 
%and assuming our readers can figure it out for themselves.
%We also want to emphasize that the heating and cooling is glossed over in this section, 
%and it's discussed more thoroughly later.
%} %end red
% GB: I agree this would be shorter, and maybe it is a better idea, but I think it's nice to be able
% GB: to include both equations and at the same time make the transformation to comoving coords
% GB: very clear -- this has engendered a lot of confusion in the past and one of the goals of writing
% GB this paper is to answer peoples questions before they ask them.

% GB: I commented out the following.  It has now been incorporated into the above discussion
% GB: (see my comments anove on why I think this is a good way to do it)
%
%\enzo\ solves the equations of ideal gas dynamics
%in a coordinate system that is comoving with the expanding universe:

%\begin{equation}
%\frac{\partial \rho_b}{\partial t} + \frac{1}{a} \vecv_b \cdot \nabla \rho_b =  \label{enzoconserve}
% - \frac{1}{a} \rho_b \nabla \cdot \vecv_b 
%\end{equation}

%\begin{equation}
%\frac{\partial \vecv_b}{\partial t} + \frac{1}{a} (\vecv_b \cdot \nabla) \vecv_b = \label{enzomomentum} 
%-\frac{\dot{a}}{a} \vecv_b - \frac{1}{a \rho_b} \nabla p - \frac{1}{a} \nabla \phi
%\end{equation}

%\begin{eqnarray}
%\frac{\partial E}{\partial t} + \frac{1}{a} \vecv_b \cdot \nabla E =  
%- \frac{\dot{a}}{a} \left( 3 \frac{p}{\rho_b} + \vecv_b^2 \right)  \nonumber \\
%- \frac{1}{a \rho_b} \nabla \cdot (p \vecv_b) \nonumber \\
%- \frac{1}{a} \vecv_b \cdot \nabla \phi + \Gamma - \Lambda
%\label{enzoenergy}
%\end{eqnarray}

%Where $\rho_b$ is the comoving baryon density, $\vecv_b$ is the baryon velocity, $p$ is the pressure, 
%$\phi$ is the modified gravitational potential (in comoving coordinates, which is related to
%the potential in proper coordinates $\Phi$ by $\phi \equiv \Phi + 0.5 $a\"{a}$ \vecx^2$) and 
%$a$ is the ``expansion parameter'' which describes the expansion of a smooth, homogeneous 
%universe as a function of time.  This expansion parameter is related to the redshift:  
%$a \equiv 1/(1+z)$.  All derivatives are in comoving coordinates.  $E$ is the specific
%energy of the gas (total energy per unit mass), and
%$\Gamma$ and $\Lambda$ represent radiative heating and cooling processes as described below.
%  Equations~\ref{enzoconserve}, \ref{enzomomentum}
%and \ref{enzoenergy} represent the conservation of mass, momentum and total (e.g.,
%kinetic plus thermal) fluid energy.

%The equations above are closed with three more equations:

%\begin{equation}
%E = p / [(\gamma - 1) \rho_b] + \vecv^2/2   \label{enzoeos}
%\end{equation}

%\begin{equation}
%\nabla^2 \phi = \frac{4 \pi G}{a} (\rho_b + \rho_{dm} - \rho_0 )
%\label{enzopoisson}
%\end{equation}

%\begin{equation}
%\frac{\ddot{a}}{a} = - \frac{4 \pi G}{3 a^3} (\rho_0 + 3 p_0 / c^2) + \Lambda/3.
%\label{enzocomove}
%\end{equation}

%where $\rho_{dm}$ is the comoving dark matter density, $\rho_0$ is the comoving
%background density 
%($\rho_0 \equiv \Omega_{matter} \rho_{crit}$) and $p_0$ is the background pressure, 
%$\gamma$ is the ratio of specific heats and $\Lambda$ is the cosmological constant.
%Equations~\ref{enzoeos}, \ref{enzopoisson} and~\ref{enzocomove} are the 
%equation of state, Poisson's equation in comoving form and an equation that 
%describes the evolution of the comoving coordinates (i.e. the formula for the 
%expansion of an isotropic, homogeneous universe). 
%All particles in the simulation are governed by Newton's equations in comoving 
%coordinates:

%\begin{equation}
%\frac{d \vecx_{part}}{dt} = \frac{1}{a} \vecv_{part}
%\label{enzopartvel}
%\end{equation}

%\begin{equation}
%\frac{d \vecv_{part}}{dt} = -\frac{\dot{a}}{a} \vecv_{part} - \frac{1}{a} \nabla \phi
%\label{enzopartmom}
%\end{equation}

%Where $\vecx_{part}$ and $\vecv_{part}$ refer to the position and peculiar velocity of any
%particles in the system.  Note that the system of equations~\ref{enzoconserve}-\ref{enzopartmom} 
%is valid only in regimes where relativistic effects are not
%important ($v_b, v_{dm} \ll c$, where c is the speed of light) 
%and where cosmological curvature effects are 
%unimportant, meaning that the simulation volume is much smaller than the radius
%of curvature of the universe, as defined as $r_{hub} \equiv c/H_0$, where $c$ is the
%speed of light and $H_0$ is the Hubble constant.

%\dcc{Fixed punctuation, added 'no mhd, rhd' caveat to Zeus statements.  This has caused
%confusion in the past.}

% ====================================================

\subsection{Overview of Numerical Methods}

In this section, we briefly describe the numerical methods which are used to solve the equations outline above.  We sweep through in the same order as will be used in section~\ref{sec.methods} so that there is a one-to-one correspondence between each of the following overviews and the complete description provided in section~\ref{sec.methods}.  The goal here is to introduce the reader to the basic idea of the methods without drowning in detail.

\subsubsection{Structured Adaptive Mesh Refinement}

The primary purpose of the \enzo\ code is its Adaptive Mesh Refinement
capability, which allows it to reach extremely large dynamical ranges
with limited computational resources, opening doors previously closed
by finite memory and computational time. Unlike moving mesh methods
\citep{1995ApJS..100..269P,1995ApJS...97..231G} or  
methods that subdivide 
individual cells \citep{Adjerid}, Berger \& Collela's AMR (also referred 
to as \emph{structured} AMR) utilizes an adaptive hierarchy of grid 
patches at varying levels of resolution.  Each rectangular grid patch 
(referred to as a ``grid'') covers some region of space in its 
\emph{parent grid} which requires higher resolution, and can itself 
become the parent grid to an even more highly resolved \emph{child grid}.

The grid hierarchy begins with the root grid which covers the entire
domain of interest with a coarse, uniform, Cartesian grid. Then, as
the solution evolves and interesting regions form, finer meshes are
placed below these regions (we use the notation `below' to refer to
finer grids and `above' for coarser grids).  We restrict the ratio
between cell sizes to be an integer, typically 2-4, and refer to a
level as all the grids with the same cell size.  In order to keep
things as simple as possible, the edges of subgrids must coincide with
the cell edge of its immediate parent (coarser) grid. Additionally,
the hierarchy can be initialized with one or more static grids if a
higher initial resolution is required.

Given the hierarchy at some time $t$, we advance the solution in the
manner of a W-cycle in a multigrid solver.  First, we determine the
maximum time step allowed for the coarsest grid based on a variety of
accuracy and stability criteria and advance the grid by that time
interval, $\Delta t_0$.  We then move down to the next level and
advance all the grids on that level by a timestep $\Delta t_1$
($\Delta t_1 \leq \Delta t_0$) which is the minimum of all the allowed
timesteps for those grids.  If there are more levels, we repeat this
procedure until the bottom level of the hierarchy has been reached.
Once there, we continue advancing the grids on the lowest level until
they have `caught up' to the next highest level above (i.e. $\sum
\Delta t_l = \Delta t_{l-1}$).  This procedure repeats itself until
all grids have been advanced by a total time of $\Delta t_0$.

Since interesting regions on the grid may move, the hierarchy must
adapt itself.  We do this whenever a level has caught up to the
coarser level above it and consists of entirely rebuilding the grids
on that level and below.  This is done by applying the grid refinement
criterion to the grids on that level, flagging zones which require
extra grids.  This criterion depends on the physical problem being
simulated.  We have implemented a number of options, including shock
and steep gradient detectors, but in the problem described here,
employ one based on the mass within a cell, imitating the Lagrangian
nature of the SPH algorithm.  Once a grid has a set of flagged cells,
we run a machine-vision based algorithm \citep{Berger91} to find edges
and determine a good placement of subgrids.  These subgrids must not
overlap one another, must cover all flagged cells and their
neighbouring cells, and be above a preset efficiency threshold, where
the efficiency of a cell is defined as the ratio of flagged cells to
total cells.  Once these new subgrids have been identified, the
solution from the next coarser grid is interpolated (see below) in
order to initialize the values on the new grids.  Finally, any overlap
between these new subgrids and the old ones is identified and the
solution within the regions of overlap is copied to the new subgrids.
The entire procedure just outlined is then repeated on the new grids
and in this way the entire hierarchy (from the original level examined
and below) is rebuilt.

\subsubsection{Hydrodynamics: PPM method}

Two different hydrodynamic methods are implemented in \enzo: the
piecewise parabolic method (PPM) developed by~\citet{1984JCoPh..54..174C}
and extended to cosmology by~\citet{1995CoPhC..89..149B},
and the hydrodynamics method used in \zeus~\citep{Stone92a,Stone92b}.
We adopt the direct Eulerian PPM method (as opposed to the Lagrange-Remap varient).

PPM is an explicit, higher order-accurate version of
Godunov's method for ideal gas dynamics with third order-accurate piecewise parabolic
monotonic interpolation and a nonlinear Riemann solver for shock
capturing.  It advances the hydrodynamic equations in the following
steps:
\begin{enumerate}
 \item Construct monotonic parabolic interpolation of cell average
 data, for each fluid quantity.
 \item Compute interface states by averaging the parabola for over the domain of dependence for
 each interface
 \item Use interface data to solve the Riemann problem.
 \item Difference the interface fluxes to update the cell average quantities.
\end{enumerate}
It does an excellent job of capturing strong shocks in at
most two cells.  Multidimensional schemes are built up by directional
splitting and produce a method that is formally second order-accurate
in space and time which explicitly conserves mass, linear momentum, and energy.

As described in \citet{Bryan95}, we modify the method for use in hypersonic flows when the thermal energy $e$ is extremely small compared to the total energy $E$.  This is a problem because in total energy method, the temperature is computed by subtracting one large number from another (i.e. the kinetic energy from the total energy), a situation which is can generate large numerical inaccuracies.  We address this situation by also solving the thermal energy equation and using $e$ from this equation when we expect the error to be large.

The advantage of the PPM method is that it conserves energy better and is higher-order accurate.  The disadvantage is that it is generally less robust than the ZEUS method.

\subsubsection{Hydrodynamics: ZEUS method}

The second hydrodynamics method implemented is the finite-different algorithm used in the ZEUS, as described in \citet{Stone92a}.  For this reason, we call this the ZEUS method, although the code is entirely independent from the ZEUS code, and only the hydrodynamical algorithm of \zeus\ is implemented in \enzo -- the MHD and RHD codes are not.

The ZEUS method uses a staggered mesh such that the velocities are face-centered, while the density and thermal energy quantities are cell-centered.  It splits the solution up into two steps, the so-called source step, in which the momentum and energy values are updated to reflect the pressure and gravity forces, as well as the effect of an artificial viscosity required for stability.  The transport step accounts for the advection of conserved quantities (mass, momentum and energy) across the grid.

\subsubsection{Gravity}

The current implementation of self-gravity in \enzo uses a Fast Fourier Technique \citep{Hockney88} to solve Poisson's equation on the top (level 0) grid.  The advantage of this is that it naturally allows both periodic and isolated boundary conditions for the gravity, choices which are very common in astrophysics and cosmology.  On subgrids, we interpolate the boundary conditions from that grid's parent (either the level 0 grid or some other subgrid).  Poisson's equation is then solved using a multigrid technique.  In \enzo\, self-gravity is optional.  There are also a number of options for static gravitational fields.

\subsubsection{N-body Dynamics}

The collisionless matter is modelled through particles.  Since the
particles follow the collapse of structure by definition, they are not
adaptively refined.  Nor are there duplicate sets of particles for
each level; instead, each particle is associated with the most refined
level available at its position in space and moved as the hierarchy is
rebuilt.  Thus, a particle has the same timestep and feels the same
gravitational force as the grid at that level.

Although the particles are fixed in mass once initialized, we are free
to create them with any set of masses and positions.  For example, in
many cosmological simulations, a static subgrid is included from the
beginning in order to improve the initial baryonic mass resolution.
On this subgrid, we also use smaller particles to improve the
collisionless mass resolution.  One particle per initial grid point
seems to provide approximately equal sampling between the dark matter
and gas.

When solving Poisson's equation, we need to compute the contribution of 
the particles to the density on the grid.  This is done using a second-order
accurate Cloud-In-Cell technique \citep{Hockney88}.

\subsubsection{Chemistry, radiation and cooling}

Enzo may operate in a number of modes with regard to radiative cooling and chemistry.  In the simplest mode, with the multi-species flag turned off (i.e. set to zero), the cooling rate is computed from a simple temperature-dependent cooling rate, taken from \citet{SW87}.

In addition to this, \enzo\ includes the capability of following up to 12 chemical species using a non equilibrium-solver.  The species can be turned on in sets, with the simplest model including just H, H$^+$, He, He$^+$, He$^++$, and $e^-$, and more complete models adding first species important for gas-phase molecular hydrogen formation: $H^-$, H$_2$ and H$_2^+$, and then $HD$ formation: $HD$, $D$, $D+$.  The cooling and heating due to these species is also included.  The solution of the rate equations is carried out using a time-backwards differencing which insures stability; to maintain accuracy we sub-cycle the rate equations with a shorter timestep such that the electron and neutral fractions do not change by more than 10\%.

\subsubsection{Star Formation and Feedback}

A simple heuristic method is used to model the formation of stars and their feedback of metals and energy into the gas.  This method is based on \citep{CO1992} and involves identify plausible cites of star formation based on a set of criteria (dense gas with a short cooling time, which is both collapsing and unstable), and computing a star formation rate based on the Schmidt-Kennicutt relation \citep{K89}.  The relating gas is converted into a star particle over a few dynamical times, and over this timespan metals and thermal energy are injected into the single, local cell occupied by the star particle.

\subsubsection{Timestep constraints}

All grids on a given level are advanced with the same timestep.  To determine this timestep, we calculate, for each cell and for each physical process (except for the chemistry step, which is sub-cycled), the minimum $\Delta t$ allowed for stability.  Then the minimum of all timesteps is taken and the level is advanced with this step.

### 2012-06-29 parallel.tex
first commit of paper.

--- 2012-06-29 numerical-overview.tex
@@
-\section{Overview of Numerical Methods}\label{sec.overview}
+\section{Parallel Strategy and Performance}
 
-%---------------------------------------------------------------------------%
-\subsection{N-body dynamics}\label{sec.ov.nbody}
-fft + multigrid
-
-There are multiple methods to compute the gravitational potential 
-(which is an elliptic equation in the Newtonian limit) in a structured 
-AMR framework.  One way would be to model the dark matter (or other
-collisionless particle-like objects, such as stars) as a second fluid 
-in addition to the baryonic fluid and solve the collisionless
-Boltzmann equation, which follows the evolution of the fluid density in
-six-dimensional phase space.  However, this is computationally prohibitive 
-owing to the large dimensionality of the problem, making this approach 
-unappealing for the cosmological AMR code.
-
-Instead, \enzo\ uses a particle-mesh N-body method to calculate 
-the dynamics of collisionless systems \citep{Hockney88}.  This method 
-follows trajectories 
-of a representative sample of individual particles and is much more 
-efficient than a direct solution of the Boltzmann equation in most 
-astrophysical situations. 
-The gravitational potential is computed by solving the elliptic 
-Poisson's equation (Eqtn.~\ref{enzopoisson}).
-
-These equations are finite-differenced and for simplicity are solved with the
-same timestep as the hydrodynamic equations, as discussed in 
-Section~\ref{sec.ov.amr}.
-The dark matter particles are distributed onto the grids
-using the cloud-in-cell (CIC) interpolation technique to form a spatially
-discretized density field (analogous to the baryon densities used to calculate
-the equations of hydrodynamics).  After sampling dark matter density onto the
-grid and adding baryon density if it exists (to get the total matter density
-in each cell), the gravitational potential is calculated on the periodic root
-grid using a fast Fourier transform.  In order to calculate more accurate
-potentials on subgrids, \enzo\ re-samples the dark matter density onto
-individual subgrids using the same CIC method as on the root grid, but at
-higher spatial resolution (and again adding baryon densities if applicable).
-Boundary conditions are then interpolated from the potential values on the
-parent grid (with adjacent grid patches on a given level communicating to
-ensure that their boundary values are the same), and then a multigrid
-relaxation technique is used to calculate the gravitational potential at every
-point within a subgrid.  Forces are computed on the mesh by
-finite-differencing potential values and are then interpolated to the particle
-positions, where they are used to update the particle's position and velocity
-information.  Potentials on child grids are computed recursively and particle
-positions are updated using the same timestep as in the hydrodynamic
-equations.  Particles are stored in the most highly refined grid patch at the
-point in space where they exist, and particles which move out of a subgrid
-patch are sent to the grid patch covering the adjacent volume with the finest
-spatial resolution, which may be of the same spatial resolution, coarser, or finer
-than the grid patch that the particles are moved from.  This takes place in a
-communication process at the end of each timestep on a level.
-
-At this point it is useful to emphasize that the effective force
-resolution of an adaptive particle-mesh calculation is approximately twice
-as coarse as the grid spacing at a given level of resolution.  The 
-potential is solved in each grid cell;
-however, the quantity of interest, namely the acceleration, is the gradient
-of the potential, and hence two potential values are required to calculate
-this.  In addition, it should be noted that the adaptive particle-mesh
-technique described here is very memory intensive: in order to ensure reasonably 
-accurate force resolution at grid edges the multigrid relaxation
-method used in the code requires a layer of ghost zones which is very deep --
-typically 6 cells in every direction around the edge of a grid patch.  This greatly
-adds to the memory requirements of the simulation, particularly because subgrids
-are typically small (on the order of $12^3 - 16^3$ real cells for a standard 
-cosmological calculation) and ghost zones can dominate the memory and computational 
-requirements of the code.
-
-%---------------------------------------------------------------------------%
-\subsection{Hydrodynamics}\label{sec.ov.hydro}
-
-\subsubsection{The PPM hydro method}\label{sec.ov.hydro.ppm}
-
-\dcc{Removed ``things to do'' list.  Mention ppm-lr and ppm-de, and
-  that ppm-de is preferred.  Briefly describe ppm-de itself.}
-
-The primary hydrodynamic method used in \enzo\ is based on the
-piecewise parabolic method (PPM) of
-\citet{1984JCoPh..54..174C}, which has been
-modified for the study of cosmological fluid flows.  
-\citet{1984JCoPh..54..174C} describe two variants on this method: a
-Lagrangian Remap variant and a Direct Eulerian variant.  Both have
-been adapted for \enzo, and the Lagrangian Remap 
-variant is described in \citet{1995CoPhC..89..149B}.  Because of its
-conservative formulation, however, the Direct Eulerian variant is
-preferred for \enzo\ AMR simulations.  We will describe the Direct
-Eulerian (PPM-DE) variant and its additions briefly here, and in
-further depth in Appendix \ref{app.ppm}
-
-PPM is an explicit, higher order-accurate version of
-Godunov's method for ideal gas dynamics with third order-accurate piecewise parabolic
-monotonic interpolation and a nonlinear Riemann solver for shock
-capturing.  It advances the hydrodynamic equations in the following
-steps:
-\begin{enumerate}
- \item Construct monotonic parabolic interpolation of cell average
- data, for each fluid quantity.
- \item Compute interface states by averaging the parabola for over the domain of dependence for
- each interface
- \item Use interface data to solve the Riemann problem.
- \item Difference the interface fluxes to update the cell average quantities.
-\end{enumerate}
-It does an excellent job of capturing strong shocks in at
-most two cells.  Multidimensional schemes are built up by directional
-splitting and produce a method that is formally second order-accurate
-in space and time which explicitly conserves mass, linear momentum, and energy.
-
-To use the above algorithm for cosmological simulations, the
-conservation laws for fluid mass, momentum and energy 
-density are written in comoving coordinates for a
-Friedman-Robertson-Walker space-time, as described previously 
-in Equations~\ref{enzoconserve} 
-through~\ref{enzoenergy}.  Both the conservation laws and
-the Riemann solver are modified to include gravity, which is solved
-using an adaptive particle-mesh (PM) technique (see Section~\ref{sec.ov.nbody}).  
-The terms due to cosmological expansion, as well as primordial chemistry and
-radiative heating and cooling, are solved in a separate step because they have
-different numerical behavior, and therefore must be treated differently to ensure
-stability.  Note that unlike the \zeus\ hydro scheme, PPM does not need to use 
-artificial viscosity to resolve shocks.
-
-The system of equations described above works well for systems with
-relatively low Mach numbers, as long as these systems are well resolved.  
-However, cosmology is replete with situations where there are bulk hypersonic
-flows.  In these situations, the ratio of kinetic to thermal energy can be
-very high -- in some situations up to $10^6 - 10^8$.  This implies that the
-thermal energy is an extremely tiny fraction of the kinetic energy, which can 
-cause numerical problems when one is interested in just the thermal energy of the gas, since
- Equation~\ref{enzoenergy} solves for the total energy.  In this system of
-equations, the
- thermal energy $E_{therm}$ is calculated as $E - E_{kin}$, 
-where E is the total specific energy as
-calculated in equation~\ref{enzoenergy} and $E_{kin}$ is the specific kinetic energy,
-$0.5 \vec{v}_b^2$.  
-In hypersonic flows $E$ and $E_{kin}$ are nearly the same, and any 
-number calculated as 
-the difference of these is going to be strongly affected by numerical error.  To 
-avoid this problem, \enzo\ also solves the internal energy equation in
-comoving coordinates:
-
-\begin{equation}
-\frac{\partial e}{\partial t} + \frac{1}{a} \vec{v}_b \cdot \nabla e = 
-- \frac{3(\gamma-1)\dot{a}}{a}e - \frac{p}{a \rho} \nabla \cdot \vec{v}_b
-\label{enzointenergy}
-\end{equation}
-
-In this equation $e$ is the internal energy (defined as $P/\rho (\gamma-1)$ and the other terms are as described
-previously.  The code still conserves total energy ($E$) as well.  In order
-to maintain consistency, both equations are solved at all times in all cells,
-with the equation for the total energy (eqtn.~\ref{enzoenergy}) being used
-for hydrodynamics routines and the internal energy $e$ being used when 
-temperature is required.  When pressure is required for dynamic purposes,
-the total energy is used if the ratio of thermal energy to total energy
-is less than some threshold value $\eta$, and the internal energy is used
-for values of the ratio larger than $\eta$.  A typical value of this parameter
-is $10^{-3}$.  This \emph{dual energy formulation} ensures that the
-method produces the correct entropy jump at strong shocks and also
-yields accurate pressures and temperatures in cosmological hypersonic
-flows.
-
-\subsubsection{The Zeus hydro method}\label{sec.ov.hydro.zeus}
-
-As a check on PPM, \enzo\ also includes an implementation of the
-finite-difference hydrodynamic algorithm employed in the compressible 
-magnetohydrodynamics code `\zeus'
-\citep{Stone92a, Stone92b}.  Fluid transport is solved on a Cartesian
-grid using the upwind, monotonic advection scheme of \citet{1977JCoPh..23..276V} 
-within a multistep 
-(operator split) solution procedure which is fully explicit in time.  
-This method is formally second order-accurate in space but 
-first order-accurate in time.  It is important to note that 
-this method conserves internal energy
-rather than total energy, so the ``dual energy
-formulation'' discussed in Section~\ref{sec.ov.hydro.ppm} is unnecessary.
-
-Operator split methods break
-the solution of the hydrodynamic equations into parts, with each part
-representing a single term in the equations.  Each part is evaluated
-successively using the results preceding it.  The individual parts of
-the solution are grouped into two steps, called the \emph{source} and
-\emph{transport} steps.
-
-The \zeus\ method uses a von Neumann-Richtmeyer artificial viscosity 
-to smooth shock discontinuities that may
-appear in fluid flows and can cause a break-down of finite-difference
-equations.  The artificial viscosity term is added in the source terms
-as
-\begin{eqnarray}
-\rho \frac{\partial\textbf{v}}{\partial t} &=& - \nabla p - \rho \nabla \phi 
-- \nabla \cdot \textbf{Q} \\
-\frac{\partial e}{\partial t} &=& -p \nabla \cdot \textbf{v} - \textbf{Q} : \nabla \textbf{v}, 
-\end{eqnarray}
-where \textbf{v} is the baryon velocity, $\rho$ is the mass density,
-$p$ is the pressure, $e$ is the internal energy density of the gas and \textbf{Q}
-is the artificial viscosity stress tensor, such that:
-
-\begin{eqnarray}
-Q_{ii} = \left\{ \begin{array}{ll}
-Q_{\rm AV} \rho_b  (a \Delta v_{i} + \dot{a} \Delta x_i)^2, 
-& \textrm{for $a \Delta v_{i} + \dot{a} \Delta x_i < 0$}  \\
-0 & \textrm{otherwise}\\
-\end{array} \right. 
-\end{eqnarray}
-and
-\begin{equation}
-Q_{ij} = 0  \;\;\;{\rm for}\;\; i \ne j .
-\end{equation}
-
-$\Delta x_i$ and $\Delta v_{i}$ refer to the comoving 
-width of the grid cell
-along the $i$-th axis and the corresponding difference in gas
-peculiar velocities across the grid cell, respectively, and $a$ is the
-cosmological scale factor.  $Q_{\rm AV}$ is a constant with a typical
-value of 2. We refer the interested reader to \citet{1994ApJ...429..434A} for more details.
-
-The limitation of a technique that uses an artificial viscosity is that,
-while the correct Rankine-Hugoniot jump conditions are achieved,
-shocks are broadened over 6-8 mesh cells and are thus not treated as true
-discontinuities. This may cause unphysical pre-heating of gas upstream
-of the shock wave, as discussed in \citet{1994ApJ...429..434A}.
-
-%---------------------------------------------------------------------------%
-\subsection{Chemistry, radiation and cooling}\label{sec.ov.chem}
-
-\red{
-As this stands it is the methods that I (brian) use, so I don't really emphasize
-either the tabular cooling functions or the various radiation background options.
-Add more discussion of this, point towards appendix.
-}
-
-Though the equations of hydrodynamics described above are a closed system,
-they are still missing a crucial piece of physics: radiative heating and cooling.
-Radiative cooling is extremely important in many astrophysical situations, as is 
-the heating of gas from some sort of radiation background.  \enzo\ has a very simple
-Sutherland and Dopita equilibrium cooling function \citep{1993ApJS...88..253S} 
-implemented, which uses a
-cooling table assuming a fixed metallicity of $Z = 0.3 Z_\odot$, and also a
-nonequilibrium heating/cooling model that assumes gas of primordial composition exposed to a
-uniform metagalactic ultraviolet (UV) background that varies with time \citep{1996ApJS..105...19K}.
-
-The simulations discussed in this thesis almost exclusively use the nonequilibrium
-routines, described in great detail by Abel et al. and Anninos et al.  
-\citep{abel97,anninos97} and summarized in Appendix~\ref{app:chemistry}. These 
-routines follow the non-equilibrium chemistry of a 
-gas of primordial composition with 9 total species:  
-$H$, $H^+$, $He$, $He^+$, $He^{++}$, $H^-$, $H_2^+$, $H_2$, and $e^-$.  The code also calculates 
-radiative heating and cooling, following atomic line excitation, recombination,
-collisional excitation, free-free transitions, molecular line excitations, and Compton
-scattering of the cosmic microwave background, as well as any of
-approximately a dozen different models for a metagalactic UV background that heat
-the gas via photoionization and photodissociation. 
-The multispecies rate equations are solved out of
-equilibrium to properly model situations where, e.g., the cooling rate of the gas
-is much shorter than the hydrogen recombination time.  The effect of this nonequilibrium
-cooling is to leave behind a much larger fraction of residual free electrons than one would
-expect if the assumption of equilibrium were being made.  The practical effect of this
-is that more $H^-$ is formed, which subsequently produces hydrogen molecules.  If large
-amounts of $H_2$ is formed it can greatly increase the cooling rate of primordial gas
-at relatively low temperatures ($T \leq 10^4$ K).  This can efficiently cool the gas 
-to approximately $200$ K, which significantly reduces the Jeans mass of the gas.  Correct
-modeling of the formation of molecular hydrogen is crucial to the study of star formation
-in a primordial gas.
-
-A total of 9 kinetic equations are solved, including 29 kinetic and radiative 
-processes, for the 9 species mentioned above.  See Table~\ref{table.collisional} 
-for the collisional processes and Table~\ref{table.radiative} for the 
-radiative processes solved.
-
-%---------------- table of collisional processes
-
-\begin{table}
-\begin{center}
-{\bfseries Collisional Processes}\\[1ex]
-\begin{tabular}{llllllll}
-(1) & H & + & e$^-$ & $\rightarrow$ & H$^+$ &+& 2e$^-$ \\
-(2) & H$^+$ &+ &e$^-$ & $\rightarrow$ & H &+ &$\gamma$ \\
-(3) & He &+& e$^-$ & $\rightarrow$ & He$^+$ &+& 2e$^-$  \\
-(4) & He$^+$ &+& e$^-$ & $\rightarrow$ & He &+ &$\gamma$  \\
-(5) & He$^{+}$ &+& e$^-$ & $\rightarrow$ & He$^{++}$ &+& 2$e^-$  \\
-(6) & He$^{++}$ &+& e$^-$ & $\rightarrow$ & He$^+$ &+& $\gamma$ \\
-(7) & H &+& e$^-$ &$\rightarrow$& H$^-$ &+& $\gamma$  \\
-(8) & H$^-$ &+& H &$\rightarrow$ & H$_2$ & +& e$^-$ \\
-(9) & H &+ &H$^+$ &$\rightarrow$ &H$_2^+$ &+ &$\gamma$ \\
-(10) & H$_2^+$ &+ &H &$\rightarrow$ &$H_2$ &+ &$H^+$ \\
-(11) & H$_2$ &+ &H$^+$ &$\rightarrow$ &H$_2^+$ & +& H \\
-(12) & H$_2$ &+ &e$^-$ & $\rightarrow$ & 2H & + & e$^-$  \\
-(13) & H$_2$ & + & H & $\rightarrow$ & 3H &   &      \\
-(14) & H$^-$ & + & e$^-$ & $\rightarrow$ & H & + & 2e$^-$ \\
-(15) & H$^-$ & + & H & $\rightarrow$ & 2H & + & e$^-$ \\ 
-(16) & H$^-$ & + & H$^+$ & $\rightarrow$ & 2H & & \\
-(17) & H$^-$ & + & H$^+$ & $\rightarrow$ & H$_2^+$ & + & e$^-$ \\
-(18) & H$_2^+$ & + & e$^-$ & $\rightarrow$ & 2H & & \\
-(19) & H$_2^+$ & + & H$^-$ & $\rightarrow$ & H$_2$ & + & H  \\
-(20) & 3H & & & $\rightarrow$ & H$_2$ & + & H
-\end{tabular}
-\caption[]{Collisional processes solved in the Enzo nonequilibrium
-primordial chemistry routines.}
-\label{table.collisional}
-\end{center}
-\end{table}
-
-\begin{table}
-\begin{center}
-{\bfseries Radiative Processes}\\[1ex]
-\begin{tabular}{llllllll}
-(21) & H & + & $\gamma$ & $\rightarrow$ & H$^+$ & + & e$^-$ \\
-(22) & He & + & $\gamma$ & $\rightarrow$ & He$^+$ & + & e$^-$ \\
-(23) & He$^+$ & + & $\gamma$ & $\rightarrow$ & He$^{++}$ & + & e$^-$ \\
-(24) & H$^-$ & + & $\gamma$ & $\rightarrow$ & H & + & e$^-$ \\
-(25) & H$_2$ & + & $\gamma$ & $\rightarrow$ & H$_2^+$ & + & e$^-$ \\
-(26) & H$_2^+$ & + & $\gamma$ & $\rightarrow$ & H & + & H$^+$ \\
-(27) & H$_2^+$ & + & $\gamma$ & $\rightarrow$ & 2H$^+$ & + & e$^-$ \\
-(28) & H$_2$ & + & $\gamma$ & $\rightarrow$ & H$_2^*$ & $\rightarrow$ & 2H \\
-(29) & H$_2$ & + & $\gamma$ & $\rightarrow$ & 2H &  & 
-\end{tabular}
-\caption[]{Radiative processes solved in the Enzo nonequilibrium
-primordial chemistry routines.}
-\label{table.radiative}
-\end{center}
-\end{table}
-
-The chemical reaction equation network is technically challenging to solve due to 
-the huge range of reaction timescales involved -- the characteristic creation
-and destruction timescales of the various species and reactions can differ by 
-many orders of magnitude.  As a result, the set of rate equations is extremely 
-stiff, and an explicit scheme for integration of the rate equations can be 
-exceptionally costly if small enough timesteps are taken to keep the network
-stable.  This makes an implicit scheme much more preferable for such a set of 
-equations.  However, an implicit scheme typically require an iterative 
-procedure to converge, and for large networks (such as this one) an implicit
-method can be very time-consuming, making it undesirable for a large, three-dimensional
-simulation.
-
-\enzo\ solves the rate equations using a method based on a backwards differencing 
-formula (BDF) in order to provide a stable and accurate solution.  This technique is
-optimized by taking the chemical intermediaries $H^-$ and $H_2^+$, which have 
-large rate coefficients and low concentrations, and grouping them into a separate
-category of equations.  Due to their fast reactions these species are very sensitive
-to small changes in the more abundant species.  However, due to their low overall
-concentrations, they do not significantly influence the abundance of species with
-higher concentrations.  Therefore, reactions involving these two species can be
-decoupled from the rest of the network and treated independently.  In fact, $H^-$ 
-and $H_2^+$ are treated as being in equilibrium at all times, independent of 
-the other species and the hydrodynamic variables.  This allows a large speedup
-in solution as the BDF scheme is then applied only to the slower 7-species network
-on timescales closer to those required by the hydrodynamics of the simulation.
-Even so, the accuracy and stability of the scheme is maintained by subcycling the 
-rate solver over a single hydrodynamic timestep.  These subcycle timesteps are 
-determined so that the maximum fractional change in the electron concentration is
-limited to no more than $10\%$ per timestep.
-
-It is important to note the regime in which this model is valid.  According to Abel et al. and
-Anninos et al. \citep{abel97,anninos97}, the reaction network is valid for temperatures
-between $10^0 - 10^8$ K.  The original model discussed in these two references is only
-valid up to $n_H \sim 10^4$~cm$^{-3}$.  However, addition of the 3-body $H_2$ formation
-process (equation 20 in Table~\ref{table.collisional}) allows 
-correct modeling of the chemistry of the gas up
-until the point where collisionally-induced emission from molecular hydrogen becomes an important
-cooling process, which occurs at $n_H \sim 10^{14}$~cm$^{-3}$.  A further concern is that
-the optically thin approximation for radiative cooling breaks down, which
-occurs before $n_H \sim 10^{16} - 10^{17}$~cm$^{-3}$.  Beyond this point, 
-modifications the cooling function that take into account the non-negligible
-opacity in the gas must be made, as discussed by \citet{2004MNRAS.348.1019R}. 
-Even with these modifications, a completely correct description of the cooling of
-this gas will require some form of radiation transport, which will greatly 
-increase the cost of the simulations.
-
-Several processes are neglected.  The deuterium atom and its processes are 
-completely ignored, which may have some effect.  Recent work shows that HD is
-a more effective coolant than previously thought~\citep{2005MNRAS.361..850L}.  However,
-the fractional abundance of HD is so low that under circumstances relevant to
-the formation of a Population III star in an un-ionized region it should be
-sub-dominant.  However, the enhanced electron fraction in fossil 
-HII regions (as discussed later in this thesis) could result in the HD molecule
-becoming a dominant cooling mechanism at relatively low ($\sim$ few hundred K)
-temperatures, and could potentially cool the gas down to below $100$~K, which can
-enhance fragmentation and could have important consequences for the IMF of 
-primordial stars forming in a relic HII region~\citep{2005MNRAS.364.1378N}.
-
-Aside from deuterium, the chemical reactions involving lithium 
-are also neglected.  According to \citet{1998A&A...335..403G}, these are 
-not important for the density and temperature regimes explored 
-by the simulations discussed in this thesis.  However, at higher densities it 
-is possible that there are regimes where lithium can be an important coolant.
-
-%---------------------------------------------------------------------------%
-\subsection{Star formation and feedback}\label{sec.ov.star}
-
-\red{By the time this paper is submitted my (BWO's) paper on the star formation
-and feedback algorithms, and the various parameters contained therein, should be
-at least written and submitted as well.  Point to this, and maybe include a couple
-interesting plots (perhaps in an appendix?).}
-
-While the physics discussed previously is all crucial to the study of
-cosmological structure formation, most cosmological observations are of
-stars and related phenomena.  Additionally, stars eject energy and metal-
-enriched gas throughout their lives, drastically modifying their own 
-environment.  The formation of galaxies and clusters cannot be completely
-modeled without including the feedback of energy and metals.  In particular,
-it is thought that feedback is crucial for suppressing the large numbers of
-dwarf-like galaxies that CDM theories predict 
-\citep{1991ApJ...381...14L,1991ApJ...379...52W}.  An early
-burst of star formation could remove a large fraction of cold gas from such systems
-\citep{1978MNRAS.183..341W,1991ApJ...367...45C}.  Also, the unexpectedly 
-low luminosities of small clusters and
-groups (relative to the richest clusters) is commonly explained through feedback
-\citep{1991ApJ...383..104K}.  Energy ejected during the formation of the cluster ellipticals 
-lowers the central density and hence the X-ray luminosity of such clusters 
- \citep{1997ApJ...484L..21C}.  Therefore, the inclusion of star formation and the 
-feedback of energy and metals into the intergalactic medium in a cosmological
-code is crucial for many reasons.
-
-We have extended the \enzo\ code to include multiple models for star formation
-and feedback.  It is difficult to directly simulate the formation of individual
-stars in the context of simulations of galaxy formation and evolution due
-to the immense computational cost.  Therefore, 
-we adopt parametric methods which attempt
-to model star forming regions.  One model is based on Kravtsov's method
- \citep{2003ApJ...590L...1K}, and the other is based on the method of 
-\citet{1992ApJ...399L.113C}, which has been modified for use in an AMR code.  The basic
-ideas behind the methods are straightforward and observationally motivated.
-Similar (though somewhat more advanced) methods have been employed for the 
-smoothed particle hydrodynamics
-(SPH) method, most recently by Springel \& Hernquist 
-\citep{sh03a,sh03b,hs03}. This 
-method assumes a
-multiphase IGM and has been shown to accurately reproduce the cosmic star
-formation rate, and will be implemented into \enzo\ in the near future.
-
-In the following sections we will detail both the Kravtsov and Cen \& Ostriker
-star formation methods separately.  Tests of the star formation and
-feedback algorithms will not be shown, since this work is currently in
-progress (and an extension of the AMR/SPH code comparison described in
-\citet{2005ApJS..160....1O}).
-
-\subsubsection{The Kravtsov star formation and feedback algorithm}
-
-Kravtsov's method of star formation is designed to reproduce
-the global Schmidt law of star 
-formation~\citep{2003ApJ...590L...1K,1959ApJ...129..243S}.
-This algorithm is deliberately minimal, and is explicitly geared
-towards modeling star formation in a phenomenological way on kiloparsec
-scales.  Stars are assume to form with a characteristic gas timescale 
-$\tau_*$ such that $\dot{\rho}_* = \rho_{gas}/\tau_*$ where $\rho_{gas}$ 
-and $\rho_*$ are the baryon gas and stellar densities, respectively.  
-This ``constant efficiency'' model on the scale of star formation 
-regions is well motivated observationally \citep{1996AJ....112.1903Y,2002ApJ...569..157W}.  
-Star formation is only allowed to take place in very dense regions
-with $\rho_{gas} \geq \rho_{SF}$, where $\rho_{SF}$ is a constant
-proper density threshold above which star formation is allowed to
-occur.  No other criteria are imposed.  Kravtsov's typical choices
-for $\tau_*$ and $\rho_{SF}$ are $\tau_* = 4$ Gyr and $\rho_{SF} =
-1.64~$M$_\odot$~pc$^{-3}$~$(n_H \sim 50$~cm$^{-3})$.  The adopted timescale
-is derived from the observationally-determined normalization of the
-Schmidt law, and the density threshold is determined by observations
-of star forming regions on $\sim 100$ pc scales.  Note that the density
-threshold is in proper, not comoving, units.
-
-Algorithmically, the star formation events in Kravtsov's code
-are assumed to occur once every global time step 
-$\Delta t_0 \leq 10^7$.  In cells where star formation is
-determined to occur (i.e. $\rho_{gas} \geq \rho_{SF}$) a
-collisionless ``star particle'' is assumed to form, with a 
-mass $m_* = \dot{\rho}_* V_{cell} \Delta t_0$, where
-$\dot{\rho}_*$ is described above and $V_{cell}$ is the 
-volume of the mesh cell.  These star particles are
-formed instantaneously, with all $m_*$ of gas being
-taken out of the cell and immediately deposited into the
-star particle.  This particle is then given the velocity
-of the gas in the cell, and thereafter treated as a 
-collisionless particle.  The \enzo\ implementation
-of this algorithm is similar, except that instead of 
-forming stars only at the root grid time step, we allow 
-stars to form at the time step of the highest level of
-resolution at any particular point in space.  As can be seen
-from the equation for $m_*$ above, this can result in very 
-small stellar masses.  To avoid memory and processor time
-issues related to having very large numbers of star particles
-we impose a threshold mass $M_{*,min}$ such that a star
-particle only forms if $m_* \geq M_{*,min}$.  An
-appropriate choice of $M_{*,min}$ does not significantly 
-change the star overall star formation history of a 
-simulation, though it may delay the onset of star formation 
-in a given cell relative to a simulation without a particle mass
-threshold.
-
-Each ``star particle'' is assumed to represent an ensemble
-of stars and is treated as a single-age stellar population.
-Kravtsov assumes that the IMF is described by a
-Miller \& Scalo functional form with stellar masses between
-$0.1$ and $100~M_\odot$ \citep{1979ApJS...41..513M}.  All stars in this
-IMF with $M_* > 8~M_\odot$ deposit $2 \times 10^{51}$ ergs of
-thermal energy and a mass $f_z M_*$ of metals into the
-cell in which they form without delay, with
-$f_z \equiv min(0.2, 0.01~M_*-0.06)$ (i.e. instantaneous
-deposition of metals).  The definition of $f_z$ is a rough
-approximation of the results of 
-\citet{1995ApJS..101..181W}.
-
-Kravtsov reports that simulations with this algorithm
-reliably reproduce the star formation rate-gas surface
-density relation of the Schmidt law, and are not particularly
-sensitive to numerical parameters \citep{2003ApJ...590L...1K}.
-He also notes that this is surprisingly insensitive to
-the presence or absence of feedback and details of the cooling
-and heating properties of the gas, which suggests that the
-global star formation rate is determined by gravitationally-
-driven supersonic turbulence (on large scales)
-rather than stellar feedback or thermal instabilities on small 
-scales.
-
-\subsubsection{The Cen \& Ostriker star formation algorithm}
-
-The Cen \& Ostriker method is a heuristic model of star formation
-on galactic scales.  This method, first described by
-\citet{1992ApJ...399L.113C}, is similar in some ways to the Kravtsov algorithm
-but has more complex criteria for determining where stars should
-be formed.  In this method, cells that form stars must have a 
-baryon overdensity higher than some threshold 
-$\rho_b/\bar{\rho}_b \geq \eta$ 
-where $\rho_b$ is the baryon density in that cell, 
-$\bar{\rho_b}$ is the mean baryon density in the simulation,
-and $\eta$ is the user-defined overdensity threshold.
-Additionally, the gas in the cells must be contracting, 
-cooling rapidly, and gravitationally unstable, e.g.:
-
-\begin{equation}
-\nabla \cdot \vec{v}_b < 0,
-\label{cencont}
-\end{equation}
-
-\begin{equation}
-t_{cool} < t_{dyn} \equiv \sqrt{3 \pi / 32G \rho_{tot}},
-\end{equation}
-
-\begin{equation}
-m_{b} > m_{J} \equiv G^{-3/2} \rho_{b}^{-1/2}C^{3}
-\left[ 1 + \frac{\delta\rho_{d}}{\delta\rho_{b}} \right]^{-3/2}
-\end{equation}
-
-where $\vec{v}$ is the velocity of the gas in the cell, $\rho_{b}$ and 
-$\rho_{d}$ are the cell's baryon and dark matter density, respectively,
-$\rho_{total} = \rho_{b} + \rho_{d}$, $m_{b}$ and $m_{j}$ are the 
-baryonic mass in the cell and jeans mass of the cell, and C is the 
-isothermal sound speed in the cell.  
-If all of these criteria are met, 
-the mass of a star particle is calculated as \(m_{*} = m_{b} 
-\frac{ \Delta t}{ t_{dyn} } f_{*eff} \), 
-where $f_{*eff}$ is the star formation efficiency parameter.
-
-If $m_{*}$ is greater than a minimum star mass $m_{*min}$, a particle 
-is created and given several attributes:  Mass, a unique index number, 
-the time of formation $t_{form}$, the local dynamical free-fall time 
-$t_{dyn}$ and the metallicity fraction of the baryon gas in the cell 
-$f_{Zb}$.  There is a user-defined minimum dynamical time
-$T_{dyn,min}$ which is observationally motivated and affects the
-feedback rates (see below).
-The particle is placed in the center of the cell and given 
-the same peculiar velocity as the gas in the cell, and is then treated 
-in the same manner as the dark matter particles.  An 
-amount of baryon gas corresponding to the new particle's mass is 
-removed from the cell.
-
-In addition, we have added a stochastic star formation algorithm that keeps 
-track of all of the sub-mass threshold stars that should have been created 
-and when the total mass of uncreated stars is greater than the minimum mass, 
-a star particle is created.
-
-The feedback of energy and metals into the baryon gas is similar to
-the Kravtsov feedback, with some important differences.  The star formation 
-algorithm creates each star particle instantaneously.  However, feedback
-should take place over a significant timescale, as all of the stars contained
-within the ``star particle'' would form over a long period of time.  Therefore,
-we assume that for the purposes of feedback that the mass of stars formed
-at a time $t$ with finite timestep $\Delta t$ is:
-
-\begin{eqnarray}  
-\Delta m_{sf} =  m_{*} \left[ \right. \left(1 + \frac{t-t_{form}}{t_{dyn}} 
-                    \right) \nonumber \\ 
-		   \times exp \left( \frac{-(t-t_{form})}{t_{dyn}} \right) 
-                    \nonumber \\ 
-	       - \left( 1 + \frac{t+ \Delta t-t_{form}}{t_{dyn}} \right) \nonumber \\ 
-                    exp \left( \frac{-(t+\Delta t-t_{form})}{t_{dyn}} \right) 
-                   \left. \right]
-\end{eqnarray}
-
-which can be represented more clearly in integral form:
-
-\begin{eqnarray}
-\int_{t}^{t+Dt} \frac{dM}{dt} dt = \int_{t}^{t+Dt} m_{*} 
-\frac{dt}{t_{tyn}} \left( \frac{t-t_{form}}{t_{dyn}} \right) \nonumber \\ 
-\times exp \left( - ~ \frac{t-t_{form}}{t_{dyn}} \right) 
-\end{eqnarray}
-
-During this timestep, we assume that the star particle feeds back
-metal-enriched gas and thermal energy from supernovae and from stellar winds.
-Since massive stars have very short lifetimes, we assume that there is an
-immediate feedback of some fraction $f_{SN}$ of the rest energy from the
-stars created that timestep into the baryon gas, such that 
-$E_{add} = f_{SN} \Delta m_{sf} c^2$, where c is the speed of light.
-In addition, a fraction $f_{Z*}$ of the metal from the star particle
-is fed back into the baryon gas, which takes into account the effects
-of metal recycling.  Finally, a fraction of the mass $f_{m*}$ from all
-stars (rather than just supernovae) is fed back into the gas along with
-momentum in order to simulate the mass ejection from non-exploding stars
-via stellar winds.
-
-There are six user-defined parameters in this algorithm:  three deal with the
-star formation ($\eta$, $m_{*min}$ and $T_{dyn,min}$), and three 
-deal with feedback ($f_{SN}$, $f_{Z*}$ and $f_{m*}$).  Some of these
-parameters are completely free, while others can be guided by observation
-or theory.  For example, the supernova feedback parameter, $f_{SN}$, can be 
-constrained assuming that, for every $200 M_\odot$ of stars created, one
-supernova occurs, and this event feeds back approximately $10^{51}$ ergs of
-thermal energy, giving:
-
-\begin{equation}
-f_{SN} = \frac{10^{51} ergs}{200~M_\odot c^2} \simeq 3 \times 10^{-6}
-\end{equation}
-
-The metal yield $f_{Z*}$, defined as the mass in metals produced per unit
-mass of stars created, can be constrained by a theoretical model of
- \citet{1995ApJS..101..181W}.  This model suggests that $f_{Z*} = 0.02$
-is an appropriate number.  The minimum dynamical time is set to be $T_{dyn,min} = 10^7$
-~years to agree with timescales seen in nearby OB associations.
-
-The other parameters, such as the overdensity threshold $\eta$, minimum star
- mass $m_{*min}$, and mass ejection fraction $f_{m*}$ are not well constrained 
-either theoretically or observationally.  Indeed, $m_{*min}$ is a purely
-numerical parameter designed to keep the code from producing too many star 
-particles, and thus has no observational or theoretical counterpart.  These
-parameters have to be set by performing parameter studies and comparing to 
-observations.  Unfortunately, the range of parameter space is large, and
-the results may be degenerate for some combinations of these parameters.
-
-%---------------------------------------------------------------------------%
-\subsection{Adaptive mesh refinement}\label{sec.ov.amr}
-\dcc{removed most of the notes.  Added intro blurb.  moved parallelism
-  stuff. Re-worded the bit on ghost zones.  Summary of 'grid as
-  independent update + syncing with neighbors and parents,' W-cycle. Added
-  sentence on subgrid creation.  Made timestepping its own subsection.}
-\red{Talk about reasonable refinement factors - which 
-ones are usable, etc?}
-
-The primary purpose of the \enzo\ code is its Adaptive Mesh Refinement
-capability, which allows it to reach extremely large dynamical ranges
-with limited computational resources, opening doors previously closed
-by finite memory and computational time. Unlike moving mesh methods
-\citep{1995ApJS..100..269P,1995ApJS...97..231G} or  
-methods that subdivide 
-individual cells \citep{Adjerid}, Berger \& Collela's AMR (also referred 
-to as \emph{structured} AMR) utilizes an adaptive hierarchy of grid 
-patches at varying levels of resolution.  Each rectangular grid patch 
-(referred to as a ``grid'') covers some region of space in its 
-\emph{parent grid} which requires higher resolution, and can itself 
-become the parent grid to an even more highly resolved \emph{child grid}. 
-ENZO's implementation of structured AMR places no restrictions on 
-the number of grids at a given level of refinement, or on the number of 
-levels of refinement.  However, owing to limited computational resources 
-it is practical to institute a maximum level of refinement $\Lmax$.
-
-The \enzo\ implementation of AMR allows arbitrary integer ratios of parent and
-child grid resolution.  However, in practice 
-refinement factors (defined as the ratio of parent grid resolution to child grid resolution) 
-of two or four are typically used, with a refinement factor
-of two being most commonly used for cosmological simulations for reasons of
-efficiency. The ratio of
-box size to the maximum grid resolution of a given simulation is
-therefore $L/e = \Nroot \times 2^{\Lmax}$, where $\Nroot$ is the
-number of cells along one edge of the root grid, $\Lmax$ is the
-maximum level of refinement allowed, and L and e are the box size
-and grid resolution of the most highly refined region, respectively.
-
-The AMR grid patches are the primary data structure in \enzo.  Each
-patch is treated as an individual object which can contain both field
-variables and particle data.  Individual grids are organized into a 
-dynamic, distributed hierarchy of mesh patches.  
-See 
-Figure~\ref{fig.2.amrhierdens} for an example of an \enzo\
-simulation performed using this AMR scheme where both the baryon density
-and AMR grid hierarchy are shown together.
-
-\begin{figure}
-\begin{center}
-\includegraphics[width=0.4\textwidth]{figures/amr_hierarchy.eps}
-\end{center}
-\caption{\emph{Left:}  Example of a simple, 
-distributed AMR hierarchy showing real and 
-ghost grids.
-\emph{Right:}  Example 2D \enzo\ grid showing real and ghost zones, as 
-needed for the PPM hydro stencil.  Image courtesy of James Bordner (Cent.
-Astrophysics and Space Sciences, UCSD).
-}
-\label{fig.2.amrhierarchy}
-\end{figure}
-
-\begin{figure}
-\begin{center}
-\includegraphics[width=0.4\textwidth]{figures/amr-hier-dens.eps}
-\caption{Example of an \enzo\ simulation showing the AMR grid hierarchy.  
-This is a simulation of the collapse of a single massive halo with a 
-$32^3$ root grid and two $32^3$ static nested grids.  AMR is only allowed
-within the most highly refined static nested grid.  Log baryon density is shown
-by colored isocontours with values corresponding to the legend at center left.  
-The rectangular solid wire frames correspond to individual 
-AMR grids, with different colors corresponding to level as indicated by the
-legend at top left.}
-\label{fig.2.amrhierdens}
-\end{center}
-\end{figure}
-
-Each grid patch in \enzo\ contains arrays of values for baryon and 
-particle quantities.   Grids are 
-partitioned into a core of \emph{real zones} and a surrounding layer 
-of \emph{ghost zones}.  The real zones store field values and ghost 
-zones are used to temporarily store values which have been obtained 
-directly from neighboring grids or interpolated from a parent grid.  
-These zones are necessary to accommodate the computational stencil of 
-the hydrodynamics solvers (Sections~\ref{sec.ov.hydro.ppm} and~\ref{sec.ov.hydro.zeus}) 
-and the gravity solver (Section~\ref{sec.ov.nbody}).  
-The PPM and Zeus hydro solvers require 3 layers of ghost zones,  and the gravity
-solver requires 6.  This can lead to significant memory 
-and computational overhead, particularly for smaller grid patches
-at high levels of refinement.  More information on filling of ghost
-zones can be found in the appendix.
-
-From the point of view of the hydro solver, each grid is solved as an
-independent CFD problem, with Dirichlet boundary conditions stored in
-the ghost zones.  At the end of each time step, cells on a given level
-fill ghost zones by interpolating from the parent and copying from
-neighbors on the same level.  Grids then correct the flux at the interface boundary between
-fine and coarse zones, and finally project their real zone data to the
-parent grid.
-
-Timesteps are determined on each level as
-described in \ref{sed.timestepping}, and the hierarchy is advanced on
-a level by level basis in what's known as a W-Cycle.    Beginning with
-the coarsest level, $L$, all grids on that level are advanced one
-timestep.  Then one timestep is taken on all grids at the next level
-of refinement, $L+1$, and so on until the finest level is resolved.
-The finest level is then advanced, using as many steps as it takes to
-reach the level immediately above.  The finest level is then synced
-to the level above it, which then proceeds forewards one more step.
-\red{Needs a picture of the W cycle.}
-
-At the end of every timestep on every level each grid updates 
-its ghost zones by exchanging information with its neighboring grid patches 
-(if any exist) and/or by interpolating from a parent grid.  
-In addition, cells are examined to see if they should be refined or 
-de-refined, and the entire grid hierarchy is rebuilt at that 
-level (including all more highly refined levels).  The timestepping and 
-hierarchy advancement/rebuilding process described here is repeated 
-recursively on every level to the specified maximum level of 
-refinement in the simulation.
-
-Since the addition of more highly refined grids is adaptive,  
-the conditions for refinement must be specified.  The criteria of 
-refinement can be set by the threshold value of the overdensity of 
-baryon gas or dark matter in a cell (which is really a refinement on
-the mass of gas or DM in a cell), the local Jeans length, the 
-local density gradient, or local pressure and energy gradients.  
-A cell reaching any or all of these criteria 
-will then be flagged for refinement.  
-Once all cells at a given level have been examined, rectangular boundaries 
-are determined which minimally encompass the flagged cells. The set of
-flagging cells is then further fractured into smaller bounding
-rectangles to ensure computational 
-efficiency.  More details on subgrid creation can be found in Appendix
-\ref{app.subgrid_creation}.  A refined grid patch is 
-introduced within each such bounding rectangle. Thus the cells needing 
-refinement, as well as adjacent cells within the patch which do not need 
-refinement, are refined. While this approach is not as memory efficient as 
-cell-splitting AMR schemes, it offers more freedom with 
-finite difference stencils. For example, PPM requires a stencil of seven 
-cells per dimension. This cannot easily be accommodated in cell-splitting 
-AMR schemes.
-
-\blue{ \emph{I think this  should be
-  re-written to reflect data presented or referenced in this paper,
-  whatever that winds up being.}
-In the simulations discussed in this thesis we typically use baryon and
-dark matter overdensities as our refinement criteria, though
-for some higher-resolution simulations we also refine on other
-criteria as needed.}
-
-\subsection{Time Stepping}\label{sec.timestepping}
-\dcc{Made this its own section (not really part of the AMR, it's more
-  general than that.)  Added stuff on harmonic average for 3d. See
-  note below on acceleration timestep criterion being crap. Moved
-  w-cycle to the previous section}
-In \enzo, resolution of the equations being solved is adaptive in time 
-as well as in space.  The timestep in \enzo\ is satisfied on a level-by-level
- basis by finding the largest timestep such that multiple criteria are
-satisfied on each level.  The timestep criteria used by Enzo are 
-(showing the one-dimensional case for clarity):
-
-\begin{equation}
-\Delta t_{hydro} = min \left( \kappa_{hydro} \frac{a \Delta x}{c_{s} + |v_x|} \right)_L ,
-\label{eqn:dthydro}
-\end{equation}
-
-\begin{equation}
-\Delta t_{dm} = min \left(\kappa_{dm} \frac{a \Delta x}{v_{dm,x}} \right)_L ,
-\label{eqn:dtdarkmatter}
-\end{equation}
-
-\begin{equation}
-\Delta t_{exp} = f_{exp} \left( \frac{a}{\dot{a}} \right) ,
-\label{eqn:dtexpand}
-\end{equation}
-
-\begin{equation}
-\Delta t_{accel} = min \left( \sqrt{\frac{\Delta x}{\vec{g}}} \right)_L 
-\label{eqn:dtaccel}
-\end{equation}
-
- In equations~\ref{eqn:dthydro}-\ref{eqn:dtaccel}, the $min ( \ldots )_L$
-formalism means that this value is calculated for all cells on a given level
-L and the minimum overall value is taken as the timestep.  
-Equation~\ref{eqn:dthydro} ensures that all cells satisfy the 
-Courant-Freidrichs-Levy (CFL) condition for accuracy and stability of an explicit
-finite difference discretization of the Euler equations.  Effectively this condition
-forces the timestep to be small enough such that any changes in the fluid propagate
-less than a single grid spacing, $\Delta x$.  In this equation, $\kappa_{hydro}$ is 
-a numerical constant with a value of $0 < \kappa_{hydro} \leq 1$ (with a typical
-value of $\kappa_{hydro} \sim 0.3-0.5$) that ensures that the CFL condition is always
-met, and $c_s$ and $v_x$ are the sound speed and peculiar baryon
-velocity in a given cell.
-
-Equation~\ref{eqn:dtdarkmatter} is analogous to Equation~\ref{eqn:dthydro} and ensures
-accuracy in the N-body solver by requiring that no dark matter particle travels more than
-one cell width.  The parameter $\kappa_{dm}$ is analogous to $\kappa_{dthydro}$, with
-an identical range of values.  Equation~\ref{eqn:dtexpand} limits the timestep such 
-that the expansion parameter a can only change by a fractional amount of $f_{exp} = \Delta a/a$, 
-where $f_{exp}$ is a user-defined parameter and has typical values of $f_{exp} = 0.01-0.02$.
-This is required for the stability of the PPM algorithm in comoving coordinates, and
-typically limits the timestep only at high redshifts when densities are relatively
-homogeneous.
-
-Equation~\ref{eqn:dtaccel} is supplementary to equation~\ref{eqn:dthydro} in that it
-takes into account the possibility of large accelerations causing numerical 
-instabilities by violating the Courant condition.  In this equation, $\vec{g}$ is the
-gravitational acceleration in each cell on level L.  \blue{ \emph{It should be
-  noted that the timestep is computed before the acceleration field is computed,
-  and the acceleration field is deleted at the end of each time step.
-  So this condition isn't ever actually enforced.  We should either
-  fix this problem, or not discuss it.
- -dcollins.} }
-
-Equation ~\ref{eqn:dthydro} is valid for one dimension.  For 2 or 3
-dimensions, it was shown by \cite{Godunov1959}  that using the
-harmonic average of the timestep found along each of the coordinate
-axes yields a maximum $\kappa_{hydro} = 0.8$.  So letting $\Delta
-t_x$, $\Delta t_y$, and $\Delta t_z$ be the analogues of
-eqn. \ref{eqn:dthydro} along the $x,y$ and $z$ axes, 
-\begin{equation}
-\Delta t_{hydro} = min ( \kappa_{hydro} /( \frac{1}{\Delta t_x}
-+\frac{1}{\Delta t_y} + \frac{1}{\Delta t_z} ) )_L
-\end{equation}
-
-For all other criteria, multiple dimensions are accounted for by
-repeating the one dimensional criterion along each axis, and taking
-the minimum.
-
-%---------------------------------------------------------------------------%
-\subsection{Parallelism}\label{sec.ov.parallel}
-\dcc{moved hierarchy description, parallism notes from AMR section
-  here. Mentioned performance/scaling study.  Still short.}
+\subsection{Parallel Strategy}
 
@@
 %Figure~\ref{fig.2.amrhierarchy} for a schematic example of this.
+
+\begin{figure}
+\begin{center}
+\includegraphics[width=0.4\textwidth]{figures/amr_hierarchy.eps}
+\end{center}
+\caption{\emph{Left:}  Example of a simple, distributed AMR hierarchy showing real and ghost grids.
+\emph{Right:}  Example 2D \enzo\ grid showing real and ghost zones, as 
+needed for the PPM hydro stencil. }
+\label{fig:amr_hierarchy}
+\end{figure}
+
+% ----------------------------------
+
+\subsection{Performance}
+\label{sec.performance}
+
+\red{
+This is really James' section.  Profiling the code is very difficult because
+of the many possible configurations that it can be run under.  In general,
+we want to show both AMR and unigrid calculations scale, for cosmology runs
+and for pure hydro (i.e. turbulence) runs.  The gravity scales far differently
+than the hydro, so perhaps we should do a gravity-only (dark matter?) vs. 
+hydro only (driven isothermal turbulence, non-gravitating) scaling?
+We should also include performance information for some of the very large
+simulations that we've done -- 2048$^3$ unigrid and 512$^3$ AMR runs.
+}
+
+\red{It seems like this section will be fiercely debated.  So have fun.}

### 2013-01-18 ms.tex
updated ms.tex

@@
 
-% Paper Outline  (last updated June 15, 2009)
+% Paper Outline  (last updated January 18, 2013)
 %
-% 1.  Introduction [GB,MN]
-% 2.  Overview [GB]
-% - including both equations and a few paragraph overview of how each method works
-% - this might have section 2.1, 2.2, etc. corresponding to very brief overviews of section 3.1, 3.2, etc.
+% 1.  Introduction
+% 2.  Physical equations and overview
+%   2.1 Physical equations
+%   2.2 Overview of numerical methods
 % 3.  Numerical methods
-%  3.1 AMR [GB]
-%  3.2. Hydro: ZEUS [GB]
-%  3.3. Hydro: PPM [GB,AK]
-%  3.4. Gravity [GB]
-%  3.5. Particle Integration [GB]
-%  3.6. Chemistry [BWO]
-%  3.7. Star formation & Feedback [BWO]
-%  3.8. Timestep constraints [BWO,GB]
-% 4. Software Implementation
-% - maybe this could be split up by into sections 4.1, 4.2, 4.3, etc. using the same scheme as section 3 [same responsibilities as 3]
+%   3.1 AMR
+%   3.2. Hydro: Godunov (PPM)
+%   3.3  Hydro: Godunov (MUSCL)
+%   3.4. Hydro: Finite difference (ZEUS)
+%   3.5. MHD: Dedner
+%   3.6. Gravity
+%   3.7. Particle Integration
+%   3.8. Chemistry
+%   3.9. Cooling
+%   3.10. Radiation: Homogeneous
+%   3.11  Radiation: ray-tracing
+ %  3.12  Radiation: FLD
+%   3.13. Star formation & Feedback
+%   3.14. Timestep constraints
+% 4. Code tests
+%   4.1 Verifying and Validating the Enzo code
+%   4.2 Representative test problems
 % 5. Parallelism 
-% 5.1. Parallel strategy [GB, RH]
-% 5.2. Parallel performance [JB, GB, RH]
-% 6. Code Tests [All]
-% 7. Summary and Conclusions
+%   5.1. Parallel strategy 
+%   5.2. Parallel performance 
+% 6. Code development methodology
+% 7. Conclusions
 %
@@
 \title{Enzo: An Adaptive Mesh Refinement Code for Astrophysics}
-\author{Greg Bryan\altaffilmark{1}, Michael L. Norman\altaffilmark{2}, 
-Brian W. O'Shea\altaffilmark{3}, Robert Harkness\altaffilmark{4}, Tom Abel\altaffilmark{5},
-David C. Collins,\altaffilmark{2} 
-Alexei Kritsuk\altaffilmark{2} and James Bordner\altaffilmark{2} }
+\author{Authors}
 
-\altaffiltext{1}{ Department of Astronomy, Columbia University,
-New York, NY, 10027, U.S.A.  Email: gbryan@astro.columbia.edu }
-\altaffiltext{2}{Center for Astrophysics and Space Sciences,
-University of California at San Diego.
-Email: mnorman,dcollins,akritsuk,jbordner@cosmos.ucsd.edu}
-\altaffiltext{3}{Theoretical Astrophysics (T-6), Los Alamos National
-Laboratories.  Email:  bwoshea@lanl.gov}
-\altaffiltext{4}{San Diego Supercomputing Center, University of California
-at San Diego.  Email: harkness@sdsc.edu}
-\altaffiltext{5}{Kavli Institute for Particle Astrophysics and Cosmology,
- Stanford University.  Email:  tabel@stanford.edu}
+%\altaffiltext{1}{ Department of Astronomy, Columbia University,
+%New York, NY, 10027, U.S.A.  Email: gbryan@astro.columbia.edu }
+%\altaffiltext{2}{Center for Astrophysics and Space Sciences,
+%University of California at San Diego.
+%Email: mnorman,dcollins,akritsuk,jbordner@cosmos.ucsd.edu}
+%\altaffiltext{3}{Theoretical Astrophysics (T-6), Los Alamos National
+%Laboratories.  Email:  bwoshea@lanl.gov}
+%\altaffiltext{4}{San Diego Supercomputing Center, University of California
+%at San Diego.  Email: harkness@sdsc.edu}
+%\altaffiltext{5}{Kavli Institute for Particle Astrophysics and Cosmology,
+% Stanford University.  Email:  tabel@stanford.edu}
 
@@
 \input{numerical-ppm}
+\input{numerical-muscl} % new
 \input{numerical-zeus}
+\input{numerical-mhd-dedner} % new
 \input{numerical-gravity}
@@
 \input{numerical-chemistry}
+\input{numerical-cooling} % new
+\input{numerical-radiation-homogeneous} % new
+\input{numerical-radiation-raytracing} % new
+\input{numerical-radiation-fld} % new
 \input{numerical-starformation}
 \input{numerical-timestep}
-%\input{numerical-units}
-
-%% Section: software implementation
-% \input{software}
+\input{numerical-analysis} % new
 
@@
 
-%% Section: discussion
-\input{discussion}
+%% Section: code development methodology
+\input{development} % new
 
-%% Section: summary
-\input{summary}
-
-\appendix
-
-%% Appendix: interpolation
-%\input{app-interpolation}
-
-%% Appendix:  Data Structures
-\input{app-datastruct}
-
-%% Appendix:  I/O
-\input{app-IO}
-
-%% Appendix:  Zeus hydro
-%\input{app-zeushydro}
-
-%% Appendix:  PPM hydro
-%\input{app-ppmhydro}
-
-%% Appendix:  gravity solver
-%\input{app-gravity}
-
-%% Appendix:  
-\input{app-chemistry}
+%% Section: conclusions
+\input{conclusions}  % new
 

### 2013-01-18 conclusions.tex
added two new sections.

--- 2012-06-29 app-datastruct.tex
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\section{Conclusions}
 
-\section{Data structures}\label{app.datastructure}
-
-The primary data structure in ENZO\ is the {\tt grid}.  
-\gtt{It has yet to be determined how much of this
-appendix will be in the Method Paper, and how much will be in the
-User/Developer Guide.  Green type denotes text that will definitely be
-in the User/Developer guide.} \red{Red denotes meta notes-- things to
-write, fix, find.}
-
-\subsection{Outline and List of Routines.}
 \begin{enumerate}
- \item Creation and Destruction
-   \begin{enumerate}
-     \item Flow of Rebuild hierarchy: 
-       \begin{enumerate}
-	 \item delete hierarchy
-	 \item build new hierarchy
-	 \item scavange from old hierarchy
+\item near-future development plans
+\item far-future development plans
+\item clear areas of improvement
 \end{enumerate}
-     \item Cell Flagging Methods
-     \item Grid Fracturing Methods
-   \end{enumerate}    
- \item Parallelism
-  \begin{enumerate}
-    \item Top Grid (block domain decomposition)
-    \item Subgrids (memory balance, level by level)(Actually, I think
-          its Parent Grid by Parent Grid.)
-  \end{enumerate}
- \item Relation to other objects: 
-   \begin{enumerate}
-     \item Parents (No refinement jump requirements.  Talk about flux
-        correction improvement to this effect here?)
-     \item Siblings
-     \item Children
-     \end{enumerate}
- \item Traversing the hierarchy (this might be rolled into the
-   previous bullet)
-  \begin{enumerate}
-    \item Level iterator 
-    \item Hierarchy iterator
-    \item Grid Array
-    \item Chaining Mesh
-  \end{enumerate}
-  \item Particle treatement
-   \begin{enumerate}
-     \item Which grid gets the particles
-     \item How particles are refined: criteria
-     \item How particles are refined: how it fits in with the Grid refinement
-   \end{enumerate}
-  \item Ghost zones
-    \begin{enumerate}
-      \item how, when filled
-    \end{enumerate}
-  \item Timestepping
-    \begin{enumerate}
-      \item How timestep is determined
-      \item Order of level operations: W cycle
-    \end{enumerate}
-\end{enumerate}
-A list of routines to mention: \red{Zeus does this in a big table that
-  references code name, what it does, and what equations it
-  solves. Might be a useful idea to steal.}
-\begin{description}
- \item[\tt{EvolveHierarchy}] Primary timestep control routine. 
- \item[\tt{EvolveLevel}] Primary evolution routine.
- \item[\tt{RebuildHierarchy}] Rebuilds the grid hierarchy
-\end{description}
-
-\subsection{Grid Intro}
-The life of an \enzo\ simulation centers around the {\tt grid}
-object and relationships between individual {\tt grid} instances.  
-Grids are independant cartesian patches in space, which contain meta data about
-that space (for instance its size, its position in the volume, 
-size of the cells in that patch) as well as the data that belongs in
-that space (fluid quntities like density and energy, as well as
-particle quantites like position and velosity.)
-
-Grids are treated, as individual instances of fluid
-dynamics problems, with Dirichlet boundary conditions.  Boundary
-conditions, stored in ghost zones, are determined by
-one of 3 methods; from user defined exterior conditions like
-outflow or reflecting, copied from active zones of 
-neighbors, or interpolated from parents.  Grids are group in two
-logical structures: a tree like hierarchy that relates parent grids to
-children grids; and a level structure that relates grids of the same
-refinement.
-
-\subsection{Creation and Destruction: \tt{RebuildHierarchy}}
-Creation and destruction of grids happens in the routine
-{\tt RebuildHierarchy}.  
-At the end of every timestep, the Hierarchy is rebuilt.
-Rebuilding starts at the coarsest level at the point in time (note
-that coarser levels may be further ahead in their integration, due to
-their larger time step and the W-cycle)  Beginning with this coarsest level, each
-grid on each level is examined for refinement.  Cells are first flagged
-based on a variety of criterion, described in section
-\ref{sec.flagging}.  The set of flagged cells, stored in the \grid\ memeber
-array {\tt FlaggingField}, is then devided into grids as described
-in section \ref{app.subgrid_creation}.  Data in these subgrids
-is then created by interpolation from the parent using \red{what kind
-  of interpolation?}  After this New Subgrid is filled with
-interpolated data, the set of Old Subgrids is examined for overlapping
-data.  All Old Cells that correspond to New Cells are coppied to the
-New Grid.
-
-\subsection{Cell Flagging Methods}\label{sec.flagging}
-\enzo\ selects cells to refine based on the physics of the simulation.
-Essentially any criterion that can be derived locally from the fluid
-data and its derivatives can, in principal, be used to flag the cells
-for refinement. Currently, \enzo\ has 8 methods of refinement.
-installed, which we will discuss here.  Any or all of these methods may be employed in
-any given \enzo\ simulation.  \gtt{Information on how to write your own
-refinement criterion can be found Somewhere, to be written}
-\begin{description}
-  \item[Field Slope] \gtt{grid::FlagCellsToBeRefinedBySlope()} For each
-  fluid quantity $u$, a cell $i$ is flagged if 
-$$ \frac{u_{i+1}-u_{i-1}}{u_i} > slope $$
-  where $slope$ \gtt{\tt MinimumSlopeForRefinement} is
-  a  user-defined
-  parameter. This is repeated for all
-  3 directions. 
-  \item[Baryon mass] \gtt{
-  grid::FlagCellsToBeRefinedByMass()}  Cells
-  are flagged if $$ \rho  V > m_{f} r^{\alpha L} $$ 
-  where $V$ is the
-  volume of the cell, $r$ is the refinement ratio, $L$ is the grid
-  level, $m_{f}$ and ${\alpha}$ are user parameters. \gtt{MinimumMassForRefinement} and 
-  \gtt{ MinimumMassForRefinementLevelExponent\[\]}
-  \item[Baryon Overdensity] As the name suggests, refines if the
-  density is greater than the mean density in the simulation.  In
-  practice, this refinement method uses the same machenery as flagging
-  by Baryon Mass, with the minimum
-  mass for flagging, $m_f$ \gtt{ = MinimumMassForRefinement}, set to the
-  input parameter \gtt{MinimumOverDensityForRefinement} times the
-  mean density.  \gtt{This is calculted differently for each problem
-  generator: consult the source for further documentation}
-  \red{This could probably use a word on the effects of
-  parameter choices.}
-  \item[Shocks] \gtt{
-  grid::FlagCellsToBeRefinedByShocks()} \red{This text should be
-  updated to reflect Alexei's improvements} A cell $i$ is flagged if it is
-  determined to be taking part in a shock.  This is determined if it
-  has each of the following:
-  \begin{enumerate}
-    \item Large relative pressure jump: 
-      $$\frac{|p_{i+1} - p_{i-1}| }{min( p_{i+1}, p_{i-1} ) }
-      > \gtt{MinimumPressureJumpForRefinement} $$
-     \item compression: $$ u_{i+1} < u_{i-1} $$
-    \item More thermal energy than Kinetic: $$ \frac{
-    \frac{p_i}{\gamma -1} }{ \frac{1}{2}\rho v^2 } > \gtt{
-    MinimumEnergyRatioForRefinement } $$ \red{why?}
-  \end{enumerate}  
-  \item[Particle Mass] Using the Cloud In Cell interpolation method,
-  described  \red{elsewhere?}, particle data is interpolated to the
-  Eulerian grid cell centers.  A cell is flagged if the resulting mass field passes
-  the ``RefineByMass'' test, described above.
-  \item[Jeans Length] If a gravitationally collapsing simulation is
-  under-resolved, artificial fragmentation can occur.  It was
-  descoverd experimentally by \red{Truelove; get citation}
-  that sufficient resolution was achieved if the Jeans Length is
-  resolved by at least 4 cells.  This is determined if 
-  $$ dx^2 >
-  \frac{\alpha}{N_{cells} \gtt{= RefineByJeansLengthSafetyFactor}}\frac{T_{i,j,k}}{\rho_{i,j,k}} $$
-
-  and $\alpha = \frac{\pi k_b }{G m_{proton} }*
-  \frac{LengthUnits^2}{DensityUnits}$
-
-  \red{dude, double check this.}
-  \item[Cooling Time]  \red{Finish this.}
-
-  \item[Shear] For highly turbulent flows, it may be useful to resolve
-  shear layers.  This is done whenever
-  $$\frac{\partial v_x}{\partial y}^2 + \frac{\partial v_x}{\partial z}^2 
-  > \eta \gtt{= MinimumShearForRefinement}
-  $$
-  In practice, $$\frac{\partial v_x}{\partial y} = 
-  u_{j+1}-u_{j-1}$$
-  The same is done for $v_y$ and $v_z$
-
-\end{description}
-
-\subsection{Grid Fracturing}\label{app.subgrid_creation}
-Once the cells in a given grid are flagged for refinement, they're
-broken into a set of subgrids by an iterative process.  This process
-first draws a minimally enclosing box around the set of flagging
-cells, and generates a {\tt ProtoSubgrid}.  A {\tt ProtoSubgrid} is
-essentially just the spatial extent and flagging field that will
-define the spatial extent of the future \grid.  Each {\tt
-  ProtoSubgrid} is then examined for efficiency (the ratio of flagged cells to
-total), total size (for cache optimization) and minimum side
-length (to avoid a flock of one zone grids).  {\tt ProtoSubgrid} that are too
-large or not efficient enough are broken into smaller grids, by a
-process to be described shortly, new minimally enclosing boxes are
-drawn around each of the two remaining flagging fields, and new {\tt
-  ProtoSubgrid}s are created.  This process is then repeated until
-all {\tt ProtoSubgrid}s are both small enough and efficient enough.
-It should be noted that \enzo\ does not glue grids back together-- the
-Minimum Side Length is enforced by not examining the other two
-conditions if the {\tt ProtoSubgrid} is too small.
-
-Maximum size, minimum side length,  and minimum efficiency are
-controlled by the parameters {\tt MaximumSubgridSize}, {\tt MinimumSubgridEdge} and {\tt
-  MinimumEfficiency}, respectivly. Typical maximum sizes are 10000 zones, minimum
-side lengths are typically 5 zones, and minimum efficiencies are  0.2.
-Choice of these parameters, especially maximum size, are
-based on the architecture of the computer in question.  
-Machines with larger cache sizes will benifit from larger values for
-{\tt MaximumSubgridSize}, as communication and memory overhead from
-ghost zones will be reduced; machines with smaller cache sizes benifit
-from lower values, in that the computational time is reduced for a
-grid that can fit entirely into cache.
-
-A {\tt ProtoSubgrid} that fails either the efficency or size
-requirement is then broken into smaller {\tt ProtoSubgrid}s by a
-method that tries to preserve physically continuous structures onto
-individual grids.  The process examines the 1d Signature of the Flagging
-field, which is the projection of the Flagging Field along two of the
-coordinate axes.  First, each of the 3 signature fields is examined
-for zeros.  Zeros in the signature field indicates disjoint
-structures, so the {\tt ProtoSubgrid} is first cut into two smaller
-ones along that line.  If no such cuts can be made, zero crossings of
-the second derivative in the signature are then looked for.  Zero
-crossings in the second derivative indicates edges of structures.  If
-more than one exists, the 'strongest crossing,' that is the largest
-jump, is selected for the cut.  If no such zero crossings can be
-found, the {\tt ProtoSubgrid} is then cut evenly along the longest
-axis.  See figure \ref{fig.signature} for a schematic of this process.
-
-\begin{figure}
-\begin{center}
-\includegraphics[width=0.4\textwidth]{figures/SignatureSchematic.eps}
-\end{center}
-\caption{
-Schematic of Flagging Field and Signature.  Cuts will be made along
-some line I'll draw.  The figure will be improved in the future.
-}
-\label{fig.signature}
-\end{figure}

### 2013-01-18 numerical-cooling.tex
adding a bunch of new sections and subsections.

\subsection{Radiative cooling}
\label{sec.num.cooling}

### 2013-01-18 numerical-mhd-dedner.tex
adding a bunch of new sections and subsections.

\subsection{MHD: Dedner}
\label{sec.num.mhd-dedner}

### 2013-01-18 numerical-radiation-fld.tex
adding a bunch of new sections and subsections.

\subsection{Radiation transport: Flux-limited diffusion}
\label{sec.num.rad-fld}

### 2013-01-18 numerical-radiation-homogeneous.tex
adding a bunch of new sections and subsections.

\subsection{Homogeneous radiation background}
\label{sec.num.rad-homogeneous}

### 2013-01-18 numerical-radiation-raytracing.tex
adding a bunch of new sections and subsections.

\subsection{Radiation transport: ray tracing}
\label{sec.num.raytracing}

### 2013-01-18 code-tests.tex
updating code tests section

@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Code tests}\label{sec.tests}
+\section{Code tests}
+\label{sec.tests}
 
-\red{
-\begin{enumerate}
-\item Things that need to be addressed:  Can we refine by non-powers of two?  What test problems to do?  Order of test problems?
-\item Some of the test problems are motivated by the FLASH code method paper \citep{flash_method}, and are referred to there.
-Others are motivated by the GADGET-2 method paper \citep{gadget_2_method}, and are referred to as such.
-\item In this paper we want to make sure to refer to the Santa Barbara Cluster comparison project \citep{SantaBarbara} and to our ENZO/GADGET-2 comparison project~\citep{2005ApJS..160....1O}.
-\end{enumerate}
-}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\subsection{Verifying and validating the \enzo\ code}
+\label{sec.tests.vandv}
 
-\red{
+\red{(Britton)} Description of test suite and test runner.
+
+\red{(Brian)} Refer to all code comparisons that Enzo has been part of.
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\subsection{Representative test problems}
+\label{sec.tests.problems}
+
 General structure for each test problem:  Outline how the test problem is constructed (initial and boundary conditions), 
@@
 a plot showing how well enzo solves said test problem, and a brief description of the plot and how awesome enzo is.
-}
 
-\red{*** WE NEED SUGGESTIONS FOR CODE TESTS***}
-
-\subsection{Advection problem}\label{sec.tests.advect}
-Problem 7.1 in the flash method paper~\citep{flash_method}.
-
-\subsection{Sod Shock Tube}\label{sec.tests.sodshock}
-Problem type 1.  Both unigrid and AMR.  Tests the hydro.
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\subsubsection{Sod Shock Tube}
+\label{sec.tests.sodshock}
+\red{(Greg)}
+Problem type 1.  AMR version.  Tests the hydro.
 This is also problem 7.2 in the FLASH method paper.
 
-\subsection{Wave pool}\label{sec.tests.wavepool}
-Problem type 2, unigrid and fixed AMR 1D mostly.  Tests reflections of waves at boundaries.
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\subsubsection{Wave pool}
+\label{sec.tests.wavepool}
+\red{(Greg)}
+Problem type 2, with AMR.  Tests reflections of waves at grid boundaries.
 
-\subsection{Shock pool}\label{sec.tests.shockpool}
-Problem type 3, unigrid and AMR 2D.  Tests passage of shock through a refinement boundary.
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\subsubsection{Shock pool}
+\label{sec.tests.shockpool}
+\red{(Greg)}
+Problem type 3, with AMR.  Tests passage of shock through a refinement boundary.
 
-\subsection{Interacting Blast-Wave problem}\label{sec.tests.interblast}
-Problem 7.5 in the FLASH method paper.
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\subsubsection{Double mach reflection}
+\label{sec.tests.doublemach}
+\red{(Brian)}
+Problem type 4.  This is one of Alexei's test problems.  Tests
+boundary conditions in hydro.
 
-\subsection{Double mach reflection}\label{sec.tests.doublemach}
-Problem type 4.  This is one of Alexei's test problems: 
-http://cosmos.ucsd.edu/~akritsuk/renzo/mach/mach.html
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\subsubsection{Sedov Explosion}
+\label{sec.tests.sedov}
+\red{(Elizabeth)}
+Problem type 7.  One of Alexei's test problems.  
+Also problem 7.4 in the FLASH method paper.
+This tests the hydro in the strong-shock limit.
 
-This tests the hydro.
-
-\subsection{Sedov Explosion}\label{sec.tests.sedov}
-Problem type 7.  One of Alexei's test problems:
-http://cosmos.ucsd.edu/~akritsuk/renzo/sedov/sedov.html
-Also problem 7.4 in the FLASH method paper.
-This tests the hydro.
-
-\subsection{Noh problem}\label{sec.tests.noh}
-Problem type 9.  One of Alexei's test problems:
-http://cosmos.ucsd.edu/~akritsuk/renzo/noh/noh.html
-This tests the hydro.
-
-\subsection{Radiating shock problem}\label{sec.test.radshock}
-In~\citet{anninos97} -- tests the energy equation.  Followon
-to Noh problem.  Original reference is cited in this paper.
-
-\subsection{Point source gravity test}\label{sec.test.gravitypointsource}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\subsubsection{Point source gravity test}
+\label{sec.test.gravitypointsource}
+\red{(Greg)}
 This is the TestGravity (problem 23) test problem.  It tests gravity around a point source, using fixed AMR.
 
-\subsection{Orbit Test}\label{sec.test.testorbit}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\subsubsection{Orbit Test}
+\label{sec.test.testorbit}
+\red{(Greg)}
 This is the TestOrbit problem (problem 29 for GB).  This tests gravity and particle integration (no hydro).
 
-\subsection{Pressureless Sphere Collapse}\label{sec.test.pressurelesscollapse}
-This is in the zeus paper and in the kronos test paper -- it tests mostly gravity and advection.  Test problem 21.
-
-\subsection{Collapse of an adiabatic gas sphere}\label{sec.tests.adiasphere}
-I think this is problem type 27, and also is problem 6.2 in the
-GADGET-2 method paper~\citep{gadget_2_method}.
-
-\subsection{Self-Similar infall test}\label{sec.tests.infall}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\subsubsection{Self-Similar infall test}
+\label{sec.tests.infall}
+\red{(Greg)}
 Problem type 24.  This is a test based on Bertschinger's 1985 3D self-similar infall
@@
 
-\subsection{Zel`Dovich Pancake}\label{sec.tests.pancake}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\subsubsection{Zel`Dovich Pancake}
+\label{sec.tests.pancake}
+\red{(Brian)}
 Problem type 20, unigrid and AMR both.  Tests gravity + hydro.
 
-\subsection{Santa Barbara Cluster comparison}\label{sec.tests.santabarbara}
-This is the classic test -- we can mostly just point to the enzo results in that paper.  Maybe update them?
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\subsubsection{Cosmology simulations}
+\label{sec.tests.}
+\red{(Brian)}
+Talk about cosmology code comparisons, where Enzo is good/weak.
 
-\subsection{Dark matter halo mass function and clustering}\label{sec.tests.dm}
-Standard problem, also problem 6.4 in GADGET-2 method paper.  Should this be here, or just a reference to another paper?
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\subsubsection{Representative MHD test}
+\label{sec.tests.mhd}
+\red{(Dave)}
+One representative MHD test.  What's the best/hardest one?
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\subsubsection{1-zone free-fall test}
+\label{sec.tests.freefall}
+\red{(Britton)}
+One zone free fall test.  Tests cooling, chemistry.
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\subsubsection{Representative ray-tracing test}
+\label{sec.tests.raytracing}
+\red{(John)}
+One test for MORAY: best/hardest one!
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\subsubsection{Representative FLD test}
+\label{sec.tests.fld}
+\red{(Dan)}
+One test for the implicit FLD: best/hardest one.

### 2013-01-18 parallel.tex
Updated.

@@
 \subsection{Parallel Strategy}
+
+% GB to work on this
+%  - expand to include better description
+%  - include load balancing options
 
@@
  is used to transfer data between processors.
-
-The performance of \enzo\ has been extensively studied using the
-software package LCAPerf.  More on LCAPerf can be found in section
-\ref{sec.performance}
 
@@
 
-\red{
-This is really James' section.  Profiling the code is very difficult because
-of the many possible configurations that it can be run under.  In general,
-we want to show both AMR and unigrid calculations scale, for cosmology runs
-and for pure hydro (i.e. turbulence) runs.  The gravity scales far differently
-than the hydro, so perhaps we should do a gravity-only (dark matter?) vs. 
-hydro only (driven isothermal turbulence, non-gravitating) scaling?
-We should also include performance information for some of the very large
-simulations that we've done -- 2048$^3$ unigrid and 512$^3$ AMR runs.
-}
+\subsubsection{Performance model \red{(Greg)}}
 
-\red{It seems like this section will be fiercely debated.  So have fun.}
+% Maybe include Greg's text on general scaling arguments
+
+\subsubsection{Unigrid scaling \red{(Brian)}}
+
+% Brian to work on this, scaling data from Harkness
+
+\subsubsection{AMR scaling \red{(Greg)}}
+
+% Greg to work on this, scaling data from his runs, or maybe from John Wise

### 2013-01-18 numerical-analysis.tex
Tweaked numerical analysis file.

\subsection{Analysis}
\label{sec.num.analysis}

\subsubsection{Inline analysis with yt}

\subsubsection{Tracer particles}

\subsubsection{Shock finding}

### 2013-01-18 numerical-mhd-ct.tex
Adding constrained transport MHD file and call from ms.tex

\subsection{MHD: Constrained transport}
\label{sec.num.mhd-ct}

### 2013-01-18 numerical-mhd-ct.tex
Adding MHD-CT text, references

@@
 \label{sec.num.mhd-ct}
+\def\Bvec{{\bf B}}
+\def\Bf{Bf}
+\def\Bc{Bc}
+\def\Evec{{\bf E}}
+\def\Divb{\ensuremath{\nabla \cdot \Bvec}}
+Physically, the divergence of the magnetic field, \Divb, is
+identically zero due to the fact that the evolution of the magnetic field is the
+curl of a vector, and the divergence of the curl of a vector is identically
+zero:
+\begin{eqnarray}
+\partial \Bvec/\partial t & =& \frac{1}{a} \nabla \times \Evec\\
+\Divb & =& 0.
+\end{eqnarray}
+The Constrained Transport (CT) method \citep{Evans88, Balsara99} of MHD evolution employs this same vector
+property to evolve the magnetic field in a manner that preserves $\Divb=0$ to
+the ability of the computer.  The electric field is computed from the fluxes
+from the Riemann solver.  The curl of that electric field is then used to update
+the  magnetic field.   The advantage of this method is that it preserves \Divb\
+to machine precision.  The drawback is twofold: increased algorithm  complexity,
+as will be described in this section; and since only the update of the magnetic
+field is divergence free, any monopoles created by other numerical sources
+persist.
+
+Enzo MHD-CT has been described in horrid detail in \citet{Collins10}, so the
+description here will be an overview only.  The base Godunov method, described
+in \citet{Li08a}, uses spatially and temporally second order reconstruction
+(both MUSCL-Hancock and Piecewise Linear Method), a selection of Riemann solvers
+including HLLC and HLLD \citep{Mignone07}.  The CT methods are the first order
+method described by \citet{Balsara99} and the second order methods described in
+\citet{Gardiner05}.  The AMR machinery is described by \citet{Balsara01} and
+\citet{Collins10}.
+
+The increased complexity comes in the form of area-averaged face and
+length-averaged edge centered variables, while the rest of Enzo employs
+predominantly volume-averaged cell centered variables.  The magnetic field is
+represented by both a face centered field, \Bf, and a cell centered field, \Bc.
+The electric field is edge centered.
+The magnetic field is updated in four steps:  first, the
+Riemann problem is solved in the traditional manner, using the cell centered
+field; second, an edge-centered
+electric field is computed using the fluxes from the Riemann solver;  third, the curl
+of that electric field is used to update the face centered field; finally,
+the cell centered magnetic field is updated with an average of the face centered
+field.
+
+Divergence free AMR is slightly more complex than the AMR employed elsewhere in
+Enzo.  First, the interpolation must be constrained to be divergence free.
+Thus, all three face-centered field components are interpolated in concert.
+Second, any magnetic field information from previous time steps must be included
+in the interpolation, making the interpolation more complex than the simple
+parent-child relation used for other fields.  Third, the flux correction
+involves more possible grid relations than traditional AMR.  In order to
+circumvent this last complexity, the electric field is projected from fine grids
+to parent grids, rather than the magnetic field, which is then used to re-update
+the parent magnetic field.  This is described in detail in \citet{Balsara99,
+Collins10}
+
+Enzo's MHD-CT has been used in a variety of contexts, from galaxy cluster
+evolution \citep{Xu11,Xu12}, synthetic observations of radio halos
+\citep{Skillman13},  primordial star formation \citep{Xu08}, 
+present day star formation \citep{Collins11,Collins12}.

### 2013-01-18 numerical-conduction.tex
fixed up some subsections.  added a subsection in 2 and 3 for thermal conductivity.

\subsection{Thermal conduction}
\label{sec.num.conductions}

We've got both kinds of conduction: isotropic AND anisotropic.  How
great is that?

### 2013-01-18 code-tests.tex
Added hooks for thermal conduction tests, tweaked the main ms.tex file, and fixed a couple of typos in the outline.

@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsubsection{Representative ray-tracing test}
+\subsubsection{Photo-evaporation of a dense clump}
 \label{sec.tests.raytracing}
 \red{(John)}
-One test for MORAY: best/hardest one!
+Photo-evaporation of a dense clump.  Tests ray tracing, specifically
+the shadowing effects of the clump, and the hydrodynamic response of
+the clump.
 
@@
 One test for the implicit FLD: best/hardest one.
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+\subsubsection{Representative thermal conduction test}
+\label{sec.tests.conduct}
+\red{(Brian)}
+One test for the thermal conduction (maybe 1 for isotropic, 1 for
+anisotropic).

### 2013-01-21 code-tests.tex
adding Orszag Tang test problem

@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsubsection{Representative MHD test}
+\subsubsection{MHD: Orszag-Tang Vortex}
 \label{sec.tests.mhd}
-\red{(Dave)}
-One representative MHD test.  What's the best/hardest one?
+Figure \ref{fig.orszag} shows the Orszag-Tang vortex problem, a typical test
+problem for MHD \citep{Orszag79}.  The left panel shows the MHD-CT result, while
+the right panel shows the Dedner result.  This test shows that small scale structure
+can be generated in MHD from large scale initial perturbations.  
+The test begins 
+with uniform density, $\rho_0=25/36 \pi$ and pressure, $P_0=5/12 \pi$.  There is
+a single rotational mode in the velocity, and two in the magnetic field: 
+${\bf v}_0 $ = (-sin(2$\pi$ y) $ \hat{x}$ , sin(2$\pi$ x) $\hat{y}$),
+${\bf B}_0$  =  (-sin(2 $\pi$ y) $ \hat{x},$ sin( 4 $\pi$ x )$\hat{y}$).  The
+simulation is evolved to t=0.48.  One can see that the structures are accurately
+represented as compared to, for instance, \citet{Toth00}, and that the
+resolution of shocks is comparable in both methods.
+
+{\emph Dave's note:  I'll probably play around with these results in future
+versions.}
+
+\begin{figure}
+\begin{center}
+\includegraphics[width=0.4\textwidth]{figures/MHDCT_OrszagTang_Density.eps}
+\includegraphics[width=0.4\textwidth]{figures/MHDDedner_OrszagTang_Density.eps}
+\caption{TANG}
+\label{fig.orszag}
+\end{center}
+\end{figure}
 
@@
 One test for the implicit FLD: best/hardest one.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-\subsubsection{Representative thermal conduction test}
-\label{sec.tests.conduct}
-\red{(Brian)}
-One test for the thermal conduction (maybe 1 for isotropic, 1 for
-anisotropic).

### 2013-01-22 code-tests-fld.tex
Added separate file for FLD test problem

--- 2013-01-18 code-tests.tex
@@
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Code tests}
-\label{sec.tests}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Verifying and validating the \enzo\ code}
-\label{sec.tests.vandv}
-
-\red{(Britton)} Description of test suite and test runner.
-
-\red{(Brian)} Refer to all code comparisons that Enzo has been part of.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Representative test problems}
-\label{sec.tests.problems}
-
-General structure for each test problem:  Outline how the test problem is constructed (initial and boundary conditions), 
-the analytical solution, why we have in the paper (what does it break, or what flaw does it expose (not in enzo of course)),
-a plot showing how well enzo solves said test problem, and a brief description of the plot and how awesome enzo is.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsubsection{Sod Shock Tube}
-\label{sec.tests.sodshock}
-\red{(Greg)}
-Problem type 1.  AMR version.  Tests the hydro.
-This is also problem 7.2 in the FLASH method paper.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsubsection{Wave pool}
-\label{sec.tests.wavepool}
-\red{(Greg)}
-Problem type 2, with AMR.  Tests reflections of waves at grid boundaries.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsubsection{Shock pool}
-\label{sec.tests.shockpool}
-\red{(Greg)}
-Problem type 3, with AMR.  Tests passage of shock through a refinement boundary.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsubsection{Double mach reflection}
-\label{sec.tests.doublemach}
-\red{(Brian)}
-Problem type 4.  This is one of Alexei's test problems.  Tests
-boundary conditions in hydro.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsubsection{Sedov Explosion}
-\label{sec.tests.sedov}
-\red{(Elizabeth)}
-Problem type 7.  One of Alexei's test problems.  
-Also problem 7.4 in the FLASH method paper.
-This tests the hydro in the strong-shock limit.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsubsection{Point source gravity test}
-\label{sec.test.gravitypointsource}
-\red{(Greg)}
-This is the TestGravity (problem 23) test problem.  It tests gravity around a point source, using fixed AMR.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsubsection{Orbit Test}
-\label{sec.test.testorbit}
-\red{(Greg)}
-This is the TestOrbit problem (problem 29 for GB).  This tests gravity and particle integration (no hydro).
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsubsection{Self-Similar infall test}
-\label{sec.tests.infall}
-\red{(Greg)}
-Problem type 24.  This is a test based on Bertschinger's 1985 3D self-similar infall
-solution, and tests gravity + hydro.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsubsection{Zel`Dovich Pancake}
-\label{sec.tests.pancake}
-\red{(Brian)}
-Problem type 20, unigrid and AMR both.  Tests gravity + hydro.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsubsection{Cosmology simulations}
-\label{sec.tests.}
-\red{(Brian)}
-Talk about cosmology code comparisons, where Enzo is good/weak.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsubsection{Representative MHD test}
-\label{sec.tests.mhd}
-\red{(Dave)}
-One representative MHD test.  What's the best/hardest one?
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsubsection{1-zone free-fall test}
-\label{sec.tests.freefall}
-\red{(Britton)}
-One zone free fall test.  Tests cooling, chemistry.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsubsection{Representative ray-tracing test}
-\label{sec.tests.raytracing}
-\red{(John)}
-One test for MORAY: best/hardest one!
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \subsubsection{Representative FLD test}
 \label{sec.tests.fld}
+
 \red{(Dan)}

### 2013-01-23 numerical-radiation-fld.tex
Updated FLD section of method paper (section 3.13)

@@
 \label{sec.num.rad-fld}
+
+%% With regards to the numerical method section, we're really looking for
+%% something that's no more than a page long: a few relevant equations,
+%% maybe a finite difference equation or two if you feel that it's
+%% relevant, but primarily a description of the algorithms used for the
+%% FLD solve and a brief explanation of how it's coupled to the
+%% chemistry, as well as references to the relevant method paper(s).
+%% Working in a reference or two to science papers is also great!
+
+In addition to the ray-tracing approach for radiation transport
+described in Section \ref{sec.num.raytracing}, Enzo currently includes
+a field-based radiation transport solver for problems posed on uniform
+(i.e.~non-SAMR grids).  Detailed explanations of the model and
+solution approach may be found in
+\cite{NBHBROW2007}, \cite{ReynoldsHayesPaschosNorman2009} and
+\cite{NRS2009}, the salient features of which are reproduced here. 
+In addition, comparisons of this solver with other astrophysical
+radiation transport solvers may be found in \cite{IlievEtAl2009}.
+
+Enzo's field-based radiation solver focuses on
+a flux-limited diffusion approximation for cosmological radiative
+transfer, with couplings to both the gas energy and chemical number
+densities, 
+\begin{eqnarray}
+  \label{eq:fld_radiation}
+  \partial_t E_r + \frac1a \nabla\cdot\left(E_r \vec{v}_b\right) &=& 
+    \nabla\cdot\left(D\nabla E_r\right) -
+    \frac{\dot{a}}{a} E_r - c\kappa E_r + \eta, \\
+  \label{eq:fld_heating}
+  \partial_t e_c &=& -\frac{2\dot{a}}{a} e_c + G - \Lambda, \\
+  \label{eq:fld_chemistry}
+  \partial_t {\mathrm n}_i + 
+    \frac1a \nabla\cdot\left({\mathrm n}_i \vec{v}_b\right) &=& 
+    \alpha_{i,j} {\mathrm n}_e {\mathrm n}_j - {\mathrm n}_i
+    \Gamma_i^{ph}, \quad i=1,\ldots,N_s,
+\end{eqnarray}
+where $E_r$ is a grey radiation energy density, $e_c$ is the
+internal energy correction due to photo-heating and chemical cooling,
+and ${\mathrm n}_i$ correspond to the chemical number densities of
+HI, HII, HeI, HeII and HeIII, and ${\mathrm n}_e$ is the electron
+number density.  Here, we define the grey radiation energy density
+through first assuming a fixed frequency spectrum, i.e.
+$E_{\nu}(\nu,\vec{x},t) = \tilde{E}_r(\vec{x},t) \chi(\nu)$, and then
+defining the integrated quantity
+\begin{equation}
+\label{eq:grey_radiation_energy}
+   E_r(\vec{x},t) \equiv \int_{\nu_0}^{\infty}
+   E_{\nu}(\nu,\vec{x},t)\,\mathrm d\nu \  = \ 
+   \tilde{E}_r(\vec{x},t) \int_{\nu_0}^{\infty} \chi(\nu)\,\mathrm d\nu.
+\end{equation}
+In addition, $D$ in (\ref{eq:fld_radiation}) is the {\em Larsen
+square-root flux-limiter} (see \cite{Morel2000}), $\kappa$ is the total
+opacity, $\eta$ is the field of radiation sources, $G$ provides the
+radiation induced photo-heating, $\Lambda$ provides the chemical
+cooling, $\alpha_{i,j}$ are the reaction rate coefficients defining
+the interactions between species ${\mathrm n}_i$ and ${\mathrm n}_j$,
+and $\Gamma_i^{ph}$ are the radiation induced photo-ionization rates.
+
+The system of equations
+(\ref{eq:fld_radiation}-\ref{eq:fld_chemistry}) is solved
+independently of Enzo's hydrodynamics, gravity and dark-matter solvers
+(sections \ref{sec.hydro.ppm}-\ref{sec.ov.nbody}), thereby allowing
+the advective portions of equations (\ref{eq:fld_radiation}) and
+(\ref{eq:fld_chemistry}) to be taken care of by the fluid solvers.
+Due to the disparate time scales between radiation transport and chemical
+ionization and heating, the remainder of these equations is solved
+using an operator-split algorithm.  Within a given time step to evolve
+$(E_r^n, e_c^n, {\mathrm n}_i^n) \to (E_r^{n+1}, e_c^{n+1}, {\mathrm
+n}_i^{n+1})$, we first evolve equation (\ref{eq:fld_radiation}):
+$(E_r^n, e_c^n, {\mathrm n}_i^n) \to (E_r^{n+1}, e_c^{n}, {\mathrm
+  n}_i^{n})$.  This uses an implicit Euler time discretization, and a
+second-order centered finite-difference spatial discretization,
+resulting in a large linear system of equations.  These are solved
+using a preconditioned conjugate gradient iteration, where the
+preconditioner consists of a geometric multigrid solver.  Both of
+these linear solvers are provided by the HYPRE linear solver library
+(see \cite{FalgoutYang2002} and \cite{hypre-manual}).
+
+We then evolve the system (\ref{eq:fld_heating}-\ref{eq:fld_chemistry}):  
+$(E_r^{n+1}, e_c^n, {\mathrm n}_i^n) \to (E_r^{n+1}, e_c^{n+1},
+{\mathrm n}_i^{n+1})$.  Due to the lack of spatial derivatives (since
+advection is handled elsewhere), this system is a coupled system of
+nonlinear ordinary differential equations.  This utilizes an implicit
+quasi-steady-state approximation, formulated as follows.  We consider
+the modified equations,
+\begin{eqnarray}
+  \label{eq:fld_heating_qss}
+  \partial_t e_c &=& -\frac{2\dot{a}}{a} e_c +
+    G\left(\bar{E}_r,\bar{\mathrm n}_i\right) - 
+    \Lambda\left(\bar{E}_r,\bar{\mathrm n}_i\right), \\
+  \label{eq:fld_chemistry_qss}
+  \partial_t {\mathrm n}_i &=& \alpha_{i,j}\left(\bar{e}_c\right)
+    {\mathrm n}_e \bar{\mathrm n}_j - {\mathrm n}_i 
+    \Gamma_i^{ph}\left(\bar{E}_r\right), \quad i=1,\ldots,N_s,
+\end{eqnarray}
+where we have defined the time-centered ``background'' states
+$\bar{E}_r = \left(E_r^{n}+E_r^{n+1}\right)/2$, 
+$\bar{\mathrm n}_i = \left({\mathrm n}_i^{n}+{\mathrm n}_i^{n+1}\right)/2$
+and $\bar{e}_c = \left(e_c^{n}+e_c^{n+1}\right)/2$.  These equations
+may each be solved {\em analytically} for their solution at the time
+$t$, which we denote by
+\begin{eqnarray}
+  \label{eq:fld_heating_qss_sol}
+  e_c(t) &=& \text{sol}_{e}\left(t,\bar{E}_r,\bar{\mathrm n}_i,e_c^n\right), \\
+  \label{eq:fld_chemistry_qss_sol}
+  {\mathrm n}_i(t) &=& \text{sol}_{\mathrm n_i}
+  \left(t,\bar{E}_r,\bar{e_c},\mathrm n_i^n\right), \quad i=1,\ldots,N_s. 
+\end{eqnarray}
+We then define a nonlinear system of equations to compute the
+time-evolved solutions $\left(E_r^{n+1}, e_c^{n+1}, 
+{\mathrm n}_i^{n+1})\right)$ as
+\begin{eqnarray}
+  \label{eq:fld_heating_qss_fe}
+  f_e(e_c^{n+1},{\mathrm n}_i^{n+1}) &\equiv& e_c^{n+1} -
+    \text{sol}_{e}\left(t^{n+1},\bar{E}_r,\bar{\mathrm n}_i,e_c^n\right)
+    = 0, \\
+  \label{eq:fld_chemistry_qss_fn}
+  f_{\mathrm n_i}(e_c^{n+1},{\mathrm n}_i^{n+1}) &\equiv& 
+    {\mathrm n}_i(t) - \text{sol}_{\mathrm n_i}
+    \left(t,\bar{E}_r,\bar{e_c},\mathrm n_i^n\right)=0, \quad
+    i=1,\ldots,N_s.  
+\end{eqnarray}
+This system of $N_s+1$ nonlinear equations is solved using a damped
+fixed point iteration, 
+\[
+   U_i = U_i - \lambda f_i(U), \quad i=1,\ldots,N_s+1,
+\]
+where $U$ is a vector containing the solutions to equations
+(\ref{eq:fld_heating_qss_fe}-\ref{eq:fld_chemistry_qss_fn}).  In this
+iteration, for the first 50 sweeps we use $\lambda=1$, and for more
+challenging problems if that does not suffice we switch to a damping
+parameter of $\lambda=0.1$.

### 2013-01-23 numerical-analysis.tex
First set of text on inline yt.

@@
 
+Analyzing the results of simulations in detail requires both tools to ask
+sophisticated questions of data, and the ability to process vast quantities of
+data at high-cadence.  As simulations grow in size and complexity, storing data
+for post-processing becomes simply intractable.  To cope with this, we have
+instrumented Enzo with the ability to conduct analysis during the course of a
+simulation.  This enables analyzing with extremely high cadence (even as often
+as every subcycle of the finest refinement level), as well as without
+attempting to write an entire checkpoint output to disk.  The current mechanism
+for conducting analysis in Enzo during the course of the simulation utilizes
+the same compute resources as used by the simulation itself; this is often
+referred to as \textit{in situ} visualization.  Utilizing a
+dynamically-scheduled second set of computing resources, often referred to as
+\textit{co-scheduled} visualization, provides greater flexibility and overall
+throughput at the expense of simplicity.
+
+We expose Enzo's mesh geometry and fluid quantities to the analysis platform
+\texttt{yt} \citep{2011ApJS..192....9T, 2011arXiv1112.4482T}.  At compile time,
+Enzo is (dynamically or statically) linked against the Python and NumPy
+libraries necessary to create proxy objects exposing the mesh geometry,
+fluid quantities and particle arrays as NumPy arrays.  This information is then
+passed to a special handler inside \texttt{yt}.  \texttt{yt} interprets the
+mesh and fluid information and, without saving data to disk except as
+explicitly requested, constructs a native representation of the in-memory state
+of the simulation that appears identical to an on-disk simulation output.
+Enzo then executes a user-provided analysis script, which is able to access the
+in-memory simulation object.  Once the analysis script has returned control to
+Enzo, the simulation proceeds.  This process can occur at either the top of the
+main ``EvolveHierarchy'' loop or at the end of a timestep at the finest level,
+and the frequency with which it is called is adjustable by a run-time
+parameter.  During the course of conducting analysis, this simulation is halted
+until the conclusion of the analysis.
+
+Most analysis operations that can be performed on simulations that reside on
+disk can be performed on in-memory simulations in \texttt{yt}.  This includes
+projections (i.e., line integrals, both on- and off-axis), slices, 1-, 2-, and
+3-D fluid phase distributions, calculation of derived quantities and arbitrary
+data selection.  As of version 2.5 of \texttt{yt},  the Rockstar phase-space
+halo finder \citep{2013ApJ...762..109B} can be executed through \texttt{yt} on
+in-memory Enzo data.  Operations that currently cannot be conducted on
+in-memory datasets are those that require spatial decomposition of data; for
+instance, calculating marching cubes on a data object with \texttt{yt} is a
+fully local operation and can be conducted \textit{in situ}.  However,
+calculating topologically-connected sets requires a spatial decomposition of
+data and can thus not be conducted \textit{in situ}.  This prohibition extends
+to most halo finding operations, most multi-level parallelism operations,
+volume rendering and connected set identification.
+
+Where microphysical solvers or other operator-split physics calculations can be
+done in Python, \texttt{yt} can serve as a driver for these calculations.  A
+major feature set that is currently being developed is to pass structured
+(i.e., non-fluid) information back from \texttt{yt} into Enzo.  For instance,
+this could be the result of semi-analytic models of the growth and evolution of
+star clusters, galaxy particle feedback parameters that have been influenced by
+merger-tree analysis, or even spectral energy distributions that are calculated
+within \texttt{yt} and provided to Enzo as input into radiative transfer
+calculations.  Future versions will include this, as well as the ability to
+dynamically allocate compute resources to \texttt{yt} such that the simulation
+may proceed asynchronously with analysis (\textit{co-scheduled} analysis).
+With this functionality will also come the ability to dynamically partition
+data, such that spatially-decomposed operations such as volume rendering become
+feasible during the course of a simulation.
+
 \subsubsection{Tracer particles}

### 2013-01-25 development.tex
added two new sections.
adding a bunch of new sections and subsections.
Couple paragraphs on development.

--- 2013-01-18 conclusions.tex
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Conclusions}
+\section{Code development methodology}
 
-\begin{enumerate}
-\item near-future development plans
-\item far-future development plans
-\item clear areas of improvement
-\end{enumerate}
+Over time, Enzo's development has followed a trajectory toward openness.
+Started as the graduate research project of Greg Bryan at the University of
+Illinois, it has been subsequently been stewarded by the Laboratory for
+Computational Astrophysics (LCA) at the University of California at San Diego,
+and has transitioned to a distributed, completely open, and community-driven
+project.  Initially, Enzo was versioned using a series of ``snapshots'' of the
+code base, usually hand-created by the developing individuals.  These were
+distributed to collaborators and colleagues, but the central ``trunk'' of
+development was updated primarily by a single person: while patches and
+technology were accepted from external developers, the relatively small number
+of individuals using the code resulted in a strong centralization of
+development.
+
+As the stewardship of the code passed to the LCA, the code was released first
+to ``friendly users'' and then as an open source release.  However, while the
+code was made available with documentation, technology developed in the broader
+community of users was typically not re-integrated.  This led to a wide
+dispersal of development, largely independent, by individuals who downloaded
+and used the version of the code developed by the LCA.
+
+Following the first public, open source release of Enzo, the code was migrated
+to using the Subversion version control system.  This is a centralized version
+control system, and the ``stable'' Enzo source code was made globally readable
+following the Enzo 1.5. release.  Access to the primary development tree
+required a password and login for each user, and providing upstream changes
+either required this password and the granting of write access or a sequence of
+patches and manually-created diffs (much like the original development system.)
+The technical friction of manually contributing patches and modifications,
+combined with Subversion's difficulty with tracking merges, resulted in further
+fragmentation of the code base.
+
+The main development streams of Enzo were eventually merged using the
+distributed version control system (DVCS) Mercurial
+(\url{http://mercurial.selenic.com/}) into a branch of the code known as
+\texttt{week-of-code}, so named after the in-person development sprint at which
+it was created.  The fundamental, and transformative, distinction between the
+previous \textit{centralized} version control system and mercurial's
+\textit{distributed} version control system is the elimination of gatekeepers.
+While there still exists a canonical, central location that stable and
+development versions of Enzo can be obtained, changesets and versions can be
+exchanged between peers without the intervention of designated gatekeepers.
+This has the direct effect of enabling local development to be versioned and
+provenance ensured, while still retaining the ability to benefit from
+``upstream'' development.  An important, even crucial, side effect is that the
+technology used for local versioning provides mechanisms for submitting
+locally-developed modifications to the community source location.  Mercurial
+internally represents all changes as nodes in a directed, acyclic graph (DAG),
+which results in the natural ability to more consistently and easily manage
+merging development streams.
+
+Currently, Enzo is developed using the hosted source control platform Bitbucket
+(\url{http://bitbucket.org/}) at \url{http://bitbucket.org/enzo/}.  There are
+two mailing lists, one for usage-focused questions and discussion, and another
+for development discussion.  Both of these lists are open and publicly
+archived.  Bitbucket provides mechanisms for inspecting source, hosting
+branches and forks of the primary source, as well as code review.  All proposed
+source code changes for Enzo are subjected to a peer review process, where
+experienced developers read, inspect, test and provide feedback on source code
+changes.  All developers, including long-time Enzo contributors and developers,
+are subject to this process for including code in the primary Enzo repository.
+By using a remote, hosted system, Enzo is now \textit{completely} open to
+contributions from the community.  Individuals, who may or may not consider
+themselves Enzo developers, are free to ``fork'' the Enzo code base, develop
+changes (signed with their own name), and submit them for review and inclusion.
+In contrast to the centralized, gatekeeper-focused technology used previously,
+this enables anyone to contribute changes to be evaluated for inclusion in the
+Enzo codebase.
+
+While peer review is able to catch many bugs and problems with source
+code changes, Enzo is also subject to ``answer'' testing.  We have created a
+set of parameter files and problem types that exercise the underlying machinery
+of Enzo.  These ``test problems'' have affiliated ``tests,'' which consist of
+scripts that use \texttt{yt} \citep{2011ApJS..192....9T} to produce results
+such as mass distribution, projections, profiles and so on.  The testing
+infrastructure then evaluates whether the variation in the new results compared
+to a ``gold standard'' of results has exceeded an acceptable threshold,
+typically set to roughly single precision.  Optionally, for those test problems
+that are deemed unsafe to change to any precision, the tests also produce
+hashes of the outputs; these hashes will only remain unchanged in the event of
+bitwise identicality between results.  The results of the gold standard are
+versioned and stored in Amazon S3, enabling remote testing to proceed.  While
+the testing process -- building, running, analyzing and comparing -- is not yet
+automated against incoming pull requests, we hope to deploy that functionality
+in the future.  The primary challenge is that of compute time; the tests are
+organized into multiple categories, including by the expected run time, but
+the full suite of tests can take several days to run.

### 2013-01-29 numerical-conduction.tex
working on thermal conductiom sections.

@@
 great is that?
+
+Enzo implements the equations of isotropic heat conduction in a manner
+similar to that of \citet{2007ApJ...664..135P}, where we model the
+flux of isotropic heat transport:
+
+\begin{equation}
+\mathbb{F} = -\kappa_{sp} \nabla T
+\end{equation}
+
+Where $\kappa_{sp}$ is the Spitzer conduction coefficient, $4.6 \times 10^{-7}$~T$^{5/2}$ erg
+s$^{-1}$~cm$^{-1}$~K$^{-1}$ \citep{1962pfig.book.....S}, and using a
+value for the Coulomb logarithm, $\log \Lambda = 37.8$, that is
+appropriate for the intracluster medium \citep{1988xrec.book.....S}.
+It is quite possible that the local heat flux can become unphysical in
+the high-temperature, low-density cluster regime when using this
+formulation; thus, we take into account the saturation of the heat
+flux at a maximum level of
+
+\begin{equation}
+j_{sat} \simeq 0.4 n_e k_b T \left( \frac{2 k_b T}{\pi m_e} \right)^{1/2}
+\end{equation}
+
+\citep{1977ApJ...211..135C}, and to ensure a smooth transition between
+the Spitzer regime and the saturated regime we define an effective
+conductivity using the formalism of \citet{1988xrec.book.....S}
+
+\begin{equation}
+\kappa_{eff} = \frac{\kappa_{sp}}{1 + 4.2 \lambda_e / \ell_T} \; ,
+\end{equation}
+
+where $\lambda_e$ is the electron mean free path and $\ell_T \equiv T
+/ |\nabla T|$ is the characteristic length scale of the local
+temperature gradient.
+
+It is important to note that the conductivity can be strongly affected
+by the presence of magnetic fields, which may strongly suppress heat
+flow.  The strength and confuction of magnetic fields in the
+intracluster medium is poorly understood, and it is unclear to what
+extent these magnetic fields will suppress conduction.  For the
+purposes of the work described in this paper, we assume that the
+conductivity of the plasma can be described in terms of an effective
+conductivity, which can be expressed as a fraction f$_{Sp}$ of the
+Spitzer conductivity (where f$_{Sp} \leq 1.0$ are considered physical
+values).

### 2013-01-29 overview.tex
working on thermal conductiom sections.

@@
 
+\subsubsection{Magnetohydrodynamics}
+
+\subsubsection{Radiation transport: ray-casting}
+
+\subsubsection{Radiation transport: moment method}
+
+\subsubsection{Thermal conduction}
+
+Enzo implements the equations of isotropic heat conduction in a manner
+similar to that of \citet{2007ApJ...664..135P}, where we model the
+flux of isotropic heat transport:
+
+\begin{equation}
+\mathbb{F} = -\kappa_{sp} \nabla T
+\end{equation}
+
+Where $\kappa_{sp}$ is the Spitzer conduction coefficient
+\citep{1962pfig.book.....S}.  Saturation of the heat flux in
+high-temperature, low-density regimes (such as the intracluster medium
+in galaxy clusters) is taken into account.  Thermal conduction in a
+plasma can be strongly affected by the presence of magnetic field
+lines, which may strongly suppress heat flow perpendicular to the
+magnetic field.  In that case, we allow for heat transport only along
+the magnetic field lines in simulations that include this physics, as follows:
+
+\begin{equation}
+\mathbb{F} = -\kappa_{sp} \^{b} (\^{b} \cdot \nabla T)
+\end{equation}
+
+Where $\^{b}$ is the unit vector denoting the direction of the
+magnetic field.  In both the isotropic and anisotropic limits, we
+assume that conductivity can be suppressed relative to the ideal
+Spitzer conductivity, and express this as a constant multiplier.
+
 % ====================================================

### 2013-01-29 numerical-amr.tex
Adding a vector graphics version of the W-cycle figure.

@@
 \begin{center}
-\includegraphics[width=0.6\textwidth]{figures/wcycle.eps}
+%\includegraphics[width=0.6\textwidth]{figures/wcycle.eps}
+\includegraphics[width=0.6\textwidth]{figures/timestepping.eps}
 \end{center}
-\caption{\emph{Left:}  Example of a simple, distributed AMR hierarchy showing real and ghost grids.
-\emph{Right:}  Example 2D \enzo\ grid showing real and ghost zones, as 
-needed for the PPM hydro stencil. }
+\caption{\emph{Left:} Example of the timesteps on a 2-level AMR
+  hierarchy.  Enzo does not restrict the timesteps on the finer levels
+  to be a factor of $1/2^n$ smaller.  \emph{Right:} The order in which
+  the AMR grids are evaulated on each level.}
 \label{fig:wcycle}

### 2013-01-30 code-tests-fld.tex
Updated FLD section of method paper (section 3.13)

@@
 
-\red{(Dan)}
-One test for the implicit FLD: best/hardest one.
+This test problem examines Enzo's FLD radiation transport and
+associated ionization chemistry solvers on a cosmological I-front
+test.  This corresponds to the test problem 4.6
+from \cite{ReynoldsHayesPaschosNorman2009}, run using the cosmological
+deceleration parameter $q_0 = 0.05$ and initial redshift $z_i=10$.
+Here, the physics of interest is the expansion of a HII region in a
+uniform gas around a single monochromatic ionizing source (with
+frequency $h\nu = 13.6$ eV).  In this problem the I-front propagates
+rapidly at first, approaching 90\% of the Str{\" o}mgren radius by a
+scaled redshift of $-\log\left[(1+z)/(1+z_i)\right] \approx 0.15$,
+before cosmological expansion overtakes ionization, pushing the Str{\"
+o}mgren radius outward faster than the I-front can propagate.  Here,
+the Str{\" o}mgren radius is analytically given by
+\[
+   r_S(t) = \left[\frac{3\dot{N}_{\gamma}}{4\pi \alpha_B
+   n_H(t)^2}\right]^{1/3}, 
+\]
+where $\dot{N}_{\gamma} = 5\cdot10^{48}$ photon sec$^{-1}$ is the
+stength of the ionizing source, $\alpha_B = 2.52\cdot10^{-13}$
+s$^{-1}$ is the HII recombination rate, and $n_H(t)$ is the proper
+number density of hydrogen.  We then define $\lambda = \alpha_B\,
+n_{H,i}\, /\, H_0\, / (1+z_i)$, where the subscript $i$ indicates the
+quantity at the initial redshift.  The analytical solution for the
+I-front position as a function of time (or actually of the
+time/redshift-dependent expansion coefficient $a(t)$) is then
+\begin{eqnarray*}
+   r_I(t) &=& r_{S,i} \left[\lambda e^{-\tau(a)} \int_1^{a(t)}
+     e^{\tau(\tilde a)} \left(1 - 2q_0
+     + \frac{2q_0(1+z_i)}{\tilde{a}}\right)^{-1/2}\mathrm
+     d\tilde{a}\right]^{1/3}, \quad\text{where} \\ 
+   \tau(a) &=& \lambda\frac{6q_0^2(1+z_i)^2}{F(a)-F(1)}, 
+     \quad\text{and} \\
+   F(a) &=& \left[2-4q_0 - \frac{2q_o(1+z_i)}{a}\right] 
+      \left[1-2q_0 + \frac{2q_0(1+z_i)}{a}\right]^{1/2}.
+\end{eqnarray*}
+The simulation parameters are a box size $L_i\approx 27$ kpc, Hubble
+constant $H_0 = 1$, and cosmological constants $\Omega_m = 0.1$,
+$\Omega_A=0$ and $\Omega_b = 0.1$.  The initial values for the
+simulation are a radiation energy density $E_{r,i} = 10^{-35}$ erg
+cm$^{-3}$, temperature $T_i = 10^4$ K, density $\rho_{b,i} =
+2.35\cdot10^{-28}$ g cm$^{-3}$, and an ionized hydrogen fraction
+${\mathrm n}_{HII}/{\mathrm n}_H = 0$.
+
+In the Figure \ref{fig.fld}, we plot profiles of the HI and HII
+fractions at the redshifts $z=\{6.24,\, 2.29,\, 1.02\}$, as well as the
+ratio of the I-front and Str{\" o}mgren radii throughout the
+simulation.  As seen in the profile figure, the I-front width starts
+out quite sharp but slowly spreads as the simulation proceeds due to
+the diffusion approximation to the radiative transfer.  However, as
+seen in the I-front propagation history plot, the computed I-front
+location very accurately captures the exact value.
+
+\begin{figure}
+\begin{center}
+\includegraphics[width=0.45\textwidth]{figures/FLDprofiles.eps}
+\includegraphics[width=0.45\textwidth]{figures/FLDhistory.eps}
+\caption{Ionized fraction profiles and I-front propgation history for
+the cosmological radiative transfer test problem in
+Section \ref{sec.tests.fld}.}
+\label{fig.fld}
+\end{center}
+\end{figure}

### 2013-01-30 numerical-amr.tex
interim update on refinement stuff.
updating refinement criteria section (3.1.x).

@@
 
-\item{\em Baryon mass.}  This refinement criteria is identified to mimic a Lagrangian method in that it tries to keep a fixed mass resolution.
+\item{\em Baryon mass.}  This refinement criterion is identified to mimic a Lagrangian method in that it tries to keep a fixed mass resolution.
 In this method, a cell is refined if the baryonic mass in the cell ($M_g = \rho \Delta_x^d$) is larger than a specified value:
@@
 
-\item{\em Particle (dark matter) mass.}  This refinement criteria is precisely the same as the baryon overdensity criteria except it uses the grided (CIC) particle density.
+\item{\em Particle (dark matter) mass.}  This refinement criterion is precisely the same as the baryon overdensity criterion except it uses the grided (CIC) particle density.
 
-\item{\em Jeans Length.}  Following \cite{Truelove98}, the code can be forced to refine the Jeans length by a fixed number of cells, or, in more specifically, whenever the following criteria is not met:
+\item{\em Jeans Length.}  Following \cite{Truelove98}, the code can be forced to refine the Jeans length by a fixed number of cells, or, in more specifically, whenever the following criterion is not met:
 \begin{equation}
@@
 where $\epsilon_s$ is a dimensionless parameter.
+
+\item{\em Optical depth.} This refinement criterion ensures that cells on the highest level of refinement have an HI optical depth lower than unity.  The optical depth of a cell is estimated to be:
+
+\begin{equation}
+\tau = \sigma_{HI} n_{HI} \Delta x 
+\end{equation}
+
+where $\sigma_{HI}$ is the neutral atomic hydrogen absorption cross section, $n_{HI}$ is the proper HI number density, and $\Delta x$ is the cell width.
+If $\tau > 1$, the cell is refined further, up to the maximum level of refinement.
+
+\item{\em Resistive length.}  In a way similar to the Jeans length criterion described above, the code can be forced to locally resolve the resistive length scale by a minimum of a fixed number of cells.
+This is done by ensuring that the resistive length scale, defined as
+
+\begin{equation}
+L_R = |B| / |\nabla \times B|
+\end{equation}
+
+is always refined by at least N cells locally (i.e., refine a given cell further if $L_R \leq N \Delta x$).
+
+\item{\em ``Must refine region''}.
+This refinement criterion is used to ensure that all cells in a given subvolume of the simulation are refined to {\it at least} some minimum level of refinement.
+The subvolume is user-specified, and may evolve over time (e.g., to follow a structure of interest that is traveling through the simulation volume).
+
+\item{\em Metallicity.}  This refinement criterion ensures that all cells with a metallicity (defined as $Z = \equiv \rho_{Z} / \rho_{b} / 0.022$, where $\rho_{Z}$ is the density of metals in the cell, $\rho_{b}$ is the total baryon density (including metals), and 0.022 is the solar metal fraction) above a user-defined value is refined to {\it at least} some minimum level of refinement.
 

### 2013-01-30 numerical-conduction.tex
First draft of thermal conduction section done.

@@
 \label{sec.num.conductions}
-
-We've got both kinds of conduction: isotropic AND anisotropic.  How
-great is that?
 
@@
 / |\nabla T|$ is the characteristic length scale of the local
-temperature gradient.
-
-It is important to note that the conductivity can be strongly affected
+temperature gradient.  Furthermore, it is important to note that the
+overall level of thermal conductivity can be strongly affected
 by the presence of magnetic fields, which may strongly suppress heat
@@
 values).
+
+Thermal conduction in a
+plasma can be strongly affected by the presence of magnetic field
+lines, which may strongly suppress heat flow perpendicular to the
+magnetic field.  In that case, we allow for heat transport only
+parallel to
+the magnetic field lines in magnetohydrodynamic simulations.
+Mathematically, this is treated as:
+
+\begin{equation}
+\mathbb{F} = -\kappa_{sp} \^{b} (\^{b} \cdot \nabla T)
+\end{equation}
+
+Where $\^{b}$ is the unit vector denoting the direction of the
+magnetic field.  As with the isotropic thermal conduction, we allow a
+multiplicative factor f$_{Sp}$ to take into account the possible
+behavior of magnetic fields below the resolution limit of the
+simulation.
+
+Both isotropic and anisotropic thermal conduction in Enzo are treated
+in an operator-split manner.  Furthermore, within the heat transport
+module, transport along the x, y, and z directions are computed in a
+directionally-split fashion, with heat flux along each direction
+calculated at the + and - faces of the cell using the arithmetic mean
+of the cell-centered temperature in cells $n$ and $n+1$ or $n-1$ and
+$n$, respectively (empirically, this is more stable than taking the
+geometric mean of the cell-centered temperatures).  The addition of
+transport along magnetic field lines requires the calculation of
+cross-terms in the temperature derivatives at cell faces, which can
+result in spurious oscillations in the temperature field in regions
+where the temperature gradient is strong in more than one spatial
+direction.  Controlling these oscillations requires the addition of a
+flux limiter for calculations of the temperature field.  In this case,
+we choose the monotonized central difference flux limiter \citep{1977JCoPh..23..263V}, which serves to maintain numerical stability without
+sacrificing substantial speed or accuracy.

### 2013-01-30 numerical-analysis.tex
Adding shock-analysis section, including one citation and one newcommand \Mach.

@@
 \subsubsection{Shock finding}
+
+Identification of shocks and their pre- and post-shock conditions can be
+accomplished through a combination of either temperature or velocity jumps with
+dimensionally split or unsplit search methods.  The primary method used in
+\enzo~is the dimensionally unsplit temperature jump method described in detail
+in \citet{2008ApJ...689.1063S}.  We briefly outline the method below.
+
+For every cell, we first determine whether is satisfies the following conditions
+necessary to be flagged as a shock:
+\begin{eqnarray}
+\nabla \cdot \vec{v} < 0 \\
+\nabla T \cdot \nabla S > 0\\
+T_2 > T_1 \\
+\rho_2 > \rho_1,
+\end{eqnarray}
+where $\vec{v}$ is the velocity field, $T$ is the temperature, $\rho$ is the
+density, and $S=T/\rho^{\gamma-1}$ is the entropy.  An optional temperature floor
+may be chosen, useful for example in cosmological simulations without a radiation
+background.
+
+Once a cell is flagged, the local temperature gradient is calculated, which is 
+then used to traverse cells parallel to the gradient to search for the first
+pre- and post-shock cell that do not satisfy the above conditions.  If during
+the search a cell satisfying the conditions is found to have a more convergent
+flow, that cell is marked as the center, and the search is started again. Using the 
+temperature values from each of these cells, the Mach number is then solved
+using the Rankine-Hugoniot temperature jump conditions:
+\begin{equation}
+\frac{T_2}{T_1} = \frac{(5\Mach^2 - 1)(\Mach^2 + 3)}{16\Mach^2},
+\end{equation}
+where $T_2$ and $T_1$ are the post-shock~(downstream) and pre-shock~(upstream)
+temperatures, respectively. $\Mach$ is the upstream Mach number.
+
+Shock finding in the context of AMR is applied grid-by-grid.  If a search for
+pre/post-shock cells goes outside the bounds of the grid ghost zones, the
+search is stopped for that particular shocked cell. In most situations this is
+adequate since the hydrodynamic shock is captured in fewer than the number of
+ghost zones.  Shock finding can be run either upon data output, or at each step
+in the evolution of an AMR level if the Mach and pre/post-shock quantities are
+needed for additional physics modules.

### 2013-01-31 overview.tex
Switched prime and non-prime variables in overview.

@@
 \begin{eqnarray} 
-\left( \frac{\partial \rhop}{\partial t} \right)_r 
-          + \vecvp \cdot \nabla_r \rhop
-    & = & - \rhop \nabla_r \cdot \vecvp,
-        \label{eq:mass_prime} \\
+\frac{\partial \rho}{\partial t} 
+          + \vecv \cdot \nabla \rho
+    & = & - \rho \nabla \cdot \vecv,
+        \label{eq:mass} \\
 %
-\left( \frac{\partial \vecvp }{ \partial t} \right)_r 
-          + ( \vecvp \cdot \nabla_r ) \vecvp 
-    & = & - \frac{1}{\rhop} \nabla_r \pp 
-          - \nabla_r \phip,
-        \label{eq:momentum_prime} \\
+\frac{\partial \vecv }{ \partial t} 
+          + ( \vecv \cdot \nabla ) \vecv 
+    & = & - \frac{1}{\rho} \nabla p
+          - \nabla \phi,
+        \label{eq:momentum} \\
 %
-\left( \frac{\partial \Ep}{\partial t} \right)_r + \vecvp \cdot \nabla_r \Ep
-    & = & - \frac{1}{\rhop} \nabla_r \cdot (\pp \vecvp) 
-          - \vecvp \cdot \nabla_r \phip
+ \frac{\partial E}{\partial t} + \vecv \cdot \nabla E
+    & = & - \frac{1}{\rho} \nabla \cdot (p \vecv) 
+          - \vecv \cdot \nabla \phi
 %\nonumber          \\  & &     % uncomment for 2-column
      + \Gamma - \Lambda
-        \label{eq:total_energy_prime}
+        \label{eq:total_energy}
 \end{eqnarray}
@@
 Here $E$, $e$, $\rho$, $\vecv$ and $p$ are the baryonic total specific fluid energy, the baryonic specific thermal energy, the baryonic density, velocity and pressure, respectively.  
-Here we use primes to indicate variables defined in the absolute frame of reference (to distinguish them from a comoving frame).
+%Here we use primes to indicate variables defined in the absolute frame of reference (to distinguish them from a comoving frame).  
 The first equation represents conservation of mass, the second, conservation of momentum, and the third, conservation of total fluid (kinetic plus thermal) energy.
-They are respectively, the first, second and third moments of the Boltzman equation.
+They are respectively, the first, second and third moments of the Boltzmann equation.
 Terms representing radiative cooling ($\Lambda$) and heating ($\Gamma$) enter on the right-hand side of the energy equation (\ref{eq:total_energy_prime}).
 
-The total fluid specific energy ($\Ep$) is given by:
+The total fluid specific energy ($E$) is given by:
 \begin{equation}
-\Ep =  e^\prime + \frac{1}{2} \vecvp \cdot \vecvp.
-        \label{eq:total_energy_def_prime}
+E =  e + \frac{1}{2} \vecv \cdot \vecv.
+        \label{eq:total_energy_def}
 \end{equation}
@@
 %
-e^\prime    & = & \pp / [(\gamma - 1) \rhop ],
-        \label{eq:eq_of_state_prime} \\
+e   & = & p / [(\gamma - 1) \rho ],
+        \label{eq:eq_of_state} \\
 %
-\nabla_r^2 \phip & = & 4 \pi G \rho_{total}^\prime 
-%                   + 3 p_{total}^\prime/c^2) - \Lambda. 
-        \label{eq:potential_prime}
+\nabla^2 \phi & = & 4 \pi G \rho_{total}
+%                   + 3 p_{total}/c^2) - \Lambda. 
+        \label{eq:potential}
 \end{eqnarray}
 %
-For simplicity, the equation of state is shown here for an ideal gas with a ratio of specific heats $\gamma$.
-Equation (\ref{eq:potential_prime}) includes contributions to the gravitational potential $\phi^\prime$ from all  populations ($\rho_{total}^\prime$).
+The equation of state is shown here for an ideal gas with a ratio of specific heats $\gamma$.
+Equation (\ref{eq:potential}) includes contributions to the gravitational potential $\phi$ from the total mass density $\rho_{\rm total}$).
 % as well as any radiation or massless neutrino background ($p_{total}^\prime$) or cosmological constant ($\Lambda_c$), if we are interested in the cosmological case.
@@
 \vecx & = & \vecr/a, \\
-\vecv & = & a \hspace{0.5mm} d\vecx/dt = 
-              \vecvp - \dot{a}\vecx, \\
-\rho    & = & a^3 \rhop,   \\
-p       & = & a^3 \pp, \\
-E       & = & \Ep - 
-              \dot{a} \vecx \cdot \vecv - 
+\vecvp & = & a \hspace{0.5mm} d\vecx/dt = 
+              \vecv - \dot{a}\vecx, \\
+\rhop    & = & a^3 \rho,   \\
+\pp       & = & a^3 p, \\
+\Ep       & = & E - 
+              \dot{a} \vecx \cdot \vecvp - 
               \frac{1}{2} \dot{a}^2 \vecx^2, \\
-\phi    & = & \phip + \frac{1}{2} a \ddot{a} \vecx^2.
+\phip    & = & \phi + \frac{1}{2} a \ddot{a} \vecx^2.
 \end{eqnarray}
@@
 These definitions result in the following comoving equivalents to eqs. 
-(\ref{eq:mass_prime})-(\ref{eq:potential_prime}).
+(\ref{eq:mass})-(\ref{eq:potential}).
 %
@@
 \begin{eqnarray} 
-\frac{\partial \rhob}{\partial t} 
-          + \frac{1}{a} \vecvb \cdot \nabla \rhob 
-    & = & - \frac{1}{a} \rhob \nabla \cdot \vecvb, 
-        \label{eq:mass} \\
+\frac{\partial \rhop}{\partial t} 
+          + \frac{1}{a} \vecvp \cdot \nabla \rhop 
+    & = & - \frac{1}{a} \rhop \nabla \cdot \vecvp, 
+        \label{eq:mass_prime} \\
 %
-\frac{\partial \vecvb}{\partial t}  
-          + \frac{1}{a} ( \vecvb \cdot \nabla ) \vecvb 
-    & = & - \frac{\dot{a}}{a} \vecvb 
-          - \frac{1}{a \rhob} \nabla p 
-          - \frac{1}{a} \nabla \phi, 
-        \label{eq:momentum} \\
+\frac{\partial \vecvp}{\partial t}  
+          + \frac{1}{a} ( \vecvp \cdot \nabla ) \vecvp 
+    & = & - \frac{\dot{a}}{a} \vecvp 
+          - \frac{1}{a \rhop} \nabla \pp 
+          - \frac{1}{a} \nabla \phip, 
+        \label{eq:momentum_prime} \\
 %
-\frac{\partial E}{\partial t} 
-         + \frac{1}{a} \vecvb \cdot \nabla E 
-   & = & - \frac{\dot{a}}{a} (3 \frac{p}{\rhob} + \vecvb^2 )
-         - \frac{1}{a \rhob } \nabla \cdot (p \vecvb) 
+\frac{\partial \Ep}{\partial t} 
+         + \frac{1}{a} \vecvp \cdot \nabla \Ep 
+   & = & - \frac{\dot{a}}{a} (3 \frac{\pp}{\rhop} + {\vecvp \cdot \vecvp} )
+         - \frac{1}{a \rhop } \nabla \cdot (p \vecvp) 
 %\nonumber   & &   %uncomment for 2-column
-         - \frac{1}{a} \vecvb \cdot \nabla \phi 
+         - \frac{1}{a} \vecvp \cdot \nabla \phip 
                 + \Gamma - \Lambda.
-       \label{eq:total_energy} \\
+       \label{eq:total_energy_prime} \\
 %
-E  & = & e + \frac{1}{2} {\vecvb}^2,
-        \label{eq:total_energy_def} \\
+\Ep  & = & \ep + \frac{1}{2} {\vecvp \cdot \vecvp},
+        \label{eq:total_energy_def_prime} \\
 %
-e                  & = & 
-        p/\left[ \left( \gamma - 1 \right) \rhob \right],
-      \label{eq:state} \\ 
-{\nabla}^2 \phi    & = & 
-        \frac{4\pi G}{a} ( \rho_{total} - \rho_0 ).
+\ep                  & = & 
+        \pp/\left[ \left( \gamma - 1 \right) \rhop \right],
+      \label{eq:state_prime} \\ 
+{\nabla}^2 \phip    & = & 
+        \frac{4\pi G}{a} ( \rhop_{total} - \rhop_0 ).
       \label{eq:poisson}
@@
 \frac{\ddot{a}}{a} & = & 
-      - \frac{4 \pi G }{3 a^3 } (\rho_0 
-      + 3p_0/c^2) 
+      - \frac{4 \pi G }{3 a^3 } (\rhop_0 
+      + 3\pp_0/c^2) 
       + \Lambda_c /3 .
@@
 %
-Here, $\rho_0$ is the mean (baryonic and dark) comoving mass density and $p_0$ is the comoving background pressure contribution, and $\Lambda_c$ is the cosmological constant.
+Here, $\rhop_0$ is the mean (baryonic and dark) comoving mass density and $\pp_0$ is the comoving background pressure contribution, and $\Lambda_c$ is the cosmological constant.
 %$\gamma$ is the ratio of specific heats, and $\Lambda$ is the cosmological constant.
@@
 \begin{equation}
-\mathbb{F} = -\kappa_{sp} \^{b} (\^{b} \cdot \nabla T)
+\mathbb{F} = -\kappa_{sp} \hat{b} (\hat{b} \cdot \nabla T)
 \end{equation}
 
-Where $\^{b}$ is the unit vector denoting the direction of the
+Where $\hat{b}$ is the unit vector denoting the direction of the
 magnetic field.  In both the isotropic and anisotropic limits, we

### 2013-01-31 code-tests.tex
Added separate file for FLD test problem
Sedov blast test added

@@
 \red{(Elizabeth)}
-Problem type 7.  One of Alexei's test problems.  
-Also problem 7.4 in the FLASH method paper.
-This tests the hydro in the strong-shock limit.
+
+\begin{figure}
+\begin{center}
+\includegraphics[width=0.4\textwidth]{figures/sedov-ppm-slice.eps}
+\includegraphics[width=0.4\textwidth]{figures/sedov-zeus-slice.eps}
+\caption{Density slice from the Sedov Blast Test at $t = 0.07$.
+Left-hand image shows the results from PPM ({\tt HydroMethod = 0}), right-hand image shows the results from Zeus ({\tt Hydromethod = 2}).
+Notably, the Zeus shock front has progressed less far than forthe PPM run.
+This is due to energy loss when conserving internal, not total, energy.}
+\label{fig.sedov1}
+\end{center}
+\end{figure}
+
+\begin{figure}
+\begin{center}
+\includegraphics[width=\textwidth]{figures/sedov-profiles.eps}
+\caption{}
+\label{fig.sedov2}
+\end{center}
+\end{figure}
+
+The Sedov Blast Test \citep{Sedov1959} models an intense explosion, initiated by depositing thermal energy into a homogenous distribution of gas.
+The result is a strong spherical shock wave centered on the point of energy injection.
+This problem is a popular test of astrophysical numerical codes for three reasons: Firstly, it is particularly appropriate to astrophysics since it represents the situation when a supernovae explosion occurs.
+Secondly, it has an exact analytical solution whereby the shock front's radial position is given by:
+
+\begin{equation}
+r(t) = \left(\frac{E_0}{\alpha\rho_0}\right)^{1/5}t^{2/5}
+\end{equation}
+
+\noindent where $E_0$ is the initial energy injection, $\rho_0$ is the background density and $\alpha = 1.0$ for cylindrical symmetry and an ideal gas with $\gamma = 1.4$.
+For the full derivation see \citet{Sedov1959}.
+This solution makes it possible to assess how well the code performs.
+Thirdly, the spherical shock is a challenging problem for numerical codes (both grid and particle) since the shock wave increases in size as the simulation progresses and its symmetrical nature highlights any directional preferences that grid codes can succumb to.
+The test presented here is the two-dimensional version that is included in the Enzo distribution.
+The three-dimensional results from this test, both for Enzo and three other leading astrophysics codes, can be founds in \citet{Tasker2008}.
+
+In the initial state, the box contains a homogenous distribution of gas at a density of 1 (note that, in the absense of gravity, this problem is scale-free and so without units).
+Thermal energy is deposited into a single cell at the center of the box with $E_0 = 10.0$.
+The problem is completed in two dimensions with reflecting boundary conditions (the default for Enzo) and uses a boxsize of side $1$.
+For this problem, we selected a top grid of $100 \times 100$ and four levels of refinement, placed based on shock location and the slope in density and total energy (this corresponds to cell flagging methods 1 and 3 in the Enzo parameter file).
+The exception to this scheme is in the initial conditions where grids are placed directly around the injection point.
+The results were assessed at $t = 0.07$, which corresponding to a time just before the shock reaches the box edge (see Figure~\ref{fig.sedov1}).
+
+Figure~\ref{fig.sedov2} shows the radial profiles for the simulation run with the PPM hydro-solver (blue dashed line) and the Zeus hydro-solver (red dash-dot-dot line) together with the analytical solution (black solid line).
+Clockwise from the top-left are density, velocity, internal energy and pressure.
+PPM matches the analytical solution extremely well for all quantities.
+However, Zeus' shock front lags behind the analytical position.
+This can also be seen in the slices shown in Figure~\ref{fig.sedov1}.
+The cause of this discrepancy is that Zeus shows a substantial energy lost during the first few timesteps and produces a diamond-shaped, rather than spherical, shockfront during this time.
+After this, the code correctly conserves energy but this intial energy loss remains clearly visible in the position of the shock at $t = 0.07$.
+This problem was addressed directly by \citet{Clarke2010}, who attributed the source of the issue to this version of Zeus solving the internal, rather than total, energy equation.
+In situations with strong energy gradients, this choice caused an energy loss and the artificial viscosity produces the direction-dependent shockfront shape.
+In their paper, \citet{Clarke2010} presents results from an alternative version of Zeus that conserves total energy.
+This problem is less marked for smaller energy gradients and it should be noted that Zeus' stability and speed make it a highly competitive choice, despite the disagreements in this test.
 
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsubsection{Representative FLD test}
-\label{sec.tests.fld}
-\red{(Dan)}
-One test for the implicit FLD: best/hardest one.
+\input{code-tests-fld}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+\subsubsection{Representative thermal conduction test}
+\label{sec.tests.conduct}
+\red{(Brian)}
+One test for the thermal conduction (maybe 1 for isotropic, 1 for
+anisotropic).

### 2013-01-31 overview.tex
Streamlined section 2.1  More work required.

@@
 
-The usual Eulerian equations of gas dynamics with gravity are given by
+The usual Eulerian equations of ideal magnetohydrodynamics with gravity in a coordinate systems comoving the cosmological expansion are given by
 
@@
 \frac{\partial \rho}{\partial t} 
-          + \vecv \cdot \nabla \rho
-    & = & - \rho \nabla \cdot \vecv,
+          + \frac{1}{a} \vecv \cdot \nabla \rho
+    & = & - \frac{1}{a} \rho \nabla \cdot \vecv, 
         \label{eq:mass} \\
 %
-\frac{\partial \vecv }{ \partial t} 
-          + ( \vecv \cdot \nabla ) \vecv 
-    & = & - \frac{1}{\rho} \nabla p
-          - \nabla \phi,
+\frac{\partial \vecv}{\partial t}  
+          + \frac{1}{a} ( \vecv \cdot \nabla ) \vecv
+    & = & - \frac{\dot{a}}{a} \vecv 
+          - \frac{1}{a \rho} \nabla p
+          - \frac{1}{a} \nabla \phi, 
         \label{eq:momentum} \\
 %
- \frac{\partial E}{\partial t} + \vecv \cdot \nabla E
-    & = & - \frac{1}{\rho} \nabla \cdot (p \vecv) 
-          - \vecv \cdot \nabla \phi
-%\nonumber          \\  & &     % uncomment for 2-column
-     + \Gamma - \Lambda
+\frac{\partial \Ep}{\partial t} 
+         + \frac{1}{a} \vecv \cdot \nabla E
+   & = & - \frac{\dot{a}}{a} (3 \frac{p}{\rho} + {\vecv \cdot \vecv} )
+         - \frac{1}{a \rho } \nabla \cdot (p \vecv) 
+%\nonumber   & &   %uncomment for 2-column
+         - \frac{1}{a} \vecv \cdot \nabla \phi 
+                + \Gamma - \Lambda.
        \label{eq:total_energy} 
 \end{eqnarray}
+
+%\begin{eqnarray} 
+%\frac{\partial \rho}{\partial t} 
+%          + \vecv \cdot \nabla \rho
+%    & = & - \rho \nabla \cdot \vecv,
+%        \label{eq:mass} \\
+%%
+%\frac{\partial \vecv }{ \partial t} 
+%          + ( \vecv \cdot \nabla ) \vecv 
+%    & = & - \frac{1}{\rho} \nabla p
+%          - \nabla \phi,
+%        \label{eq:momentum} \\
+%%
+% \frac{\partial E}{\partial t} + \vecv \cdot \nabla E
+%    & = & - \frac{1}{\rho} \nabla \cdot (p \vecv) 
+%          - \vecv \cdot \nabla \phi
+%%\nonumber          \\  & &     % uncomment for 2-column
+%     + \Gamma - \Lambda
+%        \label{eq:total_energy}
+%\end{eqnarray}
 %
-Here $E$, $e$, $\rho$, $\vecv$ and $p$ are the baryonic total specific fluid energy, the baryonic specific thermal energy, the baryonic density, velocity and pressure, respectively.  
+Here $E$, $e$, $\rho$, $\vecv$ and $p$ are the comoving total specific fluid energy, specific thermal energy, baryonic density, velocity and pressure, respectively.  
 %Here we use primes to indicate variables defined in the absolute frame of reference (to distinguish them from a comoving frame).  
@@
 
+For completeness, we note here that as part of the usual transformation to comoving coordinates, we have defined the following comoving quantities (primes indicate variables defined in the usual, proper, coordinate system).
+
 This set of coupled hydrodynamic plus gravity equations is the standard one that \enzo solves.
@@
 \vecx & = & \vecr/a, \\
-\vecvp & = & a \hspace{0.5mm} d\vecx/dt = 
-              \vecv - \dot{a}\vecx, \\
-\rhop    & = & a^3 \rho,   \\
-\pp       & = & a^3 p, \\
-\Ep       & = & E - 
+\vecv & = & a \hspace{0.5mm} d\vecx/dt = 
+              \vecvp - \dot{a}\vecx, \\
+\rho    & = & a^3 \rhop,   \\
+p       & = & a^3 \pp, \\
+E       & = & \Ep - 
               \dot{a} \vecx \cdot \vecvp - 
               \frac{1}{2} \dot{a}^2 \vecx^2, \\
-\phip    & = & \phi + \frac{1}{2} a \ddot{a} \vecx^2.
+\phi    & = & \phip + \frac{1}{2} a \ddot{a} \vecx^2.
 \end{eqnarray}
 
-These definitions result in the following comoving equivalents to eqs. 
-(\ref{eq:mass})-(\ref{eq:potential}).
+%These definitions result in the following comoving equivalents to eqs. 
+%(\ref{eq:mass})-(\ref{eq:potential}).
+%%
 %
-
-\begin{eqnarray} 
-\frac{\partial \rhop}{\partial t} 
-          + \frac{1}{a} \vecvp \cdot \nabla \rhop 
-    & = & - \frac{1}{a} \rhop \nabla \cdot \vecvp, 
-        \label{eq:mass_prime} \\
-%
-\frac{\partial \vecvp}{\partial t}  
-          + \frac{1}{a} ( \vecvp \cdot \nabla ) \vecvp 
-    & = & - \frac{\dot{a}}{a} \vecvp 
-          - \frac{1}{a \rhop} \nabla \pp 
-          - \frac{1}{a} \nabla \phip, 
-        \label{eq:momentum_prime} \\
-%
-\frac{\partial \Ep}{\partial t} 
-         + \frac{1}{a} \vecvp \cdot \nabla \Ep 
-   & = & - \frac{\dot{a}}{a} (3 \frac{\pp}{\rhop} + {\vecvp \cdot \vecvp} )
-         - \frac{1}{a \rhop } \nabla \cdot (p \vecvp) 
-%\nonumber   & &   %uncomment for 2-column
-         - \frac{1}{a} \vecvp \cdot \nabla \phip 
-                + \Gamma - \Lambda.
-       \label{eq:total_energy_prime} \\
-%
-\Ep  & = & \ep + \frac{1}{2} {\vecvp \cdot \vecvp},
-        \label{eq:total_energy_def_prime} \\
-%
-\ep                  & = & 
-        \pp/\left[ \left( \gamma - 1 \right) \rhop \right],
-      \label{eq:state_prime} \\ 
-{\nabla}^2 \phip    & = & 
-        \frac{4\pi G}{a} ( \rhop_{total} - \rhop_0 ).
-      \label{eq:poisson}
-\end{eqnarray}
-%
-%We use the proper peculiar baryonic velocity $\vecv_b \equiv a(t) d\vecx / dt$, proper pressure $p$, and modified gravitational potential $ \phi $ which is related to the potential in proper coordinates $\Phi$ by
-%$$\phi \equiv \Phi + \frac{1}{2} a \ddot{a} \vecx^2. $$
-%The density, however, is comoving:
-%$$\rho_b \equiv \rho_{b,proper}a(t)^3, $$
+%\begin{eqnarray} 
+%\frac{\partial \rhop}{\partial t} 
+%          + \frac{1}{a} \vecvp \cdot \nabla \rhop 
+%    & = & - \frac{1}{a} \rhop \nabla \cdot \vecvp, 
+%        \label{eq:mass_prime} \\
+%%
+%\frac{\partial \vecvp}{\partial t}  
+%          + \frac{1}{a} ( \vecvp \cdot \nabla ) \vecvp 
+%    & = & - \frac{\dot{a}}{a} \vecvp 
+%          - \frac{1}{a \rhop} \nabla \pp 
+%          - \frac{1}{a} \nabla \phip, 
+%        \label{eq:momentum_prime} \\
+%%
+%\frac{\partial \Ep}{\partial t} 
+%         + \frac{1}{a} \vecvp \cdot \nabla \Ep 
+%   & = & - \frac{\dot{a}}{a} (3 \frac{\pp}{\rhop} + {\vecvp \cdot \vecvp} )
+%         - \frac{1}{a \rhop } \nabla \cdot (p \vecvp) 
+%%\nonumber   & &   %uncomment for 2-column
+%         - \frac{1}{a} \vecvp \cdot \nabla \phip 
+%                + \Gamma - \Lambda.
+%       \label{eq:total_energy_prime} \\
+%%
+%\Ep  & = & \ep + \frac{1}{2} {\vecvp \cdot \vecvp},
+%        \label{eq:total_energy_def_prime} \\
+%%
+%\ep                  & = & 
+%        \pp/\left[ \left( \gamma - 1 \right) \rhop \right],
+%      \label{eq:state_prime} \\ 
+%{\nabla}^2 \phip    & = & 
+%        \frac{4\pi G}{a} ( \rhop_{total} - \rhop_0 ).
+%      \label{eq:poisson}
+%\end{eqnarray}
+%%
+%%We use the proper peculiar baryonic velocity $\vecv_b \equiv a(t) d\vecx / dt$, proper pressure $p$, and modified gravitational potential $ \phi $ which is related to the potential in proper coordinates $\Phi$ by
+%%$$\phi \equiv \Phi + \frac{1}{2} a \ddot{a} \vecx^2. $$
+%%The density, however, is comoving:
+%%$$\rho_b \equiv \rho_{b,proper}a(t)^3, $$
 The expansion parameter $a \equiv 1/(1 + z)$ follows the expansion of a smooth, homogeneous background, where $z$, the redshift, is a function only of $t$.
@@
 
-It is clear that these equation are equivalent to the absolute coordinate version in the limit in which $a = 1$ and $\dot{a} = 0$.
-In the rest of this paper we will employ only the unprimed variables with the understanding that we are not restricting ourselves to cosmological applications.
+It is clear that the evolution equation are equivalent to the fixed coordinate version in the limit in which $a = 1$ and $\dot{a} = 0$.
+We include the cosmological terms with the understanding that we are not restricting ourselves to cosmological applications.
 

### 2013-02-14 numerical-analysis.tex
adding section on tracer particles

@@
 
+One of the inherent drawbacks of a grid-based fluid method is the
+inability to follow the evolution of a single parcel of fluid as it
+travels through the simulation volume.  To address this, Enzo has the
+capability to introduce Lagrangian ``tracer particles'' into a
+calculation either at the beginning of the simulation or when
+restarting the calculation.  These tracer particles are put into the
+simulation in a rectangular solid volume with uniform, user-specified
+spacing.  Each particle's position and velocity is updated over the
+course of a single timestep $\Delta t$ as follows:
+
+\begin{eqnarray}
+\label{eqn.drifttrace}
+x^{n+1/2} & = & x^n + (\Delta t/2) v^{interp,n} \nonumber \\
+v^{n+1} & = & v^{interp,n+1} \\
+x^{n+1} & = & x^{n+1/2} + (\Delta t/2) v^{n+1} \nonumber
+\end{eqnarray}
+
+This is essentially a drift-kick-drift particle update from time $n$
+to time $n+1$ -- however, instead of computing an acceleration at the
+half-timestep $t^n + \Delta t/2$ (as is done for massive particles --
+see Equation~\ref{eqn.driftkick}), the particle's velocity is updated
+both at the beginning of each timestep and at the half-timestep
+by linearly interpolating the cell-centered baryon velocity to the
+position of the particle ($v^{interp}$), and assigning it to that value.
+
+At user-specified intervals, the tracer particles output their data
+in a streaming format into a binary file.  The data written out
+typically includes the tracer particle's unique ID, position, velocity, density,
+and temperature, but is easily extensible to output any grid-based
+quantity that a user requires.  This capability has been used quite
+effectively in several papers, including \citep{2010ApJ...715.1575S, 2012ApJ...748...12S}.
+
 \subsubsection{Shock finding}

### 2013-02-15 numerical-timestep.tex
working on timesteps.

@@
 \begin{equation}
+\Delta t_{MHD} = min \left( \kappa_{MHD} \frac{a \Delta x}{v_{f} + |v_x|} \right)_L ,
+\label{eqn:dtMHD}
+\end{equation}
+
+\begin{equation}
 \Delta t_{dm} = min \left(\kappa_{dm} \frac{a \Delta x}{v_{dm,x}} \right)_L ,
 \label{eqn:dtdarkmatter}
+\end{equation}
+
+\begin{equation}
+\Delta t_{accel} = min \left( \sqrt{\frac{\Delta x}{\vec{g}}} \right)_L ,
+\label{eqn:dtaccel}
+\end{equation}
+
+\begin{equation}
+\Delta t_{rad} = min \left(  \sqrt{\frac{\Delta x}{\vec{a}_{rad}} \right)_L,
+\label{eqn:dtrad}
+\end{equation}
+
+\begin{equation}
+\Delta t_{cond} = min \left(  \kappa_{cond} \frac{\Delta x^2
+    n}{\kappa(T)} \right)_L,
+\label{eqn:dtcond}
 \end{equation}
@@
 
-\begin{equation}
-\Delta t_{accel} = min \left( \sqrt{\frac{\Delta x}{\vec{g}}} \right)_L 
-\label{eqn:dtaccel}
-\end{equation}
+% \begin{equation}
+% \Delta t_{} = 
+% \label{eqn:dt}
+% \end{equation}
 
- In equations~\ref{eqn:dthydro}-\ref{eqn:dtaccel}, the $min ( \ldots )_L$
-formalism means that this value is calculated for all cells on a given level
-L and the minimum overall value is taken as the timestep.  
+% \begin{equation}
+% \Delta t_{} = 
+% \label{eqn:dt}
+% \end{equation}
+
+ In equations~\ref{eqn:dthydro}-\ref{eqn:dtcond}, the $min ( \ldots
+)_L$ formalism means that this value is calculated for all cells on a
+given level L and the minimum overall value is taken as the timestep.
 Equation~\ref{eqn:dthydro} ensures that all cells satisfy the
-Courant-Freidrichs-Levy (CFL) condition for accuracy and stability of an explicit
-finite difference discretization of the Euler equations.  Effectively this condition
-forces the timestep to be small enough such that any changes in the fluid propagate
-less than a single grid spacing, $\Delta x$.  In this equation, $\kappa_{hydro}$ is 
-a numerical constant with a value of $0 < \kappa_{hydro} \leq 1$ (with a typical
-value of $\kappa_{hydro} \sim 0.3-0.5$) that ensures that the CFL condition is always
-met, and $c_s$ and $v_x$ are the sound speed and peculiar baryon
-velocity in a given cell.
-
-Equation~\ref{eqn:dtdarkmatter} is analogous to Equation~\ref{eqn:dthydro} and ensures
-accuracy in the N-body solver by requiring that no dark matter particle travels more than
-one cell width.  The parameter $\kappa_{dm}$ is analogous to $\kappa_{dthydro}$, with
-an identical range of values.  Equation~\ref{eqn:dtexpand} limits the timestep such 
-that the expansion parameter a can only change by a fractional amount of $f_{exp} = \Delta a/a$, 
-where $f_{exp}$ is a user-defined parameter and has typical values of $f_{exp} = 0.01-0.02$.
-This is required for the stability of the PPM algorithm in comoving coordinates, and
-typically limits the timestep only at high redshifts when densities are relatively
-homogeneous.
-
-Equation~\ref{eqn:dtaccel} is supplementary to equation~\ref{eqn:dthydro} in that it
-takes into account the possibility of large accelerations causing numerical 
-instabilities by violating the Courant condition.  In this equation, $\vec{g}$ is the
-gravitational acceleration in each cell on level L.  \blue{ \emph{It should be
-  noted that the timestep is computed before the acceleration field is computed,
-  and the acceleration field is deleted at the end of each time step.
-  So this condition isn't ever actually enforced.  We should either
-  fix this problem, or not discuss it.
- -dcollins.} }
+Courant-Freidrichs-Levy (CFL) condition for accuracy and stability of
+an explicit finite difference discretization of the Euler equations.
+Effectively this condition forces the timestep to be small enough such
+that any changes in the fluid propagate less than a single grid
+spacing, $\Delta x$.  In this equation, $\kappa_{hydro}$ is a
+numerical constant with a value of $0 < \kappa_{hydro} \leq 1$ (with a
+typical value of $\kappa_{hydro} \sim 0.3-0.5$) that ensures that the
+CFL condition is always met, and $c_s$ and $v_x$ are the sound speed
+and peculiar baryon velocity in a given cell.
 
@@
 the minimum.
+
+Equation~\ref{eqn:dtMHD} is only enforced when the equations of
+magnetohydrodynamics are being solved, and is directly analogous to
+Equation~\ref{eqn:dthydro} in that it ensures that the CLF condition
+is being enforced at all times.  In this equation, $\kappa_{MHD}$ is a
+numerical constant with a value of $0 < \kappa_{MHD} \leq 1$ (with a
+typical value of $\kappa_{MHD} \sim 0.5$) that ensures that the CFL
+condition is always met, and $v_f$ and $v_x$ are the ``fast wave
+speed'' and peculiar baryon velocity in a given cell.  The fast wave
+speed comes from a stability analysis of the MHD equations, and is
+expressed as:
+
+\begin{equation}
+v_f = \sqrt{ \frac{1}{2} \left(  v_A^2 + c_s^2 + \sqrt{(v_A^2 +
+      c_s^2)^2 - 4 v_A^2 c_s^2}  \right)  },
+\label{eqn:vfastmhd}
+\end{equation}
+
+Where $c_s$ is the sound speed and $v_A$ is the Alfven speed, calculated
+as $v_A = \sqrt{B^2/\rho_B}$ in units where $\mu_0 = 1$.
+
+Equation~\ref{eqn:dtdarkmatter} is analogous to
+Equation~\ref{eqn:dthydro} and ensures accuracy in the N-body solver
+by requiring that no dark matter particle travels more than one cell
+width.  The parameter $\kappa_{dm}$ is analogous to
+$\kappa_{dthydro}$, with an identical range of values.
+Equation~\ref{eqn:dtexpand} limits the timestep such that the
+expansion parameter a can only change by a fractional amount of
+$f_{exp} = \Delta a/a$, where $f_{exp}$ is a user-defined parameter
+and has typical values of $f_{exp} = 0.01-0.02$.  This is required for
+the stability of the PPM algorithm in comoving coordinates, and
+typically limits the timestep only at high redshifts when densities
+are relatively homogeneous.
+
+Equation~\ref{eqn:dtaccel} and~\ref{eqn:dtrad} are supplementary to equation~\ref{eqn:dthydro} in that they
+take into account the possibility of large accelerations due to either
+gravity (Eqn.~\ref{eqn:dtaccel} or radiation pressure
+(Eqn.~\ref{eqn:dtrad}), which may cause numerical 
+instabilities by violating the Courant condition.  In Equation~\ref{eqn:dtaccel}, $\vec{g}$ is the
+gravitational acceleration in each cell on level L.  In
+Equation~\ref{eqn:dtrad}, $\vec{a}_{rad}$ is the estimated
+acceleration due to radiation pressure in each cell on level L,
+defined as
+
+\begin{equation}
+\vec{a}_{rad} = \frac{ \sum_i \frac{\dot{E_i}}{c} \hat{r_i} }{m_b} 
+\end{equation}
+
+Where the sum calculates the energy deposited in a cell during the
+previous timestep due to \textit{all photon packets} that crossed that
+cell, with $\hat{r_i}$ being a unit vector that accounting for the
+packet direction.
+
+Equation~\ref{}

### 2013-02-15 numerical-timestep.tex
Finished seciton on timesteps.

@@
 \begin{equation}
-\Delta t_{cond} = min \left(  \kappa_{cond} \frac{\Delta x^2
-    n}{\kappa(T)} \right)_L,
+\Delta t_{cond} = min \left(  \frac{ \kappa_{cond}}{f_{Sp}} \frac{\Delta x^2
+    n_b}{\kappa(T)} \right)_L,
 \label{eqn:dtcond}
@@
 
-Equation~\ref{}
+Equation~\ref{eqn:dtcond} is the CFL condition for an explicit
+solution to the equation of heat conduction.  In this expression,
+$n_b$ is the baryon number density, defined as $n_b = \rho_b / \mu
+m_p$, where $\mu$ is the mean molecular weight, $\kappa(T)$ is the
+Spitzer thermal conductivity, and f$_{Sp}$ is a user-defined
+conduction suppression factor whose value must be f$_{Sp} \leq 1$.  Both are defined in
+Section~\ref{sec.num.conductions}.  \kappa_{cond} is a prefactor whose
+value must be $0 < \kappa_{cond}  < 0.5$, and is exactly 0.5 for the
+implemention in Enzo.
+From a practical
+perspective, it is useful to note that, unlike other timestep criteria
+discussed above (which effectively scale as $\Delta x / \sqrt{T}$,
+the timestep criterion due to thermal conduction scales as $\Delta x^2
+/ T^{2.5}$, which can result in a rapid decrease in timestep in
+regions of high resolution and/or temperature.
+
+Finally, Equation~\ref{eqn:dtexpand} is a cosmological constraint that
+limits the timestep so that the simulated universe only expands by
+some fractional amount, f$_{exp}$, during a single step.  In this
+equation, $a$ and $\dot{a}$ refer to the scale factor of the universe
+and its rate of change, respectively.  This
+criteria is necessary because the expansion of the universe and its
+first derivative with respect to time both appear in the equations of cosmological (magneto)hydrodynamics and particle motion,
+and some limit is required in order to model the change of energy and
+momentum due to expansion with a reasonable level of accuracy.  In
+reality, this is only a significant limitation at very high redshift;
+at later times, other timestep criteria are much more important.

### 2013-03-09 parallel.tex
Adding description of perfromance measurement implementation.

@@
 
+\subsubsection{Performance Measurement \& Instrumentation \red{(Sam)}}
+
+Because of the wide variety of simulations, methods, and uses of \bb{enzo}, it
+is difficult to define exactly which routines are most costly during a given
+simulation.  As such, we have designed a lightweight registering system that
+has been implemented for the most commonly used routines (such as the
+hydrodynamic and gravity solvers) as well as refinement level timers that
+measure the (exclusive) time spent on each level.  Beyond this minimal set of
+routines, we have designed a simple way for the user to modify the source
+by adding TIMER_START(``Your Routine Name'') and TIMER_END(``Your Routine
+Name'').  These timers are automatically registered in a
+std::map\footnote{http://www.cplusplus.com/reference/map/map/}.  These timers
+are created and managed individually on each processor in an asynchronous
+fashion.
+
+At each complete EvolveHierarchy (or less often if specified), each timer is
+then communicated to the root processor where it calculates the mean, standard
+deviation, minimum, and maximum for each of the timers across all processors. 
+For level timers, there are also attributes such as the number of cell updates,
+the current number of grids, and the average cells/s/MPI process.  This
+information is then output to a ``performance.out'' logfile.  This provides a 
+simplified interface to the user that can be used to diagnose performance 
+issues as well as estimate a given problem type's scalability.  In addition to 
+the logfile, we have developed a plotting interface for quickly producing 
+figures that process the data from the logfile.  These capabilities are 
+described in the online documentation, along with a further discussion of the 
+performance measurement implementation.
+
 \subsubsection{Unigrid scaling \red{(Brian)}}

### 2013-03-09 introduction.tex

@@
 
-The numerical simulation of astronomical phenomena now plays a key role, along with observations and analytic theory, in pushing forward our understanding of the cosmos 
+The numerical simulation of astronomical phenomena now plays a key role, along with observations and analytic theory, in pushing forward our understanding of the cosmos.
 %\citep[e.g.,][]{DecadalSurvey01}.  
@@
 Those developing simulation tools must fulfill two key obligations: the first is to make those tools available to the community as a whole, much in the way that astronomical data is usually made publicly available.
-The second is document, test and refinement those methods so that they can be critically evaluated and expanded by others.
+The second is document, test and refineme those methods so that they can be critically evaluated and expanded by others.
 This paper represents our attempt to meet our obligations with respect to the \enzo\ code.

### 2013-03-09 parallel.tex
Removing the dedner texfile from the Makefile, and fixing up a few missing $'s.

@@
 routines, we have designed a simple way for the user to modify the source
-by adding TIMER_START(``Your Routine Name'') and TIMER_END(``Your Routine
-Name'').  These timers are automatically registered in a
+by adding $TIMER_START(``Your Routine Name'')$ and $TIMER_END(``Your Routine
+Name'')$.  These timers are automatically registered in a
 std::map\footnote{http://www.cplusplus.com/reference/map/map/}.  These timers

### 2013-03-12 numerical-amr.tex
Adding ray tracing method section (a little more than a page).  Fixed a lot of missing references and citations.

@@
 
-Given the grid and it's boundary conditions, the process of solving equations~(\ref{eq:mass_prime})--(\ref{eq:total_energy_prime}) can be detached from the rest of the SAMR technique, so that it is possible to experiment with a number of different methods.
-The restrictions are that the method should be flux-conservative, or at least produce an accurate estimate of the flux of conserved quantities at cell interfaces.
-This is required in the flux-correction step (see \ref{sec:flux_correction}).
-There is also the issue of centering, which refers to the position of physical quantities within a cell (the primary types are cell-centered, face-centered, and node-centered).
-We will employ both the direct Eulerian version of the piecewise parabolic method \citep{1984JCoPh..54..174C,1995CoPhC..89..149B} which assumes cell-centering for all quantities (actually they are cell-averages).
-However, it is also possible to use other schemes, which employ different centering for different variables \citep[e.g.,][]{Stone92a}.
-The interpolation step, described in section~\ref{sec:interpolation}, must correctly interpolate with the selected centering.
+Given the grid and its boundary conditions, the process of solving
+equations~(\ref{eq:mass})--(\ref{eq:total_energy}) can be detached
+from the rest of the SAMR technique, so that it is possible to
+experiment with a number of different methods.  The restrictions are
+that the method should be flux-conservative, or at least produce an
+accurate estimate of the flux of conserved quantities at cell
+interfaces.  This is required in the flux-correction step (see
+\ref{sec:flux_correction}).  There is also the issue of centering,
+which refers to the position of physical quantities within a cell (the
+primary types are cell-centered, face-centered, and node-centered). We
+will employ both the direct Eulerian version of the piecewise
+parabolic method \citep{1984JCoPh..54..174C,1995CoPhC..89..149B} which
+assumes cell-centering for all quantities (actually they are
+cell-averages).  However, it is also possible to use other schemes,
+which employ different centering for different variables
+\citep[e.g.,][]{Stone92a}.  The interpolation step, described in
+section~\ref{sec:interpolation}, must correctly interpolate with the
+selected centering.
 
@@
 
-\item{\em Shear} As described in \citet{KritsukXX}, turbulence calculations can be successfully modeled with AMR if the shear is used as a refinement parameter; in particular, refinement occurs when the finite difference version of the following inequality holds:
+\item{\em Shear} As described in \citet{Kritsuk06}, turbulence calculations can be successfully modeled with AMR if the shear is used as a refinement parameter; in particular, refinement occurs when the finite difference version of the following inequality holds:
 \begin{equation}
@@
 %AMR schemes.
+
+%%% Local Variables: 
+%%% mode: latex
+%%% TeX-master: "ms"
+%%% End:

### 2013-03-12 numerical-gravity.tex
Adding ray tracing method section (a little more than a page).  Fixed a lot of missing references and citations.

@@
 After sampling dark matter density onto the grid and adding baryon density if it exists (to get the total matter density in each cell), the gravitational potential is calculated on the periodic root grid using a fast Fourier transform.
-For periodic boundary conditions, we can use either a simple Greens function kernel of $-k^{-2}$, or the finite-difference equivalent \citep{HockneyEastwoord1980}:
+For periodic boundary conditions, we can use either a simple Greens function kernel of $-k^{-2}$, or the finite-difference equivalent \citep{Hockney88}:
 \begin{equation}

### 2013-03-12 numerical-overview.tex
Adding ray tracing method section (a little more than a page).  Fixed a lot of missing references and citations.

@@
 The gravitational potential is computed by solving the elliptic 
-Poisson's equation (Eqtn.~\ref{enzopoisson}).
+Poisson's equation (Eq.~\ref{eq:potential}).
 
@@
  \citep{2003ApJ...590L...1K}, and the other is based on the method of 
-\citet{1992ApJ...399L.113C}, which has been modified for use in an AMR code.  The basic
+\citet{CO1992}, which has been modified for use in an AMR code.  The basic
 ideas behind the methods are straightforward and observationally motivated.
@@
 on galactic scales.  This method, first described by
-\citet{1992ApJ...399L.113C}, is similar in some ways to the Kravtsov algorithm
+\citet{CO1992}, is similar in some ways to the Kravtsov algorithm
 but has more complex criteria for determining where stars should

### 2013-03-12 numerical-radiation-raytracing.tex
Adding ray tracing method section (a little more than a page).  Fixed a lot of missing references and citations.

@@
 \label{sec.num.raytracing}
+
+Local radiation from stars and black holes strongly affects their
+surroundings.  Radiation transport is a well-studied problem; however,
+its treatment in multidimensional calculations is difficult because of
+the dependence on seven variables -- three spatial, two angular,
+frequency, and time.  The non-local nature of the thermal and
+hydrodynamical response to radiation sources further adds to the
+difficulty.  The radiative transfer equation in comoving coordinates
+\citep{Gnedin97} reads
+%
+\begin{equation}
+  \label{eq:rteqn}
+  \frac{1}{c} \; \frac{\partial I_\nu}{\partial t} + 
+  \frac{\hat{n} \cdot \nabla I_\nu}{\bar{a}} -
+  \frac{H}{c} \; \left( \nu \frac{\partial I_\nu}{\partial \nu} -
+  3 I_\nu \right) = -\kappa_\nu I_\nu + j_\nu .
+\end{equation}
+%
+Here $I_\nu \equiv I(\nu, \mathbf{x}, \Omega, t)$ is the radiation
+specific intensity in units of energy per time $t$ per solid angle per
+unit area per frequency $\nu$.  $H = \dot{a}/a$ is the Hubble
+constant, where $a$ is the scale factor.  $\bar{a} = a/a_{em}$ is the
+ratio of scale factors at the current time and time of emission.  The
+second term represents the propagation of radiation, where the factor
+$1/\bar{a}$ accounts for cosmic expansion.  The third term describes
+both the cosmological redshift and dilution of radiation.  On the
+right hand side, the first term considers the absorption coefficient
+$\kappa_\nu \equiv \kappa_\nu(\mathbf{x},\nu,t)$, and the second term
+$j_\nu \equiv j_\nu(\mathbf{x},\nu,t)$ is the emission coefficient
+that includes any point sources of radiation or diffuse radiation.
+
+We can make some appropriate approximations to reduce the complexity
+of Equation (\ref{eq:rteqn}) in order to include radiation transport
+in numerical calculations.  Typically timesteps in dynamic
+calculations are small enough so that $\Delta a/a \ll 1$, therefore
+$\bar{a} = 1$ in any given timestep, reducing the second term to
+$\hat{n} \partial I_\nu/\partial \mathbf{x}$.  To determine the
+importance of the third term, we evaluate the ratio of the third term
+to the second term.  This is $HL/c$, where $L$ is the simulation box
+length.  If this ratio is $\ll 1$, we can ignore the third term.  For
+example at $z=5$, this ratio is 0.1 when $L = c/H(z=5)$ = 53 proper
+Mpc.  In large boxes where the light crossing time is comparable to
+the Hubble time, then it becomes important to consider cosmological
+redshifting and dilution of the radiation.  Thus equation
+(\ref{eq:rteqn}) reduces to the non-cosmological form in this local
+approximation,
+%
+\begin{equation}
+  \frac{1}{c} \frac{\partial I_\nu}{\partial t} + 
+  \hat{n} \frac{\partial I_\nu}{\partial \mathbf{x}} =
+  -\kappa_\nu I_\nu + j_\nu .
+\end{equation}
+%
+We choose to represent the source term $j_\nu$ as point sources of
+radiation (e.g. stars, quasars) that emit radial rays that are
+propagated along the direction $\hat{n}$.
+
+Ray tracing is an accurate method to propagate radiation from point
+sources on a computational grid, given that there are a sufficient
+number of rays passing through each cell.  Along a ray, the radiation
+transfer equation reduces to
+%
+\begin{equation}
+\label{eqn:rtray}
+\frac{1}{c} \frac{\partial P}{\partial t} + \frac{\partial P}{\partial
+  r} = -\kappa P,
+\end{equation}
+where $P$ is the photon number flux along the ray.  To sample the
+radiation field at large radii, ray tracing requires at least $N_{ray}
+= 4\pi R^2 / (\Delta x)^2$ rays to sample each cell with one ray,
+where $R$ is the radius from the source to the cell and $\Delta x$ is
+the cell width.  If one were to trace $N_{ray}$ rays out to $R$, the
+cells at a smaller radius $r$ would be sampled by, on average,
+$(r/R)^2$ rays, which is computationally wasteful because only a few
+rays per cell, as we will show later, provides an accurate calculation
+of the radiation field.
+
+We avoid this inefficiency by utilizing adaptive ray tracing
+\citep{Abel02_RT}, which progressively splits rays when the sampling
+becomes too coarse and is based on Hierarchical Equal Area isoLatitude
+Pixelation \citep[HEALPix;][]{HEALPix}.  In this scheme, the rays are
+traced along normal directions of the centers of HEALPix pixels, which
+evenly divides a sphere into equal areas.  The rays are initialized at
+each point source with the photon luminosity (ph s$^{-1}$) equally
+spread across $N_{\rm pix} = 12 \times 4^l$ rays, where $l$ is the
+initial HEALPix level.  We usually find $l$ = 0 or 1 is sufficient
+because these coarse rays will usually be split before traversing
+the first cell.
+
+The rays are traced through the grid in a typical fashion
+\citep[e.g.][]{Abel99_RT}, in which we calculate the next cell
+boundary crossing.  The ray segment length crossing the cell is
+%
+\begin{equation}
+  \label{eqn:trace}
+  dr = R_0 - \min_{i=1 \rightarrow 3} \left[(x_{\rm cell,i} - x_{\rm src,i}) /
+    \hat{n}_{\rm ray,i} \right],
+\end{equation}
+%
+where $R_0$, $\hat{n}_{\rm ray}$, $x_{\rm cell,i}$, and $x_{\rm
+  src,i}$ are the initial distance traveled by the ray, normal
+direction of the ray, the next cell boundary crossing in the $i$-th
+dimension, and the position of the point source that emitted the ray,
+respectively.  However before the ray travels across the cell, we
+evaluate the ratio of the face area $A_{\rm cell}$ of the current cell
+and the solid angle $\Omega_{\rm ray}$ of the ray,
+%
+\begin{equation}
+  \label{eqn:split}
+  \Phi_c = \frac{A_{\rm cell}} {\Omega_{\rm ray}} = 
+  \frac{N_{\rm pix} (\Delta x)^2} {4\pi R_0^2}.
+\end{equation}
+%
+If $\Phi_c$ is less than a pre-determined value (usually $>3$), the
+ray is split into 4 child rays.  The pixel numbers of the child rays
+$p^\prime$ are given by the ``nested'' scheme of HEALPix at the next
+level, i.e. $p^\prime = 4 \times p + [0,1,2,3]$, where $p$ is the
+original pixel number.  The child rays (1) acquire the new normal
+vectors of the pixels, (2) retain the same radius of the parent ray,
+and (3) gets a quarter of the photon flux of the parent ray.
+Afterward the parent ray is discontinued.
+
+A ray propagates and splits until (1) the photon has traveled $c
+\times dt_P$, where $dt_P$ is the radiative transfer timestep; (2) its
+photon flux is almost fully absorbed ($>99.9\%$) in a single cell,
+which significantly reduces the computational time if the radiation
+volume filling fraction is small; (3) the photon leaves the
+computational domain with isolated boundary conditions; or (4) the
+photon travels $\sqrt{3}$ of the simulation box length with periodic
+boundary conditions.
+
+In the first case, the photon is halted at that position and saved,
+where it will be considered in the solution of $I_\nu$ at the next
+timestep.  In the next timestep, the photon will encounter a different
+hydrodynamical and ionization state, hence $\kappa$, in its path.
+Furthermore any time variations of the luminosities will be retained
+in the radiation field.  This is how this method retains the time
+derivative of the radiative transfer equation.  The last restriction
+prevents our method from considering sources external to the
+computational domain, but a uniform radiation background can be used
+in conjunction with ray tracing that adds the local radiation field to
+the background intensity.
+
+The radiation field is calculated by integrating equation
+(\ref{eqn:rtray}) along each ray, which is done by considering the
+discretization of the ray into segments.  In the following
+description, we assume the rays are monochromatic for simplicity.  For
+convenience, we express the integration in terms of optical depth
+$\tau = \int \kappa(r,t) \; dr$, and for a ray segment,
+%
+\begin{equation}
+  \label{eqn:dtau}
+  d\tau = \sigma_{\rm abs}(\nu) n_{\rm abs} dr.
+\end{equation}
+Here $\sigma_{\rm abs}$ and $n_{\rm abs}$ are the cross section and
+number density of the absorbing medium, respectively.  In the static
+case, equation (\ref{eqn:rtray}) has a simple exponential analytic
+solution, and the photon flux of a ray is reduced by
+%
+\begin{equation}
+  \label{eqn:flux}
+  dP = P \times (1 - e^{-\tau})
+\end{equation}
+as it crosses a cell.  We equate the photo-ionization rate to the
+absorption rate, resulting in photon conservation \citep{Abel99_RT,
+  Mellema06}.  Thus the photo-ionization $k_{\rm ph}$ and
+photo-heating $\Gamma_{\rm ph}$ rates associated with a single ray are
+%
+\begin{equation}
+  \label{eqn:kph}
+  k_{\rm ph} = \frac{P (1 - e^{-\tau})}{n_{\rm abs} \; V_{\rm cell} \; dt_P},
+\end{equation}
+\begin{equation}
+  \label{eqn:gamma}
+  \Gamma_{\rm ph} = k_{\rm ph} \; (E_{\rm ph} - E_i),
+\end{equation}
+where $V_{\rm cell}$ is the cell volume, $E_{\rm ph}$ is the photon
+energy, and $E_i$ is the ionization energy of the absorbing material.
+In each cell, the photo-ionization and photo-heating rates from each
+ray in the calculation are summed, and after the ray tracing is
+complete, these rates are used to update the ionization state and
+energy of the cells with the solver described in \S\ref{sec.ov.chem}.
+
+%%% Local Variables: 
+%%% mode: latex
+%%% TeX-master: "ms"
+%%% End:

### 2013-03-12 overview.tex
Adding ray tracing method section (a little more than a page).  Fixed a lot of missing references and citations.

@@
 They are respectively, the first, second and third moments of the Boltzmann equation.
-Terms representing radiative cooling ($\Lambda$) and heating ($\Gamma$) enter on the right-hand side of the energy equation (\ref{eq:total_energy_prime}).
+Terms representing radiative cooling ($\Lambda$) and heating ($\Gamma$) enter on the right-hand side of the energy equation (\ref{eq:total_energy}).
 
@@
 
-\subsubsection{Hydrodynamics: PPM method}
+\subsubsection{Godunov PPM method (Hydrodynamics)}
 
@@
 
-\subsubsection{Hydrodynamics: ZEUS method}
+\subsubsection{Godonuv MUSCL with Dednet (HD and MHD)}
+
+\subsubsection{MHDCT}
+
+\subsubsection{Second order finite difference method (HD)}
 
@@
 Then the minimum of all timesteps is taken and the level is advanced with this step.
+
+%%% Local Variables: 
+%%% mode: latex
+%%% TeX-master: "ms"
+%%% End:

### 2013-03-12 parallel.tex
Adding ray tracing method section (a little more than a page).  Fixed a lot of missing references and citations.

--- 2013-03-09 parallel.tex
@@
 
-Because of the wide variety of simulations, methods, and uses of \bb{enzo}, it
-is difficult to define exactly which routines are most costly during a given
-simulation.  As such, we have designed a lightweight registering system that
-has been implemented for the most commonly used routines (such as the
-hydrodynamic and gravity solvers) as well as refinement level timers that
-measure the (exclusive) time spent on each level.  Beyond this minimal set of
-routines, we have designed a simple way for the user to modify the source
-by adding TIMER_START(``Your Routine Name'') and TIMER_END(``Your Routine
-Name'').  These timers are automatically registered in a
-std::map\footnote{http://www.cplusplus.com/reference/map/map/}.  These timers
-are created and managed individually on each processor in an asynchronous
-fashion.
+Because of the wide variety of simulations, methods, and uses of Enzo,
+it is difficult to define exactly which routines are most costly
+during a given simulation.  As such, we have designed a lightweight
+registering system that has been implemented for the most commonly
+used routines (such as the hydrodynamic and gravity solvers) as well
+as refinement level timers that measure the (exclusive) time spent on
+each level.  Beyond this minimal set of routines, we have designed a
+simple way for the user to modify the source by adding
+\texttt{TIMER\_START(``Your Routine Name'')} and
+\texttt{TIMER\_END(``Your Routine Name'')}.  These timers are
+automatically registered in a
+std::map\footnote{http://www.cplusplus.com/reference/map/map/}.  These
+timers are created and managed individually on each processor in an
+asynchronous fashion.
 
@@
 % Greg to work on this, scaling data from his runs, or maybe from John Wise
+
+%%% Local Variables: 
+%%% mode: latex
+%%% TeX-master: "ms"
+%%% End:

### 2013-03-14 code-tests.tex
Adding section on one-zone collapse test.

@@
 \includegraphics[width=\textwidth]{figures/sedov-profiles.eps}
-\caption{}
+\caption{Radially averaged profiles for the Sedov Blast Test at $t = 0.07$. Clockwise from top-left shows density, velocity, internal energy and pressure.}
 \label{fig.sedov2}
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsubsection{1-zone free-fall test}
-\label{sec.tests.freefall}
-\red{(Britton)}
-One zone free fall test.  Tests cooling, chemistry.
+\subsubsection{One-zone collapse test}
+\label{sec.tests.1-zone}
+The one-zone collapse test simulates the collapse of a
+self-gravitating gas cloud using a semi-analytic model for the
+evolution of gas density and adiabatic heat input as a function of
+time.  It is designed to test the chemistry and cooling modules over a
+wide range in densities and over physically motivated timescales.
+Because this test disables the hydrodynamic and gravity solvers and
+uses a simple model for the density evolution, it is far faster than
+running a true collapse simulation.  The density evolution is based on
+the self-similar Larson-Penston \citep{1969MNRAS.145..271L,
+  1969MNRAS.144..425P} solution for isothermal collapse with a
+modification to account for the efficiency with which the heat
+introduced by compression can be radiated away
+\citep{1983ApJ...265.1047Y}.  Our implementation, described briefly
+here, follows the work of \citet{2005ApJ...626..627O}, which one
+should consult for further details.  The density evolution is given by 
+\begin{equation}
+\frac{d\rho}{dt} = \frac{\rho}{t_{col}},
+\end{equation}
+where the collapse timescale, $t_{col}$, is
+\begin{equation} \label{eqn.tcol}
+t_{col} = \frac{t_{dyn}}{\sqrt{1 - f}},
+\end{equation}
+and $t_{dyn}$ is the dynamical time, expressed as
+\begin{equation}
+t_{dyn} = \sqrt{\frac{3 \pi}{32 G \rho}}.
+\end{equation}
+The collapse timescale is altered from the dynamical time by the
+factor, $f$ in Equation \ref{eqn.tcol}, which an approximation of the
+ratio of the pressure to the force of gravity.  The value of $f$
+depends on the effective adiabatic index, $\gamma_{ef} \equiv (\partial
+ln\ p / \partial ln\ \rho)$, which we calculate by linearly extrapolating
+from derivative values for the two previous timesteps.  For the value
+of $f$, we use the piecewise function of \citet{2005ApJ...626..627O},
+given by
+\begin{equation}
+f = \left\{
+  \begin{array}{ll}
+  0, & \gamma_{ef} < 0.83,\\
+  0.6 + 2.5 (\gamma_{ef} - 1) - 6.0 (\gamma_{ef} - 1)^{2}, & 0.83 <
+  \gamma_{ef} < 1,\\
+  1.0 + 0.2 (\gamma_{ef} - 4/3) - 2.9 (\gamma_{ef} - 4/3)^{2}, & \gamma_{ef} > 1.}
+\end{array} \right.
+\end{equation}
+The specific energy evolves as
+\begin{equation}
+\frac{de}{dt} = -p \frac{d}{dt} \frac{1}{\rho} - \Lambda,
+\end{equation}
+where $\Lambda$ is the cooling rate in units of erg s$^{-1}$ g$^{-1}$
+and energy, temperature, density, and pressure are related by the
+ideal gas law.  Figure \ref{fig.onezone} shows an example of the
+one-zone collapse test performed with an initial number density of 1
+cm$^{-3}$ and temperature of 100 K using the 12 species chemistry
+network with H, D, and He species and metal cooling rates calculated
+with Cloudy.
+
+\begin{figure} \label{fig.onezone}
+  \begin{center}
+    \includegraphics[width=1.0\textwidth]{figures/OneZoneCollapseTest.eps}
+    \caption{Evolution of temperature versus number density for a one-zone
+      collapse test for gas at metallicities from 0 to 10$^{-2} Z_{\odot}$
+      using the primordial chemistry network with tabulated metal cooling
+      rates calculated with Cloudy.}
+    \label{fig.1zone}
+  \end{center}
+\end{figure}
 

### 2013-03-14 code-tests.tex
Adding section on test suite and test runner.

@@
 
-\red{(Britton)} Description of test suite and test runner.
+\enzo\ is capable of simulating a large variety of problems types, with
+all but a small few requiring only a parameter file as sole input.
+The notable exception is the cosmology simulation, which takes as
+input initial conditions created by other codes.  These test problems
+span a wide range in complexity.  At one end of this spectrum are
+simple problems that utilize only a single component of \enzo\ and for
+which analytic solutions exist for comparison with the simulation
+results.  At the opposite end are problems that exercise a large
+portion of \enzo's machinery in concert.  Together, the available
+problem types fully cover all of \enzo's functionality.  This enables
+them to serve as a vehicle for verifying that the code's behavior
+remains stable over time.
+
+\enzo\ uses an automated testing framework that allows a user to run,
+with a single command, a set of test problems and compare the results
+against results produced by any other version of the code.  Within the
+\enzo\ source distribution, the test problem parameter files are stored
+in a nested directory structure organized according to the primary
+functionality tested, e.g., hydrodynamics, gravity, cooling, etc.
+Each parameter file is accompanied by a text file containing various
+descriptive keywords, such as the machinery tested, the
+dimensionality, and the approximate run time.  A test runner script is
+responsible for taking as input from the user a series of keywords
+with which to select a subset of all available test problems.  The
+test problems are also grouped into three suites: the quick, push, and
+full suites, each a superset of the ones named before.  The quick
+suite is considered to minimally cover the primary functionality in
+\enzo\ and is designed to be run repeatedly during the development
+process.  The push suite has slightly increased feature coverage and
+is mandated to be run before code changes are accepted into the main
+repository.  The full suite consists of all test problems which can be
+run with no additional input.  Approximate run times for the quick,
+push, and full suite are 15 minutes, 1 hour, and 60 hours,
+respectively.
+
+After the test problems are selected by the test runner script, they
+are run in succession either by with \enzo\ executable contained within
+that distribution or an external executable built from another version
+and specified by the user.  This allows for any version of the code to
+be tested with an identical set of test problems.  After running the
+test problem simulations, the test runner then performs a series of
+basic analysis tasks using the \texttt{yt} analysis toolkit
+\citep{2011ApJS..192....9T, 2011arXiv1112.4482T}.  The default
+analysis performed on all test problems includes calculation of
+various statistics, such as extrema; mean; and variance, on the fields
+present in the output data.  Custom analysis provided by scripts that
+accompany the test problem parameter files is run for special cases,
+such as when an analytical solution exists that can be compared
+against the simulation data.  After the analysis is performed, the
+results are compared against a set of gold standard results which are
+maintained on a website and downloaded on the fly by the test runner
+script.  Alternately, results from any version of \enzo\ can be stored
+locally and compared to any other version of the code.  Below, we
+describe some of the key test problems which are used to verify proper 
+behavior.
 

### 2013-03-15 numerical-cooling.tex
Added first half of cooling section and a cooling rate figure.

@@
-\subsection{Radiative cooling}
+\subsection{Radiative Cooling and Heating}
 \label{sec.num.cooling}
+
+\enzo has multiple methods for computing the energy change from
+radiative cooling and heating.  All of them assume the gas is totally
+optically thin.  Below we describe the methods for computing the
+cooling rates from metal-free and metal-enriched gas.  Sample cooling
+curves for each of \enzo's primary cooling methods are shown in Figure
+\ref{fig.cooling}.
+
+\subsubsection{Primordial Cooling}
+
+As discussed in \S\ref{sec.ov.chem}, the set of reactions that
+characterize a metal-free gas is simple enough to be computed in
+non-equilibrium during the simulation.  Similarly, the radiative
+cooling of metal-free gas is solved by directly computing the cooling
+and heating rates from the following individual processes for atomic
+H and He: collisional excitation and ionization, recombination,
+free-free emission, Compton scattering off of the cosmic microwave
+background (CMB), and photo-heating from a UV metagalactic background.
+If the H$_{2}$ chemistry network is enabled, the following H$_{2}$
+cooling processes are also considered: ro-vibrational transitions 
+\citep{1998A&A...335..403G}, heating and cooling from molecular
+formation and destruction \citep{2009Sci...325..601T}, and 
+collision-induced emission \citep{2004MNRAS.348.1019R}.  If Deuterium
+chemistry is enabled, then rotational transitions of HD
+\citep{1998A&A...335..403G} are treated as well.  The radiative
+cooling calculation is coupled to the update of the chemistry network
+such that they both occur within the same subcycling loop.  In
+addition to the subcycle timestepping contraints mentioned in
+\S\ref{sec.ov.chem}, the subcycle timestep is also not permitted to
+exceed 10\% of the cooling time, $e/\dot{e}$.  A metagalactic
+background affects the gas through both photo-heating and
+photo-ionization.  These are treated by including redshift-dependent 
+photo-ionization and photo-heating rate terms in the chemistry and
+cooling equations for HI, HeI, and HeII.  More detail on the specific
+UV backgrounds present in \enzo is given below.
+
+\subsubsection{Metal Cooling}
+
+\begin{figure} \label{fig.cooling}
+  \begin{center}
+    \includegraphics[width=1.0\textwidth]{figures/cooling_rate.eps}
+    \caption{Radiative cooling rates from the various cooling methods
+      available in \enzo.  The black curves show cooling rates from a gas
+      with primordial composition using the non-equilibrium chemistry
+      network.  The blue curves are for a gas with metallicity of 0.5
+      $Z_{\odot}$ computed with the Cloudy cooling method.  The solid
+      black and blue lines assume collisional processes only while the
+      dashed lines include photo-ionization and photo-heating from a UV
+      metagalactic background at $z = 0$ with a gas number density of
+      10$^{-4}$ cm$^{-3}$.  The low temperature rates shown by the
+      dashed lines indicate a net heating.  The red curve is the
+      tabulated cooling function of \citet{SW87} which assumes a fully
+      ionized gas with metallicity of 0.5 $Z_{\odot}$.}
+  \end{center}
+\end{figure}

### 2013-03-15 numerical-starformation.tex
Adding ray tracing method section (a little more than a page).  Fixed a lot of missing references and citations.
working on star formation section, 1st checkin.

@@
 
-%\red{By the time this paper is submitted my (BWO's) paper on the star formation
-%and feedback algorithms, and the various parameters contained therein, should be
-%at least written and submitted as well.  Point to this, and maybe include a couple
-%interesting plots (perhaps in an appendix?).}
-%GB: can we really point to that paper?
-
-While the physics discussed previously is all crucial to the study of
-cosmological structure formation, most cosmological observations are of
-stars and related phenomena.  Additionally, stars eject energy and metal-
+While the physics discussed in other sections of this paper is crucial to the study of
+cosmological structure formation, most cosmological observations in
+the infrared, optical, and ultraviolet wavelengths are of
+stellar populations.  Furthermore, stars eject energy and metal-
 enriched gas throughout their lives, drastically modifying their own 
-environment.  The formation of galaxies and clusters cannot be completely
+environments' cooling properties.  Broadly speaking, the formation of galaxies and clusters cannot be completely
 modeled without including the feedback of energy and metals.  In particular,
@@
 lowers the central density and hence the X-ray luminosity of such clusters 
- \citep{1997ApJ...484L..21C}.  Therefore, the inclusion of star formation and the 
-feedback of energy and metals into the intergalactic medium in a cosmological
-code is crucial for many reasons.
+ \citep{1997ApJ...484L..21C}.  For all of these reasons, it is
+ critical to include star formation and the feedback of energy and
+ metals into the interstellar and intergalactic media in cosmological simulations.
 

### 2013-03-15 numerical-starformation.tex
intermediate checkin.

@@
 
-While the physics discussed in other sections of this paper is crucial to the study of
-cosmological structure formation, most cosmological observations in
-the infrared, optical, and ultraviolet wavelengths are of
+\subsubsection{Overview}
+% Overview of methodology
+
+While the physics discussed in other sections of this paper is crucial
+to the study of cosmological structure formation, most observations of
+galaxies in the infrared, optical, and ultraviolet wavelengths are of
 stellar populations.  Furthermore, stars eject energy and metal-
 enriched gas throughout their lives, drastically modifying their own
-environments' cooling properties.  Broadly speaking, the formation of galaxies and clusters cannot be completely
-modeled without including the feedback of energy and metals.  In particular,
-it is thought that feedback is crucial for suppressing the large numbers of
-dwarf-like galaxies that CDM theories predict 
-\citep{1991ApJ...381...14L,1991ApJ...379...52W}.  An early
-burst of star formation could remove a large fraction of cold gas from such systems
-\citep{1978MNRAS.183..341W,1991ApJ...367...45C}.  Also, the unexpectedly 
-low luminosities of small clusters and
-groups (relative to the richest clusters) is commonly explained through feedback
-\citep{1991ApJ...383..104K}.  Energy ejected during the formation of the cluster ellipticals 
-lowers the central density and hence the X-ray luminosity of such clusters 
- \citep{1997ApJ...484L..21C}.  For all of these reasons, it is
- critical to include star formation and the feedback of energy and
- metals into the interstellar and intergalactic media in cosmological simulations.
+environments' cooling properties.  Broadly speaking, the formation of
+galaxies and galaxy clusters cannot be completely modeled without
+including the feedback of energy and metals from stellar populations.
+In particular, it is thought that feedback is crucial for suppressing
+the large numbers of dwarf-like galaxies that CDM theories predict
+\citep{1991ApJ...381...14L,1991ApJ...379...52W}.  An early burst of
+star formation could remove a large fraction of cold gas from such
+systems \citep{1978MNRAS.183..341W,1991ApJ...367...45C}.  Also, the
+unexpectedly low luminosities of small clusters and groups (relative
+to the richest clusters) is commonly explained through feedback
+\citep{1991ApJ...383..104K}.  Energy ejected during the formation of
+the cluster ellipticals lowers the central density and hence the X-ray
+luminosity of such clusters \citep{1997ApJ...484L..21C}.  For all of
+these reasons, it is critical to include star formation and the
+feedback of energy and metals into the interstellar and intergalactic
+media in cosmological simulations.
 
-We have extended the \enzo\ code to include multiple models for star formation
-and feedback.  It is difficult to directly simulate the formation of individual
-stars in the context of simulations of galaxy formation and evolution due
-to the immense computational cost.  Therefore, we adopt a parametric method which attempts
-to model star forming regions in a plausible but heuristic way.  
-%One model is based on Kravtsov's method  \citep{2003ApJ...590L...1K}, and the other is 
-Our model is based on the method of  \citet{CO1992}, which 
-has been modified for use in an AMR code.  The basic
-ideas behind the methods are straightforward and observationally motivated.
-Similar (though somewhat more advanced) methods have been employed for the 
-smoothed particle hydrodynamics
-(SPH) method, most recently by Springel \& Hernquist 
-\citep{sh03a,sh03b,hs03}. This 
-method assumes a multiphase IGM and has been shown to accurately reproduce the cosmic star
-formation rate, and will be implemented into \enzo\ in the near future.
+Modeling star formation in the context of galaxy formation forces one
+to confront two fundamental problems. The dynamical range required to
+study galaxy formation and evolution is substantial, and attempting to
+directly resolve the formation of individual stars in the context of a
+Milky Way-type galaxy is computationally unfeasible.  Also, a strong
+theoretical understanding of the details of star formation is
+currently lacking.  Due to these challenges, models of star formation
+in cosmological simulations are largely phenomenological, and the
+``stars'' that are formed are in most cases meant to represent an
+ensemble stellar population (although not all; see
+Section~\ref{sec:starform_pop3}).
+% BWO: I use 'phenomenological' right above this, but could probably
+% go for 'parametric' or 'heuristic' if people object to my usage.
 
-%In the following sections we will detail both the Kravtsov and Cen \& Ostriker star formation methods separately.  
-Tests of the star formation and
-feedback algorithms will not be shown, since this work is currently in
-progress (and an extension of the AMR/SPH code comparison described in
-\citet{2005ApJS..160....1O}).
-%
-%\subsubsection{The Kravtsov star formation and feedback algorithm}
+The \enzo\ code includes approximately ten models for star formation
+and feedback.  Broadly speaking, these methods all work in similar
+ways: At a specified time interval, all grid cells that are at the
+highest local level of refinement (i.e., have no child cells) are
+examined to see if they meet a set of criteria for star formation.
+This may simply be a baryon density threshold, but may include more
+complex tests, such as an examination of local cooling and dynamical
+time scales, molecular hydrogen fraction, metallicity, and converging
+gas flows.  If the star formation criteria are met, some mass of gas
+is taken away from the cell in question and a ``star particle'' with the
+same mass is placed in the center of that cell, with the same momentum
+vector as the removed gas.  This star particle is then allowed to
+inject mass, momentum, thermal energy, metals, and possibly magnetic
+fields and/or cosmic ray populations into its local environment.  In
+general, the particle is treated as an ensemble of stars, with
+feedback properties occuring over time according to the assumed
+initial mass function of the stellar population.
 
-% GB: Is this going to be in the code????
-% GB: I assume not, so I am commenting it out.  Text to be used in another paper?
+In the following sections, we describe some of the more widely-used star formation
+and feedback methods used in \enzo.  We note that similar methods have
+been employed in many other codes used for galaxy formation, with
+comparable implementations for both star formation and feedback in
+other grid-based codes.  With regards to particle-based codes, star
+formation is broadly similar in implementation, though feedback is
+typically implemented in a very different way due to the Lagrangian
+nature of the method \citep[see, e.g.,][]{sh03a,sh03b,hs03}.
 
-%Kravtsov's method of star formation is designed to reproduce
-%the global Schmidt law of star 
-%formation~\citep{2003ApJ...590L...1K,1959ApJ...129..243S}.
-%This algorithm is deliberately minimal, and is explicitly geared
-%towards modeling star formation in a phenomenological way on kiloparsec
-%scales.  Stars are assume to form with a characteristic gas timescale 
-%$\tau_*$ such that $\dot{\rho}_* = \rho_{gas}/\tau_*$ where $\rho_{gas}$ 
-%and $\rho_*$ are the baryon gas and stellar densities, respectively.  
-%This ``constant efficiency'' model on the scale of star formation 
-%regions is well motivated observationally \citep{1996AJ....112.1903Y,2002ApJ...569..157W}.  
-%Star formation is only allowed to take place in very dense regions
-%with $\rho_{gas} \geq \rho_{SF}$, where $\rho_{SF}$ is a constant
-%proper density threshold above which star formation is allowed to
-%occur.  No other criteria are imposed.  Kravtsov's typical choices
-%for $\tau_*$ and $\rho_{SF}$ are $\tau_* = 4$ Gyr and $\rho_{SF} =
-%1.64~$M$_\odot$~pc$^{-3}$~$(n_H \sim 50$~cm$^{-3})$.  The adopted timescale
-%is derived from the observationally-determined normalization of the
-%Schmidt law, and the density threshold is determined by observations
-%of star forming regions on $\sim 100$ pc scales.  Note that the density
-%threshold is in proper, not comoving, units.
-
-%Algorithmically, the star formation events in Kravtsov's code
-%are assumed to occur once every global time step 
-%$\Delta t_0 \leq 10^7$.  In cells where star formation is
-%determined to occur (i.e. $\rho_{gas} \geq \rho_{SF}$) a
-%collisionless ``star particle'' is assumed to form, with a 
-%mass $m_* = \dot{\rho}_* V_{cell} \Delta t_0$, where
-%$\dot{\rho}_*$ is described above and $V_{cell}$ is the 
-%volume of the mesh cell.  These star particles are
-%formed instantaneously, with all $m_*$ of gas being
-%taken out of the cell and immediately deposited into the
-%star particle.  This particle is then given the velocity
-%of the gas in the cell, and thereafter treated as a 
-%collisionless particle.  The \enzo\ implementation
-%of this algorithm is similar, except that instead of 
-%forming stars only at the root grid time step, we allow 
-%stars to form at the time step of the highest level of
-%resolution at any particular point in space.  As can be seen
-%from the equation for $m_*$ above, this can result in very 
-%small stellar masses.  To avoid memory and processor time
-%issues related to having very large numbers of star particles
-%we impose a threshold mass $M_{*,min}$ such that a star
-%particle only forms if $m_* \geq M_{*,min}$.  An
-%appropriate choice of $M_{*,min}$ does not significantly 
-%change the star overall star formation history of a 
-%simulation, though it may delay the onset of star formation 
-%in a given cell relative to a simulation without a particle mass
-%threshold.
-
-%Each ``star particle'' is assumed to represent an ensemble
-%of stars and is treated as a single-age stellar population.
-%Kravtsov assumes that the IMF is described by a
-%Miller \& Scalo functional form with stellar masses between
-%$0.1$ and $100~M_\odot$ \citep{1979ApJS...41..513M}.  All stars in this
-%IMF with $M_* > 8~M_\odot$ deposit $2 \times 10^{51}$ ergs of
-%thermal energy and a mass $f_z M_*$ of metals into the
-%cell in which they form without delay, with
-%$f_z \equiv min(0.2, 0.01~M_*-0.06)$ (i.e. instantaneous
-%deposition of metals).  The definition of $f_z$ is a rough
-%approximation of the results of 
-%\citet{1995ApJS..101..181W}.
-
-%Kravtsov reports that simulations with this algorithm
-%reliably reproduce the star formation rate-gas surface
-%density relation of the Schmidt law, and are not particularly
-%sensitive to numerical parameters \citep{2003ApJ...590L...1K}.
-%He also notes that this is surprisingly insensitive to
-%the presence or absence of feedback and details of the cooling
-%and heating properties of the gas, which suggests that the
-%global star formation rate is determined by gravitationally-
-%driven supersonic turbulence (on large scales)
-%rather than stellar feedback or thermal instabilities on small 
-%scales.
-
-%\subsubsection{The Cen \& Ostriker star formation algorithm}
+% Specific example: Cen & Ostriker
+\subsubsection{Cen \& Ostriker}
+\label{sec:starform_cen}
 
@@
 on galactic scales.  This method, first described by
-\citet{CO1992}, is similar in some ways to the Kravtsov algorithm
+\citet{CO1992}, assumes that star particles
+
+s similar in some ways to the Kravtsov algorithm
 but has more complex criteria for determining where stars should
@@
 the results may be degenerate for some combinations of these parameters.
+
+% Specific example: Kravtsov
+\subsubsection{Kravtsov}
+\label{sec:starform_kravtsov}
+
+% GB: Is this going to be in the code????
+% GB: I assume not, so I am commenting it out.  Text to be used in another paper?
+
+%Kravtsov's method of star formation is designed to reproduce
+%the global Schmidt law of star 
+%formation~\citep{2003ApJ...590L...1K,1959ApJ...129..243S}.
+%This algorithm is deliberately minimal, and is explicitly geared
+%towards modeling star formation in a phenomenological way on kiloparsec
+%scales.  Stars are assume to form with a characteristic gas timescale 
+%$\tau_*$ such that $\dot{\rho}_* = \rho_{gas}/\tau_*$ where $\rho_{gas}$ 
+%and $\rho_*$ are the baryon gas and stellar densities, respectively.  
+%This ``constant efficiency'' model on the scale of star formation 
+%regions is well motivated observationally \citep{1996AJ....112.1903Y,2002ApJ...569..157W}.  
+%Star formation is only allowed to take place in very dense regions
+%with $\rho_{gas} \geq \rho_{SF}$, where $\rho_{SF}$ is a constant
+%proper density threshold above which star formation is allowed to
+%occur.  No other criteria are imposed.  Kravtsov's typical choices
+%for $\tau_*$ and $\rho_{SF}$ are $\tau_* = 4$ Gyr and $\rho_{SF} =
+%1.64~$M$_\odot$~pc$^{-3}$~$(n_H \sim 50$~cm$^{-3})$.  The adopted timescale
+%is derived from the observationally-determined normalization of the
+%Schmidt law, and the density threshold is determined by observations
+%of star forming regions on $\sim 100$ pc scales.  Note that the density
+%threshold is in proper, not comoving, units.
+
+%Algorithmically, the star formation events in Kravtsov's code
+%are assumed to occur once every global time step 
+%$\Delta t_0 \leq 10^7$.  In cells where star formation is
+%determined to occur (i.e. $\rho_{gas} \geq \rho_{SF}$) a
+%collisionless ``star particle'' is assumed to form, with a 
+%mass $m_* = \dot{\rho}_* V_{cell} \Delta t_0$, where
+%$\dot{\rho}_*$ is described above and $V_{cell}$ is the 
+%volume of the mesh cell.  These star particles are
+%formed instantaneously, with all $m_*$ of gas being
+%taken out of the cell and immediately deposited into the
+%star particle.  This particle is then given the velocity
+%of the gas in the cell, and thereafter treated as a 
+%collisionless particle.  The \enzo\ implementation
+%of this algorithm is similar, except that instead of 
+%forming stars only at the root grid time step, we allow 
+%stars to form at the time step of the highest level of
+%resolution at any particular point in space.  As can be seen
+%from the equation for $m_*$ above, this can result in very 
+%small stellar masses.  To avoid memory and processor time
+%issues related to having very large numbers of star particles
+%we impose a threshold mass $M_{*,min}$ such that a star
+%particle only forms if $m_* \geq M_{*,min}$.  An
+%appropriate choice of $M_{*,min}$ does not significantly 
+%change the star overall star formation history of a 
+%simulation, though it may delay the onset of star formation 
+%in a given cell relative to a simulation without a particle mass
+%threshold.
+
+%Each ``star particle'' is assumed to represent an ensemble
+%of stars and is treated as a single-age stellar population.
+%Kravtsov assumes that the IMF is described by a
+%Miller \& Scalo functional form with stellar masses between
+%$0.1$ and $100~M_\odot$ \citep{1979ApJS...41..513M}.  All stars in this
+%IMF with $M_* > 8~M_\odot$ deposit $2 \times 10^{51}$ ergs of
+%thermal energy and a mass $f_z M_*$ of metals into the
+%cell in which they form without delay, with
+%$f_z \equiv min(0.2, 0.01~M_*-0.06)$ (i.e. instantaneous
+%deposition of metals).  The definition of $f_z$ is a rough
+%approximation of the results of 
+%\citet{1995ApJS..101..181W}.
+
+%Kravtsov reports that simulations with this algorithm
+%reliably reproduce the star formation rate-gas surface
+%density relation of the Schmidt law, and are not particularly
+%sensitive to numerical parameters \citep{2003ApJ...590L...1K}.
+%He also notes that this is surprisingly insensitive to
+%the presence or absence of feedback and details of the cooling
+%and heating properties of the gas, which suggests that the
+%global star formation rate is determined by gravitationally-
+%driven supersonic turbulence (on large scales)
+%rather than stellar feedback or thermal instabilities on small 
+%scales.
+
+% Specific example: H2-regulated
+\subsubsection{H2-regulated}
+\label{sec:starform_H2reg}
+
+% Specific example: Pop III
+\subsubsection{Population III}
+\label{sec:starform_pop3}
+
+%\subsubsection{The Kravtsov star formation and feedback algorithm}

### 2013-03-15 introduction.tex

@@
 
-The numerical simulation of astronomical phenomena now plays a key role, along with observations and analytic theory, in pushing forward our understanding of the cosmos.
-%\citep[e.g.,][]{DecadalSurvey01}.  
-But we believe that with this role comes responsibility.
-Those developing simulation tools must fulfill two key obligations: the first is to make those tools available to the community as a whole, much in the way that astronomical data is usually made publicly available.
-The second is document, test and refineme those methods so that they can be critically evaluated and expanded by others.
-This paper represents our attempt to meet our obligations with respect to the \enzo\ code.
+There are many diverse phenomena in astrophysics and cosmology which require a high spatial and temporal dynamical range. There are a number of numerical techniques which have been used to
+model such problems.
+The most common is a gridless, fully Lagrangian method call Smoothed Particle Hydrodynamcs \citep[SPH;][]{Lucy77, SPH} which has been very successful, particularly for those situations dominated by gravity.
+Its development, however, is still at a relatively early stage when
+compared to the effort put into Eulerian, grid-based hydrodynamic schemes \citep[e.g.,][]{Woodward84}.
+
+Therefore, we would like to make use of this expertise; unfortunately, none of these schemes provide an easy way to adaptively and dynamically increase the resolution in small volumes, as required by, for example, gravitational instability.
+We have instead, adapted an idea first suggested by \citet{Berger89} in the Computational Fluid Dynamics (CFD) community, which has become known as Structured Adaptive Mesh Refinement (SAMR).
+The principle is to adaptively add (and modify) additional, finer meshes, or patches, over regions which require higher resolution.
+Since each patch has a simple geometry (here, cartesian) the hyperbolic fluid-dynamical equations can be solved on that patch with a well-developed `off-the-shelf' numerical method.
+In addition, it is relatively straightforward to add other advanced physics (e.g.
+MHD, radiation transport, etc.).
+This paper is devoted to describing how we have implemented this idea for astrophysics, including the addition of self-gravity, collisionless fluids, MHD, radiation transport and a range of other physical effects.
+
+There have been many other numerical methods described in the astronomical literature which contain elements of this idea, or have a similar aim.
+For example, an N-body solver \cite{Villumsen89} used non-adaptive meshing to increase the resolution in pre-selected regions.
+This static approach, applied to hydrodynamics, has been used extensively \citep[e.g.,]{Ruffert94, Anninos94}.
+Adding adaptivity is a more recent enhancement, and there are now a number of N-body codes which possess this feature, both with and without  hydrodynamics  \cite{Couchman91, Jessop94, Suisalu95, Splinter96, Gelato97, ART97, Truelove98, flash_method, MLAPM01,  Yahagi01, RAMSES, Quilis04, Ziegler05, Zhang06}.
+Of these, perhaps the most comparable are FLASH \citep{flash_method}, which uses grid blocks of fixed size, and RAMSES \citep{RAMSES} and ART \citep{ART97}, both of which refine individual cells.
+It is also possible to deform the grid to obtain high resolution \citep[e.g.,][]{Gnedin95, Xu97, Pen98}, and more recently
 
@@
 
-There are many diverse phenomena in astrophysics and cosmology which require a high spatial and temporal dynamical range. There are a number of numerical techniques which have been used to
-model such problems.
-The most common is a gridless, fully Lagrangian method call Smoothed Particle Hydrodynamcs \citep[SPH;][]{Lucy77, SPH} which has been very successful, particularly for those situations dominated by gravity.
-Its development, however, is still at a relatively early stage when
-compared to the effort put into Eulerian, grid-based hydrodynamic schemes \citep{Woodward84}.
+The \enzo\ code has been widely used over the lasts decade in a wide variety of problems in topics ranging from star formation to turbulence to large-scale structure.
+More than 100 peer-reviewed papers have been published based on results obtained with the code.
 
-Therefore, we would like to make use of this expertise; unfortunately, none of these schemes provide an easy way to adaptively and dynamically increase the resolution in small volumes, as required by, for example, gravitational instability.
-We have instead, adapted an idea first suggested by \citet{Berger89} in the Computational Fluid Dynamics (CFD) community, which has become known as Structured Adaptive Mesh Refinement (SAMR).
-The principle is to adaptively add (and modify) additional, finer meshes, or patches, over regions which require higher resolution.
-Since each patch has a simple geometry (here, cartesian) the hyperbolic fluid-dynamical equations can be solved on that patch with a well-developed `off-the-shelf' numerical method.
-In addition, it will be relatively straightforward to add other advanced physics (e.g.
-MHD, radiation transport beyond flux limited diffusion, etc.).
-This paper is devoted to describing how we have implemented this idea for astrophysics, including the addition of self-gravity and collisionless fluids.
-
-There have been a few other numerical methods described in the astronomical literature which contain elements of this idea, or have a similar aim.
-For example, an N-body solver \cite{Villumsen89} used
-non-adaptive meshing to increase the resolution in pre-selected.
-This static approach, applied to hydrodynamics, has been used extensively \citep[e.g.,]{Ruffert94, Anninos94}.
-Adding adaptivity is a more recent enhancement, and there are now a number of N-body codes which possess this feature, both with and without  hydrodynamics  \cite{Couchman91, Jessop94, Suisalu95, Splinter96, Gelato97, ART97, Truelove98, flash_method, MLAPM01,  Yahagi01, RAMSES, Quilis04, Ziegler05, Zhang06}.
-It is also possible to deform the grid to obtain high resolution \citep[e.g.,][]{Gnedin95, Xu97, Pen98}.
-
-The \enzo\ code has been widely used over the lasts decade in a wide variety of problems in topics ranging from star formation to turbulence to large-scale structure.
-More than 50 peer-reviewed papers have been published based on results obtained with the code.
+Finally, we note that the numerical simulation of astronomical phenomena now plays a key role, along with observations and analytic theory, in pushing forward our understanding of the cosmos.
+%\citep[e.g.,][]{DecadalSurvey01}.  
+But along with this role comes responsibility.
+We believe that those developing simulation tools must fulfill two key obligations: the first is to make those tools available to the community as a whole, much in the way that astronomical data is now regularly made publicly available.
+The second is document, test and refine those methods so that they can be critically evaluated and expanded by others.
+This paper represents one part of our attempt to meet our obligations with respect to the \enzo\ code.
 

### 2013-03-15 numerical-cooling.tex

@@
 curves for each of \enzo's primary cooling methods are shown in Figure
-\ref{fig.cooling}.
+\ref{fig.cooling_rate}.
 
@@
 cooling calculation is coupled to the update of the chemistry network
-such that they both occur within the same subcycling loop.  In
+such that they both occur within the same subcycling loop.  This is
+necessary in regimes where rapid cooling or change in the ionization
+state occur.  In
 addition to the subcycle timestepping contraints mentioned in
@@
 photo-ionization and photo-heating rate terms in the chemistry and
-cooling equations for HI, HeI, and HeII.  More detail on the specific
-UV backgrounds present in \enzo is given below.
+cooling equations for HI, HeI, and HeII.  The black curves in Figure
+\ref{fig.cooling_rate} show cooling rates for a metal-free gas with
+number density $n_{H} = 10^{-4}$ cm$^{-3}$ both with and without a
+radiation background.  More detail on the specific UV backgrounds
+present in \enzo is presented below in
+\S\ref{sec.num.rad-homogeneous}.
 
@@
 
-\begin{figure} \label{fig.cooling}
+A proper treatment of the cooling from metals is significantly more
+challenging due to the large number of chemical reactions and energy
+transitions that must be taken into account for each element.  Because
+of this, most metal cooling methods employ significant assumptions in
+order to seek out a balance between accuracy and speed.  There are two
+primary metal cooling methods available in \enzo.  The simpler of the
+two uses the analytic cooling function of \citet{SW87}, which assumes
+a fully ionized gas with a constant metallicity of 0.5 Z$_{\odot}$.
+The cooling rate produced by this cooling function is shown by the red 
+curve in Figure \ref{fig.cooling_rate}.
+
+A more sophisticated method makes use of multidimensional cooling and
+heating rate tables computed with the photo-ionization code,
+\texttt{Cloudy} \citep{1998PASP..110..761F}.  This method, detailed in
+\citet{2008MNRAS.385.1443S} and \citet{2011ApJ...731....6S}, works by
+using \texttt{Cloudy} to compute the cooling and heating rates from
+the metal species only.  The primary assumption made is that of
+ionization equilibrium.  The tables can have up to five dimensions,
+where those dimensions are density, metallicity, electron fraction,
+redshift (for an evolving metagalactic UV backgroun), and
+temperature.  Tables can be created for any abundance pattern for
+elements up to atomic number 30 (Zn) and for any incident radiation
+field.  The cooling from H and He is calculated by the non-equilibrium
+primordial chemistry and cooling solver.  The contribution from metals
+to the cooling is computed within the same subcycling loop as the
+coupled primordial chemistry and cooling solver.  The blue curves in
+Figure \ref{fig.cooling_rate} show cooling rates calculated with this
+method for a gas with number density $n_{H} = 10^{-4}$ cm$^{-3}$ and
+metallicity of 0.5 Z$_{\odot}$.
+
+\begin{figure}
   \begin{center}
     \includegraphics[width=1.0\textwidth]{figures/cooling_rate.eps}
+  \end{center}
   \caption{Radiative cooling rates from the various cooling methods
@@
     ionized gas with metallicity of 0.5 $Z_{\odot}$.}
-  \end{center}
+  \label{fig.cooling_rate}
 \end{figure}

### 2013-03-15 numerical-starformation.tex
finished star formation algorithm section (well, maybe).

@@
 ensemble stellar population (although not all; see
-Section~\ref{sec:starform_pop3}).
-% BWO: I use 'phenomenological' right above this, but could probably
-% go for 'parametric' or 'heuristic' if people object to my usage.
+Section~\ref{sec:starform_pop3}).  % BWO: I use 'phenomenological'
+right above this, but could probably % go for 'parametric' or
+'heuristic' if people object to my usage.
 
@@
 gas flows.  If the star formation criteria are met, some mass of gas
-is taken away from the cell in question and a ``star particle'' with the
-same mass is placed in the center of that cell, with the same momentum
-vector as the removed gas.  This star particle is then allowed to
-inject mass, momentum, thermal energy, metals, and possibly magnetic
-fields and/or cosmic ray populations into its local environment.  In
-general, the particle is treated as an ensemble of stars, with
-feedback properties occuring over time according to the assumed
-initial mass function of the stellar population.
+is taken away from the cell in question and a ``star particle'' with
+the same mass is placed in the center of that cell, with the same
+momentum vector as the removed gas.  This star particle is then
+allowed to inject mass, momentum, thermal energy, metals, and possibly
+magnetic fields and/or cosmic ray populations into its local
+environment.  In general, the particle is treated as an ensemble of
+stars, with feedback properties occuring over time according to the
+assumed initial mass function of the stellar population.
 
-In the following sections, we describe some of the more widely-used star formation
-and feedback methods used in \enzo.  We note that similar methods have
-been employed in many other codes used for galaxy formation, with
-comparable implementations for both star formation and feedback in
-other grid-based codes.  With regards to particle-based codes, star
-formation is broadly similar in implementation, though feedback is
-typically implemented in a very different way due to the Lagrangian
-nature of the method \citep[see, e.g.,][]{sh03a,sh03b,hs03}.
+In the following sections, we describe some of the more widely-used
+star formation and feedback methods used in \enzo.  We note that
+similar methods have been employed in many other codes used for galaxy
+formation, with comparable implementations for both star formation and
+feedback in other grid-based codes.  With regards to particle-based
+codes, star formation is broadly similar in implementation, though
+feedback is typically implemented in a very different way due to the
+Lagrangian nature of the method \citep[see,
+e.g.,][]{sh03a,sh03b,hs03}.
 
@@
 
-The Cen \& Ostriker method is a heuristic model of star formation
-on galactic scales.  This method, first described by
-\citet{CO1992}, assumes that star particles
+The Cen \& Ostriker method is a heuristic model of star formation on
+galactic scales.  This method, first described by \citet{CO1992},
+assumes that stars form in substantially overdense, converging, and
+gravitationally unstable gas.  Algorithmically, every cell at the
+locally highest level of refinement are examined at every timestep to
+see if they meet the critiera for star formation, and star particles
+are allowed to form in a cell if the following criteria are met:
 
-s similar in some ways to the Kravtsov algorithm
-but has more complex criteria for determining where stars should
-be formed.  In this method, cells that form stars must have a 
-baryon overdensity higher than some threshold 
-$\rho_b/\bar{\rho}_b \geq \eta$ 
-where $\rho_b$ is the baryon density in that cell, 
-$\bar{\rho_b}$ is the mean baryon density in the simulation,
-and $\eta$ is the user-defined overdensity threshold.
-Additionally, the gas in the cells must be contracting, 
-cooling rapidly, and gravitationally unstable, e.g.:
+\begin{equation}
+\rho_b/\bar{\rho}_b \geq \eta
+\label{cendens}
+\end{equation}
 
@@
 
-where $\vec{v}$ is the velocity of the gas in the cell, $\rho_{b}$ and 
-$\rho_{d}$ are the cell's baryon and dark matter density, respectively,
-$\rho_{total} = \rho_{b} + \rho_{d}$, $m_{b}$ and $m_{j}$ are the 
-baryonic mass in the cell and jeans mass of the cell, and C is the 
-isothermal sound speed in the cell.  
-If all of these criteria are met, 
-the mass of a star particle is calculated as \(m_{*} = m_{b} 
-\frac{ \Delta t}{ t_{dyn} } f_{*eff} \), 
-where $f_{*eff}$ is the star formation efficiency parameter.
+Where $\rho_b$ is the baryon density in that cell, $\bar{\rho_b}$ is
+the mean baryon density in the simulation, $\eta$ is the user-defined
+overdensity threshold, $\vec{v}$ is the velocity of the gas in the
+cell, $\rho_{d}$ is the dark matter density, $\rho_{total} = \rho_{b}
++ \rho_{d}$, $m_{b}$ and $m_{j}$ are the baryonic mass in the cell and
+jeans mass of the cell, and C is the isothermal sound speed in the
+cell.  If all of these criteria are met, the mass of a star particle
+is calculated as \(m_{*} = m_{b} \frac{ \Delta t}{ t_{dyn} } f_{*eff}
+\), where $f_{*eff}$ is the star formation efficiency parameter.
 
 If $m_{*}$ is greater than a minimum star mass $m_{*min}$, a particle
-is created and given several attributes:  Mass, a unique index number, 
+is created and given several attributes: mass, a unique index number,
 the time of formation $t_{form}$, the local dynamical free-fall time
@@
 $T_{dyn,min}$ which is observationally motivated and affects the
-feedback rates (see below).
-The particle is placed in the center of the cell and given 
-the same peculiar velocity as the gas in the cell, and is then treated 
-in the same manner as the dark matter particles.  An 
-amount of baryon gas corresponding to the new particle's mass is 
+feedback rates (see below).  The particle is placed in the center of
+the cell and given the same peculiar velocity as the gas in the cell,
+and is then treated in the same manner as the dark matter particles.
+An amount of baryon gas corresponding to the new particle's mass is
 removed from the cell.
 
-In addition, we have added a stochastic star formation algorithm that keeps 
-track of all of the sub-mass threshold stars that should have been created 
-and when the total mass of uncreated stars is greater than the minimum mass, 
-a star particle is created.
+% In addition, we have added a stochastic star formation algorithm that
+% keeps track of all of the sub-mass threshold stars that should have
+% been created and when the total mass of uncreated stars is greater
+% than the minimum mass, a star particle is created.
 
 The feedback of energy and metals into the baryon gas is similar to
-the Kravtsov feedback, with some important differences.  The star formation 
-algorithm creates each star particle instantaneously.  However, feedback
-should take place over a significant timescale, as all of the stars contained
-within the ``star particle'' would form over a long period of time.  Therefore,
-we assume that for the purposes of feedback that the mass of stars formed
-at a time $t$ with finite timestep $\Delta t$ is:
+the Kravtsov feedback, with some important differences.
+
+The star formation algorithm creates each star particle
+instantaneously.  However, feedback should take place over a
+significant timescale, as all of the stars contained within the ``star
+particle'' would in reality form (and massive stars would die) over a
+substantial period of time.  Therefore, we assume that for the
+purposes of feedback that the mass of stars formed at a time $t$ with
+finite timestep $\Delta t$ is:
 
@@
 During this timestep, we assume that the star particle feeds back
-metal-enriched gas and thermal energy from supernovae and from stellar winds.
-Since massive stars have very short lifetimes, we assume that there is an
-immediate feedback of some fraction $f_{SN}$ of the rest energy from the
-stars created that timestep into the baryon gas, such that 
-$E_{add} = f_{SN} \Delta m_{sf} c^2$, where c is the speed of light.
-In addition, a fraction $f_{Z*}$ of the metal from the star particle
-is fed back into the baryon gas, which takes into account the effects
-of metal recycling.  Finally, a fraction of the mass $f_{m*}$ from all
-stars (rather than just supernovae) is fed back into the gas along with
-momentum in order to simulate the mass ejection from non-exploding stars
-via stellar winds.
+metal-enriched gas and thermal energy from supernovae and from stellar
+winds.  Since massive stars have very short lifetimes, we assume that
+there is an immediate feedback of some fraction $f_{SN}$ of the rest
+energy from the stars created that timestep into the baryon gas, such
+that $E_{add} = f_{SN} \Delta m_{sf} c^2$, where c is the speed of
+light.  In addition, a fraction $f_{Z*}$ of the metal from the star
+particle is fed back into the baryon gas, which takes into account the
+effects of metal recycling.  Finally, a fraction of the mass $f_{m*}$
+from all stars (rather than just supernovae) is fed back into the gas
+along with momentum in order to simulate the mass ejection from
+non-exploding stars via stellar winds.
 
-There are six user-defined parameters in this algorithm:  three deal with the
-star formation ($\eta$, $m_{*min}$ and $T_{dyn,min}$), and three 
-deal with feedback ($f_{SN}$, $f_{Z*}$ and $f_{m*}$).  Some of these
-parameters are completely free, while others can be guided by observation
-or theory.  For example, the supernova feedback parameter, $f_{SN}$, can be 
-constrained assuming that, for every $200 M_\odot$ of stars created, one
-supernova occurs, and this event feeds back approximately $10^{51}$ ergs of
-thermal energy, giving:
+There are six user-defined parameters in this algorithm: three deal
+with the star formation ($\eta$, $m_{*min}$ and $T_{dyn,min}$), and
+three deal with feedback ($f_{SN}$, $f_{Z*}$ and $f_{m*}$).  Some of
+these parameters are completely free, while others can be guided by
+observation or theory.  For example, the supernova feedback parameter,
+$f_{SN}$, can be constrained assuming that, for every $200 M_\odot$ of
+stars created, one supernova occurs, and this event feeds back
+approximately $10^{51}$ ergs of thermal energy, giving:
 
@@
 
-The metal yield $f_{Z*}$, defined as the mass in metals produced per unit
-mass of stars created, can be constrained by a theoretical model of
- \citet{1995ApJS..101..181W}.  This model suggests that $f_{Z*} = 0.02$
-is an appropriate number.  The minimum dynamical time is set to be $T_{dyn,min} = 10^7$
-~years to agree with timescales seen in nearby OB associations.
+The metal yield $f_{Z*}$, defined as the mass in metals produced per
+unit mass of stars created, can be constrained by, e.g., the
+theoretical model of \citet{1995ApJS..101..181W}.  This model suggests
+that $f_{Z*} = 0.02$ is an appropriate number.  The minimum dynamical
+time is set to be $T_{dyn,min} = 10^7$ ~years to agree with timescales
+seen in nearby OB associations.
 
-The other parameters, such as the overdensity threshold $\eta$, minimum star
- mass $m_{*min}$, and mass ejection fraction $f_{m*}$ are not well constrained 
-either theoretically or observationally.  Indeed, $m_{*min}$ is a purely
-numerical parameter designed to keep the code from producing too many star 
-particles, and thus has no observational or theoretical counterpart.  These
-parameters have to be set by performing parameter studies and comparing to 
-observations.  Unfortunately, the range of parameter space is large, and
-the results may be degenerate for some combinations of these parameters.
+The other parameters, such as the overdensity threshold $\eta$,
+minimum star mass $m_{*min}$, and mass ejection fraction $f_{m*}$ are
+not well constrained either theoretically or observationally.  Indeed,
+$m_{*min}$ is a purely numerical parameter designed to keep the code
+from producing too many star particles, and thus has no observational
+or theoretical counterpart.  The $\eta$ parameter, on the other hand,
+nominally has physical meaning (i.e., the density above which star
+formation must occur on a very short timescale in a self-gravitating
+cloud); however, in the vast majority of simulations the densities
+that are reachable are nowhere near the densities of protostellar
+clouds, and thus this parameter becomes a rough proxy for finding the
+densest environments in a given simulation.  Given the
+simulation-dependent nature of these parameters, it is critical to set
+them carefully, and with some sensitivity testing.
 
 % Specific example: Kravtsov
-\subsubsection{Kravtsov}
+\subsubsection{Schmidt-Law method}
 \label{sec:starform_kravtsov}
 
-% GB: Is this going to be in the code????
-% GB: I assume not, so I am commenting it out.  Text to be used in another paper?
+The Kravtsov method of star particle creation is designed to reproduce
+the global Schmidt law of star formation~\citep{2003ApJ...590L...1K,
+1959ApJ...129..243S}.  This algorithm is deliberately minimal, and is
+explicitly geared towards modeling star formation in a
+phenomenological way on kiloparsec scales.  Stars are assume to form
+with a characteristic gas timescale $\tau_*$ such that $\dot{\rho}_* =
+\rho_{gas}/\tau_*$ where $\rho_{gas}$ and $\rho_*$ are the baryon gas
+and stellar densities, respectively.  This ``constant efficiency''
+model on the scale of star formation regions is well motivated
+observationally \citep{1996AJ....112.1903Y,2002ApJ...569..157W}.  Star
+formation is only allowed to take place in very dense regions with
+$\rho_{gas} \geq \rho_{SF}$, where $\rho_{SF}$ is a constant proper
+(as opposed to comoving) density threshold.  No other criteria are
+imposed.  Typical choices for $\tau_*$ and $\rho_{SF}$ are $\tau_* =
+4$ Gyr and $\rho_{SF} = 1.64~$M$_\odot$~pc$^{-3}$~$(n_H \sim
+50$~cm$^{-3})$.  The adopted timescale is derived from the
+observationally-determined normalization of the Schmidt law, and the
+density threshold is determined by observations of star forming
+regions on $\sim 100$ pc scales.
 
-%Kravtsov's method of star formation is designed to reproduce
-%the global Schmidt law of star 
-%formation~\citep{2003ApJ...590L...1K,1959ApJ...129..243S}.
-%This algorithm is deliberately minimal, and is explicitly geared
-%towards modeling star formation in a phenomenological way on kiloparsec
-%scales.  Stars are assume to form with a characteristic gas timescale 
-%$\tau_*$ such that $\dot{\rho}_* = \rho_{gas}/\tau_*$ where $\rho_{gas}$ 
-%and $\rho_*$ are the baryon gas and stellar densities, respectively.  
-%This ``constant efficiency'' model on the scale of star formation 
-%regions is well motivated observationally \citep{1996AJ....112.1903Y,2002ApJ...569..157W}.  
-%Star formation is only allowed to take place in very dense regions
-%with $\rho_{gas} \geq \rho_{SF}$, where $\rho_{SF}$ is a constant
-%proper density threshold above which star formation is allowed to
-%occur.  No other criteria are imposed.  Kravtsov's typical choices
-%for $\tau_*$ and $\rho_{SF}$ are $\tau_* = 4$ Gyr and $\rho_{SF} =
-%1.64~$M$_\odot$~pc$^{-3}$~$(n_H \sim 50$~cm$^{-3})$.  The adopted timescale
-%is derived from the observationally-determined normalization of the
-%Schmidt law, and the density threshold is determined by observations
-%of star forming regions on $\sim 100$ pc scales.  Note that the density
-%threshold is in proper, not comoving, units.
+Algorithmically, the star formation events in Kravtsov's code are
+assumed to occur once every global time step $\Delta t_0 \leq 10^7$.
+In cells where star formation is determined to occur (i.e. $\rho_{gas}
+\geq \rho_{SF}$), star particles with a mass of $m_* = \dot{\rho}_*
+V_{cell} \Delta t_0$ (where $V_{cell}$ is the volume of the mesh cell)
+are assumed to form instantaneously in a manner similar to that
+described in Section~\ref{sec:starform_cen}.  The \enzo\
+implementation of this algorithm is similar, except that instead of
+forming stars only at the root grid time step, we allow stars to form
+at the time step of the highest level of resolution at any particular
+point in space.  As can be seen from the equation for $m_*$ above,
+this can result in very small stellar masses.  To avoid memory and
+processor time issues related to having very large numbers of star
+particles we impose a threshold mass $M_{*,min}$ such that a star
+particle only forms if $m_* \geq M_{*,min}$.  An appropriate choice of
+$M_{*,min}$ does not significantly change the star overall star
+formation history of a simulation, though it may delay the onset of
+star formation in a given cell relative to a simulation without a
+particle mass threshold.
 
-%Algorithmically, the star formation events in Kravtsov's code
-%are assumed to occur once every global time step 
-%$\Delta t_0 \leq 10^7$.  In cells where star formation is
-%determined to occur (i.e. $\rho_{gas} \geq \rho_{SF}$) a
-%collisionless ``star particle'' is assumed to form, with a 
-%mass $m_* = \dot{\rho}_* V_{cell} \Delta t_0$, where
-%$\dot{\rho}_*$ is described above and $V_{cell}$ is the 
-%volume of the mesh cell.  These star particles are
-%formed instantaneously, with all $m_*$ of gas being
-%taken out of the cell and immediately deposited into the
-%star particle.  This particle is then given the velocity
-%of the gas in the cell, and thereafter treated as a 
-%collisionless particle.  The \enzo\ implementation
-%of this algorithm is similar, except that instead of 
-%forming stars only at the root grid time step, we allow 
-%stars to form at the time step of the highest level of
-%resolution at any particular point in space.  As can be seen
-%from the equation for $m_*$ above, this can result in very 
-%small stellar masses.  To avoid memory and processor time
-%issues related to having very large numbers of star particles
-%we impose a threshold mass $M_{*,min}$ such that a star
-%particle only forms if $m_* \geq M_{*,min}$.  An
-%appropriate choice of $M_{*,min}$ does not significantly 
-%change the star overall star formation history of a 
-%simulation, though it may delay the onset of star formation 
-%in a given cell relative to a simulation without a particle mass
-%threshold.
+Each ``star particle'' is assumed to represent an ensemble of stars
+and is treated as a single-age stellar population (as in the previous
+section).  Kravtsov assumes that the IMF is described by a Miller \&
+Scalo functional form with stellar masses between $0.1$ and
+$100~M_\odot$ \citep{1979ApJS...41..513M}.  All stars in this IMF with
+$M_* > 8~M_\odot$ deposit $2 \times 10^{51}$ ergs of thermal energy
+and a mass $f_z M_*$ of metals into the cell in which they form
+without delay, with $f_z \equiv min(0.2, 0.01~M_*-0.06)$
+(i.e. instantaneous deposition of metals).  The definition of $f_z$ is
+a rough approximation of the results of \citet{1995ApJS..101..181W}.
 
-%Each ``star particle'' is assumed to represent an ensemble
-%of stars and is treated as a single-age stellar population.
-%Kravtsov assumes that the IMF is described by a
-%Miller \& Scalo functional form with stellar masses between
-%$0.1$ and $100~M_\odot$ \citep{1979ApJS...41..513M}.  All stars in this
-%IMF with $M_* > 8~M_\odot$ deposit $2 \times 10^{51}$ ergs of
-%thermal energy and a mass $f_z M_*$ of metals into the
-%cell in which they form without delay, with
-%$f_z \equiv min(0.2, 0.01~M_*-0.06)$ (i.e. instantaneous
-%deposition of metals).  The definition of $f_z$ is a rough
-%approximation of the results of 
-%\citet{1995ApJS..101..181W}.
-
-%Kravtsov reports that simulations with this algorithm
-%reliably reproduce the star formation rate-gas surface
-%density relation of the Schmidt law, and are not particularly
-%sensitive to numerical parameters \citep{2003ApJ...590L...1K}.
-%He also notes that this is surprisingly insensitive to
-%the presence or absence of feedback and details of the cooling
-%and heating properties of the gas, which suggests that the
-%global star formation rate is determined by gravitationally-
-%driven supersonic turbulence (on large scales)
-%rather than stellar feedback or thermal instabilities on small 
-%scales.
+Kravtsov reports that simulations with this algorithm reliably
+reproduce the star formation rate-gas surface density relation of the
+Schmidt law, and are not particularly sensitive to numerical
+parameters \citep{2003ApJ...590L...1K}.  He also notes that this
+algorithm is surprisingly insensitive to the presence or absence of
+feedback and details of the cooling and heating properties of the gas,
+which suggests that the global star formation rate is determined by
+gravitationally- driven supersonic turbulence (on large scales) rather
+than stellar feedback or thermal instabilities on small scales.
 
 % Specific example: H2-regulated
-\subsubsection{H2-regulated}
+\subsubsection{H2-regulated method}
 \label{sec:starform_H2reg}
+
+The methods described in the previous two sections are generally used
+in simulations that have very poor resolution, $\Delta x \ga 1$~kpc.
+At this physical scale, individual star forming regions are not
+resolved, so all uncertainty about the behavior of molecular clouds is
+folded into a density threshold for star formation.  In calculations
+with much higher resolution, however ( $\la$~few pc), individual
+molecular clouds can be resolved, thus rendering these approximations
+irrelevant.  To that end, \citet{2012ApJ...749...36K} have implemented
+a star formation algorithm that is specifically geared to
+high-resolution cosmological simulations of galaxy formation, where
+the formation of molecular hydrogen is followed directly, and stars
+are allowed to form at the highest level of refinement when the local
+H$_2$ fraction exceeds a pre-determined threshold.  Cells are examined
+every root grid timestep, $\Delta t_0$, and cells at the highest level
+that exceed the H$_2$ fraction limit form star particles with masses
+proportional to the inferred mass of molecular hydrogen in the
+star-forming region
+\citep{2008ApJ...689..865K,2009ApJ...693..216K,2010ApJ...709..308M}.
+The mass of the particle is calculated as:
+
+\begin{equation} 
+m_p = \epsilon \rho_{gas} (\Delta x_{m})^3 \frac{\Delta t_0}{t_*}
+\end{equation}
+
+where $\Delta x_{m}$ is the resolution of the maximum level of
+refinement, $\epsilon$ is an efficiency parameter with a standard
+value of 0.01 \citep[as motivated by][]{2007ApJ...654..304K}, and
+t$_*$ is the local free-fall timescale.
+
+The feedback method used in this method is identical to that described
+in Section~\ref{sec:starform_cen}.
 
@@
 
-%\subsubsection{The Kravtsov star formation and feedback algorithm}
+Unlike in the previous sections, under some circumstances it is both
+possible and desirable to simulate stars individually, rather than
+treating particles as ensembles.  One particular example of this is
+Population III star formation
+\citep{ABN02,2007ApJ...654...66O,2008ApJ...685...40W,2009Sci...325..601T},
+where a given halo may only form one star of primordial composition.
+To accommodate this, \citet{2007ApJ...659L..87A,2008ApJ...685...40W, 2012MNRAS.427..311W} have created an
+algorithm within the \enzo\ star formation framework that forms
+Population III stars directly.  Using criteria similar
+to~\ref{CO1992}, a star particle forms when a cell has all of the
+following criteria:
+
+\begin{enumerate}
+\item a baryon overdensity of $5 \times 10^5$ (corresponding to
+  roughly $10^3$ g cm$^{-3}$ at $z=10$),
+\item a converging gas flow (\nabla \cdot \vec{v} < 0), and
+\item a molecular hydrogen fraction f$_{H2} > 5 \times 10^{-4}$,
+\end{enumerate}
+
+conditions typical of a collapsing metal-free molecular cloud roughly
+10 million years before the birth of a Pop III main-sequence star.  If
+multiple neighboring cells are flagged as being able to form stars, a
+single star is created instead.  This star has a mass that is randomly
+sampled from a stellar IMF with a functional form of
+
+\begin{equation}
+f(log M) dM = M^{-1.3} \exp \Big[ -\Big( \frac{M_{char}}{M}  \Big)^{1.6} \Big]
+\end{equation}
+
+Feedback from the Population III stars comes in multiple forms.
+Radiative feedback using the MORAY radiation transport algorithm
+\citep{Wise11_Moray} is available, using the mass-dependent hydrogen
+ionizing and Lyman-Werner photon luminosities and lifetimes of the
+Population III stars from \citet{2002A&A...382...28S}.  At the end of
+their main-sequence lifetimes, explosion energies, ejected gas mass,
+and ejected metal are caclulated using a variety of sources that
+depend on the mass of the star, and are described in detail in Section
+3.2.1 of \citet{2012MNRAS.427..311W}.

### 2013-03-15 introduction.tex

@@
 
-There are many diverse phenomena in astrophysics and cosmology which require a high spatial and temporal dynamical range. There are a number of numerical techniques which have been used to
-model such problems.
-The most common is a gridless, fully Lagrangian method call Smoothed Particle Hydrodynamcs \citep[SPH;][]{Lucy77, SPH} which has been very successful, particularly for those situations dominated by gravity.
-Its development, however, is still at a relatively early stage when
-compared to the effort put into Eulerian, grid-based hydrodynamic schemes \citep[e.g.,][]{Woodward84}.
+Many phenomena in astrophysics and cosmology require a high spatial and temporal dynamical range and there are a number of numerical techniques which have been used to model such problems.
+At the current time, the most common is a gridless, fully Lagrangian method call Smoothed Particle Hydrodynamcs \citep[SPH;][]{Lucy77, SPH} which has been very successful, particularly for those situations dominated by gravity.
+Its development, however, is still at a relatively early stage when compared to the effort put into Eulerian, grid-based hydrodynamic schemes \citep[e.g.,][]{, laney-1998, toro-1997, Woodward84}.
 
-Therefore, we would like to make use of this expertise; unfortunately, none of these schemes provide an easy way to adaptively and dynamically increase the resolution in small volumes, as required by, for example, gravitational instability.
-We have instead, adapted an idea first suggested by \citet{Berger89} in the Computational Fluid Dynamics (CFD) community, which has become known as Structured Adaptive Mesh Refinement (SAMR).
+Therefore, we would like to make use of this expertise that has been invested in Eulerian solvers; unfortunately, as originally developed, none of these schemes provide an easy way to adaptively and dynamically increase the resolution in small volumes, as required by, for example, gravitational instability.
+We have instead, adopted an idea first suggested by \citet{Berger89} in the Computational Fluid Dynamics (CFD) community, which has become known as Structured Adaptive Mesh Refinement (SAMR).
 The principle is to adaptively add (and modify) additional, finer meshes, or patches, over regions which require higher resolution.
-Since each patch has a simple geometry (here, cartesian) the hyperbolic fluid-dynamical equations can be solved on that patch with a well-developed `off-the-shelf' numerical method.
+Since each patch has a simple geometry (here, Cartesian) the hyperbolic fluid-dynamical equations can be solved on that patch with a well-developed `off-the-shelf' numerical method.
 In addition, it is relatively straightforward to add other advanced physics (e.g.
 MHD, radiation transport, etc.).
-This paper is devoted to describing how we have implemented this idea for astrophysics, including the addition of self-gravity, collisionless fluids, MHD, radiation transport and a range of other physical effects.
+This paper is devoted to describing how we have implemented this idea for astrophysics, including the addition of comoving coordinates, self-gravity, radiative cooling, chemistry, heat conduction, collisionless fluids, MHD, radiation transport, star formation and a range of other physical effects.
 
@@
 This static approach, applied to hydrodynamics, has been used extensively \citep[e.g.,]{Ruffert94, Anninos94}.
-Adding adaptivity is a more recent enhancement, and there are now a number of N-body codes which possess this feature, both with and without  hydrodynamics  \cite{Couchman91, Jessop94, Suisalu95, Splinter96, Gelato97, ART97, Truelove98, flash_method, MLAPM01,  Yahagi01, RAMSES, Quilis04, Ziegler05, Zhang06}.
-Of these, perhaps the most comparable are FLASH \citep{flash_method}, which uses grid blocks of fixed size, and RAMSES \citep{RAMSES} and ART \citep{ART97}, both of which refine individual cells.
-It is also possible to deform the grid to obtain high resolution \citep[e.g.,][]{Gnedin95, Xu97, Pen98}, and more recently
+Adding adaptivity is a more recent enhancement, and there are now a number of N-body codes which possess this feature, both with and without  hydrodynamics  \citep{Couchman91, Jessop94, Suisalu95, Splinter96, Gelato97, ART97, Truelove98, flash_method, MLAPM01,  Yahagi01, RAMSES, Quilis04, Ziegler05, Zhang06, Astrobear09, Pluto-amr, GAMER, Nyx}.
+Of these, perhaps the most comparable and widely used are FLASH \citep{flash_method}, which uses grid blocks of fixed size, and RAMSES \citep{RAMSES} and ART \citep{ART97}, both of which refine individual cells.
+It is also possible to deform the grid to obtain high resolution \citep[e.g.,][]{Gnedin95, Xu97, Pen98}, and more recently a few grid codes have adopted an unstructured approach based on a moving Voronoi mesh \citep{Arepo10, Tess11}.
 
@@
 
-The \enzo\ code has been widely used over the lasts decade in a wide variety of problems in topics ranging from star formation to turbulence to large-scale structure.
-More than 100 peer-reviewed papers have been published based on results obtained with the code.
+The \enzo\ code has been widely used over the last two decades in a wide variety of problems, and more than 100 peer-reviewed papers have been published based on results obtained with the code.
+The code has previously been used to study a wide range of astrophysical systems, including X-ray clusters \citep{Loken02}, the ISM \citep{Slyz05}, the intergalactic medium \citep{Fang01}, cooling flows \citep{Li12}, turbulence \citep{Kritsuk04}, and the formation of the first stars \citep{ABN02}.
 
-Finally, we note that the numerical simulation of astronomical phenomena now plays a key role, along with observations and analytic theory, in pushing forward our understanding of the cosmos.
-%\citep[e.g.,][]{DecadalSurvey01}.  
+Finally, we note that the numerical simulation of astronomical phenomena now plays a key role, along with observations and analytic theory, in pushing forward our understanding of the cosmos \citep[e.g.,][]{DecadalSurvey01, DecadalSurvey10}.
 But along with this role comes responsibility.
@@
 The code tests are described in section~\ref{sec.tests} and the parallelism strategy and scaling results in section~\ref{sec.performance}.
-
-%\red{
-%\begin{enumerate}
-%  \item Motivate why we're writing the paper now: (because there isn't one, lots of people use it for all sorts of things, given that it's publicly available we should really have a method paper and some general proof that the code works, and a list of what it does at this point in time)
-%  \item Why have the \enzo\ code at all?
-Many diverse phenomena in astrophysics and cosmology where you need a high spatial and temporal dynamical range.
-Regular (i.e.
-monolithic) grids don't have the required range.
-While SPH is the most commonly used tool, and has been quite successful, the lack of a grid structure complicates the usage of advanced hydrodynamical techniques and other advanced physics (e.g.
-full solution of the MHD equations, radiation transport beyond flux limited diffusion, etc.).
-There has been 50(ish) years of advancement in the fields of grid-based solutions to these methods, and the use of adaptive mesh on structured grids allows us to take advantage of this for a wide range of astrophysical and cosmological applications.
-%  \item List the subjects that the code has been used for, citing appropriate papers.
-Emphasize heavily landmark contributions (Greg's 1998 galaxy cluster paper and various Motl/Burns/Hallman cluster things, various lyman alpha forest stuff, Pop III papers, alexei's turbulence stuff, etc.).
-This code has been used for approximately 10 years, and is publicly available.
-List papers by non-collaborators that have already used the code.
-%  \item Explain structure of the paper.
-Emphasize that we pass through the enzo code in multiple levels of (increasing) detail: The ``physical equations'' section provides bare-bones ``what we're solving'' information, the ``overview of numerical methods'' section briefly explains (i.e.
-no pseudocode or finite difference stencils) how the various important subsystems of the code work, and the appendices of the paper have the real nitty-gritty details that most people don't want to know, but a select few will find highly useful.
-%\end{enumerate}
-%} %end red

### 2013-03-15 numerical-gravity.tex

@@
-\subsection{Gravity: Solving the Poisson Equations}
+\subsection{Gravity}
 \label{sec.gravity}
 
-There are multiple methods to compute the gravitational potential (which is an elliptic equation in the Newtonian limit) in a structured AMR framework.
-One way would be to model the dark matter (or other collisionless particle-like objects, such as stars) as a second fluid in addition to the baryonic fluid and solve the collisionless Boltzmann equation, which follows the evolution of the fluid density in six-dimensional phase space.
-However, this is computationally prohibitive owing to the large dimensionality of the problem, making this approach unappealing for the cosmological AMR code.
+Solving for the accelerations for the cells and particles in a grid due to self-gravity involves three steps: first, computing the total gravitating mass; second, solving for the potential field with the appropriate boundary conditions, and finally, differencing the potential to get the acceleration and, if necessary, interpolating the potential back to the particles.
+These steps are described in detail below.
 
-The dark matter particles are distributed onto the grids using the cloud-in-cell (CIC) interpolation technique to form a spatially discretized density field (analogous to the baryon densities used to calculate the equations of hydrodynamics).
-After sampling dark matter density onto the grid and adding baryon density if it exists (to get the total matter density in each cell), the gravitational potential is calculated on the periodic root grid using a fast Fourier transform.
+First, the dark matter particles are distributed onto the grids using the cloud-in-cell (CIC) interpolation technique \citet{Hockney88} to form a spatially discretized density field $\rho_{\rm DM}$.
+When doing the CIC interpolation, particle positions are (temporarily) advanced by $0.5 \Delta t v^n$ so that we generate an estimate of the time-centered density field.
+Particles on sub-grids are also added in this way.
+In addition, since the gravitating field for a grid is defined beyond the grid edges (see below), dark matter particles from sibling grids and sub grids which lie within the entire gravitating field are used.
+
+Next, we add the baryonic grid densities in a similar fashion.
+In particular, we treat baryonic cells as virtual CIC particles which are are placed at the grid center but are advanced by $0.5 \Delta t v^n$ in order to approximately time-center the gravitating mass field.
+Cells which are covered by further refined grids are treated in a similar way (i.e.
+we use the sub grid cells as virtual CIC particles).
+This procedure results in a total gravitating mass field $\rho_{\rm total}^{n+1/2}$.
+
+To compute the potential field from this gravitating mass field on the root grid, we use a a fast Fourier transform.
 For periodic boundary conditions, we can use either a simple Greens function kernel of $-k^{-2}$, or the finite-difference equivalent \citep{Hockney88}:
@@
 In this case, the Greens function is generated in real-space so as to have the correct zero-padding properties and then transformed into the Fourier domain.
+In either case, the potential is then transformed back into the real domain to get potential values at the cell centers.
+These are differenced with a two-point centered difference to get accelerations at the cell centers (except if we are using the staggered Zeus-like solver, in which case the accelerations are computed at the cell faces to match the velocities).
 

### 2013-03-15 numerical-chemistry.tex
Added note to chemistry section.
Changing section title to just chemistry.

@@
-\subsection{Chemistry, radiation and cooling}
-\label{sec.ov.chem}
+\subsection{Chemistry}
+\label{sec.num.chemistry}
 
-\red{
-As this stands it is the methods that I (brian) use, so I don't really emphasize
-either the tabular cooling functions or the various radiation background options.
-Add more discussion of this, point towards appendix.
-}
+While it is often safe to assume that species (both chemical and
+ionization) within a gas can be treated as being in 
+equilibrium, in some regimes that are found in astrophysics this is not true.  For example, the cooling and collapse of primordial gas in Population III
+star formation is dominated by molecular hydrogen, which in the
+absence of dust forms via an inefficient pair of collisional processes
+that depend heavily on the local population of free electrons.  
+As a
+result, when modeling primordial star formation, it is critical to
+follow the non-equilibrium evolution of the chemical species of
+hydrogen, including molecular hydrogen and deuterium.
 
-Though the equations of hydrodynamics described above are a closed system,
-they are still missing a crucial piece of physics: radiative heating and cooling.
-Radiative cooling is extremely important in many astrophysical situations, as is 
-the heating of gas from some sort of radiation background.  \enzo\ has a very simple
-Sutherland and Dopita equilibrium cooling function \citep{1993ApJS...88..253S} 
-implemented, which uses a
-cooling table assuming a fixed metallicity of $Z = 0.3 Z_\odot$, and also a
-nonequilibrium heating/cooling model that assumes gas of primordial composition exposed to a
-uniform metagalactic ultraviolet (UV) background that varies with time \citep{1996ApJS..105...19K}.
-
-The simulations discussed in this thesis almost exclusively use the nonequilibrium
-routines, described in great detail by Abel et al. and Anninos et al.  
-\citep{abel97,anninos97} and summarized in Appendix~\ref{app:chemistry}. These 
+The primordial non-equilibrium chemistry routines used in \enzo\ were
+first described by Abel et al. and Anninos et
+al. \citep{abel97,anninos97}, but have been extended since then with
+updated reaction rates and the inclusion of deuterium species.  These 
 routines follow the non-equilibrium chemistry of a 
-gas of primordial composition with 9 total species:  
-$H$, $H^+$, $He$, $He^+$, $He^{++}$, $H^-$, $H_2^+$, $H_2$, and $e^-$.  The code also calculates 
-radiative heating and cooling, following atomic line excitation, recombination,
-collisional excitation, free-free transitions, molecular line excitations, and Compton
-scattering of the cosmic microwave background, as well as any of
-approximately a dozen different models for a metagalactic UV background that heat
-the gas via photoionization and photodissociation. 
-The multispecies rate equations are solved out of
-equilibrium to properly model situations where, e.g., the cooling rate of the gas
-is much shorter than the hydrogen recombination time.  The effect of this nonequilibrium
-cooling is to leave behind a much larger fraction of residual free electrons than one would
-expect if the assumption of equilibrium were being made.  The practical effect of this
-is that more $H^-$ is formed, which subsequently produces hydrogen molecules.  If large
-amounts of $H_2$ is formed it can greatly increase the cooling rate of primordial gas
-at relatively low temperatures ($T \leq 10^4$ K).  This can efficiently cool the gas 
-to approximately $200$ K, which significantly reduces the Jeans mass of the gas.  Correct
-modeling of the formation of molecular hydrogen is crucial to the study of star formation
-in a primordial gas.
-
-A total of 9 kinetic equations are solved, including 29 kinetic and radiative 
+gas of primordial composition with 12 total species:  
+$H$, $H^+$, $He$, $He^+$, $He^{++}$, $H^-$, $H_2^+$, $H_2$, $e^-$,
+$D$, $D^+$, and $HD$.  (A separate, but related, module follows the
+radiative heating and cooling of the gas from atomic and molecular line excitation,
+recombination, collisional excitation, free-free transitions, Compton
+scattering of the cosmic microwave background, as well as several
+models for a metagalactic UV background that heat the gas via
+photoionization and discussion -- see Section~\ref{sec.num.cooling}
+for more details).  Depending on user parameters, one can follow
+simply atomic species ($H$, $H^+$, $He$, $He^+$, $He^{++}$,  and
+$e^-$), include those relevant for molecular hydrogen formation
+($H_2$, $H_2^+$, and $H^-$), and further include deuterium and its
+species ($D$, $D^+$, and $HD$).  
+\red{A total of 9 kinetic equations are solved, including 29 kinetic and radiative 
 processes, for the 9 species mentioned above.  See Table~\ref{table.collisional} 
 for the collisional processes and Table~\ref{table.radiative} for the 
-radiative processes solved.
+radiative processes solved.}
+
+The chemical reaction equation network is technically challenging to solve due to 
+the huge range of reaction timescales involved -- the characteristic creation
+and destruction timescales of the various species and reactions can differ by 
+many orders of magnitude.  As a result, the set of rate equations is extremely 
+stiff, and an explicit scheme for integration of the rate equations can be 
+exceptionally costly if small enough timesteps are taken to keep the network
+stable.  This makes an implicit scheme much more preferable for such a set of 
+equations.  However, an implicit scheme typically require an iterative 
+procedure to converge, and for large networks (such as this one) an implicit
+method can be very time-consuming, making it undesirable for a large, three-dimensional
+simulation.
+
+\enzo\ solves the rate equations using a method based on a backwards differencing 
+formula (BDF) in order to provide a stable and accurate solution.  This technique is
+optimized by taking the chemical intermediaries $H^-$ and $H_2^+$, which have 
+large rate coefficients and low concentrations, and grouping them into a separate
+category of equations.  Due to their fast reactions these species are very sensitive
+to small changes in the more abundant species.  However, due to their low overall
+concentrations, they do not significantly influence the abundance of species with
+higher concentrations.  Therefore, reactions involving these two species can be
+decoupled from the rest of the network and treated independently.  In fact, $H^-$ 
+and $H_2^+$ are treated as being in equilibrium at all times, independent of 
+the other species and the hydrodynamic variables.  This allows a large speedup
+in solution as the BDF scheme is then applied only to the slower 7- or
+10-species network (depending on whether deuterium is included or not)
+on timescales closer to those required by the hydrodynamics of the simulation.
+Even so, the accuracy and stability of the scheme is maintained by subcycling the 
+rate solver over a single hydrodynamic timestep.  These subcycle timesteps are 
+determined so that the maximum fractional change in the electron concentration is
+limited to no more than $10\%$ per timestep.
+
+It is important to note the regime in which this model is valid.  According to Abel et al. and
+Anninos et al. \citep{abel97,anninos97}, the reaction network is valid for temperatures
+between $10^0 - 10^8$ K.  The original model discussed in these two references is only
+valid up to $n_H \sim 10^4$~cm$^{-3}$.  However, addition of the 3-body $H_2$ formation
+process (equation 20 in Table~\ref{table.collisional}) allows 
+correct modeling of the chemistry of the gas up
+until the point where collisionally-induced emission from molecular hydrogen becomes an important
+cooling process, which occurs at $n_H \sim 10^{14}$~cm$^{-3}$.  A further concern is that
+the optically thin approximation for radiative cooling breaks down, which
+occurs before $n_H \sim 10^{16} - 10^{17}$~cm$^{-3}$.  Beyond this point, 
+modifications the cooling function that take into account the non-negligible
+opacity in the gas must be made, as discussed by \citet{2004MNRAS.348.1019R}. 
+Even with these modifications, a completely correct description of the cooling of
+this gas will require some form of radiation transport, which will greatly 
+increase the cost of the simulations.
+
+\red{Several processes are neglected.  The deuterium atom and its processes are 
+completely ignored, which may have some effect.  Recent work shows that HD is
+a more effective coolant than previously thought~\citep{2005MNRAS.361..850L}.  However,
+the fractional abundance of HD is so low that under circumstances relevant to
+the formation of a Population III star in an un-ionized region it should be
+sub-dominant.  However, the enhanced electron fraction in fossil 
+HII regions (as discussed later in this thesis) could result in the HD molecule
+becoming a dominant cooling mechanism at relatively low ($\sim$ few hundred K)
+temperatures, and could potentially cool the gas down to below $100$~K, which can
+enhance fragmentation and could have important consequences for the IMF of 
+primordial stars forming in a relic HII region~\citep{2005MNRAS.364.1378N}.}
+
+Aside from deuterium, the chemical reactions involving lithium 
+are also neglected.  According to \citet{1998A&A...335..403G}, these are 
+not important for the density and temperature regimes explored 
+by the simulations discussed in this thesis.  However, at higher densities it 
+is possible that there are regimes where lithium can be an important coolant.
 
@@
 \end{table}
-
-The chemical reaction equation network is technically challenging to solve due to 
-the huge range of reaction timescales involved -- the characteristic creation
-and destruction timescales of the various species and reactions can differ by 
-many orders of magnitude.  As a result, the set of rate equations is extremely 
-stiff, and an explicit scheme for integration of the rate equations can be 
-exceptionally costly if small enough timesteps are taken to keep the network
-stable.  This makes an implicit scheme much more preferable for such a set of 
-equations.  However, an implicit scheme typically require an iterative 
-procedure to converge, and for large networks (such as this one) an implicit
-method can be very time-consuming, making it undesirable for a large, three-dimensional
-simulation.
-
-\enzo\ solves the rate equations using a method based on a backwards differencing 
-formula (BDF) in order to provide a stable and accurate solution.  This technique is
-optimized by taking the chemical intermediaries $H^-$ and $H_2^+$, which have 
-large rate coefficients and low concentrations, and grouping them into a separate
-category of equations.  Due to their fast reactions these species are very sensitive
-to small changes in the more abundant species.  However, due to their low overall
-concentrations, they do not significantly influence the abundance of species with
-higher concentrations.  Therefore, reactions involving these two species can be
-decoupled from the rest of the network and treated independently.  In fact, $H^-$ 
-and $H_2^+$ are treated as being in equilibrium at all times, independent of 
-the other species and the hydrodynamic variables.  This allows a large speedup
-in solution as the BDF scheme is then applied only to the slower 7-species network
-on timescales closer to those required by the hydrodynamics of the simulation.
-Even so, the accuracy and stability of the scheme is maintained by subcycling the 
-rate solver over a single hydrodynamic timestep.  These subcycle timesteps are 
-determined so that the maximum fractional change in the electron concentration is
-limited to no more than $10\%$ per timestep.
-
-It is important to note the regime in which this model is valid.  According to Abel et al. and
-Anninos et al. \citep{abel97,anninos97}, the reaction network is valid for temperatures
-between $10^0 - 10^8$ K.  The original model discussed in these two references is only
-valid up to $n_H \sim 10^4$~cm$^{-3}$.  However, addition of the 3-body $H_2$ formation
-process (equation 20 in Table~\ref{table.collisional}) allows 
-correct modeling of the chemistry of the gas up
-until the point where collisionally-induced emission from molecular hydrogen becomes an important
-cooling process, which occurs at $n_H \sim 10^{14}$~cm$^{-3}$.  A further concern is that
-the optically thin approximation for radiative cooling breaks down, which
-occurs before $n_H \sim 10^{16} - 10^{17}$~cm$^{-3}$.  Beyond this point, 
-modifications the cooling function that take into account the non-negligible
-opacity in the gas must be made, as discussed by \citet{2004MNRAS.348.1019R}. 
-Even with these modifications, a completely correct description of the cooling of
-this gas will require some form of radiation transport, which will greatly 
-increase the cost of the simulations.
-
-Several processes are neglected.  The deuterium atom and its processes are 
-completely ignored, which may have some effect.  Recent work shows that HD is
-a more effective coolant than previously thought~\citep{2005MNRAS.361..850L}.  However,
-the fractional abundance of HD is so low that under circumstances relevant to
-the formation of a Population III star in an un-ionized region it should be
-sub-dominant.  However, the enhanced electron fraction in fossil 
-HII regions (as discussed later in this thesis) could result in the HD molecule
-becoming a dominant cooling mechanism at relatively low ($\sim$ few hundred K)
-temperatures, and could potentially cool the gas down to below $100$~K, which can
-enhance fragmentation and could have important consequences for the IMF of 
-primordial stars forming in a relic HII region~\citep{2005MNRAS.364.1378N}.
-
-Aside from deuterium, the chemical reactions involving lithium 
-are also neglected.  According to \citet{1998A&A...335..403G}, these are 
-not important for the density and temperature regimes explored 
-by the simulations discussed in this thesis.  However, at higher densities it 
-is possible that there are regimes where lithium can be an important coolant.

### 2013-03-15 numerical-chemistry.tex
updating chemistry!

@@
 While it is often safe to assume that species (both chemical and
-ionization) within a gas can be treated as being in 
-equilibrium, in some regimes that are found in astrophysics this is not true.  For example, the cooling and collapse of primordial gas in Population III
+ionization) within a gas can be treated as being in equilibrium, in
+some regimes that are found in astrophysics this is not true.  For
+example, the cooling and collapse of primordial gas in Population III
 star formation is dominated by molecular hydrogen, which in the
 absence of dust forms via an inefficient pair of collisional processes
-that depend heavily on the local population of free electrons.  
-As a
+that depend heavily on the local population of free electrons.  As a
 result, when modeling primordial star formation, it is critical to
@@
 al. \citep{abel97,anninos97}, but have been extended since then with
-updated reaction rates and the inclusion of deuterium species.  These 
-routines follow the non-equilibrium chemistry of a 
-gas of primordial composition with 12 total species:  
-$H$, $H^+$, $He$, $He^+$, $He^{++}$, $H^-$, $H_2^+$, $H_2$, $e^-$,
-$D$, $D^+$, and $HD$.  (A separate, but related, module follows the
-radiative heating and cooling of the gas from atomic and molecular line excitation,
-recombination, collisional excitation, free-free transitions, Compton
-scattering of the cosmic microwave background, as well as several
-models for a metagalactic UV background that heat the gas via
-photoionization and discussion -- see Section~\ref{sec.num.cooling}
-for more details).  Depending on user parameters, one can follow
-simply atomic species ($H$, $H^+$, $He$, $He^+$, $He^{++}$,  and
-$e^-$), include those relevant for molecular hydrogen formation
-($H_2$, $H_2^+$, and $H^-$), and further include deuterium and its
-species ($D$, $D^+$, and $HD$).  
-\red{A total of 9 kinetic equations are solved, including 29 kinetic and radiative 
-processes, for the 9 species mentioned above.  See Table~\ref{table.collisional} 
-for the collisional processes and Table~\ref{table.radiative} for the 
-radiative processes solved.}
+updated reaction rates and the inclusion of deuterium species
+\citep{2009PhDT.........5T}.  These routines follow the
+non-equilibrium chemistry of a gas of primordial composition with 12
+total species: $H$, $H^+$, $He$, $He^+$, $He^{++}$, $H^-$, $H_2^+$,
+$H_2$, $e^-$, $D$, $D^+$, and $HD$.  (A separate, but related, module
+follows the radiative heating and cooling of the gas from atomic and
+molecular line excitation, recombination, collisional excitation,
+free-free transitions, Compton scattering of the cosmic microwave
+background, as well as several models for a metagalactic UV background
+that heat the gas via photoionization and discussion -- see
+Section~\ref{sec.num.cooling} for more details).  Depending on user
+parameters, one can follow simply atomic species ($H$, $H^+$, $He$,
+$He^+$, $He^{++}$, and $e^-$), include those relevant for molecular
+hydrogen formation ($H_2$, $H_2^+$, and $H^-$), and further include
+deuterium and its species ($D$, $D^+$, and $HD$.)  \red{A total of 9
+kinetic equations are solved, including 29 kinetic and radiative
+processes, for the 9 species mentioned above.  See
+Table~\ref{table.collisional} for the collisional processes and
+Table~\ref{table.radiative} for the radiative processes solved.}
 

### 2013-03-15 overview.tex
partial commit

@@
 
-%GB: The division of this subsection into subsubsections is somewhat arbitrary.
-%GB: I wanted to do it at first in the same divisions as section 3, but this is not really
-%GB: possible because there is no "AMR" equation.  Maybe somebody can think 
-%GB: of a clever way to fix this.
-
 % ----------------------------------------------------------------------------------------
 
-\subsubsection{Hydrodynamics and Gravity}
+\subsubsection{Magnetohydrodynamics and Gravity}
 
@@
 \frac{\partial \rho}{\partial t} 
-          + \frac{1}{a} \vecv \cdot \nabla \rho
-    & = & - \frac{1}{a} \rho \nabla \cdot \vecv, 
+          + \frac{1}{a} \nabla \cdot (\rho \vecv) & = & 0
         \label{eq:mass} \\
 %
-\frac{\partial \vecv}{\partial t}  
-          + \frac{1}{a} ( \vecv \cdot \nabla ) \vecv
-    & = & - \frac{\dot{a}}{a} \vecv 
-          - \frac{1}{a \rho} \nabla p
-          - \frac{1}{a} \nabla \phi, 
+\frac{\partial \rho \vecv}{\partial t}  
+   + \frac{1}{a} \nabla \cdot \left(\rho \vecv \vecv + \myvec{I} p - \vecB \vecB \right) & = &
+    - \frac{\dot{a}}{a} \rho \vecv - \frac{1}{a} \rho \nabla \phi
+    %
+%          + \frac{1}{a} ( \vecv \cdot \nabla ) \vecv
+%    & = & - \frac{\dot{a}}{a} \vecv 
+%          - \frac{1}{a \rho} \nabla p
+%          - \frac{1}{a \rho} (\nalba \times \vecB) \times \vecB
+%          - \frac{1}{a} \nabla \phi, 
         \label{eq:momentum} \\
 %
-\frac{\partial \Ep}{\partial t} 
-         + \frac{1}{a} \vecv \cdot \nabla E
-   & = & - \frac{\dot{a}}{a} (3 \frac{p}{\rho} + {\vecv \cdot \vecv} )
-         - \frac{1}{a \rho } \nabla \cdot (p \vecv) 
-%\nonumber   & &   %uncomment for 2-column
-         - \frac{1}{a} \vecv \cdot \nabla \phi 
-                + \Gamma - \Lambda.
-       \label{eq:total_energy} 
+\frac{\partial E \rho}{\partial t} 
++ \frac{1}{a} \nabla \cdot \left[ (E + p) \vecv - (\vecB \cdot \vecv) \vecB \right] & = &
+     - \frac{2\dot{a}}{a} E - \frac{1}{a} \vecv \cdot \nabla \phi + \frac{\dot{a}}{a} \frac{\vecB \cdot \vecB}{2}
+%
+%         + \frac{1}{a} \vecv \cdot \nabla E
+ %  & = & - \frac{\dot{a}}{a} (3 \frac{p}{\rho} + {\vecv \cdot \vecv} )
+  %       - \frac{1}{a \rho } \nabla \cdot (p \vecv) 
+%%\nonumber   & &   %uncomment for 2-column
+ %        - \frac{1}{a} \vecv \cdot \nabla \phi 
+ %               + \Gamma - \Lambda.
+       \label{eq:total_energy}  \\
+%
+\frac{\partial \vecB}{\partial t} & = & \frac{1}{a}  \nabla \times (\vecv \times \vecB) - \frac{\dot{a}}{2a} \vecB \label{eq:induction}
 \end{eqnarray}
@@
 %
-Here $E$, $e$, $\rho$, $\vecv$ and $p$ are the comoving total specific fluid energy, specific thermal energy, baryonic density, velocity and pressure, respectively.  
+Here $E$, $e$, $\rho$, $\vecv$, $\vecB$, and $p$ are the comoving total fluid energy, specific thermal energy, baryonic density, velocity, magnetic field strength and pressure, respectively.
+$I$ is the identify matrix and we use a set of units for the magnetic field such that magnetic xxx is unity.
 %Here we use primes to indicate variables defined in the absolute frame of reference (to distinguish them from a comoving frame).  
@@
 \begin{equation}
-E =  e + \frac{1}{2} \vecv \cdot \vecv.
+E =  e + \frac{1}{2} \vecv \cdot \vecv + \vecB \cdot \vecB
         \label{eq:total_energy_def}
@@
 
-\subsubsection{Godonuv MUSCL with Dednet (HD and MHD)}
+\subsubsection{Godonov MUSCL with Dedner (HD and MHD)}
 
@@
 On subgrids, we interpolate the boundary conditions from that grid's parent (either the level 0 grid or some other subgrid).
-Poisson's equation is then solved using a multigrid technique.
+The Poisson equation is then solved using a multigrid technique.
 In \enzo\, self-gravity is optional.

### 2013-03-17 overview.tex
partial checkin

@@
 \frac{\partial \rho \vecv}{\partial t}  
-   + \frac{1}{a} \nabla \cdot \left(\rho \vecv \vecv + \myvec{I} p - \vecB \vecB \right) & = &
+   + \frac{1}{a} \nabla \cdot \left(\rho \vecv \vecv + p^* - \vecB \vecB \right) & = &
     - \frac{\dot{a}}{a} \rho \vecv - \frac{1}{a} \rho \nabla \phi
@@
 %
-\frac{\partial E \rho}{\partial t} 
-+ \frac{1}{a} \nabla \cdot \left[ (E + p) \vecv - (\vecB \cdot \vecv) \vecB \right] & = &
-     - \frac{2\dot{a}}{a} E - \frac{1}{a} \vecv \cdot \nabla \phi + \frac{\dot{a}}{a} \frac{\vecB \cdot \vecB}{2}
+\frac{\partial E}{\partial t} 
++ \frac{1}{a} \nabla \cdot \left[ (E + p^*) \vecv -  \vecB (\vecB \cdot \vecv) \right] & = &
+     - \frac{\dot{a}}{a} \left( 2E - \frac{\vecB^2}{2} \right)  - \frac{1}{a} \vecv \cdot \nabla \phi 
      %
@@
 %
-\frac{\partial \vecB}{\partial t} & = & \frac{1}{a}  \nabla \times (\vecv \times \vecB) - \frac{\dot{a}}{2a} \vecB \label{eq:induction}
+\frac{\partial \vecB}{\partial t} - \frac{1}{a}  \nabla \times (\vecv \times \vecB) & = & 0 \label{eq:induction}
 \end{eqnarray}
@@
 %
-Here $E$, $e$, $\rho$, $\vecv$, $\vecB$, and $p$ are the comoving total fluid energy, specific thermal energy, baryonic density, velocity, magnetic field strength and pressure, respectively.
-$I$ is the identify matrix and we use a set of units for the magnetic field such that magnetic xxx is unity.
-%Here we use primes to indicate variables defined in the absolute frame of reference (to distinguish them from a comoving frame).  
+Here $E$, $\rho$, $\vecv$, and $\vecB$ are the comoving total fluid energy, comoving baryonic density, peculiar velocity, comoving magnetic field strength, respectively.  
 The first equation represents conservation of mass, the second, conservation of momentum, and the third, conservation of total fluid (kinetic plus thermal) energy.
@@
 
-The total fluid specific energy ($E$) is given by:
+The total fluid energy ($E$) is given by:
 \begin{equation}
-E =  e + \frac{1}{2} \vecv \cdot \vecv + \vecB \cdot \vecB
-        \label{eq:total_energy_def}
+E =  \rho e + \frac{1}{2} \vecv \cdot \vecv + \vecB \cdot \vecB
+        \label{eq:total_energy_def}.
 \end{equation}
-It is useful to solve the equation of total energy because doing so enforces local energy conservation.
-
+where $e$ is the comoving specific thermal energy, and the total isotropic pressure $p^*$ is given by
+\begin{equation}
+p^* = p + \frac{1}{2} \vecB^2
+\end{equation}
 The equations are closed with the equation of state and Poisson's equation (we work in the Newtonian limit):
@@
               \frac{1}{2} \dot{a}^2 \vecx^2, \\
-\phi    & = & \phip + \frac{1}{2} a \ddot{a} \vecx^2.
+\phi    & = & \phip + \frac{1}{2} a \ddot{a} \vecx^2. \\
+\vecB & = & a^2 \vecB^\prime
 \end{eqnarray}
+where primes indicate the usual proper quantities.
 
@@
 \frac{\ddot{a}}{a} & = & 
-      - \frac{4 \pi G }{3 a^3 } (\rhop_0 
-      + 3\pp_0/c^2) 
+      - \frac{4 \pi G }{3 a^3 } (\rho_0 
+      + 3p_0/c^2) 
       + \Lambda_c /3 .
@@
 %
-Here, $\rhop_0$ is the mean (baryonic and dark) comoving mass density and $\pp_0$ is the comoving background pressure contribution, and $\Lambda_c$ is the cosmological constant.
+Here, $\rho_0$ is the mean (baryonic and dark) comoving mass density and $p_0$ is the comoving background pressure contribution, and $\Lambda_c$ is the cosmological constant.
 %$\gamma$ is the ratio of specific heats, and $\Lambda$ is the cosmological constant.

### 2013-03-18 numerical-chemistry.tex
finishing first pass at numerical chemistry section

@@
 
-The chemical reaction equation network is technically challenging to solve due to 
-the huge range of reaction timescales involved -- the characteristic creation
-and destruction timescales of the various species and reactions can differ by 
-many orders of magnitude.  As a result, the set of rate equations is extremely 
-stiff, and an explicit scheme for integration of the rate equations can be 
-exceptionally costly if small enough timesteps are taken to keep the network
-stable.  This makes an implicit scheme much more preferable for such a set of 
+The chemical reaction equation network is technically challenging to
+solve due to the huge range of reaction timescales involved -- the
+characteristic creation and destruction timescales of the various
+species and reactions can differ by many orders of magnitude.  As a
+result, the set of rate equations is extremely stiff, and an explicit
+scheme for integration of the rate equations can be exceptionally
+costly if small enough timesteps are taken to keep the network stable.
+This makes an implicit scheme much more preferable for such a set of
 equations.  However, an implicit scheme typically require an iterative
-procedure to converge, and for large networks (such as this one) an implicit
-method can be very time-consuming, making it undesirable for a large, three-dimensional
-simulation.
+procedure to converge, and for large networks (such as this one) an
+implicit method can be very time-consuming, making it undesirable for
+a large, three-dimensional simulation.
 
-\enzo\ solves the rate equations using a method based on a backwards differencing 
-formula (BDF) in order to provide a stable and accurate solution.  This technique is
-optimized by taking the chemical intermediaries $H^-$ and $H_2^+$, which have 
-large rate coefficients and low concentrations, and grouping them into a separate
-category of equations.  Due to their fast reactions these species are very sensitive
-to small changes in the more abundant species.  However, due to their low overall
-concentrations, they do not significantly influence the abundance of species with
-higher concentrations.  Therefore, reactions involving these two species can be
-decoupled from the rest of the network and treated independently.  In fact, $H^-$ 
-and $H_2^+$ are treated as being in equilibrium at all times, independent of 
-the other species and the hydrodynamic variables.  This allows a large speedup
-in solution as the BDF scheme is then applied only to the slower 7- or
-10-species network (depending on whether deuterium is included or not)
-on timescales closer to those required by the hydrodynamics of the simulation.
-Even so, the accuracy and stability of the scheme is maintained by subcycling the 
-rate solver over a single hydrodynamic timestep.  These subcycle timesteps are 
-determined so that the maximum fractional change in the electron concentration is
-limited to no more than $10\%$ per timestep.
+\enzo\ solves the rate equations using a method based on a backwards
+differencing formula (BDF) in order to provide a stable and accurate
+solution. This technique is optimized by taking the chemical
+intermediaries $H^-$ and $H_2^+$, which have large rate coefficients
+and low concentrations, and grouping them into a separate category of
+equations.  Due to their fast reactions these species are very
+sensitive to small changes in the more abundant species.  However, due
+to their low overall concentrations, they do not significantly
+influence the abundance of species with higher concentrations.
+Therefore, reactions involving these two species can be decoupled from
+the rest of the network and treated independently.  In fact, $H^-$ and
+$H_2^+$ are treated as being in equilibrium at all times, independent
+of the other species and the hydrodynamic variables.  This allows a
+large speedup in solution as the BDF scheme is then applied only to
+the slower 7- or 10-species network (depending on whether deuterium is
+included or not) on timescales closer to those required by the
+hydrodynamics of the simulation.  Even so, the accuracy and stability
+of the scheme is maintained by subcycling the rate solver over a
+single hydrodynamic timestep.  These subcycle timesteps are determined
+so that the maximum fractional change in the electron concentration is
+limited to no more than $10\%$ per timestep.  Given this subcycling,
+the method is not truly implicit over an entire hydrodynamical
+timestep, and thus not truly a BDF method -- however, it uses the
+formalism and is implicit over subcycle timesteps.
 
-It is important to note the regime in which this model is valid.  According to Abel et al. and
-Anninos et al. \citep{abel97,anninos97}, the reaction network is valid for temperatures
-between $10^0 - 10^8$ K.  The original model discussed in these two references is only
-valid up to $n_H \sim 10^4$~cm$^{-3}$.  However, addition of the 3-body $H_2$ formation
-process (equation 20 in Table~\ref{table.collisional}) allows 
-correct modeling of the chemistry of the gas up
-until the point where collisionally-induced emission from molecular hydrogen becomes an important
-cooling process, which occurs at $n_H \sim 10^{14}$~cm$^{-3}$.  A further concern is that
-the optically thin approximation for radiative cooling breaks down, which
-occurs before $n_H \sim 10^{16} - 10^{17}$~cm$^{-3}$.  Beyond this point, 
-modifications the cooling function that take into account the non-negligible
-opacity in the gas must be made, as discussed by \citet{2004MNRAS.348.1019R}. 
-Even with these modifications, a completely correct description of the cooling of
-this gas will require some form of radiation transport, which will greatly 
+It is important to note the regime in which this model is valid.
+According to Abel et al. and Anninos et al. \citep{abel97,anninos97},
+the reaction network is valid for temperatures between $10^0 - 10^8$
+K.  The original model discussed in these two references is only valid
+up to $n_H \sim 10^4$~cm$^{-3}$.  However, addition of the 3-body
+$H_2$ formation process (equation 20 in Table~\ref{table.collisional})
+allows correct modeling of the chemistry of the gas up until the point
+where collisionally-induced emission from molecular hydrogen becomes
+an important cooling process, which occurs at $n_H \sim
+10^{14}$~cm$^{-3}$.  A further concern is that the optically thin
+approximation for radiative cooling breaks down, which occurs before
+$n_H \sim 10^{16} - 10^{17}$~cm$^{-3}$.  Beyond this point,
+modifications the cooling function that take into account the
+non-negligible opacity in the gas must be made, as discussed by
+\citet{2004MNRAS.348.1019R}, and was put into \enzo\ for the work
+published in \citep{2009Sci...325..601T,2009PhDT.........5T}.  Even with these
+modifications, a completely correct description of the cooling of this
+gas will require some form of radiation transport, which will greatly
 increase the cost of the simulations.
-
-\red{Several processes are neglected.  The deuterium atom and its processes are 
-completely ignored, which may have some effect.  Recent work shows that HD is
-a more effective coolant than previously thought~\citep{2005MNRAS.361..850L}.  However,
-the fractional abundance of HD is so low that under circumstances relevant to
-the formation of a Population III star in an un-ionized region it should be
-sub-dominant.  However, the enhanced electron fraction in fossil 
-HII regions (as discussed later in this thesis) could result in the HD molecule
-becoming a dominant cooling mechanism at relatively low ($\sim$ few hundred K)
-temperatures, and could potentially cool the gas down to below $100$~K, which can
-enhance fragmentation and could have important consequences for the IMF of 
-primordial stars forming in a relic HII region~\citep{2005MNRAS.364.1378N}.}
-
-Aside from deuterium, the chemical reactions involving lithium 
-are also neglected.  According to \citet{1998A&A...335..403G}, these are 
-not important for the density and temperature regimes explored 
-by the simulations discussed in this thesis.  However, at higher densities it 
-is possible that there are regimes where lithium can be an important coolant.
 
@@
 (6) & He$^{++}$ &+& e$^-$ & $\rightarrow$ & He$^+$ &+& $\gamma$ \\
+\hline
 (7) & H &+& e$^-$ &$\rightarrow$& H$^-$ &+& $\gamma$  \\
@@
 (19) & H$_2^+$ & + & H$^-$ & $\rightarrow$ & H$_2$ & + & H  \\
-(20) & 3H & & & $\rightarrow$ & H$_2$ & + & H
+(20) & 2H & + & H$_2$ & $\rightarrow$ & 2H$_2$ &  &   \\
+(21) & 2H & + & H & $\rightarrow$ & H$_2$ & + & H  \\
+(22) & H$_2$ & + & H$_2$ & $\rightarrow$ & H$_2$ & + & 2H  \\
+(23) & 3H & & & $\rightarrow$ & H$_2$ & + & H \\
+\hline
+(24) & D & + & e$^-$ & $\rightarrow$ & D$^+$ &+& 2e$^-$ \\
+(25) & D$^+$ &+ &e$^-$ & $\rightarrow$ & D &+ &$\gamma$ \\
+(26) & H$^+$ &+ &D & $\rightarrow$ & H &+ &D$^+$ \\
+(27) & H &+ &D$^+$ & $\rightarrow$ & H$^+$ &+ &D \\
+(28) & H$_2$ &+ &D$^+$ & $\rightarrow$ & HD &+ &D$^+$ \\
+(29) & HD &+ &H$^+$ & $\rightarrow$ & H$_2$ &+ &D$^+$ \\
+(30) & H$_2$ &+ &D & $\rightarrow$ & HD &+ &H \\
+(31) & HD &+ &H & $\rightarrow$ & H$_2$ &+ &D \\
+
 \end{tabular}
 \caption[]{Collisional processes solved in the Enzo nonequilibrium
-primordial chemistry routines.}
+primordial chemistry routines. \red{Missing 3-body rate for D$_2$?}}
 \label{table.collisional}
@@
 \caption[]{Radiative processes solved in the Enzo nonequilibrium
-primordial chemistry routines.}
+primordial chemistry routines. \red{Missing rates for D$_2$?}}
 \label{table.radiative}

### 2013-03-18 code-tests.tex
corrected tex typo

@@
   \gamma_{ef} < 1,\\
-  1.0 + 0.2 (\gamma_{ef} - 4/3) - 2.9 (\gamma_{ef} - 4/3)^{2}, & \gamma_{ef} > 1.}
+  1.0 + 0.2 (\gamma_{ef} - 4/3) - 2.9 (\gamma_{ef} - 4/3)^{2}, & \gamma_{ef} > 1.
 \end{array} \right.
@@
 
-\begin{figure} \label{fig.onezone}
+\begin{figure}
   \begin{center}
     \includegraphics[width=1.0\textwidth]{figures/OneZoneCollapseTest.eps}
+  \end{center}
   \caption{Evolution of temperature versus number density for a one-zone
@@
     rates calculated with Cloudy.}
-    \label{fig.1zone}
-  \end{center}
+  \label{fig.onezone}
 \end{figure}
@@
 \label{sec.tests.raytracing}
-\red{(John)}
-Photo-evaporation of a dense clump.  Tests ray tracing, specifically
-the shadowing effects of the clump, and the hydrodynamic response of
-the clump.
+
+The photo-evaporation of dense clumps is prevalent in radiation
+hydrodynamics simulations, and this test problem examines the
+ionization front propagation into the dense clump, shadowing effects
+behind the clump, and the hydrodynamic response on the clump from
+photo-heating.  The problem setup is the same as Test 7 in the
+Cosmological Radiative Transfer Comparison Project
+\citep{IlievEtAl2009} and \citet{Wise11_Moray}.  The simulation domain
+is 6.6~kpc on a side with an ambient medium of pure neutral hydrogen
+$n_{\rm H} = 2 \times 10^{-4}\; \cubecm$ and $T = 8000 \unit{K}$.  We
+place a spherical overdensity in hydrostatic equilibrium with the
+ambient medium.  It has a radius $r = 0.8 \unit{kpc}$, hydrogen
+density $n_{\rm H} = 0.04 \cubecm$, and temperature $T = 40$ K, and it
+is centered at $(x,y,z) = (5, 3.3, 3.3) \unit{kpc}$.  In
+\citet{IlievEtAl2009}, all of the codes used a fixed $128^3$ grid to
+ease the comparison, but in this test to demonstrate a higher
+resolution AMR solution, we employ a $128^3$ grid with two additional
+levels of refinement with cells with a baryon mass (method 2 in
+\S\ref{sec:refinement_criteria}) greater than 1.5 being flagged for
+refinement.
+
+This cloud is subject to radiation from a point source at the center
+of the $y=0$ boundary with an ionizing photon luminosity
+$\dot{N}_\gamma = 3 \times 10^{51}$ photons s$^{-1}$, corresponding to
+a flux $F_0 = 10^6 \unit{photons s}^{-1} \unit{cm}^{-2}$ at the clump
+surface closest to the radiation source.  The radiation source has a
+spectrum of a $T = 10^5 \unit{K}$ blackbody, and we use four energy
+groups with the following mean energies and relative luminosities:
+$E_i = (17.98, 31.15, 49.09, 76.98) \unit{eV}, L_i/L = (0.23, 0.36,
+0.24, 0.06)$ that are optimized to reduce errors in the solutions with
+a full spectrum and energy discretization \citep{Mirocha12}.  Note
+that this choice of energy groups is different from
+\citet{Wise11_Moray}.
 
@@
 anisotropic).
+
+%%% Local Variables: 
+%%% mode: latex
+%%% TeX-master: "ms"
+%%% End:

### 2013-03-18 overview.tex

@@
 \frac{\partial \rho \vecv}{\partial t}  
-   + \frac{1}{a} \nabla \cdot \left(\rho \vecv \vecv + p^* - \vecB \vecB \right) & = &
+   + \frac{1}{a} \nabla \cdot \left(\rho \vecv \vecv + \myvec{I}p^* - \frac{\vecB \vecB}{a} \right) & = &
     - \frac{\dot{a}}{a} \rho \vecv - \frac{1}{a} \rho \nabla \phi
@@
 \frac{\partial E}{\partial t} 
-+ \frac{1}{a} \nabla \cdot \left[ (E + p^*) \vecv -  \vecB (\vecB \cdot \vecv) \right] & = &
-     - \frac{\dot{a}}{a} \left( 2E - \frac{\vecB^2}{2} \right)  - \frac{1}{a} \vecv \cdot \nabla \phi 
++ \frac{1}{a} \nabla \cdot \left[ (E + p^*) \vecv -  \frac{1}{a} \vecB (\vecB \cdot \vecv) \right] & = &
+     - \frac{\dot{a}}{a} \left( 2E - \frac{B^2}{2a} \right)  - \frac{1}{a} \vecv \cdot \nabla \phi 
      %
@@
 %
-Here $E$, $\rho$, $\vecv$, and $\vecB$ are the comoving total fluid energy, comoving baryonic density, peculiar velocity, comoving magnetic field strength, respectively.  
+Here $E$, $\rho$, $\vecv$, and $\vecB$ are the comoving total fluid energy, comoving baryonic density, peculiar velocity, and comoving magnetic field strength, respectively.
+$\myvec{I}$ is the identity matrix.
 The first equation represents conservation of mass, the second, conservation of momentum, and the third, conservation of total fluid (kinetic plus thermal) energy.
 They are respectively, the first, second and third moments of the Boltzmann equation.
+The fourth equation is the magnetic induction equation.
 Terms representing radiative cooling ($\Lambda$) and heating ($\Gamma$) enter on the right-hand side of the energy equation (\ref{eq:total_energy}).
 
-The total fluid energy ($E$) is given by:
+The total fluid energy $E$ is given by:
 \begin{equation}
-E =  \rho e + \frac{1}{2} \vecv \cdot \vecv + \vecB \cdot \vecB
+E =  \rho e + \frac{\rho v^2}{2}  + \frac{B^2}{2a}
         \label{eq:total_energy_def}.
@@
 \begin{equation}
-p^* = p + \frac{1}{2} \vecB^2
+p^* = p + \frac{B^2}{2a}
 \end{equation}
@@
 The equation of state is shown here for an ideal gas with a ratio of specific heats $\gamma$.
-Equation (\ref{eq:potential}) includes contributions to the gravitational potential $\phi$ from the total mass density $\rho_{\rm total}$).
-% as well as any radiation or massless neutrino background ($p_{total}^\prime$) or cosmological constant ($\Lambda_c$), if we are interested in the cosmological case.
+Equation (\ref{eq:potential}) includes contributions to the gravitational potential $\phi$ from the total mass density $\rho_{\rm total}$.
+Although we write the equations including both MHD and comoving coordinates, the code can operate both in the pure hydrodynamic limit, and without cosmological expansion.
 
@@
 
-For completeness, we note here that as part of the usual transformation to comoving coordinates, we have defined the following comoving quantities (primes indicate variables defined in the usual, proper, coordinate system).
-
-This set of coupled hydrodynamic plus gravity equations is the standard one that \enzo solves.
-However, in the cosmological case it is useful to transform these to a coordinate system that is comoving with the expanding universe \citep{Peebles93} (but note that \enzo\ can also operate in a fixed coordinate system).
+For completeness, we note here that as part of the usual transformation to comoving coordinates, we have defined a set of comoving quantities.
+See, for example, \citet{Peebles93} for more details.
 Specifically, we define:
@@
 p       & = & a^3 \pp, \\
-E       & = & \Ep - 
+E       & = & a^3 (\Ep - 
               \dot{a} \vecx \cdot \vecvp - 
-              \frac{1}{2} \dot{a}^2 \vecx^2, \\
+              \frac{1}{2} \dot{a}^2 \vecx^2), \\
 \phi    & = & \phip + \frac{1}{2} a \ddot{a} \vecx^2. \\
@@
 \end{eqnarray}
-where primes indicate the usual proper quantities.
+where primes indicate the usual quantities defined in a fixed frame frame, $\vecr$ is the proper distance and $\vecx$ is the comoving distance.
 
@@
 
-\subsubsection{Magnetohydrodynamics}
+%\subsubsection{Magnetohydrodynamics}
 
-\subsubsection{Radiation transport: ray-casting}
+\subsubsection{Radiation transport: ray-tracing}
 

### 2013-03-18 numerical-chemistry.tex
updated chemistry to reflect matt's feedback re: D2 and HD.
small updates to correct numbering.

@@
 \caption[]{Collisional processes solved in the Enzo nonequilibrium
-primordial chemistry routines. \red{Missing 3-body rate for D$_2$?}}
+primordial chemistry routines.}
 \label{table.collisional}
@@
 \begin{tabular}{llllllll}
-(21) & H & + & $\gamma$ & $\rightarrow$ & H$^+$ & + & e$^-$ \\
-(22) & He & + & $\gamma$ & $\rightarrow$ & He$^+$ & + & e$^-$ \\
-(23) & He$^+$ & + & $\gamma$ & $\rightarrow$ & He$^{++}$ & + & e$^-$ \\
-(24) & H$^-$ & + & $\gamma$ & $\rightarrow$ & H & + & e$^-$ \\
-(25) & H$_2$ & + & $\gamma$ & $\rightarrow$ & H$_2^+$ & + & e$^-$ \\
-(26) & H$_2^+$ & + & $\gamma$ & $\rightarrow$ & H & + & H$^+$ \\
-(27) & H$_2^+$ & + & $\gamma$ & $\rightarrow$ & 2H$^+$ & + & e$^-$ \\
-(28) & H$_2$ & + & $\gamma$ & $\rightarrow$ & H$_2^*$ & $\rightarrow$ & 2H \\
-(29) & H$_2$ & + & $\gamma$ & $\rightarrow$ & 2H &  & 
+(32) & H & + & $\gamma$ & $\rightarrow$ & H$^+$ & + & e$^-$ \\
+(33) & He & + & $\gamma$ & $\rightarrow$ & He$^+$ & + & e$^-$ \\
+(34) & He$^+$ & + & $\gamma$ & $\rightarrow$ & He$^{++}$ & + & e$^-$ \\
+(35) & H$^-$ & + & $\gamma$ & $\rightarrow$ & H & + & e$^-$ \\
+(36) & H$_2$ & + & $\gamma$ & $\rightarrow$ & H$_2^+$ & + & e$^-$ \\
+(37) & H$_2^+$ & + & $\gamma$ & $\rightarrow$ & H & + & H$^+$ \\
+(38) & H$_2^+$ & + & $\gamma$ & $\rightarrow$ & 2H$^+$ & + & e$^-$ \\
+(39) & H$_2$ & + & $\gamma$ & $\rightarrow$ & H$_2^*$ & $\rightarrow$ & 2H \\
+(40) & H$_2$ & + & $\gamma$ & $\rightarrow$ & 2H &  & 
 \end{tabular}
 \caption[]{Radiative processes solved in the Enzo nonequilibrium
-primordial chemistry routines. \red{Missing rates for D$_2$?}}
+primordial chemistry routines. \red{Missing rates for HD destruction/ionization?}}
 \label{table.radiative}

### 2013-03-18 code-tests.tex
first crack at section on code comparisons.

@@
 
+Ensuring that a complex piece of software is behaving correctly is a
+non-trivial task.  While there are a range of techniques that one can
+apply to ensure correctness, the \enzo\ code uses two primary methods:
+a suite of test problems that can be compared to previous versions of
+the code, to ensure that \enzo\ is, e.g., running correctly on a new
+computer, with a new compiler, or after substantial changes have been
+made to the code base; and by direct comparisons to other
+astrophysical fluid dynamics codes.  We describe the \enzo\ test
+methodology in Section~\ref{sec.tests.suite}, code comparisons in
+Section~\ref{sec.tests.compare}, and show a small set of
+representative test problems in Section~\ref{sec.tests.problems}.
+
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 \label{sec.tests.vandv}
+
+\subsubsection{The \enzo\ test suite}
+\label{sec.tests.suite}
 
@@
 After the test problems are selected by the test runner script, they
-are run in succession either by with \enzo\ executable contained within
-that distribution or an external executable built from another version
-and specified by the user.  This allows for any version of the code to
-be tested with an identical set of test problems.  After running the
-test problem simulations, the test runner then performs a series of
-basic analysis tasks using the \texttt{yt} analysis toolkit
+are run in succession either by with \enzo\ executable contained
+within that distribution or an external executable built from another
+version and specified by the user.  This allows for any version of the
+code to be tested with an identical set of test problems.  After
+running the test problem simulations, the test runner then performs a
+series of basic analysis tasks using the \texttt{yt} analysis toolkit
 \citep{2011ApJS..192....9T, 2011arXiv1112.4482T}.  The default
@@
 script.  Alternately, results from any version of \enzo\ can be stored
-locally and compared to any other version of the code.  Below, we
-describe some of the key test problems which are used to verify proper 
-behavior.
+locally and compared to any other version of the code.  In
+Section~\ref{sec.tests.problems}, we describe some of the key test
+problems which are used to verify proper behavior.  All of these test
+problems, as well as the scripts to generate the figures from them,
+are available in the \enzo\ test suite.
 
-\red{(Brian)} Refer to all code comparisons that Enzo has been part of.
+\subsubsection{Code comparisons}
+\label{sec.tests.compare}
+
+Over the course of its existence, \enzo\ has been involved in several
+comparisons with other astrophysical codes used for self-gravitating
+fluid dynamics.
+
+Cosmology: 
+Santa Barbara Clusters, \citep{SantaBarbara}.
+Enzo-gadget cosmology comparison, \citep{2005ApJS..160....1O}.
+Regan et al. Enzo/Gadget Lyman Alpha Forest comparison,
+\citep{2007MNRAS.374..196R}.
+Heitmann n-body comparison, \citep{2008CS&D....1a5003H}
+
+Hydro:
+Agertz hydro comparison, \citep{2007MNRAS.380..963A}.
+Tasker test suite, \citep{Tasker2008}.
+
+Turbulence:
+Kritsuk supersonic MHD turbulence comparison,
+\citep{2011ApJ...737...13K}.  Compares MHD-CT version.
+
+Radiation transport:
+Iliev et al. radiation transport comparison, paper II,
+\citep{IlievEtAl2009}.  Compares enzo-FLD.
 

### 2013-03-18 code-tests.tex
added more comparison papers.
Finishing shadowing test section along with a figure.

@@
 \citep{IlievEtAl2009}.  Compares enzo-FLD.
+
+Other:
+Turk et al. primordial star formation comparison,
+\citep{2011ApJ...726...55T}.  Passy et al. common envelope binary
+simulations (SPH-uniform grid comparison), \citep{2012ApJ...744...52P}.
 
@@
 \S\ref{sec:refinement_criteria}) greater than 1.5 being flagged for
-refinement.
+refinement.  This test is run for 15 Myr.
 
@@
 that this choice of energy groups is different from
-\citet{Wise11_Moray}.
+\citet{Wise11_Moray}.  We use a minimum angular resolution of 10 rays
+per cell and a constant radiative transfer timestep of 25 kyr.  Figure
+\ref{fig:shadowing} depicts the clump after at $t = 15$ Myr with the
+outer layers expanding after being photo-heated.  It also shows the
+sharp shadowing effects of the dense clump in the neutral fraction
+plot that is representative of ray tracing techniques.
+
+\begin{figure}
+  \centering
+  \includegraphics[width=1.0\textwidth]{figures/code-test-shadowing.eps}
+  \caption{Photo-evaporation of a dense clump.  Clockwise from the
+    upper left-hand side: slices through the clump center of the
+    density, temperature, neutral fraction, and pressure.}
+  \label{fig:shadowing}
+\end{figure}
 

### 2013-03-19 code-tests.tex
updating code test section

@@
 Section~\ref{sec.tests.compare}, and show a small set of
-representative test problems in Section~\ref{sec.tests.problems}.
+representative test problems in Section~\ref{sec.tests.problems}.  We
+further note that the \enzo\ test suite
+(Section~\ref{sec.tests.suite}) contains hundreds of test problems, as
+well as the ability to compare to a ``gold standard'' solution, and
+thus all of the tests shown here are easily reproducible by the
+reader.
 
@@
 \citep{2007MNRAS.374..196R}.
-Heitmann n-body comparison, \citep{2008CS&D....1a5003H}
+Heitmann n-body comparison, \citep{2008CS&D....1a5003H}.
+Vazza shock comparison (galaxy clusters), \citep{2011MNRAS.418..960V}.
+Voit et al. cluster comparison, \citep{2005MNRAS.364..909V}.
 
@@
 Tasker test suite, \citep{Tasker2008}.
+McNally et al. Kelvin-Helmholz test, \citep{2012ApJS..201...18M}.
+Robertson et al. Galilean invariance, \citep{2010MNRAS.401.2463R}.
 
 Turbulence:
+Kitsionas et al. decaying isothermal supersonic turbulence, \citep{2009A&A...508..541K}.
 Kritsuk supersonic MHD turbulence comparison,
@@
 Radiation transport:
-Iliev et al. radiation transport comparison, paper II,
-\citep{IlievEtAl2009}.  Compares enzo-FLD.
+Iliev et al. radiation transport comparison, paper II, \citep{IlievEtAl2009}.  Compares enzo-FLD.
 
 Other:
-Turk et al. primordial star formation comparison,
-\citep{2011ApJ...726...55T}.  Passy et al. common envelope binary
+Turk et al. primordial star formation comparison, \citep{2011ApJ...726...55T}.  
+Passy et al. common envelope binary
 simulations (SPH-uniform grid comparison), \citep{2012ApJ...744...52P}.

### 2013-03-19 numerical-amr.tex
Formatting police passing through.

@@
 \begin{equation}
-q^{\rm coarse} = \tilde{q}^{\rm coarse} - \Delta t \left( F^{\rm coarse} - \sum_{j,k} F^{fine}_{j,k} \right)
+  q^{\rm coarse} = \tilde{q}^{\rm coarse} - \Delta t \left( F^{\rm
+      coarse} - \sum_{j,k} F^{\rm fine}_{j,k} \right)
 \end{equation}
@@
 \begin{equation}
-\Delta x < \left( \frac{\pi k_B T}{J^2 G \rho m_H} \right)^{1/2}
+\Delta x < \left( \frac{\pi k_B T}{J^2 G \rho \mh} \right)^{1/2}
 \end{equation}
@@
 However, in some cases it is desirable to resolve the cooling process, in which case we flag on cells which have a cooling time shorter than their sound crossing time.
-The cooling time is given by $t_{\rm cool} = 1.5 kT / (n \Lambda(T)$ and the sound crossing time is $t_{\rm cross} = \Delta x / c_s$, where $c_s$ is the sound speed.
+The cooling time is given by $t_{\rm cool} = 1.5 kT / n \Lambda(T)$ and the sound crossing time is $t_{\rm cross} = \Delta x / c_s$, where $c_s$ is the sound speed.
 
@@
 
-\item{\em Optical depth.} This refinement criterion ensures that cells on the highest level of refinement have an HI optical depth lower than unity.  The optical depth of a cell is estimated to be:
+\item{\em Optical depth.} This refinement criterion ensures that cells on the highest level of refinement have an \ion{H}{1} optical depth lower than unity.
+The optical depth of a cell is estimated to be:
 
 \begin{equation}
-\tau = \sigma_{HI} n_{HI} \Delta x 
+\tau = \sigma_{\rm HI} n_{\rm HI} \Delta x 
 \end{equation}
 
-where $\sigma_{HI}$ is the neutral atomic hydrogen absorption cross section, $n_{HI}$ is the proper HI number density, and $\Delta x$ is the cell width.
-If $\tau > 1$, the cell is refined further, up to the maximum level of refinement.
+where $\sigma_{\rm HI}$ is the neutral atomic hydrogen absorption
+cross section, $n_{\rm HI}$ is the proper \ion{H}{1} number density,
+and $\Delta x$ is the cell width.  If $\tau > 1$, the cell is refined
+further, up to the maximum level of refinement.
 

### 2013-03-19 numerical-analysis.tex
Formatting police passing through.

@@
 \label{eqn.drifttrace}
-x^{n+1/2} & = & x^n + (\Delta t/2) v^{interp,n} \nonumber \\
-v^{n+1} & = & v^{interp,n+1} \\
+x^{n+1/2} & = & x^n + (\Delta t/2) v^{\rm interp,n} \nonumber \\
+v^{n+1} & = & v^{\rm interp,n+1} \\
 x^{n+1} & = & x^{n+1/2} + (\Delta t/2) v^{n+1} \nonumber
@@
 by linearly interpolating the cell-centered baryon velocity to the
-position of the particle ($v^{interp}$), and assigning it to that value.
+position of the particle ($v^{\rm interp}$), and assigning it to that value.
 

### 2013-03-19 numerical-overview.tex
Formatting police passing through.

@@
 sub-dominant.  However, the enhanced electron fraction in fossil 
-HII regions (as discussed later in this thesis) could result in the HD molecule
+\ion{H}{2} regions (as discussed later in this thesis) could result in the HD molecule
 becoming a dominant cooling mechanism at relatively low ($\sim$ few hundred K)
@@
 enhance fragmentation and could have important consequences for the IMF of 
-primordial stars forming in a relic HII region~\citep{2005MNRAS.364.1378N}.
+primordial stars forming in a relic \ion{H}{2} region~\citep{2005MNRAS.364.1378N}.
 

### 2013-03-19 numerical-radiation-fld.tex
Formatting police passing through.

@@
 \end{eqnarray}
-where $E_r$ is a grey radiation energy density, $e_c$ is the
-internal energy correction due to photo-heating and chemical cooling,
-and ${\mathrm n}_i$ correspond to the chemical number densities of
-HI, HII, HeI, HeII and HeIII, and ${\mathrm n}_e$ is the electron
-number density.  Here, we define the grey radiation energy density
-through first assuming a fixed frequency spectrum, i.e.
-$E_{\nu}(\nu,\vec{x},t) = \tilde{E}_r(\vec{x},t) \chi(\nu)$, and then
-defining the integrated quantity
+where $E_r$ is a grey radiation energy density, $e_c$ is the internal
+energy correction due to photo-heating and chemical cooling, and
+${\mathrm n}_i$ correspond to the chemical number densities of
+\ion{H}{1}, \ion{H}{2}, \ion{He}{1}, \ion{He}{2} and \ion{He}{3}, and
+${\mathrm n}_e$ is the electron number density.  Here, we define the
+grey radiation energy density through first assuming a fixed frequency
+spectrum, i.e.  $E_{\nu}(\nu,\vec{x},t) = \tilde{E}_r(\vec{x},t)
+\chi(\nu)$, and then defining the integrated quantity
 \begin{equation}

### 2013-03-19 numerical-timestep.tex
Formatting police passing through.

@@
 \begin{equation}
-\Delta t_{hydro} = min \left( \kappa_{hydro} \frac{a \Delta x}{c_{s} + |v_x|} \right)_L ,
+\Delta t_{\rm hydro} = \min \left( \kappa_{\rm hydro} \frac{a \Delta x}{c_{s} + |v_x|} \right)_L ,
 \label{eqn:dthydro}
@@
 \begin{equation}
-\Delta t_{MHD} = min \left( \kappa_{MHD} \frac{a \Delta x}{v_{f} + |v_x|} \right)_L ,
+\Delta t_{\rm MHD} = \min \left( \kappa_{\rm MHD} \frac{a \Delta x}{v_{f} + |v_x|} \right)_L ,
 \label{eqn:dtMHD}
@@
 \begin{equation}
-\Delta t_{dm} = min \left(\kappa_{dm} \frac{a \Delta x}{v_{dm,x}} \right)_L ,
+\Delta t_{\rm dm} = \min \left(\kappa_{\rm dm} \frac{a \Delta x}{v_{dm,x}} \right)_L ,
 \label{eqn:dtdarkmatter}
@@
 \begin{equation}
-\Delta t_{accel} = min \left( \sqrt{\frac{\Delta x}{\vec{g}}} \right)_L ,
+\Delta t_{\rm accel} = \min \left( \sqrt{\frac{\Delta x}{\vec{g}}} \right)_L ,
 \label{eqn:dtaccel}
@@
 \begin{equation}
-\Delta t_{rad} = min \left(  \sqrt{\frac{\Delta x}{\vec{a}_{rad}} \right)_L,
+\Delta t_{\rm rad} = \min \left(  \sqrt{\frac{\Delta x}{\vec{a}_{rad}}} \right)_L,
 \label{eqn:dtrad}
@@
 \begin{equation}
-\Delta t_{cond} = min \left(  \frac{ \kappa_{cond}}{f_{Sp}} \frac{\Delta x^2
+\Delta t_{\rm cond} = \min \left(  \frac{ \kappa_{\rm cond}}{f_{\rm Sp}} \frac{\Delta x^2
     n_b}{\kappa(T)} \right)_L,
@@
 \begin{equation}
-\Delta t_{exp} = f_{exp} \left( \frac{a}{\dot{a}} \right) ,
+\Delta t_{\rm exp} = f_{\rm exp} \left( \frac{a}{\dot{a}} \right) ,
 \label{eqn:dtexpand}
@@
 
- In equations~\ref{eqn:dthydro}-\ref{eqn:dtcond}, the $min ( \ldots
+In equations~\ref{eqn:dthydro}-\ref{eqn:dtcond}, the $\min ( \ldots
 )_L$ formalism means that this value is calculated for all cells on a
@@
 that any changes in the fluid propagate less than a single grid
-spacing, $\Delta x$.  In this equation, $\kappa_{hydro}$ is a
-numerical constant with a value of $0 < \kappa_{hydro} \leq 1$ (with a
-typical value of $\kappa_{hydro} \sim 0.3-0.5$) that ensures that the
-CFL condition is always met, and $c_s$ and $v_x$ are the sound speed
-and peculiar baryon velocity in a given cell.
+spacing, $\Delta x$.  In this equation, $\kappa_{\rm hydro}$ is a
+numerical constant with a value of $0 < \kappa_{\rm hydro} \leq 1$
+(with a typical value of $\kappa_{\rm hydro} \sim 0.3-0.5$) that
+ensures that the CFL condition is always met, and $c_s$ and $v_x$ are
+the sound speed and peculiar baryon velocity in a given cell.
 
@@
 harmonic average of the timestep found along each of the coordinate
-axes yields a maximum $\kappa_{hydro} = 0.8$.  So letting $\Delta
+axes yields a maximum $\kappa_{\rm hydro} = 0.8$.  So letting $\Delta
 t_x$, $\Delta t_y$, and $\Delta t_z$ be the analogues of
@@
 \begin{equation}
-\Delta t_{hydro} = min ( \kappa_{hydro} /( \frac{1}{\Delta t_x}
-+\frac{1}{\Delta t_y} + \frac{1}{\Delta t_z} ) )_L
+  \Delta t_{\rm hydro} = \min \left( \frac{\kappa_{\rm hydro}} {1/\Delta t_x
+  +1/\Delta t_y + 1/\Delta t_z} \right)_L
 \end{equation}
@@
 magnetohydrodynamics are being solved, and is directly analogous to
-Equation~\ref{eqn:dthydro} in that it ensures that the CLF condition
-is being enforced at all times.  In this equation, $\kappa_{MHD}$ is a
-numerical constant with a value of $0 < \kappa_{MHD} \leq 1$ (with a
-typical value of $\kappa_{MHD} \sim 0.5$) that ensures that the CFL
+Equation~\ref{eqn:dthydro} in that it ensures that the CFL condition
+is being enforced at all times.  In this equation, $\kappa_{\rm MHD}$ is a
+numerical constant with a value of $0 < \kappa_{\rm MHD} \leq 1$ (with a
+typical value of $\kappa_{\rm MHD} \sim 0.5$) that ensures that the CFL
 condition is always met, and $v_f$ and $v_x$ are the ``fast wave
@@
 expressed as:
-
+%
 \begin{equation}
@@
 \end{equation}
-
-Where $c_s$ is the sound speed and $v_A$ is the Alfven speed, calculated
+%
+where $c_s$ is the sound speed and $v_A$ is the Alfven speed, calculated
 as $v_A = \sqrt{B^2/\rho_B}$ in units where $\mu_0 = 1$.
@@
 by requiring that no dark matter particle travels more than one cell
-width.  The parameter $\kappa_{dm}$ is analogous to
-$\kappa_{dthydro}$, with an identical range of values.
+width.  The parameter $\kappa_{\rm dm}$ is analogous to $\kappa_{\rm
+  hydro}$, with an identical range of values.
 Equation~\ref{eqn:dtexpand} limits the timestep such that the
 expansion parameter a can only change by a fractional amount of
-$f_{exp} = \Delta a/a$, where $f_{exp}$ is a user-defined parameter
-and has typical values of $f_{exp} = 0.01-0.02$.  This is required for
-the stability of the PPM algorithm in comoving coordinates, and
-typically limits the timestep only at high redshifts when densities
-are relatively homogeneous.
+$f_{\rm exp} = \Delta a/a$, where $f_{\rm exp}$ is a user-defined
+parameter and has typical values of $f_{\rm exp} = 0.01-0.02$.  This
+is required for the stability of the PPM algorithm in comoving
+coordinates, and typically limits the timestep only at high redshifts
+when densities are relatively homogeneous.
 
@@
 gravitational acceleration in each cell on level L.  In
-Equation~\ref{eqn:dtrad}, $\vec{a}_{rad}$ is the estimated
+Equation~\ref{eqn:dtrad}, $\vec{a}_{\rm rad}$ is the estimated
 acceleration due to radiation pressure in each cell on level L,
@@
 conduction suppression factor whose value must be f$_{Sp} \leq 1$.  Both are defined in
-Section~\ref{sec.num.conductions}.  \kappa_{cond} is a prefactor whose
-value must be $0 < \kappa_{cond}  < 0.5$, and is exactly 0.5 for the
+Section~\ref{sec.num.conductions}.  $\kappa_{\rm cond}$ is a prefactor whose
+value must be $0 < \kappa_{\rm cond}  < 0.5$, and is exactly 0.5 for the
 implemention in Enzo.
@@
 limits the timestep so that the simulated universe only expands by
-some fractional amount, f$_{exp}$, during a single step.  In this
+some fractional amount, f$_{\rm exp}$, during a single step.  In this
 equation, $a$ and $\dot{a}$ refer to the scale factor of the universe
@@
 at later times, other timestep criteria are much more important.
+
+%%% Local Variables: 
+%%% mode: latex
+%%% TeX-master: "ms"
+%%% End:

### 2013-03-19 overview.tex
Formatting police passing through.

@@
 \begin{equation}
-\mathbb{F} = -\kappa_{sp} \nabla T
+\mathbb{F} = -\kappa_{\rm sp} \nabla T
 \end{equation}
 
-Where $\kappa_{sp}$ is the Spitzer conduction coefficient
+Where $\kappa_{\rm sp}$ is the Spitzer conduction coefficient
 \citep{1962pfig.book.....S}.  Saturation of the heat flux in
@@
 \begin{equation}
-\mathbb{F} = -\kappa_{sp} \hat{b} (\hat{b} \cdot \nabla T)
+\mathbb{F} = -\kappa_{\rm sp} \hat{b} (\hat{b} \cdot \nabla T)
 \end{equation}
@@
 
-Enzo may operate in a number of modes with regard to radiative cooling and chemistry.
-In the simplest mode, with the multi-species flag turned off (i.e.
-set to zero), the cooling rate is computed from a simple temperature-dependent cooling rate, taken from \citet{SW87}.
+Enzo may operate in a number of modes with regard to radiative cooling
+and chemistry.  In the simplest mode, with the multi-species flag
+turned off (i.e. set to zero), the cooling rate is computed from a
+simple temperature-dependent cooling rate, taken from \citet{SW87}.
 
-In addition to this, \enzo\ includes the capability of following up to 12 chemical species using a non equilibrium-solver.
-The species can be turned on in sets, with the simplest model including just H, H$^+$, He, He$^+$, He$^++$, and $e^-$, and more complete models adding first species important for gas-phase molecular hydrogen formation: $H^-$, H$_2$ and H$_2^+$, and then $HD$ formation: $HD$, $D$, $D+$.
-The cooling and heating due to these species is also included.
-The solution of the rate equations is carried out using a time-backwards differencing which insures stability; to maintain accuracy we sub-cycle the rate equations with a shorter timestep such that the electron and neutral fractions do not change by more than 10\%.
+In addition to this, \enzo\ includes the capability of following up to
+12 chemical species using a non equilibrium-solver.  The species can
+be turned on in sets, with the simplest model including just H, \Hp,
+He, \Hep, \Hepp, e$^-$, and more complete models adding first species
+important for gas-phase molecular hydrogen formation: \Hm, \HH~and
+\HHp, and then HD formation: HD, D, \Dp.  The cooling and heating due
+to these species is also included.  The solution of the rate equations
+is carried out using a time-backwards differencing which insures
+stability; to maintain accuracy we sub-cycle the rate equations with a
+shorter timestep such that the electron and neutral fractions do not
+change by more than 10\%.
 

### 2013-03-19 numerical-ppm.tex
Many minor changes.  No more undefined references!  Added appendix file.

@@
 for each of $p$, $\rho$, and $v$  The pressure is determined from
-eq.~(\ref{eq:state}), the equation of state.
+eq.~(\ref{eq:eq_of_state}), the equation of state.
 %We have dropped the subscript b in this and the following two sections because we will be referring solely to baryonic quantities.  

### 2013-03-19 code-tests.tex
finished section 4.1, including all of the code comparisons Enzo has been part of.  I decided to move the cosmological code comparisons to this section, rather than putting them down in the test problems in section 4.2 - it just didn't seem to make a ton of sense to do it the other way after I got started.

--- 2013-03-18 code-tests.tex
@@
 comparisons with other astrophysical codes used for self-gravitating
-fluid dynamics.
+fluid dynamics.  In general, \enzo\ behaves in a manner similar to other
+grid-based (and particularly AMR-based) codes, as we will summarize below.
 
-Cosmology: 
-Santa Barbara Clusters, \citep{SantaBarbara}.
-Enzo-gadget cosmology comparison, \citep{2005ApJS..160....1O}.
-Regan et al. Enzo/Gadget Lyman Alpha Forest comparison,
-\citep{2007MNRAS.374..196R}.
-Heitmann n-body comparison, \citep{2008CS&D....1a5003H}
+\enzo\ has been involved in several cosmological code comparisons,
+including the Santa Barbara Cluster code comparison project
+\citep{SantaBarbara}, a large comparison of N-body simulations
+\citep{2008CS&D....1a5003H}, as well as several direct comparisons
+between \enzo\ and the GADGET SPH code under a variety of
+circumstances, including N-body and adiabatic hydrodynamics
+\citep{2005ApJS..160....1O,2005MNRAS.364..909V, 2011MNRAS.418..960V}
+and simulations of the Lyman-alpha forest \citep{2007MNRAS.374..196R}.
+The \enzo\ code typically has a more difficult time resolving
+small-scale self-gravitating structures (for an equivalent dark matter
+particle count and nominally equivalent force resolution), but does
+comparably well as a tree-based code for larger structure, and is
+typically superior in terms of resolving fluid features due to its
+higher-order (and artificial viscosity-free) hydrodynamics solver.
+When examining classical test cases such as the Santa Barbara
+project \citep{SantaBarbara}, \enzo\ forms galaxy clusters with very similar density,
+temperature, and entropy profiles to other grid-based codes that use
+Godunov-type hydro methods, which
+systematically differ from particle-based codes using SPH.  Similarly,
+in galaxy cluster simulations that look specifically at the properties
+of cosmological shocks \citep[e.g.][]{2011MNRAS.418..960V}, \enzo\
+produces results that are similar to other high-order grid-based
+hydrodynamics codes, and far superior behavior of \enzo\ in
+low-density regions when compared to a particle-based code.  In tests
+of the Lyman-alpha forest that include radiative cooling and a uniform
+metagalactic ultraviolet background, \enzo\ and Gadget provide results
+on metrics such as the matter power spectrum that are comparable to
+within 5\% \citep{2007MNRAS.374..196R}.
 
-Hydro:
-Agertz hydro comparison, \citep{2007MNRAS.380..963A}.
-Tasker test suite, \citep{Tasker2008}.
+Several comparisons have been made that focus specifically on
+hydrodynamics solvers.  In particular, the work of
+\citet{2007MNRAS.380..963A} and \citet{Tasker2008} perform
+head-to-head comparisons between several grid- and particle-based
+codes for a variety of test problems (including shocked gas clouds,
+self-gravitating, translating clouds, and Sedov-Taylor explosions),
+and show that \enzo\ is comparable or superior in behavior to the
+other grid-based hydrodynamics codes involved in the comparison, and
+provide useful information on whether an AMR code such as \enzo\
+experiences challenges.  More specific comparisons, including one
+specifically testing the linear and nonlinear growth of the
+Kelvin-Helmholz instability \citep{2012ApJS..201...18M} and one more
+broadly examining Galilean invariance in grid-based codes
+\citep{2010MNRAS.401.2463R} show that \enzo, and in particular its
+implementation of the Piecewise Parabolic Method hydrodynamics solver,
+converge to the correct solution as expected, and generally provide
+less diffusive solutions than lower-order codes, including those that
+use artificial viscosity.  Finally, two code comparison projects have
+been undertaken to examine turbulence in hydro codes -- one studied
+the behavior of decaying isothermal supersonic turbulence
+\citep{2009A&A...508..541K}, and the other looked at supersonic
+magnetohydrodynamical turbulence \citep{2011ApJ...737...13K}.  Both
+included the \enzo\ code, with the former testing the PPM
+hydrodynamics and the latter the constrained transport MHD
+implementation.  In both cases, \enzo\ performed similarly to other
+grid-based codes that use Godunov-based fluid solvers, and typically
+had better effective resolution than particle-based codes using the
+same number of particles as the \enzo\ simulation used grid cells.
 
-Turbulence:
-Kritsuk supersonic MHD turbulence comparison,
-\citep{2011ApJ...737...13K}.  Compares MHD-CT version.
-
-Radiation transport:
-Iliev et al. radiation transport comparison, paper II,
-\citep{IlievEtAl2009}.  Compares enzo-FLD.
-
-Other:
-Turk et al. primordial star formation comparison,
-\citep{2011ApJ...726...55T}.  Passy et al. common envelope binary
-simulations (SPH-uniform grid comparison), \citep{2012ApJ...744...52P}.
+Three other comparisons between portions of the \enzo\ code and other
+simulation tools have been performed in recent years.  The
+flux-limited diffusion radiation transport scheme was measured against
+several test problems by \citet{IlievEtAl2009}, producing results that
+are similar to both expected results and other methods, although there
+are some minor differences (although all codes in the study exhibited
+differences from the majority of other codes on at least a subset of
+the tests).  \citet{2011ApJ...726...55T} shows the result of varied
+reaction rate coefficients for the formation of molecular hydrogen via
+the 3-body process in both the \enzo\ and Gadget-2 codes.  Similar
+trends were visible in both the \enzo\ and Gadget simulations, though
+at a nominally equivalent resolution (i.e., with particle and cell gas
+masses being comparable) \enzo\ simulations typically displayed a much
+higher level of gas structure.  This is unsurprising due to \enzo's
+higher-order hydro solver.  Finally, \citet{2012ApJ...744...52P} show
+the results of comparing \enzo\ in its non-AMR mode to a
+smoothed-particle hydrodynamics code, SNSPH, in the context of
+common-envelope binary stellar interactions.  The authors show that
+the codes display reasonable convergence properties as a function of
+simulation resolution, and also agree quite well with each other --
+however, the observed mass-loss rates do not agree particularly well
+with observations.
 
@@
 Problem type 20, unigrid and AMR both.  Tests gravity + hydro.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsubsection{Cosmology simulations}
-\label{sec.tests.}
-\red{(Brian)}
-Talk about cosmology code comparisons, where Enzo is good/weak.
 

### 2013-03-19 code-tests-fld.tex

@@
 deceleration parameter $q_0 = 0.05$ and initial redshift $z_i=10$.
-Here, the physics of interest is the expansion of a HII region in a
+Here, the physics of interest is the expansion of a \ion{H}{2} region in a
 uniform gas around a single monochromatic ionizing source (with
@@
 \[
-   r_S(t) = \left[\frac{3\dot{N}_{\gamma}}{4\pi \alpha_B
-   n_H(t)^2}\right]^{1/3}, 
+   r_S(t) = \left[\frac{3\dot{N}_{\gamma}}{4\pi \alpha_{\rm B}
+   n_{\rm H}(t)^2}\right]^{1/3}, 
 \]
 where $\dot{N}_{\gamma} = 5\cdot10^{48}$ photon sec$^{-1}$ is the
-stength of the ionizing source, $\alpha_B = 2.52\cdot10^{-13}$
-s$^{-1}$ is the HII recombination rate, and $n_H(t)$ is the proper
-number density of hydrogen.  We then define $\lambda = \alpha_B\,
-n_{H,i}\, /\, H_0\, / (1+z_i)$, where the subscript $i$ indicates the
-quantity at the initial redshift.  The analytical solution for the
-I-front position as a function of time (or actually of the
-time/redshift-dependent expansion coefficient $a(t)$) is then
+stength of the ionizing source, $\alpha_{\rm B} = 2.52\cdot10^{-13}$
+s$^{-1}$ is the \ion{H}{2} recombination rate, and $n_H(t)$ is the
+proper number density of hydrogen.  We then define $\lambda =
+\alpha_{\rm B}\, n_{H,i}\, /\, H_0\, / (1+z_i)$, where the subscript $i$
+indicates the quantity at the initial redshift.  The analytical
+solution for the I-front position as a function of time (or actually
+of the time/redshift-dependent expansion coefficient $a(t)$) is then
 \begin{eqnarray*}
@@
 2.35\cdot10^{-28}$ g cm$^{-3}$, and an ionized hydrogen fraction
-${\mathrm n}_{HII}/{\mathrm n}_H = 0$.
+$n_{\rm HII}/n_{\rm H} = 0$.
 
-In the Figure \ref{fig.fld}, we plot profiles of the HI and HII
-fractions at the redshifts $z=\{6.24,\, 2.29,\, 1.02\}$, as well as the
-ratio of the I-front and Str{\" o}mgren radii throughout the
-simulation.  As seen in the profile figure, the I-front width starts
-out quite sharp but slowly spreads as the simulation proceeds due to
-the diffusion approximation to the radiative transfer.  However, as
-seen in the I-front propagation history plot, the computed I-front
-location very accurately captures the exact value.
+In the Figure \ref{fig.fld}, we plot profiles of the \ion{H}{1} and
+\ion{H}{2} fractions at the redshifts $z=\{6.24,\, 2.29,\, 1.02\}$, as
+well as the ratio of the I-front and Str{\" o}mgren radii throughout
+the simulation.  As seen in the profile figure, the I-front width
+starts out quite sharp but slowly spreads as the simulation proceeds
+due to the diffusion approximation to the radiative transfer.
+However, as seen in the I-front propagation history plot, the computed
+I-front location very accurately captures the exact value.
 
@@
 \end{figure}
+
+%%% Local Variables: 
+%%% mode: latex
+%%% TeX-master: "ms"
+%%% End:

### 2013-03-19 numerical-amr.tex

@@
 donor cell, Van-Leer and piecewise parabolic interpolation, which are
-first, second and third-order accurate respectively \citep[e.g.,][]{Stone92a}.
-However, we require fully multi-dimensional interpolation functions, and so
-focus on second-order schemes since they are tractable and
-relatively accurate. 
-\footnote{We could, in principle, build a multi-dimensional interpolator out
-of a one-dimensional interpolation function; however, at higher order, this would not necessarily
-satisfy our constraints on not introducing new extrema.}
-In Appendix~\ref{app:interpolation} we explicitely write down
-functions in one, two and three dimensions, but here we discuss the
-general design considerations.
+first, second and third-order accurate respectively
+\citep[e.g.,][]{Stone92a}.  However, we require fully
+multi-dimensional interpolation functions, and so focus on
+second-order schemes since they are tractable and relatively accurate.
+\footnote{We could, in principle, build a multi-dimensional
+  interpolator out of a one-dimensional interpolation function;
+  however, at higher order, this would not necessarily satisfy our
+  constraints on not introducing new extrema.}  In the Appendix, we
+explicitely write down functions in one, two and three dimensions, but
+here we discuss the general design considerations.
 

### 2013-03-19 numerical-chemistry.tex

@@
 non-equilibrium chemistry of a gas of primordial composition with 12
-total species: $H$, $H^+$, $He$, $He^+$, $He^{++}$, $H^-$, $H_2^+$,
-$H_2$, $e^-$, $D$, $D^+$, and $HD$.  (A separate, but related, module
-follows the radiative heating and cooling of the gas from atomic and
-molecular line excitation, recombination, collisional excitation,
-free-free transitions, Compton scattering of the cosmic microwave
-background, as well as several models for a metagalactic UV background
-that heat the gas via photoionization and discussion -- see
+total species: H, \Hp, He, \Hep, \Hepp, \Hm, \HHp, \HH, e$^-$, D, \Dp,
+and HD.  (A separate, but related, module follows the radiative
+heating and cooling of the gas from atomic and molecular line
+excitation, recombination, collisional excitation, free-free
+transitions, Compton scattering of the cosmic microwave background, as
+well as several models for a metagalactic UV background that heat the
+gas via photoionization and discussion -- see
 Section~\ref{sec.num.cooling} for more details).  Depending on user
-parameters, one can follow simply atomic species ($H$, $H^+$, $He$,
-$He^+$, $He^{++}$, and $e^-$), include those relevant for molecular
-hydrogen formation ($H_2$, $H_2^+$, and $H^-$), and further include
-deuterium and its species ($D$, $D^+$, and $HD$.)  \red{A total of 9
-kinetic equations are solved, including 29 kinetic and radiative
-processes, for the 9 species mentioned above.  See
-Table~\ref{table.collisional} for the collisional processes and
-Table~\ref{table.radiative} for the radiative processes solved.}
+parameters, one can follow simply atomic species (H, \Hp, He, \Hep,
+\Hepp, and e$^-$), include those relevant for molecular hydrogen
+formation (\HH, \HHp, and \Hm), and further include deuterium and its
+species (D, \Dp, and HD).  \red{A total of 9 kinetic equations are
+  solved, including 29 kinetic and radiative processes, for the 9
+  species mentioned above.  See Table~\ref{table.collisional} for the
+  collisional processes and Table~\ref{table.radiative} for the
+  radiative processes solved.}
 
@@
 solution. This technique is optimized by taking the chemical
-intermediaries $H^-$ and $H_2^+$, which have large rate coefficients
+intermediaries \Hm and \HHp, which have large rate coefficients
 and low concentrations, and grouping them into a separate category of
@@
 Therefore, reactions involving these two species can be decoupled from
-the rest of the network and treated independently.  In fact, $H^-$ and
-$H_2^+$ are treated as being in equilibrium at all times, independent
+the rest of the network and treated independently.  In fact, \Hm~and
+\HHp~are treated as being in equilibrium at all times, independent
 of the other species and the hydrodynamic variables.  This allows a
@@
 It is important to note the regime in which this model is valid.
-According to Abel et al. and Anninos et al. \citep{abel97,anninos97},
+According to \citet{abel97} and \citet{anninos97},
 the reaction network is valid for temperatures between $10^0 - 10^8$
@@
 up to $n_H \sim 10^4$~cm$^{-3}$.  However, addition of the 3-body
-$H_2$ formation process (equation 20 in Table~\ref{table.collisional})
+\HH~formation process (equation 20 in Table~\ref{table.collisional})
 allows correct modeling of the chemistry of the gas up until the point
 where collisionally-induced emission from molecular hydrogen becomes
-an important cooling process, which occurs at $n_H \sim
+an important cooling process, which occurs at $n_{\rm H} \sim
 10^{14}$~cm$^{-3}$.  A further concern is that the optically thin
 approximation for radiative cooling breaks down, which occurs before
-$n_H \sim 10^{16} - 10^{17}$~cm$^{-3}$.  Beyond this point,
+$n_{\rm H} \sim 10^{16} - 10^{17}$~cm$^{-3}$.  Beyond this point,
 modifications the cooling function that take into account the
@@
 \citet{2004MNRAS.348.1019R}, and was put into \enzo\ for the work
-published in \citep{2009Sci...325..601T,2009PhDT.........5T}.  Even with these
-modifications, a completely correct description of the cooling of this
-gas will require some form of radiation transport, which will greatly
-increase the cost of the simulations.
+published in \citep{2009Sci...325..601T,2009PhDT.........5T}.  Even
+with these modifications, a completely correct description of the
+cooling of this gas will require some form of radiation transport,
+which will greatly increase the cost of the simulations.
 
@@
 \end{table}
+
+%%% Local Variables: 
+%%% mode: latex
+%%% TeX-master: "ms"
+%%% End:

### 2013-03-19 numerical-cooling.tex

@@
 
-\enzo has multiple methods for computing the energy change from
+\enzo\ has multiple methods for computing the energy change from
 radiative cooling and heating.  All of them assume the gas is totally
@@
 cooling of metal-free gas is solved by directly computing the cooling
-and heating rates from the following individual processes for atomic
-H and He: collisional excitation and ionization, recombination,
+and heating rates from the following individual processes for atomic H
+and He: collisional excitation and ionization, recombination,
 free-free emission, Compton scattering off of the cosmic microwave
@@
 necessary in regimes where rapid cooling or change in the ionization
-state occur.  In
-addition to the subcycle timestepping contraints mentioned in
-\S\ref{sec.ov.chem}, the subcycle timestep is also not permitted to
-exceed 10\% of the cooling time, $e/\dot{e}$.  A metagalactic
-background affects the gas through both photo-heating and
+state occur.  In addition to the subcycle timestepping contraints
+mentioned in \S\ref{sec.ov.chem}, the subcycle timestep is also not
+permitted to exceed 10\% of the cooling time, $e/\dot{e}$.  A
+metagalactic background affects the gas through both photo-heating and
 photo-ionization.  These are treated by including redshift-dependent
 photo-ionization and photo-heating rate terms in the chemistry and
-cooling equations for HI, HeI, and HeII.  The black curves in Figure
-\ref{fig.cooling_rate} show cooling rates for a metal-free gas with
-number density $n_{H} = 10^{-4}$ cm$^{-3}$ both with and without a
-radiation background.  More detail on the specific UV backgrounds
-present in \enzo is presented below in
+cooling equations for \ion{H}{1}, \ion{He}{1}, and \ion{He}{2}.  The
+black curves in Figure \ref{fig.cooling_rate} show cooling rates for a
+metal-free gas with number density $n_{\rm H} = 10^{-4}$ cm$^{-3}$ both
+with and without a radiation background.  More detail on the specific
+UV backgrounds present in \enzo is presented below in
 \S\ref{sec.num.rad-homogeneous}.
@@
 Figure \ref{fig.cooling_rate} show cooling rates calculated with this
-method for a gas with number density $n_{H} = 10^{-4}$ cm$^{-3}$ and
+method for a gas with number density $n_{\rm H} = 10^{-4}$ cm$^{-3}$ and
 metallicity of 0.5 Z$_{\odot}$.

### 2013-03-19 numerical-particle.tex
Adding ray tracing method section (a little more than a page).  Fixed a lot of missing references and citations.

@@
 
-The codes needs to be able to model collisionless 
-One way would be to model the dark matter (or other collisionless particle-like objects, such as stars) as a second fluid in addition to the baryonic fluid and solve the collisionless Boltzmann equation, which follows the evolution of the fluid density in six-dimensional phase space.
-However, this is computationally prohibitive owing to the large dimensionality of the problem, making this approach  unappealing for the cosmological AMR code.
+% The code needs to be able to model collisionless matter.  One way
+% would be to model the dark matter (or other collisionless
+% particle-like objects, such as stars) as a second fluid in addition to
+% the baryonic fluid and solve the collisionless Boltzmann equation,
+% which follows the evolution of the fluid density in six-dimensional
+% phase space.  However, this is computationally prohibitive owing to
+% the large dimensionality of the problem, making this approach
+% unappealing for the cosmological AMR code.
 %GB: removed the above (good discussion but not really appropriate in this paper I think, or maybe at the beginning)
@@
 The gravitational potential is computed by solving the elliptic 
-Poisson's equation (Eqtn.~\ref{enzopoisson}).
+Poisson's equation (Eq.~\ref{eq:potential}).
 
 \begin{eqnarray}
+\label{eqn.driftkick}
 x^{n+1/2} & = & x^n + (\Delta t/2) v^{n} \nonumber \\

### 2013-03-19 numerical-starformation.tex

@@
 \begin{equation}
-t_{cool} < t_{dyn} \equiv \sqrt{3 \pi / 32G \rho_{tot}},
+t_{\rm cool} < t_{\rm dyn} \equiv \sqrt{3 \pi / 32G \rho_{\rm tot}},
 \end{equation}
@@
 overdensity threshold, $\vec{v}$ is the velocity of the gas in the
-cell, $\rho_{d}$ is the dark matter density, $\rho_{total} = \rho_{b}
-+ \rho_{d}$, $m_{b}$ and $m_{j}$ are the baryonic mass in the cell and
-jeans mass of the cell, and C is the isothermal sound speed in the
-cell.  If all of these criteria are met, the mass of a star particle
-is calculated as \(m_{*} = m_{b} \frac{ \Delta t}{ t_{dyn} } f_{*eff}
-\), where $f_{*eff}$ is the star formation efficiency parameter.
+cell, $\rho_{d}$ is the dark matter density, $\rho_{\rm tot} =
+\rho_{b} + \rho_{d}$, $m_{b}$ and $m_{j}$ are the baryonic mass in the
+cell and jeans mass of the cell, and C is the isothermal sound speed
+in the cell.  If all of these criteria are met, the mass of a star
+particle is calculated as \(m_{*} = m_{b} \frac{ \Delta t}{ t_{\rm
+    dyn} } f_{\rm *eff} \), where $f_{\rm *eff}$ is the star formation
+efficiency parameter.
 
-If $m_{*}$ is greater than a minimum star mass $m_{*min}$, a particle
+If $m_{*}$ is greater than a minimum star mass $m_{\rm *min}$, a particle
 is created and given several attributes: mass, a unique index number,
-the time of formation $t_{form}$, the local dynamical free-fall time
-$t_{dyn}$ and the metallicity fraction of the baryon gas in the cell
-$f_{Zb}$.  There is a user-defined minimum dynamical time
-$T_{dyn,min}$ which is observationally motivated and affects the
+the time of formation $t_{\rm form}$, the local dynamical free-fall time
+$t_{\rm dyn}$ and the metallicity fraction of the baryon gas in the cell
+$f_{\rm Zb}$.  There is a user-defined minimum dynamical time
+$T_{\rm dyn,min}$ which is observationally motivated and affects the
 feedback rates (see below).  The particle is placed in the center of
@@
 \begin{eqnarray}  
-\Delta m_{sf} =  m_{*} \left[ \left(1 + \frac{t-t_{form}}{t_{dyn}}  \right)
+\Delta m_{\rm sf} =  m_{*} \left[ \left(1 + \frac{t-t_{\rm
+        form}}{t_{\rm dyn}}  \right)
 %\nonumber \times \\ % uncomment for 2-column
-		   \exp{\left( \frac{-(t-t_{form})}{t_{dyn}} \right) }
+		   \exp{\left( \frac{-(t-t_{\rm form})}{t_{\rm dyn}} \right) }
 %                    \nonumber \\ 
-	       - \left( 1 + \frac{t+ \Delta t-t_{form}}{t_{dyn}} \right) 
+	       - \left( 1 + \frac{t+ \Delta t-t_{\rm form}}{t_{\rm dyn}} \right) 
 %	       \nonumber \\ 
-                    \exp{ \left( \frac{-(t+\Delta t-t_{form})}{t_{dyn}} \right)  }
+                    \exp{ \left( \frac{-(t+\Delta t-t_{\rm
+                            form})}{t_{\rm dyn}} \right)  }
                    \right]
@@
 \int_{t}^{t+Dt} \frac{dM}{dt} dt = \int_{t}^{t+Dt} m_{*} 
-\frac{dt}{t_{tyn}} \left( \frac{t-t_{form}}{t_{dyn}} \right) 
+\frac{dt}{t_{\rm dyn}} \left( \frac{t-t_{\rm form}}{t_{\rm dyn}} \right) 
 %\nonumber \times \\ 
-\exp{ \left( - ~ \frac{t-t_{form}}{t_{dyn}} \right) }
+\exp{ \left( - ~ \frac{t-t_{\rm form}}{t_{\rm dyn}} \right) }
 \end{eqnarray}
@@
 winds.  Since massive stars have very short lifetimes, we assume that
-there is an immediate feedback of some fraction $f_{SN}$ of the rest
+there is an immediate feedback of some fraction $f_{\rm SN}$ of the rest
 energy from the stars created that timestep into the baryon gas, such
-that $E_{add} = f_{SN} \Delta m_{sf} c^2$, where c is the speed of
+that $E_{\rm add} = f_{\rm SN} \Delta m_{\rm sf} c^2$, where c is the speed of
 light.  In addition, a fraction $f_{Z*}$ of the metal from the star
@@
 There are six user-defined parameters in this algorithm: three deal
-with the star formation ($\eta$, $m_{*min}$ and $T_{dyn,min}$), and
-three deal with feedback ($f_{SN}$, $f_{Z*}$ and $f_{m*}$).  Some of
-these parameters are completely free, while others can be guided by
-observation or theory.  For example, the supernova feedback parameter,
-$f_{SN}$, can be constrained assuming that, for every $200 M_\odot$ of
-stars created, one supernova occurs, and this event feeds back
-approximately $10^{51}$ ergs of thermal energy, giving:
+with the star formation ($\eta$, $m_{\rm *min}$ and $T_{\rm
+  dyn,min}$), and three deal with feedback ($f_{\rm SN}$, $f_{Z*}$ and
+$f_{m*}$).  Some of these parameters are completely free, while others
+can be guided by observation or theory.  For example, the supernova
+feedback parameter, $f_{\rm SN}$, can be constrained assuming that, for
+every $200 M_\odot$ of stars created, one supernova occurs, and this
+event feeds back approximately $10^{51}$ ergs of thermal energy,
+giving:
 
 \begin{equation}
-f_{SN} = \frac{10^{51} ergs}{200~M_\odot c^2} \simeq 3 \times 10^{-6}
+f_{\rm SN} = \frac{10^{51}\, \mathrm{erg}}{200~M_\odot\, c^2} \simeq 3 \times 10^{-6}
 \end{equation}
@@
 that $f_{Z*} = 0.02$ is an appropriate number.  The minimum dynamical
-time is set to be $T_{dyn,min} = 10^7$ ~years to agree with timescales
+time is set to be $T_{\rm dyn,min} = 10^7$ ~years to agree with timescales
 seen in nearby OB associations.
@@
 The other parameters, such as the overdensity threshold $\eta$,
-minimum star mass $m_{*min}$, and mass ejection fraction $f_{m*}$ are
+minimum star mass $m_{\rm *min}$, and mass ejection fraction $f_{m*}$ are
 not well constrained either theoretically or observationally.  Indeed,
-$m_{*min}$ is a purely numerical parameter designed to keep the code
+$m_{\rm *min}$ is a purely numerical parameter designed to keep the code
 from producing too many star particles, and thus has no observational
@@
 the global Schmidt law of star formation~\citep{2003ApJ...590L...1K,
-1959ApJ...129..243S}.  This algorithm is deliberately minimal, and is
-explicitly geared towards modeling star formation in a
+  1959ApJ...129..243S}.  This algorithm is deliberately minimal, and
+is explicitly geared towards modeling star formation in a
 phenomenological way on kiloparsec scales.  Stars are assume to form
 with a characteristic gas timescale $\tau_*$ such that $\dot{\rho}_* =
-\rho_{gas}/\tau_*$ where $\rho_{gas}$ and $\rho_*$ are the baryon gas
-and stellar densities, respectively.  This ``constant efficiency''
-model on the scale of star formation regions is well motivated
-observationally \citep{1996AJ....112.1903Y,2002ApJ...569..157W}.  Star
-formation is only allowed to take place in very dense regions with
-$\rho_{gas} \geq \rho_{SF}$, where $\rho_{SF}$ is a constant proper
-(as opposed to comoving) density threshold.  No other criteria are
-imposed.  Typical choices for $\tau_*$ and $\rho_{SF}$ are $\tau_* =
-4$ Gyr and $\rho_{SF} = 1.64~$M$_\odot$~pc$^{-3}$~$(n_H \sim
-50$~cm$^{-3})$.  The adopted timescale is derived from the
+\rho_{\rm gas}/\tau_*$ where $\rho_{\rm gas}$ and $\rho_*$ are the
+baryon gas and stellar densities, respectively.  This ``constant
+efficiency'' model on the scale of star formation regions is well
+motivated observationally
+\citep{1996AJ....112.1903Y,2002ApJ...569..157W}.  Star formation is
+only allowed to take place in very dense regions with $\rho_{\rm gas}
+\geq \rho_{\rm SF}$, where $\rho_{\rm SF}$ is a constant proper (as
+opposed to comoving) density threshold.  No other criteria are
+imposed.  Typical choices for $\tau_*$ and $\rho_{\rm SF}$ are $\tau_*
+= 4$ Gyr and $\rho_{\rm SF} = 1.64~$M$_\odot$~pc$^{-3}$~$(n_{\rm H}
+\sim 50$~cm$^{-3})$.  The adopted timescale is derived from the
 observationally-determined normalization of the Schmidt law, and the
@@
 assumed to occur once every global time step $\Delta t_0 \leq 10^7$.
-In cells where star formation is determined to occur (i.e. $\rho_{gas}
-\geq \rho_{SF}$), star particles with a mass of $m_* = \dot{\rho}_*
-V_{cell} \Delta t_0$ (where $V_{cell}$ is the volume of the mesh cell)
-are assumed to form instantaneously in a manner similar to that
-described in Section~\ref{sec:starform_cen}.  The \enzo\
-implementation of this algorithm is similar, except that instead of
-forming stars only at the root grid time step, we allow stars to form
-at the time step of the highest level of resolution at any particular
-point in space.  As can be seen from the equation for $m_*$ above,
-this can result in very small stellar masses.  To avoid memory and
-processor time issues related to having very large numbers of star
-particles we impose a threshold mass $M_{*,min}$ such that a star
-particle only forms if $m_* \geq M_{*,min}$.  An appropriate choice of
-$M_{*,min}$ does not significantly change the star overall star
-formation history of a simulation, though it may delay the onset of
-star formation in a given cell relative to a simulation without a
-particle mass threshold.
+In cells where star formation is determined to occur (i.e. $\rho_{\rm
+  gas} \geq \rho_{\rm SF}$), star particles with a mass of $m_* =
+\dot{\rho}_* V_{\rm cell} \Delta t_0$ (where $V_{\rm cell}$ is the
+volume of the mesh cell) are assumed to form instantaneously in a
+manner similar to that described in Section~\ref{sec:starform_cen}.
+The \enzo\ implementation of this algorithm is similar, except that
+instead of forming stars only at the root grid time step, we allow
+stars to form at the time step of the highest level of resolution at
+any particular point in space.  As can be seen from the equation for
+$m_*$ above, this can result in very small stellar masses.  To avoid
+memory and processor time issues related to having very large numbers
+of star particles we impose a threshold mass $M_{\rm *,min}$ such that
+a star particle only forms if $m_* \geq M_{\rm *,min}$.  An
+appropriate choice of $M_{\rm *,min}$ does not significantly change
+the star overall star formation history of a simulation, though it may
+delay the onset of star formation in a given cell relative to a
+simulation without a particle mass threshold.
 
@@
 $100~M_\odot$ \citep{1979ApJS...41..513M}.  All stars in this IMF with
-$M_* > 8~M_\odot$ deposit $2 \times 10^{51}$ ergs of thermal energy
+$M_* > 8~M_\odot$ deposit $2 \times 10^{51}$ erg of thermal energy
 and a mass $f_z M_*$ of metals into the cell in which they form
-without delay, with $f_z \equiv min(0.2, 0.01~M_*-0.06)$
+without delay, with $f_z \equiv \min(0.2, 0.01~M_*-0.06)$
 (i.e. instantaneous deposition of metals).  The definition of $f_z$ is
@@
 % Specific example: H2-regulated
-\subsubsection{H2-regulated method}
+\subsubsection{\HH-regulated method}
 \label{sec:starform_H2reg}
@@
 \begin{equation} 
-m_p = \epsilon \rho_{gas} (\Delta x_{m})^3 \frac{\Delta t_0}{t_*}
+m_p = \epsilon \rho_{\rm gas} (\Delta x_{m})^3 \frac{\Delta t_0}{t_*}
 \end{equation}
@@
 % Specific example: Pop III
-\subsubsection{Population III}
+\subsubsection{Population III star formation}
 \label{sec:starform_pop3}
@@
 where a given halo may only form one star of primordial composition.
-To accommodate this, \citet{2007ApJ...659L..87A,2008ApJ...685...40W, 2012MNRAS.427..311W} have created an
-algorithm within the \enzo\ star formation framework that forms
-Population III stars directly.  Using criteria similar
-to~\ref{CO1992}, a star particle forms when a cell has all of the
+To accommodate this, Enzo contains a star formation algorithm that
+forms Population III stars directly \citep{2007ApJ...659L..87A,
+  2008ApJ...685...40W, 2012MNRAS.427..311W}.  Using criteria similar
+to~\citet{CO1992}, a star particle forms when a cell has all of the
 following criteria:
@@
   roughly $10^3$ g cm$^{-3}$ at $z=10$),
-\item a converging gas flow (\nabla \cdot \vec{v} < 0), and
-\item a molecular hydrogen fraction f$_{H2} > 5 \times 10^{-4}$,
+
+\item a converging gas flow ($\nabla \cdot \vec{v} < 0$), and
+
+\item a molecular hydrogen fraction f$_{\rm H2} > 5 \times 10^{-4}$,
 \end{enumerate}
@@
 \begin{equation}
-f(log M) dM = M^{-1.3} \exp \Big[ -\Big( \frac{M_{char}}{M}  \Big)^{1.6} \Big]
+f(\log M) dM = M^{\alpha} \exp \Big[ -\Big( \frac{M_{\rm char}}{M}
+\Big)^{\beta} \Big]
 \end{equation}
@@
 3.2.1 of \citet{2012MNRAS.427..311W}.
+
+%%% Local Variables: 
+%%% mode: latex
+%%% TeX-master: "ms"
+%%% End:

### 2013-03-19 conclusions.tex
adding a bunch of new sections and subsections.
Tossed some ideas into the conclusions section.  Made a couple of small tweaks in the test problem section.

@@
 
-\begin{enumerate}
-\item near-future development plans
-\item far-future development plans
-\item clear areas of improvement
-\end{enumerate}
+In this paper, we have presented the algorithms underlying \enzo, an
+open-source adaptive mesh refinement code designed for
+self-gravitating compressible fluid dynamics, including the effects of
+magnetic fields, radiation transport, and a variety of microphysical
+and subgrid processes.  In addition, we have described the \enzo\ code
+development process, have shown the outputs of some representative
+test problems, and have provided information about its performance and
+scaling on current-generation (circa 2013) supercomputing platforms.
+The \enzo\ code, its test suite, and all of the scripts used to
+generate plots and figures for this paper are open-source and are
+available at the \enzo\ website, http://enzo-project.org.
+Furthermore, the yt toolkit, which is designed to analyze \enzo\ data
+(as well as data from a wide variety of other simulation tools), can
+be found at its website, http://yt-project.org.  Both of these codes
+have active user and developer communities, extensive documentation
+and user suppport, and strong mechanisms for
+users to contribute their changes and fixes to the codebase.
+
+The developers of the \enzo\ code are currently working on several
+projects that will extend the functionality, scalability, or overall
+performance of the code in the near future.  Work is
+underway to create a hybrid-parallel version of \enzo, combining MPI
+for communication between nodes of a supercomputer and OpenMP for
+thread-based parallelism within a node.  This will reduce on-node
+memory usage and will in principle help with scaling.  \red{other stuff here!}.
+This work is currently underway, and will be released in the stable
+branch of \enzo\ in the near
+future.
+
+In the longer term, we plan to restructure \enzo's treatment of
+particles to accommodate a wider range of ``active'' particles that
+can easily interact with each other and with grids.  We also plan to
+use the HYPRE library to create a new gravity solver for the code,
+which will improve its scalability and correctness.  \red{other stuff
+  here!}
+
+\red{Do we want a separate paragraph for areas of improvement outside
+  of short-term and longer-term?}

### 2013-03-20 introduction.tex
more changes to intro, overview

@@
 At the current time, the most common is a gridless, fully Lagrangian method call Smoothed Particle Hydrodynamcs \citep[SPH;][]{Lucy77, SPH} which has been very successful, particularly for those situations dominated by gravity.
-Its development, however, is still at a relatively early stage when compared to the effort put into Eulerian, grid-based hydrodynamic schemes \citep[e.g.,][]{, laney-1998, toro-1997, Woodward84}.
+Its development, however, is still at a relatively early stage when compared to the effort put into Eulerian, grid-based hydrodynamic schemes \citep[e.g.,][]{laney-1998, toro-1997, Woodward84}.
 

### 2013-03-20 numerical-radiation-fld.tex
more changes to intro, overview

--- 2013-01-23 numerical-radiation-fld.tex
@@
 transfer, with couplings to both the gas energy and chemical number
-densities, 
-\begin{eqnarray}
-  \label{eq:fld_radiation}
-  \partial_t E_r + \frac1a \nabla\cdot\left(E_r \vec{v}_b\right) &=& 
-    \nabla\cdot\left(D\nabla E_r\right) -
-    \frac{\dot{a}}{a} E_r - c\kappa E_r + \eta, \\
-  \label{eq:fld_heating}
-  \partial_t e_c &=& -\frac{2\dot{a}}{a} e_c + G - \Lambda, \\
-  \label{eq:fld_chemistry}
-  \partial_t {\mathrm n}_i + 
-    \frac1a \nabla\cdot\left({\mathrm n}_i \vec{v}_b\right) &=& 
-    \alpha_{i,j} {\mathrm n}_e {\mathrm n}_j - {\mathrm n}_i
-    \Gamma_i^{ph}, \quad i=1,\ldots,N_s,
-\end{eqnarray}
-where $E_r$ is a grey radiation energy density, $e_c$ is the
-internal energy correction due to photo-heating and chemical cooling,
-and ${\mathrm n}_i$ correspond to the chemical number densities of
-HI, HII, HeI, HeII and HeIII, and ${\mathrm n}_e$ is the electron
-number density.  Here, we define the grey radiation energy density
-through first assuming a fixed frequency spectrum, i.e.
-$E_{\nu}(\nu,\vec{x},t) = \tilde{E}_r(\vec{x},t) \chi(\nu)$, and then
-defining the integrated quantity
-\begin{equation}
-\label{eq:grey_radiation_energy}
-   E_r(\vec{x},t) \equiv \int_{\nu_0}^{\infty}
-   E_{\nu}(\nu,\vec{x},t)\,\mathrm d\nu \  = \ 
-   \tilde{E}_r(\vec{x},t) \int_{\nu_0}^{\infty} \chi(\nu)\,\mathrm d\nu.
-\end{equation}
-In addition, $D$ in (\ref{eq:fld_radiation}) is the {\em Larsen
-square-root flux-limiter} (see \cite{Morel2000}), $\kappa$ is the total
-opacity, $\eta$ is the field of radiation sources, $G$ provides the
-radiation induced photo-heating, $\Lambda$ provides the chemical
-cooling, $\alpha_{i,j}$ are the reaction rate coefficients defining
-the interactions between species ${\mathrm n}_i$ and ${\mathrm n}_j$,
-and $\Gamma_i^{ph}$ are the radiation induced photo-ionization rates.
+densities.
+%\begin{eqnarray}
+%  \label{eq:fld_radiation}
+%  \partial_t E_r + \frac1a \nabla\cdot\left(E_r \vec{v}_b\right) &=& 
+%    \nabla\cdot\left(D\nabla E_r\right) -
+%    \frac{\dot{a}}{a} E_r - c\kappa E_r + \eta, \\
+%  \label{eq:fld_heating}
+%  \partial_t e_c &=& -\frac{2\dot{a}}{a} e_c + G - \Lambda, \\
+%  \label{eq:fld_chemistry}
+%  \partial_t {\mathrm n}_i + 
+%    \frac1a \nabla\cdot\left({\mathrm n}_i \vec{v}_b\right) &=& 
+%    \alpha_{i,j} {\mathrm n}_e {\mathrm n}_j - {\mathrm n}_i
+%    \Gamma_i^{ph}, \quad i=1,\ldots,N_s,
+%\end{eqnarray}
+%where $E_r$ is a grey radiation energy density, $e_c$ is the
+%internal energy correction due to photo-heating and chemical cooling,
+%and ${\mathrm n}_i$ correspond to the chemical number densities of
+%HI, HII, HeI, HeII and HeIII, and ${\mathrm n}_e$ is the electron
+%number density.  Here, we define the grey radiation energy density
+%through first assuming a fixed frequency spectrum, i.e.
+%$E_{\nu}(\nu,\vec{x},t) = \tilde{E}_r(\vec{x},t) \chi(\nu)$, and then
+%defining the integrated quantity
+%\begin{equation}
+%\label{eq:grey_radiation_energy}
+%   E_r(\vec{x},t) \equiv \int_{\nu_0}^{\infty}
+%   E_{\nu}(\nu,\vec{x},t)\,\mathrm d\nu \  = \ 
+%   \tilde{E}_r(\vec{x},t) \int_{\nu_0}^{\infty} \chi(\nu)\,\mathrm d\nu.
+%\end{equation}
+%In addition, $D$ in (\ref{eq:fld_radiation}) is the {\em Larsen
+%square-root flux-limiter} (see \cite{Morel2000}), $\kappa$ is the total
+%opacity, $\eta$ is the field of radiation sources, $G$ provides the
+%radiation induced photo-heating, $\Lambda$ provides the chemical
+%cooling, $\alpha_{i,j}$ are the reaction rate coefficients defining
+%the interactions between species ${\mathrm n}_i$ and ${\mathrm n}_j$,
+%and $\Gamma_i^{ph}$ are the radiation induced photo-ionization rates.
 
 The system of equations
-(\ref{eq:fld_radiation}-\ref{eq:fld_chemistry}) is solved
+(\ref{eq:fld_radiation}-\ref{eq:fld_heating}) along with the chemical network 
+(eq.~\ref{eq:species_evolution} is solved
 independently of Enzo's hydrodynamics, gravity and dark-matter solvers
@@
   \label{eq:fld_heating_qss}
-  \partial_t e_c &=& -\frac{2\dot{a}}{a} e_c +
-    G\left(\bar{E}_r,\bar{\mathrm n}_i\right) - 
+  \frac{\partial e_c}{\partial t} &=& -\frac{2\dot{a}}{a} e_c +
+    \Gamma\left(\bar{E}_r,\bar{\mathrm n}_i\right) - 
     \Lambda\left(\bar{E}_r,\bar{\mathrm n}_i\right), \\
   \label{eq:fld_chemistry_qss}
-  \partial_t {\mathrm n}_i &=& \alpha_{i,j}\left(\bar{e}_c\right)
+  \frac{\partial {\mathrm n}_i}{\partial t} &=& k_{i,j}\left(\bar{e}_c\right)
     {\mathrm n}_e \bar{\mathrm n}_j - {\mathrm n}_i 

### 2013-03-20 numerical-radiation-raytracing.tex
Adding description of photo-evaporation test and some minor changes to the Moray section.
more changes to intro, overview

@@
 hydrodynamical response to radiation sources further adds to the
-difficulty.  The radiative transfer equation in comoving coordinates
-\citep{Gnedin97} reads
-%
-\begin{equation}
-  \label{eq:rteqn}
-  \frac{1}{c} \; \frac{\partial I_\nu}{\partial t} + 
-  \frac{\hat{n} \cdot \nabla I_\nu}{\bar{a}} -
-  \frac{H}{c} \; \left( \nu \frac{\partial I_\nu}{\partial \nu} -
-  3 I_\nu \right) = -\kappa_\nu I_\nu + j_\nu .
-\end{equation}
-%
-Here $I_\nu \equiv I(\nu, \mathbf{x}, \Omega, t)$ is the radiation
-specific intensity in units of energy per time $t$ per solid angle per
-unit area per frequency $\nu$.  $H = \dot{a}/a$ is the Hubble
-constant, where $a$ is the scale factor.  $\bar{a} = a/a_{em}$ is the
-ratio of scale factors at the current time and time of emission.  The
-second term represents the propagation of radiation, where the factor
-$1/\bar{a}$ accounts for cosmic expansion.  The third term describes
-both the cosmological redshift and dilution of radiation.  On the
-right hand side, the first term considers the absorption coefficient
-$\kappa_\nu \equiv \kappa_\nu(\mathbf{x},\nu,t)$, and the second term
-$j_\nu \equiv j_\nu(\mathbf{x},\nu,t)$ is the emission coefficient
-that includes any point sources of radiation or diffuse radiation.
+difficulty.  Here we briefly describe the ray tracing implementation \moray,
+which is presented in full detail in \citet{Wise11_Moray} with seven
+code tests and six applications.
 
+%The radiative transfer equation in comoving coordinates
+%\citep{Gnedin97} reads
+%%
+%\begin{equation}
+%  \label{eq:rteqn}
+%  \frac{1}{c} \; \frac{\partial I_\nu}{\partial t} + 
+%  \frac{\hat{n} \cdot \nabla I_\nu}{\bar{a}} -
+%  \frac{H}{c} \; \left( \nu \frac{\partial I_\nu}{\partial \nu} -
+%  3 I_\nu \right) = -\kappa_\nu I_\nu + j_\nu .
+%\end{equation}
+%%
+%Here $I_\nu \equiv I(\nu, \mathbf{x}, \Omega, t)$ is the radiation
+%specific intensity in units of energy per time $t$ per solid angle per
+%unit area per frequency $\nu$.  $H = \dot{a}/a$ is the Hubble
+%constant, where $a$ is the scale factor.  $\bar{a} = a/a_{em}$ is the
+%ratio of scale factors at the current time and time of emission.  The
+%second term represents the propagation of radiation, where the factor
+%$1/\bar{a}$ accounts for cosmic expansion.  The third term describes
+%both the cosmological redshift and dilution of radiation.  On the
+%right hand side, the first term considers the absorption coefficient
+%$\kappa_\nu \equiv \kappa_\nu(\mathbf{x},\nu,t)$, and the second term
+%$j_\nu \equiv j_\nu(\mathbf{x},\nu,t)$ is the emission coefficient
+%that includes any point sources of radiation or diffuse radiation.
+%GB: this was moved to the equations section.
+
+We solve the radiative transfer equation in comoving coordinates (given by eq.~\ref{eq:rteqn}).
 We can make some appropriate approximations to reduce the complexity
-of Equation (\ref{eq:rteqn}) in order to include radiation transport
+of this equation in order to include radiation transport
 in numerical calculations.  Typically timesteps in dynamic

### 2013-03-20 overview.tex
Many minor changes.  No more undefined references!  Added appendix file.
more changes to intro, overview

--- 2013-03-18 overview.tex
@@
 
-\subsubsection{Magnetohydrodynamics and Gravity}
+%\subsubsection{Magnetohydrodynamics and Gravity}
 
@@
      - \frac{\dot{a}}{a} \left( 2E - \frac{B^2}{2a} \right)  - \frac{1}{a} \vecv \cdot \nabla \phi 
+     - \Lambda + \Gamma + \frac{1}{a^2} \nabla \cdot \fcond
      %
@@
 The fourth equation is the magnetic induction equation.
-Terms representing radiative cooling ($\Lambda$) and heating ($\Gamma$) enter on the right-hand side of the energy equation (\ref{eq:total_energy}).
+Terms representing radiative cooling ($\Lambda$) and heating ($\Gamma$) enter on the right-hand side of the energy equation (\ref{eq:total_energy}), as well as the flux due to thermal heat conduction ($\fcond$).
 
@@
 \begin{equation}
-E =  \rho e + \frac{\rho v^2}{2}  + \frac{B^2}{2a}
+E =  e + \frac{\rho v^2}{2}  + \frac{B^2}{2a}
         \label{eq:total_energy_def}.
 \end{equation}
-where $e$ is the comoving specific thermal energy, and the total isotropic pressure $p^*$ is given by
+where $e$ is the comoving thermal energy density, and the total isotropic pressure $p^*$ is given by
 \begin{equation}
@@
 %
-\nabla^2 \phi & = & 4 \pi G \rho_{total}
+\nabla^2 \phi & = & 4 \pi G \rho_{\rm total}
 %                   + 3 p_{total}/c^2) - \Lambda. 
@@
 
-\subsubsection{Comoving Coordinates}
+%\subsubsection{Comoving Coordinates}
 
@@
 
-\subsubsection{Particles and Chemistry}
+%\subsubsection{Particles}
 
@@
 
-In addition, we solve mass conservation equations for any chemical species.  For any species $i$, these equations have the form:
+%\subsubsection{Chemistry}
 
-\begin{eqnarray}
-\frac{\partial \rho_i}{\partial t} 
-          + \frac{1}{a} \vecvb \cdot \nabla \rho_i 
-     & = &  - \frac{1}{a} \rho_i \nabla \cdot \vecvb 
-        + \sum_j \sum_l k_{jl}(T) \rho_j \rho_l 
-% \nonumber \\    & &    %uncomment for 2-column
-      + \sum_j I_j \rho_j 
-        \label{eq:mass_species}
-\end{eqnarray}
+In addition, we solve mass conservation equations for any chemical species, including a variety of chemical networks.
+For any species $i$ with comoving number density $n_i$, these equations have the form:
+\begin{equation}
+\frac{\partial n_i}{\partial t} 
+          + \frac{1}{a} \nabla \cdot (n_i \vecv) = 
+        k_{ij}(T) n_i n_j 
+      + \Gamma^{ph}_j n_j 
+        \label{eq:species_evolution}
+\end{equation}
 where $k_{ij}$ are the rate coefficients for the two-body reactions and are usually functions of only the gas species (we will specifically note the cases where we either include three-body reactions or have density-dependent rates).
-The $I_j$ are destruction/creation rates due to photoionizations and/or photodissociations.
+The $\Gamma^{ph}_j$ are destruction/creation rates due to photoionizations and/or photodissociations.
 
@@
 
-\subsubsection{Radiation transport: ray-tracing}
+%\subsubsection{Radiation transport: ray-tracing}
 
-\subsubsection{Radiation transport: moment method}
+The code can include either a homogeneous radiation background (with possible local self-shielding corrections), or evolve an inhomogeneous radiation field either by directly solving the radiative transfer equation along rays, or by solving a set of moment equations derived from the radiative transfer equation.
 
-\subsubsection{Thermal conduction}
+The radiative transfer equation in comoving coordinates 
+\citep[e.g.,][]{Gnedin97} reads
+%
+\begin{equation}
+  \label{eq:rteqn}
+  \frac{1}{c} \; \frac{\partial I_\nu}{\partial t} + 
+  \frac{\hat{n} \cdot \nabla I_\nu}{\bar{a}} -
+  \frac{H}{c} \; \left( \nu \frac{\partial I_\nu}{\partial \nu} -
+  3 I_\nu \right) = -\kappa_\nu I_\nu + j_\nu .
+\end{equation}
+%
+Here $I_\nu \equiv I(\nu, \mathbf{x}, \Omega, t)$ is the radiation
+specific intensity in units of energy per time $t$ per solid angle per
+unit area per frequency $\nu$.  $H = \dot{a}/a$ is the Hubble
+constant, where $a$ is the scale factor.  $\bar{a} = a/a_{em}$ is the
+ratio of scale factors at the current time and time of emission.  The
+second term represents the propagation of radiation, where the factor
+$1/\bar{a}$ accounts for cosmic expansion.  The third term describes
+both the cosmological redshift and dilution of radiation.  On the
+right hand side, the first term considers the absorption coefficient
+$\kappa_\nu \equiv \kappa_\nu(\mathbf{x},\nu,t)$, and the second term
+$j_\nu \equiv j_\nu(\mathbf{x},\nu,t)$ is the emission coefficient
+that includes any point sources of radiation or diffuse radiation.
+
+%\subsubsection{Radiation transport: moment method}
+
+In the other limit, we solve an equation derived from integrating
+the equation of radiative transfer over angles.  In particular, we use a
+flux-limited diffusion approximation for cosmological radiative
+transfer, with couplings to both the gas energy and chemical number
+densities. The equations solved are:
+\begin{eqnarray}
+  \label{eq:fld_radiation}
+  \frac{\partial E_r}{\partial t} + \frac1a \nabla\cdot\left(E_r \vecv \right) &=& 
+    \nabla\cdot\left(D\nabla E_r\right) -
+    \frac{\dot{a}}{a} E_r - c\kappa E_r + \eta, \\
+  \label{eq:fld_heating}
+  \frac{\partial e_c}{\partial t} &=& -\frac{2\dot{a}}{a} e_c + \Gamma - \Lambda,
+%  \label{eq:fld_chemistry}
+%  \partial_t {\mathrm n}_i + 
+%    \frac1a \nabla\cdot\left({\mathrm n}_i \vec{v}_b\right) &=& 
+%    \alpha_{i,j} {\mathrm n}_e {\mathrm n}_j - {\mathrm n}_i
+%    \Gamma_i^{ph}, \quad i=1,\ldots,N_s,
+\end{eqnarray}
+where $E_r$ is a grey radiation energy density, $e_c$ is the
+internal energy correction due to photo-heating and chemical cooling.
+%and ${\mathrm n}_i$ correspond to the chemical number densities of
+%HI, HII, HeI, HeII and HeIII, and ${\mathrm n}_e$ is the electron
+%number density.  
+Here, we define the grey radiation energy density
+through first assuming a fixed frequency spectrum, i.e.
+$E_{\nu}(\nu,\vec{x},t) = \tilde{E}_r(\vec{x},t) \chi(\nu)$, and then
+defining the integrated quantity
+\begin{equation}
+\label{eq:grey_radiation_energy}
+   E_r(\vec{x},t) \equiv \int_{\nu_0}^{\infty}
+   E_{\nu}(\nu,\vec{x},t)\,\mathrm d\nu \  = \ 
+   \tilde{E}_r(\vec{x},t) \int_{\nu_0}^{\infty} \chi(\nu)\,\mathrm d\nu.
+\end{equation}
+In addition, $D$ in (\ref{eq:fld_radiation}) is the {\em Larsen
+square-root flux-limiter} (see \cite{Morel2000}), $\kappa$ is the total
+opacity, $\eta$ is the field of radiation sources, $\Gamma$ provides the
+radiation induced photo-heating and, $\Lambda$ the chemical
+cooling.
+%, $\alpha_{i,j}$ are the reaction rate coefficients defining
+%the interactions between species ${\mathrm n}_i$ and ${\mathrm n}_j$,
+%and $\Gamma_i^{ph}$ are the radiation induced photo-ionization rates.
+
+%\subsubsection{Thermal conduction}
 
@@
 \begin{equation}
-\mathbb{F} = -\kappa_{sp} \nabla T
+\fcond = -\kappa_{sp} \nabla T
 \end{equation}
@@
 \begin{equation}
-\mathbb{F} = -\kappa_{sp} \hat{b} (\hat{b} \cdot \nabla T)
+\fcond= -\kappa_{sp} \hat{b} (\hat{b} \cdot \nabla T)
 \end{equation}
@@
 
-In this section, we briefly describe the numerical methods which are used to solve the equations outline above.
+In this section, we briefly describe the numerical methods which are used to solve the equations outlined above.
 We sweep through in the same order as will be used in section~\ref{sec.methods} so that there is a one-to-one correspondence between each of the following overviews and the complete description provided in section~\ref{sec.methods}.
@@
 
-\subsubsection{Godunov PPM method (Hydrodynamics)}
+\subsubsection{Godunov PPM method (HD only)}
 
@@
 steps: 
-\begin{enumerate}
- \item Construct monotonic parabolic interpolation of cell average
- data, for each fluid quantity.
- \item Compute interface states by averaging the parabola for over the domain of dependence for
- each interface
- \item Use interface data to solve the Riemann problem.
- \item Difference the interface fluxes to update the cell average quantities.
-\end{enumerate}
+%\begin{enumerate}
+% \item 
+ (i) Construct monotonic parabolic interpolation of cell average data, for each fluid quantity.
+% \item 
+(ii) Compute interface states by averaging the parabola for over the domain of dependence for each interface
+% \item 
+(iii) Use interface data to solve the Riemann problem.
+ %\item 
+ (iv) Difference the interface fluxes to update the cell average quantities.
+%\end{enumerate}
+
 It does an excellent job of capturing strong shocks in at
@@
 
-\subsubsection{Godonov MUSCL with Dedner (HD and MHD)}
+\subsubsection{Godunov MUSCL (HD) with Dender divergence cleaning (MHD)}
 
-\subsubsection{MHDCT}
+\subsubsection{MHD with Constrained Transport (MHD)}
 
-\subsubsection{Second order finite difference method (HD)}
+\subsubsection{Second order finite difference method (HD only)}
 
@@
 
-\subsubsection{Chemistry, radiation and cooling}
+\subsubsection{Chemistry}
+\label{sec.ov.chem}
+
+In addition to this, \enzo\ includes the capability of following up to 12 chemical species using a non equilibrium-solver.
+The species can be turned on in sets, with the simplest model including just H, H$^+$, He, He$^+$, He$^++$, and $e^-$, and more complete models adding first species important for gas-phase molecular hydrogen formation: $H^-$, H$_2$ and H$_2^+$, and then $HD$ formation: $HD$, $D$, $D+$.
+The cooling and heating due to these species is also included.
+The solution of the rate equations is carried out using a time-backwards differencing which insures stability; to maintain accuracy we sub-cycle the rate equations with a shorter timestep such that the electron and neutral fractions do not change by more than 10\%.
+
+\subsubsection{Radiative Cooling and Heating}
 
@@
 
-In addition to this, \enzo\ includes the capability of following up to 12 chemical species using a non equilibrium-solver.
-The species can be turned on in sets, with the simplest model including just H, H$^+$, He, He$^+$, He$^++$, and $e^-$, and more complete models adding first species important for gas-phase molecular hydrogen formation: $H^-$, H$_2$ and H$_2^+$, and then $HD$ formation: $HD$, $D$, $D+$.
-The cooling and heating due to these species is also included.
-The solution of the rate equations is carried out using a time-backwards differencing which insures stability; to maintain accuracy we sub-cycle the rate equations with a shorter timestep such that the electron and neutral fractions do not change by more than 10\%.
+\subsubsection{Homogeneous radiation backgrounds}
+
+\subsubsection{Radiation transport: ray tracing}
+
+\subsubsection{Radiation transport: Flux-limited diffusion}
+
+\subsubsection{Thermal Heat Conduction}
 

### 2013-03-20 overview.tex
Even more changes to overview (getting there...)

@@
 
-We begin by laying out the equations to be solved and then discuss briefly the techniques used to solve each part.
+We begin this section by first writing down the entire set of physical equations which \enzo\ can solve, and then briefly describe the numerical algorithms that we use to solve them.
+This section is intended both to place all of the equations in a single place as well as to provide a brief introduction to the numerics of the code.
+Detailed descriptions of the individual components are then provided in the following section.
 
@@
 
-The usual Eulerian equations of ideal magnetohydrodynamics with gravity in a coordinate systems comoving the cosmological expansion are given by
+The Eulerian equations of ideal magnetohydrodynamics including gravity in a coordinate systems comoving with the cosmological expansion are given by
 
@@
 
-In addition, we solve mass conservation equations for any chemical species, including a variety of chemical networks.
-For any species $i$ with comoving number density $n_i$, these equations have the form:
+In addition, we can solve the mass conservation equations for a set of chemical species and reactions.  For any species $i$ with comoving number density $n_i$, these equations have the form:
 \begin{equation}
@@
 The $\Gamma^{ph}_j$ are destruction/creation rates due to photoionizations and/or photodissociations.
+Currently the species we can follow include H, H$^+$, He, He$^+$, He$^++$, and $e^-$, as well as additional options that add H$^-$, H$_2$ H$_2^+$ and  HD, D, D$^+$.
 
@@
 In this section, we briefly describe the numerical methods which are used to solve the equations outlined above.
-We sweep through in the same order as will be used in section~\ref{sec.methods} so that there is a one-to-one correspondence between each of the following overviews and the complete description provided in section~\ref{sec.methods}.
+We sweep through in the same order as will be used in section~\ref{sec.methods} so that there is a one-to-one correspondence between each of the following overviews and the complete description provided in that section.
 The goal here is to introduce the reader to the basic idea of the methods without drowning in detail.
@@
 
-Two different hydrodynamic methods are implemented in \enzo: the
-piecewise parabolic method (PPM) developed by~\citet{1984JCoPh..54..174C}
-and extended to cosmology by~\citet{1995CoPhC..89..149B},
-and the hydrodynamics method used in \zeus~\citep{Stone92a,Stone92b}.
-We adopt the direct Eulerian PPM method (as opposed to the Lagrange-Remap varient).
+Four different (magneto)-hydrodynamic methods are implemented in \enzo: (i) the
+HD-only piecewise parabolic method (PPM) developed by~\citet{1984JCoPh..54..174C}
+and extended to cosmology by~\citet{1995CoPhC..89..149B};
+(ii) the MUSCL-like Gudonov scheme \citep{MUSCL} with or without magnetic fields
+and Dedner-based divergence cleansing and described in more detail in 
+\citet{WangAbelZhang08} and \citet{WangAbel09};
+(iii) a constrained transport staggered MHD scheme \citet{Collins10}, and
+(iv) the second-order finite-difference hydrodynamics method described 
+in \zeus~\citep{Stone92a,Stone92b}.
 
-PPM is an explicit, higher order-accurate version of
+We begin with the direct-Eulerian PPM implementation.  
+This is an explicit, higher order-accurate version of
 Godunov's method for ideal gas dynamics with third order-accurate piecewise parabolic
@@
 
-It does an excellent job of capturing strong shocks in at
-most two cells.  Multidimensional schemes are built up by directional
+It does an excellent job of capturing strong shocks in a few cells; Multidimensional schemes are built up by directional
 splitting and produce a method that is formally second order-accurate
-in space and time which explicitly conserves mass, linear momentum, and energy.
+in space and time which explicitly conserves mass, linear momentum, and energy.  A variety of
+reconstruction methods and Riemann solvers have been implemented.
 
@@
 
-The advantage of the PPM method is that it conserves energy better and is higher-order accurate.  The disadvantage is that it is generally less robust than the ZEUS method.
+%The advantage of the PPM method is that it conserves energy better and is higher-order accurate.  The disadvantage is that it is generally less robust than the ZEUS method.
 
@@
 
-The second hydrodynamics method implemented is the finite-different algorithm used in the ZEUS, as described in \citet{Stone92a}.
-For this reason, we call this the ZEUS method, although the code is entirely independent from the ZEUS code, and only the hydrodynamical algorithm of \zeus\ is implemented in \enzo -- the MHD and RHD codes are not.
+The last hydrodynamics method we briefly describe is the finite-different algorithm used in the ZEUS, as given in more detai in \citet{Stone92a}.
+For this reason, we call this the ZEUS method, although the code is entirely independent of the ZEUS code, and only the hydrodynamical algorithm of \zeus\ is implemented in \enzo -- the MHD and RHD schemes are not.
 
@@
 
-The collisionless matter is modelled through particles.  Since the
-particles follow the collapse of structure by definition, they are not
-adaptively refined.  Nor are there duplicate sets of particles for
-each level; instead, each particle is associated with the most refined
-level available at its position in space and moved as the hierarchy is
-rebuilt.  Thus, a particle has the same timestep and feels the same
-gravitational force as the grid at that level.
+Collisionless matter (dark matter, stars, etc.) is modeled with particles which interact with the baryons only via gravity.
+These particles are advanced in a single time step using a drift-kick-drift algorithm to provide second-order accuracy even in the presence of changing timesteps.
+Since the particles follow the collapse of structure, they are not adaptively refined.
+Nor are there duplicate sets of particles for each level; instead, each particle is associated with the highest refined
+level available at its position in space, and moved as the hierarchy is rebuilt.  Thus, a particle has the same timestep and feels the same gravitational force as the grid at that level.
 
-Although the particles are fixed in mass once initialized, we are free
-to create them with any set of masses and positions.  For example, in
+Although the particles are fixed in mass once initialized, we can
+create them with any initial set of masses and positions.  For example, in
 many cosmological simulations, a static subgrid is included from the
@@
 
-When solving Poisson's equation, we need to compute the contribution of 
-the particles to the density on the grid.  This is done using a second-order
-accurate Cloud-In-Cell technique \citep{Hockney88}.
-
 \subsubsection{Chemistry}
@@
 The species can be turned on in sets, with the simplest model including just H, H$^+$, He, He$^+$, He$^++$, and $e^-$, and more complete models adding first species important for gas-phase molecular hydrogen formation: $H^-$, H$_2$ and H$_2^+$, and then $HD$ formation: $HD$, $D$, $D+$.
-The cooling and heating due to these species is also included.
+The cooling and heating due to these species is also included (see the next section).
 The solution of the rate equations is carried out using a time-backwards differencing which insures stability; to maintain accuracy we sub-cycle the rate equations with a shorter timestep such that the electron and neutral fractions do not change by more than 10\%.
@@
 
-Enzo may operate in a number of modes with regard to radiative cooling and chemistry.
+Enzo may operate in a number of modes with regard to radiative cooling and heating.
 In the simplest mode, with the multi-species flag turned off (i.e.
 set to zero), the cooling rate is computed from a simple temperature-dependent cooling rate, taken from \citet{SW87}.
+If chemistry is turned on, then the code can include cooling from all species of H and He (including H$_2$) -- this primordial cooling is also treated using a backwards-time differencing scheme with sub-cycling to maintain accuracy.
+Finally, we have also recently added metal cooling based on a set of multi-dimensional lookup tables computed with the \texttt{Cloudy} code \citep{1998PASP..110..761F} as described in \citet{2008MNRAS.385.1443S} and \citet{2011ApJ...731....6S}.
+Finally, we note that the cooling and heating is (almost always) treated in the optically-thin limit, although the code can also follow radiative transfer in a limited set of energy bins (see below).
 
@@
 
+The chemical networks and heating rates can be affected by external radiation fields, and the code includes a number of pre-set radiation backgrounds that are uniform in space but can very in time.
+These are generally based on the redshift-dependent rates given in \citet{1996ApJ...461...20H} and \citet{2012ApJ...746..125H}, but can also include a uniform H$_2$ photo-dissociating background, either constant in time or varying as in \citet{WiseAbel05}.
+
 \subsubsection{Radiation transport: ray tracing}
+
+The code includes a photon-conserving radiative transfer algorithm that is based on an adaptive ray-tracing method that utilizes the HEALPix pixilation of the sphere \citep{AbelWandelt02}.
+Photons are integrated outwards from sources using an adaptive time-stepping scheme that preserves accuracy in ionization fronts even in the optically-thin limit.
+This has been integrated with the chemistry and cooling network to provide ionization and heating rates on a cell-by-cell basis.
+The method is described in more detail in our main methodology section but is fully detailed and tested in \citet{Wise11_Moray}.
 
@@
 
-\subsubsection{Thermal Heat Conduction}
+The other option for radiative transfer is a moment-based method which adds an addition field describing the energy-density of radiation.
+This field is evolved using the flux-limited diffusion method which transitions smoothly between streaming (optically thin) and opaque limits and is coupled to the hydrogen ionization network.
+The resulting set of linear equations is solved using the parallel HYPRE framework.
+Full details for the enzo implementation can be found in \citet{ReynoldsHayesPaschosNorman2009}.
+
+\subsubsection{Heat Conduction}
+
+Heat conduction, both isotropic and anisotropic, can be included using an sub-cycled, operator-split method.
+The heat fluxes are computed with simple second-order accurate finite differences and stability is ensured by restricting the time step, and using flux-limiters.
+This is described in more detail below.
 

### 2013-03-20 numerical-gravity.tex
Formatting police passing through.
updated gravity and particles in section 3

@@
 
-Solving for the accelerations for the cells and particles in a grid due to self-gravity involves three steps: first, computing the total gravitating mass; second, solving for the potential field with the appropriate boundary conditions, and finally, differencing the potential to get the acceleration and, if necessary, interpolating the potential back to the particles.
+Solving for the accelerations of the cells and particles on the grid due to self-gravity involves three steps: (i) computing the total gravitating mass, (ii) solving for the potential field with the appropriate boundary conditions, and (iii) differencing the potential to get the acceleration, and, if necessary, interpolating the potential back to the particles.
 These steps are described in detail below.
 
-First, the dark matter particles are distributed onto the grids using the cloud-in-cell (CIC) interpolation technique \citet{Hockney88} to form a spatially discretized density field $\rho_{\rm DM}$.
+First, the dark matter particles are distributed onto the grids using the second-order cloud-in-cell (CIC) interpolation technique \citet{Hockney88} to form a spatially discretized density field $\rho_{\rm DM}$.
 When doing the CIC interpolation, particle positions are (temporarily) advanced by $0.5 \Delta t v^n$ so that we generate an estimate of the time-centered density field.
@@
 In addition, since the gravitating field for a grid is defined beyond the grid edges (see below), dark matter particles from sibling grids and sub grids which lie within the entire gravitating field are used.
+This step can involve communication.
 
@@
 Cells which are covered by further refined grids are treated in a similar way (i.e.
-we use the sub grid cells as virtual CIC particles).
+we also use the sub grid cells as virtual CIC particles).
 This procedure results in a total gravitating mass field $\rho_{\rm total}^{n+1/2}$.
@@
 \begin{equation}
-G(\vec{k}) = - \frac{\Delta x}{2 \left( sin(k_x \Delta x/2)^2 + sin(k_y \Delta y/2)^2 + sin(k_z \Delta z/2)^2 \right) }
+G(\vec{k}) = - \frac{\Delta x}{2 \left( \sin(k_x \Delta x/2)^2 + \sin(k_y \Delta y/2)^2 + \sin(k_z \Delta z/2)^2 \right) }
 \end{equation}
@@
 
-For isolated boundary conditions, we use James method (REF).
+For isolated boundary conditions, we use James method \citep{James77}.
 In this case, the Greens function is generated in real-space so as to have the correct zero-padding properties and then transformed into the Fourier domain.
 In either case, the potential is then transformed back into the real domain to get potential values at the cell centers.
-These are differenced with a two-point centered difference to get accelerations at the cell centers (except if we are using the staggered Zeus-like solver, in which case the accelerations are computed at the cell faces to match the velocities).
+These are differenced with a two-point centered difference to obtain accelerations at the cell centers (except if we are using the staggered Zeus-like solver, in which case the accelerations are computed at the cell faces to match the velocities).
+Particle accelerations are obtained using a (linear) CIC interpolation from the grid.
 
-In order to calculate more accurate potentials on subgrids, \enzo\ re-samples the dark matter density onto individual subgrids using the same CIC method as on the root grid, but at higher spatial resolution (and again adding baryon densities if applicable).
+In order to calculate more accurate potentials on subgrids, \enzo\ uses a similar but slightly different technique on the root grids.
+The generation of the total gravitating mass field is essentially identical, using CIC interpolation for both the particles and baryons, including sub grids as before.
+To compute the potential on sub grids; however, we use the standard seven-point (in three dimensions) second-order finite-difference approximation to Poisson's equation.
 Boundary conditions are then interpolated from the potential values on the parent grid.
-We use either tri-linear interpolation or a natural second-order spline: both methods seem to give similar results.
-The potential equation on each subgrid is then solved with the given Dirichlet boundary conditions.
-This solution is done with a multigrid relaxation technique (but only applied to each subgrid).
+We use either tri-linear interpolation or a natural second-order spline for this: both methods give similar results.
+The potential equation on each subgrid is then solved with the given Dirichlet boundary conditions with a multigrid relaxation technique.
+This is applied to each sub grid separately.
 
-The region immediately next to the boundary can contain unwanted oscillations (Anninos \etal REF), and so we use an expanded buffer zone around the grid, of size three root grid boundary zones (so typically six refined zones).
-The density is computed in this region and the potential solved, but only the region which overlaps with the grid itself is used to calculate accelerations.
+The region immediately next to the boundary can contain unwanted oscillations \citep[e.g.,][]{Anninos94}, and so we use an expanded buffer zone around the grid, of size three root grid boundary zones (so typically six refined zones for a refinement factor of 2).
+The density is computed in this region and the potential solved, but only the region which overlaps with the active region of grid itself is used to calculate accelerations.
 
@@
 To prevent overshoot, we average the potential on the boundary and allow for (typically) 4 iterations.
-This procedure can help in many cases, but does not produce a coherent solution across all grids and so does not solve the problem; we are working on a slower but more accurate method which does a multi-grid solve across the whole grid (Reynolds \etal, in preparation).
-
-Forces are computed on the mesh by finite-differencing potential values and are then interpolated to the particle positions, where they are used to update the particle's position and velocity information.
-Potentials on child grids are computed recursively and particle positions are updated using the same timestep as in the hydrodynamic equations.
+This procedure can help in many cases, but does not necessarily produce a coherent solution across all grids and so does not completely solve the problem; we are working on a slower but more accurate method which does a multi-grid solve across the whole grid (Reynolds \etal, in preparation).
 
 At this point it is useful to emphasize that the effective force resolution of an adaptive particle-mesh calculation is approximately twice as coarse as the grid spacing at a given level of resolution.
-%The potential is solved in each grid cell; however, the quantity of interest, namely the acceleration, is the gradient of the potential, and hence two potential values are required to calculate this.
-In addition, it should be noted that the adaptive particle-mesh technique described here is very memory intensive: in order to ensure reasonably accurate force resolution at grid edges the multigrid relaxation method used in the code requires a layer of ghost zones which is very deep -- typically 6 cells in every direction around the edge of a grid patch.
-This greatly adds to the memory requirements of the simulation, particularly because subgrids are typically small (on the order of $12^3 - 16^3$ real cells for a standard cosmological calculation) and ghost zones can dominate the memory and computational
-requirements of the code.

### 2013-03-20 numerical-particle.tex
updated gravity and particles in section 3

@@
 
-To solve the, \enzo\ uses a particle-mesh N-body method to calculate 
-the dynamics of collisionless systems \citep{Hockney88}.  This method 
-follows trajectories 
-of a representative sample of individual particles and is much more 
-efficient than a direct solution of the Boltzmann equation in most 
-astrophysical situations. 
-The gravitational potential is computed by solving the elliptic 
-Poisson's equation (Eq.~\ref{eq:potential}).
-
+\enzo\ uses a particle-mesh N-body method to calculate the 
+dynamics of collisionless systems \citep{Hockney88}.  This method 
+follows trajectories of a representative sample of individual particles 
+and is much more efficient than a direct solution of the Boltzmann 
+equation in essentially all astrophysical situations. 
+As described earlier, the gravitational potential is computed by solving the elliptic 
+Poisson's equation (Eq.~\ref{eq:potential}), and differencing the potential
+to find accelerations, which are then interpolated back to particles.  This
+acceleration is time-centered (because the underlying gravitating mass
+field is approximately time-centered), and so we have accelerations $\myvec{a}^{n+1/2}$
+for each particle.  These are used to update the particle positions and velocities
+starting from $\myvec{x}^n$ and $\myvec{v}^n$ using a standard drift-kick-drift technique:
 \begin{eqnarray}
 \label{eqn.driftkick}
-x^{n+1/2} & = & x^n + (\Delta t/2) v^{n} \nonumber \\
-v^{n+1} & = & v^n + \Delta t a^{n+1/2} \\
-x^{n+1} & = & x^{n+1/2} + (\Delta t/2) v^{n+1} \nonumber
+\myvec{x}^{n+1/2} & = & \myvec{x}^n + \frac{\Delta t}{2 a^n} \myvec{v}^{n} \nonumber \\
+\myvec{v}^{n+1} & = & \myvec{v}^n \left(1 - \frac{\dot{a}^{n+1/2}}{a^{n+1/2}}\right) + \frac{\Delta t}{ a^{n+1/2}} \myvec{a}^{n+1/2} \\
+\myvec{x}^{n+1} & = & \myvec{x}^{n+1/2} + \frac{\Delta t}{2 a^{n+1}} \myvec{v}^{n+1} \nonumber
 \end{eqnarray}

### 2013-03-23 numerical-radiation-homogeneous.tex
Intermediate checkin for homogeneous radiation field data.

@@
 \label{sec.num.rad-homogeneous}
+
+Enzo supports the use of a set of spatially uniform (but possibly time-varying) radiation fields that can interface with the chemistry and cooling/heating routines described elsewhere.
+Many of these use fits to the H, He and He$^+$ ionizing and photo-ionization heating rates that are of the form
+\begin{equation}
+\rm{rate} = k_0 * (1+z)^{\alpha} * \exp{\left( \frac{\beta *(z-z_0)^2 }{1 + \gamma * (z+z_1)^2} \right)}
+\label{eq:homo_rad_template}
+\end{equation}
+where the constant coefficients $\alpha$, $\beta$, $\gamma$, $z_0$ and $z_1$ are fits from the literature.
+
+\begin{table}
+\begin{center}
+\caption{Homogeneous radiation field coefficients}
+\begin{tabular}{lrrrrrr}
+\tableline\tableline
+{Element} & {$k_0$} &  {$\alpha$} & {$\beta$} & {$\gamma$} & {$z_0$} & {$z_1$}  \\
+\tableline
+\multicolumn{7}{c}{Radiation Type 1 \citep{1996ApJ...461...20H} for case $\alpha_q = 1.5$} \\
+\tableline
+H ionization & $6.7 \times 10^{-13}$ & 0.43 & 1/1.95 & 0 & 2.3 & 0 \\
+He ionization & $6.3 \times 10^{-15}$ & 0.51 & 1/2.35 & 0 & 2.3 & 0 \\
+He$^+$ ionization & $3.2 \times 10^{-13}$ & 0.50 & 1/2.00 & 0 & 2.3 & 0 \\
+H heating & $4.7 \times 10^{-24}$ & 0.43 & 1/1.95 & 0 & 2.3 & 0 \\
+He heating & $8.2 \times 10^{-24}$ & 0.50 & 1/2.00 & 0 & 2.3 & 0 \\
+He$^+$ heating & $1.6 \times 10^{-25}$ & 0.51 & 1/2.35 & 0 & 2.3 & 0 \\
+\tableline
+\multicolumn{7}{c}{Radiation Type 2 \citep{1996ApJ...461...20H} for case $\alpha_q = 1.8$} \\
+\tableline
+H ionization & $5.6 \times 10^{-13}$ & 0.43 & 1/1.95 & 0 & 2.3 & 0 \\
+He ionization & $3.2 \times 10^{-15}$ & 0.30 & 1/2.60 & 0 & 2.3 & 0 \\
+He$^+$ ionization & $4.8 \times 10^{-13}$ & 0.43 & 1/1.95 & 0 & 2.3 & 0 \\
+H heating & $3.9 \times 10^{-24}$ & 0.43 & 1/1.95 & 0 & 2.3 & 0 \\
+He heating & $6.4 \times 10^{-24}$ & 0.43 & 1/2.10 & 0 & 2.3 & 0 \\
+He$^+$ heating & $8.7 \times 10^{-26}$ & 0.30 & 1/2.70 & 0 & 2.3 & 0 \\
+\tableline
+\multicolumn{7}{c}{Radiation Type 3 \citep{Kirkman05}} \\
+\tableline
+H ionization &        $1.04 \times 10^{-12}$ & 0.231 & -0.6818 & 0.1646 & 1.855 & 0.3097 \\
+He ionization &       $1.84 \times 10^{-14}$ & -1.038 & -1.1640 & 0.1940 & 1.973 & -0.6561 \\
+He$^+$ ionization & $5.79 \times 10^{-13}$ & 0.278 & -0.8260 & 0.1730 & 1.973 & 0.2880 \\
+H heating &            $8.86 \times 10^{-25}$ & -0.0290 & -0.7055 & 0.1884 & 2.003 & 0.2888 \\
+He heating &         $5.86 \times 10^{-24}$ & 0.1764 & -0.8029 & 0.1732 & 2.088 & 0.1362 \\
+He$^+$ heating &   $2.17 \times 10^{-25}$ & -0.2196 & -1.070 & 0.2124 & 1.1782 & -0.9213 \\
+\end{tabular}
+%\tablecomments{}
+\end{center}
+\end{table}

### 2013-03-24 numerical-radiation-homogeneous.tex
finished homogeneous radiation section

@@
 
-Enzo supports the use of a set of spatially uniform (but possibly time-varying) radiation fields that can interface with the chemistry and cooling/heating routines described elsewhere.
+\enzo\ supports the use of a set of spatially uniform (but possibly time-varying) radiation fields that can interface with the chemistry and cooling/heating routines described elsewhere.
 Many of these use fits to the H, He and He$^+$ ionizing and photo-ionization heating rates that are of the form
@@
 where the constant coefficients $\alpha$, $\beta$, $\gamma$, $z_0$ and $z_1$ are fits from the literature.
+For radiation field types 1-3, we give the coefficients for the photoionization (and photo-heating) rates of H, He, and He$^+$ in Table~\ref{table:homo_coefs}.
+Types 1 and 2 are based on \citet{1996ApJ...461...20H} with two different intrinsic quasar spectra slopes, while Type 3 is from \citet{2012ApJ...746..125H}, modified to match the normalized field found in \citet{Kirkman05}.
+Type 4 is the same as Type 3 but also adds X-ray Compton heating from \citet{MadauEfstathiou99}, using equations (4) and (11) of that paper.
+
+Homogeneous radiation field types 5 and 6 start with a spectral shape which is then integrated against the appropriate H, He, He$^+$ cross-sections to compute the ionization and photo-ioniziation heating rates (we use 400 bins logarithmically specced from 0.74 eV to $7.24 \times 10^9$ eV).
+In particular, Type 5 has a hard, featureless quasar-like power law spectrum $f_{\nu} = f_{\rm HI} \nu^{\alpha_0}$, where $\alpha_0 = 1.5$ by default and the spectrum is normalized at the HI ionization edge.
+Type 6 has the same spectrum by is attenuated by a column density of $10^{22}$ cm$^{-3}$ neutral hydrogen.
+
+Types 8, 9, and 14 have only photo-dissociating Lyman-Werner flux, with Type 8 being constant and Type 9 using the redshift-dependent results of \citet{TrentiStiavelli09}.
+Type 14 uses a fit from \citet{WiseAbel05} for the range $6 < z < 50$, constant for $z<6$, and zero for $z > 50$.
+Other types are undefined or unused.
 
@@
 \tableline
-\multicolumn{7}{c}{Radiation Type 1 \citep{1996ApJ...461...20H} for case $\alpha_q = 1.5$} \\
+\multicolumn{7}{c}{Radiation Type 1 \citep{1996ApJ...461...20H} for the case $\alpha_q = 1.5$} \\
 \tableline
@@
 \tableline
-\multicolumn{7}{c}{Radiation Type 2 \citep{1996ApJ...461...20H} for case $\alpha_q = 1.8$} \\
+\multicolumn{7}{c}{Radiation Type 2 \citep{1996ApJ...461...20H} for the case $\alpha_q = 1.8$} \\
 \tableline
@@
 \tableline
-\multicolumn{7}{c}{Radiation Type 3 \citep{Kirkman05}} \\
+\multicolumn{7}{c}{Radiation Type 3 modified \citep{2012ApJ...746..125H}} \\
 \tableline
@@
 %\tablecomments{}
+\label{table:homo_coefs}
 \end{center}

### 2013-03-24 parallel.tex
putting in draft of unigrid scaling plot + documents.

@@
 % Brian to work on this, scaling data from Harkness
+Figure~\ref{fig.uniscale} shows \enzo\ weak scaling for a sequence of
+scaled non-AMR Lyman alpha forest calculations on NICS Kraken XT4 and
+ORNL Jaguar XT4 using decomposition with $128^3$ root grid tiles.  The
+grid resolution is of constant physical size.  Scaling up to 98,304
+cores has been achieved on the NICS Kraken XT5 using a hybrid-parallel
+version of \enzo\.  Excellent scaling in unigrid mode is seen as long
+as each core is given enough work -- somewhere around $128^3$
+cells/core seems to be about right.
+
+\begin{figure}
+\begin{center}
+\includegraphics[width=0.4\textwidth]{figures/enzo_unigrid_weak_scaling.eps}
+\caption{Unigrid scaling.  ENZO weak scaling performance for a set of
+  Lyman Alpha Forest cosmology sim- ulations with constant physical
+  resolution per mesh point, showing cell updates per second per
+  computational core plotted as a function of the number of root grid
+  tiles of dimension $128^3$ (R) in each dimension. The number of MPI
+  tasks is N$ = R^3$, so R$ = 16$ on this plot corresponds to a $2048^3$ computational mesh running on 4096 MPI tasks. }
+\label{fig.uniscale}
+\end{center}
+\end{figure}
 

### 2013-03-25 parallel.tex
working on unigrid scaling section

@@
 
-% Brian to work on this, scaling data from Harkness
-Figure~\ref{fig.uniscale} shows \enzo\ weak scaling for a sequence of
-scaled non-AMR Lyman alpha forest calculations on NICS Kraken XT4 and
-ORNL Jaguar XT4 using decomposition with $128^3$ root grid tiles.  The
-grid resolution is of constant physical size.  Scaling up to 98,304
+It is advantageous to use \enzo\ in its ``unigrid'' (i.e.,
+non-adaptive mesh) mode for a variety of problems, including
+non-gravitating turbulence
+\citep[e.g.,][]{2002ApJ...569L.127K,Kritsuk04}, the Lyman-alpha forest
+\citep{2005MNRAS.361...70J,2009MNRAS.399.1934P}, or feedback of
+metal-enriched gas from galaxies into the intergalactic medium
+\citep{2004ApJ...601L.115N,2011ApJ...731....6S}.  Achieving good
+scaling of the code in unigrid mode is relatively straightforward --
+upon initialization, unigrid simulations are decomposed such that each
+MPI process has a roughly equal subvolume (and thus number of grid
+cells), meaning that work is evenly distributed.  Communication
+patterns for both the gravity solve (which uses a fast Fourier
+transform) and the fluid solves (which transfer boundary information
+between subvolumes) are predictable and straightforward, and
+rebuilding of the grid hierarchy does not take place, removing a
+substantial global operation and a great deal of communication.
+
+Figure~\ref{fig.uniscale} shows \enzo\ weak scaling results for a
+sequence of scaled unigrid Lyman alpha forest calculations. These
+calculations include dark matter dynamics, hydrodynamics using the
+piecewise parabolic method, six-species non-equilibrium chemistry and
+radiative cooling, and a uniform metagalactic ultraviolet background.
+In this sequence of test calculations, we perform a weak scaling test
+on up to 13,824 MPI tasks on the NICS Kraken XT4 and ORNL Jaguar XT4
+supercomputers\footnote{these simulations were performed in 2008,
+prior to conversion of both machines to the current-generation
+systems}.  In this test, each MPI task was given a $128^3$ root grid
+tile (i.e., $128^3$ grid cells containing baryon quantities) and,
+initially, approximately $128^3$ dark matter particles.  The number of
+grid cells was constant throughout the calculation; the number of dark
+matter particles varies as they are moved from subvolume to subvolume
+as structure evolves.  The grid resolution was kept at a constant
+comoving size of $\simeq 40$~kpc/h, and as the core count was
+increased, so was the simulation volume.  On each machine, a compute
+node contained a single AMD Opteron quad-core chip (2.1 Ghz on Jaguar;
+2.3 Ghz on Kraken) with 2 GB/memory per core (8 GB/total per node).
+Both machines used the SeaStar2 interconnect.  In the scaling study,
+calculations were run with 1, 2 or 4 MPI tasks per node.  The figure
+shows cell updates per second per MPI process; perfect scaling would
+be a horizontal line.
+
+As can be seen in Figure~\ref{fig.uniscale}, the unigrid weak scaling
+performance of the code is extremely good for this problem, with only
+a 20\% decrease in cell updates/second/task as the code is scaled from
+8 to 4,096 MPI tasks, and a 40\% decrease in performance overall going
+from 8 to 13,824 (or $24^3$) MPI tasks.  We speculate that this
+decrease is likely to be partially due to global MPI communications
+used to, e.g., calculate the overall timestep of the simulation, and
+also likely due to load imbalances due to increasing cosmological
+power (and thus an increasingly uneven distribution of dark matter
+particles between MPI tasks at late times) as the simulation volume
+grows.  We also observe that a systematic difference in speed can be
+seen between the two machines, which can be attributed primarily to
+the slightly faster CPUs on Kraken (2.3 Ghz, vs. 2.1 Ghz on Jaguar).
+The difference in speed when using different numbers of MPI tasks per
+node can be attributed primarily due to differences in competing usage
+of shared cache on the quad-core chips used on this machine.
+
+Broadly, excellent scaling in \enzo's unigrid
+mode is seen for a variety of problems as long as each compute core is
+given an adequate amount of work to do.  For cosmological simulations, this seems to be
+roughly $128^3$ cells per core.  If fewer cells per core are used, the
+CPU is essentially data-starved, and poor scaling is observed due to
+computing units being idle while waiting for information to be
+communicated from other processes (for, e.g., boundary information or
+gravity solves).  Substantially larger cell counts per core  would in
+principle help scaling, but are typically impractical on
+most machines due to memory limits.
+
+As a final point, we 
+up to 98,304
 cores has been achieved on the NICS Kraken XT5 using a hybrid-parallel
+(MPI + OpenMP)
 version of \enzo\.  Excellent scaling in unigrid mode is seen as long

### 2013-03-25 parallel.tex
final checkin of weak scaling section, including some new references.

@@
 
-\subsubsection{Unigrid scaling \red{(Brian)}}
+\subsubsection{Unigrid scaling}
 
@@
 on up to 13,824 MPI tasks on the NICS Kraken XT4 and ORNL Jaguar XT4
-supercomputers\footnote{these simulations were performed in 2008,
+supercomputers\footnote{These simulations were performed in 2008,
 prior to conversion of both machines to the current-generation
@@
 
-Broadly, excellent scaling in \enzo's unigrid
-mode is seen for a variety of problems as long as each compute core is
-given an adequate amount of work to do.  For cosmological simulations, this seems to be
+Broadly, excellent scaling in \enzo's unigrid mode is seen for a
+variety of problems as long as each compute core is given an adequate
+amount of work to do.  For cosmological simulations, this seems to be
 roughly $128^3$ cells per core.  If fewer cells per core are used, the
@@
 gravity solves).  Substantially larger cell counts per core would in
-principle help scaling, but are typically impractical on
-most machines due to memory limits.
+principle help scaling by reducing the amount of inter-process
+communication needed, but larger cell counts are typically impractical
+on most machines due to memory limits.
 
-As a final point, we 
-up to 98,304
-cores has been achieved on the NICS Kraken XT5 using a hybrid-parallel
-(MPI + OpenMP)
-version of \enzo\.  Excellent scaling in unigrid mode is seen as long
-as each core is given enough work -- somewhere around $128^3$
-cells/core seems to be about right.
+As a final point, we observe that scaling at larger core counts has
+been measured, but only with an experimental hybrid-parallel (MPI +
+OpenMP) version of \enzo.  Using this version, scaling comparable to
+that shown in Figure~\ref{fig.uniscale} was seen on up to 98,304 cores
+on the NICS Kraken XT5 (an upgraded version of the XT4 machine used
+for the scaling study shown in the figure), using 2-8 OpenMP threads
+per MPI process.  Hybrid parallelism has the potential to
+substantially improve scaling by reducing the amount of communication
+per grid tile (as described in the previous paragraph); however, the
+optimal ratio of OpenMP threads per MPI task seems to vary
+substantially between computational platforms and astrophysical
+problems/required \enzo\ physics modules, and thus we hesitate to
+provide strict guidelines here.
 
@@
 \includegraphics[width=0.4\textwidth]{figures/enzo_unigrid_weak_scaling.eps}
-\caption{Unigrid scaling.  ENZO weak scaling performance for a set of
-  Lyman Alpha Forest cosmology sim- ulations with constant physical
-  resolution per mesh point, showing cell updates per second per
-  computational core plotted as a function of the number of root grid
-  tiles of dimension $128^3$ (R) in each dimension. The number of MPI
-  tasks is N$ = R^3$, so R$ = 16$ on this plot corresponds to a $2048^3$ computational mesh running on 4096 MPI tasks. }
+\caption{\enzo\ weak scaling performance for a set of Lyman alpha
+forest cosmology simulations with constant comoving spatial resolution
+per grid cell, showing cell updates per second per computational core
+plotted as a function of the number of root grid tiles of dimension
+$128^3$ (R) in each dimension.  The number of MPI tasks is N$ = R^3$,
+so R$ = 16$ on this plot corresponds to a $2048^3$ computational mesh
+running on 4096 MPI tasks.  This plot goes from R$ = 2$ (8 MPI tasks)
+to R$ = 24$ (13,824 MPI tasks) on two supercomputers -- NICS Kraken
+and ORNL Jaguar when they were Cray XT4 systems -- and using 1, 2 or 4
+MPI processes per node, where each compute node contained a single
+quad-core AMD Opteron CPU having a speed of 2.1 GHz on Jaguar and 2.3
+GHz on Kraken.}
 \label{fig.uniscale}

### 2013-03-25 conclusions.tex
finishing up conclusions.

--- 2013-01-18 conclusions.tex
@@
 
-\begin{enumerate}
-\item near-future development plans
-\item far-future development plans
-\item clear areas of improvement
-\end{enumerate}
+In this paper, we have presented the algorithms underlying \enzo, an
+open-source adaptive mesh refinement code designed for
+self-gravitating compressible fluid dynamics, including the effects of
+magnetic fields, radiation transport, and a variety of microphysical
+and subgrid processes.  In addition, we have described the \enzo\ code
+development process, have shown the outputs of some representative
+test problems, and have provided information about \enzo's performance
+and parallel scaling on recent supercomputing platforms.  The \enzo\
+code, its test suite, and all of the scripts used to generate plots
+and figures for this paper are open source and are available at the
+\enzo\ website, http://enzo-project.org.  Furthermore, the yt toolkit,
+which is designed to analyze \enzo\ data (as well as data from a wide
+variety of other simulation tools), can be found at its website,
+http://yt-project.org.  Both of these codes have active user and
+developer communities, extensive documentation and user suppport, and
+strong mechanisms for users to contribute their changes and fixes to
+the codebase.
+
+The developers of the \enzo\ code are currently working on several
+projects that will extend the functionality, scalability, or overall
+performance of the code in the near future.  Projects that will appear
+in forthcoming releases of the \enzo\ code include:
+
+\begin{itemize}
+\item The creation of a hybrid-parallel version of \enzo, combining
+MPI for communication between nodes of a supercomputer and OpenMP for
+thread-based parallelism within a node.  This will reduce on-node
+memory usage and overall scaling behavior.
+\item The restructuring of \enzo's treatment of particles to
+accommodate a wider range of ``active'' particles that can easily
+interact with each other and with multiple grids, and include sink,
+source, and particle creating, splitting, and merging functionality.
+\item A new HYPRE-based AMR gravity solver that is faster, more
+accurate, and more scalable than the current multigrid solver.
+\item New infrastructure for problem initialization, enabling
+  users to more quickly and easily create new types of simulations.
+\end{itemize}
+
+In the longer term, it seems likely that trends in supercomputing
+hardware will require substantial modifications to \enzo's core
+infrastructure, and very possibly some of the core algorithsms.  More
+specifically, these trends involve the usage of specialized
+large-core-count, vectorized computing units such as graphics
+processuing units (GPUs) or chips like the Intel Xeon Phi, as well as
+precipitously decreasing amounts of RAM per computing core.  The
+former trend means that the amount of processing power per compute
+node will continue to trend upward, likely much faster than the
+bandwidth between nodes, and will require tremendous reduction in
+inter-node (and possibly inter-CPU) communication in order to maintain
+code scalability.  Also, much of the current code will need to be
+rewritten to take advantage of the vector nature of these CPUs, making
+assumptions that are quite unlike those made in much of the current
+codebase.  The latter trend means that duplication of data -- for
+example, the grid hierarchy -- must be effectively eliminated to save
+memory, and all inter-core and inter-node communication must be
+carefully thought through to minimize the amount of data moved.  An
+additional challenge as one goes to core counts in the tens to
+hundreds of millions (or more) is that the reliability of individual
+computing elements will become much more of an issue, requiring
+robustness to hardware failure to be built into the code at some
+level.  Furthermore, we are nearing the physical limits of transistor
+speed and latency~\citep{feynman1999feynman}, meaning that simple
+hardware improvements will not make these challenges disappear, and
+careful thought (and the rewriting of a great deal of code) must take
+place!  These challenges are not unique to the \enzo\ code, and in
+fact are faced by effectively all applications that wish to take
+advantage of new computational architectures.  Thus, we suspect that
+\enzo\ (or a code that has the capabilities of \enzo, from a user's
+point of view) will continue to be usable at the largest scales on
+such machines.

### 2013-03-26 code-tests.tex
Tossed some ideas into the conclusions section.  Made a couple of small tweaks in the test problem section.
working on test problems.
working on test problems, adding a couple of figures.
updating code tests.

@@
 \red{(Brian)}
+
+The double mach reflection test is a classic test of hydrodynamical
+algorithms, and was originally described in
+\citet{1984JCoPh..54..115W} \citep[and more recently in][]{2008ApJS..178..137S}.
+
 Problem type 4.  This is one of Alexei's test problems.  Tests
-boundary conditions in hydro.
+boundary conditions in hydro.  See Figure~\ref{fig.doublemach}.
+
+\begin{figure}
+\begin{center}
+\includegraphics[width=0.8\textwidth]{figures/DoubleMachTest.eps}
+\caption{Double mach test.}
+\label{fig.doublemach}
+\end{center}
+\end{figure}
 
@@
 \label{sec.tests.sedov}
-\red{(Elizabeth)}
 
@@
 \red{(Brian)}
-Problem type 20, unigrid and AMR both.  Tests gravity + hydro.
+Problem type 20, unigrid and AMR both.  Tests gravity + hydro.  Figure~\ref{fig.pancake}.
+
+\begin{figure}
+\begin{center}
+\includegraphics[width=0.8\textwidth]{figures/AMRZeldovichPancake.eps}
+\caption{Zel`Dovich Pancake test.}
+\label{fig.pancake}
+\end{center}
+\end{figure}
 
@@
 One test for the thermal conduction (maybe 1 for isotropic, 1 for
-anisotropic).
+anisotropic).  Figure~\ref{fig.conduct}.
+
+\begin{figure}
+\begin{center}
+\includegraphics[width=0.4\textwidth]{figures/iso_conduction_final_output.eps}
+\includegraphics[width=0.4\textwidth]{figures/aniso_conduction_final_output.eps}
+\caption{Temperature blah blah.  Conduction tests!}
+\label{fig.conduct}
+\end{center}
+\end{figure}
 

### 2013-03-26 numerical-radiation-fld.tex

@@
 (\ref{eq:fld_radiation}-\ref{eq:fld_heating}) along with the chemical network 
-(eq.~\ref{eq:species_evolution} is solved
+(eq.~\ref{eq:species_evolution}) is solved
 independently of Enzo's hydrodynamics, gravity and dark-matter solvers
@@
 the advective portions of equations (\ref{eq:fld_radiation}) and
-(\ref{eq:fld_chemistry}) to be taken care of by the fluid solvers.
+(\ref{eq:species_evolution}) to be taken care of by the fluid solvers.
 Due to the disparate time scales between radiation transport and chemical
@@
 
-We then evolve the system (\ref{eq:fld_heating}-\ref{eq:fld_chemistry}):  
+We then evolve the heating and chemistry system (\ref{eq:fld_heating})
+and (\ref{eq:species_evolution}):  
 $(E_r^{n+1}, e_c^n, {\mathrm n}_i^n) \to (E_r^{n+1}, e_c^{n+1},

### 2013-03-26 code-tests.tex
double mach test.

@@
 
-The double mach reflection test is a classic test of hydrodynamical
-algorithms, and was originally described in
-\citet{1984JCoPh..54..115W} \citep[and more recently in][]{2008ApJS..178..137S}.
+The double mach reflection test is a classic two-dimensional test of
+hydrodynamic algorithms, and was originally described in
+\citet{1984JCoPh..54..115W} \citep[and more recently
+in][]{2008ApJS..178..137S}.  In this problem (shown in
+Figure~\ref{fig.doublemach}), a shock is injected at an angle to a
+reflecting surface, and a jet appears along the reflecting surface.
+The ideal solution is self-similar, and the appearance of this
+solution is highly sensitive to numerical diffusion.  If numerical
+noise is present, a Kelvin-Helmholz instability develops along this
+jet and breaks the self-similarity.
 
-Problem type 4.  This is one of Alexei's test problems.  Tests
-boundary conditions in hydro.  See Figure~\ref{fig.doublemach}.
+In the test problem shown in Figure~\ref{fig.doublemach}, a 2D
+simulation with $960 \times 240$ cells was formed with a domain of $x
+= [0, 4]$ and $y = [0, 1]$.  We use an ideal gas equation of state of
+$\gamma = 1.4$, a pre-shock density of 1.4, and a pre-shock specific
+internal energy of $2.5/1.4$ (all in arbitrary units).  A Mach 10
+shock is initialized with a shock normal that is 30$^\circ$ from the
+x-axis and an initial position on the lower boundary of x$ = 1/6$.
+The lower y boundary and right x boundary are reflecting; the left x
+boundary is inflowing, and the upper y boundary has a time-dependent
+boundary condition that allows the shock to propagate into the domain
+as if it extends to infinity.  The simulation starts at t$ = 0$ and
+runs until t$ = 0.205$ (arbitrary units), at which point the rightmost
+extent of the shock should be at roughly x$ = 3$.  In this run, we use
+the direct Eulerian implementation of the piecewise parabolic hydro
+method, with diffusion, flattening, and shock steepening all turned
+on.
+
+It is instructive to compare Figure~\ref{fig.doublemach} to Figure 9
+in~\citet{1984JCoPh..54..115W}.  By the end of the simulation, a dense
+jet is apparent at the leading edge of the shock, propagating along
+the x-axis.  The shape of this jet is sensitive to numerical
+diffusion, and our figure compares favorably to the higher-order
+images from Woodward \& Colella.
 

### 2013-03-26 code-tests.tex
doing a bit of cleanup in the code test file

@@
 \includegraphics[width=0.4\textwidth]{figures/sedov-zeus-slice.eps}
-\caption{Density slice from the Sedov Blast Test at $t = 0.07$.
-Left-hand image shows the results from PPM ({\tt HydroMethod = 0}), right-hand image shows the results from Zeus ({\tt Hydromethod = 2}).
-Notably, the Zeus shock front has progressed less far than forthe PPM run.
-This is due to energy loss when conserving internal, not total, energy.}
+\caption{Density slice from the Sedov Blast Test at $t =
+0.07$. Left-hand image shows the results from PPM ({\tt HydroMethod =
+0}), right-hand image shows the results from Zeus ({\tt Hydromethod =
+2}). Notably, the Zeus shock front has progressed less far than forthe
+PPM run. This is due to energy loss when conserving internal, not
+total, energy.}
 \label{fig.sedov1}
@@
 \includegraphics[width=\textwidth]{figures/sedov-profiles.eps}
-\caption{Radially averaged profiles for the Sedov Blast Test at $t = 0.07$. Clockwise from top-left shows density, velocity, internal energy and pressure.}
+\caption{Radially averaged profiles for the Sedov Blast Test at $t =
+0.07$. Clockwise from top-left shows density, velocity, internal
+energy and pressure.}
 \label{fig.sedov2}
@@
 
-The Sedov Blast Test \citep{Sedov1959} models an intense explosion, initiated by depositing thermal energy into a homogenous distribution of gas.
-The result is a strong spherical shock wave centered on the point of energy injection.
-This problem is a popular test of astrophysical numerical codes for three reasons: Firstly, it is particularly appropriate to astrophysics since it represents the situation when a supernovae explosion occurs.
-Secondly, it has an exact analytical solution whereby the shock front's radial position is given by:
+The Sedov Blast Test \citep{Sedov1959} models an intense explosion,
+initiated by depositing thermal energy into a homogenous distribution
+of gas. The result is a strong spherical shock wave centered on the
+point of energy injection.  This problem is a popular test of
+astrophysical numerical codes for three reasons: Firstly, it is
+particularly appropriate to astrophysics since it represents the
+situation when a supernovae explosion occurs. Secondly, it has an
+exact analytical solution whereby the shock front's radial position is
+given by:
 
-\begin{equation}
-r(t) = \left(\frac{E_0}{\alpha\rho_0}\right)^{1/5}t^{2/5}
+\begin{equation} r(t) =
+\left(\frac{E_0}{\alpha\rho_0}\right)^{1/5}t^{2/5}
 \end{equation}
 
-\noindent where $E_0$ is the initial energy injection, $\rho_0$ is the background density and $\alpha = 1.0$ for cylindrical symmetry and an ideal gas with $\gamma = 1.4$.
-For the full derivation see \citet{Sedov1959}.
-This solution makes it possible to assess how well the code performs.
-Thirdly, the spherical shock is a challenging problem for numerical codes (both grid and particle) since the shock wave increases in size as the simulation progresses and its symmetrical nature highlights any directional preferences that grid codes can succumb to.
-The test presented here is the two-dimensional version that is included in the Enzo distribution.
-The three-dimensional results from this test, both for Enzo and three other leading astrophysics codes, can be founds in \citet{Tasker2008}.
+\noindent where $E_0$ is the initial energy injection, $\rho_0$ is the
+background density and $\alpha = 1.0$ for cylindrical symmetry and an
+ideal gas with $\gamma = 1.4$. For the full derivation see
+\citet{Sedov1959}. This solution makes it possible to assess how well
+the code performs. Thirdly, the spherical shock is a challenging
+problem for numerical codes (both grid and particle) since the shock
+wave increases in size as the simulation progresses and its
+symmetrical nature highlights any directional preferences that grid
+codes can succumb to. The test presented here is the two-dimensional
+version that is included in the Enzo distribution. The
+three-dimensional results from this test, both for Enzo and three
+other leading astrophysics codes, can be founds in \citet{Tasker2008}.
 
-In the initial state, the box contains a homogenous distribution of gas at a density of 1 (note that, in the absense of gravity, this problem is scale-free and so without units).
-Thermal energy is deposited into a single cell at the center of the box with $E_0 = 10.0$.
-The problem is completed in two dimensions with reflecting boundary conditions (the default for Enzo) and uses a boxsize of side $1$.
-For this problem, we selected a top grid of $100 \times 100$ and four levels of refinement, placed based on shock location and the slope in density and total energy (this corresponds to cell flagging methods 1 and 3 in the Enzo parameter file).
-The exception to this scheme is in the initial conditions where grids are placed directly around the injection point.
-The results were assessed at $t = 0.07$, which corresponding to a time just before the shock reaches the box edge (see Figure~\ref{fig.sedov1}).
+In the initial state, the box contains a homogenous distribution of
+gas at a density of 1 (note that, in the absense of gravity, this
+problem is scale-free and so without units). Thermal energy is
+deposited into a single cell at the center of the box with $E_0 =
+10.0$. The problem is completed in two dimensions with reflecting
+boundary conditions (the default for Enzo) and uses a boxsize of side
+$1$. For this problem, we selected a top grid of $100 \times 100$ and
+four levels of refinement, placed based on shock location and the
+slope in density and total energy (this corresponds to cell flagging
+methods 1 and 3 in the Enzo parameter file). The exception to this
+scheme is in the initial conditions where grids are placed directly
+around the injection point. The results were assessed at $t = 0.07$,
+which corresponding to a time just before the shock reaches the box
+edge (see Figure~\ref{fig.sedov1}).
 
-Figure~\ref{fig.sedov2} shows the radial profiles for the simulation run with the PPM hydro-solver (blue dashed line) and the Zeus hydro-solver (red dash-dot-dot line) together with the analytical solution (black solid line).
-Clockwise from the top-left are density, velocity, internal energy and pressure.
-PPM matches the analytical solution extremely well for all quantities.
-However, Zeus' shock front lags behind the analytical position.
-This can also be seen in the slices shown in Figure~\ref{fig.sedov1}.
-The cause of this discrepancy is that Zeus shows a substantial energy lost during the first few timesteps and produces a diamond-shaped, rather than spherical, shockfront during this time.
-After this, the code correctly conserves energy but this intial energy loss remains clearly visible in the position of the shock at $t = 0.07$.
-This problem was addressed directly by \citet{Clarke2010}, who attributed the source of the issue to this version of Zeus solving the internal, rather than total, energy equation.
-In situations with strong energy gradients, this choice caused an energy loss and the artificial viscosity produces the direction-dependent shockfront shape.
-In their paper, \citet{Clarke2010} presents results from an alternative version of Zeus that conserves total energy.
-This problem is less marked for smaller energy gradients and it should be noted that Zeus' stability and speed make it a highly competitive choice, despite the disagreements in this test.
+Figure~\ref{fig.sedov2} shows the radial profiles for the simulation
+run with the PPM hydro-solver (blue dashed line) and the Zeus
+hydro-solver (red dash-dot-dot line) together with the analytical
+solution (black solid line). Clockwise from the top-left are density,
+velocity, internal energy and pressure. PPM matches the analytical
+solution extremely well for all quantities. However, Zeus' shock front
+lags behind the analytical position. This can also be seen in the
+slices shown in Figure~\ref{fig.sedov1}. The cause of this discrepancy
+is that Zeus shows a substantial energy lost during the first few
+timesteps and produces a diamond-shaped, rather than spherical,
+shockfront during this time. After this, the code correctly conserves
+energy but this intial energy loss remains clearly visible in the
+position of the shock at $t = 0.07$. This problem was addressed
+directly by \citet{Clarke2010}, who attributed the source of the issue
+to this version of Zeus solving the internal, rather than total,
+energy equation. In situations with strong energy gradients, this
+choice caused an energy loss and the artificial viscosity produces the
+direction-dependent shockfront shape. In their paper,
+\citet{Clarke2010} presents results from an alternative version of
+Zeus that conserves total energy. This problem is less marked for
+smaller energy gradients and it should be noted that Zeus' stability
+and speed make it a highly competitive choice, despite the
+disagreements in this test.
 
@@
 \label{sec.tests.mhd}
-Figure \ref{fig.orszag} shows the Orszag-Tang vortex problem, a typical test
-problem for MHD \citep{Orszag79}.  The left panel shows the MHD-CT result, while
-the right panel shows the Dedner result.  This test shows that small scale structure
-can be generated in MHD from large scale initial perturbations.  
-The test begins 
-with uniform density, $\rho_0=25/36 \pi$ and pressure, $P_0=5/12 \pi$.  There is
-a single rotational mode in the velocity, and two in the magnetic field: 
+Figure \ref{fig.orszag} shows the Orszag-Tang vortex problem, a
+typical test problem for MHD \citep{Orszag79}.  The left panel shows
+the MHD-CT result, while the right panel shows the Dedner result.
+This test shows that small scale structure can be generated in MHD
+from large scale initial perturbations.  The test begins with uniform
+density, $\rho_0=25/36 \pi$ and pressure, $P_0=5/12 \pi$.  There is a
+single rotational mode in the velocity, and two in the magnetic field:
 ${\bf v}_0 $ = (-sin(2$\pi$ y) $ \hat{x}$ , sin(2$\pi$ x) $\hat{y}$),
-${\bf B}_0$  =  (-sin(2 $\pi$ y) $ \hat{x},$ sin( 4 $\pi$ x )$\hat{y}$).  The
-simulation is evolved to t=0.48.  One can see that the structures are accurately
-represented as compared to, for instance, \citet{Toth00}, and that the
-resolution of shocks is comparable in both methods.
+${\bf B}_0$ = (-sin(2 $\pi$ y) $ \hat{x},$ sin( 4 $\pi$ x )$\hat{y}$).
+The simulation is evolved to t=0.48.  One can see that the structures
+are accurately represented as compared to, for instance,
+\citet{Toth00}, and that the resolution of shocks is comparable in
+both methods.
 
-{\emph Dave's note:  I'll probably play around with these results in future
-versions.}
+{\emph Dave's note: I'll probably play around with these results in
+future versions.}
 

### 2013-03-26 code-tests.tex
adding text for the zel'dovich pancake.

@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsubsection{Zel`Dovich Pancake}
+\subsubsection{Zel`dovich Pancake}
 \label{sec.tests.pancake}
-\red{(Brian)}
-Problem type 20, unigrid and AMR both.  Tests gravity + hydro.  Figure~\ref{fig.pancake}.
+
+The Zel`dovich pancake \citep{1970A&A.....5...84Z} is particularly
+relevant to cosmological simulations, because it includes many
+features that are critical to structure formation: hydrodynamics,
+expansion, and self gravity.  This problem represents an ideal,
+isolated caustic formation, and is thus a useful proxy for much more
+complicated structures in full 3-dimensional cosmological simulations
+(such as the collapse of gas onto a cosmological halo or filament).
+
+The initial conditions are simple, and we follow the prescription of
+\citet{Anninos94}.  Assuming a flat cosmology, the density
+perturbation is given by
+
+\begin{equation}
+\rho(x_l) = \rho_0 \big[ 1 - \frac{1+z_c}{1+z} \cos(k x_l) \big]
+\end{equation}
+
+with the internal energy of the gas set so that the entropy of the gas
+maintains a constant value throughout.  The velocity perturbation is given by
+
+\begin{equation}
+v(x_l) = -H_0 \frac{1 + z_c}{(1+z)^{1/2}} \frac{\sin(k x_l}{k}
+\end{equation}
+
+In the equations above, $\rho_0$ is the background density, z$_c$ is a
+free parameter and is the redshift at which the sheet forms a caustic
+(i.e., 'pancakes'), z is the redshift of initialization, x$_l$ is the
+Lagrangian mass coordinate, $k = 2 \pi / \lambda$ (where $\lambda$ is
+the perturbation wavelength), and H$_0$ is the $z = 0$ value of the
+Hubble constant.  Note that this solution is expressed in terms of
+Lagrangian positions, so one needs to convert this into the Eulerian
+coordinates, x$_e$, that are more useful to \enzo:
+
+\begin{equation}
+x_e = x_l - \frac{1 + z_c}{1 + z} \frac{\sin(k x_l)}{k}
+\end{equation}
+
+We note that the solution described above is exact up to the point of
+caustic formation.  In Figure~\ref{fig.pancake}, we show the results
+of a test of the AMR version of \enzo's Zel'dovich Pancake test.  A
+one-dimensional box of length $64$~Mpc/h is initialized at z$ = 20$ in
+an $\Omega_M = 1$ universe with h$ = 0.5$ and a background temperature
+of 100~K, $\rho_0 = \rho_c$.  The simulation is initialized with 64
+grid cells, refining by factors of four using criteria based on cell
+mass and the presence of shocks, for a maximum of 2 levels (i.e., an
+equivalent maximum resolution of 1024 grid cells).  The simulation was
+evolved to z$ = 0$ using the PPM hydro method, with the final output
+of the calculation being shown in Figure~\ref{fig.pancake}.
+
+Figure~\ref{fig.pancake} shows the key features of this test problem.
+The strong shocks and large density gradients are well-resolved, with
+density and velocity jumps being well-delineated.  The key features of
+this test problem can be resolved with far fewer cells -- simulations
+including a mere 8 cells resolve the key features, as shown in
+sections 3.3-3.5 of~\citet{1996PhDT........80B} -- but we choose a
+higher resolution here for illustrative purposes.
 
@@
 \includegraphics[width=0.8\textwidth]{figures/AMRZeldovichPancake.eps}
-\caption{Zel`Dovich Pancake test.}
+\caption{Zel`Dovich Pancake test shown at z$ = 0$, initialized in a
+$\Omega_m = 1$ universe at z$ = 20$ on a one-dimensional grid having
+64 cells, and further refined by factors of four based on cell mass
+and the presence of shocks for up to two additional levels of mesh,
+having a maximum effective resolution of 1024 grid cells.}
 \label{fig.pancake}

### 2013-03-26 code-tests.tex
one additional update of ZP test

@@
 including a mere 8 cells resolve the key features, as shown in
-sections 3.3-3.5 of~\citet{1996PhDT........80B} -- but we choose a
+sections 3.3.4-3.3.5 of~\citet{1996PhDT........80B} -- but we choose a
 higher resolution here for illustrative purposes.

### 2013-03-26 code-tests.tex
working on conduction test - anisotropic only.

@@
 
-\subsubsection{Representative thermal conduction test}
+\subsubsection{Anisotropic Thermal conduction}
 \label{sec.tests.conduct}
-\red{(Brian)}
-One test for the thermal conduction (maybe 1 for isotropic, 1 for
-anisotropic).  Figure~\ref{fig.conduct}.
+
+In a plasma, conduction takes place primarily due to electron motion.
+In astrophysical environments such as the intergalactic and
+intracluster media, with typical magnetic strengths of $\sim 1 \mu$G
+and temperatures in the millions of Kelvin, the electron gyroradius is
+very small compared to the scapes of interest in a cosmological
+simulation.  As a result, electrons can only move \textit{along}
+magnetic field lines, not perpendicular to them, which can result in
+very interesting magnetohydrodynamical instabilities
+\citep[e.g.,][]{2008ApJ...677L...9P,2008ApJ...688..905P}. 
+ The correct modeling of this behavior can be quite challenging, and is described in 
+Section~\ref{sec.num.conductions}.  Conduction in \enzo\ is calculated
+in both an operator-split and directionally-split manner, and the
+addition of heat transport solely along magnetic field lines requires
+computation of cross-terms in the temperature derivative at cell
+faces, which can add spurious oscillations in the temperature field in
+regions where the temperature gradient is strong in more than one
+spatial direction unless an appropriate flux-limiter \citep[such as that
+of][]{1977JCoPh..23..263V} is chosen.
+
+Figure~\ref{fig.conduct} shows a test that demonstrates the
+correct behavior of anisotropic thermal conduction in \enzo.  We
+initialize a two-dimensional, $256^2$ cell simulation having a physical
+scale of 1 kpc on a side, a uniform density of 1 proton/cc, and a
+background temperature of $10^6$ Kelvin.  Magnetic fields with a
+strength of  B$_0 = 1 \mu$G are initialized such that the field lines form
+circles around the center of the simulation volume, such that B$_x =
+-B_0\sin(\theta)$ and B$_y = B_0\cos(\theta)$, where $\theta$ is the
+angle measured from the $+x$ direction in a counterclockwise manner.
+A Gaussian temperature pulse is injected at $(0.75, 0.5)$ (in units of the box
+size), with a peak temperature of $10^8$ K and a FWHM of $0.015625$ of
+the box size.  This initial setup is shown in the left panel of
+Figure~\ref{fig.conduct}.  The simulation is then allowed to evolve
+with \textit{only} anisotropic conduction turned on (e.g.,
+no hydrodynamics or radiative cooling), and with a Spitzer fraction of
+f$_{sp} = 1$.
+
+The right panel of Figure~\ref{fig.conduct} shows the state of the
+simulation after 300 Myr.  Heat has clearly been transported only along
+field lines -- there has been no diffusion perpendicular to the
+magnetic field setup, which is critical for many studies involving
+anisotropic thermal conduction.  No oscillations are seen in the
+temperature field in regions where the fields are not aligned with the
+grid, suggesting that the flux-limiter is operating as expected.
 
@@
 \begin{center}
-\includegraphics[width=0.4\textwidth]{figures/iso_conduction_final_output.eps}
+\includegraphics[width=0.42\textwidth]{figures/aniso_conduction_initial_output.eps}
 \includegraphics[width=0.4\textwidth]{figures/aniso_conduction_final_output.eps}
-\caption{Temperature blah blah.  Conduction tests!}
+\caption{Two-dimensional anisotropic conduction test in a uniform,
+constant temperature background with circular magnetic fields centered
+on (0.5, 0.5).  The background medium has a density of 1 proton/cc and
+a temperature of $10^6$~K.  At t$ = 0$ (left panel), a Gaussian heat
+pulse is injected at (0.75, 0.5) with a FWHM of $0.015625$ (with all
+numbers given in units of the box size) and a peak temperature of
+$10^8$~K, and allowed to evolve without hydrodynamical motion (i.e.,
+static gas) and no radiative cooling for 300 Myr.  At t$ = 300$~Myr
+(right panel), heat has been transported along magnetic field lines
+with no significant diffusion perpendicular to field
+lines. Furthermore, there is no detectable oscillations in the
+temperature in regions where the field is not parallel with the grids.}
 \label{fig.conduct}

### 2013-03-26 numerical-conduction.tex
working on conduction test - anisotropic only.

@@
 \begin{equation}
-\mathbb{F} = -\kappa_{sp} \^{b} (\^{b} \cdot \nabla T)
+\mathbb{F} = -\kappa_{sp} \hat{b} (\hat{b} \cdot \nabla T)
 \end{equation}
 
-Where $\^{b}$ is the unit vector denoting the direction of the
+Where $\hat{b}$ is the unit vector denoting the direction of the
 magnetic field.  As with the isotropic thermal conduction, we allow a
@@
 flux limiter for calculations of the temperature field.  In this case,
-we choose the monotonized central difference flux limiter \citep{1977JCoPh..23..263V}, which serves to maintain numerical stability without
-sacrificing substantial speed or accuracy.
+we choose the monotonized central difference flux limiter
+\citep{1977JCoPh..23..263V}, which serves to maintain numerical
+stability without sacrificing substantial speed or accuracy.

### 2013-04-03 code-tests.tex
Add Sod shock tube figure and results.

@@
 \label{sec.tests.sodshock}
-\red{(Greg)}
-Problem type 1.  AMR version.  Tests the hydro.
-This is also problem 7.2 in the FLASH method paper.
+
+We begin the paper test suite with the classic Sod shock tube problem \citep{Sod78}, which provides a good test of a scheme's ability to resolve a clean Riemann problem with clear separation between the three waves, in particular a rarefaction fan, a contact discontinuity and a moderate shock.
+In Figure~\ref{fig.sodshocktube}, we show the solution for three of our hydro solves, the spatially third-order PPM, as well as the two second-order Zeus and MUSCL schemes.
+We use 100 cells across the domain, which is a relatively standard choice and show solutions both with and without AMR.
+Without AMR, it is clear that the PPM scheme produces by far the cleanest solution, with all wave families crisply reproduced.
+Zeus and MUSCL produce similar results, with MUSCL doing a slightly better job on the rarefaction fan.
+In the figure, we also show the mean absolute deviation from the exact solution, as well as the maximum deviation.
+These numbers confirm the qualitative differences noted previously.
+
+We also run the same simulation but with two levels of AMR (using a refinement factor of 2), triggered based on a normalized slope greater than 10\% in the density.
+This is a minimal level of refinement, and does not include the rarefaction fan at late times.
+The results are shown in the bottom panels of Figure~\ref{fig.sodshocktube}, and the results are much better for all three methods, with much sharper shocks and contact discontinuities, and even a better representation of the rarefaction wave, which is only refined at early times.
+Although the results are improved for all methods, the PPM still produces the best result, as shown clearly by the compute error norms, shown again in the figure.
+
+\begin{figure}
+\begin{center}
+\includegraphics[width=\textwidth]{figures/SodShockTube.eps}
+\caption{The density distribution of the classic Sod Shock Tube for three different solvers (from left to right) and with (bottom) and without (top) AMR.
+In each case 100 zones are used on the root level and the results are shown at $t=0.25$.
+All cells are plotted and color-coded by level (at the time shown, only small region surrounding the contact discontinuity and the shock are refined).
+In each panel, we show the exact solution as a solid line and the $L_1$ and $L_\infty$ error norms in the upper right.}
+\label{fig.sodshocktube}
+\end{center}
+\end{figure}
 

### 2013-04-22 ms.tex
Adding constrained transport MHD file and call from ms.tex
fixed up some subsections.  added a subsection in 2 and 3 for thermal conductivity.
Combining MUSCL and Dedner MHD
Intermediate checkin
Revised abstract.

@@
 %   3.3   Hydro: Godunov (MUSCL)
-%   3.4. Hydro: Finite difference (ZEUS)
-%   3.5. MHD: Dedner
-%   3.6. Gravity
-%   3.7. Particle Integration
-%   3.8. Chemistry
-%   3.9. Cooling
-%   3.10. Radiation: Homogeneous
-%   3.11  Radiation: ray-tracing
- %  3.12  Radiation: FLD
-%   3.13. Star formation & Feedback
-%   3.14. Timestep constraints
+%   3.4   Hydro: Finite difference (ZEUS)
+%   3.5   MHD: Dedner
+%   3.6   MHD: CT
+%   3.7   Gravity
+%   3.8   Particle Integration
+%   3.9.  Chemistry
+%   3.10 Cooling
+%   3.11 Radiation: Homogeneous
+%   3.12 Radiation: ray-tracing
+%   3.13 Radiation: FLD
+%   3.14 Thermal conduction
+%   3.15 Star formation & Feedback
+%   3.16 Timestep constraints
+%   3.17 Analysis
 % 4. Code tests
@@
 % 5. Parallelism 
-%   5.1. Parallel strategy 
-%   5.2. Parallel performance 
+%   5.1 Parallel strategy 
+%   5.2 Parallel performance 
 % 6. Code development methodology
@@
 \begin{abstract}
-We describe the \enzo\ code, which uses a structured adaptive mesh refinement technique to provide high spatial and temporal resolution for modeling astrophysical flows.
-The code includes support for the following physics: multi-dimensional hydrodynamics, self-gravity, N-body dynamics, primordial gas chemistry, optically-thin radiative cooling, and cosmological expansion.
-Two Eulerian hydrodynamic solvers are provided.
-This paper describes the method as well as the code's performance on a wide range of test problems.
+This paper describes the open-source \enzo\ code, which uses a
+block-structured adaptive mesh refinement to provide high spatial and
+temporal resolution for modeling astrophysical fluid flows.  The code
+is Cartesian and can be run in 1, 2, and 3 dimensions, and supports a
+large variety of physics, including hydrodynamics, ideal and non-ideal
+magnetohydrodynamics, N-body dynamics (and, more broadly, self-gravity
+of fluids and particles), primordial gas chemistry, optically-thin
+radiative cooling of primordial and metal-enriched plasmas, radiation
+transport, and cosmological expansion.  In addition to explaining the
+algorithms implemented, we present solutions for a wide range of test
+problems, demonstrate the code's parallel scaling properties, and
+discuss the \enzo\ consortium's code development methodology.
 \end{abstract}
@@
 \input{numerical-muscl} % new
+%\input{numerical-mhd-dedner} % new (combined with muscl)
+\input{numerical-mhd-ct}  % new
 \input{numerical-zeus}
-\input{numerical-mhd-dedner} % new
 \input{numerical-gravity}
@@
 \input{numerical-radiation-fld} % new
+\input{numerical-conduction} % new
 \input{numerical-starformation}
@@
 
+\input{appendix}  % new
+
 \bibliographystyle{apj}

### 2013-04-22 overview.tex
updated section 2.1 (physical equations)

@@
 
-We begin this section by first writing down the entire set of physical equations which \enzo\ can solve, and then briefly describe the numerical algorithms that we use to solve them.
-This section is intended both to place all of the equations in a single place as well as to provide a brief introduction to the numerics of the code.
-Detailed descriptions of the individual components are then provided in the following section.
+We begin this section by first writing down the complete set of
+physical equations that be solved by \enzo, and then briefly describe
+the numerical algorithms that we use to solve these equations.  This
+section is intended to be an overview of the \enzo\ code's
+capabilities: thus, we gather all of the equations solved into a
+single place and provide a brief and high-level introduction to the
+numerics of the code.  Detailed descriptions of the individual
+components are then provided in Section~\ref{sec.methods}.
 
@@
 
-The Eulerian equations of ideal magnetohydrodynamics including gravity in a coordinate systems comoving with the cosmological expansion are given by
+The Eulerian equations of ideal magnetohydrodynamics including gravity
+in a coordinate systems comoving with the cosmological expansion are
+given by:
 
@@
 %
-Here $E$, $\rho$, $\vecv$, and $\vecB$ are the comoving total fluid energy, comoving baryonic density, peculiar velocity, and comoving magnetic field strength, respectively.
-$\myvec{I}$ is the identity matrix.
+In these equations, $E$, $\rho$, $\vecv$, and $\vecB$ are the comoving
+total fluid energy, comoving baryonic density, peculiar velocity, and
+comoving magnetic field strength, respectively.  $\myvec{I}$ is the
+identity matrix, and $a$ is the cosmological expansion parameter
+(which is discussed in more detail below).
 The first equation represents conservation of mass, the second, conservation of momentum, and the third, conservation of total fluid (kinetic plus thermal) energy.
@@
 The equation of state is shown here for an ideal gas with a ratio of specific heats $\gamma$.
-Equation (\ref{eq:potential}) includes contributions to the gravitational potential $\phi$ from the total mass density $\rho_{\rm total}$.
-Although we write the equations including both MHD and comoving coordinates, the code can operate both in the pure hydrodynamic limit, and without cosmological expansion.
+Equation (\ref{eq:potential}) includes contributions to the
+gravitational potential $\phi$ from the total mass density $\rho_{\rm
+  total}$.  Although we write the equations including both MHD and
+comoving coordinates, the code can operate both in the pure
+hydrodynamic limit ($\vecB = 0$) and without cosmological expansion
+($a = 1$, $\dot{a} = 0$).
 
@@
 %
-Here, $\rho_0$ is the mean (baryonic and dark) comoving mass density and $p_0$ is the comoving background pressure contribution, and $\Lambda_c$ is the cosmological constant.
+Here, $\rho_0$ is the mean comoving mass density (including both baryonic and dark matter), $p_0$ is the comoving background pressure contribution, and $\Lambda_c$ is the cosmological constant.
 %$\gamma$ is the ratio of specific heats, and $\Lambda$ is the cosmological constant.
@@
 
-It is clear that the evolution equation are equivalent to the fixed coordinate version in the limit in which $a = 1$ and $\dot{a} = 0$.
+It is clear that the evolution equations are equivalent to the fixed coordinate version in the limit in which $a = 1$ and $\dot{a} = 0$.
 We include the cosmological terms with the understanding that we are not restricting ourselves to cosmological applications.
@@
 
-Any collisionless component (e.g. dark matter or stars) are modeled by particles which are governed by Newton's equations in comoving coordinates:
+Any collisionless component (e.g. dark matter, stars) is modeled by particles which are governed by Newton's equations in comoving coordinates:
 %
@@
 %
-The particles also make a contribution to the potential through their contribution to Poisson's equations.
+The particles also make a contribution to the potential through their contribution to Poisson's equation.
 
@@
 \end{equation}
-where $k_{ij}$ are the rate coefficients for the two-body reactions and are usually functions of only the gas species (we will specifically note the cases where we either include three-body reactions or have density-dependent rates).
-The $\Gamma^{ph}_j$ are destruction/creation rates due to photoionizations and/or photodissociations.
-Currently the species we can follow include H, H$^+$, He, He$^+$, He$^++$, and $e^-$, as well as additional options that add H$^-$, H$_2$ H$_2^+$ and  HD, D, D$^+$.
+where $k_{ij}$ are the rate coefficients for the two-body reactions
+and are usually functions of only the gas species (we will
+specifically note the cases where we either include three-body
+reactions or have density-dependent rates).  The $\Gamma^{ph}_j$ are
+destruction/creation rates due to photoionizations and/or
+photodissociations.  Currently the species we can follow include H,
+H$^+$, He, He$^+$, He$^{++}$, and $e^-$, as well as additional options
+that add H$^-$, H$_2$, H$_2^+$, and  HD, D, and D$^+$.
 
@@
 
-The code can include either a homogeneous radiation background (with possible local self-shielding corrections), or evolve an inhomogeneous radiation field either by directly solving the radiative transfer equation along rays, or by solving a set of moment equations derived from the radiative transfer equation.
+The code can include either a homogeneous radiation background (with possible local self-shielding corrections), or evolve an inhomogeneous radiation field either by directly solving the radiative transfer equation along rays or by solving a set of moment equations derived from the radiative transfer equation.
 
@@
 \end{eqnarray}
-where $E_r$ is a grey radiation energy density, $e_c$ is the
+where $E_r$ is a grey radiation energy density and $e_c$ is the
 internal energy correction due to photo-heating and chemical cooling.
@@
 In addition, $D$ in (\ref{eq:fld_radiation}) is the {\em Larsen
-square-root flux-limiter} (see \cite{Morel2000}), $\kappa$ is the total
+square-root flux-limiter} \citep[see][]{Morel2000}, $\kappa$ is the total
 opacity, $\eta$ is the field of radiation sources, $\Gamma$ provides the
-radiation induced photo-heating and, $\Lambda$ the chemical
+radiation induced photo-heating, and $\Lambda$ the chemical
 cooling.
@@
 Where $\kappa_{sp}$ is the Spitzer conduction coefficient
-\citep{1962pfig.book.....S}.  Saturation of the heat flux in
+\citep{1962pfig.book.....S} and T is the baryon temperature (with
+fluids explicitly assumed to be single-temperature).  Saturation of the heat flux in
 high-temperature, low-density regimes (such as the intracluster medium
@@
 \begin{equation}
-\fcond= -\kappa_{sp} \hat{b} (\hat{b} \cdot \nabla T)
+\fcond= -\kappa_{sp} \hat{\textbf{b}} (\hat{\textbf{b}} \cdot \nabla T)
 \end{equation}
 
-Where $\hat{b}$ is the unit vector denoting the direction of the
+Where $\hat{\textbf{b}}$ is the unit vector denoting the direction of the
 magnetic field.  In both the isotropic and anisotropic limits, we
 assume that conductivity can be suppressed relative to the ideal
-Spitzer conductivity, and express this as a constant multiplier.
+Spitzer conductivity, and express this as a multiplicative constant.
 

### 2013-04-23 overview.tex
Finished editing section 2.

@@
 \subsection{Physical Equations}
+\label{sec.equations}
 
@@
 \subsection{Overview of Numerical Methods}
+\label{sec.method_overview}
 
-In this section, we briefly describe the numerical methods which are used to solve the equations outlined above.
-We sweep through in the same order as will be used in section~\ref{sec.methods} so that there is a one-to-one correspondence between each of the following overviews and the complete description provided in that section.
-The goal here is to introduce the reader to the basic idea of the methods without drowning in detail.
+In this section, we briefly describe the numerical methods which are
+used to solve the equations outlined in Section~\ref{sec.equations}.  We
+proceed through the numerical methods in the same order as will be used in
+section~\ref{sec.methods} so that there is a one-to-one correspondence
+between each of the following overviews and the complete description
+provided in that section.  The goal here is to introduce the reader to
+the basic principles of the methods without drowning in detail.
 
@@
 
-The primary purpose of the \enzo\ code is its Adaptive Mesh Refinement
+The primary purpose of the \enzo\ code is its Adaptive Mesh Refinement (AMR)
 capability, which allows it to reach extremely large dynamical ranges
@@
 
-The grid hierarchy begins with the root grid which covers the entire
+The grid hierarchy begins with the root grid, which covers the entire
 domain of interest with a coarse, uniform, Cartesian grid. Then, as
@@
 finer grids and `above' for coarser grids).  We restrict the ratio
-between cell sizes to be an integer, typically 2-4, and refer to a
-level as all the grids with the same cell size.  In order to keep
+between cell sizes to be an integer, typically 2 or 4, and refer to a
+level as all grids with the same cell size.  In order to keep
 things as simple as possible, the edges of subgrids must coincide with
@@
 advance all the grids on that level by a timestep $\Delta t_1$
-($\Delta t_1 \leq \Delta t_0$) which is the minimum of all the allowed
+($\Delta t_1 \leq \Delta t_0$), which is the minimum of all the allowed
 timesteps for those grids.  If there are more levels, we repeat this
@@
 on that level and below.  This is done by applying the grid refinement
-criterion to the grids on that level, flagging zones which require
-extra grids.  This criterion depends on the physical problem being
+criteria to the grids on that level, flagging zones that require
+extra grids.  These criteria depend on the physical problem being
 simulated.  We have implemented a number of options, including shock
@@
 
-Four different (magneto)-hydrodynamic methods are implemented in \enzo: (i) the
-HD-only piecewise parabolic method (PPM) developed by~\citet{1984JCoPh..54..174C}
-and extended to cosmology by~\citet{1995CoPhC..89..149B};
-(ii) the MUSCL-like Gudonov scheme \citep{MUSCL} with or without magnetic fields
-and Dedner-based divergence cleansing and described in more detail in 
-\citet{WangAbelZhang08} and \citet{WangAbel09};
-(iii) a constrained transport staggered MHD scheme \citet{Collins10}, and
-(iv) the second-order finite-difference hydrodynamics method described 
-in \zeus~\citep{Stone92a,Stone92b}.
+Four different (magneto)-hydrodynamic methods are implemented in
+\enzo: (i) the hydrodynamic-only piecewise parabolic method (PPM)
+developed by~\citet{1984JCoPh..54..174C} and extended to cosmology
+by~\citet{1995CoPhC..89..149B}; (ii) the MUSCL-like Gudonov scheme
+\citep{MUSCL} with or without magnetic fields and Dedner-based
+divergence cleaning, described in more detail in
+\citet{WangAbelZhang08} and \citet{WangAbel09}; (iii) a constrained
+transport staggered MHD scheme \citep{Collins10}, and (iv) the
+second-order finite-difference hydrodynamics method described in
+\zeus~\citep{Stone92a,Stone92b}.
 
-We begin with the direct-Eulerian PPM implementation.  
-This is an explicit, higher order-accurate version of
-Godunov's method for ideal gas dynamics with third order-accurate piecewise parabolic
-monotonic interpolation and a nonlinear Riemann solver for shock
-capturing.  It advances the hydrodynamic equations in the following
-steps: 
+We begin with the direct-Eulerian PPM implementation.  This is an
+explicit, higher-order-accurate version of Godunov's method for ideal
+gas dynamics with third order-accurate piecewise parabolic monotonic
+interpolation and a nonlinear Riemann solver for shock capturing.  It
+advances the hydrodynamic equations in the following steps:
 %\begin{enumerate}
@@
 % \item 
-(ii) Compute interface states by averaging the parabola for over the domain of dependence for each interface
+(ii) Compute interface states by averaging the parabola over the domain of dependence for each interface
 % \item 
@@
 
-It does an excellent job of capturing strong shocks in a few cells; Multidimensional schemes are built up by directional
-splitting and produce a method that is formally second order-accurate
-in space and time which explicitly conserves mass, linear momentum, and energy.  A variety of
-reconstruction methods and Riemann solvers have been implemented.
+This PPM implementation does an excellent job of capturing strong
+shocks in a few cells.  Multidimensional schemes are built up by
+directional splitting and produce a method that is formally second
+order-accurate in space and time and which explicitly conserves mass,
+linear momentum, and energy.  A variety of reconstruction methods and
+Riemann solvers have been implemented.
 
-As described in \citet{Bryan95}, we modify the method for use in hypersonic flows when the thermal energy $e$ is extremely small compared to the total energy $E$.
-This is a problem because in total energy method, the temperature is computed by subtracting one large number from another (i.e.
-the kinetic energy from the total energy), a situation which is can generate large numerical inaccuracies.
-We address this situation by also solving the thermal energy equation and using $e$ from this equation when we expect the error to be large.
+As described in \citet{Bryan95}, we modify the method for use in
+hypersonic flows when the thermal energy $e$ is extremely small
+compared to the total energy $E$.  This is a problem because in the
+total energy method, the temperature is computed by subtracting one
+large number from another (i.e. the kinetic energy from the total
+energy), a situation which can generate large numerical inaccuracies.
+We address this situation by also solving the thermal energy equation
+and using $e$ from this equation when we expect the error to be large.
 
@@
 
-The last hydrodynamics method we briefly describe is the finite-different algorithm used in the ZEUS, as given in more detai in \citet{Stone92a}.
-For this reason, we call this the ZEUS method, although the code is entirely independent of the ZEUS code, and only the hydrodynamical algorithm of \zeus\ is implemented in \enzo -- the MHD and RHD schemes are not.
+The last hydrodynamics method we briefly describe is the
+finite-difference algorithm used in the \zeus\ code, as given in more
+detail in \citet{Stone92a}.  For this reason, we call this the \zeus\
+method, although the code is entirely independent of the \zeus\ code,
+and only the hydrodynamical algorithm of \zeus\ is implemented in
+\enzo -- the MHD and RHD schemes are not.
 
-The ZEUS method uses a staggered mesh such that the velocities are face-centered, while the density and thermal energy quantities are cell-centered.
-It splits the solution up into two steps, the so-called source step, in which the momentum and energy values are updated to reflect the pressure and gravity forces, as well as the effect of an artificial viscosity required for stability.
-The transport step accounts for the advection of conserved quantities (mass, momentum and energy) across the grid.
+The ZEUS method uses a staggered mesh such that the velocities are
+face-centered, while the density and internal energy quantities are
+cell-centered.  It splits the solution up into two steps, the
+so-called source step, in which the momentum and energy values are
+updated to reflect the pressure and gravity forces, as well as the
+effect of an artificial viscosity required for stability.  The second
+step (known as the transport step) accounts for the advection of
+conserved quantities (mass, momentum and energy) across the grid.
 
@@
 
-The current implementation of self-gravity in \enzo uses a Fast Fourier Technique \citep{Hockney88} to solve Poisson's equation on the top (level 0) grid.
-The advantage of this is that it naturally allows both periodic and isolated boundary conditions for the gravity, choices which are very common in astrophysics and cosmology.
-On subgrids, we interpolate the boundary conditions from that grid's parent (either the level 0 grid or some other subgrid).
-The Poisson equation is then solved using a multigrid technique.
-In \enzo\, self-gravity is optional.
-There are also a number of options for static gravitational fields.
+The current implementation of self-gravity in \enzo\ uses a Fast
+Fourier Technique \citep{Hockney88} to solve Poisson's equation on the
+top (level 0) grid.  The advantage of using this method is that it
+naturally allows both periodic and isolated boundary conditions for
+the gravity, choices which are very common in astrophysics and
+cosmology.  On subgrids, we interpolate the boundary conditions from
+that grid's parent (either the level 0 grid or some other subgrid).
+The Poisson equation is then solved using a multigrid technique.  In
+\enzo, self-gravity is optional.  Aside from calculating the
+gravitational potential from the baryon fields and particles existing
+in the simulation, there are also a number of options for static gravitational fields.
 
@@
 
-Collisionless matter (dark matter, stars, etc.) is modeled with particles which interact with the baryons only via gravity.
-These particles are advanced in a single time step using a drift-kick-drift algorithm to provide second-order accuracy even in the presence of changing timesteps.
-Since the particles follow the collapse of structure, they are not adaptively refined.
-Nor are there duplicate sets of particles for each level; instead, each particle is associated with the highest refined
-level available at its position in space, and moved as the hierarchy is rebuilt.  Thus, a particle has the same timestep and feels the same gravitational force as the grid at that level.
+Collisionless matter (dark matter, stars, etc.) is modeled with
+particles that interact with the baryons only via gravity.  These
+particles are advanced in a single time step using a drift-kick-drift
+algorithm to provide second-order accuracy even in the presence of
+changing timesteps.  Since the particles follow the collapse of
+structure, they are not adaptively refined.  Nor are there duplicate
+sets of particles for each level; instead, each particle is associated
+with the highest refined level available at its position in space, and is
+moved as the hierarchy is rebuilt.  Thus, a particle has the same
+timestep and feels the same gravitational force as a grid at that refinement
+level.
 
@@
 
-In addition to this, \enzo\ includes the capability of following up to 12 chemical species using a non equilibrium-solver.
-The species can be turned on in sets, with the simplest model including just H, H$^+$, He, He$^+$, He$^++$, and $e^-$, and more complete models adding first species important for gas-phase molecular hydrogen formation: $H^-$, H$_2$ and H$_2^+$, and then $HD$ formation: $HD$, $D$, $D+$.
-The cooling and heating due to these species is also included (see the next section).
-The solution of the rate equations is carried out using a time-backwards differencing which insures stability; to maintain accuracy we sub-cycle the rate equations with a shorter timestep such that the electron and neutral fractions do not change by more than 10\%.
+\enzo\ includes the capability of following up to
+12 chemical species using a non-equilibrium solver.  The species can
+be turned on in sets, with the simplest model including just H, H$^+$,
+He, He$^+$, He$^++$, and $e^-$, and more complete models adding first
+species important for gas-phase molecular hydrogen formation: $H^-$,
+H$_2$ and H$_2^+$, and then $HD$ formation: $HD$, $D$, $D+$.  The
+cooling and heating due to these species is also included (see the
+next section).  The solution of the rate equations is carried out
+using a time-backwards differencing which insures stability; to
+maintain accuracy we sub-cycle the rate equations with a shorter
+timestep such that the electron and neutral fractions do not change by
+more than 10\%.
 
@@
 
-Enzo may operate in a number of modes with regard to radiative cooling and heating.
-In the simplest mode, with the multi-species flag turned off (i.e.
-set to zero), the cooling rate is computed from a simple temperature-dependent cooling rate, taken from \citet{SW87}.
-If chemistry is turned on, then the code can include cooling from all species of H and He (including H$_2$) -- this primordial cooling is also treated using a backwards-time differencing scheme with sub-cycling to maintain accuracy.
-Finally, we have also recently added metal cooling based on a set of multi-dimensional lookup tables computed with the \texttt{Cloudy} code \citep{1998PASP..110..761F} as described in \citet{2008MNRAS.385.1443S} and \citet{2011ApJ...731....6S}.
-Finally, we note that the cooling and heating is (almost always) treated in the optically-thin limit, although the code can also follow radiative transfer in a limited set of energy bins (see below).
+Enzo may operate in a number of modes with regard to radiative cooling
+and heating.  In the simplest mode, with the multi-species flag turned
+off (i.e. set to zero), the cooling rate is computed from a simple
+temperature-dependent cooling rate, taken from \citet{SW87}.  If
+chemistry is turned on, then the code can include cooling from all
+species of Hydrogen and Helium (including H$_2$) -- this primordial cooling is
+also treated using a backwards-time differencing scheme with
+sub-cycling to maintain accuracy.  Finally, we have also recently
+added metal cooling based on a set of multi-dimensional lookup tables
+computed with the \texttt{Cloudy} code \citep{1998PASP..110..761F} as
+described in \citet{2008MNRAS.385.1443S} and
+\citet{2011ApJ...731....6S}.  Finally, we note that the cooling and
+heating is (almost always) treated in the optically-thin limit,
+although the code can also follow radiative transfer in a limited set
+of energy bins (see below).
 
@@
 
-The chemical networks and heating rates can be affected by external radiation fields, and the code includes a number of pre-set radiation backgrounds that are uniform in space but can very in time.
-These are generally based on the redshift-dependent rates given in \citet{1996ApJ...461...20H} and \citet{2012ApJ...746..125H}, but can also include a uniform H$_2$ photo-dissociating background, either constant in time or varying as in \citet{WiseAbel05}.
+The chemical networks and heating rates described in the previous
+section can be affected by external radiation fields, and the code
+includes a number of pre-set radiation backgrounds that are uniform in
+space but can very in time.  These are generally based on the
+redshift-dependent rates given in \citet{1996ApJ...461...20H} and
+\citet{2012ApJ...746..125H}, but can also include a uniform H$_2$
+photo-dissociating background that is either constant in time or
+varying as in \citet{WiseAbel05}.
 
@@
 
-The code includes a photon-conserving radiative transfer algorithm that is based on an adaptive ray-tracing method that utilizes the HEALPix pixilation of the sphere \citep{AbelWandelt02}.
-Photons are integrated outwards from sources using an adaptive time-stepping scheme that preserves accuracy in ionization fronts even in the optically-thin limit.
-This has been integrated with the chemistry and cooling network to provide ionization and heating rates on a cell-by-cell basis.
-The method is described in more detail in our main methodology section but is fully detailed and tested in \citet{Wise11_Moray}.
+The code includes a photon-conserving radiative transfer algorithm
+that is based on an adaptive ray-tracing method that utilizes the
+HEALPix pixilation of the sphere \citep{AbelWandelt02}.  Photons are
+integrated outwards from sources using an adaptive time-stepping
+scheme that preserves accuracy in ionization fronts even in the
+optically-thin limit.  This has been integrated with the chemistry and
+cooling network to provide ionization and heating rates on a
+cell-by-cell basis.  The method is described in more detail in our
+main methodology section but is fully detailed and tested in
+\citet{Wise11_Moray}.
 
@@
 
-The other option for radiative transfer is a moment-based method which adds an addition field describing the energy-density of radiation.
-This field is evolved using the flux-limited diffusion method which transitions smoothly between streaming (optically thin) and opaque limits and is coupled to the hydrogen ionization network.
-The resulting set of linear equations is solved using the parallel HYPRE framework.
-Full details for the enzo implementation can be found in \citet{ReynoldsHayesPaschosNorman2009}.
+A second option for radiative transfer is a moment-based method that
+adds an additional field tracking the radiant energy density
+This field is evolved using the flux-limited diffusion method, which
+transitions smoothly between streaming (optically thin) and opaque
+limits and is coupled to the hydrogen ionization network.  The
+resulting set of linear equations is solved using the parallel HYPRE
+framework.  Full details for the enzo implementation can be found in
+\citet{ReynoldsHayesPaschosNorman2009}.
 
@@
 
-Heat conduction, both isotropic and anisotropic, can be included using an sub-cycled, operator-split method.
-The heat fluxes are computed with simple second-order accurate finite differences and stability is ensured by restricting the time step, and using flux-limiters.
-This is described in more detail below.
+Heat conduction, both isotropic and anisotropic, can be included using
+an sub-cycled, operator-split method.  The heat fluxes are computed
+with simple second-order accurate finite differences and stability is
+ensured by restricting the time step, and using flux-limiters where
+appropriate.
 
@@
 
-A simple heuristic method is used to model the formation of stars and their feedback of metals and energy into the gas.
-This method is based on \citep{CO1992} and involves identify plausible cites of star formation based on a set of criteria (dense gas with a short cooling time, which is both collapsing and unstable), and computing a star formation rate based on the Schmidt-Kennicutt relation \citep{K89}.
-The relating gas is converted into a star particle over a few dynamical times, and over this timespan metals and thermal energy are injected into the single, local cell occupied by the star particle.
+A family of simple heuristic methods are used to model the formation
+of stars and their feedback of metals and energy into the gas.  These
+methods are based on the work of \citet{CO1992}, and involve the
+identification of plausible sites of star formation based on a set of
+criteria (for example, dense gas with a short cooling time, which is
+both collapsing and unstable).  The local star formation rate is
+computed using a range of methods, such as a density-dependent method
+based on the Schmidt-Kennicutt relation \citep{K89}.  The affected gas
+is converted into a star particle over a few dynamical times, and
+metals and thermal energy are injected into the region surrounding the
+star particle.
 
@@
 
-All grids on a given level are advanced with the same timestep.
-To determine this timestep, we calculate, for each cell and for each physical process (except for the chemistry step, which is sub-cycled), the minimum $\Delta t$ allowed for stability.
-Then the minimum of all timesteps is taken and the level is advanced with this step.
+All grids on a given level are advanced with the same timestep.  To
+determine this timestep, we calculate, for each cell and for each
+physical process (except for the chemistry step, which is sub-cycled),
+the minimum $\Delta t$ allowed for stability.  Then the minimum of all
+timesteps is taken and the level is advanced with this step.
 

### 2013-04-24 development.tex
Edits to introduction (also adding section labels to some other files).
Tweaking development section.

--- 2013-01-18 conclusions.tex
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Conclusions}
+\section{Code development methodology}
+\label{sec.development}
 
-\begin{enumerate}
-\item near-future development plans
-\item far-future development plans
-\item clear areas of improvement
-\end{enumerate}
+Over time, \enzo's development has followed a trajectory toward
+openness.  Started as the graduate research project of Greg Bryan at
+the University of Illinois, it has been subsequently been stewarded by
+the Laboratory for Computational Astrophysics (LCA) at the University
+of California at San Diego, and has transitioned to a distributed,
+completely open, and community-driven project.  Initially, \enzo\ was
+versioned using a series of ``snapshots'' of the code base, usually
+hand-created by the developing individuals.  These were distributed to
+collaborators and colleagues, but the central ``trunk'' of development
+was updated primarily by a single person: while patches and technology
+were accepted from external developers, the relatively small number of
+individuals using the code resulted in a strong centralization of
+development.
+
+As the stewardship of the code passed to the LCA, the code was
+released first to ``friendly users'' and then as an open source
+release.  However, while the code was made available with
+documentation, technology developed in the broader community of users
+was typically not re-integrated.  This led to a wide dispersal of
+development, largely independent, by individuals who downloaded and
+used the version of the code developed by the LCA.
+
+Following the first public, open source release of \enzo, the code was
+migrated to using the Subversion version control system.  This is a
+centralized version control system, and the ``stable'' \enzo\ source
+code was made globally readable following the \enzo\ 1.5. release.
+Access to the primary development tree required a password and login
+for each user, and providing upstream changes either required this
+password and the granting of write access or a sequence of patches and
+manually-created diffs (much like the original development system).
+The technical friction of manually contributing patches and
+modifications, combined with Subversion's difficulty with tracking
+merges, resulted in further fragmentation of the code base.
+
+The main development streams of \enzo\ were eventually merged using
+the distributed version control system (DVCS) Mercurial
+(\url{http://mercurial.selenic.com/}) into a branch of the code known
+as \texttt{week-of-code}, so named after the in-person development
+sprint at which it was created.  The fundamental, and transformative,
+distinction between the previous \textit{centralized} version control
+system and mercurial's \textit{distributed} version control system is
+the elimination of gatekeepers.  While there still exists a canonical,
+central location where stable and development versions of \enzo\ can
+be obtained, changesets and versions can be exchanged between peers
+without the intervention of designated gatekeepers.  This has the
+direct effect of enabling local development to be versioned and its
+provenance ensured, while still retaining the ability to benefit from
+``upstream'' development.  An important, even crucial, side effect is
+that the technology used for local versioning provides mechanisms for
+easily submitting locally-developed modifications to the community
+source location.  Mercurial internally represents all changes as nodes
+in a directed, acyclic graph (DAG), which results in the natural
+ability to more consistently and easily manage merging development
+streams.
+
+Currently, \enzo\ is developed using the hosted source control
+platform Bitbucket (\url{http://bitbucket.org/}) at
+\url{http://bitbucket.org/enzo/}.  There are two mailing lists, one
+for usage-focused questions and discussion, and another for
+development discussion.  Both of these lists are open and publicly
+archived.  Bitbucket provides mechanisms for inspecting source,
+hosting branches and forks of the primary source, and for code review.
+All proposed source code changes for \enzo\ are subjected to a peer
+review process, where experienced developers read, inspect, test, and
+provide feedback on source code changes.  All developers, including
+long-time \enzo\ contributors and developers, are subject to this
+process before their code is included in the primary \enzo\
+repository.  By using a remote, hosted system, \enzo\ is now
+\textit{completely} open to contributions from the community.
+Individuals, who may or may not consider themselves \enzo\ developers,
+are free to ``fork'' the \enzo\ code base, develop changes (signed
+with their own name), and submit them for review and inclusion.  In
+contrast to the centralized, gatekeeper-focused technology used
+previously, this enables anyone to contribute changes to be evaluated
+for inclusion in the \enzo\ codebase.
+
+While peer review is able to catch many bugs and problems with source
+code changes, \enzo\ is also subject to ``answer'' testing.  We have
+created a set of parameter files and problem types that exercise the
+underlying machinery of \enzo.  These ``test problems'' have
+affiliated ``tests,'' which consist of scripts that use \texttt{yt}
+\citep{2011ApJS..192....9T} to produce results such as mass
+distribution, projections, profiles and so on.  The testing
+infrastructure then evaluates whether the variation in the new results
+compared to a ``gold standard'' of results has exceeded an acceptable
+threshold, typically set to roughly single precision.  Optionally, for
+those test problems that are deemed unsafe to change to any precision,
+the tests also produce hashes of the outputs; these hashes will only
+remain unchanged in the event of bitwise identicality between results.
+The results of the gold standard are versioned and stored in Amazon
+S3, enabling remote testing to proceed.  While the testing process --
+building, running, analyzing and comparing -- is not yet automated
+against incoming pull requests, we hope to deploy that functionality
+in the future.  The primary challenge is that of compute time; the
+tests are organized into multiple categories, including by the
+expected run time, but the full suite of tests can take several days
+to run on multiple processors.

### 2013-04-24 development.tex
ah, second thoughts!

@@
 expected run time, but the full suite of tests can take several days
-to run on multiple processors.
+to run.

### 2013-04-24 conclusions.tex
Edits to introduction (also adding section labels to some other files).
Tweaked conclusions.

@@
 \section{Conclusions}
+\label{sec.conclusions}
 
@@
 and subgrid processes.  In addition, we have described the \enzo\ code
-development process, have shown the outputs of some representative
+development process, have shown the outputs of a set of representative
 test problems, and have provided information about \enzo's performance
@@
 and figures for this paper are open source and are available at the
-\enzo\ website, http://enzo-project.org.  Furthermore, the yt toolkit,
+\enzo\ website, \url{http://enzo-project.org}.  Furthermore, the yt toolkit,
 which is designed to analyze \enzo\ data (as well as data from a wide
 variety of other simulation tools), can be found at its website,
-http://yt-project.org.  Both of these codes have active user and
+\url{http://yt-project.org}.  Both of these codes have active user and
 developer communities, extensive documentation and user suppport, and
@@
 interact with each other and with multiple grids, and include sink,
-source, and particle creating, splitting, and merging functionality.
+source, and particle creation, destruction, splitting, and merging functionality.
 \item A new HYPRE-based AMR gravity solver that is faster, more

### 2013-04-24 acknowledgments.tex
added acknowledgments file, put some basic stuff in there.

\acknowledgments

Development of \enzo\ has been supported by the National Science Foundation, the National Aeronautics and Space Administration, the Department of Energy, the University of Illinois and National Center for Supercomputing Applications, UC San Diego and the San Diego Supercomputing Center, Columbia University, Georgia Tech, Michigan State University, ...  \red{We all need to add to the acknowledgments!}

### 2013-04-26 numerical-amr.tex
Edited AMR section.

@@
 
-The back-bone of the SAMR idea is the patch, or grid, which we take to be rectilinear to simplify booking (in practice, the lower level routines all assume Cartesian coordinates).
+The back-bone of the SAMR idea is the patch, or grid, which we take to be rectilinear to simplify bookkeeping (in practice, the lower-level routines such as the hydro solvers all assume Cartesian coordinates).
 The single root grid covers the entire computational domain and plays the role of the root node in the grid hierarchy.
 Subgrids are placed such that they cover (hopefully small) sub-volumes of the root grid at higher resolution.
-This structure forms a tree may be extended to arbitrary depths by adding additional grids to refine regions in the subgrids.
+This structure forms a tree that may be extended to arbitrary depths by adding additional grids to refine regions in the subgrids.
 We use the following notation to describe grid relationships.
-A grid's parent is that coarse grid which completely contains it.
-The root grid has no parent.
-A child grid is the opposite (note that a grid may have many children but only one parent), while sibling grids are those which have the same resolution, but not necessarily the same parent.
+A grid's \textit{parent} is the coarse grid that completely contains it.
+The \textit{root grid} is the most coarsely-resolved grid, and has no parent.
+A \textit{child grid} is grid completely contained by its parent grid (note that a grid may have many children but only one parent), while sibling grids are those that have the same resolution, but not necessarily the same parent.
 We also refer to levels of the hierarchy, where the root grid is labeled level zero and all grids with the same resolution are on the
@@
  \item The refinement factor $r$, or the ratio of the coarse cell width to the fine cell width, must be an integer.
-In practice, we also always use $r=2$, since tests have shown this provides the best performance.
+In practice, \enzo\ almost always uses $r=2$, since tests have shown this typically provides the best performance.
  \item Each subgrid must begin and end on the boundary of a coarse cell.  This means that the number of cells in a patch will be a multiple of the refinement factor.
@@
 The AMR grid patches are the primary data structure in \enzo.
-Each patch is treated as an individual object which can contain both field variables and particle data.
+Each patch is treated as an individual object that can contain both field variables and particle data.
 Individual grids are organized into a  dynamic, distributed hierarchy of mesh patches.
@@
 Grids are partitioned into a core of \emph{real zones} and a surrounding layer of \emph{ghost zones}, as shown in Figure~\ref{fig:amr_hierarchy}.
-The real zones store field values and ghost zones are used to temporarily store values which have been obtained directly from neighboring grids or interpolated from a parent grid.
-These zones are necessary to accommodate the computational stencil of the hydrodynamics solvers (sections~\ref{sec.hydro.ppm} and~\ref{sec.hydro.zeus}) and the gravity solver (section~\ref{sec.gravity}).
+The real zones store field values and ghost zones are used to temporarily store values that have been obtained directly from neighboring grids or interpolated from a parent grid.
+These zones are necessary to accommodate the computational stencil of the (magneto)hydrodynamics solvers (Sections~\ref{sec.hydro.ppm} and~\ref{sec.hydro.zeus}) and the gravity solver (section~\ref{sec.gravity}).
 The PPM and Zeus hydro solvers require 3 layers of ghost zones,  and the gravity solver requires 6 (although only for the density and potential fields).
@@
 The finest level is then advanced, using as many steps as it takes to reach the level immediately above.
-The finest level is then synced to the level above it, which then proceeds forewards one more step.
+The finest level is then synched to the level above it, which then proceeds foreward in time one more step.
 This is shown graphically in Figure~\ref{fig:wcycle}
@@
 
-The basic control algorithm looks very much like any used in a single
-grid scheme, and can be written schematically as:
+The basic control algorithm looks very much like any used in a grid-based code of a single resolution, and can be written schematically as:
 \begin{verbatim}
@@
 \begin{enumerate}
-  \item{\em reflecting.} In this case, the boundary is like a mirror, with the solution on the boundary reflecting the solution in the computational domain ($q(-x) = q(+x)$) but with the velocity normal to the boundary reversed: $v_x(-x) = -v_x(+x)$.
-  \item{\em outflow.} True outflow conditions are extremely difficult (REF?).  We approximate them by duplicating the solution at the edge of the computational domain: $q(-x) = q(0)$.
-  \item{\em inflow.} In this case, the boundary values are fixed by a pre-determined function ($q(-x) = q0(x,t)$).
-The code provides a simple way to set inflow values which are constant in both time and space over an entire face, or hooks are provided for special cases where the boundary must be time-varying.
-  \item{\em periodic.} The boundary solution is obtained from the other side of the grid: $q(-x) = q(x_{\rm{max}}-x)$.
+  \item{\em Reflecting.} In this case, the boundary behaves like a mirror, with the solution on the boundary reflecting the solution in the computational domain ($q(-x) = q(+x)$) but with the velocity normal to the boundary reversed: $v_x(-x) = -v_x(+x)$.
+  \item{\em Outflow.} True outflow conditions are extremely difficult \red{(REF?)}.  We approximate them by duplicating the solution at the edge of the computational domain: $q(-x) = q(0)$.
+  \item{\em Inflow.} In this case, the boundary values are fixed by a pre-determined function ($q(-x) = q0(x,t)$).
+The code provides a simple way to set inflow values that are constant in both time and space over an entire face, or hooks are provided for special cases where the boundary must be time-varying.
+  \item{\em Periodic.} The boundary solution is obtained from the other side of the grid: $q(-x) = q(x_{\rm{max}}-x)$.
 \end{enumerate}
@@
 In one dimension, such functions are well known; examples include
-donor cell, Van-Leer and piecewise parabolic interpolation, which are
+donor cell, Van Leer and piecewise parabolic interpolation, which are
 first, second and third-order accurate respectively
@@
   however, at higher order, this would not necessarily satisfy our
-  constraints on not introducing new extrema.}  In the Appendix, we
+  constraint regarding not introducing new extrema in interpolated quantities.}  In Appendix~\ref{app:interpolation}, we
 explicitely write down functions in one, two and three dimensions, but
@@
 to this value), and second, by treating the cell-diagonals as
-one-dimensional Van-Leer type functions.  The reason to focus on the
+one-dimensional Van Leer-type functions.  The reason to focus on the
 cell diagonals is that for a linear function, the extrema will occur at the cell 
@@
 a least-squares fit under the given constraints, however, this is
-computationally too expensive.  Therefore, in the appendix we describe
-a heuristic approach which satisfies four constaints analytically and
+computationally too expensive.  Therefore, in Appendix~\ref{app:interpolation} we describe
+a heuristic approach that satisfies four constaints analytically and
 then adjusts the interpolation parameters in such a way as to satisfy the fifth.
-While the result is not necessarily optimal, it does seem to perform
+While the result is not necessarily optimal, it does appear to perform
 adequately in our tests.
@@
 \citep[e.g.,][]{Stone92a}.  The interpolation step, described in
-section~\ref{sec:interpolation}, must correctly interpolate with the
+Section~\ref{sec:interpolation}, must correctly interpolate with the
 selected centering.
@@
 more ``solutions'' (i.e.
-values of density, velocity, etc.) which represent the same volume of space.
-This is one important aspect of the SAMR method -- some physical regions are represented by grids at various resolutions.
+values of density, velocity, etc.) that represent the same volume of space.
+This is one important aspect of the SAMR method -- some physical regions are represented by grids at multiple resolutions.
 To
@@
 \end{equation}
-where $r$ is the refinement factor, $d$ the dimensionality, and the sum is over all fine cells which ``cover" the coarse cell.
+where $r$ is the refinement factor, $d$ the dimensionality, and the sum is over all fine cells that ``cover" the coarse cell.
 
@@
 
-One of the big advantages a flux-conservative methods is, of course, their ability to maintain conserved-quantities to machine precision.
+One of the significant advantages of flux-conservative methods is, of course, their ability to maintain conserved quantities to machine precision.
 Unfortunately, the projection step upsets this flux conservation since, at the boundary of a refined region, the coarse and fine cells on either side of the border are updated with different estimates of the flux across the boundary.
 In one case this comes from the coarse grid computation, and in the other from the fine grid solution.
-Conservation can be restored by correcting all coarse cells which lie at the boundary of a fine region
+Conservation can be restored by correcting all coarse cells that lie at the boundary of a fine region
 with the difference between the coarse and fine estimates of the flux across the boundary:
+
 \begin{equation}
@@
 \end{equation}
+
 where $F$ is the flux of quantity $q_i$ in the $i$-direction through the cell interface.
-We use $\tilde{q}$ to indicate the uncorrected quantity on the coarse grid and the sum is over the $r^{d-1}$ fine cell interfaces in the perpendicular dimensions which share a face with the coarse cell being corrected.
-This corrected is carried out for all coarse cells which share a face with a refined cell (and are not themselves covered with more refined cells).
+We use $\tilde{q}$ to indicate the uncorrected quantity on the coarse grid and the sum is over the $r^{d-1}$ fine cell interfaces in the perpendicular dimensions that share a face with the coarse cell being corrected.
+This correction is carried out for all coarse cells that share a face with a refined cell (and are not themselves covered with more refined cells).
 
@@
 Since the addition of more highly refined grids is adaptive, the conditions for refinement must be specified.
-Up to five different criteria can be specified, and these refinement criteria are discussed in more detail in the following section (section~\ref{sec:refinement_criteria}).
-Once all cells in a given grid that require additional refinement have been flagged, rectangular boundaries are determined which minimally encompass the flagged cells.
-This is done using an iterative machine-vesion technique as described in (REF).
+Up to seven different criteria can be specified for a given simulation, and these refinement criteria are discussed in more detail in the following section (Section~\ref{sec:refinement_criteria}).
+Once all cells in a given grid that require additional refinement have been flagged, rectangular boundaries are determined that minimally encompass the flagged cells.
+This is done using an iterative machine-vesion technique as described in \red{(REF)}.
 
-We first find the smallest region (or ``proto-subgrid") which covers all of the refined cells on a grid.
+We first find the smallest region (or ``proto-subgrid") that covers all of the refined cells on a grid.
 We compute an efficiency $\epsilon$ for this grid, defined as the ratio of flagged cells to total cells.
@@
 First, traditional grid-based hydro method can be simply added with only minor modifications.
-Second, it is easier to accommodate larger stencils (as in PPM, which requires three boundary zones), and finally it can be more efficient for some machines since there are more predictable and linear memory accesses.
+Second, it is easier to accommodate larger stencils (as in PPM, which requires three boundary zones), and finally it can be more efficient for some machines since there are more predictable and linear memory accesses, which more efficiently use a CPU's cache.
 
@@
 Finally, we turn to the refinement criteria themselves.
-Although generalized refinement criteria are available based on estimators for the error, such as the Richardson method (REF), we have focused generally on methods which use physical estimators on the assumption that the simulator has a specific quantity to the indicates refinement.
+Although generalized refinement criteria are available based on estimators for the error, such as the Richardson method \red{(REF)}, we have focused generally on methods that use physical estimators based on the assumption that the simulator has a specific quantity or quantities that indicate the need for refinement.
 The flagging methods currently built into \enzo\ are:
@@
 
-\item{\em Slope.}  If the normalized slope $(q_{i+1} - q_{i-1})/ (2 q_i)$ is larger than a specified value (typically 0.3), then the cell is flagged.
+\item{\em Slope.}  If the normalized slope $(q_{i+1} - q_{i-1})/ (2 q_i)$ of baryon quantities between two adjacent cells is larger than a specified value (typically 0.3), then the cell is flagged.
 
-\item{\em Baryon mass.}  This refinement criterion is identified to mimic a Lagrangian method in that it tries to keep a fixed mass resolution.
+\item{\em Baryon mass.}  This refinement criterion is designed to mimic a Lagrangian method in that it tries to keep a fixed mass resolution.
 In this method, a cell is refined if the baryonic mass in the cell ($M_g = \rho \Delta_x^d$) is larger than a specified value:
@@
 \end{equation}
-where $\rho_{\rm flag}$ is the equivalent density on the root grid required for refinement, $\Delta x_{\rm root}$ is the root grid cell spacing, $r$ is the refinement factor, $l$ is the level and $\epsilon_l$ is a parameter which can be used to make the refinement faster ($\epsilon_l < 0$) or slower than pure Lagrangian refinement ($\epsilon_l > 0$).
+where $\rho_{\rm flag}$ is the equivalent density on the root grid required for refinement, $\Delta x_{\rm root}$ is the root grid cell spacing, $r$ is the refinement factor, $l$ is the level and $\epsilon_l$ is a parameter that can be used to make the refinement more aggressive ($\epsilon_l < 0$) or less aggressive than pure Lagrangian refinement ($\epsilon_l > 0$) -- in other words, super-Lagrangian or sub-Lagrangian in refinement.
 
-\item{\em Shocks.}  In this method, we identify and refine shocks with the following criteria:
+\item{\em Shocks.}  In this method, we identify and refine cells that contain shocks using the following criteria:
 \begin{eqnarray}
@@
 
-\item{\em Particle (dark matter) mass.}  This refinement criterion is precisely the same as the baryon overdensity criterion except it uses the grided (CIC) particle density.
+\item{\em Particle mass.}  This refinement criterion is precisely the same as the baryon overdensity criterion except that it uses the gridded (CIC) particle density.
+In general, this criteria is used in cosmological simulations, and thus refines on dark matter and stellar density.
 
-\item{\em Jeans Length.}  Following \cite{Truelove98}, the code can be forced to refine the Jeans length by a fixed number of cells, or, in more specifically, whenever the following criterion is not met:
+\item{\em Jeans length.}  Following \cite{Truelove98}, the code can be forced to refine the Jeans length by a fixed number of cells, or, in more specifically, whenever the following criterion is not met:
 \begin{equation}
@@
 \item{\em Cooling time.}  Often in cosmological simulations, cooling is a rapid process and we do not necessarily want to be limited by the cooling time.
-However, in some cases it is desirable to resolve the cooling process, in which case we flag on cells which have a cooling time shorter than their sound crossing time.
+However, in some cases it is desirable to resolve the cooling process, in which case we flag cells that have a cooling time shorter than their sound crossing time.
 The cooling time is given by $t_{\rm cool} = 1.5 kT / n \Lambda(T)$ and the sound crossing time is $t_{\rm cross} = \Delta x / c_s$, where $c_s$ is the sound speed.
 
-\item{\em Must-refine particles}  In order to allow for refinement in moving regions, it is possible to generate ``must-refine" particles which for local refinement over the 8 nearest cells to a specific level.
-The particles feel on gravitational forces and their trajectories are followed as for any other collisionless particle.
+\item{\em Must-refine particles.}  In order to allow for refinement in moving regions, it is possible to generate ``must-refine" particles that ensure local refinement of the 8 nearest cells to a specific level.
+The particles feel gravitational forces and their trajectories are followed as for any other collisionless particle.
 This feature can be used to ensure that a specific Lagrangian region stays refined to a specified level of resolution.
 
-\item{\em Shear} As described in \citet{Kritsuk06}, turbulence calculations can be successfully modeled with AMR if the shear is used as a refinement parameter; in particular, refinement occurs when the finite difference version of the following inequality holds:
+\item{\em Shear.} As described in \citet{Kritsuk06}, turbulence calculations can be successfully modeled with AMR if the shear is used as a refinement parameter; in particular, refinement occurs when the finite difference version of the following inequality holds:
 \begin{equation}
@@
 
-\item{\em Optical depth.} This refinement criterion ensures that cells on the highest level of refinement have an \ion{H}{1} optical depth lower than unity.
+\item{\em Optical depth.} This refinement criterion, typically used for simulations using radiation transport, ensures that cells on the highest level of refinement have an \ion{H}{1} optical depth lower than unity.
 The optical depth of a cell is estimated to be:
@@
 
-\item{\em ``Must refine region''}.
-This refinement criterion is used to ensure that all cells in a given subvolume of the simulation are refined to {\it at least} some minimum level of refinement.
+\item{\em ``Must refine region.''}  This refinement criterion is used to ensure that all cells in a given subvolume of the simulation are refined to {\it at least} some minimum level of refinement.
 The subvolume is user-specified, and may evolve over time (e.g., to follow a structure of interest that is traveling through the simulation volume).
 
-\item{\em Metallicity.}  This refinement criterion ensures that all cells with a metallicity (defined as $Z = \equiv \rho_{Z} / \rho_{b} / 0.022$, where $\rho_{Z}$ is the density of metals in the cell, $\rho_{b}$ is the total baryon density (including metals), and 0.022 is the solar metal fraction) above a user-defined value is refined to {\it at least} some minimum level of refinement.
+\item{\em Metallicity.}  This refinement criterion ensures that all cells with a metallicity (defined as $Z \equiv \rho_{Z} / \rho_{b} / 0.022$, where $\rho_{Z}$ is the density of metals in the cell, $\rho_{b}$ is the total baryon density (including metals), and 0.022 is the solar mass fraction of metals) above a user-defined value is refined to {\it at least} some minimum level of refinement.
 
@@
 
-Finally, it is also possible to specify a specific rectangular region which is statically refined at a certain level.
+Finally, it is also possible to specify a rectangular region that is statically refined to a certain level.
 These static refined regions are in place throughout the simulation regardless of the grid quantities and can be used for testing purposes, or for multiply-zoomed initial conditions (for example in cosmology simulations) that would not otherwise be refined.

### 2013-04-26 numerical-amr.tex
one additional edit.

@@
 The real zones store field values and ghost zones are used to temporarily store values that have been obtained directly from neighboring grids or interpolated from a parent grid.
-These zones are necessary to accommodate the computational stencil of the (magneto)hydrodynamics solvers (Sections~\ref{sec.hydro.ppm} and~\ref{sec.hydro.zeus}) and the gravity solver (section~\ref{sec.gravity}).
+These zones are necessary to accommodate the computational stencil of the (magneto)hydrodynamics solvers (Sections~\ref{sec.hydro.ppm} and~\ref{sec.hydro.zeus}) and the gravity solver (Section~\ref{sec.gravity}).
 The PPM and Zeus hydro solvers require 3 layers of ghost zones,  and the gravity solver requires 6 (although only for the density and potential fields).

### 2013-04-26 numerical-mhd-ct.tex
Editing MHD-CT section.  Collins writes like a madman.

@@
 \def\Divb{\ensuremath{\nabla \cdot \Bvec}}
+
 Physically, the divergence of the magnetic field, \Divb, is
@@
 \end{eqnarray}
-The Constrained Transport (CT) method \citep{Evans88, Balsara99} of MHD evolution employs this same vector
-property to evolve the magnetic field in a manner that preserves $\Divb=0$ to
-the ability of the computer.  The electric field is computed from the fluxes
-from the Riemann solver.  The curl of that electric field is then used to update
-the  magnetic field.   The advantage of this method is that it preserves \Divb\
-to machine precision.  The drawback is twofold: increased algorithm  complexity,
-as will be described in this section; and since only the update of the magnetic
-field is divergence free, any monopoles created by other numerical sources
+
+The Constrained Transport (CT) method \citep{Evans88, Balsara99} for
+magnetohydrodynamic (MHD) evolution employs this same vector property
+to evolve the magnetic field in a manner that preserves $\Divb=0$ to
+to limits of machine precision.  The electric field is computed using
+the fluxes from the Riemann solver.  The curl of that electric field
+is then used to update the magnetic field.  The advantage of this
+method is that it preserves \Divb\ to machine precision.  The drawback
+is twofold: increased algorithm complexity, as will be described in
+this section; and since only the update of the magnetic field is
+divergence free, any monopoles created by other numerical sources
 persist.
 
-Enzo MHD-CT has been described in horrid detail in \citet{Collins10}, so the
-description here will be an overview only.  The base Godunov method, described
-in \citet{Li08a}, uses spatially and temporally second order reconstruction
-(both MUSCL-Hancock and Piecewise Linear Method), a selection of Riemann solvers
-including HLLC and HLLD \citep{Mignone07}.  The CT methods are the first order
-method described by \citet{Balsara99} and the second order methods described in
-\citet{Gardiner05}.  The AMR machinery is described by \citet{Balsara01} and
-\citet{Collins10}.
+Enzo MHD-CT has been described in great detail in \citet{Collins10},
+and we direct the interested reader to that reference.  The base
+Godunov method, described in \citet{Li08a}, uses spatially and
+temporally second order reconstruction (both MUSCL-Hancock and
+Piecewise Linear Method), a selection of Riemann solvers including
+HLLC and HLLD \citep{Mignone07}.  The constrainted transport methods
+are the first-order method described by \citet{Balsara99} and the
+second order methods described in \citet{Gardiner05}.  The AMR
+machinery is described by \citet{Balsara01} and \citet{Collins10}.
 
-The increased complexity comes in the form of area-averaged face and
-length-averaged edge centered variables, while the rest of Enzo employs
-predominantly volume-averaged cell centered variables.  The magnetic field is
-represented by both a face centered field, \Bf, and a cell centered field, \Bc.
-The electric field is edge centered.
-The magnetic field is updated in four steps:  first, the
-Riemann problem is solved in the traditional manner, using the cell centered
-field; second, an edge-centered
-electric field is computed using the fluxes from the Riemann solver;  third, the curl
-of that electric field is used to update the face centered field; finally,
-the cell centered magnetic field is updated with an average of the face centered
-field.
+The increased complexity of the constrained transport scheme comes in
+the form of area-averaged face and length-averaged edge-centered
+variables, while the rest of Enzo employs predominantly
+volume-averaged cell centered variables.  The magnetic field is
+represented by both a face centered field, B$_f$, and a cell centered
+field, B$_c$.  The electric field is edge centered.  The magnetic
+field is updated in four steps: first, the Riemann problem is solved
+in the traditional manner, using the cell centered field; second, an
+edge-centered electric field is computed using the fluxes from the
+Riemann solver; third, the curl of that electric field is used to
+update the face centered field; finally, the cell centered magnetic
+field is updated with an average of the face centered field.
 
-Divergence free AMR is slightly more complex than the AMR employed elsewhere in
+Divergence-free AMR is somewhat more complex than the AMR employed elsewhere in
 Enzo.  First, the interpolation must be constrained to be divergence free.
@@
 circumvent this last complexity, the electric field is projected from fine grids
-to parent grids, rather than the magnetic field, which is then used to re-update
-the parent magnetic field.  This is described in detail in \citet{Balsara99,
-Collins10}
+to parent grids (rather than the magnetic field), and is then used to re-update
+the parent magnetic field.  This is described in detail in
+\citet{Balsara99} and \citet{Collins10}.
 
@@
 evolution \citep{Xu11,Xu12}, synthetic observations of radio halos
-\citep{Skillman13},  primordial star formation \citep{Xu08}, 
-present day star formation \citep{Collins11,Collins12}.
+\citep{Skillman13},  primordial star formation \citep{Xu08}, and
+present-day star formation \citep{Collins11,Collins12}.

### 2013-04-26 numerical-particle.tex
edited n-body dynamics section

@@
 
-\enzo\ uses a particle-mesh N-body method to calculate the 
-dynamics of collisionless systems \citep{Hockney88}.  This method 
-follows trajectories of a representative sample of individual particles 
-and is much more efficient than a direct solution of the Boltzmann 
-equation in essentially all astrophysical situations. 
-As described earlier, the gravitational potential is computed by solving the elliptic 
-Poisson's equation (Eq.~\ref{eq:potential}), and differencing the potential
-to find accelerations, which are then interpolated back to particles.  This
-acceleration is time-centered (because the underlying gravitating mass
-field is approximately time-centered), and so we have accelerations $\myvec{a}^{n+1/2}$
-for each particle.  These are used to update the particle positions and velocities
-starting from $\myvec{x}^n$ and $\myvec{v}^n$ using a standard drift-kick-drift technique:
+\enzo\ uses a particle-mesh N-body method to calculate the dynamics of
+collisionless systems \citep{Hockney88}.  This method follows
+trajectories of a representative sample of individual particles that
+sample the phase space of the dark matter distribution, and is much
+more efficient than a direct solution of the Boltzmann equation in
+essentially all astrophysical situations for the levels of accuracy
+that are required for simulations of structure formation.  As
+described earlier, the gravitational potential is computed by solving
+the elliptic Poisson's equation (Eq.~\ref{eq:potential}), and
+differencing the potential to find accelerations, which are then
+interpolated back to particles.  This acceleration is time-centered
+(because the underlying gravitating mass field is approximately
+time-centered), and so we have accelerations $\myvec{a}^{n+1/2}$ for
+each particle.  These are used to update the particle positions and
+velocities starting from $\myvec{x}^n$ and $\myvec{v}^n$ using a
+standard drift-kick-drift technique:
+
 \begin{eqnarray}
@@
 
-Particles are stored in the most highly refined grid patch at the point in space where they exist, and particles which move out of a subgrid patch are sent to the grid patch covering the adjacent volume with the finest spatial resolution, which may be of the same spatial resolution, coarser, or finer than the grid patch that the particles are moved from.
-This takes place in a communication process at the end of each timestep on a level.
+Particles are stored in the most highly refined grid patch at the
+point in space where they exist, and particles that move out of a
+subgrid patch are sent to the grid patch covering the adjacent volume
+with the finest spatial resolution, which may be of the same spatial
+resolution, coarser, or finer than the grid patch that the particles
+are moved from.  This takes place in a communication process at the
+end of each timestep on a level.

### 2013-04-26 numerical-radiation-homogeneous.tex
edited uniform radiation background section (3.10)

@@
 
-\enzo\ supports the use of a set of spatially uniform (but possibly time-varying) radiation fields that can interface with the chemistry and cooling/heating routines described elsewhere.
-Many of these use fits to the H, He and He$^+$ ionizing and photo-ionization heating rates that are of the form
+\enzo\ supports the use of a set of spatially uniform (but possibly
+time-varying) radiation fields that can interface with the chemistry
+and cooling/heating routines described elsewhere.  Many of these use
+fits to the H, He and He$^+$ ionizing and photo-ionization heating
+rates that are of the form
+
 \begin{equation}
-\rm{rate} = k_0 * (1+z)^{\alpha} * \exp{\left( \frac{\beta *(z-z_0)^2 }{1 + \gamma * (z+z_1)^2} \right)}
+\rm{rate} = k_0  (1+z)^{\alpha}  \exp{\left( \frac{\beta (z-z_0)^2 }{1 + \gamma  (z+z_1)^2} \right)}
 \label{eq:homo_rad_template}
 \end{equation}
-where the constant coefficients $\alpha$, $\beta$, $\gamma$, $z_0$ and $z_1$ are fits from the literature.
-For radiation field types 1-3, we give the coefficients for the photoionization (and photo-heating) rates of H, He, and He$^+$ in Table~\ref{table:homo_coefs}.
-Types 1 and 2 are based on \citet{1996ApJ...461...20H} with two different intrinsic quasar spectra slopes, while Type 3 is from \citet{2012ApJ...746..125H}, modified to match the normalized field found in \citet{Kirkman05}.
-Type 4 is the same as Type 3 but also adds X-ray Compton heating from \citet{MadauEfstathiou99}, using equations (4) and (11) of that paper.
 
-Homogeneous radiation field types 5 and 6 start with a spectral shape which is then integrated against the appropriate H, He, He$^+$ cross-sections to compute the ionization and photo-ioniziation heating rates (we use 400 bins logarithmically specced from 0.74 eV to $7.24 \times 10^9$ eV).
-In particular, Type 5 has a hard, featureless quasar-like power law spectrum $f_{\nu} = f_{\rm HI} \nu^{\alpha_0}$, where $\alpha_0 = 1.5$ by default and the spectrum is normalized at the HI ionization edge.
-Type 6 has the same spectrum by is attenuated by a column density of $10^{22}$ cm$^{-3}$ neutral hydrogen.
+where the constant coefficients $\alpha$, $\beta$, $\gamma$, $z_0$ and
+$z_1$ are fits from the literature.  For radiation field types 1-3
+(numbered as they appear in the \enzo\ source code and simulation
+parameter values), we
+give the coefficients for the photoionization (and photo-heating)
+rates of H, He, and He$^+$ in Table~\ref{table:homo_coefs}.  Radiation
+background types 1
+and 2 are based on \citet{1996ApJ...461...20H} with two different
+intrinsic quasar spectra slopes, while Type 3 is from
+\citet{2012ApJ...746..125H}, modified to match the normalized field
+found in \citet{Kirkman05}.  Type 4 is the same as Type 3 but also
+adds X-ray Compton heating from \citet{MadauEfstathiou99}, using
+equations (4) and (11) of that paper.
 
-Types 8, 9, and 14 have only photo-dissociating Lyman-Werner flux, with Type 8 being constant and Type 9 using the redshift-dependent results of \citet{TrentiStiavelli09}.
-Type 14 uses a fit from \citet{WiseAbel05} for the range $6 < z < 50$, constant for $z<6$, and zero for $z > 50$.
-Other types are undefined or unused.
+Homogeneous radiation field types 5 and 6 start with a spectral shape
+that is then integrated against the appropriate H, He, and He$^+$
+cross-sections to compute the ionization and photo-ioniziation heating
+rates (we use 400 bins logarithmically spaced from 0.74 eV to $7.24
+\times 10^9$ eV).  In particular, Type 5 has a hard, featureless
+quasar-like power law spectrum $f_{\nu} = f_{\rm HI} \nu^{\alpha_0}$,
+where $\alpha_0 = 1.5$ by default and the spectrum is normalized at
+the HI ionization edge.  Type 6 has the same spectrum by is attenuated
+by a column density of $10^{22}$ cm$^{-3}$ neutral hydrogen.
+
+Types 8, 9, and 14 have only photo-dissociating Lyman-Werner flux,
+with Type 8 being constant and Type 9 using the redshift-dependent
+results of \citet{TrentiStiavelli09}.  Type 14 uses a fit from
+\citet{WiseAbel05} for the range $6 < z < 50$, constant for $z<6$, and
+zero for $z > 50$.  Other types are undefined or unused.
 

### 2013-04-28 numerical-starformation.tex
Edited star formation section.  This needs to be looked at by somebody else, please!

@@
 \label{sec.ov.star}
+
+\red{( (BWO) I am a little worried that this section is too
+  long. The editorial hatchet may be called for, but I think somebody
+  else (who didn't write the section in the first place) needs to
+  think it through. )}
 
@@
 
+\red{(BWO: can this whole paragraph go?)}
 While the physics discussed in other sections of this paper is crucial
@@
 ensemble stellar population (although not all; see
-Section~\ref{sec:starform_pop3}).  % BWO: I use 'phenomenological'
-right above this, but could probably % go for 'parametric' or
-'heuristic' if people object to my usage.
+Section~\ref{sec:starform_pop3}).
 
@@
 ways: At a specified time interval, all grid cells that are at the
-highest local level of refinement (i.e., have no child cells) are
+highest local level of refinement (i.e., that have no child cells) are
 examined to see if they meet a set of criteria for star formation.
@@
 $f_{\rm Zb}$.  There is a user-defined minimum dynamical time
-$T_{\rm dyn,min}$ which is observationally motivated and affects the
+$T_{\rm dyn,min}$, which is observationally motivated and affects the
 feedback rates (see below).  The particle is placed in the center of
@@
 and is then treated in the same manner as the dark matter particles.
-An amount of baryon gas corresponding to the new particle's mass is
+An amount of gas corresponding to the new particle's mass is
 removed from the cell.
@@
 
-The feedback of energy and metals into the baryon gas is similar to
-the Kravtsov feedback, with some important differences.
+% The feedback of energy and metals into the baryon gas is similar to
+% the Kravtsov feedback, with some important differences.
 
@@
 Algorithmically, the star formation events in Kravtsov's code are
-assumed to occur once every global time step $\Delta t_0 \leq 10^7$.
+assumed to occur once every global time step (with the constraint $\Delta t_0 \leq
+10^7$ years).
 In cells where star formation is determined to occur (i.e. $\rho_{\rm
@@
 folded into a density threshold for star formation.  In calculations
-with much higher resolution, however ( $\la$~few pc), individual
+with much higher resolution, however (on the order of a few pc), individual
 molecular clouds can be resolved, thus rendering these approximations
-irrelevant.  To that end, \citet{2012ApJ...749...36K} have implemented
+invalid.  To that end, \citet{2012ApJ...749...36K} have implemented
 a star formation algorithm that is specifically geared to
@@
 where a given halo may only form one star of primordial composition.
-To accommodate this, Enzo contains a star formation algorithm that
-forms Population III stars directly \citep{2007ApJ...659L..87A,
+To accommodate this, \enzo\ contains a star formation algorithm that
+forms individual Population III stars directly \citep{2007ApJ...659L..87A,
   2008ApJ...685...40W, 2012MNRAS.427..311W}.  Using criteria similar
-to~\citet{CO1992}, a star particle forms when a cell has all of the
-following criteria:
+to~\citet{CO1992}, a star particle forms when a cell meets all of the
+following conditions:
 
 \begin{enumerate}
-\item a baryon overdensity of $5 \times 10^5$ (corresponding to
-  roughly $10^3$ g cm$^{-3}$ at $z=10$),
+\item A baryon overdensity of $5 \times 10^5$ (corresponding to a
+  hydrogen number density of 
+  roughly $10^3$ cm$^{-3}$ at $z=10$),
 
-\item a converging gas flow ($\nabla \cdot \vec{v} < 0$), and
+\item A converging gas flow ($\nabla \cdot \vec{v} < 0$), and
 
-\item a molecular hydrogen fraction f$_{\rm H2} > 5 \times 10^{-4}$,
+\item A molecular hydrogen mass fraction f$_{\rm H2} > 5 \times 10^{-4}$,
 \end{enumerate}
 
-conditions typical of a collapsing metal-free molecular cloud roughly
+These are comparable to the conditions typical of a collapsing metal-free molecular cloud roughly
 10 million years before the birth of a Pop III main-sequence star.  If
@@
 
-Feedback from the Population III stars comes in multiple forms.
-Radiative feedback using the MORAY radiation transport algorithm
+Feedback from Population III stars created using this algorithm comes in multiple forms.
+Radiative feedback using the \moray\ radiation transport algorithm
 \citep{Wise11_Moray} is available, using the mass-dependent hydrogen

### 2013-04-28 numerical-starformation.tex
Minor star formation tweak.
ok, one more tweak.

@@
 
-\red{(BWO: can this whole paragraph go?)}
-While the physics discussed in other sections of this paper is crucial
-to the study of cosmological structure formation, most observations of
-galaxies in the infrared, optical, and ultraviolet wavelengths are of
-stellar populations.  Furthermore, stars eject energy and metal-
-enriched gas throughout their lives, drastically modifying their own
-environments' cooling properties.  Broadly speaking, the formation of
-galaxies and galaxy clusters cannot be completely modeled without
-including the feedback of energy and metals from stellar populations.
-In particular, it is thought that feedback is crucial for suppressing
-the large numbers of dwarf-like galaxies that CDM theories predict
-\citep{1991ApJ...381...14L,1991ApJ...379...52W}.  An early burst of
-star formation could remove a large fraction of cold gas from such
-systems \citep{1978MNRAS.183..341W,1991ApJ...367...45C}.  Also, the
-unexpectedly low luminosities of small clusters and groups (relative
-to the richest clusters) is commonly explained through feedback
-\citep{1991ApJ...383..104K}.  Energy ejected during the formation of
-the cluster ellipticals lowers the central density and hence the X-ray
-luminosity of such clusters \citep{1997ApJ...484L..21C}.  For all of
-these reasons, it is critical to include star formation and the
-feedback of energy and metals into the interstellar and intergalactic
-media in cosmological simulations.
+% \red{(BWO: can this whole paragraph go?)}
+% While the physics discussed in other sections of this paper is crucial
+% to the study of cosmological structure formation, most observations of
+% galaxies in the infrared, optical, and ultraviolet wavelengths are of
+% stellar populations.  Furthermore, stars eject energy and metal-
+% enriched gas throughout their lives, drastically modifying their own
+% environments' cooling properties.  Broadly speaking, the formation of
+% galaxies and galaxy clusters cannot be completely modeled without
+% including the feedback of energy and metals from stellar populations.
+% In particular, it is thought that feedback is crucial for suppressing
+% the large numbers of dwarf-like galaxies that CDM theories predict
+% \citep{1991ApJ...381...14L,1991ApJ...379...52W}.  An early burst of
+% star formation could remove a large fraction of cold gas from such
+% systems \citep{1978MNRAS.183..341W,1991ApJ...367...45C}.  Also, the
+% unexpectedly low luminosities of small clusters and groups (relative
+% to the richest clusters) is commonly explained through feedback
+% \citep{1991ApJ...383..104K}.  Energy ejected during the formation of
+% the cluster ellipticals lowers the central density and hence the X-ray
+% luminosity of such clusters \citep{1997ApJ...484L..21C}.  For all of
+% these reasons, it is critical to include star formation and the
+% feedback of energy and metals into the interstellar and intergalactic
+% media in cosmological simulations.
 
@@
 Feedback from Population III stars created using this algorithm comes in multiple forms.
-Radiative feedback using the \moray\ radiation transport algorithm
+Radiative feedback using the Moray radiation transport algorithm
 \citep{Wise11_Moray} is available, using the mass-dependent hydrogen

### 2013-04-28 numerical-timestep.tex
ok, one more tweak.
editing timestep section (3.15)

@@
 \label{sec.timestepping}
-
-%\dcc{Made this its own section (not really part of the AMR, it's more
-%  general than that.)  Added stuff on harmonic average for 3d. See
-%  note below on acceleration timestep criterion being crap. Moved
-%  w-cycle to the previous section}
 
@@
  basis by finding the largest timestep such that multiple criteria are
-satisfied on each level.  The timestep criteria used by Enzo are 
+satisfied on each level.  The timestep criteria used by \enzo\ are 
 (showing the one-dimensional case for clarity):
@@
 
-In equations~\ref{eqn:dthydro}-\ref{eqn:dtcond}, the $\min ( \ldots
-)_L$ formalism means that this value is calculated for all cells on a
+In Equations~\ref{eqn:dthydro}-\ref{eqn:dtcond}, the $\min ( \ldots
+)_L$ formalism means that this value is calculated for all cells or
+particles on a
 given level L and the minimum overall value is taken as the timestep.
@@
 
-Equation ~\ref{eqn:dthydro} is valid for one dimension.  For 2 or 3
+Equation~\ref{eqn:dthydro} is valid for one dimension.  For 2 or 3
 dimensions, it was shown by \cite{Godunov1959}  that using the
@@
 t_x$, $\Delta t_y$, and $\Delta t_z$ be the analogues of
-eqn. \ref{eqn:dthydro} along the $x,y$ and $z$ axes, 
+Equation \ref{eqn:dthydro} along the $x,y$ and $z$ axes, 
 \begin{equation}
@@
 
-For all other criteria, multiple dimensions are accounted for by
+For all other criteria except for Equation~\ref{eqn:dtMHD}, multiple dimensions are accounted for by
 repeating the one dimensional criterion along each axis, and taking
@@
   hydro}$, with an identical range of values.
-Equation~\ref{eqn:dtexpand} limits the timestep such that the
-expansion parameter a can only change by a fractional amount of
-$f_{\rm exp} = \Delta a/a$, where $f_{\rm exp}$ is a user-defined
-parameter and has typical values of $f_{\rm exp} = 0.01-0.02$.  This
-is required for the stability of the PPM algorithm in comoving
-coordinates, and typically limits the timestep only at high redshifts
-when densities are relatively homogeneous.
 
-Equation~\ref{eqn:dtaccel} and~\ref{eqn:dtrad} are supplementary to equation~\ref{eqn:dthydro} in that they
+Equations~\ref{eqn:dtaccel} and~\ref{eqn:dtrad} are supplementary to Equation~\ref{eqn:dthydro} in that they
 take into account the possibility of large accelerations due to either
-gravity (Eqn.~\ref{eqn:dtaccel} or radiation pressure
-(Eqn.~\ref{eqn:dtrad}), which may cause numerical 
+gravity (Equation~\ref{eqn:dtaccel} or radiation pressure
+(Equation~\ref{eqn:dtrad}), which may cause numerical 
 instabilities by violating the Courant condition.  In Equation~\ref{eqn:dtaccel}, $\vec{g}$ is the
@@
 
-Where the sum calculates the energy deposited in a cell during the
+where the sum calculates the energy deposited in a cell during the
 previous timestep due to \textit{all photon packets} that crossed that
-cell, with $\hat{r_i}$ being a unit vector that accounting for the
+cell, with $\hat{r_i}$ being a unit vector that accounts for the
 packet direction.
@@
 value must be $0 < \kappa_{\rm cond}  < 0.5$, and is exactly 0.5 for the
-implemention in Enzo.
+implemention in \enzo.
 From a practical
 perspective, it is useful to note that, unlike other timestep criteria
-discussed above (which effectively scale as $\Delta x / \sqrt{T}$,
+discussed above (which effectively scale as $\Delta x / \sqrt{T}$),
 the timestep criterion due to thermal conduction scales as $\Delta x^2
@@
 equation, $a$ and $\dot{a}$ refer to the scale factor of the universe
-and its rate of change, respectively.  This
-criteria is necessary because the expansion of the universe and its
-first derivative with respect to time both appear in the equations of cosmological (magneto)hydrodynamics and particle motion,
-and some limit is required in order to model the change of energy and
-momentum due to expansion with a reasonable level of accuracy.  In
-reality, this is only a significant limitation at very high redshift;
-at later times, other timestep criteria are much more important.
+and its rate of change, respectively.  This criteria is necessary
+because the expansion of the universe and its first derivative with
+respect to time both appear in the equations of cosmological
+(magneto)hydrodynamics and particle motion, and some limit is required
+for the stability of the PPM algorithm in comoving coordinates.  This
+criterion typically limits the timestep (moreso than other criteria)
+only at high redshifts, when densities are relatively homogeneous.
 

### 2013-04-28 code-tests.tex
Updated test suite section (4.1).  The only remaining section to be edited is 4.2, which is the actual tests themselves.

@@
 a suite of test problems that can be compared to previous versions of
-the code, to ensure that \enzo\ is, e.g., running correctly on a new
+the code, to ensure that \enzo\ is running correctly on a new
 computer, with a new compiler, or after substantial changes have been
@@
 
-\enzo\ is capable of simulating a large variety of problems types, with
-all but a small few requiring only a parameter file as sole input.
-The notable exception is the cosmology simulation, which takes as
-input initial conditions created by other codes.  These test problems
-span a wide range in complexity.  At one end of this spectrum are
-simple problems that utilize only a single component of \enzo\ and for
-which analytic solutions exist for comparison with the simulation
+\enzo\ is capable of simulating a large variety of problems types,
+with all but a few of these types requiring only a parameter file as an
+input.  The notable exception is the cosmology simulation, which takes
+as input initial conditions created by other codes.  The suite of test
+problems spans a wide range in complexity.  At one end of this spectrum
+are simple problems that utilize only a single component of \enzo\ and
+for which analytic solutions exist for comparison with the simulation
 results.  At the opposite end are problems that exercise a large
@@
 them to serve as a vehicle for verifying that the code's behavior
-remains stable over time.
+remains stable over time and after modifcation of the codebase.
 
@@
 responsible for taking as input from the user a series of keywords
-with which to select a subset of all available test problems.  The
+that are used to select a subset of all available test problems.  The
 test problems are also grouped into three suites: the quick, push, and
@@
 is mandated to be run before code changes are accepted into the main
-repository.  The full suite consists of all test problems which can be
+repository.  The full suite consists of all test problems that can be
 run with no additional input.  Approximate run times for the quick,
 push, and full suite are 15 minutes, 1 hour, and 60 hours,
-respectively.
+respectively, on a relatively new desktop computer (circa 2013).
 
 After the test problems are selected by the test runner script, they
-are run in succession either by with \enzo\ executable contained
-within that distribution or an external executable built from another
+are run in succession using either the \enzo\ executable contained
+within that distribution or an external executable built from another \enzo\
 version and specified by the user.  This allows for any version of the
@@
 against the simulation data.  After the analysis is performed, the
-results are compared against a set of gold standard results which are
+results are compared against a set of gold standard results that are
 maintained on a website and downloaded on the fly by the test runner
@@
 Section~\ref{sec.tests.problems}, we describe some of the key test
-problems which are used to verify proper behavior.  All of these test
+problems that are used to verify proper behavior.  All of these test
 problems, as well as the scripts to generate the figures from them,
-are available in the \enzo\ test suite.
+are available in the \enzo\ test suite, allowing the interested reader
+to regenerate all of the figures in Section~\ref{sec.tests.compare}
+using the \textit{'methodpaper'} suite of test problems.
 

### 2013-04-29 code-tests.tex
Oh just kidding, I forgot to edit the code comparison section (4.1.2).  Onward and upward, to the test problem section!

@@
 
-Over the course of its existence, \enzo\ has been involved in several
+Over the course of its existence, \enzo\ has been involved in numerous
 comparisons with other astrophysical codes used for self-gravitating
@@
 
-\enzo\ has been involved in several cosmological code comparisons,
+\enzo\ has been involved in multiple cosmological code comparisons,
 including the Santa Barbara Cluster code comparison project
@@
 typically superior in terms of resolving fluid features due to its
-higher-order (and artificial viscosity-free) hydrodynamics solver.
-When examining classical test cases such as the Santa Barbara
-project \citep{SantaBarbara}, \enzo\ forms galaxy clusters with very similar density,
-temperature, and entropy profiles to other grid-based codes that use
-Godunov-type hydro methods, which
-systematically differ from particle-based codes using SPH.  Similarly,
-in galaxy cluster simulations that look specifically at the properties
-of cosmological shocks \citep[e.g.][]{2011MNRAS.418..960V}, \enzo\
-produces results that are similar to other high-order grid-based
-hydrodynamics codes, and far superior behavior of \enzo\ in
-low-density regions when compared to a particle-based code.  In tests
-of the Lyman-alpha forest that include radiative cooling and a uniform
-metagalactic ultraviolet background, \enzo\ and Gadget provide results
-on metrics such as the matter power spectrum that are comparable to
-within 5\% \citep{2007MNRAS.374..196R}.
+higher-order (and artificial viscosity-free) Piecewise Parabolic
+Method hydrodynamics solver.  When examining classical test cases such
+as the Santa Barbara project \citep{SantaBarbara}, \enzo\ forms galaxy
+clusters with very similar density, temperature, and entropy profiles
+to other grid-based codes that use Godunov-type hydro methods, which
+systematically differed from particle-based codes using SPH in this
+code comparison.  Similarly, in galaxy cluster simulations that look
+specifically at the properties of cosmological shocks
+\citep[e.g.][]{2011MNRAS.418..960V}, \enzo\ produces results that are
+similar to other high-order grid-based hydrodynamics codes, and far
+superior performance of \enzo\ is observed (in terms of resolution of
+fluid features and shock detection) in low-density regions when
+compared to a particle-based code.  In tests of the Lyman-alpha forest
+that include radiative cooling and a uniform metagalactic ultraviolet
+background, \enzo\ and Gadget provide results on metrics such as the
+matter power spectrum that are comparable to within 5\%
+\citep{2007MNRAS.374..196R}.
 
 Several comparisons have been made that focus specifically on
-hydrodynamics solvers.  In particular, the work of
+hydrodynamics solvers and fluid behavior.  In particular, the work of
 \citet{2007MNRAS.380..963A} and \citet{Tasker2008} perform
 head-to-head comparisons between several grid- and particle-based
-codes for a variety of test problems (including shocked gas clouds,
+codes for a variety of fluid-centric test problems (including shocked gas clouds,
 self-gravitating, translating clouds, and Sedov-Taylor explosions),
@@
 other grid-based hydrodynamics codes involved in the comparison, and
-provide useful information on whether an AMR code such as \enzo\
-experiences challenges.  More specific comparisons, including one
+provide useful information on the sort of practical challenges that a
+user of an
+AMR code such as \enzo\ may experience.  More specific comparisons, including one
 specifically testing the linear and nonlinear growth of the
-Kelvin-Helmholz instability \citep{2012ApJS..201...18M} and one more
-broadly examining Galilean invariance in grid-based codes
-\citep{2010MNRAS.401.2463R} show that \enzo, and in particular its
+Kelvin-Helmholz instability \citep{2012ApJS..201...18M}, as well as a
+comparison that more
+broadly examines Galilean invariance in grid-based codes
+\citep{2010MNRAS.401.2463R}, show that \enzo, and in particular its
 implementation of the Piecewise Parabolic Method hydrodynamics solver,
@@
 hydrodynamics and the latter the constrained transport MHD
-implementation.  In both cases, \enzo\ performed similarly to other
+implementation of \citet{Collins10}.  In both cases, \enzo\ performed similarly to other
 grid-based codes that use Godunov-based fluid solvers, and typically
@@
 
-Three other comparisons between portions of the \enzo\ code and other
-simulation tools have been performed in recent years.  The
-flux-limited diffusion radiation transport scheme was measured against
-several test problems by \citet{IlievEtAl2009}, producing results that
-are similar to both expected results and other methods, although there
-are some minor differences (although all codes in the study exhibited
-differences from the majority of other codes on at least a subset of
-the tests).  \citet{2011ApJ...726...55T} shows the result of varied
-reaction rate coefficients for the formation of molecular hydrogen via
-the 3-body process in both the \enzo\ and Gadget-2 codes.  Similar
-trends were visible in both the \enzo\ and Gadget simulations, though
-at a nominally equivalent resolution (i.e., with particle and cell gas
-masses being comparable) \enzo\ simulations typically displayed a much
-higher level of gas structure.  This is unsurprising due to \enzo's
-higher-order hydro solver.  Finally, \citet{2012ApJ...744...52P} show
-the results of comparing \enzo\ in its non-AMR mode to a
-smoothed-particle hydrodynamics code, SNSPH, in the context of
-common-envelope binary stellar interactions.  The authors show that
-the codes display reasonable convergence properties as a function of
-simulation resolution, and also agree quite well with each other --
-however, the observed mass-loss rates do not agree particularly well
-with observations.
+Three other comparisons between the \enzo\ code and other simulation
+tools have been performed that focus on physics other than gravity and
+fluid flow.  The flux-limited diffusion radiation transport scheme was
+measured against several test problems by \citet{IlievEtAl2009},
+producing results that are similar to both expected results (i.e.,
+analytic test problems) and other methods (i.e., codes using other
+methods, written by other groups), although there are some minor
+differences (although all codes in the study exhibited differences
+from the majority of other codes on at least a subset of the tests).
+\citet{2011ApJ...726...55T} shows the results of varied reaction rate
+coefficients for the formation of molecular hydrogen via the 3-body
+process in both the \enzo\ and Gadget-2 codes.  Similar trends were
+visible in both the \enzo\ and Gadget simulations, though at a
+nominally equivalent resolution (i.e., with particle and cell gas
+masses being comparable) \enzo\ simulations typically displayed a
+substantially higher level of gas structure.  This is unsurprising due
+to \enzo's higher-order hydro solver.  Finally,
+\citet{2012ApJ...744...52P} show the results of comparing \enzo\ in
+its non-AMR mode to a smoothed-particle hydrodynamics code, SNSPH, in
+the context of common-envelope binary stellar interactions.  The
+authors show that the codes display reasonable convergence properties
+as a function of simulation resolution, and also agree quite well with
+each other -- however, the observed mass-loss rates do not agree
+particularly well with observations.
 

### 2013-04-29 code-tests.tex
test problems through sod shock tube.

@@
 
-General structure for each test problem:  Outline how the test problem is constructed (initial and boundary conditions), 
-the analytical solution, why we have in the paper (what does it break, or what flaw does it expose (not in enzo of course)),
-a plot showing how well enzo solves said test problem, and a brief description of the plot and how awesome enzo is.
+The \enzo\ test suite (described in Section~\ref{sec.tests.suite})
+contains hundreds of test problems that probe the code's behavior in a
+wide range of physical circumstances, and explicitly test each physics
+package in the \enzo\ code, both individually and in combination.  It
+is impractical to include a substantial fraction of these problems in
+a method paper; as a result, we have chosen to publish the results of
+only a small subset of particularly crucial test problems here.  If
+the interested reader desires, they can download \enzo\ and
+\texttt{yt}, run the test suite, and see the results of any other test
+problem and its comparison to an analytic solution (if available).
+
+The general structure of each test problem description is as follows.
+We will describe the construction of the test problem (including its
+initial and boundary conditions), the analytical or expected
+solutions, and motivate why we have included it in the paper.  After
+that, we will show and describe \enzo's solution to the test problem.
+
+% General structure for each test problem: Outline how the test problem
+% is constructed (initial and boundary conditions), the analytical
+% solution, why we have in the paper (what does it break, or what flaw
+% does it expose (not in enzo of course)), a plot showing how well enzo
+% solves said test problem, and a brief description of the plot and how
+% awesome enzo is.
 
@@
 
-We begin the paper test suite with the classic Sod shock tube problem \citep{Sod78}, which provides a good test of a scheme's ability to resolve a clean Riemann problem with clear separation between the three waves, in particular a rarefaction fan, a contact discontinuity and a moderate shock.
-In Figure~\ref{fig.sodshocktube}, we show the solution for three of our hydro solves, the spatially third-order PPM, as well as the two second-order Zeus and MUSCL schemes.
-We use 100 cells across the domain, which is a relatively standard choice and show solutions both with and without AMR.
-Without AMR, it is clear that the PPM scheme produces by far the cleanest solution, with all wave families crisply reproduced.
-Zeus and MUSCL produce similar results, with MUSCL doing a slightly better job on the rarefaction fan.
-In the figure, we also show the mean absolute deviation from the exact solution, as well as the maximum deviation.
-These numbers confirm the qualitative differences noted previously.
+\red{(BWO: this is missing an explanation of the ICs)}
 
-We also run the same simulation but with two levels of AMR (using a refinement factor of 2), triggered based on a normalized slope greater than 10\% in the density.
-This is a minimal level of refinement, and does not include the rarefaction fan at late times.
-The results are shown in the bottom panels of Figure~\ref{fig.sodshocktube}, and the results are much better for all three methods, with much sharper shocks and contact discontinuities, and even a better representation of the rarefaction wave, which is only refined at early times.
-Although the results are improved for all methods, the PPM still produces the best result, as shown clearly by the compute error norms, shown again in the figure.
+We begin the paper test suite with the classic Sod shock tube problem
+\citep{Sod78}, which provides a good test of a hydrodynamical solver's
+ability to resolve a clean Riemann problem with clear separation
+between the three waves.  These waves include a rarefaction fan, a
+contact discontinuity, and a moderate shock.  In
+Figure~\ref{fig.sodshocktube}, we show the density solution at the
+final time for three of our hydro solvers -- the spatially third-order
+PPM, as well as the two second-order \zeus\ and MUSCL schemes.  We use
+100 cells across the domain, (which is a relatively standard choice in
+code method papers) and show solutions both with and without adaptive
+mesh refinement.  Without AMR (in the top row of the figure), it is
+clear that the PPM scheme produces by far the cleanest solution, with
+all wave families crisply reproduced (in particular, the contact
+discontinuity and shock).  \zeus\ and MUSCL produce similar results,
+with MUSCL doing a slightly better job on the rarefaction fan.  In the
+figure, we also show the mean absolute deviation from the exact
+solution ($||E_\infty||$), as well as the maximum deviation
+($||E_1||$).  These numbers confirm the qualitative differences noted
+previously.
+
+We also run the same simulation but with two levels of AMR (using a
+refinement factor of 2), triggered based on a normalized slope greater
+than 10\% in the density.  This is a minimal level of refinement, and
+does not include the rarefaction fan at late times.  The results are
+shown in the bottom row of Figure~\ref{fig.sodshocktube}.  Using AMR,
+the results are much better for all three methods, with much sharper
+shocks and contact discontinuities and even a better representation of
+the rarefaction wave, which is only refined beyond the root grid at
+early times.  Although the results are improved for all methods, PPM
+still produces the best result, as is shown clearly by the computed
+error norms (displayed again in each of the individual panels).
 
@@
 \includegraphics[width=\textwidth]{figures/SodShockTube.eps}
-\caption{The density distribution of the classic Sod Shock Tube for three different solvers (from left to right) and with (bottom) and without (top) AMR.
-In each case 100 zones are used on the root level and the results are shown at $t=0.25$.
-All cells are plotted and color-coded by level (at the time shown, only small region surrounding the contact discontinuity and the shock are refined).
-In each panel, we show the exact solution as a solid line and the $L_1$ and $L_\infty$ error norms in the upper right.}
+\caption{The density distribution of the classic Sod Shock Tube for
+three different solvers (from left to right column) and with (bottom
+row) and without (top row) AMR.  In each case 100 zones are used on
+the root level and the results are shown at $t=0.25$.  All cells are
+plotted and color-coded by level (at the time shown, only small region
+surrounding the contact discontinuity and the shock are refined).  In
+each panel, we show the analytic solution as a solid line and the $E_1$
+and $E_\infty$ error norms in the upper right.}
 \label{fig.sodshocktube}

### 2013-04-29 code-tests.tex
edits done through 1-zone collapse test.

@@
 \label{sec.tests.doublemach}
-\red{(Brian)}
 
@@
 \citet{1984JCoPh..54..115W} \citep[and more recently
-in][]{2008ApJS..178..137S}.  In this problem (shown in
-Figure~\ref{fig.doublemach}), a shock is injected at an angle to a
-reflecting surface, and a jet appears along the reflecting surface.
-The ideal solution is self-similar, and the appearance of this
-solution is highly sensitive to numerical diffusion.  If numerical
-noise is present, a Kelvin-Helmholz instability develops along this
-jet and breaks the self-similarity.
+in][]{2008ApJS..178..137S}.  In this problem, shown in
+Figure~\ref{fig.doublemach}, a shock is injected at an angle to a
+reflecting surface (the -y boundary), and a jet appears along the
+reflecting surface.  The ideal solution is self-similar, and the
+appearance of this solution is highly sensitive to numerical
+diffusion.  If numerical noise is present, a Kelvin-Helmholz
+instability develops along this jet and breaks the self-similarity.
 
@@
 the direct Eulerian implementation of the piecewise parabolic hydro
-method, with diffusion, flattening, and shock steepening all turned
-on.
+method, with the diffusion, flattening, and shock steepening all
+turned on.
 
@@
 \includegraphics[width=0.8\textwidth]{figures/DoubleMachTest.eps}
-\caption{Double mach test.}
+\caption{Density field at the final time in the Double Mach test.  A
+Mach 10 shock is injected into the domain with a shock normal that is
+30$^\circ$ from the x-axis with a time-dependent +y boundary condition
+that mimics a shock of infinite length.  The solution is self-similar,
+with the jet and whorls at $x \simeq 2.5-3.0$ in this figure being very sensitive
+to numerical diffusion.  Our results compare favorably to the
+higher-order images from \citet{1984JCoPh..54..115W}.}
 \label{fig.doublemach}
@@
 \includegraphics[width=0.4\textwidth]{figures/sedov-zeus-slice.eps}
-\caption{Density slice from the Sedov Blast Test at $t =
-0.07$. Left-hand image shows the results from PPM ({\tt HydroMethod =
-0}), right-hand image shows the results from Zeus ({\tt Hydromethod =
-2}). Notably, the Zeus shock front has progressed less far than forthe
-PPM run. This is due to energy loss when conserving internal, not
-total, energy.}
+\caption{Density slice from the Sedov Blast test at $t = 0.07$. Left:
+results using the Piecewise Parabolic Method hydro scheme.  Right: results using the \zeus\
+hydro method. Notably, the \zeus\ shock front has progressed less far
+than in the PPM run. This is due to energy loss when conserving only
+internal, and not total, energy.}
 \label{fig.sedov1}
@@
 \includegraphics[width=\textwidth]{figures/sedov-profiles.eps}
-\caption{Radially averaged profiles for the Sedov Blast Test at $t =
+\caption{Radially averaged profiles for the Sedov Blast test at $t =
 0.07$. Clockwise from top-left shows density, velocity, internal
-energy and pressure.}
+energy and pressure.  The black solid line shows the analytic
+solution.  The blue dashed line shows the simulation using the PPM
+method, and the red dot-dashed line using \zeus.  The \zeus\ result
+substantially lags the true result due to total energy not being
+explicitly converged.}
 \label{fig.sedov2}
@@
 point of energy injection.  This problem is a popular test of
-astrophysical numerical codes for three reasons: Firstly, it is
+astrophysical numerical codes for three reasons: First, it is
 particularly appropriate to astrophysics since it represents the
-situation when a supernovae explosion occurs. Secondly, it has an
+situation when a supernovae explosion occurs. Second, it has an
 exact analytical solution whereby the shock front's radial position is
@@
 \citet{Sedov1959}. This solution makes it possible to assess how well
-the code performs. Thirdly, the spherical shock is a challenging
+the code performs.  Third, the spherical shock is a challenging
 problem for numerical codes (both grid and particle) since the shock
@@
 codes can succumb to. The test presented here is the two-dimensional
-version that is included in the Enzo distribution. The
-three-dimensional results from this test, both for Enzo and three
+version that is included in the \enzo\ distribution. The
+three-dimensional results from this test, both for \enzo\ and three
 other leading astrophysics codes, can be founds in \citet{Tasker2008}.
@@
 gas at a density of 1 (note that, in the absense of gravity, this
-problem is scale-free and so without units). Thermal energy is
+problem is scale-free and thus without units). Thermal energy is
 deposited into a single cell at the center of the box with $E_0 =
 10.0$. The problem is completed in two dimensions with reflecting
-boundary conditions (the default for Enzo) and uses a boxsize of side
-$1$. For this problem, we selected a top grid of $100 \times 100$ and
-four levels of refinement, placed based on shock location and the
-slope in density and total energy (this corresponds to cell flagging
-methods 1 and 3 in the Enzo parameter file). The exception to this
-scheme is in the initial conditions where grids are placed directly
-around the injection point. The results were assessed at $t = 0.07$,
-which corresponding to a time just before the shock reaches the box
-edge (see Figure~\ref{fig.sedov1}).
+boundary conditions (the default for \enzo) and uses a box size of side
+$1$. For this problem, we selected a top grid of $100 \times 100$
+cells and a maximum of four levels of refinement, placed based on
+shock location and the slope in density and total energy. The
+exception to this scheme is in the initial conditions, where grids are
+placed directly around the injection point. The results were assessed
+at $t = 0.07$, which corresponding to a time just before the shock
+reaches the box edge (see Figure~\ref{fig.sedov1}).
 
 Figure~\ref{fig.sedov2} shows the radial profiles for the simulation
-run with the PPM hydro-solver (blue dashed line) and the Zeus
+run with the PPM hydro-solver (blue dashed line) and the \zeus\
 hydro-solver (red dash-dot-dot line) together with the analytical
@@
 velocity, internal energy and pressure. PPM matches the analytical
-solution extremely well for all quantities. However, Zeus' shock front
-lags behind the analytical position. This can also be seen in the
-slices shown in Figure~\ref{fig.sedov1}. The cause of this discrepancy
-is that Zeus shows a substantial energy lost during the first few
-timesteps and produces a diamond-shaped, rather than spherical,
-shockfront during this time. After this, the code correctly conserves
-energy but this intial energy loss remains clearly visible in the
-position of the shock at $t = 0.07$. This problem was addressed
-directly by \citet{Clarke2010}, who attributed the source of the issue
-to this version of Zeus solving the internal, rather than total,
-energy equation. In situations with strong energy gradients, this
-choice caused an energy loss and the artificial viscosity produces the
-direction-dependent shockfront shape. In their paper,
+solution extremely well for all quantities. However, the shock front
+in the \zeus\ simulation lags behind the analytical position. This can
+also be seen in the slices shown in Figure~\ref{fig.sedov1}. The cause
+of this discrepancy is that \zeus\ shows a substantial energy loss
+during the first few timesteps and produces a diamond-shaped, rather
+than spherical, shockfront during this time. After this, the code
+correctly conserves energy but this intial energy loss remains clearly
+visible in the position of the shock at $t = 0.07$. This problem was
+addressed directly by \citet{Clarke2010}, who attributed the source of
+the issue to this version of \zeus\ solving the internal, rather than
+total, energy equation. In situations with strong energy gradients,
+this choice caused an energy loss and the artificial viscosity
+produces the direction-dependent shockfront shape. In their paper,
 \citet{Clarke2010} presents results from an alternative version of
-Zeus that conserves total energy. This problem is less marked for
-smaller energy gradients and it should be noted that Zeus' stability
-and speed make it a highly competitive choice, despite the
-disagreements in this test.
+\zeus\ that conserves total energy. This problem is less marked for
+smaller energy gradients and it should be noted that the \zeus\ hydro
+algorithm's stability and speed make it a highly competitive choice,
+despite the disagreements in this test.
 
@@
 features that are critical to structure formation: hydrodynamics,
-expansion, and self gravity.  This problem represents an ideal,
-isolated caustic formation, and is thus a useful proxy for much more
+expansion, and self gravity.  This problem represents the formation of an ideal,
+isolated caustic, and is thus a useful proxy for much more
 complicated structures in full 3-dimensional cosmological simulations
@@
 The initial conditions are simple, and we follow the prescription of
-\citet{Anninos94}.  Assuming a flat cosmology, the density
+\citet{Anninos94}.  Assuming a geometrically flat cosmology, the density
 perturbation is given by
@@
 \begin{equation}
-v(x_l) = -H_0 \frac{1 + z_c}{(1+z)^{1/2}} \frac{\sin(k x_l}{k}
+v(x_l) = -H_0 \frac{1 + z_c}{(1+z)^{1/2}} \frac{\sin(k x_l)}{k}
 \end{equation}
@@
 In the equations above, $\rho_0$ is the background density, z$_c$ is a
-free parameter and is the redshift at which the sheet forms a caustic
-(i.e., 'pancakes'), z is the redshift of initialization, x$_l$ is the
+free parameter and is the redshift where the sheet forms a caustic
+(i.e., where it `pancakes'), z is the redshift of initialization, x$_l$ is the
 Lagrangian mass coordinate, $k = 2 \pi / \lambda$ (where $\lambda$ is
@@
 caustic formation.  In Figure~\ref{fig.pancake}, we show the results
-of a test of the AMR version of \enzo's Zel'dovich Pancake test.  A
+of a test of the adaptive mesh version of \enzo's Zel'dovich Pancake test.  A
 one-dimensional box of length $64$~Mpc/h is initialized at z$ = 20$ in
@@
 The strong shocks and large density gradients are well-resolved, with
-density and velocity jumps being well-delineated.  The key features of
+density and velocity jumps being well-delineated and at the correct locations.  The key features of
 this test problem can be resolved with far fewer cells -- simulations
 including a mere 8 cells resolve the key features, as shown in
-sections 3.3.4-3.3.5 of~\citet{1996PhDT........80B} -- but we choose a
+Sections 3.3.4-3.3.5 of~\citet{1996PhDT........80B} -- but we choose a
 higher resolution here for illustrative purposes.
@@
 and the presence of shocks for up to two additional levels of mesh,
-having a maximum effective resolution of 1024 grid cells.}
+having a maximum effective resolution of 1024 grid cells. The top,
+middle, and bottom rows show density, temperature, and velocity of the
+gas, respectively; all are a function of position in units of the box
+size.  The central region ($x \simeq 0.45-0.55$) has bene adaptively
+refined, as can be seen by the locations of grid points.  Shocks and
+the central density peak are clearly resolved, with well-delineated
+jumps at the appropriate locations.}
 \label{fig.pancake}
@@
 \label{sec.tests.mhd}
-Figure \ref{fig.orszag} shows the Orszag-Tang vortex problem, a
-typical test problem for MHD \citep{Orszag79}.  The left panel shows
-the MHD-CT result, while the right panel shows the Dedner result.
-This test shows that small scale structure can be generated in MHD
-from large scale initial perturbations.  The test begins with uniform
-density, $\rho_0=25/36 \pi$ and pressure, $P_0=5/12 \pi$.  There is a
+Figure~\ref{fig.orszag} shows the Orszag-Tang vortex problem, a
+standard magnetohydrodynamic test problem \citep{Orszag79}.  The left panel shows
+the result using \enzo's constrained transport MHD method, while the right panel shows the result using
+\enzo's implementation of Dedner MHD.
+The Orszag-Tang vortex test shows that significant small scale structure can be generated in MHD
+from large scale initial perturbations, and is used to compare the
+effective resolution of different MHD schemes.  The test begins with uniform
+density, $\rho_0=25/36 \pi$ and pressure, $P_0=5/12 \pi$ (as with
+other tests in this section, in the absence of gravity or chemistry we use
+arbitrary units).  There is a
 single rotational mode in the velocity, and two in the magnetic field:
@@
 ${\bf B}_0$ = (-sin(2 $\pi$ y) $ \hat{x},$ sin( 4 $\pi$ x )$\hat{y}$).
-The simulation is evolved to t=0.48.  One can see that the structures
+The simulation is evolved to $t=0.48$ (in arbitrary units).  One can see that the structures
 are accurately represented as compared to, for instance,
@@
 both methods.
-
-{\emph Dave's note: I'll probably play around with these results in
-future versions.}
 
@@
 \includegraphics[width=0.4\textwidth]{figures/MHDDedner_OrszagTang_Density.eps}
-\caption{TANG}
+\caption{Density field from the Orszag-Tang vortex test, at $t=0.48$.
+Left: solution using constrained transport MHD.  Right: solution using
+Dedner MHD. The initial conditions are uniform density, with a single
+rotating velocity structure and two circular magnetic structures.
+These initial conditions generate significant small-scale structure in
+both the CT and Dedner schemes, which have approximately equal
+effective resolution.}
 \label{fig.orszag}
@@
 \citep{1983ApJ...265.1047Y}.  Our implementation, described briefly
-here, follows the work of \citet{2005ApJ...626..627O}, which one
-should consult for further details.  The density evolution is given by 
+here, follows the work of \citet{2005ApJ...626..627O}, and we direct
+the interested reader to this paper for further details.  The
+evolution of the gas density, $\rho$, is given by 
 \begin{equation}
@@
 \end{equation}
-and $t_{dyn}$ is the dynamical time, expressed as
+In this equation, $t_{dyn}$ is the dynamical time and is expressed as
 \begin{equation}
@@
 \end{equation}
-The collapse timescale is altered from the dynamical time by the
-factor, $f$ in Equation \ref{eqn.tcol}, which an approximation of the
-ratio of the pressure to the force of gravity.  The value of $f$
+
+with G being the gravitational constant.  The collapse timescale is altered from the dynamical time by a
+factor $f$ in Equation \ref{eqn.tcol}, which is an approximation of the
+ratio of the gas pressure to the force of gravity.  The value of $f$
 depends on the effective adiabatic index, $\gamma_{ef} \equiv (\partial
@@
 from derivative values for the two previous timesteps.  For the value
-of $f$, we use the piecewise function of \citet{2005ApJ...626..627O},
+of $f$ in this test problem, we use the piecewise function of \citet{2005ApJ...626..627O},
 given by
@@
 one-zone collapse test performed with an initial number density of 1
-cm$^{-3}$ and temperature of 100 K using the 12 species chemistry
+hydrogen atom per
+cm$^{-3}$ and an initial temperature of 100 K using the 12 species chemistry
 network with H, D, and He species and metal cooling rates calculated
-with Cloudy.
+with the \texttt{Cloudy} code.  The effects of metal cooling can be
+clearly seen; as the metallicity increases from zero to
+$10^{-2}$~Z$_\odot$, the gas deviates from the primoridal result
+(black line) both more rapidly and more significantly.  Our primordial results
+compare very well to those shown in \citet{2005ApJ...626..627O}; however,
+we use a different cooling method than Omukai, so the lines describing
+the evolution of the metal-enriched gas are not directly comparable.
 
@@
   \caption{Evolution of temperature versus number density for a one-zone
-    collapse test for gas at metallicities from 0 to 10$^{-2} Z_{\odot}$
-    using the primordial chemistry network with tabulated metal cooling
-    rates calculated with Cloudy.}
+    collapse test for gas at metallicities from 0 to 10$^{-2}
+    Z_{\odot}$.  This test is based on the results of
+    \citet{2005ApJ...626..627O}, and approximates the collapse of a
+    self-gravitating, cooling gas cloud.  This test problem uses \enzo's primordial chemistry network with tabulated metal cooling
+    rates calculated with the \texttt{Cloudy} code, and compares
+    favorably to Omukai's results.}
   \label{fig.onezone}

### 2013-04-29 code-tests.tex
oops, needed to do a bit of cleanup in 1-zone test.

@@
 \label{sec.tests.1-zone}
+
 The one-zone collapse test simulates the collapse of a
@@
 evolution of the gas density, $\rho$, is given by
+
 \begin{equation}
@@
 \end{equation}
+
 where the collapse timescale, $t_{col}$, is
+
 \begin{equation} \label{eqn.tcol}
@@
 
-with G being the gravitational constant.  The collapse timescale is altered from the dynamical time by a
-factor $f$ in Equation \ref{eqn.tcol}, which is an approximation of the
-ratio of the gas pressure to the force of gravity.  The value of $f$
-depends on the effective adiabatic index, $\gamma_{ef} \equiv (\partial
-ln\ p / \partial ln\ \rho)$, which we calculate by linearly extrapolating
+with G being the gravitational constant.  The collapse timescale is
+altered from the dynamical time by a factor $f$ in Equation
+\ref{eqn.tcol}, which is an approximation of the ratio of the gas
+pressure to the force of gravity.  The value of $f$ depends on the
+effective adiabatic index, $\gamma_{ef} \equiv (\partial ln\ p
+/ \partial ln\ \rho)$, which we calculate by linearly extrapolating
 from derivative values for the two previous timesteps.  For the value
-of $f$ in this test problem, we use the piecewise function of \citet{2005ApJ...626..627O},
-given by
+of $f$ in this test problem, we use the piecewise function of
+\citet{2005ApJ...626..627O}, given by
+
 \begin{equation}
@@
 \end{equation}
+
 The specific energy evolves as
+
 \begin{equation}
@@
 \end{equation}
+
 where $\Lambda$ is the cooling rate in units of erg s$^{-1}$ g$^{-1}$
@@
 one-zone collapse test performed with an initial number density of 1
-hydrogen atom per
-cm$^{-3}$ and an initial temperature of 100 K using the 12 species chemistry
-network with H, D, and He species and metal cooling rates calculated
-with the \texttt{Cloudy} code.  The effects of metal cooling can be
-clearly seen; as the metallicity increases from zero to
-$10^{-2}$~Z$_\odot$, the gas deviates from the primoridal result
-(black line) both more rapidly and more significantly.  Our primordial results
-compare very well to those shown in \citet{2005ApJ...626..627O}; however,
-we use a different cooling method than Omukai, so the lines describing
-the evolution of the metal-enriched gas are not directly comparable.
+hydrogen atom per cm$^{-3}$ and an initial temperature of 100 K using
+the 12 species chemistry network with H, D, and He species and metal
+cooling rates calculated with the \texttt{Cloudy} code.  The effects
+of metal cooling can be clearly seen; as the metallicity increases
+from zero to $10^{-2}$~Z$_\odot$, the gas deviates from the primoridal
+result (black line) both more rapidly and more significantly.  Our
+primordial results compare very well to those shown in
+\citet{2005ApJ...626..627O}; however, we use a different cooling
+method than Omukai, so the lines describing the evolution of the
+metal-enriched gas are not directly comparable.
 
@@
   \end{center}
-  \caption{Evolution of temperature versus number density for a one-zone
-    collapse test for gas at metallicities from 0 to 10$^{-2}
+  \caption{Evolution of temperature versus number density for a
+one-zone collapse test for gas at metallicities from 0 to 10$^{-2}
 Z_{\odot}$.  This test is based on the results of
 \citet{2005ApJ...626..627O}, and approximates the collapse of a
-    self-gravitating, cooling gas cloud.  This test problem uses \enzo's primordial chemistry network with tabulated metal cooling
-    rates calculated with the \texttt{Cloudy} code, and compares
-    favorably to Omukai's results.}
+self-gravitating, cooling gas cloud.  This test problem uses \enzo's
+primordial chemistry network with tabulated metal cooling rates
+calculated with the \texttt{Cloudy} code, and compares favorably to
+Omukai's results.}
   \label{fig.onezone}

### 2013-04-29 code-tests.tex
edited test problems through moray ray-tracing test.

@@
 
-The photo-evaporation of dense clumps is prevalent in radiation
+The photo-evaporation of dense clumps of gas is prevalent in radiation
 hydrodynamics simulations, and this test problem examines the
@@
 behind the clump, and the hydrodynamic response on the clump from
-photo-heating.  The problem setup is the same as Test 7 in the
-Cosmological Radiative Transfer Comparison Project
-\citep{IlievEtAl2009} and \citet{Wise11_Moray}.  The simulation domain
-is 6.6~kpc on a side with an ambient medium of pure neutral hydrogen
-$n_{\rm H} = 2 \times 10^{-4}\; \cubecm$ and $T = 8000 \unit{K}$.  We
-place a spherical overdensity in hydrostatic equilibrium with the
-ambient medium.  It has a radius $r = 0.8 \unit{kpc}$, hydrogen
-density $n_{\rm H} = 0.04 \cubecm$, and temperature $T = 40$ K, and it
-is centered at $(x,y,z) = (5, 3.3, 3.3) \unit{kpc}$.  In
-\citet{IlievEtAl2009}, all of the codes used a fixed $128^3$ grid to
-ease the comparison, but in this test to demonstrate a higher
-resolution AMR solution, we employ a $128^3$ grid with two additional
-levels of refinement with cells with a baryon mass (method 2 in
-\S\ref{sec:refinement_criteria}) greater than 1.5 being flagged for
-refinement.  This test is run for 15 Myr.
+photo-heating, all using \enzo's Moray ray-tracing module.  The
+problem setup is the same as Test 7 in the Cosmological Radiative
+Transfer Comparison Project \citep{IlievEtAl2009} and
+\citet{Wise11_Moray}.  The simulation domain is 6.6~kpc on a side with
+an ambient medium of pure neutral hydrogen $n_{\rm H} = 2 \times
+10^{-4}\; \cubecm$ and $T = 8000 \unit{K}$.  We place a spherical
+overdensity in hydrostatic equilibrium with the ambient medium.  It
+has a radius $r = 0.8 \unit{kpc}$, hydrogen density $n_{\rm H} = 0.04
+\cubecm$, and temperature $T = 40$ K, and it is centered at $(x,y,z) =
+(5, 3.3, 3.3) \unit{kpc}$.  In \citet{IlievEtAl2009}, all of the codes
+used a fixed $128^3$ grid to ease the comparison, but in this test to
+demonstrate a higher resolution AMR solution, we employ a $128^3$ grid
+with two additional levels of refinement with cells with a baryon mass
+(method 2 in Section~\ref{sec:refinement_criteria}) greater than 1.5
+being flagged for refinement.  This test is run for 15 Myr.
 
@@
 0.24, 0.06)$ that are optimized to reduce errors in the solutions with
-a full spectrum and energy discretization \citep{Mirocha12}.  Note
-that this choice of energy groups is different from
-\citet{Wise11_Moray}.  We use a minimum angular resolution of 10 rays
+a full spectrum and energy discretization \citep{Mirocha12}.
+\citep[Note that this choice of energy groups is different
+from][]{Wise11_Moray}.  We use a minimum angular resolution of 10 rays
 per cell and a constant radiative transfer timestep of 25 kyr.  Figure
@@
   \includegraphics[width=1.0\textwidth]{figures/code-test-shadowing.eps}
-  \caption{Photo-evaporation of a dense clump.  Clockwise from the
-    upper left-hand side: slices through the clump center of the
-    density, temperature, neutral fraction, and pressure.}
+  \caption{Photo-evaporation of a dense clump using the Moray
+ray-tracing module.  Clockwise from the upper left corner: slices
+through the clump center of the density, temperature, neutral hydrogen
+fraction, and pressure at $t=15$~Myr after the initialization of the
+simulation.  A point source of radiation is at the center of the -x
+boundary and illuminates the clump with a constant luminosity, casting
+a clear shadow behind the clump, as seen in the neutral hydrogen
+fraction.}
   \label{fig:shadowing}

### 2013-04-29 code-tests.tex
Edited anisotropic thermal conduction test problem.  And with that, I am done editing this paper.  VICTORY IS MINE.

@@
 In astrophysical environments such as the intergalactic and
-intracluster media, with typical magnetic strengths of $\sim 1 \mu$G
+intracluster media, with typical magnetic strengths of $\sim 1$~$\mu$G
 and temperatures in the millions of Kelvin, the electron gyroradius is
-very small compared to the scapes of interest in a cosmological
-simulation.  As a result, electrons can only move \textit{along}
-magnetic field lines, not perpendicular to them, which can result in
-very interesting magnetohydrodynamical instabilities
-\citep[e.g.,][]{2008ApJ...677L...9P,2008ApJ...688..905P}. 
- The correct modeling of this behavior can be quite challenging, and is described in 
-Section~\ref{sec.num.conductions}.  Conduction in \enzo\ is calculated
-in both an operator-split and directionally-split manner, and the
-addition of heat transport solely along magnetic field lines requires
-computation of cross-terms in the temperature derivative at cell
-faces, which can add spurious oscillations in the temperature field in
-regions where the temperature gradient is strong in more than one
-spatial direction unless an appropriate flux-limiter \citep[such as that
-of][]{1977JCoPh..23..263V} is chosen.
+very small compared to the scales of interest in a typical
+cosmological simulation.  As a result, electrons can only move
+\textit{along} magnetic field lines, not perpendicular to them, which
+can result in very interesting magnetohydrodynamical instabilities
+\citep[e.g.,][]{2008ApJ...677L...9P,2008ApJ...688..905P}.  The correct
+modeling of this behavior can be quite challenging, and is described
+in Section~\ref{sec.num.conductions}.  Conduction in \enzo\ is
+calculated in both an operator-split and directionally-split manner,
+and the addition of heat transport solely along magnetic field lines
+requires computation of cross-terms in the temperature derivative at
+cell faces, which can add spurious oscillations in the temperature
+field in regions where the temperature gradient is strong in more than
+one spatial direction unless an appropriate flux-limiter \citep[such
+as that of][]{1977JCoPh..23..263V} is chosen.
 
-Figure~\ref{fig.conduct} shows a test that demonstrates the
-correct behavior of anisotropic thermal conduction in \enzo.  We
-initialize a two-dimensional, $256^2$ cell simulation having a physical
+Figure~\ref{fig.conduct} shows a test that demonstrates the correct
+behavior of anisotropic thermal conduction in \enzo.  We initialize a
+two-dimensional, $256 \times 256$ cell simulation having a physical
 scale of 1 kpc on a side, a uniform density of 1 proton/cc, and a
 background temperature of $10^6$ Kelvin.  Magnetic fields with a
-strength of  B$_0 = 1 \mu$G are initialized such that the field lines form
-circles around the center of the simulation volume, such that B$_x =
--B_0\sin(\theta)$ and B$_y = B_0\cos(\theta)$, where $\theta$ is the
-angle measured from the $+x$ direction in a counterclockwise manner.
-A Gaussian temperature pulse is injected at $(0.75, 0.5)$ (in units of the box
-size), with a peak temperature of $10^8$ K and a FWHM of $0.015625$ of
-the box size.  This initial setup is shown in the left panel of
-Figure~\ref{fig.conduct}.  The simulation is then allowed to evolve
-with \textit{only} anisotropic conduction turned on (e.g.,
-no hydrodynamics or radiative cooling), and with a Spitzer fraction of
-f$_{sp} = 1$.
+strength of B$_0 = 1$~$\mu$G are initialized such that the field lines
+form circles around the center of the simulation volume, such that
+B$_x = -B_0\sin(\theta)$ and B$_y = B_0\cos(\theta)$, where $\theta$
+is the angle measured from the $+x$ direction in a counterclockwise
+manner.  A Gaussian temperature pulse is injected at $(0.75, 0.5)$ (in
+units of the box size), with a peak temperature of $10^8$ K and a FWHM
+of $1/64$ of the box size.  This initial setup is shown in the left
+panel of Figure~\ref{fig.conduct}.  The simulation is then allowed to
+evolve with \textit{only} anisotropic conduction turned on (e.g., no
+hydrodynamics, radiative cooling, or cosmological expansion), and with
+a Spitzer fraction of f$_{sp} = 1$.
 
@@
 constant temperature background with circular magnetic fields centered
-on (0.5, 0.5).  The background medium has a density of 1 proton/cc and
-a temperature of $10^6$~K.  At t$ = 0$ (left panel), a Gaussian heat
-pulse is injected at (0.75, 0.5) with a FWHM of $0.015625$ (with all
-numbers given in units of the box size) and a peak temperature of
-$10^8$~K, and allowed to evolve without hydrodynamical motion (i.e.,
-static gas) and no radiative cooling for 300 Myr.  At t$ = 300$~Myr
-(right panel), heat has been transported along magnetic field lines
-with no significant diffusion perpendicular to field
-lines. Furthermore, there is no detectable oscillations in the
-temperature in regions where the field is not parallel with the grids.}
+on (0.5, 0.5) in a $256 \times 256$ grid.  The background medium has a
+density of 1 proton/cc and a temperature of $10^6$~K.  At t$ = 0$
+(left panel), a Gaussian heat pulse is injected at (0.75, 0.5) with a
+FWHM of $1/64$ (with all numbers given in units of the box size) and a
+peak temperature of $10^8$~K, and allowed to evolve without
+hydrodynamical motion (i.e., static gas) and no radiative cooling for
+300 Myr.  At t$ = 300$~Myr (right panel), heat has been transported
+along magnetic field lines with no significant diffusion perpendicular
+to field lines. Furthermore, there are no detectable oscillations in
+the temperature in regions where the magnetic field is not parallel
+with the grids.}
 \label{fig.conduct}

### 2013-05-01 overview.tex
Added description of Dedner MHD solver to the overview section.

--- 2013-03-20 overview.tex
@@
 
-\subsubsection{Godunov MUSCL (HD) with Dender divergence cleaning (MHD)}
+\subsubsection{Godunov MUSCL (HD) with Dedner divergence cleaning (MHD)}
+This solver was developed by Wang \& Abel to attack problems in
+magnetic field amplification during the formation of galaxies
+\citep{Wang:2009a} and to understand the role proto-stellar jets for
+the theory of star formation \citep{Wang:2009b}. It combines the standard
+approach of Godunov \citep{Godunov1959} for finite volume techniques
+with the method of lines as described as e.g. by
+\cite{leveque2002finite} and \cite{toro-1997}. In addition, however,
+it implements the hyperbolic divergence cleaning algorithm of
+\cite{2002JCoPh.175..645D}. It supports multiple approximate Riemann
+solvers and supports non-ideal equations of states. Consequently,
+these suite of solvers can be used for hydro and magneto-hydrodynamic
+simulations. It is also this class of solvers as well as a version of
+the PPM hydro solver which have been ported to Nvidia's CUDA framework
+allowing enzo to take advantage of modern graphics hardware \citep{Wang:2010}.
 

### 2013-05-01 overview.tex
use \enzo instead of enzo.

@@
 the PPM hydro solver which have been ported to Nvidia's CUDA framework
-allowing enzo to take advantage of modern graphics hardware \citep{Wang:2010}.
+allowing \enzo to take advantage of modern graphics hardware \citep{Wang:2010}.
 

### 2013-05-01 introduction.tex

@@
 
-Many phenomena in astrophysics and cosmology require a high spatial and temporal dynamical range and there are a number of numerical techniques which have been used to model such problems.
-At the current time, the most common is a gridless, fully Lagrangian method call Smoothed Particle Hydrodynamcs \citep[SPH;][]{Lucy77, SPH} which has been very successful, particularly for those situations dominated by gravity.
+Many phenomena in astrophysics and cosmology require a high spatial and temporal dynamical range, and there are a number of numerical techniques that have been used to model such problems.
+At the current time, the most common is a gridless, fully Lagrangian method call Smoothed Particle Hydrodynamcs \citep[SPH;][]{Lucy77, SPH} which has been very successful -- particularly for those situations dominated by gravity, where systems evolve over large spatial and temporal dynamic ranges.
 Its development, however, is still at a relatively early stage when compared to the effort put into Eulerian, grid-based hydrodynamic schemes \citep[e.g.,][]{laney-1998, toro-1997, Woodward84}.
 
-Therefore, we would like to make use of this expertise that has been invested in Eulerian solvers; unfortunately, as originally developed, none of these schemes provide an easy way to adaptively and dynamically increase the resolution in small volumes, as required by, for example, gravitational instability.
-We have instead, adopted an idea first suggested by \citet{Berger89} in the Computational Fluid Dynamics (CFD) community, which has become known as Structured Adaptive Mesh Refinement (SAMR).
-The principle is to adaptively add (and modify) additional, finer meshes, or patches, over regions which require higher resolution.
-Since each patch has a simple geometry (here, Cartesian) the hyperbolic fluid-dynamical equations can be solved on that patch with a well-developed `off-the-shelf' numerical method.
+Therefore, we would like to make use of the expertise that has been invested in Eulerian solvers; unfortunately, as originally developed, none of these schemes provide an easy way to adaptively and dynamically increase the spatial and temporal resolution in small volumes, as required by, for example, gravitational instability.
+We have instead adopted an idea first suggested by \citet{Berger89} in the Computational Fluid Dynamics (CFD) community, which has become known as Structured Adaptive Mesh Refinement (SAMR).
+The principle is to adaptively add and modify additional, finer meshes (``patches'') over regions that require higher resolution.
+Since each patch has a simple geometry (Cartesian, in our case) the hyperbolic fluid-dynamical equations can be solved on that patch with a well-developed `off-the-shelf' numerical method.
 In addition, it is relatively straightforward to add other advanced physics (e.g.
-MHD, radiation transport, etc.).
+magnetohydrodynamics, radiation transport, chemistry, etc.).
 This paper is devoted to describing how we have implemented this idea for astrophysics, including the addition of comoving coordinates, self-gravity, radiative cooling, chemistry, heat conduction, collisionless fluids, MHD, radiation transport, star formation and a range of other physical effects.
@@
 There have been many other numerical methods described in the astronomical literature which contain elements of this idea, or have a similar aim.
-For example, an N-body solver \cite{Villumsen89} used non-adaptive meshing to increase the resolution in pre-selected regions.
-This static approach, applied to hydrodynamics, has been used extensively \citep[e.g.,]{Ruffert94, Anninos94}.
-Adding adaptivity is a more recent enhancement, and there are now a number of N-body codes which possess this feature, both with and without  hydrodynamics  \citep{Couchman91, Jessop94, Suisalu95, Splinter96, Gelato97, ART97, Truelove98, flash_method, MLAPM01,  Yahagi01, RAMSES, Quilis04, Ziegler05, Zhang06, Astrobear09, Pluto-amr, GAMER, Nyx}.
+For example, an N-body solver \citep{Villumsen89} used non-adaptive meshing to increase the resolution in pre-selected regions.
+This static approach, applied to hydrodynamics, has been used extensively \citep[e.g.,][]{Ruffert94, Anninos94}.
+Adding adaptivity is a more recent enhancement, and there are now a number of N-body codes that possess this feature, both with and without  hydrodynamics  \citep{Couchman91, Jessop94, Suisalu95, Splinter96, Gelato97, ART97, Truelove98, flash_method, MLAPM01,  Yahagi01, RAMSES, Quilis04, Ziegler05, Zhang06, Astrobear09, Pluto-amr, GAMER, Nyx}.
 Of these, perhaps the most comparable and widely used are FLASH \citep{flash_method}, which uses grid blocks of fixed size, and RAMSES \citep{RAMSES} and ART \citep{ART97}, both of which refine individual cells.
@@
 It has been expanded as a general tool for astrophysical fluid dynamics and is intended to be efficient, accurate and easily expanded.
-Although many of the components of the code have been described in previous publications \citep{1995CoPhC..89..149B, BryanThesis96, Bryan97a, Bryan97b, Norman99, BryanCompSci99, Bryan01, Oshea04}, there has been no systematic and complete description of the code.
-In this paper we provide that description, filling in many gaps and showing the result of the code on a wide variety of test problems.
+Although many of the components of the code have been described in previous publications \citep{1995CoPhC..89..149B, BryanThesis96, Bryan97a, Bryan97b, Norman99, BryanCompSci99, Bryan01, Oshea04, 2007arXiv0705.1556N}, there has been no systematic and complete description of the code.
+In this paper we provide that description, filling in many gaps and showing the code's performance for a wide variety of test problems.
 
-The \enzo\ code has been widely used over the last two decades in a wide variety of problems, and more than 100 peer-reviewed papers have been published based on results obtained with the code.
-The code has previously been used to study a wide range of astrophysical systems, including X-ray clusters \citep{Loken02}, the ISM \citep{Slyz05}, the intergalactic medium \citep{Fang01}, cooling flows \citep{Li12}, turbulence \citep{Kritsuk04}, and the formation of the first stars \citep{ABN02}.
+The \enzo\ code has been extensively used over the last two decades in a wide variety of problems, and more than 100 peer-reviewed papers have been published based on results obtained with the code.
+The code has previously been used to study a wide range of astrophysical systems, including galaxy clusters \citep{Loken02}, the interstellar medium \citep{Slyz05}, the intergalactic medium \citep{Fang01}, cooling flows \citep{Li12}, turbulence \citep{Kritsuk04}, and the formation of the first stars \citep{ABN02}.
 
@@
 The structure of the code is as follows.
-In section~\ref{sec.overview}, we first provide a top-level overview of the code method and structure.
+In Section~\ref{sec.overview}, we first provide a top-level overview of the code method and structure.
 This is designed to give a first broad-brush picture of the equations we aim to solve and the methods used to solve them.
-Next, in section~\ref{sec.methods}, we describe the methods we use in detail, reserving some of the longer descriptions of particular components for the appendix in order not to interrupt the flow of the paper.
-The code tests are described in section~\ref{sec.tests} and the parallelism strategy and scaling results in section~\ref{sec.performance}.
+Next, in Section~\ref{sec.methods}, we describe the methods we use in detail, reserving some of the longer descriptions of particular components for the appendix in order not to interrupt the flow of the paper.
+The \enzo\ testing framework and code tests are described in Section~\ref{sec.tests}.
+The parallelism strategy and scaling results are described in Section~\ref{sec.parallel}.
+Finally, we discuss the code's development methodology (which is as far as we know unique in the astrophysics community) in Section~\ref{sec.development}.

### 2013-05-01 numerical-analysis.tex

@@
 Analyzing the results of simulations in detail requires both tools to ask
-sophisticated questions of data, and the ability to process vast quantities of
-data at high-cadence.  As simulations grow in size and complexity, storing data
-for post-processing becomes simply intractable.  To cope with this, we have
+sophisticated questions of data and the ability to process vast quantities of
+data at high time-cadence.  As simulations grow in size and complexity, storing data
+for post-processing simply becomes intractable.  To cope with this, we have
 instrumented Enzo with the ability to conduct analysis during the course of a
-simulation.  This enables analyzing with extremely high cadence (even as often
+simulation.  This enables analysis with extremely high time cadence (even as often
 as every subcycle of the finest refinement level), as well as without
@@
 the same compute resources as used by the simulation itself; this is often
-referred to as \textit{in situ} visualization.  Utilizing a
+referred to as \textit{in situ} analysis or visualization.  Utilizing a
 dynamically-scheduled second set of computing resources, often referred to as
-\textit{co-scheduled} visualization, provides greater flexibility and overall
+\textit{co-scheduled} analysis or visualization, provides greater flexibility and overall
 throughput at the expense of simplicity.
@@
 and the frequency with which it is called is adjustable by a run-time
-parameter.  During the course of conducting analysis, this simulation is halted
+parameter.  During the course of conducting analysis, the simulation is halted
 until the conclusion of the analysis.
@@
 quantity that a user requires.  This capability has been used quite
-effectively in several papers, including \citep{2010ApJ...715.1575S, 2012ApJ...748...12S}.
+effectively in several papers, including \citet{2010ApJ...715.1575S}
+and \citet{2012ApJ...748...12S}.
 
@@
 \enzo~is the dimensionally unsplit temperature jump method described in detail
-in \citet{2008ApJ...689.1063S}.  We briefly outline the method below.
+in \citet{2008ApJ...689.1063S}.  We briefly outline the method here.
 
@@
 necessary to be flagged as a shock:
+
 \begin{eqnarray}
@@
 \end{eqnarray}
+
 where $\vec{v}$ is the velocity field, $T$ is the temperature, $\rho$ is the
 density, and $S=T/\rho^{\gamma-1}$ is the entropy.  An optional temperature floor
-may be chosen, useful for example in cosmological simulations without a radiation
+may be chosen, which is useful for, e.g., cosmological simulations without a radiation
 background.

### 2013-05-01 numerical-chemistry.tex

@@
 ionization) within a gas can be treated as being in equilibrium, in
-some regimes that are found in astrophysics this is not true.  For
-example, the cooling and collapse of primordial gas in Population III
-star formation is dominated by molecular hydrogen, which in the
-absence of dust forms via an inefficient pair of collisional processes
-that depend heavily on the local population of free electrons.  As a
-result, when modeling primordial star formation, it is critical to
-follow the non-equilibrium evolution of the chemical species of
-hydrogen, including molecular hydrogen and deuterium.
+some regimes that are found in astrophysics this assumption leads to
+considerable error.  For example, the cooling and collapse of
+primordial gas in Population III star formation is dominated by
+molecular hydrogen, which in the absence of dust forms via an
+inefficient pair of collisional processes that depend heavily on the
+local population of free electrons.  As a result, when modeling
+primordial star formation, it is critical to follow the
+non-equilibrium evolution of the chemical species of hydrogen,
+including molecular hydrogen and deuterium.
 
@@
 well as several models for a metagalactic UV background that heat the
-gas via photoionization and discussion -- see
+gas via photoionization and photodissociation; see
 Section~\ref{sec.num.cooling} for more details).  Depending on user
-parameters, one can follow simply atomic species (H, \Hp, He, \Hep,
-\Hepp, and e$^-$), include those relevant for molecular hydrogen
-formation (\HH, \HHp, and \Hm), and further include deuterium and its
-species (D, \Dp, and HD).  \red{A total of 9 kinetic equations are
-  solved, including 29 kinetic and radiative processes, for the 9
-  species mentioned above.  See Table~\ref{table.collisional} for the
-  collisional processes and Table~\ref{table.radiative} for the
-  radiative processes solved.}
+parameters, one can follow only atomic species (H, \Hp, He, \Hep,
+\Hepp, and e$^-$), include those species relevant for molecular
+hydrogen formation (\HH, \HHp, and \Hm), and can further include
+deuterium and its species (D, \Dp, and HD).  A total of 9 kinetic
+equations are solved, including 29 kinetic and radiative processes,
+for the 12 species mentioned above.  See Table~\ref{table.collisional}
+for the collisional processes and Table~\ref{table.radiative} for the
+radiative processes solved.
 
@@
 costly if small enough timesteps are taken to keep the network stable.
-This makes an implicit scheme much more preferable for such a set of
+This makes an implicit scheme strongly preferable for such a set of
 equations.  However, an implicit scheme typically require an iterative
@@
 solution. This technique is optimized by taking the chemical
-intermediaries \Hm and \HHp, which have large rate coefficients
-and low concentrations, and grouping them into a separate category of
-equations.  Due to their fast reactions these species are very
+intermediaries \Hm and \HHp, which have large rate coefficients and
+low concentrations, and grouping them into a separate category of
+equations.  Due to their fast reactions, these species are very
 sensitive to small changes in the more abundant species.  However, due
@@
 the rest of the network and treated independently.  In fact, \Hm~and
-\HHp~are treated as being in equilibrium at all times, independent
-of the other species and the hydrodynamic variables.  This allows a
-large speedup in solution as the BDF scheme is then applied only to
-the slower 7- or 10-species network (depending on whether deuterium is
+\HHp~are treated as being in equilibrium at all times, independent of
+the other species and the hydrodynamic variables.  This allows a large
+speedup in solution as the BDF scheme is then applied only to the
+slower 7- or 10-species network (depending on whether deuterium is
 included or not) on timescales closer to those required by the
@@
 single hydrodynamic timestep.  These subcycle timesteps are determined
-so that the maximum fractional change in the electron concentration is
-limited to no more than $10\%$ per timestep.  Given this subcycling,
-the method is not truly implicit over an entire hydrodynamical
-timestep, and thus not truly a BDF method -- however, it uses the
-formalism and is implicit over subcycle timesteps.
+so that the estimated fractional change in the electron concentration
+is limited to no more than $10\%$ per timestep.  Given this
+subcycling, the method is not truly implicit over an entire
+hydrodynamical timestep, and thus not truly a BDF method -- however,
+it uses the formalism and is implicit over subcycle timesteps.
 
@@
 10^{14}$~cm$^{-3}$.  A further concern is that the optically thin
-approximation for radiative cooling breaks down, which occurs before
-$n_{\rm H} \sim 10^{16} - 10^{17}$~cm$^{-3}$.  Beyond this point,
+approximation for radiative cooling eventually breaks down, which occurs before
+$n_{\rm H} \sim 10^{16} - 10^{17}$~cm$^{-3}$ in gas of primordial composition.  Beyond this point,
 modifications the cooling function that take into account the
@@
 with these modifications, a completely correct description of the
-cooling of this gas will require some form of radiation transport,
+cooling of primordial gas requires some form of radiation transport,
 which will greatly increase the cost of the simulations.

### 2013-05-01 numerical-conduction.tex

@@
 
-Enzo implements the equations of isotropic heat conduction in a manner
+\enzo\ implements the equations of isotropic heat conduction in a manner
 similar to that of \citet{2007ApJ...664..135P}, where we model the
@@
 by the presence of magnetic fields, which may strongly suppress heat
-flow.  The strength and confuction of magnetic fields in the
+flow.  The strength and distribution of magnetic fields in the
 intracluster medium is poorly understood, and it is unclear to what
@@
 
-Both isotropic and anisotropic thermal conduction in Enzo are treated
+Both isotropic and anisotropic thermal conduction in \enzo\ are treated
 in an operator-split manner.  Furthermore, within the heat transport

### 2013-05-01 numerical-cooling.tex

@@
 
-As discussed in \S\ref{sec.ov.chem}, the set of reactions that
+As discussed in Section~\ref{sec.num.chemistry}, the set of reactions that
 characterize a metal-free gas is simple enough to be computed in
@@
 state occur.  In addition to the subcycle timestepping contraints
-mentioned in \S\ref{sec.ov.chem}, the subcycle timestep is also not
-permitted to exceed 10\% of the cooling time, $e/\dot{e}$.  A
+mentioned in Section~\ref{sec.num.chemistry}, the subcycle timestep is also not
+permitted to exceed 10\% of the cooling time, $t_{cool} = e/\dot{e}$.  A
 metagalactic background affects the gas through both photo-heating and
@@
 UV backgrounds present in \enzo is presented below in
-\S\ref{sec.num.rad-homogeneous}.
+Section~\ref{sec.num.rad-homogeneous}.
 
@@
 A more sophisticated method makes use of multidimensional cooling and
-heating rate tables computed with the photo-ionization code,
+heating rate tables computed with the photo-ionization code
 \texttt{Cloudy} \citep{1998PASP..110..761F}.  This method, detailed in
@@
 where those dimensions are density, metallicity, electron fraction,
-redshift (for an evolving metagalactic UV backgroun), and
+redshift (for an evolving metagalactic UV background), and
 temperature.  Tables can be created for any abundance pattern for
@@
 field.  The cooling from H and He is calculated by the non-equilibrium
-primordial chemistry and cooling solver.  The contribution from metals
+primordial chemistry and cooling solver.  The contribution of metals
 to the cooling is computed within the same subcycling loop as the
@@
     metagalactic background at $z = 0$ with a gas number density of
-    10$^{-4}$ cm$^{-3}$.  The low temperature rates shown by the
-    dashed lines indicate a net heating.  The red curve is the
-    tabulated cooling function of \citet{SW87} which assumes a fully
+    10$^{-4}$ cm$^{-3}$.  The rates shown by the
+    dashed lines indicate a net heating below $T \sim 10^{4.5}$, where
+    the rapid change in rate is evident (the curve is an absolute
+    value so it can be shown on a log plot).  The red curve is the
+    tabulated cooling function of \citet{SW87}, which assumes a fully
     ionized gas with metallicity of 0.5 $Z_{\odot}$.}

### 2013-05-01 numerical-gravity.tex

@@
 
-Solving for the accelerations of the cells and particles on the grid due to self-gravity involves three steps: (i) computing the total gravitating mass, (ii) solving for the potential field with the appropriate boundary conditions, and (iii) differencing the potential to get the acceleration, and, if necessary, interpolating the potential back to the particles.
+Solving for the accelerations of the cells and particles on the grid due to self-gravity involves three steps: (i) computing the total gravitating mass, (ii) solving for the gravitational potential field with the appropriate boundary conditions, and (iii) differencing the potential to get the acceleration, and, if necessary, interpolating the acceleration back to the particles.
 These steps are described in detail below.
 
-First, the dark matter particles are distributed onto the grids using the second-order cloud-in-cell (CIC) interpolation technique \citet{Hockney88} to form a spatially discretized density field $\rho_{\rm DM}$.
+First, the dark matter particles are distributed onto the grids using the second-order cloud-in-cell (CIC) interpolation technique \citep{Hockney88} to form a spatially-discretized density field $\rho_{\rm DM}$.
 When doing the CIC interpolation, particle positions are (temporarily) advanced by $0.5 \Delta t v^n$ so that we generate an estimate of the time-centered density field.
 Particles on sub-grids are also added in this way.
-In addition, since the gravitating field for a grid is defined beyond the grid edges (see below), dark matter particles from sibling grids and sub grids which lie within the entire gravitating field are used.
+In addition, since the gravitating field for a grid is defined beyond the grid edges (see below), dark matter particles from sibling grids and sub grids whhatich lie within the entire gravitating field are used.
 This step can involve communication.
@@
 Next, we add the baryonic grid densities in a similar fashion.
-In particular, we treat baryonic cells as virtual CIC particles which are are placed at the grid center but are advanced by $0.5 \Delta t v^n$ in order to approximately time-center the gravitating mass field.
-Cells which are covered by further refined grids are treated in a similar way (i.e.
-we also use the sub grid cells as virtual CIC particles).
+In particular, we treat baryonic cells as virtual CIC particles that are are placed at the grid center but are advanced by $0.5 \Delta t v^n$ in order to approximately time-center the gravitating mass field.
+Cells that are covered by further-refined grids are treated in a similar way (i.e., we also use the subgrid cells as virtual CIC particles).
 This procedure results in a total gravitating mass field $\rho_{\rm total}^{n+1/2}$.
@@
 
-For isolated boundary conditions, we use James method \citep{James77}.
+For isolated boundary conditions, we use the James method \citep{James77}.
 In this case, the Greens function is generated in real-space so as to have the correct zero-padding properties and then transformed into the Fourier domain.
 In either case, the potential is then transformed back into the real domain to get potential values at the cell centers.
-These are differenced with a two-point centered difference to obtain accelerations at the cell centers (except if we are using the staggered Zeus-like solver, in which case the accelerations are computed at the cell faces to match the velocities).
+These are differenced with a two-point centered difference scheme to obtain accelerations at the cell centers (except if we are using the staggered Zeus-like solver, in which case the accelerations are computed at the cell faces to match the velocities).
 Particle accelerations are obtained using a (linear) CIC interpolation from the grid.
 
-In order to calculate more accurate potentials on subgrids, \enzo\ uses a similar but slightly different technique on the root grids.
+In order to calculate more accurate potentials on subgrids, \enzo\ uses a similar but slightly different technique from the root grid.
 The generation of the total gravitating mass field is essentially identical, using CIC interpolation for both the particles and baryons, including sub grids as before.
-To compute the potential on sub grids; however, we use the standard seven-point (in three dimensions) second-order finite-difference approximation to Poisson's equation.
+To compute the potential on sub grids, however, we use the standard seven-point (in three dimensions) second-order finite-difference approximation to Poisson's equation.
 Boundary conditions are then interpolated from the potential values on the parent grid.
@@
 
-The region immediately next to the boundary can contain unwanted oscillations \citep[e.g.,][]{Anninos94}, and so we use an expanded buffer zone around the grid, of size three root grid boundary zones (so typically six refined zones for a refinement factor of 2).
-The density is computed in this region and the potential solved, but only the region which overlaps with the active region of grid itself is used to calculate accelerations.
+The region immediately next to the boundary can contain unwanted oscillations \citep[e.g.,][]{Anninos94}, and so we use an expanded buffer zone around the grid, of size three parent grid boundary zones (so typically six refined zones for a refinement factor of 2).
+The density is computed in this region and the potential solved, but only the region that overlaps with the active region of the grid itself is used to calculate accelerations.
 
-It is known (REF) that simply interpolating the potential without feeding it back to higher levels leads to errors in the potential at more refined levels because of the build-up of errors during the interpolation of coarse boundary values.
+It is known \red{(REF)} that simply interpolating the potential without feeding it back to higher levels leads to errors in the potential at more refined levels due to the build-up of errors during the interpolation of coarse boundary values.
 In addition, neighboring subgrids are not guaranteed to generate the same potential values because of the lack of a coherent potential solve across the whole hierarchy.
@@
 To prevent overshoot, we average the potential on the boundary and allow for (typically) 4 iterations.
-This procedure can help in many cases, but does not necessarily produce a coherent solution across all grids and so does not completely solve the problem; we are working on a slower but more accurate method which does a multi-grid solve across the whole grid (Reynolds \etal, in preparation).
+This procedure can help in many cases, but does not necessarily produce a coherent solution across all grids and so does not completely solve the problem; we are working on a slower but more accurate method that does a multi-grid solve across the whole grid (Reynolds \etal, in preparation).
 

### 2013-05-01 numerical-mhd-ct.tex

@@
 
-Divergence-free AMR is somewhat more complex than the AMR employed elsewhere in
-Enzo.  First, the interpolation must be constrained to be divergence free.
-Thus, all three face-centered field components are interpolated in concert.
-Second, any magnetic field information from previous time steps must be included
-in the interpolation, making the interpolation more complex than the simple
-parent-child relation used for other fields.  Third, the flux correction
-involves more possible grid relations than traditional AMR.  In order to
-circumvent this last complexity, the electric field is projected from fine grids
-to parent grids (rather than the magnetic field), and is then used to re-update
+Divergence-free AMR is somewhat more complex than the AMR employed
+elsewhere in Enzo.  First, the interpolation must be constrained to be
+divergence free.  Thus, all three face-centered field components are
+interpolated in concert.  Second, any magnetic field information from
+previous time steps must be included in the interpolation, making the
+interpolation more complex than the simple parent-child relation used
+for other fields.  Third, the flux correction involves more possible
+grid relations than traditional AMR.  In order to circumvent this last
+complexity, the electric field is projected from fine grids to parent
+grids (rather than the magnetic field), and is then used to re-update
 the parent magnetic field.  This is described in detail in
@@
 
-Enzo's MHD-CT has been used in a variety of contexts, from galaxy cluster
-evolution \citep{Xu11,Xu12}, synthetic observations of radio halos
-\citep{Skillman13},  primordial star formation \citep{Xu08}, and
+Enzo's MHD-CT has been used in a variety of contexts, from galaxy
+cluster evolution \citep{Xu11,Xu12}, synthetic observations of radio
+halos \citep{Skillman13}, primordial star formation \citep{Xu08}, and
 present-day star formation \citep{Collins11,Collins12}.

### 2013-05-01 numerical-ppm.tex

--- 2012-06-29 numerical-ppm.tex
@@
 
-We can examine the structure of eqs.~(\ref{eq:mass})-(\ref{eq:total_energy})
+We can examine the structure of Equations~(\ref{eq:mass})-(\ref{eq:total_energy})
 to obtain a better understanding of the effect of each term.
@@
 
-The first term on the right hand side of eqs. (\ref{eq:momentum}) and
+The first term on the right hand side of Equations~(\ref{eq:momentum}) and
 (\ref{eq:total_energy}) comes from our choice of comoving coordinates
-(a similar term also appears in eq.~(\ref{eq:dm_velocity}), the velocity 
+(a similar term also appears in Equation~(\ref{eq:dm_velocity}), the velocity 
 relation for the dark matter particles).  It does not
@@
 but now apply it spatially.   We can rewrite the one-dimensional (Eulerian) 
-versions of eqs.~(\ref{eq:mass})--(\ref{eq:total_energy}) without expansion terms
+versions of Equations~(\ref{eq:mass})--(\ref{eq:total_energy}) without expansion terms
 in conservative form as,
@@
 peculiar velocity of the baryonic gas, and $g$ is the acceleration at the cell center.
-We write this including the cosmological scale-factors but it reduces to the non-cosmological case with $a = 1$.
+We write this including the cosmological scale factors, but it reduces to the non-cosmological case with $a = 1$.
 These equations are now in a form that can be solved by the split PPM scheme.
@@
 
-We now restrict ourselves to the solution of eqs.
-(\ref{eq:mass1d})--(\ref{eq:total_energy1d}), plus transport, in one dimension.
-In the Direct Eulerian version of PPM, this is accomplished by a three-step process.
-First, we compute effective left and right states at each grid boundary by constructing a piecewise parabolic description of the primary variables ($\rho$, $u$ and $E$) and then averaging over the regions corresponding to the distance each characteristic wave can travel ($u$, $u-c$ and $u+c$).
-Second, a Riemann problem is solved using these effective left and right states, and finally the fluxes are computed based on the solution to this Riemann problem.
-This is described in detail in \citet{1984JCoPh..54..174C}, but we will briefly outline the procedure here both for completeness and to put the changes we will make in context.
+We now restrict ourselves to the solution of Equations
+(\ref{eq:mass1d})--(\ref{eq:total_energy1d}), plus transport, in one
+dimension.  In the Direct Eulerian version of PPM, this is
+accomplished by a three-step process.  First, we compute effective
+left and right states at each grid boundary by constructing a
+piecewise parabolic description of the primary variables ($\rho$, $u$
+and $E$) and then averaging over the regions corresponding to the
+distance each characteristic wave can travel ($u$, $u-c$ and $u+c$).
+Second, a Riemann problem is solved using these effective left and
+right states, and finally the fluxes are computed based on the
+solution to this Riemann problem.  This is described in detail in
+\citet{1984JCoPh..54..174C}, but we will briefly outline the procedure
+here both for completeness and to put the changes we will make in
+context.
 
@@
 for each of $p$, $\rho$, and $v$  The pressure is determined from
-eq.~(\ref{eq:state}), the equation of state.
+Equation~(\ref{eq:eq_of_state}), the equation of state.
 %We have dropped the subscript b in this and the following two sections because we will be referring solely to baryonic quantities.  
@@
 %
-This is eq.~(1.4) of CW84 (in the radial rather than mass coordinate as is appropriate for the direct rulerian approach). 
+This is Equation~(1.4) of CW84 (in the radial rather than mass coordinate as is appropriate for the direct rulerian approach). 
 %The zones edges in the mass coordinates are $m_{j-1/2}$ and $m_{j+1/2}$ for the left and right edges, respectively.  
@@
 The resulting formulae are straightforward but complicated and are not
-reproduced here, but see eqs. 1.7 to 1.10 of \citet{1984JCoPh..54..174C}.
+reproduced here, but see Equations 1.7 to 1.10 of \citet{1984JCoPh..54..174C}.
 We also optionally allow steepening as described in this reference.
@@
 
-In the Eulerian approach, this is more complex because the Eulerian zone-boundary does not naturally correspond to one the three characteristic waves and so a more complicated approach is required.
+In the Eulerian approach, this is more complex because the Eulerian
+zone boundary does not naturally correspond to one the three characteristic waves and so a more complicated approach is required.
 First, the primary quantities ($p$, $\rho$, $v$ and $E$) are averaged over the domains corresponding to the three characteristics $C = u-c$, $u$ or $u+c$ (where $c$ is the sound speed in a cell):
@@
 \end{eqnarray}
-The linearized gas dynamics equations are then used to compute second-order correct left and right states which take into account the multiple wave families.
+The linearized gas dynamics equations are then used to compute second-order correct left and right states that take into account the multiple wave families.
 This process is described in \citet{1984JCoPh..54..174C} and we use their equations (3.6) and (3.7).
@@
 $\overline{p}_{j\pm 1/2}$, $\overline{\rho}_{j\pm1/2}$, and $\overline{v}_{j\pm 1/2}$ that are third order
-accurate in space and second order accurate in time.  These are then used to solve the difference eqs. (\ref{eq:mass_diff})--(\ref{eq:energy_diff}) for $\rho^{n+1}$, $v^{n+1}$, and $E^{n+1}$.
+accurate in space and second order accurate in time.  These are then
+used to solve the difference Equations (\ref{eq:mass_diff})--(\ref{eq:energy_diff}) for $\rho^{n+1}$, $v^{n+1}$, and $E^{n+1}$.
 
-We include an optional diffusive flux (and flattening for the parabolic curves) which can improve the solution in some cases.
+We include an optional diffusive flux (and flattening for the parabolic curves) that can improve the solution in some cases.
 Our implementation follows that in the appendix of \citet{1984JCoPh..54..174C}.
@@
 In addition, as discussed earlier, the three dimensional scheme is achieved by operator splitting and repeating the above procedure in the other two orthogonal directions.
-The transverse velocities and any additional passive quantities are naturally and easily added to this system (see eq.
-3.6 of \citet{1984JCoPh..54..174C}).
+The transverse velocities and any additional passive quantities are naturally and easily added to this system (see Equation 3.6 of \citet{1984JCoPh..54..174C}).
 
-We note that the acceleration required in eq.~(\ref{eq:momentum_diff}) is actually the
+We note that the acceleration required in Equation~(\ref{eq:momentum_diff}) is actually the
 acceleration felt by the entire zone and not just at the zone center.
@@
 
-The system described so far works well for gravitating systems with
+The system described thus far works well for gravitating systems with
 reasonable Mach numbers ($<100$) as long as the structures are well resolved.
 This section and the next detail changes that are required
-to correctly account for situations in which one or both of these
+to correctly account for situations where one or both of these
 requirements are not met.
@@
 if the temperature distribution is required for other reasons
-(e.g. radiative processes), a remedy is required.
+(e.g., for calculating radiative processes), a remedy is required.
 
@@
 \end{equation}
-in comoving co-ordinates.  The structure is similar to the total energy
+in comoving coordinates.  The structure is similar to the total energy
 equation; the second term on the left hand side represents transport, while
@@
 %
-Note that because of the structure of this equation, is not in flux-conservative form.
-In particular, the pressure is evaluated at the cell-center.
-Unfortunately, time-centering of this pressure has proved difficult to do without generating large errors in the internal energy and so we conservatively leave the pressure at the old time in this difference equation.
-This leads to spreading of shocks; however, we note that this equation is only used in hypersonic flows.
+Note that because of the structure of this equation, it is not in
+flux-conservative form.  In particular, the pressure is evaluated at
+the cell center.  Unfortunately, time-centering of this pressure has
+proved difficult to do without generating large errors in the internal
+energy and so we conservatively leave the pressure at the old time in
+this difference equation.  This leads to spreading of shocks; however,
+we note that this equation is only used in hypersonic flows.
 
@@
 the dual energy method {\it will have no dynamical effect}.  We use
-${\eta}_1 = 10^{-3}$ which is consistent with the truncation error of the
+${\eta}_1 = 10^{-3}$, which is consistent with the truncation error of the
 scheme.  We are now free to select the method by which the internal energy

### 2013-05-01 numerical-radiation-fld.tex

--- 2013-03-20 numerical-radiation-fld.tex
@@
 In addition to the ray-tracing approach for radiation transport
-described in Section \ref{sec.num.raytracing}, Enzo currently includes
+described in Section~\ref{sec.num.raytracing}, \enzo\ currently includes
 a field-based radiation transport solver for problems posed on uniform
-(i.e.~non-SAMR grids).  Detailed explanations of the model and
+(i.e.~non-AMR grids).  Detailed explanations of the model and
 solution approach may be found in
@@
 (\ref{eq:fld_radiation}-\ref{eq:fld_heating}) along with the chemical network 
-(eq.~\ref{eq:species_evolution} is solved
+(Equation~\ref{eq:species_evolution}) is solved
 independently of Enzo's hydrodynamics, gravity and dark-matter solvers
-(sections \ref{sec.hydro.ppm}-\ref{sec.ov.nbody}), thereby allowing
-the advective portions of equations (\ref{eq:fld_radiation}) and
-(\ref{eq:fld_chemistry}) to be taken care of by the fluid solvers.
+(Sections \ref{sec.hydro.ppm}-\ref{sec.ov.nbody}), thereby allowing
+the advective portions of Equations (\ref{eq:fld_radiation}) and
+(\ref{eq:species_evolution}) to be taken care of by the fluid solvers.
 Due to the disparate time scales between radiation transport and chemical
@@
 
-We then evolve the system (\ref{eq:fld_heating}-\ref{eq:fld_chemistry}):  
+We then evolve the heating and chemistry system, Equations~(\ref{eq:fld_heating})
+and (\ref{eq:species_evolution}):  
 $(E_r^{n+1}, e_c^n, {\mathrm n}_i^n) \to (E_r^{n+1}, e_c^{n+1},

### 2013-05-01 numerical-radiation-raytracing.tex

@@
 
-We solve the radiative transfer equation in comoving coordinates (given by eq.~\ref{eq:rteqn}).
+We solve the radiative transfer equation in comoving coordinates (given by Equation~\ref{eq:rteqn}).
 We can make some appropriate approximations to reduce the complexity
@@
 \end{equation}
+
 where $P$ is the photon number flux along the ray.  To sample the
@@
 $(r/R)^2$ rays, which is computationally wasteful because only a few
-rays per cell, as we will show later, provides an accurate calculation
-of the radiation field.
+rays per cel  are required to provide an accurate calculation
+of the radiation field (as we will show later).
 
@@
 evenly divides a sphere into equal areas.  The rays are initialized at
-each point source with the photon luminosity (ph s$^{-1}$) equally
+each point source with the photon luminosity (photon s$^{-1}$) equally
 spread across $N_{\rm pix} = 12 \times 4^l$ rays, where $l$ is the
@@
 and (3) gets a quarter of the photon flux of the parent ray.
-Afterward the parent ray is discontinued.
+Afterward, the parent ray is discontinued.
 
-A ray propagates and splits until (1) the photon has traveled $c
+A ray propagates and splits until at least one of the following conditions is met: (1) the photon has traveled $c
 \times dt_P$, where $dt_P$ is the radiative transfer timestep; (2) its
@@
 
-The radiation field is calculated by integrating equation
+The radiation field is calculated by integrating Equation
 (\ref{eqn:rtray}) along each ray, which is done by considering the
@@
 absorption rate, resulting in photon conservation \citep{Abel99_RT,
-  Mellema06}.  Thus the photo-ionization $k_{\rm ph}$ and
-photo-heating $\Gamma_{\rm ph}$ rates associated with a single ray are
+  Mellema06}.  Thus the photo-ionization and
+photo-heating rates associated with a single ray ($k_{\rm ph}$ and
+$\Gamma_{\rm ph}$, respectively)  are
 %
@@
 complete, these rates are used to update the ionization state and
-energy of the cells with the solver described in \S\ref{sec.ov.chem}.
+energy of the cells with the solver described in Section~\ref{sec.ov.chem}.
 

### 2013-05-01 numerical-zeus.tex

@@
-\subsection{The Zeus hydro method}
+\subsection{The \zeus\ hydro method}
 \label{sec.hydro.zeus}
 
-As an alternative to PPM, \enzo\ also includes an implementation of the finite-difference hydrodynamic algorithm employed in the compressible magnetohydrodynamics code \zeus \citep{Stone92a, Stone92b}.
-Fluid transport is solved on a Cartesian grid using the upwind, monotonic advection scheme of \citet{1977JCoPh..23..276V} within a multistep (operator split) solution procedure which is fully explicit in time.
+As an alternative to PPM, \enzo\ also includes an implementation of the finite-difference hydrodynamic algorithm employed in the compressible magnetohydrodynamics code \zeus\ \citep{Stone92a, Stone92b}.
+Fluid transport is solved on a Cartesian grid using the upwind, monotonic advection scheme of \citet{1977JCoPh..23..276V} within a multistep (operator split) solution procedure that is fully explicit in time.
 This method is formally second order-accurate in space but first order-accurate in time.
 
-As discussed in the section describing PPM, operator split methods break the solution of the hydrodynamic equations into parts, with each part representing a single term in the equations.
+As discussed in the section describing the Piecewise Parabolic Method (Section~\ref{sec.hydro.ppm}), operator split methods break the solution of the hydrodynamic equations into parts, with each part representing a single term in the equations.
 Each part is evaluated successively using the results preceding it.
 In this method, in addition to operator splitting the expansion terms (i.e.
-those terms in eqs.~(\ref{eq:mass}) -- (\ref{eq:total_energy}) which depend on $\dot{a}$), we divide the remaining terms into \emph{source} and \emph{transport} steps.
+those terms in eqs.~(\ref{eq:mass}) -- (\ref{eq:total_energy}) that depend on $\dot{a}$), we divide the remaining terms into \emph{source} and \emph{transport} steps.
 The terms to be solved in the source step are those on the right-hand side of eqs.~(\ref{eq:mass}) -- (\ref{eq:total_energy}), while the transport terms are on the left-hand side of these equations and are responsible for the advection of mass, momentum and energy across the grid.
 
-The \zeus\ method uses a von Neumann-Richtmeyer artificial viscosity to smooth shock discontinuities that may appear in fluid flows and can cause a break-down of finite-difference equations.
+The \zeus\ method uses a von Neumann-Richtmeyer artificial viscosity to smooth shock discontinuities that may appear in fluid flows and can cause a breakdown of the finite-difference equations.
 The artificial viscosity term is added in the source terms as:
+
 \begin{eqnarray}
@@
 %where \textbf{v} is the baryon velocity, $\rho$ is the mass density, $p$ is the pressure, $e$ is the internal energy density of the gas and
+
 Here \textbf{Q} is the artificial viscosity stress tensor, which we take to be diagonal with on-axis terms given by $l^2 \rho (\partial v / \partial x)^2$ as proposed by von Neumann \& Richtmyer.
@@
 In our Cartesian coordinate system, the finite difference version of these equations is particularly simple, although there is one important complication.
-In the \zeus\ formalism, the velocity is a face-centered quantity -- that is the velocity is recorded on a grid which is staggered as compared to the density, pressure and energy, which are at the cell center.
-Therefore we must remember that $v_j$ is at position $x_{j-1/2}$ (we use notation, rather than writing $v_{j+1/2}$ both to match the original \zeus\ paper and also to make it easier to compare these equations to what is actually in the code).
+In the \zeus\ formalism, the velocity is a face-centered quantity -- that is, the velocity is recorded on a grid that is staggered as compared to the density, pressure and energy, which are at the cell center.
+Therefore we must remember that $v_j$ is at position $x_{j-1/2}$ (we use this notation, rather than writing $v_{j+1/2}$ both to match the original \zeus\ paper and also to make it easier to compare these equations to what is actually in the code).
 
@@
 Partial updates are denoted by the $n+a$, $n+b$ notation.
-We show only one-dimension as the extension to the multi-dimensional case is straightforward (note that all dimensions are carried out for each substep before progressing to the next substep).
-Then we add the artificial viscosity:
+We show updates in one dimension as the extension to the multi-dimensional case is straightforward (note that all dimensions are carried out for each substep before progressing to the next substep).
+We then add the artificial viscosity:
+
 \begin{eqnarray}
@@
 \end{eqnarray}
+
 The artificial viscosity coefficient $q_j$ is given by:
@@
 Next, we examine the transport step, which is conservative.
-Once again, we dimensionally split the equations and so present only the one-dimensional version.
+Once again, we dimensionally split the equations and present only the one-dimensional version.
 The finite-difference equations actually solved are:
@@
 Note that we advect the specific energy and specific momenta using the mass flux, as dictated by the principle of consistent transport.
-This requires appropriate averaging for the momenta in the perpendicular directions as outline in equations (57)-(72) of \citet{Stone92a}.
+This requires appropriate averaging for the momenta in the perpendicular directions as outlined in equations (57)-(72) of \citet{Stone92a}.
 

### 2013-05-01 overview.tex

--- 2013-04-23 overview.tex
@@
 
-\subsubsection{Godunov MUSCL (HD) with Dender divergence cleaning (MHD)}
+\subsubsection{Godunov MUSCL (HD) with Dedner divergence cleaning (MHD)}
+This solver was developed by Wang \& Abel to attack problems in
+magnetic field amplification during the formation of galaxies
+\citep{Wang:2009a} and to understand the role proto-stellar jets for
+the theory of star formation \citep{Wang:2009b}. It combines the standard
+approach of Godunov \citep{Godunov1959} for finite volume techniques
+with the method of lines as described as e.g. by
+\cite{leveque2002finite} and \cite{toro-1997}. In addition, however,
+it implements the hyperbolic divergence cleaning algorithm of
+\cite{2002JCoPh.175..645D}. It supports multiple approximate Riemann
+solvers and supports non-ideal equations of states. Consequently,
+these suite of solvers can be used for hydro and magneto-hydrodynamic
+simulations. It is also this class of solvers as well as a version of
+the PPM hydro solver which have been ported to Nvidia's CUDA framework
+allowing \enzo to take advantage of modern graphics hardware \citep{Wang:2010}.
 

### 2013-05-02 code-tests.tex
Added WavePool

@@
 
-\red{(BWO: this is missing an explanation of the ICs)}
-
 We begin the paper test suite with the classic Sod shock tube problem
@@
 between the three waves.  These waves include a rarefaction fan, a
-contact discontinuity, and a moderate shock.  In
+contact discontinuity, and a moderate shock.  The initial state is 
+$\rho_{\rm L}, P_{\rm L} = 1.0, 1.0$ on the left of the boundary at 
+$x=0.5$ and $\rho_{\rm R}, P_{\rm R} = 0.125, 0.1$ on the right.
 Figure~\ref{fig.sodshocktube}, we show the density solution at the
-final time for three of our hydro solvers -- the spatially third-order
+final time ($t=0.25$) for three of our hydro solvers -- the spatially third-order
 PPM, as well as the two second-order \zeus\ and MUSCL schemes.  We use
@@
 \label{sec.tests.wavepool}
-\red{(Greg)}
-Problem type 2, with AMR.  Tests reflections of waves at grid boundaries.
+
+In this one-dimensional test, we pass a short wavelength, linear wave through a static refined region.
+The entire domain is from 0 to 1 with a region region with a factor of from 0.25 to 0.75.
+We modify the left boundary consistent with a linear pressure wave $\rho(x,t) = \rho_0 (1 + A \sin(kx - \omega t))$, where the amplitude is $A = 0.01$ and the wavelength $\lambda = k/2\pi = 0.1$.
+The initial density and pressure are set to unity and we use $\gamma = 1.4$.
+The (unrefined) root grid is covered with 100 cells making the wavelength of the linear wave only 10 cells.
+This is, therefore, a challenging problem for hydro methods.
+We are particularly interested in any reflection or artifacts introduced by the wave entering and exiting the refined regions.
+
+In Figure~\ref{fig.wavepool}, we show a 9x9 set of plots showing the evolution of the wave at three different times ($t = 0.2, 0.3$ and 0.8) for three of our solvers (PPM, Zeus, and MUSCL).
+The leftmost column show the wave just before it enters the refined region so that we can gauge how the solver is operating in the absence of AMR, the center column shows the wave after it has fully entered the refined region, and the rightmost column shows the wave well after it has exited the refined region.
+
+Beginning with the PPM, we see that even before it has entered the refine region, the wave is getting clipped, which is not unexpected for such a short wavelength mode, even for PPM.
+The remaining panels demonstrate that the wave cleanly enters and exits the refined region.
+No reflection is seen, either on entry or exit, and the amount of damping is mild.
+
+For the Zeus solver, we see that even before it has entered the refine region, there are small oscillations excited behind the wave; although it is also worth noting that the wave itself is beautifully propagated without significant damping or phase errors.
+In this test, we use only the standard, low amount of quadratic viscosity -- these waves can be damped by additional viscosity, but we do not add any in order to be sensitive to any artifacts at the grid boundary.
+The remaining panels show that the trailing oscillations continue, but do not generate any additional noise -- the end result is very similar to the case without any refined region.
+
+Finally, the MUSCL solver shows also shows a very clear entry and exit from the refined region without any oscillations, although (for the PLM reconstruction used here).
+The wave is spread more than with the other methods, although very little energy is lost.
+
+\begin{figure}
+\begin{center}
+\includegraphics[width=\textwidth]{figures/WavePool.eps}
+\caption{This plot shows, for each column, three snap shots (at $t=0.2, 0.3$ and 0.8) of a linear wave as it propagates through the grid.
+The static refined region extends from $x = 0.25$ to 0.75 and is shown in grey in each panel.
+The individual cells are also color-coded by level (blue is the root grid, and red is for the refined region)  Each row shows the result for a different solver (top is PPM, middle is Zeus, and the bottom is MUSCL).
+The solid line shows the analytic solution for a linear, undamped wave.
+Note that we focus each panel on a small region of the entire domain to better show the wave itself.}
+\label{fig.wavepool}
+\end{center}
+\end{figure}
 

### 2013-05-03 code-tests.tex
Edited WavePool text

--- 2013-04-29 code-tests.tex
@@
 between the three waves.  These waves include a rarefaction fan, a
-contact discontinuity, and a moderate shock.  In
+contact discontinuity, and a moderate shock.  The initial state is 
+$\rho_{\rm L}, P_{\rm L} = 1.0, 1.0$ on the left of the boundary at 
+$x=0.5$ and $\rho_{\rm R}, P_{\rm R} = 0.125, 0.1$ on the right.
+All velocities are initially zero.  In
 Figure~\ref{fig.sodshocktube}, we show the density solution at the
-final time for three of our hydro solvers -- the spatially third-order
+final time ($t=0.25$) for three of our hydro solvers -- the spatially third-order
 PPM, as well as the two second-order \zeus\ and MUSCL schemes.  We use
@@
 refinement factor of 2), triggered based on a normalized slope greater
-than 10\% in the density.  This is a minimal level of refinement, and
+than 10\% in the density.  This refinement criterion results in only
+the refinement of strong gradients, and
 does not include the rarefaction fan at late times.  The results are
@@
 \label{sec.tests.wavepool}
-\red{(Greg)}
-Problem type 2, with AMR.  Tests reflections of waves at grid boundaries.
+
+In this one-dimensional test, we pass a short wavelength, linear wave through a static refined region.
+The full domain is from 0 to 1 and the refined region is set from 0.25 to 0.75 at all times.
+We modify the left boundary consistent with a single linear sound wave with density given by $\rho(x,t) = \rho_0 (1 + A \sin(kx - \omega t))$, where the amplitude is $A = 0.01$ and the wavelength $\lambda = k/2\pi = 0.1$, and similar expressions exist for the pressure and velocity.
+The initial, unperturbed density and pressure are set to unity and we adopt $\gamma = 1.4$.
+The (unrefined) root grid is covered with 100 cells, resulting in a wavelength for the linear wave of only 10 cells.
+This is, therefore, a challenging problem for hydro methods.
+We are particularly interested in any reflection or artifacts introduced by the wave entering and exiting the refined regions.
+
+Figure~\ref{fig.wavepool} shows the evolution of the wave at three different times ($t = 0.2, 0.3$ and 0.8) for three of our solvers (PPM, Zeus, and MUSCL).
+The leftmost column show the wave just before it enters the refined region so that we can gauge how the solver is operating in the absence of AMR, the center column shows the wave after it has fully entered the refined region, and the rightmost column shows the wave well after it has exited the refined region.
+
+Beginning with the PPM, we see that even before it has entered the refine region, the wave is slightly damped, which is not unexpected for such a short wavelength mode, even for PPM.
+The remaining panels demonstrate that the wave cleanly enters and exits the refined region.
+No significant reflection is seen, either on entry or exit, and the amount of damping is mild.
+
+For the Zeus solver, we see that even before it has entered the refined region, there are small oscillations excited behind the wave; although it is also worth noting that the wave itself is beautifully propagated without significant damping or phase errors.
+In this test, we use only our standard, low amount of quadratic artificial viscosity -- these waves can be damped by additional viscosity, but we do not add any in order to be sensitive to any artifacts at the grid boundary.
+The remaining panels show that the trailing oscillations continue, but do not generate any additional noise -- the end result is similar to the case without any refined region.
+
+Finally, the MUSCL solver shows also shows a very clean entry and exit from the refined region without any oscillations, although (for the PLM reconstruction used here) the wave is spread more than with the other methods.
+
+\begin{figure}
+\begin{center}
+\includegraphics[width=0.8\textwidth]{figures/WavePool.eps}
+\caption{This plot shows, for each column, three snap shots (at $t=0.2, 0.3$ and 0.8) of a linear wave as it propagates through the grid.
+The static refined region extends from $x = 0.25$ to 0.75 and is shown in grey in each panel.
+The individual cells are also color-coded by level (blue indicates the root grid, and red is for the refined region)  Each row shows the result for a different solver (top is PPM, middle is Zeus, and the bottom is MUSCL).
+The solid line shows the analytic solution for a linear, undamped wave.
+Note that we focus each panel on a small region of the entire domain to better show the wave itself.}
+\label{fig.wavepool}
+\end{center}
+\end{figure}
 

### 2013-05-03 code-tests.tex
Added ShockPool

@@
 The static refined region extends from $x = 0.25$ to 0.75 and is shown in grey in each panel.
-The individual cells are also color-coded by level (blue indicates the root grid, and red is for the refined region)  Each row shows the result for a different solver (top is PPM, middle is Zeus, and the bottom is MUSCL).
+The individual cells are also color-coded by level: blue indicates the root grid, and red is for the refined region.
+Each row shows the result for a different solver (top is PPM, middle is Zeus, and the bottom is MUSCL).
 The solid line shows the analytic solution for a linear, undamped wave.
@@
 \label{sec.tests.shockpool}
-\red{(Greg)}
-Problem type 3, with AMR.  Tests passage of shock through a refinement boundary.
+
+The next problem is similar to the last one, but examines how a shock wave with a Mach number of 2 passes into and out of a static refined region (again defined from $z=0.25$ to 0.75).
+The density and pressure in the domain is initially set equal to 1.0, with zero velocity, and then at $t=0$, the left boundary is set with the density, pressure and velocity appropriate for $M=2$ shock wave.
+
+In Figure~\ref{fig.shockpool}, we show the evolution of the shock wave at three times, again corresponding to just before entering the refine region (left column), after entering the refined region (center), and after exiting the refined region (right).
+We examine the same set of three solvers as in the previous test problem.
+
+Beginning with PPM (top row), we see that this method captures the shock in a small number of zones with only a small amount of oscillation.
+Open entering the refined region, the shock finds itself broader than the natural width of the scheme (since the cell spacing is decreased by a factor of 2), and so the shock front contracts, causing a slight entropy perturbation in the post-shock gas.
+
+For Zeus (middle row of Figure~\ref{fig.shockpool}, the shock is broadened because of artificial viscosity, and there are slightly more post-shock oscillations, although again quite mild.
+The impact of entering and exiting the refined region is somewhat larger than in PPM; however, the most noticeable difference is the incorrect position of the shock front, which is due to the fact that the scheme is not energy-conserving (see also the Sedov problem, below), and has essentially nothing to do with the passage through the refined region.
+
+Finally, the MUSCL scheme produces a shock which is intermediate in width between the two previous cases.
+This method is energy conserving and so correctly reproduces the shock speed.
+The oscillations are mild except for the cell immediately outside of the refined region upon exiting.
+
+\begin{figure}
+\begin{center}
+\includegraphics[width=0.8\textwidth]{figures/ShockPool}
+\caption{This plot shows, for each column, three snap shots (at $t=0.1, 0.15$ and 0.35) of a $M=2$ shock wave as it propagates through the grid with a static refined region extends from $x = 0.25$ to 0.75 (shown in grey in each panel).
+Each row shows the result for a different solver (PPM/Zeus/MUSCL from top to bottom).
+The solid line shows the analytic solution.
+Note that we focus each panel on a small region of the entire domain to better show the shock front.}
+\label{fig.shockpool}
+\end{center}
+\end{figure}
 

### 2013-05-06 test_problems/TestOrbit/make_energy_plot.py
adding test orbit parameter file and scripts

@@ -1,0 +1,83 @@
+#
+#  This script makes the total energy plot for the TestOrbit parameter type.
+#
+# Note: this script assumes enzo has run the TestOrbit problem with the baryon
+#  fields turned on and the gravitational potential written out.  To do this,
+#  make sure that the following parameters are set to 1 in the enzo parameter
+#  file:
+#
+#    TestOrbitUseBaryons = 1
+#    ComputePotential = 1
+#    WritePotential = 1
+#
+from yt.mods import *
+from yt.utilities.linear_interpolators import *
+import numpy as na
+import matplotlib.pyplot as plt
+
+KE = []
+PE = []
+TE = []
+time = []
+
+it = 0
+
+ts = TimeSeriesData.from_filenames("*/*.hierarchy")
+for pf in ts:
+    
+    # particle position
+    xp = pf.h.grids[0]["particle_position_x"][1]
+    yp = pf.h.grids[0]["particle_position_y"][1]
+    zp = pf.h.grids[0]["particle_position_z"][1]
+    # particle mass
+    xv = pf.h.grids[0]["particle_velocity_x"][1]
+    yv = pf.h.grids[0]["particle_velocity_y"][1]
+    zv = pf.h.grids[0]["particle_velocity_z"][1]
+
+    myKE = 0.5 * (xv**2 + yv**2 + zv**2)
+
+    # calculate potential energy (somewhat laborious; need to use potential value
+    # in grids, hence the gymnastics)
+    fv = {'x':np.array([xp],ndmin=3), 'y':np.array([yp],ndmin=3), 'z':np.array([zp],ndmin=3)}
+    pot = TrilinearFieldInterpolator(pf.h.grids[0]["Grav_Potential"],(0.0, 1.0, 0.0, 1.0, 0.0, 1.0),"xyz",True)
+    myPE = float(pot(fv)[0][0][0])
+
+    myTE = myKE + myPE
+
+    #print myKE, myPE, myTE
+
+    # ignore first value: the PE is garbage since the potential energy field is not
+    # actually calculated before this is written out.
+    if it > 0:
+        KE.append(myKE)
+        PE.append(myPE)
+        TE.append(myTE)
+        time.append(pf.current_time)
+
+    it += 1
+
+# turn into NumPy arrays for convenient manipulation
[…]

### 2013-05-06 test_problems/TestOrbit/make_fig_plot.py

@@ -1,0 +1,116 @@
+#
+#  This script makes most of the plots for the TestOrbit problem type.  The
+#   one that is not made is the total energy plot, which is in another file
+#   for the sake of clarity.
+#
+# Note: this script assumes enzo has been run with "enzo.exe TestOrbit.exe > TestOrbit.out"
+#   (since we need some stuff from TestOrbit.out)
+#
+import numpy as na
+import matplotlib.pyplot as plt
+import os
+from math import *
+
+FileIn = open("TestOrbit.out", "r")
+
+xpos = []
+ypos = []
+zpos = []
+xvel = []
+yvel = []
+zvel = []
+
+# read stuff in from file
+for line in FileIn:
+   if 'id=1' in line:
+      lst = line.split()
+      xpos.append(lst[1])
+      ypos.append(lst[2])
+      zpos.append(lst[3])
+      xvel.append(lst[4])
+      yvel.append(lst[5])
+      zvel.append(lst[6])
+      
+FileIn.close()
+
+# convert to numpy array: easier to manipulate
+x = na.array(xpos)
+y = na.array(ypos)
+z = na.array(zpos)
+xv = na.array(xvel)
+yv = na.array(yvel)
+zv = na.array(zvel)
+
+rad_diff = []
+counter = []
+energy = []
+
+diffweight=0.0
+radsum=0.0
+energysum=0.0
+
+# go through and calculate radius error, also kinetic energy
+for i in range(len(x)):
+   radius = ( ( float(x[i])-0.5)**2 + ( float(y[i])-0.5)**2 + ( float(z[i])-0.5)**2 )**0.5
+   raddiff = (radius-0.3)/0.03125 
+   radsum = radsum + raddiff
+   diffweight = diffweight + 1.0
+   kinetice = 0.5*(float(xv[i])**2 + float(yv[i])**2 + float(zv[i])**2)
+   vmag = (float(xv[i])**2 + float(yv[i])**2 + float(zv[i])**2)**0.5
+   rad_diff.append(raddiff)
+   energy.append(kinetice)
+   energysum += kinetice
+   counter.append(i)
+
+radsum = radsum/diffweight
+energysum /= diffweight
+
+#print "radsum is ", radsum
+#print "energysum is ", energysum
+
+# tweak radius to mean
+rd = na.array(rad_diff)-radsum
+# convert counter to orbit number
[…]

### 2013-05-06 code-tests.tex
Figure shenanigans - changing png to eps.
Added text + figure for the Orbit Test.

@@
 \label{sec.test.testorbit}
-\red{(Greg)}
-This is the TestOrbit problem (problem 29 for GB).  This tests gravity and particle integration (no hydro).
+
+\begin{figure}
+\begin{center}
+\includegraphics[width=0.45\textwidth]{figures/TestOrbit_xy.eps}
+\includegraphics[width=0.45\textwidth]{figures/TestOrbit_TotalEnergy.eps}
+\caption{Test particle behavior in the Orbit problem, which is used to
+test the gravity solver and particle integration.  An effectively
+massless test particle is set in circular orbit around a massive
+central particle, with the gravitational potential calculated on a 3D,
+$32^3$ grid, and evolved forward in time for 200 orbits.  Left panel:
+path of the test particle in the orbital plane plotted for the
+duration of the simulation.  Right panel: total specific energy
+(kinetic plus potential) for the test particle for the duration of the
+simulation.  The particle's mean energy (integrated over the length of
+the simulation) is represented by a horizontal dashed line, and +/-
+one standard deviation by horizontal dotted lines.  All quantities are
+in arbitrary units.}
+\label{fig.orbittest}
+\end{center}
+\end{figure}
+
+The ``orbit test'' simulates the orbit of an effectively massless
+particle around a massive primary particle, and is used to test the
+accuracy of both the gravity solver and particle integration.  In this
+problem (shown in Figure~\ref{fig.orbittest}), a 3D box with unit size
+and $32^3$ grid is created, and a particle with mass 1 is placed at
+the center of the box (both box length and particle mass are in
+arbitrary units).  A second ``test'' particle, with mass of $10^{-6}$,
+is placed at a distance of 0.3 L$_{box}$ from the primary particle in
+the x-y plane (at $z=0.5$~L$_{box}$).  Both particles are given
+velocities such that they will orbit their common centers of mass in
+the x-y plane.  The gravitational potential is calculated using a fast
+Fourier Transform on the $32^3$ grid, and the simulation is integrated
+forward in time for 200 particle orbits, writing out a dataset
+approximately once per orbit (with the particle positions being
+written out several hundred times per orbit).
+
+The left panel of Figure~\ref{fig.orbittest} shows the path of the
+test particle in the x-y plane (at z=$0.5$), plotted over the duration
+of the simulation (approximately 200 orbits).  The test particle
+maintains a path that has effectively unnoticeable deviation from
+circularity, and with no discernable drift in either net radius or
+orbit center over time.  The right panel displays the total specific
+energy of the test particle (defined as $\frac{1}{2} {\bf v}^2 +
+\phi$, where $\phi$ is the gravitational potential at the position of
+the test particle) as a function of time, plotted once per orbit.  The
+mean total specific energy is $-1.76797$ in arbitrary units, with a
+standard deviation of $0.004885$ ($0.276\%$ of the mean value), and
+with a maximum deviation from the mean of $0.546\%$ of the mean value.
+Given that the potential is calculated on $32^3$ cells, the two
+particles are separated by between 10 and 14 cells depending on their
+relative positioning in the grid, and the particles are integrated
+using a second-order leapfrog method, this level of accuracy is
+roughly what one might expect.
 

### 2013-05-07 test_problems/SedovAMR/sedov.py
added files for Sedov test problem

@@ -1,0 +1,113 @@
+#!/usr/bin/python
+
+import numpy as na
+
+time   = 0.07       # problem time 
+npoints= 200
+nu     = 2.0       # cylindrical symmetry       
+gamma  = 1.4       # adiabatic index                                                                                                      
+alpha  = 1.0       # alpha(gamma, nu), see p. 231                                                                      
+                   # alpha(1.4, 3) = 0.851                                                       
+                   # alpha(1.4, 2) = 1.0 (an approximate                                                            
+                   # value taken from a plot in Sedov 1959)                                                            
+                   # alpha(1.4, 1) = 1.1 (also an approximation)                                                                   
+
+E_0    = 10.0      # explosion energy [per unit length (area) if nu=2 (=1)]                                                      
+E      = E_0/alpha # dimensional constant #1                                                                                   
+rho1   = 1.0       # ambient density (dimensional const #2)                                                                      
+
+endpoints = 50
+
+# useful combinations involving gamma and nu                                                                                      
+
+gm1    = gamma - 1.
+gp1    = gamma + 1.
+nup2   = nu + 2.
+
+#;;;;;;;;;;;;;;;;;;;;;;;;;;;;;;;;;;;;;;;;;;;;;;;;;;;;;;;;;;;;;;;;;;;                                                              
+
+#; shock wave position and velocity [Sedov 1959, p. 213, eqs (11.3)]                                                              
+
+#;;;;;;;;;;;;;;;;;;;;;;;;;;;;;;;;;;;;;;;;;;;;;;;;;;;;;;;;;;;;;;;;;;;                                                              
+
+r2     = (E/rho1)**(1./nup2)*time**(2./nup2)
+v_shock= 2./nup2*(E/rho1)**(1./nup2)*time**(-nu/nup2)
+
[…]

### 2013-05-07 test_problems/SedovAMR/sedov_outputdata.py

@@ -1,0 +1,35 @@
+from yt.mods import *
+import matplotlib.pyplot as plt
+
+# Load the dataset.
+#pf = load("../100amr4/ppm/DD0007/sb_L2x2_amr4_0007")
+pf = load("../100amr4/zeus/DD0007/sb_L2x2_amr4_zeus_0007")
+
+pc = PlotCollection(pf, center=[0.5, 0.5, 0.5])
+
+pc.add_profile_sphere(0.5, "cm", ["Radius", "CellMass"], weight=None, x_bins=100, x_log=False)
+mass_data = pc.plots[-1].data["CellMass"]
+radius_data =  pc.plots[-1].data["Radius"]
+
+pc.add_profile_sphere(0.5, "cm", ["Radius", "CellVolume"], weight=None, x_bins=100, x_log=False)
+volume_data = pc.plots[-1].data["CellVolume"]
+
+density_data = mass_data/volume_data
+
+pc.add_profile_sphere(0.5, "cm", ["Radius", "RadialVelocity"], weight="CellMass", x_bins=100, x_log=False)
+velocity_data = pc.plots[-1].data["RadialVelocity"]
+
+pc.add_profile_sphere(0.5, "cm", ["Radius", "Pressure"], weight="CellMass", x_bins=100, x_log=False)
+pressure_data = pc.plots[-1].data["Pressure"]
+
+pc.add_profile_sphere(0.5, "cm", ["Radius", "ThermalEnergy"], weight="CellMass", x_bins=100, x_log=False)
+internalenergy_data = pc.plots[-1].data["ThermalEnergy"]
+
+
+# output profiles
+
+file = open("sedov_zeus_output.dat", 'w')
+print >> file, "# radius density internal-energy pressure velocity"
+for i in range(len(radius_data)):
+    print >> file, "%f\t%f\t%f\t%f\t%f" % (radius_data[i], density_data[i], internalenergy_data[i], pressure_data[i], velocity_data[i])
+file.close()

### 2013-05-07 test_problems/SedovAMR/sedov_plot.py

@@ -1,0 +1,42 @@
+from yt.mods import *
+import matplotlib.pyplot as plt
+
+# Load the dataset.
+#pf = load("../100amr4/ppm/DD0007/sb_L2x2_amr4_0007")
+pf = load("../100amr4/zeus/DD0007/sb_L2x2_amr4_zeus_0007")
+
+pc = PlotCollection(pf, center=[0.5, 0.5, 0.5])
+
+pc.add_profile_sphere(0.5, "cm", ["Radius", "CellMass"], weight=None, x_bins=100, x_log=False)
+mass_data = pc.plots[-1].data["CellMass"]
+radius_data =  pc.plots[-1].data["Radius"]
+
+pc.add_profile_sphere(0.5, "cm", ["Radius", "CellVolume"], weight=None, x_bins=100, x_log=False)
+volume_data = pc.plots[-1].data["CellVolume"]
+
+density_data = mass_data/volume_data
+
+
+# read in anlytical solution
+sedovsol_radius = []
+sedovsol_density = []
+file = open("sedov.in", "r")
+for line in file:
+    linesplit = line.split('\t')
+    if linesplit[0] != '#':
+        sedovsol_radius.append(float(linesplit[0]))
+        sedovsol_density.append(float(linesplit[2]))
+        
+
+# python plot                                                                                                                     
+plotfig = plt.figure()
+
+# normalised mass by total mass                                                                                                   
+plt.plot(radius_data, density_data, 'ro', sedovsol_radius, sedovsol_density)
+
+plt.xlabel('Radius')
+plt.ylabel('Density')
+
+# save python plot as png file                                                                                                    
+plt.savefig("sedov_density_plot")
+

### 2013-05-07 test_problems/SedovAMR/sedov_slice.py

@@ -1,0 +1,13 @@
+from yt.mods import * 
+
+#fn = "../100amr4/ppm/DD0007/sb_L2x2_amr4_0007"
+fn = "../100amr4/zeus/DD0007/sb_L2x2_amr4_zeus_0007"
+
+pf = load(fn)
+pc = PlotCollection(pf, center=[0.5, 0.5, 0.5])
+
+p = pc.add_slice("Density", 2)
+#p.modify["grids"]()
+
+pc.save(fn)
+

### 2013-05-07 test_problems/CosmoIFront/makeplots.py
test problem data for the cosmological i-front test

@@ -1,0 +1,289 @@
+# matplotlib-based plotting script for Shapiro & Giroux q=0.5 test
+# Daniel R. Reynolds, reynolds@smu.edu
+
+# imports
+from pylab import *
+
+
+# set the total number of snapshots
+te = 81
+
+# set the graphics output type
+#pictype = '.png'
+#pictype = '.pdf'
+pictype = '.eps'
+
+# set some constants
+q0 = 0.05              # deceleration parameter
+Nph = 5.0e48           # ionization source strength [photons/sec]
+alpha2 = 2.52e-13      # recombination rate coefficient
+mp = 1.67262171e-24    # proton mass [g]
+Myr = 3.15576e13       # duration of a Megayear [sec]
+
+# initialize time-history outputs
+#    row 1: i-front radius
+#    row 2: stromgren sphere radius (rs)
+#    row 3: redshift (z)
+#    row 4: time (t)
+#    row 5: i-front radius (analytical)
+rdata = zeros( (5, te+1), dtype=float);
+
+
+##########
+# define some helpful functions
+def get_params(file):
+    """Returns z0, z, xR, t0, H0, dUnit, tUnit, lUnit from a parameter files"""
+    import shlex
+    f = open(file)
+    for line in f:
+        text = shlex.split(line)
+        if ("CosmologyInitialRedshift" in text):
+            z0 = float(text[len(text)-1])
+        elif ("CosmologyCurrentRedshift" in text):
+            z = float(text[len(text)-1])
+        elif ("InitialTime" in text):
+            t0 = float(text[len(text)-1])
+        elif ("CosmologyHubbleConstantNow" in text):
+            H0 = float(text[len(text)-1])
+    f.close()
+    f = open(file + '.rtmodule' )
+    for line in f:
+        text = shlex.split(line)
+        if ("DensityUnits" in text):
+            dUnit = float(text[len(text)-1])
+        elif ("TimeUnits" in text):
+            tUnit = float(text[len(text)-1])
+        elif ("LengthUnits" in text):
+            lUnit = float(text[len(text)-1])
+    f.close()
+    xR = lUnit
[…]

### 2013-05-07 test_problems/DoubleMachReflection/newplot.py
files for double mach reflection test problem

@@ -1,0 +1,16 @@
+from yt.mods import *
+
+for n in range(103,104):
+
+    pf = load("DD%04i/DD%04i" % (n,n))
+
+    c = [2.0,0.5,0]
+
+    buff = [120,480]  # this is backward because matplotlib transposes buffers.
+
+    slice = SlicePlot(pf, 'z', "Density", center=c, width=((4.0, '1'), (1.0, '1')), origin='domain', axes_unit=('1','1')) 
+    slice.set_buff_size(buff)
+    #slice.set_window_size(4.8)
+    slice.save()
+
+    #slice.save(mpl_kwargs={'bbox_inches':'tight'})

### 2013-05-07 test_problems/AnisoConduction/newplot.py
anisotropic conduction test problem

@@ -1,0 +1,27 @@
+from yt.mods import *
+
+for n in range(300,301):
+    
+    pf = load("DD%04i/DD%04i" % (n,n))
+
+    dd = pf.h.all_data()
+
+    print "----------------------------------------------------"
+    print "output number", n
+    print " "
+    print "Temperature:"
+    print dd["Temperature"].min()
+    print dd["Temperature"].max()
+
+    #print "Density:"
+    #print dd["Density"].min()
+    #print dd["Density"].max()
+
+    slice = SlicePlot(pf, 'z', "Temperature", width = (1.0, '1'), origin='domain', axes_unit=('1','1'))  #.save()
+    #slice = SlicePlot(pf, 'z', "Temperature", width = (1.0, '1'), origin='domain', axes_unit=None)  #.save()
+
+    slice.save(mpl_kwargs={'bbox_inches':'tight'})
+
+    #slice.save("testplot.png")
+
+# the end

### 2013-05-07 development.tex
More description of the development process.

@@
 
-The main development streams of \enzo\ were eventually merged using
+A version developed at Penn State and Stanford forked from a version
+prior to the LCA version and was the one in which MHD with Dedener
+divergence cleaning, the MUSCL hydro solvers, the ray tracing
+radiative transfer module, relativistic fluid dynamics, the shearing
+box boundaries and updates to the multi-species chemistry were
+included. This version was the one that was eventually merged using
 the distributed version control system (DVCS) Mercurial
@@
 as \texttt{week-of-code}, so named after the in-person development
-sprint at which it was created.  The fundamental, and transformative,
-distinction between the previous \textit{centralized} version control
-system and mercurial's \textit{distributed} version control system is
-the elimination of gatekeepers.  While there still exists a canonical,
-central location where stable and development versions of \enzo\ can
-be obtained, changesets and versions can be exchanged between peers
-without the intervention of designated gatekeepers.  This has the
-direct effect of enabling local development to be versioned and its
-provenance ensured, while still retaining the ability to benefit from
-``upstream'' development.  An important, even crucial, side effect is
-that the technology used for local versioning provides mechanisms for
-easily submitting locally-developed modifications to the community
-source location.  Mercurial internally represents all changes as nodes
-in a directed, acyclic graph (DAG), which results in the natural
-ability to more consistently and easily manage merging development
-streams.
+sprint at KIPAC in June 2009 at which it was created.  The
+fundamental, and transformative, distinction between the previous
+\textit{centralized} version control system and mercurial's
+\textit{distributed} version control system is the elimination of
+gatekeepers.  While there still exists a canonical, central location
+where stable and development versions of \enzo\ can be obtained,
+changesets and versions can be exchanged between peers without the
+intervention of designated gatekeepers.  This has the direct effect of
+enabling local development to be versioned and its provenance ensured,
+while still retaining the ability to benefit from ``upstream''
+development.  An important, even crucial, side effect is that the
+technology used for local versioning provides mechanisms for easily
+submitting locally-developed modifications to the community source
+location.  Mercurial internally represents all changes as nodes in a
+directed, acyclic graph (DAG), which results in the natural ability to
+more consistently and easily manage merging development streams.
 

### 2013-05-07 acknowledgments.tex
added some acknowledgements. Should we also try to list all the grant numbers?

@@
 
-Development of \enzo\ has been supported by the National Science Foundation, the National Aeronautics and Space Administration, the Department of Energy, the University of Illinois and National Center for Supercomputing Applications, UC San Diego and the San Diego Supercomputing Center, Columbia University, Georgia Tech, Michigan State University, ...
-\red{We all need to add to the acknowledgments!}
+Development of \enzo\ has been supported by the National Science
+Foundation, the National Aeronautics and Space Administration, the
+Department of Energy, the University of Illinois and National Center
+for Supercomputing Applications, UC San Diego and the San Diego
+Supercomputing Center, Columbia University, Georgia Tech, Michigan
+State University, Pennsylvania State University, Stanford University,
+SLAC National Accelerator Laboratory, Kavli Institute for Particle
+Astrophysics and Cosmology, the Kavli Institute for Theoretical
+Physics at Santa Barbara, ...  \red{We all need to add to the acknowledgments!}

### 2013-05-07 test_problems/PhotonShadowing/make-plot.py
Adding photon shadowing parameter file and plotting script.

@@ -1,0 +1,22 @@
+from yt.mods import *
+import yt.visualization.eps_writer as EPS
+
+output_num = 150
+#fields = ["Density", "HI_Fraction", "Pressure", "Temperature"]
+fields = ["Pressure", "Neutral_Fraction", "Density", "Temperature"]
+cmap = ["algae", "gist_stern", "algae", "hot"]
+zlim = [(2e-15, 2e-13), (1e-4, 1), (1e-28, 1e-25), (1e4, 1e5)]
+
+fn = "DD%4.4d/data%4.4d" % (output_num, output_num)
+pf = load(fn)
+
+pc = PlotCollection(pf, center=[0.5, 0.5, 0.4990234375])
+for i,f in enumerate(fields):
+    p = pc.add_slice(f, 'y', use_colorbar=False)
+    p.set_cmap(cmap[i])
+    p.set_zlim(zlim[i][0], zlim[i][1])
+
+ep = EPS.multiplot_yt(2,2,pc, margins=(0.2,0.2), bare_axes=True)
+ep.scale_line(1.0/6.6, '1 kpc')
+ep.title_box("15 Myr")
+ep.save_fig("code-test-shadowing")

### 2013-05-09 test_problems/OneZoneFreefallTest/plot.py
Adding scripts and par files for OneZoneFreefallTest.

@@ -1,0 +1,70 @@
+from collections import defaultdict
+from matplotlib import pyplot
+import numpy as np
+import sys
+
+from yt.mods import *
+
+def get_tracks(filenames, fields=None):
+    if fields is None: fields = ['NumberDensity', 'Temperature']
+    ts = TimeSeriesData.from_filenames(filenames)
+    data = defaultdict(list)
+    for pf in ts:
+        print pf.fullpath
+        for field in fields:
+            data[field].append(pf.h.grids[0][field][0, 0, 0])
+    for field in fields:
+        data[field] = np.array(data[field])
+    del ts
+    return data
+
+dirs = ['mf', 'Z-5', 'Z-4', 'Z-3', 'Z-2']
+labels = ['0', '10$^{-5}$',  '10$^{-4}$',  '10$^{-3}$',  '10$^{-2}$']
+colors = ['black', 'blue', 'green', 'orange', 'red']
+
+my_data = {}
+for my_dir in dirs:
+    my_data[my_dir] = get_tracks("%s/DD*/*.hierarchy" % my_dir)
+
+fontsize = 14
+
+n_rows = 1
+n_columns = 1
+
+# blank space between edge of figure and active plot area
+top_buffer = 0.01
+bottom_buffer = 0.1
+left_buffer = 0.09
+right_buffer = 0.04
+
+# blank space between plots
+hor_buffer = 0.05
+vert_buffer = 0.05
+
+# calculate the height and width of each panel
+panel_width = ((1.0 - left_buffer - right_buffer - 
+                ((n_columns-1)*hor_buffer)) / n_columns)
+panel_height = ((1.0 - top_buffer - bottom_buffer - 
+                 ((n_rows-1)*vert_buffer)) / n_rows)
+
+# create a figure (figsize is in inches
+pyplot.figure()
+
+my_axes = pyplot.axes((left_buffer, bottom_buffer, 
+                       panel_width, panel_height))
+    
+for i, my_dir in enumerate(dirs):
+    my_axes.loglog(my_data[my_dir]['NumberDensity'],
+                   my_data[my_dir]['Temperature'],
+                   color=colors[i], linestyle='-',
+                   linewidth=1.5, alpha=0.7, label=labels[i])
+my_axes.set_xlim(1, 1e12)
+my_axes.set_ylim(10, 2e3)
+my_axes.yaxis.set_label_text("T [K]", fontsize=fontsize)
[…]

### 2013-05-10 acknowledgments.tex
updated acknowledgments.

--- 2013-04-24 acknowledgments.tex
@@
 
-Development of \enzo\ has been supported by the National Science Foundation, the National Aeronautics and Space Administration, the Department of Energy, the University of Illinois and National Center for Supercomputing Applications, UC San Diego and the San Diego Supercomputing Center, Columbia University, Georgia Tech, Michigan State University, ...
-\red{We all need to add to the acknowledgments!}
+Development of \enzo\ has been ongoing since 1994 by a wide range of
+agencies and institutions.  In all grants listed, we put the
+initials of the PI (if an \enzo\ developer) or the \enzo\ developer
+funded by the grant (if the PI is not a developer of \enzo).
+
+This work has been supported by the National Science Foundation by
+grants
+AAG-0808184 (DRR),
+AAG-1109008 (DRR),
+ACI-9619019 (MLN),
+ASC-9313135 (MLN),
+AST-9803137 (MLN), 
+AST-0307690 (MLN), 
+AST-0407176 (RC),
+AST-0407368 (SS, EJH),
+AST-0507521 (RC) 
+AST-0507717 (MLN), 
+AST-0507768 (AK),
+AST-0607675 (AK)
+AST-0702923 (EJH),
+AST-0707474 (BDS), 
+AST-0708960 (MLN), 
+AST-0807215 (JB),
+AST-0808184 (MLN, AK),
+AST-0908819 (BWO) 
+AST-0955300 (NG),
+AST-1008134 (GB) 
+AST-1009802 (JSO), 
+AST-1102943 (MLN)
+AST-1106437 (JB),
+AST-1210890 (GB),
+AST-1211626 (JW),
+OCI-0832662 (BWO, MLN),
+OCI-0941373 (BWO),
+PHY-1104819 (MLN, JB),
+the CI TraCS fellowship (OCI-1048505; MJT),
+and the Graduate Research Fellowship program (NG; SJS).
+
+This work has been supported by the National Aeronautics and Space
+Administration through grants
+NAGW-3152 (MLN),
+NAG5-3923 (MLN),
+NNX08AH26G (MLN),
+NNX09AD80G (BWO),
+NNX12AH41G (GB),
+NNX12AC98G (BWO),
+NNZ07-AG77G (BDS),
+NNG05GK10G (RC),
+ATP09-0094 (SVL),
+Chandra Theory grant \#TM9-0008X (BWO),
+Hubble Space Telescope Theory Grant HST-AR-10978.01 (BDS),
+Fermi Guest Investigator Program (\#21077; BWO),
+and the Hubble Postdoctoral Fellowship through the Space Telescope Science
+Insititue (120-6370; JW).
+
+This work has been supported by the Department of Energy via
+Los Alamos National Laboratory (LANL) Laboratory Directed Research and
+Development Program (BWO, DCC, HX), 
+the LANL Institute for Geophysics and Planetary Physics (BWO, DCC),
+Los Alamos National Lab Director's Fellowship (No. DE-AC52-06NA25396;
+BWO and DCC), and the
+DOE Computational Science Graduate Fellowship (DE-FG02-97ER25308; SJS)
+
+Additional financial support for the \enzo\ code has come from
+Canada's NSERC through the USRA and CGS programs (EL) and through a 
+Japan MEXT grant for the Tenure Track System (EJT).
+
+We acknowledge the  many institutions that have supported \enzo\
+development, including (in alphabetical order)
+the
+Columbia University,
+Georgia Tech,
+Michigan State University and the MSU Institute for Cyber-Enabled
+Research, 
+the National Center for Supercomputing Applications, 
+the Pennsylvania State University
+ the San Diego Supercomputing Center (through the Strategic Applications
+Partner program and the Director’s office),
+SLAC National Accelerator Laboratory,
+the SLAC/Stanford Kavli Institute for Particle
+Astrophysics and Cosmology,  Stanford University,
+University of Arizona,
+University of Califoria at San Diego, 
+ University of Florida,
+the University of Illinois.
+We acknowledge support from the Kavli Institute for Theoretical
+Physics at Santa Barbara and the Aspen Center for Physics, who
+generously hosted \enzo\ developers.
+
+Computational resources for \enzo\ development have come from the NSF
+XSEDE program, NASA high performance computing, the DOE INCITE
+program, and the DOE Advanced Simulation and Computing (ASC) program.
+
+The \enzo\ collaboration would also like to acknowledge the significant contributions to
+\enzo\ development made by the late Dr. Robert P. Harkness.

### 2013-05-10 acknowledgments.tex
Some udpates to the acknowledgments.

--- 2013-04-24 acknowledgments.tex
@@
+%  If you are editing this file to add acknowledgments, please note that
+%  some of the grant number formats have been edited a bit (from, say,
+%  08-08184 to 0808184) to ensure consistency between different
+%  grants.  I ask that you please adhere to the current formatting,
+%  etc. as well.  Thanks!  --Brian
+
 \acknowledgments
 
-Development of \enzo\ has been supported by the National Science Foundation, the National Aeronautics and Space Administration, the Department of Energy, the University of Illinois and National Center for Supercomputing Applications, UC San Diego and the San Diego Supercomputing Center, Columbia University, Georgia Tech, Michigan State University, ...
-\red{We all need to add to the acknowledgments!}
+Development of \enzo\ has been ongoing since 1994 by a wide range of
+agencies and institutions.  In all grants listed, we put the
+initials of the PI (if an \enzo\ developer) or the \enzo\ developer
+funded by the grant (if the PI is not a developer of \enzo).
+
+This work has been supported by the National Science Foundation by
+grants
+AAG-0808184 (DRR),
+AAG-1109008 (DRR),
+ACI-9619019 (MLN),
+ASC-9313135 (MLN),
+AST-9803137 (MLN), 
+AST-0307690 (MLN), 
+AST-0407176 (RC),
+AST-0407368 (SS, EJH),
+AST-0507521 (RC) 
+AST-0507717 (MLN), 
+AST-0507768 (AK),
+AST-0607675 (AK)
+AST-0702923 (EJH),
+AST-0707474 (BDS), 
+AST-0708960 (MLN), 
+AST-0807215 (JB),
+AST-0808184 (MLN, AK),
+AST-0908819 (BWO) 
+AST-0955300 (NG),
+AST-1008134 (GB) 
+AST-1009802 (JSO), 
+AST-1102943 (MLN)
+AST-1106437 (JB),
+AST-1210890 (GB),
+AST-1211626 (JW),
+OCI-0832662 (BWO, MLN),
+OCI-0941373 (BWO),
+PHY-1104819 (MLN, JB),
+the CI TraCS fellowship (OCI-1048505; MJT),
+and the Graduate Research Fellowship program (NG; SJS).
+
+This work has been supported by the National Aeronautics and Space
+Administration through grants
+NAGW-3152 (MLN),
+NAG5-3923 (MLN),
+NNX08AH26G (MLN),
+NNX09AD80G (BWO),
+NNX12AH41G (GB),
+NNX12AC98G (BWO),
+NNZ07-AG77G (BDS),
+NNG05GK10G (RC),
+ATP09-0094 (SVL),
+Chandra Theory grant \#TM9-0008X (BWO),
+Hubble Space Telescope Theory Grant HST-AR-10978.01 (BDS),
+the Fermi Guest Investigator Program (\#21077; BWO),
+and the Hubble Postdoctoral Fellowship through the Space Telescope Science
+Insititue, \#120-6370 (JW).
+
+This work has been supported by the Department of Energy via the
+Los Alamos National Laboratory (LANL) Laboratory Directed Research and
+Development Program (BWO, DCC, HX, SJS), 
+the LANL Institute for Geophysics and Planetary Physics (BWO, DCC, CP,
+BC),
+the Los Alamos National Laboratory Director's Postdoctoral Fellowship
+program (No. DE-AC52-06NA25396;
+BWO and DCC), and the
+DOE Computational Science Graduate Fellowship (DE-FG02-97ER25308; SJS)
+
+Additional financial support for the \enzo\ code has come from
+Canada's NSERC through the USRA and CGS programs (EL) and through a 
+Japan MEXT grant for the Tenure Track System (EJT).
+
+We acknowledge the  many academic institutions that have supported \enzo\
+development, including (in alphabetical order)
+Columbia University,
+Georgia Tech,
+Michigan State University and the MSU Institute for Cyber-Enabled
+Research, 
+the National Center for Supercomputing Applications, 
+the Pennsylvania State University,
+ the San Diego Supercomputing Center (through the Strategic Applications
+Partner program and the Director’s office),
+SLAC National Accelerator Laboratory,
+the SLAC/Stanford Kavli Institute for Particle
+Astrophysics and Cosmology,  
+Stanford University,
+the University of Arizona,
+the University of Califoria at San Diego, 
+the University of Florida,
+and the University of Illinois.
+We acknowledge support from the Kavli Institute for Theoretical
+Physics at Santa Barbara, the Aspen Center for Physics, and the UCLA
+Institute for Pura and Applied Mathematics, which have
+generously hosted \enzo\ developers through their conference and
+workshop programs.
+
+Computational resources for \enzo\ development have come from the NSF
+XSEDE program, the NASA High Performance Computing program, the DOE INCITE
+program, and the DOE Advanced Simulation and Computing (ASC) program.
+
+The \enzo\ collaboration would also like to acknowledge the significant contributions to
+\enzo\ development made by the late Dr. Robert P. Harkness.

### 2013-05-10 acknowledgments.tex
updated nathan's initials

@@
 AST-0908819 (BWO) 
-AST-0955300 (NG),
+AST-0955300 (NJG),
 AST-1008134 (GB) 
@@
 the CI TraCS fellowship (OCI-1048505; MJT),
-and the Graduate Research Fellowship program (NG; SJS).
+and the Graduate Research Fellowship program (NJG; SJS).
 

### 2013-05-10 acknowledgments.tex
fixed typos

@@
 the Pennsylvania State University,
- the San Diego Supercomputing Center (through the Strategic Applications
+ the San Diego Supercomputer Center (through the Strategic Applications
 Partner program and the Director’s office),
@@
 Physics at Santa Barbara, the Aspen Center for Physics, and the UCLA
-Institute for Pura and Applied Mathematics, which have
+Institute for Pure and Applied Mathematics, which have
 generously hosted \enzo\ developers through their conference and

### 2013-05-10 acknowledgments.tex
added CU/Boulder.

@@
 the University of Califoria at San Diego, 
+the University of Colorado at Boulder,
 the University of Florida,

### 2013-05-10 acknowledgments.tex
updated acknowledgments to include grants and fix many tragic and hilarious typos.

--- 2013-05-10 acknowledgments.tex
@@
 AST-0507768 (AK),
-AST-0607675 (AK)
+AST-0529734 (TA),
+AST-0607675 (AK),
 AST-0702923 (EJH),
@@
 AST-0708960 (MLN), 
+AST-0807075 (TA),
 AST-0807215 (JB),
 AST-0808184 (MLN, AK),
+AST-0808398 (TA),
 AST-0908819 (BWO) 
@@
 AST-1210890 (GB),
-AST-1211626 (JW),
+AST-1211626 (JHW),
 OCI-0832662 (BWO, MLN),
@@
 the CI TraCS fellowship (OCI-1048505; MJT),
-and the Graduate Research Fellowship program (NJG; SJS).
+and the Graduate Research Fellowship program (NJG; SWS).
 
@@
 NAG5-3923 (MLN),
-NNX08AH26G (MLN),
+NNX08AH26G (MLN, TA),
 NNX09AD80G (BWO),
@@
 and the Hubble Postdoctoral Fellowship through the Space Telescope Science
-Insititue, \#120-6370 (JW).
+Insititue, \#120-6370 (JHW).
 
@@
 Los Alamos National Laboratory (LANL) Laboratory Directed Research and
-Development Program (BWO, DCC, HX, SJS), 
+Development Program (BWO, DCC, HX, SWS), 
 the LANL Institute for Geophysics and Planetary Physics (BWO, DCC, CP,
@@
 BWO and DCC), and the
-DOE Computational Science Graduate Fellowship (DE-FG02-97ER25308; SJS)
+DOE Computational Science Graduate Fellowship (DE-FG02-97ER25308; SWS)
 
@@
 Columbia University,
-Georgia Tech,
+Georgia Institute of Technology,
 Michigan State University and the MSU Institute for Cyber-Enabled
@@
 Astrophysics and Cosmology,  
+Southern Methodist University,
 Stanford University,

### 2013-05-11 acknowledgments.tex
Adding a bunch of commas and another grant.

@@
 AST-0407368 (SS, EJH),
-AST-0507521 (RC) 
+AST-0507521 (RC), 
 AST-0507717 (MLN), 
@@
 AST-0808398 (TA),
-AST-0908819 (BWO) 
+AST-0908740 (AK, DCC),
+AST-0908819 (BWO), 
 AST-0955300 (NJG),
-AST-1008134 (GB) 
+AST-1008134 (GB), 
 AST-1009802 (JSO), 
-AST-1102943 (MLN)
+AST-1102943 (MLN),
 AST-1106437 (JB),

### 2013-05-14 overview.tex
Small change.
dcollins: added MHD-CT overview

@@
 \cite{2002JCoPh.175..645D}. It supports multiple approximate Riemann
-solvers and supports non-ideal equations of states. Consequently,
+solvers and non-ideal equations of states. Consequently,
 these suite of solvers can be used for hydro and magneto-hydrodynamic
@@
 \subsubsection{MHD with Constrained Transport (MHD)}
+This MHD method is second order in time and space, and preserves the divergence constraint,
+$\nabla \cdot \vecB = 0$, to machine precision through the Constrained Transport
+(CT) method.  CT, originally described by \citet{Evans88}, updates the
+magnetic field with the curl of an electric field, suitably formulated to
+preserve $\nabla \cdot \vecB$.  We employ the CT methods
+described by \citet{Balsara99} and \citet{Gardiner05}.  Further, we employ the
+second order hyperbolic solver of \citet{Li08} and the divergence free AMR
+scheme of \citet{Balsara01}.
 

### 2013-05-14 numerical-mhd-ct.tex
adding dual energy, cosmology, and a better intro

@@
 
-Physically, the divergence of the magnetic field, \Divb, is
+This MHD method was developed by \citet{Collins10} in order to study Population
+III stars \citep{Xu08}, molecular clouds \citep{collins12}, and galaxy clusters
+\citep{Xu12, Skillman13}.  A full description and suite of test problems can be
+found in \citet{Collins10}
+
+The divergence of the magnetic field, \Divb, is
 identically zero due to the fact that the evolution of the magnetic field is the
@@
 
+Cosmological MHD can be treated in a number of ways.  As described in Section
+\red{reference}, the MUSCL with Dedner treats the cosmological dilution of
+$\vecB$ with a source term, 
+$$ 
+\frac{\partial \vecB}{\partial t} - \frac{1}{a}  \nabla \times (\vecv \times
+\vecB)  =  -\frac{\dot{a}}{2 a} \vecB. \label{eq:induction_cosmo_source}
+$$
+However, one can eliminate the right hand side of Equation 
+\ref{eq:induction_cosmo_source} by defining
+$\vecB^\prime = \vecB / \sqrt{a}.$
+Then Equation \ref{eq:induction_cosmo_source} becomes
+$$ 
+\frac{\partial \vecB^\prime}{\partial t} - \frac{1}{a}  \nabla \times (\vecv \times
+\vecB^\prime)  = 0, \label{eq:induction_cosmo_prime_nosource}
+$$
+and we can reuse each part of the non-cosmological algorithm (hyperbolic solver,
+electric field creation, interpolation, projection, and flux correction) without
+modification.
+
+Dual energy formalism has also been incorporated in two methods, one internal
+energy, similar to that described in Section \ref{sec.hydro.ppm}, and one that
+uses entropy.  The additional equations are
+\begin{eqnarray}
+\end{eqnarray}
+\begin{eqnarray}
+  \frac{\partial S}{\partial t} +\frac{1}{a} \nabla \cdot (S \vecv) & = &  -
+  \frac{ 3 (\gamma - 1) \dot{a}}{a} S \\
+  \frac{\partial \rho e}{\partial t} + \frac{1}{a} \nabla \cdot (\rho e \vecv) &
+  = & -\frac{ 3(\gamma -1 ) \dot{a}}{a} + \frac{p}{a} \nabla \cdot \vecv,
+\end{eqnarray}
+where $e$ is the internal energy and $S=p/\rho^{\gamma-1}$ is the entropy.  If
+the ratio of thermal to kinetic energy, $\eta$, is below a certain value, the
+alternative energy is used.  As in \citet{Li08}, we use $\eta=0.008$.
+
 Enzo's MHD-CT has been used in a variety of contexts, from galaxy

### 2013-05-15 numerical-mhd-dedner.tex
First cobbled together text for the dedner description.

@@
-\subsection{MHD: Dedner}
-\label{sec.num.mhd-dedner}
+\subsection{Hydrodynamics: MUSCL}
+\label{sec.num.hydro-muscl}
+
+The equations of compressible inviscid hydrodynamics can be written in the form of conservation laws,
+\begin{equation}
+ \frac{\partial{U}}{\partial{t}} +
+ \frac{\partial{F^x}}{\partial{x}} + \frac{\partial{F^y}}{\partial{y}} + \frac{\partial{F^z}}{\partial{z}}= 0, \label{hydro}
+\end{equation}
+The conserved variable $U$ is given by
+\begin{equation}
+ U = (\rho, \rho v_x, \rho v_y, \rho v_z, \rho E)^{T},
+\end{equation} 
+where $\rho$ is density, $v_i$ are the three components of velocity
+for $i={x,y,z}$, $E=v^2/2 + e$ denotes the specific total energy and $e$ the
+specific internal energy.
+
+For the generalized Lagrange multiplier formulation of the MHD
+equations \citep{}{}{2002JCoPh.175..645D}, we consider these 
+conserved variables
+\begin{equation}
+ U = (\rho, \rho v_x, \rho v_y, \rho v_z, \rho E+B^2/2, B_x, B_y, B_z, \psi)^{T},
+\end{equation} 
+where $B_i$ with $i={x,y,z}$ are the three components of magnetic
+fields and $\psi$ is the additional scalar field introduced in the GLM
+formulation for the divergence cleaning.  The fluxes then are
+\begin{eqnarray}
+ F^x &=& (\rho v_x, \rho v_x^2+p+B^2/2-B_x^2, \rho v_yv_x-B_yB_x, \cr
+ && \rho v_zv_x-B_zB_x, \rho ({v^2\over2} + h)v_x+B^2v_x-B_xB\cdot v, \cr
+&& \psi, v_xB_y-v_yB_x, -v_zB_x+v_xB_z, c_h^2B_x)^T,\\
+ F^y &=& (\rho v_y, \rho v_xv_y-B_xB_y, \rho v_y^2+p+B^2/2-B_y^2, \cr
+ && \rho v_zv_y-B_zB_y, \rho ({v^2\over2} + h)v_y+B^2v_y-B_yB\cdot v, \cr
+ && v_yB_z-v_zB_y,\psi,-v_xB_y+v_yB_x, c_h^2B_y)^{T}, \\
+ F^z &=& (\rho v_z, \rho v_xv_z-B_xB_z, \rho v_yv_z-B_yB_z, \rho v_z^2+p+B^2/2-B_z^2, \cr
+ && \rho ({v^2\over2} + h)v_z+B^2v_z-B_zB\cdot v, \cr
+    &&  -v_yB_z+v_zB_y, v_zB_x-v_xB_z,\psi, c_h^2B_z)^{T},
+\end{eqnarray}
+where $c_h$ is a constant controlling the propagation speed and
+damping rate of $\nabla\cdot B$. $h=e+p/\rho$ denotes the enthalpy.
+
+As a reconstruction scheme we here implemented only the simple
+piecewise linear reconstruction. And a number of approximate Riemann
+solvers.

### 2013-05-15 test_problems/OrszagTang/PlotOrszagTang.py
adding OT test scripts

@@ -1,0 +1,11 @@
+from yt.mods import *
+name1 = '/Users/dcollins/scratch/Paper31/ot04_large_2d/DD0056/data0056'
+name2 = '/Users/dcollins/scratch/Paper31/ot08_large_dedner_hlld/DD0032/data0032'
+pf1 = load(name1)
+pf2 = load(name2)
+names=['mhdct_0056','mhddedner_0035']
+sl=[]
+for i, pf in enumerate([pf1,pf2]):
+  sl.append( SlicePlot( pf, 2, 'Density') )
+  sl[-1].set_zlim('Density',0.1,1)
+  print sl[-1].save(names[i]+".eps")

### 2013-05-15 ms.tex
added acknowledgments file, put some basic stuff in there.
minor tweak: author list idea.

@@
 \title{Enzo: An Adaptive Mesh Refinement Code for Astrophysics}
-\author{Authors}
+\author{The Enzo Collaboration:}
+\author{author 1}
+\author{author 2}
+\author{author 3...}
 
@@
 
-We now describe each of the numerical methods in detail.  The structure of this section mirrors the previous one, in which each method was briefly described; here we present the details.
+We now describe each of the numerical methods in detail.  The
+structure of this section mirrors that of Section~\ref{sec.overview},
+in which each method was briefly described.
 
@@
 \input{numerical-muscl} % new
-%\input{numerical-mhd-dedner} % new (combined with muscl)
+\input{numerical-mhd-dedner} % new (combined with muscl)
 \input{numerical-mhd-ct}  % new
@@
 
+\input{acknowledgments}
+
 \input{appendix}  % new

### 2013-05-15 numerical-analysis.tex

@@
 effectively in several papers, including \citet{2010ApJ...715.1575S}
-and \citet{2012ApJ...748...12S}.
+and \citet{2012ApJ...748...12S}. It is crucial to keep in mind that
+tracer particles model some Lagrangian fluid trajectories. The fluid
+on the grid, however, models the motion of all the mass and represents
+the average quantity in a grid cells volume. Consequently, after some
+evolution tracer particles, which even if they initially had the same
+density distribution as the gas, will not have the same density
+distribution as the fluid. They tend to accumulate at stagnation points
+of the flow, e.g., and care has to be taken in using these particles
+appropriately. They can be very useful, e.g. to study a variety of
+histories of the hydrodynamic quantities in Lagrangian fluid elements
+and evaluate complex chemical and cooling models in regard to the
+simpler ones used in the actual numerical evolution.
 

### 2013-05-16 strong_scaling/plot.py
Adding strong scaling AMR tests and figures

@@ -1,0 +1,91 @@
+import glob
+import numpy as np
+import matplotlib.pyplot as plt
+
+from matplotlib import rc
+rc('font', size=24)
+from matplotlib.font_manager import fontManager, FontProperties
+font= FontProperties(size=16)
+
+files = glob.glob('performance_*.out')
+nfiles = len(files)
+
+MAX_LEVEL = 6
+
+timing = {'cores': np.zeros(nfiles, dtype='int'),
+          'total_time': np.zeros(nfiles),
+          'level_time': np.zeros((nfiles, MAX_LEVEL+1)),
+          'routines': [],
+          'routine_time': None}
+
+# Determine the routine names
+lines = open(files[0], 'r').readlines()
+for l in lines:
+    if not l.startswith('#') and len(l.split()) == 5:
+        timing['routines'].append(l.split()[0])
+nroutines = len(timing['routines'])
+timing['routine_time'] = np.zeros((nfiles, nroutines, 4))
+ 
+# Parse performance files         
+for i, f in enumerate(files):
+    lines = open(f, 'r').readlines()
+    for l in lines:
+        if l.find('MPI processes') >= 0:
+            timing['cores'][i] = int(l.split(':')[1])
+        elif l.startswith('Level'):
+            level = int(l[7:9])
+            time = float(l.split()[1])
+            timing['level_time'][i,level] = time
+        elif l.startswith('Total'):
+            timing['total_time'][i] = float(l.split()[1])
+        elif len(l) > 1:
+            str0 = l.split()[0]
+            if str0 in timing['routines']:
+                ipl = timing['routines'].index(str0)
+                timing['routine_time'][i,ipl,:] = \
+                    map(float, l.split()[1:5])
+        
+
+# Sort by cores
+isort = np.argsort(timing['cores'])
+timing['cores'] = timing['cores'][isort]
+timing['total_time'] = timing['total_time'][isort]
+timing['level_time'] = timing['level_time'][isort,:]
+timing['routine_time'] = timing['routine_time'][isort,:,:]
+
+colors = ['k', 'b', 'r', 'g', 'm', 'c']
+styles = ['-', '--', ':', '-.']
+markers = ['v', '^', 's', '*', '+', 'D']
+########################################################################
+# Plot of time spent: total and on each level
+########################################################################
[…]

### 2013-05-16 parallel.tex
Edits to introduction (also adding section labels to some other files).
Adding strong scaling figures to the text.

@@
 \section{Parallel Strategy and Performance}
+\label{sec.parallel}
 
@@
 \subsubsection{Unigrid scaling}
+
+\begin{figure}
+\begin{center}
+\includegraphics[width=0.6\textwidth]{figures/enzo_unigrid_weak_scaling.eps}
+\caption{\enzo\ weak scaling performance for a set of Lyman alpha
+forest cosmology simulations with constant comoving spatial resolution
+per grid cell, showing cell updates per second per computational core
+plotted as a function of the number of root grid tiles of dimension
+$128^3$ (R) in each dimension.  The number of MPI tasks is N$ = R^3$,
+so R$ = 16$ on this plot corresponds to a $2048^3$ computational mesh
+running on 4096 MPI tasks.  This plot goes from R$ = 2$ (8 MPI tasks)
+to R$ = 24$ (13,824 MPI tasks) on two supercomputers -- NICS Kraken
+and ORNL Jaguar when they were Cray XT4 systems -- and using 1, 2 or 4
+MPI processes per node, where each compute node contained a single
+quad-core AMD Opteron CPU having a speed of 2.1 GHz on Jaguar and 2.3
+GHz on Kraken.}
+\label{fig.uniscale}
+\end{center}
+\end{figure}
 
@@
 
+\subsubsection{AMR scaling \red{(John)}}
+
 \begin{figure}
 \begin{center}
-\includegraphics[width=0.4\textwidth]{figures/enzo_unigrid_weak_scaling.eps}
-\caption{\enzo\ weak scaling performance for a set of Lyman alpha
-forest cosmology simulations with constant comoving spatial resolution
-per grid cell, showing cell updates per second per computational core
-plotted as a function of the number of root grid tiles of dimension
-$128^3$ (R) in each dimension.  The number of MPI tasks is N$ = R^3$,
-so R$ = 16$ on this plot corresponds to a $2048^3$ computational mesh
-running on 4096 MPI tasks.  This plot goes from R$ = 2$ (8 MPI tasks)
-to R$ = 24$ (13,824 MPI tasks) on two supercomputers -- NICS Kraken
-and ORNL Jaguar when they were Cray XT4 systems -- and using 1, 2 or 4
-MPI processes per node, where each compute node contained a single
-quad-core AMD Opteron CPU having a speed of 2.1 GHz on Jaguar and 2.3
-GHz on Kraken.}
-\label{fig.uniscale}
+\includegraphics[width=0.48\textwidth]{figures/strong_scaling_levels.eps}
+\hfill
+\includegraphics[width=0.48\textwidth]{figures/strong_scaling_routines.eps}
 \end{center}
+\caption{\emph{Left:} Strong scaling test of a 512$^3$ AMR
+  cosmological calculation.  The root grid scaling is not
+  representative of the true strong scaling of Enzo because the root
+  grid tiles are not repartitioned when a simulation is restarted with
+  additional cores.  The weak scaling test in Figure
+  \ref{fig.uniscale} is more representative of the scaling on the root
+  level.  The performance in the refined levels show good strong
+  scaling up to 4096 cores.  \emph{Right:} Time spent in the major
+  routines in the same AMR calculation.}
+\label{fig:amr_hierarchy}
 \end{figure}
-
-\subsubsection{AMR scaling \red{(Greg)}}
 

### 2013-05-16 numerical-mhd-ct.tex
fixing some references.

@@
 the ratio of thermal to kinetic energy, $\eta$, is below a certain value, the
-alternative energy is used.  As in \citet{Li08}, we use $\eta=0.008$.
+alternative energy is used.  As in \citet{Li08a}, we use $\eta=0.008$.
 
@@
 halos \citep{Skillman13}, primordial star formation \citep{Xu08}, and
-present-day star formation \citep{Collins11,Collins12}.
+present-day star formation \citep{Collins11, Collins12a}.

### 2013-05-16 numerical-mhd-ct.tex
fixing refs. again.

@@
 This MHD method was developed by \citet{Collins10} in order to study Population
-III stars \citep{Xu08}, molecular clouds \citep{collins12}, and galaxy clusters
+III stars \citep{Xu08}, molecular clouds \citep{Collins12a}, and galaxy clusters
 \citep{Xu12, Skillman13}.  A full description and suite of test problems can be

### 2013-05-16 numerical-mhd-ct.tex
Adding AMR scaling section with an associated table.  I still need to add some references in this section.  Fixing some undefined and multiple labels and making the UVB table wider and spaced nicely.

--- 2013-05-14 numerical-mhd-ct.tex
@@
 This MHD method was developed by \citet{Collins10} in order to study Population
-III stars \citep{Xu08}, molecular clouds \citep{collins12}, and galaxy clusters
+III stars \citep{Xu08}, molecular clouds \citep{Collins12}, and galaxy clusters
 \citep{Xu12, Skillman13}.  A full description and suite of test problems can be

### 2013-05-16 parallel.tex
Adding AMR scaling section with an associated table.  I still need to add some references in this section.  Fixing some undefined and multiple labels and making the UVB table wider and spaced nicely.

@@
 \subsubsection{Unigrid scaling}
+\label{sec:weak_scaling}
 
@@
 so R$ = 16$ on this plot corresponds to a $2048^3$ computational mesh
-running on 4096 MPI tasks.  This plot goes from R$ = 2$ (8 MPI tasks)
+running on 4,096 MPI tasks.  This plot goes from R$ = 2$ (8 MPI tasks)
 to R$ = 24$ (13,824 MPI tasks) on two supercomputers -- NICS Kraken
@@
 
-\subsubsection{AMR scaling \red{(John)}}
+\subsubsection{AMR scaling}
 
@@
   level.  The performance in the refined levels show good strong
-  scaling up to 4096 cores.  \emph{Right:} Time spent in the major
-  routines in the same AMR calculation.}
-\label{fig:amr_hierarchy}
+  scaling up to 2,048 cores.  \emph{Right:} Time spent in
+  representative major routines in the same AMR calculation.}
+\label{fig:strong_scaling}
 \end{figure}
 
-% Greg to work on this, scaling data from his runs, or maybe from John Wise
+Many astrophysical problems, such as cosmological galaxy formation
+\citep{}, high-resolution disk galaxy simulations \citep{},
+high-redshift \citep{} and present-day star formation \citep{},
+involve multi-scale physics that span several orders of magnitude in
+both space and time.  In these situations, using Enzo in its adaptive
+mesh refinement mode is beneficial.  Because the adaptive grid
+hierarchy is dynamic and not static, grid boundaries and, thus,
+communication patterns can be unpredictable, hindering good strong
+scaling at high core counts.
+
+Figure \ref{fig:strong_scaling} shows strong scaling results from a
+single 50 Mpc/h cosmology simulation run on $N_{\rm core}$ compute
+cores, ranging from 128 to 16,384 at power-of-two intervals.  It was
+run on the NICS Kraken XT5 supercomputer that has two AMD Opteron
+hexa-core processors with 16 GB of memory per compute node.  The
+simulations that utilized 128, 256, and 512 cores were executed on 128
+nodes because of the memory requirements.  The higher core count jobs
+were run with 8 cores per node.  This simulation would not run with 12
+cores per node because of the memory overhead associated with the grid
+hierarchy being duplicated on each MPI process.  However, this
+overhead is greatly diminished if a hybrid-parallel (MPI + OpenMP)
+approach is used.
+
+This simulation includes dark matter dynamics, hydrodynamics using the
+piecewise parabolic method, six-species non-equilibrium chemistry and
+radiative cooling, and a uniform metagalactic ultraviolet background.
+The simulation uses the space-filling curve method for load balancing
+the AMR grids.  It has a $512^3$ root grid that is divided into 512
+tiles, and 6 additional AMR levels are used.  We perform these scaling
+tests when the simulation reaches $z=4$.  The results shown in Figure
+\ref{fig:strong_scaling} come from a single root-level timestep
+$\Delta t = 2.1\, \textrm{Myr}$.  At this time, there are $3.04 \times
+10^5$ AMR grids, $9.03 \times 10^8$ ($\sim 967^3$) computational
+cells, and $1.34 \times 10^8$ dark matter particles in total.  The
+breakdown of the number of AMR grids, cells, timesteps, and number of
+cell updates on each level is shown in Table \ref{tab:amr_scale}.
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\begin{table*}
+  \begin{center}
+  \caption{Strong scaling test computational details}
+  \begin{tabular*}{0.9\textwidth}{@{\extracolsep{\fill}}c c c c c}
+    \tableline\tableline
+    {Level} & {$N_{\rm grid}$} & {$N_{\rm cells}$} & {$N_{\rm up}$} &
+    {$N_{\rm timesteps}$}\\
+    \tableline
+    0 & 512 & $1.34 \times 10^8$ & $1.34 \times 10^8$ & 1\\
+    1 & 61,868 & $4.01 \times 10^8$ & $4.01 \times 10^8$ & 1\\
+    2 & 91,182 & $1.99 \times 10^8$ & $5.96 \times 10^8$ & 3\\
+    3 & 59,932 & $7.62 \times 10^7$ & $5.34 \times 10^8$ & 7\\
+    4 & 40,700 & $3.32 \times 10^7$ & $5.65 \times 10^8$ & 17\\
+    5 & 28,346 & $2.76 \times 10^7$ & $1.36 \times 10^9$ & 49\\
+    6 & 19,771 & $2.80 \times 10^7$ & $5.25 \times 10^9$ & 187\\
+    \tableline
+    Total & 302,311 & $9.03 \times 10^8$ & $8.83 \times 10^9$ & --\\
+  \end{tabular*}
+  \parbox[t]{0.9\textwidth}{\textbf{Note.} --- Data shown at $z=4$ for
+    a root grid timestep of 2.1~Myr.  Col. (1): AMR Level. Col. (2):
+    Number of grids. Col. (3): Number of computational
+    cells. Col. (4): Number of cell updates. Col. (5): Number of
+    timesteps.}
+  \label{tab:amr_scale}
+  \end{center}
+\end{table*}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+The left panel in Figure \ref{fig:strong_scaling} shows the
+computational and communication time spent on each level.  In the AMR
+levels, there exists good strong scaling up to 2,048 cores, and
+marginal speed-ups are found at 4,096 cores.  On the root-grid level,
+there exists good scaling up to 1,024 cores, but the performance
+dramatically decreases at higher core counts.  This occurs because the
+root grid is not re-partitioned into $N_{\rm core}$ tiles when the
+simulation is restarted with a different core count.  This feature can
+be easily implemented and is planned in the next major release of
+Enzo, where scaling results would be similar to the weak scaling shown
+in \S\ref{sec:weak_scaling}.  The right panel in Figure
+\ref{fig:strong_scaling} shows the time spent in some representative
+major routines in Enzo.  The local physics routines, for example
+\texttt{SolveHydroEquations}, exhibit perfect strong scaling because
+they involve no communication.  By investigating the scaling behavior
+in each routine, it is clear that the communication in the
+\texttt{SetBoundaryConditions}, \texttt{SolveForPotential} (multi-grid
+solver in AMR levels), and \texttt{RebuildHierarchy} are responsible
+for the lack of strong scaling at $N_{\rm core} \ga$~4,096 in this
+particular simulation.  The transpose of the root grid tiles are
+responsible for the performance decrease because it is not optimized
+for situations where the number of tiles is greater than the number of
+MPI processes.  These results are directly applicable to simulations
+with similar computational demands.  In simulations with less
+computational cells, strong scaling will cease at a smaller core count
+because the CPUs will become data-starved faster, and the opposite
+occurs with larger simulations.
 

### 2013-05-17 ms.tex
Changing to the newest emulateapj.
Adding AMR scaling section with an associated table.  I still need to add some references in this section.  Fixing some undefined and multiple labels and making the UVB table wider and spaced nicely.
Rearranging sections.

@@
 %\documentclass[12pt,preprint2]{aastex}  % for e-submission to ApJ - two column
-\documentclass[onecolumn]{emulateapj}  % this makes everything look like ApJ
+\documentclass[onecolumn,iop]{emulateapj}  % this makes everything look like ApJ
 
@@
 %   2.2 Overview of numerical methods
-% 3.  Numerical methods
+% 3.  Numerical methods   %%%%%% THIS IS NOW A BUNCH OF DIFFERENT SECTIONS
 %   3.1   AMR
@@
 %% Section: numerical methods
-\section{Numerical Methods}
-\label{sec.methods}
+% \section{Numerical Methods}
+% \label{sec.methods}
 
-We now describe each of the numerical methods in detail.  The
-structure of this section mirrors that of Section~\ref{sec.overview},
-in which each method was briefly described.
+% We now describe each of the numerical methods in detail.  The
+% structure of this section mirrors that of Section~\ref{sec.overview},
+% in which each method was briefly described.
 
 \input{numerical-amr}
+
+\section{Fluid methods}
+\label{sec:fluids}
+
 \input{numerical-ppm}
-\input{numerical-muscl} % new
-\input{numerical-mhd-dedner} % new (combined with muscl)
-\input{numerical-mhd-ct}  % new
+%\input{numerical-muscl} 
+\input{numerical-mhd-dedner}  % includes MUSCL and Dedner
+\input{numerical-mhd-ct}  
 \input{numerical-zeus}
+
+\section{Gravity and N-body}
+\label{sec:allgrav}
 \input{numerical-gravity}
 \input{numerical-particle}
+
+\section{Microphysics}
+\label{sec:microphysics}
 \input{numerical-chemistry}
-\input{numerical-cooling} % new
-\input{numerical-radiation-homogeneous} % new
-\input{numerical-radiation-raytracing} % new
-\input{numerical-radiation-fld} % new
-\input{numerical-conduction} % new
+\input{numerical-cooling}
+
+\section{Radiation}
+\label{sec:allrad}
+\input{numerical-radiation-homogeneous} 
+\input{numerical-radiation-raytracing} 
+\input{numerical-radiation-fld}
+
+\section{Other physics}
+\label{sec:otherphys}
+\input{numerical-conduction} 
 \input{numerical-starformation}
+
+% this is its own section
 \input{numerical-timestep}
-\input{numerical-analysis} % new
+
+% this is its own section
+\input{numerical-analysis}
 
@@
 %% Section: code development methodology
-\input{development} % new
+\input{development}
 
 %% Section: conclusions
-\input{conclusions}  % new
+\input{conclusions}
 
@@
 
-\input{appendix}  % new
+\input{appendix}
 

### 2013-05-17 introduction.tex
Rearranging sections - cleaned this up a bit.

@@
 This is designed to give a first broad-brush picture of the equations we aim to solve and the methods used to solve them.
-Next, in Section~\ref{sec.methods}, we describe the methods we use in detail, reserving some of the longer descriptions of particular components for the appendix in order not to interrupt the flow of the paper.
+Next, in Sections~\ref{sec.amr} through~\ref{sec.num.analysis}, we describe the methods we use in detail, reserving some of the longer descriptions of particular components for the appendix in order not to interrupt the flow of the paper.
 The \enzo\ testing framework and code tests are described in Section~\ref{sec.tests}.

### 2013-05-17 overview.tex
Adding AMR scaling section with an associated table.  I still need to add some references in this section.  Fixing some undefined and multiple labels and making the UVB table wider and spaced nicely.
Rearranging sections - cleaned this up a bit.

@@
 numerics of the code.  Detailed descriptions of the individual
-components are then provided in Section~\ref{sec.methods}.
+components are then provided in Sections~\ref{sec.amr}-\ref{sec.num.analysis}.
 
@@
 proceed through the numerical methods in the same order as will be used in
-section~\ref{sec.methods} so that there is a one-to-one correspondence
+Sections~\ref{sec.amr} through~\ref{sec.num.analysis} so that there is a one-to-one correspondence
 between each of the following overviews and the complete description
-provided in that section.  The goal here is to introduce the reader to
+provided in the later sections.  The goal in this section is to introduce the reader to
 the basic principles of the methods without drowning in detail.
@@
 described by \citet{Balsara99} and \citet{Gardiner05}.  Further, we employ the
-second order hyperbolic solver of \citet{Li08} and the divergence free AMR
+second order hyperbolic solver of \citet{Li08a} and the divergence free AMR
 scheme of \citet{Balsara01}.

### 2013-05-19 overview.tex
Edits due to Tasker suggestions.

@@
 
-The Eulerian equations of ideal magnetohydrodynamics including gravity
-in a coordinate systems comoving with the cosmological expansion are
+The Eulerian equations of ideal magnetohydrodynamics including gravity,
+in a coordinate systems comoving with the cosmological expansion, are
 given by:
@@
   total}$.  Although we write the equations including both MHD and
-comoving coordinates, the code can operate both in the pure
+comoving coordinates, the code is frequently used both in the pure
 hydrodynamic limit ($\vecB = 0$) and without cosmological expansion
@@
 
-For completeness, we note here that as part of the usual transformation to comoving coordinates, we have defined a set of comoving quantities.
-See, for example, \citet{Peebles93} for more details.
+For completeness, we note here that we have defined several key quantities
+in the comoving frame to make the previous equations more readable.
 Specifically, we define:
+
 \begin{eqnarray}
@@
 
-It is clear that the evolution equations are equivalent to the fixed coordinate version in the limit in which $a = 1$ and $\dot{a} = 0$.
-We include the cosmological terms with the understanding that we are not restricting ourselves to cosmological applications.
+It is clear that the evolution equations are equivalent to the fixed
+coordinate version in the limit in which $a = 1$ and $\dot{a} = 0$.
+We include the cosmological terms with the understanding that \enzo\
+is not restricted to cosmological applications -- while historically
+the code was written with cosmological situations in mind (e.g.,
+galaxy clusters), in more recent years it has been used to simulate a much
+broader range of astrophysical environments.
 
@@
 the equation of radiative transfer over angles.  In particular, we use a
-flux-limited diffusion approximation for cosmological radiative
+flux-limited diffusion approximation for radiative
 transfer, with couplings to both the gas energy and chemical number

### 2013-05-19 introduction.tex
Added most of Dave Collins' suggested edits (except for a bunch of stuff having to do with the hydro).

@@
 We have instead adopted an idea first suggested by \citet{Berger89} in the Computational Fluid Dynamics (CFD) community, which has become known as Structured Adaptive Mesh Refinement (SAMR).
-The principle is to adaptively add and modify additional, finer meshes (``patches'') over regions that require higher resolution.
+The principle is to adaptively add and modify additional, finer meshes (``grids'') over regions that require higher resolution.
 Since each patch has a simple geometry (Cartesian, in our case) the hyperbolic fluid-dynamical equations can be solved on that patch with a well-developed `off-the-shelf' numerical method.
-In addition, it is relatively straightforward to add other advanced physics (e.g.
+In addition, it is possible to add other advanced physics (e.g.
 magnetohydrodynamics, radiation transport, chemistry, etc.).
@@
 It has been expanded as a general tool for astrophysical fluid dynamics and is intended to be efficient, accurate and easily expanded.
-Although many of the components of the code have been described in previous publications \citep{1995CoPhC..89..149B, BryanThesis96, Bryan97a, Bryan97b, Norman99, BryanCompSci99, Bryan01, Oshea04, 2007arXiv0705.1556N}, there has been no systematic and complete description of the code.
+Although many of the components of the code have been described in previous publications \citep{1995CoPhC..89..149B, BryanThesis96, Bryan97a, Bryan97b, Norman99, BryanCompSci99, Bryan01, Oshea04, 2007arXiv0705.1556N, WangAbelZhang08,ReynoldsHayesPaschosNorman2009,Collins10,Wise11_Moray}, there has been no systematic and complete description of the code.
 In this paper we provide that description, filling in many gaps and showing the code's performance for a wide variety of test problems.
@@
 The \enzo\ code has been extensively used over the last two decades in a wide variety of problems, and more than 100 peer-reviewed papers have been published based on results obtained with the code.
-The code has previously been used to study a wide range of astrophysical systems, including galaxy clusters \citep{Loken02}, the interstellar medium \citep{Slyz05}, the intergalactic medium \citep{Fang01}, cooling flows \citep{Li12}, turbulence \citep{Kritsuk04}, and the formation of the first stars \citep{ABN02}.
+The code has previously been used to study a wide range of astrophysical systems, including galaxies \citep{2003ApJ...587...13T, 2012ApJ...749..140H}, galaxy clusters \citep{Loken02, Skillman13}, the interstellar medium \citep{Slyz05}, the intergalactic medium \citep{Fang01, 2011ApJ...731....6S}, cooling flows \citep{Li12, 2013ApJ...763...38S}, turbulence \citep{Kritsuk04, 2009JPhCS.180a2020K, Collins11}, the formation of the first stars \citep{ABN02, 2007ApJ...654...66O, 2009Sci...325..601T}, and the formation of stars in our own Galaxy \citep{Collins12a}.
 

### 2013-05-19 ms.tex
Rearranging sections - cleaned this up a bit.
Added most of Dave Collins' suggested edits (except for a bunch of stuff having to do with the hydro).

@@
 
-% Paper Outline  (last updated January 18, 2013)
+% Paper Outline  (last updated May 17, 2013)
 %
 % 1.  Introduction
-% 2.  Physical equations and overview
+% 2.  Physical equations and overview of numerical methodology
 %   2.1 Physical equations
 %   2.2 Overview of numerical methods
-% 3.  Numerical methods   %%%%%% THIS IS NOW A BUNCH OF DIFFERENT SECTIONS
-%   3.1   AMR
-%   3.2.  Hydro: Godunov (PPM)
-%   3.3   Hydro: Godunov (MUSCL)
-%   3.4   Hydro: Finite difference (ZEUS)
-%   3.5   MHD: Dedner
-%   3.6   MHD: CT
-%   3.7   Gravity
-%   3.8   Particle Integration
-%   3.9.  Chemistry
-%   3.10 Cooling
-%   3.11 Radiation: Homogeneous
-%   3.12 Radiation: ray-tracing
-%   3.13 Radiation: FLD
-%   3.14 Thermal conduction
-%   3.15 Star formation & Feedback
-%   3.16 Timestep constraints
-%   3.17 Analysis
-% 4. Code tests
-%   4.1 Verifying and Validating the Enzo code
-%   4.2 Representative test problems
-% 5. Parallelism 
-%   5.1 Parallel strategy 
-%   5.2 Parallel performance 
-% 6. Code development methodology
-% 7. Conclusions
+% 3.  The structured AMR method
+% 4.  Fluid methods
+%   4.1 PPM
+%   4.2 MUSCL
+%   4.3 MHD-CT
+%   4.4 ZEUS 
+% 5.  Gravity and N-body
+%   5.1 Gravity
+%   5.2 N-body dynamics
+% 6. Microphysics
+%   6.1  Chemistry
+%   6.2  Radiative Cooling and heating
+% 7. Radiation
+%   7.1 Radiation: Homogeneous
+%   7.2 Radiation: ray-tracing
+%   7.3 Radiation: FLD
+% 8. Other physics
+%   8.1 Thermal conduction
+%   8.2 Star formation & Feedback
+% 9. Time stepping
+% 10.  Analysis
+% 11. Code tests
+%   11.1 Verifying and Validating the Enzo code
+%   11.2 Representative test problems
+% 12. Parallelism 
+%   12.1 Parallel strategy 
+%   12.2 Parallel performance 
+% 13. Code development methodology
+% 14. Conclusions
 %
@@
 \begin{abstract}
-This paper describes the open-source \enzo\ code, which uses a
+This paper describes the open-source code \enzo, which uses a
 block-structured adaptive mesh refinement to provide high spatial and
@@
 problems, demonstrate the code's parallel scaling properties, and
-discuss the \enzo\ consortium's code development methodology.
+discuss the \enzo\ collaboration's code development methodology.
 \end{abstract}
@@
 \section{Fluid methods}
-\label{sec:fluids}
+\label{sec.fluids}
 
@@
 \section{Gravity and N-body}
-\label{sec:allgrav}
+\label{sec.allgrav}
 \input{numerical-gravity}
@@
 \section{Microphysics}
-\label{sec:microphysics}
+\label{sec.microphysics}
 \input{numerical-chemistry}
@@
 \section{Radiation}
-\label{sec:allrad}
+\label{sec.allrad}
 \input{numerical-radiation-homogeneous} 
@@
 \section{Other physics}
-\label{sec:otherphys}
+\label{sec.otherphys}
 \input{numerical-conduction} 

### 2013-05-19 numerical-amr.tex
Rearranging sections.
Added most of Dave Collins' suggested edits (except for a bunch of stuff having to do with the hydro).

@@
-\subsection{The Structured Adaptive Mesh Refinement (SAMR) Method}
+\section{The Structured Adaptive Mesh Refinement (SAMR) Method}
 \label{sec.amr}
@@
 Each grid patch in \enzo\ contains arrays of values for baryon and particle quantities.
-Grids are partitioned into a core of \emph{real zones} and a surrounding layer of \emph{ghost zones}, as shown in Figure~\ref{fig:amr_hierarchy}.
-The real zones store field values and ghost zones are used to temporarily store values that have been obtained directly from neighboring grids or interpolated from a parent grid.
+Grids are partitioned into a core of \emph{active zones} and a surrounding layer of \emph{ghost zones}, as shown in Figure~\ref{fig:amr_hierarchy}.
+The active zones store field values and ghost zones are used to temporarily store values that have been obtained directly from neighboring grids or interpolated from a parent grid.
 These zones are necessary to accommodate the computational stencil of the (magneto)hydrodynamics solvers (Sections~\ref{sec.hydro.ppm} and~\ref{sec.hydro.zeus}) and the gravity solver (Section~\ref{sec.gravity}).
@@
 At the end of each time step, cells on a given level fill ghost zones by interpolating from the parent and copying from neighbors on the same level.
-Grids then correct the flux at the interface boundary between fine and coarse zones, and finally project their real zone data to the parent grid.
+Grids then correct the flux at the interface boundary between fine and coarse zones, and finally project their active zone data to the parent grid.
 
@@
 
-\subsubsection{Setting the boundary values}
+\subsection{Setting the boundary values}
 \label{sec:interpolation}
@@
   \item{\em Reflecting.} In this case, the boundary behaves like a mirror, with the solution on the boundary reflecting the solution in the computational domain ($q(-x) = q(+x)$) but with the velocity normal to the boundary reversed: $v_x(-x) = -v_x(+x)$.
-  \item{\em Outflow.} True outflow conditions are extremely difficult \red{(REF?)}.  We approximate them by duplicating the solution at the edge of the computational domain: $q(-x) = q(0)$.
+  \item{\em Outflow.}  We approximate outflow boundary conditions by duplicating the solution at the edge of the computational domain: $q(-x) = q(0)$.
   \item{\em Inflow.} In this case, the boundary values are fixed by a pre-determined function ($q(-x) = q0(x,t)$).
@@
 
-\subsubsection{Solving the hydrodynamic equations on a grid}
+\subsection{Solving the hydrodynamic equations on a grid}
 \label{sec:solve_hydro}
@@
 primary types are cell-centered, face-centered, and node-centered). We
-will employ both the direct Eulerian version of the piecewise
-parabolic method \citep{1984JCoPh..54..174C,1995CoPhC..89..149B} which
+will employ the direct Eulerian version of the piecewise
+parabolic method \citep{1984JCoPh..54..174C,1995CoPhC..89..149B}, which
 assumes cell-centering for all quantities (actually they are
@@
 
-\subsubsection{Projection}
+\subsection{Projection}
 \label{sec:projection}
@@
 \end{equation}
-where $r$ is the refinement factor, $d$ the dimensionality, and the sum is over all fine cells that ``cover" the coarse cell.
+where $r$ is the refinement factor and the sum is over all fine cells that ``cover" the coarse cell.
+The exponent, $d$, refers to the weight of the average.
+For cell-centered quantities, this is the dimensionality of the problem.
+For face-centered quantities, this is one less than the dimensionality, and for edge-centered quantities this is one.
 
-\subsubsection{Flux correction}
+\subsection{Flux correction}
 \label{sec:flux_correction}
@@
 Unfortunately, the projection step upsets this flux conservation since, at the boundary of a refined region, the coarse and fine cells on either side of the border are updated with different estimates of the flux across the boundary.
-In one case this comes from the coarse grid computation, and in the other from the fine grid solution.
-Conservation can be restored by correcting all coarse cells that lie at the boundary of a fine region
+At one refinement level this comes from the coarse grid computation, and in the other from the fine grid solution.
+Conservation can be restored by correcting all coarse cells that lie outside the boundary of a fine region
 with the difference between the coarse and fine estimates of the flux across the boundary:
@@
 
-\subsubsection{Hierarchy reconstruction}
+\subsection{Hierarchy reconstruction}
 \label{sec:hierarchy_reconstruction}
@@
 Since the addition of more highly refined grids is adaptive, the conditions for refinement must be specified.
-Up to seven different criteria can be specified for a given simulation, and these refinement criteria are discussed in more detail in the following section (Section~\ref{sec:refinement_criteria}).
+Many different criteria can be simultaneously specified for a given simulation, and these refinement criteria are discussed in more detail in the following section (Section~\ref{sec:refinement_criteria}).
 Once all cells in a given grid that require additional refinement have been flagged, rectangular boundaries are determined that minimally encompass the flagged cells.
@@
 We compute an efficiency $\epsilon$ for this grid, defined as the ratio of flagged cells to total cells.
-If this efficiency is sufficiently larger (typically larger than 30\%), or if the grid is small enough (typically less $4^d$ coarse cells) then we adopt the region as a new grid and move on to the next region (if any).
-If the region is not acceptable, then we try to split it into two using the following algorithm.
+If this efficiency is sufficiently larger (typically larger than 30\%), or if the grid is small enough (typically less than $4^d$ coarse cells, where d is the dimensionality of the simulation) then we adopt the region as a new grid and move on to the next region (if any).
+If the region is not acceptable, then we split it into two using the following algorithm.
 First, we determine the longest dimension and compute the signature along that dimension.
@@
 Edges are typically indicates by inflection points in this signature, so we calculate the second derivative using a simple three-point finite difference estimator and look for the maximum in the second derivative, with ties going to the index closest to the center.
+If no inflection points are found along the longest exis, the next-longest axis is searched.
+If no inflection points are found, the longest axis is cut in half.
 The original region is then split into two new regions, or proto-subgrids.
@@
 First, traditional grid-based hydro method can be simply added with only minor modifications.
-Second, it is easier to accommodate larger stencils (as in PPM, which requires three boundary zones), and finally it can be more efficient for some machines since there are more predictable and linear memory accesses, which more efficiently use a CPU's cache.
+Second, it is easier to accommodate larger stencils (as in PPM, which requires three boundary zones), and third, it can be more efficient for some machines since there are more predictable and linear memory accesses, which more efficiently use a CPU's cache.
 
-\subsubsection{Refinement criteria}
+\subsection{Refinement criteria}
 \label{sec:refinement_criteria}
 
-Finally, we turn to the refinement criteria themselves.
+In this section, we turn to the refinement criteria themselves.
 Although generalized refinement criteria are available based on estimators for the error, such as the Richardson method \red{(REF)}, we have focused generally on methods that use physical estimators based on the assumption that the simulator has a specific quantity or quantities that indicate the need for refinement.
@@
 \end{eqnarray}
+We note that the values $0.33$ and $0.1$ in the equations above were determined empirically, and can be modified by the user to refine more or less aggressively, as the situation warrants.
 
-\item{\em Particle mass.}  This refinement criterion is precisely the same as the baryon overdensity criterion except that it uses the gridded (CIC) particle density.
+\item{\em Particle mass.}  This refinement criterion is precisely the same as the baryon overdensity criterion except that it uses the gridded (cloud-in-cell, or CIC) particle density.
 In general, this criteria is used in cosmological simulations, and thus refines on dark matter and stellar density.
@@
 cross section, $n_{\rm HI}$ is the proper \ion{H}{1} number density,
-and $\Delta x$ is the cell width.  If $\tau > 1$, the cell is refined
+and $\Delta x$ is the cell width (which is assumed to be the same along each axis).  If $\tau > 1$, the cell is refined
 further, up to the maximum level of refinement.
@@
 
-\item{\em ``Must refine region.''}  This refinement criterion is used to ensure that all cells in a given subvolume of the simulation are refined to {\it at least} some minimum level of refinement.
+\item{\em Must refine region.}  This refinement criterion is used to ensure that all cells in a given subvolume of the simulation are refined to {\it at least} some minimum level of refinement.
 The subvolume is user-specified, and may evolve over time (e.g., to follow a structure of interest that is traveling through the simulation volume).

### 2013-05-19 numerical-analysis.tex
Rearranging sections.
Added most of Dave Collins' suggested edits (except for a bunch of stuff having to do with the hydro).

@@
-\subsection{Analysis}
+\section{Analysis}
 \label{sec.num.analysis}
 
-\subsubsection{Inline analysis with yt}
+\subsection{Inline analysis with yt}
 
@@
 simulation.  This enables analysis with extremely high time cadence (even as often
-as every subcycle of the finest refinement level), as well as without
+as every subcycle of the finest refinement level), without
 attempting to write an entire checkpoint output to disk.  The current mechanism
@@
 
-\subsubsection{Tracer particles}
+\subsection{Tracer particles}
 
@@
 
-At user-specified intervals, the tracer particles output their data
-in a streaming format into a binary file.  The data written out
+At user-specified intervals, the tracer particles output their data.  The data written out
 typically includes the tracer particle's unique ID, position, velocity, density,
@@
 
-\subsubsection{Shock finding}
+\subsection{Shock finding}
 

### 2013-05-19 numerical-conduction.tex
Added most of Dave Collins' suggested edits (except for a bunch of stuff having to do with the hydro).

@@
 / |\nabla T|$ is the characteristic length scale of the local
-temperature gradient.  Furthermore, it is important to note that the
-overall level of thermal conductivity can be strongly affected
-by the presence of magnetic fields, which may strongly suppress heat
-flow.  The strength and distribution of magnetic fields in the
-intracluster medium is poorly understood, and it is unclear to what
-extent these magnetic fields will suppress conduction.  For the
-purposes of the work described in this paper, we assume that the
-conductivity of the plasma can be described in terms of an effective
-conductivity, which can be expressed as a fraction f$_{Sp}$ of the
-Spitzer conductivity (where f$_{Sp} \leq 1.0$ are considered physical
-values).
+temperature gradient.  We also assume that the conductivity of the
+plasma can be described in terms of an effective conductivity, which
+can be expressed as a fraction f$_{Sp}$ of the Spitzer conductivity
+(where f$_{Sp} \leq 1.0$ are considered physical values).  This takes
+into account physical processes below the resolution limit of the
+simulation (such as tangled magnetic fields) that can suppress heat
+transport.
 
-Thermal conduction in a
-plasma can be strongly affected by the presence of magnetic field
-lines, which may strongly suppress heat flow perpendicular to the
-magnetic field.  In that case, we allow for heat transport only
-parallel to
-the magnetic field lines in magnetohydrodynamic simulations.
-Mathematically, this is treated as:
+Thermal conduction in a plasma can be strongly affected by the
+presence of magnetic field lines, which may strongly suppress heat
+flow perpendicular to the magnetic field.  In that case, we allow for
+heat transport only parallel to the magnetic field lines in
+magnetohydrodynamic simulations.  Mathematically, this is treated as:
 

### 2013-05-19 numerical-gravity.tex
Added most of Dave Collins' suggested edits (except for a bunch of stuff having to do with the hydro).

@@
 Particles on sub-grids are also added in this way.
-In addition, since the gravitating field for a grid is defined beyond the grid edges (see below), dark matter particles from sibling grids and sub grids whhatich lie within the entire gravitating field are used.
+In addition, since the gravitating field for a grid is defined beyond the grid edges (see below), dark matter particles from sibling grids and sub grids that lie within the entire gravitating field are used.
 This step can involve communication.

### 2013-05-19 numerical-particle.tex
Added most of Dave Collins' suggested edits (except for a bunch of stuff having to do with the hydro).

@@
 (because the underlying gravitating mass field is approximately
-time-centered), and so we have accelerations $\myvec{a}^{n+1/2}$ for
+time-centered), and so we have accelerations $\myvec{g}^{n+1/2}$ for
 each particle.  These are used to update the particle positions and
@@
 \myvec{x}^{n+1/2} & = & \myvec{x}^n + \frac{\Delta t}{2 a^n} \myvec{v}^{n} \nonumber \\
-\myvec{v}^{n+1} & = & \myvec{v}^n \left(1 - \frac{\dot{a}^{n+1/2}}{a^{n+1/2}}\right) + \frac{\Delta t}{ a^{n+1/2}} \myvec{a}^{n+1/2} \\
+\myvec{v}^{n+1} & = & \myvec{v}^n \left(1 - \frac{\dot{g}^{n+1/2}}{g^{n+1/2}}\right) + \frac{\Delta t}{ a^{n+1/2}} \myvec{g}^{n+1/2} \\
 \myvec{x}^{n+1} & = & \myvec{x}^{n+1/2} + \frac{\Delta t}{2 a^{n+1}} \myvec{v}^{n+1} \nonumber

### 2013-05-19 numerical-starformation.tex
Added most of Dave Collins' suggested edits (except for a bunch of stuff having to do with the hydro).

@@
 
-Modeling star formation in the context of galaxy formation forces one
-to confront two fundamental problems. The dynamical range required to
-study galaxy formation and evolution is substantial, and attempting to
-directly resolve the formation of individual stars in the context of a
-Milky Way-type galaxy is computationally unfeasible.  Also, a strong
-theoretical understanding of the details of star formation is
-currently lacking.  Due to these challenges, models of star formation
-in cosmological simulations are largely phenomenological, and the
-``stars'' that are formed are in most cases meant to represent an
-ensemble stellar population (although not all; see
-Section~\ref{sec:starform_pop3}).
+% Modeling star formation in the context of galaxy formation forces one
+% to confront two fundamental problems. The dynamical range required to
+% study galaxy formation and evolution is substantial, and attempting to
+% directly resolve the formation of individual stars in the context of a
+% Milky Way-type galaxy is computationally unfeasible.  Also, a strong
+% theoretical understanding of the details of star formation is
+% currently lacking.  Due to these challenges, models of star formation
+% in cosmological simulations are largely phenomenological, and the
+% ``stars'' that are formed are in most cases meant to represent an
+% ensemble stellar population (although not all; see
+% Section~\ref{sec:starform_pop3}).
 
+Due to the computationally unfeasible number of stars in a galaxy
+(10^11) and the lack of detailed understanding of star formation, a
+number of phenomenological star formation models are included in
+\enzo.
 The \enzo\ code includes approximately ten models for star formation
@@
 
-Algorithmically, the star formation events in Kravtsov's code are
+Algorithmically, the star formation events in the Scmidt-law algorithm are
 assumed to occur once every global time step (with the constraint $\Delta t_0 \leq
@@
 
-Kravtsov reports that simulations with this algorithm reliably
-reproduce the star formation rate-gas surface density relation of the
-Schmidt law, and are not particularly sensitive to numerical
-parameters \citep{2003ApJ...590L...1K}.  He also notes that this
-algorithm is surprisingly insensitive to the presence or absence of
-feedback and details of the cooling and heating properties of the gas,
-which suggests that the global star formation rate is determined by
-gravitationally- driven supersonic turbulence (on large scales) rather
-than stellar feedback or thermal instabilities on small scales.
+% Kravtsov reports that simulations with this algorithm reliably
+% reproduce the star formation rate-gas surface density relation of the
+% Schmidt law, and are not particularly sensitive to numerical
+% parameters \citep{2003ApJ...590L...1K}.  He also notes that this
+% algorithm is surprisingly insensitive to the presence or absence of
+% feedback and details of the cooling and heating properties of the gas,
+% which suggests that the global star formation rate is determined by
+% gravitationally- driven supersonic turbulence (on large scales) rather
+% than stellar feedback or thermal instabilities on small scales.
 

### 2013-05-19 numerical-timestep.tex
Rearranging sections.
Added most of Dave Collins' suggested edits (except for a bunch of stuff having to do with the hydro).

@@
-\subsection{Time Stepping}
+\section{Time Stepping}
 \label{sec.timestepping}
@@
 spacing, $\Delta x$.  In this equation, $\kappa_{\rm hydro}$ is a
-numerical constant with a value of $0 < \kappa_{\rm hydro} \leq 1$
+numerical constant with a value of $0 < \kappa_{\rm hydro} \leq 0.5$
 (with a typical value of $\kappa_{\rm hydro} \sim 0.3-0.5$) that

### 2013-05-19 overview.tex
Added most of Dave Collins' suggested edits (except for a bunch of stuff having to do with the hydro).

@@
 gravitational potential $\phi$ from the total mass density $\rho_{\rm
-  total}$.  Although we write the equations including both MHD and
+  total}$, which is defined as $\rho_{\rm gas} + \rho_{\rm dm} + \rho_{\rm stars}
+- \rho_{\rm average}$.  Although we write the equations including both MHD and
 comoving coordinates, the code is frequently used both in the pure
@@
 $j_\nu \equiv j_\nu(\mathbf{x},\nu,t)$ is the emission coefficient
-that includes any point sources of radiation or diffuse radiation.
+that includes any sources of radiation (whether point or diffuse).
 
@@
 
-In the other limit, we solve an equation derived from integrating
+As an alternative to the ray-casting strategy, we solve an equation derived from integrating
 the equation of radiative transfer over angles.  In particular, we use a
@@
 %  \partial_t {\mathrm n}_i + 
-%    \frac1a \nabla\cdot\left({\mathrm n}_i \vec{v}_b\right) &=& 
+%    \frac1a \nabla\cdot\left({\mathrm n}_i \vecv_b\right) &=& 
 %    \alpha_{i,j} {\mathrm n}_e {\mathrm n}_j - {\mathrm n}_i
@@
 \end{eqnarray}
-where $E_r$ is a grey radiation energy density and $e_c$ is the
+where $E_r = E_r(x,t)$ is a grey radiation energy density and $e_c$ is the
 internal energy correction due to photo-heating and chemical cooling.
@@
 through first assuming a fixed frequency spectrum, i.e.
-$E_{\nu}(\nu,\vec{x},t) = \tilde{E}_r(\vec{x},t) \chi(\nu)$, and then
+$E_{\nu}(\nu,\vecx,t) = \tilde{E}_r(\vecx,t) \chi(\nu)$, and then
 defining the integrated quantity
@@
 \label{eq:grey_radiation_energy}
-   E_r(\vec{x},t) \equiv \int_{\nu_0}^{\infty}
-   E_{\nu}(\nu,\vec{x},t)\,\mathrm d\nu \  = \ 
-   \tilde{E}_r(\vec{x},t) \int_{\nu_0}^{\infty} \chi(\nu)\,\mathrm d\nu.
+   E_r(\vecx,t) \equiv \int_{\nu_0}^{\infty}
+   E_{\nu}(\nu,\vecx,t)\,\mathrm d\nu \  = \ 
+   \tilde{E}_r(\vecx,t) \int_{\nu_0}^{\infty} \chi(\nu)\,\mathrm d\nu.
 \end{equation}
@@
 the solution evolves and interesting regions form, finer meshes are
-placed below these regions (we use the notation `below' to refer to
-finer grids and `above' for coarser grids).  We restrict the ratio
+placed within these coarse regions.
+% (we use the notation `below' to refer to
+% finer grids and `above' for coarser grids). 
+ We restrict the ratio
 between cell sizes to be an integer, typically 2 or 4, and refer to a
@@
 Once there, we continue advancing the grids on the lowest level until
-they have `caught up' to the next highest level above (i.e. $\sum
+they have `caught up' to the coarser parent grids (i.e. $\sum
 \Delta t_l = \Delta t_{l-1}$).  This procedure repeats itself until
-all grids have been advanced by a total time of $\Delta t_0$.
+all grids have been advanced by a total time of $\Delta t_0$ (see
+Figure~\ref{fig:wcycle} for a schematic).
 
@@
 adapt itself.  We do this whenever a level has caught up to the
-coarser level above it and consists of entirely rebuilding the grids
-on that level and below.  This is done by applying the grid refinement
+coarser parent level and consists of entirely rebuilding the grids
+on that level and at finer resolutions.  This is done by applying the grid refinement
 criteria to the grids on that level, flagging zones that require
@@
 we run a machine-vision based algorithm \citep{Berger91} to find edges
-and determine a good placement of subgrids.  These subgrids must not
-overlap one another, must cover all flagged cells and their
-neighbouring cells, and be above a preset efficiency threshold, where
-the efficiency of a cell is defined as the ratio of flagged cells to
-total cells.  Once these new subgrids have been identified, the
+and determine a good placement of subgrids. 
+%  These subgrids must not overlap one another, must cover all flagged
+% cells and their neighbouring cells, and be above a preset efficiency
+% threshold, where the efficiency of a cell is defined as the ratio of
+% flagged cells to total cells. 
+Once these new subgrids have been identified, the
 solution from the next coarser grid is interpolated (see below) in
@@
 and in this way the entire hierarchy (from the original level examined
-and below) is rebuilt.
+and continuing on to finer levels) is rebuilt.
 
@@
 Fourier Technique \citep{Hockney88} to solve Poisson's equation on the
-top (level 0) grid.  The advantage of using this method is that it
+coarsest (Level 0) grid.  The advantage of using this method is that
+it is fast, accurate, and 
 naturally allows both periodic and isolated boundary conditions for
@@
 A second option for radiative transfer is a moment-based method that
-adds an additional field tracking the radiant energy density
+adds an additional field tracking the radiant energy density.
 This field is evolved using the flux-limited diffusion method, which

### 2013-05-19 code-tests.tex
Reynolds edits.

@@
 This cloud is subject to radiation from a point source at the center
-of the $y=0$ boundary with an ionizing photon luminosity
+of the $x=0$ boundary with an ionizing photon luminosity
 $\dot{N}_\gamma = 3 \times 10^{51}$ photons s$^{-1}$, corresponding to

### 2013-05-19 numerical-amr.tex
Reynolds edits.

@@
 Grids are partitioned into a core of \emph{active zones} and a surrounding layer of \emph{ghost zones}, as shown in Figure~\ref{fig:amr_hierarchy}.
-The active zones store field values and ghost zones are used to temporarily store values that have been obtained directly from neighboring grids or interpolated from a parent grid.
+\red{(Move this figure to here?)}  The active zones store field values and ghost zones are used to temporarily store values that have been obtained directly from neighboring grids or interpolated from a parent grid.
 These zones are necessary to accommodate the computational stencil of the (magneto)hydrodynamics solvers (Sections~\ref{sec.hydro.ppm} and~\ref{sec.hydro.zeus}) and the gravity solver (Section~\ref{sec.gravity}).
@@
 
-Timesteps are determined on each level as described in \ref{sec.timestepping}, and the hierarchy is advanced on
+Timesteps are determined on each level as described in Section~\ref{sec.timestepping}, and the hierarchy is advanced on
 a level-by-level basis in a W-Cycle.    Beginning with the coarsest level, $l$, all grids on that level are advanced one
@@
 The finest level is then synched to the level above it, which then proceeds foreward in time one more step.
-This is shown graphically in Figure~\ref{fig:wcycle}
+This is shown graphically in Figure~\ref{fig:wcycle}.
 
@@
     SolveHydroEquations(dt)
+    SolveOtherEquations(dt)
     Time = Time + dt
@@
   constraint regarding not introducing new extrema in interpolated quantities.}  In Appendix~\ref{app:interpolation}, we
-explicitely write down functions in one, two and three dimensions, but
+explicitly write down functions in one, two and three dimensions, but
 here we discuss the general design considerations.
@@
 conservative function by requiring, first, that the function at the
-center of the cell return the cell-centered quantitiy (and since it is
+center of the cell return the cell-centered quantity (and since it is
 linear, this implies that an average over the cell will also be equal
@@
 which refers to the position of physical quantities within a cell (the
-primary types are cell-centered, face-centered, and node-centered). We
+primary types are cell-centered, face-centered, and node-centered). In this description we 
 will employ the direct Eulerian version of the piecewise
@@
 Once all cells in a given grid that require additional refinement have been flagged, rectangular boundaries are determined that minimally encompass the flagged cells.
-This is done using an iterative machine-vesion technique as described in \red{(REF)}.
+This is done using an iterative machine-vision technique, as described in \red{(REF)}.
 
@@
 The signature is found by summing the number of refined cells with the same index along the longest dimension.
-Edges are typically indicates by inflection points in this signature, so we calculate the second derivative using a simple three-point finite difference estimator and look for the maximum in the second derivative, with ties going to the index closest to the center.
+Edges are typically indicated by inflection points in this signature, so we calculate the second derivative using a simple three-point finite difference estimator and look for the maximum in the second derivative, with ties going to the index closest to the center.
 If no inflection points are found along the longest exis, the next-longest axis is searched.
@@
 However, it has a number of advantages.
-First, traditional grid-based hydro method can be simply added with only minor modifications.
+First, traditional grid-based hydro methods can be simply added with only minor modifications.
 Second, it is easier to accommodate larger stencils (as in PPM, which requires three boundary zones), and third, it can be more efficient for some machines since there are more predictable and linear memory accesses, which more efficiently use a CPU's cache.

### 2013-05-19 numerical-chemistry.tex
Reynolds edits.

@@
 This makes an implicit scheme strongly preferable for such a set of
-equations.  However, an implicit scheme typically require an iterative
+equations.  However, an implicit scheme typically requires an iterative
 procedure to converge, and for large networks (such as this one) an
@@
 
-\enzo\ solves the rate equations using a method based on a backwards
-differencing formula (BDF) in order to provide a stable and accurate
+\enzo\ solves the rate equations using a method based on a first-order-accurate, implicit
+Euler formulation in order to provide a stable and accurate
 solution. This technique is optimized by taking the chemical
@@
 the other species and the hydrodynamic variables.  This allows a large
-speedup in solution as the BDF scheme is then applied only to the
+speedup in solution as the implicit Euler scheme is then applied only to the
 slower 7- or 10-species network (depending on whether deuterium is
@@
 subcycling, the method is not truly implicit over an entire
-hydrodynamical timestep, and thus not truly a BDF method -- however,
+hydrodynamical timestep  -- however,
 it uses the formalism and is implicit over subcycle timesteps.
@@
 $n_{\rm H} \sim 10^{16} - 10^{17}$~cm$^{-3}$ in gas of primordial composition.  Beyond this point,
-modifications the cooling function that take into account the
+modifications to the cooling function that take into account the
 non-negligible opacity in the gas must be made, as discussed by

### 2013-05-19 numerical-mhd-ct.tex
Reynolds edits.

--- 2013-05-16 numerical-mhd-ct.tex
@@
 to evolve the magnetic field in a manner that preserves $\Divb=0$ to
-to limits of machine precision.  The electric field is computed using
+limits of machine precision.  The electric field is computed using
 the fluxes from the Riemann solver.  The curl of that electric field
@@
 interpolated in concert.  Second, any magnetic field information from
-previous time steps must be included in the interpolation, making the
+the previous time step must be included in the interpolation, making the
 interpolation more complex than the simple parent-child relation used

### 2013-05-19 numerical-overview.tex
Many minor changes.  No more undefined references!  Added appendix file.
Reynolds edits.

@@
 %---------------------------------------------------------------------------%
-\subsection{Chemistry, radiation and cooling}\label{sec.ov.chem}
+\subsection{Chemistry, radiation and cooling}
+\label{sec.ov.chem}
 
@@
 
-\enzo\ solves the rate equations using a method based on a backwards differencing 
-formula (BDF) in order to provide a stable and accurate solution.  This technique is
+\enzo\ solves the rate equations using a method based on an implicit
+Euler formulation in order to provide a stable and accurate solution.  This technique is
 optimized by taking the chemical intermediaries $H^-$ and $H_2^+$, which have 
@@
 the other species and the hydrodynamic variables.  This allows a large speedup
-in solution as the BDF scheme is then applied only to the slower 7-species network
+in solution as the implicit scheme is then applied only to the slower 7-species network
 on timescales closer to those required by the hydrodynamics of the simulation.
@@
 %Figure~\ref{fig.2.amrhierarchy} for a schematic example of this.
+
+%%% Local Variables: 
+%%% mode: latex
+%%% TeX-master: "ms"
+%%% End:

### 2013-05-19 numerical-particle.tex
Reynolds edits.

@@
 velocities starting from $\myvec{x}^n$ and $\myvec{v}^n$ using a
-standard drift-kick-drift technique:
+standard drift-kick-drift technique \citep{Hockney88}:
 

### 2013-05-19 numerical-ppm.tex
Reynolds edits.

@@
 and $E$) and then averaging over the regions corresponding to the
-distance each characteristic wave can travel ($u$, $u-c$ and $u+c$).
+distance each characteristic wave can travel ($u$, $u-c_s$ and
+$u+c_s$, where $c_s$ is the sound speed).
 Second, a Riemann problem is solved using these effective left and
@@
 %
-This is Equation~(1.4) of CW84 (in the radial rather than mass coordinate as is appropriate for the direct rulerian approach). 
+This is Equation~(1.4) of CW84 (in the radial rather than mass coordinate as is appropriate for the direct Eulerian approach). 
 %The zones edges in the mass coordinates are $m_{j-1/2}$ and $m_{j+1/2}$ for the left and right edges, respectively.  
@@
 ${\eta}_1 = 10^{-3}$, which is consistent with the truncation error of the
-scheme.  We are now free to select the method by which the internal energy
+scheme for grid sizes that are typically used in our simulations.  We are now free to select the method by which the internal energy
 field variable $e$ is updated so that it will not become contaminated with
@@
 This scheme is optional and is generally only required in large-scale 
-cosmological simulations where the gas cools to the expansion of the universe
+cosmological simulations where the gas cools due to the expansion of the universe
 but large bulk flows develop due to the formation of structure.

### 2013-05-19 numerical-radiation-fld.tex
Reynolds edits.

@@
 a field-based radiation transport solver for problems posed on uniform
-(i.e.~non-AMR grids).  Detailed explanations of the model and
+(i.e.~non-AMR) grids, which has been tuned for large-scale simulations
+involving many ionizing sources.  Detailed explanations of the model and
 solution approach may be found in
@@
 and $\bar{e}_c = \left(e_c^{n}+e_c^{n+1}\right)/2$.  These equations
-may each be solved {\em analytically} for their solution at the time
+may each be solved analytically for their solution at the time
 $t$, which we denote by

### 2013-05-19 numerical-radiation-homogeneous.tex
Adding AMR scaling section with an associated table.  I still need to add some references in this section.  Fixing some undefined and multiple labels and making the UVB table wider and spaced nicely.
Reynolds edits.

@@
 where $\alpha_0 = 1.5$ by default and the spectrum is normalized at
-the HI ionization edge.  Type 6 has the same spectrum by is attenuated
+the HI ionization edge.  Type 6 has the same spectrum, but attenuated
 by a column density of $10^{22}$ cm$^{-3}$ neutral hydrogen.
@@
 
-\begin{table}
+\red{Note: we talk about several radiation field types that are not in
+Table~\ref{table:homo_coefs} (4...14).  Do we want to increase the
+length of the table, remove the text from the section, or perhaps just
+say something in this section about how the table is for illustrative purposes? }
+
+\begin{table*}
 \begin{center}
 \caption{Homogeneous radiation field coefficients}
-\begin{tabular}{lrrrrrr}
+\begin{tabular*}{0.9\textwidth}{@{\extracolsep{\fill}}lrrrrrr}
 \tableline\tableline
@@
 He$^+$ heating &   $2.17 \times 10^{-25}$ & -0.2196 & -1.070 & 0.2124 & 1.1782 & -0.9213 \\
-\end{tabular}
+\end{tabular*}
 %\tablecomments{}
@@
 \end{center}
-\end{table}
+\end{table*}

### 2013-05-19 numerical-radiation-raytracing.tex
Reynolds edits.

@@
 $(r/R)^2$ rays, which is computationally wasteful because only a few
-rays per cel  are required to provide an accurate calculation
+rays per cell  are required to provide an accurate calculation
 of the radiation field (as we will show later).

### 2013-05-19 numerical-starformation.tex
Reynolds edits.

@@
 
-\red{( (BWO) I am a little worried that this section is too
-  long. The editorial hatchet may be called for, but I think somebody
-  else (who didn't write the section in the first place) needs to
-  think it through. )}
+% \red{( (BWO) I am a little worried that this section is too
+%   long. The editorial hatchet may be called for, but I think somebody
+%   else (who didn't write the section in the first place) needs to
+%   think it through. )}
 
@@
 is explicitly geared towards modeling star formation in a
-phenomenological way on kiloparsec scales.  Stars are assume to form
+phenomenological way on kiloparsec scales.  Stars are assumed to form
 with a characteristic gas timescale $\tau_*$ such that $\dot{\rho}_* =
@@
 and is treated as a single-age stellar population (as in the previous
-section).  Kravtsov assumes that the IMF is described by a Miller \&
+section).  Kravtsov assumes that the stellar initial mass function (IMF) is described by a Miller \&
 Scalo functional form with stellar masses between $0.1$ and
@@
 their main-sequence lifetimes, explosion energies, ejected gas mass,
-and ejected metal are caclulated using a variety of sources that
+and ejected metal are calculated using a variety of sources that
 depend on the mass of the star, and are described in detail in Section

### 2013-05-19 overview.tex
Reynolds edits.

@@
 We begin this section by first writing down the complete set of
-physical equations that be solved by \enzo, and then briefly describe
+physical equations that can be solved by \enzo, and then briefly describe
 the numerical algorithms that we use to solve these equations.  This
@@
 %
-The particles also make a contribution to the potential through their contribution to Poisson's equation.
+The particles also make a contribution to the gravitational potential through their contribution to Poisson's equation.
 
@@
 extra grids.  These criteria depend on the physical problem being
-simulated.  We have implemented a number of options, including shock
-and steep gradient detectors, but in the problem described here,
-employ one based on the mass within a cell, imitating the Lagrangian
-nature of the SPH algorithm.  Once a grid has a set of flagged cells,
+simulated -- see Section~\ref{sec:refinement_criteria} for more details.  Once a grid has a set of flagged cells,
 we run a machine-vision based algorithm \citep{Berger91} to find edges
@@
 % \item 
-(ii) Compute interface states by averaging the parabola over the domain of dependence for each interface
+(ii) Compute interface states by averaging the parabola over the domain of dependence for each interface.
 % \item 
@@
 order-accurate in space and time and which explicitly conserves mass,
-linear momentum, and energy.  A variety of reconstruction methods and
-Riemann solvers have been implemented.
+linear momentum, and energy.  \red{(REF?)}  A variety of reconstruction methods and
+Riemann solvers have been implemented.  \red{(These are never described!)}
 
@@
 magnetic field amplification during the formation of galaxies
-\citep{Wang:2009a} and to understand the role proto-stellar jets for
+\citep{Wang:2009a} and to understand the role of proto-stellar jets for
 the theory of star formation \citep{Wang:2009b}. It combines the standard
 approach of Godunov \citep{Godunov1959} for finite volume techniques
-with the method of lines as described as e.g. by
-\cite{leveque2002finite} and \cite{toro-1997}. In addition, however,
+with the method of lines as described by
+\cite{leveque2002finite} and \cite{toro-1997}. In addition, 
 it implements the hyperbolic divergence cleaning algorithm of
@@
 the PPM hydro solver which have been ported to Nvidia's CUDA framework
-allowing \enzo to take advantage of modern graphics hardware \citep{Wang:2010}.
+allowing \enzo\ to take advantage of modern graphics hardware \citep{Wang:2010}.
 
@@
 and only the hydrodynamical algorithm of \zeus\ is implemented in
-\enzo -- the MHD and RHD schemes are not.
+\enzo\ -- the MHD and RHD schemes are not.
 
@@
 Fourier Technique \citep{Hockney88} to solve Poisson's equation on the
-coarsest (Level 0) grid.  The advantage of using this method is that
+coarsest (Level 0) grid on each time step.  The advantage of using this method is that
 it is fast, accurate, and 
@@
 that grid's parent (either the level 0 grid or some other subgrid).
-The Poisson equation is then solved using a multigrid technique.  In
+The Poisson equation is then solved on every timestep using a multigrid technique on one
+subgrid at a time.  In
 \enzo, self-gravity is optional.  Aside from calculating the
@@
 particles are advanced in a single time step using a drift-kick-drift
-algorithm to provide second-order accuracy even in the presence of
+algorithm\citep{Hockney88} to provide second-order accuracy even in the presence of
 changing timesteps.  Since the particles follow the collapse of
@@
 species important for gas-phase molecular hydrogen formation: $H^-$,
-H$_2$ and H$_2^+$, and then $HD$ formation: $HD$, $D$, $D+$.  The
+H$_2$ and H$_2^+$, and then $HD$ formation: $HD$, $D$, $D+$.  (We note
+that a reduced version of the chemistry, used only with the
+flux-limited diffusion radiation transport, includes only hydrogen
+species.)  The
 cooling and heating due to these species is also included (see the
 next section).  The solution of the rate equations is carried out
-using a time-backwards differencing which insures stability; to
-maintain accuracy we sub-cycle the rate equations with a shorter
+using one Jacobi iteration of an implicit Euler time discretization to
+ensure stability; to
+ensure accuracy we sub-cycle the rate equations with a shorter
 timestep such that the electron and neutral fractions do not change by
@@
 species of Hydrogen and Helium (including H$_2$) -- this primordial cooling is
-also treated using a backwards-time differencing scheme with
-sub-cycling to maintain accuracy.  Finally, we have also recently
+also treated using 
+one Jacobi iteration of an implicit Euler time discretization to maintain accuracy.  Finally, we have also recently
 added metal cooling based on a set of multi-dimensional lookup tables
@@
 includes a number of pre-set radiation backgrounds that are uniform in
-space but can very in time.  These are generally based on the
+space but can vary in time.  These are generally based on the
 redshift-dependent rates given in \citet{1996ApJ...461...20H} and
@@
 A second option for radiative transfer is a moment-based method that
-adds an additional field tracking the radiant energy density.
+adds an additional field tracking the radiation energy density.
 This field is evolved using the flux-limited diffusion method, which
@@
 resulting set of linear equations is solved using the parallel HYPRE
-framework.  Full details for the enzo implementation can be found in
+framework.  Full details for the \enzo\ implementation can be found in
 \citet{ReynoldsHayesPaschosNorman2009}.
@@
 Heat conduction, both isotropic and anisotropic, can be included using
-an sub-cycled, operator-split method.  The heat fluxes are computed
+a sub-cycled, operator-split method.  The heat fluxes are computed
 with simple second-order accurate finite differences and stability is

### 2013-05-19 parallel.tex
Reynolds edits.

@@
 and any chemistry data) and particle data are stored on only one
-processor.  Storing the entire hierarchy on all processor does create
+processor.  Storing the entire hierarchy on all processors does create
 some overhead, but in practice this is not an issue until one reaches
-the extremely large scale.
+extremely large scales.
   The code handles load balancing 
@@
 necessarily maximum efficiency).  
-The MPI message passing library\footnote{http://www-unix.mcs.anl.gov/mpi/}
+The MPI message passing library
  is used to transfer data between processors.

### 2013-05-21 code-tests.tex
Added gravity and self-similar tests

@@
 \label{sec.test.gravitypointsource}
-\red{(Greg)}
-This is the TestGravity (problem 23) test problem.  It tests gravity around a point source, using fixed AMR.
+
+This test is a simple test of the ability of the Poisson solver to correctly reproduce the acceleration around a point source without any dynamics.
+We place a single particle in the center of the unit domain with root grid dimensions $32^3$ and add a static AMR grid with refinement factor two that covers the central $1/8^3$ of the domain (i.e.
+so the refined grid is only $8^3$).
+We then place 5000 particles throughout the domain distributed uniformly in angle and with the logarithm of the radius (from the central point).
+The measured accelerations are shown in Figure~\ref{fig.gravitytest}.
+As is typical for Particle-Mesh based methods, the errors decrease at large distances and peak on roughly the grid scale.
+The accelerations are relatively smooth across the top grid/fine grid boundary with the most noticeable impact being a drop in the amplitude of tangential errors on the fine grid.
+The force law rolls over at about 1.6 fine cell widths (equivalent to 0.8 root grid widths in this figure), consistent with the CIC-deposition and interpolation used.
+
+\begin{figure}
+\begin{center}
+\includegraphics[width=0.6\textwidth]{figures/GravityTest.eps}
+\caption{The radial and tangential accelerations measured around a point source placed in a static AMR grid.  The solid line shows the analytic radial profile.}
+\label{fig.gravitytest}
+\end{center}
+\end{figure}
 
@@
 \label{sec.tests.infall}
-\red{(Greg)}
-Problem type 24.  This is a test based on Bertschinger's 1985 3D self-similar infall
-solution, and tests gravity + hydro.
+
+The next test problem is based on the self-similar solutions to the cosmological spherical infall problem found by \citet{Bertschinger1985}.
+It features a small density perturbation in a homogeneous $\Omega = 1$ universe and is a strong test of the cosmological evolution, gravity, and hydrodynamic portions of the code.
+It is close analogue to cosmological cluster formation and has an analytic solution.
+
+For initial conditions, we adopt a $32^3$ top grid with a single, initial sub grid (covering $1/8^3$ of the domain) with a refinement factor of 2.
+An overdensity $\delta = 40$ is placed in a single cell near the center of the domain.
+We begin at $z=199$ and evolve to $z=0$, which is a sufficient time for the evolution to largely forget its initial conditions and approach the self-similar result.
+An over density refinement criteria ($\delta_{\rm crit} = 1.1$ on the top grid) is used to add additional grids, going up to 5 additional levels beyond the root grid.
+We use the PPM solver without radiative cooling; only baryons are used for this problem.
+An ideal gas law with $\gamma = 5/3$ is adopted.
+
+In Figure~\ref{fig.sphericalinfall}, we show the results, scaled according to the dimensionless variables as defined in equations (2.9) and (3.2) of \citet{Bertschinger1985}.
+This demonstrates that we can quickly and easily obtain a good solution with only a fairly modest initial grid.
+The shock is sharply resolved and the asymptotic profiles as small $\lambda$ are recovered.
+At very small values of $\lambda$, the initial conditions have not been fully forgotten and the self-similar solution is not recovered -- this can be seen most clearly in the dimensionless mass; however, this is simply because of the limited amount of time for which we evolved the solution.
+
+\begin{figure}
+\begin{center}
+\includegraphics[width=0.7\textwidth]{figures/SphericalInfall.eps}
+\caption{The results of the self-similar spherical infall test.
+The panels show, as a function of the dimensionless radius $\lambda$, the dimensionless density $D$ (top left), velocity $V$ (top right), pressure $P$ (bottom left), and enclose mass $M$ (bottom right).
+In each case we show azimuthally averaged profiles from the simulation as circles and the analytic solution for for $\gamma = 5/3$ as solid lines.}
+\label{fig.sphericalinfall}
+\end{center}
+\end{figure}
 

### 2013-05-21 numerical-starformation.tex
Added gravity and self-similar tests

@@
 Due to the computationally unfeasible number of stars in a galaxy
-(10^11) and the lack of detailed understanding of star formation, a
+($10^{11}$) and the lack of detailed understanding of star formation, a
 number of phenomenological star formation models are included in

### 2013-05-21 test_problems/SedovAMR/sedov_outputdata.py

@@ -1,4 +1,4 @@
-from yt.mods import *
+:from yt.mods import *
 import matplotlib.pyplot as plt
 
 # Load the dataset.

### 2013-05-21 test_problems/GravityTest/make_plot.py
Adding my test problem parameter files and scripts.

@@ -1,0 +1,32 @@
+import numpy as na
+import matplotlib.pyplot as plt
+
+Data = na.loadtxt("TestGravityCheckResults.out")
+
+# Read in data from text file produced by code
+
+radius = Data[:,0]
+ForceTangentialComputed = Data[:,1]
+ForceRadialComputed = Data[:,2]
+ForceRadialTrue = Data[:,3]
+
+Error = (ForceRadialComputed-ForceRadialTrue)/ForceRadialTrue
+indices = na.where((radius > 1.0) & (radius < 8.0))
+rmsError = na.std(Error[indices])
+print "rms error = "+str(rmsError)
+
+# Plot the computed radial force again the r-2 profile 
+#  (which should be equal except for outer part where the periodic 
+#   boundary conditions are important, and inner part, where grid
+#   softening is important).
+# Also plot tangential component of force, which should be zero
+#  (or at least small compared to radial component).
+
+plt.loglog(radius, ForceRadialComputed, label=r'F$_{\rm rad}$', ls='None', marker='.')
+plt.loglog(radius, ForceRadialTrue, label=r'F$_{\rm rad}$ (true)')
+plt.loglog(radius, ForceTangentialComputed, label=r'F$_{\rm tang}$', ls='None', marker='+')
+plt.xlabel('r (root grid cells)')
+plt.ylabel('Acceleration')
+plt.axis([0.2,14.0,1.0e-7,1.0e0])
+plt.legend()
+plt.savefig('GravityTest.eps')

### 2013-05-21 test_problems/SelfSimilarInfall/make_plots.py

@@ -1,0 +1,109 @@
+from yt.mods import *
+import pylab
+
+# This script generates a plot for the self-similar infall test
+
+### define problem name
+problem_name = 'SphericalInfall'
+
+### define simulation output directory and filename base
+output_dir_base = 'DD'
+datafile_base = 'data'
+
+### define output to be plotted
+dumpid = '0024'
+
+# This is the effective resolution of the initial conditions
+Ninitial = 64
+
+# This is the initial perturbation amplitude of a single cell in the IC's
+delta_init = 40.0
+
+### define some filenames
+exact_solution_filename = './exact_solution_gamma_53.txt'
+plot_filename = './' + problem_name + '.eps'
+filename = output_dir_base + dumpid + '/' + datafile_base + dumpid
+
+### read exact solution
+exact = pylab.csv2rec( exact_solution_filename, delimiter=' ', names=('lambda', 'V', 'D', 'P', 'M') )
+
+# open output
+pf = load(filename)
+
+
+# Compute radial profile
+r_min = pf.h.get_smallest_dx()*pf.units["cm"]
+r_max = 0.4*pf.units["cm"]
+center = pf.domain_center*(1.0+1.0/Ninitial)
+print "center = ", center
+center = pf.h.all_data().quantities["MaxLocation"]('Density')[2:5]
+print "center = ", center
+
+slc = SlicePlot(pf, 'z', ['Density','x-velocity', 'Temperature'], center='max')
+slc.save()
+
+sphere = pf.h.sphere(center, r_max)
+profile = BinnedProfile1D(sphere, 40, "Radius", r_min, r_max)
+profile.add_fields("Density", weight="CellVolume")
+profile.add_fields("RadialVelocity", weight="CellVolume")
+profile.add_fields("Pressure", weight="CellVolume")
+profile.add_fields("CellMass", weight=None, accumulation=True)
+#profile.write_out('test.txt', '%0.6e', 'left')
+
+
+# Convert to Bertschinger dimensionless variables (eq. 2.9 and 3.2 of Bertschinger 1985)
+G          = 6.67e-8
[…]

### 2013-05-21 test_problems/ShockPool/make_plots.py

@@ -1,0 +1,103 @@
+from yt.mods import *
+import pylab
+
+# This script generates a 3x2 panel of plots for the Shock Pool
+#   for three different hydro solvers at three different times.
+#   It assumes the simulations have been run and are in the
+#   directory structure PPM, Zeus, MUSCL
+
+### define problem name
+problem_name = 'ShockPool'
+
+### define simulation output directory and filename base
+output_dir_base = 'DD'
+datafile_base = 'data'
+
+
+### define some filenames
+plot_filename = './' + problem_name + '.pdf'
+
+### generate exact solution
+def make_exact(time):
+    dens1 = 1.0
+    Gamma = 1.4
+    Mach  = 2.0
+    dens2 = dens1*(Gamma + 1.0)*Mach**2 / ((Gamma - 1.0) * Mach**2 + 2.0)
+    speed = Mach*pylab.sqrt(Gamma*1.0/dens1)
+    xexact = pylab.arange(0.0, 1.0, 0.001)
+    deltime = time - xexact/speed
+    exact = pylab.ones(shape=deltime.shape)*dens1
+    index = pylab.all([deltime > 0],axis=0)
+    exact[index] = dens2
+    return (xexact, exact)
+
+
+#define plot a single frame
+def plot_one(filename, name, axes, xbegin1, xwidth):
+    print "opening " + filename
+
+    # open output and create ray
+    pf = load(filename)
+    ray = pf.h.ortho_ray(0, [0.5, 0.5])
+
+    # Get exact result
+    (xexact, exact) = make_exact(pf.current_time)
+
+    # rough computation of level for each point in ray (probably better way)
+    level = -pylab.log(ray['dx']/ray['dx'][0])/pylab.log(2.0)
+
+    #get axis
+    ax = pylab.axes(axes)
+
+    #plot grey for refined region
+    x1 = pylab.array([0.25, 0.75])
+    y1 = pylab.array([0.0, 0.0])
+    ax.fill_between(x1, y1, y1+3, facecolor='grey', alpha=0.2)
+
+    # plot
+    pylab.axhline(0,color='k',linestyle='dotted')
+    pylab.scatter(ray['x']+0.5*ray['dx'],ray['Density'], c=level, s=40, linewidths=(0,))
+    pylab.plot(xexact, exact)
+
[…]

### 2013-05-21 test_problems/ShockTube/make_plots.py

@@ -1,0 +1,107 @@
+from yt.mods import *
+import pylab
+
+# This script generates a 3x2 panel of plots for the SodShockTube
+#   for three different hydro solvers with and without AMR.
+#   It assumes the simulations have been run and are in the
+#   directory structure PPM/AMR, PPM/noAMR, Zeus/AMR, etc.
+
+### define problem name
+problem_name = 'SodShockTube'
+
+### define simulation output directory and filename base
+output_dir_base = 'DD'
+datafile_base = 'data'
+
+### define output to be plotted
+dumpid = '0001'
+
+### define some filenames
+exact_solution_filename = './SodShockTube_t=0.25_exact.txt'
+plot_filename = './' + problem_name + '.eps'
+
+### read exact solution
+exact = pylab.csv2rec( exact_solution_filename, delimiter=' ', names=('x', 'Density', 'x-velocity', 'Pressure', 'InternalEnergy') )
+
+# define error norm text label function
+def error_label(name, norm, maxnorm, xpos, ypos):
+    thistext = name + '\n'
+    vexp =  pylab.floor( pylab.log10( norm ) )
+    if norm == 0: vexp = 0
+    vmant = norm / 10**vexp
+    thistext += r'$\Vert$E$\Vert_1 = \, %3.1f' % vmant
+    if vexp != 0:
+        thistext += r' \times \, 10^{%2d}' % vexp
+    thistext += '$\n'
+
+    vexp =  pylab.floor( pylab.log10( maxnorm ) )
+    if maxnorm == 0: vexp = 0
+    vmant = maxnorm / 10**vexp
+    thistext += r'$\Vert$E$\Vert_\infty \! = \, %3.1f' % vmant
+    if vexp != 0:
+        thistext += r' \times \, 10^{%2d}' % vexp
+    thistext += '$'
+
+    pylab.text(xpos, ypos, thistext, va='top')  
+
+
+#define plot a single frame
+def plot_one(filename, name, axes):
+    field = 'Density'
+    print "opening " + filename
+
+    # open output and create ray
+    pf = load(filename)
+    ray = pf.h.ortho_ray(0, [0.5, 0.5])
+
+    # first interpolate the exact solution onto the ray
+    ray_exact = {'x': ray['x'], 
[…]

### 2013-05-21 test_problems/WavePool/make_plots.py

@@ -1,0 +1,103 @@
+from yt.mods import *
+import pylab
+
+# This script generates a 3x2 panel of plots for the Wave Pool
+#   for three different hydro solvers at three different times.
+#   It assumes the simulations have been run and are in the
+#   directory structure PPM/AMR, PPM/noAMR, Zeus/AMR, etc.
+
+### define problem name
+problem_name = 'WavePool'
+
+### define simulation output directory and filename base
+output_dir_base = 'DD'
+datafile_base = 'data'
+
+
+### define some filenames
+plot_filename = './' + problem_name + '.eps'
+
+### generate exact solution
+def make_exact(time):
+    dens0 = 1.0
+    wavelength = 0.1
+    Gamma = 1.4
+    speed = pylab.sqrt(Gamma*1.0/dens0)
+    omega = 2.0*pylab.pi*speed/wavelength
+    xexact = pylab.arange(0.0, 1.0, 0.001)
+    deltime = time - xexact/speed
+    exact = pylab.zeros(shape=deltime.shape)
+    index = pylab.all([deltime*speed < wavelength, deltime > 0],axis=0)
+    exact[index] = 0.01*pylab.sin(omega*deltime[index])
+    exact = dens0*(1.0 + exact)
+    return (xexact, exact)
+
+
+#define plot a single frame
+def plot_one(filename, name, axes, xbegin1, xwidth):
+    print "opening " + filename
+
+    # open output and create ray
+    pf = load(filename)
+    ray = pf.h.ortho_ray(0, [0.5, 0.5])
+
+    # Get exact result
+    (xexact, exact) = make_exact(pf.current_time)
+
+    # rough computation of level for each point in ray (probably better way)
+    level = -pylab.log(ray['dx']/ray['dx'][0])/pylab.log(2.0)
+
+    # plot
+    ax = pylab.axes(axes)
+    pylab.axhline(0,color='k',linestyle='dotted')
+    pylab.scatter(ray['x']+0.5*ray['dx'],ray['Density'], c=level, s=40, linewidths=(0,))
+    pylab.plot(xexact, exact)
+
+    #plot dashed lines for refined region
+#    pylab.plot([0.25, 0.25], [0.9, 1.1], linestyle="--") 
[…]

### 2013-05-21 code-tests.tex
Splitting up test problem section into subsections.

--- 2013-05-19 code-tests.tex
@@
 
-\subsubsection{The \enzo\ test suite}
-\label{sec.tests.suite}
-
-\enzo\ is capable of simulating a large variety of problems types,
-with all but a few of these types requiring only a parameter file as an
-input.  The notable exception is the cosmology simulation, which takes
-as input initial conditions created by other codes.  The suite of test
-problems spans a wide range in complexity.  At one end of this spectrum
-are simple problems that utilize only a single component of \enzo\ and
-for which analytic solutions exist for comparison with the simulation
-results.  At the opposite end are problems that exercise a large
-portion of \enzo's machinery in concert.  Together, the available
-problem types fully cover all of \enzo's functionality.  This enables
-them to serve as a vehicle for verifying that the code's behavior
-remains stable over time and after modifcation of the codebase.
-
-\enzo\ uses an automated testing framework that allows a user to run,
-with a single command, a set of test problems and compare the results
-against results produced by any other version of the code.  Within the
-\enzo\ source distribution, the test problem parameter files are stored
-in a nested directory structure organized according to the primary
-functionality tested, e.g., hydrodynamics, gravity, cooling, etc.
-Each parameter file is accompanied by a text file containing various
-descriptive keywords, such as the machinery tested, the
-dimensionality, and the approximate run time.  A test runner script is
-responsible for taking as input from the user a series of keywords
-that are used to select a subset of all available test problems.  The
-test problems are also grouped into three suites: the quick, push, and
-full suites, each a superset of the ones named before.  The quick
-suite is considered to minimally cover the primary functionality in
-\enzo\ and is designed to be run repeatedly during the development
-process.  The push suite has slightly increased feature coverage and
-is mandated to be run before code changes are accepted into the main
-repository.  The full suite consists of all test problems that can be
-run with no additional input.  Approximate run times for the quick,
-push, and full suite are 15 minutes, 1 hour, and 60 hours,
-respectively, on a relatively new desktop computer (circa 2013).
-
-After the test problems are selected by the test runner script, they
-are run in succession using either the \enzo\ executable contained
-within that distribution or an external executable built from another \enzo\
-version and specified by the user.  This allows for any version of the
-code to be tested with an identical set of test problems.  After
-running the test problem simulations, the test runner then performs a
-series of basic analysis tasks using the \texttt{yt} analysis toolkit
-\citep{2011ApJS..192....9T, 2011arXiv1112.4482T}.  The default
-analysis performed on all test problems includes calculation of
-various statistics, such as extrema; mean; and variance, on the fields
-present in the output data.  Custom analysis provided by scripts that
-accompany the test problem parameter files is run for special cases,
-such as when an analytical solution exists that can be compared
-against the simulation data.  After the analysis is performed, the
-results are compared against a set of gold standard results that are
-maintained on a website and downloaded on the fly by the test runner
-script.  Alternately, results from any version of \enzo\ can be stored
-locally and compared to any other version of the code.  In
-Section~\ref{sec.tests.problems}, we describe some of the key test
-problems that are used to verify proper behavior.  All of these test
-problems, as well as the scripts to generate the figures from them,
-are available in the \enzo\ test suite, allowing the interested reader
-to regenerate all of the figures in Section~\ref{sec.tests.compare}
-using the \textit{'methodpaper'} suite of test problems.
-
-\subsubsection{Code comparisons}
-\label{sec.tests.compare}
-
-Over the course of its existence, \enzo\ has been involved in numerous
-comparisons with other astrophysical codes used for self-gravitating
-fluid dynamics.  In general, \enzo\ behaves in a manner similar to other
-grid-based (and particularly AMR-based) codes, as we will summarize below.
-
-\enzo\ has been involved in multiple cosmological code comparisons,
-including the Santa Barbara Cluster code comparison project
-\citep{SantaBarbara}, a large comparison of N-body simulations
-\citep{2008CS&D....1a5003H}, as well as several direct comparisons
-between \enzo\ and the GADGET SPH code under a variety of
-circumstances, including N-body and adiabatic hydrodynamics
-\citep{2005ApJS..160....1O,2005MNRAS.364..909V, 2011MNRAS.418..960V}
-and simulations of the Lyman-alpha forest \citep{2007MNRAS.374..196R}.
-The \enzo\ code typically has a more difficult time resolving
-small-scale self-gravitating structures (for an equivalent dark matter
-particle count and nominally equivalent force resolution), but does
-comparably well as a tree-based code for larger structure, and is
-typically superior in terms of resolving fluid features due to its
-higher-order (and artificial viscosity-free) Piecewise Parabolic
-Method hydrodynamics solver.  When examining classical test cases such
-as the Santa Barbara project \citep{SantaBarbara}, \enzo\ forms galaxy
-clusters with very similar density, temperature, and entropy profiles
-to other grid-based codes that use Godunov-type hydro methods, which
-systematically differed from particle-based codes using SPH in this
-code comparison.  Similarly, in galaxy cluster simulations that look
-specifically at the properties of cosmological shocks
-\citep[e.g.][]{2011MNRAS.418..960V}, \enzo\ produces results that are
-similar to other high-order grid-based hydrodynamics codes, and far
-superior performance of \enzo\ is observed (in terms of resolution of
-fluid features and shock detection) in low-density regions when
-compared to a particle-based code.  In tests of the Lyman-alpha forest
-that include radiative cooling and a uniform metagalactic ultraviolet
-background, \enzo\ and Gadget provide results on metrics such as the
-matter power spectrum that are comparable to within 5\%
-\citep{2007MNRAS.374..196R}.
-
-Several comparisons have been made that focus specifically on
-hydrodynamics solvers and fluid behavior.  In particular, the work of
-\citet{2007MNRAS.380..963A} and \citet{Tasker2008} perform
-head-to-head comparisons between several grid- and particle-based
-codes for a variety of fluid-centric test problems (including shocked gas clouds,
-self-gravitating, translating clouds, and Sedov-Taylor explosions),
-and show that \enzo\ is comparable or superior in behavior to the
-other grid-based hydrodynamics codes involved in the comparison, and
-provide useful information on the sort of practical challenges that a
-user of an
-AMR code such as \enzo\ may experience.  More specific comparisons, including one
-specifically testing the linear and nonlinear growth of the
-Kelvin-Helmholz instability \citep{2012ApJS..201...18M}, as well as a
-comparison that more
-broadly examines Galilean invariance in grid-based codes
-\citep{2010MNRAS.401.2463R}, show that \enzo, and in particular its
-implementation of the Piecewise Parabolic Method hydrodynamics solver,
-converge to the correct solution as expected, and generally provide
-less diffusive solutions than lower-order codes, including those that
-use artificial viscosity.  Finally, two code comparison projects have
-been undertaken to examine turbulence in hydro codes -- one studied
-the behavior of decaying isothermal supersonic turbulence
-\citep{2009A&A...508..541K}, and the other looked at supersonic
-magnetohydrodynamical turbulence \citep{2011ApJ...737...13K}.  Both
-included the \enzo\ code, with the former testing the PPM
-hydrodynamics and the latter the constrained transport MHD
-implementation of \citet{Collins10}.  In both cases, \enzo\ performed similarly to other
-grid-based codes that use Godunov-based fluid solvers, and typically
-had better effective resolution than particle-based codes using the
-same number of particles as the \enzo\ simulation used grid cells.
-
-Three other comparisons between the \enzo\ code and other simulation
-tools have been performed that focus on physics other than gravity and
-fluid flow.  The flux-limited diffusion radiation transport scheme was
-measured against several test problems by \citet{IlievEtAl2009},
-producing results that are similar to both expected results (i.e.,
-analytic test problems) and other methods (i.e., codes using other
-methods, written by other groups), although there are some minor
-differences (although all codes in the study exhibited differences
-from the majority of other codes on at least a subset of the tests).
-\citet{2011ApJ...726...55T} shows the results of varied reaction rate
-coefficients for the formation of molecular hydrogen via the 3-body
-process in both the \enzo\ and Gadget-2 codes.  Similar trends were
-visible in both the \enzo\ and Gadget simulations, though at a
-nominally equivalent resolution (i.e., with particle and cell gas
-masses being comparable) \enzo\ simulations typically displayed a
-substantially higher level of gas structure.  This is unsurprising due
-to \enzo's higher-order hydro solver.  Finally,
-\citet{2012ApJ...744...52P} show the results of comparing \enzo\ in
-its non-AMR mode to a smoothed-particle hydrodynamics code, SNSPH, in
-the context of common-envelope binary stellar interactions.  The
-authors show that the codes display reasonable convergence properties
-as a function of simulation resolution, and also agree quite well with
-each other -- however, the observed mass-loss rates do not agree
-particularly well with observations.
+\input{tests-vandv}
 
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsubsection{Sod Shock Tube}
-\label{sec.tests.sodshock}
+\input{tests-sodshocktube}
 
-We begin the paper test suite with the classic Sod shock tube problem
-\citep{Sod78}, which provides a good test of a hydrodynamical solver's
-ability to resolve a clean Riemann problem with clear separation
-between the three waves.  These waves include a rarefaction fan, a
-contact discontinuity, and a moderate shock.  The initial state is 
-$\rho_{\rm L}, P_{\rm L} = 1.0, 1.0$ on the left of the boundary at 
-$x=0.5$ and $\rho_{\rm R}, P_{\rm R} = 0.125, 0.1$ on the right.
-All velocities are initially zero.  In
-Figure~\ref{fig.sodshocktube}, we show the density solution at the
-final time ($t=0.25$) for three of our hydro solvers -- the spatially third-order
-PPM, as well as the two second-order \zeus\ and MUSCL schemes.  We use
-100 cells across the domain, (which is a relatively standard choice in
-code method papers) and show solutions both with and without adaptive
-mesh refinement.  Without AMR (in the top row of the figure), it is
-clear that the PPM scheme produces by far the cleanest solution, with
-all wave families crisply reproduced (in particular, the contact
-discontinuity and shock).  \zeus\ and MUSCL produce similar results,
-with MUSCL doing a slightly better job on the rarefaction fan.  In the
-figure, we also show the mean absolute deviation from the exact
-solution ($||E_\infty||$), as well as the maximum deviation
-($||E_1||$).  These numbers confirm the qualitative differences noted
-previously.
+\input{tests-wavepool}
 
-We also run the same simulation but with two levels of AMR (using a
-refinement factor of 2), triggered based on a normalized slope greater
-than 10\% in the density.  This refinement criterion results in only
-the refinement of strong gradients, and
-does not include the rarefaction fan at late times.  The results are
-shown in the bottom row of Figure~\ref{fig.sodshocktube}.  Using AMR,
-the results are much better for all three methods, with much sharper
-shocks and contact discontinuities and even a better representation of
-the rarefaction wave, which is only refined beyond the root grid at
-early times.  Although the results are improved for all methods, PPM
-still produces the best result, as is shown clearly by the computed
-error norms (displayed again in each of the individual panels).
+\input{tests-shockpool}
 
-\begin{figure}
-\begin{center}
-\includegraphics[width=\textwidth]{figures/SodShockTube.eps}
-\caption{The density distribution of the classic Sod Shock Tube for
-three different solvers (from left to right column) and with (bottom
-row) and without (top row) AMR.  In each case 100 zones are used on
-the root level and the results are shown at $t=0.25$.  All cells are
-plotted and color-coded by level (at the time shown, only small region
-surrounding the contact discontinuity and the shock are refined).  In
-each panel, we show the analytic solution as a solid line and the $E_1$
-and $E_\infty$ error norms in the upper right.}
-\label{fig.sodshocktube}
-\end{center}
-\end{figure}
+\input{tests-doublemach}
+
+\input{tests-sedov}
+
+\input{tests-pointsourcegravity}
+
+\input{tests-orbit}
+
+\input{tests-selfsiminfall.tex}
+
+\input{tests-zeldovichpancake}
+
+\input{tests-orszagtang}
+
+\input{tests-onezonecollapse}
+
+\input{tests-photoevaporation}
+
+\input{tests-fld-ifront}
+
+\input{tests-aniso-conduct}
 
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsubsection{Wave pool}
-\label{sec.tests.wavepool}
-
-In this one-dimensional test, we pass a short wavelength, linear wave through a static refined region.
-The full domain is from 0 to 1 and the refined region is set from 0.25 to 0.75 at all times.
-We modify the left boundary consistent with a single linear sound wave with density given by $\rho(x,t) = \rho_0 (1 + A \sin(kx - \omega t))$, where the amplitude is $A = 0.01$ and the wavelength $\lambda = k/2\pi = 0.1$, and similar expressions exist for the pressure and velocity.
-The initial, unperturbed density and pressure are set to unity and we adopt $\gamma = 1.4$.
-The (unrefined) root grid is covered with 100 cells, resulting in a wavelength for the linear wave of only 10 cells.
-This is, therefore, a challenging problem for hydro methods.
-We are particularly interested in any reflection or artifacts introduced by the wave entering and exiting the refined regions.
-
-Figure~\ref{fig.wavepool} shows the evolution of the wave at three different times ($t = 0.2, 0.3$ and 0.8) for three of our solvers (PPM, Zeus, and MUSCL).
-The leftmost column show the wave just before it enters the refined region so that we can gauge how the solver is operating in the absence of AMR, the center column shows the wave after it has fully entered the refined region, and the rightmost column shows the wave well after it has exited the refined region.
-
-Beginning with the PPM, we see that even before it has entered the refine region, the wave is slightly damped, which is not unexpected for such a short wavelength mode, even for PPM.
-The remaining panels demonstrate that the wave cleanly enters and exits the refined region.
-No significant reflection is seen, either on entry or exit, and the amount of damping is mild.
-
-For the Zeus solver, we see that even before it has entered the refined region, there are small oscillations excited behind the wave; although it is also worth noting that the wave itself is beautifully propagated without significant damping or phase errors.
-In this test, we use only our standard, low amount of quadratic artificial viscosity -- these waves can be damped by additional viscosity, but we do not add any in order to be sensitive to any artifacts at the grid boundary.
-The remaining panels show that the trailing oscillations continue, but do not generate any additional noise -- the end result is similar to the case without any refined region.
-
-Finally, the MUSCL solver shows also shows a very clean entry and exit from the refined region without any oscillations, although (for the PLM reconstruction used here) the wave is spread more than with the other methods.
-
-\begin{figure}
-\begin{center}
-\includegraphics[width=0.8\textwidth]{figures/WavePool.eps}
-\caption{This plot shows, for each column, three snap shots (at $t=0.2, 0.3$ and 0.8) of a linear wave as it propagates through the grid.
-The static refined region extends from $x = 0.25$ to 0.75 and is shown in grey in each panel.
-The individual cells are also color-coded by level: blue indicates the root grid, and red is for the refined region.
-Each row shows the result for a different solver (top is PPM, middle is Zeus, and the bottom is MUSCL).
-The solid line shows the analytic solution for a linear, undamped wave.
-Note that we focus each panel on a small region of the entire domain to better show the wave itself.}
-\label{fig.wavepool}
-\end{center}
-\end{figure}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsubsection{Shock pool}
-\label{sec.tests.shockpool}
-
-The next problem is similar to the last one, but examines how a shock wave with a Mach number of 2 passes into and out of a static refined region (again defined from $z=0.25$ to 0.75).
-The density and pressure in the domain is initially set equal to 1.0, with zero velocity, and then at $t=0$, the left boundary is set with the density, pressure and velocity appropriate for $M=2$ shock wave.
-
-In Figure~\ref{fig.shockpool}, we show the evolution of the shock wave at three times, again corresponding to just before entering the refine region (left column), after entering the refined region (center), and after exiting the refined region (right).
-We examine the same set of three solvers as in the previous test problem.
-
-Beginning with PPM (top row), we see that this method captures the shock in a small number of zones with only a small amount of oscillation.
-Open entering the refined region, the shock finds itself broader than the natural width of the scheme (since the cell spacing is decreased by a factor of 2), and so the shock front contracts, causing a slight entropy perturbation in the post-shock gas.
-
-For Zeus (middle row of Figure~\ref{fig.shockpool}, the shock is broadened because of artificial viscosity, and there are slightly more post-shock oscillations, although again quite mild.
-The impact of entering and exiting the refined region is somewhat larger than in PPM; however, the most noticeable difference is the incorrect position of the shock front, which is due to the fact that the scheme is not energy-conserving (see also the Sedov problem, below), and has essentially nothing to do with the passage through the refined region.
-
-Finally, the MUSCL scheme produces a shock which is intermediate in width between the two previous cases.
-This method is energy conserving and so correctly reproduces the shock speed.
-The oscillations are mild except for the cell immediately outside of the refined region upon exiting.
-
-\begin{figure}
-\begin{center}
-\includegraphics[width=0.8\textwidth]{figures/ShockPool}
-\caption{This plot shows, for each column, three snap shots (at $t=0.1, 0.15$ and 0.35) of a $M=2$ shock wave as it propagates through the grid with a static refined region extends from $x = 0.25$ to 0.75 (shown in grey in each panel).
-Each row shows the result for a different solver (PPM/Zeus/MUSCL from top to bottom).
-The solid line shows the analytic solution.
-Note that we focus each panel on a small region of the entire domain to better show the shock front.}
-\label{fig.shockpool}
-\end{center}
-\end{figure}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsubsection{Double mach reflection}
-\label{sec.tests.doublemach}
-
-The double mach reflection test is a classic two-dimensional test of
-hydrodynamic algorithms, and was originally described in
-\citet{1984JCoPh..54..115W} \citep[and more recently
-in][]{2008ApJS..178..137S}.  In this problem, shown in
-Figure~\ref{fig.doublemach}, a shock is injected at an angle to a
-reflecting surface (the -y boundary), and a jet appears along the
-reflecting surface.  The ideal solution is self-similar, and the
-appearance of this solution is highly sensitive to numerical
-diffusion.  If numerical noise is present, a Kelvin-Helmholz
-instability develops along this jet and breaks the self-similarity.
-
-In the test problem shown in Figure~\ref{fig.doublemach}, a 2D
-simulation with $960 \times 240$ cells was formed with a domain of $x
-= [0, 4]$ and $y = [0, 1]$.  We use an ideal gas equation of state of
-$\gamma = 1.4$, a pre-shock density of 1.4, and a pre-shock specific
-internal energy of $2.5/1.4$ (all in arbitrary units).  A Mach 10
-shock is initialized with a shock normal that is 30$^\circ$ from the
-x-axis and an initial position on the lower boundary of x$ = 1/6$.
-The lower y boundary and right x boundary are reflecting; the left x
-boundary is inflowing, and the upper y boundary has a time-dependent
-boundary condition that allows the shock to propagate into the domain
-as if it extends to infinity.  The simulation starts at t$ = 0$ and
-runs until t$ = 0.205$ (arbitrary units), at which point the rightmost
-extent of the shock should be at roughly x$ = 3$.  In this run, we use
-the direct Eulerian implementation of the piecewise parabolic hydro
-method, with the diffusion, flattening, and shock steepening all
-turned on.
-
-It is instructive to compare Figure~\ref{fig.doublemach} to Figure 9
-in~\citet{1984JCoPh..54..115W}.  By the end of the simulation, a dense
-jet is apparent at the leading edge of the shock, propagating along
-the x-axis.  The shape of this jet is sensitive to numerical
-diffusion, and our figure compares favorably to the higher-order
-images from Woodward \& Colella.
-
-\begin{figure}
-\begin{center}
-\includegraphics[width=0.8\textwidth]{figures/DoubleMachTest.eps}
-\caption{Density field at the final time in the Double Mach test.  A
-Mach 10 shock is injected into the domain with a shock normal that is
-30$^\circ$ from the x-axis with a time-dependent +y boundary condition
-that mimics a shock of infinite length.  The solution is self-similar,
-with the jet and whorls at $x \simeq 2.5-3.0$ in this figure being very sensitive
-to numerical diffusion.  Our results compare favorably to the
-higher-order images from \citet{1984JCoPh..54..115W}.}
-\label{fig.doublemach}
-\end{center}
-\end{figure}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsubsection{Sedov Explosion}
-\label{sec.tests.sedov}
-
-\begin{figure}
-\begin{center}
-\includegraphics[width=0.4\textwidth]{figures/sedov-ppm-slice.eps}
-\includegraphics[width=0.4\textwidth]{figures/sedov-zeus-slice.eps}
-\caption{Density slice from the Sedov Blast test at $t = 0.07$. Left:
-results using the Piecewise Parabolic Method hydro scheme.  Right: results using the \zeus\
-hydro method. Notably, the \zeus\ shock front has progressed less far
-than in the PPM run. This is due to energy loss when conserving only
-internal, and not total, energy.}
-\label{fig.sedov1}
-\end{center}
-\end{figure}
-
-\begin{figure}
-\begin{center}
-\includegraphics[width=\textwidth]{figures/sedov-profiles.eps}
-\caption{Radially averaged profiles for the Sedov Blast test at $t =
-0.07$. Clockwise from top-left shows density, velocity, internal
-energy and pressure.  The black solid line shows the analytic
-solution.  The blue dashed line shows the simulation using the PPM
-method, and the red dot-dashed line using \zeus.  The \zeus\ result
-substantially lags the true result due to total energy not being
-explicitly converged.}
-\label{fig.sedov2}
-\end{center}
-\end{figure}
-
-The Sedov Blast Test \citep{Sedov1959} models an intense explosion,
-initiated by depositing thermal energy into a homogenous distribution
-of gas. The result is a strong spherical shock wave centered on the
-point of energy injection.  This problem is a popular test of
-astrophysical numerical codes for three reasons: First, it is
-particularly appropriate to astrophysics since it represents the
-situation when a supernovae explosion occurs. Second, it has an
-exact analytical solution whereby the shock front's radial position is
-given by:
-
-\begin{equation} r(t) =
-\left(\frac{E_0}{\alpha\rho_0}\right)^{1/5}t^{2/5}
-\end{equation}
-
-\noindent where $E_0$ is the initial energy injection, $\rho_0$ is the
-background density and $\alpha = 1.0$ for cylindrical symmetry and an
-ideal gas with $\gamma = 1.4$. For the full derivation see
-\citet{Sedov1959}. This solution makes it possible to assess how well
-the code performs.  Third, the spherical shock is a challenging
-problem for numerical codes (both grid and particle) since the shock
-wave increases in size as the simulation progresses and its
-symmetrical nature highlights any directional preferences that grid
-codes can succumb to. The test presented here is the two-dimensional
-version that is included in the \enzo\ distribution. The
-three-dimensional results from this test, both for \enzo\ and three
-other leading astrophysics codes, can be founds in \citet{Tasker2008}.
-
-In the initial state, the box contains a homogenous distribution of
-gas at a density of 1 (note that, in the absense of gravity, this
-problem is scale-free and thus without units). Thermal energy is
-deposited into a single cell at the center of the box with $E_0 =
-10.0$. The problem is completed in two dimensions with reflecting
-boundary conditions (the default for \enzo) and uses a box size of side
-$1$. For this problem, we selected a top grid of $100 \times 100$
-cells and a maximum of four levels of refinement, placed based on
-shock location and the slope in density and total energy. The
-exception to this scheme is in the initial conditions, where grids are
-placed directly around the injection point. The results were assessed
-at $t = 0.07$, which corresponding to a time just before the shock
-reaches the box edge (see Figure~\ref{fig.sedov1}).
-
-Figure~\ref{fig.sedov2} shows the radial profiles for the simulation
-run with the PPM hydro-solver (blue dashed line) and the \zeus\
-hydro-solver (red dash-dot-dot line) together with the analytical
-solution (black solid line). Clockwise from the top-left are density,
-velocity, internal energy and pressure. PPM matches the analytical
-solution extremely well for all quantities. However, the shock front
-in the \zeus\ simulation lags behind the analytical position. This can
-also be seen in the slices shown in Figure~\ref{fig.sedov1}. The cause
-of this discrepancy is that \zeus\ shows a substantial energy loss
-during the first few timesteps and produces a diamond-shaped, rather
-than spherical, shockfront during this time. After this, the code
-correctly conserves energy but this intial energy loss remains clearly
-visible in the position of the shock at $t = 0.07$. This problem was
-addressed directly by \citet{Clarke2010}, who attributed the source of
-the issue to this version of \zeus\ solving the internal, rather than
-total, energy equation. In situations with strong energy gradients,
-this choice caused an energy loss and the artificial viscosity
-produces the direction-dependent shockfront shape. In their paper,
-\citet{Clarke2010} presents results from an alternative version of
-\zeus\ that conserves total energy. This problem is less marked for
-smaller energy gradients and it should be noted that the \zeus\ hydro
-algorithm's stability and speed make it a highly competitive choice,
-despite the disagreements in this test.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsubsection{Point source gravity test}
-\label{sec.test.gravitypointsource}
-\red{(Greg)}
-This is the TestGravity (problem 23) test problem.  It tests gravity around a point source, using fixed AMR.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsubsection{Orbit Test}
-\label{sec.test.testorbit}
-
-\begin{figure}
-\begin{center}
-\includegraphics[width=0.45\textwidth]{figures/TestOrbit_xy.eps}
-\includegraphics[width=0.45\textwidth]{figures/TestOrbit_TotalEnergy.eps}
-\caption{Test particle behavior in the Orbit problem, which is used to
-test the gravity solver and particle integration.  An effectively
-massless test particle is set in circular orbit around a massive
-central particle, with the gravitational potential calculated on a 3D,
-$32^3$ grid, and evolved forward in time for 200 orbits.  Left panel:
-path of the test particle in the orbital plane plotted for the
-duration of the simulation.  Right panel: total specific energy
-(kinetic plus potential) for the test particle for the duration of the
-simulation.  The particle's mean energy (integrated over the length of
-the simulation) is represented by a horizontal dashed line, and +/-
-one standard deviation by horizontal dotted lines.  All quantities are
-in arbitrary units.}
-\label{fig.orbittest}
-\end{center}
-\end{figure}
-
-The ``orbit test'' simulates the orbit of an effectively massless
-particle around a massive primary particle, and is used to test the
-accuracy of both the gravity solver and particle integration.  In this
-problem (shown in Figure~\ref{fig.orbittest}), a 3D box with unit size
-and $32^3$ grid is created, and a particle with mass 1 is placed at
-the center of the box (both box length and particle mass are in
-arbitrary units).  A second ``test'' particle, with mass of $10^{-6}$,
-is placed at a distance of 0.3 L$_{box}$ from the primary particle in
-the x-y plane (at $z=0.5$~L$_{box}$).  Both particles are given
-velocities such that they will orbit their common centers of mass in
-the x-y plane.  The gravitational potential is calculated using a fast
-Fourier Transform on the $32^3$ grid, and the simulation is integrated
-forward in time for 200 particle orbits, writing out a dataset
-approximately once per orbit (with the particle positions being
-written out several hundred times per orbit).
-
-The left panel of Figure~\ref{fig.orbittest} shows the path of the
-test particle in the x-y plane (at z=$0.5$), plotted over the duration
-of the simulation (approximately 200 orbits).  The test particle
-maintains a path that has effectively unnoticeable deviation from
-circularity, and with no discernable drift in either net radius or
-orbit center over time.  The right panel displays the total specific
-energy of the test particle (defined as $\frac{1}{2} {\bf v}^2 +
-\phi$, where $\phi$ is the gravitational potential at the position of
-the test particle) as a function of time, plotted once per orbit.  The
-mean total specific energy is $-1.76797$ in arbitrary units, with a
-standard deviation of $0.004885$ ($0.276\%$ of the mean value), and
-with a maximum deviation from the mean of $0.546\%$ of the mean value.
-Given that the potential is calculated on $32^3$ cells, the two
-particles are separated by between 10 and 14 cells depending on their
-relative positioning in the grid, and the particles are integrated
-using a second-order leapfrog method, this level of accuracy is
-roughly what one might expect.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsubsection{Self-Similar infall test}
-\label{sec.tests.infall}
-\red{(Greg)}
-Problem type 24.  This is a test based on Bertschinger's 1985 3D self-similar infall
-solution, and tests gravity + hydro.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsubsection{Zel`dovich Pancake}
-\label{sec.tests.pancake}
-
-The Zel`dovich pancake \citep{1970A&A.....5...84Z} is particularly
-relevant to cosmological simulations, because it includes many
-features that are critical to structure formation: hydrodynamics,
-expansion, and self gravity.  This problem represents the formation of an ideal,
-isolated caustic, and is thus a useful proxy for much more
-complicated structures in full 3-dimensional cosmological simulations
-(such as the collapse of gas onto a cosmological halo or filament).
-
-The initial conditions are simple, and we follow the prescription of
-\citet{Anninos94}.  Assuming a geometrically flat cosmology, the density
-perturbation is given by
-
-\begin{equation}
-\rho(x_l) = \rho_0 \big[ 1 - \frac{1+z_c}{1+z} \cos(k x_l) \big]
-\end{equation}
-
-with the internal energy of the gas set so that the entropy of the gas
-maintains a constant value throughout.  The velocity perturbation is given by
-
-\begin{equation}
-v(x_l) = -H_0 \frac{1 + z_c}{(1+z)^{1/2}} \frac{\sin(k x_l)}{k}
-\end{equation}
-
-In the equations above, $\rho_0$ is the background density, z$_c$ is a
-free parameter and is the redshift where the sheet forms a caustic
-(i.e., where it `pancakes'), z is the redshift of initialization, x$_l$ is the
-Lagrangian mass coordinate, $k = 2 \pi / \lambda$ (where $\lambda$ is
-the perturbation wavelength), and H$_0$ is the $z = 0$ value of the
-Hubble constant.  Note that this solution is expressed in terms of
-Lagrangian positions, so one needs to convert this into the Eulerian
-coordinates, x$_e$, that are more useful to \enzo:
-
-\begin{equation}
-x_e = x_l - \frac{1 + z_c}{1 + z} \frac{\sin(k x_l)}{k}
-\end{equation}
-
-We note that the solution described above is exact up to the point of
-caustic formation.  In Figure~\ref{fig.pancake}, we show the results
-of a test of the adaptive mesh version of \enzo's Zel'dovich Pancake test.  A
-one-dimensional box of length $64$~Mpc/h is initialized at z$ = 20$ in
-an $\Omega_M = 1$ universe with h$ = 0.5$ and a background temperature
-of 100~K, $\rho_0 = \rho_c$.  The simulation is initialized with 64
-grid cells, refining by factors of four using criteria based on cell
-mass and the presence of shocks, for a maximum of 2 levels (i.e., an
-equivalent maximum resolution of 1024 grid cells).  The simulation was
-evolved to z$ = 0$ using the PPM hydro method, with the final output
-of the calculation being shown in Figure~\ref{fig.pancake}.
-
-Figure~\ref{fig.pancake} shows the key features of this test problem.
-The strong shocks and large density gradients are well-resolved, with
-density and velocity jumps being well-delineated and at the correct locations.  The key features of
-this test problem can be resolved with far fewer cells -- simulations
-including a mere 8 cells resolve the key features, as shown in
-Sections 3.3.4-3.3.5 of~\citet{1996PhDT........80B} -- but we choose a
-higher resolution here for illustrative purposes.
-
-\begin{figure}
-\begin{center}
-\includegraphics[width=0.8\textwidth]{figures/AMRZeldovichPancake.eps}
-\caption{Zel`Dovich Pancake test shown at z$ = 0$, initialized in a
-$\Omega_m = 1$ universe at z$ = 20$ on a one-dimensional grid having
-64 cells, and further refined by factors of four based on cell mass
-and the presence of shocks for up to two additional levels of mesh,
-having a maximum effective resolution of 1024 grid cells. The top,
-middle, and bottom rows show density, temperature, and velocity of the
-gas, respectively; all are a function of position in units of the box
-size.  The central region ($x \simeq 0.45-0.55$) has bene adaptively
-refined, as can be seen by the locations of grid points.  Shocks and
-the central density peak are clearly resolved, with well-delineated
-jumps at the appropriate locations.}
-\label{fig.pancake}
-\end{center}
-\end{figure}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsubsection{MHD: Orszag-Tang Vortex}
-\label{sec.tests.mhd}
-Figure~\ref{fig.orszag} shows the Orszag-Tang vortex problem, a
-standard magnetohydrodynamic test problem \citep{Orszag79}.  The left panel shows
-the result using \enzo's constrained transport MHD method, while the right panel shows the result using
-\enzo's implementation of Dedner MHD.
-The Orszag-Tang vortex test shows that significant small scale structure can be generated in MHD
-from large scale initial perturbations, and is used to compare the
-effective resolution of different MHD schemes.  The test begins with uniform
-density, $\rho_0=25/36 \pi$ and pressure, $P_0=5/12 \pi$ (as with
-other tests in this section, in the absence of gravity or chemistry we use
-arbitrary units).  There is a
-single rotational mode in the velocity, and two in the magnetic field:
-${\bf v}_0 $ = (-sin(2$\pi$ y) $ \hat{x}$ , sin(2$\pi$ x) $\hat{y}$),
-${\bf B}_0$ = (-sin(2 $\pi$ y) $ \hat{x},$ sin( 4 $\pi$ x )$\hat{y}$).
-The simulation is evolved to $t=0.48$ (in arbitrary units).  One can see that the structures
-are accurately represented as compared to, for instance,
-\citet{Toth00}, and that the resolution of shocks is comparable in
-both methods.
-
-\begin{figure}
-\begin{center}
-\includegraphics[width=0.4\textwidth]{figures/MHDCT_OrszagTang_Density.eps}
-\includegraphics[width=0.4\textwidth]{figures/MHDDedner_OrszagTang_Density.eps}
-\caption{Density field from the Orszag-Tang vortex test, at $t=0.48$.
-Left: solution using constrained transport MHD.  Right: solution using
-Dedner MHD. The initial conditions are uniform density, with a single
-rotating velocity structure and two circular magnetic structures.
-These initial conditions generate significant small-scale structure in
-both the CT and Dedner schemes, which have approximately equal
-effective resolution.}
-\label{fig.orszag}
-\end{center}
-\end{figure}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsubsection{One-zone collapse test}
-\label{sec.tests.1-zone}
-
-The one-zone collapse test simulates the collapse of a
-self-gravitating gas cloud using a semi-analytic model for the
-evolution of gas density and adiabatic heat input as a function of
-time.  It is designed to test the chemistry and cooling modules over a
-wide range in densities and over physically motivated timescales.
-Because this test disables the hydrodynamic and gravity solvers and
-uses a simple model for the density evolution, it is far faster than
-running a true collapse simulation.  The density evolution is based on
-the self-similar Larson-Penston \citep{1969MNRAS.145..271L,
-1969MNRAS.144..425P} solution for isothermal collapse with a
-modification to account for the efficiency with which the heat
-introduced by compression can be radiated away
-\citep{1983ApJ...265.1047Y}.  Our implementation, described briefly
-here, follows the work of \citet{2005ApJ...626..627O}, and we direct
-the interested reader to this paper for further details.  The
-evolution of the gas density, $\rho$, is given by
-
-\begin{equation}
-\frac{d\rho}{dt} = \frac{\rho}{t_{col}},
-\end{equation}
-
-where the collapse timescale, $t_{col}$, is
-
-\begin{equation} \label{eqn.tcol}
-t_{col} = \frac{t_{dyn}}{\sqrt{1 - f}},
-\end{equation}
-In this equation, $t_{dyn}$ is the dynamical time and is expressed as
-\begin{equation}
-t_{dyn} = \sqrt{\frac{3 \pi}{32 G \rho}}.
-\end{equation}
-
-with G being the gravitational constant.  The collapse timescale is
-altered from the dynamical time by a factor $f$ in Equation
-\ref{eqn.tcol}, which is an approximation of the ratio of the gas
-pressure to the force of gravity.  The value of $f$ depends on the
-effective adiabatic index, $\gamma_{ef} \equiv (\partial ln\ p
-/ \partial ln\ \rho)$, which we calculate by linearly extrapolating
-from derivative values for the two previous timesteps.  For the value
-of $f$ in this test problem, we use the piecewise function of
-\citet{2005ApJ...626..627O}, given by
-
-\begin{equation}
-f = \left\{
-  \begin{array}{ll}
-  0, & \gamma_{ef} < 0.83,\\
-  0.6 + 2.5 (\gamma_{ef} - 1) - 6.0 (\gamma_{ef} - 1)^{2}, & 0.83 <
-  \gamma_{ef} < 1,\\
-  1.0 + 0.2 (\gamma_{ef} - 4/3) - 2.9 (\gamma_{ef} - 4/3)^{2}, & \gamma_{ef} > 1.
-\end{array} \right.
-\end{equation}
-
-The specific energy evolves as
-
-\begin{equation}
-\frac{de}{dt} = -p \frac{d}{dt} \frac{1}{\rho} - \Lambda,
-\end{equation}
-
-where $\Lambda$ is the cooling rate in units of erg s$^{-1}$ g$^{-1}$
-and energy, temperature, density, and pressure are related by the
-ideal gas law.  Figure \ref{fig.onezone} shows an example of the
-one-zone collapse test performed with an initial number density of 1
-hydrogen atom per cm$^{-3}$ and an initial temperature of 100 K using
-the 12 species chemistry network with H, D, and He species and metal
-cooling rates calculated with the \texttt{Cloudy} code.  The effects
-of metal cooling can be clearly seen; as the metallicity increases
-from zero to $10^{-2}$~Z$_\odot$, the gas deviates from the primoridal
-result (black line) both more rapidly and more significantly.  Our
-primordial results compare very well to those shown in
-\citet{2005ApJ...626..627O}; however, we use a different cooling
-method than Omukai, so the lines describing the evolution of the
-metal-enriched gas are not directly comparable.
-
-\begin{figure}
-  \begin{center}
-    \includegraphics[width=1.0\textwidth]{figures/OneZoneCollapseTest.eps}
-  \end{center}
-  \caption{Evolution of temperature versus number density for a
-one-zone collapse test for gas at metallicities from 0 to 10$^{-2}
-Z_{\odot}$.  This test is based on the results of
-\citet{2005ApJ...626..627O}, and approximates the collapse of a
-self-gravitating, cooling gas cloud.  This test problem uses \enzo's
-primordial chemistry network with tabulated metal cooling rates
-calculated with the \texttt{Cloudy} code, and compares favorably to
-Omukai's results.}
-  \label{fig.onezone}
-\end{figure}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsubsection{Photo-evaporation of a dense clump}
-\label{sec.tests.raytracing}
-
-The photo-evaporation of dense clumps of gas is prevalent in radiation
-hydrodynamics simulations, and this test problem examines the
-ionization front propagation into the dense clump, shadowing effects
-behind the clump, and the hydrodynamic response on the clump from
-photo-heating, all using \enzo's Moray ray-tracing module.  The
-problem setup is the same as Test 7 in the Cosmological Radiative
-Transfer Comparison Project \citep{IlievEtAl2009} and
-\citet{Wise11_Moray}.  The simulation domain is 6.6~kpc on a side with
-an ambient medium of pure neutral hydrogen $n_{\rm H} = 2 \times
-10^{-4}\; \cubecm$ and $T = 8000 \unit{K}$.  We place a spherical
-overdensity in hydrostatic equilibrium with the ambient medium.  It
-has a radius $r = 0.8 \unit{kpc}$, hydrogen density $n_{\rm H} = 0.04
-\cubecm$, and temperature $T = 40$ K, and it is centered at $(x,y,z) =
-(5, 3.3, 3.3) \unit{kpc}$.  In \citet{IlievEtAl2009}, all of the codes
-used a fixed $128^3$ grid to ease the comparison, but in this test to
-demonstrate a higher resolution AMR solution, we employ a $128^3$ grid
-with two additional levels of refinement with cells with a baryon mass
-(method 2 in Section~\ref{sec:refinement_criteria}) greater than 1.5
-being flagged for refinement.  This test is run for 15 Myr.
-
-This cloud is subject to radiation from a point source at the center
-of the $x=0$ boundary with an ionizing photon luminosity
-$\dot{N}_\gamma = 3 \times 10^{51}$ photons s$^{-1}$, corresponding to
-a flux $F_0 = 10^6 \unit{photons s}^{-1} \unit{cm}^{-2}$ at the clump
-surface closest to the radiation source.  The radiation source has a
-spectrum of a $T = 10^5 \unit{K}$ blackbody, and we use four energy
-groups with the following mean energies and relative luminosities:
-$E_i = (17.98, 31.15, 49.09, 76.98) \unit{eV}, L_i/L = (0.23, 0.36,
-0.24, 0.06)$ that are optimized to reduce errors in the solutions with
-a full spectrum and energy discretization \citep{Mirocha12}.
-\citep[Note that this choice of energy groups is different
-from][]{Wise11_Moray}.  We use a minimum angular resolution of 10 rays
-per cell and a constant radiative transfer timestep of 25 kyr.  Figure
-\ref{fig:shadowing} depicts the clump after at $t = 15$ Myr with the
-outer layers expanding after being photo-heated.  It also shows the
-sharp shadowing effects of the dense clump in the neutral fraction
-plot that is representative of ray tracing techniques.
-
-\begin{figure}
-  \centering
-  \includegraphics[width=1.0\textwidth]{figures/code-test-shadowing.eps}
-  \caption{Photo-evaporation of a dense clump using the Moray
-ray-tracing module.  Clockwise from the upper left corner: slices
-through the clump center of the density, temperature, neutral hydrogen
-fraction, and pressure at $t=15$~Myr after the initialization of the
-simulation.  A point source of radiation is at the center of the -x
-boundary and illuminates the clump with a constant luminosity, casting
-a clear shadow behind the clump, as seen in the neutral hydrogen
-fraction.}
-  \label{fig:shadowing}
-\end{figure}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\input{code-tests-fld}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-\subsubsection{Anisotropic Thermal conduction}
-\label{sec.tests.conduct}
-
-In a plasma, conduction takes place primarily due to electron motion.
-In astrophysical environments such as the intergalactic and
-intracluster media, with typical magnetic strengths of $\sim 1$~$\mu$G
-and temperatures in the millions of Kelvin, the electron gyroradius is
-very small compared to the scales of interest in a typical
-cosmological simulation.  As a result, electrons can only move
-\textit{along} magnetic field lines, not perpendicular to them, which
-can result in very interesting magnetohydrodynamical instabilities
-\citep[e.g.,][]{2008ApJ...677L...9P,2008ApJ...688..905P}.  The correct
-modeling of this behavior can be quite challenging, and is described
-in Section~\ref{sec.num.conductions}.  Conduction in \enzo\ is
-calculated in both an operator-split and directionally-split manner,
-and the addition of heat transport solely along magnetic field lines
-requires computation of cross-terms in the temperature derivative at
-cell faces, which can add spurious oscillations in the temperature
-field in regions where the temperature gradient is strong in more than
-one spatial direction unless an appropriate flux-limiter \citep[such
-as that of][]{1977JCoPh..23..263V} is chosen.
-
-Figure~\ref{fig.conduct} shows a test that demonstrates the correct
-behavior of anisotropic thermal conduction in \enzo.  We initialize a
-two-dimensional, $256 \times 256$ cell simulation having a physical
-scale of 1 kpc on a side, a uniform density of 1 proton/cc, and a
-background temperature of $10^6$ Kelvin.  Magnetic fields with a
-strength of B$_0 = 1$~$\mu$G are initialized such that the field lines
-form circles around the center of the simulation volume, such that
-B$_x = -B_0\sin(\theta)$ and B$_y = B_0\cos(\theta)$, where $\theta$
-is the angle measured from the $+x$ direction in a counterclockwise
-manner.  A Gaussian temperature pulse is injected at $(0.75, 0.5)$ (in
-units of the box size), with a peak temperature of $10^8$ K and a FWHM
-of $1/64$ of the box size.  This initial setup is shown in the left
-panel of Figure~\ref{fig.conduct}.  The simulation is then allowed to
-evolve with \textit{only} anisotropic conduction turned on (e.g., no
-hydrodynamics, radiative cooling, or cosmological expansion), and with
-a Spitzer fraction of f$_{sp} = 1$.
-
-The right panel of Figure~\ref{fig.conduct} shows the state of the
-simulation after 300 Myr.  Heat has clearly been transported only along
-field lines -- there has been no diffusion perpendicular to the
-magnetic field setup, which is critical for many studies involving
-anisotropic thermal conduction.  No oscillations are seen in the
-temperature field in regions where the fields are not aligned with the
-grid, suggesting that the flux-limiter is operating as expected.
-
-\begin{figure}
-\begin{center}
-\includegraphics[width=0.42\textwidth]{figures/aniso_conduction_initial_output.eps}
-\includegraphics[width=0.4\textwidth]{figures/aniso_conduction_final_output.eps}
-\caption{Two-dimensional anisotropic conduction test in a uniform,
-constant temperature background with circular magnetic fields centered
-on (0.5, 0.5) in a $256 \times 256$ grid.  The background medium has a
-density of 1 proton/cc and a temperature of $10^6$~K.  At t$ = 0$
-(left panel), a Gaussian heat pulse is injected at (0.75, 0.5) with a
-FWHM of $1/64$ (with all numbers given in units of the box size) and a
-peak temperature of $10^8$~K, and allowed to evolve without
-hydrodynamical motion (i.e., static gas) and no radiative cooling for
-300 Myr.  At t$ = 300$~Myr (right panel), heat has been transported
-along magnetic field lines with no significant diffusion perpendicular
-to field lines. Furthermore, there are no detectable oscillations in
-the temperature in regions where the magnetic field is not parallel
-with the grids.}
-\label{fig.conduct}
-\end{center}
-\end{figure}
 

### 2013-05-21 tests-aniso-conduct.tex
Splitting up test problem section into subsections.

\subsubsection{Anisotropic Thermal conduction}
\label{sec.tests.conduct}

In a plasma, conduction takes place primarily due to electron motion.
In astrophysical environments such as the intergalactic and
intracluster media, with typical magnetic strengths of $\sim 1$~$\mu$G
and temperatures in the millions of Kelvin, the electron gyroradius is
very small compared to the scales of interest in a typical
cosmological simulation.  As a result, electrons can only move
\textit{along} magnetic field lines, not perpendicular to them, which
can result in very interesting magnetohydrodynamical instabilities
\citep[e.g.,][]{2008ApJ...677L...9P,2008ApJ...688..905P}.  The correct
modeling of this behavior can be quite challenging, and is described
in Section~\ref{sec.num.conductions}.  Conduction in \enzo\ is
calculated in both an operator-split and directionally-split manner,
and the addition of heat transport solely along magnetic field lines
requires computation of cross-terms in the temperature derivative at
cell faces, which can add spurious oscillations in the temperature
field in regions where the temperature gradient is strong in more than
one spatial direction unless an appropriate flux-limiter \citep[such
as that of][]{1977JCoPh..23..263V} is chosen.

Figure~\ref{fig.conduct} shows a test that demonstrates the correct
behavior of anisotropic thermal conduction in \enzo.  We initialize a
two-dimensional, $256 \times 256$ cell simulation having a physical
scale of 1 kpc on a side, a uniform density of 1 proton/cc, and a
background temperature of $10^6$ Kelvin.  Magnetic fields with a
strength of B$_0 = 1$~$\mu$G are initialized such that the field lines
form circles around the center of the simulation volume, such that
B$_x = -B_0\sin(\theta)$ and B$_y = B_0\cos(\theta)$, where $\theta$
is the angle measured from the $+x$ direction in a counterclockwise
manner.  A Gaussian temperature pulse is injected at $(0.75, 0.5)$ (in
units of the box size), with a peak temperature of $10^8$ K and a FWHM
of $1/64$ of the box size.  This initial setup is shown in the left
panel of Figure~\ref{fig.conduct}.  The simulation is then allowed to
evolve with \textit{only} anisotropic conduction turned on (e.g., no
hydrodynamics, radiative cooling, or cosmological expansion), and with
a Spitzer fraction of f$_{sp} = 1$.

The right panel of Figure~\ref{fig.conduct} shows the state of the
simulation after 300 Myr.  Heat has clearly been transported only along
field lines -- there has been no diffusion perpendicular to the
magnetic field setup, which is critical for many studies involving
anisotropic thermal conduction.  No oscillations are seen in the
temperature field in regions where the fields are not aligned with the
grid, suggesting that the flux-limiter is operating as expected.

\begin{figure}
\begin{center}
\includegraphics[width=0.42\textwidth]{figures/aniso_conduction_initial_output.eps}
\includegraphics[width=0.4\textwidth]{figures/aniso_conduction_final_output.eps}
\caption{Two-dimensional anisotropic conduction test in a uniform,
constant temperature background with circular magnetic fields centered
on (0.5, 0.5) in a $256 \times 256$ grid.  The background medium has a
density of 1 proton/cc and a temperature of $10^6$~K.  At t$ = 0$
(left panel), a Gaussian heat pulse is injected at (0.75, 0.5) with a
FWHM of $1/64$ (with all numbers given in units of the box size) and a
peak temperature of $10^8$~K, and allowed to evolve without
hydrodynamical motion (i.e., static gas) and no radiative cooling for
300 Myr.  At t$ = 300$~Myr (right panel), heat has been transported
along magnetic field lines with no significant diffusion perpendicular
to field lines. Furthermore, there are no detectable oscillations in
the temperature in regions where the magnetic field is not parallel
with the grids.}
\label{fig.conduct}
\end{center}
\end{figure}

### 2013-05-21 tests-doublemach.tex
Splitting up test problem section into subsections.

\subsubsection{Double mach reflection}
\label{sec.tests.doublemach}

The double mach reflection test is a classic two-dimensional test of
hydrodynamic algorithms, and was originally described in
\citet{1984JCoPh..54..115W} \citep[and more recently
in][]{2008ApJS..178..137S}.  In this problem, shown in
Figure~\ref{fig.doublemach}, a shock is injected at an angle to a
reflecting surface (the -y boundary), and a jet appears along the
reflecting surface.  The ideal solution is self-similar, and the
appearance of this solution is highly sensitive to numerical
diffusion.  If numerical noise is present, a Kelvin-Helmholz
instability develops along this jet and breaks the self-similarity.

In the test problem shown in Figure~\ref{fig.doublemach}, a 2D
simulation with $960 \times 240$ cells was formed with a domain of $x
= [0, 4]$ and $y = [0, 1]$.  We use an ideal gas equation of state of
$\gamma = 1.4$, a pre-shock density of 1.4, and a pre-shock specific
internal energy of $2.5/1.4$ (all in arbitrary units).  A Mach 10
shock is initialized with a shock normal that is 30$^\circ$ from the
x-axis and an initial position on the lower boundary of x$ = 1/6$.
The lower y boundary and right x boundary are reflecting; the left x
boundary is inflowing, and the upper y boundary has a time-dependent
boundary condition that allows the shock to propagate into the domain
as if it extends to infinity.  The simulation starts at t$ = 0$ and
runs until t$ = 0.205$ (arbitrary units), at which point the rightmost
extent of the shock should be at roughly x$ = 3$.  In this run, we use
the direct Eulerian implementation of the piecewise parabolic hydro
method, with the diffusion, flattening, and shock steepening all
turned on.

It is instructive to compare Figure~\ref{fig.doublemach} to Figure 9
in~\citet{1984JCoPh..54..115W}.  By the end of the simulation, a dense
jet is apparent at the leading edge of the shock, propagating along
the x-axis.  The shape of this jet is sensitive to numerical
diffusion, and our figure compares favorably to the higher-order
images from Woodward \& Colella.

\begin{figure}
\begin{center}
\includegraphics[width=0.8\textwidth]{figures/DoubleMachTest.eps}
\caption{Density field at the final time in the Double Mach test.  A
Mach 10 shock is injected into the domain with a shock normal that is
30$^\circ$ from the x-axis with a time-dependent +y boundary condition
that mimics a shock of infinite length.  The solution is self-similar,
with the jet and whorls at $x \simeq 2.5-3.0$ in this figure being very sensitive
to numerical diffusion.  Our results compare favorably to the
higher-order images from \citet{1984JCoPh..54..115W}.}
\label{fig.doublemach}
\end{center}
\end{figure}

### 2013-05-21 tests-fld-ifront.tex
Splitting up test problem section into subsections.

--- 2013-03-19 code-tests-fld.tex
@@
-\subsubsection{Representative FLD test}
+\subsubsection{Cosmological I-front propagation}
 \label{sec.tests.fld}
 
-This test problem examines Enzo's FLD radiation transport and
-associated ionization chemistry solvers on a cosmological I-front
-test.  This corresponds to the test problem 4.6
-from \cite{ReynoldsHayesPaschosNorman2009}, run using the cosmological
+This test problem examines Enzo's flux-limited diffusion radiation
+transport and associated ionization chemistry solvers on a
+cosmological I-front test.  This corresponds to test problem 4.6 from
+\cite{ReynoldsHayesPaschosNorman2009}, run using the cosmological
 deceleration parameter $q_0 = 0.05$ and initial redshift $z_i=10$.
-Here, the physics of interest is the expansion of a \ion{H}{2} region in a
-uniform gas around a single monochromatic ionizing source (with
+Here, the physics of interest is the expansion of a \ion{H}{2} region
+in a uniform gas around a single monochromatic ionizing source (with
 frequency $h\nu = 13.6$ eV).  In this problem the I-front propagates
@@
 \]
-where $\dot{N}_{\gamma} = 5\cdot10^{48}$ photon sec$^{-1}$ is the
-stength of the ionizing source, $\alpha_{\rm B} = 2.52\cdot10^{-13}$
+where $\dot{N}_{\gamma} = 5 \times 10^{48}$ photon sec$^{-1}$ is the
+stength of the ionizing source, $\alpha_{\rm B} = 2.52 \times 10^{-13}$
 s$^{-1}$ is the \ion{H}{2} recombination rate, and $n_H(t)$ is the
@@
 of the time/redshift-dependent expansion coefficient $a(t)$) is then
+
 \begin{eqnarray*}
@@
 \end{eqnarray*}
+
 The simulation parameters are a box size $L_i\approx 27$ kpc, Hubble
-constant $H_0 = 1$, and cosmological constants $\Omega_m = 0.1$,
-$\Omega_A=0$ and $\Omega_b = 0.1$.  The initial values for the
-simulation are a radiation energy density $E_{r,i} = 10^{-35}$ erg
-cm$^{-3}$, temperature $T_i = 10^4$ K, density $\rho_{b,i} =
-2.35\cdot10^{-28}$ g cm$^{-3}$, and an ionized hydrogen fraction
-$n_{\rm HII}/n_{\rm H} = 0$.
+constant $H_0 = 1$ (in units of 100 km/s/Mpc), and cosmological
+constants $\Omega_m = 0.1$, $\Omega_A=0$ and $\Omega_b = 0.1$.  The
+initial values for the simulation are a radiation energy density
+$E_{r,i} = 10^{-35}$ erg cm$^{-3}$, temperature $T_i = 10^4$ K,
+density $\rho_{b,i} = 2.35 \times 10^{-28}$ g cm$^{-3}$, and an
+ionized hydrogen fraction $n_{\rm HII}/n_{\rm H} = 0$.
 
-In the Figure \ref{fig.fld}, we plot profiles of the \ion{H}{1} and
+In Figure \ref{fig.fld}, we plot profiles of the \ion{H}{1} and
 \ion{H}{2} fractions at the redshifts $z=\{6.24,\, 2.29,\, 1.02\}$, as
 well as the ratio of the I-front and Str{\" o}mgren radii throughout
-the simulation.  As seen in the profile figure, the I-front width
-starts out quite sharp but slowly spreads as the simulation proceeds
-due to the diffusion approximation to the radiative transfer.
-However, as seen in the I-front propagation history plot, the computed
-I-front location very accurately captures the exact value.
+the simulation.  As seen in the left panel (species fractions as a
+function of radius), the I-front is initially quite narrow, but slowly
+becomes wider as the simulation proceeds due to the diffusion
+approximation used by the radiative transfer scheme.  However, as seen
+in the I-front propagation history plot (right panel), the computed
+I-front location very accurately matches the exact value.
 
@@
 \includegraphics[width=0.45\textwidth]{figures/FLDhistory.eps}
-\caption{Ionized fraction profiles and I-front propgation history for
-the cosmological radiative transfer test problem in
-Section \ref{sec.tests.fld}.}
+\caption{Ionized fraction profiles and I-front location history for
+the propagation of an ionization front in an expanding universe, using
+\enzo's flux-limited diffusion radiation transport method.  Left
+panel: profiles of the \ion{H}{1} and \ion{H}{2} fractions (blue and
+red lines, respectively) at redshifts $z=\{6.24,\, 2.29,\, 1.02\}$
+(solid, dashed, dot-dashed lines).  Right panel: ratio of the I-front
+and Str{\" o}mgren radii as a function of scaled redshift in the test
+problem (blue solid line) and analytical solution (red dashed line).}
 \label{fig.fld}

### 2013-05-21 tests-onezonecollapse.tex
Splitting up test problem section into subsections.

\subsubsection{One-zone collapse test}
\label{sec.tests.1-zone}

The one-zone collapse test simulates the collapse of a
self-gravitating gas cloud using a semi-analytic model for the
evolution of gas density and adiabatic heat input as a function of
time.  It is designed to test the chemistry and cooling modules over a
wide range in densities and over physically motivated timescales.
Because this test disables the hydrodynamic and gravity solvers and
uses a simple model for the density evolution, it is far faster than
running a true collapse simulation.  The density evolution is based on
the self-similar Larson-Penston \citep{1969MNRAS.145..271L,
1969MNRAS.144..425P} solution for isothermal collapse with a
modification to account for the efficiency with which the heat
introduced by compression can be radiated away
\citep{1983ApJ...265.1047Y}.  Our implementation, described briefly
here, follows the work of \citet{2005ApJ...626..627O}, and we direct
the interested reader to this paper for further details.  The
evolution of the gas density, $\rho$, is given by

\begin{equation}
\frac{d\rho}{dt} = \frac{\rho}{t_{col}},
\end{equation}

where the collapse timescale, $t_{col}$, is

\begin{equation} \label{eqn.tcol}
t_{col} = \frac{t_{dyn}}{\sqrt{1 - f}},
\end{equation}
In this equation, $t_{dyn}$ is the dynamical time and is expressed as
\begin{equation}
t_{dyn} = \sqrt{\frac{3 \pi}{32 G \rho}}.
\end{equation}

with G being the gravitational constant.  The collapse timescale is
altered from the dynamical time by a factor $f$ in Equation
\ref{eqn.tcol}, which is an approximation of the ratio of the gas
pressure to the force of gravity.  The value of $f$ depends on the
effective adiabatic index, $\gamma_{ef} \equiv (\partial ln\ p
/ \partial ln\ \rho)$, which we calculate by linearly extrapolating
from derivative values for the two previous timesteps.  For the value
of $f$ in this test problem, we use the piecewise function of
\citet{2005ApJ...626..627O}, given by

\begin{equation}
f = \left\{
  \begin{array}{ll}
  0, & \gamma_{ef} < 0.83,\\
  0.6 + 2.5 (\gamma_{ef} - 1) - 6.0 (\gamma_{ef} - 1)^{2}, & 0.83 <
  \gamma_{ef} < 1,\\
  1.0 + 0.2 (\gamma_{ef} - 4/3) - 2.9 (\gamma_{ef} - 4/3)^{2}, & \gamma_{ef} > 1.
\end{array} \right.
\end{equation}

The specific energy evolves as

\begin{equation}
\frac{de}{dt} = -p \frac{d}{dt} \frac{1}{\rho} - \Lambda,
\end{equation}

where $\Lambda$ is the cooling rate in units of erg s$^{-1}$ g$^{-1}$
and energy, temperature, density, and pressure are related by the
ideal gas law.  Figure \ref{fig.onezone} shows an example of the
one-zone collapse test performed with an initial number density of 1
hydrogen atom per cm$^{-3}$ and an initial temperature of 100 K using
the 12 species chemistry network with H, D, and He species and metal
cooling rates calculated with the \texttt{Cloudy} code.  The effects
of metal cooling can be clearly seen; as the metallicity increases
from zero to $10^{-2}$~Z$_\odot$, the gas deviates from the primoridal
result (black line) both more rapidly and more significantly.  Our
primordial results compare very well to those shown in
\citet{2005ApJ...626..627O}; however, we use a different cooling
method than Omukai, so the lines describing the evolution of the
metal-enriched gas are not directly comparable.

\begin{figure}
  \begin{center}
    \includegraphics[width=1.0\textwidth]{figures/OneZoneCollapseTest.eps}
  \end{center}
  \caption{Evolution of temperature versus number density for a
one-zone collapse test for gas at metallicities from 0 to 10$^{-2}
Z_{\odot}$.  This test is based on the results of
\citet{2005ApJ...626..627O}, and approximates the collapse of a
self-gravitating, cooling gas cloud.  This test problem uses \enzo's
primordial chemistry network with tabulated metal cooling rates
calculated with the \texttt{Cloudy} code, and compares favorably to
Omukai's results.}
  \label{fig.onezone}
\end{figure}

### 2013-05-21 tests-orbit.tex
Splitting up test problem section into subsections.

\subsubsection{Orbit Test}
\label{sec.test.testorbit}

\begin{figure}
\begin{center}
\includegraphics[width=0.45\textwidth]{figures/TestOrbit_xy.eps}
\includegraphics[width=0.45\textwidth]{figures/TestOrbit_TotalEnergy.eps}
\caption{Test particle behavior in the Orbit problem, which is used to
test the gravity solver and particle integration.  An effectively
massless test particle is set in circular orbit around a massive
central particle, with the gravitational potential calculated on a 3D,
$32^3$ grid, and evolved forward in time for 200 orbits.  Left panel:
path of the test particle in the orbital plane plotted for the
duration of the simulation.  Right panel: total specific energy
(kinetic plus potential) for the test particle for the duration of the
simulation.  The particle's mean energy (integrated over the length of
the simulation) is represented by a horizontal dashed line, and +/-
one standard deviation by horizontal dotted lines.  All quantities are
in arbitrary units.}
\label{fig.orbittest}
\end{center}
\end{figure}

The ``orbit test'' simulates the orbit of an effectively massless
particle around a massive primary particle, and is used to test the
accuracy of both the gravity solver and particle integration.  In this
problem (shown in Figure~\ref{fig.orbittest}), a 3D box with unit size
and $32^3$ grid is created, and a particle with mass 1 is placed at
the center of the box (both box length and particle mass are in
arbitrary units).  A second ``test'' particle, with mass of $10^{-6}$,
is placed at a distance of 0.3 L$_{box}$ from the primary particle in
the x-y plane (at $z=0.5$~L$_{box}$).  Both particles are given
velocities such that they will orbit their common centers of mass in
the x-y plane.  The gravitational potential is calculated using a fast
Fourier Transform on the $32^3$ grid, and the simulation is integrated
forward in time for 200 particle orbits, writing out a dataset
approximately once per orbit (with the particle positions being
written out several hundred times per orbit).

The left panel of Figure~\ref{fig.orbittest} shows the path of the
test particle in the x-y plane (at z=$0.5$), plotted over the duration
of the simulation (approximately 200 orbits).  The test particle
maintains a path that has effectively unnoticeable deviation from
circularity, and with no discernable drift in either net radius or
orbit center over time.  The right panel displays the total specific
energy of the test particle (defined as $\frac{1}{2} {\bf v}^2 +
\phi$, where $\phi$ is the gravitational potential at the position of
the test particle) as a function of time, plotted once per orbit.  The
mean total specific energy is $-1.76797$ in arbitrary units, with a
standard deviation of $0.004885$ ($0.276\%$ of the mean value), and
with a maximum deviation from the mean of $0.546\%$ of the mean value.
Given that the potential is calculated on $32^3$ cells, the two
particles are separated by between 10 and 14 cells depending on their
relative positioning in the grid, and the particles are integrated
using a second-order leapfrog method, this level of accuracy is
roughly what one might expect.

### 2013-05-21 tests-orszagtang.tex
Splitting up test problem section into subsections.

\subsubsection{MHD: Orszag-Tang Vortex}
\label{sec.tests.mhd}
Figure~\ref{fig.orszag} shows the Orszag-Tang vortex problem, a
standard magnetohydrodynamic test problem \citep{Orszag79}.  The left panel shows
the result using \enzo's constrained transport MHD method, while the right panel shows the result using
\enzo's implementation of Dedner MHD.
The Orszag-Tang vortex test shows that significant small scale structure can be generated in MHD
from large scale initial perturbations, and is used to compare the
effective resolution of different MHD schemes.  The test begins with uniform
density, $\rho_0=25/36 \pi$ and pressure, $P_0=5/12 \pi$ (as with
other tests in this section, in the absence of gravity or chemistry we use
arbitrary units).  There is a
single rotational mode in the velocity, and two in the magnetic field:
${\bf v}_0 $ = (-sin(2$\pi$ y) $ \hat{x}$ , sin(2$\pi$ x) $\hat{y}$),
${\bf B}_0$ = (-sin(2 $\pi$ y) $ \hat{x},$ sin( 4 $\pi$ x )$\hat{y}$).
The simulation is evolved to $t=0.48$ (in arbitrary units).  One can see that the structures
are accurately represented as compared to, for instance,
\citet{Toth00}, and that the resolution of shocks is comparable in
both methods.

\begin{figure}
\begin{center}
\includegraphics[width=0.4\textwidth]{figures/MHDCT_OrszagTang_Density.eps}
\includegraphics[width=0.4\textwidth]{figures/MHDDedner_OrszagTang_Density.eps}
\caption{Density field from the Orszag-Tang vortex test, at $t=0.48$.
Left: solution using constrained transport MHD.  Right: solution using
Dedner MHD. The initial conditions are uniform density, with a single
rotating velocity structure and two circular magnetic structures.
These initial conditions generate significant small-scale structure in
both the CT and Dedner schemes, which have approximately equal
effective resolution.}
\label{fig.orszag}
\end{center}
\end{figure}

### 2013-05-21 tests-photoevaporation.tex
Splitting up test problem section into subsections.

\subsubsection{Photo-evaporation of a dense clump}
\label{sec.tests.raytracing}

The photo-evaporation of dense clumps of gas is prevalent in radiation
hydrodynamics simulations, and this test problem examines the
ionization front propagation into the dense clump, shadowing effects
behind the clump, and the hydrodynamic response on the clump from
photo-heating, all using \enzo's Moray ray-tracing module.  The
problem setup is the same as Test 7 in the Cosmological Radiative
Transfer Comparison Project \citep{IlievEtAl2009} and
\citet{Wise11_Moray}.  The simulation domain is 6.6~kpc on a side with
an ambient medium of pure neutral hydrogen $n_{\rm H} = 2 \times
10^{-4}\; \cubecm$ and $T = 8000 \unit{K}$.  We place a spherical
overdensity in hydrostatic equilibrium with the ambient medium.  It
has a radius $r = 0.8 \unit{kpc}$, hydrogen density $n_{\rm H} = 0.04
\cubecm$, and temperature $T = 40$ K, and it is centered at $(x,y,z) =
(5, 3.3, 3.3) \unit{kpc}$.  In \citet{IlievEtAl2009}, all of the codes
used a fixed $128^3$ grid to ease the comparison, but in this test to
demonstrate a higher resolution AMR solution, we employ a $128^3$ grid
with two additional levels of refinement with cells with a baryon mass
(method 2 in Section~\ref{sec:refinement_criteria}) greater than 1.5
being flagged for refinement.  This test is run for 15 Myr.

This cloud is subject to radiation from a point source at the center
of the $x=0$ boundary with an ionizing photon luminosity
$\dot{N}_\gamma = 3 \times 10^{51}$ photons s$^{-1}$, corresponding to
a flux $F_0 = 10^6 \unit{photons s}^{-1} \unit{cm}^{-2}$ at the clump
surface closest to the radiation source.  The radiation source has a
spectrum of a $T = 10^5 \unit{K}$ blackbody, and we use four energy
groups with the following mean energies and relative luminosities:
$E_i = (17.98, 31.15, 49.09, 76.98) \unit{eV}, L_i/L = (0.23, 0.36,
0.24, 0.06)$ that are optimized to reduce errors in the solutions with
a full spectrum and energy discretization \citep{Mirocha12}.
\citep[Note that this choice of energy groups is different
from][]{Wise11_Moray}.  We use a minimum angular resolution of 10 rays
per cell and a constant radiative transfer timestep of 25 kyr.  Figure
\ref{fig:shadowing} depicts the clump after at $t = 15$ Myr with the
outer layers expanding after being photo-heated.  It also shows the
sharp shadowing effects of the dense clump in the neutral fraction
plot that is representative of ray tracing techniques.

\begin{figure}
  \centering
  \includegraphics[width=1.0\textwidth]{figures/code-test-shadowing.eps}
  \caption{Photo-evaporation of a dense clump using the Moray
ray-tracing module.  Clockwise from the upper left corner: slices
through the clump center of the density, temperature, neutral hydrogen
fraction, and pressure at $t=15$~Myr after the initialization of the
simulation.  A point source of radiation is at the center of the -x
boundary and illuminates the clump with a constant luminosity, casting
a clear shadow behind the clump, as seen in the neutral hydrogen
fraction.}
  \label{fig:shadowing}
\end{figure}

### 2013-05-21 tests-pointsourcegravity.tex
Splitting up test problem section into subsections.

\subsubsection{Point source gravity test}
\label{sec.test.gravitypointsource}

This test is a simple test of the ability of the Poisson solver to correctly reproduce the acceleration around a point source without any dynamics.  We place a single particle in the center of the unit domain with root grid dimensions $32^3$ and add a static AMR grid with refinement factor two that covers the central $1/8^3$ of the domain (i.e. so the refined grid is only $8^3$).  We then place 5000 particles throughout the domain distributed uniformly in angle and with the logarithm of the radius (from the central point).  The measured accelerations are shown in Figure~\ref{fig.gravitytest}.  As is typical for Particle-Mesh based methods, the errors decrease at large distances and peak on roughly the grid scale.  The accelerations are relatively smooth across the top grid/fine grid boundary with the most noticeable impact being a drop in the amplitude of tangential errors on the fine grid.  The force law rolls over at about 1.6 fine cell widths (equivalent to 0.8 root grid widths in this figure), consistent with the CIC-deposition and interpolation used.

\begin{figure}
\begin{center}
\includegraphics[width=0.6\textwidth]{figures/GravityTest.eps}
\caption{The radial and tangential accelerations measured around a point source placed in a static AMR grid.  The solid line shows the analytic radial profile.}
\label{fig.gravitytest}
\end{center}
\end{figure}

### 2013-05-21 tests-sedov.tex
Splitting up test problem section into subsections.

\subsubsection{Sedov Explosion}
\label{sec.tests.sedov}

\begin{figure}
\begin{center}
\includegraphics[width=0.4\textwidth]{figures/sedov-ppm-slice.eps}
\includegraphics[width=0.4\textwidth]{figures/sedov-zeus-slice.eps}
\caption{Density slice from the Sedov Blast test at $t = 0.07$. Left:
results using the Piecewise Parabolic Method hydro scheme.  Right: results using the \zeus\
hydro method. Notably, the \zeus\ shock front has progressed less far
than in the PPM run. This is due to energy loss when conserving only
internal, and not total, energy.}
\label{fig.sedov1}
\end{center}
\end{figure}

\begin{figure}
\begin{center}
\includegraphics[width=\textwidth]{figures/sedov-profiles.eps}
\caption{Radially averaged profiles for the Sedov Blast test at $t =
0.07$. Clockwise from top-left shows density, velocity, internal
energy and pressure.  The black solid line shows the analytic
solution.  The blue dashed line shows the simulation using the PPM
method, and the red dot-dashed line using \zeus.  The \zeus\ result
substantially lags the true result due to total energy not being
explicitly converged.}
\label{fig.sedov2}
\end{center}
\end{figure}

The Sedov Blast Test \citep{Sedov1959} models an intense explosion,
initiated by depositing thermal energy into a homogenous distribution
of gas. The result is a strong spherical shock wave centered on the
point of energy injection.  This problem is a popular test of
astrophysical numerical codes for three reasons: First, it is
particularly appropriate to astrophysics since it represents the
situation when a supernovae explosion occurs. Second, it has an
exact analytical solution whereby the shock front's radial position is
given by:

\begin{equation} r(t) =
\left(\frac{E_0}{\alpha\rho_0}\right)^{1/5}t^{2/5}
\end{equation}

\noindent where $E_0$ is the initial energy injection, $\rho_0$ is the
background density and $\alpha = 1.0$ for cylindrical symmetry and an
ideal gas with $\gamma = 1.4$. For the full derivation see
\citet{Sedov1959}. This solution makes it possible to assess how well
the code performs.  Third, the spherical shock is a challenging
problem for numerical codes (both grid and particle) since the shock
wave increases in size as the simulation progresses and its
symmetrical nature highlights any directional preferences that grid
codes can succumb to. The test presented here is the two-dimensional
version that is included in the \enzo\ distribution. The
three-dimensional results from this test, both for \enzo\ and three
other leading astrophysics codes, can be founds in \citet{Tasker2008}.

In the initial state, the box contains a homogenous distribution of
gas at a density of 1 (note that, in the absense of gravity, this
problem is scale-free and thus without units). Thermal energy is
deposited into a single cell at the center of the box with $E_0 =
10.0$. The problem is completed in two dimensions with reflecting
boundary conditions (the default for \enzo) and uses a box size of side
$1$. For this problem, we selected a top grid of $100 \times 100$
cells and a maximum of four levels of refinement, placed based on
shock location and the slope in density and total energy. The
exception to this scheme is in the initial conditions, where grids are
placed directly around the injection point. The results were assessed
at $t = 0.07$, which corresponding to a time just before the shock
reaches the box edge (see Figure~\ref{fig.sedov1}).

Figure~\ref{fig.sedov2} shows the radial profiles for the simulation
run with the PPM hydro-solver (blue dashed line) and the \zeus\
hydro-solver (red dash-dot-dot line) together with the analytical
solution (black solid line). Clockwise from the top-left are density,
velocity, internal energy and pressure. PPM matches the analytical
solution extremely well for all quantities. However, the shock front
in the \zeus\ simulation lags behind the analytical position. This can
also be seen in the slices shown in Figure~\ref{fig.sedov1}. The cause
of this discrepancy is that \zeus\ shows a substantial energy loss
during the first few timesteps and produces a diamond-shaped, rather
than spherical, shockfront during this time. After this, the code
correctly conserves energy but this intial energy loss remains clearly
visible in the position of the shock at $t = 0.07$. This problem was
addressed directly by \citet{Clarke2010}, who attributed the source of
the issue to this version of \zeus\ solving the internal, rather than
total, energy equation. In situations with strong energy gradients,
this choice caused an energy loss and the artificial viscosity
produces the direction-dependent shockfront shape. In their paper,
\citet{Clarke2010} presents results from an alternative version of
\zeus\ that conserves total energy. This problem is less marked for
smaller energy gradients and it should be noted that the \zeus\ hydro
algorithm's stability and speed make it a highly competitive choice,
despite the disagreements in this test.

### 2013-05-21 tests-selfsiminfall.tex
Splitting up test problem section into subsections.

\subsubsection{Self-Similar infall test}
\label{sec.tests.infall}

The next test problem is based on the self-similar solutions to the cosmological spherical infall problem found by \citet{Bertschinger1985}.  It features a small density perturbation in a homogeneous $\Omega = 1$ universe and is a strong test of the cosmological evolution, gravity, and hydrodynamic portions of the code.  It is close analogue to cosmological cluster formation and has an analytic solution.

For initial conditions, we adopt a $32^3$ top grid with a single, initial sub grid (covering $1/8^3$ of the domain) with a refinement factor of 2.  An overdensity $\delta = 40$ is placed in a single cell near the center of the domain.  We begin at $z=199$ and evolve to $z=0$, which is a sufficient time for the evolution to largely forget its initial conditions and approach the self-similar result.  An over density refinement criteria ($\delta_{\rm crit} = 1.1$ on the top grid) is used to add additional grids, going up to 5 additional levels beyond the root grid.  We use the PPM solver without radiative cooling; only baryons are used for this problem.  An ideal gas law with $\gamma = 5/3$ is adopted.

In Figure~\ref{fig.sphericalinfall}, we show the results, scaled according to the dimensionless variables as defined in equations (2.9) and (3.2) of \citet{Bertschinger1985}.  This demonstrates that we can quickly and easily obtain a good solution with only a fairly modest initial grid.  The shock is sharply resolved and the asymptotic profiles as small $\lambda$ are recovered.  At very small values of $\lambda$, the initial conditions have not been fully forgotten and the self-similar solution is not recovered -- this can be seen most clearly in the dimensionless mass; however, this is simply because of the limited amount of time for which we evolved the solution.

\begin{figure}
\begin{center}
\includegraphics[width=0.7\textwidth]{figures/SphericalInfall.eps}
\caption{The results of the self-similar spherical infall test.  The panels show, as a function of the dimensionless radius $\lambda$, the dimensionless density $D$ (top left), velocity $V$ (top right), pressure $P$ (bottom left), and enclose mass $M$ (bottom right).  In each case we show azimuthally averaged profiles from the simulation as circles and the analytic solution for for $\gamma = 5/3$ as solid lines.}
\label{fig.sphericalinfall}
\end{center}
\end{figure}

### 2013-05-21 tests-shockpool.tex
Splitting up test problem section into subsections.

\subsubsection{Shock pool}
\label{sec.tests.shockpool}

The next problem is similar to the last one, but examines how a shock wave with a Mach number of 2 passes into and out of a static refined region (again defined from $z=0.25$ to 0.75).  The density and pressure in the domain is initially set equal to 1.0, with zero velocity, and then at $t=0$, the left boundary is set with the density, pressure and velocity appropriate for $M=2$ shock wave.

In Figure~\ref{fig.shockpool}, we show the evolution of the shock wave at three times, again corresponding to just before entering the refine region (left column), after entering the refined region (center), and after exiting the refined region (right).  We examine the same set of three solvers as in the previous test problem.

Beginning with PPM (top row), we see that this method captures the shock in a small number of zones with only a small amount of oscillation.  Open entering the refined region, the shock finds itself broader than the natural width of the scheme (since the cell spacing is decreased by a factor of 2), and so the shock front contracts, causing a slight entropy perturbation in the post-shock gas.

For Zeus (middle row of Figure~\ref{fig.shockpool}, the shock is broadened because of artificial viscosity, and there are slightly more post-shock oscillations, although again quite mild.  The impact of entering and exiting the refined region is somewhat larger than in PPM; however, the most noticeable difference is the incorrect position of the shock front, which is due to the fact that the scheme is not energy-conserving (see also the Sedov problem, below), and has essentially nothing to do with the passage through the refined region.

Finally, the MUSCL scheme produces a shock which is intermediate in width between the two previous cases.  This method is energy conserving and so correctly reproduces the shock speed.  The oscillations are mild except for the cell immediately outside of the refined region upon exiting.

\begin{figure}
\begin{center}
\includegraphics[width=0.8\textwidth]{figures/ShockPool}
\caption{This plot shows, for each column, three snap shots (at $t=0.1, 0.15$ and 0.35) of a $M=2$ shock wave as it propagates through the grid with a static refined region extends from $x = 0.25$ to 0.75 (shown in grey in each panel).   Each row shows the result for a different solver (PPM/Zeus/MUSCL from top to bottom).  The solid line shows the analytic solution.  Note that we focus each panel on a small region of the entire domain to better show the shock front.}
\label{fig.shockpool}
\end{center}
\end{figure}

### 2013-05-21 tests-sodshocktube.tex
Splitting up test problem section into subsections.

\subsubsection{Sod Shock Tube}
\label{sec.tests.sodshock}

We begin the paper test suite with the classic Sod shock tube problem
\citep{Sod78}, which provides a good test of a hydrodynamical solver's
ability to resolve a clean Riemann problem with clear separation
between the three waves.  These waves include a rarefaction fan, a
contact discontinuity, and a moderate shock.  The initial state is 
$\rho_{\rm L}, P_{\rm L} = 1.0, 1.0$ on the left of the boundary at 
$x=0.5$ and $\rho_{\rm R}, P_{\rm R} = 0.125, 0.1$ on the right.
All velocities are initially zero.  In
Figure~\ref{fig.sodshocktube}, we show the density solution at the
final time ($t=0.25$) for three of our hydro solvers -- the spatially third-order
PPM, as well as the two second-order \zeus\ and MUSCL schemes.  We use
100 cells across the domain, (which is a relatively standard choice in
code method papers) and show solutions both with and without adaptive
mesh refinement.  Without AMR (in the top row of the figure), it is
clear that the PPM scheme produces by far the cleanest solution, with
all wave families crisply reproduced (in particular, the contact
discontinuity and shock).  \zeus\ and MUSCL produce similar results,
with MUSCL doing a slightly better job on the rarefaction fan.  In the
figure, we also show the mean absolute deviation from the exact
solution ($||E_\infty||$), as well as the maximum deviation
($||E_1||$).  These numbers confirm the qualitative differences noted
previously.

We also run the same simulation but with two levels of AMR (using a
refinement factor of 2), triggered based on a normalized slope greater
than 10\% in the density.  This refinement criterion results in only
the refinement of strong gradients, and
does not include the rarefaction fan at late times.  The results are
shown in the bottom row of Figure~\ref{fig.sodshocktube}.  Using AMR,
the results are much better for all three methods, with much sharper
shocks and contact discontinuities and even a better representation of
the rarefaction wave, which is only refined beyond the root grid at
early times.  Although the results are improved for all methods, PPM
still produces the best result, as is shown clearly by the computed
error norms (displayed again in each of the individual panels).

\begin{figure}
\begin{center}
\includegraphics[width=\textwidth]{figures/SodShockTube.eps}
\caption{The density distribution of the classic Sod Shock Tube for
three different solvers (from left to right column) and with (bottom
row) and without (top row) AMR.  In each case 100 zones are used on
the root level and the results are shown at $t=0.25$.  All cells are
plotted and color-coded by level (at the time shown, only small region
surrounding the contact discontinuity and the shock are refined).  In
each panel, we show the analytic solution as a solid line and the $E_1$
and $E_\infty$ error norms in the upper right.}
\label{fig.sodshocktube}
\end{center}
\end{figure}

### 2013-05-21 tests-vandv.tex
Splitting up test problem section into subsections.

\subsubsection{The \enzo\ test suite}
\label{sec.tests.suite}

\enzo\ is capable of simulating a large variety of problems types,
with all but a few of these types requiring only a parameter file as an
input.  The notable exception is the cosmology simulation, which takes
as input initial conditions created by other codes.  The suite of test
problems spans a wide range in complexity.  At one end of this spectrum
are simple problems that utilize only a single component of \enzo\ and
for which analytic solutions exist for comparison with the simulation
results.  At the opposite end are problems that exercise a large
portion of \enzo's machinery in concert.  Together, the available
problem types fully cover all of \enzo's functionality.  This enables
them to serve as a vehicle for verifying that the code's behavior
remains stable over time and after modifcation of the codebase.

\enzo\ uses an automated testing framework that allows a user to run,
with a single command, a set of test problems and compare the results
against results produced by any other version of the code.  Within the
\enzo\ source distribution, the test problem parameter files are stored
in a nested directory structure organized according to the primary
functionality tested, e.g., hydrodynamics, gravity, cooling, etc.
Each parameter file is accompanied by a text file containing various
descriptive keywords, such as the machinery tested, the
dimensionality, and the approximate run time.  A test runner script is
responsible for taking as input from the user a series of keywords
that are used to select a subset of all available test problems.  The
test problems are also grouped into three suites: the quick, push, and
full suites, each a superset of the ones named before.  The quick
suite is considered to minimally cover the primary functionality in
\enzo\ and is designed to be run repeatedly during the development
process.  The push suite has slightly increased feature coverage and
is mandated to be run before code changes are accepted into the main
repository.  The full suite consists of all test problems that can be
run with no additional input.  Approximate run times for the quick,
push, and full suite are 15 minutes, 1 hour, and 60 hours,
respectively, on a relatively new desktop computer (circa 2013).

After the test problems are selected by the test runner script, they
are run in succession using either the \enzo\ executable contained
within that distribution or an external executable built from another \enzo\
version and specified by the user.  This allows for any version of the
code to be tested with an identical set of test problems.  After
running the test problem simulations, the test runner then performs a
series of basic analysis tasks using the \texttt{yt} analysis toolkit
\citep{2011ApJS..192....9T, 2011arXiv1112.4482T}.  The default
analysis performed on all test problems includes calculation of
various statistics, such as extrema; mean; and variance, on the fields
present in the output data.  Custom analysis provided by scripts that
accompany the test problem parameter files is run for special cases,
such as when an analytical solution exists that can be compared
against the simulation data.  After the analysis is performed, the
results are compared against a set of gold standard results that are
maintained on a website and downloaded on the fly by the test runner
script.  Alternately, results from any version of \enzo\ can be stored
locally and compared to any other version of the code.  In
Section~\ref{sec.tests.problems}, we describe some of the key test
problems that are used to verify proper behavior.  All of these test
problems, as well as the scripts to generate the figures from them,
are available in the \enzo\ test suite, allowing the interested reader
to regenerate all of the figures in Section~\ref{sec.tests.compare}
using the \textit{'methodpaper'} suite of test problems.

\subsubsection{Code comparisons}
\label{sec.tests.compare}

Over the course of its existence, \enzo\ has been involved in numerous
comparisons with other astrophysical codes used for self-gravitating
fluid dynamics.  In general, \enzo\ behaves in a manner similar to other
grid-based (and particularly AMR-based) codes, as we will summarize below.

\enzo\ has been involved in multiple cosmological code comparisons,
including the Santa Barbara Cluster code comparison project
\citep{SantaBarbara}, a large comparison of N-body simulations
\citep{2008CS&D....1a5003H}, as well as several direct comparisons
between \enzo\ and the GADGET SPH code under a variety of
circumstances, including N-body and adiabatic hydrodynamics
\citep{2005ApJS..160....1O,2005MNRAS.364..909V, 2011MNRAS.418..960V}
and simulations of the Lyman-alpha forest \citep{2007MNRAS.374..196R}.
The \enzo\ code typically has a more difficult time resolving
small-scale self-gravitating structures (for an equivalent dark matter
particle count and nominally equivalent force resolution), but does
comparably well as a tree-based code for larger structure, and is
typically superior in terms of resolving fluid features due to its
higher-order (and artificial viscosity-free) Piecewise Parabolic
Method hydrodynamics solver.  When examining classical test cases such
as the Santa Barbara project \citep{SantaBarbara}, \enzo\ forms galaxy
clusters with very similar density, temperature, and entropy profiles
to other grid-based codes that use Godunov-type hydro methods, which
systematically differed from particle-based codes using SPH in this
code comparison.  Similarly, in galaxy cluster simulations that look
specifically at the properties of cosmological shocks
\citep[e.g.][]{2011MNRAS.418..960V}, \enzo\ produces results that are
similar to other high-order grid-based hydrodynamics codes, and far
superior performance of \enzo\ is observed (in terms of resolution of
fluid features and shock detection) in low-density regions when
compared to a particle-based code.  In tests of the Lyman-alpha forest
that include radiative cooling and a uniform metagalactic ultraviolet
background, \enzo\ and Gadget provide results on metrics such as the
matter power spectrum that are comparable to within 5\%
\citep{2007MNRAS.374..196R}.

Several comparisons have been made that focus specifically on
hydrodynamics solvers and fluid behavior.  In particular, the work of
\citet{2007MNRAS.380..963A} and \citet{Tasker2008} perform
head-to-head comparisons between several grid- and particle-based
codes for a variety of fluid-centric test problems (including shocked gas clouds,
self-gravitating, translating clouds, and Sedov-Taylor explosions),
and show that \enzo\ is comparable or superior in behavior to the
other grid-based hydrodynamics codes involved in the comparison, and
provide useful information on the sort of practical challenges that a
user of an
AMR code such as \enzo\ may experience.  More specific comparisons, including one
specifically testing the linear and nonlinear growth of the
Kelvin-Helmholz instability \citep{2012ApJS..201...18M}, as well as a
comparison that more
broadly examines Galilean invariance in grid-based codes
\citep{2010MNRAS.401.2463R}, show that \enzo, and in particular its
implementation of the Piecewise Parabolic Method hydrodynamics solver,
converge to the correct solution as expected, and generally provide
less diffusive solutions than lower-order codes, including those that
use artificial viscosity.  Finally, two code comparison projects have
been undertaken to examine turbulence in hydro codes -- one studied
the behavior of decaying isothermal supersonic turbulence
\citep{2009A&A...508..541K}, and the other looked at supersonic
magnetohydrodynamical turbulence \citep{2011ApJ...737...13K}.  Both
included the \enzo\ code, with the former testing the PPM
hydrodynamics and the latter the constrained transport MHD
implementation of \citet{Collins10}.  In both cases, \enzo\ performed similarly to other
grid-based codes that use Godunov-based fluid solvers, and typically
had better effective resolution than particle-based codes using the
same number of particles as the \enzo\ simulation used grid cells.

Three other comparisons between the \enzo\ code and other simulation
tools have been performed that focus on physics other than gravity and
fluid flow.  The flux-limited diffusion radiation transport scheme was
measured against several test problems by \citet{IlievEtAl2009},
producing results that are similar to both expected results (i.e.,
analytic test problems) and other methods (i.e., codes using other
methods, written by other groups), although there are some minor
differences (although all codes in the study exhibited differences
from the majority of other codes on at least a subset of the tests).
\citet{2011ApJ...726...55T} shows the results of varied reaction rate
coefficients for the formation of molecular hydrogen via the 3-body
process in both the \enzo\ and Gadget-2 codes.  Similar trends were
visible in both the \enzo\ and Gadget simulations, though at a
nominally equivalent resolution (i.e., with particle and cell gas
masses being comparable) \enzo\ simulations typically displayed a
substantially higher level of gas structure.  This is unsurprising due
to \enzo's higher-order hydro solver.  Finally,
\citet{2012ApJ...744...52P} show the results of comparing \enzo\ in
its non-AMR mode to a smoothed-particle hydrodynamics code, SNSPH, in
the context of common-envelope binary stellar interactions.  The
authors show that the codes display reasonable convergence properties
as a function of simulation resolution, and also agree quite well with
each other -- however, the observed mass-loss rates do not agree
particularly well with observations.

### 2013-05-21 tests-wavepool.tex
Splitting up test problem section into subsections.

\subsubsection{Wave pool}
\label{sec.tests.wavepool}

In this one-dimensional test, we pass a short wavelength, linear wave through a static refined region.  The full domain is from 0 to 1 and the refined region is set from 0.25 to 0.75 at all times.  We modify the left boundary consistent with a single linear sound wave with density given by $\rho(x,t) = \rho_0 (1 + A \sin(kx - \omega t))$, where the amplitude is $A = 0.01$ and the wavelength $\lambda = k/2\pi = 0.1$, and similar expressions exist for the pressure and velocity.  The initial, unperturbed density and pressure are set to unity and we adopt $\gamma = 1.4$.  The (unrefined) root grid is covered with 100 cells, resulting in a wavelength for the linear wave of only 10 cells.  This is, therefore, a challenging problem for hydro methods.  We are particularly interested in any reflection or artifacts introduced by the wave entering and exiting the refined regions.

Figure~\ref{fig.wavepool} shows the evolution of the wave at three different times ($t = 0.2, 0.3$ and 0.8) for three of our solvers (PPM, Zeus, and MUSCL).  The leftmost column show the wave just before it enters the refined region so that we can gauge how the solver is operating in the absence of AMR, the center column shows the wave after it has fully entered the refined region, and the rightmost column shows the wave well after it has exited the refined region.

Beginning with the PPM, we see that even before it has entered the refine region, the wave is slightly damped, which is not unexpected for such a short wavelength mode, even for PPM.  The remaining panels demonstrate that the wave cleanly enters and exits the refined region.  No significant reflection is seen, either on entry or exit, and the amount of damping is mild.

For the Zeus solver, we see that even before it has entered the refined region, there are small oscillations excited behind the wave; although it is also worth noting that the wave itself is beautifully propagated without significant damping or phase errors.  In this test, we use only our standard, low amount of quadratic artificial viscosity -- these waves can be damped by additional viscosity, but we do not add any in order to be sensitive to any artifacts at the grid boundary.  The remaining panels show that the trailing oscillations continue, but do not generate any additional noise -- the end result is similar to the case without any refined region.

Finally, the MUSCL solver shows also shows a very clean entry and exit from the refined region without any oscillations, although (for the PLM reconstruction used here) the wave is spread more than with the other methods.

\begin{figure}
\begin{center}
\includegraphics[width=0.8\textwidth]{figures/WavePool.eps}
\caption{This plot shows, for each column, three snap shots (at $t=0.2, 0.3$ and 0.8) of a linear wave as it propagates through the grid.  The static refined region extends from $x = 0.25$ to 0.75 and is shown in grey in each panel.  The individual cells are also color-coded by level: blue indicates the root grid, and red is for the refined region.  Each row shows the result for a different solver (top is PPM, middle is Zeus, and the bottom is MUSCL).   The solid line shows the analytic solution for a linear, undamped wave.  Note that we focus each panel on a small region of the entire domain to better show the wave itself.}
\label{fig.wavepool}
\end{center}
\end{figure}

### 2013-05-21 tests-zeldovichpancake.tex
Splitting up test problem section into subsections.

\subsubsection{Zel`dovich Pancake}
\label{sec.tests.pancake}

The Zel`dovich pancake \citep{1970A&A.....5...84Z} is particularly
relevant to cosmological simulations, because it includes many
features that are critical to structure formation: hydrodynamics,
expansion, and self gravity.  This problem represents the formation of an ideal,
isolated caustic, and is thus a useful proxy for much more
complicated structures in full 3-dimensional cosmological simulations
(such as the collapse of gas onto a cosmological halo or filament).

The initial conditions are simple, and we follow the prescription of
\citet{Anninos94}.  Assuming a geometrically flat cosmology, the density
perturbation is given by

\begin{equation}
\rho(x_l) = \rho_0 \big[ 1 - \frac{1+z_c}{1+z} \cos(k x_l) \big]
\end{equation}

with the internal energy of the gas set so that the entropy of the gas
maintains a constant value throughout.  The velocity perturbation is given by

\begin{equation}
v(x_l) = -H_0 \frac{1 + z_c}{(1+z)^{1/2}} \frac{\sin(k x_l)}{k}
\end{equation}

In the equations above, $\rho_0$ is the background density, z$_c$ is a
free parameter and is the redshift where the sheet forms a caustic
(i.e., where it `pancakes'), z is the redshift of initialization, x$_l$ is the
Lagrangian mass coordinate, $k = 2 \pi / \lambda$ (where $\lambda$ is
the perturbation wavelength), and H$_0$ is the $z = 0$ value of the
Hubble constant.  Note that this solution is expressed in terms of
Lagrangian positions, so one needs to convert this into the Eulerian
coordinates, x$_e$, that are more useful to \enzo:

\begin{equation}
x_e = x_l - \frac{1 + z_c}{1 + z} \frac{\sin(k x_l)}{k}
\end{equation}

We note that the solution described above is exact up to the point of
caustic formation.  In Figure~\ref{fig.pancake}, we show the results
of a test of the adaptive mesh version of \enzo's Zel'dovich Pancake test.  A
one-dimensional box of length $64$~Mpc/h is initialized at z$ = 20$ in
an $\Omega_M = 1$ universe with h$ = 0.5$ and a background temperature
of 100~K, $\rho_0 = \rho_c$.  The simulation is initialized with 64
grid cells, refining by factors of four using criteria based on cell
mass and the presence of shocks, for a maximum of 2 levels (i.e., an
equivalent maximum resolution of 1024 grid cells).  The simulation was
evolved to z$ = 0$ using the PPM hydro method, with the final output
of the calculation being shown in Figure~\ref{fig.pancake}.

Figure~\ref{fig.pancake} shows the key features of this test problem.
The strong shocks and large density gradients are well-resolved, with
density and velocity jumps being well-delineated and at the correct locations.  The key features of
this test problem can be resolved with far fewer cells -- simulations
including a mere 8 cells resolve the key features, as shown in
Sections 3.3.4-3.3.5 of~\citet{1996PhDT........80B} -- but we choose a
higher resolution here for illustrative purposes.

\begin{figure}
\begin{center}
\includegraphics[width=0.8\textwidth]{figures/AMRZeldovichPancake.eps}
\caption{Zel`Dovich Pancake test shown at z$ = 0$, initialized in a
$\Omega_m = 1$ universe at z$ = 20$ on a one-dimensional grid having
64 cells, and further refined by factors of four based on cell mass
and the presence of shocks for up to two additional levels of mesh,
having a maximum effective resolution of 1024 grid cells. The top,
middle, and bottom rows show density, temperature, and velocity of the
gas, respectively; all are a function of position in units of the box
size.  The central region ($x \simeq 0.45-0.55$) has bene adaptively
refined, as can be seen by the locations of grid points.  Shocks and
the central density peak are clearly resolved, with well-delineated
jumps at the appropriate locations.}
\label{fig.pancake}
\end{center}
\end{figure}

### 2013-05-21 numerical-radiation-homogeneous.tex
clean up

@@
 \citet{2012ApJ...746..125H}, modified to match the normalized field
-found in \citet{Kirkman05}.  Type 4 is the same as Type 3 but also
+found in \citet{Kirkman05}.
+
+The remainder of the radiation field types either build on the fits in Types
+1-3 or use a completely different form and therefore we describe them
+in the text rather than in Table~\ref{table:homo_coefs}.
+Type 4 is the same as Type 3 but also
 adds X-ray Compton heating from \citet{MadauEfstathiou99}, using
@@
 
-Types 8, 9, and 14 have only photo-dissociating Lyman-Werner flux,
+Types 8, 9, and 14 have only a photo-dissociating Lyman-Werner flux,
 with Type 8 being constant and Type 9 using the redshift-dependent
@@
 \citet{WiseAbel05} for the range $6 < z < 50$, constant for $z<6$, and
-zero for $z > 50$.  Other types are undefined or unused.
-
-\red{Note: we talk about several radiation field types that are not in
-Table~\ref{table:homo_coefs} (4...14).  Do we want to increase the
-length of the table, remove the text from the section, or perhaps just
-say something in this section about how the table is for illustrative purposes? }
+zero for $z > 50$.  Other types are either undefined or currently unused.
 

### 2013-05-22 parallel.tex
Added parallel section

@@
 
-% GB to work on this
-%  - expand to include better description
-%  - include load balancing options
+The current version of \enzo\ has been parallelized for distributed memory platforms using the Message Passing Interface (MPI).
+This is done using a single grid object as the basic unit of parallelization.
+Each grid object -- including all cell and grid data -- is fully contained in a single processor\footnote{We use the word {\it processor} to mean a basic distribution unit; this could by either a core or a node, depending on details of the system.
+Note that the current version of Enzo is not threaded, although the development of a hybrid version is in process}.
+Parallelization is accomplished by distributing grids amongst processors.
+This is done on the root grid using a simple tiling system, where the root grid is split up into $N_{\rm root}$ tiles, where $N_{\rm root}$ is typically equal to the number of processors $N_p$.
 
-When \enzo\ is run on a distrubuted platform, some of the data is
-replicated to all processors, while some remains on a single processor.
-A description of the entire hierarchy (the position and size of each grid, as well
-some other meta data) is stored on each processor.  This enables each
-grid to know what grids are near it in the hierarchy, regardless of
-where the data is.  The actual Baryon data (density, velocity, energy,
-and any chemistry data) and particle data are stored on only one
-processor.  Storing the entire hierarchy on all processors does create
-some overhead, but in practice this is not an issue until one reaches
-extremely large scales.
-  The code handles load balancing 
-on a level-by-level basis such that the workload on each level is 
-distributed as uniformly as possible across all processors.  Spatial locality of 
-grids is not forced during message passing, for maximum flexibility (though not
-necessarily maximum efficiency).  
-The MPI message passing library
- is used to transfer data between processors.
+Load balancing on levels other than the root level (i.e.
+grids for which the level $l > 0$) is different as the refined patches are not generally uniformly distributed.
+Grids on refined levels are first placed on the same processor as their parent to minimize communication; however, this generally does not result in a well-balanced load.
+Therefore, the code has a number of options for load balancing the grids on a given level.
+Each grid is assigned a load, generally equal to the number of cells in the grid.
+The first load-balancing option is to move a grid from the processor with the highest load to the processor with the lowest load, with the proviso that only grids with load factors less than half the difference between the highest and lowest loaded processors will be moved.
+This continues until the load ratio between the most-to-least loaded processor is below 1.05 or until no suitable grid can be found to transfer.
+A second load balancing option uses a space-filling Hilbert curve to order the grids by their approximate spatial position.
+Then, once the grids have a specific one-dimensional ordering, we can divide up the grids into $N_p$ groups (as equally as possible).
+Load balancing is done separately for each level.
+Clearly load balancing is most successful if there are significantly more grids than processors; however, small grids are less efficient (because of their ghost zones), and so the code uses a simple heuristic in order to split up grids until there are of order 10 grids per processor.
+This generally results in good load-balancing while not producing grids which are wastefully small.
 
-%old text
-%Every processor 
-%keeps a description of the entire grid hierarchy at all times, so that 
-%each processor knows where all grids are.  However, baryon and particle 
-%data for a given grid only exists on a single processor.  See 
-%Figure~\ref{fig.2.amrhierarchy} for a schematic example of this.
+Communication between processors is done using a non-blocking communication strategy that allows overlap of communication and computation.
+This can be done efficiently because each processor retains a copy of the entire hierarchy of grids, except that grids which do not `live' on a given processor only contain meta data (essentially location and size of the grid; such grids are denoted as `ghost' grids).
+Replicating the hierarchy means that all communication of data from one grid to another can be identified by each processor independently.
+The metadata for `ghost' grids are quite small and so the extra memory required is generally not onerous unless very large numbers of grids are used (more than a few hundred thousand grids).
+A schematic of this distribution is shown in Figure~\ref{fig.amr_hierarchy}.
+
+Data is transferred through a three-step procedure that takes advantage of MPI capabilities: (i) as the code progresses, and data is needed from another grid on another processor, the receiving processor posts an MPI non-blocking receive indicating that it is expecting data; this outstanding receive is recorded in a table , (ii) the sending processor calls the MPI non-blocking send function, and then finally (iii) the receiving processor, after it has carried out all the computation it can, waits for any MPI message to arrive.
+Each message is coded so that it can be matched with appropriate receive posted in the first step, and based on that, the appropriate routine is called to processes the data.
+Step (iii) is repeated until there are no outstanding receives.
 
@@
 \begin{center}
-\includegraphics[width=0.4\textwidth]{figures/amr_hierarchy.eps}
+\includegraphics[width=0.5\textwidth]{figures/amr_hierarchy.eps}
 \end{center}
@@
 needed for the PPM hydro stencil. }
-\label{fig:amr_hierarchy}
+\label{fig.amr_hierarchy}
 \end{figure}
@@
 
-\subsubsection{Performance model \red{(Greg)}}
-
-% Maybe include Greg's text on general scaling arguments
-
-\subsubsection{Performance Measurement \& Instrumentation \red{(Sam)}}
+\subsubsection{Performance Measurement \& Instrumentation}
 
@@
 
-%%% Local Variables: 
-%%% mode: latex
-%%% TeX-master: "ms"
-%%% End:
+\subsubsection{An Approximate Time and Memory Scaling Model}
+
+In this section, we develop a simple, approximate model to estimate the compute time and memory required to complete a given calculation.
+Note that we are {\it not} trying to model parallel performance (which we discussed in the previous section), but have an even simpler goal: to determine how to scale compute time estimates as we increase the resolution.
+This is a straightforward thing to do in codes which use static grids, as the computational effort per time step is constant.
+However, for an AMR calculation, the number of grids that will be generated during the run is not known in advance, and so the cpu time per problem time can vary drastically throughout the calculation.
+%For example, at the beginning, when the densities are nearly uniform, only the static grid is required and the calculation progresses rapidly.
+However, as structure forms and dense clumps are generated, the number of grid points swells by orders of magnitude (an increase by $10^3$ is not uncommon) and most of the cpu time is consumed at late times.
+Therefore, simply performing a few steps does produce a good estimate of the cpu time.
+
+Nevertheless, we can try to determine how the compute time scales for a given run, as we increase the resolution.
+First, we neglect particles and concentrate on the time taken by the grids, which can be justified both theoretically and empirically\footnote{Since the potential is calculated on the grid, the only particle costs are: depositing mass to the grid, interpolating accelerations, and updating particle positions and velocities.
+These are relatively cheap operations.}.
+Second, we break the problem down slightly, and examine the scaling over a short enough period of time that the grid structure does not change significantly (i.e.
+the number of grids at each level).
+We then assume that the whole run scales in the same way, or in other words, that changes in the resolution affect the simulation in the same way at each time step.
+This is usually a good approximation as the most costly part of the calculation typically don't change their characteristics substantially.
+
+To make progress, we assume that the cost to advance a single cell by one timestep is a constant.
+This can be incorrect if the chemistry solver requires many iterations, but is usually fairly accurate.
+For a unigrid calculation, the time would by proportional to $N_{\rm root}^4$ since the number of cells scales as $N_{\rm root}^3$ and, assuming the Courant condition is the time limiting factor, then the number of steps to advance a given time is proportional to $N_{\rm root}.$  Therefore, to advance a hierarchy a given time, we find, accounting for all levels and using a refinement factor of 2.
+\begin{equation}
+t_{\rm SU} = C_1 \sum_l f_l N_{\rm root}^4 2^{4l}
+\end{equation}
+where $C_1$ is a constant, which can be thought of as the time taken to advance a single cell.
+The factor $f_l$ is the fraction of the volume on a given level that is actually refined.
+By definition $f_0 = 1$, and $f_l \le 1$.
+In writing down this equation, we neglect a number of costs which are not directly proportional to cell count, including communication between processors, the $\log{N}$ factor for the root grid FFT, Cache misses, optimizations, and other costs associated with processing the hierarchy.
+The first item on this list, in particular, is clearly not true for large processor counts; however, we neglect parallel considerations in this section.
+
+Note that unless $f_l/f_{l+1} < 1/16$ (i.e.
+if less then about 6\% of a given level is further refined), the cost per level will increase with level.
+A key question therefore, is the values of $f_l$.
+Unfortunately, this depends strongly on the simulation being run.
+In Figure~\ref{fig:scaling}, we show the values of $f_l$ for three simulations: two cosmological runs with varying box size, and a third simulation focusing on a single disk galaxy.
+
+\begin{figure}
+\centerline{\includegraphics[width=0.7\textwidth]{figures/scaling_plot2}}
+\caption{This figure show (clockwise from upper left), the covering fraction $f_l$ of the grids on a given level $l$, the ratio of $f_{l}/f_{l+1}$, the relative memory usage of each level (normalized by the memory usage of level 0), and, in the lower-left panel, the relative cpu usage of each level, again normalized by $l=0$.
+In each panel, three curves are shown: the solid black and long dashed red are two cosmological simulations with box sizes of 30 h$^{-1}$ Mpc and 80h$^{-1}$ Mpc respectively.
+The dot-dashed blue line is for a non-cosmological simulation of a ram-pressure stripped galaxy.
+In the top-right panel, there is a line at the critical ratio of 0.06, which determines if the next finer level ($l+1)$ takes more (above) or less (below) cpu time than level $l$.}
+\label{fig:scaling}
+\end{figure}
+
+As, the figure demonstrates, the $f_l$ values all show a sharp decline with level, dropping as power-laws with the level $l$.
+Focusing first on the ``refine-everywhere" cosmological runs, which both show similar behavior despite the different box sizes and redshifts at which we collect the data.
+In fact, we find the results are very typical for cosmological runs, with only slight variations depending on if we are simulating first stars or clusters of galaxies.
+The top-left panel shows the ratio of $f_{l}/f_{l+1}$ which, remarkably, hovers around the critical value of 0.0625 determined earlier.
+This implies that the relative cpu usage of each level, shown in the lower-left panel, is nearly flat, and that going to high levels of refinement only increases the amount of calculation by a factor of roughly $1/l_{\rm max}$.
+Even in the most extreme case, the L80 run, the last level adds only 25\% to the computation time.
+
+The memory used by each level, on the other hand scales as:
+\begin{equation}
+{\rm memory} = C_2 \sum_l f_l N_{\rm root}^3 2^{3l}
+\end{equation}
+and the terms in this sum are shown in the bottom-right panel of Figure~\ref{fig:scaling}.
+This demonstrates that the memory usage is dominated by the top three levels, and adding additional levels only adds minimally to the memory usage of the run.
+We have empirically tested this scaling for cosmological simulations and found it to be reasonably accurate (although the increase is typically slightly larger than found here, usually 30-50\%, probably because of less-than-ideal parallel scaling).
+
+The physical reason for the $f_{l+1}/f_l$ ratios found in the cosmological simulations appears to be due to a combination of the density structure of individual clouds, and the distribution of the clouds themselves.
+In particular, we note that for a $\rho \propto r^{-2}$ profile, the Lagrangian refinement criteria typically used in such simulations produces an $f_{l+1}/f_l$ ratio of 1/8 for a single, resolved cloud.
+Of course, at some point we would resolve the flat density of individual clouds and the ratio would climb, making further refinement more costly; however, we are not yet in this regime in this example.
+
+These conclusions are, however, highly dependent on the type of run being performed.
+We contrast this cosmological case to the simulation of a galaxy formation run, as shown by the dot-dashed curves in Figure~\ref{fig:scaling}.
+Note that the levels 1 and 2 are statically refined to ensure the Lagrangian volume of the galaxy halo is resolved to at least level 2 at all times.
+More importantly, the $f_{l+1}/f_{l}$ ratio of the highest levels are systematically slightly larger than the critical 0.0625 value, indicating that the cpu time is dominated by the most refined levels, as shown in the bottom-left panel.
+The memory usage is dominated by levels 2 and 3, the static levels, as the bottom-right panel demonstrates.
+In this case, adding more levels of refinement would somewhat increase the cost of the simulation.
+
+Finally, we have so far focused on scaling when increasing $l_{\rm max}$, but without changing the refinement criteria.
+However, we can also keep the levels fixed and modify the refinement criteria.
+For cosmological runs, this is typically done by increasing the mass resolution (and simultaneously decreasing the dark matter particle mass), and adding additional linear small-scale modes to the initial conditions.
+The effect of this is to boost the $f_l$ values at all levels (except for the root grid where $f_0$ is already 1) by a factor linearly proportional to the increase in the mass resolution.
+This is because the refinement criteria typically boosts the number of refined cells on the first and subsequent levels by this factor.
+Again, we have empirically demonstrated that this scaling is approximately valid (provided we keep the maximum level constant).
+
+Putting this together, we obtain an approximate scaling for cosmological simulations as 
+\begin{equation}
+t_{\rm SU} \propto M_{\rm res}^{-1} l_{\rm max}.
+\end{equation}
+which at least gives users some way to estimate compute times.

### 2013-05-23 test_problems/AnisoConduction/make_plots.py
updated anisotropic conduction problem

@@ -1,6 +1,6 @@
 from yt.mods import *
 
-for n in range(300,301):
+for n in range(11,12):
     
     pf = load("DD%04i/DD%04i" % (n,n))
 
@@ -21,7 +21,5 @@ for n in range(300,301):
     #slice = SlicePlot(pf, 'z', "Temperature", width = (1.0, '1'), origin='domain', axes_unit=None)  #.save()
 
     slice.save(mpl_kwargs={'bbox_inches':'tight'})
-
-    #slice.save("testplot.png")
 
 # the end

### 2013-05-23 test_problems/DoubleMachReflection/make_plots.py
updated double mach reflection problem

@@ -1,6 +1,6 @@
 from yt.mods import *
 
-for n in range(103,104):
+for n in range(1,2):
 
     pf = load("DD%04i/DD%04i" % (n,n))
 
@@ -11,6 +11,6 @@ for n in range(103,104):
     slice = SlicePlot(pf, 'z', "Density", center=c, width=((4.0, '1'), (1.0, '1')), origin='domain', axes_unit=('1','1')) 
     slice.set_buff_size(buff)
     #slice.set_window_size(4.8)
-    slice.save()
+    #slice.save()
 
-    #slice.save(mpl_kwargs={'bbox_inches':'tight'})
+    slice.save(mpl_kwargs={'bbox_inches':'tight'})

### 2013-05-23 test_problems/test_problem_runner.py
Added first draft of script to run test problems.

@@ -1,0 +1,135 @@
+"""
+This script makes all of the figures in the 'Representative test problems' section
+of the Enzo method paper (Section 11.2).  It requires you to have built an Enzo executable,
+named enzo.exe, ***using the HYPRE library***.  If you do NOT use the HYPRE library,
+remove the 'CosmoIFront' problem from the testnames list below.  You also need to set
+two paths (edit the variables directly below this comment block), and you can turn on
+and off some options.  
+
+Note that running this script can take quite a bit of time - some of the simulations
+take a while to run (hours).  If you're worried that it's not doing anything, go into
+the simulation directory for the simulation that is currently running (which will be
+printed out by the script) and look to see when the file output.log was last modified.
+This file contains stdout and stderr from the simulation.
+
+    --Brian O'Shea, oshea@msu.edu, May 2013
+    
+"""
+import os as os
+
+############ USER SETS THESE PARAMETERS #############
+
+# directory where the enzo.exe binary is stored
+enzodir = '/Users/bwoshea/Desktop/SYNCHED/Current\ Projects/Papers/enzo-method-paper/test_problems'
+
+# directory where this file is located (need absolute path name)
+topleveldir = '/Users/bwoshea/Desktop/SYNCHED/Current Projects/Papers/enzo-method-paper/test_problems'
+
+# the following options allow you to articulate whether you want the simulations to be run (runsim),
+#   figures to be made (makefig), and directories to be cleaned out (cleandata/cleanfigs).  Note that 
+#   this is so you can separately run the various simulations, make the figures, and/or clean out all
+#   of the various subdirectories.
+
+# if cleandata == 1, delete all simulation files in the directory.  If 0, don't touch them.
+#    Note: output log is always deleted before a simulation is run.  Also, if you clean out
+#    the data, you can't make figures without rerunning the simulations!
+cleandata = 1
+
+# if cleanfigs == 1, delete all figures in the directory.  If 0, don't touch them.
+cleanfigs = 1
+
+# if runsim > 0, run simulations; if 0, don't.
+runsim = 1
+
+# if makefig > 0, make figures.  if 0, don't.  Figures require simulation data!
+makefig = 1
+
+############ END OF USER-SET PARAMETERS #############
+
+# do some preemptive cleaning of directory names
[…]

### 2013-05-23 test_problems/AnisoConduction/make_plots.py
updating more plots

@@ -20,6 +20,8 @@ for n in range(11,12):
     slice = SlicePlot(pf, 'z', "Temperature", width = (1.0, '1'), origin='domain', axes_unit=('1','1'))  #.save()
     #slice = SlicePlot(pf, 'z', "Temperature", width = (1.0, '1'), origin='domain', axes_unit=None)  #.save()
 
-    slice.save(mpl_kwargs={'bbox_inches':'tight'})
+    filename = "DD%04i" % n
+
+    slice.save(mpl_kwargs={'bbox_inches':'tight'}, name="%s_slice_Temperature_z.eps" % filename)
 
 # the end

### 2013-05-23 test_problems/DoubleMachReflection/make_plots.py

@@ -1,6 +1,6 @@
 from yt.mods import *
 
-for n in range(1,2):
+for n in range(11,12):
 
     pf = load("DD%04i/DD%04i" % (n,n))
 
@@ -10,7 +10,7 @@ for n in range(1,2):
 
     slice = SlicePlot(pf, 'z', "Density", center=c, width=((4.0, '1'), (1.0, '1')), origin='domain', axes_unit=('1','1')) 
     slice.set_buff_size(buff)
-    #slice.set_window_size(4.8)
-    #slice.save()
 
-    slice.save(mpl_kwargs={'bbox_inches':'tight'})
+    filename = "DD%04i" % n
+
+    slice.save(mpl_kwargs={'bbox_inches':'tight'},name="%s_slice_Density_z.eps" % filename)

### 2013-05-23 test_problems/test_problem_runner.py

@@ -39,10 +39,10 @@ cleandata = 1
 cleanfigs = 1
 
 # if runsim > 0, run simulations; if 0, don't.
-runsim = 1
+runsim = 0
 
 # if makefig > 0, make figures.  if 0, don't.  Figures require simulation data!
-makefig = 1
+makefig = 0
 
 ############ END OF USER-SET PARAMETERS #############
 
@@ -62,7 +62,7 @@ testnames = ['AnisoConduction', 'CosmoIFront', 'DoubleMachReflection',
             'WavePool', 'ZeldovichPancake']
 """
 #testnames = ['ZeldovichPancake',  'OneZoneFreefallTest']
-testnames =  ['DoubleMachReflection']
+testnames =  ['GravityTest']
 
 # goes through directory dirname and its subdirs and makes list of directories that
 #   contains enzo parameter files (which end in '.enzo')

### 2013-05-23 test_problems/test_problem_runner.py
updating plots

@@ -62,7 +62,7 @@ testnames = ['AnisoConduction', 'CosmoIFront', 'DoubleMachReflection',
             'WavePool', 'ZeldovichPancake']
 """
 #testnames = ['ZeldovichPancake',  'OneZoneFreefallTest']
-testnames =  ['GravityTest']
+testnames =  ['SelfSimilarInfall']
 
 # goes through directory dirname and its subdirs and makes list of directories that
 #   contains enzo parameter files (which end in '.enzo')
@@ -100,7 +100,7 @@ for simdir in all_sim_dirs:
     os.chdir(simdir)  # go into simulation directory
      
     if cleandata == 1:
-        os.system('rm -rf DD* Enzo_Build* Evtime *out *log *Log RunFinished')
+        os.system('rm -rf DD* Enzo_Build* Evtime *out *log *Log RunFinished SphericalInfallReport')
         
     if runsim > 0:  # now we actually run the simulation!
         dirlisting = os.listdir('.')  # get listing of all of the files in the directory

### 2013-05-23 test_problems/TestOrbit/make_fig_plot.py
updating TestOrbit.

@@ -1,116 +1,0 @@
-#
-#  This script makes most of the plots for the TestOrbit problem type.  The
-#   one that is not made is the total energy plot, which is in another file
-#   for the sake of clarity.
-#
-# Note: this script assumes enzo has been run with "enzo.exe TestOrbit.exe > TestOrbit.out"
-#   (since we need some stuff from TestOrbit.out)
-#
-import numpy as na
-import matplotlib.pyplot as plt
-import os
-from math import *
-
-FileIn = open("TestOrbit.out", "r")
-
-xpos = []
-ypos = []
-zpos = []
-xvel = []
-yvel = []
-zvel = []
-
-# read stuff in from file
-for line in FileIn:
-   if 'id=1' in line:
-      lst = line.split()
-      xpos.append(lst[1])
-      ypos.append(lst[2])
-      zpos.append(lst[3])
-      xvel.append(lst[4])
-      yvel.append(lst[5])
-      zvel.append(lst[6])
-      
-FileIn.close()
-
-# convert to numpy array: easier to manipulate
-x = na.array(xpos)
-y = na.array(ypos)
-z = na.array(zpos)
-xv = na.array(xvel)
-yv = na.array(yvel)
-zv = na.array(zvel)
-
-rad_diff = []
-counter = []
-energy = []
-
-diffweight=0.0
-radsum=0.0
-energysum=0.0
-
-# go through and calculate radius error, also kinetic energy
-for i in range(len(x)):
-   radius = ( ( float(x[i])-0.5)**2 + ( float(y[i])-0.5)**2 + ( float(z[i])-0.5)**2 )**0.5
-   raddiff = (radius-0.3)/0.03125 
-   radsum = radsum + raddiff
-   diffweight = diffweight + 1.0
-   kinetice = 0.5*(float(xv[i])**2 + float(yv[i])**2 + float(zv[i])**2)
-   vmag = (float(xv[i])**2 + float(yv[i])**2 + float(zv[i])**2)**0.5
-   rad_diff.append(raddiff)
-   energy.append(kinetice)
-   energysum += kinetice
-   counter.append(i)
-
-radsum = radsum/diffweight
-energysum /= diffweight
-
-#print "radsum is ", radsum
-#print "energysum is ", energysum
-
-# tweak radius to mean
-rd = na.array(rad_diff)-radsum
-# convert counter to orbit number
[…]

### 2013-05-23 test_problems/test_problem_runner.py

@@ -62,7 +62,7 @@ testnames = ['AnisoConduction', 'CosmoIFront', 'DoubleMachReflection',
             'WavePool', 'ZeldovichPancake']
 """
 #testnames = ['ZeldovichPancake',  'OneZoneFreefallTest']
-testnames =  ['SelfSimilarInfall']
+testnames =  ['TestOrbit']
 
 # goes through directory dirname and its subdirs and makes list of directories that
 #   contains enzo parameter files (which end in '.enzo')

### 2013-05-23 test_problems/test_problem_runner.py
working on Sedov

@@ -62,7 +62,7 @@ testnames = ['AnisoConduction', 'CosmoIFront', 'DoubleMachReflection',
             'WavePool', 'ZeldovichPancake']
 """
 #testnames = ['ZeldovichPancake',  'OneZoneFreefallTest']
-testnames =  ['TestOrbit']
+testnames =  ['OneZoneFreefallTest']
 
 # goes through directory dirname and its subdirs and makes list of directories that
 #   contains enzo parameter files (which end in '.enzo')

### 2013-05-24 test_problems/SedovAMR/sedov_outputdata.py
updating sedov test

@@ -1,34 +1,68 @@
-:from yt.mods import *
+# Calculates radial profiles for simulations, writes into
+# files sedov_ppm_output.dat and sedov_zeus_output.dat
+
+from yt.mods import *
 import matplotlib.pyplot as plt
 
-# Load the dataset.
-#pf = load("../100amr4/ppm/DD0007/sb_L2x2_amr4_0007")
-pf = load("../100amr4/zeus/DD0007/sb_L2x2_amr4_zeus_0007")
+################# ZEUS plots
+pf1 = load("Zeus/DD0007/DD0007")
+pc1 = PlotCollection(pf1, center=[0.5, 0.5, 0.5])
+pc1.add_profile_sphere(0.5, "cm", ["Radius", "CellMass"], weight=None, x_bins=100, x_log=False)
 
-pc = PlotCollection(pf, center=[0.5, 0.5, 0.5])
+mass_data = pc1.plots[-1].data["CellMass"]
+radius_data =  pc1.plots[-1].data["Radius"]
 
-pc.add_profile_sphere(0.5, "cm", ["Radius", "CellMass"], weight=None, x_bins=100, x_log=False)
-mass_data = pc.plots[-1].data["CellMass"]
-radius_data =  pc.plots[-1].data["Radius"]
-
-pc.add_profile_sphere(0.5, "cm", ["Radius", "CellVolume"], weight=None, x_bins=100, x_log=False)
-volume_data = pc.plots[-1].data["CellVolume"]
+pc1.add_profile_sphere(0.5, "cm", ["Radius", "CellVolume"], weight=None, x_bins=100, x_log=False)
+volume_data = pc1.plots[-1].data["CellVolume"]
 
 density_data = mass_data/volume_data
 
-pc.add_profile_sphere(0.5, "cm", ["Radius", "RadialVelocity"], weight="CellMass", x_bins=100, x_log=False)
-velocity_data = pc.plots[-1].data["RadialVelocity"]
+pc1.add_profile_sphere(0.5, "cm", ["Radius", "RadialVelocity"], weight="CellMass", x_bins=100, x_log=False)
+velocity_data = pc1.plots[-1].data["RadialVelocity"]
 
-pc.add_profile_sphere(0.5, "cm", ["Radius", "Pressure"], weight="CellMass", x_bins=100, x_log=False)
-pressure_data = pc.plots[-1].data["Pressure"]
[…]

### 2013-05-24 test_problems/SedovAMR/sedov_slice.py

@@ -1,13 +1,12 @@
+# Makes slices of density for both simulations (PPM, Zeus)
+
 from yt.mods import * 
 
-#fn = "../100amr4/ppm/DD0007/sb_L2x2_amr4_0007"
-fn = "../100amr4/zeus/DD0007/sb_L2x2_amr4_zeus_0007"
+fn1 = "PPM/DD0007/DD0007"
+pf1 = load(fn1)
+slice1 = SlicePlot(pf1, 'z', "Density", center=[0.5, 0.5, 0.0]).save("PPM_Density_slice.eps")
 
-pf = load(fn)
-pc = PlotCollection(pf, center=[0.5, 0.5, 0.5])
-
-p = pc.add_slice("Density", 2)
-#p.modify["grids"]()
-
-pc.save(fn)
+fn2 = "Zeus/DD0007/DD0007"
+pf2 = load(fn2)
+slice2 = SlicePlot(pf2, 'z', "Density", center=[0.5, 0.5, 0.0]).save("Zeus_Density_slice.eps")
 

### 2013-05-24 test_problems/test_problem_runner.py

@@ -1,7 +1,8 @@
 """
 This script makes all of the figures in the 'Representative test problems' section
 of the Enzo method paper (Section 11.2).  It requires you to have built an Enzo executable,
-named enzo.exe, ***using the HYPRE library***.  If you do NOT use the HYPRE library,
+named enzo.exe, ***using the HYPRE library (http://acts.nersc.gov/hypre/)***, and also yt built
+with SciPy.  If you do NOT use the HYPRE library, or do not have SciPy installed along with yt, 
 remove the 'CosmoIFront' problem from the testnames list below.  You also need to set
 two paths (edit the variables directly below this comment block), and you can turn on
 and off some options.  

### 2013-05-25 test_problems/test_problem_runner.py
Added README file, updated test_problem_runner.py.

@@ -1,17 +1,24 @@
 """
-This script makes all of the figures in the 'Representative test problems' section
-of the Enzo method paper (Section 11.2).  It requires you to have built an Enzo executable,
-named enzo.exe, ***using the HYPRE library (http://acts.nersc.gov/hypre/)***, and also yt built
-with SciPy.  If you do NOT use the HYPRE library, or do not have SciPy installed along with yt, 
-remove the 'CosmoIFront' problem from the testnames list below.  You also need to set
-two paths (edit the variables directly below this comment block), and you can turn on
-and off some options.  
+         Enzo method paper figure generator
 
-Note that running this script can take quite a bit of time - some of the simulations
-take a while to run (hours).  If you're worried that it's not doing anything, go into
-the simulation directory for the simulation that is currently running (which will be
-printed out by the script) and look to see when the file output.log was last modified.
-This file contains stdout and stderr from the simulation.
+This script makes all of the figures in the 'Representative test
+problems' section of the Enzo method paper (Section 11.2).  It
+requires you to have compiled an Enzo executable, named enzo.exe,
+***using the HYPRE library (http://acts.nersc.gov/hypre/)***, and also
+have yt (http://yt-project.org) built with SciPy, which requires a
+non-default option to be set in the install script.  If you do NOT
+compile Enzo with the HYPRE library, or do not have SciPy installed
+along with yt, remove the 'CosmoIFront' problem from the testnames
+list below.  You also need to set two paths -- the location of the
+Enzo binary and the path to the test problem directory -- and you
+can turn on and off some options.  This is all done directly below
+this comment block.
+
+Note: some of the simulations used to create the method paper figures
+can take several hours to run, or should be run on several CPUs. See
+the README file in this directory for specifics.  You can change the
+simulations that are run with this script by editing the variable
+'testnames', below.
 
     --Brian O'Shea, oshea@msu.edu, May 2013
     
@@ -20,10 +27,10 @@ import os as os
 
 ############ USER SETS THESE PARAMETERS #############
 
-# directory where the enzo.exe binary is stored
[…]

### 2013-05-25 test_problems/PhotonShadowing/make_plots.py
tweaked a plot to write out eps.

@@ -19,4 +19,4 @@ for i,f in enumerate(fields):
 ep = EPS.multiplot_yt(2,2,pc, margins=(0.2,0.2), bare_axes=True)
 ep.scale_line(1.0/6.6, '1 kpc')
 ep.title_box("15 Myr")
-ep.save_fig("code-test-shadowing")
+ep.save_fig("code-test-shadowing.eps")

### 2013-05-26 tests-aniso-conduct.tex
Edited subsections 11.2.11 through 11.2.14.

@@
 \textit{along} magnetic field lines, not perpendicular to them, which
-can result in very interesting magnetohydrodynamical instabilities
-\citep[e.g.,][]{2008ApJ...677L...9P,2008ApJ...688..905P}.  The correct
+can result in very interesting magnetohydrodynamical instabilities \protect \citep[e.g.,][]{2008ApJ...677L...9P,2008ApJ...688..905P}. The correct
 modeling of this behavior can be quite challenging, and is described
-in Section~\ref{sec.num.conductions}.  Conduction in \enzo\ is
-calculated in both an operator-split and directionally-split manner,
+in Section~\protect\ref{sec.num.conductions}. Conduction in \protect\enzo\ is calculated in both an operator-split and directionally-split manner,
 and the addition of heat transport solely along magnetic field lines
@@
 field in regions where the temperature gradient is strong in more than
-one spatial direction unless an appropriate flux-limiter \citep[such
-as that of][]{1977JCoPh..23..263V} is chosen.
+one spatial direction unless an appropriate flux-limiter \protect\citep[such as that of][]{1977JCoPh..23..263V} is chosen.
+\mqk{The previous paragraph more or less repeats what's in Section~\ref{sec.num.conductions}, I think it should be removed (or merged with that section).}
 
@@
 two-dimensional, $256 \times 256$ cell simulation having a physical
-scale of 1 kpc on a side, a uniform density of 1 proton/cc, and a
+scale of 1 kpc on a side, a uniform density of 1 proton cm$^{-3}$ and a
 background temperature of $10^6$ Kelvin.  Magnetic fields with a
@@
 hydrodynamics, radiative cooling, or cosmological expansion), and with
-a Spitzer fraction of f$_{sp} = 1$.
+a Spitzer fraction of f$_{\rm sp} = 1$.
 
 The right panel of Figure~\ref{fig.conduct} shows the state of the
-simulation after 300 Myr.  Heat has clearly been transported only along
-field lines -- there has been no diffusion perpendicular to the
+simulation after 300 Myr.  Heat has clearly been transported only
+along field lines -- there has been no diffusion perpendicular to the
 magnetic field setup, which is critical for many studies involving
@@
 temperature field in regions where the fields are not aligned with the
-grid, suggesting that the flux-limiter is operating as expected.
+grid, suggesting that the flux-limiter is operating as expected (see
+discussion in Section~\ref{sec.num.conductions}).
 
@@
 constant temperature background with circular magnetic fields centered
-on (0.5, 0.5) in a $256 \times 256$ grid.  The background medium has a
-density of 1 proton/cc and a temperature of $10^6$~K.  At t$ = 0$
-(left panel), a Gaussian heat pulse is injected at (0.75, 0.5) with a
-FWHM of $1/64$ (with all numbers given in units of the box size) and a
-peak temperature of $10^8$~K, and allowed to evolve without
-hydrodynamical motion (i.e., static gas) and no radiative cooling for
-300 Myr.  At t$ = 300$~Myr (right panel), heat has been transported
-along magnetic field lines with no significant diffusion perpendicular
-to field lines. Furthermore, there are no detectable oscillations in
-the temperature in regions where the magnetic field is not parallel
-with the grids.}
+on (0.5, 0.5) in a $256 \times 256$ grid. \mqk{Maybe show some field
+lines in the figure?} The background medium has a density of 1
+proton/cc and a temperature of $10^6$~K.  At t$ = 0$ (left panel), a
+Gaussian heat pulse is injected at (0.75, 0.5) with a FWHM of $1/64$
+(with all numbers given in units of the box size) and a peak
+temperature of $10^8$~K, and allowed to evolve without hydrodynamical
+motion (i.e., static gas) and no radiative cooling for 300 Myr.  At t$
+= 300$~Myr (right panel), heat has been transported along magnetic
+field lines with no significant diffusion perpendicular to field
+lines. Furthermore, there are no detectable oscillations in the
+temperature in regions where the magnetic field is not parallel with
+the grids.}
 \label{fig.conduct}

### 2013-05-26 tests-onezonecollapse.tex
Edited subsections 11.2.11 through 11.2.14.

@@
 \begin{equation}
-\frac{d\rho}{dt} = \frac{\rho}{t_{col}},
+\frac{d\rho}{dt} = \frac{\rho}{t_{\rm col}},
 \end{equation}
 
-where the collapse timescale, $t_{col}$, is
+where the collapse timescale, $t_{\rm col}$, is
 
 \begin{equation} \label{eqn.tcol}
-t_{col} = \frac{t_{dyn}}{\sqrt{1 - f}},
+t_{\rm col} = \frac{t_{\rm dyn}}{\sqrt{1 - f}},
 \end{equation}
-In this equation, $t_{dyn}$ is the dynamical time and is expressed as
+In this equation, $t_{\rm dyn}$ is the dynamical time and is expressed as
 \begin{equation}
-t_{dyn} = \sqrt{\frac{3 \pi}{32 G \rho}}.
+t_{\rm dyn} = \sqrt{\frac{3 \pi}{32 G \rho}}.
 \end{equation}
 
-with G being the gravitational constant.  The collapse timescale is
-altered from the dynamical time by a factor $f$ in Equation
+with G being the gravitational constant.
+\mqk{Instead of explaining what $G$ is here, would it make sense to have a table at the beginning of the paper listing all the natural constants and variable symbols that are used throughout the paper?} The collapse timescale is
+altered from the dynamical time by a factor $1/\sqrt{1-f}$ in Equation
 \ref{eqn.tcol}, which is an approximation of the ratio of the gas
 pressure to the force of gravity.  The value of $f$ depends on the
-effective adiabatic index, $\gamma_{ef} \equiv (\partial ln\ p
-/ \partial ln\ \rho)$, which we calculate by linearly extrapolating
-from derivative values for the two previous timesteps.  For the value
+effective adiabatic index, $\gamma_{\rm ef} \equiv (\partial \ln p
+/ \partial \ln \rho)$, which we linearly extrapolate
+from derivative values at the two previous timesteps.  For the value
 of $f$ in this test problem, we use the piecewise function of
@@
   \begin{array}{ll}
-  0, & \gamma_{ef} < 0.83,\\
-  0.6 + 2.5 (\gamma_{ef} - 1) - 6.0 (\gamma_{ef} - 1)^{2}, & 0.83 <
-  \gamma_{ef} < 1,\\
-  1.0 + 0.2 (\gamma_{ef} - 4/3) - 2.9 (\gamma_{ef} - 4/3)^{2}, & \gamma_{ef} > 1.
+  0, & \gamma_{\rm ef} < 0.83,\\
+  0.6 + 2.5 (\gamma_{\rm ef} - 1) - 6.0 (\gamma_{\rm ef} - 1)^{2}, & 0.83 <
+  \gamma_{\rm ef} < 1,\\
+  1.0 + 0.2 (\gamma_{\rm ef} - 4/3) - 2.9 (\gamma_{\rm ef} - 4/3)^{2}, & \gamma_{\rm ef} > 1.
 \end{array} \right.
@@
 of metal cooling can be clearly seen; as the metallicity increases
-from zero to $10^{-2}$~Z$_\odot$, the gas deviates from the primoridal
-result (black line) both more rapidly and more significantly.  Our
-primordial results compare very well to those shown in
-\citet{2005ApJ...626..627O}; however, we use a different cooling
-method than Omukai, so the lines describing the evolution of the
+from zero to $10^{-2}$~Z$_\odot$, the gas rapidly and significantly
+deviates from the primordial result (black line).  Our primordial
+results compare very well to those shown in
+\citet{2005ApJ...626..627O}; however, we use a different metal cooling
+method, so the lines describing the evolution of the
 metal-enriched gas are not directly comparable.
@@
 calculated with the \texttt{Cloudy} code, and compares favorably to
-Omukai's results.}
+the results of \citet{2005ApJ...626..627O}.}
   \label{fig.onezone}

### 2013-05-26 tests-photoevaporation.tex
Edited subsections 11.2.11 through 11.2.14.

@@
 hydrodynamics simulations, and this test problem examines the
-ionization front propagation into the dense clump, shadowing effects
+ionization front propagation into a dense clump, shadowing effects
 behind the clump, and the hydrodynamic response on the clump from
@@
 \citet{Wise11_Moray}.  The simulation domain is 6.6~kpc on a side with
-an ambient medium of pure neutral hydrogen $n_{\rm H} = 2 \times
-10^{-4}\; \cubecm$ and $T = 8000 \unit{K}$.  We place a spherical
-overdensity in hydrostatic equilibrium with the ambient medium.  It
-has a radius $r = 0.8 \unit{kpc}$, hydrogen density $n_{\rm H} = 0.04
-\cubecm$, and temperature $T = 40$ K, and it is centered at $(x,y,z) =
-(5, 3.3, 3.3) \unit{kpc}$.  In \citet{IlievEtAl2009}, all of the codes
-used a fixed $128^3$ grid to ease the comparison, but in this test to
-demonstrate a higher resolution AMR solution, we employ a $128^3$ grid
-with two additional levels of refinement with cells with a baryon mass
-(method 2 in Section~\ref{sec:refinement_criteria}) greater than 1.5
-being flagged for refinement.  This test is run for 15 Myr.
+an ambient medium of pure neutral hydrogen of density $n_{\rm H} =
+2 \times 10^{-4}\; \cubecm$ and temperature $T = 8000 \unit{K}$.  We
+place a spherical overdensity in hydrostatic equilibrium with the
+ambient medium.  It has a radius $r = 0.8 \unit{kpc}$, hydrogen
+density $n_{\rm H} = 0.04\ \cubecm$, and temperature $T = 40$ K, and
+is centered at $(x,y,z) = (5, 3.3, 3.3) \unit{kpc}$.
+In \citet{IlievEtAl2009} all of the codes used a fixed $128^3$ grid to
+ease the comparison, but in this test to demonstrate a higher
+resolution AMR solution, we employ a $128^3$ grid with two additional
+levels of refinement for cells with a baryon mass greater than 1.5
+(method 2 in Section~\ref{sec:refinement_criteria}). This test is run
+for 15 Myr.
 
@@
 a full spectrum and energy discretization \citep{Mirocha12}.
-\citep[Note that this choice of energy groups is different
-from][]{Wise11_Moray}.  We use a minimum angular resolution of 10 rays
+(Note that this choice of energy groups is different from Wise \& Abel 2011.) \nocite{Wise11_Moray}  We use a minimum angular resolution of 10 rays
 per cell and a constant radiative transfer timestep of 25 kyr.  Figure
-\ref{fig:shadowing} depicts the clump after at $t = 15$ Myr with the
+\ref{fig:shadowing} depicts the clump at $t = 15$ Myr with the
 outer layers expanding after being photo-heated.  It also shows the
@@
   \caption{Photo-evaporation of a dense clump using the Moray
-ray-tracing module.  Clockwise from the upper left corner: slices
-through the clump center of the density, temperature, neutral hydrogen
+ray-tracing module.  Clockwise from the upper left: slices
+through the clump center of density, temperature, neutral hydrogen
 fraction, and pressure at $t=15$~Myr after the initialization of the

### 2013-05-28 tests-orbit.tex
Minor revision to orbit test section.

@@
 one standard deviation by horizontal dotted lines.  All quantities are
-in arbitrary units.}
+in dimensionless units.}
 \label{fig.orbittest}
@@
 The ``orbit test'' simulates the orbit of an effectively massless
-particle around a massive primary particle, and is used to test the
+particle around a massive primary particle and is used to test the
 accuracy of both the gravity solver and particle integration.  In this
@@
 the center of the box (both box length and particle mass are in
-arbitrary units).  A second ``test'' particle, with mass of $10^{-6}$,
+dimensionless units).  A second ``test'' particle, with mass of $10^{-6}$,
 is placed at a distance of 0.3 L$_{box}$ from the primary particle in
 the x-y plane (at $z=0.5$~L$_{box}$).  Both particles are given
-velocities such that they will orbit their common centers of mass in
-the x-y plane.  The gravitational potential is calculated using a fast
-Fourier Transform on the $32^3$ grid, and the simulation is integrated
-forward in time for 200 particle orbits, writing out a dataset
+velocities such that they will orbit their common center of mass in
+the x-y plane.  The simulation is run for 200 orbits, writing
+out a dataset 
 approximately once per orbit (with the particle positions being
@@
 test particle in the x-y plane (at z=$0.5$), plotted over the duration
-of the simulation (approximately 200 orbits).  The test particle
+of the simulation.  The test particle
 maintains a path that has effectively unnoticeable deviation from
@@
 using a second-order leapfrog method, this level of accuracy is
-roughly what one might expect.
+expected.

### 2013-05-28 tests-orbit.tex
Minor revision to orszag-tang section.

@@
 one standard deviation by horizontal dotted lines.  All quantities are
-in dimensionless units.}
+in arbitrary units.}
 \label{fig.orbittest}
@@
 the center of the box (both box length and particle mass are in
-dimensionless units).  A second ``test'' particle, with mass of $10^{-6}$,
+arbitrary units).  A second ``test'' particle, with mass of $10^{-6}$,
 is placed at a distance of 0.3 L$_{box}$ from the primary particle in

### 2013-05-28 numerical-cooling.tex
Added most of Dave Collins' suggested edits (except for a bunch of stuff having to do with the hydro).
We definitely use Glover & Abel 2008 by default for H2, and the code suggests we use Lepp & Shull as well as GA98 for HD.

@@
 
-\enzo\ has multiple methods for computing the energy change from
-radiative cooling and heating.  All of them assume the gas is totally
-optically thin.  Below we describe the methods for computing the
-cooling rates from metal-free and metal-enriched gas.  Sample cooling
-curves for each of \enzo's primary cooling methods are shown in Figure
-\ref{fig.cooling_rate}.
+\enzo\ has multiple methods for computing the energy change from radiative
+cooling and heating.  All of them assume the gas can be modeled either as
+completely optically thin or with a simple, local approximation to optical
+thickness.  Below we describe the methods for computing the cooling rates from
+metal-free and metal-enriched gas.  Sample cooling curves for each of \enzo's
+primary cooling methods are shown in Figure \ref{fig.cooling_rate}.
 
@@
 characterize a metal-free gas is simple enough to be computed in
-non-equilibrium during the simulation.  Similarly, the radiative
+non-equilibrium during even very simulations.  Similarly, the radiative
 cooling of metal-free gas is solved by directly computing the cooling
@@
 cooling processes are also considered: ro-vibrational transitions
-\citep{1998A&A...335..403G}, heating and cooling from molecular
+\citep{2008MNRAS.388.1627G,1998A&A...335..403G}, heating and cooling from molecular
 formation and destruction \citep{2009Sci...325..601T}, and
@@
 chemistry is enabled, then rotational transitions of HD
-\citep{1998A&A...335..403G} are treated as well.  The radiative
-cooling calculation is coupled to the update of the chemistry network
-such that they both occur within the same subcycling loop.  This is
-necessary in regimes where rapid cooling or change in the ionization
-state occur.  In addition to the subcycle timestepping contraints
-mentioned in Section~\ref{sec.num.chemistry}, the subcycle timestep is also not
-permitted to exceed 10\% of the cooling time, $t_{cool} = e/\dot{e}$.  A
-metagalactic background affects the gas through both photo-heating and
-photo-ionization.  These are treated by including redshift-dependent
-photo-ionization and photo-heating rate terms in the chemistry and
-cooling equations for \ion{H}{1}, \ion{He}{1}, and \ion{He}{2}.  The
-black curves in Figure \ref{fig.cooling_rate} show cooling rates for a
-metal-free gas with number density $n_{\rm H} = 10^{-4}$ cm$^{-3}$ both
-with and without a radiation background.  More detail on the specific
-UV backgrounds present in \enzo is presented below in
-Section~\ref{sec.num.rad-homogeneous}.
+\citep{1998A&A...335..403G, 1983ApJ...270..578L} are treated as well.  The
+radiative cooling calculation is coupled to the update of the chemistry network
+such that they both occur within the same subcycling loop.  This is necessary
+in regimes where rapid cooling or change in the ionization state occur, as this
+will influence the kinetic rate coefficients through both changes in energy and
+the equation of state of the gas.  In addition to the subcycle timestepping
+contraints mentioned in Section~\ref{sec.num.chemistry}, the subcycle timestep
+is also not permitted to exceed 10\% of the cooling time, $t_{cool} =
+e/\dot{e}$.  A metagalactic background affects the gas through both
+photo-heating and photo-ionization.  These are treated by including
+redshift-dependent photo-ionization and photo-heating rate terms in the
+chemistry and cooling equations for \ion{H}{1}, \ion{He}{1}, and \ion{He}{2}.
+The black curves in Figure \ref{fig.cooling_rate} show cooling rates for a
+metal-free gas with number density $n_{\rm H} = 10^{-4}$ cm$^{-3}$ both with
+and without a radiation background.  More detail on the specific UV backgrounds
+present in \enzo\ is presented below in Section~\ref{sec.num.rad-homogeneous}.
 

### 2013-05-28 numerical-cooling.tex
Change a bit of wording and usages of macros.

@@
 characterize a metal-free gas is simple enough to be computed in
-non-equilibrium during even very simulations.  Similarly, the radiative
+non-equilibrium during even very large simulations.  Similarly, the radiative
 cooling of metal-free gas is solved by directly computing the cooling
@@
 background (CMB), and photo-heating from a UV metagalactic background.
-If the H$_{2}$ chemistry network is enabled, the following H$_{2}$
+If the \HH chemistry network is enabled, the following \HH
 cooling processes are also considered: ro-vibrational transitions
@@
 in regimes where rapid cooling or change in the ionization state occur, as this
-will influence the kinetic rate coefficients through both changes in energy and
-the equation of state of the gas.  In addition to the subcycle timestepping
-contraints mentioned in Section~\ref{sec.num.chemistry}, the subcycle timestep
-is also not permitted to exceed 10\% of the cooling time, $t_{cool} =
-e/\dot{e}$.  A metagalactic background affects the gas through both
+will influence the chemical kinetic rate coefficients through both changes in
+energy and the equation of state of the gas.  In addition to the subcycle
+timestepping contraints mentioned in Section~\ref{sec.num.chemistry}, the
+subcycle timestep is also not permitted to exceed 10\% of the cooling time,
+$t_{cool} = e/\dot{e}$.  A metagalactic background affects the gas through both
 photo-heating and photo-ionization.  These are treated by including
@@
 the metal species only.  The primary assumption made is that of
-ionization equilibrium.  The tables can have up to five dimensions,
-where those dimensions are density, metallicity, electron fraction,
+ionization equilibrium.  The tables can vary along up to five dimensions,
+of density, metallicity, electron fraction,
 redshift (for an evolving metagalactic UV background), and
@@
 elements up to atomic number 30 (Zn) and for any incident radiation
-field.  The cooling from H and He is calculated by the non-equilibrium
-primordial chemistry and cooling solver.  The contribution of metals
+field.  Cooling from the standard \enzo\ non-equilibrium cooling module is
+applied on top of the metal contributions.  The contribution of metals
 to the cooling is computed within the same subcycling loop as the

### 2013-05-29 introduction.tex
minor changes to gravity section

--- 2013-05-17 introduction.tex
@@
 
-Many phenomena in astrophysics and cosmology require a high spatial and temporal dynamical range, and there are a number of numerical techniques that have been used to model such problems.
-At the current time, the most common is a gridless, fully Lagrangian method call Smoothed Particle Hydrodynamcs \citep[SPH;][]{Lucy77, SPH} which has been very successful -- particularly for those situations dominated by gravity, where systems evolve over large spatial and temporal dynamic ranges.
-Its development, however, is still at a relatively early stage when compared to the effort put into Eulerian, grid-based hydrodynamic schemes \citep[e.g.,][]{laney-1998, toro-1997, Woodward84}.
+Due to the high spatial and temporal dynamical ranges involved, astrophysics and cosmological phenomena present a taxing challenge for simulators.
+To tackle these, a number of numerical techniques have been developed which can be broadly split into the gridlesss,fully Lagrangian methods and grid-based, Eulerian schemes.
+The most commonly used is an example of the first typeand is known as Smoothed Particle Hydrodynamics \citep[SPH;][]{Lucy77, SPH}.
+It has achived much success, particularly in regimes dominated by gravity.
+%where systems evolve over large spatial and temporal dynamic ranges (cut because we've already said that in the first sentense - EJT)
+However, its development to include an increasing number of sort-after physical processes is still at a relatively early stage when compared with the effort put into the latter type of Eulerian grid-based hydrodynamic schemes \citep[e.g.,][]{laney-1998, toro-1997, Woodward84}.
+% is that what was meant by `` Its development, however, is still at a relatively early stage''? Since SPH is older than AMR, I found this a little confusing as written.
 
-Therefore, we would like to make use of the expertise that has been invested in Eulerian solvers; unfortunately, as originally developed, none of these schemes provide an easy way to adaptively and dynamically increase the spatial and temporal resolution in small volumes, as required by, for example, gravitational instability.
-We have instead adopted an idea first suggested by \citet{Berger89} in the Computational Fluid Dynamics (CFD) community, which has become known as Structured Adaptive Mesh Refinement (SAMR).
-The principle is to adaptively add and modify additional, finer meshes (``patches'') over regions that require higher resolution.
-Since each patch has a simple geometry (Cartesian, in our case) the hyperbolic fluid-dynamical equations can be solved on that patch with a well-developed `off-the-shelf' numerical method.
-In addition, it is relatively straightforward to add other advanced physics (e.g.
-magnetohydrodynamics, radiation transport, chemistry, etc.).
-This paper is devoted to describing how we have implemented this idea for astrophysics, including the addition of comoving coordinates, self-gravity, radiative cooling, chemistry, heat conduction, collisionless fluids, MHD, radiation transport, star formation and a range of other physical effects.
+Despite this invested expertise, in their original form the Eulerian solvers have a serious drawback: they do not provide an easy method of adaptively increasing the spatial and temporal resolution in small volumes of the simulation.
+Such flexibility is essential for following physical processes such as gravitational instability.
+%Yet, despite the expertise invested in Eulerian solvers, in their original form they do not provide an easy method of adaptively increasing the spatial and temporal resolution in small volumes of the simulation.
+Such flexibility is essential for following physical processes such as gravitational instability.
+A solution to this was first proposed by \citet{Berger89} in the Computational Fluid Dynamics (CFD) community, and became known as Structured Adaptive Mesh Refinement (SAMR).
+The pricipal is to adaptively add and modify additional, finer meshes (``grids'') over regions that require higher resolution.
+In addition, it is possible to add other more advanced physics including --for the implementation in the astrophysics code presented in this paper-- comoving coordinates, self-gravity, radiative cooling, chemistry, heat conduction, collisionless fluids, magnetohydrodynamics, radiation transport, star formation and a range of other physical effects.
+% I was concerned the original text sounded a little like this was a brand new technique in a brand new code (despite the references below) which would be confusing to a reader familiar with the field and in particular, Enzo.
 
-There have been many other numerical methods described in the astronomical literature which contain elements of this idea, or have a similar aim.
-For example, an N-body solver \citep{Villumsen89} used non-adaptive meshing to increase the resolution in pre-selected regions.
-This static approach, applied to hydrodynamics, has been used extensively \citep[e.g.,][]{Ruffert94, Anninos94}.
+There have been a number of numerical methods described in the astronomical literature which contain elements of SAMR or have a similar aim.
+For example, the N-body solver developed by \citet{Villumsen89} used non-adaptive meshing to increase the resolution in pre-selected regions.
+This static approach was later used extensively when applied to hydrodynamics \citep[e.g.,][]{Ruffert94, Anninos94}.
 Adding adaptivity is a more recent enhancement, and there are now a number of N-body codes that possess this feature, both with and without  hydrodynamics  \citep{Couchman91, Jessop94, Suisalu95, Splinter96, Gelato97, ART97, Truelove98, flash_method, MLAPM01,  Yahagi01, RAMSES, Quilis04, Ziegler05, Zhang06, Astrobear09, Pluto-amr, GAMER, Nyx}.
@@
 
-\enzo\ is a structured adaptive mesh refinement (SAMR) code that was originally developed for cosmological hydrodynamics and has been used on a wide variety of problems.
-It has been expanded as a general tool for astrophysical fluid dynamics and is intended to be efficient, accurate and easily expanded.
-Although many of the components of the code have been described in previous publications \citep{1995CoPhC..89..149B, BryanThesis96, Bryan97a, Bryan97b, Norman99, BryanCompSci99, Bryan01, Oshea04, 2007arXiv0705.1556N}, there has been no systematic and complete description of the code.
-In this paper we provide that description, filling in many gaps and showing the code's performance for a wide variety of test problems.
+In this paper, we present \enzo\ ; a structured adaptive mesh refinement (SAMR) code.
+Originally developed for cosmological hydrodynamics, \enzo\ has since been used on a wide variety of problems.
+It has grown to become a general tool for astrophysical fluid dynamics and is intended to be efficient, accurate and easily extended to include new capabilities.
+Although many of the components of the \enzo\ have been described in previous publications \citep{1995CoPhC..89..149B, BryanThesis96, Bryan97a, Bryan97b, Norman99, BryanCompSci99, Bryan01, Oshea04, 2007arXiv0705.1556N, WangAbelZhang08,ReynoldsHayesPaschosNorman2009,Collins10,Wise11_Moray}, there has previously been no systematic and complete description of the code.
+In this paper we provide that description, filling in many previous ommissions and showing the code's performance for a wide variety of test problems.
 
-The \enzo\ code has been extensively used over the last two decades in a wide variety of problems, and more than 100 peer-reviewed papers have been published based on results obtained with the code.
-The code has previously been used to study a wide range of astrophysical systems, including galaxy clusters \citep{Loken02}, the interstellar medium \citep{Slyz05}, the intergalactic medium \citep{Fang01}, cooling flows \citep{Li12}, turbulence \citep{Kritsuk04}, and the formation of the first stars \citep{ABN02}.
-
-Finally, we note that the numerical simulation of astronomical phenomena now plays a key role, along with observations and analytic theory, in pushing forward our understanding of the cosmos \citep[e.g.,][]{DecadalSurvey01, DecadalSurvey10}.
+The \enzo\ code has been extensively used over the last two decades in a wide variety of problems, resulting in the publication of more than 100 peer-reviewed papers based on results obtained with the code.
+The variety of astrophysical systems that \enzo\ has been used for include galaxies \citep{2003ApJ...587...13T, 2012ApJ...749..140H}, galaxy clusters \citep{Loken02, Skillman13}, the interstellar medium \citep{Slyz05}, the intergalactic medium \citep{Fang01, 2011ApJ...731....6S}, cooling flows \citep{Li12, 2013ApJ...763...38S}, turbulence \citep{Kritsuk04, 2009JPhCS.180a2020K, Collins11}, the formation of the first stars \citep{ABN02, 2007ApJ...654...66O, 2009Sci...325..601T}, and the formation of stars in our own Galaxy \citep{Collins12a}.
+% 'finally' here sounded strange, since this is really the key point
+Numerical simulations of astronomical phenomena now play a key role, along with observations and analytic theory, in pushing forward our understanding of the cosmos \citep[e.g.,][]{DecadalSurvey01, DecadalSurvey10}.
 But along with this role comes responsibility.
 We believe that those developing simulation tools must fulfill two key obligations: the first is to make those tools available to the community as a whole, much in the way that astronomical data is now regularly made publicly available.
-The second is document, test and refine those methods so that they can be critically evaluated and expanded by others.
+The second is to document, test and refine those methods so that they can be critically evaluated and expanded by others.
 This paper represents one part of our attempt to meet our obligations with respect to the \enzo\ code.

### 2013-05-29 numerical-gravity.tex
minor changes to gravity section

@@
+% EJT: small changes only.
+% dark matter particles -> massive (dark matter and star) particles
+
+% Changed:  ``Particles from sub-grids are also added in this way'' to ``Particles on sub-grids within the grid's volume are also added to its gravitating field using the same method.'' I expanded because I missed the point and had to clarify with Brian when I read the text.
+But if it's too long, please just cut it!
+
+% added: ``Each scheme can be selected via the parameter file.'' I felt it sounded like a dice roll otherwise, but again, feel free to overule me!
+
+% removed blank reference [REF]
+
 \subsection{Gravity}
@@
 
-First, the dark matter particles are distributed onto the grids using the second-order cloud-in-cell (CIC) interpolation technique \citep{Hockney88} to form a spatially-discretized density field $\rho_{\rm DM}$.
-When doing the CIC interpolation, particle positions are (temporarily) advanced by $0.5 \Delta t v^n$ so that we generate an estimate of the time-centered density field.
-Particles on sub-grids are also added in this way.
-In addition, since the gravitating field for a grid is defined beyond the grid edges (see below), dark matter particles from sibling grids and sub grids that lie within the entire gravitating field are used.
+First, the massive (dark matter and star) particles are distributed onto the grids using the second-order cloud-in-cell (CIC) interpolation technique \citep{Hockney88} to form a spatially-discretized density field $\rho_{\rm DM}$.
+During the CIC interpolation, particle positions are (temporarily) advanced by $0.5 \Delta t v^n$ so that we generate an estimate of the time-centered density field.
+Particles on sub-grids within the grid's volume are also added to its gravitating field using the same method.
+In addition, since the gravitating field for a grid is defined beyond the grid edges (see below), massive particles from sibling grids and sub grids that lie within the entire gravitating field are used.
 This step can involve communication.
@@
 In this case, the Greens function is generated in real-space so as to have the correct zero-padding properties and then transformed into the Fourier domain.
-In either case, the potential is then transformed back into the real domain to get potential values at the cell centers.
+In both cases, the potential is then transformed back into the real domain to get potential values at the cell centers.
 These are differenced with a two-point centered difference scheme to obtain accelerations at the cell centers (except if we are using the staggered Zeus-like solver, in which case the accelerations are computed at the cell faces to match the velocities).
@@
 
-In order to calculate more accurate potentials on subgrids, \enzo\ uses a similar but slightly different technique from the root grid.
-The generation of the total gravitating mass field is essentially identical, using CIC interpolation for both the particles and baryons, including sub grids as before.
-To compute the potential on sub grids, however, we use the standard seven-point (in three dimensions) second-order finite-difference approximation to Poisson's equation.
+In order to calculate more accurate potentials on the subgrids, \enzo\ uses a similar but slightly different technique from the root grid.
+The generation of the total gravitating mass field is essentially identical, using CIC interpolation for both the particles and baryons, including sub-grids as before.
+To compute the potential on subgrids, however, we use the standard seven-point (in three dimensions) second-order finite-difference approximation to Poisson's equation.
 Boundary conditions are then interpolated from the potential values on the parent grid.
-We use either tri-linear interpolation or a natural second-order spline for this: both methods give similar results.
+We use either tri-linear interpolation or a natural second-order spline for this: both methods give similar results but we generally choose to use tri-linear interpolation, which empirically provides a resonable compromise between speed and accuracy.
+Each scheme can be selected via the parameter file.
 The potential equation on each subgrid is then solved with the given Dirichlet boundary conditions with a multigrid relaxation technique.
-This is applied to each sub grid separately.
+This is applied to each subgrid separately.
 
@@
 
-It is known \red{(REF)} that simply interpolating the potential without feeding it back to higher levels leads to errors in the potential at more refined levels due to the build-up of errors during the interpolation of coarse boundary values.
+Simply interpolating the potential without feeding it back to higher levels leads to errors in the potential at more refined levels, due to the build-up of errors during the interpolation of coarse boundary values.
 In addition, neighboring subgrids are not guaranteed to generate the same potential values because of the lack of a coherent potential solve across the whole hierarchy.
@@
 To prevent overshoot, we average the potential on the boundary and allow for (typically) 4 iterations.
-This procedure can help in many cases, but does not necessarily produce a coherent solution across all grids and so does not completely solve the problem; we are working on a slower but more accurate method that does a multi-grid solve across the whole grid (Reynolds \etal, in preparation).
+This procedure can help in many cases, but does not necessarily produce a coherent solution across all grids and so does not completely solve the problem; we are working on a slower but more accurate method that does a multigrid solve across the whole grid (Reynolds \etal, in preparation).
 

### 2013-05-29 numerical-analysis.tex
Reynolds edits.
numerical-analysis adjusted. Found this section very clear, minor changes only

@@
+% EJT: general changes
+% Enzo to \enzo\
+% changed ``some Lagrangian fluid trajectories'' to ``a fixed number of Lagrangian fluid trajectories.'' Easier to understand.
+% changed ``some evolution'' to ``a period of evolution'' ... I basically don't like 'some'. It sounds vague.
+% rephrased to remove ``e.g.'' as they read slightly awkwardly
+% moved definition of T_1 and T_2 slightly higher in shock text to where they first appear.
+
 \section{Analysis}
@@
 
-Analyzing the results of simulations in detail requires both tools to ask
-sophisticated questions of data and the ability to process vast quantities of
+Detailed analsis of simualtion results requires both the tools to ask
+sophisticated questions of the data and the ability to process vast quantities of
 data at high time-cadence.  As simulations grow in size and complexity, storing data
 for post-processing simply becomes intractable.  To cope with this, we have
-instrumented Enzo with the ability to conduct analysis during the course of a
+instrumented \enzo\ with the ability to conduct analysis during the course of a
 simulation.  This enables analysis with extremely high time cadence (even as often
@@
 attempting to write an entire checkpoint output to disk.  The current mechanism
-for conducting analysis in Enzo during the course of the simulation utilizes
-the same compute resources as used by the simulation itself; this is often
+for conducting analysis in \enzo\ during the course of the simulation utilizes
+the same compute resources as used by the simulation itself by transferring their usage 
+from \enzo\ to the analysis routines; this is often
 referred to as \textit{in situ} analysis or visualization.  Utilizing a
@@
 
-We expose Enzo's mesh geometry and fluid quantities to the analysis platform
+We expose \enzo\ 's mesh geometry and fluid quantities to the analysis platform
 \texttt{yt} \citep{2011ApJS..192....9T, 2011arXiv1112.4482T}.  At compile time,
-Enzo is (dynamically or statically) linked against the Python and NumPy
+\enzo\ is (dynamically or statically) linked against the Python and NumPy
 libraries necessary to create proxy objects exposing the mesh geometry,
@@
 of the simulation that appears identical to an on-disk simulation output.
-Enzo then executes a user-provided analysis script, which is able to access the
+\enzo\ then executes a user-provided analysis script, which is able to access the
 in-memory simulation object.  Once the analysis script has returned control to
-Enzo, the simulation proceeds.  This process can occur at either the top of the
+\enzo\ , the simulation proceeds.  This process can occur at either the top of the
 main ``EvolveHierarchy'' loop or at the end of a timestep at the finest level,
@@
 halo finder \citep{2013ApJ...762..109B} can be executed through \texttt{yt} on
-in-memory Enzo data.  Operations that currently cannot be conducted on
+in-memory \enzo\ data.  Operations that currently cannot be conducted on
 in-memory datasets are those that require spatial decomposition of data; for
@@
 calculating topologically-connected sets requires a spatial decomposition of
-data and can thus not be conducted \textit{in situ}.  This prohibition extends
+data and thus cannot be conducted \textit{in situ}.  This prohibition extends
 to most halo finding operations, most multi-level parallelism operations,
@@
 major feature set that is currently being developed is to pass structured
-(i.e., non-fluid) information back from \texttt{yt} into Enzo.  For instance,
+(i.e., non-fluid) information back from \texttt{yt} into \enzo\ .  For instance,
 this could be the result of semi-analytic models of the growth and evolution of
@@
 merger-tree analysis, or even spectral energy distributions that are calculated
-within \texttt{yt} and provided to Enzo as input into radiative transfer
+within \texttt{yt} and provided to \enzo\ as input into radiative transfer
 calculations.  Future versions will include this, as well as the ability to
@@
 and \citet{2012ApJ...748...12S}. It is crucial to keep in mind that
-tracer particles model some Lagrangian fluid trajectories. The fluid
+tracer particles model a fixed number of Lagrangian fluid trajectories. The fluid
 on the grid, however, models the motion of all the mass and represents
-the average quantity in a grid cells volume. Consequently, after some
-evolution tracer particles, which even if they initially had the same
-density distribution as the gas, will not have the same density
-distribution as the fluid. They tend to accumulate at stagnation points
-of the flow, e.g., and care has to be taken in using these particles
-appropriately. They can be very useful, e.g. to study a variety of
+the average quantity in a grid cell's volume. Consequently, after a period of
+evolution, tracer particles --even if they initially had the same
+density distribution as the gas-- will not have the same density
+distribution as the fluid. For example, they tend to accumulate at stagnation points
+of the flow, and care has to be taken in using these particles
+appropriately. Tracer particles are very useful in studies such as the variety of
 histories of the hydrodynamic quantities in Lagrangian fluid elements
-and evaluate complex chemical and cooling models in regard to the
+and when evaluating complex chemical and cooling models in regard to the
 simpler ones used in the actual numerical evolution.
@@
 
-For every cell, we first determine whether is satisfies the following conditions
+For every cell, we first determine whether it satisfies the following conditions
 necessary to be flagged as a shock:
@@
 where $\vec{v}$ is the velocity field, $T$ is the temperature, $\rho$ is the
-density, and $S=T/\rho^{\gamma-1}$ is the entropy.  An optional temperature floor
-may be chosen, which is useful for, e.g., cosmological simulations without a radiation
+density, and $S=T/\rho^{\gamma-1}$ is the entropy.  $T_2$ and $T_1$ are the post-shock~(downstream)
+and pre-shock~(upstream) temperatures, respectively. An optional temperature floor
+may be chosen, which is useful for situations such as cosmological simulations without a radiation
 background.
@@
 \end{equation}
-where $T_2$ and $T_1$ are the post-shock~(downstream) and pre-shock~(upstream)
-temperatures, respectively. $\Mach$ is the upstream Mach number.
+where $\Mach$ is the upstream Mach number.
 

### 2013-05-29 numerical-analysis.tex
small text changed to conclusions

@@
 
-Detailed analsis of simualtion results requires both the tools to ask
+Detailed analysis of simulation results requires both the tools to ask
 sophisticated questions of the data and the ability to process vast quantities of
@@
 
-We expose \enzo\ 's mesh geometry and fluid quantities to the analysis platform
+We expose \enzo's mesh geometry and fluid quantities to the analysis platform
 \texttt{yt} \citep{2011ApJS..192....9T, 2011arXiv1112.4482T}.  At compile time,
@@
 in-memory simulation object.  Once the analysis script has returned control to
-\enzo\ , the simulation proceeds.  This process can occur at either the top of the
+\enzo, the simulation proceeds.  This process can occur at either the top of the
 main ``EvolveHierarchy'' loop or at the end of a timestep at the finest level,
@@
 major feature set that is currently being developed is to pass structured
-(i.e., non-fluid) information back from \texttt{yt} into \enzo\ .  For instance,
+(i.e., non-fluid) information back from \texttt{yt} into \enzo.  For instance,
 this could be the result of semi-analytic models of the growth and evolution of
@@
 inability to follow the evolution of a single parcel of fluid as it
-travels through the simulation volume.  To address this, Enzo has the
+travels through the simulation volume.  To address this, \enzo\ has the
 capability to introduce Lagrangian ``tracer particles'' into a

### 2013-05-29 numerical-ppm.tex

@@
 
+\subsubsection{Riemann Solvers and Fallback}
+
+While most numerical hydrodynamics schemes center around finding approximate
+solutions to the hydrodynamics equations, the core of philosophy of a Godunov's
+method \citep{Godunov1959} is to find the exact solution to an
+approximation of the physical setup.  The approximate problem in this case is
+the Riemann problem, wherein two volumes are separated by a thin membrane that
+is instantaneously removed at time $t=0$.  The subsequent evolution has an exact 
+analytic solution.  This solution is
+described in detail in many texts on computational fluid dynamics
+\citep[e.g.,][]{toro-1997}.   In brief, there are three waves that propagate
+away from the initial discontinuity.  The central wave, characterized by a
+density jump but not a pressure jump, is called the contact discontinuity.  The
+waves traveling to the left and right of the contact discontinuity can be either
+shocks, if characteristics converge on the wave front, or  rarefaction fans if
+characteristics diverge.
+
+While there exists an exact solution to this problem, finding it is expensive.  There are four
+possible combination of left- and right- traveling shocks and rarefactions,
+only one of which is fully consistent with the initial conditions.  Once the
+correct physical state is determined, the pressure in the central region can
+only be found by finding the root to an algebraic equation, which is necessarily
+an iterative process.  Thus a series of \emph{approximate} Riemann solvers are
+typically used.  There are four approximate Riemann solvers in Enzo: two-shock
+\citep{toro-1997},
+Harten-Lax-van Leer \citep[HLL,][]{toro-1997}, HLL with a contact discontinuity
+ \citep[HLLC,][]{toro-1997}, and HLL with
+multiple discontinuities \citep[HLLD,][]{Miyoshi05}.  Two-shock is used only
+with the PPM method.  HLL and HLLC are used with PPM, MUSCL (both with and
+without MHD) and MHD-CT.  HLLD is exclusively an MHD solver, and works with both
+MUSCL and MHD-CT methods.
+
+The only approximation that two-shock makes is that both left- and right- moving
+waves are shocks.  This solution still requires an iterative method for finding
+the pressure in between the two waves.  The HLL method alleviates this iteration
+by assuming that there is no central contact discontinuity, and the signal speed
+in the central region is approximated by an average over the left- and right-
+moving waves.  This method is significantly faster than the two-shock method,
+but also quite a bit more dissipative.  The HLLC is a three-wave method that improves upon the HLL
+method by also including the third wave, the contact discontinuity.
+
+For the MHD equations, there are seven waves, instead of three.  This makes the
+exact solution to the Riemann quite a bit more expensive.  Both the HLL and HLLC
+approximations can be formulated for the MHD equations, and are employed in both
+Dedner and CT solvers in Enzo.  The HLLD
+solver includes two of the additional waves, the rotational discontinuities,
+making it a five-wave solver.
+
+On rare occasions, high order
+solutions can cause negative densities or energies.  Both PPM and MUSCL solvers employ a Riemann
+solver fallback mechanism \citep{Lemaster09}.  If a negative density is found at
+a particular interface, the more diffusive HLL Riemann solver is used instead.
+
 % ------------------------------------------------------------

### 2013-05-29 tests-fld-ifront.tex

@@
 
-This test problem examines Enzo's flux-limited diffusion radiation
+This test problem examines \enzo's flux-limited diffusion radiation
 transport and associated ionization chemistry solvers on a
@@
 deceleration parameter $q_0 = 0.05$ and initial redshift $z_i=10$.
-Here, the physics of interest is the expansion of a \ion{H}{2} region
-in a uniform gas around a single monochromatic ionizing source (with
-frequency $h\nu = 13.6$ eV).  In this problem the I-front propagates
-rapidly at first, approaching 90\% of the Str{\" o}mgren radius by a
-scaled redshift of $-\log\left[(1+z)/(1+z_i)\right] \approx 0.15$,
-before cosmological expansion overtakes ionization, pushing the Str{\"
-o}mgren radius outward faster than the I-front can propagate.  Here,
-the Str{\" o}mgren radius is analytically given by
+Here the physics of interest is the expansion of an \ion{H}{2} region
+in a uniform medium around a single monochromatic ionizing source
+(with frequency $h\nu = 13.6$ eV).  The I-front propagates rapidly at
+first, approaching 90\% of the Str{\" o}mgren radius by a scaled
+redshift of $-\log\left[(1+z)/(1+z_i)\right] \approx 0.15$, before
+cosmological expansion overtakes ionization, pushing the Str{\"
+  o}mgren radius outward faster than the I-front can propagate.  The Str{\" o}mgren radius is analytically given by
 \[
@@
 where $\dot{N}_{\gamma} = 5 \times 10^{48}$ photon sec$^{-1}$ is the
-stength of the ionizing source, $\alpha_{\rm B} = 2.52 \times 10^{-13}$
+strength of the ionizing source, $\alpha_{\rm B} = 2.52 \times 10^{-13}$
 s$^{-1}$ is the \ion{H}{2} recombination rate, and $n_H(t)$ is the
@@
 \alpha_{\rm B}\, n_{H,i}\, /\, H_0\, / (1+z_i)$, where the subscript $i$
-indicates the quantity at the initial redshift.  The analytical
-solution for the I-front position as a function of time (or actually
-of the time/redshift-dependent expansion coefficient $a(t)$) is then
+indicates quantities at the initial redshift.  The analytical
+solution for the I-front position as a function of time is then
 
@@
      d\tilde{a}\right]^{1/3}, \quad\text{where} \\ 
-   \tau(a) &=& \lambda\frac{6q_0^2(1+z_i)^2}{F(a)-F(1)}, 
-     \quad\text{and} \\
+   \tau(a) &=& \lambda\frac{6q_0^2(1+z_i)^2}{F(a)-F(1)}, \\
    F(a) &=& \left[2-4q_0 - \frac{2q_o(1+z_i)}{a}\right] 
-      \left[1-2q_0 + \frac{2q_0(1+z_i)}{a}\right]^{1/2}.
+      \left[1-2q_0 + \frac{2q_0(1+z_i)}{a}\right]^{1/2},
 \end{eqnarray*}
+and $a(t)$ is the cosmological expansion coefficient.
 
-The simulation parameters are a box size $L_i\approx 27$ kpc, Hubble
-constant $H_0 = 1$ (in units of 100 km/s/Mpc), and cosmological
-constants $\Omega_m = 0.1$, $\Omega_A=0$ and $\Omega_b = 0.1$.  The
+The simulation parameters are: box size $L_i\approx 27$ kpc, Hubble
+constant $H_0 = 100$ km s$^{-1}$ Mpc$^{-1}$, and cosmological constants $\Omega_m = 0.1$, $\Omega_\Lambda=0$ and $\Omega_b = 0.1$.  The
 initial values for the simulation are a radiation energy density
@@
 In Figure \ref{fig.fld}, we plot profiles of the \ion{H}{1} and
-\ion{H}{2} fractions at the redshifts $z=\{6.24,\, 2.29,\, 1.02\}$, as
+\ion{H}{2} fractions at redshifts $z=\{6.24,\, 2.29,\, 1.02\}$, as
 well as the ratio of the I-front and Str{\" o}mgren radii throughout

### 2013-05-29 tests-selfsiminfall.tex

@@
 
-The next test problem is based on the self-similar solutions to the cosmological spherical infall problem found by \citet{Bertschinger1985}.
-It features a small density perturbation in a homogeneous $\Omega = 1$ universe and is a strong test of the cosmological evolution, gravity, and hydrodynamic portions of the code.
-It is close analogue to cosmological cluster formation and has an analytic solution.
+This test problem is based on the self-similar solutions to the
+cosmological spherical infall problem found
+by \citet{Bertschinger1985}.  It features a small density perturbation
+in a homogeneous $\Omega = 1$ universe and is a strong test of the
+cosmological evolution, gravity, and hydrodynamic portions of the
+code.  It is a close analogue to cosmological halo formation.
 
-For initial conditions, we adopt a $32^3$ top grid with a single, initial sub grid (covering $1/8^3$ of the domain) with a refinement factor of 2.
-An overdensity $\delta = 40$ is placed in a single cell near the center of the domain.
-We begin at $z=199$ and evolve to $z=0$, which is a sufficient time for the evolution to largely forget its initial conditions and approach the self-similar result.
-An over density refinement criteria ($\delta_{\rm crit} = 1.1$ on the top grid) is used to add additional grids, going up to 5 additional levels beyond the root grid.
-We use the PPM solver without radiative cooling; only baryons are used for this problem.
-An ideal gas law with $\gamma = 5/3$ is adopted.
+For the initial conditions, we adopt a $32^3$ top grid with a single,
+initial sub grid (covering $1/8^3$ of the domain) with a refinement
+factor of 2.  An overdensity $\delta = 40$ is placed in a single cell
+near the center of the domain.  We begin at $z=199$ and evolve to
+$z=0$, which is a sufficient time for the evolution to largely forget
+its initial conditions and approach the self-similar result.  An
+overdensity refinement criteria ($\delta_{\rm crit} = 1.1$ on the top
+grid) is used to add additional grids, going up to 5 additional levels
+beyond the root grid.  We use the PPM solver without radiative
+cooling; only baryons are used for this problem.  An ideal gas law
+with $\gamma = 5/3$ is adopted.
 
-In Figure~\ref{fig.sphericalinfall}, we show the results, scaled according to the dimensionless variables as defined in equations (2.9) and (3.2) of \citet{Bertschinger1985}.
-This demonstrates that we can quickly and easily obtain a good solution with only a fairly modest initial grid.
-The shock is sharply resolved and the asymptotic profiles as small $\lambda$ are recovered.
-At very small values of $\lambda$, the initial conditions have not been fully forgotten and the self-similar solution is not recovered -- this can be seen most clearly in the dimensionless mass; however, this is simply because of the limited amount of time for which we evolved the solution.
+In Figure~\ref{fig.sphericalinfall}, we show the results, scaled
+according to the dimensionless variables as defined in equations (2.9)
+and (3.2) of \citet{Bertschinger1985}.  This demonstrates that we can
+quickly and easily obtain a good solution with only a fairly modest
+initial grid.  The shock is sharply resolved and the asymptotic
+profiles at small $\lambda$ are recovered.  At very small values of
+$\lambda$, the initial conditions have not been fully forgotten and
+the self-similar solution is not recovered.  This can be seen most
+clearly in the dimensionless mass.  However, this is simply because of
+the limited amount of time for which we evolve the solution.
 
@@
 \includegraphics[width=0.7\textwidth]{figures/SphericalInfall.eps}
-\caption{The results of the self-similar spherical infall test.
-The panels show, as a function of the dimensionless radius $\lambda$, the dimensionless density $D$ (top left), velocity $V$ (top right), pressure $P$ (bottom left), and enclose mass $M$ (bottom right).
-In each case we show azimuthally averaged profiles from the simulation as circles and the analytic solution for for $\gamma = 5/3$ as solid lines.}
+\caption{The results of the self-similar spherical infall test.  The
+panels show, as a function of the dimensionless radius $\lambda$, the
+dimensionless density $D$ (top left), velocity $V$ (top right),
+pressure $P$ (bottom left), and enclosed mass $M$ (bottom right).  In
+each case we show azimuthally averaged profiles from the simulation as
+circles and the analytic solution for $\gamma = 5/3$ as solid lines.}
 \label{fig.sphericalinfall}

### 2013-05-29 tests-zeldovichpancake.tex

@@
 isolated caustic, and is thus a useful proxy for much more
-complicated structures in full 3-dimensional cosmological simulations
-(such as the collapse of gas onto a cosmological halo or filament).
+complicated structures in full 3-dimensional cosmological simulations,
+such as the collapse of gas onto a cosmological halo or filament.
 
@@
 
-with the internal energy of the gas set so that the entropy of the gas
-maintains a constant value throughout.  The velocity perturbation is given by
+with the internal energy of the gas set so that the entropy 
+is constant throughout.  The velocity perturbation is given by
 
@@
 Lagrangian mass coordinate, $k = 2 \pi / \lambda$ (where $\lambda$ is
-the perturbation wavelength), and H$_0$ is the $z = 0$ value of the
-Hubble constant.  Note that this solution is expressed in terms of
+the perturbation wavelength), and H$_0$ is the value of the Hubble
+constant at $z = 0$.  Note that this solution is expressed in terms of 
 Lagrangian positions, so one needs to convert this into the Eulerian
@@
 of a test of the adaptive mesh version of \enzo's Zel'dovich Pancake test.  A
-one-dimensional box of length $64$~Mpc/h is initialized at z$ = 20$ in
-an $\Omega_M = 1$ universe with h$ = 0.5$ and a background temperature
+one-dimensional box of length $64$~Mpc/h is initialized at $z = 20$ in
+an $\Omega_M = 1$ universe with $h = 0.5$ and a background temperature
 of 100~K, $\rho_0 = \rho_c$.  The simulation is initialized with 64
@@
 mass and the presence of shocks, for a maximum of 2 levels (i.e., an
-equivalent maximum resolution of 1024 grid cells).  The simulation was
+equivalent maximum resolution of 1024 grid cells).  The simulation is
 evolved to z$ = 0$ using the PPM hydro method, with the final output
-of the calculation being shown in Figure~\ref{fig.pancake}.
+of the calculation shown in Figure~\ref{fig.pancake}.
 
@@
 The strong shocks and large density gradients are well-resolved, with
-density and velocity jumps being well-delineated and at the correct locations.  The key features of
+density and velocity jumps being well-delineated and at the correct
+locations.  The key features of 
 this test problem can be resolved with far fewer cells -- simulations
@@
 gas, respectively; all are a function of position in units of the box
-size.  The central region ($x \simeq 0.45-0.55$) has bene adaptively
+size.  The central region ($x \simeq 0.45-0.55$) has been adaptively
 refined, as can be seen by the locations of grid points.  Shocks and

### 2013-05-29 parallel.tex
working on section 12.

@@
 
-The current version of \enzo\ has been parallelized for distributed memory platforms using the Message Passing Interface (MPI).
-This is done using a single grid object as the basic unit of parallelization.
-Each grid object -- including all cell and grid data -- is fully contained in a single processor\footnote{We use the word {\it processor} to mean a basic distribution unit; this could by either a core or a node, depending on details of the system.
-Note that the current version of Enzo is not threaded, although the development of a hybrid version is in process}.
-Parallelization is accomplished by distributing grids amongst processors.
-This is done on the root grid using a simple tiling system, where the root grid is split up into $N_{\rm root}$ tiles, where $N_{\rm root}$ is typically equal to the number of processors $N_p$.
+The current version of \enzo\ has been parallelized for distributed
+memory platforms using the Message Passing Interface (MPI).  This is
+done using a single grid object as the basic unit of parallelization.
+Each grid object -- including all cell and grid data -- is fully
+contained in a single processor\footnote{We use the word {\it
+processor} to mean a basic distribution unit; this could by either a
+core or a node, depending on details of the system.  Note that the
+current version of Enzo is not threaded, although the development of a
+hybrid version is in process}.  Parallelization is accomplished by
+distributing grids amongst processors.  This is done on the root grid
+using a simple tiling system, where the root grid is split up into
+$N_{\rm root}$ tiles, where $N_{\rm root}$ is typically equal to the
+number of processors $N_p$.
 
-Load balancing on levels other than the root level (i.e.
-grids for which the level $l > 0$) is different as the refined patches are not generally uniformly distributed.
-Grids on refined levels are first placed on the same processor as their parent to minimize communication; however, this generally does not result in a well-balanced load.
-Therefore, the code has a number of options for load balancing the grids on a given level.
-Each grid is assigned a load, generally equal to the number of cells in the grid.
-The first load-balancing option is to move a grid from the processor with the highest load to the processor with the lowest load, with the proviso that only grids with load factors less than half the difference between the highest and lowest loaded processors will be moved.
-This continues until the load ratio between the most-to-least loaded processor is below 1.05 or until no suitable grid can be found to transfer.
-A second load balancing option uses a space-filling Hilbert curve to order the grids by their approximate spatial position.
-Then, once the grids have a specific one-dimensional ordering, we can divide up the grids into $N_p$ groups (as equally as possible).
-Load balancing is done separately for each level.
-Clearly load balancing is most successful if there are significantly more grids than processors; however, small grids are less efficient (because of their ghost zones), and so the code uses a simple heuristic in order to split up grids until there are of order 10 grids per processor.
-This generally results in good load-balancing while not producing grids which are wastefully small.
+Load balancing on levels other than the root level (i.e., grids for
+which the level $l > 0$) is different as the refined patches are not
+generally uniformly distributed.  Grids on refined levels are first
+placed on the same processor as their parent to minimize
+communication; however, this generally does not result in a
+well-balanced load.  Therefore, the code has a number of options for
+load balancing the grids on a given level. Each grid is assigned a
+load, generally equal to the number of cells in the grid (which,
+empirically, is a good estimate of computational cost).  The first
+load-balancing option is to move a grid from the processor with the
+highest load to the processor with the lowest load, with the proviso
+that only grids with load factors less than half the difference
+between the highest and lowest loaded processors will be moved.  This
+continues until the load ratio between the most-to-least loaded
+processor is below 1.05 or until no suitable grid can be found to
+transfer.  A second load balancing option uses a space-filling Hilbert
+curve to order the grids by their approximate spatial position.  Then,
+once the grids have a specific one-dimensional ordering, we can divide
+up the grids into $N_p$ groups (with the division taking place as
+equally as possible).  Load balancing is done separately for each
+level.  Clearly load balancing is most successful if there are
+significantly more grids than processors; however, small grids are
+less efficient (because of their ghost zones), and so the code uses a
+simple heuristic in order to split up grids until there are of order
+10 grids per processor.  This generally results in good load-balancing
+while not producing grids that are wastefully small.
 
-Communication between processors is done using a non-blocking communication strategy that allows overlap of communication and computation.
-This can be done efficiently because each processor retains a copy of the entire hierarchy of grids, except that grids which do not `live' on a given processor only contain meta data (essentially location and size of the grid; such grids are denoted as `ghost' grids).
-Replicating the hierarchy means that all communication of data from one grid to another can be identified by each processor independently.
-The metadata for `ghost' grids are quite small and so the extra memory required is generally not onerous unless very large numbers of grids are used (more than a few hundred thousand grids).
-A schematic of this distribution is shown in Figure~\ref{fig.amr_hierarchy}.
+Communication between processors is done using a non-blocking
+communication strategy that allows overlap of communication and
+computation.  This can be done efficiently because each processor
+retains a copy of the entire hierarchy of grids, except that grids
+that do not `live' on a given processor only contain meta data
+(essentially location and size of the grid; such grids are denoted as
+`ghost' grids).  Replicating the hierarchy means that all
+communication of data from one grid to another can be identified by
+each processor independently.  The metadata for `ghost' grids are
+quite small and so the extra memory required is generally not onerous
+unless very large numbers of grids are used (more than a few hundred
+thousand grids).  A schematic of this distribution is shown in
+Figure~\ref{fig.amr_hierarchy}.
 
-Data is transferred through a three-step procedure that takes advantage of MPI capabilities: (i) as the code progresses, and data is needed from another grid on another processor, the receiving processor posts an MPI non-blocking receive indicating that it is expecting data; this outstanding receive is recorded in a table , (ii) the sending processor calls the MPI non-blocking send function, and then finally (iii) the receiving processor, after it has carried out all the computation it can, waits for any MPI message to arrive.
-Each message is coded so that it can be matched with appropriate receive posted in the first step, and based on that, the appropriate routine is called to processes the data.
-Step (iii) is repeated until there are no outstanding receives.
+Data is transferred through a three-step procedure that takes
+advantage of the capabilities of the MPI library: (i) as the code
+progresses, and data is needed from another grid on another processor,
+the receiving processor posts an MPI non-blocking receive indicating
+that it is expecting data; this outstanding receive is recorded in a
+table, (ii) the sending processor calls the MPI non-blocking send
+function, and then finally (iii) the receiving processor, after it has
+carried out all the computation it can, waits for any MPI message to
+arrive.  Each message is coded so that it can be matched with the
+appropriate receive posted in the first step, and based on that, the
+appropriate routine is called to processes the data.  Step (iii) is
+repeated until there are no outstanding receives.
 
@@
 \end{center}
-\caption{\emph{Left:}  Example of a simple, distributed AMR hierarchy showing real and ghost grids.
-\emph{Right:}  Example 2D \enzo\ grid showing real and ghost zones, as 
-needed for the PPM hydro stencil. }
+\caption{\emph{Left:} Example of a simple, distributed AMR hierarchy
+showing real and ghost grids.  \emph{Right:} Example 2D \enzo\ grid
+showing real and ghost zones, as needed for the PPM hydro stencil. }
 \label{fig.amr_hierarchy}
@@
 Because of the wide variety of simulations, methods, and uses of Enzo,
-it is difficult to define exactly which routines are most costly
+it is difficult to state generally which routines will be most costly
 during a given simulation.  As such, we have designed a lightweight
@@
 used routines (such as the hydrodynamic and gravity solvers) as well
-as refinement level timers that measure the (exclusive) time spent on
-each level.  Beyond this minimal set of routines, we have designed a
-simple way for the user to modify the source by adding
-\texttt{TIMER\_START(``Your Routine Name'')} and
-\texttt{TIMER\_END(``Your Routine Name'')}.  These timers are
-automatically registered in a
-std::map\footnote{http://www.cplusplus.com/reference/map/map/}.  These
-timers are created and managed individually on each processor in an
-asynchronous fashion.
+as refinement level timers that measure the time spent on each level.
+Beyond this minimal set of routines, we have designed a simple way for
+the user to modify the source by adding \texttt{TIMER\_START(``Your
+Routine Name'')} and \texttt{TIMER\_END(``Your Routine Name'')}.
+These timers are created and managed individually on each processor in
+an asynchronous fashion, and contribute minimal computational, memory,
+and IO overhead.
 
-At each complete EvolveHierarchy (or less often if specified), each timer is
-then communicated to the root processor where it calculates the mean, standard
-deviation, minimum, and maximum for each of the timers across all processors. 
-For level timers, there are also attributes such as the number of cell updates,
-the current number of grids, and the average cells/s/MPI process.  This
-information is then output to a ``performance.out'' logfile.  This provides a 
-simplified interface to the user that can be used to diagnose performance 
-issues as well as estimate a given problem type's scalability.  In addition to 
-the logfile, we have developed a plotting interface for quickly producing 
-figures that process the data from the logfile.  These capabilities are 
-described in the online documentation, along with a further discussion of the 
+At each complete root grid time step (or less often if specified),
+each timer is then communicated to the root processor where it
+calculates the mean, standard deviation, minimum, and maximum for each
+of the timers across all processors.  For level timers, there are also
+attributes such as the number of cell updates, the current number of
+grids, and the average cells/s/MPI process.  This information is then
+output to a logfile.  This provides a simplified interface to the user
+that can be used to diagnose performance issues as well as estimate a
+given problem type's scalability.  In addition to the logfile, we have
+developed a plotting interface for quickly producing figures that
+process the data from the logfile.  These capabilities are described
+in the online documentation, along with a further discussion of the
 performance measurement implementation.
@@
 on up to 13,824 MPI tasks on the NICS Kraken XT4 and ORNL Jaguar XT4
-supercomputers\footnote{These simulations were performed in 2008,
-prior to conversion of both machines to the current-generation
-systems}.  In this test, each MPI task was given a $128^3$ root grid
-tile (i.e., $128^3$ grid cells containing baryon quantities) and,
-initially, approximately $128^3$ dark matter particles.  The number of
-grid cells was constant throughout the calculation; the number of dark
-matter particles varies as they are moved from subvolume to subvolume
-as structure evolves.  The grid resolution was kept at a constant
+supercomputers\footnote{These simulations were performed prior to
+conversion of both machines to the current-generation systems}.  In
+this test, each MPI task was given a $128^3$ root grid tile (i.e.,
+$128^3$ grid cells containing baryon quantities) and, initially,
+approximately $128^3$ dark matter particles.  The number of grid cells
+was constant throughout the calculation; the number of dark matter
+particles varies as they are moved from subvolume to subvolume as
+structure evolves.  The grid resolution was kept at a constant
 comoving size of $\simeq 40$~kpc/h, and as the core count was
@@
 Both machines used the SeaStar2 interconnect.  In the scaling study,
-calculations were run with 1, 2 or 4 MPI tasks per node.  The figure
+calculations were run with 1, 2, or 4 MPI tasks per node.  The figure
 shows cell updates per second per MPI process; perfect scaling would
@@
 seen between the two machines, which can be attributed primarily to
-the slightly faster CPUs on Kraken (2.3 Ghz, vs. 2.1 Ghz on Jaguar).
-The difference in speed when using different numbers of MPI tasks per
-node can be attributed primarily due to differences in competing usage
-of shared cache on the quad-core chips used on this machine.
+the slightly faster CPUs on Kraken at th etime (2.3 Ghz, vs. 2.1 Ghz
+on Jaguar).  The difference in speed when using different numbers of
+MPI tasks per node can be attributed primarily due to differences in
+competing usage of shared cache on the quad-core chips used on this
+machine.
 
@@
 variety of problems as long as each compute core is given an adequate
-amount of work to do.  For cosmological simulations, this seems to be
-roughly $128^3$ cells per core.  If fewer cells per core are used, the
-CPU is essentially data-starved, and poor scaling is observed due to
-computing units being idle while waiting for information to be
-communicated from other processes (for, e.g., boundary information or
-gravity solves).  Substantially larger cell counts per core would in
-principle help scaling by reducing the amount of inter-process
-communication needed, but larger cell counts are typically impractical
-on most machines due to memory limits.
+amount of work to do.  For cosmological simulations, this value has
+been empirically determined to be roughly $128^3$ cells per core.  If
+fewer cells per core are used, the CPU is essentially data-starved,
+and poor scaling is observed due to computing units being idle while
+waiting for information to be communicated from other processes (for,
+e.g., boundary information or gravity solves).  Substantially larger
+cell counts per core would in principle help scaling by reducing the
+amount of inter-process communication needed, but larger cell counts
+are typically impractical on most machines due to memory limits.
 
@@
 \caption{\emph{Left:} Strong scaling test of a 512$^3$ AMR
-  cosmological calculation.  The root grid scaling is not
-  representative of the true strong scaling of Enzo because the root
-  grid tiles are not repartitioned when a simulation is restarted with
-  additional cores.  The weak scaling test in Figure
-  \ref{fig.uniscale} is more representative of the scaling on the root
-  level.  The performance in the refined levels show good strong
-  scaling up to 2,048 cores.  \emph{Right:} Time spent in
-  representative major routines in the same AMR calculation.}
+cosmological calculation.  The root grid scaling is not representative
+of the true strong scaling of Enzo because the root grid tiles are not
+repartitioned when a simulation is restarted with additional cores.
+The weak scaling test in Figure \ref{fig.uniscale} is more
+representative of the scaling on the root level.  The performance in
+the refined levels show good strong scaling up to 2,048 cores.
+\emph{Right:} Time spent in representative major routines in the same
+AMR calculation.}
 \label{fig:strong_scaling}
@@
 Many astrophysical problems, such as cosmological galaxy formation
-\citep{}, high-resolution disk galaxy simulations \citep{},
-high-redshift \citep{} and present-day star formation \citep{},
-involve multi-scale physics that span several orders of magnitude in
-both space and time.  In these situations, using Enzo in its adaptive
-mesh refinement mode is beneficial.  Because the adaptive grid
-hierarchy is dynamic and not static, grid boundaries and, thus,
-communication patterns can be unpredictable, hindering good strong
-scaling at high core counts.
+\citep{2012ApJ...749..140H}, high-resolution disk galaxy simulations
+\citep{}, high-redshift \citep{2009Sci...325..601T}, and present-day
+star formation \citep{Collins12a}, involve multi-scale physics that
+span several orders of magnitude in both space and time.  In these
+situations, using Enzo in its adaptive mesh refinement mode is
+beneficial.  Because the adaptive grid hierarchy is dynamic, grid
+boundaries and, thus, communication patterns can be unpredictable,
+hindering strong scaling behavior to high core counts.
 
@@
 single 50 Mpc/h cosmology simulation run on $N_{\rm core}$ compute
-cores, ranging from 128 to 16,384 at power-of-two intervals.  It was
-run on the NICS Kraken XT5 supercomputer that has two AMD Opteron
-hexa-core processors with 16 GB of memory per compute node.  The
-simulations that utilized 128, 256, and 512 cores were executed on 128
-nodes because of the memory requirements.  The higher core count jobs
-were run with 8 cores per node.  This simulation would not run with 12
-cores per node because of the memory overhead associated with the grid
-hierarchy being duplicated on each MPI process.  However, this
-overhead is greatly diminished if a hybrid-parallel (MPI + OpenMP)
-approach is used.
+cores, ranging from 128 to 16,384 cores at power-of-two intervals.  It
+was run on the NICS Kraken XT5 supercomputer, which has two AMD
+Opteron hexa-core processors with 16 GB of memory per compute node.
+The simulations that utilized 128, 256, and 512 cores were executed on
+128 nodes because of the memory requirements.  The higher core count
+simulations were run with 8 cores per node.  This simulation would not
+run with 12 cores per node because of the memory overhead associated
+with the grid hierarchy being duplicated on each MPI process.
+However, this overhead is greatly diminished if a hybrid-parallel (MPI
++ OpenMP) approach is used.
 
@@
 tiles, and 6 additional AMR levels are used.  We perform these scaling
-tests when the simulation reaches $z=4$.  The results shown in Figure
+tests when the simulation reaches $z=4$, where the AMR grid hierarchy
+is well-developed and is thus a reasonable representation of AMR
+simulation behavior in general.  The results shown in Figure
 \ref{fig:strong_scaling} come from a single root-level timestep
-$\Delta t = 2.1\, \textrm{Myr}$.  At this time, there are $3.04 \times
+of $\Delta t = 2.1\, \textrm{Myr}$.  At this time, there are $3.04 \times
 10^5$ AMR grids, $9.03 \times 10^8$ ($\sim 967^3$) computational
@@
   \parbox[t]{0.9\textwidth}{\textbf{Note.} --- Data shown at $z=4$ for
-    a root grid timestep of 2.1~Myr.  Col. (1): AMR Level. Col. (2):
+    a root grid timestep of 2.1~Myr.  
+    Col. (1): AMR Level. Col. (2):
     Number of grids. Col. (3): Number of computational
@@
 MPI processes.  These results are directly applicable to simulations
-with similar computational demands.  In simulations with less
+with similar computational demands.  In simulations with fewer
 computational cells, strong scaling will cease at a smaller core count
-because the CPUs will become data-starved faster, and the opposite
-occurs with larger simulations.
+because the CPUs will become data-starved more quickly, and the
+opposite occurs with larger simulations.
 
@@
 
-In this section, we develop a simple, approximate model to estimate the compute time and memory required to complete a given calculation.
-Note that we are {\it not} trying to model parallel performance (which we discussed in the previous section), but have an even simpler goal: to determine how to scale compute time estimates as we increase the resolution.
-This is a straightforward thing to do in codes which use static grids, as the computational effort per time step is constant.
-However, for an AMR calculation, the number of grids that will be generated during the run is not known in advance, and so the cpu time per problem time can vary drastically throughout the calculation.
-%For example, at the beginning, when the densities are nearly uniform, only the static grid is required and the calculation progresses rapidly.
-However, as structure forms and dense clumps are generated, the number of grid points swells by orders of magnitude (an increase by $10^3$ is not uncommon) and most of the cpu time is consumed at late times.
-Therefore, simply performing a few steps does produce a good estimate of the cpu time.
+In this section, we develop a simple, approximate model to estimate
+the compute time and memory required to complete a given calculation.
+Note that we are {\it not} trying to model parallel performance (which
+was discussed in the previous section), but have an even simpler goal:
+to determine how to scale compute time estimates as we increase the simulation
+resolution.  This is a straightforward thing to do in codes that use
+static grids, as the computational effort per time step is constant.
+However, for an AMR calculation, the number of grids that will be
+generated during the run is not known in advance, and so the CPU time
+per problem time can vary drastically throughout the calculation.
+For example, at the beginning of the simulation when the densities are nearly uniform,
+only the static grid is required and the calculation progresses
+rapidly.  However, as structure forms and dense clumps are generated,
+the number of grid points swells by orders of magnitude (an increase
+of $10^3$ is not uncommon) and most of the cpu time is consumed at
+late times.  Therefore, simply performing a few steps does produce a
+good estimate of the cpu time.
 
-Nevertheless, we can try to determine how the compute time scales for a given run, as we increase the resolution.
-First, we neglect particles and concentrate on the time taken by the grids, which can be justified both theoretically and empirically\footnote{Since the potential is calculated on the grid, the only particle costs are: depositing mass to the grid, interpolating accelerations, and updating particle positions and velocities.
-These are relatively cheap operations.}.
-Second, we break the problem down slightly, and examine the scaling over a short enough period of time that the grid structure does not change significantly (i.e.
-the number of grids at each level).
-We then assume that the whole run scales in the same way, or in other words, that changes in the resolution affect the simulation in the same way at each time step.
-This is usually a good approximation as the most costly part of the calculation typically don't change their characteristics substantially.
+Nevertheless, we can try to determine how the compute time scales for
+a given run as we increase the resolution.  First, we neglect
+particles and concentrate on the time taken by the grids, which can be
+justified both theoretically and empirically\footnote{Since the
+potential is calculated on the grid, the only particle costs are:
+depositing mass to the grid, interpolating accelerations, and updating
+particle positions and velocities.  These are relatively
+computationally inexpensive operations.}.  Second, we break the
+problem down slightly, and examine the scaling over a short enough
+period of time that the grid structure does not change significantly
+(i.e. the number of grids at each level remains approximately
+constant).  We then assume that the whole run scales in the same way,
+or in other words, that changes in the resolution affect the
+simulation in the same way at each time step.  This is usually a good
+approximation, as the most costly parts of the calculation typically
+don't change their characteristics substantially between timesteps.
 
-To make progress, we assume that the cost to advance a single cell by one timestep is a constant.
-This can be incorrect if the chemistry solver requires many iterations, but is usually fairly accurate.
-For a unigrid calculation, the time would by proportional to $N_{\rm root}^4$ since the number of cells scales as $N_{\rm root}^3$ and, assuming the Courant condition is the time limiting factor, then the number of steps to advance a given time is proportional to $N_{\rm root}.$  Therefore, to advance a hierarchy a given time, we find, accounting for all levels and using a refinement factor of 2.
-\begin{equation}
-t_{\rm SU} = C_1 \sum_l f_l N_{\rm root}^4 2^{4l}
-\end{equation}
-where $C_1$ is a constant, which can be thought of as the time taken to advance a single cell.
-The factor $f_l$ is the fraction of the volume on a given level that is actually refined.
-By definition $f_0 = 1$, and $f_l \le 1$.
-In writing down this equation, we neglect a number of costs which are not directly proportional to cell count, including communication between processors, the $\log{N}$ factor for the root grid FFT, Cache misses, optimizations, and other costs associated with processing the hierarchy.
-The first item on this list, in particular, is clearly not true for large processor counts; however, we neglect parallel considerations in this section.
+To make progress, we assume that the computational cost to advance a
+single cell by one timestep is a constant.  This can be incorrect if
+the chemistry solver requires many iterations, but is usually fairly
+accurate.  For a unigrid calculation, the time would by proportional
+to $N_{\rm root}^4$ since the number of cells scales as $N_{\rm
+root}^3$ and, assuming the Courant condition is the time limiting
+factor, then the number of steps to advance a given time is
+proportional to $N_{\rm root}.$ Therefore, to advance a hierarchy a
+given time, we find, accounting for all levels and using a refinement
+factor of 2.
+\begin{equation} t_{\rm SU} = C_1 \sum_l f_l N_{\rm root}^4 2^{4l}
+\end{equation} where $C_1$ is a constant, which can be thought of as
+the time taken to advance a single cell.  The factor $f_l$ is the
+fraction of the volume on a given level that is actually refined.  By
+definition $f_0 = 1$, and $f_l \le 1$.  In writing down this equation,
+we neglect a number of costs which are not directly proportional to
+cell count, including communication between processors, the $\log{N}$
+factor for the root grid FFT, Cache misses, optimizations, and other
+costs associated with processing the hierarchy.  The first item on
+this list, in particular, is clearly not true for large processor
+counts; however, we neglect parallel considerations in this section.
 
-Note that unless $f_l/f_{l+1} < 1/16$ (i.e.
-if less then about 6\% of a given level is further refined), the cost per level will increase with level.
-A key question therefore, is the values of $f_l$.
-Unfortunately, this depends strongly on the simulation being run.
-In Figure~\ref{fig:scaling}, we show the values of $f_l$ for three simulations: two cosmological runs with varying box size, and a third simulation focusing on a single disk galaxy.
+Note that unless $f_l/f_{l+1} < 1/16$ (i.e. if less then about 6\% of
+a given level is further refined), the cost per level will increase
+with level.  A key question, therefore, is the value of $f_l$ for each
+level.  Unfortunately, this depends strongly on the simulation being
+run.  In Figure~\ref{fig:scaling}, we show the values of $f_l$ for
+three simulations: two cosmological runs with varying box size, and a
+third simulation focusing on a single disk galaxy.
 
@@
 \centerline{\includegraphics[width=0.7\textwidth]{figures/scaling_plot2}}
-\caption{This figure show (clockwise from upper left), the covering fraction $f_l$ of the grids on a given level $l$, the ratio of $f_{l}/f_{l+1}$, the relative memory usage of each level (normalized by the memory usage of level 0), and, in the lower-left panel, the relative cpu usage of each level, again normalized by $l=0$.
-In each panel, three curves are shown: the solid black and long dashed red are two cosmological simulations with box sizes of 30 h$^{-1}$ Mpc and 80h$^{-1}$ Mpc respectively.
-The dot-dashed blue line is for a non-cosmological simulation of a ram-pressure stripped galaxy.
-In the top-right panel, there is a line at the critical ratio of 0.06, which determines if the next finer level ($l+1)$ takes more (above) or less (below) cpu time than level $l$.}
+\caption{This figure show (clockwise from upper left), the covering
+fraction $f_l$ of the grids on a given level $l$, the ratio of
+$f_{l}/f_{l+1}$, the relative memory usage of each level (normalized
+by the memory usage of level 0), and, in the lower-left panel, the
+relative cpu usage of each level, again normalized by $l=0$.  In each
+panel, three curves are shown: the solid black and long dashed red are
+two cosmological simulations with box sizes of 30 h$^{-1}$ Mpc and
+80h$^{-1}$ Mpc respectively.  The dot-dashed blue line is for a
+non-cosmological simulation of a ram-pressure stripped galaxy.  In the
+top-right panel, there is a line at the critical ratio of 0.06, which
+determines if the next finer level ($l+1)$ takes more (above) or less
+(below) cpu time than level $l$.}
 \label{fig:scaling}
@@
 
-As, the figure demonstrates, the $f_l$ values all show a sharp decline with level, dropping as power-laws with the level $l$.
-Focusing first on the ``refine-everywhere" cosmological runs, which both show similar behavior despite the different box sizes and redshifts at which we collect the data.
-In fact, we find the results are very typical for cosmological runs, with only slight variations depending on if we are simulating first stars or clusters of galaxies.
-The top-left panel shows the ratio of $f_{l}/f_{l+1}$ which, remarkably, hovers around the critical value of 0.0625 determined earlier.
-This implies that the relative cpu usage of each level, shown in the lower-left panel, is nearly flat, and that going to high levels of refinement only increases the amount of calculation by a factor of roughly $1/l_{\rm max}$.
-Even in the most extreme case, the L80 run, the last level adds only 25\% to the computation time.
+As, the figure demonstrates, the $f_l$ values all show a sharp decline
+with level, dropping as power-laws with the level $l$.  Focusing first
+on the ``refine-everywhere" cosmological runs, which both show similar
+behavior despite the different box sizes and redshifts at which we
+collect the data.  In fact, we find the results are very typical for
+cosmological runs, with only slight variations depending on if we are
+simulating first stars or clusters of galaxies.  The top-left panel
+shows the ratio of $f_{l}/f_{l+1}$ which, remarkably, hovers around
+the critical value of 0.0625 determined earlier.  This implies that
+the relative cpu usage of each level, shown in the lower-left panel,
+is nearly flat, and that going to additional levels of refinement only
+increases the amount of calculation by a factor of roughly $1/l_{\rm
+max}$.  Even in the most extreme case, the L80 run, the last level
+adds only 25\% to the computation time.
 
@@
 \end{equation}
-and the terms in this sum are shown in the bottom-right panel of Figure~\ref{fig:scaling}.
-This demonstrates that the memory usage is dominated by the top three levels, and adding additional levels only adds minimally to the memory usage of the run.
-We have empirically tested this scaling for cosmological simulations and found it to be reasonably accurate (although the increase is typically slightly larger than found here, usually 30-50\%, probably because of less-than-ideal parallel scaling).
+and the terms in this sum are shown in the bottom-right panel of
+Figure~\ref{fig:scaling}.  This demonstrates that the memory usage is
+dominated by the top three levels, and adding additional levels only
+adds minimally to the memory usage of the run.  We have empirically
+tested this scaling for cosmological simulations and found it to be
+reasonably accurate (although the increase is typically slightly
+larger than found here, usually 30-50\%, probably because of
+less-than-ideal parallel scaling).
 
-The physical reason for the $f_{l+1}/f_l$ ratios found in the cosmological simulations appears to be due to a combination of the density structure of individual clouds, and the distribution of the clouds themselves.
-In particular, we note that for a $\rho \propto r^{-2}$ profile, the Lagrangian refinement criteria typically used in such simulations produces an $f_{l+1}/f_l$ ratio of 1/8 for a single, resolved cloud.
-Of course, at some point we would resolve the flat density of individual clouds and the ratio would climb, making further refinement more costly; however, we are not yet in this regime in this example.
+The physical reason for the $f_{l+1}/f_l$ ratios found in the
+cosmological simulations appears to be due to a combination of the
+density structure of individual clouds, and the distribution of the
+clouds themselves.  In particular, we note that for a $\rho \propto
+r^{-2}$ density profile, the Lagrangian refinement criteria typically used in
+such simulations produces an $f_{l+1}/f_l$ ratio of 1/8 for a single,
+resolved cloud.  Of course, at some point we would resolve the flat
+density of individual clouds and the ratio would climb, making further
+refinement more costly; however, we are not yet in this regime in this
+example.
 
-These conclusions are, however, highly dependent on the type of run being performed.
-We contrast this cosmological case to the simulation of a galaxy formation run, as shown by the dot-dashed curves in Figure~\ref{fig:scaling}.
-Note that the levels 1 and 2 are statically refined to ensure the Lagrangian volume of the galaxy halo is resolved to at least level 2 at all times.
-More importantly, the $f_{l+1}/f_{l}$ ratio of the highest levels are systematically slightly larger than the critical 0.0625 value, indicating that the cpu time is dominated by the most refined levels, as shown in the bottom-left panel.
-The memory usage is dominated by levels 2 and 3, the static levels, as the bottom-right panel demonstrates.
-In this case, adding more levels of refinement would somewhat increase the cost of the simulation.
+These conclusions are, however, highly dependent on the type of run
+being performed.  We contrast this cosmological case to the simulation
+of a galaxy formation run, as shown by the dot-dashed curves in
+Figure~\ref{fig:scaling}.  Note that the levels 1 and 2 are statically
+refined to ensure the Lagrangian volume of the galaxy halo is resolved
+to at least level 2 at all times. More importantly, the
+$f_{l+1}/f_{l}$ ratio of the highest levels are systematically
+slightly larger than the critical 0.0625 value, indicating that the
+cpu time is dominated by the most refined levels, as shown in the
+bottom-left panel.  The memory usage is dominated by levels 2 and 3,
+the static levels, as the bottom-right panel demonstrates.  In this
+case, adding more levels of refinement would somewhat increase the
+cost of the simulation.
 
-Finally, we have so far focused on scaling when increasing $l_{\rm max}$, but without changing the refinement criteria.
-However, we can also keep the levels fixed and modify the refinement criteria.
-For cosmological runs, this is typically done by increasing the mass resolution (and simultaneously decreasing the dark matter particle mass), and adding additional linear small-scale modes to the initial conditions.
-The effect of this is to boost the $f_l$ values at all levels (except for the root grid where $f_0$ is already 1) by a factor linearly proportional to the increase in the mass resolution.
-This is because the refinement criteria typically boosts the number of refined cells on the first and subsequent levels by this factor.
-Again, we have empirically demonstrated that this scaling is approximately valid (provided we keep the maximum level constant).
+We have focused thus far on scaling when increasing $l_{\rm max}$, but
+without changing the refinement criteria.  However, we can also keep
+the levels fixed and modify the refinement criteria.  For cosmological
+runs, this is typically done by increasing the mass resolution (and
+simultaneously decreasing the dark matter particle mass), and adding
+additional linear small-scale modes to the initial conditions.  The
+effect of this is to boost the $f_l$ values at all levels (except for
+the root grid where $f_0$ is already 1) by a factor linearly
+proportional to the increase in the mass resolution.  This is because
+the refinement criteria typically boosts the number of refined cells
+on the first and subsequent levels by this factor.  Again, we have
+empirically demonstrated that this scaling is approximately valid,
+provided that we keep the maximum level constant.
 
-Putting this together, we obtain an approximate scaling for cosmological simulations as 
+Putting this together, we find an approximate scaling for the
+computational time required for cosmological simulations:
 \begin{equation}
@@
 \end{equation}
-which at least gives users some way to estimate compute times.
+This is approximately accurate, and gives users some way to estimate
+compute times for cosmological calculations.  Similar scaling
+relations can be obtained for other types of \enzo\ AMR simulations.

### 2013-05-29 parallel.tex
further edits to parallel section.

@@
 Each grid object -- including all cell and grid data -- is fully
-contained in a single processor\footnote{We use the word {\it
-processor} to mean a basic distribution unit; this could by either a
-core or a node, depending on details of the system.  Note that the
-current version of Enzo is not threaded, although the development of a
-hybrid version is in process}.  Parallelization is accomplished by
-distributing grids amongst processors.  This is done on the root grid
-using a simple tiling system, where the root grid is split up into
-$N_{\rm root}$ tiles, where $N_{\rm root}$ is typically equal to the
-number of processors $N_p$.
+contained in a single processor\footnote{In this context, we use the
+word {\it processor} to mean a basic distribution unit; this could by
+either a core or a node, depending on details of the system.  Note
+that the current version of Enzo is not threaded, although the
+development of a hybrid version is in process}.  Parallelization is
+accomplished by distributing grids amongst processors.  This is done
+on the root grid using a simple tiling system, where the root grid is
+split up into $N_{\rm root}$ tiles, with $N_{\rm root}$ typically
+equal to the number of processors $N_p$.
 
 Load balancing on levels other than the root level (i.e., grids for
-which the level $l > 0$) is different as the refined patches are not
+which the level $l > 0$) is different, as the refined patches are not
 generally uniformly distributed.  Grids on refined levels are first
@@
 communication; however, this generally does not result in a
-well-balanced load.  Therefore, the code has a number of options for
-load balancing the grids on a given level. Each grid is assigned a
-load, generally equal to the number of cells in the grid (which,
+well-balanced computational load.  Therefore, the code has a number of options for
+load balancing the grids on a given level. Each grid is assigned an
+estimated computational load, generally equal to the number of cells in the grid (which,
 empirically, is a good estimate of computational cost).  The first
 load-balancing option is to move a grid from the processor with the
-highest load to the processor with the lowest load, with the proviso
+highest computational load to the processor with the lowest load, with the proviso
 that only grids with load factors less than half the difference
@@
 Because of the wide variety of simulations, methods, and uses of Enzo,
-it is difficult to state generally which routines will be most costly
-during a given simulation.  As such, we have designed a lightweight
-registering system that has been implemented for the most commonly
-used routines (such as the hydrodynamic and gravity solvers) as well
-as refinement level timers that measure the time spent on each level.
-Beyond this minimal set of routines, we have designed a simple way for
-the user to modify the source by adding \texttt{TIMER\_START(``Your
-Routine Name'')} and \texttt{TIMER\_END(``Your Routine Name'')}.
-These timers are created and managed individually on each processor in
-an asynchronous fashion, and contribute minimal computational, memory,
-and IO overhead.
+it is difficult to state in general terms which routines within the
+code will be most costly during a given simulation.  As such, we have
+designed a lightweight registering system that has been implemented
+for the most commonly used routines (such as the hydrodynamic and
+gravity solvers) as well as refinement level timers that measure the
+time spent on each level.  Beyond this minimal set of routines, we
+have designed a simple way for the user to modify the source by adding
+\texttt{TIMER\_START(``Your_Routine_Name'')} and
+\texttt{TIMER\_END(``Your_Routine_Name'')}.  These timers are created
+and managed individually on each processor in an asynchronous fashion,
+and contribute minimal computational, memory, and IO overhead.
 
@@
 calculates the mean, standard deviation, minimum, and maximum for each
-of the timers across all processors.  For level timers, there are also
+of the timers across all processors.  For level timers, there are additional
 attributes such as the number of cell updates, the current number of
@@
 In this section, we develop a simple, approximate model to estimate
-the compute time and memory required to complete a given calculation.
-Note that we are {\it not} trying to model parallel performance (which
-was discussed in the previous section), but have an even simpler goal:
-to determine how to scale compute time estimates as we increase the simulation
-resolution.  This is a straightforward thing to do in codes that use
-static grids, as the computational effort per time step is constant.
-However, for an AMR calculation, the number of grids that will be
-generated during the run is not known in advance, and so the CPU time
-per problem time can vary drastically throughout the calculation.
-For example, at the beginning of the simulation when the densities are nearly uniform,
-only the static grid is required and the calculation progresses
-rapidly.  However, as structure forms and dense clumps are generated,
-the number of grid points swells by orders of magnitude (an increase
-of $10^3$ is not uncommon) and most of the cpu time is consumed at
-late times.  Therefore, simply performing a few steps does produce a
-good estimate of the cpu time.
+the computational time and memory required to complete a given
+calculation.  Note that we are {\it not} trying to model parallel
+performance (which was discussed in the previous section), but have an
+even simpler goal: to determine how to scale computational time
+estimates as we increase the simulation resolution.  This is a
+straightforward thing to do in codes that use static grids, as the
+computational effort per time step is constant.  However, for an AMR
+calculation, the number of grids that will be generated during the run
+is not known in advance, and so the CPU time per problem time can vary
+drastically throughout the calculation.  For example, at the beginning
+of the simulation when the densities are nearly uniform, only the
+static grid is required and the calculation progresses rapidly.
+However, as structure forms and dense clumps are generated, the number
+of grid points swells by orders of magnitude (an increase of $10^3$ is
+not uncommon) and most of the CPU time is consumed at late times.
+Therefore, simply performing a few steps at the beginning of the
+calculation does produce a good estimate of the required CPU time.
 
@@
 to $N_{\rm root}^4$ since the number of cells scales as $N_{\rm
-root}^3$ and, assuming the Courant condition is the time limiting
-factor, then the number of steps to advance a given time is
+root}^3$ and, assuming the Courant condition is the 
+factor that controls the time step, the number of steps to advance the
+calculation over a given time interval is
 proportional to $N_{\rm root}.$ Therefore, to advance a hierarchy a
-given time, we find, accounting for all levels and using a refinement
-factor of 2.
+given time interval, we find, accounting for all levels and using a refinement
+factor of 2,
 \begin{equation} t_{\rm SU} = C_1 \sum_l f_l N_{\rm root}^4 2^{4l}
@@
 definition $f_0 = 1$, and $f_l \le 1$.  In writing down this equation,
-we neglect a number of costs which are not directly proportional to
+we neglect a number of costs that are not directly proportional to
 cell count, including communication between processors, the $\log{N}$
-factor for the root grid FFT, Cache misses, optimizations, and other
+factor for the root grid FFT, cache misses, optimizations, and other
 costs associated with processing the hierarchy.  The first item on
-this list, in particular, is clearly not true for large processor
+this list, in particular, is clearly important for large processor
 counts; however, we neglect parallel considerations in this section.
@@
 by the memory usage of level 0), and, in the lower-left panel, the
-relative cpu usage of each level, again normalized by $l=0$.  In each
+relative CPU usage of each level, again normalized by $l=0$.  In each
 panel, three curves are shown: the solid black and long dashed red are
@@
 determines if the next finer level ($l+1)$ takes more (above) or less
-(below) cpu time than level $l$.}
+(below) CPU time than level $l$.}
 \label{fig:scaling}
@@
 behavior despite the different box sizes and redshifts at which we
-collect the data.  In fact, we find the results are very typical for
-cosmological runs, with only slight variations depending on if we are
-simulating first stars or clusters of galaxies.  The top-left panel
+collect the data.  In fact, we find these results are very typical for
+cosmological runs, with only slight variations depending on the exact
+system that is being simulated.  The top-left panel
 shows the ratio of $f_{l}/f_{l+1}$ which, remarkably, hovers around
 the critical value of 0.0625 determined earlier.  This implies that
-the relative cpu usage of each level, shown in the lower-left panel,
+the relative CPU usage of each level, shown in the lower-left panel,
 is nearly flat, and that going to additional levels of refinement only
@@
 
-The memory used by each level, on the other hand scales as:
+The memory used by each level, on the other hand, scales as
 \begin{equation}
@@
 of a galaxy formation run, as shown by the dot-dashed curves in
-Figure~\ref{fig:scaling}.  Note that the levels 1 and 2 are statically
-refined to ensure the Lagrangian volume of the galaxy halo is resolved
+Figure~\ref{fig:scaling}.  Note that grid levels 1 and 2 in this simulation are statically
+refined to ensure that the Lagrangian volume of the galaxy halo is resolved
 to at least level 2 at all times. More importantly, the
@@
 slightly larger than the critical 0.0625 value, indicating that the
-cpu time is dominated by the most refined levels, as shown in the
+CPU time is dominated by the most refined levels, as shown in the
 bottom-left panel.  The memory usage is dominated by levels 2 and 3,
@@
 This is approximately accurate, and gives users some way to estimate
-compute times for cosmological calculations.  Similar scaling
-relations can be obtained for other types of \enzo\ AMR simulations.
+compute times for \enzo\ calculations.

### 2013-05-29 numerical-radiation-homogeneous.tex
Minor edits to radiation sections.

@@
 and cooling/heating routines described elsewhere.  Many of these use
-fits to the H, He and He$^+$ ionizing and photo-ionization heating
+fits to the H, He and \Hep ionizing and photo-ionization heating
 rates that are of the form
@@
 give the coefficients for the photoionization (and photo-heating)
-rates of H, He, and He$^+$ in Table~\ref{table:homo_coefs}.  Radiation
+rates of H, He, and \Hep in Table~\ref{table:homo_coefs}.  Radiation
 background types 1
@@
 Homogeneous radiation field types 5 and 6 start with a spectral shape
-that is then integrated against the appropriate H, He, and He$^+$
+that is then integrated against the appropriate H, He, and \Hep,
 cross-sections to compute the ionization and photo-ioniziation heating

### 2013-05-29 conclusions.tex

@@
+% EJT added intro sentense to section on future developement and tweaked final sentense
+
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 
-In this paper, we have presented the algorithms underlying \enzo, an
+In this paper, we have presented the algorithms underlying \enzo: an
 open-source adaptive mesh refinement code designed for
@@
 and figures for this paper are open source and are available at the
-\enzo\ website, \url{http://enzo-project.org}.  Furthermore, the yt toolkit,
+\enzo\ website, \url{http://enzo-project.org}.  Furthermore, the \texttt{yt} toolkit,
 which is designed to analyze \enzo\ data (as well as data from a wide
@@
 \url{http://yt-project.org}.  Both of these codes have active user and
-developer communities, extensive documentation and user suppport, and
+developer communities, extensive documentation and user support, and
 strong mechanisms for users to contribute their changes and fixes to
@@
 
-In the longer term, it seems likely that trends in supercomputing
-hardware will require substantial modifications to \enzo's core
-infrastructure, and very possibly some of the core algorithsms.  More
-specifically, these trends involve the usage of specialized
+With the continual rapid development of computer hardware, it makes sense to review
+not only at \enzo's current capabilities, but to look towards its future development 
+in view of predicted technological trends. These trends in supercomputing hardware suggest that  
+substantial modifications to \enzo's core infrastructure, and very possibly some of 
+the core algorithms will be required. More
+specifically, the progression involves the usage of specialized
 large-core-count, vectorized computing units such as graphics
-processuing units (GPUs) or chips like the Intel Xeon Phi, as well as
+processing units (GPUs) or chips like the Intel Xeon Phi, as well as
 precipitously decreasing amounts of RAM per computing core.  The
 former trend means that the amount of processing power per compute
-node will continue to trend upward, likely much faster than the
+node will continue to increase, likely much faster than the
 bandwidth between nodes, and will require tremendous reduction in
@@
 fact are faced by effectively all applications that wish to take
-advantage of new computational architectures.  Thus, we suspect that
+advantage of new computational architectures. We therefore anticipated that 
 \enzo\ (or a code that has the capabilities of \enzo, from a user's
-point of view) will continue to be usable at the largest scales on
+point of view) will continue to develop to be usable at the largest scales on
 such machines.

### 2013-05-29 parallel.tex

@@
 \citep{2012ApJ...749..140H}, high-resolution disk galaxy simulations
-\citep{}, high-redshift \citep{2009Sci...325..601T}, and present-day
+\citep{2011ApJ...738...54K}, high-redshift \citep{2009Sci...325..601T}, and present-day
 star formation \citep{Collins12a}, involve multi-scale physics that

### 2013-05-29 tests-orszagtang.tex

@@
 Figure~\ref{fig.orszag} shows the Orszag-Tang vortex problem, a
-standard magnetohydrodynamic test problem \citep{Orszag79}.  The left panel shows
+classic MHD test problem described by \citep{Orszag79}.  The left panel shows
 the result using \enzo's constrained transport MHD method, while the right panel shows the result using
@@
 The simulation is evolved to $t=0.48$ (in arbitrary units).  One can see that the structures
-are accurately represented as compared to, for instance,
+are accurately represented as compared to, for example,
 \citet{Toth00}, and that the resolution of shocks is comparable in

### 2013-05-29 tests-pointsourcegravity.tex

@@
 
-This test is a simple test of the ability of the Poisson solver to correctly reproduce the acceleration around a point source without any dynamics.
-We place a single particle in the center of the unit domain with root grid dimensions $32^3$ and add a static AMR grid with refinement factor two that covers the central $1/8^3$ of the domain (i.e.
-so the refined grid is only $8^3$).
-We then place 5000 particles throughout the domain distributed uniformly in angle and with the logarithm of the radius (from the central point).
-The measured accelerations are shown in Figure~\ref{fig.gravitytest}.
-As is typical for Particle-Mesh based methods, the errors decrease at large distances and peak on roughly the grid scale.
-The accelerations are relatively smooth across the top grid/fine grid boundary with the most noticeable impact being a drop in the amplitude of tangential errors on the fine grid.
-The force law rolls over at about 1.6 fine cell widths (equivalent to 0.8 root grid widths in this figure), consistent with the CIC-deposition and interpolation used.
+This is a simple test of the ability of the Poisson solver to
+correctly reproduce the acceleration around a point source without any
+dynamics.  We place a single particle in the center of the unit domain
+with root grid dimensions $32^3$ and add a static, nested grid 
+refinemed by a factor of two that covers the central $1/8^3$ of the
+domain (i.e. so the refined grid is only $8^3$).  We then place 5000
+particles throughout the domain distributed uniformly in angle and
+in the logarithm of the radius (from the central point).  The
+measured accelerations are shown in Figure~\ref{fig.gravitytest}.  As
+is typical for Particle-Mesh based methods, the errors decrease at
+large distances and peak on roughly the grid scale.  The accelerations
+are relatively smooth across the top grid/fine grid boundary with the
+most noticeable impact being a drop in the amplitude of tangential
+errors on the fine grid.  The force law rolls over at about 1.6 fine
+cell widths (equivalent to 0.8 root grid widths in this figure),
+consistent with the CIC-deposition and interpolation used.
 

### 2013-05-29 tests-sedov.tex

@@
 \includegraphics[width=0.4\textwidth]{figures/sedov-zeus-slice.eps}
-\caption{Density slice from the Sedov Blast test at $t = 0.07$. Left:
+\caption{Density slices from the Sedov Blast test at $t = 0.07$. Left:
 results using the Piecewise Parabolic Method hydro scheme.  Right: results using the \zeus\
@@
 \includegraphics[width=\textwidth]{figures/sedov-profiles.eps}
-\caption{Radially averaged profiles for the Sedov Blast test at $t =
+\caption{Radial profiles for the Sedov Blast test at $t =
 0.07$. Clockwise from top-left shows density, velocity, internal
@@
 point of energy injection.  This problem is a popular test of
-astrophysical numerical codes for three reasons: First, it is
-particularly appropriate to astrophysics since it represents the
-situation when a supernovae explosion occurs. Second, it has an
-exact analytical solution whereby the shock front's radial position is
-given by:
+astrophysical simulation codes for three reasons.  First, this problem
+is representative of the astrophysical phenomenon of a supernova
+explosion.   Second, an analytical solution to the evolution of the
+Sedov blast-wave exists, allowing for a direct test of the accuracy of
+the code.  The radius of the shock front as a function of time is
+given by
 
@@
 ideal gas with $\gamma = 1.4$.  For the full derivation see
-\citet{Sedov1959}. This solution makes it possible to assess how well
-the code performs.  Third, the spherical shock is a challenging
-problem for numerical codes (both grid and particle) since the shock
-wave increases in size as the simulation progresses and its
-symmetrical nature highlights any directional preferences that grid
-codes can succumb to. The test presented here is the two-dimensional
+\citet{Sedov1959}.  Third, as the spherical shock expands, its
+symmetry, or lack thereof, serves to highlight any directional
+preferences of the hydrodynamics solvers.  The test presented here is
+the two-dimensional 
 version that is included in the \enzo\ distribution. The
 three-dimensional results from this test, both for \enzo\ and three
-other leading astrophysics codes, can be founds in \citet{Tasker2008}.
+other leading astrophysics codes, can be found in \citet{Tasker2008}.
 
@@
 deposited into a single cell at the center of the box with $E_0 =
-10.0$. The problem is completed in two dimensions with reflecting
-boundary conditions (the default for \enzo) and uses a box size of side
-$1$. For this problem, we selected a top grid of $100 \times 100$
-cells and a maximum of four levels of refinement, placed based on
-shock location and the slope in density and total energy. The
-exception to this scheme is in the initial conditions, where grids are
-placed directly around the injection point. The results were assessed
-at $t = 0.07$, which corresponding to a time just before the shock
+10.0$. The problem is run in two dimensions with reflecting
+boundary conditions and a box size of side
+$1$.  We selecte a top grid of $100 \times 100$
+cells and a maximum of four levels of adaptive refinement, refining on
+shocks and the slopes of the density and total energy fields. The
+exception to this scheme is in the initial conditions, where
+additional refinement is placed directly around the injection point to
+better resolve spherical geometry. The results are assessed
+at $t = 0.07$, which corresponds to a time just before the shock
 reaches the box edge (see Figure~\ref{fig.sedov1}).
@@
 produces the direction-dependent shockfront shape. In their paper,
-\citet{Clarke2010} presents results from an alternative version of
+\citet{Clarke2010} present results from an alternative version of
 \zeus\ that conserves total energy. This problem is less marked for

### 2013-05-29 numerical-amr.tex
Adding my edits to the overview section.

@@
  \item Each subgrid must begin and end on the boundary of a coarse cell.  This means that the number of cells in a patch will be a multiple of the refinement factor.
- \item A subgrid most be completely enclosed by its parent.
+ \item A subgrid must be completely enclosed by its parent.
 \end{itemize}
@@
 Each grid patch in \enzo\ contains arrays of values for baryon and particle quantities.
-Grids are partitioned into a core of \emph{active zones} and a surrounding layer of \emph{ghost zones}, as shown in Figure~\ref{fig:amr_hierarchy}.
-\red{(Move this figure to here?)}  The active zones store field values and ghost zones are used to temporarily store values that have been obtained directly from neighboring grids or interpolated from a parent grid.
+Grids are partitioned into a core of \emph{active zones} and a surrounding layer of \emph{ghost zones}, as shown in Figure~\ref{fig.amr_hierarchy}.
+The active zones store field values and ghost zones are used to temporarily store values that have been obtained directly from neighboring grids or interpolated from a parent grid.
 These zones are necessary to accommodate the computational stencil of the (magneto)hydrodynamics solvers (Sections~\ref{sec.hydro.ppm} and~\ref{sec.hydro.zeus}) and the gravity solver (Section~\ref{sec.gravity}).
@@
 The finest level is then advanced, using as many steps as it takes to reach the level immediately above.
-The finest level is then synched to the level above it, which then proceeds foreward in time one more step.
+The finest level is then synched to the level above it, which then proceeds forward in time one more step.
 This is shown graphically in Figure~\ref{fig:wcycle}.
-
-\begin{figure}
-\begin{center}
-%\includegraphics[width=0.6\textwidth]{figures/wcycle.eps}
-\includegraphics[width=0.6\textwidth]{figures/timestepping.eps}
-\end{center}
-\caption{\emph{Left:} Example of the timesteps on a 2-level AMR
-  hierarchy.  Enzo does not restrict the timesteps on the finer levels
-  to be a factor of $1/2^n$ smaller.  \emph{Right:} The order in which
-  the AMR grids are evaulated on each level.}
-\label{fig:wcycle}
-\end{figure}
 
@@
 
-In the first case, we allow for four types of boundary conditions:
+In the first case, we allow for four types of boundary conditions
+(represented here as being at $x=0$, with positive $x$ interior and
+negative $x$ exterior to the domain):
 \begin{enumerate}
@@
   \item{\em Outflow.}  We approximate outflow boundary conditions by duplicating the solution at the edge of the computational domain: $q(-x) = q(0)$.
-  \item{\em Inflow.} In this case, the boundary values are fixed by a pre-determined function ($q(-x) = q0(x,t)$).
+  \item{\em Inflow.} In this case, the boundary values are fixed by a pre-determined function ($q(-x) = q0(-x,t)$).
 The code provides a simple way to set inflow values that are constant in both time and space over an entire face, or hooks are provided for special cases where the boundary must be time-varying.
@@
 multi-dimensional interpolation functions, and so focus on
-second-order schemes since they are tractable and relatively accurate.
-\footnote{We could, in principle, build a multi-dimensional
+second-order schemes since they are tractable and relatively 
+accurate\footnote{We could, in principle, build a multi-dimensional
   interpolator out of a one-dimensional interpolation function;
   however, at higher order, this would not necessarily satisfy our
-  constraint regarding not introducing new extrema in interpolated quantities.}  In Appendix~\ref{app:interpolation}, we
+  constraint regarding not introducing new extrema in interpolated quantities.}.  In Appendix~\ref{app:interpolation}, we
 explicitly write down functions in one, two and three dimensions, but
@@
 Since a refined region is simulated with (at least) two different resolutions (one fine and one coarse), it has two or
-more ``solutions'' (i.e.
-values of density, velocity, etc.) that represent the same volume of space.
+more ``solutions'' (i.e.~values of density, velocity, etc.) that represent the same volume of space.
 This is one important aspect of the SAMR method -- some physical regions are represented by grids at multiple resolutions.
@@
 \end{equation}
-where $r$ is the refinement factor and the sum is over all fine cells that ``cover" the coarse cell.
+where $r$ is the refinement factor, and the sum is over all fine cells that ``cover" the coarse cell.
 The exponent, $d$, refers to the weight of the average.
@@
 Conservation can be restored by correcting all coarse cells that lie outside the boundary of a fine region
-with the difference between the coarse and fine estimates of the flux across the boundary:
-
+with the difference between the coarse and fine estimates of the flux
+across the boundary:
 \begin{equation}
@@
 \end{equation}
-
 where $F$ is the flux of quantity $q_i$ in the $i$-direction through the cell interface.
@@
 Once all cells in a given grid that require additional refinement have been flagged, rectangular boundaries are determined that minimally encompass the flagged cells.
-This is done using an iterative machine-vision technique, as described in \red{(REF)}.
+This is done using an iterative machine-vision technique, as described in \cite{Berger91}.
 
@@
 We compute an efficiency $\epsilon$ for this grid, defined as the ratio of flagged cells to total cells.
-If this efficiency is sufficiently larger (typically larger than 30\%), or if the grid is small enough (typically less than $4^d$ coarse cells, where d is the dimensionality of the simulation) then we adopt the region as a new grid and move on to the next region (if any).
+If this efficiency is sufficiently large (typically over 30\%), or if the grid is small enough (typically less than $4^d$ coarse cells) then we adopt the region as a new grid and move on to the next region (if any).
 If the region is not acceptable, then we split it into two using the following algorithm.
@@
 Edges are typically indicated by inflection points in this signature, so we calculate the second derivative using a simple three-point finite difference estimator and look for the maximum in the second derivative, with ties going to the index closest to the center.
-If no inflection points are found along the longest exis, the next-longest axis is searched.
+If no inflection points are found along the longest axis, the next-longest axis is searched.
 If no inflection points are found, the longest axis is cut in half.
@@
 In this section, we turn to the refinement criteria themselves.
-Although generalized refinement criteria are available based on estimators for the error, such as the Richardson method \red{(REF)}, we have focused generally on methods that use physical estimators based on the assumption that the simulator has a specific quantity or quantities that indicate the need for refinement.
+Although generalized refinement criteria are available based on
+estimators for the error, such as the Richardson method \citep{AtkinsonHan2004}, we have focused generally on methods that use physical estimators based on the assumption that the simulator has a specific quantity or quantities that indicate the need for refinement.
 The flagging methods currently built into \enzo\ are:
@@
 
-\item{\em Slope.}  If the normalized slope $(q_{i+1} - q_{i-1})/ (2 q_i)$ of baryon quantities between two adjacent cells is larger than a specified value (typically 0.3), then the cell is flagged.
+\item{\em Slope.}  If the normalized slope $(q_{i+1} - q_{i-1})/ (2 q_i)$ of baryon quantities between adjacent cells is larger than a specified value (typically 0.3), then the cell is flagged.
 
-\item{\em Baryon mass.}  This refinement criterion is designed to mimic a Lagrangian method in that it tries to keep a fixed mass resolution.
-In this method, a cell is refined if the baryonic mass in the cell ($M_g = \rho \Delta_x^d$) is larger than a specified value:
+\item{\em Baryon mass.}  This refinement criterion is designed to
+  mimic a Lagrangian method in that it tries to keep a fixed mass
+  resolution.  In this method, a cell is refined if the baryonic mass
+  in the cell ($M_g = \rho\, (\Delta x)^d$) is larger than a specified value:
 \begin{equation}
-M_g > \rho_{\rm flag} {\Delta x^d_{\rm root}} r^{\epsilon_l l},
+M_g > \rho_{\rm flag}\, (\Delta x_{\rm root})^d r^{\epsilon_l l},
 \end{equation}
@@
 \begin{eqnarray}
-(p_{i+1} - p_{i-1})/\min(p_{i+1}, p_{i-1}) & > & 0.33  \nonumber \\
-u_{i-1} - u_{i+1} & > & 0 \\
-e_i / E_i & > & 0.1  \nonumber
+%% (p_{i+1} - p_{i-1})/\min(p_{i+1}, p_{i-1}) & > & 0.33,  \nonumber \\
+%% u_{i-1} - u_{i+1} & > & 0, \quad\text{and} \\
+%% e_i / E_i & > & 0.1.  \nonumber
+(p_{i+1} - p_{i-1})/\min(p_{i+1}, p_{i-1}) > 0.33,  \qquad
+u_{i-1} - u_{i+1} > 0, \quad\text{and} \qquad
+e_i / E_i > 0.1.  \nonumber
 \end{eqnarray}
-We note that the values $0.33$ and $0.1$ in the equations above were determined empirically, and can be modified by the user to refine more or less aggressively, as the situation warrants.
+We note that the parameters $0.33$ and $0.1$ in the equations above were determined empirically, and can be modified by the user to refine more or less aggressively, as the situation warrants.
 
-\item{\em Particle mass.}  This refinement criterion is precisely the same as the baryon overdensity criterion except that it uses the gridded (cloud-in-cell, or CIC) particle density.
+\item{\em Particle mass.}  This refinement criterion is precisely the same as the baryon mass criterion except that it uses the gridded (cloud-in-cell, or CIC) particle density.
 In general, this criteria is used in cosmological simulations, and thus refines on dark matter and stellar density.
 
-\item{\em Jeans length.}  Following \cite{Truelove98}, the code can be forced to refine the Jeans length by a fixed number of cells, or, in more specifically, whenever the following criterion is not met:
+\item{\em Jeans length.}  Following \cite{Truelove98}, the code can be forced to refine the Jeans length by a fixed number of cells, or more specifically, whenever the following criterion is not met:
 \begin{equation}
-\Delta x < \left( \frac{\pi k_B T}{J^2 G \rho \mh} \right)^{1/2}
+\Delta x < \left( \frac{\pi k_B T}{J^2 G \rho \mh} \right)^{1/2},
 \end{equation}
@@
 However, in some cases it is desirable to resolve the cooling process, in which case we flag cells that have a cooling time shorter than their sound crossing time.
-The cooling time is given by $t_{\rm cool} = 1.5 kT / n \Lambda(T)$ and the sound crossing time is $t_{\rm cross} = \Delta x / c_s$, where $c_s$ is the sound speed.
+The cooling time is given by $t_{\rm cool} = 1.5 kT / (n\, \Lambda(T))$ and the sound crossing time is $t_{\rm cross} = \Delta x / c_s$, where $c_s$ is the sound speed.
 
@@
 The optical depth of a cell is estimated to be:
-
 \begin{equation}
@@
 \end{equation}
-
 where $\sigma_{\rm HI}$ is the neutral atomic hydrogen absorption
@@
 This is done by ensuring that the resistive length scale, defined as
-
 \begin{equation}
@@
 \end{equation}
-
-is always refined by at least N cells locally (i.e., refine a given cell further if $L_R \leq N \Delta x$).
+is always refined by at least N cells locally (i.e.~we refine a given cell further if $L_R \leq N \Delta x$).
 

### 2013-05-29 overview.tex
removing an obsolete note
Adding my edits to the overview section.

@@
 We begin this section by first writing down the complete set of
-physical equations that can be solved by \enzo, and then briefly describe
-the numerical algorithms that we use to solve these equations.  This
-section is intended to be an overview of the \enzo\ code's
-capabilities: thus, we gather all of the equations solved into a
-single place and provide a brief and high-level introduction to the
+physical equations that can be solved by \enzo, and then briefly
+describe the numerical algorithms that we use to solve these
+equations.  This section is intended to be an overview of the \enzo\
+code's capabilities: thus, we gather all of the equations solved into
+a single place and provide a brief and high-level introduction to the
 numerics of the code.  Detailed descriptions of the individual
-components are then provided in Sections~\ref{sec.amr}-\ref{sec.num.analysis}.
+components are then provided in
+Sections~\ref{sec.amr}-\ref{sec.num.analysis}.
 
@@
 
-The Eulerian equations of ideal magnetohydrodynamics including gravity,
-in a coordinate systems comoving with the cosmological expansion, are
-given by:
+The Eulerian equations of ideal magnetohydrodynamics including
+gravity, in a coordinate systems comoving with the cosmological
+expansion, are given by:
 
@@
   \frac{\partial \rho \vecv}{\partial t}  
-   + \frac{1}{a} \nabla \cdot \left(\rho \vecv \vecv + \myvec{I}p^* - \frac{\vecB \vecB}{a} \right) & = &
+  + \frac{1}{a} \nabla \cdot \left(\rho \vecv \vecv + \myvec{I}p^* - 
+    \frac{\vecB \vecB}{a} \right) & = &
   - \frac{\dot{a}}{a} \rho \vecv - \frac{1}{a} \rho \nabla \phi
@@
   \frac{\partial E}{\partial t} 
-+ \frac{1}{a} \nabla \cdot \left[ (E + p^*) \vecv -  \frac{1}{a} \vecB (\vecB \cdot \vecv) \right] & = &
-     - \frac{\dot{a}}{a} \left( 2E - \frac{B^2}{2a} \right)  - \frac{1}{a} \vecv \cdot \nabla \phi 
+  + \frac{1}{a} \nabla \cdot \left[ (E + p^*) \vecv - 
+    \frac{1}{a} \vecB (\vecB \cdot \vecv) \right] & = &
+  - \frac{\dot{a}}{a} \left( 2E - \frac{B^2}{2a} \right) - 
+  \frac{1}{a} \vecv \cdot \nabla \phi 
   - \Lambda + \Gamma + \frac{1}{a^2} \nabla \cdot \fcond
@@
 %
-\frac{\partial \vecB}{\partial t} - \frac{1}{a}  \nabla \times (\vecv \times \vecB) & = & 0 \label{eq:induction}
+  \frac{\partial \vecB}{\partial t} - 
+  \frac{1}{a}  \nabla \times (\vecv \times \vecB) & = & 
+  -\frac{\dot{a}}{2 a} \vecB \label{eq:induction}
 \end{eqnarray}
@@
 identity matrix, and $a$ is the cosmological expansion parameter
-(which is discussed in more detail below).
-The first equation represents conservation of mass, the second, conservation of momentum, and the third, conservation of total fluid (kinetic plus thermal) energy.
-They are respectively, the first, second and third moments of the Boltzmann equation.
-The fourth equation is the magnetic induction equation.
-Terms representing radiative cooling ($\Lambda$) and heating ($\Gamma$) enter on the right-hand side of the energy equation (\ref{eq:total_energy}), as well as the flux due to thermal heat conduction ($\fcond$).
+(which is discussed in more detail below).  The first equation
+represents conservation of mass, the second, conservation of momentum,
+and the third, conservation of total fluid (kinetic plus thermal)
+energy.  They are respectively, the first, second and third moments of
+the Boltzmann equation.  The fourth equation is the magnetic induction
+equation.  Terms representing radiative cooling ($\Lambda$) and
+heating ($\Gamma$) enter on the right-hand side of the energy equation
+(\ref{eq:total_energy}), as well as the flux due to thermal heat
+conduction ($\fcond$).
 
@@
 \end{equation}
-where $e$ is the comoving thermal energy density, and the total isotropic pressure $p^*$ is given by
+where $e$ is the comoving thermal energy density, and the total
+isotropic pressure $p^*$ is given by
 \begin{equation}
-p^* = p + \frac{B^2}{2a}
+p^* = p + \frac{B^2}{2a},
 \end{equation}
-The equations are closed with the equation of state and Poisson's equation (we work in the Newtonian limit):
+where $p$ is the gas pressure.  The equations are closed with the
+equation of state and Poisson's equation (we work in the Newtonian
+limit):
 %
@@
 %
-e   & = & p / [(\gamma - 1) \rho ],
+  e &=& p / [(\gamma - 1) \rho ],
   \label{eq:eq_of_state} \\
 %
-\nabla^2 \phi & = & 4 \pi G \rho_{\rm total}
+  \nabla^2 \phi &=& 4 \pi G \rho_{\rm total}
 %                   + 3 p_{total}/c^2) - \Lambda. 
@@
 %
-The equation of state is shown here for an ideal gas with a ratio of specific heats $\gamma$.
-Equation (\ref{eq:potential}) includes contributions to the
-gravitational potential $\phi$ from the total mass density $\rho_{\rm
-  total}$, which is defined as $\rho_{\rm gas} + \rho_{\rm dm} + \rho_{\rm stars}
-- \rho_{\rm average}$.  Although we write the equations including both MHD and
-comoving coordinates, the code is frequently used both in the pure
-hydrodynamic limit ($\vecB = 0$) and without cosmological expansion
-($a = 1$, $\dot{a} = 0$).
+The equation of state is shown here for an ideal gas with a ratio of
+specific heats $\gamma$.  Equation (\ref{eq:potential}) includes
+contributions to the gravitational potential $\phi$ from the total
+mass density $\rho_{\rm total}$, which is defined as $\rho_{\rm gas} +
+\rho_{\rm dm} + \rho_{\rm stars} - \rho_{\rm average}$.  Although we
+write the equations including both MHD and comoving coordinates, the
+code is frequently used both in the pure hydrodynamic limit ($\vecB =
+0$) and without cosmological expansion ($a = 1$, $\dot{a} = 0$).
 
@@
 
-For completeness, we note here that we have defined several key quantities
-in the comoving frame to make the previous equations more readable.
-Specifically, we define:
+For completeness, we note here that we have defined several key
+quantities in the comoving frame to make the previous equations more
+readable.  Specifically, we define:
 
@@
 p     & = & a^3 \pp, \\
-E       & = & a^3 (\Ep - 
+E     & = & a^3 \left(\Ep - 
             \dot{a} \vecx \cdot \vecvp - 
-              \frac{1}{2} \dot{a}^2 \vecx^2), \\
+            \frac{1}{2} \dot{a}^2 \vecx^2\right), \\
 \phi  & = & \phip + \frac{1}{2} a \ddot{a} \vecx^2. \\
-\vecB & = & a^2 \vecB^\prime
+\vecB & = & a^2 \vecB^\prime,
 \end{eqnarray}
-where primes indicate the usual quantities defined in a fixed frame frame, $\vecr$ is the proper distance and $\vecx$ is the comoving distance.
+where primes indicate the usual quantities defined in a fixed frame
+frame, $\vecr$ is the proper distance and $\vecx$ is the comoving
+distance.
 
@@
 %%$$\rho_b \equiv \rho_{b,proper}a(t)^3, $$
-The expansion parameter $a \equiv 1/(1 + z)$ follows the expansion of a smooth, homogeneous background, where $z$, the redshift, is a function only of $t$.
-All derivatives are determined with respect to the comoving position $\vecx$, which is defined simply to remove the universal expansion from the coordinate system.
+The expansion parameter $a \equiv 1/(1 + z)$ follows the expansion of
+a smooth, homogeneous background, where $z$, the redshift, is a
+function only of $t$.  All derivatives are determined with respect to
+the comoving position $\vecx$, which is defined simply to remove the
+universal expansion from the coordinate system.
 
-The evolution of $a(t)$ is governed by the formula for the expansion of an isotropic, homogeneous universe:
+The evolution of $a(t)$ is governed by the formula for the expansion
+of an isotropic, homogeneous universe:
 %
@@
 %
-Here, $\rho_0$ is the mean comoving mass density (including both baryonic and dark matter), $p_0$ is the comoving background pressure contribution, and $\Lambda_c$ is the cosmological constant.
+Here, $\rho_0$ is the mean comoving mass density (including both
+baryonic and dark matter), $p_0$ is the comoving background pressure
+contribution, and $\Lambda_c$ is the cosmological constant.
 %$\gamma$ is the ratio of specific heats, and $\Lambda$ is the cosmological constant.
@@
 %eq.~(\ref{eq:state}).
-This system of equations is limited to the non-relativistic regime and assumes that curvature effects are not important --- both assumptions are reasonable as long as the  size of the simulated region is small compared to the radius of curvature and the Hubble length $c/H$ ($c$ is the speed of light and $H$ is the Hubble constant).
+This system of equations is limited to the non-relativistic regime and
+assumes that curvature effects are not important --- both assumptions
+are reasonable as long as the size of the simulated region is small
+compared to the radius of curvature and the Hubble length $c/H$, where
+$c$ is the speed of light and $H = \dot{a}/a$ is the Hubble constant.
 
@@
 the code was written with cosmological situations in mind (e.g.,
-galaxy clusters), in more recent years it has been used to simulate a much
-broader range of astrophysical environments.
+galaxy clusters), in more recent years it has been used to simulate a
+much broader range of astrophysical environments.
 
@@
 
-Any collisionless component (e.g. dark matter, stars) is modeled by particles which are governed by Newton's equations in comoving coordinates:
+Any collisionless component (e.g., dark matter, stars) is modeled by
+particles which are governed by Newton's equations in comoving
+coordinates:
 %
@@
 %
-The particles also make a contribution to the gravitational potential through their contribution to Poisson's equation.
+The particles also make a contribution to the gravitational potential
+through their contribution to Poisson's equation (Equation
+\ref{eq:potential}).
 
@@
 
-In addition, we can solve the mass conservation equations for a set of chemical species and reactions.  For any species $i$ with comoving number density $n_i$, these equations have the form:
+In addition, we can solve the mass conservation equations for a set of
+chemical species and reactions.  For any species $i$ with comoving
+number density $n_i$, these equations have the form:
 \begin{equation}
@@
   + \frac{1}{a} \nabla \cdot (n_i \vecv) = 
-        k_{ij}(T) n_i n_j 
-      + \Gamma^{ph}_j n_j 
+  k_{ij}(T) n_i n_j + \Gamma^{\rm ph}_j n_j 
   \label{eq:species_evolution}
@@
 specifically note the cases where we either include three-body
-reactions or have density-dependent rates).  The $\Gamma^{ph}_j$ are
-destruction/creation rates due to photoionizations and/or
+reactions or have density-dependent rates).  The $\Gamma^{\rm ph}_j$
+are destruction/creation rates due to photoionizations and/or
 photodissociations.  Currently the species we can follow include H,
-H$^+$, He, He$^+$, He$^{++}$, and $e^-$, as well as additional options
-that add H$^-$, H$_2$, H$_2^+$, and  HD, D, and D$^+$.
+H$^+$, He, He$^+$, He$^{++}$, and e$^-$, as well as additional options
+that add H$^-$, H$_2$, H$_2^+$, and HD, D, and D$^+$.  We also can
+track the advection of a comoving metal density that may contribute to
+the radiative cooling.
 
@@
 
-The code can include either a homogeneous radiation background (with possible local self-shielding corrections), or evolve an inhomogeneous radiation field either by directly solving the radiative transfer equation along rays or by solving a set of moment equations derived from the radiative transfer equation.
+The code can include either a homogeneous radiation background (with
+possible local self-shielding corrections), or evolve an inhomogeneous
+radiation field either by directly solving the radiative transfer
+equation along rays or by solving a set of moment equations derived
+from the radiative transfer equation.
 
@@
 specific intensity in units of energy per time $t$ per solid angle per
-unit area per frequency $\nu$.  $H = \dot{a}/a$ is the Hubble
-constant, where $a$ is the scale factor.  $\bar{a} = a/a_{em}$ is the
-ratio of scale factors at the current time and time of emission.  The
-second term represents the propagation of radiation, where the factor
+unit area per frequency $\nu$.  $\bar{a} = a/a_{em}$ is the ratio of
+scale factors at the current time and time of emission.  The second
+term represents the propagation of radiation, where the factor
 $1/\bar{a}$ accounts for cosmic expansion.  The third term describes
@@
 
-As an alternative to the ray-casting strategy, we solve an equation derived from integrating
-the equation of radiative transfer over angles.  In particular, we use a
-flux-limited diffusion approximation for radiative
-transfer, with couplings to both the gas energy and chemical number
-densities. The equations solved are:
+As an alternative to the ray-casting strategy, we solve an equation
+derived from integrating the equation of radiative transfer over
+angles.  In particular, we use a flux-limited diffusion approximation
+for radiative transfer, with couplings to both the gas energy and
+chemical number densities. The equations solved are:
 \begin{eqnarray}
@@
 \end{eqnarray}
-where $E_r = E_r(x,t)$ is a grey radiation energy density and $e_c$ is the
-internal energy correction due to photo-heating and chemical cooling.
+where $E_r = E_r(\vecx,t)$ is a grey radiation energy density and
+$e_c$ is the internal energy correction due to photo-heating and
+chemical cooling.
 %and ${\mathrm n}_i$ correspond to the chemical number densities of
@@
 %number density.  
-Here, we define the grey radiation energy density
-through first assuming a fixed frequency spectrum, i.e.
-$E_{\nu}(\nu,\vecx,t) = \tilde{E}_r(\vecx,t) \chi(\nu)$, and then
-defining the integrated quantity
+Here, we define the grey radiation energy density through first
+assuming a fixed frequency spectrum, i.e.  $E_{\nu}(\nu,\vecx,t) =
+\tilde{E}_r(\vecx,t) \chi(\nu)$, and then defining the integrated
+quantity
 \begin{equation}
@@
 \end{equation}
-In addition, $D$ in (\ref{eq:fld_radiation}) is the {\em Larsen
-square-root flux-limiter} \citep[see][]{Morel2000}, $\kappa$ is the total
-opacity, $\eta$ is the field of radiation sources, $\Gamma$ provides the
-radiation induced photo-heating, and $\Lambda$ the chemical
-cooling.
+In addition, $D$ in Equation (\ref{eq:fld_radiation}) is the Larsen
+square-root flux-limiter \citep[see][]{Morel2000}; $\kappa$ is the
+total opacity; $\eta$ is the field of radiation sources; $\Gamma$
+provides the radiation induced photo-heating, and $\Lambda$ the
+chemical cooling.
 %, $\alpha_{i,j}$ are the reaction rate coefficients defining
@@
 similar to that of \citet{2007ApJ...664..135P}, where we model the
-flux of isotropic heat transport:
-
+flux of isotropic heat transport,
+%
 \begin{equation}
-\fcond = -\kappa_{sp} \nabla T
+\fcond = -\kappa_{sp} \nabla T.
 \end{equation}
-
-Where $\kappa_{sp}$ is the Spitzer conduction coefficient
+%
+Here $\kappa_{sp}$ is the Spitzer conduction coefficient
 \citep{1962pfig.book.....S} and T is the baryon temperature (with
-fluids explicitly assumed to be single-temperature).  Saturation of the heat flux in
-high-temperature, low-density regimes (such as the intracluster medium
-in galaxy clusters) is taken into account.  Thermal conduction in a
-plasma can be strongly affected by the presence of magnetic field
-lines, which may strongly suppress heat flow perpendicular to the
-magnetic field.  In that case, we allow for heat transport only along
-the magnetic field lines in simulations that include this physics, as follows:
+fluids explicitly assumed to be single-temperature).  Saturation of
+the heat flux in high-temperature, low-density regimes (such as the
+intracluster medium in galaxy clusters) is taken into account.
+Thermal conduction in a plasma can be strongly affected by the
+presence of magnetic field lines, which may strongly suppress heat
+flow perpendicular to the magnetic field.  In that case, we allow for
+heat transport only along the magnetic field lines in simulations that
+include this physics, as follows:
 
@@
 
-Where $\hat{\textbf{b}}$ is the unit vector denoting the direction of the
-magnetic field.  In both the isotropic and anisotropic limits, we
+Where $\hat{\textbf{b}}$ is the unit vector denoting the direction of
+the magnetic field.  In both the isotropic and anisotropic limits, we
 assume that conductivity can be suppressed relative to the ideal
@@
 In this section, we briefly describe the numerical methods which are
-used to solve the equations outlined in Section~\ref{sec.equations}.  We
-proceed through the numerical methods in the same order as will be used in
-Sections~\ref{sec.amr} through~\ref{sec.num.analysis} so that there is a one-to-one correspondence
-between each of the following overviews and the complete description
-provided in the later sections.  The goal in this section is to introduce the reader to
-the basic principles of the methods without drowning in detail.
+used to solve the equations outlined in Section~\ref{sec.equations}.
+We proceed through the numerical methods in the same order as will be
+used in Sections~\ref{sec.amr} through~\ref{sec.num.analysis} so that
+there is a one-to-one correspondence between each of the following
+overviews and the complete description provided in the later sections.
+The goal in this section is to introduce the reader to the basic
+principles of the methods without drowning in detail.
 
@@
 
-The primary purpose of the \enzo\ code is its Adaptive Mesh Refinement (AMR)
-capability, which allows it to reach extremely large dynamical ranges
-with limited computational resources, opening doors previously closed
-by finite memory and computational time. Unlike moving mesh methods
-\citep{1995ApJS..100..269P,1995ApJS...97..231G} or  
-methods that subdivide 
-individual cells \citep{Adjerid}, Berger \& Collela's AMR (also referred 
-to as \emph{structured} AMR) utilizes an adaptive hierarchy of grid 
-patches at varying levels of resolution.  Each rectangular grid patch 
-(referred to as a ``grid'') covers some region of space in its 
-\emph{parent grid} which requires higher resolution, and can itself 
-become the parent grid to an even more highly resolved \emph{child grid}.
+The primary purpose of the \enzo\ code is its Adaptive Mesh Refinement
+(AMR) capability, which allows it to reach extremely large dynamical
+ranges with limited computational resources, opening doors previously
+closed by finite memory and computational time. Unlike moving mesh
+methods \citep{1995ApJS..100..269P,1995ApJS...97..231G} or methods
+that subdivide individual cells \citep{Adjerid}, Berger \& Collela's
+AMR (also referred to as \emph{structured} AMR) utilizes an adaptive
+hierarchy of grid patches at varying levels of resolution.  Each
+rectangular grid patch (referred to as a ``grid'') covers some region
+of space in its \emph{parent grid} which requires higher resolution,
+and can itself become the parent grid to an even more highly resolved
+\emph{child grid}.
 
@@
 % finer grids and `above' for coarser grids). 
- We restrict the ratio
-between cell sizes to be an integer, typically 2 or 4, and refer to a
-level as all grids with the same cell size.  In order to keep
-things as simple as possible, the edges of subgrids must coincide with
-the cell edge of its immediate parent (coarser) grid. Additionally,
-the hierarchy can be initialized with one or more static grids if a
-higher initial resolution is required.
+We restrict the ratio between cell sizes to be an integer, typically 2
+or 4, and refer to a level as all grids with the same cell size.  In
+order to keep things as simple as possible, the edges of subgrids must
+coincide with the cell edge of its immediate parent (coarser)
+grid. Additionally, the hierarchy can be initialized with one or more
+static grids if a higher initial resolution is required.
 
@@
 advance all the grids on that level by a timestep $\Delta t_1$
-($\Delta t_1 \leq \Delta t_0$), which is the minimum of all the allowed
-timesteps for those grids.  If there are more levels, we repeat this
-procedure until the bottom level of the hierarchy has been reached.
-Once there, we continue advancing the grids on the lowest level until
-they have `caught up' to the coarser parent grids (i.e. $\sum
-\Delta t_l = \Delta t_{l-1}$).  This procedure repeats itself until
-all grids have been advanced by a total time of $\Delta t_0$ (see
-Figure~\ref{fig:wcycle} for a schematic).
+($\Delta t_1 \leq \Delta t_0$), which is the minimum of all the
+allowed timesteps for those grids.  If there are more levels, we
+repeat this procedure until the bottom level of the hierarchy has been
+reached.  Once there, we continue advancing the grids on the lowest
+level until they have ``caught up'' to the coarser parent grids
+(i.e. $\sum \Delta t_l = \Delta t_{l-1}$).  This procedure repeats
+itself until all grids have been advanced by a total time of $\Delta
+t_0$ (see Figure~\ref{fig:wcycle} for a schematic).
+
+\begin{figure}
+\begin{center}
+%\includegraphics[width=0.6\textwidth]{figures/wcycle.eps}
+\includegraphics[width=0.6\textwidth]{figures/timestepping.eps}
+\end{center}
+\caption{\emph{Left:} Example of the timesteps on a 2-level AMR
+  hierarchy.  Enzo does not restrict the timesteps on the finer levels
+  to be a factor of $1/2^n$ smaller.  \emph{Right:} The order in which
+  the AMR grids are evaluated on each level.\vspace{1ex}}
+\label{fig:wcycle}
+\end{figure}
 
@@
 adapt itself.  We do this whenever a level has caught up to the
-coarser parent level and consists of entirely rebuilding the grids
-on that level and at finer resolutions.  This is done by applying the grid refinement
-criteria to the grids on that level, flagging zones that require
-extra grids.  These criteria depend on the physical problem being
-simulated -- see Section~\ref{sec:refinement_criteria} for more details.  Once a grid has a set of flagged cells,
-we run a machine-vision based algorithm \citep{Berger91} to find edges
-and determine a good placement of subgrids. 
+coarser parent level and consists of entirely rebuilding the grids on
+that level and at finer resolutions.  This is done by applying the
+grid refinement criteria to the grids on that level, flagging zones
+that require extra grids.  These criteria depend on the physical
+problem being simulated -- see Section~\ref{sec:refinement_criteria}
+for more details.  Once a grid has a set of flagged cells, we run a
+machine-vision based algorithm \citep{Berger91} to find edges and
+determine a good placement of subgrids.
 %  These subgrids must not overlap one another, must cover all flagged
@@
 % flagged cells to total cells. 
-Once these new subgrids have been identified, the
-solution from the next coarser grid is interpolated (see below) in
-order to initialize the values on the new grids.  Finally, any overlap
-between these new subgrids and the old ones is identified and the
-solution within the regions of overlap is copied to the new subgrids.
-The entire procedure just outlined is then repeated on the new grids
-and in this way the entire hierarchy (from the original level examined
-and continuing on to finer levels) is rebuilt.
+Once these new subgrids have been identified, the solution from the
+next coarser grid is interpolated in order to initialize the values on
+the new grids.  Finally, any overlap between these new subgrids and
+the old ones is identified, and the solution within the regions of
+overlap is copied to the new subgrids.  The entire procedure just
+outlined is then repeated on the new grids and in this way the entire
+hierarchy (from the original level examined and continuing on to finer
+levels) is rebuilt.
 
@@
 developed by~\citet{1984JCoPh..54..174C} and extended to cosmology
-by~\citet{1995CoPhC..89..149B}; (ii) the MUSCL-like Gudonov scheme
+by~\citet{1995CoPhC..89..149B}; (ii) the MUSCL-like Godunov scheme
 \citep{MUSCL} with or without magnetic fields and Dedner-based
@@
 % \item 
- (i) Construct monotonic parabolic interpolation of cell average data, for each fluid quantity.
+(i) Construct monotonic parabolic interpolation of cell average data,
+for each fluid quantity.
 % \item 
-(ii) Compute interface states by averaging the parabola over the domain of dependence for each interface.
+(ii) Compute interface states by averaging the parabola over the
+domain of dependence for each interface.
 % \item 
@@
  %\item 
- (iv) Difference the interface fluxes to update the cell average quantities.
+(iv) Difference the interface fluxes to update the cell average
+quantities.
 %\end{enumerate}
@@
 order-accurate in space and time and which explicitly conserves mass,
-linear momentum, and energy.  \red{(REF?)}  A variety of reconstruction methods and
-Riemann solvers have been implemented.  \red{(These are never described!)}
+linear momentum, and energy.  \red{(REF?)}  A variety of Riemann
+solvers have been implemented.
 
@@
 \subsubsection{Godunov MUSCL (HD) with Dedner divergence cleaning (MHD)}
-This solver was developed by Wang \& Abel to attack problems in
-magnetic field amplification during the formation of galaxies
-\citep{Wang:2009a} and to understand the role of proto-stellar jets for
-the theory of star formation \citep{Wang:2009b}. It combines the standard
-approach of Godunov \citep{Godunov1959} for finite volume techniques
-with the method of lines as described by
-\cite{leveque2002finite} and \cite{toro-1997}. In addition, 
-it implements the hyperbolic divergence cleaning algorithm of
-\cite{2002JCoPh.175..645D}. It supports multiple approximate Riemann
-solvers and non-ideal equations of states. Consequently,
-these suite of solvers can be used for hydro and magneto-hydrodynamic
-simulations. It is also this class of solvers as well as a version of
-the PPM hydro solver which have been ported to Nvidia's CUDA framework
-allowing \enzo\ to take advantage of modern graphics hardware \citep{Wang:2010}.
+This solver was developed to attack problems in magnetic field
+amplification during the formation of galaxies \citep{Wang:2009a} and
+to understand the role of proto-stellar jets for the theory of star
+formation \citep{Wang:2009b}. It combines the standard approach of
+Godunov \citep{Godunov1959} for finite volume techniques with the
+method of lines as described by \cite{leveque2002finite} and
+\cite{toro-1997}. In addition, it implements the hyperbolic divergence
+cleaning algorithm of \cite{2002JCoPh.175..645D}. It supports multiple
+approximate Riemann solvers and non-ideal equations of
+states. Consequently, these suite of solvers can be used for hydro and
+magneto-hydrodynamic simulations. It is also this class of solvers as
+well as a version of the PPM hydro solver which have been ported to
+Nvidia's CUDA framework allowing \enzo\ to take advantage of modern
+graphics hardware \citep{Wang:2010}.
 
 \subsubsection{MHD with Constrained Transport (MHD)}
-This MHD method is second order in time and space, and preserves the divergence constraint,
-$\nabla \cdot \vecB = 0$, to machine precision through the Constrained Transport
-(CT) method.  CT, originally described by \citet{Evans88}, updates the
-magnetic field with the curl of an electric field, suitably formulated to
-preserve $\nabla \cdot \vecB$.  We employ the CT methods
-described by \citet{Balsara99} and \citet{Gardiner05}.  Further, we employ the
-second order hyperbolic solver of \citet{Li08a} and the divergence free AMR
-scheme of \citet{Balsara01}.
+This MHD method is second order in time and space, and preserves the
+divergence constraint, $\nabla \cdot \vecB = 0$, to machine precision
+through the Constrained Transport (CT) method.  CT, originally
+described by \citet{Evans88}, updates the magnetic field with the curl
+of an electric field, suitably formulated to preserve $\nabla \cdot
+\vecB$.  We employ the CT methods described by \citet{Balsara99} and
+\citet{Gardiner05}.  Further, we employ the second order hyperbolic
+solver of \citet{Li08a} and the divergence free AMR scheme of
+\citet{Balsara01}.
 
@@
 
-The ZEUS method uses a staggered mesh such that the velocities are
+The \zeus\ method uses a staggered mesh such that the velocities are
 face-centered, while the density and internal energy quantities are
@@
 Fourier Technique \citep{Hockney88} to solve Poisson's equation on the
-coarsest (Level 0) grid on each time step.  The advantage of using this method is that
-it is fast, accurate, and 
-naturally allows both periodic and isolated boundary conditions for
-the gravity, choices which are very common in astrophysics and
-cosmology.  On subgrids, we interpolate the boundary conditions from
-that grid's parent (either the level 0 grid or some other subgrid).
-The Poisson equation is then solved on every timestep using a multigrid technique on one
-subgrid at a time.  In
-\enzo, self-gravity is optional.  Aside from calculating the
-gravitational potential from the baryon fields and particles existing
-in the simulation, there are also a number of options for static gravitational fields.
+coarsest (Level 0) grid on each time step.  The advantage of using
+this method is that it is fast, accurate, and naturally allows both
+periodic and isolated boundary conditions for the gravity, choices
+which are very common in astrophysics and cosmology.  On subgrids, we
+interpolate the boundary conditions from that grid's parent (either
+the level 0 grid or some other subgrid).  The Poisson equation is then
+solved on every timestep using a multigrid technique on one subgrid at
+a time.  In \enzo, self-gravity is optional.  Aside from calculating
+the gravitational potential from the baryon fields and particles
+existing in the simulation, there are also a number of options for
+static gravitational fields.
 
@@
 
-Collisionless matter (dark matter, stars, etc.) is modeled with
+Collisionless matter (e.g., dark matter, stars, etc.) is modeled with
 particles that interact with the baryons only via gravity.  These
 particles are advanced in a single time step using a drift-kick-drift
-algorithm\citep{Hockney88} to provide second-order accuracy even in the presence of
-changing timesteps.  Since the particles follow the collapse of
-structure, they are not adaptively refined.  Nor are there duplicate
-sets of particles for each level; instead, each particle is associated
-with the highest refined level available at its position in space, and is
-moved as the hierarchy is rebuilt.  Thus, a particle has the same
-timestep and feels the same gravitational force as a grid at that refinement
-level.
+algorithm \citep{Hockney88} to provide second-order accuracy even in
+the presence of changing timesteps.  Since the particles follow the
+collapse of structure, they are not adaptively refined.  Nor are there
+duplicate sets of particles for each level; instead, each particle is
+associated with the highest refined level available at its position in
+space and is moved as the hierarchy is rebuilt.  Thus, a particle has
+the same timestep and feels the same gravitational force as a grid at
+that refinement level.
 
 Although the particles are fixed in mass once initialized, we can
-create them with any initial set of masses and positions.  For example, in
-many cosmological simulations, a static subgrid is included from the
-beginning in order to improve the initial baryonic mass resolution.
-On this subgrid, we also use smaller particles to improve the
-collisionless mass resolution.  One particle per initial grid point
-seems to provide approximately equal sampling between the dark matter
-and gas.
+create them with any initial set of masses and positions.  For
+example, in many cosmological simulations, a static subgrid is
+included from the beginning in order to improve the initial baryonic
+mass resolution.  On this subgrid, we also use smaller particles to
+improve the collisionless mass resolution.  One particle per initial
+grid point seems to provide approximately equal sampling between the
+dark matter and gas.
 
@@
 
-\enzo\ includes the capability of following up to
-12 chemical species using a non-equilibrium solver.  The species can
-be turned on in sets, with the simplest model including just H, H$^+$, 
-He, He$^+$, He$^++$, and $e^-$, and more complete models adding first
-species important for gas-phase molecular hydrogen formation: $H^-$,
-H$_2$ and H$_2^+$, and then $HD$ formation: $HD$, $D$, $D+$.  (We note
-that a reduced version of the chemistry, used only with the
-flux-limited diffusion radiation transport, includes only hydrogen
-species.)  The
-cooling and heating due to these species is also included (see the
-next section).  The solution of the rate equations is carried out
-using one Jacobi iteration of an implicit Euler time discretization to
-ensure stability; to
-ensure accuracy we sub-cycle the rate equations with a shorter
-timestep such that the electron and neutral fractions do not change by
-more than 10\%.
+\enzo\ includes the capability of following up to 12 chemical species
+using a non-equilibrium solver.  The species can be turned on in sets,
+with the simplest model including just H, H$^+$, He, He$^+$, He$^{++}$,
+and e$^-$, and more complete models adding first species important for
+gas-phase molecular hydrogen formation: H$^-$, H$_2$ and H$_2^+$, and
+then HD formation: HD, D, D$^+$.  (We note that a reduced version
+of the chemistry, used only with the flux-limited diffusion radiation
+transport, includes only hydrogen species.)  The cooling and heating
+due to these species is also included (see the next section).  The
+solution of the rate equations is carried out using one Jacobi
+iteration of an implicit Euler time discretization to ensure
+stability; to ensure accuracy we sub-cycle the rate equations with a
+shorter timestep such that the electron and neutral fractions do not
+change by more than 10\%.
 
@@
 chemistry is turned on, then the code can include cooling from all
-species of Hydrogen and Helium (including H$_2$) -- this primordial cooling is
-also treated using 
-one Jacobi iteration of an implicit Euler time discretization to maintain accuracy.  Finally, we have also recently
-added metal cooling based on a set of multi-dimensional lookup tables
-computed with the \texttt{Cloudy} code \citep{1998PASP..110..761F} as
-described in \citet{2008MNRAS.385.1443S} and
-\citet{2011ApJ...731....6S}.  Finally, we note that the cooling and
-heating is (almost always) treated in the optically-thin limit,
-although the code can also follow radiative transfer in a limited set
-of energy bins (see below).
+species of hydrogen and helium (including H$_2$) -- this primordial
+cooling is treated in the same Jacobi iteration as the chemistry.
+Finally, we have also recently added metal cooling based on a set of
+multi-dimensional lookup tables computed with the \textsc{Cloudy} code
+\citep{1998PASP..110..761F} as described in
+\citet{2008MNRAS.385.1443S} and \citet{2011ApJ...731....6S}.  Finally,
+we note that the cooling and heating is (almost always) treated in the
+optically-thin limit, although the code can also follow radiative
+transfer in a limited set of energy bins.
 
@@
 that is based on an adaptive ray-tracing method that utilizes the
-HEALPix pixilation of the sphere \citep{AbelWandelt02}.  Photons are
+HEALPix pixelization of a sphere \citep{AbelWandelt02}.  Photons are
 integrated outwards from sources using an adaptive time-stepping

### 2013-05-29 tests-shockpool.tex
Editing the shockpool test. Mostly small fixes.

@@
 
-The next problem is similar to the last one, but examines how a shock wave with a Mach number of 2 passes into and out of a static refined region (again defined from $z=0.25$ to 0.75).
-The density and pressure in the domain is initially set equal to 1.0, with zero velocity, and then at $t=0$, the left boundary is set with the density, pressure and velocity appropriate for $M=2$ shock wave.
+The next problem is similar to the last one but instead examines how a shock with a Mach number of 2 passes in and out of a static refined region (again defined from $z=0.25$ to 0.75).
+The density and pressure in the domain is initially set equal to 1.0 with zero velocity.
+At $t=0$ the left boundary is set with the density, pressure and velocity appropriate for $\mathcal{M}=2$ shock wave.
 
@@
 
-Beginning with PPM (top row), we see that this method captures the shock in a small number of zones with only a small amount of oscillation.
-Open entering the refined region, the shock finds itself broader than the natural width of the scheme (since the cell spacing is decreased by a factor of 2), and so the shock front contracts, causing a slight entropy perturbation in the post-shock gas.
+Beginning with PPM (top row of Figure \ref{fig.shockpool}), we see that this method captures the shock in a small number of zones with only a small amount of oscillation.
+Upon entering the refined region, the shock finds itself broader than the natural width of the scheme (since the cell spacing is decreased by a factor of 2), and so the shock front contracts, causing a slight entropy perturbation in the post-shock gas.
 
-For Zeus (middle row of Figure~\ref{fig.shockpool}, the shock is broadened because of artificial viscosity, and there are slightly more post-shock oscillations, although again quite mild.
-The impact of entering and exiting the refined region is somewhat larger than in PPM; however, the most noticeable difference is the incorrect position of the shock front, which is due to the fact that the scheme is not energy-conserving (see also the Sedov problem, below), and has essentially nothing to do with the passage through the refined region.
+For Zeus (middle row of Figure~\ref{fig.shockpool}, the shock is broadened because of artificial viscosity and there are slightly more post-shock oscillations, although again quite mild.
+The impact of entering and exiting the refined region is somewhat larger than in PPM; however, the most noticeable difference is the incorrect position of the shock front.
+This is due to the fact that the scheme is not energy-conserving (see also the Sedov problem, below), and has very little to do with the passage through the refined region.
 
@@
 \includegraphics[width=0.8\textwidth]{figures/ShockPool}
-\caption{This plot shows, for each column, three snap shots (at $t=0.1, 0.15$ and 0.35) of a $M=2$ shock wave as it propagates through the grid with a static refined region extends from $x = 0.25$ to 0.75 (shown in grey in each panel).
+\caption{This plot shows, for each column, three snapshots at $t=0.1, 0.15$ and 0.35 of a $\mathcal{M}=2$ shock as it propagates through the grid with a static refined region extends from $x = 0.25$ to 0.75 (shown in grey in each panel).
 Each row shows the result for a different solver (PPM/Zeus/MUSCL from top to bottom).

### 2013-05-29 tests-aniso-conduct.tex
Removed introductory paragraph.

@@
 
-In a plasma, conduction takes place primarily due to electron motion.
-In astrophysical environments such as the intergalactic and
-intracluster media, with typical magnetic strengths of $\sim 1$~$\mu$G
-and temperatures in the millions of Kelvin, the electron gyroradius is
-very small compared to the scales of interest in a typical
-cosmological simulation.  As a result, electrons can only move
-\textit{along} magnetic field lines, not perpendicular to them, which
-can result in very interesting magnetohydrodynamical instabilities \protect \citep[e.g.,][]{2008ApJ...677L...9P,2008ApJ...688..905P}. The correct
-modeling of this behavior can be quite challenging, and is described
-in Section~\protect\ref{sec.num.conductions}. Conduction in \protect\enzo\ is calculated in both an operator-split and directionally-split manner,
-and the addition of heat transport solely along magnetic field lines
-requires computation of cross-terms in the temperature derivative at
-cell faces, which can add spurious oscillations in the temperature
-field in regions where the temperature gradient is strong in more than
-one spatial direction unless an appropriate flux-limiter \protect\citep[such as that of][]{1977JCoPh..23..263V} is chosen.
-\mqk{The previous paragraph more or less repeats what's in Section~\ref{sec.num.conductions}, I think it should be removed (or merged with that section).}
-
 Figure~\ref{fig.conduct} shows a test that demonstrates the correct
-behavior of anisotropic thermal conduction in \enzo.  We initialize a
+behavior of anisotropic thermal conduction (see
+Section~\ref{sec.num.conductions}) in \enzo. We initialize a
 two-dimensional, $256 \times 256$ cell simulation having a physical
-scale of 1 kpc on a side, a uniform density of 1 proton cm$^{-3}$ and a
-background temperature of $10^6$ Kelvin.  Magnetic fields with a
+scale of 1 kpc on a side, a uniform density of 1 proton cm$^{-3}$ and
+a background temperature of $10^6$ Kelvin.  Magnetic fields with a
 strength of B$_0 = 1$~$\mu$G are initialized such that the field lines

### 2013-05-29 numerical-chemistry.tex

@@
 
-While it is often safe to assume that species (both chemical and
-ionization) within a gas can be treated as being in equilibrium, in
-some regimes that are found in astrophysics this assumption leads to
-considerable error.  For example, the cooling and collapse of
-primordial gas in Population III star formation is dominated by
-molecular hydrogen, which in the absence of dust forms via an
-inefficient pair of collisional processes that depend heavily on the
-local population of free electrons.  As a result, when modeling
-primordial star formation, it is critical to follow the
-non-equilibrium evolution of the chemical species of hydrogen,
-including molecular hydrogen and deuterium.
+While it is often safe to assume that species (both chemical and ionization)
+within a gas can be treated as being in equilibrium, in some regimes that are
+found in astrophysics this assumption leads to considerable error.  For
+example, the cooling and collapse of primordial gas in Population III star
+formation is dominated by molecular hydrogen, which in the absence of dust
+forms via an inefficient pair of collisional processes that depend heavily on
+the local, highly non-equilibrium population of free electrons.  As a result,
+when modeling primordial star formation, it is critical to follow the
+non-equilibrium evolution of the chemical species of hydrogen, including
+molecular hydrogen and deuterium.
 
-The primordial non-equilibrium chemistry routines used in \enzo\ were
-first described by Abel et al. and Anninos et
-al. \citep{abel97,anninos97}, but have been extended since then with
-updated reaction rates and the inclusion of deuterium species
-\citep{2009PhDT.........5T}.  These routines follow the
-non-equilibrium chemistry of a gas of primordial composition with 12
-total species: H, \Hp, He, \Hep, \Hepp, \Hm, \HHp, \HH, e$^-$, D, \Dp,
-and HD.  (A separate, but related, module follows the radiative
-heating and cooling of the gas from atomic and molecular line
-excitation, recombination, collisional excitation, free-free
-transitions, Compton scattering of the cosmic microwave background, as
-well as several models for a metagalactic UV background that heat the
-gas via photoionization and photodissociation; see
-Section~\ref{sec.num.cooling} for more details).  Depending on user
-parameters, one can follow only atomic species (H, \Hp, He, \Hep,
-\Hepp, and e$^-$), include those species relevant for molecular
-hydrogen formation (\HH, \HHp, and \Hm), and can further include
-deuterium and its species (D, \Dp, and HD).  A total of 9 kinetic
-equations are solved, including 29 kinetic and radiative processes,
-for the 12 species mentioned above.  See Table~\ref{table.collisional}
-for the collisional processes and Table~\ref{table.radiative} for the
-radiative processes solved.
+The primordial non-equilibrium chemistry routines used in \enzo\ were first
+described by Abel et al. and Anninos et al. \citep{abel97,anninos97}, but have
+been extended since then with updated reaction rates and the inclusion of
+deuterium species \citep{2009PhDT.........5T}.  These routines follow the
+non-equilibrium chemistry of a gas of primordial composition with 12 total
+species: H, \Hp, He, \Hep, \Hepp, \Hm, \HHp, \HH, e$^-$, D, \Dp, and HD.  Enzo
+also computes the radiative heating and cooling of the gas from atomic and
+molecular line excitation, recombination, collisional excitation, free-free
+transitions, Compton scattering of the cosmic microwave background, as well as
+several models for a metagalactic UV background that heat the gas via
+photoionization and photodissociation; see Section~\ref{sec.num.cooling} for
+more details.  The chemical and thermal states of the gas can be updated either
+at the same hydrodynamical timestep (i.e., de-coupled and operator split) or
+through the same subcycling system (i.e., a coupled chemical and thermal
+system).  The default behavior of \enzo\ is to couple these two systems at
+subcycles of the hydrodynamic timestep; this results in updates to both the
+chemical and thermal states of the gas (which also inform the temperature, the
+reaction rate coefficients and the cooling functions of the gas) on timescales
+that faster than those of the gas dynamics.
 
-The chemical reaction equation network is technically challenging to
-solve due to the huge range of reaction timescales involved -- the
-characteristic creation and destruction timescales of the various
-species and reactions can differ by many orders of magnitude.  As a
-result, the set of rate equations is extremely stiff, and an explicit
-scheme for integration of the rate equations can be exceptionally
-costly if small enough timesteps are taken to keep the network stable.
-This makes an implicit scheme strongly preferable for such a set of
-equations.  However, an implicit scheme typically requires an iterative
-procedure to converge, and for large networks (such as this one) an
-implicit method can be very time-consuming, making it undesirable for
-a large, three-dimensional simulation.
+Input parameters to \enzo\ govern the chemical species that are updated during
+the course of the simulation.  This can include only the atomic species (H,
+\Hp, He, \Hep, \Hepp, and e$^-$), those species relevant for molecular hydrogen
+formation (\HH, \HHp, and \Hm), and can further include deuterium and its
+species (D, \Dp, and HD).  A total of 9 kinetic equations are solved, including
+29 kinetic and radiative processes, for the 12 species mentioned above.  See
+Table~\ref{table.collisional} for the collisional processes and
+Table~\ref{table.radiative} for the radiative processes solved.
 
-\enzo\ solves the rate equations using a method based on a first-order-accurate, implicit
-Euler formulation in order to provide a stable and accurate
-solution. This technique is optimized by taking the chemical
-intermediaries \Hm and \HHp, which have large rate coefficients and
-low concentrations, and grouping them into a separate category of
-equations.  Due to their fast reactions, these species are very
-sensitive to small changes in the more abundant species.  However, due
-to their low overall concentrations, they do not significantly
-influence the abundance of species with higher concentrations.
-Therefore, reactions involving these two species can be decoupled from
-the rest of the network and treated independently.  In fact, \Hm~and
-\HHp~are treated as being in equilibrium at all times, independent of
-the other species and the hydrodynamic variables.  This allows a large
-speedup in solution as the implicit Euler scheme is then applied only to the
-slower 7- or 10-species network (depending on whether deuterium is
-included or not) on timescales closer to those required by the
-hydrodynamics of the simulation.  Even so, the accuracy and stability
-of the scheme is maintained by subcycling the rate solver over a
-single hydrodynamic timestep.  These subcycle timesteps are determined
-so that the estimated fractional change in the electron concentration
-is limited to no more than $10\%$ per timestep.  Given this
-subcycling, the method is not truly implicit over an entire
-hydrodynamical timestep  -- however,
-it uses the formalism and is implicit over subcycle timesteps.
+The chemical reaction equation network is technically challenging to solve due
+to the huge range of reaction timescales involved.  The characteristic times
+for creation and destruction of the various species and reactions can differ by
+many orders of magnitude and are often very sensitive to the chemical and
+thermal state of the gas.  This makes a fully-implicit scheme, with convergence
+criteria and error tolerance, strongly preferable for such a set of equations.
+However, most implicit schemes require an iterative procedure to converge, and
+for large networks (such as this one) an iterative, fully-implicit method can
+be very time-consuming and computationally costly for a relatively small
+increase in accuracy.  At this present time, this makes fully-implicit methods
+somewhat undesirable for a large, three-dimensional simulation.
+
+\enzo\ solves the rate equations using a method based on a semi-implicit
+formulation in order to provide a stable, positive definite and first-order
+accurate solution.  The update discretization splits chemical changes into
+formation components and destruction components and updated with a mixed set of
+time states, as described in \citet{anninos97}.  The formation components of
+species $S_i$ are computed at the current subcycle time, where the contribution
+of species $S_i$ to its own destruction components are computed at the updated
+time; all other contributions to the destruction component are computed at the
+current time.  This mixed state improves accuracy and ensures species values
+are positive definite, and is equivalent to one Jacobi iteration of an implicit
+Euler solve.  This technique is optimized by taking the chemical intermediaries
+\Hm and \HHp, which have large rate coefficients and low concentrations, and
+grouping them into a separate category of equations.  Due to their fast
+reactions, these species are very sensitive to small changes in the more
+abundant species and are at all times in astrophysical calculations close to
+equilibrium values.  Attempting to resolve their formation and destruction
+times would necessitate extremely small timesteps, and while these species
+provide necessary channels to facilitate the formation of molecular hydrogen at
+low densities, they do not otherwise significantly influence the abundance of
+species with higher concentrations.  Therefore, reactions governing these two
+species can be decoupled from the rest of the network and treated independently
+through analytic solutions for equilibrium values.  This allows a large speedup
+in solution as the timestepping scheme is applied only to the slower 7- or
+10-species network (depending on whether deuterium is included or not), which
+will be much closer to the overall hydrodynamic timestep of the simulation.
+
+Even so, the accuracy and stability of the scheme is maintained by subcycling
+the rate solver within a single hydrodynamic timestep.  These subcycle
+timesteps are determined so that the estimated fractional change in the
+electron concentration is limited to no more than $10\%$ per timestep;
+additional criteria may be applied based on the expected change from radiative
+cooling, chemical heating from the formation of molecular hydrogen.
 
@@
 \citet{2004MNRAS.348.1019R}, and was put into \enzo\ for the work
-published in \citep{2009Sci...325..601T,2009PhDT.........5T}.  Even
-with these modifications, a completely correct description of the
-cooling of primordial gas requires some form of radiation transport,
-which will greatly increase the cost of the simulations.
+published in \citep{2009Sci...325..601T,2009PhDT.........5T}.   The formation
+of molecular hydrogen as catalyzed by dust was recently added to \enzo\ to
+enable studies of low-metallicity gas, as well as the inclusion of appropriate
+timestepping criteria to account for the input of ionizing radiation.  Even
+with these modifications, a completely correct description of the cooling of
+primordial gas at very high densities requires some form of radiation
+transport, which will greatly increase the cost of the simulations.
+Furthermore, at very high densities, the stiffness of the molecular hydrogen
+reaction rates may require better than a first-order accurate solution; as
+such, the transition to this regime will likely necessitate a fully-implicit,
+iterative solver.
 

### 2013-05-29 numerical-radiation-fld.tex

@@
 a field-based radiation transport solver for problems posed on uniform
-(i.e.~non-AMR) grids, which has been tuned for large-scale simulations
+(i.e.~non-AMR, non-SMR) grids, which has been tuned for large-scale simulations
 involving many ionizing sources.  Detailed explanations of the model and

### 2013-05-29 tests-doublemach.tex

@@
 The double mach reflection test is a classic two-dimensional test of
-hydrodynamic algorithms, and was originally described in
+hydrodynamic algorithms originally described in
 \citet{1984JCoPh..54..115W} \citep[and more recently
@@
 In the test problem shown in Figure~\ref{fig.doublemach}, a 2D
-simulation with $960 \times 240$ cells was formed with a domain of $x
+simulation with $960 \times 240$ cells was created with a domain of $x
 = [0, 4]$ and $y = [0, 1]$.  We use an ideal gas equation of state of
@@
 runs until t$ = 0.205$ (arbitrary units), at which point the rightmost
-extent of the shock should be at roughly x$ = 3$.  In this run, we use
-the direct Eulerian implementation of the piecewise parabolic hydro
-method, with the diffusion, flattening, and shock steepening all
-turned on.
+extent of the shock should be at roughly x$ = 3$.  In this simulation we use
+the direct Eulerian implementation of the piecewise parabolic hydrodynamic
+method with the diffusion, flattening, and shock steepening all enabled.
 

### 2013-05-29 tests-vandv.tex

@@
 \citet{2007MNRAS.380..963A} and \citet{Tasker2008} perform
-head-to-head comparisons between several grid- and particle-based
+direct comparisons between several grid- and particle-based
 codes for a variety of fluid-centric test problems (including shocked gas clouds,
@@
 less diffusive solutions than lower-order codes, including those that
-use artificial viscosity.  Finally, two code comparison projects have
-been undertaken to examine turbulence in hydro codes -- one studied
-the behavior of decaying isothermal supersonic turbulence
-\citep{2009A&A...508..541K}, and the other looked at supersonic
+use artificial viscosity.  Finally, there have been two code comparison
+projects that focus on turbulence simulations.  The first studied the
+behavior of decaying isothermal supersonic turbulence
+\citep{2009A&A...508..541K}, and the second examined supersonic
 magnetohydrodynamical turbulence \citep{2011ApJ...737...13K}.  Both
@@
 grid-based codes that use Godunov-based fluid solvers, and typically
-had better effective resolution than particle-based codes using the
-same number of particles as the \enzo\ simulation used grid cells.
+had better effective resolution than particle-based codes when using the
+same number of particles as the number of grid cells in the \enzo\ simulation.
 
@@
 measured against several test problems by \citet{IlievEtAl2009},
-producing results that are similar to both expected results (i.e.,
-analytic test problems) and other methods (i.e., codes using other
-methods, written by other groups), although there are some minor
-differences (although all codes in the study exhibited differences
-from the majority of other codes on at least a subset of the tests).
+which involved tests with and without analytical solutions.  \enzo\ produced
+results similar to both the analytical solutions and results obtained by 
+other codes.  We note, however, that there were minor differences throughout
+the comparison between various codes, and the majority of the codes differred
+in at least a subset of the tests.
 \citet{2011ApJ...726...55T} shows the results of varied reaction rate
 coefficients for the formation of molecular hydrogen via the 3-body
-process in both the \enzo\ and Gadget-2 codes.  Similar trends were
-visible in both the \enzo\ and Gadget simulations, though at a
-nominally equivalent resolution (i.e., with particle and cell gas
-masses being comparable) \enzo\ simulations typically displayed a
+process in both the \enzo\ and Gadget-2 codes, observing similar trends between
+the two codes. However, at nominally equivalend resolution
+where the particle and cell gas masses are comparable, 
+\enzo\ simulations typically displayed a
 substantially higher level of gas structure.  This is unsurprising due
@@
 as a function of simulation resolution, and also agree quite well with
-each other -- however, the observed mass-loss rates do not agree
+each other. However, the observed mass-loss rates do not agree
 particularly well with observations.

### 2013-05-29 tests-wavepool.tex

@@
 
-In this one-dimensional test, we pass a short wavelength, linear wave through a static refined region.
+In this one-dimensional test we pass a short wavelength linear wave through a static refined region.
 The full domain is from 0 to 1 and the refined region is set from 0.25 to 0.75 at all times.
@@
 Figure~\ref{fig.wavepool} shows the evolution of the wave at three different times ($t = 0.2, 0.3$ and 0.8) for three of our solvers (PPM, Zeus, and MUSCL).
-The leftmost column show the wave just before it enters the refined region so that we can gauge how the solver is operating in the absence of AMR, the center column shows the wave after it has fully entered the refined region, and the rightmost column shows the wave well after it has exited the refined region.
+In all panels, the shaded regions denote the statically refined region.
+The leftmost column show the wave just before it enters the refined region so that we can gauge how the solver is operating in the absence of AMR.
+The center column shows the wave after it has fully entered the refined region.
+The rightmost column shows the wave well after it has exited the refined region.
 
-Beginning with the PPM, we see that even before it has entered the refine region, the wave is slightly damped, which is not unexpected for such a short wavelength mode, even for PPM.
+With the PPM solver, we notice that even before the wave reaches the refined region it has been slightly damped.
+This is not unexpected for such a short wavelength mode, even for high-order solvers such as PPM.
 The remaining panels demonstrate that the wave cleanly enters and exits the refined region.
-No significant reflection is seen, either on entry or exit, and the amount of damping is mild.
+No significant reflection is seen on either entry or exit, and the amount of damping is mild.
 
-For the Zeus solver, we see that even before it has entered the refined region, there are small oscillations excited behind the wave; although it is also worth noting that the wave itself is beautifully propagated without significant damping or phase errors.
-In this test, we use only our standard, low amount of quadratic artificial viscosity -- these waves can be damped by additional viscosity, but we do not add any in order to be sensitive to any artifacts at the grid boundary.
-The remaining panels show that the trailing oscillations continue, but do not generate any additional noise -- the end result is similar to the case without any refined region.
+For the Zeus solver, we see that even before it has entered the refined region there are small oscillations excited behind the wave; although it is also worth noting that the wave itself is beautifully propagated without significant damping or phase errors.
+In this test we use only our standard, low amount of quadratic artificial viscosity.
+These waves could be damped by additional viscosity, but we do not add any in order to be sensitive to any artifacts at the grid boundary.
+The remaining panels show that the trailing oscillations continue, but do not generate any additional noise.
+We note that the end result is similar to the case without any refined region.
 
-Finally, the MUSCL solver shows also shows a very clean entry and exit from the refined region without any oscillations, although (for the PLM reconstruction used here) the wave is spread more than with the other methods.
+Finally, the MUSCL solver shows also shows a very clean entry and exit from the refined region without any oscillations. However,
+because we use a piecewise linear reconstruction, the wave is spread more than with the other methods.
 
@@
 \includegraphics[width=0.8\textwidth]{figures/WavePool.eps}
-\caption{This plot shows, for each column, three snap shots (at $t=0.2, 0.3$ and 0.8) of a linear wave as it propagates through the grid.
-The static refined region extends from $x = 0.25$ to 0.75 and is shown in grey in each panel.
+\caption{This plot shows, for each column, three snapshots at $t=0.2, 0.3$ and 0.8 of a linear wave as it propagates through the domain.
+A static refined region extends from $x = 0.25$ to 0.75 and is shown in grey in each panel.
 The individual cells are also color-coded by level: blue indicates the root grid, and red is for the refined region.
-Each row shows the result for a different solver (top is PPM, middle is Zeus, and the bottom is MUSCL).
+Each row shows the result for a different hydrodynamic solver (top is PPM, middle is Zeus, and the bottom is MUSCL).
 The solid line shows the analytic solution for a linear, undamped wave.

### 2013-05-29 test_problems/AnisoConduction/make_plots.py
Added field lines to the figure of the anisotropic conduction test.

@@ -1,6 +1,6 @@
 from yt.mods import *
 
-for n in range(11,12):
+for n in [0,11]:
     
     pf = load("DD%04i/DD%04i" % (n,n))
 
@@ -20,8 +20,23 @@ for n in range(11,12):
     slice = SlicePlot(pf, 'z', "Temperature", width = (1.0, '1'), origin='domain', axes_unit=('1','1'))  #.save()
     #slice = SlicePlot(pf, 'z', "Temperature", width = (1.0, '1'), origin='domain', axes_unit=None)  #.save()
 
-    filename = "DD%04i" % n
 
-    slice.save(mpl_kwargs={'bbox_inches':'tight'}, name="%s_slice_Temperature_z.eps" % filename)
+    x = (dd['x'].reshape((256,256)))[:,0]
+    y = (dd['y'].reshape((256,256)))[0,:]
+    Bx = dd['Bx'].reshape((256,256))
+    By = dd['By'].reshape((256,256))
+    ax = slice.plots['Temperature'].axes
+
+    ## I found that I had to reverse Bx and By to get the correct
+    ## closed circles. Something to do with the plot's origin being in
+    ## the lower left?
+    streams = ax.streamplot(x,y,By,Bx,color=(1,1,1,0.5),linewidth=1.5,density=1)
+
+    ax.set_rasterization_zorder(5.0)
+    
+    ## filename = "DD%04i_slice_Temperature_z.png" % n
+    filename = "DD%04i_slice_Temperature_z.eps" % n
+
+    slice.save(mpl_kwargs={'bbox_inches':'tight'}, name=filename)
 
 # the end

### 2013-05-29 overview.tex
Adding missing references that I couldn't look up in the plane!

@@
 order-accurate in space and time and which explicitly conserves mass,
-linear momentum, and energy.  \red{(REF?)}  A variety of Riemann
-solvers have been implemented.
+linear momentum, and energy \citep{Hawley84, Norman86}.  A variety of
+Riemann solvers have been implemented.
 
@@
 well as a version of the PPM hydro solver which have been ported to
-Nvidia's CUDA framework allowing \enzo\ to take advantage of modern
+nVidia's CUDA framework allowing \enzo\ to take advantage of modern
 graphics hardware \citep{Wang:2010}.

### 2013-05-29 development.tex
Reynolds edits.

@@
 openness.  Started as the graduate research project of Greg Bryan at
-the University of Illinois, it has been subsequently been stewarded by
+the University of Illinois, it has subsequently been stewarded by
 the Laboratory for Computational Astrophysics (LCA) at the University
@@
 centralized version control system, and the ``stable'' \enzo\ source
-code was made globally readable following the \enzo\ 1.5. release.
+code was made globally readable following the \enzo\ 1.5 release.
 Access to the primary development tree required a password and login

### 2013-05-29 introduction.tex

@@
+% EJT: in code comments point to changes
+
 \section{Introduction}\label{sec.intro}
@@
 Due to the high spatial and temporal dynamical ranges involved, astrophysics and cosmological phenomena present a taxing challenge for simulators.
-To tackle these, a number of numerical techniques have been developed which can be broadly split into the gridlesss,fully Lagrangian methods and grid-based, Eulerian schemes.
-The most commonly used is an example of the first typeand is known as Smoothed Particle Hydrodynamics \citep[SPH;][]{Lucy77, SPH}.
-It has achived much success, particularly in regimes dominated by gravity.
+To tackle such situations, a number of numerical techniques have been developed which can be broadly split into the gridless, fully Lagrangian methods and grid-based, Eulerian schemes.
+The most commonly used is an example of the first type known as Smoothed Particle Hydrodynamics \citep[SPH;][]{Lucy77, SPH}.
+It has achieved much success, particularly in regimes dominated by gravity.
+% I felt metioning SPH before anything else was a little confusing for the article's focus so I added an earlier mention of Eulerian. 
 %where systems evolve over large spatial and temporal dynamic ranges (cut because we've already said that in the first sentense - EJT)
@@
 
-Despite this invested expertise, in their original form the Eulerian solvers have a serious drawback: they do not provide an easy method of adaptively increasing the spatial and temporal resolution in small volumes of the simulation.
-Such flexibility is essential for following physical processes such as gravitational instability.
-%Yet, despite the expertise invested in Eulerian solvers, in their original form they do not provide an easy method of adaptively increasing the spatial and temporal resolution in small volumes of the simulation.
+Despite this invested expertise, the Eulerian solvers in their original form have a serious drawback: they do not provide an easy method of adaptively increasing the spatial and temporal resolution in small volumes of the simulation.
 Such flexibility is essential for following physical processes such as gravitational instability.
 A solution to this was first proposed by \citet{Berger89} in the Computational Fluid Dynamics (CFD) community, and became known as Structured Adaptive Mesh Refinement (SAMR).
-The pricipal is to adaptively add and modify additional, finer meshes (``grids'') over regions that require higher resolution.
+The principal is to adaptively add and modify additional, finer meshes (``grids'') over regions that require higher resolution.
 In addition, it is possible to add other more advanced physics including --for the implementation in the astrophysics code presented in this paper-- comoving coordinates, self-gravity, radiative cooling, chemistry, heat conduction, collisionless fluids, magnetohydrodynamics, radiation transport, star formation and a range of other physical effects.
@@
 
-In this paper, we present \enzo\ ; a structured adaptive mesh refinement (SAMR) code.
+In this paper, we present \enzo; a structured adaptive mesh refinement (SAMR) code.
 Originally developed for cosmological hydrodynamics, \enzo\ has since been used on a wide variety of problems.
 It has grown to become a general tool for astrophysical fluid dynamics and is intended to be efficient, accurate and easily extended to include new capabilities.
-Although many of the components of the \enzo\ have been described in previous publications \citep{1995CoPhC..89..149B, BryanThesis96, Bryan97a, Bryan97b, Norman99, BryanCompSci99, Bryan01, Oshea04, 2007arXiv0705.1556N, WangAbelZhang08,ReynoldsHayesPaschosNorman2009,Collins10,Wise11_Moray}, there has previously been no systematic and complete description of the code.
-In this paper we provide that description, filling in many previous ommissions and showing the code's performance for a wide variety of test problems.
+Although many of the components of the \enzo\ code have been described in previous publications \citep{1995CoPhC..89..149B, BryanThesis96, Bryan97a, Bryan97b, Norman99, BryanCompSci99, Bryan01, Oshea04, 2007arXiv0705.1556N, WangAbelZhang08,ReynoldsHayesPaschosNorman2009,Collins10,Wise11_Moray}, there has previously been no systematic and complete description of the code.
+In this paper we provide that description, filling in many previous omissions and showing the code's performance for a wide variety of test problems.
 
@@
 The variety of astrophysical systems that \enzo\ has been used for include galaxies \citep{2003ApJ...587...13T, 2012ApJ...749..140H}, galaxy clusters \citep{Loken02, Skillman13}, the interstellar medium \citep{Slyz05}, the intergalactic medium \citep{Fang01, 2011ApJ...731....6S}, cooling flows \citep{Li12, 2013ApJ...763...38S}, turbulence \citep{Kritsuk04, 2009JPhCS.180a2020K, Collins11}, the formation of the first stars \citep{ABN02, 2007ApJ...654...66O, 2009Sci...325..601T}, and the formation of stars in our own Galaxy \citep{Collins12a}.
+
 % 'finally' here sounded strange, since this is really the key point

### 2013-05-29 numerical-gravity.tex

@@
 But if it's too long, please just cut it!
-
-% added: ``Each scheme can be selected via the parameter file.'' I felt it sounded like a dice roll otherwise, but again, feel free to overule me!
 
@@
 Boundary conditions are then interpolated from the potential values on the parent grid.
-We use either tri-linear interpolation or a natural second-order spline for this: both methods give similar results but we generally choose to use tri-linear interpolation, which empirically provides a resonable compromise between speed and accuracy.
-Each scheme can be selected via the parameter file.
+We use either tri-linear interpolation or a natural second-order spline for this: both methods give similar results but the default is the tri-linear interpolation, which empirically provides a resonable compromise between speed and accuracy.
 The potential equation on each subgrid is then solved with the given Dirichlet boundary conditions with a multigrid relaxation technique.

### 2013-05-29 numerical-radiation-raytracing.tex

@@
 
-Local radiation from stars and black holes strongly affects their
-surroundings.  Radiation transport is a well-studied problem; however,
-its treatment in multidimensional calculations is difficult because of
-the dependence on seven variables -- three spatial, two angular,
-frequency, and time.  The non-local nature of the thermal and
-hydrodynamical response to radiation sources further adds to the
-difficulty.  Here we briefly describe the ray tracing implementation \moray,
-which is presented in full detail in \citet{Wise11_Moray} with seven
-code tests and six applications.
+Stars and black holes strongly affect their surroundings through radiation.
+Radiation transport is a well-studied problem; however, its treatment in
+multidimensional calculations is difficult because of the dependence on seven
+variables -- three spatial, two angular, frequency, and time.  The non-local
+nature of the thermal and hydrodynamical response to radiation sources further
+adds to the difficulty.  Here we briefly describe \enzo's ray tracing
+implementation \moray, which is presented in full detail in
+\citet{Wise11_Moray} with seven code tests and six applications.
 
@@
 
-Ray tracing is an accurate method to propagate radiation from point
+yay tracing is an accurate method to propagate radiation from point
 sources on a computational grid, given that there are a sufficient
@@
 ray in the calculation are summed, and after the ray tracing is
-complete, these rates are used to update the ionization state and
-energy of the cells with the solver described in Section~\ref{sec.ov.chem}.
+complete, these rates are used as inputs to the solver described in
+Section~\ref{sec.ov.chem} to update the ionization, chemical, and energy states
+of the gas in each cell.
 

### 2013-05-29 tests-sodshocktube.tex

@@
 code method papers) and show solutions both with and without adaptive
-mesh refinement.  Without AMR (in the top row of the figure), it is
-clear that the PPM scheme produces by far the cleanest solution, with
+mesh refinement.  Without AMR (top row in Figure \ref{fig.sodshocktube}), it is
+clear that the PPM scheme produces by far the cleanest solution with
 all wave families crisply reproduced (in particular, the contact
 discontinuity and shock).  \zeus\ and MUSCL produce similar results,
-with MUSCL doing a slightly better job on the rarefaction fan.  In the
-figure, we also show the mean absolute deviation from the exact
-solution ($||E_\infty||$), as well as the maximum deviation
-($||E_1||$).  These numbers confirm the qualitative differences noted
+with MUSCL doing a slightly better job on the rarefaction fan.  In Figure
+\ref{fig.sodshocktube}, we also show the integrated absolute deviation from 
+the exact solution, $||E_1|| = \sum_i \Delta x_i |F(x_i)-Exact(x_i)|$ 
+, as well as the maximum deviation,
+$||E_\infty|| = \mathrm{max_i}\left\{ \Delta x_i|F(x_i)-Exact(x_i)| \right\}$.  
+These numbers confirm the qualitative differences noted
 previously.

### 2013-05-29 overview.tex
Added a table of variables and constants used in the paper.

@@
 describe the numerical algorithms that we use to solve these
-equations.  This section is intended to be an overview of the \enzo\
-code's capabilities: thus, we gather all of the equations solved into
-a single place and provide a brief and high-level introduction to the
-numerics of the code.  Detailed descriptions of the individual
-components are then provided in
-Sections~\ref{sec.amr}-\ref{sec.num.analysis}.
+equations.  This section is intended to be an overview of the
+\enzo\ code's capabilities: thus, we gather all of the equations
+solved into a single place and provide a brief and high-level
+introduction to the numerics of the code.  Detailed descriptions of
+the individual components are then provided in
+Sections~\ref{sec.amr}-\ref{sec.num.analysis}. In
+Table~\ref{tab:variables} we provide a convenient summary of all
+variables and physical constants used throughout this manuscript.
+
+\mqk{Move this table to the Appendix?}
+\input{table_of_variables_and_constants}
 
@@
 In these equations, $E$, $\rho$, $\vecv$, and $\vecB$ are the comoving
-total fluid energy, comoving baryonic density, peculiar velocity, and
+total fluid energy density, comoving baryonic density, peculiar velocity, and
 comoving magnetic field strength, respectively.  $\myvec{I}$ is the
@@
 
-The total fluid energy $E$ is given by:
+The total fluid energy density $E$ is given by:
 \begin{equation}

### 2013-05-29 table_of_variables_and_constants.tex
Added a table of variables and constants used in the paper.

\begin{center}

%%% Use this to change the fontsize of the table
%\tabletypesize{\scriptsize}

%%% \LongTables command may not be necessary (allowed?) outside of
%%% emulateapj.cls...
\LongTables

\begin{deluxetable}{cl}
  \tablecaption{Summary of symbols used in this manuscript\label{tab:variables}}
  \tablehead{\colhead{Symbol} & \colhead{Description}}

  \startdata
  \multicolumn{2}{l}{\underline{Fluid Quantities}} \\[5pt]
  $E$ & comoving total fluid energy density \\
  $e$ & comoving thermal energy density \\
  $T$ & baryon temperature \\
  $S$ & fluid entropy \\
  $\rho$ & comoving baryonic density \\
  $\rho_{\!X}$ & mass density of component $X$ (gas, dark matter, stars, metals, total, etc.) \\
  $t$ & time \\
  $\vecv$ & peculiar velocity \\
  $\vecB$ & comoving magnetic field strength \\
  $p$ & thermal pressure \\
  $p^*$ & total (thermal + magnetic) isotropic pressure \\
  $\gamma$ & ideal gas ratio of specific heats \\
  $\phi$ & gravitational potential \\
  $\Lambda$ & radiative/chemical cooling rate \\
  $\Gamma$ & radiative heating rate \\
  $Q$ & artificial viscosity stress tensor (only for ZEUS hydro) \\[3pt]

  \multicolumn{2}{l}{\underline{Code Quantities}} \\[5pt]
  $\Delta x_l$ & cell size on level $l$ \\
  $\Delta t_l$ & timestep on level $l$ \\
  $d$ & dimensionality (rank) of the simulation \\
  $q$ & conserved quantity \\
  $F$ & flux of quantity $q$ \\
  $r$ & refinement factor \\
  $\epsilon_l$ & refinement criterion parameter \\
  $J$ & number of cells per Jeans length (for Jeans length based refinement) \\[3pt]

  \multicolumn{2}{l}{\underline{Cosmology}} \\[5pt]
  $a$ & cosmological expansion factor \\
  $\Lambda_c$ & cosmological constant \\
  $\rho_0$ & mean comoving mass density (including both baryonic and dark matter) of the universe \\
  $p_0$ & comoving background pressure \\[3pt]

  \multicolumn{2}{l}{\underline{Chemistry}} \\[5pt]
  $n_i$ & comoving number density of species $i$ \\
  $k_{ij}$ & two-body reaction rate coefficient \\[3pt]

  \multicolumn{2}{l}{\underline{Ray-tracing Radiation Transfer}} \\[5pt]
  $\Gamma_j^{ph}$ & photo-ionization/-dissociation destruction/creation rates \\
  $I_\nu$ & specific radiation intensity (energy per time per solid angle per unit area per frequency) \\
  $\kappa_\nu$ & radiation absorption coefficient \\
  $j_\nu$ & radiation emission coefficient \\
  $P$ & photon number flux along ray \\
  $\sigma_{\rm abs}$ & absorption cross section \\
  $n_{\rm abs}$ & number density of absorbing medium \\
  $dt_P$ & radiative transfer timestep \\
  $k_{ph}$ & photo-ionization rate of a single ray \\
  $\Gamma_{ph}$ & photo-heating rate of a single ray \\[3pt]

  \multicolumn{2}{l}{\underline{Flux Limited Diffusion Radiation Transfer}} \\[5pt]
  $E_r$ & grey radiation energy density \\
  $e_c$ & internal energy correction due to photo-heating and chemical cooling \\
  $D$ & Larsen square-root flux-limiter \\
  $\kappa$ & total opacity \\
  $\eta$ & field of radiation sources \\[3pt]

  \multicolumn{2}{l}{\underline{Heat Conduction}} \\[5pt]
  $\fcond$ & thermal conduction heat flux \\
  $\kappa_{sp}$ & Spitzer conduction coefficient \\
  $f_{Sp}$ & fraction of Spitzer conductivity \\[3pt]

  \multicolumn{2}{l}{\underline{Star Formation And Feedback}} \\[5pt]
  $t_{\rm cool}$ & cooling time \\
  $t_{\rm dyn}$ & local dynamical time \\
  $t_{\rm form}$ & star particle formation time \\
  $m_b$ & baryonic mass of cell \\
  $m_J$ & Jeans mass of cell \\
  $m_*$ & star particle mass \\
  $m_{\rm *min}$ & minimum star particle mass \\
  $f_{\rm *eff}$ & star formation efficiency parameter \\
  $f_{\rm Zb}$ & metallicity fraction of baryon gas in cell \\
  $f_{\rm m*}$ & fraction of star particle's mass returned to cell \\
  $f_{\rm Z*}$ & fraction of star particle's metal mass returned to cell \\
  $f_{\rm SN}$ & fraction of stellar rest mass energy deposited as feedback \\
  $\Delta m_{\rm sf}$ & mass of stars formed in current timestep \\
  $E_{\rm add}$ & feedback energy added to cell \\
  $T_{\rm dyn,min}$ & user-defined minimum dynamical time \\
  $\eta$ & overdensity threshold for star formation \\
  $\rho_{\rm SF}$ & constant proper density threshold for star formation \\
  $\tau_*$ & characteristic star formation timescale (for Schmidt-Law method) \\[3pt]

  \multicolumn{2}{l}{\underline{Natural Constants}} \\[5pt]
  $G$ & gravitational constant \\
  $H$ & Hubble constant \\
  $k_B$ & Boltzmann constant \\
  $M_\odot$ & mass of Sun \\
  $m_H$ & mass of hydrogen atom \\
  $m_e$ & mass of electron \\
  $c$ & speed of light \\
  $c_s$ & sound speed \\
  \enddata

  %% \tablecomments{Optional comments}
\end{deluxetable}

\end{center}

### 2013-05-29 tests-onezonecollapse.tex
Added a table of variables and constants used in the paper.

@@
 
-with G being the gravitational constant.
-\mqk{Instead of explaining what $G$ is here, would it make sense to have a table at the beginning of the paper listing all the natural constants and variable symbols that are used throughout the paper?} The collapse timescale is
-altered from the dynamical time by a factor $1/\sqrt{1-f}$ in Equation
+ The collapse timescale is altered from the dynamical time by a factor
+$1/\sqrt{1-f}$ in Equation
 \ref{eqn.tcol}, which is an approximation of the ratio of the gas

### 2013-05-30 appendix.tex
Added appendix

--- 2013-01-18 numerical-radiation-raytracing.tex
@@
-\subsection{Radiation transport: ray tracing}
-\label{sec.num.raytracing}
+\appendix
+\section{Interpolation methods}
+\label{app:interpolation}
+
+In this appendix, we provide details for the various interpolation methods available in the code.
+We assume throughout that we are dealing with Cartesian coordinates and equally spaced cells, although we allow for arbitrary refinement factor $r$.
+
+\vspace{0.5cm}\noindent
+{\bf SecondOrderA}
+
+This interpolation algorithm is generally second-order but has monotonicity constraints as described below.
+
+In one dimension, we define the parent values as $Q_{-1}$, $Q_0$, and $Q_{+1}$, where the central parent cell has a left edge at $x_0$ and width $\Delta x$.
+We first linearly interpolate the parent values to the cell edges: $Q_{-1/2} = (Q_0 + Q_{-1})/2$, and similarly for $Q_{1/2} = (Q_0 + Q_1)/2$, and then compute a monotonic slope: $\Delta Q_0 = \minmod{Q_{1/2} - Q_0}{Q_0 - Q_{-1/2}}$ where
+\begin{equation}
+\minmod{a}{b} = \left\{ \begin{array}{ll}
+0 & {\rm if} \quad ab < 0 \\
+\min{ \left( | a |, | b | \right) } {\rm sign} (a) & {\rm otherwise}
+\end{array}\right.
+\end{equation}
+This slope is then used to compute the interpolated subgrid values.
+Defining $q_i$ to be the subgrid values at cell centers for the $r$ cells corresponding to parent cell $Q_0$ (for refinement factor $r$), we can write:
+\begin{equation}
+q_i = Q_0 + \frac{i+(1-r)/2}{r} f_x
+\end{equation}
+where $f_x = 2 \Delta Q_0$ (this notation is used for consistency with the 2 and 3 dimensional cases).
+
+In two dimensions, the procedure is very similar in that we linearly interpolate parent values to the four cell corners: $Q_{-1/2, -1/2}$, $Q_{-1/2, 1/2}$, $Q_{1/2, -1/2}$, $Q_{1/2, 1/2}$ (by averaging parent cell-centered parent values).
+We then compute monotonic slopes across the two diagonals, since for a linear function, extrema occur at corners:
+\begin{eqnarray}
+\Delta Q_0 & = & \minmod{Q_{1/2, 1/2} - Q_0}{Q_0 - Q_{-1/2, -1/2}} \\
+\Delta Q_1 & = & \minmod{Q_{-1/2, 1/2} - Q_0}{Q_0 - Q_{1/2, -1/2}}
+\end{eqnarray}
+We then translate these slopes into the grid axes with $f_x = \Delta Q_0 -  \Delta Q_1$ and $f_y = \Delta Q_0 + \Delta Q_1$ so that the interpolation itself can be written simply as
+\begin{equation}
+q_{i,j} = Q_0 + \frac{i+(1-r)/2}{r} f_x + \frac{j+(1-r)/2}{r} f_y
+\end{equation}
+
+Finally, we write down the three-dimensional version -- unfortunately, here the number of monotonicity constraints is four (the 4 diagonals across the 8 opposing corners of the cube), while the number of slopes is three, so the problem is over constrained.
+Somewhat arbitrarily, we adopt the following procedure.
+As before, we compute $\Delta Q_0$, $\Delta Q_1$, $\Delta Q_2$, and $\Delta Q_3$ with the minmod limiter across the 4 diagonals.
+We then define
+\begin{equation}
+s =\frac{\Delta Q_1 + \Delta Q_2 + \Delta Q_3}{\Delta Q_0}
+\end{equation}
+which is the value of $\Delta Q_0$ that the other slopes ($\Delta Q_1$, $\Delta Q_2$, and $\Delta Q_3$) imply, normalized by the desired value of $\Delta Q_0$ itself.
+If $0 < s < 1$, then no adjustment needs to be made, as the monotonicity constraints are met.
+If these equalities are not met, then, if $s<0$ we define $\chi_n = 1$ if $\Delta Q_n/\Delta Q_0 < 0$, and 0 otherwise (if $s > 1$, this is reversed, so $\chi_n = 1$ if $\Delta Q_n/\Delta Q_0 > 0$, and 0 otherwise).
+These weights are used to determine which slopes to modify to match the $\Delta Q_0$ constraint.
+We then compute the amount of adjustment required:
+\begin{equation}
+f = - \frac{\Delta Q_0 s^\prime + \sum_n (1-\chi_n) \Delta Q_n}{\sum \chi_n \Delta Q_n + \epsilon}
+\end{equation}
+where $s^\prime = \min(\max(s,0), 1)$ and $\epsilon$ is a small number to prevent numerical errors.
+We then use this adjustment fraction to compute the new adjusted $\Delta Q_n$ (for $n = 1, 2, 3$) that match the $\Delta Q_0$ constraint as closely as possible with
+\begin{equation}
+\Delta Q_n = f \chi_n \Delta Q_n + (1-\chi_n) \Delta Q_n
+\end{equation}
+Finally, these are converted to the grid axes with $f_x = \Delta Q_2 + \Delta Q_3$, $f_y = \Delta Q_1 + \Delta Q_3$, $f_z = \Delta Q_1 + \Delta Q_2$ and then we do the interpolation itself with,
+\begin{equation}
+q_{i,j,k} = Q_0 + \frac{i+(1-r)/2}{r} f_x + \frac{j+(1-r)/2}{r} f_y + \frac{k+(1-r)/2}{r} f_z
+\end{equation}
+
+% --------------
+
+\vspace{0.3cm}\noindent
+{\bf SecondOrderB}
+
+We also provide a variant on the above procedure, with two changes.
+The first is that the slopes across the diagonals ($\Delta Q_0$, etc.) are computed directly rather than with the minmod limiter (e.g.
+$\Delta Q_0 = (Q_{1/2,1/2,1/2} - Q_{-1/2,-1/2,-1/2})/2$).
+To ensure positivity in the resulting interpolation when applied to positive conserved quantities, the slopes are limited so that the smallest corner value is 0.2 the cell center value.
+The same procedure as described above is then applied to turn these four slopes into a linear interpolation.
+
+% --------------
+
+\vspace{0.3cm}\noindent
+{\bf SecondOrderC}
+
+This is a completely different second-order interpolation scheme, which is based on Cloud-In-Cell (CIC) interpolation \citep{Hockney88}.
+In one dimension, we define the parent values as $Q_0$, and $Q_{+1}$, where the left parent cell has a cell center at $x_0$ and width $\Delta x$.
+Then, the interpolated value for a subgrid cell $q_i$ with a cell left edges at $x_i = x_0 + i \Delta x^p/r$ where $i$ runs from 0 to $r-1$ for refinement factor $r$, is given by:
+\begin{equation}
+q_i =  \frac{2r - 1 - 2i}{2r} Q_0 + \frac{1+2i}{2r} Q_{+1}
+\end{equation}
+The extension to two and three dimensions is straightforward, with weights in the other dimensions computed in the same way and then multiplied to get a total of 4 and 8 weights for the 2 and 3 dimensional cases, respectively.
+This scheme is monotonicity preserving, but is not conservative.
+
+% --------------
+
+\vspace{0.3cm}\noindent
+{\bf ThirdOrderA}
+
+This interpolation method provides a third-order accuracy based on the Triangular Shaped Cloud (TSC) methodology \citep{Hockney88}.
+As usual, in one dimension, we define the parent values as $Q_{-1}$, $Q_0$, and $Q_{+1}$, where the central parent cell has a left edge at $x_0$ and width $\Delta x$.
+Then, the interpolated value for a subgrid cell $q_i$ with a cell left edges at $x_i = x_0 + i \Delta x^p/r$ where $i$ runs from 0 to $r-1$ for refinement factor $r$, is given by:
+\begin{equation}
+q_i = a_i  Q_{-1} + b_i Q_0 + c_i Q_{+1}
+\end{equation}
+and the weights are given by:
+\begin{equation}
+a(i) =  \frac{(r-i)^3 - (r-i-1)^3}{6r^3}; \qquad c(i) = \frac{3i^2 + 3i + 1}{6r^3}
+\end{equation}
+and $b(i) = 1/r - a(i) - c(i)$.
+The extension to two and three dimensions is straightforward with the weights computed in the same way as for the one-dimensional case but then multiplied to determine the 9 and 27 factors necessary for the 2 and 3 dimensional cases, respectively.
+
+One problem with this interpolation technique is that it is not conservative.  In particular, the sum of the interpolated subgrid values:
+\begin{equation}
+\tilde{Q_0} = \sum_{i=0}^{r-1} q_i
+\end{equation}
+is not, in general, equal to $Q_0$.
+We can retain conservation by adding the factor $(Q_0 - \tilde{Q_0})/r$ to the interpolated values (or by multiplying the interpolated values by the ratio $Q_0/\tilde{Q_0}$).
+Unfortunately, the result of this procedure does not preserve monotonicity -- it can introduce local minima and maxima at at parent cell boundaries.
+We can then attempt to correct that by taking a weighted average between the interpolated values on the subgrid and the parent value with weights computed such that the interpolated values at the edge of the parent cell are not local maxima compared to the interpolated values in the neighboring parent cell.
+
+% --------------
+
+\vspace{0.3cm}\noindent
+{\bf FirstOrderA}
+
+Finally, for completeness, we include a first-order accurate piecewise constant interpolator, for which, using the same definitions in the previous case, we take $q_i = Q_0$, for $i = 0$ to $r-1$.
+
+\vspace{1cm}
+
+%%% Local Variables: 
+%%% mode: latex
+%%% TeX-master: "ms"
+%%% End:

### 2013-05-30 tests-wavepool.tex
Fixed typo in wavepool.  In answer to Sam, there are no red points outside the refined region (which is purely static, no AMR).  The figure quality is not ideal, so perhaps that is where the confusion comes from -- my eps backend on matplotlib doesnt render the grey properly, so I generated png and converted.

@@
 In all panels, the shaded regions denote the statically refined region.
-The leftmost column show the wave just before it enters the refined region so that we can gauge how the solver is operating in the absence of AMR.
+The leftmost column shows the wave just before it enters the refined region so that we can gauge how the solver is operating in the absence of AMR.
 The center column shows the wave after it has fully entered the refined region.
@@
 With the PPM solver, we notice that even before the wave reaches the refined region it has been slightly damped.
-This is not unexpected for such a short wavelength mode, even for high-order solvers such as PPM.
+This is not unexpected for such a short wavelength mode, even for higher-order solvers such as PPM.
 The remaining panels demonstrate that the wave cleanly enters and exits the refined region.

### 2013-05-30 numerical-timestep.tex
Edited timestepping section

--- 2013-04-28 numerical-timestep.tex
@@
-\subsection{Time Stepping}
+\section{Time Stepping}
 \label{sec.timestepping}
 
-In \enzo, resolution of the equations being solved is adaptive in time 
-as well as in space.  The timestep in \enzo\ is satisfied on a level-by-level
- basis by finding the largest timestep such that multiple criteria are
-satisfied on each level.  The timestep criteria used by \enzo\ are 
-(showing the one-dimensional case for clarity):
+In \enzo, the integration of the equations being solved is generally adaptive in time 
+as well as in space.  The timestep $\Delta t$ is set on a level-by-level
+basis by finding the largest timestep such that all of the criteria listed below are
+satisfied.  The timestep criteria are given by the following expressions
+(where we show the one-dimensional case for clarity):
 
@@
 
-In Equations~\ref{eqn:dthydro}-\ref{eqn:dtcond}, the $\min ( \ldots
-)_L$ formalism means that this value is calculated for all cells or
-particles on a
-given level L and the minimum overall value is taken as the timestep.
-Equation~\ref{eqn:dthydro} ensures that all cells satisfy the
+In equations~(\ref{eqn:dthydro})-(\ref{eqn:dtcond}), the $\min ( \ldots)_L$ 
+formalism means that this value is calculated for all cells or
+particles on a given level l and the minimum overall value is taken as the timestep.
+
+Equation~(\ref{eqn:dthydro}) ensures that all cells satisfy the
 Courant-Freidrichs-Levy (CFL) condition for accuracy and stability of
 an explicit finite difference discretization of the Euler equations.
-Effectively this condition forces the timestep to be small enough such
-that any changes in the fluid propagate less than a single grid
-spacing, $\Delta x$.  In this equation, $\kappa_{\rm hydro}$ is a
+%Effectively this condition forces the timestep to be small enough such
+%that any changes in the fluid propagate less than a single grid
+%spacing, $\Delta x$.  
+In this equation, $\kappa_{\rm hydro}$ is a
 numerical constant with a value of $0 < \kappa_{\rm hydro} \leq 1$
@@
 
-Equation~\ref{eqn:dthydro} is valid for one dimension.  For 2 or 3
+Equation~(\ref{eqn:dthydro}) is valid for one dimension.  For 2 or 3
 dimensions, it was shown by \cite{Godunov1959}  that using the
@@
 t_x$, $\Delta t_y$, and $\Delta t_z$ be the analogues of
-Equation \ref{eqn:dthydro} along the $x,y$ and $z$ axes, 
+equation~(\ref{eqn:dthydro}) along the $x,y$ and $z$ axes, 
 \begin{equation}
@@
 
-For all other criteria except for Equation~\ref{eqn:dtMHD}, multiple dimensions are accounted for by
-repeating the one dimensional criterion along each axis, and taking
-the minimum.
+For all other criteria except for equation~(\ref{eqn:dtMHD}), multiple dimensions are accounted for by
+repeating the one dimensional criterion along each axis, and taking the minimum.
 
-Equation~\ref{eqn:dtMHD} is only enforced when the equations of
+Equation~(\ref{eqn:dtMHD}) is only enforced when the equations of
 magnetohydrodynamics are being solved, and is directly analogous to
-Equation~\ref{eqn:dthydro} in that it ensures that the CFL condition
+equation~(\ref{eqn:dthydro}) in that it ensures that the CFL condition
 is being enforced at all times.  In this equation, $\kappa_{\rm MHD}$ is a
@@
 speed comes from a stability analysis of the MHD equations, and is
-expressed as:
+given by:
 %
@@
 
-Equation~\ref{eqn:dtdarkmatter} is analogous to
-Equation~\ref{eqn:dthydro} and ensures accuracy in the N-body solver
-by requiring that no dark matter particle travels more than one cell
-width.  The parameter $\kappa_{\rm dm}$ is analogous to $\kappa_{\rm
-  hydro}$, with an identical range of values.
+Equation~(\ref{eqn:dtdarkmatter}) is analogous to
+equation~(\ref{eqn:dthydro}) and helps to ensure accuracy in the N-body solver
+by requiring that no particle travels more than one cell
+width.  The parameter $\kappa_{\rm dm}$ is like $\kappa_{\rm
+hydro}$, with a similar range of values.
 
-Equations~\ref{eqn:dtaccel} and~\ref{eqn:dtrad} are supplementary to Equation~\ref{eqn:dthydro} in that they
+Equations~(\ref{eqn:dtaccel}) and~(\ref{eqn:dtrad}) are supplementary to equation~(\ref{eqn:dthydro}) in that they
 take into account the possibility of large accelerations due to either
-gravity (Equation~\ref{eqn:dtaccel} or radiation pressure
-(Equation~\ref{eqn:dtrad}), which may cause numerical 
-instabilities by violating the Courant condition.  In Equation~\ref{eqn:dtaccel}, $\vec{g}$ is the
-gravitational acceleration in each cell on level L.  In
-Equation~\ref{eqn:dtrad}, $\vec{a}_{\rm rad}$ is the estimated
-acceleration due to radiation pressure in each cell on level L,
+gravity (equation~\ref{eqn:dtaccel}) or radiation pressure
+(equation~\ref{eqn:dtrad}).  In equation~(\ref{eqn:dtaccel}), $\vec{g}$ is the
+gravitational acceleration in each cell on level l.  In
+equation~(\ref{eqn:dtrad}), $\vec{a}_{\rm rad}$ is the estimated
+acceleration due to radiation pressure in each cell on level l,
 defined as
-
 \begin{equation}
@@
 \end{equation}
-
 where the sum calculates the energy deposited in a cell during the
@@
 
-Equation~\ref{eqn:dtcond} is the CFL condition for an explicit
+Equation~(\ref{eqn:dtcond}) is the stability condition for an explicit
 solution to the equation of heat conduction.  In this expression,
-$n_b$ is the baryon number density, defined as $n_b = \rho_b / \mu
-m_p$, where $\mu$ is the mean molecular weight, $\kappa(T)$ is the
-Spitzer thermal conductivity, and f$_{Sp}$ is a user-defined
-conduction suppression factor whose value must be f$_{Sp} \leq 1$.  Both are defined in
-Section~\ref{sec.num.conductions}.  $\kappa_{\rm cond}$ is a prefactor whose
+$n_b$ is the baryon number density. 
+%defined as $n_b = \rho_b / \mu
+%m_p$, where $\mu$ is the mean molecular weight, 
+$\kappa(T)$ is the Spitzer thermal conductivity, and f$_{Sp}$ is a user-defined
+conduction suppression factor whose value must be f$_{Sp} \leq 1$.  
+%Both are defined in Section~\ref{sec.num.conductions}.  
+$\kappa_{\rm cond}$ is a prefactor whose
 value must be $0 < \kappa_{\rm cond}  < 0.5$, and is exactly 0.5 for the
@@
 
-Finally, Equation~\ref{eqn:dtexpand} is a cosmological constraint that
+Finally, equation~(\ref{eqn:dtexpand}) is a cosmological constraint that
 limits the timestep so that the simulated universe only expands by
@@
 for the stability of the PPM algorithm in comoving coordinates.  This
-criterion typically limits the timestep (moreso than other criteria)
+criterion typically limits the timestep
 only at high redshifts, when densities are relatively homogeneous.

### 2013-05-30 numerical-conduction.tex
Reynolds edits.
Edited conduction section

@@
 \enzo\ implements the equations of isotropic heat conduction in a manner
-similar to that of \citet{2007ApJ...664..135P}, where we model the
-flux of isotropic heat transport:
-
+similar to that of \citet{2007ApJ...664..135P}.  
+The isotropic flux of heat is given by:
 \begin{equation}
-\mathbb{F} = -\kappa_{sp} \nabla T
+\myvec{F} = -\kappa_{sp} \nabla T
 \end{equation}
-
-Where $\kappa_{sp}$ is the Spitzer conduction coefficient, $4.6 \times 10^{-7}$~T$^{5/2}$ erg
-s$^{-1}$~cm$^{-1}$~K$^{-1}$ \citep{1962pfig.book.....S}, and using a
+where $\kappa_{sp}$ is the Spitzer conduction coefficient, $4.6 \times 10^{-7}$~T$^{5/2}$ erg
+s$^{-1}$~cm$^{-1}$~K$^{-1}$ \citep{1962pfig.book.....S}, here we are using a
 value for the Coulomb logarithm, $\log \Lambda = 37.8$, that is
 appropriate for the intracluster medium \citep{1988xrec.book.....S}.
-It is quite possible that the local heat flux can become unphysical in
+It is quite possible that the local heat flux computed in this way 
+can become unphysically large in
 the high-temperature, low-density cluster regime when using this
-formulation; thus, we take into account the saturation of the heat
-flux at a maximum level of
-
+formulation; therefore, we take into account the saturation of the heat
+flux \citep{1977ApJ...211..135C} at a maximum level of
 \begin{equation}
-j_{sat} \simeq 0.4 n_e k_b T \left( \frac{2 k_b T}{\pi m_e} \right)^{1/2}
+F_{sat} \simeq 0.4 n_e k_b T \left( \frac{2 k_b T}{\pi m_e} \right)^{1/2}.
 \end{equation}
-
-\citep{1977ApJ...211..135C}, and to ensure a smooth transition between
-the Spitzer regime and the saturated regime we define an effective
+To ensure a smooth transition between
+the Spitzer and saturated regimes, we define an effective
 conductivity using the formalism of \citet{1988xrec.book.....S}
-
 \begin{equation}
@@
 \end{equation}
-
 where $\lambda_e$ is the electron mean free path and $\ell_T \equiv T
@@
 Thermal conduction in a plasma can be strongly affected by the
-presence of magnetic field lines, which may strongly suppress heat
+presence of magnetic field lines, which may suppress heat
 flow perpendicular to the magnetic field.  In that case, we allow for
@@
 magnetohydrodynamic simulations.  Mathematically, this is treated as:
-
 \begin{equation}
-\mathbb{F} = -\kappa_{sp} \hat{b} (\hat{b} \cdot \nabla T)
+\myvec{F} = -\kappa_{sp} \hat{b} (\hat{b} \cdot \nabla T)
 \end{equation}
-
-Where $\hat{b}$ is the unit vector denoting the direction of the
+where $\hat{b}$ is the unit vector denoting the direction of the
 magnetic field.  As with the isotropic thermal conduction, we allow a

### 2013-05-30 numerical-starformation.tex
Edited star formation - tightened up algorithm description and SF integral expression.

@@
 \enzo.
-The \enzo\ code includes approximately ten models for star formation
-and feedback.  Broadly speaking, these methods all work in similar
+%The \enzo\ code includes approximately ten models for star formation
+%and feedback.  
+Broadly speaking, these methods all work in similar
 ways: At a specified time interval, all grid cells that are at the
@@
 the same mass is placed in the center of that cell, with the same
-momentum vector as the removed gas.  This star particle is then
+velocity as the removed gas.  This star particle is then
 allowed to inject mass, momentum, thermal energy, metals, and possibly
@@
 are allowed to form in a cell if the following criteria are met:
-
-\begin{equation}
-\rho_b/\bar{\rho}_b \geq \eta
+\begin{eqnarray}
+\rho_b/\bar{\rho}_b & \geq & \eta,  \\[2mm]
 \label{cendens}
-\end{equation}
-
-\begin{equation}
-\nabla \cdot \vec{v}_b < 0,
+%\end{equation} 
+%\begin{equation}
+\nabla \cdot \vec{v}_b & < & 0, \\
 \label{cencont}
-\end{equation}
-
-\begin{equation}
-t_{\rm cool} < t_{\rm dyn} \equiv \sqrt{3 \pi / 32G \rho_{\rm tot}},
-\end{equation}
-
-\begin{equation}
-m_{b} > m_{J} \equiv G^{-3/2} \rho_{b}^{-1/2}C^{3}
+%\end{equation}
+%\begin{equation}
+t_{\rm cool} & < & t_{\rm dyn} \equiv \sqrt{3 \pi / 32G \rho_{\rm tot}}, \\
+%\end{equation}
+%\begin{equation}
+m_{b} & > & m_{J} \equiv G^{-3/2} \rho_{b}^{-1/2}c^{3} 
 \left[ 1 + \frac{\delta\rho_{d}}{\delta\rho_{b}} \right]^{-3/2}
-\end{equation}
-
-Where $\rho_b$ is the baryon density in that cell, $\bar{\rho_b}$ is
-the mean baryon density in the simulation, $\eta$ is the user-defined
-overdensity threshold, $\vec{v}$ is the velocity of the gas in the
-cell, $\rho_{d}$ is the dark matter density, $\rho_{\rm tot} =
-\rho_{b} + \rho_{d}$, $m_{b}$ and $m_{j}$ are the baryonic mass in the
-cell and jeans mass of the cell, and C is the isothermal sound speed
+\end{eqnarray}
+where 
+%$\rho_b$ is the baryon density in that cell, $\bar{\rho_b}$ is
+%the mean baryon density in the simulation, 
+$\eta$ is the user-defined
+overdensity threshold, 
+%$\vec{v}$ is the velocity of the gas in the
+%cell, $\rho_{d}$ is the dark matter density, $\rho_{\rm tot} =
+%\rho_{b} + \rho_{d}$, 
+and $m_{b}$ and $m_{j}$ are the baryonic mass in the
+cell and the Jeans mass of the cell, and c is the isothermal sound speed
 in the cell.  If all of these criteria are met, the mass of a star
@@
 purposes of feedback that the mass of stars formed at a time $t$ with
-finite timestep $\Delta t$ is:
+timestep $\Delta t$ is:
+\begin{equation}
+\Delta m_{\rm sf} = \int_t^{t+\Delta t} \frac{dM}{dt} dt =
+ \int_{\tau_0}^{\tau_1} m_* \tau e^{-\tau} d\tau =
+ m_* \left[(1+\tau_0) e^{-\tau_0} - (1 + \tau_1) e^{-\tau_1} \right]
+ \end{equation}
+where $\tau_0 = (t-t_{\rm form})/t_{\rm dyn}$ and $\tau_1 = (t+\Delta t-t_{\rm form})/t_{\rm dyn}$.
 
-\begin{eqnarray}  
-\Delta m_{\rm sf} =  m_{*} \left[ \left(1 + \frac{t-t_{\rm
-        form}}{t_{\rm dyn}}  \right)
-%\nonumber \times \\ % uncomment for 2-column
-		   \exp{\left( \frac{-(t-t_{\rm form})}{t_{\rm dyn}} \right) }
-%                    \nonumber \\ 
-	       - \left( 1 + \frac{t+ \Delta t-t_{\rm form}}{t_{\rm dyn}} \right) 
-%	       \nonumber \\ 
-                    \exp{ \left( \frac{-(t+\Delta t-t_{\rm
-                            form})}{t_{\rm dyn}} \right)  }
-                   \right]
-\end{eqnarray}
+%\begin{eqnarray}  
+%\Delta m_{\rm sf} =  m_{*} \left[ \left(1 + \frac{t-t_{\rm
+%        form}}{t_{\rm dyn}}  \right)
+%%\nonumber \times \\ % uncomment for 2-column
+%		   \exp{\left( \frac{-(t-t_{\rm form})}{t_{\rm dyn}} \right) }
+%%                    \nonumber \\ 
+%	       - \left( 1 + \frac{t+ \Delta t-t_{\rm form}}{t_{\rm dyn}} \right) 
+%%	       \nonumber \\ 
+%                    \exp{ \left( \frac{-(t+\Delta t-t_{\rm
+%                            form})}{t_{\rm dyn}} \right)  }
+%                   \right]
+%\end{eqnarray}
+%which can be represented more clearly in integral form:
+%
+%\begin{eqnarray}
+%\int_{t}^{t+Dt} \frac{dM}{dt} dt = \int_{t}^{t+Dt} m_{*} 
+%\frac{dt}{t_{\rm dyn}} \left( \frac{t-t_{\rm form}}{t_{\rm dyn}} \right) 
+%%\nonumber \times \\ 
+%\exp{ \left( - ~ \frac{t-t_{\rm form}}{t_{\rm dyn}} \right) }
+%\end{eqnarray}
 
-which can be represented more clearly in integral form:
-
-\begin{eqnarray}
-\int_{t}^{t+Dt} \frac{dM}{dt} dt = \int_{t}^{t+Dt} m_{*} 
-\frac{dt}{t_{\rm dyn}} \left( \frac{t-t_{\rm form}}{t_{\rm dyn}} \right) 
-%\nonumber \times \\ 
-\exp{ \left( - ~ \frac{t-t_{\rm form}}{t_{\rm dyn}} \right) }
-\end{eqnarray}
-
-During this timestep, we assume that the star particle feeds back
+During this timestep, the star particle returns
 metal-enriched gas and thermal energy from supernovae and from stellar
@@
 there is an immediate feedback of some fraction $f_{\rm SN}$ of the rest
-energy from the stars created that timestep into the baryon gas, such
+energy from the stars into the baryon gas, such
 that $E_{\rm add} = f_{\rm SN} \Delta m_{\rm sf} c^2$, where c is the speed of
-light.  In addition, a fraction $f_{Z*}$ of the metal from the star
-particle is fed back into the baryon gas, which takes into account the
-effects of metal recycling.  Finally, a fraction of the mass $f_{m*}$
-from all stars (rather than just supernovae) is fed back into the gas
+light.  In addition, a fraction $f_{Z*}$ of the stellar mass is fed back 
+in the form of metals.  Finally, a fraction of the mass $f_{m*}$
+is added back into the gas
 along with momentum in order to simulate the mass ejection from
-non-exploding stars via stellar winds.
+all stars (not just supernovae).
 
 There are six user-defined parameters in this algorithm: three deal
-with the star formation ($\eta$, $m_{\rm *min}$ and $T_{\rm
+with the star formation ($\eta$, $m_{\rm *min}$ and $t_{\rm
   dyn,min}$), and three deal with feedback ($f_{\rm SN}$, $f_{Z*}$ and
@@
 that $f_{Z*} = 0.02$ is an appropriate number.  The minimum dynamical
-time is set to be $T_{\rm dyn,min} = 10^7$ ~years to agree with timescales
+time is set to be $t_{\rm dyn,min} = 10^7$~years to agree with the SN timescales
 seen in nearby OB associations.
@@
 clouds, and thus this parameter becomes a rough proxy for finding the
-densest environments in a given simulation.  Given the
-simulation-dependent nature of these parameters, it is critical to set
-them carefully, and with some sensitivity testing.
+densest environments in a given simulation.  
+%Given the
+%simulation-dependent nature of these parameters, it is critical to set
+%them carefully, and with some sensitivity testing.
 
@@
 
-The Kravtsov method of star particle creation is designed to reproduce
+This method of star particle creation is designed to reproduce
 the global Schmidt law of star formation~\citep{2003ApJ...590L...1K,
@@
 with a characteristic gas timescale $\tau_*$ such that $\dot{\rho}_* =
-\rho_{\rm gas}/\tau_*$ where $\rho_{\rm gas}$ and $\rho_*$ are the
-baryon gas and stellar densities, respectively.  This ``constant
+\rho_{\rm gas}/\tau_*$. 
+%where $\rho_{\rm gas}$ and $\rho_*$ are the
+%baryon gas and stellar densities, respectively.  
+This ``constant
 efficiency'' model on the scale of star formation regions is well
@@
 
-Algorithmically, the star formation events in the Scmidt-law algorithm are
+Algorithmically, the star formation events in the Schmidt-law algorithm are
 assumed to occur once every global time step (with the constraint $\Delta t_0 \leq
@@
 every root grid timestep, $\Delta t_0$, and cells at the highest level
-that exceed the H$_2$ fraction limit form star particles with masses
+that exceed the H$_2$ limit form star particles with masses
 proportional to the inferred mass of molecular hydrogen in the
@@
 The mass of the particle is calculated as:
-
 \begin{equation} 
@@
 \end{equation}
-
 where $\Delta x_{m}$ is the resolution of the maximum level of

### 2013-05-31 numerical-analysis.tex
Minor revision to analysis section.

@@
 
-Most analysis operations that can be performed on simulations that reside on
-disk can be performed on in-memory simulations in \texttt{yt}.  This includes
+Most analysis operations that can be performed on data sets that reside on
+disk can be performed on in-memory data sets in \texttt{yt}.  This includes
 projections (i.e., line integrals, both on- and off-axis), slices, 1-, 2-, and
@@
 in-memory \enzo\ data.  Operations that currently cannot be conducted on
-in-memory datasets are those that require spatial decomposition of data; for
+in-memory datasets are those that require spatial decomposition of
+data.  For 
 instance, calculating marching cubes on a data object with \texttt{yt} is a
@@
 data and thus cannot be conducted \textit{in situ}.  This prohibition extends
-to most halo finding operations, most multi-level parallelism operations,
-volume rendering and connected set identification.
+to halo finding operations other than Rockstar, most multi-level
+parallelism operations, and volume rendering.
 
@@
 
-At user-specified intervals, the tracer particles output their data.  The data written out
+\enzo saves tracer particle data at user-specified intervals,
+independent of the intervals at which regular data sets are written.  The data written out
 typically includes the tracer particle's unique ID, position, velocity, density,
@@
 \begin{eqnarray}
-\nabla \cdot \vec{v} < 0 \\
-\nabla T \cdot \nabla S > 0\\
-T_2 > T_1 \\
+\nabla \cdot \vec{v} < 0,\\
+\nabla T \cdot \nabla S > 0,\\
+T_2 > T_1,\\
 \rho_2 > \rho_1,
@@
 may be chosen, which is useful for situations such as cosmological simulations without a radiation
-background.
+background where underdense gas in the intergalactic medium (IGM) can cool adiabatically to
+very low temperatures.
 

### 2013-05-31 numerical-overview.tex
Minor revision to analysis section.

@@
 method assumes a
-multiphase IGM and has been shown to accurately reproduce the cosmic star
+multiphase interstellar medium (ISM) and has been shown to accurately reproduce the cosmic star
 formation rate, and will be implemented into \enzo\ in the near future.

### 2013-05-31 tests-sodshocktube.tex
sweep 1: typo changed

@@
 
-We begin the paper test suite with the classic Sod shock tube problem
+We begin the paper test suite with the classic one-dimensional Sod shock tube problem
 \citep{Sod78}, which provides a good test of a hydrodynamical solver's
 ability to resolve a clean Riemann problem with clear separation
-between the three waves.  These waves include a rarefaction fan, a
+between the three resultant waves.  These waves consist of a rarefaction fan, a
 contact discontinuity, and a moderate shock.  The initial state is 
@@
 \ref{fig.sodshocktube}, we also show the integrated absolute deviation from 
-the exact solution, $||E_1|| = \sum_i \Delta x_i |F(x_i)-Exact(x_i)|$ 
+the exact solution, $||E_1|| = \sum_i \Delta x_i |F(x_i)-{\rm Exact}(x_i)|$ 
 , as well as the maximum deviation,
-$||E_\infty|| = \mathrm{max_i}\left\{ \Delta x_i|F(x_i)-Exact(x_i)| \right\}$.  
+$||E_\infty|| = \mathrm{max_i}\left\{ \Delta x_i|F(x_i)-{\rm Exact}(x_i)| \right\}$.  
 These numbers confirm the qualitative differences noted

### 2013-05-31 tests-vandv.tex
sweep 1: typo changed

@@
 them to serve as a vehicle for verifying that the code's behavior
-remains stable over time and after modifcation of the codebase.
+remains stable over time and after modification of the codebase.
 
@@
 analysis performed on all test problems includes calculation of
-various statistics, such as extrema; mean; and variance, on the fields
+various statistics, such as extrema, mean and variance, on the fields
 present in the output data.  Custom analysis provided by scripts that
@@
 to regenerate all of the figures in Section~\ref{sec.tests.compare}
-using the \textit{'methodpaper'} suite of test problems.
+using the \textit{`methodpaper'} suite of test problems.
 
@@
 \citep{2008CS&D....1a5003H}, as well as several direct comparisons
-between \enzo\ and the GADGET SPH code under a variety of
-circumstances, including N-body and adiabatic hydrodynamics
+between \enzo\ and the GADGET SPH code in a variety of
+set-ups, including N-body and adiabatic hydrodynamics
 \citep{2005ApJS..160....1O,2005MNRAS.364..909V, 2011MNRAS.418..960V}
 and simulations of the Lyman-alpha forest \citep{2007MNRAS.374..196R}.
-The \enzo\ code typically has a more difficult time resolving
+Compared to the other codes involved in these projects, the \enzo\ code typically has a more difficult time resolving
 small-scale self-gravitating structures (for an equivalent dark matter
@@
 \citep[e.g.][]{2011MNRAS.418..960V}, \enzo\ produces results that are
-similar to other high-order grid-based hydrodynamics codes, and far
+similar to other high-order grid-based hydrodynamics codes, and a far
 superior performance of \enzo\ is observed (in terms of resolution of
@@
 other codes.  We note, however, that there were minor differences throughout
-the comparison between various codes, and the majority of the codes differred
+the comparison between various codes, and the majority of the codes differed
 in at least a subset of the tests.
@@
 process in both the \enzo\ and Gadget-2 codes, observing similar trends between
-the two codes. However, at nominally equivalend resolution
+the two codes. However, at nominally equivalent resolution
 where the particle and cell gas masses are comparable, 

### 2013-05-31 parallel.tex
Fixing formatting issues with underscores in texttt.
Added a table of variables and constants used in the paper.
Minor revision to parallelism section.

@@
 Each grid object -- including all cell and grid data -- is fully
-contained in a single processor\footnote{In this context, we use the
-word {\it processor} to mean a basic distribution unit; this could by
+contained on a single processor\footnote{In this context, we use the
+word {\it processor} to mean a basic distribution unit; this could be
 either a core or a node, depending on details of the system.  Note
-that the current version of Enzo is not threaded, although the
-development of a hybrid version is in process}.  Parallelization is
+that the current version of Enzo is not threaded, although a hybrid
+MPI + OpenMP version is currently under development.}.  
+Parallelization is
 accomplished by distributing grids amongst processors.  This is done
@@
 split up into $N_{\rm root}$ tiles, with $N_{\rm root}$ typically
-equal to the number of processors $N_p$.
+equal to the number of processors, $N_p$.
 
@@
 have designed a simple way for the user to modify the source by adding
-\texttt{TIMER\_START(``Your_Routine_Name'')} and
-\texttt{TIMER\_END(``Your_Routine_Name'')}.  These timers are created
+\texttt{TIMER\_START("Your\_Routine\_Name")} and
+\texttt{TIMER\_END("Your\_Routine\_Name")}.  These timers are created
 and managed individually on each processor in an asynchronous fashion,
@@
 attributes such as the number of cell updates, the current number of
-grids, and the average cells/s/MPI process.  This information is then
+grids, and the average cells updates per second per MPI process.  This
+information is then 
 output to a logfile.  This provides a simplified interface to the user
@@
 process the data from the logfile.  These capabilities are described
-in the online documentation, along with a further discussion of the
+in the online documentation, along with further discussion of the
 performance measurement implementation.
@@
 \includegraphics[width=0.6\textwidth]{figures/enzo_unigrid_weak_scaling.eps}
-\caption{\enzo\ weak scaling performance for a set of Lyman alpha
+\caption{\enzo\ weak scaling performance for a set of Lyman-$\alpha$
 forest cosmology simulations with constant comoving spatial resolution
-per grid cell, showing cell updates per second per computational core
+per grid cell, showing cell updates per second per processor 
 plotted as a function of the number of root grid tiles of dimension
@@
 MPI processes per node, where each compute node contained a single
-quad-core AMD Opteron CPU having a speed of 2.1 GHz on Jaguar and 2.3
+quad-core AMD Opteron CPU with a speed of 2.1 GHz on Jaguar and 2.3
 GHz on Kraken.}
@@
 non-gravitating turbulence
-\citep[e.g.,][]{2002ApJ...569L.127K,Kritsuk04}, the Lyman-alpha forest
+\citep[e.g.,][]{2002ApJ...569L.127K,Kritsuk04}, the Lyman-$\alpha$ forest
 \citep{2005MNRAS.361...70J,2009MNRAS.399.1934P}, or feedback of
@@
 upon initialization, unigrid simulations are decomposed such that each
-MPI process has a roughly equal subvolume (and thus number of grid
+MPI process has a roughly equal subvolume (and thus an equal number of grid
 cells), meaning that work is evenly distributed.  Communication
@@
 Figure~\ref{fig.uniscale} shows \enzo\ weak scaling results for a
-sequence of scaled unigrid Lyman alpha forest calculations. These
+sequence of scaled unigrid Lyman-$\alpha$ forest calculations. These
 calculations include dark matter dynamics, hydrodynamics using the
-piecewise parabolic method, six-species non-equilibrium chemistry and
+PPM solver, six-species non-equilibrium chemistry and
 radiative cooling, and a uniform metagalactic ultraviolet background.
@@
 approximately $128^3$ dark matter particles.  The number of grid cells
+on each processor 
 was constant throughout the calculation; the number of dark matter
-particles varies as they are moved from subvolume to subvolume as
+particles on each processor varies as they are moved from subvolume to subvolume as
 structure evolves.  The grid resolution was kept at a constant
@@
 performance of the code is extremely good for this problem, with only
-a 20\% decrease in cell updates/second/task as the code is scaled from
+a 20\% decrease in cell updates per second per MPI task as the code is scaled from
 8 to 4,096 MPI tasks, and a 40\% decrease in performance overall going
@@
 boundaries and, thus, communication patterns can be unpredictable,
-hindering strong scaling behavior to high core counts.
+hindering strong scaling to high core counts.
 
@@
 \centerline{\includegraphics[width=0.7\textwidth]{figures/scaling_plot2}}
-\caption{This figure show (clockwise from upper left), the covering
+\caption{This figure shows (clockwise from upper left), the covering
 fraction $f_l$ of the grids on a given level $l$, the ratio of
@@
 cosmological runs, with only slight variations depending on the exact
-system that is being simulated.  The top-left panel
+problem that is being simulated.  The top-left panel
 shows the ratio of $f_{l}/f_{l+1}$ which, remarkably, hovers around
@@
 Figure~\ref{fig:scaling}.  This demonstrates that the memory usage is
-dominated by the top three levels, and adding additional levels only
+dominated by the top three levels, and adding additional levels 
 adds minimally to the memory usage of the run.  We have empirically

### 2013-05-31 code-tests.tex

@@
 a suite of test problems that can be compared to previous versions of
-the code, to ensure that \enzo\ is running correctly on a new
+the code, used to ensure that \enzo\ is running correctly on a new
 computer, with a new compiler, or after substantial changes have been

### 2013-05-31 tests-doublemach.tex

@@
 Figure~\ref{fig.doublemach}, a shock is injected at an angle to a
-reflecting surface (the -y boundary), and a jet appears along the
+reflecting surface (the -y-boundary), and a jet appears along the
 reflecting surface.  The ideal solution is self-similar, and the
@@
 x-axis and an initial position on the lower boundary of x$ = 1/6$.
-The lower y boundary and right x boundary are reflecting; the left x
-boundary is inflowing, and the upper y boundary has a time-dependent
+The lower y-boundary and right x-boundary are reflecting; the left x
+boundary is inflowing, and the upper y-boundary has a time-dependent
 boundary condition that allows the shock to propagate into the domain
@@
 Mach 10 shock is injected into the domain with a shock normal that is
-30$^\circ$ from the x-axis with a time-dependent +y boundary condition
+30$^\circ$ from the x-axis with a time-dependent +y-boundary condition
 that mimics a shock of infinite length.  The solution is self-similar,

### 2013-05-31 tests-shockpool.tex

@@
 The density and pressure in the domain is initially set equal to 1.0 with zero velocity.
-At $t=0$ the left boundary is set with the density, pressure and velocity appropriate for $\mathcal{M}=2$ shock wave.
+At $t=0$ the left boundary is set with the density, pressure and velocity appropriate for a $\mathcal{M}=2$ shock wave.
 
@@
 
-For Zeus (middle row of Figure~\ref{fig.shockpool}, the shock is broadened because of artificial viscosity and there are slightly more post-shock oscillations, although again quite mild.
+For Zeus (middle row of Figure~\ref{fig.shockpool}), the shock is broadened because of artificial viscosity and there are slightly more post-shock oscillations, although again quite mild.
 The impact of entering and exiting the refined region is somewhat larger than in PPM; however, the most noticeable difference is the incorrect position of the shock front.

### 2013-05-31 appendix.tex
reformatted appendix (no changes made yet).

@@
 
-In this appendix, we provide details for the various interpolation methods available in the code.
-We assume throughout that we are dealing with Cartesian coordinates and equally spaced cells, although we allow for arbitrary refinement factor $r$.
+In this appendix, we provide details for the various interpolation
+methods available in the code.  We assume throughout that we are
+dealing with Cartesian coordinates and equally spaced cells, although
+we allow for arbitrary refinement factor $r$.
 
@@
 
-This interpolation algorithm is generally second-order but has monotonicity constraints as described below.
+This interpolation algorithm is generally second-order but has
+monotonicity constraints as described below.
 
-In one dimension, we define the parent values as $Q_{-1}$, $Q_0$, and $Q_{+1}$, where the central parent cell has a left edge at $x_0$ and width $\Delta x$.
-We first linearly interpolate the parent values to the cell edges: $Q_{-1/2} = (Q_0 + Q_{-1})/2$, and similarly for $Q_{1/2} = (Q_0 + Q_1)/2$, and then compute a monotonic slope: $\Delta Q_0 = \minmod{Q_{1/2} - Q_0}{Q_0 - Q_{-1/2}}$ where
+In one dimension, we define the parent values as $Q_{-1}$, $Q_0$, and
+$Q_{+1}$, where the central parent cell has a left edge at $x_0$ and
+width $\Delta x$. We first linearly interpolate the parent values to
+the cell edges: $Q_{-1/2} = (Q_0 + Q_{-1})/2$, and similarly for
+$Q_{1/2} = (Q_0 + Q_1)/2$, and then compute a monotonic slope: $\Delta
+Q_0 = \minmod{Q_{1/2} - Q_0}{Q_0 - Q_{-1/2}}$ where
+
 \begin{equation}
@@
 \end{equation}
+
 This slope is then used to compute the interpolated subgrid values.
-Defining $q_i$ to be the subgrid values at cell centers for the $r$ cells corresponding to parent cell $Q_0$ (for refinement factor $r$), we can write:
+Defining $q_i$ to be the subgrid values at cell centers for the $r$
+cells corresponding to parent cell $Q_0$ (for refinement factor $r$),
+we can write:
+
 \begin{equation}
@@
 \end{equation}
-where $f_x = 2 \Delta Q_0$ (this notation is used for consistency with the 2 and 3 dimensional cases).
 
-In two dimensions, the procedure is very similar in that we linearly interpolate parent values to the four cell corners: $Q_{-1/2, -1/2}$, $Q_{-1/2, 1/2}$, $Q_{1/2, -1/2}$, $Q_{1/2, 1/2}$ (by averaging parent cell-centered parent values).
-We then compute monotonic slopes across the two diagonals, since for a linear function, extrema occur at corners:
+where $f_x = 2 \Delta Q_0$ (this notation is used for consistency with
+the 2 and 3 dimensional cases).
+
+In two dimensions, the procedure is very similar in that we linearly
+interpolate parent values to the four cell corners: $Q_{-1/2, -1/2}$,
+$Q_{-1/2, 1/2}$, $Q_{1/2, -1/2}$, $Q_{1/2, 1/2}$ (by averaging parent
+cell-centered parent values).  We then compute monotonic slopes across
+the two diagonals, since for a linear function, extrema occur at
+corners:
+
 \begin{eqnarray}
@@
 \end{eqnarray}
-We then translate these slopes into the grid axes with $f_x = \Delta Q_0 -  \Delta Q_1$ and $f_y = \Delta Q_0 + \Delta Q_1$ so that the interpolation itself can be written simply as
+
+We then translate these slopes into the grid axes with $f_x = \Delta
+Q_0 -  \Delta Q_1$ and $f_y = \Delta Q_0 + \Delta Q_1$ so that the
+interpolation itself can be written simply as
+
 \begin{equation}
@@
 
-Finally, we write down the three-dimensional version -- unfortunately, here the number of monotonicity constraints is four (the 4 diagonals across the 8 opposing corners of the cube), while the number of slopes is three, so the problem is over constrained.
-Somewhat arbitrarily, we adopt the following procedure.
-As before, we compute $\Delta Q_0$, $\Delta Q_1$, $\Delta Q_2$, and $\Delta Q_3$ with the minmod limiter across the 4 diagonals.
-We then define
+Finally, we write down the three-dimensional version -- unfortunately,
+here the number of monotonicity constraints is four (the 4 diagonals
+across the 8 opposing corners of the cube), while the number of slopes
+is three, so the problem is over constrained.  Somewhat arbitrarily,
+we adopt the following procedure.  As before, we compute $\Delta Q_0$,
+$\Delta Q_1$, $\Delta Q_2$, and $\Delta Q_3$ with the minmod limiter
+across the 4 diagonals.  We then define
+
 \begin{equation}
@@
 \end{equation}
-which is the value of $\Delta Q_0$ that the other slopes ($\Delta Q_1$, $\Delta Q_2$, and $\Delta Q_3$) imply, normalized by the desired value of $\Delta Q_0$ itself.
-If $0 < s < 1$, then no adjustment needs to be made, as the monotonicity constraints are met.
-If these equalities are not met, then, if $s<0$ we define $\chi_n = 1$ if $\Delta Q_n/\Delta Q_0 < 0$, and 0 otherwise (if $s > 1$, this is reversed, so $\chi_n = 1$ if $\Delta Q_n/\Delta Q_0 > 0$, and 0 otherwise).
-These weights are used to determine which slopes to modify to match the $\Delta Q_0$ constraint.
-We then compute the amount of adjustment required:
+
+which is the value of $\Delta Q_0$ that the other slopes ($\Delta
+Q_1$, $\Delta Q_2$, and $\Delta Q_3$) imply, normalized by the desired
+value of $\Delta Q_0$ itself.  If $0 < s < 1$, then no adjustment
+needs to be made, as the monotonicity constraints are met.  If these
+equalities are not met, then, if $s<0$ we define $\chi_n = 1$ if
+$\Delta Q_n/\Delta Q_0 < 0$, and 0 otherwise (if $s > 1$, this is
+reversed, so $\chi_n = 1$ if $\Delta Q_n/\Delta Q_0 > 0$, and 0
+otherwise).  These weights are used to determine which slopes to
+modify to match the $\Delta Q_0$ constraint.  We then compute the
+amount of adjustment required:
+
 \begin{equation}
@@
 \end{equation}
-where $s^\prime = \min(\max(s,0), 1)$ and $\epsilon$ is a small number to prevent numerical errors.
-We then use this adjustment fraction to compute the new adjusted $\Delta Q_n$ (for $n = 1, 2, 3$) that match the $\Delta Q_0$ constraint as closely as possible with
+
+where $s^\prime = \min(\max(s,0), 1)$ and $\epsilon$ is a small number
+to prevent numerical errors. We then use this adjustment fraction to
+compute the new adjusted $\Delta Q_n$ (for $n = 1, 2, 3$) that match
+the $\Delta Q_0$ constraint as closely as possible with
+
 \begin{equation}
@@
 \end{equation}
-Finally, these are converted to the grid axes with $f_x = \Delta Q_2 + \Delta Q_3$, $f_y = \Delta Q_1 + \Delta Q_3$, $f_z = \Delta Q_1 + \Delta Q_2$ and then we do the interpolation itself with,
+
+Finally, these are converted to the grid axes with $f_x = \Delta Q_2 +
+\Delta Q_3$, $f_y = \Delta Q_1 + \Delta Q_3$, $f_z = \Delta Q_1 +
+\Delta Q_2$ and then we do the interpolation itself with,
+
 \begin{equation}
@@
 We also provide a variant on the above procedure, with two changes.
-The first is that the slopes across the diagonals ($\Delta Q_0$, etc.) are computed directly rather than with the minmod limiter (e.g.
-$\Delta Q_0 = (Q_{1/2,1/2,1/2} - Q_{-1/2,-1/2,-1/2})/2$).
-To ensure positivity in the resulting interpolation when applied to positive conserved quantities, the slopes are limited so that the smallest corner value is 0.2 the cell center value.
-The same procedure as described above is then applied to turn these four slopes into a linear interpolation.
+The first is that the slopes across the diagonals ($\Delta Q_0$, etc.)
+are computed directly rather than with the minmod limiter
+(e.g. $\Delta Q_0 = (Q_{1/2,1/2,1/2} - Q_{-1/2,-1/2,-1/2})/2$).  To
+ensure positivity in the resulting interpolation when applied to
+positive conserved quantities, the slopes are limited so that the
+smallest corner value is 0.2 the cell center value.  The same
+procedure as described above is then applied to turn these four slopes
+into a linear interpolation.
 
@@
 
-This is a completely different second-order interpolation scheme, which is based on Cloud-In-Cell (CIC) interpolation \citep{Hockney88}.
-In one dimension, we define the parent values as $Q_0$, and $Q_{+1}$, where the left parent cell has a cell center at $x_0$ and width $\Delta x$.
-Then, the interpolated value for a subgrid cell $q_i$ with a cell left edges at $x_i = x_0 + i \Delta x^p/r$ where $i$ runs from 0 to $r-1$ for refinement factor $r$, is given by:
+This is a completely different second-order interpolation scheme,
+which is based on Cloud-In-Cell (CIC) interpolation \citep{Hockney88}.
+In one dimension, we define the parent values as $Q_0$, and $Q_{+1}$,
+where the left parent cell has a cell center at $x_0$ and width
+$\Delta x$.  Then, the interpolated value for a subgrid cell $q_i$
+with a cell left edges at $x_i = x_0 + i \Delta x^p/r$ where $i$ runs
+from 0 to $r-1$ for refinement factor $r$, is given by:
+
 \begin{equation}
@@
 \end{equation}
-The extension to two and three dimensions is straightforward, with weights in the other dimensions computed in the same way and then multiplied to get a total of 4 and 8 weights for the 2 and 3 dimensional cases, respectively.
-This scheme is monotonicity preserving, but is not conservative.
+
+The extension to two and three dimensions is straightforward, with
+weights in the other dimensions computed in the same way and then
+multiplied to get a total of 4 and 8 weights for the 2 and 3
+dimensional cases, respectively.  This scheme is monotonicity
+preserving, but is not conservative.
 
@@
 
-This interpolation method provides a third-order accuracy based on the Triangular Shaped Cloud (TSC) methodology \citep{Hockney88}.
-As usual, in one dimension, we define the parent values as $Q_{-1}$, $Q_0$, and $Q_{+1}$, where the central parent cell has a left edge at $x_0$ and width $\Delta x$.
-Then, the interpolated value for a subgrid cell $q_i$ with a cell left edges at $x_i = x_0 + i \Delta x^p/r$ where $i$ runs from 0 to $r-1$ for refinement factor $r$, is given by:
+This interpolation method provides a third-order accuracy based on the
+Triangular Shaped Cloud (TSC) methodology \citep{Hockney88}.  As
+usual, in one dimension, we define the parent values as $Q_{-1}$,
+$Q_0$, and $Q_{+1}$, where the central parent cell has a left edge at
+$x_0$ and width $\Delta x$.  Then, the interpolated value for a
+subgrid cell $q_i$ with a cell left edges at $x_i = x_0 + i \Delta
+x^p/r$ where $i$ runs from 0 to $r-1$ for refinement factor $r$, is
+given by:
+
 \begin{equation}
@@
 \end{equation}
+
 and the weights are given by:
+
 \begin{equation}
@@
 \end{equation}
-and $b(i) = 1/r - a(i) - c(i)$.
-The extension to two and three dimensions is straightforward with the weights computed in the same way as for the one-dimensional case but then multiplied to determine the 9 and 27 factors necessary for the 2 and 3 dimensional cases, respectively.
 
-One problem with this interpolation technique is that it is not conservative.  In particular, the sum of the interpolated subgrid values:
+and $b(i) = 1/r - a(i) - c(i)$.  The extension to two and three
+dimensions is straightforward with the weights computed in the same
+way as for the one-dimensional case but then multiplied to determine
+the 9 and 27 factors necessary for the 2 and 3 dimensional cases,
+respectively.
+
+One problem with this interpolation technique is that it is not
+conservative.  In particular, the sum of the interpolated subgrid
+values:
+
 \begin{equation}
@@
 \end{equation}
-is not, in general, equal to $Q_0$.
-We can retain conservation by adding the factor $(Q_0 - \tilde{Q_0})/r$ to the interpolated values (or by multiplying the interpolated values by the ratio $Q_0/\tilde{Q_0}$).
-Unfortunately, the result of this procedure does not preserve monotonicity -- it can introduce local minima and maxima at at parent cell boundaries.
-We can then attempt to correct that by taking a weighted average between the interpolated values on the subgrid and the parent value with weights computed such that the interpolated values at the edge of the parent cell are not local maxima compared to the interpolated values in the neighboring parent cell.
+
+is not, in general, equal to $Q_0$.  We can retain conservation by
+adding the factor $(Q_0 - \tilde{Q_0})/r$ to the interpolated values
+(or by multiplying the interpolated values by the ratio
+$Q_0/\tilde{Q_0}$).  Unfortunately, the result of this procedure does
+not preserve monotonicity -- it can introduce local minima and maxima
+at at parent cell boundaries.  We can then attempt to correct that by
+taking a weighted average between the interpolated values on the
+subgrid and the parent value with weights computed such that the
+interpolated values at the edge of the parent cell are not local
+maxima compared to the interpolated values in the neighboring parent
+cell.
 
@@
 
-Finally, for completeness, we include a first-order accurate piecewise constant interpolator, for which, using the same definitions in the previous case, we take $q_i = Q_0$, for $i = 0$ to $r-1$.
+Finally, for completeness, we include a first-order accurate piecewise
+constant interpolator, for which, using the same definitions in the
+previous case, we take $q_i = Q_0$, for $i = 0$ to $r-1$.
 

### 2013-05-31 appendix.tex
edited appendix

@@
 methods available in the code.  We assume throughout that we are
-dealing with Cartesian coordinates and equally spaced cells, although
-we allow for arbitrary refinement factor $r$.
+dealing with Cartesian coordinates and cells of equal sizes, although
+we allow for arbitrary an refinement factor $r$ between cells at
+different levels.
 
@@
 
-This interpolation algorithm is generally second-order but has
+This interpolation algorithm is generally second-order, but has
 monotonicity constraints as described below.
-
 In one dimension, we define the parent values as $Q_{-1}$, $Q_0$, and
@@
 to prevent numerical errors. We then use this adjustment fraction to
-compute the new adjusted $\Delta Q_n$ (for $n = 1, 2, 3$) that match
+compute the new, adjusted $\Delta Q_n$ (for $n = 1, 2, 3$) that match
 the $\Delta Q_0$ constraint as closely as possible with
@@
 \Delta Q_3$, $f_y = \Delta Q_1 + \Delta Q_3$, $f_z = \Delta Q_1 +
-\Delta Q_2$ and then we do the interpolation itself with,
+\Delta Q_2$.  We then do the interpolation itself with
 
@@
 positive conserved quantities, the slopes are limited so that the
-smallest corner value is 0.2 the cell center value.  The same
-procedure as described above is then applied to turn these four slopes
-into a linear interpolation.
+smallest corner value is 0.2 the cell center value.  The procedure
+described above is then applied to turn these four slopes into a
+linear interpolation.
 
@@
 $\Delta x$.  Then, the interpolated value for a subgrid cell $q_i$
-with a cell left edges at $x_i = x_0 + i \Delta x^p/r$ where $i$ runs
+with a cell left edges at $x_i = x_0 + i \Delta x^p/r$, where $i$ runs
 from 0 to $r-1$ for refinement factor $r$, is given by:
@@
 multiplied to get a total of 4 and 8 weights for the 2 and 3
-dimensional cases, respectively.  This scheme is monotonicity
-preserving, but is not conservative.
+dimensional cases, respectively.  This scheme preserves monotonicity,
+but is not conservative.
 

### 2013-06-03 tests-sodshocktube.tex
SodShockTube text change to match figure

@@
 \ref{fig.sodshocktube}, we also show the integrated absolute deviation from 
-the exact solution, $||E_1|| = \sum_i \Delta x_i |F(x_i)-{\rm Exact}(x_i)|$ 
-, as well as the maximum deviation,
-$||E_\infty|| = \mathrm{max_i}\left\{ \Delta x_i|F(x_i)-{\rm Exact}(x_i)| \right\}$.  
+the exact solution, $||E_1|| = \sum_i \Delta x_i |F(x_i)-{\rm Exact}(x_i)|$.
+%, as well as the maximum deviation,
+%$||E_\infty|| = \mathrm{max_i}\left\{ \Delta x_i|F(x_i)-{\rm Exact}(x_i)| \right\}$.  
 These numbers confirm the qualitative differences noted
@@
 each panel, we show the analytic solution as a solid line and the $E_1$
-and $E_\infty$ error norms in the upper right.}
+error norm in the upper right.}
 \label{fig.sodshocktube}

### 2013-06-05 app-chemistry.tex
Edits to the overview section. Trying to unify notations, so several other sections are also affected.

@@
 
-The primordial chemistry network implemented in Enzo is discussed in 
+The primordial chemistry network implemented in \enzo\ is discussed in 
 Section~\ref{sec.ov.chem} and also in
@@
 Tables~\ref{table.collisional} and~\ref{table.radiative} summarize the collisional
-and radiative processes solved in the Enzo nonequilibrium chemistry routines.  
+and radiative processes solved in the \enzo\ nonequilibrium chemistry routines.  
 \citet{abel97} show that accurate results can be obtained if several unnecessary

### 2013-06-05 numerical-conduction.tex
Edits to the overview section. Trying to unify notations, so several other sections are also affected.

@@
 \begin{equation}
-\myvec{F} = -\kappa_{sp} \nabla T
+\myvec{F} = -\kappa_{\rm sp} \grad T
 \end{equation}
-where $\kappa_{sp}$ is the Spitzer conduction coefficient, $4.6 \times 10^{-7}$~T$^{5/2}$ erg
+where $\kappa_{\rm sp}$ is the Spitzer conduction coefficient, $4.6 \times 10^{-7}$~T$^{5/2}$ erg
 s$^{-1}$~cm$^{-1}$~K$^{-1}$ \citep{1962pfig.book.....S}, here we are using a
@@
 \begin{equation}
-\kappa_{eff} = \frac{\kappa_{sp}}{1 + 4.2 \lambda_e / \ell_T} \; ,
+\kappa_{eff} = \frac{\kappa_{\rm sp}}{1 + 4.2 \lambda_e / \ell_T} \; ,
 \end{equation}
 where $\lambda_e$ is the electron mean free path and $\ell_T \equiv T
-/ |\nabla T|$ is the characteristic length scale of the local
+/ |\grad T|$ is the characteristic length scale of the local
 temperature gradient.  We also assume that the conductivity of the
 plasma can be described in terms of an effective conductivity, which
-can be expressed as a fraction f$_{Sp}$ of the Spitzer conductivity
-(where f$_{Sp} \leq 1.0$ are considered physical values).  This takes
+can be expressed as a fraction f$_{\rm sp}$ of the Spitzer conductivity
+(where f$_{\rm sp} \leq 1.0$ are considered physical values).  This takes
 into account physical processes below the resolution limit of the
@@
 \begin{equation}
-\myvec{F} = -\kappa_{sp} \hat{b} (\hat{b} \cdot \nabla T)
+\myvec{F} = -\kappa_{\rm sp} \hat{b} (\hat{b} \cdot \grad T)
 \end{equation}
@@
 magnetic field.  As with the isotropic thermal conduction, we allow a
-multiplicative factor f$_{Sp}$ to take into account the possible
+multiplicative factor f$_{\rm sp}$ to take into account the possible
 behavior of magnetic fields below the resolution limit of the

### 2013-06-05 numerical-mhd-ct.tex
Edits to the overview section. Trying to unify notations, so several other sections are also affected.

@@
 \def\Evec{{\bf E}}
-\def\Divb{\ensuremath{\nabla \cdot \Bvec}}
+\def\Divb{\ensuremath{\div \Bvec}}
 
@@
 \begin{eqnarray}
-\partial \Bvec/\partial t & =& \frac{1}{a} \nabla \times \Evec\\
+\partial \Bvec/\partial t & =& \frac{1}{a} \curl \Evec\\
 \Divb & =& 0.
@@
 
-Enzo MHD-CT has been described in great detail in \citet{Collins10},
+\enzo\ MHD-CT has been described in great detail in \citet{Collins10},
 and we direct the interested reader to that reference.  The base
@@
 the form of area-averaged face and length-averaged edge-centered
-variables, while the rest of Enzo employs predominantly
+variables, while the rest of \enzo\ employs predominantly
 volume-averaged cell centered variables.  The magnetic field is
@@
 Divergence-free AMR is somewhat more complex than the AMR employed
-elsewhere in Enzo.  First, the interpolation must be constrained to be
+elsewhere in \enzo.  First, the interpolation must be constrained to be
 divergence free.  Thus, all three face-centered field components are
@@
 $$ 
-\frac{\partial \vecB}{\partial t} - \frac{1}{a}  \nabla \times (\vecv \times
+\frac{\partial \vecB}{\partial t} - \frac{1}{a}  \curl (\vecv \times
 \vecB)  =  -\frac{\dot{a}}{2 a} \vecB. \label{eq:induction_cosmo_source}
@@
 $$ 
-\frac{\partial \vecB^\prime}{\partial t} - \frac{1}{a}  \nabla \times (\vecv \times
+\frac{\partial \vecB^\prime}{\partial t} - \frac{1}{a}  \curl (\vecv \times
 \vecB^\prime)  = 0, \label{eq:induction_cosmo_prime_nosource}
@@
 \begin{eqnarray}
-  \frac{\partial S}{\partial t} +\frac{1}{a} \nabla \cdot (S \vecv) & = &  -
+  \frac{\partial S}{\partial t} +\frac{1}{a} \div (S \vecv) & = &  -
   \frac{ 3 (\gamma - 1) \dot{a}}{a} S \\
-  \frac{\partial \rho e}{\partial t} + \frac{1}{a} \nabla \cdot (\rho e \vecv) &
-  = & -\frac{ 3(\gamma -1 ) \dot{a}}{a} + \frac{p}{a} \nabla \cdot \vecv,
+  \frac{\partial \rho e}{\partial t} + \frac{1}{a} \div (\rho e \vecv) &
+  = & -\frac{ 3(\gamma -1 ) \dot{a}}{a} + \frac{p}{a} \div \vecv,
 \end{eqnarray}
@@
 
-Enzo's MHD-CT has been used in a variety of contexts, from galaxy
+\enzo's MHD-CT has been used in a variety of contexts, from galaxy
 cluster evolution \citep{Xu11,Xu12}, synthetic observations of radio

### 2013-06-05 numerical-mhd-dedner.tex
a little more text.
Edits to the overview section. Trying to unify notations, so several other sections are also affected.

@@
 where $c_h$ is a constant controlling the propagation speed and
-damping rate of $\nabla\cdot B$. $h=e+p/\rho$ denotes the enthalpy.
+damping rate of $\div B$. $h=e+p/\rho$ denotes the enthalpy.
 
@@
 solvers.
+
+The method is dimensionally un-split in that the fluxes are computed
+for all dimensions first and the conserved quantities are updated in
+one go in contrast to the Strang splitting employed in the Zeus and
+PPM solvers described above.

### 2013-06-05 numerical-overview.tex
Edits to the overview section. Trying to unify notations, so several other sections are also affected.

@@
 \begin{equation}
-\frac{\partial e}{\partial t} + \frac{1}{a} \vec{v}_b \cdot \nabla e = 
-- \frac{3(\gamma-1)\dot{a}}{a}e - \frac{p}{a \rho} \nabla \cdot \vec{v}_b
+\frac{\partial e}{\partial t} + \frac{1}{a} \vec{v}_b \cdot \grad e = 
+- \frac{3(\gamma-1)\dot{a}}{a}e - \frac{p}{a \rho} \div \vec{v}_b
 \label{enzointenergy}
@@
 \begin{eqnarray}
-\rho \frac{\partial\textbf{v}}{\partial t} &=& - \nabla p - \rho \nabla \phi 
-- \nabla \cdot \textbf{Q} \\
-\frac{\partial e}{\partial t} &=& -p \nabla \cdot \textbf{v} - \textbf{Q} : \nabla \textbf{v}, 
+\rho \frac{\partial\textbf{v}}{\partial t} &=& - \grad p - \rho \grad \phi 
+- \div \textbf{Q} \\
+\frac{\partial e}{\partial t} &=& -p \div \textbf{v} - \textbf{Q} : \grad \textbf{v}, 
 \end{eqnarray}
@@
 \end{tabular}
-\caption[]{Collisional processes solved in the Enzo nonequilibrium
+\caption[]{Collisional processes solved in the \enzo\ nonequilibrium
 primordial chemistry routines.}
@@
 \end{tabular}
-\caption[]{Radiative processes solved in the Enzo nonequilibrium
+\caption[]{Radiative processes solved in the \enzo\ nonequilibrium
 primordial chemistry routines.}
@@
 \begin{equation}
-\nabla \cdot \vec{v}_b < 0,
+\div \vec{v}_b < 0,
 \label{cencont}

### 2013-06-05 numerical-ppm.tex
Edits to the overview section. Trying to unify notations, so several other sections are also affected.

@@
  \frac{\partial e}{\partial t} 
-           + \frac{1}{a} \vecv \cdot \nabla e
+           + \frac{1}{a} \vecv \cdot \grad e
 %         = - \frac{3(\gamma -1)\dot{a}}{a} e
          = - 3 \frac{\dot{a}}{a} \frac{p}{\rho}
-           - \frac{p}{a\rho} \nabla \cdot \vecv
+           - \frac{p}{a\rho} \div \vecv
 \end{equation}
@@
 an iterative process.  Thus a series of \emph{approximate} Riemann solvers are
-typically used.  There are four approximate Riemann solvers in Enzo: two-shock
+typically used.  There are four approximate Riemann solvers in \enzo: two-shock
 \citep{toro-1997},
@@
 approximations can be formulated for the MHD equations, and are employed in both
-Dedner and CT solvers in Enzo.  The HLLD
+Dedner and CT solvers in \enzo.  The HLLD
 solver includes two of the additional waves, the rotational discontinuities,
@@
 %\begin{equation}
-%\frac{\partial e}{\partial t} + \frac{1}{a} \vecv_b \cdot \nabla e = 
-%- \frac{3(\gamma-1)\dot{a}}{a}e - \frac{p}{a \rho} \nabla \cdot \vecv_b
+%\frac{\partial e}{\partial t} + \frac{1}{a} \vecv_b \cdot \grad e = 
+%- \frac{3(\gamma-1)\dot{a}}{a}e - \frac{p}{a \rho} \div \vecv_b
 %\label{enzointenergy}

### 2013-06-05 numerical-radiation-fld.tex
Edits to the overview section. Trying to unify notations, so several other sections are also affected.

@@
 
-Enzo's field-based radiation solver focuses on
+\enzo's field-based radiation solver focuses on
 a flux-limited diffusion approximation for cosmological radiative
@@
 %  \label{eq:fld_radiation}
-%  \partial_t E_r + \frac1a \nabla\cdot\left(E_r \vec{v}_b\right) &=& 
-%    \nabla\cdot\left(D\nabla E_r\right) -
+%  \partial_t E_r + \frac1a \div \left(E_r \vec{v}_b\right) &=& 
+%    \div\left(D\grad E_r\right) -
 %    \frac{\dot{a}}{a} E_r - c\kappa E_r + \eta, \\
@@
 %  \partial_t {\mathrm n}_i + 
-%    \frac1a \nabla\cdot\left({\mathrm n}_i \vec{v}_b\right) &=& 
+%    \frac1a \div\left({\mathrm n}_i \vec{v}_b\right) &=& 
 %    \alpha_{i,j} {\mathrm n}_e {\mathrm n}_j - {\mathrm n}_i
@@
 (Equation~\ref{eq:species_evolution}) is solved
-independently of Enzo's hydrodynamics, gravity and dark-matter solvers
+independently of \enzo's hydrodynamics, gravity and dark-matter solvers
 (Sections \ref{sec.hydro.ppm}-\ref{sec.ov.nbody}), thereby allowing

### 2013-06-05 numerical-timestep.tex
Edits to the overview section. Trying to unify notations, so several other sections are also affected.

@@
 \begin{equation}
-\Delta t_{\rm cond} = \min \left(  \frac{ \kappa_{\rm cond}}{f_{\rm Sp}} \frac{\Delta x^2
+\Delta t_{\rm cond} = \min \left(  \frac{ \kappa_{\rm cond}}{f_{\rm sp}} \frac{\Delta x^2
     n_b}{\kappa(T)} \right)_L,
@@
 %m_p$, where $\mu$ is the mean molecular weight, 
-$\kappa(T)$ is the Spitzer thermal conductivity, and f$_{Sp}$ is a user-defined
-conduction suppression factor whose value must be f$_{Sp} \leq 1$.  
+$\kappa(T)$ is the Spitzer thermal conductivity, and f$_{\rm sp}$ is a user-defined
+conduction suppression factor whose value must be f$_{\rm sp} \leq 1$.  
 %Both are defined in Section~\ref{sec.num.conductions}.  

### 2013-06-05 overview.tex
Edits to the overview section. Trying to unify notations, so several other sections are also affected.

@@
 describe the numerical algorithms that we use to solve these
-equations.  This section is intended to be an overview of the
-\enzo\ code's capabilities: thus, we gather all of the equations
+equations.  This section is intended to be an overview of
+\enzo's capabilities: thus, we gather all of the equations
 solved into a single place and provide a brief and high-level
@@
 
-The Eulerian equations of ideal magnetohydrodynamics including
+The Eulerian equations of ideal magnetohydrodynamics (MHD) including
 gravity, in a coordinate systems comoving with the cosmological
-expansion, are given by:
+expansion, are given by
 
@@
   \frac{\partial \rho}{\partial t} 
-  + \frac{1}{a} \nabla \cdot (\rho \vecv) & = & 0
+  + \frac{1}{a} \div (\rho \vecv) & = & 0,
   \label{eq:mass} \\
@@
   \frac{\partial \rho \vecv}{\partial t}  
-  + \frac{1}{a} \nabla \cdot \left(\rho \vecv \vecv + \myvec{I}p^* - 
+  + \frac{1}{a} \div \left(\rho \vecv \vecv + \myvec{I}p^* - 
     \frac{\vecB \vecB}{a} \right) & = &
-  - \frac{\dot{a}}{a} \rho \vecv - \frac{1}{a} \rho \nabla \phi
+  - \frac{\dot{a}}{a} \rho \vecv - \frac{1}{a} \rho \grad \phi,
 % 
-%          + \frac{1}{a} ( \vecv \cdot \nabla ) \vecv
+%          + \frac{1}{a} ( \vecv \cdot \grad ) \vecv
 %    & = & - \frac{\dot{a}}{a} \vecv 
-%          - \frac{1}{a \rho} \nabla p
+%          - \frac{1}{a \rho} \grad p
 %          - \frac{1}{a \rho} (\nalba \times \vecB) \times \vecB
-%          - \frac{1}{a} \nabla \phi, 
+%          - \frac{1}{a} \grad \phi, 
   \label{eq:momentum} \\
@@
   \frac{\partial E}{\partial t} 
-  + \frac{1}{a} \nabla \cdot \left[ (E + p^*) \vecv - 
+  + \frac{1}{a} \div \left[ (E + p^*) \vecv - 
     \frac{1}{a} \vecB (\vecB \cdot \vecv) \right] & = &
   - \frac{\dot{a}}{a} \left( 2E - \frac{B^2}{2a} \right) - 
-  \frac{1}{a} \vecv \cdot \nabla \phi 
-  - \Lambda + \Gamma + \frac{1}{a^2} \nabla \cdot \fcond
+  \frac{1}{a} \vecv \cdot \grad \phi 
+  - \Lambda + \Gamma + \frac{1}{a^2} \div \fcond,
 % 
-%         + \frac{1}{a} \vecv \cdot \nabla E
+%         + \frac{1}{a} \vecv \cdot \grad E
 %  & = & - \frac{\dot{a}}{a} (3 \frac{p}{\rho} + {\vecv \cdot \vecv} )
-%       - \frac{1}{a \rho } \nabla \cdot (p \vecv) 
+%       - \frac{1}{a \rho } \div (p \vecv) 
 %%\nonumber   & &   %uncomment for 2-column
-%        - \frac{1}{a} \vecv \cdot \nabla \phi 
+%        - \frac{1}{a} \vecv \cdot \grad \phi 
 %               + \Gamma - \Lambda.
@@
   \frac{\partial \vecB}{\partial t} - 
-  \frac{1}{a}  \nabla \times (\vecv \times \vecB) & = & 
-  -\frac{\dot{a}}{2 a} \vecB \label{eq:induction}
+  \frac{1}{a}  \grad \times (\vecv \times \vecB) & = & 
+  -\frac{\dot{a}}{2 a} \vecB. \label{eq:induction}
 \end{eqnarray}
@@
 %\frac{\partial \rho}{\partial t} 
-%          + \vecv \cdot \nabla \rho
-%    & = & - \rho \nabla \cdot \vecv,
+%          + \vecv \cdot \grad \rho
+%    & = & - \rho \div \vecv,
 %        \label{eq:mass} \\
@@
 %\frac{\partial \vecv }{ \partial t} 
-%          + ( \vecv \cdot \nabla ) \vecv 
-%    & = & - \frac{1}{\rho} \nabla p
-%          - \nabla \phi,
+%          + ( \vecv \cdot \grad ) \vecv 
+%    & = & - \frac{1}{\rho} \grad p
+%          - \grad \phi,
 %        \label{eq:momentum} \\
 %%
-% \frac{\partial E}{\partial t} + \vecv \cdot \nabla E
-%    & = & - \frac{1}{\rho} \nabla \cdot (p \vecv) 
-%          - \vecv \cdot \nabla \phi
+% \frac{\partial E}{\partial t} + \vecv \cdot \grad E
+%    & = & - \frac{1}{\rho} \div (p \vecv) 
+%          - \vecv \cdot \grad \phi
 %%\nonumber          \\  & &     % uncomment for 2-column
@@
 In these equations, $E$, $\rho$, $\vecv$, and $\vecB$ are the comoving
-total fluid energy density, comoving baryonic density, peculiar velocity, and
-comoving magnetic field strength, respectively.  $\myvec{I}$ is the
-identity matrix, and $a$ is the cosmological expansion parameter
-(which is discussed in more detail below).  The first equation
-represents conservation of mass, the second, conservation of momentum,
-and the third, conservation of total fluid (kinetic plus thermal)
-energy.  They are respectively, the first, second and third moments of
-the Boltzmann equation.  The fourth equation is the magnetic induction
-equation.  Terms representing radiative cooling ($\Lambda$) and
-heating ($\Gamma$) enter on the right-hand side of the energy equation
-(\ref{eq:total_energy}), as well as the flux due to thermal heat
-conduction ($\fcond$).
+total fluid energy density, comoving gas density, peculiar
+velocity, and comoving magnetic field strength, respectively.
+$\myvec{I}$ is the identity matrix, and $a$ is the cosmological
+expansion parameter (discussed in more detail below).  The first
+equation represents conservation of mass, the second conservation of
+momentum, and the third conservation of total (kinetic plus thermal)
+fluid energy.  They are respectively, the first, second, and third
+moments of the Boltzmann equation.  The fourth equation is the
+magnetic induction equation.  Terms representing radiative cooling
+($\Lambda$) and heating ($\Gamma$) enter on the right-hand side of the
+energy equation (\ref{eq:total_energy}), as does the flux due to
+thermal heat conduction ($\fcond$).
 
-The total fluid energy density $E$ is given by:
+The total fluid energy density $E$ is given by
 \begin{equation}
 E =  e + \frac{\rho v^2}{2}  + \frac{B^2}{2a}
-        \label{eq:total_energy_def}.
+        \label{eq:total_energy_def},
 \end{equation}
-where $e$ is the comoving thermal energy density, and the total
+where $e$ is the comoving thermal energy density, the total
 isotropic pressure $p^*$ is given by
@@
 \end{equation}
-where $p$ is the gas pressure.  The equations are closed with the
-equation of state and Poisson's equation (we work in the Newtonian
-limit):
+and $p$ is the thermal pressure.  The equations are closed by an
+equation of state and Poisson's equation (in the Newtonian limit) for
+the gravitational potential $\phi$:
 %
@@
 %
-  e &=& p / [(\gamma - 1) \rho ],
+  e &=& \frac{p}{(\gamma - 1) \rho},
   \label{eq:eq_of_state} \\
 %
-  \nabla^2 \phi &=& 4 \pi G \rho_{\rm total}
+  \nabla^2 \phi &=& 4 \pi G \left( \rho_{\rm total} - \rho_0 \right)
 %                   + 3 p_{total}/c^2) - \Lambda. 
@@
 The equation of state is shown here for an ideal gas with a ratio of
-specific heats $\gamma$.  Equation (\ref{eq:potential}) includes
-contributions to the gravitational potential $\phi$ from the total
-mass density $\rho_{\rm total}$, which is defined as $\rho_{\rm gas} +
-\rho_{\rm dm} + \rho_{\rm stars} - \rho_{\rm average}$.  Although we
-write the equations including both MHD and comoving coordinates, the
-code is frequently used both in the pure hydrodynamic limit ($\vecB =
-0$) and without cosmological expansion ($a = 1$, $\dot{a} = 0$).
+specific heats $\gamma$.  The gravitational potential $\phi$ is
+sourced by the total mass density contrast, where $\rho_{\rm total} =
+\rho_{\rm gas} + \rho_{\rm dm} + \rho_{\rm stars}$ and $\rho_0$ is the
+mean density.  Although we write the equations including both magnetic
+field terms and comoving coordinates, the code is frequently used both
+in the purely hydrodynamic limit ($\vecB = 0$, referred to as HD
+below) and without cosmological expansion ($a = 1$, $\dot{a} = 0$).
 
@@
 \begin{eqnarray}
-\vecx & = & \vecr/a, \\
-\vecv & = & a \hspace{0.5mm} d\vecx/dt = 
+\vecx & = & \vecx' / a, \\
+\vecv & = & a \hspace{0.5mm} \dot{\vecx} = 
               \vecvp - \dot{a}\vecx, \\
@@
 E     & = & a^3 \left(\Ep - 
-            \dot{a} \vecx \cdot \vecvp - 
+            \dot{a} \, \vecx \cdot \vecvp - 
             \frac{1}{2} \dot{a}^2 \vecx^2\right), \\
-\phi  & = & \phip + \frac{1}{2} a \ddot{a} \vecx^2. \\
+\phi  & = & \phip + \frac{1}{2} \, a \ddot{a} \, \vecx^2, \\
 \vecB & = & a^2 \vecB^\prime,
 \end{eqnarray}
-where primes indicate the usual quantities defined in a fixed frame
-frame, $\vecr$ is the proper distance and $\vecx$ is the comoving
-distance.
-
+where primes indicate quantities in proper coordinates (fixed frame).
 %These definitions result in the following comoving equivalents to eqs. 
@@
 %\frac{\partial \rhop}{\partial t} 
-%          + \frac{1}{a} \vecvp \cdot \nabla \rhop 
-%    & = & - \frac{1}{a} \rhop \nabla \cdot \vecvp, 
+%          + \frac{1}{a} \vecvp \cdot \grad \rhop 
+%    & = & - \frac{1}{a} \rhop \div \vecvp, 
 %        \label{eq:mass_prime} \\
@@
 %\frac{\partial \vecvp}{\partial t}  
-%          + \frac{1}{a} ( \vecvp \cdot \nabla ) \vecvp 
+%          + \frac{1}{a} ( \vecvp \cdot \grad ) \vecvp 
 %    & = & - \frac{\dot{a}}{a} \vecvp 
-%          - \frac{1}{a \rhop} \nabla \pp 
-%          - \frac{1}{a} \nabla \phip, 
+%          - \frac{1}{a \rhop} \grad \pp 
+%          - \frac{1}{a} \grad \phip, 
 %        \label{eq:momentum_prime} \\
@@
 %\frac{\partial \Ep}{\partial t} 
-%         + \frac{1}{a} \vecvp \cdot \nabla \Ep 
+%         + \frac{1}{a} \vecvp \cdot \grad \Ep 
 %   & = & - \frac{\dot{a}}{a} (3 \frac{\pp}{\rhop} + {\vecvp \cdot \vecvp} )
-%         - \frac{1}{a \rhop } \nabla \cdot (p \vecvp) 
+%         - \frac{1}{a \rhop } \div (p \vecvp) 
 %%\nonumber   & &   %uncomment for 2-column
-%         - \frac{1}{a} \vecvp \cdot \nabla \phip 
+%         - \frac{1}{a} \vecvp \cdot \grad \phip 
 %                + \Gamma - \Lambda.
@@
 a smooth, homogeneous background, where $z$, the redshift, is a
-function only of $t$.  All derivatives are determined with respect to
-the comoving position $\vecx$, which is defined simply to remove the
-universal expansion from the coordinate system.
-
-The evolution of $a(t)$ is governed by the formula for the expansion
-of an isotropic, homogeneous universe:
+function only of $t$.  All spatial derivatives are determined with
+respect to the comoving position $\vecx$, which removes the universal
+expansion from the coordinate system. The evolution of $a(t)$ is
+governed by the second Friedmann equation for the expansion of a
+spatially homogeneous and isotropic universe
 %
@@
       + 3p_0/c^2) 
-      + \Lambda_c /3 .
+      + \Lambda_c c^2 / 3 .
       \label{eq:expansion} 
@@
 %
-Here, $\rho_0$ is the mean comoving mass density (including both
+Here $\rho_0$ is the mean comoving mass density (including both
 baryonic and dark matter), $p_0$ is the comoving background pressure
@@
 
-It is clear that the evolution equations are equivalent to the fixed
-coordinate version in the limit in which $a = 1$ and $\dot{a} = 0$.
-We include the cosmological terms with the understanding that \enzo\
-is not restricted to cosmological applications -- while historically
-the code was written with cosmological situations in mind (e.g.,
-galaxy clusters), in more recent years it has been used to simulate a
-much broader range of astrophysical environments.
+The comoving evolution equations are equivalent to the fixed
+coordinate version in the non-cosmological limit ($a = 1$, $\dot{a} =
+0$).  We include the cosmological terms with the understanding that
+\enzo\ is not restricted to cosmological applications -- while
+historically the code was written with cosmological situations in mind
+(e.g. galaxy clusters), in more recent years it has been used to
+simulate a much broader range of astrophysical environments.
 
@@
 
-Any collisionless component (e.g., dark matter, stars) is modeled by
-particles which are governed by Newton's equations in comoving
-coordinates:
+Any collisionless components (e.g. dark matter and stars) are modeled
+by N-body particles, whose dynamics are governed by Newton's equations
+in comoving coordinates:
 %
@@
     & = & - \frac{\dot{a}}{a} \vecv
-          - \frac{1}{a} \nabla \phi, 
+          - \frac{1}{a} \grad \phi, 
           \label{eq:dm_velocity} 
@@
 %
-The particles also make a contribution to the gravitational potential
-through their contribution to Poisson's equation (Equation
-\ref{eq:potential}).
+The particles contribute to the gravitational potential through
+Poisson's equation (Equation \ref{eq:potential}).
 
@@
 
-In addition, we can solve the mass conservation equations for a set of
-chemical species and reactions.  For any species $i$ with comoving
+In addition, \enzo\ can solve the mass conservation equations for a set of
+chemical species and their reactions.  For any species $i$ with comoving
 number density $n_i$, these equations have the form:
@@
   \frac{\partial n_i}{\partial t} 
-  + \frac{1}{a} \nabla \cdot (n_i \vecv) = 
-  k_{ij}(T) n_i n_j + \Gamma^{\rm ph}_j n_j 
+  + \frac{1}{a} \div (n_i \vecv) = 
+  \sum_j k_{ij}(T) \, n_i n_j + \sum_j \Gamma^{\rm ph}_j n_j 
   \label{eq:species_evolution}
@@
 are destruction/creation rates due to photoionizations and/or
-photodissociations.  Currently the species we can follow include H,
-H$^+$, He, He$^+$, He$^{++}$, and e$^-$, as well as additional options
-that add H$^-$, H$_2$, H$_2^+$, and HD, D, and D$^+$.  We also can
-track the advection of a comoving metal density that may contribute to
-the radiative cooling.
+photodissociations.  Currently the species \enzo\ can follow include
+H, H$^+$, He, He$^+$, He$^{++}$, and e$^-$, and through additional
+options also H$^-$, H$_2$, H$_2^+$, and HD, D, and D$^+$. Lastly,
+\enzo\ can also track the advection of a comoving metal density field,
+which can contribute to the radiative cooling and star formation
+processes.
 
@@
 %\begin{equation}
-%\frac{\partial \rho_b}{\partial t} + \frac{1}{a} \vecv_b \cdot \nabla \rho_b =  \label{enzoconserve}
-% - \frac{1}{a} \rho_b \nabla \cdot \vecv_b 
+%\frac{\partial \rho_b}{\partial t} + \frac{1}{a} \vecv_b \cdot \grad \rho_b =  \label{enzoconserve}
+% - \frac{1}{a} \rho_b \div \vecv_b 
 %\end{equation}
@@
 %\begin{equation}
-%\frac{\partial \vecv_b}{\partial t} + \frac{1}{a} (\vecv_b \cdot \nabla) \vecv_b = \label{enzomomentum} 
-%-\frac{\dot{a}}{a} \vecv_b - \frac{1}{a \rho_b} \nabla p - \frac{1}{a} \nabla \phi
+%\frac{\partial \vecv_b}{\partial t} + \frac{1}{a} (\vecv_b \cdot \grad) \vecv_b = \label{enzomomentum} 
+%-\frac{\dot{a}}{a} \vecv_b - \frac{1}{a \rho_b} \grad p - \frac{1}{a} \grad \phi
 %\end{equation}
@@
 %\begin{eqnarray}
-%\frac{\partial E}{\partial t} + \frac{1}{a} \vecv_b \cdot \nabla E =  
+%\frac{\partial E}{\partial t} + \frac{1}{a} \vecv_b \cdot \grad E =  
 %- \frac{\dot{a}}{a} \left( 3 \frac{p}{\rho_b} + \vecv_b^2 \right)  \nonumber \\
-%- \frac{1}{a \rho_b} \nabla \cdot (p \vecv_b) \nonumber \\
-%- \frac{1}{a} \vecv_b \cdot \nabla \phi + \Gamma - \Lambda
+%- \frac{1}{a \rho_b} \div (p \vecv_b) \nonumber \\
+%- \frac{1}{a} \vecv_b \cdot \grad \phi + \Gamma - \Lambda
 %\label{enzoenergy}
@@
 %\begin{equation}
-%\frac{d \vecv_{part}}{dt} = -\frac{\dot{a}}{a} \vecv_{part} - \frac{1}{a} \nabla \phi
+%\frac{d \vecv_{part}}{dt} = -\frac{\dot{a}}{a} \vecv_{part} - \frac{1}{a} \grad \phi
 %\label{enzopartmom}
@@
 equation along rays or by solving a set of moment equations derived
-from the radiative transfer equation.
-
-The radiative transfer equation in comoving coordinates
-\citep[e.g.,][]{Gnedin97} reads
+from the radiative transfer equation. In comoving coordinates
+\citep[e.g.,][]{Gnedin97} the radiative transfer equation reads
 %
@@
   \frac{1}{c} \; \frac{\partial I_\nu}{\partial t} + 
-  \frac{\hat{n} \cdot \nabla I_\nu}{\bar{a}} -
+  \frac{a_{\rm em}}{a} \, \hat{n} \cdot \grad I_\nu -
   \frac{H}{c} \; \left( \nu \frac{\partial I_\nu}{\partial \nu} -
@@
 %
-Here $I_\nu \equiv I(\nu, \mathbf{x}, \Omega, t)$ is the radiation
-specific intensity in units of energy per time $t$ per solid angle per
-unit area per frequency $\nu$.  $\bar{a} = a/a_{em}$ is the ratio of
-scale factors at the current time and time of emission.  The second
-term represents the propagation of radiation, where the factor
-$1/\bar{a}$ accounts for cosmic expansion.  The third term describes
-both the cosmological redshift and dilution of radiation.  On the
-right hand side, the first term considers the absorption coefficient
-$\kappa_\nu \equiv \kappa_\nu(\mathbf{x},\nu,t)$, and the second term
-$j_\nu \equiv j_\nu(\mathbf{x},\nu,t)$ is the emission coefficient
-that includes any sources of radiation (whether point or diffuse).
+Here $I_\nu \equiv I(\nu, \mathbf{x}, \Omega, t)$ is specific
+intensity of the radiation, with dimensions of energy per time $t$ per
+solid angle $\Omega$ per unit area per frequency $\nu$.  The second
+term represents the propagation of radiation, where the factor $a_{\rm
+  em}/a$ accounts for cosmic expansion since the time of emission.
+The third term describes both the cosmological redshift and dilution
+of radiation.  On the right hand side, the first term captures
+absorption ($\kappa_\nu \equiv \kappa_\nu(\mathbf{x},\nu,t)$), and the
+second term emission ($j_\nu \equiv j_\nu(\mathbf{x},\nu,t)$ from any
+sources of radiation (whether point or diffuse).
 
@@
 
-As an alternative to the ray-casting strategy, we solve an equation
-derived from integrating the equation of radiative transfer over
-angles.  In particular, we use a flux-limited diffusion approximation
-for radiative transfer, with couplings to both the gas energy and
-chemical number densities. The equations solved are:
+As an alternative to the ray-casting strategy, \enzo\ can also solve
+the angle-averaged radiative transfer equation, through a
+flux-limited diffusion approximation with couplings to both the gas
+energy and chemical number densities. The equations solved are
 \begin{eqnarray}
   \label{eq:fld_radiation}
-  \frac{\partial E_r}{\partial t} + \frac1a \nabla\cdot\left(E_r \vecv \right) &=& 
-  \nabla\cdot\left(D\nabla E_r\right) -
+  \frac{\partial E_r}{\partial t} + \frac1a \div\left(E_r \vecv \right) &=& 
+  \div\left(D\grad E_r\right) -
   \frac{\dot{a}}{a} E_r - c\kappa E_r + \eta, \\
@@
 %  \partial_t {\mathrm n}_i + 
-%    \frac1a \nabla\cdot\left({\mathrm n}_i \vecv_b\right) &=& 
+%    \frac1a \div\left({\mathrm n}_i \vecv_b\right) &=& 
 %    \alpha_{i,j} {\mathrm n}_e {\mathrm n}_j - {\mathrm n}_i
@@
 %number density.  
-Here, we define the grey radiation energy density through first
-assuming a fixed frequency spectrum, i.e.  $E_{\nu}(\nu,\vecx,t) =
-\tilde{E}_r(\vecx,t) \chi(\nu)$, and then defining the integrated
-quantity
+Here, we define $E_r$ through first assuming a fixed frequency
+spectrum, i.e.  $E_{\nu}(\nu,\vecx,t) = \tilde{E}_r(\vecx,t)
+\chi(\nu)$, and then defining the integrated quantity
 \begin{equation}
@@
 \end{equation}
-In addition, $D$ in Equation (\ref{eq:fld_radiation}) is the Larsen
-square-root flux-limiter \citep[see][]{Morel2000}; $\kappa$ is the
-total opacity; $\eta$ is the field of radiation sources; $\Gamma$
-provides the radiation induced photo-heating, and $\Lambda$ the
-chemical cooling.
+$D$ in Equation (\ref{eq:fld_radiation}) is the Larsen
+square-root flux-limiter \citep[see][]{Morel2000}, $\kappa$ the
+total opacity, $\eta$ the field of radiation sources, $\Gamma$
+the radiation induced photo-heating rate, and $\Lambda$ the
+chemical cooling rate.
 %, $\alpha_{i,j}$ are the reaction rate coefficients defining
@@
 
-Enzo implements the equations of isotropic heat conduction in a manner
-similar to that of \citet{2007ApJ...664..135P}, where we model the
-flux of isotropic heat transport,
+Finally, \enzo\ implements the equations of isotropic heat conduction
+in a manner similar to that of \citet{2007ApJ...664..135P}, where the isotropic heat flux is given by
 %
 \begin{equation}
-\fcond = -\kappa_{sp} \nabla T.
+\fcond = -\kappa_{\rm cond} \grad T.
 \end{equation}
 %
-Here $\kappa_{sp}$ is the Spitzer conduction coefficient
-\citep{1962pfig.book.....S} and T is the baryon temperature (with
-fluids explicitly assumed to be single-temperature).  Saturation of
-the heat flux in high-temperature, low-density regimes (such as the
-intracluster medium in galaxy clusters) is taken into account.
+Here $\kappa_{\rm cond} = f_{\rm sp} \, \kappa_{\rm sp}$ is the heat
+conduction coefficient, given as a fraction $f_{\rm sp}$ of the
+Spitzer conductivity $\kappa_{\rm sp}$ \citep{1962pfig.book.....S},
+and T is the gas temperature (with fluids explicitly assumed to be
+single-temperature).  Saturation of the heat flux in high temperature,
+low density regimes (such as the intracluster medium in galaxy
+clusters) is taken into account.
+
 Thermal conduction in a plasma can be strongly affected by the
 presence of magnetic field lines, which may strongly suppress heat
-flow perpendicular to the magnetic field.  In that case, we allow for
-heat transport only along the magnetic field lines in simulations that
-include this physics, as follows:
+flow perpendicular to the magnetic field.  To include this physics, it
+is possible to allow heat transport only along (not across) magnetic
+field lines, as follows:
 
 \begin{equation}
-\fcond= -\kappa_{sp} \hat{\textbf{b}} (\hat{\textbf{b}} \cdot \nabla T)
+\fcond= -\kappa_{\rm cond} \, {\hat \textbf{b}} (\hat{\textbf{b}} \cdot \grad T),
 \end{equation}
-
-Where $\hat{\textbf{b}}$ is the unit vector denoting the direction of
-the magnetic field.  In both the isotropic and anisotropic limits, we
-assume that conductivity can be suppressed relative to the ideal
-Spitzer conductivity, and express this as a multiplicative constant.
+where $\hat{\textbf{b}}$ is the unit vector denoting the direction of
+the magnetic field.
 
@@
 
-In this section, we briefly describe the numerical methods which are
+In this section we briefly describe the numerical methods which are
 used to solve the equations outlined in Section~\ref{sec.equations}.
@@
 (AMR) capability, which allows it to reach extremely large dynamical
-ranges with limited computational resources, opening doors previously
+ranges with limited computational resources, opening doors otherwise
 closed by finite memory and computational time. Unlike moving mesh
@@
 rectangular grid patch (referred to as a ``grid'') covers some region
-of space in its \emph{parent grid} which requires higher resolution,
+of space in its \emph{parent grid} that requires higher resolution,
 and can itself become the parent grid to an even more highly resolved
@@
 The grid hierarchy begins with the root grid, which covers the entire
-domain of interest with a coarse, uniform, Cartesian grid. Then, as
-the solution evolves and interesting regions form, finer meshes are
+domain of interest with a coarse uniform Cartesian grid. Then, as
+the solution evolves and interesting regions start to develop, finer meshes are
 placed within these coarse regions.
@@
 \caption{\emph{Left:} Example of the timesteps on a 2-level AMR
-  hierarchy.  Enzo does not restrict the timesteps on the finer levels
+  hierarchy.  \enzo\ does not restrict the timesteps on the finer levels
   to be a factor of $1/2^n$ smaller.  \emph{Right:} The order in which
@@
 Since interesting regions on the grid may move, the hierarchy must
-adapt itself.  We do this whenever a level has caught up to the
-coarser parent level and consists of entirely rebuilding the grids on
-that level and at finer resolutions.  This is done by applying the
-grid refinement criteria to the grids on that level, flagging zones
-that require extra grids.  These criteria depend on the physical
-problem being simulated -- see Section~\ref{sec:refinement_criteria}
-for more details.  Once a grid has a set of flagged cells, we run a
+adapt itself. This happens whenever a level has caught up to its
+parent level, by entirely rebuilding the grids on that level and at
+finer resolutions. Rebuilding is achieved by applying the grid
+refinement criteria to the grids on that level, flagging zones that
+require extra grids.  These criteria depend on the physical problem
+being simulated -- see Section~\ref{sec:refinement_criteria} for more
+details.  Once a grid has a set of flagged cells, we apply a
 machine-vision based algorithm \citep{Berger91} to find edges and
@@
 next coarser grid is interpolated in order to initialize the values on
-the new grids.  Finally, any overlap between these new subgrids and
-the old ones is identified, and the solution within the regions of
-overlap is copied to the new subgrids.  The entire procedure just
-outlined is then repeated on the new grids and in this way the entire
-hierarchy (from the original level examined and continuing on to finer
-levels) is rebuilt.
+the new grids.  Finally, any overlap between new subgrids and old ones
+is identified, and the prior solution within the regions of overlap is
+copied to the new subgrids. This procedure is then repeated on the new
+grids and in this way the entire hierarchy (from the original level
+examined and continuing on to all finer levels) is rebuilt.
 
-\subsubsection{Godunov PPM method (HD only)}
+\subsubsection{(Magneto)-hydrodynamic solver methods}
 
@@
 
+\paragraph{Godunov PPM method (HD only)}
+
 We begin with the direct-Eulerian PPM implementation.  This is an
 explicit, higher-order-accurate version of Godunov's method for ideal
-gas dynamics with third order-accurate piecewise parabolic monotonic
+gas dynamics with third order accurate piecewise parabolic monotonic
 interpolation and a nonlinear Riemann solver for shock capturing.  It
@@
 (i) Construct monotonic parabolic interpolation of cell average data,
-for each fluid quantity.
+for each fluid quantity;
 % \item 
 (ii) Compute interface states by averaging the parabola over the
-domain of dependence for each interface.
+domain of dependence for each interface;
 % \item 
-(iii) Use interface data to solve the Riemann problem.
+(iii) Use interface data to solve the Riemann problem;
  %\item 
@@
 
-This PPM implementation does an excellent job of capturing strong
-shocks in a few cells.  Multidimensional schemes are built up by
+The PPM implementation does an excellent job of capturing strong
+shocks across a few cells.  Multidimensional schemes are built up by
 directional splitting and produce a method that is formally second
@@
 hypersonic flows when the thermal energy $e$ is extremely small
-compared to the total energy $E$.  This is a problem because in the
-total energy method, the temperature is computed by subtracting one
-large number from another (i.e. the kinetic energy from the total
-energy), a situation which can generate large numerical inaccuracies.
-We address this situation by also solving the thermal energy equation
-and using $e$ from this equation when we expect the error to be large.
+compared to the total energy $E$.  This situation presents a problem
+because in the total energy method the temperature is computed by
+subtracting one large number from another (i.e. the kinetic energy
+from the total energy), which tends to generate large numerical
+inaccuracies. We overcome this difficulty by additionally solving a
+thermal energy equation and using $e$ from this equation when we
+expect the error to be large.
 
@@
 
-\subsubsection{Godunov MUSCL (HD) with Dedner divergence cleaning (MHD)}
+%%% If you prefer the following \paragraph{} sections to be numbered,
+%%% then you can add \setcounter{secnumdepth}{5} to the preamble...
+
+\paragraph{Godunov MUSCL (HD) with Dedner divergence cleaning (MHD)}
 This solver was developed to attack problems in magnetic field
@@
 states. Consequently, these suite of solvers can be used for hydro and
-magneto-hydrodynamic simulations. It is also this class of solvers as
-well as a version of the PPM hydro solver which have been ported to
-nVidia's CUDA framework allowing \enzo\ to take advantage of modern
+magneto-hydrodynamic simulations. This class of solvers, as
+well as a version of the PPM hydro solver, has been ported to
+nVidia's CUDA framework, allowing \enzo\ to take advantage of modern
 graphics hardware \citep{Wang:2010}.
 
-\subsubsection{MHD with Constrained Transport (MHD)}
+\paragraph{MHD with Constrained Transport (MHD)}
 This MHD method is second order in time and space, and preserves the
-divergence constraint, $\nabla \cdot \vecB = 0$, to machine precision
+divergence constraint, $\div \vecB = 0$, to machine precision
 through the Constrained Transport (CT) method.  CT, originally
 described by \citet{Evans88}, updates the magnetic field with the curl
-of an electric field, suitably formulated to preserve $\nabla \cdot
+of an electric field, suitably formulated to preserve $\div
 \vecB$.  We employ the CT methods described by \citet{Balsara99} and
-\citet{Gardiner05}.  Further, we employ the second order hyperbolic
+\citet{Gardiner05} with the second order hyperbolic
 solver of \citet{Li08a} and the divergence free AMR scheme of
@@
 
-\subsubsection{Second order finite difference method (HD only)}
+\paragraph{Second order finite difference method (HD only)}
 
-The last hydrodynamics method we briefly describe is the
-finite-difference algorithm used in the \zeus\ code, as given in more
-detail in \citet{Stone92a}.  For this reason, we call this the \zeus\
-method, although the code is entirely independent of the \zeus\ code,
-and only the hydrodynamical algorithm of \zeus\ is implemented in
-\enzo\ -- the MHD and RHD schemes are not.
-
-The \zeus\ method uses a staggered mesh such that the velocities are
-face-centered, while the density and internal energy quantities are
-cell-centered.  It splits the solution up into two steps, the
-so-called source step, in which the momentum and energy values are
-updated to reflect the pressure and gravity forces, as well as the
-effect of an artificial viscosity required for stability.  The second
-step (known as the transport step) accounts for the advection of
-conserved quantities (mass, momentum and energy) across the grid.
+Lastly, we briefly describe the \zeus\ method, a finite-difference
+algorithm originally used in the \zeus\ code \citep{Stone92a}. Note
+that \enzo\ is entirely independent of the \zeus\ code, and only the
+hydrodynamical algorithm of \zeus\ has been implemented in \enzo\ --
+the MHD and RHD schemes have not. The \zeus\ method uses a staggered
+mesh such that the velocities are face-centered, while the density and
+internal energy quantities are cell-centered.  It splits the solution
+up into two steps. The first is the so-called source step, in which
+the momentum and energy values are updated to reflect the pressure and
+gravity forces, including an artificial viscosity required for
+stability. The second step, known as the transport step, accounts for
+the advection of conserved quantities (mass, momentum and energy)
+across the grid.
 
@@
 Fourier Technique \citep{Hockney88} to solve Poisson's equation on the
-coarsest (Level 0) grid on each time step.  The advantage of using
+coarsest (level 0) grid on each time step.  The advantage of using
 this method is that it is fast, accurate, and naturally allows both
@@
 which are very common in astrophysics and cosmology.  On subgrids, we
-interpolate the boundary conditions from that grid's parent (either
-the level 0 grid or some other subgrid).  The Poisson equation is then
+interpolate the boundary conditions from the parent grid (either the
+level 0 grid or some other subgrid). The Poisson equation is then
 solved on every timestep using a multigrid technique on one subgrid at
-a time.  In \enzo, self-gravity is optional.  Aside from calculating
-the gravitational potential from the baryon fields and particles
-existing in the simulation, there are also a number of options for
-static gravitational fields.
+a time. Aside from calculating the gravitational potential arising
+from the baryon fields and particles in the simulation, there are also
+a number of options for specifying static gravitational fields, and in
+fact including self-gravity is optional.
 
@@
 associated with the highest refined level available at its position in
-space and is moved as the hierarchy is rebuilt.  Thus, a particle has
-the same timestep and feels the same gravitational force as a grid at
-that refinement level.
+the domain and particles are moved between grids as the hierarchy is
+rebuilt. Thus, a particle has the same timestep and feels the same
+gravitational force as a grid at that refinement level.
 
-Although the particles are fixed in mass once initialized, we can
-create them with any initial set of masses and positions.  For
-example, in many cosmological simulations, a static subgrid is
-included from the beginning in order to improve the initial baryonic
-mass resolution.  On this subgrid, we also use smaller particles to
-improve the collisionless mass resolution.  One particle per initial
-grid point seems to provide approximately equal sampling between the
-dark matter and gas.
+Although the particles are fixed in mass once initialized (with the
+exception of star particles, which can lose mass in the feedback
+process), we can create them with any initial set of masses and
+positions.  For example, in many cosmological simulations static
+subgrids are included from the beginning in order to improve the
+initial spatial and baryonic mass resolution, and these subgrid are
+populated with lower mass particles in order to also improve the
+collisionless mass resolution. 
+%% One particle per initial grid point seems to provide approximately
+%% equal sampling between the dark matter and gas. \mqk{This
+%% contradicts the conclusions of O'Shea et al. 2005 that an initial
+%% grid twice as fine as the mean particle separation is necessary to
+%% properly capture the abundance of low mass halos.}
 
@@
 and e$^-$, and more complete models adding first species important for
-gas-phase molecular hydrogen formation: H$^-$, H$_2$ and H$_2^+$, and
-then HD formation: HD, D, D$^+$.  (We note that a reduced version
-of the chemistry, used only with the flux-limited diffusion radiation
-transport, includes only hydrogen species.)  The cooling and heating
-due to these species is also included (see the next section).  The
-solution of the rate equations is carried out using one Jacobi
-iteration of an implicit Euler time discretization to ensure
-stability; to ensure accuracy we sub-cycle the rate equations with a
-shorter timestep such that the electron and neutral fractions do not
-change by more than 10\%.
+gas-phase molecular hydrogen formation, H$^-$, H$_2$ and H$_2^+$, and
+then HD formation, HD, D, D$^+$. 
+%(We note that a reduced version
+%of the chemistry, used only with the flux-limited diffusion radiation
+%transport, includes only hydrogen species.)
+The cooling and heating due to these species is included (see the next
+section). The solution of the rate equations is carried out using one
+Jacobi iteration of an implicit Euler time discretization to ensure
+stability. To ensure accuracy the rate equations are sub-cycled
+within one hydrodynamic timestep, subject to the constraint that the
+electron and neutral fractions do not change by more than 10\% in one
+sub-cycle.
 
@@
 
-Enzo may operate in a number of modes with regard to radiative cooling
-and heating.  In the simplest mode, with the multi-species flag turned
-off (i.e. set to zero), the cooling rate is computed from a simple
-temperature-dependent cooling rate, taken from \citet{SW87}.  If
-chemistry is turned on, then the code can include cooling from all
-species of hydrogen and helium (including H$_2$) -- this primordial
-cooling is treated in the same Jacobi iteration as the chemistry.
-Finally, we have also recently added metal cooling based on a set of
-multi-dimensional lookup tables computed with the \textsc{Cloudy} code
-\citep{1998PASP..110..761F} as described in
-\citet{2008MNRAS.385.1443S} and \citet{2011ApJ...731....6S}.  Finally,
-we note that the cooling and heating is (almost always) treated in the
-optically-thin limit, although the code can also follow radiative
-transfer in a limited set of energy bins.
+\enzo\ can operate in a number of different modes with regard to
+radiative cooling and heating. In the simplest mode, with the
+multi-species flag turned off, the cooling rate is computed from a
+simple temperature-dependent cooling function, taken from
+\citet{SW87}.  If chemistry is turned on, then the code can include
+cooling from all species of hydrogen and helium (including H$_2$) --
+this primordial cooling is treated in the same Jacobi iteration as the
+chemistry.  It is also possible to include metal cooling based on a
+set of multi-dimensional lookup tables computed with the
+\textsc{Cloudy} code \citep{1998PASP..110..761F} as described in
+\citet{2008MNRAS.385.1443S} and \citet{2011ApJ...731....6S}. Note that
+the cooling and heating is most commonly treated in the optically-thin
+limit, but the code can also follow radiative transfer in a limited
+set of energy bins.
 
@@
 The chemical networks and heating rates described in the previous
-section can be affected by external radiation fields, and the code
+sections can be affected by external radiation fields, and the code
 includes a number of pre-set radiation backgrounds that are uniform in
@@
 
-The code includes a photon-conserving radiative transfer algorithm
-that is based on an adaptive ray-tracing method that utilizes the
-HEALPix pixelization of a sphere \citep{AbelWandelt02}.  Photons are
-integrated outwards from sources using an adaptive time-stepping
-scheme that preserves accuracy in ionization fronts even in the
-optically-thin limit.  This has been integrated with the chemistry and
-cooling network to provide ionization and heating rates on a
-cell-by-cell basis.  The method is described in more detail in our
-main methodology section but is fully detailed and tested in
+\enzo\ includes a photon-conserving radiative transfer algorithm that
+is based on an adaptive ray-tracing method utilizing the HEALPix
+pixelization of a sphere \citep{AbelWandelt02}. Photons are integrated
+outwards from sources using an adaptive time-stepping scheme that
+preserves accuracy in ionization fronts even in the optically-thin
+limit. This has been coupled to the chemistry and cooling network to
+provide ionization and heating rates on a cell-by-cell basis. The
+method is described in detail (including numerous code tests) in
 \citet{Wise11_Moray}.
@@
 A second option for radiative transfer is a moment-based method that
-adds an additional field tracking the radiation energy density.
-This field is evolved using the flux-limited diffusion method, which
+adds an additional field tracking the radiation energy density.  This
+field is evolved using the flux-limited diffusion method, which
 transitions smoothly between streaming (optically thin) and opaque
-limits and is coupled to the hydrogen ionization network.  The
+limits and is coupled to a hydrogen-only ionization network.  The
 resulting set of linear equations is solved using the parallel HYPRE
-framework.  Full details for the \enzo\ implementation can be found in
-\citet{ReynoldsHayesPaschosNorman2009}.
+framework.  Full details on the \enzo\ implementation of this method
+can be found in \citet{ReynoldsHayesPaschosNorman2009}.
 
@@
 a sub-cycled, operator-split method.  The heat fluxes are computed
-with simple second-order accurate finite differences and stability is
-ensured by restricting the time step, and using flux-limiters where
+with simple second-order accurate finite differences, and stability is
+ensured by restricting the time step and using flux-limiters where
 appropriate.
@@
 
-All grids on a given level are advanced with the same timestep.  To
-determine this timestep, we calculate, for each cell and for each
-physical process (except for the chemistry step, which is sub-cycled),
-the minimum $\Delta t$ allowed for stability.  Then the minimum of all
-timesteps is taken and the level is advanced with this step.
+All grids on a given level are advanced with the same timestep,
+determined by calculating, for each cell and for each physical process
+(except for the chemistry step, which is sub-cycled) separately, the
+minimum $\Delta t$ allowed for stability. The level is advanced with a
+timestep equal to the minimum over all of these $\Delta t$.
 

### 2013-06-05 parallel.tex
Edits to the overview section. Trying to unify notations, so several other sections are also affected.

@@
 either a core or a node, depending on details of the system.  Note
-that the current version of Enzo is not threaded, although a hybrid
+that the current version of \enzo\ is not threaded, although a hybrid
 MPI + OpenMP version is currently under development.}.  
@@
 
-Because of the wide variety of simulations, methods, and uses of Enzo,
+Because of the wide variety of simulations, methods, and uses of \enzo,
 it is difficult to state in general terms which routines within the
@@
 cosmological calculation.  The root grid scaling is not representative
-of the true strong scaling of Enzo because the root grid tiles are not
+of the true strong scaling of \enzo\ because the root grid tiles are not
 repartitioned when a simulation is restarted with additional cores.
@@
 span several orders of magnitude in both space and time.  In these
-situations, using Enzo in its adaptive mesh refinement mode is
+situations, using \enzo\ in its adaptive mesh refinement mode is
 beneficial.  Because the adaptive grid hierarchy is dynamic, grid
@@
 be easily implemented and is planned in the next major release of
-Enzo, where scaling results would be similar to the weak scaling shown
+\enzo, where scaling results would be similar to the weak scaling shown
 in \S\ref{sec:weak_scaling}.  The right panel in Figure
 \ref{fig:strong_scaling} shows the time spent in some representative
-major routines in Enzo.  The local physics routines, for example
+major routines in \enzo.  The local physics routines, for example
 \texttt{SolveHydroEquations}, exhibit perfect strong scaling because

### 2013-06-05 tests-orszagtang.tex
Updates to sections 11.2.5 through 11.2.10 (minor changes)

@@
 Figure~\ref{fig.orszag} shows the Orszag-Tang vortex problem, a
-classic MHD test problem described by \citep{Orszag79}.  The left panel shows
+classic MHD test problem described by \citet{Orszag79}.  The left panel shows
 the result using \enzo's constrained transport MHD method, while the right panel shows the result using
@@
 effective resolution of different MHD schemes.  The test begins with uniform
-density, $\rho_0=25/36 \pi$ and pressure, $P_0=5/12 \pi$ (as with
+density, $\rho_0=25/(36\pi)$, and pressure, $P_0=5/(12\pi)$ (as with
 other tests in this section, in the absence of gravity or chemistry we use
@@
 single rotational mode in the velocity, and two in the magnetic field:
-${\bf v}_0 $ = (-sin(2$\pi$ y) $ \hat{x}$ , sin(2$\pi$ x) $\hat{y}$),
-${\bf B}_0$ = (-sin(2 $\pi$ y) $ \hat{x},$ sin( 4 $\pi$ x )$\hat{y}$).
+${\bf v}_0 = (-\sin(2\pi y) \hat{x},\, \sin(2\pi x) \hat{y})$,
+${\bf B}_0 = (-\sin(2\pi y) \hat{x},\, \sin(4\pi x) \hat{y})$.
 The simulation is evolved to $t=0.48$ (in arbitrary units).  One can see that the structures

### 2013-06-05 tests-pointsourcegravity.tex
Updates to sections 11.2.5 through 11.2.10 (minor changes)

@@
 with root grid dimensions $32^3$ and add a static, nested grid 
-refinemed by a factor of two that covers the central $1/8^3$ of the
+refined by a factor of two that covers the central $1/8^3$ of the
 domain (i.e. so the refined grid is only $8^3$).  We then place 5000

### 2013-06-05 tests-zeldovichpancake.tex
Updates to sections 11.2.5 through 11.2.10 (minor changes)

@@
 perturbation is given by
-
 \begin{equation}
-\rho(x_l) = \rho_0 \big[ 1 - \frac{1+z_c}{1+z} \cos(k x_l) \big]
+\rho(x_l) = \rho_0 \left[ 1 - \frac{1+z_c}{1+z} \cos(k x_l) \right]
 \end{equation}
-
 with the internal energy of the gas set so that the entropy 
 is constant throughout.  The velocity perturbation is given by
-
 \begin{equation}
@@
 coordinates, x$_e$, that are more useful to \enzo:
-
 \begin{equation}
@@
 equivalent maximum resolution of 1024 grid cells).  The simulation is
-evolved to z$ = 0$ using the PPM hydro method, with the final output
-of the calculation shown in Figure~\ref{fig.pancake}.
+evolved to z$ = 0$ using the PPM hydro method.
 
-Figure~\ref{fig.pancake} shows the key features of this test problem.
+The final output of the calculation, with the key features of this
+test problem, is shown in Figure~\ref{fig.pancake}.
 The strong shocks and large density gradients are well-resolved, with

### 2013-06-06 app-datastruct.tex
A few more edites to section 2. Also unified "time step" -> "timestep" throughout the whole manuscript.

@@
 that coarser levels may be further ahead in their integration, due to
-their larger time step and the W-cycle)  Beginning with this coarsest level, each
+their larger timestep and the W-cycle)  Beginning with this coarsest level, each
 grid on each level is examined for refinement.  Cells are first flagged

### 2013-06-06 numerical-overview.tex
A few more edites to section 2. Also unified "time step" -> "timestep" throughout the whole manuscript.

@@
 Algorithmically, the star formation events in Kravtsov's code
-are assumed to occur once every global time step 
+are assumed to occur once every global timestep 
 $\Delta t_0 \leq 10^7$.  In cells where star formation is
@@
 of this algorithm is similar, except that instead of 
-forming stars only at the root grid time step, we allow 
-stars to form at the time step of the highest level of
+forming stars only at the root grid timestep, we allow 
+stars to form at the timestep of the highest level of
 resolution at any particular point in space.  As can be seen
@@
 independent CFD problem, with Dirichlet boundary conditions stored in
-the ghost zones.  At the end of each time step, cells on a given level
+the ghost zones.  At the end of each timestep, cells on a given level
 fill ghost zones by interpolating from the parent and copying from
@@
 
-\subsection{Time Stepping}\label{sec.timestepping}
+\subsection{Timestepping}\label{sec.timestepping}
 \dcc{Made this its own section (not really part of the AMR, it's more
@@
   noted that the timestep is computed before the acceleration field is computed,
-  and the acceleration field is deleted at the end of each time step.
+  and the acceleration field is deleted at the end of each timestep.
   So this condition isn't ever actually enforced.  We should either

### 2013-06-06 numerical-radiation-fld.tex
A few more edites to section 2. Also unified "time step" -> "timestep" throughout the whole manuscript.

@@
 ionization and heating, the remainder of these equations is solved
-using an operator-split algorithm.  Within a given time step to evolve
+using an operator-split algorithm.  Within a given timestep to evolve
 $(E_r^n, e_c^n, {\mathrm n}_i^n) \to (E_r^{n+1}, e_c^{n+1}, {\mathrm

### 2013-06-06 overview.tex
A few more edites to section 2. Also unified "time step" -> "timestep" throughout the whole manuscript.

@@
 rectangular grid patch (referred to as a ``grid'') covers some region
-of space in its \emph{parent grid} that requires higher resolution,
+of space in its \emph{parent grid} requiring higher resolution,
 and can itself become the parent grid to an even more highly resolved
@@
 manner of a W-cycle in a multigrid solver.  First, we determine the
-maximum time step allowed for the coarsest grid based on a variety of
+maximum timestep allowed for the coarsest grid based on a variety of
 accuracy and stability criteria and advance the grid by that time
 interval, $\Delta t_0$.  We then move down to the next level and
-advance all the grids on that level by a timestep $\Delta t_1$
-($\Delta t_1 \leq \Delta t_0$), which is the minimum of all the
-allowed timesteps for those grids.  If there are more levels, we
-repeat this procedure until the bottom level of the hierarchy has been
-reached.  Once there, we continue advancing the grids on the lowest
-level until they have ``caught up'' to the coarser parent grids
-(i.e. $\sum \Delta t_l = \Delta t_{l-1}$).  This procedure repeats
-itself until all grids have been advanced by a total time of $\Delta
-t_0$ (see Figure~\ref{fig:wcycle} for a schematic).
+advance all the grids on that level by a timestep $\Delta t_1 \leq
+\Delta t_0$, which is the minimum of all the allowed timesteps for
+those grids.  If there are more levels, we repeat this procedure until
+the bottom level of the hierarchy has been reached.  Once there, we
+continue advancing the grids on the lowest level until they have
+``caught up'' to the coarser parent grids (i.e. $\sum \Delta t_l =
+\Delta t_{l-1}$).  This procedure repeats itself until all grids have
+been advanced by a total time of $\Delta t_0$ (see
+Figure~\ref{fig:wcycle} for a schematic).
 
@@
 require extra grids.  These criteria depend on the physical problem
-being simulated -- see Section~\ref{sec:refinement_criteria} for more
+being simulated; see Section~\ref{sec:refinement_criteria} for more
 details.  Once a grid has a set of flagged cells, we apply a
@@
 approximate Riemann solvers and non-ideal equations of
-states. Consequently, these suite of solvers can be used for hydro and
+states. Consequently, this suite of solvers can be used for hydro and
 magneto-hydrodynamic simulations. This class of solvers, as
@@
 that \enzo\ is entirely independent of the \zeus\ code, and only the
-hydrodynamical algorithm of \zeus\ has been implemented in \enzo\ --
-the MHD and RHD schemes have not. The \zeus\ method uses a staggered
-mesh such that the velocities are face-centered, while the density and
-internal energy quantities are cell-centered.  It splits the solution
+hydrodynamical algorithm of \zeus\ has been implemented in \enzo; the
+MHD and radiation-hydrodynamics schemes have not. The \zeus\ method
+uses a staggered mesh such that velocities are face-centered, while
+density and internal energy are cell-centered.  It splits the solution
 up into two steps. The first is the so-called source step, in which
@@
 stability. The second step, known as the transport step, accounts for
-the advection of conserved quantities (mass, momentum and energy)
+the advection of conserved quantities (mass, momentum, and energy)
 across the grid.
@@
 Fourier Technique \citep{Hockney88} to solve Poisson's equation on the
-coarsest (level 0) grid on each time step.  The advantage of using
-this method is that it is fast, accurate, and naturally allows both
-periodic and isolated boundary conditions for the gravity, choices
-which are very common in astrophysics and cosmology.  On subgrids, we
-interpolate the boundary conditions from the parent grid (either the
-level 0 grid or some other subgrid). The Poisson equation is then
-solved on every timestep using a multigrid technique on one subgrid at
-a time. Aside from calculating the gravitational potential arising
-from the baryon fields and particles in the simulation, there are also
-a number of options for specifying static gravitational fields, and in
+root grid on each timestep.  The advantage of using this method is
+that it is fast, accurate, and naturally allows both periodic and
+isolated boundary conditions for the gravity, choices which are very
+common in astrophysics and cosmology.  On subgrids, we interpolate the
+boundary conditions from the parent grid (either the root grid or some
+other subgrid). The Poisson equation is then solved on every timestep
+using a multigrid technique on one subgrid at a time. Aside from
+self-consistently calculating the gravitational potential arising from
+the baryon fields and particles in the simulation, there are also a
+number of options for specifying static gravitational fields, and in
 fact including self-gravity is optional.
@@
 
-Collisionless matter (e.g., dark matter, stars, etc.) is modeled with
+Collisionless matter (e.g. dark matter, stars, etc.) is modeled with
 particles that interact with the baryons only via gravity.  These
-particles are advanced in a single time step using a drift-kick-drift
+particles are advanced through a single timestep using a drift-kick-drift
 algorithm \citep{Hockney88} to provide second-order accuracy even in
-the presence of changing timesteps.  Since the particles follow the
+the presence of varying timesteps.  Since the particles follow the
 collapse of structure, they are not adaptively refined.  Nor are there
@@
 the domain and particles are moved between grids as the hierarchy is
-rebuilt. Thus, a particle has the same timestep and feels the same
+rebuilt. Thus a particle has the same timestep and feels the same
 gravitational force as a grid at that refinement level.
@@
 initial spatial and baryonic mass resolution, and these subgrid are
-populated with lower mass particles in order to also improve the
+populated with lower mass particles to correspondingly improve the
 collisionless mass resolution.
@@
 
-\enzo\ includes the capability of following up to 12 chemical species
-using a non-equilibrium solver.  The species can be turned on in sets,
-with the simplest model including just H, H$^+$, He, He$^+$, He$^{++}$,
-and e$^-$, and more complete models adding first species important for
-gas-phase molecular hydrogen formation, H$^-$, H$_2$ and H$_2^+$, and
-then HD formation, HD, D, D$^+$. 
+\enzo\ includes the capability of following up to 12 particle species
+using a non-equilibrium chemical network.  The species can be turned
+on in sets, with the simplest model including just atomic hydrogen and
+helium (H, H$^+$, He, He$^+$, He$^{++}$, e$^-$), and more complete
+models adding first species important for gas-phase molecular hydrogen
+formation (H$^-$, H$_2$ and H$_2^+$), and then HD formation (HD, D,
+D$^+$).
 %(We note that a reduced version
@@
 sections can be affected by external radiation fields, and the code
-includes a number of pre-set radiation backgrounds that are uniform in
-space but can vary in time.  These are generally based on the
-redshift-dependent rates given in \citet{1996ApJ...461...20H} and
-\citet{2012ApJ...746..125H}, but can also include a uniform H$_2$
-photo-dissociating background that is either constant in time or
-varying as in \citet{WiseAbel05}.
+includes a number of pre-calculated meta-galactic UV radiation
+backgrounds that are uniform in space but can vary in time.  These are
+generally based on the redshift-dependent rates given in
+\citet{1996ApJ...461...20H} and \citet{2012ApJ...746..125H}, but can
+also include a uniform H$_2$ photo-dissociating background that is
+either constant in time or varying as in \citet{WiseAbel05}.
 
@@
 pixelization of a sphere \citep{AbelWandelt02}. Photons are integrated
-outwards from sources using an adaptive time-stepping scheme that
+outwards from sources using an adaptive timestepping scheme that
 preserves accuracy in ionization fronts even in the optically-thin
@@
 with simple second-order accurate finite differences, and stability is
-ensured by restricting the time step and using flux-limiters where
+ensured by restricting the timestep and using flux-limiters where
 appropriate.
@@
 determined by calculating, for each cell and for each physical process
-(except for the chemistry step, which is sub-cycled) separately, the
-minimum $\Delta t$ allowed for stability. The level is advanced with a
-timestep equal to the minimum over all of these $\Delta t$.
+(except for chemistry and heat conduction, which are sub-cycled)
+separately, the largest $\Delta t$ allowed for stability. The level is
+advanced with a timestep equal to the minimum over all of these
+$\Delta t$.
 

### 2013-06-07 tests-sodshocktube.tex
minor text edits

--- 2013-05-31 tests-sodshocktube.tex
@@
 the root level and the results are shown at $t=0.25$.  All cells are
-plotted and color-coded by level (at the time shown, only small region
+plotted and color-coded by level with purple 
+indicating level 0, green level 1 and red level 2 (at the time shown, only small region
 surrounding the contact discontinuity and the shock are refined).  In

### 2013-06-07 tests-vandv.tex
minor text edits

@@
 that include radiative cooling and a uniform metagalactic ultraviolet
-background, \enzo\ and Gadget provide results on metrics such as the
+background, \enzo\ and GADGET provide results on metrics such as the
 matter power spectrum that are comparable to within 5\%
@@
 coefficients for the formation of molecular hydrogen via the 3-body
-process in both the \enzo\ and Gadget-2 codes, observing similar trends between
+process in both the \enzo\ and GADGET-2 codes, observing similar trends between
 the two codes. However, at nominally equivalent resolution

### 2013-06-07 tests-wavepool.tex
minor text edits

@@
 
-In this one-dimensional test we pass a short wavelength linear wave through a static refined region.
+In this one-dimensional test we pass a short wavelength linear wave through a static singley-refined region.
 The full domain is from 0 to 1 and the refined region is set from 0.25 to 0.75 at all times.
@@
 
-Finally, the MUSCL solver shows also shows a very clean entry and exit from the refined region without any oscillations. However,
-because we use a piecewise linear reconstruction, the wave is spread more than with the other methods.
+Finally, the MUSCL solver shows also shows a very clean entry and exit from the refined region without any oscillations, although with mild damping on exit.
+However, because we use a piecewise linear reconstruction, the wave is spread more than with the other methods.
 

### 2013-06-07 conclusions.tex

@@
 
-In this paper, we have presented the algorithms underlying \enzo: an
+In this paper, we have presented the algorithms underlying \enzo, an
 open-source adaptive mesh refinement code designed for
@@
 and figures for this paper are open source and are available at the
-\enzo\ website, \url{http://enzo-project.org}.  Furthermore, the \texttt{yt} toolkit,
-which is designed to analyze \enzo\ data (as well as data from a wide
-variety of other simulation tools), can be found at its website,
-\url{http://yt-project.org}.  Both of these codes have active user and
-developer communities, extensive documentation and user support, and
-strong mechanisms for users to contribute their changes and fixes to
-the codebase.
+\enzo\ website, \url{http://enzo-project.org}.  Furthermore, the
+\texttt{yt} toolkit, which is designed to analyze \enzo\ data (as well
+as data from a wide variety of other simulation tools), can be found
+at its website, \url{http://yt-project.org}.  Both of these codes have
+active user and developer communities, extensive documentation and
+user support, and strong mechanisms for users to contribute their
+changes and fixes to the codebase.
 
@@
 thread-based parallelism within a node.  This will reduce on-node
-memory usage and overall scaling behavior.
+memory usage and improve overall scaling behavior.
 \item The restructuring of \enzo's treatment of particles to
@@
 
-With the continual rapid development of computer hardware, it makes sense to review
-not only at \enzo's current capabilities, but to look towards its future development 
-in view of predicted technological trends. These trends in supercomputing hardware suggest that  
-substantial modifications to \enzo's core infrastructure, and very possibly some of 
-the core algorithms will be required. More
+With the continual rapid development of computer hardware, it makes
+sense to not only review \enzo's current capabilities, but to look
+toward its future development in view of predicted technological
+trends. These trends in supercomputing hardware suggest that
+substantial modifications to \enzo's core infrastructure, and very
+possibly some of the core algorithms, will be required. More
 specifically, the progression involves the usage of specialized
 large-core-count, vectorized computing units such as graphics
-processing units (GPUs) or chips like the Intel Xeon Phi, as well as
+processing units or chips like the Intel Xeon Phi, as well as
 precipitously decreasing amounts of RAM per computing core.  The
 former trend means that the amount of processing power per compute
-node will continue to increase, likely much faster than the
-bandwidth between nodes, and will require tremendous reduction in
-inter-node (and possibly inter-CPU) communication in order to maintain
-code scalability.  Also, much of the current code will need to be
-rewritten to take advantage of the vector nature of these CPUs, making
+node will continue to increase, likely much faster than the bandwidth
+between nodes, and will require tremendous reduction in inter-node
+(and possibly inter-CPU) communication in order to maintain code
+scalability.  Also, much of the current code will need to be rewritten
+to take advantage of the vector nature of these CPUs, making
 assumptions that are quite unlike those made in much of the current
@@
 fact are faced by effectively all applications that wish to take
-advantage of new computational architectures. We therefore anticipated that 
-\enzo\ (or a code that has the capabilities of \enzo, from a user's
-point of view) will continue to develop to be usable at the largest scales on
-such machines.
+advantage of new computational architectures. We therefore anticipated
+that \enzo\ (or a code that has the capabilities of \enzo, from a
+user's point of view) will continue to be usable at the
+largest scales on such machines.

### 2013-06-07 development.tex

@@
 openness.  Started as the graduate research project of Greg Bryan at
-the University of Illinois, it has subsequently been stewarded by
-the Laboratory for Computational Astrophysics (LCA) at the University
-of California at San Diego, and has transitioned to a distributed,
+the University of Illinois, it was subsequently stewarded by the
+Laboratory for Computational Astrophysics (LCA) at the University of
+California at San Diego, and has transitioned to a distributed,
 completely open, and community-driven project.  Initially, \enzo\ was
 versioned using a series of ``snapshots'' of the code base, usually
-hand-created by the developing individuals.  These were distributed to
-collaborators and colleagues, but the central ``trunk'' of development
-was updated primarily by a single person: while patches and technology
-were accepted from external developers, the relatively small number of
-individuals using the code resulted in a strong centralization of
-development.
+hand-created by the individuals doing development work.  These were
+distributed to collaborators and colleagues, but the central ``trunk''
+of development was updated primarily by a single person: while patches
+and technology were accepted from external developers, the relatively
+small number of individuals using the code resulted in a strong
+centralization of development.
 
@@
 Following the first public, open source release of \enzo, the code was
-migrated to using the Subversion version control system.  This is a
+migrated to the Subversion version control system.  This is a
 centralized version control system, and the ``stable'' \enzo\ source
@@
 development discussion.  Both of these lists are open and publicly
-archived.  Bitbucket provides mechanisms for inspecting source,
+archived.  Bitbucket provides mechanisms for inspecting source code,
 hosting branches and forks of the primary source, and for code review.
@@
 While peer review is able to catch many bugs and problems with source
-code changes, \enzo\ is also subject to ``answer'' testing.  We have
-created a set of parameter files and problem types that exercise the
-underlying machinery of \enzo.  These ``test problems'' have
-affiliated ``tests,'' which consist of scripts that use \texttt{yt}
+code changes, \enzo\ is also subject to ``answer'' testing (described
+in detail in Section~\ref{sec.tests.suite}).  We have created a set of
+parameter files and problem types that exercise the underlying
+machinery of \enzo.  These ``test problems'' have affiliated
+``tests,'' which consist of scripts that use \texttt{yt}
 \citep{2011ApJS..192....9T} to produce results such as mass
@@
 compared to a ``gold standard'' of results has exceeded an acceptable
-threshold, typically set to roughly single precision.  Optionally, for
-those test problems that are deemed unsafe to change to any precision,
-the tests also produce hashes of the outputs; these hashes will only
+threshold, typically set to roughly the level of roundoff error in
+single-precision floating point arithmetic.  Optionally, for those
+test problems that are deemed unsafe to change to any precision, the
+tests also produce hashes of the outputs; these hashes will only
 remain unchanged in the event of bitwise identicality between results.

### 2013-06-07 introduction.tex

@@
-% EJT: in code comments point to changes
-
 \section{Introduction}\label{sec.intro}
 
-Due to the high spatial and temporal dynamical ranges involved, astrophysics and cosmological phenomena present a taxing challenge for simulators.
-To tackle such situations, a number of numerical techniques have been developed which can be broadly split into the gridless, fully Lagrangian methods and grid-based, Eulerian schemes.
-The most commonly used is an example of the first type known as Smoothed Particle Hydrodynamics \citep[SPH;][]{Lucy77, SPH}.
-It has achieved much success, particularly in regimes dominated by gravity.
-% I felt metioning SPH before anything else was a little confusing for the article's focus so I added an earlier mention of Eulerian. 
-%where systems evolve over large spatial and temporal dynamic ranges (cut because we've already said that in the first sentense - EJT)
-However, its development to include an increasing number of sort-after physical processes is still at a relatively early stage when compared with the effort put into the latter type of Eulerian grid-based hydrodynamic schemes \citep[e.g.,][]{laney-1998, toro-1997, Woodward84}.
-% is that what was meant by `` Its development, however, is still at a relatively early stage''? Since SPH is older than AMR, I found this a little confusing as written.
+Due to the high spatial and temporal dynamical ranges involved,
+astrophysics and cosmological phenomena present a taxing challenge for
+simulators. To tackle such situations, a number of numerical
+techniques have been developed that can be broadly split into the
+gridless, Lagrangian methods and grid-based, Eulerian schemes. The
+most commonly used is an example of the first type known as Smoothed
+Particle Hydrodynamics \citep[SPH;][]{Lucy77, SPH}. It has achieved
+much success, particularly in regimes dominated by gravity. However,
+its development to include an increasing number of sought-after
+physical processes is still at a relatively early stage when compared
+with the effort put into the latter type of Eulerian grid-based
+hydrodynamic schemes \citep[e.g.,][]{laney-1998, toro-1997,
+Woodward84}.
 
-Despite this invested expertise, the Eulerian solvers in their original form have a serious drawback: they do not provide an easy method of adaptively increasing the spatial and temporal resolution in small volumes of the simulation.
-Such flexibility is essential for following physical processes such as gravitational instability.
-A solution to this was first proposed by \citet{Berger89} in the Computational Fluid Dynamics (CFD) community, and became known as Structured Adaptive Mesh Refinement (SAMR).
-The principal is to adaptively add and modify additional, finer meshes (``grids'') over regions that require higher resolution.
-In addition, it is possible to add other more advanced physics including --for the implementation in the astrophysics code presented in this paper-- comoving coordinates, self-gravity, radiative cooling, chemistry, heat conduction, collisionless fluids, magnetohydrodynamics, radiation transport, star formation and a range of other physical effects.
-% I was concerned the original text sounded a little like this was a brand new technique in a brand new code (despite the references below) which would be confusing to a reader familiar with the field and in particular, Enzo.
+Despite this invested expertise, the Eulerian solvers in their
+original form have a serious drawback: they do not provide an easy
+method of adaptively increasing the spatial and temporal resolution in
+small volumes of the simulation. Such flexibility is essential for
+following physical processes such as gravitational instability. A
+solution to this was first proposed by \citet{Berger89} in the
+Computational Fluid Dynamics (CFD) community, and became known as
+Structured Adaptive Mesh Refinement (SAMR). The principle is to
+adaptively add and modify additional, finer meshes (``grids'') over
+regions that require higher resolution. In addition, it is possible to
+add other, more advanced physics including -- for the AMR
+implementation in the astrophysics code presented in this paper --
+comoving coordinates, self-gravity, radiative cooling, chemistry, heat
+conduction, collisionless fluids, magnetohydrodynamics, radiation
+transport, star formation and a range of other physical effects.
 
-There have been a number of numerical methods described in the astronomical literature which contain elements of SAMR or have a similar aim.
-For example, the N-body solver developed by \citet{Villumsen89} used non-adaptive meshing to increase the resolution in pre-selected regions.
-This static approach was later used extensively when applied to hydrodynamics \citep[e.g.,][]{Ruffert94, Anninos94}.
-Adding adaptivity is a more recent enhancement, and there are now a number of N-body codes that possess this feature, both with and without  hydrodynamics  \citep{Couchman91, Jessop94, Suisalu95, Splinter96, Gelato97, ART97, Truelove98, flash_method, MLAPM01,  Yahagi01, RAMSES, Quilis04, Ziegler05, Zhang06, Astrobear09, Pluto-amr, GAMER, Nyx}.
-Of these, perhaps the most comparable and widely used are FLASH \citep{flash_method}, which uses grid blocks of fixed size, and RAMSES \citep{RAMSES} and ART \citep{ART97}, both of which refine individual cells.
-It is also possible to deform the grid to obtain high resolution \citep[e.g.,][]{Gnedin95, Xu97, Pen98}, and more recently a few grid codes have adopted an unstructured approach based on a moving Voronoi mesh \citep{Arepo10, Tess11}.
+There have been a number of numerical methods described in the
+astronomical literature that contain elements of SAMR or have a
+similar aim. For example, the N-body solver developed by
+\citet{Villumsen89} used non-adaptive meshing to increase the
+resolution in pre-selected regions. This static approach was later
+used extensively when applied to hydrodynamics
+\citep[e.g.,][]{Ruffert94, Anninos94}. Adding adaptivity is a more
+recent enhancement, and there are now a number of N-body codes that
+possess this feature, both with and without hydrodynamics
+\citep{Couchman91, Jessop94, Suisalu95, Splinter96, Gelato97, ART97,
+Truelove98, flash_method, MLAPM01, Yahagi01, RAMSES, Quilis04,
+Ziegler05, Zhang06, Astrobear09, Pluto-amr, GAMER, Nyx}. Of these,
+perhaps the most comparable and widely used are FLASH
+\citep{flash_method}, which uses grid blocks of fixed size, and RAMSES
+\citep{RAMSES} and ART \citep{ART97}, both of which refine individual
+cells. It is also possible to deform the grid to obtain high
+resolution \citep[e.g.,][]{Gnedin95, Xu97, Pen98}, and more recently a
+few grid codes have adopted an unstructured approach based on a moving
+Voronoi mesh \citep{Arepo10, Tess11}.
 
-In this paper, we present \enzo; a structured adaptive mesh refinement (SAMR) code.
-Originally developed for cosmological hydrodynamics, \enzo\ has since been used on a wide variety of problems.
-It has grown to become a general tool for astrophysical fluid dynamics and is intended to be efficient, accurate and easily extended to include new capabilities.
-Although many of the components of the \enzo\ code have been described in previous publications \citep{1995CoPhC..89..149B, BryanThesis96, Bryan97a, Bryan97b, Norman99, BryanCompSci99, Bryan01, Oshea04, 2007arXiv0705.1556N, WangAbelZhang08,ReynoldsHayesPaschosNorman2009,Collins10,Wise11_Moray}, there has previously been no systematic and complete description of the code.
-In this paper we provide that description, filling in many previous omissions and showing the code's performance for a wide variety of test problems.
+In this paper, we present \enzo; a structured adaptive mesh refinement
+(SAMR) code. Originally developed for cosmological hydrodynamics,
+\enzo\ has since been used on a wide variety of problems.  It has
+grown to become a general tool for astrophysical fluid dynamics and is
+intended to be efficient, accurate and easily extended to include new
+capabilities.  Although many of the components of the \enzo\ code have
+been described in previous publications \citep{1995CoPhC..89..149B,
+BryanThesis96, Bryan97a, Bryan97b, Norman99, BryanCompSci99, Bryan01,
+Oshea04, 2007arXiv0705.1556N,
+WangAbelZhang08,ReynoldsHayesPaschosNorman2009,Collins10,Wise11_Moray},
+there has previously been no systematic and complete description of
+the code.  In this paper we provide that description, filling in many
+previous omissions and showing the code's performance for a wide
+variety of test problems.
 
-The \enzo\ code has been extensively used over the last two decades in a wide variety of problems, resulting in the publication of more than 100 peer-reviewed papers based on results obtained with the code.
-The variety of astrophysical systems that \enzo\ has been used for include galaxies \citep{2003ApJ...587...13T, 2012ApJ...749..140H}, galaxy clusters \citep{Loken02, Skillman13}, the interstellar medium \citep{Slyz05}, the intergalactic medium \citep{Fang01, 2011ApJ...731....6S}, cooling flows \citep{Li12, 2013ApJ...763...38S}, turbulence \citep{Kritsuk04, 2009JPhCS.180a2020K, Collins11}, the formation of the first stars \citep{ABN02, 2007ApJ...654...66O, 2009Sci...325..601T}, and the formation of stars in our own Galaxy \citep{Collins12a}.
+The \enzo\ code has been extensively used over the last two decades in
+a wide variety of problems, resulting in the publication of more than
+100 peer-reviewed papers based on results obtained with the code. The
+variety of astrophysical systems that \enzo\ has been used for include
+galaxies \citep{2003ApJ...587...13T, 2012ApJ...749..140H}, galaxy
+clusters \citep{Loken02, Skillman13}, the interstellar medium
+\citep{Slyz05}, the intergalactic medium \citep{Fang01,
+2011ApJ...731....6S}, cooling flows \citep{Li12, 2013ApJ...763...38S},
+turbulence \citep{Kritsuk04, 2009JPhCS.180a2020K, Collins11}, the
+formation of the first stars \citep{ABN02, 2007ApJ...654...66O,
+2009Sci...325..601T}, and the formation of stars in our own Galaxy
+\citep{Collins12a}.
 
-% 'finally' here sounded strange, since this is really the key point
-Numerical simulations of astronomical phenomena now play a key role, along with observations and analytic theory, in pushing forward our understanding of the cosmos \citep[e.g.,][]{DecadalSurvey01, DecadalSurvey10}.
-But along with this role comes responsibility.
-We believe that those developing simulation tools must fulfill two key obligations: the first is to make those tools available to the community as a whole, much in the way that astronomical data is now regularly made publicly available.
-The second is to document, test and refine those methods so that they can be critically evaluated and expanded by others.
-This paper represents one part of our attempt to meet our obligations with respect to the \enzo\ code.
+Numerical simulations of astronomical phenomena now play a key role,
+along with observations and analytic theory, in pushing forward our
+understanding of the cosmos \citep[e.g.,][]{DecadalSurvey01,
+DecadalSurvey10}.  But along with this role comes responsibility.  We
+believe that those developing simulation tools must fulfill two key
+obligations: the first is to make those tools available to the
+community as a whole, much in the way that astronomical data is now
+regularly made publicly available.  The second is to document, test
+and refine those methods so that they can be critically evaluated and
+expanded upon by others.  This paper represents one component of our
+attempt to meet these obligations with respect to the \enzo\ code.
 
-The structure of the code is as follows.
-In Section~\ref{sec.overview}, we first provide a top-level overview of the code method and structure.
-This is designed to give a first broad-brush picture of the equations we aim to solve and the methods used to solve them.
-Next, in Sections~\ref{sec.amr} through~\ref{sec.num.analysis}, we describe the methods we use in detail, reserving some of the longer descriptions of particular components for the appendix in order not to interrupt the flow of the paper.
-The \enzo\ testing framework and code tests are described in Section~\ref{sec.tests}.
-The parallelism strategy and scaling results are described in Section~\ref{sec.parallel}.
-Finally, we discuss the code's development methodology (which is as far as we know unique in the astrophysics community) in Section~\ref{sec.development}.
+The structure of this paper is as follows.  In
+Section~\ref{sec.overview}, we first provide a top-level overview of
+the code method and structure.  This is designed to give a broad-brush
+picture of the equations solved by \enzo\ and the methods used to
+solve them.  Next, in Sections~\ref{sec.amr}
+through~\ref{sec.num.analysis}, we describe the methods we use in
+detail, reserving some of the longer descriptions of particular
+components for the appendix in order to not interrupt the flow of the
+paper.  The \enzo\ testing framework and code tests are described in
+Section~\ref{sec.tests}.  The parallelism strategy and scaling results
+are described in Section~\ref{sec.parallel}.  Finally, we discuss the
+code's development methodology (which is, as far as we know, unique in
+the astrophysics community) in Section~\ref{sec.development}.

### 2013-06-07 numerical-chemistry.tex

@@
 non-equilibrium chemistry of a gas of primordial composition with 12 total
-species: H, \Hp, He, \Hep, \Hepp, \Hm, \HHp, \HH, e$^-$, D, \Dp, and HD.  Enzo
+species: H, \Hp, He, \Hep, \Hepp, \Hm, \HHp, \HH, e$^-$, D, \Dp, and HD.  \enzo\
 also computes the radiative heating and cooling of the gas from atomic and
@@
 \end{tabular}
-\caption[]{Radiative processes solved in the Enzo nonequilibrium
+\caption[]{Radiative processes solved in the \enzo\ nonequilibrium
 primordial chemistry routines. \red{Missing rates for HD destruction/ionization?}}

### 2013-06-07 numerical-mhd-ct.tex

@@
 interpolated in concert.  Second, any magnetic field information from
-the previous time step must be included in the interpolation, making the
+the previous timestep must be included in the interpolation, making the
 interpolation more complex than the simple parent-child relation used

### 2013-06-07 numerical-starformation.tex
Edits to the overview section. Trying to unify notations, so several other sections are also affected.

@@
 %\begin{equation}
-\nabla \cdot \vec{v}_b & < & 0, \\
+\div \vec{v}_b & < & 0, \\
 \label{cencont}
@@
 Algorithmically, the star formation events in the Schmidt-law algorithm are
-assumed to occur once every global time step (with the constraint $\Delta t_0 \leq
+assumed to occur once every global timestep (with the constraint $\Delta t_0 \leq
 10^7$ years).
@@
 The \enzo\ implementation of this algorithm is similar, except that
-instead of forming stars only at the root grid time step, we allow
-stars to form at the time step of the highest level of resolution at
+instead of forming stars only at the root grid timestep, we allow
+stars to form at the timestep of the highest level of resolution at
 any particular point in space.  As can be seen from the equation for
@@
 
-\item A converging gas flow ($\nabla \cdot \vec{v} < 0$), and
+\item A converging gas flow ($\div \vec{v} < 0$), and
 

### 2013-06-07 numerical-zeus.tex

@@
 \begin{eqnarray}
-\rho \frac{\partial\vecv}{\partial t} &=& - \nabla p - \rho \nabla \phi 
-- \nabla \cdot \textbf{Q} \\
-\frac{\partial e}{\partial t} &=& -p \nabla \cdot \vecv - \textbf{Q} : \nabla \vecv, 
+\rho \frac{\partial\vecv}{\partial t} &=& - \grad p - \rho \grad \phi 
+- \div \textbf{Q} \\
+\frac{\partial e}{\partial t} &=& -p \div \vecv - \textbf{Q} : \grad \vecv, 
 \end{eqnarray}
@@
 \begin{equation}
-e^{n+c}_j = e^{n+b}_j \left( \frac{1 - (\Delta t/2) (\gamma - 1) (\nabla \cdot \vecv)_j }
-                           {1 + (\Delta t/2) (\gamma - 1) (\nabla \cdot \vecv)_j } \right)
+e^{n+c}_j = e^{n+b}_j \left( \frac{1 - (\Delta t/2) (\gamma - 1) (\div \vecv)_j }
+                           {1 + (\Delta t/2) (\gamma - 1) (\div \vecv)_j } \right)
 \end{equation}
-We have used the notation $(\nabla \cdot \vecv)_j$ to indicate the (potentially) multi-dimensional velocity divergence evaluated at the cell center position $x_j$.
+We have used the notation $(\div \vecv)_j$ to indicate the (potentially) multi-dimensional velocity divergence evaluated at the cell center position $x_j$.
 This equation differs from the previous ones in that in the multi-dimensional case, still only one finite difference equation is evaluated, but the divergence becomes multi-dimensional.

### 2013-06-07 parallel.tex

@@
 
-At each complete root grid time step (or less often if specified),
+At each complete root grid timestep (or less often if specified),
 each timer is then communicated to the root processor where it
@@
 straightforward thing to do in codes that use static grids, as the
-computational effort per time step is constant.  However, for an AMR
+computational effort per timestep is constant.  However, for an AMR
 calculation, the number of grids that will be generated during the run
@@
 or in other words, that changes in the resolution affect the
-simulation in the same way at each time step.  This is usually a good
+simulation in the same way at each timestep.  This is usually a good
 approximation, as the most costly parts of the calculation typically
@@
 root}^3$ and, assuming the Courant condition is the 
-factor that controls the time step, the number of steps to advance the
+factor that controls the timestep, the number of steps to advance the
 calculation over a given time interval is

### 2013-06-07 table_of_variables_and_constants.tex

@@
   $\fcond$ & thermal conduction heat flux \\
-  $\kappa_{sp}$ & Spitzer conduction coefficient \\
-  $f_{Sp}$ & fraction of Spitzer conductivity \\[3pt]
+  $\kappa_{\rm cond}$ & thermal conduction coefficient \\
+  $\kappa_{\rm sp}$ & Spitzer conductivity \\
+  $f_{\rm sp}$ & fraction of Spitzer conductivity \\[3pt]
 

### 2013-06-07 tests-sedov.tex

@@
 background density and $\alpha = 1.0$ for cylindrical symmetry and an
-ideal gas with $\gamma = 1.4$.  For the full derivation see
+ideal gas with $\gamma = 1.4$; for the full derivation see
 \citet{Sedov1959}.  Third, as the spherical shock expands, its

### 2013-06-07 numerical-particle.tex
Minor change to note that PotentialIterations is user-facing and add comment about MaximumParticleRefinementLevel.

@@
 end of each timestep on a level.
+
+To avoid unphysical point-mass effects, \enzo\ provides a parameter that
+governs the maximum level at which particles will be regarded at point masses.
+At higher levels, contributions from particles to the gravitating mass field
+will be smoothed over a spherical region centered at each particle's center.

### 2013-06-07 tests-onezonecollapse.tex
Minor note to emphasize we're not using constant gamma.

@@
 time.  It is designed to test the chemistry and cooling modules over a
-wide range in densities and over physically motivated timescales.
+wide range in densities and over physically-motivated timescales.
 Because this test disables the hydrodynamic and gravity solvers and
@@
 and energy, temperature, density, and pressure are related by the
-ideal gas law.  Figure \ref{fig.onezone} shows an example of the
-one-zone collapse test performed with an initial number density of 1
-hydrogen atom per cm$^{-3}$ and an initial temperature of 100 K using
-the 12 species chemistry network with H, D, and He species and metal
-cooling rates calculated with the \texttt{Cloudy} code.  The effects
-of metal cooling can be clearly seen; as the metallicity increases
-from zero to $10^{-2}$~Z$_\odot$, the gas rapidly and significantly
-deviates from the primordial result (black line).  Our primordial
-results compare very well to those shown in
-\citet{2005ApJ...626..627O}; however, we use a different metal cooling
-method, so the lines describing the evolution of the
-metal-enriched gas are not directly comparable.
+ideal gas law, including effects from molecular hydrogen as appropriate.
+Figure \ref{fig.onezone} shows an example of the one-zone collapse test
+performed with an initial number density of 1 hydrogen atom per cm$^{-3}$ and
+an initial temperature of 100 K using the 12 species chemistry network with H,
+D, and He species and metal cooling rates calculated with the \texttt{Cloudy}
+code.  The effects of metal cooling can be clearly seen; as the metallicity
+increases from zero to $10^{-2}$~Z$_\odot$, the gas rapidly and significantly
+deviates from the primordial result (black line).  Our primordial results
+compare very well to those shown in \citet{2005ApJ...626..627O}; however, we
+use a different metal cooling method, so the lines describing the evolution of
+the metal-enriched gas are not directly comparable.
 

### 2013-06-07 numerical-chemistry.tex
Three minor changes to table:
 * H2 + D+ => HD + D+ has become H2 + D+ => HD + H+
 * Added H- + D => HD + e- which is in the code
 * Added D photoionization rate.

@@
 (27) & H &+ &D$^+$ & $\rightarrow$ & H$^+$ &+ &D \\
-(28) & H$_2$ &+ &D$^+$ & $\rightarrow$ & HD &+ &D$^+$ \\
+(28) & H$_2$ &+ &D$^+$ & $\rightarrow$ & HD &+ &H$^+$ \\
 (29) & HD &+ &H$^+$ & $\rightarrow$ & H$_2$ &+ &D$^+$ \\
@@
 (31) & HD &+ &H & $\rightarrow$ & H$_2$ &+ &D \\
+(32) & H$^-$ &+ &D & $\rightarrow$ & HD &+ &e$^-$ \\
 
@@
 \begin{tabular}{llllllll}
-(32) & H & + & $\gamma$ & $\rightarrow$ & H$^+$ & + & e$^-$ \\
-(33) & He & + & $\gamma$ & $\rightarrow$ & He$^+$ & + & e$^-$ \\
-(34) & He$^+$ & + & $\gamma$ & $\rightarrow$ & He$^{++}$ & + & e$^-$ \\
-(35) & H$^-$ & + & $\gamma$ & $\rightarrow$ & H & + & e$^-$ \\
-(36) & H$_2$ & + & $\gamma$ & $\rightarrow$ & H$_2^+$ & + & e$^-$ \\
-(37) & H$_2^+$ & + & $\gamma$ & $\rightarrow$ & H & + & H$^+$ \\
-(38) & H$_2^+$ & + & $\gamma$ & $\rightarrow$ & 2H$^+$ & + & e$^-$ \\
-(39) & H$_2$ & + & $\gamma$ & $\rightarrow$ & H$_2^*$ & $\rightarrow$ & 2H \\
-(40) & H$_2$ & + & $\gamma$ & $\rightarrow$ & 2H &  & 
+(33) & H & + & $\gamma$ & $\rightarrow$ & H$^+$ & + & e$^-$ \\
+(34) & He & + & $\gamma$ & $\rightarrow$ & He$^+$ & + & e$^-$ \\
+(35) & He$^+$ & + & $\gamma$ & $\rightarrow$ & He$^{++}$ & + & e$^-$ \\
+(36) & H$^-$ & + & $\gamma$ & $\rightarrow$ & H & + & e$^-$ \\
+(37) & H$_2$ & + & $\gamma$ & $\rightarrow$ & H$_2^+$ & + & e$^-$ \\
+(38) & H$_2^+$ & + & $\gamma$ & $\rightarrow$ & H & + & H$^+$ \\
+(39) & H$_2^+$ & + & $\gamma$ & $\rightarrow$ & 2H$^+$ & + & e$^-$ \\
+(40) & H$_2$ & + & $\gamma$ & $\rightarrow$ & H$_2^*$ & $\rightarrow$ & 2H \\
+(41) & H$_2$ & + & $\gamma$ & $\rightarrow$ & 2H &  & \\
+(42) & D & + & $\gamma$ & $\rightarrow$ & D$^+$ & + & e$^-$ \\
 \end{tabular}

### 2013-06-07 numerical-chemistry.tex
Round 2 edits of microphysics & radiation.  There are still missing rates for HD destruction/ionization. Radiation background text is out of sync with the enzo documentation.

@@
 the local, highly non-equilibrium population of free electrons.  As a result,
-when modeling primordial star formation, it is critical to follow the
+when modeling primordial star formation it is critical to follow the
 non-equilibrium evolution of the chemical species of hydrogen, including
@@
 The primordial non-equilibrium chemistry routines used in \enzo\ were first
-described by Abel et al. and Anninos et al. \citep{abel97,anninos97}, but have
-been extended since then with updated reaction rates and the inclusion of
+described by \citet{abel97} and \citet{anninos97}, but have
+since been extended with updated reaction rates and the inclusion of
 deuterium species \citep{2009PhDT.........5T}.  These routines follow the
@@
 reaction rate coefficients and the cooling functions of the gas) on timescales
-that faster than those of the gas dynamics.
+that are faster than those of the gas dynamics.
 
@@
 formation (\HH, \HHp, and \Hm), and can further include deuterium and its
-species (D, \Dp, and HD).  A total of 9 kinetic equations are solved, including
-29 kinetic and radiative processes, for the 12 species mentioned above.  See
+species (D, \Dp, and HD).  A total of 9 kinetic equations are solved for the
+12 species mentioned above, including
+29 kinetic and radiative processes.  See
 Table~\ref{table.collisional} for the collisional processes and
@@
 be very time-consuming and computationally costly for a relatively small
-increase in accuracy.  At this present time, this makes fully-implicit methods
-somewhat undesirable for a large, three-dimensional simulation.
+increase in accuracy.  At the present time, this makes fully-implicit methods
+somewhat undesirable for large, three-dimensional simulations.
 
@@
 time; all other contributions to the destruction component are computed at the
-current time.  This mixed state improves accuracy and ensures species values
+current time.  This mixed state improves accuracy, ensures species values
 are positive definite, and is equivalent to one Jacobi iteration of an implicit

### 2013-06-07 numerical-cooling.tex
Round 2 edits of microphysics & radiation.  There are still missing rates for HD destruction/ionization. Radiation background text is out of sync with the enzo documentation.

@@
 background (CMB), and photo-heating from a UV metagalactic background.
-If the \HH chemistry network is enabled, the following \HH
+If the \HH~chemistry network is enabled, the following \HH~
 cooling processes are also considered: ro-vibrational transitions
@@
 the metal species only.  The primary assumption made is that of
-ionization equilibrium.  The tables can vary along up to five dimensions,
-of density, metallicity, electron fraction,
+ionization equilibrium.  The tables can vary along up to five dimensions:
+density, metallicity, electron fraction,
 redshift (for an evolving metagalactic UV background), and

### 2013-06-07 numerical-radiation-fld.tex
Round 2 edits of microphysics & radiation.  There are still missing rates for HD destruction/ionization. Radiation background text is out of sync with the enzo documentation.

@@
 radiation transport solvers may be found in \cite{IlievEtAl2009}.  
-
 \enzo's field-based radiation solver focuses on
@@
 (\ref{eq:fld_heating_qss_fe}-\ref{eq:fld_chemistry_qss_fn}).  In this
-iteration, for the first 50 sweeps we use $\lambda=1$, and for more
-challenging problems if that does not suffice we switch to a damping
+iteration, for the first 50 sweeps we use $\lambda=1$. For more 
+challenging problems where this does not converge, we switch to a damping
 parameter of $\lambda=0.1$.

### 2013-06-07 numerical-radiation-raytracing.tex
Round 2 edits of microphysics & radiation.  There are still missing rates for HD destruction/ionization. Radiation background text is out of sync with the enzo documentation.

@@
 %  \frac{1}{c} \; \frac{\partial I_\nu}{\partial t} + 
-%  \frac{\hat{n} \cdot \nabla I_\nu}{\bar{a}} -
+%  \frac{\hat{n} \cdot \grad I_\nu}{\bar{a}} -
 %  \frac{H}{c} \; \left( \nu \frac{\partial I_\nu}{\partial \nu} -
@@
 calculations are small enough so that $\Delta a/a \ll 1$, therefore
-$\bar{a} = 1$ in any given timestep, reducing the second term to
+$a_{em}/a \approx 1$ in any given timestep, reducing the second term to
 $\hat{n} \partial I_\nu/\partial \mathbf{x}$.  To determine the
@@
 
-yay tracing is an accurate method to propagate radiation from point
-sources on a computational grid, given that there are a sufficient
+Ray tracing is an accurate method to propagate radiation from point
+sources on a computational grid as long as there are a sufficient
 number of rays passing through each cell.  Along a ray, the radiation
@@
 \end{equation}
-
 where $P$ is the photon number flux along the ray.  To sample the
@@
 We avoid this inefficiency by utilizing adaptive ray tracing
-\citep{Abel02_RT}, which progressively splits rays when the sampling
-becomes too coarse and is based on Hierarchical Equal Area isoLatitude
-Pixelation \citep[HEALPix;][]{HEALPix}.  In this scheme, the rays are
-traced along normal directions of the centers of HEALPix pixels, which
-evenly divides a sphere into equal areas.  The rays are initialized at
+\citep{Abel02_RT}, which is based on Hierarchical Equal Area isoLatitude
+Pixelation \citep[HEALPix;][]{HEALPix} and progressively splits rays when the sampling
+becomes too coarse.  In this approach, the rays are
+traced along normal directions of the centers of the HEALPix pixels that 
+evenly divide a sphere into equal areas.  The rays are initialized at
 each point source with the photon luminosity (photon s$^{-1}$) equally
@@
 boundary conditions.
-
 In the first case, the photon is halted at that position and saved,
@@
 prevents our method from considering sources external to the
-computational domain, but a uniform radiation background can be used
-in conjunction with ray tracing that adds the local radiation field to
-the background intensity.
+computational domain. However, a uniform radiation background can be used
+in conjunction with ray tracing that adds the background intensity to the
+local radiation field.
 
@@
 In each cell, the photo-ionization and photo-heating rates from each
-ray in the calculation are summed, and after the ray tracing is
+ray in the calculation are summed. After the ray tracing is
 complete, these rates are used as inputs to the solver described in

### 2013-06-07 numerical-ppm.tex
Fixed up PPM section somewhat.  Still needs more work.

@@
 
-The primary hydrodynamic method used in \enzo\ is based on the
+One (purely) hydrodynamic method used in \enzo\ is closely based on the
 piecewise parabolic method (PPM) of
@@
 modified for the study of cosmological fluid flows.  
-\citet{1984JCoPh..54..174C} describe two variants on this method: a
-Lagrangian Remap variant and a Direct Eulerian variant.  Both have
-been adapted for \enzo, and the Lagrangian Remap 
-variant is described in \citet{1995CoPhC..89..149B}.  Because of its
-conservative formulation, however, the Direct Eulerian variant is
-preferred for \enzo\ AMR simulations.  We will describe the Direct
-Eulerian (PPM-DE) variant here.
+\citet{1984JCoPh..54..174C} describe two variants on this method:
+Lagrangian Remap and Direct Eulerian; here we use the Direct Eulerian
+version which is better suited for AMR simulations.  The Lagrangian
+Remap version has previously been adapted for cosmological use
+as described in \citet{1995CoPhC..89..149B} and the implementation
+we use here is closely based on that work.
 
-We can examine the structure of Equations~(\ref{eq:mass})-(\ref{eq:total_energy})
-to obtain a better understanding of the effect of each term.
-The derivatives are first order in time
-and provide us with the dependent variables that will be evaluated each
-time step: $\rhob$, $\vecvb$, and $E$.
+%We can examine the structure of Equations~(\ref{eq:mass})-(\ref{eq:total_energy})
+%to obtain a better understanding of the effect of each term.
+%The derivatives are first order in time
+%and provide us with the dependent variables that will be evaluated each
+%timestep: $\rhob$, $\vecvb$, and $E$.
+%
+%The second term on the left hand side of each relation
+%is due to our choice of Eulerian coordinates and corresponds to
+%advection, the flow of conserved quantities 
+%($\rhob$, $\rhob \vecvb$, and $\rhob E$) 
+%across the mesh.  These will be termed the transport terms.
 
-The second term on the left hand side of each relation
-is due to our choice of Eulerian coordinates and corresponds to
-advection, the flow of conserved quantities 
-($\rhob$, $\rhob \vecvb$, and $\rhob E$) 
-across the mesh.  These will be termed the transport terms.
-
-The first term on the right hand side of Equations~(\ref{eq:momentum}) and
-(\ref{eq:total_energy}) comes from our choice of comoving coordinates
-(a similar term also appears in Equation~(\ref{eq:dm_velocity}), the velocity 
-relation for the dark matter particles).  It does not
+The first term on the right hand side of equations~(\ref{eq:momentum}) and
+(\ref{eq:total_energy}) come from our choice of comoving coordinates
+(a similar term also appears in equation~(\ref{eq:dm_velocity}), the velocity 
+relation for the dark matter particles).  A similar term does not
 appear in the mass conservation equation (\ref{eq:mass}) because
@@
 source of error.  
+We solve these terms (for all hydro methods)
+using the same technique -- they are split from the rest of the terms and
+solved using an (implicit) time-centered method, which is straightforward
+as there are no spatial gradients.
 
-The remainder of the terms (due to pressure and gravity) are denoted as
-source terms.  They drive instabilities and discontinuities in the flow.
-We make these distinctions because each of the three
-groups of terms have different numerical behaviour and are therefore
-treated differently.  The source and transport steps will be solved in
-a self-consistent fashion by the PPM scheme while the expansion terms,
-which are spatially localized, are split off and solved in a separate step
-(as are radiative cooling and heating, if present).
+%
+%The remainder of the right-hand side terms (due to pressure and gravity) are denoted as
+%source terms.  We make these distinctions because the 
+%source and transport steps will be solved in
+%a self-consistent fashion by the PPM scheme while the expansion terms,
+%which are spatially localized, are split off and solved in a separate step
+%(as are radiative cooling, heating and other processes, if present).
 
-The remainder of the terms (transport and source) in the fluid equations
+The remainder of the terms in the fluid equations
 involve derivatives.  
@@
 but now apply it spatially.   We can rewrite the one-dimensional (Eulerian) 
-versions of Equations~(\ref{eq:mass})--(\ref{eq:total_energy}) without expansion terms
+versions of equations~(\ref{eq:mass})--(\ref{eq:total_energy}) without expansion terms
 in conservative form as,
@@
 
-\subsubsection{The piecewise parabolic method in one dimension} % 2.2
+%\subsubsection{The piecewise parabolic method in one dimension} % 2.2
 
-We now restrict ourselves to the solution of Equations
-(\ref{eq:mass1d})--(\ref{eq:total_energy1d}), plus transport, in one
+We now restrict ourselves to the solution of equations
+(\ref{eq:mass1d})--(\ref{eq:total_energy1d}), in one
 dimension.  In the Direct Eulerian version of PPM, this is
@@
 left and right states at each grid boundary by constructing a
-piecewise parabolic description of the primary variables ($\rho$, $u$
+piecewise parabolic description of the primative variables ($\rho$, $u$
 and $E$) and then averaging over the regions corresponding to the
@@
 describe the steps required to compute these quantities.
-%Note that these difference equations are conservative.
 
@@
 parabolic (third order) interpolations in one dimension
-for each of $p$, $\rho$, and $v$  The pressure is determined from
-Equation~(\ref{eq:eq_of_state}), the equation of state.
+for each of $p$, $\rho$, and $v$.  The pressure is determined from
+equation~(\ref{eq:eq_of_state}), the equation of state.
 %We have dropped the subscript b in this and the following two sections because we will be referring solely to baryonic quantities.  
@@
 %
-This is Equation~(1.4) of CW84 (in the radial rather than mass coordinate as is appropriate for the direct Eulerian approach). 
+This is Equation~(1.4) of CW84 (in the spatial rather than mass coordinate as is appropriate for the direct Eulerian approach). 
 %The zones edges in the mass coordinates are $m_{j-1/2}$ and $m_{j+1/2}$ for the left and right edges, respectively.  
@@
 reproduced here, but see Equations 1.7 to 1.10 of \citet{1984JCoPh..54..174C}.
-We also optionally allow steepening as described in this reference.
+We also optionally allow steepening as described in that reference.
 
-In the Lagrangian case, once we have the interpolation constants we compute the characteristic domain for each zone edge.
-This is simply the farthest a sound wave could travel in order to reach the interface by the end of a timestep.
-The characteristic domain encompasses all of the information that can reach the (Lagrangian) zone edge during the current timestep.
-This information would then be fed into the Riemann solver to find the flux at each Lagrangian cell boundary.
-
-In the Eulerian approach, this is more complex because the Eulerian
-zone boundary does not naturally correspond to one the three characteristic waves and so a more complicated approach is required.
-First, the primary quantities ($p$, $\rho$, $v$ and $E$) are averaged over the domains corresponding to the three characteristics $C = u-c$, $u$ or $u+c$ (where $c$ is the sound speed in a cell):
+Once we have the reconstruction, the primary quantities ($p$, $\rho$, $v$ and $E$) are averaged over the domains corresponding to the three characteristics $C = u-c$, $u$ or $u+c$ (where $c$ is the sound speed in a cell):
 \begin{eqnarray}
-      q^{C+}_{j+1/2} & = & f_{j+1/2,L}(\Delta t C^n_j), \\
-      q^{C-}_{j+1/2} & = & f_{j+1/2,R}(-\Delta t C^n_{j+1}), \nonumber
+      q^{C+}_{j+1/2} & = &
+          {1 \over \Delta t C^n_j} \int^{x_{j+1/2}}_{x_{j+1/2}-\Delta t C^n_j} q_j(x) dx, \\
+      q^{C-}_{j+1/2} & = & 
+         {1 \over -\Delta t C^n_{j+1}} \int^{x_{j+1/2}+\Delta t C^n_{j+1}}_{x_{j+1/2}} q_{j+1}(x) dx \nonumber
 \end{eqnarray}
-where $q^{C+}$ ($q^{C-}$) refers to the left (right) characteristic $C$.  The functions $f$ are just integrals
-over the parabolic interpolation formula:
-%
-\begin{eqnarray}
-f_{j+1/2,L}(y) & = &
-       {1 \over y} \int^{x_{j+1/2}}_{x_{j+1/2}-y} q_j(x) dx, \\
-f_{j+1/2,R}(y) & = &
-       {1 \over y} \int^{x_{j+1/2}+y}_{x_{j+1/2}} q_{j+1}(x) dx. \nonumber
-\end{eqnarray}
-The linearized gas dynamics equations are then used to compute second-order correct left and right states that take into account the multiple wave families.
+where $q^{C+}$ ($q^{C-}$) refers to the left (right) characteristic $C$.
+The linearized gas dynamics equations are then used to compute second-order accurate left and right states that take into account the multiple wave families.
 This process is described in \citet{1984JCoPh..54..174C} and we use their equations (3.6) and (3.7).
@@
 With these effective states, an approximation to the Riemann problem is found
-with an iterative approach \citep[see][]{Woodward86}, producing estimates for 
+with an iterative approach \citep[see][and section~\ref{sec.riemann}]{Woodward86}, producing estimates for 
 $\overline{p}_{j\pm 1/2}$, $\overline{\rho}_{j\pm1/2}$, and $\overline{v}_{j\pm 1/2}$ that are third order
@@
 proved difficult to do without generating large errors in the internal
-energy and so we conservatively leave the pressure at the old time in
-this difference equation.  This leads to spreading of shocks; however,
+energy and so we leave the pressure at the old time in
+this difference equation.  This leads to some spreading of shocks; however,
 we note that this equation is only used in hypersonic flows.
@@
 After some experimentation,
-we have chosen ${\eta}_2 = 0.08$, a somewhat conservative value.
+we have chosen ${\eta}_2 = 0.1$, a somewhat conservative value.
 This scheme is optional and is generally only required in large-scale 
@@
 \subsubsection{Riemann Solvers and Fallback}
+\label{sec.riemann}
 
 While most numerical hydrodynamics schemes center around finding approximate
-solutions to the hydrodynamics equations, the core of philosophy of a Godunov's
+solutions to the hydrodynamics equations, the core philosophy of a Godunov
 method \citep{Godunov1959} is to find the exact solution to an
@@
 approximations can be formulated for the MHD equations, and are employed in both
-Dedner and CT solvers in \enzo.  The HLLD
+the Dedner and CT solvers in \enzo.  The HLLD
 solver includes two of the additional waves, the rotational discontinuities,
@@
 solver fallback mechanism \citep{Lemaster09}.  If a negative density is found at
-a particular interface, the more diffusive HLL Riemann solver is used instead.
-
-% ------------------------------------------------------------
-
-% GB: I commented the following out because it has been incorporated into the above
-%
-%PPM is an explicit, higher order-accurate version of
-%Godunov's method for ideal gas dynamics with third order-accurate piecewise parabolic
-%monotonic interpolation and a nonlinear Riemann solver for shock
-%capturing.  It advances the hydrodynamic equations in the following
-%steps:
-%\begin{enumerate}
-% \item Construct monotonic parabolic interpolation of cell average
-% data, for each fluid quantity.
-% \item Compute interface states by averaging the parabola for over the domain of dependence for
-% each interface
-% \item Use interface data to solve the Riemann problem.
-% \item Difference the interface fluxes to update the cell average quantities.
-%\end{enumerate}
-%It does an excellent job of capturing strong shocks in at
-%most two cells.  Multidimensional schemes are built up by directional
-%splitting and produce a method that is formally second order-accurate
-%in space and time which explicitly conserves mass, linear momentum, and energy.
-
-%To use the above algorithm for cosmological simulations, the
-%conservation laws for fluid mass, momentum and energy 
-%density are written in comoving coordinates for a
-%Friedman-Robertson-Walker space-time, as described previously 
-%in Equations~\ref{enzoconserve} 
-%through~\ref{enzoenergy}.  Both the conservation laws and
-%the Riemann solver are modified to include gravity, which is solved
-%using an adaptive particle-mesh (PM) technique (see Section~\ref{sec.ov.nbody}).  
-%The terms due to cosmological expansion, as well as primordial chemistry and
-%radiative heating and cooling, are solved in a separate step because they have
-%different numerical behavior, and therefore must be treated differently to ensure
-%stability.  Note that unlike the \zeus\ hydro scheme, PPM does not need to use 
-%artificial viscosity to resolve shocks.
-
-%The system of equations described above works well for systems with
-%relatively low Mach numbers, as long as these systems are well resolved.  
-%However, cosmology is replete with situations where there are bulk hypersonic
-%flows.  In these situations, the ratio of kinetic to thermal energy can be
-%very high -- in some situations up to $10^6 - 10^8$.  This implies that the
-%thermal energy is an extremely tiny fraction of the kinetic energy, which can 
-%cause numerical problems when one is interested in just the thermal energy of the gas, since
-% Equation~\ref{enzoenergy} solves for the total energy.  In this system of
-%equations, the
-% thermal energy $E_{therm}$ is calculated as $E - E_{kin}$, 
-%where E is the total specific energy as
-%calculated in equation~\ref{enzoenergy} and $E_{kin}$ is the specific kinetic energy,
-%$0.5 \vecv_b^2$.  
-%In hypersonic flows $E$ and $E_{kin}$ are nearly the same, and any 
-%number calculated as 
-%the difference of these is going to be strongly affected by numerical error.  To 
-%avoid this problem, \enzo\ also solves the internal energy equation in
-%comoving coordinates:
-
-%\begin{equation}
-%\frac{\partial e}{\partial t} + \frac{1}{a} \vecv_b \cdot \grad e = 
-%- \frac{3(\gamma-1)\dot{a}}{a}e - \frac{p}{a \rho} \div \vecv_b
-%\label{enzointenergy}
-%\end{equation}
-
-%In this equation $e$ is the internal energy (defined as $P/\rho (\gamma-1)$ and the other terms are as described
-%previously.  The code still conserves total energy ($E$) as well.  In order
-%to maintain consistency, both equations are solved at all times in all cells,
-%with the equation for the total energy (eqtn.~\ref{enzoenergy}) being used
-%for hydrodynamics routines and the internal energy $e$ being used when 
-%temperature is required.  When pressure is required for dynamic purposes,
-%the total energy is used if the ratio of thermal energy to total energy
-%is less than some threshold value $\eta$, and the internal energy is used
-%for values of the ratio larger than $\eta$.  A typical value of this parameter
-%is $10^{-3}$.  This \emph{dual energy formulation} ensures that the
-%method produces the correct entropy jump at strong shocks and also
-%yields accurate pressures and temperatures in cosmological hypersonic
-%flows.
+a particular interface, the more diffusive HLL Riemann solver is used to compute the fluxes associated with that cell.

### 2013-06-08 numerical-ppm.tex
More changes to PPM - cleaned up and made more general to be useful for other Godunov methods.

@@
+In this section, we describe the four solvers that we have implemented for
+solving the fluid equations.  We describe the PPM method in considerably
+more detail than the other methods in part
+because its implementation in Enzo has not previously been described,
+but mostly because it introduces many of the ideas and methods used for
+the MUSCL-Dedner and MHDCT schemes (including expansion terms,
+reconstruction, Riemann solvers, and dual energy formalism);
+
 \subsection{Hydrodynamics: The PPM hydro method}
@@
 
-%We can examine the structure of Equations~(\ref{eq:mass})-(\ref{eq:total_energy})
-%to obtain a better understanding of the effect of each term.
-%The derivatives are first order in time
-%and provide us with the dependent variables that will be evaluated each
-%timestep: $\rhob$, $\vecvb$, and $E$.
-%
-%The second term on the left hand side of each relation
-%is due to our choice of Eulerian coordinates and corresponds to
-%advection, the flow of conserved quantities 
-%($\rhob$, $\rhob \vecvb$, and $\rhob E$) 
-%across the mesh.  These will be termed the transport terms.
-
 The first term on the right hand side of equations~(\ref{eq:momentum}) and
@@
 although we have not done so here, as they do not constitute a major
-source of error.  
-We solve these terms (for all hydro methods)
+source of error.  We solve these terms
 using the same technique -- they are split from the rest of the terms and
 solved using an (implicit) time-centered method, which is straightforward
-as there are no spatial gradients.
-
-%
-%The remainder of the right-hand side terms (due to pressure and gravity) are denoted as
-%source terms.  We make these distinctions because the 
-%source and transport steps will be solved in
-%a self-consistent fashion by the PPM scheme while the expansion terms,
-%which are spatially localized, are split off and solved in a separate step
-%(as are radiative cooling, heating and other processes, if present).
+as there are no spatial gradients.  Note that we use this method for
+all four of the hydro solvers.
 
@@
 restrict discussion to Cartesian coordinates.
-Once again we use the concept of operator splitting,
-but now apply it spatially.   We can rewrite the one-dimensional (Eulerian) 
+We can dimensionally split the equations and rewrite the one-dimensional (Eulerian) 
 versions of equations~(\ref{eq:mass})--(\ref{eq:total_energy}) without expansion terms
@@
 \end{eqnarray}
-
 %
-Note that $x$ and $v$ refer to the one dimensional comoving position and 
+Here $x$ and $v$ refer to the one dimensional comoving position and 
 peculiar velocity of the baryonic gas, and $g$ is the acceleration at the cell center.
-We write this including the cosmological scale factors, but it reduces to the non-cosmological case with $a = 1$.
 These equations are now in a form that can be solved by the split PPM scheme.
@@
 right states, and finally the fluxes are computed based on the
-solution to this Riemann problem.  This is described in detail in
+solution to this Riemann problem and the conserved quantities are
+updated.  This is described in detail in
 \citet{1984JCoPh..54..174C}, but we will briefly outline the procedure
@@
              \right)
-     + \frac{ \Delta }{2} g_j^{n+1/2} (\rho_j^n + \rho_j^{n+1}), 
+     + \frac{ \Delta t }{2} g_j^{n+1/2} (\rho_j^n + \rho_j^{n+1}), 
      \label{eq:momentum_diff}
@@
 equation~(\ref{eq:eq_of_state}), the equation of state.
-%We have dropped the subscript b in this and the following two sections because we will be referring solely to baryonic quantities.  
 The interpolation formula for some quantity $q$ is given by:
@@
 %
-This is Equation~(1.4) of CW84 (in the spatial rather than mass coordinate as is appropriate for the direct Eulerian approach). 
-%The zones edges in the mass coordinates are $m_{j-1/2}$ and $m_{j+1/2}$ for the left and right edges, respectively.  
+This is equation~(1.4) of CW84 (in the spatial rather than mass coordinate as is appropriate for the direct Eulerian approach). 
 The quantities $q_{L,j}$, $\Delta q_j$,
@@
 Once we have the reconstruction, the primary quantities ($p$, $\rho$, $v$ and $E$) are averaged over the domains corresponding to the three characteristics $C = u-c$, $u$ or $u+c$ (where $c$ is the sound speed in a cell):
-\begin{eqnarray}
-      q^{C+}_{j+1/2} & = &
-          {1 \over \Delta t C^n_j} \int^{x_{j+1/2}}_{x_{j+1/2}-\Delta t C^n_j} q_j(x) dx, \\
-      q^{C-}_{j+1/2} & = & 
-         {1 \over -\Delta t C^n_{j+1}} \int^{x_{j+1/2}+\Delta t C^n_{j+1}}_{x_{j+1/2}} q_{j+1}(x) dx \nonumber
-\end{eqnarray}
-where $q^{C+}$ ($q^{C-}$) refers to the left (right) characteristic $C$.
+\begin{equation}
+      q^{C+}_{j+1/2} =
+          {1 \over \Delta t C^n_j} \int^{x_{j+1/2}}_{x_{j+1/2}-\Delta t C^n_j} q_j(x) dx
+\end{equation}
+where $q^{C+}$ refers to the left characteristic $C$ (and a similar integral is done for the right characteristic $q^{C-}$).
 The linearized gas dynamics equations are then used to compute second-order accurate left and right states that take into account the multiple wave families.
@@
 With these effective states, an approximation to the Riemann problem is found
-with an iterative approach \citep[see][and section~\ref{sec.riemann}]{Woodward86}, producing estimates for 
+(see below for more detail about the Riemann solvers used), producing estimates for 
 $\overline{p}_{j\pm 1/2}$, $\overline{\rho}_{j\pm1/2}$, and $\overline{v}_{j\pm 1/2}$ that are third order
@@
 
-While most numerical hydrodynamics schemes center around finding approximate
-solutions to the hydrodynamics equations, the core philosophy of a Godunov
-method \citep{Godunov1959} is to find the exact solution to an
-approximation of the physical setup.  The approximate problem in this case is
-the Riemann problem, wherein two volumes are separated by a thin membrane that
-is instantaneously removed at time $t=0$.  The subsequent evolution has an exact 
-analytic solution.  This solution is
+In this section, we describe the methods we adopt to solving the Riemann
+problem, which is generally required to compute the fluxes in 
+any Godunov.  This section therefore applies to all three of our 
+Godunov-based schemes.  The Riemann problem we are solving
+involves two constant states separated by a single discontinuity at $t=0$.
+The subsequent evolution has an exact analytic solution.  This solution is
 described in detail in many texts on computational fluid dynamics
@@
 
-On rare occasions, high order
-solutions can cause negative densities or energies.  Both PPM and MUSCL solvers employ a Riemann
+On rare occasions, high order solutions can cause negative densities or energies.  
+Both our PPM and MUSCL solvers employ a Riemann
 solver fallback mechanism \citep{Lemaster09}.  If a negative density is found at
-a particular interface, the more diffusive HLL Riemann solver is used to compute the fluxes associated with that cell.
+a particular interface, the more diffusive HLL Riemann solver is used to compute the fluxes associated with that cell, and the flux update is repeated.

### 2013-06-08 numerical-mhd-dedner.tex
First changes to MHD Dedner method.

@@
 
-The equations of compressible inviscid hydrodynamics can be written in the form of conservation laws,
+The second method we describe is a MUSCL-based solver than can be used in both HD and MHD modes.
+To more clearly include the Dedner divergence cleaning modifications, we write the equations of compressible inviscid hydrodynamics can be written in the form of conservation laws as,
 \begin{equation}
@@
 For the generalized Lagrange multiplier formulation of the MHD
-equations \citep{}{}{2002JCoPh.175..645D}, we consider these 
+equations \citep{2002JCoPh.175..645D}, we consider these 
 conserved variables
@@
 where $c_h$ is a constant controlling the propagation speed and
-damping rate of $\div B$. $h=e+p/\rho$ denotes the enthalpy.
+damping rate of $\div B$, and $h=e+p/\rho$ denotes the enthalpy.
+All quantities are cell-centered.
 
-As a reconstruction scheme we here implemented only the simple
-piecewise linear reconstruction. And a number of approximate Riemann
+Much like the PPM solver, we have three basic steps: the first is
+reconstruction of the variables, the second is a solution of the
+Riemann problem, and the third is is updating the conserved quanties
+with the fluxes as written above.
+
+For the reconstruction scheme we here implemented only the simple
+piecewise linear reconstruction \citep{
+And a number of approximate Riemann
 solvers.

### 2013-06-08 numerical-mhd-dedner.tex
Fixed some refs

@@
 For the reconstruction scheme we here implemented only the simple
-piecewise linear reconstruction \citep{
+piecewise linear reconstruction \citep{1979JCoPh..32..101V}.
 And a number of approximate Riemann

### 2013-06-08 numerical-mhd-ct.tex
Finished up my edits on the fluid equations.

@@
 
-This MHD method was developed by \citet{Collins10} in order to study Population
+The third solver we describe is an MHD method developed by 
+\citet{Collins10} in order to study Population
 III stars \citep{Xu08}, molecular clouds \citep{Collins12a}, and galaxy clusters
 \citep{Xu12, Skillman13}.  A full description and suite of test problems can be
-found in \citet{Collins10}
+found in \citet{Collins10} and therefore we only describe the method briefly.
 
@@
 curl of a vector, and the divergence of the curl of a vector is identically
-zero:
-\begin{eqnarray}
-\partial \Bvec/\partial t & =& \frac{1}{a} \curl \Evec\\
-\Divb & =& 0.
-\end{eqnarray}
-
+zero.
+%\begin{eqnarray}
+%\partial \Bvec/\partial t & =& \frac{1}{a} \curl \Evec\\
+%\Divb & =& 0.
+%\end{eqnarray}
 The Constrained Transport (CT) method \citep{Evans88, Balsara99} for
 magnetohydrodynamic (MHD) evolution employs this same vector property
-to evolve the magnetic field in a manner that preserves $\Divb=0$ to
-limits of machine precision.  The electric field is computed using
+to evolve the magnetic field in a manner that preserves $\Divb=0$.  
+The electric field is computed using
 the fluxes from the Riemann solver.  The curl of that electric field
 is then used to update the magnetic field.  The advantage of this
-method is that it preserves \Divb\ to machine precision.  The drawback
-is twofold: increased algorithm complexity, as will be described in
-this section; and since only the update of the magnetic field is
-divergence free, any monopoles created by other numerical sources
-persist.
+method is that it preserves \Divb\ to machine precision.  The primary
+drawback is increased algorithm complexity.  Note also that since 
+only the update of the magnetic field is divergence free, any monopoles 
+created by other numerical sources persist.
 
-\enzo\ MHD-CT has been described in great detail in \citet{Collins10},
-and we direct the interested reader to that reference.  The base
-Godunov method, described in \citet{Li08a}, uses spatially and
+The base Godunov method, described in \citet{Li08a}, uses spatially and
 temporally second order reconstruction (both MUSCL-Hancock and
-Piecewise Linear Method), a selection of Riemann solvers including
-HLLC and HLLD \citep{Mignone07}.  The constrainted transport methods
+Piecewise Linear Method), and a selection of Riemann solvers including
+HLLC and HLLD \citep{Mignone07}, as described earlier.  
+The constrained transport methods
 are the first-order method described by \citet{Balsara99} and the
@@
 
-Cosmological MHD can be treated in a number of ways.  As described in Section
-\red{reference}, the MUSCL with Dedner treats the cosmological dilution of
-$\vecB$ with a source term, 
-$$ 
-\frac{\partial \vecB}{\partial t} - \frac{1}{a}  \curl (\vecv \times
-\vecB)  =  -\frac{\dot{a}}{2 a} \vecB. \label{eq:induction_cosmo_source}
-$$
-However, one can eliminate the right hand side of Equation 
-\ref{eq:induction_cosmo_source} by defining
+Cosmological MHD can be treated in a number of ways.  As described in 
+the previous section, and equation~\ref{eq:induction} the MUSCL with Dedner
+method treats the cosmological dilution of
+$\vecB$ with a source term,  However, one can eliminate the right hand side of Equation 
+\ref{eq:induction} by defining
 $\vecB^\prime = \vecB / \sqrt{a}.$
-Then Equation \ref{eq:induction_cosmo_source} becomes
+Then equation~\ref{eq:induction} becomes
 $$ 
@@
 
-Dual energy formalism has also been incorporated in two methods, one internal
-energy, similar to that described in Section \ref{sec.hydro.ppm}, and one that
-uses entropy.  The additional equations are
-\begin{eqnarray}
-\end{eqnarray}
-\begin{eqnarray}
-  \frac{\partial S}{\partial t} +\frac{1}{a} \div (S \vecv) & = &  -
-  \frac{ 3 (\gamma - 1) \dot{a}}{a} S \\
-  \frac{\partial \rho e}{\partial t} + \frac{1}{a} \div (\rho e \vecv) &
-  = & -\frac{ 3(\gamma -1 ) \dot{a}}{a} + \frac{p}{a} \div \vecv,
-\end{eqnarray}
-where $e$ is the internal energy and $S=p/\rho^{\gamma-1}$ is the entropy.  If
-the ratio of thermal to kinetic energy, $\eta$, is below a certain value, the
-alternative energy is used.  As in \citet{Li08a}, we use $\eta=0.008$.
+The dual energy formalism has also been incorporated in two possible ways, one internal
+energy, as described in Section \ref{sec.hydro.ppm}, and one that
+uses entropy \cite{TVD93, Collins10}..  
+%The additional equations are
+%\begin{eqnarray}
+%\end{eqnarray}
+%\begin{eqnarray}
+%  \frac{\partial S}{\partial t} +\frac{1}{a} \div (S \vecv) & = &  -
+%  \frac{ 3 (\gamma - 1) \dot{a}}{a} S \\
+%  \frac{\partial \rho e}{\partial t} + \frac{1}{a} \div (\rho e \vecv) &
+%  = & -\frac{ 3(\gamma -1 ) \dot{a}}{a} + \frac{p}{a} \div \vecv,
+%\end{eqnarray}
+%where $e$ is the internal energy and $S=p/\rho^{\gamma-1}$ is the entropy.  If
+%the ratio of thermal to kinetic energy, $\eta$, is below a certain value, the
+%alternative energy is used.  As in \citet{Li08a}, we use $\eta=0.008$.
 
-\enzo's MHD-CT has been used in a variety of contexts, from galaxy
-cluster evolution \citep{Xu11,Xu12}, synthetic observations of radio
-halos \citep{Skillman13}, primordial star formation \citep{Xu08}, and
-present-day star formation \citep{Collins11, Collins12a}.
+%
+%\enzo's MHD-CT has been used in a variety of contexts, from galaxy
+%cluster evolution \citep{Xu11,Xu12}, synthetic observations of radio
+%halos \citep{Skillman13}, primordial star formation \citep{Xu08}, and
+%present-day star formation \citep{Collins11, Collins12a}.
+% GB NOTE: incorporate these elsewhere

### 2013-06-08 numerical-mhd-dedner.tex
Finished up my edits on the fluid equations.

--- 2013-06-08 numerical-mhd-dedner.tex
@@
 The second method we describe is a MUSCL-based solver than can be used in both HD and MHD modes.
-To more clearly include the Dedner divergence cleaning modifications, we write the equations of compressible inviscid hydrodynamics can be written in the form of conservation laws as,
+The description here will be very brief both because the ideas are similar to those described in the previous section, and because this implementation has previously
+been described in more detail elsewhere \citep{WangAbelZhang08, WangAbel09}.
+
+Much like the PPM solver, we have three basic steps: the first is
+reconstruction of the variables, the second is a solution of the
+Riemann problem, and the third is is updating the conserved quanties
+with the fluxes as written above.
+For the reconstruction scheme we have implemented only the simple
+piecewise linear reconstruction \citep{1979JCoPh..32..101V, 1985JCoPh..59..264C},
+with options for both primitive and conservative variable reconstruction.
+The available Riemann solvers are HLL, HLLC, and HLLD, as described earlier.
+
+To more clearly describe the Dedner divergence cleaning modifications, we write the equations of compressible inviscid hydrodynamics can be written in the form of conservation laws as,
 \begin{equation}
@@
 
-Much like the PPM solver, we have three basic steps: the first is
-reconstruction of the variables, the second is a solution of the
-Riemann problem, and the third is is updating the conserved quanties
-with the fluxes as written above.
-
-For the reconstruction scheme we here implemented only the simple
-piecewise linear reconstruction \citep{
-And a number of approximate Riemann
-solvers.
-
 The method is dimensionally un-split in that the fluxes are computed
 for all dimensions first and the conserved quantities are updated in
-one go in contrast to the Strang splitting employed in the Zeus and
-PPM solvers described above.
+one step, in contrast to the Strang splitting employed in the other solvers.
+Also unlike the other schemes, time-integration is done with a second-order
+Runge-Kutta scheme \citep{1988JCoPh..77..439S}.

### 2013-06-08 numerical-zeus.tex
Finished up my edits on the fluid equations.

@@
 
-As an alternative to PPM, \enzo\ also includes an implementation of the finite-difference hydrodynamic algorithm employed in the compressible magnetohydrodynamics code \zeus\ \citep{Stone92a, Stone92b}.
+As an alternative to the previous Godunov methods, \enzo\ also includes an implementation of the finite-difference hydrodynamic algorithm employed in the compressible magnetohydrodynamics code \zeus\ \citep{Stone92a, Stone92b}.
 Fluid transport is solved on a Cartesian grid using the upwind, monotonic advection scheme of \citet{1977JCoPh..23..276V} within a multistep (operator split) solution procedure that is fully explicit in time.
@@
 The artificial viscosity term is added in the source terms as:
-
 \begin{eqnarray}
@@
 \end{eqnarray}
-%where \textbf{v} is the baryon velocity, $\rho$ is the mass density, $p$ is the pressure, $e$ is the internal energy density of the gas and
-
 Here \textbf{Q} is the artificial viscosity stress tensor, which we take to be diagonal with on-axis terms given by $l^2 \rho (\partial v / \partial x)^2$ as proposed by von Neumann \& Richtmyer.
@@
 We then add the artificial viscosity:
-
 \begin{eqnarray}
@@
 \end{eqnarray}
-
 The artificial viscosity coefficient $q_j$ is given by:
@@
 On the other hand, it is much more robust than PPM and is easy to add additional physics.
-We also note that this method solves only the internal energy equation rather than total energy, so the ``dual energy formulation'' discussed in Section~\ref{sec.hydro.ppm} is unnecessary.
+We also note that this method solves only the internal energy equation rather than total energy, so the `dual energy formulation' discussed in Section~\ref{sec.hydro.ppm} is unnecessary.

### 2013-06-08 overview.tex
Fixed some refs
Finished up my edits on the fluid equations.
More edits to fluid section - done?

@@
 
-\mqk{Move this table to the Appendix?}
 \input{table_of_variables_and_constants}
@@
 
-%\red{
-%There is some debate over whether to put in the hydro equations without cosmology, in addition
-%to the equations that are already here.  Would this
-%be more or less clear?  We're leaning towards saying ``when cosmology is turned off,
-% $\dot{a}$ goes to zero and $a$ goes to 1, and the poisson equation changes to blah blah blah'' 
-%and assuming our readers can figure it out for themselves.
-%We also want to emphasize that the heating and cooling is glossed over in this section, 
-%and it's discussed more thoroughly later.
-%} %end red
-% GB: I agree this would be shorter, and maybe it is a better idea, but I think it's nice to be able
-% GB: to include both equations and at the same time make the transformation to comoving coords
-% GB: very clear -- this has engendered a lot of confusion in the past and one of the goals of writing
-% GB this paper is to answer peoples questions before they ask them.
-
-% GB: I commented out the following.  It has now been incorporated into the above discussion
-% GB: (see my comments anove on why I think this is a good way to do it)
-%
-%\enzo\ solves the equations of ideal gas dynamics
-%in a coordinate system that is comoving with the expanding universe:
-
-%\begin{equation}
-%\frac{\partial \rho_b}{\partial t} + \frac{1}{a} \vecv_b \cdot \grad \rho_b =  \label{enzoconserve}
-% - \frac{1}{a} \rho_b \div \vecv_b 
-%\end{equation}
-
-%\begin{equation}
-%\frac{\partial \vecv_b}{\partial t} + \frac{1}{a} (\vecv_b \cdot \grad) \vecv_b = \label{enzomomentum} 
-%-\frac{\dot{a}}{a} \vecv_b - \frac{1}{a \rho_b} \grad p - \frac{1}{a} \grad \phi
-%\end{equation}
-
-%\begin{eqnarray}
-%\frac{\partial E}{\partial t} + \frac{1}{a} \vecv_b \cdot \grad E =  
-%- \frac{\dot{a}}{a} \left( 3 \frac{p}{\rho_b} + \vecv_b^2 \right)  \nonumber \\
-%- \frac{1}{a \rho_b} \div (p \vecv_b) \nonumber \\
-%- \frac{1}{a} \vecv_b \cdot \grad \phi + \Gamma - \Lambda
-%\label{enzoenergy}
-%\end{eqnarray}
-
-%Where $\rho_b$ is the comoving baryon density, $\vecv_b$ is the baryon velocity, $p$ is the pressure, 
-%$\phi$ is the modified gravitational potential (in comoving coordinates, which is related to
-%the potential in proper coordinates $\Phi$ by $\phi \equiv \Phi + 0.5 $a\"{a}$ \vecx^2$) and 
-%$a$ is the ``expansion parameter'' which describes the expansion of a smooth, homogeneous 
-%universe as a function of time.  This expansion parameter is related to the redshift:  
-%$a \equiv 1/(1+z)$.  All derivatives are in comoving coordinates.  $E$ is the specific
-%energy of the gas (total energy per unit mass), and
-%$\Gamma$ and $\Lambda$ represent radiative heating and cooling processes as described below.
-%  Equations~\ref{enzoconserve}, \ref{enzomomentum}
-%and \ref{enzoenergy} represent the conservation of mass, momentum and total (e.g.,
-%kinetic plus thermal) fluid energy.
-
-%The equations above are closed with three more equations:
-
-%\begin{equation}
-%E = p / [(\gamma - 1) \rho_b] + \vecv^2/2   \label{enzoeos}
-%\end{equation}
-
-%\begin{equation}
-%\nabla^2 \phi = \frac{4 \pi G}{a} (\rho_b + \rho_{dm} - \rho_0 )
-%\label{enzopoisson}
-%\end{equation}
-
-%\begin{equation}
-%\frac{\ddot{a}}{a} = - \frac{4 \pi G}{3 a^3} (\rho_0 + 3 p_0 / c^2) + \Lambda/3.
-%\label{enzocomove}
-%\end{equation}
-
-%where $\rho_{dm}$ is the comoving dark matter density, $\rho_0$ is the comoving
-%background density 
-%($\rho_0 \equiv \Omega_{matter} \rho_{crit}$) and $p_0$ is the background pressure, 
-%$\gamma$ is the ratio of specific heats and $\Lambda$ is the cosmological constant.
-%Equations~\ref{enzoeos}, \ref{enzopoisson} and~\ref{enzocomove} are the 
-%equation of state, Poisson's equation in comoving form and an equation that 
-%describes the evolution of the comoving coordinates (i.e. the formula for the 
-%expansion of an isotropic, homogeneous universe). 
-%All particles in the simulation are governed by Newton's equations in comoving 
-%coordinates:
-
-%\begin{equation}
-%\frac{d \vecx_{part}}{dt} = \frac{1}{a} \vecv_{part}
-%\label{enzopartvel}
-%\end{equation}
-
-%\begin{equation}
-%\frac{d \vecv_{part}}{dt} = -\frac{\dot{a}}{a} \vecv_{part} - \frac{1}{a} \grad \phi
-%\label{enzopartmom}
-%\end{equation}
-
-%Where $\vecx_{part}$ and $\vecv_{part}$ refer to the position and peculiar velocity of any
-%particles in the system.  Note that the system of equations~\ref{enzoconserve}-\ref{enzopartmom} 
-%is valid only in regimes where relativistic effects are not
-%important ($v_b, v_{dm} \ll c$, where c is the speed of light) 
-%and where cosmological curvature effects are 
-%unimportant, meaning that the simulation volume is much smaller than the radius
-%of curvature of the universe, as defined as $r_{hub} \equiv c/H_0$, where $c$ is the
-%speed of light and $H_0$ is the Hubble constant.
-
-%\dcc{Fixed punctuation, added 'no mhd, rhd' caveat to Zeus statements.  This has caused
-%confusion in the past.}
-
-%\subsubsection{Magnetohydrodynamics}
-
-%\subsubsection{Radiation transport: ray-tracing}
-
 The code can include either a homogeneous radiation background (with
@@
 by~\citet{1995CoPhC..89..149B}; (ii) the MUSCL-like Godunov scheme
-\citep{MUSCL} with or without magnetic fields and Dedner-based
+\citep{1977JCoPh..23..276V} with or without magnetic fields and Dedner-based
 divergence cleaning, described in more detail in
@@
 
-%The advantage of the PPM method is that it conserves energy better and is higher-order accurate.  The disadvantage is that it is generally less robust than the ZEUS method.
-
-%%% If you prefer the following \paragraph{} sections to be numbered,
-%%% then you can add \setcounter{secnumdepth}{5} to the preamble...
-
 \paragraph{Godunov MUSCL (HD) with Dedner divergence cleaning (MHD)}
@@
 
-\paragraph{MHD with Constrained Transport (MHD)}
+\paragraph{Godunov MHD with Constrained Transport (MHD)}
 This MHD method is second order in time and space, and preserves the

### 2013-06-08 numerical-mhd-ct.tex
More edits to fluid section - done?
Minor change to MHDCT

@@
-\subsection{MHD: Constrained transport}
+\subsection{Magnetohydrodynamics: Constrained transport}
 \label{sec.num.mhd-ct}
@@
 The third solver we describe is an MHD method developed by 
-\citet{Collins10} in order to study Population
-III stars \citep{Xu08}, molecular clouds \citep{Collins12a}, and galaxy clusters
-\citep{Xu12, Skillman13}.  A full description and suite of test problems can be
-found in \citet{Collins10} and therefore we only describe the method briefly.
+\citet{Collins10} and since a full description and suite of test problems can be
+found in that paper, we only describe the method briefly here.
 
@@
 method treats the cosmological dilution of
-$\vecB$ with a source term,  However, one can eliminate the right hand side of Equation 
+$\vecB$ with a source term,  However, one can eliminate the right hand side of equation 
 \ref{eq:induction} by defining

### 2013-06-08 ms.tex
Edits to the overview section. Trying to unify notations, so several other sections are also affected.
A few more edites to section 2. Also unified "time step" -> "timestep" throughout the whole manuscript.
Fixed up PPM section somewhat.  Still needs more work.
Added author list

@@
 
-\usepackage{graphicx, natbib, color}
+\usepackage{graphicx, natbib, color, bm}
 
@@
 %   8.2 Star formation & Feedback
-% 9. Time stepping
+% 9. Timestepping
 % 10.  Analysis
@@
 \author{The Enzo Collaboration:}
-\author{author 1}
-\author{author 2}
-\author{author 3...}
+\author{Greg L. Bryan\altaffilmark{1}}
+\author{Michael L. Norman\altaffilmark{2,3}}
+\author{Brian W. O'Shea\altaffilmark{4,5}}
+\author{Tom Abel\altaffilmark{6, 19}}
+\author{John H. Wise\altaffilmark{7}}
+\author{Matthew J. Turk\altaffilmark{1}}
+\author{Daniel R. Reynolds\altaffilmark{8}}
+\author{David C. Collins\altaffilmark{9}}
+\author{Peng Wang\altaffilmark{6}}
+\author{Samuel W. Skillman\altaffilmark{10,11}}
+\author{Britton Smith\altaffilmark{4}}
+\author{Robert P. Harkness\altaffilmark{12}}
+\author{James Bordner\altaffilmark{2}}
+\author{Ji-hoon Kim\altaffilmark{13}}
+\author{Michael Kuhlen\altaffilmark{14,15}}
+\author{Hao Xu\altaffilmark{2}}
+\author{Nathan Goldbaum\altaffilmark{15}}
+\author{Cameron Hummels\altaffilmark{16}}
+\author{Alexei Kritsuk\altaffilmark{2}}
+\author{Elizabeth Tasker\altaffilmark{17}}
+\author{Stephen Skory\altaffilmark{2}}
+\author{Christine Simpson\altaffilmark{1}}
+\author{Oliver Hahn\altaffilmark{18}}
+\author{Jeffrey S. Oishi\altaffilmark{6}}
+\author{Geoffrey C So\altaffilmark{2}}
+\author{Fen Zhao\altaffilmark{19}}
+\author{Renyue Cen\altaffilmark{20}}
+\author{Yuan Li\altaffilmark{1}}
 
-%\altaffiltext{1}{ Department of Astronomy, Columbia University,
-%New York, NY, 10027, U.S.A.  Email: gbryan@astro.columbia.edu }
-%\altaffiltext{2}{Center for Astrophysics and Space Sciences,
-%University of California at San Diego.
-%Email: mnorman,dcollins,akritsuk,jbordner@cosmos.ucsd.edu}
-%\altaffiltext{3}{Theoretical Astrophysics (T-6), Los Alamos National
-%Laboratories.  Email:  bwoshea@lanl.gov}
-%\altaffiltext{4}{San Diego Supercomputing Center, University of California
-%at San Diego.  Email: harkness@sdsc.edu}
-%\altaffiltext{5}{Kavli Institute for Particle Astrophysics and Cosmology,
-% Stanford University.  Email:  tabel@stanford.edu}
+\altaffiltext{1}{Columbia University, Department of Astronomy, New York, NY, 10025, USA}
+\altaffiltext{2}{CASS, University of California, San Diego, 9500 Gilman Drive La Jolla, CA 92093-0424, USA}
+\altaffiltext{3}{SDSC, University of California, San Diego, 9500 Gilman Drive La Jolla, CA 92093-0505, USA}
+\altaffiltext{4}{Department of Physics and Astronomy, Michigan State University, East Lansing, MI 48824, USA}
+\altaffiltext{5}{Lyman Briggs College and Institute for Cyber-Enabled Research, Michigan State University, East Lansing, MI 48824, USA}
+\altaffiltext{6}{Kavli Institute for Particle Astrophysics and Cosmology, Stanford University, Menlo Park, CA 94025, USA}
+\altaffiltext{7}{Center for Relativistic Astrophysics, School of Physics, Georgia Institute of Technology, 837 State Street, Atlanta, GA, USA}
+\altaffiltext{8}{Southern Methodist University, 6425 Boaz Ln, Dallas, TX 75205, USA}
+\altaffiltext{9}{Theoretical Division, Los Alamos National Laboratory, Los Alamos, NM 87544}
+\altaffiltext{10}{Center for Astrophysics and Space Astronomy, Department of Astrophysical and Planetary Science, University of Colorado, Boulder, CO 80309, USA}
+\altaffiltext{11}{DOE Computational Science Graduate Fellow}
+\altaffiltext{12}{NICS, Oak Ridge National Laboratory, 1 Bethel Valley Rd, Oak Ridge, TN 37831, USA}
+\altaffiltext{13}{Department of Astronomy and Astrophysics, University of California, Santa Cruz, CA 95064, USA}
+\altaffiltext{14}{Theoretical Astrophysics Center, University of California Berkeley, Hearst Field Annex, Berkeley, CA 94720, USA}
+\altaffiltext{15}{Department of Astronomy and Astrophysics, University of California, 1156 High Street, Santa Cruz, CA 95064, USA}
+\altaffiltext{16}{Department of Astronomy/Steward Observatory, University of Arizona, 933 N. Cherry Ave., Tucson, AZ, 85721, USA}
+\altaffiltext{17}{Physics Department, Faculty of Science, Hokkaido University, Kita-10 Nishi 8, Kita-ku, Sapporo 060-0810, Japan}
+\altaffiltext{18}{ETH Zurich Institute for Astronomy. CH-8093 Zurich. Switzerland}
+\altaffiltext{19}{Department of Physics, Stanford University, Stanford, CA 94305, USA}
+\altaffiltext{20}{Department of Astrophysical Sciences, Princeton University, Peyton Hall, Princeton, NJ 08544, USA}
 
@@
 
-% We now describe each of the numerical methods in detail.  The
-% structure of this section mirrors that of Section~\ref{sec.overview},
-% in which each method was briefly described.
-
+% We now describe each of the numerical methods in detail.  
 \input{numerical-amr}
@@
 \input{numerical-ppm}
-%\input{numerical-muscl} 
 \input{numerical-mhd-dedner}  % includes MUSCL and Dedner

### 2013-06-08 test_problems/AnisoConduction/make_plots.py
Adding .hgignore for test outputs and latex files.

@@ -1,5 +1,8 @@
 from yt.mods import *
 
+filenames = {0: 'aniso_conduction_initial_output.eps',
+             11: 'aniso_conduction_final_output.eps'}
+
 for n in [0,11]:
     
     pf = load("DD%04i/DD%04i" % (n,n))
@@ -13,30 +16,15 @@ for n in [0,11]:
     print dd["Temperature"].min()
     print dd["Temperature"].max()
 
-    #print "Density:"
-    #print dd["Density"].min()
-    #print dd["Density"].max()
+    slice = SlicePlot(pf, 'z', "Temperature", width = (1.05, '1'), origin='domain', axes_unit=('1','1'))  #.save()
 
-    slice = SlicePlot(pf, 'z', "Temperature", width = (1.0, '1'), origin='domain', axes_unit=('1','1'))  #.save()
-    #slice = SlicePlot(pf, 'z', "Temperature", width = (1.0, '1'), origin='domain', axes_unit=None)  #.save()
-
-
-    x = (dd['x'].reshape((256,256)))[:,0]
-    y = (dd['y'].reshape((256,256)))[0,:]
-    Bx = dd['Bx'].reshape((256,256))
-    By = dd['By'].reshape((256,256))
-    ax = slice.plots['Temperature'].axes
-
-    ## I found that I had to reverse Bx and By to get the correct
-    ## closed circles. Something to do with the plot's origin being in
-    ## the lower left?
-    streams = ax.streamplot(x,y,By,Bx,color=(1,1,1,0.5),linewidth=1.5,density=1)
-
-    ax.set_rasterization_zorder(5.0)
+    slice.annotate_streamlines('Bx', 'By', plot_args={'color':(1,1,1,0.5),'linewidth':1.5})
     
-    ## filename = "DD%04i_slice_Temperature_z.png" % n
-    filename = "DD%04i_slice_Temperature_z.eps" % n
+    slice.set_zlim('Temperature', 7e5, 2e7)
 
-    slice.save(mpl_kwargs={'bbox_inches':'tight'}, name=filename)
+    # See matplotlib issue #1188
+    slice.plots['Temperature'].cb.solids.set_edgecolor('face')
+    
+    slice.save(name=filenames[n], mpl_kwargs={'bbox_inches':'tight'})
 
 # the end

### 2013-06-08 test_problems/test_problem_runner.py
Using __file__ to avoid having to set one of the paths.  Adjusting line fill for some of the comments.

@@ -9,10 +9,9 @@ have yt (http://yt-project.org) built with SciPy, which requires a
 non-default option to be set in the install script.  If you do NOT
 compile Enzo with the HYPRE library, or do not have SciPy installed
 along with yt, remove the 'CosmoIFront' problem from the testnames
-list below.  You also need to set two paths -- the location of the
-Enzo binary and the path to the test problem directory -- and you
-can turn on and off some options.  This is all done directly below
-this comment block.
+list below.  You also need to set a path -- the location of the
+Enzo binary -- and you can turn on and off some options.  This is
+all done directly below this comment block.
 
 Note: some of the simulations used to create the method paper figures
 can take several hours to run, or should be run on several CPUs. See
@@ -31,16 +30,18 @@ import os as os
 enzodir = '/Users/bwoshea/Desktop/SYNCHED/Current\ Projects/Papers/enzo-method-paper/test_problems'
 
 # directory where this file is located (needs absolute path name)
-topleveldir = '/Users/bwoshea/Desktop/SYNCHED/Current Projects/Papers/enzo-method-paper/test_problems'
+topleveldir = os.path.dirname(os.path.realpath(__file__))
 
-# the following options allow you to articulate whether you want the simulations to be run (runsim),
-#   figures to be made (makefig), and directories to be cleaned out (cleandata/cleanfigs).  Note that 
-#   this is so you can separately run the various simulations, make the figures, and/or clean out all
-#   of the various subdirectories.
+# the following options allow you to articulate whether you want the simulations
+#   to be run (runsim), figures to be made (makefig), and directories to be
+#   cleaned out (cleandata/cleanfigs).  Note that this is so you can separately
+#   run the various simulations, make the figures, and/or clean out all of the
+#   various subdirectories.
 
-# if cleandata == 1, delete all simulation files in the directory.  If 0, don't touch them.
-#    Note: output log is always deleted before a simulation is run.  Also, if you clean out
-#    the data, you can't make figures without rerunning the simulations!
+# if cleandata == 1, delete all simulation files in the directory.  If 0, don't
+#    touch them.  Note: output log is always deleted before a simulation is run.
+#    Also, if you clean out the data, you can't make figures without rerunning
+#    the simulations!
 cleandata = 1
 
 # if cleanfigs == 1, delete all figures in the directory.  If 0, don't touch them.

### 2013-06-08 introduction.tex

@@
 galaxies \citep{2003ApJ...587...13T, 2012ApJ...749..140H}, galaxy
-clusters \citep{Loken02, Skillman13}, the interstellar medium
+clusters \citep{Loken02, Xu11, Skillman13}, the interstellar medium
 \citep{Slyz05}, the intergalactic medium \citep{Fang01,
@@
 formation of the first stars \citep{ABN02, 2007ApJ...654...66O,
-2009Sci...325..601T}, and the formation of stars in our own Galaxy
-\citep{Collins12a}.
+2009Sci...325..601T, Xu08}, and the formation of stars in our own Galaxy
+\citep{Collins11, Collins12a}.
 

### 2013-06-08 tests-aniso-conduct.tex
Slight adjustment for the aniso conduction figure size.

@@
 \includegraphics[width=0.42\textwidth]{figures/aniso_conduction_initial_output.eps}
-\includegraphics[width=0.4\textwidth]{figures/aniso_conduction_final_output.eps}
+\includegraphics[width=0.42\textwidth]{figures/aniso_conduction_final_output.eps}
 \caption{Two-dimensional anisotropic conduction test in a uniform,
 constant temperature background with circular magnetic fields centered
-on (0.5, 0.5) in a $256 \times 256$ grid. \mqk{Maybe show some field
-lines in the figure?} The background medium has a density of 1
-proton/cc and a temperature of $10^6$~K.  At t$ = 0$ (left panel), a
-Gaussian heat pulse is injected at (0.75, 0.5) with a FWHM of $1/64$
-(with all numbers given in units of the box size) and a peak
-temperature of $10^8$~K, and allowed to evolve without hydrodynamical
-motion (i.e., static gas) and no radiative cooling for 300 Myr.  At t$
-= 300$~Myr (right panel), heat has been transported along magnetic
-field lines with no significant diffusion perpendicular to field
-lines. Furthermore, there are no detectable oscillations in the
-temperature in regions where the magnetic field is not parallel with
-the grids.}
+on (0.5, 0.5) in a $256 \times 256$ grid. The background medium has a
+density of 1 proton/cc and a temperature of $10^6$~K.  At t$ = 0$
+(left panel), a Gaussian heat pulse is injected at (0.75, 0.5) with a
+FWHM of $1/64$ (with all numbers given in units of the box size) and a
+peak temperature of $10^8$~K, and allowed to evolve without
+hydrodynamical motion (i.e., static gas) and no radiative cooling for
+300 Myr.  At t$ = 300$~Myr (right panel), heat has been transported
+along magnetic field lines with no significant diffusion perpendicular
+to field lines. Furthermore, there are no detectable oscillations in
+the temperature in regions where the magnetic field is not parallel
+with the grids.}
 \label{fig.conduct}

### 2013-06-08 tests-aniso-conduct.tex
Clarifying the streamlines in the aniso conduction plot.

@@
 \includegraphics[width=0.42\textwidth]{figures/aniso_conduction_final_output.eps}
-\caption{Two-dimensional anisotropic conduction test in a uniform,
-constant temperature background with circular magnetic fields centered
-on (0.5, 0.5) in a $256 \times 256$ grid. The background medium has a
-density of 1 proton/cc and a temperature of $10^6$~K.  At t$ = 0$
-(left panel), a Gaussian heat pulse is injected at (0.75, 0.5) with a
-FWHM of $1/64$ (with all numbers given in units of the box size) and a
-peak temperature of $10^8$~K, and allowed to evolve without
-hydrodynamical motion (i.e., static gas) and no radiative cooling for
-300 Myr.  At t$ = 300$~Myr (right panel), heat has been transported
-along magnetic field lines with no significant diffusion perpendicular
-to field lines. Furthermore, there are no detectable oscillations in
-the temperature in regions where the magnetic field is not parallel
-with the grids.}
+\caption{Two-dimensional anisotropic conduction test in a uniform, constant
+temperature background with circular magnetic fields (indicated by white
+streamlines) centered on (0.5, 0.5) in a $256 \times 256$ grid. The background
+medium has a density of 1 proton/cc and a temperature of $10^6$~K.  At t$ = 0$
+(left panel), a Gaussian heat pulse is injected at (0.75, 0.5) with a FWHM of
+$1/64$ (with all numbers given in units of the box size) and a peak
+temperature of $10^8$~K, and allowed to evolve without hydrodynamical motion
+(i.e., static gas) and no radiative cooling for 300 Myr.  At t$ = 300$~Myr
+(right panel), heat has been transported along magnetic field lines with no
+significant diffusion perpendicular to field lines. Furthermore, there are no
+detectable oscillations in the temperature in regions where the magnetic field
+is not parallel with the grids.}
 \label{fig.conduct}

### 2013-06-08 test_problems/OrszagTang/make_plots.py
Updating the Orszag-Tang figures.

@@ -3,9 +3,13 @@ name1 = './CT/DD0056/data0056'
 name2 = './Dedner/DD0032/data0032'
 pf1 = load(name1)
 pf2 = load(name2)
-names=['mhdct_0056','mhddedner_0035']
-sl=[]
+names=[
+    'MHDCT_OrszagTang_Density.eps',
+    'MHDDedner_OrszagTang_Density.eps'
+    ]
 for i, pf in enumerate([pf1,pf2]):
-  sl.append( SlicePlot( pf, 2, 'Density') )
-  sl[-1].set_zlim('Density',0.1,1)
-  print sl[-1].save(names[i]+".eps")
+  sl = SlicePlot( pf, 2, 'Density', origin='domain', width=(1.05, 'unitary'))
+  sl.set_zlim('Density',0.1,1)
+  # See matplotlib issue #1188
+  sl.plots['Density'].cb.solids.set_edgecolor('face')
+  print sl.save(names[i])

### 2013-06-08 test_problems/DoubleMachReflection/make_plots.py
Updating the DoubleMach figures.

@@ -6,11 +6,12 @@ for n in range(11,12):
 
     c = [2.0,0.5,0]
 
-    buff = [120,480]  # this is backward because matplotlib transposes buffers.
+    slice = SlicePlot(pf, 'z', "Density", center=c, width=((4.15, '1'), (1.15, '1')),
+                      origin='domain', axes_unit=('1','1'))
 
-    slice = SlicePlot(pf, 'z', "Density", center=c, width=((4.0, '1'), (1.0, '1')), origin='domain', axes_unit=('1','1')) 
-    slice.set_buff_size(buff)
+    slice.set_zlim('Density',.5, 20)
 
-    filename = "DD%04i" % n
+    # See matplotlib issue #1188
+    slice.plots['Density'].cb.solids.set_edgecolor('face')
 
-    slice.save(mpl_kwargs={'bbox_inches':'tight'},name="%s_slice_Density_z.eps" % filename)
+    slice.save(mpl_kwargs={'bbox_inches':'tight'},name="DoubleMachTest.eps")

### 2013-06-09 test_problems/SedovAMR/sedov_slice.py
Updating the sedov blast plots.

@@ -4,9 +4,17 @@ from yt.mods import *
 
 fn1 = "PPM/DD0007/DD0007"
 pf1 = load(fn1)
-slice1 = SlicePlot(pf1, 'z', "Density", center=[0.5, 0.5, 0.0]).save("PPM_Density_slice.eps")
+slice1 = SlicePlot(pf1, 'z', "Density", center=[0.5, 0.5, 0.0], origin='domain')
+slice1.set_width((1.05,'unitary'))
+# See matplotlib issue #1188
+slice1.plots['Density'].cb.solids.set_edgecolor('face')
+slice1.save("sedov-ppm-slice.eps", mpl_kwargs={'bbox_inches':'tight'})
 
 fn2 = "Zeus/DD0007/DD0007"
 pf2 = load(fn2)
-slice2 = SlicePlot(pf2, 'z', "Density", center=[0.5, 0.5, 0.0]).save("Zeus_Density_slice.eps")
+slice2 = SlicePlot(pf2, 'z', "Density", center=[0.5, 0.5, 0.0], origin='domain')
+slice2.set_width((1.05,'unitary'))
+# See matplotlib issue #1188
+slice2.plots['Density'].cb.solids.set_edgecolor('face')
+slice2.save("sedov-zeus-slice.eps", mpl_kwargs={'bbox_inches':'tight'})
 

### 2013-06-09 test_problems/CosmoIFront/make_plots.py
Big figure update.  Fonts are now more consistent.
Many small adjustments as well.  TestOrbit has seen a big change - we now use
plt.scatter() to generate the plot.  If you zoom in, you can see the individual
points and the small scatter from a perfectly circular orbit.

My results for SelfSimilarInfall do not match what's in the paper so I've left
it alone.  This might have something to do with the gravity bug Greg has fixed
in his branch.

Still need to update GravityTest.

@@ -2,6 +2,9 @@
 # Daniel R. Reynolds, reynolds@smu.edu
 
 # imports
+import matplotlib as mpl
+mpl.rcParams['font.family'] = 'STIXGeneral'
+
 from pylab import *
 
 
@@ -252,14 +255,13 @@ figure()
 plot(Hradii1, HIprof1, 'b-',  Hradii1, HIIprof1, 'r-',
      Hradii2, HIprof2, 'b--', Hradii2, HIIprof2, 'r--',
      Hradii3, HIprof3, 'b-.', Hradii3, HIIprof3, 'r-.')
-grid()
-xlabel('$r/L_{box}$')
-ylabel('log(xHI), log(xHII)')
-title('HI, HII Fraction Profiles')
-legend( ('xHI z=6.24','xHII z=6.24','xHI z=2.29','xHII z=2.29','xHI z=1.02','xHII z=1.02'), 
-        loc=4 )
-axis([ 0.0, 1.0, -7.0, 1.0 ])
-savefig('profiles' + pictype)
+xlabel('$r/L_{\mathrm{box}}$')
+ylabel('$\log(x_{\mathrm{HI}}),\/\log(x_{\mathrm{HII}})$')
+ylim(-6.5, 0.5)
+xlim(0,1) 
+legend( ('$x_{\mathrm{HI}}\/z=6.24$','$x_{\mathrm{HII}}\/z=6.24$','$x_{\mathrm{HI}}\/z=2.29$',
+         '$x_{\mathrm{HII}}\/z=2.29$','$x_{\mathrm{HI}}\/z=1.02$','$x_{\mathrm{HII}}\/z=1.02$'),loc=4 )
+savefig('FLDprofiles' + pictype, bbox_inches='tight')
 
 
 
@@ -278,12 +280,11 @@ z_ratio2 = (1.0 + rdata[2][2:te+1])/(1.0+zi)
 figure()
 xdata = -log10(z_ratio)
 plot(xdata,r_ratio,'b-',xdata,ranal_ratio,'r--')
-xlabel('$-log[(1+z)/(1+z_i)]$')
+xlabel('$-\log[(1+z)/(1+z_i)]$')
 ylabel('$r_I/r_S$')
-title('Scaled I-front Position vs Redshift')
+xlim(-.05,.85)
+ylim(-.05,1.05)
 legend( ('computed', 'analytical'), loc=4 )
-grid()
-axis([ 0.0, 1.0, 0.0, 1.0 ])
[…]

### 2013-06-09 test_problems/OneZoneFreefallTest/make_plots.py

@@ -1,6 +1,9 @@
 from collections import defaultdict
+import matplotlib as mpl
+mpl.rcParams['font.family'] = 'STIXGeneral'
 from matplotlib import pyplot
 import numpy as np
+import cPickle as pickle
 import sys
 
 from yt.mods import *
@@ -19,14 +22,22 @@ def get_tracks(filenames, fields=None):
     return data
 
 dirs = ['mf', 'Z-5', 'Z-4', 'Z-3', 'Z-2']
-labels = ['0', '10$^{-5}$',  '10$^{-4}$',  '10$^{-3}$',  '10$^{-2}$']
+labels = ['$Z=0$', '10$^{-5}\/Z_\odot$',  '10$^{-4}\/Z_\odot$',
+          '10$^{-3}\/Z_\odot$',  '10$^{-2}\/Z_\odot$']
 colors = ['black', 'blue', 'green', 'orange', 'red']
 
-my_data = {}
-for my_dir in dirs:
-    my_data[my_dir] = get_tracks("%s/DD*/*.hierarchy" % my_dir)
+if os.path.exists('data.p'):
+    with open('data.p', 'rb') as fp:
+        my_data = pickle.load(fp)
+else:
+    my_data = {}
+    for my_dir in dirs:
+        my_data[my_dir] = get_tracks("%s/DD*/*.hierarchy" % my_dir)
 
-fontsize = 14
+    with open('data.p', 'wb') as fp:
+        pickle.dump(my_data, fp)
+    
+fontsize = 12
 
 n_rows = 1
 n_columns = 1
@@ -67,4 +78,4 @@ tick_labels = my_axes.xaxis.get_ticklabels() + \
 for tick_label in tick_labels:
     tick_label.set_size(fontsize)
 my_axes.legend(loc='lower right', prop=dict(size=fontsize))
-pyplot.savefig('n-T.eps')
+pyplot.savefig('OneZoneCollapseTest.eps')

### 2013-06-09 test_problems/SedovAMR/sedov_slice.py

@@ -8,7 +8,9 @@ slice1 = SlicePlot(pf1, 'z', "Density", center=[0.5, 0.5, 0.0], origin='domain')
 slice1.set_width((1.05,'unitary'))
 # See matplotlib issue #1188
 slice1.plots['Density'].cb.solids.set_edgecolor('face')
+slice1.set_font({'size':15})
 slice1.save("sedov-ppm-slice.eps", mpl_kwargs={'bbox_inches':'tight'})
+
 
 fn2 = "Zeus/DD0007/DD0007"
 pf2 = load(fn2)

### 2013-06-09 test_problems/SelfSimilarInfall/make_plots.py

@@ -1,3 +1,5 @@
+import matplotlib as mpl
+mpl.rcParams['font.family'] = 'STIXGeneral'
 from yt.mods import *
 import pylab
 
@@ -29,7 +31,6 @@ exact = pylab.csv2rec( exact_solution_filename, delimiter=' ', names=('lambda', 
 
 # open output
 pf = load(filename)
-
 
 # Compute radial profile
 r_min = pf.h.get_smallest_dx()*pf.units["cm"]

### 2013-06-09 test_problems/ShockPool/make_plots.py

@@ -1,3 +1,5 @@
+import matplotlib as mpl
+mpl.rcParams['font.family'] = 'STIXGeneral'
 from yt.mods import *
 import pylab
 
@@ -100,4 +102,4 @@ for i in range(3):
 
 
 ### Save plot
-pylab.savefig(plot_filename)
+pylab.savefig(plot_filename, bbox_inches='tight')

### 2013-06-09 test_problems/ShockTube/make_plots.py

@@ -1,3 +1,5 @@
+import matplotlib as mpl
+mpl.rcParams['font.family'] = 'STIXGeneral'
 from yt.mods import *
 import pylab
 

### 2013-06-09 test_problems/TestOrbit/make_plots.py

@@ -5,6 +5,8 @@
 #   (since we need some stuff from output.log)
 #
 import numpy as na
+import matplotlib as mpl
+mpl.rcParams['font.family'] = 'STIXGeneral'
 import matplotlib.pyplot as plt
 import os
 from math import *
@@ -78,7 +80,7 @@ en = na.array(energy)/energysum
 
 
 # plots orbit in x-y plane (should be a circle)
-plt.plot(x[::2], y[::2])
+plt.scatter(x[::8], y[::8], s=1, lw=0, marker='.')
 plt.axis("equal")
 plt.xlabel('x')
 plt.ylabel('y')
@@ -87,7 +89,7 @@ plt.close()
 
 # plots orbit in x-z plane (should be a straight line)
 # NOT in paper.
-plt.plot(x[::2], z[::2])
+plt.scatter(x[::8], z[::8], s=1, lw=0, marker='.')
 plt.xlim(0.1,0.9)
 plt.ylim(0.4999999,0.5000001)
 plt.xlabel('x')
@@ -189,9 +191,9 @@ plt.xlabel('Orbit Number')
 plt.ylabel('Total Energy')
 plt.xlim(1,200)
 plt.ylim(-1.78, -1.75)
-plt.figtext(0.17, 0.82, '$<TE> = -1.76797$', size=15)
-plt.figtext(0.17, 0.77, '$\sigma_{TE} = 0.004885$', size=15)
-plt.figtext(0.17, 0.72, '$\sigma_{TE}/|<TE>| = 0.002763$', size=15)
-plt.savefig('TotalEnergy.eps')
+plt.figtext(0.17, 0.82, '$<\mathrm{TE}> = -1.76797$', size=15)
+plt.figtext(0.17, 0.77, '$\sigma_{\mathrm{TE}} = 0.004885$', size=15)
+plt.figtext(0.17, 0.72, '$\sigma_{\mathrm{TE}}/|<\mathrm{TE}>| = 0.002763$', size=15)
+plt.savefig('TestOrbit_TotalEnergy.eps')
 plt.close()
 

### 2013-06-09 test_problems/WavePool/make_plots.py

@@ -1,3 +1,5 @@
+import matplotlib as mpl
+mpl.rcParams['font.family'] = 'STIXGeneral'
 from yt.mods import *
 import pylab
 
@@ -49,6 +51,9 @@ def plot_one(filename, name, axes, xbegin1, xwidth):
 
     # plot
     ax = pylab.axes(axes)
+    x1 = pylab.array([0.25, 0.75])
+    y1 = pylab.array([1.0, 1.0])
+    ax.fill_between(x1, y1-0.01, y1+0.01, facecolor='grey', alpha=0.2, zorder=1)
     pylab.axhline(0,color='k',linestyle='dotted')
     pylab.scatter(ray['x']+0.5*ray['dx'],ray['Density'], c=level, s=40, linewidths=(0,))
     pylab.plot(xexact, exact)
@@ -56,9 +61,6 @@ def plot_one(filename, name, axes, xbegin1, xwidth):
     #plot dashed lines for refined region
 #    pylab.plot([0.25, 0.25], [0.9, 1.1], linestyle="--") 
 #    pylab.plot([0.75, 0.75], [0.9, 1.1], linestyle="--") 
-    x1 = pylab.array([0.25, 0.75])
-    y1 = pylab.array([1.0, 1.0])
-    ax.fill_between(x1, y1-0.01, y1+0.01, facecolor='grey', alpha=0.2)
 
     #set boundaries
     pylab.axis([xbegin1,xbegin1+xwidth,0.99,1.009999])
@@ -100,4 +102,4 @@ for i in range(3):
 
 
 ### Save plot
-pylab.savefig(plot_filename)
+pylab.savefig(plot_filename, bbox_inches='tight')

### 2013-06-09 test_problems/ZeldovichPancake/make_plots.py

@@ -1,21 +1,27 @@
-from matplotlib import use; use('Agg')
+import matplotlib as mpl
+mpl.rcParams['font.family'] = 'STIXGeneral'
 import matplotlib.pyplot as plt
 from yt.mods import *
 
 pf = load('DD0001/data0001')
 ray = pf.h.ortho_ray(0,(0,0))
 
-plt.subplot(311)
+p = plt.subplot(311)
 plt.semilogy(ray['x'], ray['Density'], c='k', ls='-', marker='.', ms=5)
 plt.ylabel(r'Density (cm$^{-3}$)')
 
-plt.subplot(312)
+p = plt.subplot(312)
 plt.semilogy(ray['x'], ray['Temperature'], c='k', ls='-', marker='.', ms=5)
 plt.ylabel('Temperature (K)')
+plt.ylim(0.3,1.5e8)
+p.yaxis.set_major_locator(mpl.ticker.FixedLocator([1,100,1e4,1e6,1e8]))
 
-plt.subplot(313)
+
+p = plt.subplot(313)
 plt.plot(ray['x'], 1e-5*ray['x-velocity'], c='k', ls='-', marker='.', ms=5)
 plt.ylabel('Velocity (km/s)')
 plt.xlabel('x')
+plt.ylim(-2250,2250)
+p.yaxis.set_major_locator(mpl.ticker.FixedLocator([-2000,-1000,0,1000,2000]))
 
 plt.savefig('AMRZeldovichPancake.eps')

### 2013-06-09 test_problems/AnisoConduction/make_plots.py
Using bigger fonts in the image plots.
Fixed an issue with the sedov plots that caused the two panels to be plotted
with slightly different colorbars.

@@ -17,8 +17,9 @@ for n in [0,11]:
     print dd["Temperature"].max()
 
     slice = SlicePlot(pf, 'z', "Temperature", width = (1.05, '1'), origin='domain', axes_unit=('1','1'))  #.save()
+    slice.set_window_size(5)
 
-    slice.annotate_streamlines('Bx', 'By', plot_args={'color':(1,1,1,0.5),'linewidth':1.5})
+    slice.annotate_streamlines('Bx', 'By', plot_args={'color':(1,1,1,0.5),'linewidth':0.7})
     
     slice.set_zlim('Temperature', 7e5, 2e7)
 

### 2013-06-09 test_problems/OrszagTang/make_plots.py

@@ -9,7 +9,8 @@ names=[
     ]
 for i, pf in enumerate([pf1,pf2]):
   sl = SlicePlot( pf, 2, 'Density', origin='domain', width=(1.05, 'unitary'))
+  sl.set_window_size(5)
   sl.set_zlim('Density',0.1,1)
   # See matplotlib issue #1188
   sl.plots['Density'].cb.solids.set_edgecolor('face')
-  print sl.save(names[i])
+  print sl.save(names[i],mpl_kwargs={'bbox_inches':'tight'})

### 2013-06-09 test_problems/SedovAMR/sedov_slice.py

@@ -5,17 +5,20 @@ from yt.mods import *
 fn1 = "PPM/DD0007/DD0007"
 pf1 = load(fn1)
 slice1 = SlicePlot(pf1, 'z', "Density", center=[0.5, 0.5, 0.0], origin='domain')
+slice1.set_window_size(5)
 slice1.set_width((1.05,'unitary'))
+slice1.set_zlim('Density',1.5e-3,7e0)
 # See matplotlib issue #1188
 slice1.plots['Density'].cb.solids.set_edgecolor('face')
-slice1.set_font({'size':15})
 slice1.save("sedov-ppm-slice.eps", mpl_kwargs={'bbox_inches':'tight'})
 
 
 fn2 = "Zeus/DD0007/DD0007"
 pf2 = load(fn2)
 slice2 = SlicePlot(pf2, 'z', "Density", center=[0.5, 0.5, 0.0], origin='domain')
+slice2.set_window_size(5)
 slice2.set_width((1.05,'unitary'))
+slice2.set_zlim('Density',1.5e-3,7e0)
 # See matplotlib issue #1188
 slice2.plots['Density'].cb.solids.set_edgecolor('face')
 slice2.save("sedov-zeus-slice.eps", mpl_kwargs={'bbox_inches':'tight'})

### 2013-06-09 numerical-amr.tex
Edits to the overview section. Trying to unify notations, so several other sections are also affected.
Minor updates to the AMR section.  Moving ghost zone figure to this section.

@@
 
+\begin{figure}
+\begin{center}
+\includegraphics[width=0.5\textwidth]{figures/amr_hierarchy.eps}
+\end{center}
+\caption{\emph{Left:} Example of a simple, distributed AMR hierarchy
+showing real and ghost grids.  \emph{Right:} Example 2D \enzo\ grid
+showing real and ghost zones, as needed for the PPM hydro stencil. }
+\label{fig.amr_hierarchy}
+\end{figure}
+
 Each grid patch in \enzo\ contains arrays of values for baryon and particle quantities.
@@
 From the point of view of the hydro solver, each grid is solved as an independent CFD problem, with Dirichlet boundary conditions stored in the ghost zones.
-At the end of each time step, cells on a given level fill ghost zones by interpolating from the parent and copying from neighbors on the same level.
+At the end of each timestep, cells on a given level fill ghost zones by interpolating from the parent and copying from neighbors on the same level.
 Grids then correct the flux at the interface boundary between fine and coarse zones, and finally project their active zone data to the parent grid.
@@
 \end{equation}
-where $\rho_{\rm flag}$ is the equivalent density on the root grid required for refinement, $\Delta x_{\rm root}$ is the root grid cell spacing, $r$ is the refinement factor, $l$ is the level and $\epsilon_l$ is a parameter that can be used to make the refinement more aggressive ($\epsilon_l < 0$) or less aggressive than pure Lagrangian refinement ($\epsilon_l > 0$) -- in other words, super-Lagrangian or sub-Lagrangian in refinement.
+where $\rho_{\rm flag}$ is the equivalent density on the root grid
+required for refinement, $\Delta x_{\rm root}$ is the root grid cell
+spacing, $r$ is the refinement factor, $l$ is the level and
+$\epsilon_l$ is a parameter that can be used to make the refinement
+more aggressive ($\epsilon_l < 0$) or less aggressive ($\epsilon_l >
+0$) than pure Lagrangian refinement --- in other words,
+super-Lagrangian or sub-Lagrangian in refinement.
 
@@
 \begin{equation}
-L_R = |B| / |\nabla \times B|
+L_R = |B| / |\curl B|
 \end{equation}
@@
 %independent CFD problem, with Dirichlet boundary conditions stored in
-%the ghost zones.  At the end of each time step, cells on a given level
+%the ghost zones.  At the end of each timestep, cells on a given level
 %fill ghost zones by interpolating from the parent and copying from

### 2013-06-09 numerical-zeus.tex
Use breakurl package to prevent the bitbucket url from overflowing into the margin.

@@
-\subsection{The \zeus\ hydro method}
+\subsection{Hydrodynamics: The \zeus\ method}
 \label{sec.hydro.zeus}
@@
 On the other hand, it is much more robust than PPM and is easy to add additional physics.
-We also note that this method solves only the internal energy equation rather than total energy, so the `dual energy formulation' discussed in Section~\ref{sec.hydro.ppm} is unnecessary.
+We also note that this method solves only the internal energy equation rather than total energy, so the ``dual energy formulation'' discussed in Section~\ref{sec.hydro.ppm} is unnecessary.

### 2013-06-09 test_problems/DoubleMachReflection/make_plots.py
Small tweaks for the DoubleMachReflection figure.
Adding ms.out to .hgignore as well.

@@ -6,11 +6,16 @@ for n in range(11,12):
 
     c = [2.0,0.5,0]
 
-    slice = SlicePlot(pf, 'z', "Density", center=c, width=((4.15, '1'), (1.15, '1')),
+    slice = SlicePlot(pf, 'z', "Density", center=c,
+                      width=((4.15, '1'), (1.15, '1')),
                       origin='domain', axes_unit=('1','1'))
 
     slice.set_zlim('Density',.5, 20)
 
+    slice.set_buff_size((1000, 300))
+    
+    slice.set_window_size(12)
+    
     # See matplotlib issue #1188
     slice.plots['Density'].cb.solids.set_edgecolor('face')
 

### 2013-06-09 test_problems/TestOrbit/make_plots.py
Adding an inset box to illustrate teh scatter in the orbit test figure.
Adjusting to compensate for the changes in the figure aspect ratios.

@@ -7,7 +7,10 @@
 import numpy as na
 import matplotlib as mpl
 mpl.rcParams['font.family'] = 'STIXGeneral'
+mpl.rcParams['figure.figsize'] = 6,6
 import matplotlib.pyplot as plt
+from mpl_toolkits.axes_grid1.inset_locator import zoomed_inset_axes
+from mpl_toolkits.axes_grid1.inset_locator import mark_inset
 import os
 from math import *
 from yt.mods import *
@@ -26,12 +29,12 @@ zvel = []
 for line in FileIn:
    if 'id=1' in line:
       lst = line.split()
-      xpos.append(lst[1])
-      ypos.append(lst[2])
-      zpos.append(lst[3])
-      xvel.append(lst[4])
-      yvel.append(lst[5])
-      zvel.append(lst[6])
+      xpos.append(float(lst[1]))
+      ypos.append(float(lst[2]))
+      zpos.append(float(lst[3]))
+      xvel.append(float(lst[4]))
+      yvel.append(float(lst[5]))
+      zvel.append(float(lst[6]))
       
 FileIn.close()
 
@@ -79,11 +82,32 @@ ct = na.array(counter)/416.
 en = na.array(energy)/energysum
 
 
-# plots orbit in x-y plane (should be a circle)
-plt.scatter(x[::8], y[::8], s=1, lw=0, marker='.')
-plt.axis("equal")
-plt.xlabel('x')
-plt.ylabel('y')
+# plots orbit in x-y plane (should be a circle with zoom-in inset)
+fig = plt.figure()
+ax = fig.add_subplot(111, aspect='equal')
+ax.scatter(x[::8], y[::8], s=1, lw=0, marker='.').figure
+ax.set_xlabel('x')
+ax.set_ylabel('y')
+ax.set_xlim([-0.05,1.05])
+ax.set_ylim([-0.05,1.05])
+ax.xaxis.set_major_locator(mpl.ticker.FixedLocator([0,0.2,0.4,0.6,0.8,1.0]))
+ax.yaxis.set_major_locator(mpl.ticker.FixedLocator([0,0.2,0.4,0.6,0.8,1.0]))
+
+xbd = [0.744, 0.754]
+ybd = [0.663, 0.673]
+
[…]

### 2013-06-09 test_problems/CosmoIFront/make_plots.py
Updating CosmoIFront to have bigger fonts.

@@ -4,7 +4,7 @@
 # imports
 import matplotlib as mpl
 mpl.rcParams['font.family'] = 'STIXGeneral'
-
+mpl.rcParams['figure.figsize'] = 4, 3
 from pylab import *
 
 
@@ -252,18 +252,24 @@ for tstep in range(te+1):
 
 # combined HI,HII profile plot
 figure()
-plot(Hradii1, HIprof1, 'b-',  Hradii1, HIIprof1, 'r-',
-     Hradii2, HIprof2, 'b--', Hradii2, HIIprof2, 'r--',
-     Hradii3, HIprof3, 'b-.', Hradii3, HIIprof3, 'r-.')
+plot(Hradii1, HIprof1, 'b-',
+     Hradii2, HIprof2, 'b--',
+     Hradii3, HIprof3, 'b-.',
+     Hradii1, HIIprof1, 'r-',
+     Hradii2, HIIprof2, 'r--',
+     Hradii3, HIIprof3, 'r-.')
 xlabel('$r/L_{\mathrm{box}}$')
 ylabel('$\log(x_{\mathrm{HI}}),\/\log(x_{\mathrm{HII}})$')
 ylim(-6.5, 0.5)
-xlim(0,1) 
-legend( ('$x_{\mathrm{HI}}\/z=6.24$','$x_{\mathrm{HII}}\/z=6.24$','$x_{\mathrm{HI}}\/z=2.29$',
-         '$x_{\mathrm{HII}}\/z=2.29$','$x_{\mathrm{HI}}\/z=1.02$','$x_{\mathrm{HII}}\/z=1.02$'),loc=4 )
+xlim(0,1.2) 
+legend( ('$x_{\mathrm{HI}}\/\/z=6.24$',
+         '$x_{\mathrm{HI}}\/\/z=2.29$',
+         '$x_{\mathrm{HI}}\/\/z=1.02$',
+         '$x_{\mathrm{HII}}\/z=6.24$',
+         '$x_{\mathrm{HII}}\/z=2.29$',
+         '$x_{\mathrm{HII}}\/z=1.02$'),
+         loc=4, fontsize=10)
 savefig('FLDprofiles' + pictype, bbox_inches='tight')
-
-
 
 # I-front radius plot vs analytical solution
 #   scaled i-front position
@@ -284,7 +290,7 @@ xlabel('$-\log[(1+z)/(1+z_i)]$')
 ylabel('$r_I/r_S$')
 xlim(-.05,.85)
[…]

### 2013-06-10 test_problems/DoubleMachReflection/make_plots.py
Removing colorbar unit annotations from plots for unitless problems

@@ -1,22 +1,28 @@
 from yt.mods import *
 
+def _DensityCode(field, data):
+    return data['Density']/data.convert('Density')
+
+add_field('DensityCode', display_name="Density", units="",
+          function=_DensityCode)
+
 for n in range(11,12):
 
     pf = load("DD%04i/DD%04i" % (n,n))
 
     c = [2.0,0.5,0]
 
-    slice = SlicePlot(pf, 'z', "Density", center=c,
+    slice = SlicePlot(pf, 'z', "DensityCode", center=c,
                       width=((4.15, '1'), (1.15, '1')),
                       origin='domain', axes_unit=('1','1'))
 
-    slice.set_zlim('Density',.5, 20)
+    slice.set_zlim('DensityCode',.5, 20)
 
     slice.set_buff_size((1000, 300))
     
     slice.set_window_size(12)
     
     # See matplotlib issue #1188
-    slice.plots['Density'].cb.solids.set_edgecolor('face')
+    slice.plots['DensityCode'].cb.solids.set_edgecolor('face')
 
     slice.save(mpl_kwargs={'bbox_inches':'tight'},name="DoubleMachTest.eps")

### 2013-06-10 test_problems/OrszagTang/make_plots.py

@@ -1,4 +1,11 @@
 from yt.mods import *
+
+def _DensityCode(field, data):
+    return data['Density']/data.convert('Density')
+
+add_field('DensityCode', display_name="Density", units="",
+          function=_DensityCode)
+
 name1 = './CT/DD0056/data0056'
 name2 = './Dedner/DD0032/data0032'
 pf1 = load(name1)
@@ -8,9 +15,9 @@ names=[
     'MHDDedner_OrszagTang_Density.eps'
     ]
 for i, pf in enumerate([pf1,pf2]):
-  sl = SlicePlot( pf, 2, 'Density', origin='domain', width=(1.05, 'unitary'))
+  sl = SlicePlot( pf, 2, 'DensityCode', origin='domain', width=(1.05, 'unitary'))
   sl.set_window_size(5)
-  sl.set_zlim('Density',0.1,1)
+  sl.set_zlim('DensityCode',0.1,1)
   # See matplotlib issue #1188
-  sl.plots['Density'].cb.solids.set_edgecolor('face')
+  sl.plots['DensityCode'].cb.solids.set_edgecolor('face')
   print sl.save(names[i],mpl_kwargs={'bbox_inches':'tight'})

### 2013-06-10 test_problems/SedovAMR/sedov_slice.py

@@ -1,25 +1,30 @@
 # Makes slices of density for both simulations (PPM, Zeus)
+from yt.mods import *
 
-from yt.mods import * 
+def _DensityCode(field, data):
+    return data['Density']/data.convert('Density')
+
+add_field('DensityCode', display_name="Density", units="",
+          function=_DensityCode)
 
 fn1 = "PPM/DD0007/DD0007"
 pf1 = load(fn1)
-slice1 = SlicePlot(pf1, 'z', "Density", center=[0.5, 0.5, 0.0], origin='domain')
+slice1 = SlicePlot(pf1, 'z', "DensityCode", center=[0.5, 0.5, 0.0], origin='domain')
 slice1.set_window_size(5)
 slice1.set_width((1.05,'unitary'))
-slice1.set_zlim('Density',1.5e-3,7e0)
+slice1.set_zlim('DensityCode',1.5e-3,7e0)
 # See matplotlib issue #1188
-slice1.plots['Density'].cb.solids.set_edgecolor('face')
+slice1.plots['DensityCode'].cb.solids.set_edgecolor('face')
 slice1.save("sedov-ppm-slice.eps", mpl_kwargs={'bbox_inches':'tight'})
 
 
 fn2 = "Zeus/DD0007/DD0007"
 pf2 = load(fn2)
-slice2 = SlicePlot(pf2, 'z', "Density", center=[0.5, 0.5, 0.0], origin='domain')
+slice2 = SlicePlot(pf2, 'z', "DensityCode", center=[0.5, 0.5, 0.0], origin='domain')
 slice2.set_window_size(5)
 slice2.set_width((1.05,'unitary'))
-slice2.set_zlim('Density',1.5e-3,7e0)
+slice2.set_zlim('DensityCode',1.5e-3,7e0)
 # See matplotlib issue #1188
-slice2.plots['Density'].cb.solids.set_edgecolor('face')
+slice2.plots['DensityCode'].cb.solids.set_edgecolor('face')
 slice2.save("sedov-zeus-slice.eps", mpl_kwargs={'bbox_inches':'tight'})
 

### 2013-06-10 numerical-amr.tex
Minor edits to overview, AMR and Zeus sections.

@@
 
-Unlike moving mesh methods \citep{1995ApJS..100..269P,1995ApJS...97..231G} or  methods that subdivide individual cells \citep{Adjerid}, Berger \& Collela's AMR (also referred to as \emph{structured} AMR) utilizes an adaptive hierarchy of grid patches at varying levels of resolution.
+%Unlike moving mesh methods \citep{1995ApJS..100..269P,1995ApJS...97..231G} or  methods that subdivide individual cells \citep{Adjerid}, Berger \& Collela's AMR (also referred to as \emph{structured} AMR) utilizes an adaptive hierarchy of grid patches at varying levels of resolution.
 The SAMR method described here is similar to \citet{Berger89}.
@@
 While (Time < StopTime)
-begin
    dt = ComputeTimeStep(0)
@@
   CheckForOutput(Time)
-end
 \end{verbatim}
@@
 EvolveLevel(level)
-begin
    SetBoundaryValues
    while (Time < ParentTime)
-  begin
       dt = ComputeTimeStep(level)
@@
       RebuildHierarchy(level+1)
-  end
-end
 \end{verbatim}
@@
 which refers to the position of physical quantities within a cell (the
-primary types are cell-centered, face-centered, and node-centered). In this description we 
-will employ the direct Eulerian version of the piecewise
-parabolic method \citep{1984JCoPh..54..174C,1995CoPhC..89..149B}, which
-assumes cell-centering for all quantities (actually they are
-cell-averages).  However, it is also possible to use other schemes,
-which employ different centering for different variables
-\citep[e.g.,][]{Stone92a}.  The interpolation step, described in
+primary types are cell-centered, face-centered, and node-centered). 
+%In this description we will employ the direct Eulerian version of the piecewise
+%parabolic method \citep{1984JCoPh..54..174C,1995CoPhC..89..149B}, which
+%assumes cell-centering for all quantities (actually they are
+%cell-averages).  
+%However, it is also possible to use other schemes,
+%which employ different centering for different variables
+%\citep[e.g.,][]{Stone92a}.  
+The interpolation step, described in
 Section~\ref{sec:interpolation}, must correctly interpolate with the
@@
 We compute an efficiency $\epsilon$ for this grid, defined as the ratio of flagged cells to total cells.
-If this efficiency is sufficiently large (typically over 30\%), or if the grid is small enough (typically less than $4^d$ coarse cells) then we adopt the region as a new grid and move on to the next region (if any).
+If this efficiency is sufficiently large (typically over 30\%), or if the grid is small enough, then we adopt the region as a new grid and move on to the next region (if any).
 If the region is not acceptable, then we split it into two using the following algorithm.

### 2013-06-10 numerical-zeus.tex
Minor edits to overview, AMR and Zeus sections.

@@
 
-A limitation of a technique that uses an artificial viscosity is that, while the correct Rankine-Hugoniot jump conditions are achieved, shocks are broadened over 3-4 mesh cells and are thus not treated as true discontinuities.
+A limitation of a technique that uses an artificial viscosity is that, while the correct Rankine-Hugoniot jump conditions are achieved, shocks are broadened over 3-4 mesh cells.
 This may cause unphysical pre-heating of gas upstream of the shock wave, as discussed in \citet{1994ApJ...429..434A}.
 On the other hand, it is much more robust than PPM and is easy to add additional physics.
-We also note that this method solves only the internal energy equation rather than total energy, so the ``dual energy formulation'' discussed in Section~\ref{sec.hydro.ppm} is unnecessary.
+We also note that this method solves only the internal energy equation rather than total energy, so the dual energy formulation discussed in Section~\ref{sec.hydro.ppm} is unnecessary.

### 2013-06-10 overview.tex
Minor edits to overview, AMR and Zeus sections.

@@
 
-The code can include either a homogeneous radiation background (with
-possible local self-shielding corrections), or evolve an inhomogeneous
+The code can include either a homogeneous radiation background, 
+or evolve an inhomogeneous
 radiation field either by directly solving the radiative transfer
@@
 explicit, higher-order-accurate version of Godunov's method for ideal
-gas dynamics with third order accurate piecewise parabolic monotonic
+gas dynamics with the spatially third order accurate piecewise parabolic monotonic
 interpolation and a nonlinear Riemann solver for shock capturing.  It

### 2013-06-11 test_problems/GravityTest/make_plots.py
Updating the GravityTest and SphericalInfall figures.
These were generated using a changeset that includes PR 120.

@@ -1,4 +1,6 @@
 import numpy as na
+import matplotlib as mpl
+mpl.rcParams['font.family'] = 'STIXGeneral'
 import matplotlib.pyplot as plt
 
 Data = na.loadtxt("TestGravityCheckResults.out")

### 2013-06-11 overview.tex
Added sentence clarifying B definition.

@@
 where primes indicate quantities in proper coordinates (fixed frame).
+We note that the definition of comoving $\vecB$ used here is such
+that a uniform field is constant in a homogeneous, uniform universe
+and is also the quantity used in the MHD-Dedner solver described
+below; however, it is not universal,
+and in particular is slightly different from that used both in \citet{Li08a}
+and that used in the MHD-CT solver.
 %These definitions result in the following comoving equivalents to eqs. 

### 2013-06-11 overview.tex
Added comoving to E and p

@@
 
-The total fluid energy density $E$ is given by
+The comoving total fluid energy density $E$ is given by
 \begin{equation}
@@
 \end{equation}
-where $e$ is the comoving thermal energy density, the total
+where $e$ is the comoving thermal energy density, the total comoving
 isotropic pressure $p^*$ is given by

### 2013-06-12 acknowledgments.tex
added renyue cen changes.

@@
 Partner program and the Director’s office),
+Princeton University,
 SLAC National Accelerator Laboratory,

### 2013-06-12 introduction.tex
Adding page break so that all affliations appear.
added renyue cen changes.

@@
 transport, star formation and a range of other physical effects.
+\newpage
 
@@
 variety of astrophysical systems that \enzo\ has been used for include
-galaxies \citep{2003ApJ...587...13T, 2012ApJ...749..140H}, galaxy
+galaxies \citep{2003ApJ...587...13T, 2012MNRAS.425..641L, 2012ApJ...749..140H}, galaxy
 clusters \citep{Loken02, Xu11, Skillman13}, the interstellar medium

### 2013-06-12 ms.tex
Adding Daegene Koh to author list.  He has made a few contributions since Enzo Days 2012.
Use breakurl package to prevent the bitbucket url from overflowing into the margin.
added renyue cen changes.

@@
 
-\usepackage{graphicx, natbib, color, bm}
+\usepackage{graphicx, natbib, color, bm, hyperref, breakurl}
 
@@
 of fluids and particles), primordial gas chemistry, optically-thin
-radiative cooling of primordial and metal-enriched plasmas, radiation
-transport, and cosmological expansion.  In addition to explaining the
+radiative cooling of primordial and metal-enriched plasmas (as well as
+some optically-thick cooling models), radiation
+transport, cosmological expansion, and models for star formation and
+feedback in a cosmological context.  In addition to explaining the
 algorithms implemented, we present solutions for a wide range of test

### 2013-06-12 numerical-gravity.tex
Minor change to note that PotentialIterations is user-facing and add comment about MaximumParticleRefinementLevel.
added renyue cen changes.

@@
 Simply interpolating the potential without feeding it back to higher levels leads to errors in the potential at more refined levels, due to the build-up of errors during the interpolation of coarse boundary values.
-In addition, neighboring subgrids are not guaranteed to generate the same potential values because of the lack of a coherent potential solve across the whole hierarchy.
+In addition, neighboring subgrids are not guaranteed to generate the same potential values because of the lack of a coherent potential solve including the whole grid hierarchy.
 In an attempt to partially alleviate this problem, we allow for an iterative procedure across sibling grids, in which the potential values on the boundary of grids can be updated with the potential in `active' regions of neighboring subgrids.
-To prevent overshoot, we average the potential on the boundary and allow for (typically) 4 iterations.
+To prevent overshoot, we average the potential on the boundary and allow for (by default) 4 iterations, with the number of iterations determined by a parameter specified at runtime.
 This procedure can help in many cases, but does not necessarily produce a coherent solution across all grids and so does not completely solve the problem; we are working on a slower but more accurate method that does a multigrid solve across the whole grid (Reynolds \etal, in preparation).

### 2013-06-12 introduction.tex
alexei kritsuk changes

@@
 2011ApJ...731....6S}, cooling flows \citep{Li12, 2013ApJ...763...38S},
-turbulence \citep{Kritsuk04, 2009JPhCS.180a2020K, Collins11}, the
+turbulence \citep{Kritsuk04, 2007ApJ...665..416K, 2009JPhCS.180a2020K, Collins11}, the
 formation of the first stars \citep{ABN02, 2007ApJ...654...66O,
 2009Sci...325..601T, Xu08}, and the formation of stars in our own Galaxy
-\citep{Collins11, Collins12a}.
+\citep{Collins11, 2011ApJ...727L..20K, Collins12a}.
 
@@
 obligations: the first is to make those tools available to the
-community as a whole, much in the way that astronomical data is now
+community as a whole, much in the way that astronomical data are now
 regularly made publicly available.  The second is to document, test
 and refine those methods so that they can be critically evaluated and
-expanded upon by others.  This paper represents one component of our
-attempt to meet these obligations with respect to the \enzo\ code.
+expanded upon by others.  Our public release of the \enzo\ code represents our
+attempt to meet the first of these obligations; this paper represents
+our attempt to meet the second.
 

### 2013-06-12 table_of_variables_and_constants.tex
alexei kritsuk changes

@@
   $t_{\rm form}$ & star particle formation time \\
-  $m_b$ & baryonic mass of cell \\
-  $m_J$ & Jeans mass of cell \\
+  $m_{\rm b}$ & baryonic mass of cell \\
+  $m_{\rm J}$ & Jeans mass of cell \\
   $m_*$ & star particle mass \\
@@
 
-  \multicolumn{2}{l}{\underline{Natural Constants}} \\[5pt]
+  \multicolumn{2}{l}{\underline{Natural Constants and Values}} \\[5pt]
   $G$ & gravitational constant \\
@@
   $c$ & speed of light \\
-  $c_s$ & sound speed \\
+  $c_{\rm s}$ & sound speed \\
   \enddata

### 2013-06-12 tests-doublemach.tex
alexei kritsuk changes

@@
-\subsubsection{Double mach reflection}
+\subsubsection{Double Mach reflection}
 \label{sec.tests.doublemach}
 
-The double mach reflection test is a classic two-dimensional test of
+The double Mach reflection test is a classic two-dimensional test of
 hydrodynamic algorithms originally described in
@@
 the x-axis.  The shape of this jet is sensitive to numerical
-diffusion, and our figure compares favorably to the higher-order
-images from Woodward \& Colella.
+diffusion, and our figure compares favorably to those shown in Woodward \& Colella.
 

### 2013-06-12 tests-vandv.tex
alexei kritsuk changes

@@
 included the \enzo\ code, with the former testing the PPM
-hydrodynamics and the latter the constrained transport MHD
-implementation of \citet{Collins10}.  In both cases, \enzo\ performed similarly to other
+hydrodynamics and the latter both the constrained transport MHD
+implementation of \citet{Collins10} and the Dedner method of \citet{WangAbelZhang08}.  In both cases, \enzo\ performed similarly to other
 grid-based codes that use Godunov-based fluid solvers, and typically

### 2013-06-12 acknowledgments.tex
acknowledged the people who weren't on the author list.

@@
 
+The \enzo\ collaboration would like to acknowledge the following
+scientists, who have made contributions to the \enzo\ codebase at some
+point during their research career: Brian Crosby, Elizabeth
+Harper-Clark, Daegene Koh, Eve Lee, Pascal Paschos, Carolyn Peruta,
+Alex Razoumov, Munier Salem, and Rick Wagner.
+
 The \enzo\ collaboration would also like to acknowledge the significant contributions to

### 2013-06-12 tests-vandv.tex
added matt's changes.

@@
 problems, as well as the scripts to generate the figures from them,
-are available in the \enzo\ test suite, allowing the interested reader
-to regenerate all of the figures in Section~\ref{sec.tests.compare}
-using the \textit{`methodpaper'} suite of test problems.
+are available at \url{http://bitbucket.org/bwoshea/enzo-method-paper},
+the BitBucket repository for the \enzo\ method paper.  The
+\texttt{test\_problems} subdirectory in this repository contains all of
+the files necessary to regenerate all of the figures in
+Section~\ref{sec.tests.compare}.  We note for completeness that all of
+the figures were generated using \texttt{yt} version 2.5.3 and the
+\url{https://bitbucket.org/enzo/enzo-dev} repository of \enzo\ with
+changeset \texttt{5d8c412}.
 

### 2013-06-13 overview.tex
a few more changes from alexei

@@
 equation represents conservation of mass, the second conservation of
-momentum, and the third conservation of total (kinetic plus thermal)
+momentum, and the third conservation of total (kinetic plus thermal
+plus magnetic)
 fluid energy.  They are respectively, the first, second, and third

### 2013-06-13 parallel.tex
Minor updates to the AMR section.  Moving ghost zone figure to this section.
a few more changes from alexei

@@
 
-\begin{figure}
-\begin{center}
-\includegraphics[width=0.5\textwidth]{figures/amr_hierarchy.eps}
-\end{center}
-\caption{\emph{Left:} Example of a simple, distributed AMR hierarchy
-showing real and ghost grids.  \emph{Right:} Example 2D \enzo\ grid
-showing real and ghost zones, as needed for the PPM hydro stencil. }
-\label{fig.amr_hierarchy}
-\end{figure}
-
 % ----------------------------------
@@
 non-gravitating turbulence
-\citep[e.g.,][]{2002ApJ...569L.127K,Kritsuk04}, the Lyman-$\alpha$ forest
+\citep[e.g.,][]{2002ApJ...569L.127K,Kritsuk04, 2007ApJ...665..416K,
+2009ASPC..406...15K}, the Lyman-$\alpha$ forest
 \citep{2005MNRAS.361...70J,2009MNRAS.399.1934P}, or feedback of

### 2013-06-13 numerical-mhd-ct.tex
I still managed to mess up the cosmological B field.  Ok this time I have it right, really!  In particular, the paper now adopts the MHD-CT B definitions throughout and explicits lists the changes in the MHD-Dedner scheme.

@@
 
-Cosmological MHD can be treated in a number of ways.  As described in 
-the previous section, and equation~\ref{eq:induction} the MUSCL with Dedner
-method treats the cosmological dilution of
-$\vecB$ with a source term,  However, one can eliminate the right hand side of equation 
-\ref{eq:induction} by defining
-$\vecB^\prime = \vecB / \sqrt{a}.$
-Then equation~\ref{eq:induction} becomes
-$$ 
-\frac{\partial \vecB^\prime}{\partial t} - \frac{1}{a}  \curl (\vecv \times
-\vecB^\prime)  = 0, \label{eq:induction_cosmo_prime_nosource}
-$$
-and we can reuse each part of the non-cosmological algorithm (hyperbolic solver,
-electric field creation, interpolation, projection, and flux correction) without
-modification.
+%Cosmological MHD can be treated in a number of ways.  As described in 
+%the previous section, and equation~\ref{eq:induction} the MUSCL with Dedner
+%method treats the cosmological dilution of
+%$\vecB$ with a source term,  However, one can eliminate the right hand side of equation 
+%\ref{eq:induction} by defining
+%$\vecB^\prime = \vecB / \sqrt{a}.$
+%Then equation~\ref{eq:induction} becomes
+%$$ 
+%\frac{\partial \vecB^\prime}{\partial t} - \frac{1}{a}  \curl (\vecv \times
+%\vecB^\prime)  = 0, \label{eq:induction_cosmo_prime_nosource}
+%$$
+%and we can reuse each part of the non-cosmological algorithm (hyperbolic solver,
+%electric field creation, interpolation, projection, and flux correction) without
+%modification.  
+%GB: I removed the above because the rest of the paper now uses the definition
+% of B in MHD-CT.
 

### 2013-06-13 numerical-mhd-dedner.tex
I still managed to mess up the cosmological B field.  Ok this time I have it right, really!  In particular, the paper now adopts the MHD-CT B definitions throughout and explicits lists the changes in the MHD-Dedner scheme.

@@
-\subsection{Hydrodynamics: MUSCL}
+\subsection{Hydrodynamics and Magnetohydrodynamics: MUSCL with Dedner cleaning}
 \label{sec.num.hydro-muscl}
@@
 Runge-Kutta scheme \citep{1988JCoPh..77..439S}.
+
+Finally, we note that for cosmological simulations, this solver uses a slightly different definition of the magnetic field than used in the rest of the paper.
+In particular, the field is defined as $\vecB = a^{3/2} \vecB^{\prime}$ (where $\vecB^{\prime}$ is the proper field strength).
+This adds a source term of $-\dot{a}/(2a) \vecB$ on the right-hand side of equation~\ref{eq:induction}, removes the factor of $a$ in the $\vecB$ term in the energy equation (eq.~\ref{eq:total_energy_def}) and in the definition of the isotropic pressure $p^*$.

### 2013-06-13 overview.tex
I still managed to mess up the cosmological B field.  Ok this time I have it right, really!  In particular, the paper now adopts the MHD-CT B definitions throughout and explicits lists the changes in the MHD-Dedner scheme.

@@
   - \frac{\dot{a}}{a} \left( 2E - \frac{B^2}{2a} \right) - 
-  \frac{1}{a} \vecv \cdot \grad \phi 
+  \frac{\rho}{a} \vecv \cdot \grad \phi 
   - \Lambda + \Gamma + \frac{1}{a^2} \div \fcond,
@@
   \frac{1}{a}  \grad \times (\vecv \times \vecB) & = & 
-  -\frac{\dot{a}}{2 a} \vecB. \label{eq:induction}
+  %-\frac{\dot{a}}{2 a} \vecB. 
+  0
+  \label{eq:induction}
 \end{eqnarray}
@@
 We note that the definition of comoving $\vecB$ used here is such
-that a uniform field is constant in a homogeneous, uniform universe
-and is also the quantity used in the MHD-Dedner solver described
+that a uniform field is constant in a homogeneous expanding universe
+and is also the quantity used in the MHD-CT solver described
 below; however, it is not universal,
 and in particular is slightly different from that used both in \citet{Li08a}
-and that used in the MHD-CT solver.
+and used in the MHD-Dedner solver (see section~\ref{sec.num.hydro-muscl}
+for more details).
 %These definitions result in the following comoving equivalents to eqs. 

### 2013-06-14 acknowledgments.tex
Adding Gabriel Altay to the list of enzo contributors.

--- 2013-06-12 acknowledgments.tex
@@
 
+The \enzo\ collaboration would like to acknowledge the following
+scientists, who have made contributions to the \enzo\ codebase at some
+point during their research career: Gabriel Altay, Brian Crosby,
+Elizabeth Harper-Clark, Daegene Koh, Eve Lee, Pascal Paschos, Carolyn
+Peruta, Alex Razoumov, Munier Salem, and Rick Wagner.
+
 The \enzo\ collaboration would also like to acknowledge the significant contributions to

### 2013-06-17 numerical-chemistry.tex
Adding note about the variable H2 gamma.

@@
 reaction rate coefficients and the cooling functions of the gas) on timescales
-that are faster than those of the gas dynamics.
+that are faster than those of the gas dynamics.  The gamma used by \enzo\ to
+compute the temperature of the gas from the energy and density characteristics
+utilizes a variable gamma that includes effects of the rotational state of
+molecular hydrogen, enabling it to vary from $5/3$ (fully-atomic) to $7/5$
+(fully molecular).  This further coupling of the chemical and thermal states of
+the gas underscores the need for coupled chemistry and radiative cooling
+solutions.
 
@@
 \caption[]{Radiative processes solved in the \enzo\ nonequilibrium
-primordial chemistry routines. \red{Missing rates for HD destruction/ionization?}}
+primordial chemistry routines.}
+%\red{Missing rates for HD destruction/ionization?}}
 \label{table.radiative}

### 2013-06-17 ms.tex
fixed alexei's name
Jeff Oishi's changes.

@@
 \author{Brian W. O'Shea\altaffilmark{4,5}}
-\author{Tom Abel\altaffilmark{6, 19}}
+\author{Tom Abel\altaffilmark{6, 20}}
 \author{John H. Wise\altaffilmark{7}}
@@
 \author{Cameron Hummels\altaffilmark{16}}
-\author{Alexei Kritsuk\altaffilmark{2}}
+\author{Alexei G. Kritsuk\altaffilmark{2}}
 \author{Elizabeth Tasker\altaffilmark{17}}
@@
 \author{Oliver Hahn\altaffilmark{18}}
-\author{Jeffrey S. Oishi\altaffilmark{6}}
+\author{Jeffrey S. Oishi\altaffilmark{19}}
 \author{Geoffrey C So\altaffilmark{2}}
-\author{Fen Zhao\altaffilmark{19}}
-\author{Renyue Cen\altaffilmark{20}}
+\author{Fen Zhao\altaffilmark{20}}
+\author{Renyue Cen\altaffilmark{21}}
 \author{Yuan Li\altaffilmark{1}}
@@
 \altaffiltext{18}{ETH Zurich Institute for Astronomy. CH-8093 Zurich. Switzerland}
-\altaffiltext{19}{Department of Physics, Stanford University, Stanford, CA 94305, USA}
-\altaffiltext{20}{Department of Astrophysical Sciences, Princeton University, Peyton Hall, Princeton, NJ 08544, USA}
+\altaffiltext{19}{Department of Astrophysics, American Museum of Natural History, Central Park West at 81st St, New York, NY 10024}
+\altaffiltext{20}{Department of Physics, Stanford University, Stanford, CA 94305, USA}
+\altaffiltext{21}{Department of Astrophysical Sciences, Princeton University, Peyton Hall, Princeton, NJ 08544, USA}
 

### 2013-06-17 numerical-analysis.tex
Jeff Oishi's changes.

@@
 passed to a special handler inside \texttt{yt}.  \texttt{yt} interprets the
-mesh and fluid information and, without saving data to disk except as
-explicitly requested, constructs a native representation of the in-memory state
+mesh and fluid information and, without saving data to disk, constructs a native representation of the in-memory state
 of the simulation that appears identical to an on-disk simulation output.
@@
 \begin{eqnarray}
-\nabla \cdot \vec{v} < 0,\\
-\nabla T \cdot \nabla S > 0,\\
+\div \myvec{v} < 0,\\
+\grad T \cdot \grad S > 0,\\
 T_2 > T_1,\\
@@
 
-where $\vec{v}$ is the velocity field, $T$ is the temperature, $\rho$ is the
+where $\myvec{v}$ is the velocity field, $T$ is the temperature, $\rho$ is the
 density, and $S=T/\rho^{\gamma-1}$ is the entropy.  $T_2$ and $T_1$ are the post-shock~(downstream)

### 2013-06-17 numerical-overview.tex
Jeff Oishi's changes.

@@
 calculated in equation~\ref{enzoenergy} and $E_{kin}$ is the specific kinetic energy,
-$0.5 \vec{v}_b^2$.  
+$0.5 \myvec{v}_b^2$.  
 In hypersonic flows $E$ and $E_{kin}$ are nearly the same, and any 
@@
 \begin{equation}
-\frac{\partial e}{\partial t} + \frac{1}{a} \vec{v}_b \cdot \grad e = 
-- \frac{3(\gamma-1)\dot{a}}{a}e - \frac{p}{a \rho} \div \vec{v}_b
+\frac{\partial e}{\partial t} + \frac{1}{a} \myvec{v}_b \cdot \grad e = 
+- \frac{3(\gamma-1)\dot{a}}{a}e - \frac{p}{a \rho} \div \myvec{v}_b
 \label{enzointenergy}
@@
 \begin{equation}
-\div \vec{v}_b < 0,
+\div \myvec{v}_b < 0,
 \label{cencont}
@@
 
-where $\vec{v}$ is the velocity of the gas in the cell, $\rho_{b}$ and 
+where $\myvec{v}$ is the velocity of the gas in the cell, $\rho_{b}$ and 
 $\rho_{d}$ are the cell's baryon and dark matter density, respectively,
@@
 \begin{equation}
-\Delta t_{accel} = min \left( \sqrt{\frac{\Delta x}{\vec{g}}} \right)_L 
+\Delta t_{accel} = min \left( \sqrt{\frac{\Delta x}{\myvec{g}}} \right)_L 
 \label{eqn:dtaccel}
@@
 takes into account the possibility of large accelerations causing numerical 
-instabilities by violating the Courant condition.  In this equation, $\vec{g}$ is the
+instabilities by violating the Courant condition.  In this equation, $\myvec{g}$ is the
 gravitational acceleration in each cell on level L.  \blue{ \emph{It should be

### 2013-06-17 numerical-timestep.tex
A few more edites to section 2. Also unified "time step" -> "timestep" throughout the whole manuscript.
Jeff Oishi's changes.

@@
-\section{Time Stepping}
+\section{Timestepping}
 \label{sec.timestepping}
@@
 \begin{equation}
-\Delta t_{\rm accel} = \min \left( \sqrt{\frac{\Delta x}{\vec{g}}} \right)_L ,
+\Delta t_{\rm accel} = \min \left( \sqrt{\frac{\Delta x}{\myvec{g}}} \right)_L ,
 \label{eqn:dtaccel}
@@
 \begin{equation}
-\Delta t_{\rm rad} = \min \left(  \sqrt{\frac{\Delta x}{\vec{a}_{rad}}} \right)_L,
+\Delta t_{\rm rad} = \min \left(  \sqrt{\frac{\Delta x}{\myvec{a}_{rad}}} \right)_L,
 \label{eqn:dtrad}
@@
 gravity (equation~\ref{eqn:dtaccel}) or radiation pressure
-(equation~\ref{eqn:dtrad}).  In equation~(\ref{eqn:dtaccel}), $\vec{g}$ is the
+(equation~\ref{eqn:dtrad}).  In equation~(\ref{eqn:dtaccel}), $\myvec{g}$ is the
 gravitational acceleration in each cell on level l.  In
-equation~(\ref{eqn:dtrad}), $\vec{a}_{\rm rad}$ is the estimated
+equation~(\ref{eqn:dtrad}), $\myvec{a}_{\rm rad}$ is the estimated
 acceleration due to radiation pressure in each cell on level l,
@@
 \begin{equation}
-\vec{a}_{rad} = \frac{ \sum_i \frac{\dot{E_i}}{c} \hat{r_i} }{m_b} 
+\myvec{a}_{rad} = \frac{ \sum_i \frac{\dot{E_i}}{c} \hat{r_i} }{m_b} 
 \end{equation}

### 2013-06-19 appendix.tex
Tweaks to appendix

@@
 \end{equation}
-
 is not, in general, equal to $Q_0$.  We can retain conservation by
@@
 not preserve monotonicity -- it can introduce local minima and maxima
-at at parent cell boundaries.  We can then attempt to correct that by
+at parent cell boundaries.  We can then attempt to correct that by
 taking a weighted average between the interpolated values on the

### 2013-06-20 introduction.tex
Christine Simpson's changes.

@@
 variety of astrophysical systems that \enzo\ has been used for include
-galaxies \citep{2003ApJ...587...13T, 2012MNRAS.425..641L, 2012ApJ...749..140H}, galaxy
+galaxies \citep{2003ApJ...587...13T, 2012MNRAS.425..641L,
+  2012ApJ...749..140H, 2013MNRAS.432.1989S}, galaxy
 clusters \citep{Loken02, Xu11, Skillman13}, the interstellar medium

### 2013-06-20 numerical-amr.tex
Jeff Oishi's changes.
Christine Simpson's changes.

@@
 
-From the point of view of the hydro solver, each grid is solved as an independent CFD problem, with Dirichlet boundary conditions stored in the ghost zones.
+From the point of view of the hydro solver, each grid is solved as an independent computational fluid dynamics (CFD) problem, with Dirichlet boundary conditions stored in the ghost zones.
 At the end of each timestep, cells on a given level fill ghost zones by interpolating from the parent and copying from neighbors on the same level.
@@
 is always refined by at least N cells locally (i.e.~we refine a given cell further if $L_R \leq N \Delta x$).
+Note that \enzo\ does not actually solve the resistive MHD equations -- this is the ``characteristic'' local resistive length, rather than one calculated within the MHD solver.
 

### 2013-06-20 overview.tex
Remove \rho from EOS definition and noted that units are such that magnetic permeability is unity.
Jeff Oishi's changes.
Christine Simpson's changes.

@@
 total fluid energy density, comoving gas density, peculiar
-velocity, and comoving magnetic field strength, respectively.
+velocity, and comoving magnetic field strength, respectively. The
+matrix 
 $\myvec{I}$ is the identity matrix, and $a$ is the cosmological
@@
 momentum, and the third conservation of total (kinetic plus thermal
-plus magnetic)
-fluid energy.  They are respectively, the first, second, and third
+plus magnetic) fluid energy.  They are respectively, the first, second, and third
 moments of the Boltzmann equation.  The fourth equation is the
@@
 \end{equation}
-where $e$ is the comoving thermal energy density, the total comoving
+where $e$ is the comoving thermal energy density. The total comoving
 isotropic pressure $p^*$ is given by
@@
 \end{equation}
-and $p$ is the thermal pressure.  The equations are closed by an
+and the quantity $p$ is the thermal pressure.  We use units such that the magnetic
+permeability is unity ($\mu_0=1$).  The equations are closed by an
 equation of state and Poisson's equation (in the Newtonian limit) for
@@
 %
-  e &=& \frac{p}{(\gamma - 1) \rho},
+  e &=& \frac{p}{(\gamma - 1)},
   \label{eq:eq_of_state} \\
@@
 \end{equation}
-$D$ in Equation (\ref{eq:fld_radiation}) is the Larsen
-square-root flux-limiter \citep[see][]{Morel2000}, $\kappa$ the
-total opacity, $\eta$ the field of radiation sources, $\Gamma$
-the radiation induced photo-heating rate, and $\Lambda$ the
+The quantity $D$ in Equation (\ref{eq:fld_radiation}) is the Larsen
+square-root flux-limiter \citep[see][]{Morel2000}, $\kappa$ is  the
+total opacity, $\eta$ is the field of radiation sources, $\Gamma$ is
+the radiation induced photo-heating rate, and $\Lambda$ is the
 chemical cooling rate.
@@
 
-The primary purpose of the \enzo\ code is its Adaptive Mesh Refinement
+The primary feature of the \enzo\ code is its Adaptive Mesh Refinement
 (AMR) capability, which allows it to reach extremely large dynamical
-ranges with limited computational resources, opening doors otherwise
+ranges with limited computational resources, opening doors to applications otherwise
 closed by finite memory and computational time. Unlike moving mesh
@@
 that subdivide individual cells \citep{Adjerid}, Berger \& Collela's
-AMR (also referred to as \emph{structured} AMR) utilizes an adaptive
+AMR (also referred to as \emph{structured} AMR, or SAMR) utilizes an adaptive
 hierarchy of grid patches at varying levels of resolution.  Each
@@
 domain of interest with a coarse uniform Cartesian grid. Then, as
-the solution evolves and interesting regions start to develop, finer meshes are
+the solution evolves and interesting regions start to develop, finer grids are
 placed within these coarse regions.
@@
 % finer grids and `above' for coarser grids). 
-We restrict the ratio between cell sizes to be an integer, typically 2
-or 4, and refer to a level as all grids with the same cell size.  In
-order to keep things as simple as possible, the edges of subgrids must
+We restrict the ratio between cell sizes of parent and child grids to be an integer, typically 2
+or 4, and refer to a level as all grids with the same cell size.  For simplicity, the edges of subgrids must
 coincide with the cell edge of its immediate parent (coarser)
@@
 finer resolutions. Rebuilding is achieved by applying the grid
-refinement criteria to the grids on that level, flagging zones that
+refinement criteria to the grids on that level and flagging zones that
 require extra grids.  These criteria depend on the physical problem
@@
 copied to the new subgrids. This procedure is then repeated on the new
-grids and in this way the entire hierarchy (from the original level
+grids and in this way, the entire hierarchy (from the original level
 examined and continuing on to all finer levels) is rebuilt.
@@
 \enzo\ can operate in a number of different modes with regard to
-radiative cooling and heating. In the simplest mode, with the
-multi-species flag turned off, the cooling rate is computed from a
+radiative cooling and heating. In the simplest mode, where the
+multi-species flag is turned off and no individual chemical species
+are tracked, the cooling rate is computed from a
 simple temperature-dependent cooling function, taken from
@@
 cooling from all species of hydrogen and helium (including H$_2$) --
-this primordial cooling is treated in the same Jacobi iteration as the
+and the primordial cooling rates are computed in the same Jacobi iteration as the
 chemistry.  It is also possible to include metal cooling based on a
@@
 
-All grids on a given level are advanced with the same timestep,
-determined by calculating, for each cell and for each physical process
-(except for chemistry and heat conduction, which are sub-cycled)
-separately, the largest $\Delta t$ allowed for stability. The level is
-advanced with a timestep equal to the minimum over all of these
-$\Delta t$.
+All grids on a given level are advanced with the same timestep.  This
+time step is determined by first calculating the largest time step
+allowed for each cell and for each physical process separately (except
+for chemistry and heat conduction, which are sub-cycled).  The level
+is then advanced with a timestep equal to the minimum over all of
+these $\Delta t$.
 

### 2013-06-20 numerical-analysis.tex
adding stephen skory changes

@@
 halo finder \citep{2013ApJ...762..109B} can be executed through \texttt{yt} on
-in-memory \enzo\ data.  Operations that currently cannot be conducted on
+in-memory \enzo\ data, and so can the Parallel HOP halo finder
+\citep{1998ApJ...498..137E, 2010ApJS..191...43S}.  Operations that currently cannot be conducted on
 in-memory datasets are those that require spatial decomposition of
@@
 
-\enzo saves tracer particle data at user-specified intervals,
+\enzo\ saves tracer particle data at user-specified intervals,
 independent of the intervals at which regular data sets are written.  The data written out

### 2013-06-20 numerical-mhd-ct.tex
adding stephen skory changes

@@
 energy, as described in Section \ref{sec.hydro.ppm}, and one that
-uses entropy \cite{TVD93, Collins10}..  
+uses entropy \cite{TVD93, Collins10}.
 %The additional equations are

### 2013-06-20 numerical-ppm.tex
More edits to fluid section - done?
adding stephen skory changes

@@
 
-\subsection{Hydrodynamics: The PPM hydro method}
+\subsection{Hydrodynamics: The PPM method}
 \label{sec.hydro.ppm}
@@
 problem, which is generally required to compute the fluxes in 
-any Godunov.  This section therefore applies to all three of our 
+any Godunov-based scheme.  This section therefore applies to all three of our 
 Godunov-based schemes.  The Riemann problem we are solving

### 2013-06-20 parallel.tex
adding stephen skory changes

@@
 seen between the two machines, which can be attributed primarily to
-the slightly faster CPUs on Kraken at th etime (2.3 Ghz, vs. 2.1 Ghz
+the slightly faster CPUs on Kraken at the time (2.3 Ghz, vs. 2.1 Ghz
 on Jaguar).  The difference in speed when using different numbers of

### 2013-06-20 parallel.tex
added GPU section, including a new figure.  Also include the script that generated the figure.

@@
 compute times for \enzo\ calculations.
+
+\subsection{GPU parallelization and scaling}
+
+Many of today's leading supercomputers use a heterogeneous structure:
+on a single node of a distributed-memory platform, a multi-core CPU is
+often paired with one or more many-core accelerators.  One programming
+model that has been shown to successfully take advantage of these
+hardware accelerators is to run the serial component of the algorithm
+on the CPU, and the vector-parallel part of the algorithm on the
+many-core hardware accelerator. Due to the much higher computational
+performance of the many-core accelerators, if a code can be ported to
+effectively utilize this heterogeneous architecture, massive speedup
+in simulation performance may be achieved.  This heterogeneous trend
+is likely to continue in future supercomputers, due to the relatively
+low energy needs of hardware accelerators as well as other
+factors. Thus, to ensure \enzo's efficiency on future high-end
+computational platforms, some of the most time-consuming parts of
+\enzo\ have been ported to many-core architecture.  Since GPUs are
+currently the most popular many-core accelerator, NVIDIA's CUDA C/C++
+programming technology was chosen to port some \enzo\ modules to
+NVIDIA GPUs.
+
+Thus far, the PPM and Dedner MHD solvers (described in
+Sections~\ref{sec.hydro.ppm} and~\ref{sec.num.hydro-muscl},
+respectively), have been ported to take advantage of GPUs using CUDA.
+Since the CPU and GPU on a node have access to separate pools of RAM,
+fields updated by other modules will be transferred to the GPU before
+calling the GPU fluid solver.  Correspondingly, after calling the GPU
+solver, the updated fields will be transferred back to CPU. This
+ensures that the GPU-parallelized solvers can work correctly with
+other parts of the code that have not yet been ported to the GPU, as
+well as with the communication infrastructure within \enzo.  In
+addition, the GPU solver supports SAMR, which is not a trivial task.
+As discussed in Section~\ref{sec.amr}, one of the key steps in \enzo's
+AMR implementation is flux correction, which is required when each
+level of resolution is allowed to take its own time step.  In the GPU
+version of the solvers, the fluxes are calculated on the GPU and only
+the fluxes required for flux correction are transferred back to the
+CPU.  This reduces data transfer overhead, which can be substantial in
+a heterogeneous architecture of this sort.
+
+The key step in porting to many-core architectures such as the GPU is
+exposing the massive parallelism inherent in the algorithm. Due to the
+explicit, directionally-split stencil pattern of both the PPM and
+Dedner MHD solvers, they are inherently massively parallel and thus
+should be good fit for hardware acceleration. Both solvers essentially
+contain basically two parts -- computation of fluxes, and cell update.
+When computing fluxes, the basic procedure is to compute the flux at
+each cell-interface given the inputs from neighboring cells. In the
+cell update part, the fundamental computation is updating the
+cell-centered values using the previously-computed fluxes. In the CPU
+serial code, a loop over the grid is used, where the loop body
+contains these basic computations. In both parts,
+algorithmically-different loop iterations are completely independent
+of each other. Thus, the parallelization scheme is naturally mapping
+one GPU thread to one iteration of the loop. However, the original CPU
+code, which operates on each grid as a serial process, does not
+completely expose this parallelism as some small temporary arrays are
+reused among loop iterations.  This re-use of arrays introduces data
+dependency among loop iterations, which is undesirable for GPU
+parallelization.  Because of this re-use of arrays, the main change in
+porting the serial CPU solvers to allow massively parallel GPU
+computation was replacing those shared temporary arrays by larger
+temporary arrays that are not shared among loop iterations. This
+change exposed the massive parallelism in the algorithm, which can
+then be accelerated in a straightforward manner using CUDA.
+
+To illustrate the speedup provided by porting two solvers to GPUs, we
+show the results of a weak scaling test of driven turbulence in
+Figure~\ref{fig:gpu_scaling}.  This \enzo\ problem type uses the
+Dedner MHD solver, which runs on both CPUs and GPUs, and has been
+tuned to maximize performance on both platforms.  The benchmarking platform is a Cray
+XK6 system where each node has a 16-core AMD Opteron 6272 CPU and a single
+NVIDIA Tesla K20 GPU, which has 2,496 CUDA cores and a maximum
+theoretical speed of 1.17 Tflops when used in double precision.
+For the CPU run, all the 16 cores in each node
+are used by launching 16 MPI processes per node. For the GPU run, 16 MPI
+processes are also launched per node, and all processes on this node share the single GPU
+on that node. This can work because NVIDIA's MPS (Multi-Process
+Service) technology allows multiple processes to concurrently use the
+same GPU.  A series of weak scaling tests were run, varying from 1
+to 8 nodes (16 to 128 MPI processes), with the number of cells per
+node fixed to $256^3$.  Figure~\ref{fig:gpu_scaling} displays the
+scaling results in terms of cell updates per second per node -- as a
+result, ideal weak scaling is a horizontal line.  In general, the
+simulations using the GPU-accelerated MHD solver perform roughly 5
+times better overall on this system than the CPU-only calculations.
+
+\begin{figure}
+\centerline{\includegraphics[width=0.5\textwidth]{figures/gpu_scaling.eps}}
+\caption{Weak scaling performance of MHD turbulence using the Dedner
+MHD algorithm, which has both CPU and GPU solvers.  Each node of the
+machine (a Cray XK6) contains one 16-core AMD Opteron 6272 CPU and 1
+NVIDIA Tesla K20 GPU.  In this test, there are $256^3$ cells/node.
+The y-axis shows cell updates per second per node, which for ideal
+weak scaling is a horizontal line.  Blue line: CPU solver.  Red line:
+GPU solver.}
+\label{fig:gpu_scaling}
+\end{figure}

### 2013-06-20 gpu_plot/gpu_plot.py

@@ -1,0 +1,26 @@
+import matplotlib.pyplot as plt
+
+X = [1.0,2.0,4.0,8.0]
+cpu = [1.92,3.94,7.73,14.0]
+gpu = [11.73,28.44,46.93,55.0]
+
+gpunode = []
+cpunode = []
+
+for n in range(len(cpu)):
+    gpunode.append(1.0e6*gpu[n]/X[n])
+    cpunode.append(1.0e6*cpu[n]/X[n])
+
+print gpunode
+print cpunode
+
+plt.plot(X,cpunode,color='blue',linewidth=2.0,marker='o')
+plt.plot(X,gpunode,color='red',linewidth=2.0,marker='o')
+plt.loglog()
+plt.axis([0,10,1e6,5e7])
+plt.xlabel('N$_{node}$')
+plt.ylabel('Cells/s/node')
+plt.text(7.0, 2.0e7, 'CPU', size=15, color='blue')
+plt.text(7.0, 3.0e7, 'GPU', size=15, color='red')
+plt.savefig('gpu_scaling.eps')
+

### 2013-06-20 conclusions.tex
Mike Norman's edits.

@@
 fact are faced by effectively all applications that wish to take
-advantage of new computational architectures. We therefore anticipated
+advantage of new computational architectures. We therefore anticipate
 that \enzo\ (or a code that has the capabilities of \enzo, from a

### 2013-06-20 introduction.tex
Mike Norman's edits.

@@
 Due to the high spatial and temporal dynamical ranges involved,
-astrophysics and cosmological phenomena present a taxing challenge for
+astrophysical and cosmological phenomena present a taxing challenge for
 simulators. To tackle such situations, a number of numerical
@@
 following physical processes such as gravitational instability. A
-solution to this was first proposed by \citet{Berger89} in the
+solution to this problem was first proposed by \citet{Berger89} in the
 Computational Fluid Dynamics (CFD) community, and became known as

### 2013-06-20 ms.tex
Fixed extra comma after Enzo Collab in author list
Updated SMU affiliation line
Christine Simpson's changes.
adding stephen skory changes
Mike Norman's edits.

@@
 \title{Enzo: An Adaptive Mesh Refinement Code for Astrophysics}
-\author{The Enzo Collaboration:}
-\author{Greg L. Bryan\altaffilmark{1}}
+\author{The Enzo Collaboration: Greg L. Bryan\altaffilmark{1}}
 \author{Michael L. Norman\altaffilmark{2,3}}
@@
 \author{Elizabeth Tasker\altaffilmark{17}}
-\author{Stephen Skory\altaffilmark{2}}
-\author{Christine Simpson\altaffilmark{1}}
+\author{Stephen Skory\altaffilmark{10}}
+\author{Christine M. Simpson\altaffilmark{1}}
 \author{Oliver Hahn\altaffilmark{18}}
@@
 \altaffiltext{7}{Center for Relativistic Astrophysics, School of Physics, Georgia Institute of Technology, 837 State Street, Atlanta, GA, USA}
-\altaffiltext{8}{Southern Methodist University, 6425 Boaz Ln, Dallas, TX 75205, USA}
+\altaffiltext{8}{Department of Mathematics, Southern Methodist University, Box 750156, Dallas, TX 75205-0156, USA}
 \altaffiltext{9}{Theoretical Division, Los Alamos National Laboratory, Los Alamos, NM 87544}
@@
 \begin{abstract}
-This paper describes the open-source code \enzo, which uses a
+This paper describes the open-source code \enzo, which uses 
 block-structured adaptive mesh refinement to provide high spatial and

### 2013-06-20 numerical-ppm.tex
Mike Norman's edits.

@@
 the MUSCL-Dedner and MHDCT schemes (including expansion terms,
-reconstruction, Riemann solvers, and dual energy formalism);
+reconstruction, Riemann solvers, and dual energy formalism).
 

### 2013-06-20 numerical-zeus.tex
Mike Norman's edits.

@@
 
-The \zeus\ method uses a von Neumann-Richtmeyer artificial viscosity to smooth shock discontinuities that may appear in fluid flows and can cause a breakdown of the finite-difference equations.
+The \zeus\ method uses a von Neumann-Richtmyer artificial viscosity to smooth shock discontinuities that may appear in fluid flows and can cause a breakdown of the finite-difference equations.
 The artificial viscosity term is added in the source terms as:
@@
 This requires interpolating each cell-centered quantity to the cell edge.
-As recomended in \citet{Stone92a}, we use the second-order van Leer scheme, which uses piecewise linear functions.
+As recommended in \citet{Stone92a}, we use the second-order van Leer scheme, which uses piecewise linear functions.
 These are given by equations (48) and (49) of \citet{Stone92a}.

### 2013-06-20 overview.tex
Mike Norman's edits.

@@
 options also H$^-$, H$_2$, H$_2^+$, and HD, D, and D$^+$. Lastly,
-\enzo\ can also track the advection of a comoving metal density field,
+\enzo\ can also track the advection of one or more comoving metal density fields,
 which can contribute to the radiative cooling and star formation
@@
 The goal in this section is to introduce the reader to the basic
-principles of the methods without drowning in detail.
+principles of the methods without drowning them in detail.
 
@@
 
-As described in \citet{Bryan95}, we modify the method for use in
+As described in \citet{1995CoPhC..89..149B}, we modify the method for use in
 hypersonic flows when the thermal energy $e$ is extremely small
@@
 divergence constraint, $\div \vecB = 0$, to machine precision
-through the Constrained Transport (CT) method.  CT, originally
+through the Constrained Transport (CT) method \citep{Collins10}.  CT, originally
 described by \citet{Evans88}, updates the magnetic field with the curl
@@
 is based on an adaptive ray-tracing method utilizing the HEALPix
-pixelization of a sphere \citep{AbelWandelt02}. Photons are integrated
+pixelization of a sphere \citep{Abel02_RT}. Photons are integrated
 outwards from sources using an adaptive timestepping scheme that
@@
 transitions smoothly between streaming (optically thin) and opaque
-limits and is coupled to a hydrogen-only ionization network.  The
+limits and is coupled to an ionization network of either purely
+hydrogen, or both hydrogen and helium.  The
 resulting set of linear equations is solved using the parallel HYPRE
@@
 metals and thermal energy are injected into the region surrounding the
-star particle.
+star particle.  A related set of methods involves the simulation of
+single Population III stars, rather than ensembles, and is calibrated
+by \textit{ab initio} simulations of primordial star formation.
 

### 2013-06-20 table_of_variables_and_constants.tex
Mike Norman's edits.

@@
   $c_{\rm s}$ & sound speed \\
+  $z$ & redshift \\
   \enddata

### 2013-06-20 tests-zeldovichpancake.tex
Mike Norman's edits.

@@
 including a mere 8 cells resolve the key features, as shown in
-Sections 3.3.4-3.3.5 of~\citet{1996PhDT........80B} -- but we choose a
+Sections 3.3.4-3.3.5 of~\citet{BryanThesis96} -- but we choose a
 higher resolution here for illustrative purposes.

### 2013-06-20 parallel.tex
updated GPU section a bit

@@
 To illustrate the speedup provided by porting two solvers to GPUs, we
-show the results of a weak scaling test of driven turbulence in
-Figure~\ref{fig:gpu_scaling}.  This \enzo\ problem type uses the
-Dedner MHD solver, which runs on both CPUs and GPUs, and has been
-tuned to maximize performance on both platforms.  The benchmarking platform is a Cray
-XK6 system where each node has a 16-core AMD Opteron 6272 CPU and a single
-NVIDIA Tesla K20 GPU, which has 2,496 CUDA cores and a maximum
-theoretical speed of 1.17 Tflops when used in double precision.
-For the CPU run, all the 16 cores in each node
-are used by launching 16 MPI processes per node. For the GPU run, 16 MPI
-processes are also launched per node, and all processes on this node share the single GPU
+show the results of a weak scaling test of driven MHD turbulence on a
+uniform mesh in Figure~\ref{fig:gpu_scaling}.  This particular problem
+type contains no physics other than solving the equations of ideal
+MHD, and thus is representative of the type of calculation than can
+benefit from the GPU-optimized solvers in \enzo.  In this scaling
+test, we use the Dedner MHD solver, which runs on both CPUs and GPUs
+and has been tuned to maximize performance on both platforms.  The
+benchmarking platform is a Cray XK6 system with a Gemini
+interconnect. Each node has a 16-core AMD Opteron 6272 CPU and a
+single NVIDIA Tesla K20 GPU, which has 2,496 CUDA cores and a maximum
+theoretical speed of 1.17 Tflops when used in double precision.  For
+the CPU run, all the 16 cores in each node are used by launching 16
+MPI processes per node. For the GPU run, 16 MPI processes are also
+launched per node, and all processes on this node share the single GPU
 on that node. This can work because NVIDIA's MPS (Multi-Process
 Service) technology allows multiple processes to concurrently use the
-same GPU.  A series of weak scaling tests were run, varying from 1
-to 8 nodes (16 to 128 MPI processes), with the number of cells per
-node fixed to $256^3$.  Figure~\ref{fig:gpu_scaling} displays the
-scaling results in terms of cell updates per second per node -- as a
-result, ideal weak scaling is a horizontal line.  In general, the
-simulations using the GPU-accelerated MHD solver perform roughly 5
-times better overall on this system than the CPU-only calculations.
+same GPU.  A series of weak scaling tests were run, varying from 1 to
+8 nodes (16 to 128 MPI processes), with each node containing a cube of
+$256^3$ cells (so, a 4-node computation would have a domain that is
+$512 \times 512 \times 256$ cells).  Figure~\ref{fig:gpu_scaling}
+displays the scaling results in terms of cell updates per second per
+node -- as a result, ideal weak scaling is a horizontal line.  In
+general, the simulations using the GPU-accelerated MHD solver perform
+roughly 5 times better overall on this system than the equivalent
+CPU-only calculation run on the same number of nodes.  We note,
+however, that this somewhat mis-represents the speedup obtained by
+porting the MHD code to GPUs -- in the GPU simulation case, the 16 CPU
+cores on each node are effectively idle, as they are used only for
+timestep calculation and boundary condition transmission (which are
+both computationally inexpensive compared to the fluid solve).  This
+means that, effectively, the GPU simulation is 80 times faster than a
+single CPU core, so a user would likely see much greater effective
+speedup on a system where the ratio of CPU cores to GPUs is lower.
 

### 2013-06-24 numerical-gravity.tex
Jeff Oishi's changes.
Changed all "X order" to "x-order" and "finite-difference" and "finite-volume" to "finite difference" and "finite volume", to ensure consistency.

@@
 To compute the potential field from this gravitating mass field on the root grid, we use a a fast Fourier transform.
-For periodic boundary conditions, we can use either a simple Greens function kernel of $-k^{-2}$, or the finite-difference equivalent \citep{Hockney88}:
+For periodic boundary conditions, we can use either a simple Greens function kernel of $-k^{-2}$, or the finite difference equivalent \citep{Hockney88}:
 \begin{equation}
-G(\vec{k}) = - \frac{\Delta x}{2 \left( \sin(k_x \Delta x/2)^2 + \sin(k_y \Delta y/2)^2 + \sin(k_z \Delta z/2)^2 \right) }
+G(\myvec{k}) = - \frac{\Delta x}{2 \left( \sin(k_x \Delta x/2)^2 + \sin(k_y \Delta y/2)^2 + \sin(k_z \Delta z/2)^2 \right) }
 \end{equation}
@@
 The generation of the total gravitating mass field is essentially identical, using CIC interpolation for both the particles and baryons, including sub-grids as before.
-To compute the potential on subgrids, however, we use the standard seven-point (in three dimensions) second-order finite-difference approximation to Poisson's equation.
+To compute the potential on subgrids, however, we use the standard seven-point (in three dimensions) second-order finite difference approximation to Poisson's equation.
 Boundary conditions are then interpolated from the potential values on the parent grid.

### 2013-06-24 numerical-mhd-ct.tex
Changed all "X order" to "x-order" and "finite-difference" and "finite-volume" to "finite difference" and "finite volume", to ensure consistency.

@@
 The base Godunov method, described in \citet{Li08a}, uses spatially and
-temporally second order reconstruction (both MUSCL-Hancock and
+temporally second-order reconstruction (both MUSCL-Hancock and
 Piecewise Linear Method), and a selection of Riemann solvers including
@@
 are the first-order method described by \citet{Balsara99} and the
-second order methods described in \citet{Gardiner05}.  The AMR
+second-order methods described in \citet{Gardiner05}.  The AMR
 machinery is described by \citet{Balsara01} and \citet{Collins10}.

### 2013-06-24 numerical-overview.tex
Changed all "X order" to "x-order" and "finite-difference" and "finite-volume" to "finite difference" and "finite volume", to ensure consistency.

@@
 
-These equations are finite-differenced and for simplicity are solved with the
+These equations are finite differenced and for simplicity are solved with the
 same timestep as the hydrodynamic equations, as discussed in 
@@
 point within a subgrid.  Forces are computed on the mesh by
-finite-differencing potential values and are then interpolated to the particle
+finite differencing potential values and are then interpolated to the particle
 positions, where they are used to update the particle's position and velocity
@@
 PPM is an explicit, higher order-accurate version of
-Godunov's method for ideal gas dynamics with third order-accurate piecewise parabolic
+Godunov's method for ideal gas dynamics with third-order-accurate piecewise parabolic
 monotonic interpolation and a nonlinear Riemann solver for shock
@@
 most two cells.  Multidimensional schemes are built up by directional
-splitting and produce a method that is formally second order-accurate
+splitting and produce a method that is formally second-order-accurate
 in space and time which explicitly conserves mass, linear momentum, and energy.
@@
 As a check on PPM, \enzo\ also includes an implementation of the
-finite-difference hydrodynamic algorithm employed in the compressible 
+finite difference hydrodynamic algorithm employed in the compressible 
 magnetohydrodynamics code `\zeus'
@@
 (operator split) solution procedure which is fully explicit in time.  
-This method is formally second order-accurate in space but 
-first order-accurate in time.  It is important to note that 
+This method is formally second-order-accurate in space but 
+first-order-accurate in time.  It is important to note that 
 this method conserves internal energy
@@
 to smooth shock discontinuities that may
-appear in fluid flows and can cause a break-down of finite-difference
+appear in fluid flows and can cause a breakdown of finite difference
 equations.  The artificial viscosity term is added in the source terms

### 2013-06-24 numerical-ppm.tex
Changed all "X order" to "x-order" and "finite-difference" and "finite-volume" to "finite difference" and "finite volume", to ensure consistency.

@@
 We first construct monotonic piecewise 
-parabolic (third order) interpolations in one dimension
+parabolic (third-order) interpolations in one dimension
 for each of $p$, $\rho$, and $v$.  The pressure is determined from
@@
 they also have more intuitive meanings.  For example, $q_{L,j}$ is the
-value of $q$ at the left edge of zone j, while $\Delta q_j$ and $q_{6,j}$ are analogous to the slope and first order correction to the slope of $q$ (see \citet{1984JCoPh..54..174C} for a complete discussion):
+value of $q$ at the left edge of zone j, while $\Delta q_j$ and $q_{6,j}$ are analogous to the slope and first-order correction to the slope of $q$ (see \citet{1984JCoPh..54..174C} for a complete discussion):
 \begin{equation}
@@
 (see below for more detail about the Riemann solvers used), producing estimates for 
-$\overline{p}_{j\pm 1/2}$, $\overline{\rho}_{j\pm1/2}$, and $\overline{v}_{j\pm 1/2}$ that are third order
-accurate in space and second order accurate in time.  These are then
+$\overline{p}_{j\pm 1/2}$, $\overline{\rho}_{j\pm1/2}$, and $\overline{v}_{j\pm 1/2}$ that are third-order
+accurate in space and second-order accurate in time.  These are then
 used to solve the difference Equations (\ref{eq:mass_diff})--(\ref{eq:energy_diff}) for $\rho^{n+1}$, $v^{n+1}$, and $E^{n+1}$.
@@
 by expanding the density and acceleration distributions and
-retaining all terms up to second order in $\Delta x$:
+retaining all terms up to second-order in $\Delta x$:
 %
@@
 
-On rare occasions, high order solutions can cause negative densities or energies.  
+On rare occasions, high-order solutions can cause negative densities or energies.  
 Both our PPM and MUSCL solvers employ a Riemann

### 2013-06-24 numerical-radiation-fld.tex
Jeff Oishi's changes.
Mike Norman's edits.
Changed all "X order" to "x-order" and "finite-difference" and "finite-volume" to "finite difference" and "finite volume", to ensure consistency.

@@
 solution approach may be found in
-\cite{NBHBROW2007}, \cite{ReynoldsHayesPaschosNorman2009} and
-\cite{NRS2009}, the salient features of which are reproduced here. 
+\citet{NBHBROW2007}, \citet{ReynoldsHayesPaschosNorman2009},
+\citet{NRS2009}, and \citet{2013arXiv1306.0645N}, the salient features of which are reproduced here. 
 In addition, comparisons of this solver with other astrophysical
@@
 %  \label{eq:fld_radiation}
-%  \partial_t E_r + \frac1a \div \left(E_r \vec{v}_b\right) &=& 
+%  \partial_t E_r + \frac1a \div \left(E_r \myvec{v}_b\right) &=& 
 %    \div\left(D\grad E_r\right) -
@@
 %  \partial_t {\mathrm n}_i + 
-%    \frac1a \div\left({\mathrm n}_i \vec{v}_b\right) &=& 
+%    \frac1a \div\left({\mathrm n}_i \myvec{v}_b\right) &=& 
 %    \alpha_{i,j} {\mathrm n}_e {\mathrm n}_j - {\mathrm n}_i
@@
 %through first assuming a fixed frequency spectrum, i.e.
-%$E_{\nu}(\nu,\vec{x},t) = \tilde{E}_r(\vec{x},t) \chi(\nu)$, and then
+%$E_{\nu}(\nu,\myvec{x},t) = \tilde{E}_r(\myvec{x},t) \chi(\nu)$, and then
 %defining the integrated quantity
@@
 %\label{eq:grey_radiation_energy}
-%   E_r(\vec{x},t) \equiv \int_{\nu_0}^{\infty}
-%   E_{\nu}(\nu,\vec{x},t)\,\mathrm d\nu \  = \ 
-%   \tilde{E}_r(\vec{x},t) \int_{\nu_0}^{\infty} \chi(\nu)\,\mathrm d\nu.
+%   E_r(\myvec{x},t) \equiv \int_{\nu_0}^{\infty}
+%   E_{\nu}(\nu,\myvec{x},t)\,\mathrm d\nu \  = \ 
+%   \tilde{E}_r(\myvec{x},t) \int_{\nu_0}^{\infty} \chi(\nu)\,\mathrm d\nu.
 %\end{equation}
@@
   n}_i^{n})$.  This uses an implicit Euler time discretization, and a
-second-order centered finite-difference spatial discretization,
+second-order centered finite difference spatial discretization,
 resulting in a large linear system of equations.  These are solved

### 2013-06-24 numerical-zeus.tex
Changed all "X order" to "x-order" and "finite-difference" and "finite-volume" to "finite difference" and "finite volume", to ensure consistency.

@@
 
-As an alternative to the previous Godunov methods, \enzo\ also includes an implementation of the finite-difference hydrodynamic algorithm employed in the compressible magnetohydrodynamics code \zeus\ \citep{Stone92a, Stone92b}.
+As an alternative to the previous Godunov methods, \enzo\ also includes an implementation of the finite difference hydrodynamic algorithm employed in the compressible magnetohydrodynamics code \zeus\ \citep{Stone92a, Stone92b}.
 Fluid transport is solved on a Cartesian grid using the upwind, monotonic advection scheme of \citet{1977JCoPh..23..276V} within a multistep (operator split) solution procedure that is fully explicit in time.
-This method is formally second order-accurate in space but first order-accurate in time.
+This method is formally second-order-accurate in space but first-order-accurate in time.
 
@@
 
-The \zeus\ method uses a von Neumann-Richtmyer artificial viscosity to smooth shock discontinuities that may appear in fluid flows and can cause a breakdown of the finite-difference equations.
+The \zeus\ method uses a von Neumann-Richtmyer artificial viscosity to smooth shock discontinuities that may appear in fluid flows and can cause a breakdown of the finite difference equations.
 The artificial viscosity term is added in the source terms as:
@@
 Once again, we dimensionally split the equations and present only the one-dimensional version.
-The finite-difference equations actually solved are:
+The finite difference equations actually solved are:
 \begin{equation}

### 2013-06-24 overview.tex
Changed all "X order" to "x-order" and "finite-difference" and "finite-volume" to "finite difference" and "finite volume", to ensure consistency.

@@
 transport staggered MHD scheme \citep{Collins10}, and (iv) the
-second-order finite-difference hydrodynamics method described in
+second-order finite difference hydrodynamics method described in
 \zeus~\citep{Stone92a,Stone92b}.
@@
 explicit, higher-order-accurate version of Godunov's method for ideal
-gas dynamics with the spatially third order accurate piecewise parabolic monotonic
+gas dynamics with the spatially third-order-accurate piecewise parabolic monotonic
 interpolation and a nonlinear Riemann solver for shock capturing.  It
@@
 shocks across a few cells.  Multidimensional schemes are built up by
-directional splitting and produce a method that is formally second
-order-accurate in space and time and which explicitly conserves mass,
+directional splitting and produce a method that is formally second-order-accurate in space and time and which explicitly conserves mass,
 linear momentum, and energy \citep{Hawley84, Norman86}.  A variety of
@@
 \paragraph{Godunov MHD with Constrained Transport (MHD)}
-This MHD method is second order in time and space, and preserves the
+This MHD method is second-order in time and space, and preserves the
 divergence constraint, $\div \vecB = 0$, to machine precision
@@
 \vecB$.  We employ the CT methods described by \citet{Balsara99} and
-\citet{Gardiner05} with the second order hyperbolic
+\citet{Gardiner05} with the second-order hyperbolic
 solver of \citet{Li08a} and the divergence free AMR scheme of
@@
 
-\paragraph{Second order finite difference method (HD only)}
+\paragraph{Second-order finite difference method (HD only)}
 
-Lastly, we briefly describe the \zeus\ method, a finite-difference
+Lastly, we briefly describe the \zeus\ method, a finite difference
 algorithm originally used in the \zeus\ code \citep{Stone92a}. Note

### 2013-06-24 introduction.tex
Changes from Cameron Hummels comments (also some other typos I found while working through his comments).

@@
 \citep{Slyz05}, the intergalactic medium \citep{Fang01, 
-2011ApJ...731....6S}, cooling flows \citep{Li12, 2013ApJ...763...38S},
+2011ApJ...731....6S}, the circumgalactic
+medium~\citep{2013MNRAS.430.1548H}, cooling flows \citep{Li12, 2013ApJ...763...38S},
 turbulence \citep{Kritsuk04, 2007ApJ...665..416K, 2009JPhCS.180a2020K, Collins11}, the

### 2013-06-24 numerical-amr.tex
Changes from Cameron Hummels comments (also some other typos I found while working through his comments).

@@
 A grid's \textit{parent} is the coarse grid that completely contains it.
-The \textit{root grid} is the most coarsely-resolved grid, and has no parent.
-A \textit{child grid} is grid completely contained by its parent grid (note that a grid may have many children but only one parent), while sibling grids are those that have the same resolution, but not necessarily the same parent.
+The \textit{root grid} is the most coarsely-resolved grid, and it has no parent.
+A \textit{child grid} is a grid completely contained by its parent grid (note that a grid may have many children but only one parent), while sibling grids are those that have the same resolution, but not necessarily the same parent.
 We also refer to levels of the hierarchy, where the root grid is labeled level zero and all grids with the same resolution are on the
@@
 
-In the second case, the value must be either (1) interpolated from a parent, or (2) copied from a sibling.
+In the second case, the value must be either (1) interpolated from a parent or (2) copied from a sibling.
 In practice, we solve this problem by first interpolating all boundaries from the grid's parent and then performing the sibling copy wherever possible, since these values will always be as good as, or superior to, interpolation.
@@
 a least-squares fit under the given constraints, however, this is
-computationally too expensive.  Therefore, in Appendix~\ref{app:interpolation} we describe
+too computationally expensive.  Therefore, in Appendix~\ref{app:interpolation} we describe
 a heuristic approach that satisfies four constaints analytically and

### 2013-06-24 numerical-chemistry.tex
Changes from Cameron Hummels comments (also some other typos I found while working through his comments).

@@
 more details.  The chemical and thermal states of the gas can be updated either
-at the same hydrodynamical timestep (i.e., de-coupled and operator split) or
+at the same hydrodynamical timestep (i.e., de-coupled and operator-split) or
 through the same subcycling system (i.e., a coupled chemical and thermal

### 2013-06-24 numerical-gravity.tex
Changes from Cameron Hummels comments (also some other typos I found while working through his comments).

@@
 
-% Changed:  ``Particles from sub-grids are also added in this way'' to ``Particles on sub-grids within the grid's volume are also added to its gravitating field using the same method.'' I expanded because I missed the point and had to clarify with Brian when I read the text.
+% Changed:  ``Particles from subgrids are also added in this way'' to ``Particles on subgrids within the grid's volume are also added to its gravitating field using the same method.'' I expanded because I missed the point and had to clarify with Brian when I read the text.
 But if it's too long, please just cut it!
@@
 During the CIC interpolation, particle positions are (temporarily) advanced by $0.5 \Delta t v^n$ so that we generate an estimate of the time-centered density field.
-Particles on sub-grids within the grid's volume are also added to its gravitating field using the same method.
+Particles on subgrids within the grid's volume are also added to its gravitating field using the same method.
 In addition, since the gravitating field for a grid is defined beyond the grid edges (see below), massive particles from sibling grids and sub grids that lie within the entire gravitating field are used.
@@
 In order to calculate more accurate potentials on the subgrids, \enzo\ uses a similar but slightly different technique from the root grid.
-The generation of the total gravitating mass field is essentially identical, using CIC interpolation for both the particles and baryons, including sub-grids as before.
+The generation of the total gravitating mass field is essentially identical, using CIC interpolation for both the particles and baryons, including subgrids as before.
 To compute the potential on subgrids, however, we use the standard seven-point (in three dimensions) second-order finite difference approximation to Poisson's equation.
 Boundary conditions are then interpolated from the potential values on the parent grid.
-We use either tri-linear interpolation or a natural second-order spline for this: both methods give similar results but the default is the tri-linear interpolation, which empirically provides a resonable compromise between speed and accuracy.
+We use either tri-linear interpolation or a natural second-order spline for this: both methods give similar results, but the default is the tri-linear interpolation, which empirically provides a resonable compromise between speed and accuracy.
 The potential equation on each subgrid is then solved with the given Dirichlet boundary conditions with a multigrid relaxation technique.

### 2013-06-24 numerical-mhd-ct.tex
Changes from Cameron Hummels comments (also some other typos I found while working through his comments).

@@
 variables, while the rest of \enzo\ employs predominantly
-volume-averaged cell centered variables.  The magnetic field is
-represented by both a face centered field, B$_f$, and a cell centered
-field, B$_c$.  The electric field is edge centered.  The magnetic
+volume-averaged cell-centered variables.  The magnetic field is
+represented by both a face-centered field, B$_f$, and a cell-centered
+field, B$_c$.  The electric field is edge-centered.  The magnetic
 field is updated in four steps: first, the Riemann problem is solved
-in the traditional manner, using the cell centered field; second, an
+in the traditional manner, using the cell-centered field; second, an
 edge-centered electric field is computed using the fluxes from the
 Riemann solver; third, the curl of that electric field is used to
-update the face centered field; finally, the cell centered magnetic
-field is updated with an average of the face centered field.
+update the face-centered field; finally, the cell-centered magnetic
+field is updated with an average of the face-centered field.
 

### 2013-06-24 numerical-mhd-dedner.tex
Changes from Cameron Hummels comments (also some other typos I found while working through his comments).

@@
 
-To more clearly describe the Dedner divergence cleaning modifications, we write the equations of compressible inviscid hydrodynamics can be written in the form of conservation laws as,
+To more clearly describe the Dedner divergence cleaning modifications, we write the equations of compressible inviscid hydrodynamics in the form of conservation laws as,
 \begin{equation}

### 2013-06-24 numerical-overview.tex
Changes from Cameron Hummels comments (also some other typos I found while working through his comments).

@@
 
-PPM is an explicit, higher order-accurate version of
-Godunov's method for ideal gas dynamics with third-order-accurate piecewise parabolic
+PPM is an explicit, higher-order accurate version of
+Godunov's method for ideal gas dynamics with third-order accurate piecewise parabolic
 monotonic interpolation and a nonlinear Riemann solver for shock
@@
 most two cells.  Multidimensional schemes are built up by directional
-splitting and produce a method that is formally second-order-accurate
+splitting and produce a method that is formally second-order accurate
 in space and time which explicitly conserves mass, linear momentum, and energy.
@@
 within a multistep 
-(operator split) solution procedure which is fully explicit in time.  
-This method is formally second-order-accurate in space but 
-first-order-accurate in time.  It is important to note that 
+(operator-split) solution procedure which is fully explicit in time.  
+This method is formally second-order accurate in space but 
+first-order accurate in time.  It is important to note that 
 this method conserves internal energy
@@
 
-Operator split methods break
+Operator-split methods break
 the solution of the hydrodynamic equations into parts, with each part

### 2013-06-24 numerical-particle.tex
Changes from Cameron Hummels comments (also some other typos I found while working through his comments).

@@
 with the finest spatial resolution, which may be of the same spatial
-resolution, coarser, or finer than the grid patch that the particles
-are moved from.  This takes place in a communication process at the
+resolution, coarser, or finer than the grid patch from which the particles
+moved.  This takes place in a communication process at the
 end of each timestep on a level.
@@
 To avoid unphysical point-mass effects, \enzo\ provides a parameter that
-governs the maximum level at which particles will be regarded at point masses.
+governs the maximum level at which particles will be regarded as point masses.
 At higher levels, contributions from particles to the gravitating mass field

### 2013-06-24 numerical-ppm.tex
Changes from Cameron Hummels comments (also some other typos I found while working through his comments).

@@
 piecewise parabolic method (PPM) of
-\citet{1984JCoPh..54..174C}, which has been
+\citet{1984JCoPh..54..174C} -- henceforth referred to as CW84 -- which has been
 modified for the study of cosmological fluid flows.  
-\citet{1984JCoPh..54..174C} describe two variants on this method:
+CW84 describe two variants on this method:
 Lagrangian Remap and Direct Eulerian; here we use the Direct Eulerian
@@
 Remap version has previously been adapted for cosmological use
-as described in \citet{1995CoPhC..89..149B} and the implementation
+as described in \citet{1995CoPhC..89..149B}, and the implementation
 we use here is closely based on that work.
@@
 The first term on the right hand side of equations~(\ref{eq:momentum}) and
-(\ref{eq:total_energy}) come from our choice of comoving coordinates
+(\ref{eq:total_energy}) comes from our choice of comoving coordinates
 (a similar term also appears in equation~(\ref{eq:dm_velocity}), the velocity 
@@
 updated.  This is described in detail in
-\citet{1984JCoPh..54..174C}, but we will briefly outline the procedure
+CW84, but we will briefly outline the procedure
 here both for completeness and to put the changes we will make in
@@
 they also have more intuitive meanings.  For example, $q_{L,j}$ is the
-value of $q$ at the left edge of zone j, while $\Delta q_j$ and $q_{6,j}$ are analogous to the slope and first-order correction to the slope of $q$ (see \citet{1984JCoPh..54..174C} for a complete discussion):
+value of $q$ at the left edge of zone j, while $\Delta q_j$ and $q_{6,j}$ are analogous to the slope and first-order correction to the slope of $q$ (see CW84 for a complete discussion):
 \begin{equation}
@@
 The resulting formulae are straightforward but complicated and are not
-reproduced here, but see Equations 1.7 to 1.10 of \citet{1984JCoPh..54..174C}.
+reproduced here, but see Equations 1.7 to 1.10 of CW84.
 We also optionally allow steepening as described in that reference.
@@
 The linearized gas dynamics equations are then used to compute second-order accurate left and right states that take into account the multiple wave families.
-This process is described in \citet{1984JCoPh..54..174C} and we use their equations (3.6) and (3.7).
+This process is described in CW84 and we use their equations (3.6) and (3.7).
 
@@
 
-We include an optional diffusive flux (and flattening for the parabolic curves) that can improve the solution in some cases.
-Our implementation follows that in the appendix of \citet{1984JCoPh..54..174C}.
+We include an optional diffusive flux (and flattening for the parabolic curves) that can improve the solution in some cases.  Our implementation follows that in the appendix of CW84.
 
-In addition, as discussed earlier, the three dimensional scheme is achieved by operator splitting and repeating the above procedure in the other two orthogonal directions.
-The transverse velocities and any additional passive quantities are naturally and easily added to this system (see Equation 3.6 of \citet{1984JCoPh..54..174C}).
+In addition, as discussed earlier, the three dimensional scheme is achieved by operator-zsplitting and repeating the above procedure in the other two orthogonal directions.
+The transverse velocities and any additional passive quantities are naturally and easily added to this system (see Equation 3.6 of CW84).
 
@@
 acceleration felt by the entire zone and not just at the zone center.
-Therefore, it is possible to find the mass weighted average acceleration over the zone
+Therefore, it is possible to find the mass-weighted average acceleration over the zone
 by expanding the density and acceleration distributions and
@@
 correct physical state is determined, the pressure in the central region can
-only be found by finding the root to an algebraic equation, which is necessarily
+only be determined by finding the root to an algebraic equation, which is necessarily
 an iterative process.  Thus a series of \emph{approximate} Riemann solvers are
@@
 moving waves.  This method is significantly faster than the two-shock method,
-but also quite a bit more dissipative.  The HLLC is a three-wave method that improves upon the HLL
+but it is also quite a bit more dissipative.  The HLLC is a three-wave method that improves upon the HLL
 method by also including the third wave, the contact discontinuity.

### 2013-06-24 numerical-radiation-fld.tex
Changes from Cameron Hummels comments (also some other typos I found while working through his comments).

@@
 a field-based radiation transport solver for problems posed on uniform
-(i.e.~non-AMR, non-SMR) grids, which has been tuned for large-scale simulations
+(i.e.~non-AMR, non-static mesh refinement) grids, which has been tuned for large-scale simulations
 involving many ionizing sources.  Detailed explanations of the model and

### 2013-06-24 numerical-starformation.tex
Jeff Oishi's changes.
Changes from Cameron Hummels comments (also some other typos I found while working through his comments).

@@
 is taken away from the cell in question and a ``star particle'' with
-the same mass is placed in the center of that cell, with the same
+the same mass is placed in the center of that cell with the same
 velocity as the removed gas.  This star particle is then
@@
 assumes that stars form in substantially overdense, converging, and
-gravitationally unstable gas.  Algorithmically, every cell at the
-locally highest level of refinement are examined at every timestep to
-see if they meet the critiera for star formation, and star particles
+gravitationally unstable gas.  Algorithmically, each cell at the
+locally highest level of refinement is examined at each timestep to
+see if it meets the critiera for star formation.  Star particles
 are allowed to form in a cell if the following criteria are met:
@@
 %\begin{equation}
-\div \vec{v}_b & < & 0, \\
+\div \myvec{v}_b & < & 0, \\
 \label{cencont}
@@
 overdensity threshold, 
-%$\vec{v}$ is the velocity of the gas in the
+%$\myvec{v}$ is the velocity of the gas in the
 %cell, $\rho_{d}$ is the dark matter density, $\rho_{\rm tot} =
@@
 
-\item A converging gas flow ($\div \vec{v} < 0$), and
+\item A converging gas flow ($\div \myvec{v} < 0$), and
 

### 2013-06-24 numerical-zeus.tex
Changes from Cameron Hummels comments (also some other typos I found while working through his comments).

@@
 As an alternative to the previous Godunov methods, \enzo\ also includes an implementation of the finite difference hydrodynamic algorithm employed in the compressible magnetohydrodynamics code \zeus\ \citep{Stone92a, Stone92b}.
-Fluid transport is solved on a Cartesian grid using the upwind, monotonic advection scheme of \citet{1977JCoPh..23..276V} within a multistep (operator split) solution procedure that is fully explicit in time.
-This method is formally second-order-accurate in space but first-order-accurate in time.
+Fluid transport is solved on a Cartesian grid using the upwind, monotonic advection scheme of \citet{1977JCoPh..23..276V} within a multistep (operator-split) solution procedure that is fully explicit in time.
+This method is formally second-order accurate in space but first-order accurate in time.
 
-As discussed in the section describing the Piecewise Parabolic Method (Section~\ref{sec.hydro.ppm}), operator split methods break the solution of the hydrodynamic equations into parts, with each part representing a single term in the equations.
+As discussed in the section describing the Piecewise Parabolic Method (Section~\ref{sec.hydro.ppm}), operator-split methods break the solution of the hydrodynamic equations into parts, with each part representing a single term in the equations.
 Each part is evaluated successively using the results preceding it.
-In this method, in addition to operator splitting the expansion terms (i.e.
+In this method, in addition to operator-splitting the expansion terms (i.e.
 those terms in eqs.~(\ref{eq:mass}) -- (\ref{eq:total_energy}) that depend on $\dot{a}$), we divide the remaining terms into \emph{source} and \emph{transport} steps.

### 2013-06-24 overview.tex
Changes from Cameron Hummels comments (also some other typos I found while working through his comments).

@@
 rectangular grid patch (referred to as a ``grid'') covers some region
-of space in its \emph{parent grid} requiring higher resolution,
+of space requiring higher resolution than its \emph{parent grid},
 and can itself become the parent grid to an even more highly resolved
@@
 We begin with the direct-Eulerian PPM implementation.  This is an
-explicit, higher-order-accurate version of Godunov's method for ideal
-gas dynamics with the spatially third-order-accurate piecewise parabolic monotonic
+explicit, higher-order accurate version of Godunov's method for ideal
+gas dynamics with the spatially third-order accurate piecewise parabolic monotonic
 interpolation and a nonlinear Riemann solver for shock capturing.  It
@@
 shocks across a few cells.  Multidimensional schemes are built up by
-directional splitting and produce a method that is formally second-order-accurate in space and time and which explicitly conserves mass,
+directional splitting and produce a method that is formally
+second-order accurate in space and time and which explicitly conserves mass,
 linear momentum, and energy \citep{Hawley84, Norman86}.  A variety of

### 2013-06-24 table_of_variables_and_constants.tex
Changes from Cameron Hummels comments (also some other typos I found while working through his comments).

@@
   $\rho$ & comoving baryonic density \\
-  $\rho_{\!X}$ & mass density of component $X$ (gas, dark matter, stars, metals, total, etc.) \\
+  $\rho_{\!X}$ & comoving mass density of component $X$ (gas, dark matter, stars, metals, total, etc.) \\
   $t$ & time \\

### 2013-06-24 introduction.tex
Typos/tweaks for sections 1 and 2 (intro, overview).

@@
 resolution \citep[e.g.,][]{Gnedin95, Xu97, Pen98}, and more recently a
-few grid codes have adopted an unstructured approach based on a moving
+few codes have adopted an unstructured approach based on a moving
 Voronoi mesh \citep{Arepo10, Tess11}.
 
-In this paper, we present \enzo; a structured adaptive mesh refinement
+In this paper, we present \enzo, a structured adaptive mesh refinement
 (SAMR) code. Originally developed for cosmological hydrodynamics,
@@
 galaxies \citep{2003ApJ...587...13T, 2012MNRAS.425..641L,
-  2012ApJ...749..140H, 2013MNRAS.432.1989S}, galaxy
+  2013MNRAS.432.1989S}, galaxy
 clusters \citep{Loken02, Xu11, Skillman13}, the interstellar medium
@@
 and refine those methods so that they can be critically evaluated and
-expanded upon by others.  Our public release of the \enzo\ code represents our
+expanded upon by others.  Our public release of the \enzo\ code (which
+can be found at \url{http://enzo-project.org}) represents our
 attempt to meet the first of these obligations; this paper represents

### 2013-06-24 ms.tex
Typos/tweaks for sections 1 and 2 (intro, overview).

@@
 temporal resolution for modeling astrophysical fluid flows.  The code
-is Cartesian and can be run in 1, 2, and 3 dimensions, and supports a
+is Cartesian, can be run in 1, 2, and 3 dimensions, and supports a
 large variety of physics, including hydrodynamics, ideal and non-ideal

### 2013-06-24 overview.tex
Typos/tweaks for sections 1 and 2 (intro, overview).

@@
 historically the code was written with cosmological situations in mind
-(e.g. galaxy clusters), in more recent years it has been used to
+(e.g., galaxy clusters), in more recent years it has been used to
 simulate a much broader range of astrophysical environments.
@@
 
-Any collisionless components (e.g. dark matter and stars) are modeled
+Any collisionless components (such as dark matter and stars) are modeled
 by N-body particles, whose dynamics are governed by Newton's equations
@@
 
-The code can include either a homogeneous radiation background, 
+The code can include either a homogeneous radiation background 
 or evolve an inhomogeneous
@@
 As an alternative to the ray-casting strategy, \enzo\ can also solve
-the angle-averaged radiative transfer equation, through a
-flux-limited diffusion approximation with couplings to both the gas
+the angle-averaged radiative transfer equation through a
+flux-limited diffusion approximation, with couplings to both the gas
 energy and chemical number densities. The equations solved are
@@
 total opacity, $\eta$ is the field of radiation sources, $\Gamma$ is
-the radiation induced photo-heating rate, and $\Lambda$ is the
+the radiation-induced photo-heating rate, and $\Lambda$ is the
 chemical cooling rate.
@@
 \begin{equation}
-\fcond= -\kappa_{\rm cond} \, {\hat \textbf{b}} (\hat{\textbf{b}} \cdot \grad T),
+\fcond= -\kappa_{\rm cond} \,\myvec{b} ( \myvec{b} \cdot \grad T),
 \end{equation}
-where $\hat{\textbf{b}}$ is the unit vector denoting the direction of
+where $\myvec{b}$ is the unit vector denoting the direction of
 the magnetic field.
@@
 
-In this section we briefly describe the numerical methods which are
+In this section we briefly describe the numerical methods that are
 used to solve the equations outlined in Section~\ref{sec.equations}.
@@
 The primary feature of the \enzo\ code is its Adaptive Mesh Refinement
-(AMR) capability, which allows it to reach extremely large dynamical
+(AMR) capability, which allows it to reach extremely large spatial and
+temporal dynamical
 ranges with limited computational resources, opening doors to applications otherwise
@@
 continue advancing the grids on the lowest level until they have
-``caught up'' to the coarser parent grids (i.e. $\sum \Delta t_l =
+``caught up'' to the coarser parent grids (i.e., $\sum \Delta t_l =
 \Delta t_{l-1}$).  This procedure repeats itself until all grids have
@@
 \citet{Gardiner05} with the second-order hyperbolic
-solver of \citet{Li08a} and the divergence free AMR scheme of
+solver of \citet{Li08a} and the divergence-free AMR scheme of
 \citet{Balsara01}.
@@
 The current implementation of self-gravity in \enzo\ uses a Fast
-Fourier Technique \citep{Hockney88} to solve Poisson's equation on the
+Fourier technique \citep{Hockney88} to solve Poisson's equation on the
 root grid on each timestep.  The advantage of using this method is
@@
 rebuilt. Thus a particle has the same timestep and feels the same
-gravitational force as a grid at that refinement level.
+gravitational force as a grid cell at that refinement level.
 
@@
 initial spatial and baryonic mass resolution, and these subgrid are
-populated with lower mass particles to correspondingly improve the
+populated with lower-mass particles to correspondingly improve the
 collisionless mass resolution.
@@
 \citet{SW87}.  If chemistry is turned on, then the code can include
-cooling from all species of hydrogen and helium (including H$_2$) --
+cooling from all species of hydrogen and helium (including H$_2$ and
+Deuterium-related species such as HD) --
 and the primordial cooling rates are computed in the same Jacobi iteration as the
@@
 star particle.  A related set of methods involves the simulation of
-single Population III stars, rather than ensembles, and is calibrated
+single Population III stars rather than ensembles, and is calibrated
 by \textit{ab initio} simulations of primordial star formation.

### 2013-06-24 numerical-amr.tex
edited section 3

@@
 The active zones store field values and ghost zones are used to temporarily store values that have been obtained directly from neighboring grids or interpolated from a parent grid.
-These zones are necessary to accommodate the computational stencil of the (magneto)hydrodynamics solvers (Sections~\ref{sec.hydro.ppm} and~\ref{sec.hydro.zeus}) and the gravity solver (Section~\ref{sec.gravity}).
+These zones are necessary to accommodate the computational stencil of the (magneto)hydrodynamics solvers (Sections~\ref{sec.hydro.ppm} through~\ref{sec.hydro.zeus}) and the gravity solver (Section~\ref{sec.gravity}).
 The PPM and Zeus hydro solvers require 3 layers of ghost zones,  and the gravity solver requires 6 (although only for the density and potential fields).
@@
 timestep.
-Then one timestep is taken on all grids at the next level of refinement, $l+1$, and so on until the finest level is resolved.
+Then, one timestep is taken on all grids at the next level of refinement, $l+1$, and so on until the finest level is resolved.
 The finest level is then advanced, using as many steps as it takes to reach the level immediately above.
@@
 The AMR control algorithm is contained within the recursive
-EvolveLevel algorithm, which takes the level on which to operate as an
+EvolveLevel algorithm, which takes the level upon which to operate as an
 argument.  This looks something like the following pseudo-code:
@@
 In one dimension, such functions are well known; examples include
-donor cell, Van Leer and piecewise parabolic interpolation, which are
-first, second and third-order accurate respectively
+donor cell, Van Leer, and piecewise parabolic interpolation, which are
+first-, second-, and third-order accurate respectively
 \citep[e.g.,][]{Stone92a}.  However, we require fully
@@
 parameters (four).  One way to solve this problem would be to perform
-a least-squares fit under the given constraints, however, this is
+a least-squares fit under the given constraints; however, this is
 too computationally expensive.  Therefore, in Appendix~\ref{app:interpolation} we describe
@@
 
-\item{\em Jeans length.}  Following \cite{Truelove98}, the code can be forced to refine the Jeans length by a fixed number of cells, or more specifically, whenever the following criterion is not met:
+\item{\em Jeans length.}  Following \cite{Truelove98}, the code can be forced to refine the Jeans length by a fixed number of cells, or, more specifically, whenever the following criterion is not met:
 \begin{equation}
-\Delta x < \left( \frac{\pi k_B T}{J^2 G \rho \mh} \right)^{1/2},
+\Delta x < \left( \frac{\pi k_B T}{N_J^2 G \rho \mh} \right)^{1/2},
 \end{equation}
-where $J$ is the required number of cells per Jeans length (4 by default).
+where $J_J$ is the required number of cells per Jeans length (4 by default).
 
@@
 However, in some cases it is desirable to resolve the cooling process, in which case we flag cells that have a cooling time shorter than their sound crossing time.
-The cooling time is given by $t_{\rm cool} = 1.5 kT / (n\, \Lambda(T))$ and the sound crossing time is $t_{\rm cross} = \Delta x / c_s$, where $c_s$ is the sound speed.
+The cooling time is given by $t_{\rm cool} = 1.5 k_BT / (n\, \Lambda(T))$ and the sound crossing time is $t_{\rm cross} = \Delta x / c_s$, where $c_s$ is the sound speed.
 
@@
 \end{equation}
-is always refined by at least N cells locally (i.e.~we refine a given cell further if $L_R \leq N \Delta x$).
+is always refined by at least N$_R$ cells locally (i.e.~we refine a given cell further if $L_R \leq N_R \Delta x$).
 Note that \enzo\ does not actually solve the resistive MHD equations -- this is the ``characteristic'' local resistive length, rather than one calculated within the MHD solver.
@@
 Finally, it is also possible to specify a rectangular region that is statically refined to a certain level.
-These static refined regions are in place throughout the simulation regardless of the grid quantities and can be used for testing purposes, or for multiply-zoomed initial conditions (for example in cosmology simulations) that would not otherwise be refined.
+These static refined regions are in place throughout the simulation regardless of the grid quantities and can be used for testing purposes, or for multiply-zoomed initial conditions (in cosmology simulations, for example) that would not otherwise be refined.
 

### 2013-06-24 numerical-amr.tex
Added one more refinement method (which Sam Skillmin *promises* me will be put into enzo-dev before the paper appears on the arXiv).

@@
 
+\item{\em Second Derivative.}  This is similar to the refinement criteria used
+  in Paramesh and was originally developed by \cite{1987nasa.reptQ....L}.  This
+  method calculates the second derivative of a user-defined field normalized by
+  the first derivative, and sets a threshold for refinement.  More background
+  information can be found in the FLASH4 user guide\footnote{\url{http://flash.uchicago.edu/site/flashcode/user\_support/flash4\_ug/}},
+  but we reproduce the relevant equation here for completeness. The
+  multi-dimensional version of the criteria takes in to account all
+  cross-derivatives, and is expressed by
+
+ \begin{equation}
+   S = \left\{ \frac{\sum_{ij}\left( \frac{\partial^2f}{\partial x_i \partial x_j} \right)^2}{\sum_{ij}\left[ \frac{1}{2\Delta x_j}\left( \abs{\frac{\partial f}{\partial x_i}}_{x_i+\Delta x_i} + \abs{\frac{\partial f}{\partial x_i}}_{x_i-\Delta x_i} \right) + \epsilon \frac{\abs{\bar{f_{ij}}}}{\Delta x_i \Delta x_j} \right]^2} \right\}^{1/2},
+   \label{eq:second-deriv_analytical}
+ \end{equation}
+
+  where $f$ is the user-defined field, $\abs{\bar{f_{ij}}}$ is an average of
+  $f$ over cells in the $i$ and $j$ dimensions, and $\epsilon$ is
+  a user-controlled value that provides a method of ignoring small
+  fluctuations, set by default to $10^{-2}$.  If $S$ is larger than a specified
+  threshold (bounded by [0,1]) a cell is marked for refinement. This threshold
+  is chosen at runtime and can vary for each of the fields under consideration.
+
 \end{enumerate}

### 2013-06-24 numerical-ppm.tex
Edits to section 4.1 (PPM)

@@
 CW84 describe two variants on this method:
-Lagrangian Remap and Direct Eulerian; here we use the Direct Eulerian
-version which is better suited for AMR simulations.  The Lagrangian
+Lagrangian Remap and Direct Eulerian.  We use the Direct Eulerian
+version in \enzo, which is better suited for AMR simulations.  The Lagrangian
 Remap version has previously been adapted for cosmological use
@@
 eliminated entirely by the proper choice of variables (including time),
-although we have not done so here, as they do not constitute a major
+although we have not done so here as they do not constitute a major
 source of error.  We solve these terms
@@
 We now restrict ourselves to the solution of equations
-(\ref{eq:mass1d})--(\ref{eq:total_energy1d}), in one
+(\ref{eq:mass1d})--(\ref{eq:total_energy1d}) in one
 dimension.  In the Direct Eulerian version of PPM, this is
@@
 %
-This is equation~(1.4) of CW84 (in the spatial rather than mass coordinate as is appropriate for the direct Eulerian approach). 
+This is equation~(1.4) of CW84 (in the spatial rather than mass coordinate, as is appropriate for the direct Eulerian approach). 
 The quantities $q_{L,j}$, $\Delta q_j$,
@@
 
-Once we have the reconstruction, the primary quantities ($p$, $\rho$, $v$ and $E$) are averaged over the domains corresponding to the three characteristics $C = u-c$, $u$ or $u+c$ (where $c$ is the sound speed in a cell):
+Once we have the reconstruction, the primary quantities ($p$, $\rho$, $v$ and $E$) are averaged over the domains corresponding to the three characteristics $C = u-c$, $u$, or $u+c$ (where $c$ is the sound speed in a cell):
 \begin{equation}
@@
 
-In addition, as discussed earlier, the three dimensional scheme is achieved by operator-zsplitting and repeating the above procedure in the other two orthogonal directions.
+In addition, as discussed earlier, the three-dimensional scheme is achieved by operator-splitting and repeating the above procedure in the other two orthogonal directions.
 The transverse velocities and any additional passive quantities are naturally and easily added to this system (see Equation 3.6 of CW84).
@@
 high as $10^8$.  Inverted, we see that the internal energy consists of
-an extremely small portion of the total energy.  The pressure then,
+an extremely small portion of the total energy.  In such a situation, the pressure,
 proportional to $E - E_{k}$, is the small difference between two large
@@
 It is necessary, however, to conserve the total energy so that the conversion
-of kinetic to thermal energy is performed properly.   We must therefore,
-combine the two formulations without allowing the separately advected
+of kinetic to thermal energy is performed properly.   We must therefore
+combine the two formulations without allowing the separately-advected
 internal energy $e$ to play a role in the gas dynamics.  This is done by
@@
 distinct scheme for dealing with this problem, which is endemic to
-methods adopting the total energy equation.  In \citet{TVD93} the 
+methods adopting the total energy equation.  In \citet{TVD93}, the 
 two variables adopted are total energy and
@@
 with the PPM method.  HLL and HLLC are used with PPM, MUSCL (both with and
-without MHD) and MHD-CT.  HLLD is exclusively an MHD solver, and works with both
+without MHD) and MHD-CT.  HLLD is exclusively an MHD solver, and works
+with both the
 MUSCL and MHD-CT methods.
 
-The only approximation that two-shock makes is that both left- and right- moving
+The only approximation that two-shock makes is that both left- and right-moving
 waves are shocks.  This solution still requires an iterative method for finding
@@
 
-For the MHD equations, there are seven waves, instead of three.  This makes the
-exact solution to the Riemann quite a bit more expensive.  Both the HLL and HLLC
+For the MHD equations, there are seven waves instead of three.  This makes the
+exact solution to the Riemann problem quite a bit more expensive.  Both the HLL and HLLC
 approximations can be formulated for the MHD equations, and are employed in both

### 2013-06-24 numerical-mhd-dedner.tex
edited dedner MHD section

@@
 for all dimensions first and the conserved quantities are updated in
-one step, in contrast to the Strang splitting employed in the other solvers.
+one step, in contrast to the Strang splitting employed in the other
+fluid methods described in this paper.
 Also unlike the other schemes, time-integration is done with a second-order
@@
 
-Finally, we note that for cosmological simulations, this solver uses a slightly different definition of the magnetic field than used in the rest of the paper.
-In particular, the field is defined as $\vecB = a^{3/2} \vecB^{\prime}$ (where $\vecB^{\prime}$ is the proper field strength).
-This adds a source term of $-\dot{a}/(2a) \vecB$ on the right-hand side of equation~\ref{eq:induction}, removes the factor of $a$ in the $\vecB$ term in the energy equation (eq.~\ref{eq:total_energy_def}) and in the definition of the isotropic pressure $p^*$.
+Finally, we note that for cosmological simulations, this solver uses a
+slightly different definition of the magnetic field than used in the
+rest of the paper.  In particular, the field is defined as $\vecB =
+a^{3/2} \vecB^{\prime}$ (where $\vecB^{\prime}$ is the proper field
+strength).  This adds a source term of $-\dot{a}/(2a) \vecB$ on the
+right-hand side of Equation~\ref{eq:induction} and removes the factor
+of $a$ in the $\vecB$ term in the energy equation
+(Eq.~\ref{eq:total_energy_def}) and in the definition of the isotropic
+pressure $p^*$.

### 2013-06-24 numerical-mhd-ct.tex
Edited section 4.3 (MHD-CT)

@@
 The third solver we describe is an MHD method developed by 
-\citet{Collins10} and since a full description and suite of test problems can be
-found in that paper, we only describe the method briefly here.
+\citet{Collins10}.  Since a full description and suite of test problems can be
+found in that reference, we only describe the method briefly here.
 
 The divergence of the magnetic field, \Divb, is
-identically zero due to the fact that the evolution of the magnetic field is the
+identically zero in reality due to the fact that the evolution of the magnetic field is the
 curl of a vector, and the divergence of the curl of a vector is identically
@@
 only the update of the magnetic field is divergence free, any monopoles 
-created by other numerical sources persist.
+created by other numerical sources (such as ill-chosen initial conditions) persist.
 
@@
 
-The dual energy formalism has also been incorporated in two possible ways, one internal
+The dual energy formalism has also been incorporated in two possible
+ways -- one that uses internal
 energy, as described in Section \ref{sec.hydro.ppm}, and one that
 uses entropy \cite{TVD93, Collins10}.
+
 %The additional equations are

### 2013-06-24 numerical-zeus.tex
Editing section 4.4 (zeus)

@@
 Each part is evaluated successively using the results preceding it.
-In this method, in addition to operator-splitting the expansion terms (i.e.
-those terms in eqs.~(\ref{eq:mass}) -- (\ref{eq:total_energy}) that depend on $\dot{a}$), we divide the remaining terms into \emph{source} and \emph{transport} steps.
+In this method, in addition to operator-splitting the expansion terms (i.e., those terms in Eqs.~(\ref{eq:mass}) -- (\ref{eq:total_energy}) that depend on $\dot{a}$), we divide the remaining terms into \emph{source} and \emph{transport} steps.
 The terms to be solved in the source step are those on the right-hand side of eqs.~(\ref{eq:mass}) -- (\ref{eq:total_energy}), while the transport terms are on the left-hand side of these equations and are responsible for the advection of mass, momentum and energy across the grid.
@@
 q_j = \left\{ \begin{array}{ll}
-              Q_{\rm AV} \rho_j (v_{j+1} - v_j)^2 \quad & \rm{if} (v_{j+1} - v_j) < 0 \\
+              Q_{\rm AV} \rho_j (v_{j+1} - v_j)^2 \quad & \rm{if}~(v_{j+1} - v_j) < 0 \\
                0 & \rm{otherwise}
@@
 
-Next, we examine the transport step, which is conservative.
+We next examine the transport step, which is conservative.
 Once again, we dimensionally split the equations and present only the one-dimensional version.
@@
 As recommended in \citet{Stone92a}, we use the second-order van Leer scheme, which uses piecewise linear functions.
-These are given by equations (48) and (49) of \citet{Stone92a}.
+These are given by Equations (48) and (49) of \citet{Stone92a}.
 The transport steps for the other variables are similar.

### 2013-06-24 numerical-gravity.tex
Section 5 edits (small)

@@
 First, the massive (dark matter and star) particles are distributed onto the grids using the second-order cloud-in-cell (CIC) interpolation technique \citep{Hockney88} to form a spatially-discretized density field $\rho_{\rm DM}$.
-During the CIC interpolation, particle positions are (temporarily) advanced by $0.5 \Delta t v^n$ so that we generate an estimate of the time-centered density field.
+During the CIC interpolation, particle positions are (temporarily) advanced by $0.5 v^n \Delta t$ so that we generate an estimate of the time-centered density field.
 Particles on subgrids within the grid's volume are also added to its gravitating field using the same method.
@@
 Next, we add the baryonic grid densities in a similar fashion.
-In particular, we treat baryonic cells as virtual CIC particles that are are placed at the grid center but are advanced by $0.5 \Delta t v^n$ in order to approximately time-center the gravitating mass field.
+In particular, we treat baryonic cells as virtual CIC particles that are are placed at the grid center but are advanced by $0.5 v^n \Delta t$ in order to approximately time-center the gravitating mass field.
 Cells that are covered by further-refined grids are treated in a similar way (i.e., we also use the subgrid cells as virtual CIC particles).

### 2013-06-24 numerical-particle.tex
Section 5 edits (small)

@@
 described earlier, the gravitational potential is computed by solving
-the elliptic Poisson's equation (Eq.~\ref{eq:potential}), and
+the elliptic Poisson's equation (Eq.~\ref{eq:potential}) and
 differencing the potential to find accelerations, which are then
@@
 At higher levels, contributions from particles to the gravitating mass field
-will be smoothed over a spherical region centered at each particle's center.
+will be smoothed over a spherical region centered at each particle's position.

### 2013-06-25 numerical-chemistry.tex
Section 6 edits.

@@
 several models for a metagalactic UV background that heat the gas via
-photoionization and photodissociation; see Section~\ref{sec.num.cooling} for
-more details.  The chemical and thermal states of the gas can be updated either
-at the same hydrodynamical timestep (i.e., de-coupled and operator-split) or
+photoionization and photodissociation (see Section~\ref{sec.num.cooling} for
+more details).  The chemical and thermal states of the gas can be updated either
+at the same hydrodynamical timestep (i.e., decoupled and operator-split) or
 through the same subcycling system (i.e., a coupled chemical and thermal
@@
 abundant species and are at all times in astrophysical calculations close to
-equilibrium values.  Attempting to resolve their formation and destruction
+their equilibrium values.  Attempting to resolve their formation and destruction
 times would necessitate extremely small timesteps, and while these species
@@
 electron concentration is limited to no more than $10\%$ per timestep;
-additional criteria may be applied based on the expected change from radiative
-cooling, chemical heating from the formation of molecular hydrogen.
+additional criteria may be applied based on the expected change in
+internal energy from radiative
+cooling and from  chemical heating due to the formation of molecular hydrogen.
 

### 2013-06-25 numerical-cooling.tex
More minor edits
Section 6 edits.

@@
 \enzo\ has multiple methods for computing the energy change from radiative
-cooling and heating.  All of them assume the gas can be modeled either as
+cooling and heating.  All of them assume that the gas can be modeled either as
 completely optically thin or with a simple, local approximation to optical
-thickness.  Below we describe the methods for computing the cooling rates from
+thickness.  In this section, we describe the methods for computing the cooling rates from
 metal-free and metal-enriched gas.  Sample cooling curves for each of \enzo's
@@
 free-free emission, Compton scattering off of the cosmic microwave
-background (CMB), and photo-heating from a UV metagalactic background.
-If the \HH~chemistry network is enabled, the following \HH~
-cooling processes are also considered: ro-vibrational transitions
+background (CMB), and photo-heating from a metagalactic ultraviolet background.
+If the \HH~chemistry network is enabled, the following \HH~cooling 
+processes are also considered: ro-vibrational transitions
 \citep{2008MNRAS.388.1627G,1998A&A...335..403G}, heating and cooling from molecular
@@
 and without a radiation background.  More detail on the specific UV backgrounds
-present in \enzo\ is presented below in Section~\ref{sec.num.rad-homogeneous}.
+present in \enzo\ is presented in Section~\ref{sec.num.rad-homogeneous}.
 
@@
 \texttt{Cloudy} \citep{1998PASP..110..761F}.  This method, detailed in
-\citet{2008MNRAS.385.1443S} and \citet{2011ApJ...731....6S}, works by
+\citet{2008MNRAS.385.1443S, 2011ApJ...731....6S}, works by
 using \texttt{Cloudy} to compute the cooling and heating rates from
@@
   \begin{center}
-    \includegraphics[width=1.0\textwidth]{figures/cooling_rate.eps}
+    \includegraphics[width=0.8\textwidth]{figures/cooling_rate.eps}
   \end{center}

### 2013-06-25 numerical-radiation-fld.tex
Section 7.

@@
 these linear solvers are provided by the HYPRE linear solver library
-(see \cite{FalgoutYang2002} and \cite{hypre-manual}).
+\citep[see][]{FalgoutYang2002, hypre-manual}.
 
@@
 time-evolved solutions $\left(E_r^{n+1}, e_c^{n+1}, 
-{\mathrm n}_i^{n+1})\right)$ as
+{\mathrm n}_i^{n+1}\right)$ as
 \begin{eqnarray}

### 2013-06-25 numerical-radiation-homogeneous.tex
Section 7.

@@
 time-varying) radiation fields that can interface with the chemistry
-and cooling/heating routines described elsewhere.  Many of these use
-fits to the H, He and \Hep ionizing and photo-ionization heating
+and cooling/heating routines described in Section~\ref{sec.microphysics}.  Many of these use
+fits to the H, He and \Hep~ionizing and photo-ionization heating
 rates that are of the form
@@
 give the coefficients for the photoionization (and photo-heating)
-rates of H, He, and \Hep in Table~\ref{table:homo_coefs}.  Radiation
+rates of H, He, and \Hep~in Table~\ref{table:homo_coefs}.  Radiation
 background types 1
@@
 \tableline
-\multicolumn{7}{c}{Radiation Type 3 modified \citep{2012ApJ...746..125H}} \\
+\multicolumn{7}{c}{Radiation Type 3 modified \citept2012ApJ...746..125H}} \\
 \tableline

### 2013-06-25 numerical-conduction.tex
Jeff Oishi's changes.
Edited section 8 (conduction, star formation + feedback routines).

@@
 similar to that of \citet{2007ApJ...664..135P}.  
-The isotropic flux of heat is given by:
+The isotropic flux of heat is given by
 \begin{equation}
@@
 where $\kappa_{\rm sp}$ is the Spitzer conduction coefficient, $4.6 \times 10^{-7}$~T$^{5/2}$ erg
-s$^{-1}$~cm$^{-1}$~K$^{-1}$ \citep{1962pfig.book.....S}, here we are using a
+s$^{-1}$~cm$^{-1}$~K$^{-1}$ \citep{1962pfig.book.....S}.  In this
+situation we are using a
 value for the Coulomb logarithm, $\log \Lambda = 37.8$, that is
-appropriate for the intracluster medium \citep{1988xrec.book.....S}.
+appropriate for the intracluster medium \citep{1988xrec.book.....S} --
+in astrophysically-relevant, fully ionized plasmas this value varies
+by no more than 50\% (see, e.g., Smith et al. 2013, submitted).
 It is quite possible that the local heat flux computed in this way 
@@
 can be expressed as a fraction f$_{\rm sp}$ of the Spitzer conductivity
-(where f$_{\rm sp} \leq 1.0$ are considered physical values).  This takes
+(where f$_{\rm sp} \leq 1.0$ are considered physically realistic values).  This takes
 into account physical processes below the resolution limit of the
-simulation (such as tangled magnetic fields) that can suppress heat
+simulation, such as tangled magnetic fields, that can suppress heat
 transport.
@@
 \begin{equation}
-\myvec{F} = -\kappa_{\rm sp} \hat{b} (\hat{b} \cdot \grad T)
+\myvec{F} = -\kappa_{\rm sp} \myvec{b} (\myvec{b} \cdot \grad T)
 \end{equation}
-where $\hat{b}$ is the unit vector denoting the direction of the
+where $\myvec{b}$ is the unit vector denoting the direction of the
 magnetic field.  As with the isotropic thermal conduction, we allow a

### 2013-06-25 numerical-starformation.tex
Edited section 8 (conduction, star formation + feedback routines).

@@
 examined to see if they meet a set of criteria for star formation.
-This may simply be a baryon density threshold, but may include more
+This may simply be a baryon density threshold, but can also include more
 complex tests, such as an examination of local cooling and dynamical
@@
 there is an immediate feedback of some fraction $f_{\rm SN}$ of the rest
-energy from the stars into the baryon gas, such
+energy from the stars into the gas, such
 that $E_{\rm add} = f_{\rm SN} \Delta m_{\rm sf} c^2$, where c is the speed of
@@
 There are six user-defined parameters in this algorithm: three deal
-with the star formation ($\eta$, $m_{\rm *min}$ and $t_{\rm
+with star formation ($\eta$, $m_{\rm *min}$ and $t_{\rm
   dyn,min}$), and three deal with feedback ($f_{\rm SN}$, $f_{Z*}$ and
@@
 can be guided by observation or theory.  For example, the supernova
-feedback parameter, $f_{\rm SN}$, can be constrained assuming that, for
+feedback parameter, $f_{\rm SN}$, can be constrained by assuming that, for
 every $200 M_\odot$ of stars created, one supernova occurs, and this
@@
 high-resolution cosmological simulations of galaxy formation, where
-the formation of molecular hydrogen is followed directly, and stars
+the formation of molecular hydrogen is followed directly and stars
 are allowed to form at the highest level of refinement when the local

### 2013-06-25 numerical-timestep.tex
Section 9 edits.

@@
 as well as in space.  The timestep $\Delta t$ is set on a level-by-level
-basis by finding the largest timestep such that all of the criteria listed below are
-satisfied.  The timestep criteria are given by the following expressions
-(where we show the one-dimensional case for clarity):
+basis by finding the largest timestep such that all of the criteria
+listed below (that are relevant for the simulation in question) are
+satisfied.  The timestep criteria are given by the following
+expressions, showing the one-dimensional case for clarity:
 
@@
 formalism means that this value is calculated for all cells or
-particles on a given level l and the minimum overall value is taken as the timestep.
+particles on a given level L and the minimum overall value is taken as the timestep.
 
@@
 %spacing, $\Delta x$.  
-In this equation, $\kappa_{\rm hydro}$ is a
+In this equation, $\kappa_{\rm hydro}$ is a dimensionless
 numerical constant with a value of $0 < \kappa_{\rm hydro} \leq 1$
@@
 
-Equation~(\ref{eqn:dthydro}) is valid for one dimension.  For 2 or 3
+Equation~(\ref{eqn:dthydro}) is valid for one dimension, and is used
+when the equations of hydrodynamics are being solved.  For 2 or 3
 dimensions, it was shown by \cite{Godunov1959}  that using the
@@
 equation~(\ref{eqn:dthydro}) in that it ensures that the CFL condition
-is being enforced at all times.  In this equation, $\kappa_{\rm MHD}$ is a
+is being enforced at all times.  In this equation, $\kappa_{\rm MHD}$
+is a dimensionless
 numerical constant with a value of $0 < \kappa_{\rm MHD} \leq 1$ (with a
@@
 width.  The parameter $\kappa_{\rm dm}$ is like $\kappa_{\rm
-hydro}$, with a similar range of values.
+hydro}$, with a similar range of values.  This criterion is used when
+massive particles are included in a simulation.
 
@@
 solution to the equation of heat conduction.  In this expression,
-$n_b$ is the baryon number density. 
+$n_b$ is the baryon number density, 
 %defined as $n_b = \rho_b / \mu
 %m_p$, where $\mu$ is the mean molecular weight, 
-$\kappa(T)$ is the Spitzer thermal conductivity, and f$_{\rm sp}$ is a user-defined
+$\kappa(T)$ is the Spitzer thermal conductivity, and f$_{\rm sp}$ is a
+user-defined, dimensionless
 conduction suppression factor whose value must be f$_{\rm sp} \leq 1$.  
 %Both are defined in Section~\ref{sec.num.conductions}.  
-$\kappa_{\rm cond}$ is a prefactor whose
-value must be $0 < \kappa_{\rm cond}  < 0.5$, and is exactly 0.5 for the
+$\kappa_{\rm cond}$ is a dimensionless prefactor whose
+value must be $0 < \kappa_{\rm cond}  \leq 0.5$, and is exactly 0.5 for the
 implemention in \enzo.
@@
 perspective, it is useful to note that, unlike other timestep criteria
-discussed above (which effectively scale as $\Delta x / \sqrt{T}$),
+discussed above (which effectively scale as $\Delta x / \sqrt{T}$ with
+the grid cell size $\Delta x$ and temperature $T$),
 the timestep criterion due to thermal conduction scales as $\Delta x^2
@@
 for the stability of the PPM algorithm in comoving coordinates.  This
-criterion typically limits the timestep
-only at high redshifts, when densities are relatively homogeneous.
+criterion typically limits the simulation's timestep
+only during the earliest phases of a cosmological simulation, before
+substantial structure has formed.
 

### 2013-06-25 numerical-analysis.tex
Edited Section 10 (analysis).  Planning to skip to sections 12-14, and then return to 11 later.

@@
 instrumented \enzo\ with the ability to conduct analysis during the course of a
-simulation.  This enables analysis with extremely high time cadence (even as often
+simulation.  This enables analysis with extremely high time cadence (as often
 as every subcycle of the finest refinement level), without
@@
 for conducting analysis in \enzo\ during the course of the simulation utilizes
-the same compute resources as used by the simulation itself by transferring their usage 
+the same computional resources as are used by the simulation itself by transferring their usage 
 from \enzo\ to the analysis routines; this is often
@@
 calculations.  Future versions will include this, as well as the ability to
-dynamically allocate compute resources to \texttt{yt} such that the simulation
+dynamically allocate computional resources to \texttt{yt} such that the simulation
 may proceed asynchronously with analysis (\textit{co-scheduled} analysis).
@@
 independent of the intervals at which regular data sets are written.  The data written out
-typically includes the tracer particle's unique ID, position, velocity, density,
-and temperature, but is easily extensible to output any grid-based
+typically includes the tracer particle's unique ID and position, as
+well as the velocity, density,
+and temperature of the gas at its location, but is easily extensible to output any grid-based
 quantity that a user requires.  This capability has been used quite
@@
 the average quantity in a grid cell's volume. Consequently, after a period of
-evolution, tracer particles --even if they initially had the same
-density distribution as the gas-- will not have the same density
+evolution, tracer particles -- even if they initially had the same
+density distribution as the gas -- will not have the same density
 distribution as the fluid. For example, they tend to accumulate at stagnation points
@@
 background where underdense gas in the intergalactic medium (IGM) can cool adiabatically to
-very low temperatures.
+unphysically low temperatures.
 
@@
 adequate since the hydrodynamic shock is captured in fewer than the number of
-ghost zones.  Shock finding can be run either upon data output, or at each step
+ghost zones.  Shock finding can be run either upon data output or at each step
 in the evolution of an AMR level if the Mach and pre/post-shock quantities are

### 2013-06-25 parallel.tex
Updated Section 12 (parallel performance)

@@
 that the current version of \enzo\ is not threaded, although a hybrid
-MPI + OpenMP version is currently under development.}.  
+MPI + OpenMP version is under development.}.  
 Parallelization is
@@
 significantly more grids than processors; however, small grids are
-less efficient (because of their ghost zones), and so the code uses a
+less efficient because of the large number of ghost zones compared to
+active zones, and so the code uses a
 simple heuristic in order to split up grids until there are of order
@@
 that do not `live' on a given processor only contain meta data
-(essentially location and size of the grid; such grids are denoted as
+(essentially just the location and size of the grid; such grids are denoted as
 `ghost' grids).  Replicating the hierarchy means that all
@@
 advantage of the capabilities of the MPI library: (i) as the code
-progresses, and data is needed from another grid on another processor,
+progresses and data is needed from another grid on another processor,
 the receiving processor posts an MPI non-blocking receive indicating
@@
 calculates the mean, standard deviation, minimum, and maximum for each
-of the timers across all processors.  For level timers, there are additional
+of the timers of that name across all processors.  For level timers, there are additional
 attributes such as the number of cell updates, the current number of
@@
 MPI process has a roughly equal subvolume (and thus an equal number of grid
-cells), meaning that work is evenly distributed.  Communication
+cells), meaning that work is evenly distributed among computational elements.  Communication
 patterns for both the gravity solve (which uses a fast Fourier
@@
 been empirically determined to be roughly $128^3$ cells per core.  If
-fewer cells per core are used, the CPU is essentially data-starved,
+fewer cells per core are used, the CPU is essentially data-starved 
 and poor scaling is observed due to computing units being idle while
@@
 amount of inter-process communication needed, but larger cell counts
-are typically impractical on most machines due to memory limits.
+are typically impractical on most machines due to the amount of memory
+available on each node.
 
@@
 substantially improve scaling by reducing the amount of communication
-per grid tile (as described in the previous paragraph); however, the
+per grid tile, as described in the previous paragraph; however, the
 optimal ratio of OpenMP threads per MPI task seems to vary
@@
 in each routine, it is clear that the communication in the
-\texttt{SetBoundaryConditions}, \texttt{SolveForPotential} (multi-grid
-solver in AMR levels), and \texttt{RebuildHierarchy} are responsible
+\texttt{SetBoundaryConditions}, \texttt{SolveForPotential} (the multi-grid
+solver in AMR subgrids), and \texttt{RebuildHierarchy} are responsible
 for the lack of strong scaling at $N_{\rm core} \ga$~4,096 in this
@@
 of grid points swells by orders of magnitude (an increase of $10^3$ is
-not uncommon) and most of the CPU time is consumed at late times.
+not uncommon) and most of the CPU time is consumed at late physical
+time in the simulation.
 Therefore, simply performing a few steps at the beginning of the
-calculation does produce a good estimate of the required CPU time.
+calculation does not produce a good estimate of the required CPU time.
 
@@
 cosmological simulations appears to be due to a combination of the
-density structure of individual clouds, and the distribution of the
+density structure of individual clouds and to the distribution of the
 clouds themselves.  In particular, we note that for a $\rho \propto
@@
 
-Many of today's leading supercomputers use a heterogeneous structure:
+Many of today's leading supercomputers use a heterogeneous
+computing platform:
 on a single node of a distributed-memory platform, a multi-core CPU is
@@
 Dedner MHD solvers, they are inherently massively parallel and thus
-should be good fit for hardware acceleration. Both solvers essentially
-contain basically two parts -- computation of fluxes, and cell update.
+should be a good fit for hardware acceleration. Both solvers essentially
+contain two parts -- computation of fluxes, and a cell update.
 When computing fluxes, the basic procedure is to compute the flux at
@@
 algorithmically-different loop iterations are completely independent
-of each other. Thus, the parallelization scheme is naturally mapping
+of each other. Thus, the natural parallelization scheme is to map
 one GPU thread to one iteration of the loop. However, the original CPU
@@
 temporary arrays that are not shared among loop iterations. This
-change exposed the massive parallelism in the algorithm, which can
+change exposed the massive parallelism in the algorithm, which could
 then be accelerated in a straightforward manner using CUDA.
@@
 uniform mesh in Figure~\ref{fig:gpu_scaling}.  This particular problem
-type contains no physics other than solving the equations of ideal
-MHD, and thus is representative of the type of calculation than can
+type contains no physics other than the equations of ideal
+MHD, and thus is representative of the type of calculation than can currently
 benefit from the GPU-optimized solvers in \enzo.  In this scaling
@@
 single NVIDIA Tesla K20 GPU, which has 2,496 CUDA cores and a maximum
-theoretical speed of 1.17 Tflops when used in double precision.  For
+theoretical speed of 1.17 Tflops for double-precision computations.  For
 the CPU run, all the 16 cores in each node are used by launching 16
@@
 porting the MHD code to GPUs -- in the GPU simulation case, the 16 CPU
-cores on each node are effectively idle, as they are used only for
+cores on each node are mostly idle, as they are used only for
 timestep calculation and boundary condition transmission (which are

### 2013-06-25 development.tex
Use breakurl package to prevent the bitbucket url from overflowing into the margin.
added matt's changes.
Section 13 edits.

@@
 Over time, \enzo's development has followed a trajectory toward
-openness.  Started as the graduate research project of Greg Bryan at
+increasing openness.  Started as the graduate research project of Greg Bryan at
 the University of Illinois, it was subsequently stewarded by the
@@
 As the stewardship of the code passed to the LCA, the code was
-released first to ``friendly users'' and then as an open source
+released first to ``friendly users'' and then as a public open source
 release.  However, while the code was made available with
@@
 are free to ``fork'' the \enzo\ code base, develop changes (signed
-with their own name), and submit them for review and inclusion.  In
+with their own name), and submit them for review and inclusion into
+the primary code repository.  In
 contrast to the centralized, gatekeeper-focused technology used
 previously, this enables anyone to contribute changes to be evaluated
-for inclusion in the \enzo\ codebase.
+for inclusion in the \enzo\ codebase.  One challenge that this
+presents is that the \enzo\ code is a moving target -- to that end, we
+recommend that users include the changeset hash of the \enzo\
+repository that generated their simulation data (and also the version
+of \texttt{yt} used to analyze the data) in their publications.
 
@@
 to run.
+
+%%% Local Variables: 
+%%% mode: latex
+%%% TeX-master: "ms"
+%%% End:

### 2013-06-25 conclusions.tex
Section 14 edited (conclusions).

@@
 level.  Furthermore, we are nearing the physical limits of transistor
-speed and latency~\citep{feynman1999feynman}, meaning that simple
+speed and interconnect latency~\citep{feynman1999feynman}, meaning that simple
 hardware improvements will not make these challenges disappear, and

### 2013-06-25 appendix.tex
Edits to appendix.  Time to move back to the test problem section, which I have been avoiding because it's very, very long.

@@
 dealing with Cartesian coordinates and cells of equal sizes, although
-we allow for arbitrary an refinement factor $r$ between cells at
+we allow for an arbitrary refinement factor $r$ between cells at
 different levels.
@@
 cell-centered parent values).  We then compute monotonic slopes across
-the two diagonals, since for a linear function, extrema occur at
+the two diagonals, since for a linear function extrema occur at
 corners:
@@
 across the 8 opposing corners of the cube), while the number of slopes
-is three, so the problem is over constrained.  Somewhat arbitrarily,
+is three, so the problem is over-constrained.  Somewhat arbitrarily,
 we adopt the following procedure.  As before, we compute $\Delta Q_0$,
@@
 positive conserved quantities, the slopes are limited so that the
-smallest corner value is 0.2 the cell center value.  The procedure
+smallest corner value is 0.2 of the cell center value.  The procedure
 described above is then applied to turn these four slopes into a
@@
 taking a weighted average between the interpolated values on the
-subgrid and the parent value with weights computed such that the
+subgrid and the parent value, with weights computed such that the
 interpolated values at the edge of the parent cell are not local

### 2013-06-25 code-tests.tex
Edited section 11.1 and the overall Section 11 tex file (code-tests.tex).

@@
 astrophysical fluid dynamics codes.  We describe the \enzo\ test
-methodology in Section~\ref{sec.tests.suite}, code comparisons in
-Section~\ref{sec.tests.compare}, and show a small set of
-representative test problems in Section~\ref{sec.tests.problems}.  We
-further note that the \enzo\ test suite
-(Section~\ref{sec.tests.suite}) contains hundreds of test problems, as
-well as the ability to compare to a ``gold standard'' solution, and
-thus all of the tests shown here are easily reproducible by the
-reader.
+methodology in Section~\ref{sec.tests.suite}, discuss code comparisons
+involving \enzo\ in Section~\ref{sec.tests.compare}, and show a small
+set of representative test problems in
+Section~\ref{sec.tests.problems}.  We further note that the \enzo\
+test suite (Section~\ref{sec.tests.suite}) contains hundreds of test
+problems, as well as the ability to compare to a ``gold standard''
+solution, and thus all of the tests shown here are easily reproducible
+by the reader.  To facilitate this, all of the test problems included
+in this paper, as well as scripts to run the test problems and
+generate the figures found in Section~\ref{sec.tests.problems}, can be
+found on the \enzo\ website.
 
@@
 contains hundreds of test problems that probe the code's behavior in a
-wide range of physical circumstances, and explicitly test each physics
+wide range of physical circumstances, and explicitly tests each physics
 package in the \enzo\ code, both individually and in combination.  It
@@
 \texttt{yt}, run the test suite, and see the results of any other test
-problem and its comparison to an analytic solution (if available).
+problem and its comparison to an analytic solution (if available), or
+to a ``gold standard'' solution from a stable version of \enzo.
 

### 2013-06-25 tests-vandv.tex
Edited section 11.1 and the overall Section 11 tex file (code-tests.tex).

--- 2013-06-12 tests-vandv.tex
@@
 with all but a few of these types requiring only a parameter file as an
-input.  The notable exception is the cosmology simulation, which takes
+input.  The most notable exception is the cosmology simulation, which takes
 as input initial conditions created by other codes.  The suite of test
@@
 them to serve as a vehicle for verifying that the code's behavior
-remains stable over time and after modification of the codebase.
+remains stable over time, on new computing platforms, and after modification of the codebase.
 
@@
 in a nested directory structure organized according to the primary
-functionality tested, e.g., hydrodynamics, gravity, cooling, etc.
+functionality tested (e.g., hydrodynamics, gravity, cooling, etc.).
 Each parameter file is accompanied by a text file containing various
@@
 analysis performed on all test problems includes calculation of
-various statistics, such as extrema, mean and variance, on the fields
+various statistics (such as extrema, mean, and variance) on the fields
 present in the output data.  Custom analysis provided by scripts that
@@
 problems, as well as the scripts to generate the figures from them,
-are available in the \enzo\ test suite, allowing the interested reader
-to regenerate all of the figures in Section~\ref{sec.tests.compare}
-using the \textit{`methodpaper'} suite of test problems.
+are available at \url{http://bitbucket.org/bwoshea/enzo-method-paper},
+the Bitbucket repository for the \enzo\ method paper.  The
+\texttt{test\_problems} subdirectory in this repository contains all of
+the files necessary to regenerate all of the figures in
+Section~\ref{sec.tests.problems}.  We note for completeness that 
+the figures in Section~\ref{sec.tests.problems} were generated using \texttt{yt} version 2.5.3 and the
+\url{https://bitbucket.org/enzo/enzo-dev} repository of \enzo\ with
+changeset \texttt{5d8c412}.
 

### 2013-06-25 table_of_variables_and_constants.tex
noticed a typo in the big table

--- 2013-06-20 table_of_variables_and_constants.tex
@@
   $\Gamma$ & radiative heating rate \\
-  $Q$ & artificial viscosity stress tensor (only for ZEUS hydro) \\[3pt]
+  $Q$ & artificial viscosity stress tensor (only for Zeus hydro) \\[3pt]
 

### 2013-06-25 code-tests.tex
test problems through the Test Orbit (11.2.7)

@@
 that, we will show and describe \enzo's solution to the test problem.
+We remind the user that, as discussed at the end of
+Section~\ref{sec.tests.suite}, they can download a Mercurial
+repository containing this paper, the \enzo\ test problem parameter
+files that produced the simulation data used to generate the figures
+in this section, and the \texttt{yt} scripts necessary to create the
+figures themselves.
 

### 2013-06-25 tests-orbit.tex
test problems through the Test Orbit (11.2.7)

@@
 path of the test particle in the orbital plane plotted for the
-duration of the simulation.  Right panel: total specific energy
+duration of the simulation, with a zoomed box showing a small portion
+of the orbit to highlight deviations from perfect circularity.  Right panel: total specific energy
 (kinetic plus potential) for the test particle for the duration of the
@@
 problem (shown in Figure~\ref{fig.orbittest}), a 3D box with unit size
-and $32^3$ grid is created, and a particle with mass 1 is placed at
+and a $32^3$ grid is created, and a particle with mass 1 is placed at
 the center of the box (both box length and particle mass are in
@@
 using a second-order leapfrog method, this level of accuracy is
-expected.
+expected and acceptable.

### 2013-06-25 tests-sedov.tex
test problems through the Test Orbit (11.2.7)

@@
 
-The Sedov Blast Test \citep{Sedov1959} models an intense explosion,
+The Sedov Blast test \citep{Sedov1959} models an intense explosion,
 initiated by depositing thermal energy into a homogenous distribution
@@
 explosion.   Second, an analytical solution to the evolution of the
-Sedov blast-wave exists, allowing for a direct test of the accuracy of
+Sedov blast wave exists, allowing for a direct test of the accuracy of
 the code.  The radius of the shock front as a function of time is
@@
 In the initial state, the box contains a homogenous distribution of
-gas at a density of 1 (note that, in the absense of gravity, this
+gas at a density of 1 (note that, in the absence of gravity or
+radiative cooling, this
 problem is scale-free and thus without units). Thermal energy is
@@
 10.0$. The problem is run in two dimensions with reflecting
-boundary conditions and a box size of side
-$1$.  We selecte a top grid of $100 \times 100$
-cells and a maximum of four levels of adaptive refinement, refining on
+boundary conditions and a box having a length of
+$1$ in both directions.  We selecte a top grid of $100 \times 100$
+cells and a maximum of four levels of adaptive refinement, refining by
+factors of two on
 shocks and the slopes of the density and total energy fields. The

### 2013-06-25 tests-shockpool.tex
test problems through the Test Orbit (11.2.7)

@@
 The impact of entering and exiting the refined region is somewhat larger than in PPM; however, the most noticeable difference is the incorrect position of the shock front.
-This is due to the fact that the scheme is not energy-conserving (see also the Sedov problem, below), and has very little to do with the passage through the refined region.
+This is due to the fact that the scheme is not energy-conserving (see also the Sedov problem in Section~\ref{sec.tests.sedov}), and has very little to do with the passage through the refined region.
 
-Finally, the MUSCL scheme produces a shock which is intermediate in width between the two previous cases.
-This method is energy conserving and so correctly reproduces the shock speed.
+Finally, the MUSCL scheme produces a shock that is intermediate in width between the two previous cases.
+This method is energy-conserving, and thus correctly reproduces the shock speed.
 The oscillations are mild except for the cell immediately outside of the refined region upon exiting.

### 2013-06-25 tests-sodshocktube.tex
test problems through the Test Orbit (11.2.7)

@@
 PPM, as well as the two second-order \zeus\ and MUSCL schemes.  We use
-100 cells across the domain, (which is a relatively standard choice in
-code method papers) and show solutions both with and without adaptive
+100 cells across the domain, which is a relatively standard choice in
+code method papers, and show solutions both with and without adaptive
 mesh refinement.  Without AMR (top row in Figure \ref{fig.sodshocktube}), it is

### 2013-06-25 tests-wavepool.tex
test problems through the Test Orbit (11.2.7)

@@
 
-In this one-dimensional test we pass a short wavelength linear wave through a static singley-refined region.
+In this one-dimensional test we pass a short wavelength linear wave through a static singly-refined region.
 The full domain is from 0 to 1 and the refined region is set from 0.25 to 0.75 at all times.
-We modify the left boundary consistent with a single linear sound wave with density given by $\rho(x,t) = \rho_0 (1 + A \sin(kx - \omega t))$, where the amplitude is $A = 0.01$ and the wavelength $\lambda = k/2\pi = 0.1$, and similar expressions exist for the pressure and velocity.
+We modify the left boundary consistent with a single linear sound wave with density given by $\rho(x,t) = \rho_0 (1 + A \sin(kx - \omega t))$, where the amplitude is $A = 0.01$ and the wavelength $\lambda = k/2\pi = 0.1$.
+Similar expressions exist for the pressure and velocity.
 The initial, unperturbed density and pressure are set to unity and we adopt $\gamma = 1.4$.
@@
 
-For the Zeus solver, we see that even before it has entered the refined region there are small oscillations excited behind the wave; although it is also worth noting that the wave itself is beautifully propagated without significant damping or phase errors.
+For the Zeus solver, we see that even before it has entered the refined region there are small oscillations excited behind the wave, although it is also worth noting that the wave itself is beautifully propagated without significant damping or phase errors.
 In this test we use only our standard, low amount of quadratic artificial viscosity.

### 2013-06-25 tests-fld-ifront.tex
Finished up all of the test problems!

@@
 deceleration parameter $q_0 = 0.05$ and initial redshift $z_i=10$.
-Here the physics of interest is the expansion of an \ion{H}{2} region
+In this test problem, the physics of interest is the expansion of an \ion{H}{2} region
 in a uniform medium around a single monochromatic ionizing source
@@
 The simulation parameters are: box size $L_i\approx 27$ kpc, Hubble
-constant $H_0 = 100$ km s$^{-1}$ Mpc$^{-1}$, and cosmological constants $\Omega_m = 0.1$, $\Omega_\Lambda=0$ and $\Omega_b = 0.1$.  The
+constant $H_0 = 100$ km s$^{-1}$ Mpc$^{-1}$, and cosmological parameters $\Omega_m = 0.1$, $\Omega_\Lambda=0$, and $\Omega_b = 0.1$.  The
 initial values for the simulation are a radiation energy density

### 2013-06-25 tests-onezonecollapse.tex
Finished up all of the test problems!

@@
 running a true collapse simulation.  The density evolution is based on
-the self-similar Larson-Penston \citep{1969MNRAS.145..271L,
-1969MNRAS.144..425P} solution for isothermal collapse with a
+the self-similar Larson-Penston solution for isothermal collapse \citep{1969MNRAS.145..271L,
+1969MNRAS.144..425P} with a
 modification to account for the efficiency with which the heat
@@
 \end{equation}
-In this equation, $t_{\rm dyn}$ is the dynamical time and is expressed as
+In this equation, $t_{\rm dyn}$ is the dynamical time for the collapse
+of a spherical cloud, and is expressed as
 \begin{equation}
@@
   \begin{center}
-    \includegraphics[width=1.0\textwidth]{figures/OneZoneCollapseTest.eps}
+    \includegraphics[width=0.8\textwidth]{figures/OneZoneCollapseTest.eps}
   \end{center}

### 2013-06-25 tests-orszagtang.tex
Finished up all of the test problems!

@@
 \label{sec.tests.mhd}
-Figure~\ref{fig.orszag} shows the Orszag-Tang vortex problem, a
-classic MHD test problem described by \citet{Orszag79}.  The left panel shows
+Figure~\ref{fig.orszag} shows the Orszag-Tang vortex problem \citep{Orszag79}.  The left panel shows
 the result using \enzo's constrained transport MHD method, while the right panel shows the result using
 \enzo's implementation of Dedner MHD.
-The Orszag-Tang vortex test shows that significant small scale structure can be generated in MHD
-from large scale initial perturbations, and is used to compare the
+The Orszag-Tang vortex test is a classig MHD test problem, and shows that significant small scale structure can be generated in MHD
+from large scale initial perturbations.  It is often used to compare the
 effective resolution of different MHD schemes.  The test begins with uniform
 density, $\rho_0=25/(36\pi)$, and pressure, $P_0=5/(12\pi)$ (as with
-other tests in this section, in the absence of gravity or chemistry we use
+other tests in this section, in the absence of gravity or chemistry/cooling we use
 arbitrary units).  There is a

### 2013-06-25 tests-photoevaporation.tex
Finished up all of the test problems!

@@
 ambient medium.  It has a radius $r = 0.8 \unit{kpc}$, hydrogen
-density $n_{\rm H} = 0.04\ \cubecm$, and temperature $T = 40$ K, and
+density $n_{\rm H} = 0.04\ \cubecm$ (i.e., overdensity of 200), and temperature $T = 40$ K, and
 is centered at $(x,y,z) = (5, 3.3, 3.3) \unit{kpc}$.
@@
 resolution AMR solution, we employ a $128^3$ grid with two additional
-levels of refinement for cells with a baryon mass greater than 1.5
+levels of refinement by factors of 2 for cells with a baryon mass greater than 1.5
 (method 2 in Section~\ref{sec:refinement_criteria}). This test is run
@@
 
-This cloud is subject to radiation from a point source at the center
+The cloud is subject to radiation from a point source at the center
 of the $x=0$ boundary with an ionizing photon luminosity
@@
 a full spectrum and energy discretization \citep{Mirocha12}.
-(Note that this choice of energy groups is different from Wise \& Abel 2011.) \nocite{Wise11_Moray}  We use a minimum angular resolution of 10 rays
+(Note that this choice of energy groups is different from those used
+in Wise \& Abel 2011.) \nocite{Wise11_Moray}  We use a minimum angular resolution of 10 rays
 per cell and a constant radiative transfer timestep of 25 kyr.  Figure

### 2013-06-25 tests-zeldovichpancake.tex
Finished up all of the test problems!

@@
-\subsubsection{Zel`dovich Pancake}
+\subsubsection{Zel'dovich Pancake}
 \label{sec.tests.pancake}
 
-The Zel`dovich pancake \citep{1970A&A.....5...84Z} is particularly
+The Zel'dovich pancake \citep{1970A&A.....5...84Z} is particularly
 relevant to cosmological simulations, because it includes many
@@
 Lagrangian positions, so one needs to convert this into the Eulerian
-coordinates, x$_e$, that are more useful to \enzo:
+coordinates, x$_e$, that are more useful to a grid-based calculation:
 \begin{equation}
@@
 an $\Omega_M = 1$ universe with $h = 0.5$ and a background temperature
-of 100~K, $\rho_0 = \rho_c$.  The simulation is initialized with 64
+of 100~K, with a background density of $\rho_0 = \rho_c$.  The simulation is initialized with 64
 grid cells, refining by factors of four using criteria based on cell

### 2013-06-25 app-datastruct.tex
Use breakurl package to prevent the bitbucket url from overflowing into the margin.
Fixed inconsistencies with reference to the ZEUS hydro method.  In some places it was 'Zeus', in others 'ZEUS'.  Now all calls are to the \zeus command (defined in local-commands.tex), and can be adjusted however the heck one wants.

@@
 \end{enumerate}
-A list of routines to mention: \red{Zeus does this in a big table that
+A list of routines to mention: \red{\zeus\ does this in a big table that
   references code name, what it does, and what equations it
@@
 crossings in the second derivative indicates edges of structures.  If
-more than one exists, the 'strongest crossing,' that is the largest
+more than one exists, the ``strongest crossing'', that is the largest
 jump, is selected for the cut.  If no such zero crossings can be

### 2013-06-25 numerical-amr.tex
Fixed inconsistencies with reference to the ZEUS hydro method.  In some places it was 'Zeus', in others 'ZEUS'.  Now all calls are to the \zeus command (defined in local-commands.tex), and can be adjusted however the heck one wants.

@@
 These zones are necessary to accommodate the computational stencil of the (magneto)hydrodynamics solvers (Sections~\ref{sec.hydro.ppm} through~\ref{sec.hydro.zeus}) and the gravity solver (Section~\ref{sec.gravity}).
-The PPM and Zeus hydro solvers require 3 layers of ghost zones,  and the gravity solver requires 6 (although only for the density and potential fields).
+The PPM and \zeus\ hydro solvers require 3 layers of ghost zones,  and the gravity solver requires 6 (although only for the density and potential fields).
 This can lead to significant memory and computational overhead, particularly for smaller grid patches at high levels of refinement.
@@
 %and the gravity solver (Section~\ref{sec.ov.nbody}).  
-%The PPM and Zeus hydro solvers require 3 layers of ghost zones,  and the gravity
+%The PPM and \zeus\ hydro solvers require 3 layers of ghost zones,  and the gravity
 %solver requires 6.  This can lead to significant memory 

### 2013-06-25 numerical-gravity.tex
Fixed inconsistencies with reference to the ZEUS hydro method.  In some places it was 'Zeus', in others 'ZEUS'.  Now all calls are to the \zeus command (defined in local-commands.tex), and can be adjusted however the heck one wants.

@@
 In both cases, the potential is then transformed back into the real domain to get potential values at the cell centers.
-These are differenced with a two-point centered difference scheme to obtain accelerations at the cell centers (except if we are using the staggered Zeus-like solver, in which case the accelerations are computed at the cell faces to match the velocities).
+These are differenced with a two-point centered difference scheme to obtain accelerations at the cell centers (except if we are using the staggered \zeus-like solver, in which case the accelerations are computed at the cell faces to match the velocities).
 Particle accelerations are obtained using a (linear) CIC interpolation from the grid.

### 2013-06-25 numerical-overview.tex
Fixed inconsistencies with reference to the ZEUS hydro method.  In some places it was 'Zeus', in others 'ZEUS'.  Now all calls are to the \zeus command (defined in local-commands.tex), and can be adjusted however the heck one wants.

@@
 
-\subsubsection{The Zeus hydro method}\label{sec.ov.hydro.zeus}
+\subsubsection{The \zeus\ hydro method}\label{sec.ov.hydro.zeus}
 
@@
 and the gravity solver (Section~\ref{sec.ov.nbody}).  
-The PPM and Zeus hydro solvers require 3 layers of ghost zones,  and the gravity
+The PPM and \zeus\ hydro solvers require 3 layers of ghost zones,  and the gravity
 solver requires 6.  This can lead to significant memory 

### 2013-06-25 table_of_variables_and_constants.tex
Fixed inconsistencies with reference to the ZEUS hydro method.  In some places it was 'Zeus', in others 'ZEUS'.  Now all calls are to the \zeus command (defined in local-commands.tex), and can be adjusted however the heck one wants.

--- 2013-06-20 table_of_variables_and_constants.tex
@@
   $\Gamma$ & radiative heating rate \\
-  $Q$ & artificial viscosity stress tensor (only for ZEUS hydro) \\[3pt]
+  $Q$ & artificial viscosity stress tensor (only for \zeus\ hydro) \\[3pt]
 

### 2013-06-25 tests-shockpool.tex
Fixed inconsistencies with reference to the ZEUS hydro method.  In some places it was 'Zeus', in others 'ZEUS'.  Now all calls are to the \zeus command (defined in local-commands.tex), and can be adjusted however the heck one wants.

@@
 
-For Zeus (middle row of Figure~\ref{fig.shockpool}), the shock is broadened because of artificial viscosity and there are slightly more post-shock oscillations, although again quite mild.
+For \zeus\ (middle row of Figure~\ref{fig.shockpool}), the shock is broadened because of artificial viscosity and there are slightly more post-shock oscillations, although again quite mild.
 The impact of entering and exiting the refined region is somewhat larger than in PPM; however, the most noticeable difference is the incorrect position of the shock front.
@@
 \caption{This plot shows, for each column, three snapshots at $t=0.1, 0.15$ and 0.35 of a $\mathcal{M}=2$ shock as it propagates through the grid with a static refined region extends from $x = 0.25$ to 0.75 (shown in grey in each panel).
-Each row shows the result for a different solver (PPM/Zeus/MUSCL from top to bottom).
+Each row shows the result for a different solver (PPM/\zeus/MUSCL from top to bottom).
 The solid line shows the analytic solution.

### 2013-06-25 tests-wavepool.tex
Fixed inconsistencies with reference to the ZEUS hydro method.  In some places it was 'Zeus', in others 'ZEUS'.  Now all calls are to the \zeus command (defined in local-commands.tex), and can be adjusted however the heck one wants.

@@
 
-Figure~\ref{fig.wavepool} shows the evolution of the wave at three different times ($t = 0.2, 0.3$ and 0.8) for three of our solvers (PPM, Zeus, and MUSCL).
+Figure~\ref{fig.wavepool} shows the evolution of the wave at three different times ($t = 0.2, 0.3$ and 0.8) for three of our solvers (PPM, \zeus, and MUSCL).
 In all panels, the shaded regions denote the statically refined region.
@@
 
-For the Zeus solver, we see that even before it has entered the refined region there are small oscillations excited behind the wave, although it is also worth noting that the wave itself is beautifully propagated without significant damping or phase errors.
+For the \zeus\ solver, we see that even before it has entered the refined region there are small oscillations excited behind the wave, although it is also worth noting that the wave itself is beautifully propagated without significant damping or phase errors.
 In this test we use only our standard, low amount of quadratic artificial viscosity.
@@
 The individual cells are also color-coded by level: blue indicates the root grid, and red is for the refined region.
-Each row shows the result for a different hydrodynamic solver (top is PPM, middle is Zeus, and the bottom is MUSCL).
+Each row shows the result for a different hydrodynamic solver (top is PPM, middle is \zeus, and the bottom is MUSCL).
 The solid line shows the analytic solution for a linear, undamped wave.

### 2013-06-25 appendix.tex
Many small changes to all files -- GB final edits

@@
 equalities are not met, then, if $s<0$ we define $\chi_n = 1$ if
-$\Delta Q_n/\Delta Q_0 < 0$, and 0 otherwise (if $s > 1$, this is
+$\Delta Q_n/\Delta Q_0 < 0$, and $\chi_n = 0$ otherwise (if $s > 1$, this is
 reversed, so $\chi_n = 1$ if $\Delta Q_n/\Delta Q_0 > 0$, and 0
@@
 
-This interpolation method provides a third-order accuracy based on the
+This interpolation method provides third-order accuracy based on the
 Triangular Shaped Cloud (TSC) methodology \citep{Hockney88}.  As
@@
 Finally, for completeness, we include a first-order accurate piecewise
-constant interpolator, for which, using the same definitions in the
+constant interpolator for which, using the same definitions as in the
 previous case, we take $q_i = Q_0$, for $i = 0$ to $r-1$.

### 2013-06-25 introduction.tex
Many small changes to all files -- GB final edits

@@
 simulators. To tackle such situations, a number of numerical
-techniques have been developed that can be broadly split into the
+techniques have been developed that can be broadly split into
 gridless, Lagrangian methods and grid-based, Eulerian schemes. The
@@
 \citep[e.g.,][]{Ruffert94, Anninos94}. Adding adaptivity is a more
-recent enhancement, and there are now a number of N-body codes that
+recent enhancement, and there are now a number of codes that
 possess this feature, both with and without hydrodynamics
@@
 
-In this paper, we present \enzo, a structured adaptive mesh refinement
+In this paper, we present \enzo\ 2.3, a structured adaptive mesh refinement
 (SAMR) code. Originally developed for cosmological hydrodynamics,
@@
 a wide variety of problems, resulting in the publication of more than
-100 peer-reviewed papers based on results obtained with the code. The
+100 peer-reviewed papers. The
 variety of astrophysical systems that \enzo\ has been used for include

### 2013-06-25 ms.tex
Many small changes to all files -- GB final edits

@@
 temporal resolution for modeling astrophysical fluid flows.  The code
-is Cartesian, can be run in 1, 2, and 3 dimensions, and supports a
-large variety of physics, including hydrodynamics, ideal and non-ideal
+is Cartesian, can be run in 1, 2, and 3 dimensions, and supports a wide
+variety of physics including hydrodynamics, ideal and non-ideal
 magnetohydrodynamics, N-body dynamics (and, more broadly, self-gravity
@@
 algorithms implemented, we present solutions for a wide range of test
-problems, demonstrate the code's parallel scaling properties, and
+problems, demonstrate the code's parallel performance, and
 discuss the \enzo\ collaboration's code development methodology.

### 2013-06-25 numerical-amr.tex
Many small changes to all files -- GB final edits

@@
 
-In order to simplify bookeeping, a number of restrictions are imposed on the subgrids:
+In order to simplify bookkeeping, a number of restrictions are imposed on the subgrids:
 \begin{itemize}
  \item The refinement factor $r$, or the ratio of the coarse cell width to the fine cell width, must be an integer.
-In practice, \enzo\ almost always uses $r=2$, since tests have shown this typically provides the best performance.
+In practice, we generally adopt $r=2$, since tests have shown this typically provides the best performance.
  \item Each subgrid must begin and end on the boundary of a coarse cell.  This means that the number of cells in a patch will be a multiple of the refinement factor.
@@
 From the point of view of the hydro solver, each grid is solved as an independent computational fluid dynamics (CFD) problem, with Dirichlet boundary conditions stored in the ghost zones.
-At the end of each timestep, cells on a given level fill ghost zones by interpolating from the parent and copying from neighbors on the same level.
+At the beginning of each timestep, cells on a given level fill ghost zones by interpolating from the parent and copying from neighbors on the same level.
 Grids then correct the flux at the interface boundary between fine and coarse zones, and finally project their active zone data to the parent grid.
+
+\begin{figure}
+\begin{center}
+%\includegraphics[width=0.6\textwidth]{figures/wcycle.eps}
+\includegraphics[width=0.6\textwidth]{figures/timestepping.eps}
+\end{center}
+\caption{\emph{Left:} Example of the timesteps on a 2-level AMR
+  hierarchy.  \enzo\ does not restrict the timesteps on the finer levels
+  to be a factor of $1/2^n$ smaller.  \emph{Right:} The order in which
+  the AMR grids are evaluated on each level.\vspace{1ex}}
+\label{fig:wcycle}
+\end{figure}
 
@@
 InitializeHierarchy
-While (Time < StopTime)
+While (Time < StopTime):
    dt = ComputeTimeStep(0)
@@
 \begin{verbatim}
-EvolveLevel(level)
+EvolveLevel(level):
    SetBoundaryValues
-   while (Time < ParentTime)
+   while (Time < ParentTime):
       dt = ComputeTimeStep(level)
@@
 In the first case, we allow for four types of boundary conditions
-(represented here as being at $x=0$, with positive $x$ interior and
+(for simplicity, represented here as being at $x=0$, with positive $x$ interior and
 negative $x$ exterior to the domain):
@@
 0$) than pure Lagrangian refinement --- in other words,
-super-Lagrangian or sub-Lagrangian in refinement.
+super-Lagrangian or sub-Lagrangian refinement.
 
@@
 
-\item{\em Optical depth.} This refinement criterion, typically used for simulations using radiation transport, ensures that cells on the highest level of refinement have an \ion{H}{1} optical depth lower than unity.
-The optical depth of a cell is estimated to be:
+\item{\em Optical depth.} This refinement criterion, typically used for simulations using radiation transport, 
+flags cells for refinement if they have an \ion{H}{1} optical depth lower than unity.  The optical depth of a cell is estimated to be:
 \begin{equation}

### 2013-06-25 numerical-chemistry.tex
Many small changes to all files -- GB final edits

@@
 since been extended with updated reaction rates and the inclusion of
-deuterium species \citep{2009PhDT.........5T}.  These routines follow the
+deuterium species \citep{McGreer2008, 2009PhDT.........5T}.  These routines follow the
 non-equilibrium chemistry of a gas of primordial composition with 12 total
@@
 accurate solution.  The update discretization splits chemical changes into
-formation components and destruction components and updated with a mixed set of
+formation components and destruction components and updates with a mixed set of
 time states, as described in \citet{anninos97}.  The formation components of
@@
 reactions, these species are very sensitive to small changes in the more
-abundant species and are at all times in astrophysical calculations close to
+abundant species and are (at almost all times in astrophysical calculations) close to
 their equilibrium values.  Attempting to resolve their formation and destruction
-times would necessitate extremely small timesteps, and while these species
-provide necessary channels to facilitate the formation of molecular hydrogen at
-low densities, they do not otherwise significantly influence the abundance of
-species with higher concentrations.  Therefore, reactions governing these two
+times would necessitate extremely small timesteps.
+%and while these species
+%provide necessary channels to facilitate the formation of molecular hydrogen at
+%low densities, they do not otherwise significantly influence the abundance of
+%species with higher concentrations.  
+Therefore, reactions governing these two
 species can be decoupled from the rest of the network and treated independently
-through analytic solutions for equilibrium values.  This allows a large speedup
-in solution as the timestepping scheme is applied only to the slower 7- or
+through analytic solutions for equilibrium values.  This allows a significant speedup
+in the solution speed, as the timestepping scheme is applied only to the slower 7- or
 10-species network (depending on whether deuterium is included or not), which
@@
 the reaction network is valid for temperatures between $10^0 - 10^8$
-K.  The original model discussed in these two references is only valid
+K.  The original model discussed in these two references was only claimed to be valid
 up to $n_H \sim 10^4$~cm$^{-3}$.  However, addition of the 3-body

### 2013-06-25 numerical-conduction.tex
Many small changes to all files -- GB final edits

@@
 similar to that of \citet{2007ApJ...664..135P}.  
-The isotropic flux of heat is given by
-\begin{equation}
-\myvec{F} = -\kappa_{\rm sp} \grad T
-\end{equation}
-where $\kappa_{\rm sp}$ is the Spitzer conduction coefficient, $4.6 \times 10^{-7}$~T$^{5/2}$ erg
+The isotropic flux of heat is given by equation~(\ref{eq:conduction}) and we
+use a value for the Spitzer conduction coefficient, 
+$\kappa_{\rm sp} = 4.6 \times 10^{-7}$~T$^{5/2}$ erg
 s$^{-1}$~cm$^{-1}$~K$^{-1}$ \citep{1962pfig.book.....S}.  In this
@@
 \begin{equation}
-\kappa_{eff} = \frac{\kappa_{\rm sp}}{1 + 4.2 \lambda_e / \ell_T} \; ,
+\kappa_{eff} = \frac{\kappa_{\rm cond}}{1 + 4.2 \lambda_e / \ell_T} \; ,
 \end{equation}
@@
 heat transport only parallel to the magnetic field lines in
-magnetohydrodynamic simulations.  Mathematically, this is treated as:
-\begin{equation}
-\myvec{F} = -\kappa_{\rm sp} \myvec{b} (\myvec{b} \cdot \grad T)
-\end{equation}
-where $\myvec{b}$ is the unit vector denoting the direction of the
-magnetic field.  As with the isotropic thermal conduction, we allow a
+magnetohydrodynamic simulations.  Mathematically, this is given
+by equation~(\ref{eq:anisotropic_conduction}.
+As with the isotropic thermal conduction, we allow a
 multiplicative factor f$_{\rm sp}$ to take into account the possible
-behavior of magnetic fields below the resolution limit of the
+suppression of magnetic fields below the resolution limit of the
 simulation.

### 2013-06-25 numerical-mhd-dedner.tex
Many small changes to all files -- GB final edits

@@
 for $i={x,y,z}$, $E=v^2/2 + e$ denotes the specific total energy and $e$ the
-specific internal energy.
+specific internal energy (note that in this section only, we use the specific
+energy).
 
-For the generalized Lagrange multiplier formulation of the MHD
+For the generalized Lagrange multiplier (GLM) formulation of the MHD
 equations \citep{2002JCoPh.175..645D}, we consider these 
@@
 where $c_h$ is a constant controlling the propagation speed and
-damping rate of $\div B$, and $h=e+p/\rho$ denotes the enthalpy.
-All quantities are cell-centered.
+damping rate of $\div B$, and $h=e+p/\rho$ denotes the specific 
+enthalpy.  All quantities are cell-centered.
 
@@
 strength).  This adds a source term of $-\dot{a}/(2a) \vecB$ on the
-right-hand side of Equation~\ref{eq:induction} and removes the factor
+right-hand side of Equation~(\ref{eq:induction}) and removes the factor
 of $a$ in the $\vecB$ term in the energy equation

### 2013-06-25 numerical-ppm.tex
Many small changes to all files -- GB final edits

@@
       \frac{1}{a} \frac{\partial p}{ \partial x} & = & 
-      \rho \frac{1}{a} \frac{\partial \phi}{ \partial x}  \label{eq:momentum1d}  \\
+      \rho \frac{g}{a}  \label{eq:momentum1d}  \\
 \frac{\partial \rho E}{\partial t}  + \frac{1}{a} \frac{\partial \rho v E}{\partial x}  & =  &
-      \rho v \frac{1}{a} \frac{\partial \phi}{\partial x}. \label{eq:total_energy1d}
+      \rho v \frac{g}{a} \label{eq:total_energy1d}
 \end{eqnarray}
@@
 by expanding the density and acceleration distributions and
-retaining all terms up to second-order in $\Delta x$:
+retaining all terms up to second-order in $\Delta x$ \citep{1995CoPhC..89..149B}, although we find that
+the potential is so slowly varying that this is unnecessary.
 %
-\begin{equation}
-g_j^{n+1/2} = 
-       \frac{1}{2 a^{n+1/2} \delta x_j} \left[ 
-             \phi_{j+1}^{n+1/2} 
-           - \phi_{j-1}^{n+1/2} 
-           + \frac{1}{12} \left(    \phi^{n+1/2}_{j+1} 
-                                 - 2\phi^{n+1/2}_j 
-                                 + \phi^{n+1/2}_{j-1} \right) 
-                                   {\delta d_j \over d_j}
-       \right].
-       \label{eq:accel}
-\end{equation}
-Generally we find that the potential is so slowly varying that this change is unnecessary.
+%\begin{equation}
+%g_j^{n+1/2} = 
+%       \frac{1}{2 a^{n+1/2} \delta x_j} \left[ 
+%             \phi_{j+1}^{n+1/2} 
+%           - \phi_{j-1}^{n+1/2} 
+%           + \frac{1}{12} \left(    \phi^{n+1/2}_{j+1} 
+%                                 - 2\phi^{n+1/2}_j 
+%                                 + \phi^{n+1/2}_{j-1} \right) 
+%                                   {\delta d_j \over d_j}
+%       \right].
+%       \label{eq:accel}
+%\end{equation}
+%Generally we find that the potential is so slowly varying that this change is unnecessary.
 
@@
 
-In this section, we describe the methods we adopt to solving the Riemann
+In this section, we describe the methods we adopt to solve the Riemann
 problem, which is generally required to compute the fluxes in 
@@
 correct physical state is determined, the pressure in the central region can
-only be determined by finding the root to an algebraic equation, which is necessarily
+only be computed by finding the root to an algebraic equation, which is necessarily
 an iterative process.  Thus a series of \emph{approximate} Riemann solvers are

### 2013-06-25 numerical-radiation-raytracing.tex
added renyue cen changes.
Many small changes to all files -- GB final edits

@@
 vectors of the pixels, (2) retain the same radius of the parent ray,
-and (3) gets a quarter of the photon flux of the parent ray.
+and (3) get a quarter of the photon flux of the parent ray.
 Afterward, the parent ray is discontinued.
@@
 convenience, we express the integration in terms of optical depth
-$\tau = \int \kappa(r,t) \; dr$, and for a ray segment,
+$\tau = \int \kappa(r,t) \; dr$, and for a ray segment
 %

### 2013-06-25 numerical-starformation.tex
Many small changes to all files -- GB final edits

@@
 Broadly speaking, these methods all work in similar
-ways: At a specified time interval, all grid cells that are at the
+ways: at a specified time interval, all grid cells that are at the
 highest local level of refinement (i.e., that have no child cells) are
@@
 The methods described in the previous two sections are generally used
-in simulations that have very poor resolution, $\Delta x \ga 1$~kpc.
+in simulations that have relatively poor resolution, $\Delta x \ga 1$~kpc.
 At this physical scale, individual star forming regions are not

### 2013-06-25 numerical-timestep.tex
Many small changes to all files -- GB final edits

@@
 \begin{equation}
-\Delta t_{\rm accel} = \min \left( \sqrt{\frac{\Delta x}{\myvec{g}}} \right)_L ,
+\Delta t_{\rm accel} = \min \left( \sqrt{\frac{\Delta x}{|\myvec{g}|}} \right)_L ,
 \label{eqn:dtaccel}
@@
 \begin{equation}
-\Delta t_{\rm rad} = \min \left(  \sqrt{\frac{\Delta x}{\myvec{a}_{rad}}} \right)_L,
+\Delta t_{\rm rad} = \min \left(  \sqrt{\frac{\Delta x}{|\myvec{a}_{rad}|}} \right)_L,
 \label{eqn:dtrad}
@@
 \begin{equation}
-\Delta t_{\rm cond} = \min \left(  \frac{ \kappa_{\rm cond}}{f_{\rm sp}} \frac{\Delta x^2
-    n_b}{\kappa(T)} \right)_L,
+\Delta t_{\rm cond} = \min \left(  \frac{ k_{\rm cond}}{f_{\rm sp}} \frac{\Delta x^2
+    n_b}{\kappa_{\rm sp}(T)} \right)_L,
 \label{eqn:dtcond}
@@
 %m_p$, where $\mu$ is the mean molecular weight, 
-$\kappa(T)$ is the Spitzer thermal conductivity, and f$_{\rm sp}$ is a
+$\kappa_{\rm sp}(T)$ is the Spitzer thermal conductivity, and f$_{\rm sp}$ is a
 user-defined, dimensionless
@@
 %Both are defined in Section~\ref{sec.num.conductions}.  
-$\kappa_{\rm cond}$ is a dimensionless prefactor whose
-value must be $0 < \kappa_{\rm cond}  \leq 0.5$, and is exactly 0.5 for the
+$k_{\rm cond}$ is a dimensionless prefactor whose
+value must be $0 < k_{\rm cond}  \leq 0.5$, and is exactly 0.5 for the
 implemention in \enzo.

### 2013-06-25 overview.tex
Many small changes to all files -- GB final edits

@@
 permeability is unity ($\mu_0=1$).  The equations are closed by an
-equation of state and Poisson's equation (in the Newtonian limit) for
-the gravitational potential $\phi$:
+equation of state and Poisson's equation for the gravitational potential $\phi$:
 %
@@
 %
-  \nabla^2 \phi &=& 4 \pi G \left( \rho_{\rm total} - \rho_0 \right)
+  \nabla^2 \phi &=& \frac{4 \pi G}{a} \left( \rho_{\rm total} - \rho_0 \right)
 %                   + 3 p_{total}/c^2) - \Lambda. 
@@
 below; however, it is not universal,
-and in particular is slightly different from that used both in \citet{Li08a}
-and used in the MHD-Dedner solver (see section~\ref{sec.num.hydro-muscl}
+and is slightly different from that used both in \citet{Li08a}
+and in the MHD-Dedner solver (see section~\ref{sec.num.hydro-muscl}
 for more details).
@@
 where $k_{ij}$ are the rate coefficients for the two-body reactions
-and are usually functions of only the gas species (we will
+and are usually functions of only temperature (we will
 specifically note the cases where we either include three-body
@@
 absorption ($\kappa_\nu \equiv \kappa_\nu(\mathbf{x},\nu,t)$), and the
-second term emission ($j_\nu \equiv j_\nu(\mathbf{x},\nu,t)$ from any
+second term emission ($j_\nu \equiv j_\nu(\mathbf{x},\nu,t)$) from any
 sources of radiation (whether point or diffuse).
@@
 \fcond = -\kappa_{\rm cond} \grad T.
+\label{eq:conduction}
 \end{equation}
@@
 presence of magnetic field lines, which may strongly suppress heat
-flow perpendicular to the magnetic field.  To include this physics, it
+flow perpendicular to the magnetic field.  To include this effect, it
 is possible to allow heat transport only along (not across) magnetic
@@
 \fcond= -\kappa_{\rm cond} \,\myvec{b} ( \myvec{b} \cdot \grad T),
+\label{eq:anisotropic_conduction}
 \end{equation}
@@
 accuracy and stability criteria and advance the grid by that time
-interval, $\Delta t_0$.  We then move down to the next level and
-advance all the grids on that level by a timestep $\Delta t_1 \leq
-\Delta t_0$, which is the minimum of all the allowed timesteps for
-those grids.  If there are more levels, we repeat this procedure until
+interval.  We then move down to the next level and
+advance all the grids on that level.  If there are more levels, we repeat this procedure until
 the bottom level of the hierarchy has been reached.  Once there, we
 continue advancing the grids on the lowest level until they have
-``caught up'' to the coarser parent grids (i.e., $\sum \Delta t_l =
-\Delta t_{l-1}$).  This procedure repeats itself until all grids have
-been advanced by a total time of $\Delta t_0$ (see
-Figure~\ref{fig:wcycle} for a schematic).
-
-\begin{figure}
-\begin{center}
-%\includegraphics[width=0.6\textwidth]{figures/wcycle.eps}
-\includegraphics[width=0.6\textwidth]{figures/timestepping.eps}
-\end{center}
-\caption{\emph{Left:} Example of the timesteps on a 2-level AMR
-  hierarchy.  \enzo\ does not restrict the timesteps on the finer levels
-  to be a factor of $1/2^n$ smaller.  \emph{Right:} The order in which
-  the AMR grids are evaluated on each level.\vspace{1ex}}
-\label{fig:wcycle}
-\end{figure}
+``caught up'' to the coarser parent grids.  This procedure repeats itself until all grids have
+been advanced by the desired time.
 
@@
 compared to the total energy $E$.  This situation presents a problem
-because in the total energy method the temperature is computed by
+because in the total energy method the thermal energy is computed by
 subtracting one large number from another (i.e. the kinetic energy
@@
 the baryon fields and particles in the simulation, there are also a
-number of options for specifying static gravitational fields, and in
-fact including self-gravity is optional.
+number of options for specifying static gravitational fields.
 
@@
 subgrids are included from the beginning in order to improve the
-initial spatial and baryonic mass resolution, and these subgrid are
+initial spatial and baryonic mass resolution, and these subgrids are
 populated with lower-mass particles to correspondingly improve the

### 2013-06-25 parallel.tex
Many small changes to all files -- GB final edits

@@
 cell counts per core would in principle help scaling by reducing the
-amount of inter-process communication needed, but larger cell counts
-are typically impractical on most machines due to the amount of memory
-available on each node.
+amount of inter-process communication needed.
 
@@
 substantially improve scaling by reducing the amount of communication
-per grid tile, as described in the previous paragraph; however, the
-optimal ratio of OpenMP threads per MPI task seems to vary
-substantially between computational platforms and astrophysical
-problems/required \enzo\ physics modules, and thus we hesitate to
-provide strict guidelines here.
+per grid tile, as described in the previous paragraph.
 
@@
 drastically throughout the calculation.  For example, at the beginning
-of the simulation when the densities are nearly uniform, only the
+of a cosmological simulation, when the densities are nearly uniform, only the
 static grid is required and the calculation progresses rapidly.
@@
 cosmological runs, with only slight variations depending on the exact
-problem that is being simulated.  The top-left panel
+problem that is being simulated.  The top-right panel
 shows the ratio of $f_{l}/f_{l+1}$ which, remarkably, hovers around
@@
 the static levels, as the bottom-right panel demonstrates.  In this
-case, adding more levels of refinement would somewhat increase the
+case, adding more levels of refinement would substantially increase the
 cost of the simulation.
@@
 on the first and subsequent levels by this factor.  Again, we have
-empirically demonstrated that this scaling is approximately valid,
+empirically tested that this scaling is approximately valid,
 provided that we keep the maximum level constant.

### 2013-06-25 table_of_variables_and_constants.tex
Many small changes to all files -- GB final edits

--- 2013-06-25 table_of_variables_and_constants.tex
@@
   $\vecB$ & comoving magnetic field strength \\
-  $p$ & thermal pressure \\
-  $p^*$ & total (thermal + magnetic) isotropic pressure \\
+  $p$ & comoving thermal pressure \\
+  $p^*$ & comoving total (thermal + magnetic) isotropic pressure \\
   $\gamma$ & ideal gas ratio of specific heats \\
@@
   $\Gamma$ & radiative heating rate \\
-  $Q$ & artificial viscosity stress tensor (only for Zeus hydro) \\[3pt]
+  $Q$ & artificial viscosity (only for \zeus\ hydro) \\[3pt]
 
@@
   $c_{\rm s}$ & sound speed \\
-  $z$ & redshift \\
+  $z$ & redshift
   \enddata

### 2013-06-25 tests-onezonecollapse.tex
Many small changes to all files -- GB final edits

@@
 \end{equation}
-
 where the collapse timescale, $t_{\rm col}$, is
@@
 \end{equation}
-
 where $\Lambda$ is the cooling rate in units of erg s$^{-1}$ g$^{-1}$

### 2013-06-25 tests-orszagtang.tex
Many small changes to all files -- GB final edits

@@
 other tests in this section, in the absence of gravity or chemistry/cooling we use
-arbitrary units).  There is a
+dimensionless units).  There is a
 single rotational mode in the velocity, and two in the magnetic field:
@@
 ${\bf B}_0 = (-\sin(2\pi y) \hat{x},\, \sin(4\pi x) \hat{y})$.
-The simulation is evolved to $t=0.48$ (in arbitrary units).  One can see that the structures
+The simulation is evolved to $t=0.48$.  One can see that the structures
 are accurately represented as compared to, for example,

### 2013-06-25 tests-vandv.tex
Many small changes to all files -- GB final edits

@@
 problems, as well as the scripts to generate the figures from them,
-are available at \url{http://bitbucket.org/bwoshea/enzo-method-paper},
+are available at \url{http://bitbucket.org/enzo/enzo-method-paper},
 the Bitbucket repository for the \enzo\ method paper.  The
@@
 \url{https://bitbucket.org/enzo/enzo-dev} repository of \enzo\ with
-changeset \texttt{5d8c412}.
+changeset \texttt{5d8c412}, which corresponds to \enzo\ 2.3.
 
@@
 typically superior in terms of resolving fluid features due to its
-higher-order (and artificial viscosity-free) Piecewise Parabolic
-Method hydrodynamics solver.  When examining classical test cases such
+higher-order (and artificial viscosity-free) PPM hydrodynamics 
+solver.  When examining classical test cases such
 as the Santa Barbara project \citep{SantaBarbara}, \enzo\ forms galaxy
@@
 AMR code such as \enzo\ may experience.  More specific comparisons, including one
-specifically testing the linear and nonlinear growth of the
+testing the linear and nonlinear growth of the
 Kelvin-Helmholz instability \citep{2012ApJS..201...18M}, as well as a
@@
 \citep{2010MNRAS.401.2463R}, show that \enzo, and in particular its
-implementation of the Piecewise Parabolic Method hydrodynamics solver,
+implementation of the PPM hydrodynamics solver,
 converge to the correct solution as expected, and generally provide
@@
 in at least a subset of the tests.
-\citet{2011ApJ...726...55T} shows the results of varied reaction rate
+\citet{2011ApJ...726...55T} show the result of varying reaction rate
 coefficients for the formation of molecular hydrogen via the 3-body

### 2013-06-25 tests-wavepool.tex
Many small changes to all files -- GB final edits

@@
 In this test we use only our standard, low amount of quadratic artificial viscosity.
-These waves could be damped by additional viscosity, but we do not add any in order to be sensitive to any artifacts at the grid boundary.
+The trailing oscillations could be damped by additional viscosity, but we do not add any in order to be sensitive to any artifacts at the grid boundary.
 The remaining panels show that the trailing oscillations continue, but do not generate any additional noise.
@@
 
-Finally, the MUSCL solver shows also shows a very clean entry and exit from the refined region without any oscillations, although with mild damping on exit.
+Finally, the MUSCL solver also shows a very clean entry and exit from the refined region without any oscillations, although with mild damping on exit.
 However, because we use a piecewise linear reconstruction, the wave is spread more than with the other methods.

### 2013-06-25 numerical-ppm.tex
One more

@@
 
-Once we have the reconstruction, the primary quantities ($p$, $\rho$, $v$ and $E$) are averaged over the domains corresponding to the three characteristics $C = u-c$, $u$, or $u+c$ (where $c$ is the sound speed in a cell):
-\begin{equation}
-      q^{C+}_{j+1/2} =
-          {1 \over \Delta t C^n_j} \int^{x_{j+1/2}}_{x_{j+1/2}-\Delta t C^n_j} q_j(x) dx
-\end{equation}
-where $q^{C+}$ refers to the left characteristic $C$ (and a similar integral is done for the right characteristic $q^{C-}$).
+Once we have the reconstruction, the primary quantities ($p$, $\rho$, $v$ and $E$) are averaged over the domains corresponding to the three characteristics $u-c$, $u$, or $u+c$ (where $c$ is the sound speed in a cell).
 The linearized gas dynamics equations are then used to compute second-order accurate left and right states that take into account the multiple wave families.

### 2013-06-25 numerical-starformation.tex
Removed commented-out text from lots of files - basically vestigial junk.

@@
 
-% \red{( (BWO) I am a little worried that this section is too
-%   long. The editorial hatchet may be called for, but I think somebody
-%   else (who didn't write the section in the first place) needs to
-%   think it through. )}
-
 \subsubsection{Overview}
-% Overview of methodology
-
-% \red{(BWO: can this whole paragraph go?)}
-% While the physics discussed in other sections of this paper is crucial
-% to the study of cosmological structure formation, most observations of
-% galaxies in the infrared, optical, and ultraviolet wavelengths are of
-% stellar populations.  Furthermore, stars eject energy and metal-
-% enriched gas throughout their lives, drastically modifying their own
-% environments' cooling properties.  Broadly speaking, the formation of
-% galaxies and galaxy clusters cannot be completely modeled without
-% including the feedback of energy and metals from stellar populations.
-% In particular, it is thought that feedback is crucial for suppressing
-% the large numbers of dwarf-like galaxies that CDM theories predict
-% \citep{1991ApJ...381...14L,1991ApJ...379...52W}.  An early burst of
-% star formation could remove a large fraction of cold gas from such
-% systems \citep{1978MNRAS.183..341W,1991ApJ...367...45C}.  Also, the
-% unexpectedly low luminosities of small clusters and groups (relative
-% to the richest clusters) is commonly explained through feedback
-% \citep{1991ApJ...383..104K}.  Energy ejected during the formation of
-% the cluster ellipticals lowers the central density and hence the X-ray
-% luminosity of such clusters \citep{1997ApJ...484L..21C}.  For all of
-% these reasons, it is critical to include star formation and the
-% feedback of energy and metals into the interstellar and intergalactic
-% media in cosmological simulations.
-
-% Modeling star formation in the context of galaxy formation forces one
-% to confront two fundamental problems. The dynamical range required to
-% study galaxy formation and evolution is substantial, and attempting to
-% directly resolve the formation of individual stars in the context of a
-% Milky Way-type galaxy is computationally unfeasible.  Also, a strong
-% theoretical understanding of the details of star formation is
-% currently lacking.  Due to these challenges, models of star formation
-% in cosmological simulations are largely phenomenological, and the
-% ``stars'' that are formed are in most cases meant to represent an
-% ensemble stellar population (although not all; see
-% Section~\ref{sec:starform_pop3}).
 
@@
 \enzo.
-%The \enzo\ code includes approximately ten models for star formation
-%and feedback.  
 Broadly speaking, these methods all work in similar
@@
 \label{cendens}
-%\end{equation} 
-%\begin{equation}
 \div \myvec{v}_b & < & 0, \\
 \label{cencont}
-%\end{equation}
-%\begin{equation}
 t_{\rm cool} & < & t_{\rm dyn} \equiv \sqrt{3 \pi / 32G \rho_{\rm tot}}, \\
-%\end{equation}
-%\begin{equation}
 m_{b} & > & m_{J} \equiv G^{-3/2} \rho_{b}^{-1/2}c^{3} 
@@
 where 
-%$\rho_b$ is the baryon density in that cell, $\bar{\rho_b}$ is
-%the mean baryon density in the simulation, 
 $\eta$ is the user-defined
 overdensity threshold, 
-%$\myvec{v}$ is the velocity of the gas in the
-%cell, $\rho_{d}$ is the dark matter density, $\rho_{\rm tot} =
-%\rho_{b} + \rho_{d}$, 
 and $m_{b}$ and $m_{j}$ are the baryonic mass in the
@@
 
-% In addition, we have added a stochastic star formation algorithm that
-% keeps track of all of the sub-mass threshold stars that should have
-% been created and when the total mass of uncreated stars is greater
-% than the minimum mass, a star particle is created.
-
-% The feedback of energy and metals into the baryon gas is similar to
-% the Kravtsov feedback, with some important differences.
-
 The star formation algorithm creates each star particle
@@
 where $\tau_0 = (t-t_{\rm form})/t_{\rm dyn}$ and $\tau_1 = (t+\Delta t-t_{\rm form})/t_{\rm dyn}$.
-
-%\begin{eqnarray}  
-%\Delta m_{\rm sf} =  m_{*} \left[ \left(1 + \frac{t-t_{\rm
-%        form}}{t_{\rm dyn}}  \right)
-%%\nonumber \times \\ % uncomment for 2-column
-%		   \exp{\left( \frac{-(t-t_{\rm form})}{t_{\rm dyn}} \right) }
-%%                    \nonumber \\ 
-%	       - \left( 1 + \frac{t+ \Delta t-t_{\rm form}}{t_{\rm dyn}} \right) 
-%%	       \nonumber \\ 
-%                    \exp{ \left( \frac{-(t+\Delta t-t_{\rm
-%                            form})}{t_{\rm dyn}} \right)  }
-%                   \right]
-%\end{eqnarray}
-%which can be represented more clearly in integral form:
-%
-%\begin{eqnarray}
-%\int_{t}^{t+Dt} \frac{dM}{dt} dt = \int_{t}^{t+Dt} m_{*} 
-%\frac{dt}{t_{\rm dyn}} \left( \frac{t-t_{\rm form}}{t_{\rm dyn}} \right) 
-%%\nonumber \times \\ 
-%\exp{ \left( - ~ \frac{t-t_{\rm form}}{t_{\rm dyn}} \right) }
-%\end{eqnarray}
 
@@
 densest environments in a given simulation.
-%Given the
-%simulation-dependent nature of these parameters, it is critical to set
-%them carefully, and with some sensitivity testing.
 
@@
 \rho_{\rm gas}/\tau_*$. 
-%where $\rho_{\rm gas}$ and $\rho_*$ are the
-%baryon gas and stellar densities, respectively.  
 This ``constant
@@
 a rough approximation of the results of \citet{1995ApJS..101..181W}.
-
-% Kravtsov reports that simulations with this algorithm reliably
-% reproduce the star formation rate-gas surface density relation of the
-% Schmidt law, and are not particularly sensitive to numerical
-% parameters \citep{2003ApJ...590L...1K}.  He also notes that this
-% algorithm is surprisingly insensitive to the presence or absence of
-% feedback and details of the cooling and heating properties of the gas,
-% which suggests that the global star formation rate is determined by
-% gravitationally- driven supersonic turbulence (on large scales) rather
-% than stellar feedback or thermal instabilities on small scales.
 
@@
 3.2.1 of \citet{2012MNRAS.427..311W}.
-
-%%% Local Variables: 
-%%% mode: latex
-%%% TeX-master: "ms"
-%%% End:

### 2013-06-25 numerical-timestep.tex
Removed commented-out text from lots of files - basically vestigial junk.

@@
 
-% \begin{equation}
-% \Delta t_{} = 
-% \label{eqn:dt}
-% \end{equation}
-
-% \begin{equation}
-% \Delta t_{} = 
-% \label{eqn:dt}
-% \end{equation}
-
 In equations~(\ref{eqn:dthydro})-(\ref{eqn:dtcond}), the $\min ( \ldots)_L$ 
@@
 an explicit finite difference discretization of the Euler equations.
-%Effectively this condition forces the timestep to be small enough such
-%that any changes in the fluid propagate less than a single grid
-%spacing, $\Delta x$.  
 In this equation, $\kappa_{\rm hydro}$ is a dimensionless
@@
 $n_b$ is the baryon number density, 
-%defined as $n_b = \rho_b / \mu
-%m_p$, where $\mu$ is the mean molecular weight, 
 $\kappa_{\rm sp}(T)$ is the Spitzer thermal conductivity, and f$_{\rm sp}$ is a
@@
 substantial structure has formed.
-
-%%% Local Variables: 
-%%% mode: latex
-%%% TeX-master: "ms"
-%%% End:

### 2013-06-25 overview.tex
Removed commented-out text from lots of files - basically vestigial junk.

@@
 
-%\subsubsection{Magnetohydrodynamics and Gravity}
-
 The Eulerian equations of ideal magnetohydrodynamics (MHD) including
@@
   - \frac{\dot{a}}{a} \rho \vecv - \frac{1}{a} \rho \grad \phi,
-% 
-%          + \frac{1}{a} ( \vecv \cdot \grad ) \vecv
-%    & = & - \frac{\dot{a}}{a} \vecv 
-%          - \frac{1}{a \rho} \grad p
-%          - \frac{1}{a \rho} (\nalba \times \vecB) \times \vecB
-%          - \frac{1}{a} \grad \phi, 
   \label{eq:momentum} \\
@@
   - \Lambda + \Gamma + \frac{1}{a^2} \div \fcond,
-% 
-%         + \frac{1}{a} \vecv \cdot \grad E
-%  & = & - \frac{\dot{a}}{a} (3 \frac{p}{\rho} + {\vecv \cdot \vecv} )
-%       - \frac{1}{a \rho } \div (p \vecv) 
-%%\nonumber   & &   %uncomment for 2-column
-%        - \frac{1}{a} \vecv \cdot \grad \phi 
-%               + \Gamma - \Lambda.
   \label{eq:total_energy}  \\
@@
 
-%\begin{eqnarray} 
-%\frac{\partial \rho}{\partial t} 
-%          + \vecv \cdot \grad \rho
-%    & = & - \rho \div \vecv,
-%        \label{eq:mass} \\
-%%
-%\frac{\partial \vecv }{ \partial t} 
-%          + ( \vecv \cdot \grad ) \vecv 
-%    & = & - \frac{1}{\rho} \grad p
-%          - \grad \phi,
-%        \label{eq:momentum} \\
-%%
-% \frac{\partial E}{\partial t} + \vecv \cdot \grad E
-%    & = & - \frac{1}{\rho} \div (p \vecv) 
-%          - \vecv \cdot \grad \phi
-%%\nonumber          \\  & &     % uncomment for 2-column
-%     + \Gamma - \Lambda
-%        \label{eq:total_energy}
-%\end{eqnarray}
-%
 In these equations, $E$, $\rho$, $\vecv$, and $\vecB$ are the comoving
@@
 for more details).
-%These definitions result in the following comoving equivalents to eqs. 
-%(\ref{eq:mass})-(\ref{eq:potential}).
-%%
-%
-%\begin{eqnarray} 
-%\frac{\partial \rhop}{\partial t} 
-%          + \frac{1}{a} \vecvp \cdot \grad \rhop 
-%    & = & - \frac{1}{a} \rhop \div \vecvp, 
-%        \label{eq:mass_prime} \\
-%%
-%\frac{\partial \vecvp}{\partial t}  
-%          + \frac{1}{a} ( \vecvp \cdot \grad ) \vecvp 
-%    & = & - \frac{\dot{a}}{a} \vecvp 
-%          - \frac{1}{a \rhop} \grad \pp 
-%          - \frac{1}{a} \grad \phip, 
-%        \label{eq:momentum_prime} \\
-%%
-%\frac{\partial \Ep}{\partial t} 
-%         + \frac{1}{a} \vecvp \cdot \grad \Ep 
-%   & = & - \frac{\dot{a}}{a} (3 \frac{\pp}{\rhop} + {\vecvp \cdot \vecvp} )
-%         - \frac{1}{a \rhop } \div (p \vecvp) 
-%%\nonumber   & &   %uncomment for 2-column
-%         - \frac{1}{a} \vecvp \cdot \grad \phip 
-%                + \Gamma - \Lambda.
-%       \label{eq:total_energy_prime} \\
-%%
-%\Ep  & = & \ep + \frac{1}{2} {\vecvp \cdot \vecvp},
-%        \label{eq:total_energy_def_prime} \\
-%%
-%\ep                  & = & 
-%        \pp/\left[ \left( \gamma - 1 \right) \rhop \right],
-%      \label{eq:state_prime} \\ 
-%{\nabla}^2 \phip    & = & 
-%        \frac{4\pi G}{a} ( \rhop_{total} - \rhop_0 ).
-%      \label{eq:poisson}
-%\end{eqnarray}
-%%
-%%We use the proper peculiar baryonic velocity $\vecv_b \equiv a(t) d\vecx / dt$, proper pressure $p$, and modified gravitational potential $ \phi $ which is related to the potential in proper coordinates $\Phi$ by
-%%$$\phi \equiv \Phi + \frac{1}{2} a \ddot{a} \vecx^2. $$
-%%The density, however, is comoving:
-%%$$\rho_b \equiv \rho_{b,proper}a(t)^3, $$
 The expansion parameter $a \equiv 1/(1 + z)$ follows the expansion of
@@
 contribution, and $\Lambda_c$ is the cosmological constant.
-%$\gamma$ is the ratio of specific heats, and $\Lambda$ is the cosmological constant.
-%We note that if $\gamma$ is not $5/3$ then $E$ in the first term on
-%the right side of eq.~(ref{eq:total_energy}) must be replaced with 
-%eq.~(\ref{eq:state}).
 This system of equations is limited to the non-relativistic regime and
@@
 simulate a much broader range of astrophysical environments.
-
-%\subsubsection{Particles}
 
@@
 Poisson's equation (Equation \ref{eq:potential}).
-
-%\subsubsection{Chemistry}
 
@@
 
-%\subsubsection{Radiation transport: moment method}
-
 As an alternative to the ray-casting strategy, \enzo\ can also solve
@@
   \frac{\partial e_c}{\partial t} &=& -\frac{2\dot{a}}{a} e_c + \Gamma - \Lambda,
-%  \label{eq:fld_chemistry}
-%  \partial_t {\mathrm n}_i + 
-%    \frac1a \div\left({\mathrm n}_i \vecv_b\right) &=& 
-%    \alpha_{i,j} {\mathrm n}_e {\mathrm n}_j - {\mathrm n}_i
-%    \Gamma_i^{ph}, \quad i=1,\ldots,N_s,
 \end{eqnarray}
@@
 chemical cooling.
-%and ${\mathrm n}_i$ correspond to the chemical number densities of
-%HI, HII, HeI, HeII and HeIII, and ${\mathrm n}_e$ is the electron
-%number density.  
 Here, we define $E_r$ through first assuming a fixed frequency
@@
 chemical cooling rate.
-%, $\alpha_{i,j}$ are the reaction rate coefficients defining
-%the interactions between species ${\mathrm n}_i$ and ${\mathrm n}_j$,
-%and $\Gamma_i^{ph}$ are the radiation induced photo-ionization rates.
-
-%\subsubsection{Thermal conduction}
 
@@
 placed within these coarse regions.
-% (we use the notation `below' to refer to
-% finer grids and `above' for coarser grids). 
 We restrict the ratio between cell sizes of parent and child grids to be an integer, typically 2
@@
 determine a good placement of subgrids.
-%  These subgrids must not overlap one another, must cover all flagged
-% cells and their neighbouring cells, and be above a preset efficiency
-% threshold, where the efficiency of a cell is defined as the ratio of
-% flagged cells to total cells. 
 Once these new subgrids have been identified, the solution from the
@@
 advances the hydrodynamic equations in the following steps:
-%\begin{enumerate}
-% \item 
 (i) Construct monotonic parabolic interpolation of cell average data,
 for each fluid quantity;
-% \item 
 (ii) Compute interface states by averaging the parabola over the
 domain of dependence for each interface;
-% \item 
 (iii) Use interface data to solve the Riemann problem;
- %\item 
 (iv) Difference the interface fluxes to update the cell average
 quantities.
-%\end{enumerate}
 
@@
 collisionless mass resolution.
-%% One particle per initial grid point seems to provide approximately
-%% equal sampling between the dark matter and gas. \mqk{This
-%% contradicts the conclusions of O'Shea et al. 2005 that an initial
-%% grid twice as fine as the mean particle separation is necessary to
-%% properly capture the abundance of low mass halos.}
 
@@
 D$^+$).
-%(We note that a reduced version
-%of the chemistry, used only with the flux-limited diffusion radiation
-%transport, includes only hydrogen species.)
 The cooling and heating due to these species is included (see the next
@@
 these $\Delta t$.
-
-%%% Local Variables:
-%%% mode: latex
-%%% TeX-master: "ms"
-%%% End:

### 2013-06-25 introduction.tex
Reformatting tex files so they're all standardized in terms of paragraph wrapping, etc. (no content changed, just formatting.)

@@
 Due to the high spatial and temporal dynamical ranges involved,
-astrophysical and cosmological phenomena present a taxing challenge for
-simulators. To tackle such situations, a number of numerical
+astrophysical and cosmological phenomena present a taxing challenge
+for simulators. To tackle such situations, a number of numerical
 techniques have been developed that can be broadly split into
@@
 \citep[e.g.,][]{Ruffert94, Anninos94}. Adding adaptivity is a more
-recent enhancement, and there are now a number of codes that
-possess this feature, both with and without hydrodynamics
-\citep{Couchman91, Jessop94, Suisalu95, Splinter96, Gelato97, ART97,
-Truelove98, flash_method, MLAPM01, Yahagi01, RAMSES, Quilis04,
-Ziegler05, Zhang06, Astrobear09, Pluto-amr, GAMER, Nyx}. Of these,
-perhaps the most comparable and widely used are FLASH
-\citep{flash_method}, which uses grid blocks of fixed size, and RAMSES
-\citep{RAMSES} and ART \citep{ART97}, both of which refine individual
-cells. It is also possible to deform the grid to obtain high
-resolution \citep[e.g.,][]{Gnedin95, Xu97, Pen98}, and more recently a
-few codes have adopted an unstructured approach based on a moving
-Voronoi mesh \citep{Arepo10, Tess11}.
+recent enhancement, and there are now a number of codes that possess
+this feature, both with and without hydrodynamics \citep{Couchman91,
+Jessop94, Suisalu95, Splinter96, Gelato97, ART97, Truelove98,
+flash_method, MLAPM01, Yahagi01, RAMSES, Quilis04, Ziegler05, Zhang06,
+Astrobear09, Pluto-amr, GAMER, Nyx}. Of these, perhaps the most
+comparable and widely used are FLASH \citep{flash_method}, which uses
+grid blocks of fixed size, and RAMSES \citep{RAMSES} and ART
+\citep{ART97}, both of which refine individual cells. It is also
+possible to deform the grid to obtain high resolution
+\citep[e.g.,][]{Gnedin95, Xu97, Pen98}, and more recently a few codes
+have adopted an unstructured approach based on a moving Voronoi mesh
+\citep{Arepo10, Tess11}.
 
-In this paper, we present \enzo\ 2.3, a structured adaptive mesh refinement
-(SAMR) code. Originally developed for cosmological hydrodynamics,
-\enzo\ has since been used on a wide variety of problems.  It has
-grown to become a general tool for astrophysical fluid dynamics and is
-intended to be efficient, accurate and easily extended to include new
-capabilities.  Although many of the components of the \enzo\ code have
-been described in previous publications \citep{1995CoPhC..89..149B,
-BryanThesis96, Bryan97a, Bryan97b, Norman99, BryanCompSci99, Bryan01,
-Oshea04, 2007arXiv0705.1556N,
+In this paper, we present \enzo\ 2.3, a structured adaptive mesh
+refinement (SAMR) code. Originally developed for cosmological
+hydrodynamics, \enzo\ has since been used on a wide variety of
+problems.  It has grown to become a general tool for astrophysical
+fluid dynamics and is intended to be efficient, accurate and easily
+extended to include new capabilities.  Although many of the components
+of the \enzo\ code have been described in previous publications
+\citep{1995CoPhC..89..149B, BryanThesis96, Bryan97a, Bryan97b,
+Norman99, BryanCompSci99, Bryan01, Oshea04, 2007arXiv0705.1556N,
 WangAbelZhang08,ReynoldsHayesPaschosNorman2009,Collins10,Wise11_Moray},
@@
 a wide variety of problems, resulting in the publication of more than
-100 peer-reviewed papers. The
-variety of astrophysical systems that \enzo\ has been used for include
-galaxies \citep{2003ApJ...587...13T, 2012MNRAS.425..641L,
-  2013MNRAS.432.1989S}, galaxy
-clusters \citep{Loken02, Xu11, Skillman13}, the interstellar medium
+100 peer-reviewed papers. The variety of astrophysical systems that
+\enzo\ has been used for include galaxies \citep{2003ApJ...587...13T,
+2012MNRAS.425..641L, 2013MNRAS.432.1989S}, galaxy clusters
+\citep{Loken02, Xu11, Skillman13}, the interstellar medium
 \citep{Slyz05}, the intergalactic medium \citep{Fang01,
 2011ApJ...731....6S}, the circumgalactic
-medium~\citep{2013MNRAS.430.1548H}, cooling flows \citep{Li12, 2013ApJ...763...38S},
-turbulence \citep{Kritsuk04, 2007ApJ...665..416K, 2009JPhCS.180a2020K, Collins11}, the
-formation of the first stars \citep{ABN02, 2007ApJ...654...66O,
-2009Sci...325..601T, Xu08}, and the formation of stars in our own Galaxy
-\citep{Collins11, 2011ApJ...727L..20K, Collins12a}.
+medium~\citep{2013MNRAS.430.1548H}, cooling flows \citep{Li12,
+2013ApJ...763...38S}, turbulence \citep{Kritsuk04,
+2007ApJ...665..416K, 2009JPhCS.180a2020K, Collins11}, the formation of
+the first stars \citep{ABN02, 2007ApJ...654...66O,
+2009Sci...325..601T, Xu08}, and the formation of stars in our own
+Galaxy \citep{Collins11, 2011ApJ...727L..20K, Collins12a}.
 
@@
 expanded upon by others.  Our public release of the \enzo\ code (which
-can be found at \url{http://enzo-project.org}) represents our
-attempt to meet the first of these obligations; this paper represents
-our attempt to meet the second.
+can be found at \url{http://enzo-project.org}) represents our attempt
+to meet the first of these obligations; this paper represents our
+attempt to meet the second.
 

### 2013-06-25 numerical-amr.tex
Removed commented-out text from lots of files - basically vestigial junk.
Reformatting tex files so they're all standardized in terms of paragraph wrapping, etc. (no content changed, just formatting.)

@@
 
-%\subsubsection{SAMR Overview}
+The back-bone of the SAMR idea is the patch, or grid, which we take to
+be rectilinear to simplify bookkeeping (in practice, the lower-level
+routines such as the hydro solvers all assume Cartesian coordinates).
+The single root grid covers the entire computational domain and plays
+the role of the root node in the grid hierarchy.  Subgrids are placed
+such that they cover (hopefully small) sub-volumes of the root grid at
+higher resolution.  This structure forms a tree that may be extended
+to arbitrary depths by adding additional grids to refine regions in
+the subgrids.  We use the following notation to describe grid
+relationships.  A grid's \textit{parent} is the coarse grid that
+completely contains it.  The \textit{root grid} is the most
+coarsely-resolved grid, and it has no parent.  A \textit{child grid}
+is a grid completely contained by its parent grid (note that a grid
+may have many children but only one parent), while sibling grids are
+those that have the same resolution, but not necessarily the same
+parent.  We also refer to levels of the hierarchy, where the root grid
+is labeled level zero and all grids with the same resolution are on
+the same level.
 
-%Unlike moving mesh methods \citep{1995ApJS..100..269P,1995ApJS...97..231G} or  methods that subdivide individual cells \citep{Adjerid}, Berger \& Collela's AMR (also referred to as \emph{structured} AMR) utilizes an adaptive hierarchy of grid patches at varying levels of resolution.
-The SAMR method described here is similar to \citet{Berger89}.
-A more complete discussion of the history and development of this type of algorithm is available in \citet{Neeman96}.
+In order to simplify bookkeeping, a number of restrictions are imposed
+on the subgrids:
 
-The back-bone of the SAMR idea is the patch, or grid, which we take to be rectilinear to simplify bookkeeping (in practice, the lower-level routines such as the hydro solvers all assume Cartesian coordinates).
-The single root grid covers the entire computational domain and plays the role of the root node in the grid hierarchy.
-Subgrids are placed such that they cover (hopefully small) sub-volumes of the root grid at higher resolution.
-This structure forms a tree that may be extended to arbitrary depths by adding additional grids to refine regions in the subgrids.
-We use the following notation to describe grid relationships.
-A grid's \textit{parent} is the coarse grid that completely contains it.
-The \textit{root grid} is the most coarsely-resolved grid, and it has no parent.
-A \textit{child grid} is a grid completely contained by its parent grid (note that a grid may have many children but only one parent), while sibling grids are those that have the same resolution, but not necessarily the same parent.
-We also refer to levels of the hierarchy, where the root grid is labeled level zero and all grids with the same resolution are on the
-same level.
-
-In order to simplify bookkeeping, a number of restrictions are imposed on the subgrids:
 \begin{itemize}
- \item The refinement factor $r$, or the ratio of the coarse cell width to the fine cell width, must be an integer.
-In practice, we generally adopt $r=2$, since tests have shown this typically provides the best performance.
- \item Each subgrid must begin and end on the boundary of a coarse cell.  This means that the number of cells in a patch will be a multiple of the refinement factor.
+ \item The refinement factor $r$, or the ratio of the coarse cell
+   width to the fine cell width, must be an integer.  In practice, we
+   generally adopt $r=2$, since tests have shown this typically
+   provides the best performance.
+ \item Each subgrid must begin and end on the boundary of a coarse
+   cell.  This means that the number of cells in a patch will be a
+   multiple of the refinement factor.
  \item A subgrid must be completely enclosed by its parent.
@@
 
-The AMR grid patches are the primary data structure in \enzo.
-Each patch is treated as an individual object that can contain both field variables and particle data.
-Individual grids are organized into a  dynamic, distributed hierarchy of mesh patches.
-%See Figure~\ref{fig.2.amrhierdens} for an example of an \enzo\ simulation performed using this AMR scheme where both the baryon density and AMR grid hierarchy are shown together. 
-% GB: I think this figure is too application-specific
+The AMR grid patches are the primary data structure in \enzo.  Each
+patch is treated as an individual object that can contain both field
+variables and particle data.  Individual grids are organized into a
+dynamic, distributed hierarchy of mesh patches.
 
@@
 
-Each grid patch in \enzo\ contains arrays of values for baryon and particle quantities.
-Grids are partitioned into a core of \emph{active zones} and a surrounding layer of \emph{ghost zones}, as shown in Figure~\ref{fig.amr_hierarchy}.
-The active zones store field values and ghost zones are used to temporarily store values that have been obtained directly from neighboring grids or interpolated from a parent grid.
-These zones are necessary to accommodate the computational stencil of the (magneto)hydrodynamics solvers (Sections~\ref{sec.hydro.ppm} through~\ref{sec.hydro.zeus}) and the gravity solver (Section~\ref{sec.gravity}).
-The PPM and \zeus\ hydro solvers require 3 layers of ghost zones,  and the gravity solver requires 6 (although only for the density and potential fields).
-This can lead to significant memory and computational overhead, particularly for smaller grid patches at high levels of refinement.
+Each grid patch in \enzo\ contains arrays of values for baryon and
+particle quantities.   Grids are partitioned into a core of
+\emph{active zones} and a surrounding layer of \emph{ghost zones}, as
+shown in Figure~\ref{fig.amr_hierarchy}.  The active zones store field
+values and ghost zones are used to temporarily store values that have
+been obtained directly from neighboring grids or interpolated from a
+parent grid.  These zones are necessary to accommodate the
+computational stencil of the (magneto)hydrodynamics solvers
+(Sections~\ref{sec.hydro.ppm} through~\ref{sec.hydro.zeus}) and the
+gravity solver (Section~\ref{sec.gravity}).  The PPM and \zeus\ hydro
+solvers require 3 layers of ghost zones,  and the gravity solver
+requires 6 (although only for the density and potential fields).  This
+can lead to significant memory and computational overhead,
+particularly for smaller grid patches at high levels of refinement.
 
-From the point of view of the hydro solver, each grid is solved as an independent computational fluid dynamics (CFD) problem, with Dirichlet boundary conditions stored in the ghost zones.
-At the beginning of each timestep, cells on a given level fill ghost zones by interpolating from the parent and copying from neighbors on the same level.
-Grids then correct the flux at the interface boundary between fine and coarse zones, and finally project their active zone data to the parent grid.
+From the point of view of the hydro solver, each grid is solved as an
+independent computational fluid dynamics (CFD) problem, with Dirichlet
+boundary conditions stored in the ghost zones.  At the beginning of
+each timestep, cells on a given level fill ghost zones by
+interpolating from the parent and copying from neighbors on the same
+level.  Grids then correct the flux at the interface boundary between
+fine and coarse zones, and finally project their active zone data to
+the parent grid.
 
@@
 \begin{center}
-%\includegraphics[width=0.6\textwidth]{figures/wcycle.eps}
 \includegraphics[width=0.6\textwidth]{figures/timestepping.eps}
@@
 
-Timesteps are determined on each level as described in Section~\ref{sec.timestepping}, and the hierarchy is advanced on
-a level-by-level basis in a W-Cycle.    Beginning with the coarsest level, $l$, all grids on that level are advanced one
-timestep.
-Then, one timestep is taken on all grids at the next level of refinement, $l+1$, and so on until the finest level is resolved.
-The finest level is then advanced, using as many steps as it takes to reach the level immediately above.
-The finest level is then synched to the level above it, which then proceeds forward in time one more step.
-This is shown graphically in Figure~\ref{fig:wcycle}.
+Timesteps are determined on each level as described in
+Section~\ref{sec.timestepping}, and the hierarchy is advanced on a
+level-by-level basis in a W-Cycle.  Beginning with the coarsest level,
+$l$, all grids on that level are advanced one timestep.  Then, one
+timestep is taken on all grids at the next level of refinement, $l+1$,
+and so on until the finest level is resolved. The finest level is then
+advanced, using as many steps as it takes to reach the level
+immediately above.  The finest level is then synched to the level
+above it, which then proceeds forward in time one more step.  This is
+shown graphically in Figure~\ref{fig:wcycle}.
 
-At the end of every timestep on every level, each grid updates its ghost zones by exchanging information with its neighboring grid patches (if any exist) and/or by interpolating from a parent grid.
-In addition, cells are examined to see where refinement is required and the entire grid hierarchy is rebuilt at that level (including all more highly refined levels).
-The timestepping and hierarchy advancement/rebuilding process described here is repeated recursively on every level to the specified maximum level of refinement in the simulation.
+At the end of every timestep on every level, each grid updates its
+ghost zones by exchanging information with its neighboring grid
+patches (if any exist) and/or by interpolating from a parent grid.  In
+addition, cells are examined to see where refinement is required and
+the entire grid hierarchy is rebuilt at that level (including all more
+highly refined levels).  The timestepping and hierarchy
+advancement/rebuilding process described here is repeated recursively
+on every level to the specified maximum level of refinement in the
+simulation.
 
-The basic control algorithm looks very much like any used in a grid-based code of a single resolution, and can be written schematically as:
+The basic control algorithm looks very much like any used in a
+grid-based code of a single resolution, and can be written
+schematically as:
+
 \begin{verbatim}
@@
 \end{verbatim}
+
 The AMR control algorithm is contained within the recursive
-EvolveLevel algorithm, which takes the level upon which to operate as an
-argument.  This looks something like the following pseudo-code:
+EvolveLevel algorithm, which takes the level upon which to operate as
+an argument.  This looks something like the following pseudo-code:
+
 \begin{verbatim}
@@
 
-Each function operates on all the grids on the given level.  In the following sections, we will address each of the major elements in turn.
+Each function operates on all the grids on the given level.  In the
+following sections, we will address each of the major elements in
+turn.
 
@@
 
-In order to solve the hydrodynamic equations on a patch, it is necessary to specify the boundary conditions on that patch.
-There are two cases.
-The first is that the boundary is external to the computational domain.
-This occurs for all the boundaries of the root grid.
-The second possibility is that the boundary is within the computational domain, which is true for most, but not all, subgrids.
+In order to solve the hydrodynamic equations on a patch, it is
+necessary to specify the boundary conditions on that patch.  There are
+two cases.  The first is that the boundary is external to the
+computational domain.  This occurs for all the boundaries of the root
+grid.  The second possibility is that the boundary is within the
+computational domain, which is true for most, but not all, subgrids.
 
-In the first case, we allow for four types of boundary conditions
-(for simplicity, represented here as being at $x=0$, with positive $x$ interior and
-negative $x$ exterior to the domain):
+In the first case, we allow for four types of boundary conditions (for
+simplicity, represented here as being at $x=0$, with positive $x$
+interior and negative $x$ exterior to the domain):
+
 \begin{enumerate}
-  \item{\em Reflecting.} In this case, the boundary behaves like a mirror, with the solution on the boundary reflecting the solution in the computational domain ($q(-x) = q(+x)$) but with the velocity normal to the boundary reversed: $v_x(-x) = -v_x(+x)$.
-  \item{\em Outflow.}  We approximate outflow boundary conditions by duplicating the solution at the edge of the computational domain: $q(-x) = q(0)$.
-  \item{\em Inflow.} In this case, the boundary values are fixed by a pre-determined function ($q(-x) = q0(-x,t)$).
-The code provides a simple way to set inflow values that are constant in both time and space over an entire face, or hooks are provided for special cases where the boundary must be time-varying.
-  \item{\em Periodic.} The boundary solution is obtained from the other side of the grid: $q(-x) = q(x_{\rm{max}}-x)$.
+  \item{\em Reflecting.} In this case, the boundary behaves like a
+    mirror, with the solution on the boundary reflecting the solution
+    in the computational domain ($q(-x) = q(+x)$) but with the
+    velocity normal to the boundary reversed: $v_x(-x) = -v_x(+x)$.
+  \item{\em Outflow.}  We approximate outflow boundary conditions by
+    duplicating the solution at the edge of the computational domain:
+    $q(-x) = q(0)$.
+  \item{\em Inflow.} In this case, the boundary values are fixed by a
+    pre-determined function ($q(-x) = q0(-x,t)$).  The code provides a
+    simple way to set inflow values that are constant in both time and
+    space over an entire face, or hooks are provided for special cases
+    where the boundary must be time-varying.
+  \item{\em Periodic.} The boundary solution is obtained from the
+    other side of the grid: $q(-x) = q(x_{\rm{max}}-x)$.
 \end{enumerate}
 
-In the second case, the value must be either (1) interpolated from a parent or (2) copied from a sibling.
-In practice, we solve this problem by first interpolating all boundaries from the grid's parent and then performing the sibling copy wherever possible, since these values will always be as good as, or superior to, interpolation.
+In the second case, the value must be either (1) interpolated from a
+parent or (2) copied from a sibling.  In practice, we solve this
+problem by first interpolating all boundaries from the grid's parent
+and then performing the sibling copy wherever possible, since these
+values will always be as good as, or superior to, interpolation.
 
-We focus first on conserved, cell-centered quantities.
-The interpolation function should be conservative in the sense that the mean over the interpolated quantity, in a given cell, must return the cell-centered value.
-It should also be monotonic, so that new extrema are not introduced into a smoothly varying field.
+We focus first on conserved, cell-centered quantities.  The
+interpolation function should be conservative in the sense that the
+mean over the interpolated quantity, in a given cell, must return the
+cell-centered value.  It should also be monotonic, so that new extrema
+are not introduced into a smoothly varying field.
 
@@
 accurate\footnote{We could, in principle, build a multi-dimensional
-  interpolator out of a one-dimensional interpolation function;
-  however, at higher order, this would not necessarily satisfy our
-  constraint regarding not introducing new extrema in interpolated quantities.}.  In Appendix~\ref{app:interpolation}, we
-explicitly write down functions in one, two and three dimensions, but
-here we discuss the general design considerations.
+interpolator out of a one-dimensional interpolation function; however,
+at higher order, this would not necessarily satisfy our constraint
+regarding not introducing new extrema in interpolated quantities.}.
+In Appendix~\ref{app:interpolation}, we explicitly write down
+functions in one, two and three dimensions, but here we discuss the
+general design considerations.
 
@@
 one-dimensional Van Leer-type functions.  The reason to focus on the
-cell diagonals is that for a linear function, the extrema will occur at the cell 
-corners, which are in turn linked by the diagonals.  This process can be done in
-two dimensions since there are three constraints (one at the center and one along
-each diagonal), and three parameters in a two-dimensional linear
-function.
+cell diagonals is that for a linear function, the extrema will occur
+at the cell corners, which are in turn linked by the diagonals.  This
+process can be done in two dimensions since there are three
+constraints (one at the center and one along each diagonal), and three
+parameters in a two-dimensional linear function.
 
@@
 parameters (four).  One way to solve this problem would be to perform
-a least-squares fit under the given constraints; however, this is
-too computationally expensive.  Therefore, in Appendix~\ref{app:interpolation} we describe
-a heuristic approach that satisfies four constaints analytically and
-then adjusts the interpolation parameters in such a way as to satisfy the fifth.
-While the result is not necessarily optimal, it does appear to perform
+a least-squares fit under the given constraints; however, this is too
+computationally expensive.  Therefore, in
+Appendix~\ref{app:interpolation} we describe a heuristic approach that
+satisfies four constaints analytically and then adjusts the
+interpolation parameters in such a way as to satisfy the fifth.  While
+the result is not necessarily optimal, it does appear to perform
 adequately in our tests.
-
-%The same procedure can be used for face-centered quantities; however, for the velocity variable in zeus, we drop the extrema constraint.
 
@@
 primary types are cell-centered, face-centered, and node-centered).
-%In this description we will employ the direct Eulerian version of the piecewise
-%parabolic method \citep{1984JCoPh..54..174C,1995CoPhC..89..149B}, which
-%assumes cell-centering for all quantities (actually they are
-%cell-averages).  
-%However, it is also possible to use other schemes,
-%which employ different centering for different variables
-%\citep[e.g.,][]{Stone92a}.  
-The interpolation step, described in
-Section~\ref{sec:interpolation}, must correctly interpolate with the
-selected centering.
+The interpolation step, described in Section~\ref{sec:interpolation},
+must correctly interpolate with the selected centering.
 
@@
 
-Since a refined region is simulated with (at least) two different resolutions (one fine and one coarse), it has two or
-more ``solutions'' (i.e.~values of density, velocity, etc.) that represent the same volume of space.
-This is one important aspect of the SAMR method -- some physical regions are represented by grids at multiple resolutions.
-To
-maintain consistency, the coarse grid quantities are updated with the finer values, using a volume-weighted average of the conserved quantities:
+Since a refined region is simulated with (at least) two different
+resolutions (one fine and one coarse), it has two or more
+``solutions'' (i.e.~values of density, velocity, etc.) that represent
+the same volume of space.  This is one important aspect of the SAMR
+method -- some physical regions are represented by grids at multiple
+resolutions.  To maintain consistency, the coarse grid quantities are
+updated with the finer values, using a volume-weighted average of the
+conserved quantities:
+
 \begin{equation}
@@
 \end{equation}
-where $r$ is the refinement factor, and the sum is over all fine cells that ``cover" the coarse cell.
-The exponent, $d$, refers to the weight of the average.
-For cell-centered quantities, this is the dimensionality of the problem.
-For face-centered quantities, this is one less than the dimensionality, and for edge-centered quantities this is one.
+
+where $r$ is the refinement factor, and the sum is over all fine cells
+that ``cover" the coarse cell. The exponent, $d$, refers to the weight
+of the average.  For cell-centered quantities, this is the
+dimensionality of the problem.  For face-centered quantities, this is
+one less than the dimensionality, and for edge-centered quantities
+this is one.
 
@@
 
-One of the significant advantages of flux-conservative methods is, of course, their ability to maintain conserved quantities to machine precision.
-Unfortunately, the projection step upsets this flux conservation since, at the boundary of a refined region, the coarse and fine cells on either side of the border are updated with different estimates of the flux across the boundary.
-At one refinement level this comes from the coarse grid computation, and in the other from the fine grid solution.
-Conservation can be restored by correcting all coarse cells that lie outside the boundary of a fine region
-with the difference between the coarse and fine estimates of the flux
-across the boundary:
+One of the significant advantages of flux-conservative methods is, of
+course, their ability to maintain conserved quantities to machine
+precision.  Unfortunately, the projection step upsets this flux
+conservation since, at the boundary of a refined region, the coarse
+and fine cells on either side of the border are updated with different
+estimates of the flux across the boundary.  At one refinement level
+this comes from the coarse grid computation, and in the other from the
+fine grid solution.  Conservation can be restored by correcting all
+coarse cells that lie outside the boundary of a fine region with the
+difference between the coarse and fine estimates of the flux across
+the boundary:
+
 \begin{equation}
@@
 \end{equation}
-where $F$ is the flux of quantity $q_i$ in the $i$-direction through the cell interface.
-We use $\tilde{q}$ to indicate the uncorrected quantity on the coarse grid and the sum is over the $r^{d-1}$ fine cell interfaces in the perpendicular dimensions that share a face with the coarse cell being corrected.
-This correction is carried out for all coarse cells that share a face with a refined cell (and are not themselves covered with more refined cells).
+
+where $F$ is the flux of quantity $q_i$ in the $i$-direction through
+the cell interface.   We use $\tilde{q}$ to indicate the uncorrected
+quantity on the coarse grid and the sum is over the $r^{d-1}$ fine
+cell interfaces in the perpendicular dimensions that share a face with
+the coarse cell being corrected.  This correction is carried out for
+all coarse cells that share a face with a refined cell (and are not
+themselves covered with more refined cells).
 
@@
 
-Since the addition of more highly refined grids is adaptive, the conditions for refinement must be specified.
-Many different criteria can be simultaneously specified for a given simulation, and these refinement criteria are discussed in more detail in the following section (Section~\ref{sec:refinement_criteria}).
-Once all cells in a given grid that require additional refinement have been flagged, rectangular boundaries are determined that minimally encompass the flagged cells.
-This is done using an iterative machine-vision technique, as described in \cite{Berger91}.
+Since the addition of more highly refined grids is adaptive, the
+conditions for refinement must be specified. Many different criteria
+can be simultaneously specified for a given simulation, and these
+refinement criteria are discussed in more detail in the following
+section (Section~\ref{sec:refinement_criteria}).   Once all cells in a
+given grid that require additional refinement have been flagged,
+rectangular boundaries are determined that minimally encompass the
+flagged cells.  This is done using an iterative machine-vision
+technique, as described in \cite{Berger91}.
 
-We first find the smallest region (or ``proto-subgrid") that covers all of the refined cells on a grid.
-We compute an efficiency $\epsilon$ for this grid, defined as the ratio of flagged cells to total cells.
-If this efficiency is sufficiently large (typically over 30\%), or if the grid is small enough, then we adopt the region as a new grid and move on to the next region (if any).
-If the region is not acceptable, then we split it into two using the following algorithm.
-First, we determine the longest dimension and compute the signature along that dimension.
-The signature is found by summing the number of refined cells with the same index along the longest dimension.
-Edges are typically indicated by inflection points in this signature, so we calculate the second derivative using a simple three-point finite difference estimator and look for the maximum in the second derivative, with ties going to the index closest to the center.
-If no inflection points are found along the longest axis, the next-longest axis is searched.
-If no inflection points are found, the longest axis is cut in half.
-The original region is then split into two new regions, or proto-subgrids.
-This process is then re-applied to each of the new regions.
+We first find the smallest region (or ``proto-subgrid") that covers
+all of the refined cells on a grid.  We compute an efficiency
+$\epsilon$ for this grid, defined as the ratio of flagged cells to
+total cells.  If this efficiency is sufficiently large (typically over
+30\%), or if the grid is small enough, then we adopt the region as a
+new grid and move on to the next region (if any).  If the region is
+not acceptable, then we split it into two using the following
+algorithm.  First, we determine the longest dimension and compute the
+signature along that dimension.  The signature is found by summing the
+number of refined cells with the same index along the longest
+dimension.  Edges are typically indicated by inflection points in this
+signature, so we calculate the second derivative using a simple
+three-point finite difference estimator and look for the maximum in
+the second derivative, with ties going to the index closest to the
+center.  If no inflection points are found along the longest axis, the
+next-longest axis is searched.  If no inflection points are found, the
+longest axis is cut in half.  The original region is then split into
+two new regions, or proto-subgrids.  This process is then re-applied
+to each of the new regions.
 
-This algorithm is not necessarily optimal, but experience shows it does a reasonable job of generating an efficient covering of the refined cells.
-Note that some cells are unnecessarily refined with the SAMR approach, resulting in wasted resources (unlike in cell-splitting AMR schemes).
-However, it has a number of advantages.
-First, traditional grid-based hydro methods can be simply added with only minor modifications.
-Second, it is easier to accommodate larger stencils (as in PPM, which requires three boundary zones), and third, it can be more efficient for some machines since there are more predictable and linear memory accesses, which more efficiently use a CPU's cache.
+This algorithm is not necessarily optimal, but experience shows it
+does a reasonable job of generating an efficient covering of the
+refined cells.  Note that some cells are unnecessarily refined with
+the SAMR approach, resulting in wasted resources (unlike in
+cell-splitting AMR schemes).  However, it has a number of advantages.
+First, traditional grid-based hydro methods can be simply added with
+only minor modifications.  Second, it is easier to accommodate larger
+stencils (as in PPM, which requires three boundary zones), and third,
+it can be more efficient for some machines since there are more
+predictable and linear memory accesses, which more efficiently use a
+CPU's cache.
 
@@
 Although generalized refinement criteria are available based on
-estimators for the error, such as the Richardson method \citep{AtkinsonHan2004}, we have focused generally on methods that use physical estimators based on the assumption that the simulator has a specific quantity or quantities that indicate the need for refinement.
+estimators for the error, such as the Richardson method
+\citep{AtkinsonHan2004}, we have focused generally on methods that use
+physical estimators based on the assumption that the simulator has a
+specific quantity or quantities that indicate the need for refinement.
 The flagging methods currently built into \enzo\ are:
@@
 
-\item{\em Slope.}  If the normalized slope $(q_{i+1} - q_{i-1})/ (2 q_i)$ of baryon quantities between adjacent cells is larger than a specified value (typically 0.3), then the cell is flagged.
+\item{\em Slope.}  If the normalized slope $(q_{i+1} - q_{i-1})/ (2
+  q_i)$ of baryon quantities between adjacent cells is larger than a
+  specified value (typically 0.3), then the cell is flagged.
 
@@
   in the cell ($M_g = \rho\, (\Delta x)^d$) is larger than a specified value:
+
 \begin{equation}
@@
 \end{equation}
+
 where $\rho_{\rm flag}$ is the equivalent density on the root grid
@@
 
-\item{\em Shocks.}  In this method, we identify and refine cells that contain shocks using the following criteria:
+\item{\em Shocks.}  In this method, we identify and refine cells that
+  contain shocks using the following criteria:
+
 \begin{eqnarray}
-%% (p_{i+1} - p_{i-1})/\min(p_{i+1}, p_{i-1}) & > & 0.33,  \nonumber \\
-%% u_{i-1} - u_{i+1} & > & 0, \quad\text{and} \\
-%% e_i / E_i & > & 0.1.  \nonumber
 (p_{i+1} - p_{i-1})/\min(p_{i+1}, p_{i-1}) > 0.33,  \qquad
@@
 \end{eqnarray}
-We note that the parameters $0.33$ and $0.1$ in the equations above were determined empirically, and can be modified by the user to refine more or less aggressively, as the situation warrants.
 
-\item{\em Particle mass.}  This refinement criterion is precisely the same as the baryon mass criterion except that it uses the gridded (cloud-in-cell, or CIC) particle density.
-In general, this criteria is used in cosmological simulations, and thus refines on dark matter and stellar density.
+We note that the parameters $0.33$ and $0.1$ in the equations above
+were determined empirically, and can be modified by the user to refine
+more or less aggressively, as the situation warrants.
 
-\item{\em Jeans length.}  Following \cite{Truelove98}, the code can be forced to refine the Jeans length by a fixed number of cells, or, more specifically, whenever the following criterion is not met:
+\item{\em Particle mass.}  This refinement criterion is precisely the
+  same as the baryon mass criterion except that it uses the gridded
+  (cloud-in-cell, or CIC) particle density.  In general, this criteria
+  is used in cosmological simulations, and thus refines on dark matter
+  and stellar density.
+
+\item{\em Jeans length.}  Following \cite{Truelove98}, the code can be
+  forced to refine the Jeans length by a fixed number of cells, or,
+  more specifically, whenever the following criterion is not met:
+
 \begin{equation}
@@
 
-\item{\em Cooling time.}  Often in cosmological simulations, cooling is a rapid process and we do not necessarily want to be limited by the cooling time.
-However, in some cases it is desirable to resolve the cooling process, in which case we flag cells that have a cooling time shorter than their sound crossing time.
-The cooling time is given by $t_{\rm cool} = 1.5 k_BT / (n\, \Lambda(T))$ and the sound crossing time is $t_{\rm cross} = \Delta x / c_s$, where $c_s$ is the sound speed.
+\item{\em Cooling time.}  Often in cosmological simulations, cooling
+  is a rapid process and we do not necessarily want to be limited by
+  the cooling time.  However, in some cases it is desirable to resolve
+  the cooling process, in which case we flag cells that have a cooling
+  time shorter than their sound crossing time.  The cooling time is
+  given by $t_{\rm cool} = 1.5 k_BT / (n\, \Lambda(T))$ and the sound
+  crossing time is $t_{\rm cross} = \Delta x / c_s$, where $c_s$ is
+  the sound speed.
 
-\item{\em Must-refine particles.}  In order to allow for refinement in moving regions, it is possible to generate ``must-refine" particles that ensure local refinement of the 8 nearest cells to a specific level.
-The particles feel gravitational forces and their trajectories are followed as for any other collisionless particle.
-This feature can be used to ensure that a specific Lagrangian region stays refined to a specified level of resolution.
+\item{\em Must-refine particles.}  In order to allow for refinement in
+  moving regions, it is possible to generate ``must-refine" particles
+  that ensure local refinement of the 8 nearest cells to a specific
+  level.  The particles feel gravitational forces and their
+  trajectories are followed as for any other collisionless particle.
+  This feature can be used to ensure that a specific Lagrangian region
+  stays refined to a specified level of resolution.
 
-\item{\em Shear.} As described in \citet{Kritsuk06}, turbulence calculations can be successfully modeled with AMR if the shear is used as a refinement parameter; in particular, refinement occurs when the finite difference version of the following inequality holds:
+\item{\em Shear.} As described in \citet{Kritsuk06}, turbulence
+  calculations can be successfully modeled with AMR if the shear is
+  used as a refinement parameter; in particular, refinement occurs
+  when the finite difference version of the following inequality
+  holds:
+
 \begin{equation}
-%\left(\frac{du}{dy}\right)^2 + \left(\frac{du}{dz}\right)^2 + 
-%\left(\frac{dv}{dx}\right)^2 + \left(\frac{dv}{dz}\right)^2 + 
-%\left(\frac{dw}{dx}\right)^2 + \left(\frac{dw}{dy}\right)^2 
 \sum_i \sum_j \left( \frac{\partial v_i}{\partial x_j} \right)^2 -  \sum_i \left( \frac{\partial v_i}{\partial x_i} \right)^2
@@
 
-\item{\em Optical depth.} This refinement criterion, typically used for simulations using radiation transport, 
-flags cells for refinement if they have an \ion{H}{1} optical depth lower than unity.  The optical depth of a cell is estimated to be:
+\item{\em Optical depth.} This refinement criterion, typically used
+for simulations using radiation transport, flags cells for refinement
+if they have an \ion{H}{1} optical depth lower than unity.  The
+optical depth of a cell is estimated to be:
+
 \begin{equation}
@@
 \end{equation}
+
 where $\sigma_{\rm HI}$ is the neutral atomic hydrogen absorption
 cross section, $n_{\rm HI}$ is the proper \ion{H}{1} number density,
-and $\Delta x$ is the cell width (which is assumed to be the same along each axis).  If $\tau > 1$, the cell is refined
-further, up to the maximum level of refinement.
+and $\Delta x$ is the cell width (which is assumed to be the same
+along each axis).  If $\tau > 1$, the cell is refined further, up to
+the maximum level of refinement.
 
-\item{\em Resistive length.}  In a way similar to the Jeans length criterion described above, the code can be forced to locally resolve the resistive length scale by a minimum of a fixed number of cells.
+\item{\em Resistive length.}  In a way similar to the Jeans length
+  criterion described above, the code can be forced to locally resolve
+  the resistive length scale by a minimum of a fixed number of cells.
   This is done by ensuring that the resistive length scale, defined as
+
 \begin{equation}
@@
 \end{equation}
-is always refined by at least N$_R$ cells locally (i.e.~we refine a given cell further if $L_R \leq N_R \Delta x$).
-Note that \enzo\ does not actually solve the resistive MHD equations -- this is the ``characteristic'' local resistive length, rather than one calculated within the MHD solver.
 
-\item{\em Must refine region.}  This refinement criterion is used to ensure that all cells in a given subvolume of the simulation are refined to {\it at least} some minimum level of refinement.
-The subvolume is user-specified, and may evolve over time (e.g., to follow a structure of interest that is traveling through the simulation volume).
+is always refined by at least N$_R$ cells locally (i.e.~we refine a
+given cell further if $L_R \leq N_R \Delta x$).  Note that \enzo\ does
+not actually solve the resistive MHD equations -- this is the
+``characteristic'' local resistive length, rather than one calculated
+within the MHD solver.
 
-\item{\em Metallicity.}  This refinement criterion ensures that all cells with a metallicity (defined as $Z \equiv \rho_{Z} / \rho_{b} / 0.022$, where $\rho_{Z}$ is the density of metals in the cell, $\rho_{b}$ is the total baryon density (including metals), and 0.022 is the solar mass fraction of metals) above a user-defined value is refined to {\it at least} some minimum level of refinement.
+\item{\em Must refine region.}  This refinement criterion is used to
+  ensure that all cells in a given subvolume of the simulation are
+  refined to {\it at least} some minimum level of refinement.  The
+  subvolume is user-specified, and may evolve over time (e.g., to
+  follow a structure of interest that is traveling through the
+  simulation volume).
+
+\item{\em Metallicity.}  This refinement criterion ensures that all
+  cells with a metallicity (defined as $Z \equiv \rho_{Z} / \rho_{b} /
+  0.022$, where $\rho_{Z}$ is the density of metals in the cell,
+  $\rho_{b}$ is the total baryon density (including metals), and 0.022
+  is the solar mass fraction of metals) above a user-defined value is
+  refined to {\it at least} some minimum level of refinement.
 
@@
 
-Finally, it is also possible to specify a rectangular region that is statically refined to a certain level.
-These static refined regions are in place throughout the simulation regardless of the grid quantities and can be used for testing purposes, or for multiply-zoomed initial conditions (in cosmology simulations, for example) that would not otherwise be refined.
-
-%GB: The following section has now been worked into the main text above and, I think, can be removed.
-%
-%\subsection{Adaptive mesh refinement}\label{sec.ov.amr}
-%\dcc{removed most of the notes.  Added intro blurb.  moved parallelism
-%  stuff. Re-worded the bit on ghost zones.  Summary of 'grid as
-%  independent update + syncing with neighbors and parents,' W-cycle. Added
-%  sentence on subgrid creation.  Made timestepping its own subsection.}
-%\red{Talk about reasonable refinement factors - which 
-%ones are usable, etc?}
-
-%The primary purpose of the \enzo\ code is its Adaptive Mesh Refinement
-%capability, which allows it to reach extremely large dynamical ranges
-%with limited computational resources, opening doors previously closed
-%by finite memory and computational time. Unlike moving mesh methods
-%\citep{1995ApJS..100..269P,1995ApJS...97..231G} or  
-%methods that subdivide 
-%individual cells \citep{Adjerid}, Berger \& Collela's AMR (also referred 
-%to as \emph{structured} AMR) utilizes an adaptive hierarchy of grid 
-%patches at varying levels of resolution.  Each rectangular grid patch 
-%(referred to as a ``grid'') covers some region of space in its 
-%\emph{parent grid} which requires higher resolution, and can itself 
-%become the parent grid to an even more highly resolved \emph{child grid}. 
-%ENZO's implementation of structured AMR places no restrictions on 
-%the number of grids at a given level of refinement, or on the number of 
-%levels of refinement.  However, owing to limited computational resources 
-%it is practical to institute a maximum level of refinement $\Lmax$.
-
-%The \enzo\ implementation of AMR allows arbitrary integer ratios of parent and
-%child grid resolution.  However, in practice 
-%refinement factors (defined as the ratio of parent grid resolution to child grid resolution) 
-%of two or four are typically used, with a refinement factor
-%of two being most commonly used for cosmological simulations for reasons of
-%efficiency. The ratio of
-%box size to the maximum grid resolution of a given simulation is
-%therefore $L/e = \Nroot \times 2^{\Lmax}$, where $\Nroot$ is the
-%number of cells along one edge of the root grid, $\Lmax$ is the
-%maximum level of refinement allowed, and L and e are the box size
-%and grid resolution of the most highly refined region, respectively.
-
-%The AMR grid patches are the primary data structure in \enzo.  Each
-%patch is treated as an individual object which can contain both field
-%variables and particle data.  Individual grids are organized into a 
-%dynamic, distributed hierarchy of mesh patches.  
-%See 
-%Figure~\ref{fig.2.amrhierdens} for an example of an \enzo\
-%simulation performed using this AMR scheme where both the baryon density
-%and AMR grid hierarchy are shown together.
-
-%\begin{figure}
-%\begin{center}
-%\includegraphics[width=0.4\textwidth]{figures/amr-hier-dens.eps}
-%\caption{Example of an \enzo\ simulation showing the AMR grid hierarchy.  
-%This is a simulation of the collapse of a single massive halo with a 
-%$32^3$ root grid and two $32^3$ static nested grids.  AMR is only allowed
-%within the most highly refined static nested grid.  Log baryon density is shown
-%by colored isocontours with values corresponding to the legend at center left.  
-%The rectangular solid wire frames correspond to individual 
-%AMR grids, with different colors corresponding to level as indicated by the
-%legend at top left.}
-%\label{fig.2.amrhierdens}
-%\end{center}
-%\end{figure}
-
-%Each grid patch in \enzo\ contains arrays of values for baryon and 
-%particle quantities.   Grids are 
-%partitioned into a core of \emph{real zones} and a surrounding layer 
-%of \emph{ghost zones}.  The real zones store field values and ghost 
-%zones are used to temporarily store values which have been obtained 
-%directly from neighboring grids or interpolated from a parent grid.  
-%These zones are necessary to accommodate the computational stencil of 
-%the hydrodynamics solvers (Sections~\ref{sec.ov.hydro.ppm} and~\ref{sec.ov.hydro.zeus}) 
-%and the gravity solver (Section~\ref{sec.ov.nbody}).  
-%The PPM and \zeus\ hydro solvers require 3 layers of ghost zones,  and the gravity
-%solver requires 6.  This can lead to significant memory 
-%and computational overhead, particularly for smaller grid patches
-%at high levels of refinement.  More information on filling of ghost
-%zones can be found in the appendix.
-
-%From the point of view of the hydro solver, each grid is solved as an
-%independent CFD problem, with Dirichlet boundary conditions stored in
-%the ghost zones.  At the end of each timestep, cells on a given level
-%fill ghost zones by interpolating from the parent and copying from
-%neighbors on the same level.  Grids then correct the flux at the interface boundary between
-%fine and coarse zones, and finally project their real zone data to the
-%parent grid.
-
-%Timesteps are determined on each level as
-%described in \ref{sed.timestepping}, and the hierarchy is advanced on
-%a level by level basis in what's known as a W-Cycle.    Beginning with
-%the coarsest level, $L$, all grids on that level are advanced one
-%timestep.  Then one timestep is taken on all grids at the next level
-%of refinement, $L+1$, and so on until the finest level is resolved.
-%The finest level is then advanced, using as many steps as it takes to
-%reach the level immediately above.  The finest level is then synced
-%to the level above it, which then proceeds forewards one more step.
-%\red{Needs a picture of the W cycle.}
-
-%At the end of every timestep on every level each grid updates 
-%its ghost zones by exchanging information with its neighboring grid patches 
-%(if any exist) and/or by interpolating from a parent grid.  
-%In addition, cells are examined to see if they should be refined or 
-%de-refined, and the entire grid hierarchy is rebuilt at that 
-%level (including all more highly refined levels).  The timestepping and 
-%hierarchy advancement/rebuilding process described here is repeated 
-%recursively on every level to the specified maximum level of 
-%refinement in the simulation.
-
-%Since the addition of more highly refined grids is adaptive,  
-%the conditions for refinement must be specified.  The criteria of 
-%refinement can be set by the threshold value of the overdensity of 
-%baryon gas or dark matter in a cell (which is really a refinement on
-%the mass of gas or DM in a cell), the local Jeans length, the 
-%local density gradient, or local pressure and energy gradients.  
-%A cell reaching any or all of these criteria 
-%will then be flagged for refinement.  
-%Once all cells at a given level have been examined, rectangular boundaries 
-%are determined which minimally encompass the flagged cells. The set of
-%flagging cells is then further fractured into smaller bounding
-%rectangles to ensure computational 
-%efficiency.  More details on subgrid creation can be found in Appendix
-%\ref{app.subgrid_creation}.  A refined grid patch is 
-%introduced within each such bounding rectangle. Thus the cells needing 
-%refinement, as well as adjacent cells within the patch which do not need 
-%refinement, are refined. While this approach is not as memory efficient as 
-%cell-splitting AMR schemes, it offers more freedom with 
-%finite difference stencils. For example, PPM requires a stencil of seven 
-%cells per dimension. This cannot easily be accommodated in cell-splitting 
-%AMR schemes.
-
-%%% Local Variables: 
-%%% mode: latex
-%%% TeX-master: "ms"
-%%% End:
+Finally, it is also possible to specify a rectangular region that is
+statically refined to a certain level.  These static refined regions
+are in place throughout the simulation regardless of the grid
+quantities and can be used for testing purposes, or for
+multiply-zoomed initial conditions (in cosmology simulations, for
+example) that would not otherwise be refined.

### 2013-06-25 numerical-analysis.tex
Many small changes to all files -- GB final edits
Removed commented-out text from lots of files - basically vestigial junk.
Reformatting tex files so they're all standardized in terms of paragraph wrapping, etc. (no content changed, just formatting.)

@@
-% EJT: general changes
-% Enzo to \enzo\
-% changed ``some Lagrangian fluid trajectories'' to ``a fixed number of Lagrangian fluid trajectories.'' Easier to understand.
-% changed ``some evolution'' to ``a period of evolution'' ... I basically don't like 'some'. It sounds vague.
-% rephrased to remove ``e.g.'' as they read slightly awkwardly
-% moved definition of T_1 and T_2 slightly higher in shock text to where they first appear.
-
 \section{Analysis}
@@
 Detailed analysis of simulation results requires both the tools to ask
-sophisticated questions of the data and the ability to process vast quantities of
-data at high time-cadence.  As simulations grow in size and complexity, storing data
-for post-processing simply becomes intractable.  To cope with this, we have
-instrumented \enzo\ with the ability to conduct analysis during the course of a
-simulation.  This enables analysis with extremely high time cadence (as often
-as every subcycle of the finest refinement level), without
-attempting to write an entire checkpoint output to disk.  The current mechanism
-for conducting analysis in \enzo\ during the course of the simulation utilizes
-the same computional resources as are used by the simulation itself by transferring their usage 
-from \enzo\ to the analysis routines; this is often
-referred to as \textit{in situ} analysis or visualization.  Utilizing a
-dynamically-scheduled second set of computing resources, often referred to as
-\textit{co-scheduled} analysis or visualization, provides greater flexibility and overall
+sophisticated questions of the data and the ability to process vast
+quantities of data at high time-cadence.  As simulations grow in size
+and complexity, storing data for post-processing simply becomes
+intractable.  To cope with this, we have instrumented \enzo\ with the
+ability to conduct analysis during the course of a simulation.  This
+enables analysis with extremely high time cadence (as often as every
+subcycle of the finest refinement level), without attempting to write
+an entire checkpoint output to disk.  The current mechanism for
+conducting analysis in \enzo\ during the course of the simulation
+utilizes the same computional resources as are used by the simulation
+itself by transferring their usage from \enzo\ to the analysis
+routines; this is often referred to as \textit{in situ} analysis or
+visualization.  Utilizing a dynamically-scheduled second set of
+computing resources, often referred to as \textit{co-scheduled}
+analysis or visualization, provides greater flexibility and overall
 throughput at the expense of simplicity.
 
-We expose \enzo's mesh geometry and fluid quantities to the analysis platform
-\texttt{yt} \citep{2011ApJS..192....9T, 2011arXiv1112.4482T}.  At compile time,
-\enzo\ is (dynamically or statically) linked against the Python and NumPy
-libraries necessary to create proxy objects exposing the mesh geometry,
-fluid quantities and particle arrays as NumPy arrays.  This information is then
-passed to a special handler inside \texttt{yt}.  \texttt{yt} interprets the
-mesh and fluid information and, without saving data to disk, constructs a native representation of the in-memory state
-of the simulation that appears identical to an on-disk simulation output.
-\enzo\ then executes a user-provided analysis script, which is able to access the
-in-memory simulation object.  Once the analysis script has returned control to
-\enzo, the simulation proceeds.  This process can occur at either the top of the
-main ``EvolveHierarchy'' loop or at the end of a timestep at the finest level,
-and the frequency with which it is called is adjustable by a run-time
-parameter.  During the course of conducting analysis, the simulation is halted
-until the conclusion of the analysis.
+We expose \enzo's mesh geometry and fluid quantities to the analysis
+platform \texttt{yt} \citep{2011ApJS..192....9T, 2011arXiv1112.4482T}.
+At compile time, \enzo\ is (dynamically or statically) linked against
+the Python and NumPy libraries necessary to create proxy objects
+exposing the mesh geometry, fluid quantities and particle arrays as
+NumPy arrays.  This information is then passed to a special handler
+inside \texttt{yt}.  \texttt{yt} interprets the mesh and fluid
+information and, without saving data to disk, constructs a native
+representation of the in-memory state of the simulation that appears
+identical to an on-disk simulation output.  \enzo\ then executes a
+user-provided analysis script, which is able to access the in-memory
+simulation object.  Once the analysis script has returned control to
+\enzo, the simulation proceeds.  This process can occur at either the
+top of the main ``EvolveHierarchy'' loop or at the end of a timestep
+at the finest level, and the frequency with which it is called is
+adjustable by a run-time parameter.  During the course of conducting
+analysis, the simulation is halted until the conclusion of the
+analysis.
 
-Most analysis operations that can be performed on data sets that reside on
-disk can be performed on in-memory data sets in \texttt{yt}.  This includes
-projections (i.e., line integrals, both on- and off-axis), slices, 1-, 2-, and
-3-D fluid phase distributions, calculation of derived quantities and arbitrary
-data selection.  As of version 2.5 of \texttt{yt},  the Rockstar phase-space
-halo finder \citep{2013ApJ...762..109B} can be executed through \texttt{yt} on
+Most analysis operations that can be performed on data sets that
+reside on disk can be performed on in-memory data sets in \texttt{yt}.
+This includes projections (i.e., line integrals, both on- and
+off-axis), slices, 1-, 2-, and 3-D fluid phase distributions,
+calculation of derived quantities and arbitrary data selection.  As of
+version 2.5 of \texttt{yt}, the Rockstar phase-space halo finder
+\citep{2013ApJ...762..109B} can be executed through \texttt{yt} on
 in-memory \enzo\ data, and so can the Parallel HOP halo finder
-\citep{1998ApJ...498..137E, 2010ApJS..191...43S}.  Operations that currently cannot be conducted on
-in-memory datasets are those that require spatial decomposition of
-data.  For 
-instance, calculating marching cubes on a data object with \texttt{yt} is a
-fully local operation and can be conducted \textit{in situ}.  However,
-calculating topologically-connected sets requires a spatial decomposition of
-data and thus cannot be conducted \textit{in situ}.  This prohibition extends
-to halo finding operations other than Rockstar, most multi-level
-parallelism operations, and volume rendering.
+\citep{1998ApJ...498..137E, 2010ApJS..191...43S}.  Operations that
+currently cannot be conducted on in-memory datasets are those that
+require spatial decomposition of data.  For instance, calculating
+marching cubes on a data object with \texttt{yt} is a fully local
+operation and can be conducted \textit{in situ}.  However, calculating
+topologically-connected sets requires a spatial decomposition of data
+and thus cannot be conducted \textit{in situ}.  This prohibition
+extends to halo finding operations other than Rockstar, most
+multi-level parallelism operations, and volume rendering.
 
-Where microphysical solvers or other operator-split physics calculations can be
-done in Python, \texttt{yt} can serve as a driver for these calculations.  A
-major feature set that is currently being developed is to pass structured
-(i.e., non-fluid) information back from \texttt{yt} into \enzo.  For instance,
-this could be the result of semi-analytic models of the growth and evolution of
-star clusters, galaxy particle feedback parameters that have been influenced by
-merger-tree analysis, or even spectral energy distributions that are calculated
-within \texttt{yt} and provided to \enzo\ as input into radiative transfer
-calculations.  Future versions will include this, as well as the ability to
-dynamically allocate computional resources to \texttt{yt} such that the simulation
-may proceed asynchronously with analysis (\textit{co-scheduled} analysis).
-With this functionality will also come the ability to dynamically partition
-data, such that spatially-decomposed operations such as volume rendering become
-feasible during the course of a simulation.
+Where microphysical solvers or other operator-split physics
+calculations can be done in Python, \texttt{yt} can serve as a driver
+for these calculations.  A major feature set that is currently being
+developed is to pass structured (i.e., non-fluid) information back
+from \texttt{yt} into \enzo.  For instance, this could be the result
+of semi-analytic models of the growth and evolution of star clusters,
+galaxy particle feedback parameters that have been influenced by
+merger-tree analysis, or even spectral energy distributions that are
+calculated within \texttt{yt} and provided to \enzo\ as input into
+radiative transfer calculations.  Future versions will include this,
+as well as the ability to dynamically allocate computional resources
+to \texttt{yt} such that the simulation may proceed asynchronously
+with analysis (\textit{co-scheduled} analysis).  With this
+functionality will also come the ability to dynamically partition
+data, such that spatially-decomposed operations such as volume
+rendering become feasible during the course of a simulation.
 
@@
 inability to follow the evolution of a single parcel of fluid as it
-travels through the simulation volume.  To address this, \enzo\ has the
-capability to introduce Lagrangian ``tracer particles'' into a
+travels through the simulation volume.  To address this, \enzo\ has
+the capability to introduce Lagrangian ``tracer particles'' into a
 calculation either at the beginning of the simulation or when
@@
 see Equation~\ref{eqn.driftkick}), the particle's velocity is updated
-both at the beginning of each timestep and at the half-timestep
-by linearly interpolating the cell-centered baryon velocity to the
-position of the particle ($v^{\rm interp}$), and assigning it to that value.
+both at the beginning of each timestep and at the half-timestep by
+linearly interpolating the cell-centered baryon velocity to the
+position of the particle ($v^{\rm interp}$), and assigning it to that
+value.
 
 \enzo\ saves tracer particle data at user-specified intervals,
-independent of the intervals at which regular data sets are written.  The data written out
-typically includes the tracer particle's unique ID and position, as
-well as the velocity, density,
-and temperature of the gas at its location, but is easily extensible to output any grid-based
-quantity that a user requires.  This capability has been used quite
-effectively in several papers, including \citet{2010ApJ...715.1575S}
-and \citet{2012ApJ...748...12S}. It is crucial to keep in mind that
-tracer particles model a fixed number of Lagrangian fluid trajectories. The fluid
-on the grid, however, models the motion of all the mass and represents
-the average quantity in a grid cell's volume. Consequently, after a period of
-evolution, tracer particles -- even if they initially had the same
-density distribution as the gas -- will not have the same density
-distribution as the fluid. For example, they tend to accumulate at stagnation points
-of the flow, and care has to be taken in using these particles
-appropriately. Tracer particles are very useful in studies such as the variety of
-histories of the hydrodynamic quantities in Lagrangian fluid elements
-and when evaluating complex chemical and cooling models in regard to the
-simpler ones used in the actual numerical evolution.
+independent of the intervals at which regular data sets are written.
+The data written out typically includes the tracer particle's unique
+ID and position, as well as the velocity, density, and temperature of
+the gas at its location, but is easily extensible to output any
+grid-based quantity that a user requires.  This capability has been
+used quite effectively in several papers, including
+\citet{2010ApJ...715.1575S} and \citet{2012ApJ...748...12S}. It is
+crucial to keep in mind that tracer particles model a fixed number of
+Lagrangian fluid trajectories. The fluid on the grid, however, models
+the motion of all the mass and represents the average quantity in a
+grid cell's volume. Consequently, after a period of evolution, tracer
+particles -- even if they initially had the same density distribution
+as the gas -- will not have the same density distribution as the
+fluid. For example, they tend to accumulate at stagnation points of
+the flow, and care has to be taken in using these particles
+appropriately. Tracer particles are very useful in studies such as the
+variety of histories of the hydrodynamic quantities in Lagrangian
+fluid elements and when evaluating complex chemical and cooling models
+in regard to the simpler ones used in the actual numerical evolution.
 
@@
 
-Identification of shocks and their pre- and post-shock conditions can be
-accomplished through a combination of either temperature or velocity jumps with
-dimensionally split or unsplit search methods.  The primary method used in
-\enzo~is the dimensionally unsplit temperature jump method described in detail
-in \citet{2008ApJ...689.1063S}.  We briefly outline the method here.
+Identification of shocks and their pre- and post-shock conditions can
+be accomplished through a combination of either temperature or
+velocity jumps with dimensionally split or unsplit search methods.
+The primary method used in \enzo~is the dimensionally unsplit
+temperature jump method described in detail in
+\citet{2008ApJ...689.1063S}.  We briefly outline the method here.
 
-For every cell, we first determine whether it satisfies the following conditions
-necessary to be flagged as a shock:
+For every cell, we first determine whether it satisfies the following
+conditions necessary to be flagged as a shock:
 
@@
 
-where $\myvec{v}$ is the velocity field, $T$ is the temperature, $\rho$ is the
-density, and $S=T/\rho^{\gamma-1}$ is the entropy.  $T_2$ and $T_1$ are the post-shock~(downstream)
-and pre-shock~(upstream) temperatures, respectively. An optional temperature floor
-may be chosen, which is useful for situations such as cosmological simulations without a radiation
-background where underdense gas in the intergalactic medium (IGM) can cool adiabatically to
-unphysically low temperatures.
+where $\myvec{v}$ is the velocity field, $T$ is the temperature,
+$\rho$ is the density, and $S=T/\rho^{\gamma-1}$ is the entropy.
+$T_2$ and $T_1$ are the post-shock~(downstream) and
+pre-shock~(upstream) temperatures, respectively. An optional
+temperature floor may be chosen, which is useful for situations such
+as cosmological simulations without a radiation background where
+underdense gas in the intergalactic medium (IGM) can cool
+adiabatically to unphysically low temperatures.
 
-Once a cell is flagged, the local temperature gradient is calculated, which is 
-then used to traverse cells parallel to the gradient to search for the first
-pre- and post-shock cell that do not satisfy the above conditions.  If during
-the search a cell satisfying the conditions is found to have a more convergent
-flow, that cell is marked as the center, and the search is started again. Using the 
-temperature values from each of these cells, the Mach number is then solved
-using the Rankine-Hugoniot temperature jump conditions:
+Once a cell is flagged, the local temperature gradient is calculated,
+which is then used to traverse cells parallel to the gradient to
+search for the first pre- and post-shock cell that do not satisfy the
+above conditions.  If during the search a cell satisfying the
+conditions is found to have a more convergent flow, that cell is
+marked as the center, and the search is started again. Using the
+temperature values from each of these cells, the Mach number is then
+solved using the Rankine-Hugoniot temperature jump conditions:
+
 \begin{equation}
@@
 \end{equation}
+
 where $\Mach$ is the upstream Mach number.
 
-Shock finding in the context of AMR is applied grid-by-grid.  If a search for
-pre/post-shock cells goes outside the bounds of the grid ghost zones, the
-search is stopped for that particular shocked cell. In most situations this is
-adequate since the hydrodynamic shock is captured in fewer than the number of
-ghost zones.  Shock finding can be run either upon data output or at each step
-in the evolution of an AMR level if the Mach and pre/post-shock quantities are
-needed for additional physics modules.
+Shock finding in the context of AMR is applied grid-by-grid.  If a
+search for pre/post-shock cells goes outside the bounds of the grid
+ghost zones, the search is stopped for that particular shocked
+cell. In most situations this is adequate since the hydrodynamic shock
+is captured in fewer than the number of ghost zones.  Shock finding
+can be run either upon data output or at each step in the evolution of
+an AMR level if the Mach and pre/post-shock quantities are needed for
+additional physics modules.

### 2013-06-25 numerical-chemistry.tex
Removed commented-out text from lots of files - basically vestigial junk.
Reformatting tex files so they're all standardized in terms of paragraph wrapping, etc. (no content changed, just formatting.)

@@
 
-While it is often safe to assume that species (both chemical and ionization)
-within a gas can be treated as being in equilibrium, in some regimes that are
-found in astrophysics this assumption leads to considerable error.  For
-example, the cooling and collapse of primordial gas in Population III star
-formation is dominated by molecular hydrogen, which in the absence of dust
-forms via an inefficient pair of collisional processes that depend heavily on
-the local, highly non-equilibrium population of free electrons.  As a result,
-when modeling primordial star formation it is critical to follow the
-non-equilibrium evolution of the chemical species of hydrogen, including
-molecular hydrogen and deuterium.
+While it is often safe to assume that species (both chemical and
+ionization) within a gas can be treated as being in equilibrium, in
+some regimes that are found in astrophysics this assumption leads to
+considerable error.  For example, the cooling and collapse of
+primordial gas in Population III star formation is dominated by
+molecular hydrogen, which in the absence of dust forms via an
+inefficient pair of collisional processes that depend heavily on the
+local, highly non-equilibrium population of free electrons.  As a
+result, when modeling primordial star formation it is critical to
+follow the non-equilibrium evolution of the chemical species of
+hydrogen, including molecular hydrogen and deuterium.
 
-The primordial non-equilibrium chemistry routines used in \enzo\ were first
-described by \citet{abel97} and \citet{anninos97}, but have
+The primordial non-equilibrium chemistry routines used in \enzo\ were
+first described by \citet{abel97} and \citet{anninos97}, but have
 since been extended with updated reaction rates and the inclusion of
-deuterium species \citep{McGreer2008, 2009PhDT.........5T}.  These routines follow the
-non-equilibrium chemistry of a gas of primordial composition with 12 total
-species: H, \Hp, He, \Hep, \Hepp, \Hm, \HHp, \HH, e$^-$, D, \Dp, and HD.  \enzo\
-also computes the radiative heating and cooling of the gas from atomic and
-molecular line excitation, recombination, collisional excitation, free-free
-transitions, Compton scattering of the cosmic microwave background, as well as
-several models for a metagalactic UV background that heat the gas via
-photoionization and photodissociation (see Section~\ref{sec.num.cooling} for
-more details).  The chemical and thermal states of the gas can be updated either
-at the same hydrodynamical timestep (i.e., decoupled and operator-split) or
-through the same subcycling system (i.e., a coupled chemical and thermal
-system).  The default behavior of \enzo\ is to couple these two systems at
-subcycles of the hydrodynamic timestep; this results in updates to both the
-chemical and thermal states of the gas (which also inform the temperature, the
-reaction rate coefficients and the cooling functions of the gas) on timescales
-that are faster than those of the gas dynamics.  The gamma used by \enzo\ to
-compute the temperature of the gas from the energy and density characteristics
-utilizes a variable gamma that includes effects of the rotational state of
-molecular hydrogen, enabling it to vary from $5/3$ (fully-atomic) to $7/5$
-(fully molecular).  This further coupling of the chemical and thermal states of
-the gas underscores the need for coupled chemistry and radiative cooling
-solutions.
+deuterium species \citep{McGreer2008, 2009PhDT.........5T}.  These
+routines follow the non-equilibrium chemistry of a gas of primordial
+composition with 12 total species: H, \Hp, He, \Hep, \Hepp, \Hm, \HHp,
+\HH, e$^-$, D, \Dp, and HD.  \enzo\ also computes the radiative
+heating and cooling of the gas from atomic and molecular line
+excitation, recombination, collisional excitation, free-free
+transitions, Compton scattering of the cosmic microwave background, as
+well as several models for a metagalactic UV background that heat the
+gas via photoionization and photodissociation (see
+Section~\ref{sec.num.cooling} for more details).  The chemical and
+thermal states of the gas can be updated either at the same
+hydrodynamical timestep (i.e., decoupled and operator-split) or
+through the same subcycling system (i.e., a coupled chemical and
+thermal system).  The default behavior of \enzo\ is to couple these
+two systems at subcycles of the hydrodynamic timestep; this results in
+updates to both the chemical and thermal states of the gas (which also
+inform the temperature, the reaction rate coefficients and the cooling
+functions of the gas) on timescales that are faster than those of the
+gas dynamics.  The gamma used by \enzo\ to compute the temperature of
+the gas from the energy and density characteristics utilizes a
+variable gamma that includes effects of the rotational state of
+molecular hydrogen, enabling it to vary from $5/3$ (fully-atomic) to
+$7/5$ (fully molecular).  This further coupling of the chemical and
+thermal states of the gas underscores the need for coupled chemistry
+and radiative cooling solutions.
 
-Input parameters to \enzo\ govern the chemical species that are updated during
-the course of the simulation.  This can include only the atomic species (H,
-\Hp, He, \Hep, \Hepp, and e$^-$), those species relevant for molecular hydrogen
-formation (\HH, \HHp, and \Hm), and can further include deuterium and its
-species (D, \Dp, and HD).  A total of 9 kinetic equations are solved for the
-12 species mentioned above, including
-29 kinetic and radiative processes.  See
+Input parameters to \enzo\ govern the chemical species that are
+updated during the course of the simulation.  This can include only
+the atomic species (H, \Hp, He, \Hep, \Hepp, and e$^-$), those species
+relevant for molecular hydrogen formation (\HH, \HHp, and \Hm), and
+can further include deuterium and its species (D, \Dp, and HD).  A
+total of 9 kinetic equations are solved for the 12 species mentioned
+above, including 29 kinetic and radiative processes.  See
 Table~\ref{table.collisional} for the collisional processes and
@@
 
-The chemical reaction equation network is technically challenging to solve due
-to the huge range of reaction timescales involved.  The characteristic times
-for creation and destruction of the various species and reactions can differ by
-many orders of magnitude and are often very sensitive to the chemical and
-thermal state of the gas.  This makes a fully-implicit scheme, with convergence
-criteria and error tolerance, strongly preferable for such a set of equations.
-However, most implicit schemes require an iterative procedure to converge, and
-for large networks (such as this one) an iterative, fully-implicit method can
-be very time-consuming and computationally costly for a relatively small
-increase in accuracy.  At the present time, this makes fully-implicit methods
-somewhat undesirable for large, three-dimensional simulations.
+The chemical reaction equation network is technically challenging to
+solve due to the huge range of reaction timescales involved.  The
+characteristic times for creation and destruction of the various
+species and reactions can differ by many orders of magnitude and are
+often very sensitive to the chemical and thermal state of the gas.
+This makes a fully-implicit scheme, with convergence criteria and
+error tolerance, strongly preferable for such a set of equations.
+However, most implicit schemes require an iterative procedure to
+converge, and for large networks (such as this one) an iterative,
+fully-implicit method can be very time-consuming and computationally
+costly for a relatively small increase in accuracy.  At the present
+time, this makes fully-implicit methods somewhat undesirable for
+large, three-dimensional simulations.
 
-\enzo\ solves the rate equations using a method based on a semi-implicit
-formulation in order to provide a stable, positive definite and first-order
-accurate solution.  The update discretization splits chemical changes into
-formation components and destruction components and updates with a mixed set of
-time states, as described in \citet{anninos97}.  The formation components of
-species $S_i$ are computed at the current subcycle time, where the contribution
-of species $S_i$ to its own destruction components are computed at the updated
-time; all other contributions to the destruction component are computed at the
-current time.  This mixed state improves accuracy, ensures species values
-are positive definite, and is equivalent to one Jacobi iteration of an implicit
-Euler solve.  This technique is optimized by taking the chemical intermediaries
-\Hm and \HHp, which have large rate coefficients and low concentrations, and
-grouping them into a separate category of equations.  Due to their fast
-reactions, these species are very sensitive to small changes in the more
-abundant species and are (at almost all times in astrophysical calculations) close to
-their equilibrium values.  Attempting to resolve their formation and destruction
-times would necessitate extremely small timesteps.
-%and while these species
-%provide necessary channels to facilitate the formation of molecular hydrogen at
-%low densities, they do not otherwise significantly influence the abundance of
-%species with higher concentrations.  
-Therefore, reactions governing these two
-species can be decoupled from the rest of the network and treated independently
-through analytic solutions for equilibrium values.  This allows a significant speedup
-in the solution speed, as the timestepping scheme is applied only to the slower 7- or
-10-species network (depending on whether deuterium is included or not), which
-will be much closer to the overall hydrodynamic timestep of the simulation.
+\enzo\ solves the rate equations using a method based on a
+semi-implicit formulation in order to provide a stable, positive
+definite and first-order accurate solution.  The update discretization
+splits chemical changes into formation components and destruction
+components and updates with a mixed set of time states, as described
+in \citet{anninos97}.  The formation components of species $S_i$ are
+computed at the current subcycle time, where the contribution of
+species $S_i$ to its own destruction components are computed at the
+updated time; all other contributions to the destruction component are
+computed at the current time.  This mixed state improves accuracy,
+ensures species values are positive definite, and is equivalent to one
+Jacobi iteration of an implicit Euler solve.  This technique is
+optimized by taking the chemical intermediaries \Hm and \HHp, which
+have large rate coefficients and low concentrations, and grouping them
+into a separate category of equations.  Due to their fast reactions,
+these species are very sensitive to small changes in the more abundant
+species and are (at almost all times in astrophysical calculations)
+close to their equilibrium values.  Attempting to resolve their
+formation and destruction times would necessitate extremely small
+timesteps.  Therefore, reactions governing these two species can be
+decoupled from the rest of the network and treated independently
+through analytic solutions for equilibrium values.  This allows a
+significant speedup in the solution speed, as the timestepping scheme
+is applied only to the slower 7- or 10-species network (depending on
+whether deuterium is included or not), which will be much closer to
+the overall hydrodynamic timestep of the simulation.
 
-Even so, the accuracy and stability of the scheme is maintained by subcycling
-the rate solver within a single hydrodynamic timestep.  These subcycle
-timesteps are determined so that the estimated fractional change in the
-electron concentration is limited to no more than $10\%$ per timestep;
-additional criteria may be applied based on the expected change in
-internal energy from radiative
-cooling and from  chemical heating due to the formation of molecular hydrogen.
+Even so, the accuracy and stability of the scheme is maintained by
+subcycling the rate solver within a single hydrodynamic timestep.
+These subcycle timesteps are determined so that the estimated
+fractional change in the electron concentration is limited to no more
+than $10\%$ per timestep; additional criteria may be applied based on
+the expected change in internal energy from radiative cooling and from
+chemical heating due to the formation of molecular hydrogen.
 
 It is important to note the regime in which this model is valid.
-According to \citet{abel97} and \citet{anninos97},
-the reaction network is valid for temperatures between $10^0 - 10^8$
-K.  The original model discussed in these two references was only claimed to be valid
-up to $n_H \sim 10^4$~cm$^{-3}$.  However, addition of the 3-body
-\HH~formation process (equation 20 in Table~\ref{table.collisional})
-allows correct modeling of the chemistry of the gas up until the point
-where collisionally-induced emission from molecular hydrogen becomes
-an important cooling process, which occurs at $n_{\rm H} \sim
-10^{14}$~cm$^{-3}$.  A further concern is that the optically thin
-approximation for radiative cooling eventually breaks down, which occurs before
-$n_{\rm H} \sim 10^{16} - 10^{17}$~cm$^{-3}$ in gas of primordial composition.  Beyond this point,
-modifications to the cooling function that take into account the
-non-negligible opacity in the gas must be made, as discussed by
+According to \citet{abel97} and \citet{anninos97}, the reaction
+network is valid for temperatures between $10^0 - 10^8$ K.  The
+original model discussed in these two references was only claimed to
+be valid up to $n_H \sim 10^4$~cm$^{-3}$.  However, addition of the
+3-body \HH~formation process (equation 20 in
+Table~\ref{table.collisional}) allows correct modeling of the
+chemistry of the gas up until the point where collisionally-induced
+emission from molecular hydrogen becomes an important cooling process,
+which occurs at $n_{\rm H} \sim 10^{14}$~cm$^{-3}$.  A further concern
+is that the optically thin approximation for radiative cooling
+eventually breaks down, which occurs before $n_{\rm H} \sim 10^{16} -
+10^{17}$~cm$^{-3}$ in gas of primordial composition.  Beyond this
+point, modifications to the cooling function that take into account
+the non-negligible opacity in the gas must be made, as discussed by
 \citet{2004MNRAS.348.1019R}, and was put into \enzo\ for the work
-published in \citep{2009Sci...325..601T,2009PhDT.........5T}.   The formation
-of molecular hydrogen as catalyzed by dust was recently added to \enzo\ to
-enable studies of low-metallicity gas, as well as the inclusion of appropriate
-timestepping criteria to account for the input of ionizing radiation.  Even
-with these modifications, a completely correct description of the cooling of
-primordial gas at very high densities requires some form of radiation
-transport, which will greatly increase the cost of the simulations.
-Furthermore, at very high densities, the stiffness of the molecular hydrogen
-reaction rates may require better than a first-order accurate solution; as
-such, the transition to this regime will likely necessitate a fully-implicit,
-iterative solver.
+published in \citep{2009Sci...325..601T,2009PhDT.........5T}.  The
+formation of molecular hydrogen as catalyzed by dust was recently
+added to \enzo\ to enable studies of low-metallicity gas, as well as
+the inclusion of appropriate timestepping criteria to account for the
+input of ionizing radiation.  Even with these modifications, a
+completely correct description of the cooling of primordial gas at
+very high densities requires some form of radiation transport, which
+will greatly increase the cost of the simulations.  Furthermore, at
+very high densities, the stiffness of the molecular hydrogen reaction
+rates may require better than a first-order accurate solution; as
+such, the transition to this regime will likely necessitate a
+fully-implicit, iterative solver.
 
@@
 \end{table}
-
-%%% Local Variables: 
-%%% mode: latex
-%%% TeX-master: "ms"
-%%% End:

### 2013-06-25 numerical-conduction.tex
Reformatting tex files so they're all standardized in terms of paragraph wrapping, etc. (no content changed, just formatting.)

@@
 
-\enzo\ implements the equations of isotropic heat conduction in a manner
-similar to that of \citet{2007ApJ...664..135P}.  
-The isotropic flux of heat is given by equation~(\ref{eq:conduction}) and we
-use a value for the Spitzer conduction coefficient, 
-$\kappa_{\rm sp} = 4.6 \times 10^{-7}$~T$^{5/2}$ erg
-s$^{-1}$~cm$^{-1}$~K$^{-1}$ \citep{1962pfig.book.....S}.  In this
-situation we are using a
-value for the Coulomb logarithm, $\log \Lambda = 37.8$, that is
-appropriate for the intracluster medium \citep{1988xrec.book.....S} --
-in astrophysically-relevant, fully ionized plasmas this value varies
-by no more than 50\% (see, e.g., Smith et al. 2013, submitted).
-It is quite possible that the local heat flux computed in this way 
-can become unphysically large in
-the high-temperature, low-density cluster regime when using this
-formulation; therefore, we take into account the saturation of the heat
-flux \citep{1977ApJ...211..135C} at a maximum level of
+\enzo\ implements the equations of isotropic heat conduction in a
+manner similar to that of \citet{2007ApJ...664..135P}.  The isotropic
+flux of heat is given by equation~(\ref{eq:conduction}) and we use a
+value for the Spitzer conduction coefficient, $\kappa_{\rm sp} = 4.6
+\times 10^{-7}$~T$^{5/2}$ erg s$^{-1}$~cm$^{-1}$~K$^{-1}$
+\citep{1962pfig.book.....S}.  In this situation we are using a value
+for the Coulomb logarithm, $\log \Lambda = 37.8$, that is appropriate
+for the intracluster medium \citep{1988xrec.book.....S} -- in
+astrophysically-relevant, fully ionized plasmas this value varies by
+no more than 50\% (see, e.g., Smith et al. 2013, submitted).  It is
+quite possible that the local heat flux computed in this way can
+become unphysically large in the high-temperature, low-density cluster
+regime when using this formulation; therefore, we take into account
+the saturation of the heat flux \citep{1977ApJ...211..135C} at a
+maximum level of
+
 \begin{equation}
@@
 \end{equation}
-To ensure a smooth transition between
-the Spitzer and saturated regimes, we define an effective
-conductivity using the formalism of \citet{1988xrec.book.....S}
+
+To ensure a smooth transition between the Spitzer and saturated
+regimes, we define an effective conductivity using the formalism of
+\citet{1988xrec.book.....S}
+
 \begin{equation}
@@
 \end{equation}
+
 where $\lambda_e$ is the electron mean free path and $\ell_T \equiv T
@@
 plasma can be described in terms of an effective conductivity, which
-can be expressed as a fraction f$_{\rm sp}$ of the Spitzer conductivity
-(where f$_{\rm sp} \leq 1.0$ are considered physically realistic values).  This takes
-into account physical processes below the resolution limit of the
-simulation, such as tangled magnetic fields, that can suppress heat
-transport.
+can be expressed as a fraction f$_{\rm sp}$ of the Spitzer
+conductivity (where f$_{\rm sp} \leq 1.0$ are considered physically
+realistic values).  This takes into account physical processes below
+the resolution limit of the simulation, such as tangled magnetic
+fields, that can suppress heat transport.
 
 Thermal conduction in a plasma can be strongly affected by the
-presence of magnetic field lines, which may suppress heat
-flow perpendicular to the magnetic field.  In that case, we allow for
-heat transport only parallel to the magnetic field lines in
-magnetohydrodynamic simulations.  Mathematically, this is given
-by equation~(\ref{eq:anisotropic_conduction}.
-As with the isotropic thermal conduction, we allow a
-multiplicative factor f$_{\rm sp}$ to take into account the possible
-suppression of magnetic fields below the resolution limit of the
-simulation.
+presence of magnetic field lines, which may suppress heat flow
+perpendicular to the magnetic field.  In that case, we allow for heat
+transport only parallel to the magnetic field lines in
+magnetohydrodynamic simulations.  Mathematically, this is given by
+equation~(\ref{eq:anisotropic_conduction}.  As with the isotropic
+thermal conduction, we allow a multiplicative factor f$_{\rm sp}$ to
+take into account the possible suppression of magnetic fields below
+the resolution limit of the simulation.
 
-Both isotropic and anisotropic thermal conduction in \enzo\ are treated
-in an operator-split manner.  Furthermore, within the heat transport
-module, transport along the x, y, and z directions are computed in a
-directionally-split fashion, with heat flux along each direction
-calculated at the + and - faces of the cell using the arithmetic mean
-of the cell-centered temperature in cells $n$ and $n+1$ or $n-1$ and
-$n$, respectively (empirically, this is more stable than taking the
-geometric mean of the cell-centered temperatures).  The addition of
-transport along magnetic field lines requires the calculation of
-cross-terms in the temperature derivatives at cell faces, which can
-result in spurious oscillations in the temperature field in regions
-where the temperature gradient is strong in more than one spatial
-direction.  Controlling these oscillations requires the addition of a
-flux limiter for calculations of the temperature field.  In this case,
-we choose the monotonized central difference flux limiter
-\citep{1977JCoPh..23..263V}, which serves to maintain numerical
-stability without sacrificing substantial speed or accuracy.
+Both isotropic and anisotropic thermal conduction in \enzo\ are
+treated in an operator-split manner.  Furthermore, within the heat
+transport module, transport along the x, y, and z directions are
+computed in a directionally-split fashion, with heat flux along each
+direction calculated at the + and - faces of the cell using the
+arithmetic mean of the cell-centered temperature in cells $n$ and
+$n+1$ or $n-1$ and $n$, respectively (empirically, this is more stable
+than taking the geometric mean of the cell-centered temperatures).
+The addition of transport along magnetic field lines requires the
+calculation of cross-terms in the temperature derivatives at cell
+faces, which can result in spurious oscillations in the temperature
+field in regions where the temperature gradient is strong in more than
+one spatial direction.  Controlling these oscillations requires the
+addition of a flux limiter for calculations of the temperature field.
+In this case, we choose the monotonized central difference flux
+limiter \citep{1977JCoPh..23..263V}, which serves to maintain
+numerical stability without sacrificing substantial speed or accuracy.

### 2013-06-25 numerical-mhd-dedner.tex
Reformatting tex files so they're all standardized in terms of paragraph wrapping, etc. (no content changed, just formatting.)

@@
 
-The second method we describe is a MUSCL-based solver than can be used in both HD and MHD modes.
-The description here will be very brief both because the ideas are similar to those described in the previous section, and because this implementation has previously
-been described in more detail elsewhere \citep{WangAbelZhang08, WangAbel09}.
+The second method we describe is a MUSCL-based solver than can be used
+in both HD and MHD modes.  The description here will be very brief
+both because the ideas are similar to those described in the previous
+section, and because this implementation has previously been described
+in more detail elsewhere \citep{WangAbelZhang08, WangAbel09}.
 
-Much like the PPM solver, we have three basic steps: the first is
-reconstruction of the variables, the second is a solution of the
-Riemann problem, and the third is is updating the conserved quanties
-with the fluxes as written above.
-For the reconstruction scheme we have implemented only the simple
-piecewise linear reconstruction \citep{1979JCoPh..32..101V, 1985JCoPh..59..264C},
-with options for both primitive and conservative variable reconstruction.
+Much like the PPM solver, we have three basic steps: the first is reconstruction of the variables, the second is a solution of the Riemann problem, and the third is is updating the conserved quanties with the fluxes as written above.
+For the reconstruction scheme we have implemented only the simple piecewise linear reconstruction \citep{1979JCoPh..32..101V, 1985JCoPh..59..264C}, with options for both primitive and conservative variable reconstruction.
 The available Riemann solvers are HLL, HLLC, and HLLD, as described earlier.
 
-To more clearly describe the Dedner divergence cleaning modifications, we write the equations of compressible inviscid hydrodynamics in the form of conservation laws as,
+To more clearly describe the Dedner divergence cleaning modifications,
+we write the equations of compressible inviscid hydrodynamics in the
+form of conservation laws as,
+
 \begin{equation}
@@
 \end{equation}
+
 The conserved variable $U$ is given by
+
 \begin{equation}
@@
 \end{equation}
+
 where $\rho$ is density, $v_i$ are the three components of velocity
@@
 conserved variables
+
 \begin{equation}
@@
 \end{equation}
+
 where $B_i$ with $i={x,y,z}$ are the three components of magnetic
@@
 formulation for the divergence cleaning.  The fluxes then are
+
 \begin{eqnarray}
@@
 \end{eqnarray}
+
 where $c_h$ is a constant controlling the propagation speed and
@@
 one step, in contrast to the Strang splitting employed in the other
-fluid methods described in this paper.
-Also unlike the other schemes, time-integration is done with a second-order
-Runge-Kutta scheme \citep{1988JCoPh..77..439S}.
+fluid methods described in this paper.  Also unlike the other schemes,
+time-integration is done with a second-order Runge-Kutta scheme
+\citep{1988JCoPh..77..439S}.
 
@@
 strength).  This adds a source term of $-\dot{a}/(2a) \vecB$ on the
-right-hand side of Equation~(\ref{eq:induction}) and removes the factor
-of $a$ in the $\vecB$ term in the energy equation
+right-hand side of Equation~(\ref{eq:induction}) and removes the
+factor of $a$ in the $\vecB$ term in the energy equation
 (Eq.~\ref{eq:total_energy_def}) and in the definition of the isotropic

### 2013-06-25 numerical-ppm.tex
Removed commented-out text from lots of files - basically vestigial junk.
Reformatting tex files so they're all standardized in terms of paragraph wrapping, etc. (no content changed, just formatting.)

@@
-In this section, we describe the four solvers that we have implemented for
-solving the fluid equations.  We describe the PPM method in considerably
-more detail than the other methods in part
-because its implementation in Enzo has not previously been described,
-but mostly because it introduces many of the ideas and methods used for
-the MUSCL-Dedner and MHDCT schemes (including expansion terms,
+In this section, we describe the four solvers that we have implemented
+for solving the fluid equations.  We describe the PPM method in
+considerably more detail than the other methods in part because its
+implementation in Enzo has not previously been described, but mostly
+because it introduces many of the ideas and methods used for the
+MUSCL-Dedner and MHDCT schemes (including expansion terms,
 reconstruction, Riemann solvers, and dual energy formalism).
@@
 
-One (purely) hydrodynamic method used in \enzo\ is closely based on the
-piecewise parabolic method (PPM) of
-\citet{1984JCoPh..54..174C} -- henceforth referred to as CW84 -- which has been
-modified for the study of cosmological fluid flows.  
-CW84 describe two variants on this method:
-Lagrangian Remap and Direct Eulerian.  We use the Direct Eulerian
-version in \enzo, which is better suited for AMR simulations.  The Lagrangian
-Remap version has previously been adapted for cosmological use
-as described in \citet{1995CoPhC..89..149B}, and the implementation
-we use here is closely based on that work.
+One (purely) hydrodynamic method used in \enzo\ is closely based on
+the piecewise parabolic method (PPM) of \citet{1984JCoPh..54..174C} --
+henceforth referred to as CW84 -- which has been modified for the
+study of cosmological fluid flows.  CW84 describe two variants on this
+method: Lagrangian Remap and Direct Eulerian.  We use the Direct
+Eulerian version in \enzo, which is better suited for AMR simulations.
+The Lagrangian Remap version has previously been adapted for
+cosmological use as described in \citet{1995CoPhC..89..149B}, and the
+implementation we use here is closely based on that work.
 
-The first term on the right hand side of equations~(\ref{eq:momentum}) and
-(\ref{eq:total_energy}) comes from our choice of comoving coordinates
-(a similar term also appears in equation~(\ref{eq:dm_velocity}), the velocity 
-relation for the dark matter particles).  A similar term does not
-appear in the mass conservation equation (\ref{eq:mass}) because
-of the comoving density definition.   
-We note that these expansion terms could be
-eliminated entirely by the proper choice of variables (including time),
-although we have not done so here as they do not constitute a major
-source of error.  We solve these terms
-using the same technique -- they are split from the rest of the terms and
-solved using an (implicit) time-centered method, which is straightforward
-as there are no spatial gradients.  Note that we use this method for
-all four of the hydro solvers.
+The first term on the right hand side of equations~(\ref{eq:momentum})
+and (\ref{eq:total_energy}) comes from our choice of comoving
+coordinates (a similar term also appears in
+equation~(\ref{eq:dm_velocity}), the velocity relation for the dark
+matter particles).  A similar term does not appear in the mass
+conservation equation (\ref{eq:mass}) because of the comoving density
+definition.  We note that these expansion terms could be eliminated
+entirely by the proper choice of variables (including time), although
+we have not done so here as they do not constitute a major source of
+error.  We solve these terms using the same technique -- they are
+split from the rest of the terms and solved using an (implicit)
+time-centered method, which is straightforward as there are no spatial
+gradients.  Note that we use this method for all four of the hydro
+solvers.
 
-The remainder of the terms in the fluid equations
-involve derivatives.  
-Because we are interested in phenomena with no special geometry, we will
-restrict discussion to Cartesian coordinates.
-We can dimensionally split the equations and rewrite the one-dimensional (Eulerian) 
-versions of equations~(\ref{eq:mass})--(\ref{eq:total_energy}) without expansion terms
-in conservative form as,
+The remainder of the terms in the fluid equations involve derivatives.
+Because we are interested in phenomena with no special geometry, we
+will restrict discussion to Cartesian coordinates.  We can
+dimensionally split the equations and rewrite the one-dimensional
+(Eulerian) versions of
+equations~(\ref{eq:mass})--(\ref{eq:total_energy}) without expansion
+terms in conservative form as,
 \begin{eqnarray}
@@
 
-%\subsubsection{The piecewise parabolic method in one dimension} % 2.2
-
 We now restrict ourselves to the solution of equations
-(\ref{eq:mass1d})--(\ref{eq:total_energy1d}) in one
-dimension.  In the Direct Eulerian version of PPM, this is
-accomplished by a three-step process.  First, we compute effective
-left and right states at each grid boundary by constructing a
-piecewise parabolic description of the primative variables ($\rho$, $u$
-and $E$) and then averaging over the regions corresponding to the
-distance each characteristic wave can travel ($u$, $u-c_s$ and
-$u+c_s$, where $c_s$ is the sound speed).
+(\ref{eq:mass1d})--(\ref{eq:total_energy1d}) in one dimension.  In the
+Direct Eulerian version of PPM, this is accomplished by a three-step
+process.  First, we compute effective left and right states at each
+grid boundary by constructing a piecewise parabolic description of the
+primative variables ($\rho$, $u$ and $E$) and then averaging over the
+regions corresponding to the distance each characteristic wave can
+travel ($u$, $u-c_s$ and $u+c_s$, where $c_s$ is the sound speed).
 Second, a Riemann problem is solved using these effective left and
@@
 solution to this Riemann problem and the conserved quantities are
-updated.  This is described in detail in
-CW84, but we will briefly outline the procedure
-here both for completeness and to put the changes we will make in
-context.
+updated.  This is described in detail in CW84, but we will briefly
+outline the procedure here both for completeness and to put the
+changes we will make in context.
 
@@
 
-We have used subscripts to indicate zone-centered ($j$)
-and face-centered ($j+1/2$) quantities, while superscripts refer to
-position in time.  The cell width is $\Delta x_j$.
-Although they have been discretized
-in space, the accuracy of the updates depend on how well we can compute
-the fluxes into and out of the cell during $\Delta t$.  
-This in turn depends on our ability to compute the time-averaged 
-values of $p$, $\rho$ and $v$ at the cell interfaces, denoted here
-by $\overline{p}_{j\pm 1/2}$, $\overline{\rho}_{j\pm 1/2}$, and $\overline{v}_{j\pm 1/2}$.  We now
-describe the steps required to compute these quantities.
+We have used subscripts to indicate zone-centered ($j$) and
+face-centered ($j+1/2$) quantities, while superscripts refer to
+position in time.  The cell width is $\Delta x_j$.  Although they have
+been discretized in space, the accuracy of the updates depend on how
+well we can compute the fluxes into and out of the cell during $\Delta
+t$.  This in turn depends on our ability to compute the time-averaged
+values of $p$, $\rho$ and $v$ at the cell interfaces, denoted here by
+$\overline{p}_{j\pm 1/2}$, $\overline{\rho}_{j\pm 1/2}$, and
+$\overline{v}_{j\pm 1/2}$.  We now describe the steps required to
+compute these quantities.
 
-We first construct monotonic piecewise 
-parabolic (third-order) interpolations in one dimension
-for each of $p$, $\rho$, and $v$.  The pressure is determined from
-equation~(\ref{eq:eq_of_state}), the equation of state.
-The interpolation formula for some quantity $q$ is given by:
+We first construct monotonic piecewise parabolic (third-order)
+interpolations in one dimension for each of $p$, $\rho$, and $v$.  The
+pressure is determined from equation~(\ref{eq:eq_of_state}), the
+equation of state.  The interpolation formula for some quantity $q$ is
+given by:
 %
@@
 %
-This is equation~(1.4) of CW84 (in the spatial rather than mass coordinate, as is appropriate for the direct Eulerian approach). 
-The quantities $q_{L,j}$, $\Delta q_j$,
-and $q_{6,j}$ can be viewed simply as interpolation constants; however,
-they also have more intuitive meanings.  For example, $q_{L,j}$ is the
-value of $q$ at the left edge of zone j, while $\Delta q_j$ and $q_{6,j}$ are analogous to the slope and first-order correction to the slope of $q$ (see CW84 for a complete discussion):
+This is equation~(1.4) of CW84 (in the spatial rather than mass
+coordinate, as is appropriate for the direct Eulerian approach).  The
+quantities $q_{L,j}$, $\Delta q_j$, and $q_{6,j}$ can be viewed simply
+as interpolation constants; however, they also have more intuitive
+meanings.  For example, $q_{L,j}$ is the value of $q$ at the left edge
+of zone j, while $\Delta q_j$ and $q_{6,j}$ are analogous to the slope
+and first-order correction to the slope of $q$ (see CW84 for a
+complete discussion):
 \begin{equation}
@@
 
-We have reduced the problem to finding $q_{L,j}$ and $q_{R,j}$.  While this
-is simple in principle, it is complicated somewhat by the requirement that
-these values be of sufficient accuracy and that the resulting distribution
-be monotonic.  That is, no new maxima or minima are introduced.
-The resulting formulae are straightforward but complicated and are not
-reproduced here, but see Equations 1.7 to 1.10 of CW84.
-We also optionally allow steepening as described in that reference.
+We have reduced the problem to finding $q_{L,j}$ and $q_{R,j}$.  While
+this is simple in principle, it is complicated somewhat by the
+requirement that these values be of sufficient accuracy and that the
+resulting distribution be monotonic.  That is, no new maxima or minima
+are introduced.  The resulting formulae are straightforward but
+complicated and are not reproduced here, but see Equations 1.7 to 1.10
+of CW84.  We also optionally allow steepening as described in that
+reference.
 
-Once we have the reconstruction, the primary quantities ($p$, $\rho$, $v$ and $E$) are averaged over the domains corresponding to the three characteristics $u-c$, $u$, or $u+c$ (where $c$ is the sound speed in a cell).
-The linearized gas dynamics equations are then used to compute second-order accurate left and right states that take into account the multiple wave families.
-This process is described in CW84 and we use their equations (3.6) and (3.7).
+Once we have the reconstruction, the primary quantities ($p$, $\rho$,
+$v$ and $E$) are averaged over the domains corresponding to the three
+characteristics $u-c$, $u$, or $u+c$ (where $c$ is the sound speed in
+a cell).  The linearized gas dynamics equations are then used to
+compute second-order accurate left and right states that take into
+account the multiple wave families.  This process is described in CW84
+and we use their equations (3.6) and (3.7).
 
-With these effective states, an approximation to the Riemann problem is found
-(see below for more detail about the Riemann solvers used), producing estimates for 
-$\overline{p}_{j\pm 1/2}$, $\overline{\rho}_{j\pm1/2}$, and $\overline{v}_{j\pm 1/2}$ that are third-order
-accurate in space and second-order accurate in time.  These are then
-used to solve the difference Equations (\ref{eq:mass_diff})--(\ref{eq:energy_diff}) for $\rho^{n+1}$, $v^{n+1}$, and $E^{n+1}$.
+With these effective states, an approximation to the Riemann problem
+is found (see below for more detail about the Riemann solvers used),
+producing estimates for $\overline{p}_{j\pm 1/2}$,
+$\overline{\rho}_{j\pm1/2}$, and $\overline{v}_{j\pm 1/2}$ that are
+third-order accurate in space and second-order accurate in time.
+These are then used to solve the difference Equations
+(\ref{eq:mass_diff})--(\ref{eq:energy_diff}) for $\rho^{n+1}$,
+$v^{n+1}$, and $E^{n+1}$.
 
-We include an optional diffusive flux (and flattening for the parabolic curves) that can improve the solution in some cases.  Our implementation follows that in the appendix of CW84.
+We include an optional diffusive flux (and flattening for the
+parabolic curves) that can improve the solution in some cases.  Our
+implementation follows that in the appendix of CW84.
 
-In addition, as discussed earlier, the three-dimensional scheme is achieved by operator-splitting and repeating the above procedure in the other two orthogonal directions.
-The transverse velocities and any additional passive quantities are naturally and easily added to this system (see Equation 3.6 of CW84).
+In addition, as discussed earlier, the three-dimensional scheme is
+achieved by operator-splitting and repeating the above procedure in
+the other two orthogonal directions.  The transverse velocities and
+any additional passive quantities are naturally and easily added to
+this system (see Equation 3.6 of CW84).
 
-We note that the acceleration required in Equation~(\ref{eq:momentum_diff}) is actually the
-acceleration felt by the entire zone and not just at the zone center.
-Therefore, it is possible to find the mass-weighted average acceleration over the zone
-by expanding the density and acceleration distributions and
-retaining all terms up to second-order in $\Delta x$ \citep{1995CoPhC..89..149B}, although we find that
-the potential is so slowly varying that this is unnecessary.
-%
-%\begin{equation}
-%g_j^{n+1/2} = 
-%       \frac{1}{2 a^{n+1/2} \delta x_j} \left[ 
-%             \phi_{j+1}^{n+1/2} 
-%           - \phi_{j-1}^{n+1/2} 
-%           + \frac{1}{12} \left(    \phi^{n+1/2}_{j+1} 
-%                                 - 2\phi^{n+1/2}_j 
-%                                 + \phi^{n+1/2}_{j-1} \right) 
-%                                   {\delta d_j \over d_j}
-%       \right].
-%       \label{eq:accel}
-%\end{equation}
-%Generally we find that the potential is so slowly varying that this change is unnecessary.
+We note that the acceleration required in
+Equation~(\ref{eq:momentum_diff}) is actually the acceleration felt by
+the entire zone and not just at the zone center.  Therefore, it is
+possible to find the mass-weighted average acceleration over the zone
+by expanding the density and acceleration distributions and retaining
+all terms up to second-order in $\Delta x$
+\citep{1995CoPhC..89..149B}, although we find that the potential is so
+slowly varying that this is unnecessary.
 
@@
 The system described thus far works well for gravitating systems with
-reasonable Mach numbers ($<100$) as long as the structures are well resolved.
-This section and the next detail changes that are required
+reasonable Mach numbers ($<100$) as long as the structures are well
+resolved.  This section and the next detail changes that are required
 to correctly account for situations where one or both of these
@@
 high as $10^8$.  Inverted, we see that the internal energy consists of
-an extremely small portion of the total energy.  In such a situation, the pressure,
-proportional to $E - E_{k}$, is the small difference between two large
-numbers: a disastrous numerical situation.  This is not as large a
-problem as it may at first appear because it only occurs when the
-pressure is negligibly small.  Therefore, even if we suffer large
-errors in the pressure distribution in these regions, the dynamics and
-total energy budget of the flow will remain unaffected.  Nevertheless,
-if the temperature distribution is required for other reasons
-(e.g., for calculating radiative processes), a remedy is required.
+an extremely small portion of the total energy.  In such a situation,
+the pressure, proportional to $E - E_{k}$, is the small difference
+between two large numbers: a disastrous numerical situation.  This is
+not as large a problem as it may at first appear because it only
+occurs when the pressure is negligibly small.  Therefore, even if we
+suffer large errors in the pressure distribution in these regions, the
+dynamics and total energy budget of the flow will remain unaffected.
+Nevertheless, if the temperature distribution is required for other
+reasons (e.g., for calculating radiative processes), a remedy is
+required.
 
@@
 in comoving coordinates.  The structure is similar to the total energy
-equation; the second term on the left hand side represents transport, while
-the first term on the right is due to expansion of the coordinate
-system.  
-It is differenced (again, in Eulerian form without the expansion term) as,
+equation; the second term on the left hand side represents transport,
+while the first term on the right is due to expansion of the
+coordinate system.  It is differenced (again, in Eulerian form without
+the expansion term) as,
 %
@@
 proved difficult to do without generating large errors in the internal
-energy and so we leave the pressure at the old time in
-this difference equation.  This leads to some spreading of shocks; however,
-we note that this equation is only used in hypersonic flows.
+energy and so we leave the pressure at the old time in this difference
+equation.  This leads to some spreading of shocks; however, we note
+that this equation is only used in hypersonic flows.
 
-It is necessary, however, to conserve the total energy so that the conversion
-of kinetic to thermal energy is performed properly.   We must therefore
-combine the two formulations without allowing the separately-advected
-internal energy $e$ to play a role in the gas dynamics.  This is done by
-carrying both terms through the simulation and using
-the total energy $E$ for hydrodynamic routines and the internal energy
-$e$ when the temperature profile is required.  One way to view this procedure
-is to treat $e$ as enhanced precision (extra digits) for $E$ that automatically
-`floats' to where it is needed.
-We only require that they be kept synchronized when the two levels of precision
-overlap.
+It is necessary, however, to conserve the total energy so that the
+conversion of kinetic to thermal energy is performed properly.  We
+must therefore combine the two formulations without allowing the
+separately-advected internal energy $e$ to play a role in the gas
+dynamics.  This is done by carrying both terms through the simulation
+and using the total energy $E$ for hydrodynamic routines and the
+internal energy $e$ when the temperature profile is required.  One way
+to view this procedure is to treat $e$ as enhanced precision (extra
+digits) for $E$ that automatically `floats' to where it is needed.  We
+only require that they be kept synchronized when the two levels of
+precision overlap.
 
@@
 
-It should be stressed that as long as the parameter ${\eta}_1$ is small enough
-the dual energy method {\it will have no dynamical effect}.  We use
-${\eta}_1 = 10^{-3}$, which is consistent with the truncation error of the
-scheme for grid sizes that are typically used in our simulations.  We are now free to select the method by which the internal energy
-field variable $e$ is updated so that it will not become contaminated with
-errors advected by the total energy formulation but still give the
-correct distribution in shocked regions. 
-Since we are concerned with the advection of errors, the selection
-criterion must look at each cell's local neighbourhood.
-In one dimension, this is done with,
+It should be stressed that as long as the parameter ${\eta}_1$ is
+small enough the dual energy method {\it will have no dynamical
+effect}.  We use ${\eta}_1 = 10^{-3}$, which is consistent with the
+truncation error of the scheme for grid sizes that are typically used
+in our simulations.  We are now free to select the method by which the
+internal energy field variable $e$ is updated so that it will not
+become contaminated with errors advected by the total energy
+formulation but still give the correct distribution in shocked
+regions.  Since we are concerned with the advection of errors, the
+selection criterion must look at each cell's local neighbourhood.  In
+one dimension, this is done with,
+
 \begin{equation}
@@
 \end{equation}
-Thus, ${\eta}_2$ determines when the synchronization (of $e$ with $E$) occurs.
-%is a measure of how much error we will allow to be passed
-%from neighboring cells.  
-Too high a value may mask relatively weak shocks, 
-while spurious heating (via contamination) may occur if it is set too low.
-After some experimentation,
-we have chosen ${\eta}_2 = 0.1$, a somewhat conservative value.
-This scheme is optional and is generally only required in large-scale 
-cosmological simulations where the gas cools due to the expansion of the universe
-but large bulk flows develop due to the formation of structure.
+
+Thus, ${\eta}_2$ determines when the synchronization (of $e$ with $E$)
+occurs.  Too high a value may mask relatively weak shocks, while
+spurious heating (via contamination) may occur if it is set too low.
+After some experimentation, we have chosen ${\eta}_2 = 0.1$, a
+somewhat conservative value.  This scheme is optional and is generally
+only required in large-scale cosmological simulations where the gas
+cools due to the expansion of the universe but large bulk flows
+develop due to the formation of structure.
 
@@
 distinct scheme for dealing with this problem, which is endemic to
-methods adopting the total energy equation.  In \citet{TVD93}, the 
-two variables adopted are total energy and
-entropy (rather than total energy and thermal energy), with an
-analogous scheme for choosing which variable to employ.
+methods adopting the total energy equation.  In \citet{TVD93}, the two
+variables adopted are total energy and entropy (rather than total
+energy and thermal energy), with an analogous scheme for choosing
+which variable to employ.
 
@@
 In this section, we describe the methods we adopt to solve the Riemann
-problem, which is generally required to compute the fluxes in 
-any Godunov-based scheme.  This section therefore applies to all three of our 
-Godunov-based schemes.  The Riemann problem we are solving
-involves two constant states separated by a single discontinuity at $t=0$.
-The subsequent evolution has an exact analytic solution.  This solution is
-described in detail in many texts on computational fluid dynamics
-\citep[e.g.,][]{toro-1997}.   In brief, there are three waves that propagate
-away from the initial discontinuity.  The central wave, characterized by a
-density jump but not a pressure jump, is called the contact discontinuity.  The
-waves traveling to the left and right of the contact discontinuity can be either
-shocks, if characteristics converge on the wave front, or rarefaction fans if
-characteristics diverge.
+problem, which is generally required to compute the fluxes in any
+Godunov-based scheme.  This section therefore applies to all three of
+our Godunov-based schemes.  The Riemann problem we are solving
+involves two constant states separated by a single discontinuity at
+$t=0$.  The subsequent evolution has an exact analytic solution.  This
+solution is described in detail in many texts on computational fluid
+dynamics \citep[e.g.,][]{toro-1997}.  In brief, there are three waves
+that propagate away from the initial discontinuity.  The central wave,
+characterized by a density jump but not a pressure jump, is called the
+contact discontinuity.  The waves traveling to the left and right of
+the contact discontinuity can be either shocks, if characteristics
+converge on the wave front, or rarefaction fans if characteristics
+diverge.
 
-While there exists an exact solution to this problem, finding it is expensive.  There are four
-possible combination of left- and right- traveling shocks and rarefactions,
-only one of which is fully consistent with the initial conditions.  Once the
-correct physical state is determined, the pressure in the central region can
-only be computed by finding the root to an algebraic equation, which is necessarily
-an iterative process.  Thus a series of \emph{approximate} Riemann solvers are
-typically used.  There are four approximate Riemann solvers in \enzo: two-shock
-\citep{toro-1997},
-Harten-Lax-van Leer \citep[HLL,][]{toro-1997}, HLL with a contact discontinuity
- \citep[HLLC,][]{toro-1997}, and HLL with
-multiple discontinuities \citep[HLLD,][]{Miyoshi05}.  Two-shock is used only
-with the PPM method.  HLL and HLLC are used with PPM, MUSCL (both with and
-without MHD) and MHD-CT.  HLLD is exclusively an MHD solver, and works
-with both the
-MUSCL and MHD-CT methods.
+While there exists an exact solution to this problem, finding it is
+expensive.  There are four possible combination of left- and right-
+traveling shocks and rarefactions, only one of which is fully
+consistent with the initial conditions.  Once the correct physical
+state is determined, the pressure in the central region can only be
+computed by finding the root to an algebraic equation, which is
+necessarily an iterative process.  Thus a series of \emph{approximate}
+Riemann solvers are typically used.  There are four approximate
+Riemann solvers in \enzo: two-shock \citep{toro-1997}, Harten-Lax-van
+Leer \citep[HLL,][]{toro-1997}, HLL with a contact discontinuity
+\citep[HLLC,][]{toro-1997}, and HLL with multiple discontinuities
+\citep[HLLD,][]{Miyoshi05}.  Two-shock is used only with the PPM
+method.  HLL and HLLC are used with PPM, MUSCL (both with and without
+MHD) and MHD-CT.  HLLD is exclusively an MHD solver, and works with
+both the MUSCL and MHD-CT methods.
 
-The only approximation that two-shock makes is that both left- and right-moving
-waves are shocks.  This solution still requires an iterative method for finding
-the pressure in between the two waves.  The HLL method alleviates this iteration
-by assuming that there is no central contact discontinuity, and the signal speed
-in the central region is approximated by an average over the left- and right-
-moving waves.  This method is significantly faster than the two-shock method,
-but it is also quite a bit more dissipative.  The HLLC is a three-wave method that improves upon the HLL
-method by also including the third wave, the contact discontinuity.
+The only approximation that two-shock makes is that both left- and
+right-moving waves are shocks.  This solution still requires an
+iterative method for finding the pressure in between the two waves.
+The HLL method alleviates this iteration by assuming that there is no
+central contact discontinuity, and the signal speed in the central
+region is approximated by an average over the left- and right- moving
+waves.  This method is significantly faster than the two-shock method,
+but it is also quite a bit more dissipative.  The HLLC is a three-wave
+method that improves upon the HLL method by also including the third
+wave, the contact discontinuity.
 
-For the MHD equations, there are seven waves instead of three.  This makes the
-exact solution to the Riemann problem quite a bit more expensive.  Both the HLL and HLLC
-approximations can be formulated for the MHD equations, and are employed in both
-the Dedner and CT solvers in \enzo.  The HLLD
-solver includes two of the additional waves, the rotational discontinuities,
-making it a five-wave solver.
+For the MHD equations, there are seven waves instead of three.  This
+makes the exact solution to the Riemann problem quite a bit more
+expensive.  Both the HLL and HLLC approximations can be formulated for
+the MHD equations, and are employed in both the Dedner and CT solvers
+in \enzo.  The HLLD solver includes two of the additional waves, the
+rotational discontinuities, making it a five-wave solver.
 
-On rare occasions, high-order solutions can cause negative densities or energies.  
-Both our PPM and MUSCL solvers employ a Riemann
-solver fallback mechanism \citep{Lemaster09}.  If a negative density is found at
-a particular interface, the more diffusive HLL Riemann solver is used to compute the fluxes associated with that cell, and the flux update is repeated.
+On rare occasions, high-order solutions can cause negative densities
+or energies.  Both our PPM and MUSCL solvers employ a Riemann solver
+fallback mechanism \citep{Lemaster09}.  If a negative density is found
+at a particular interface, the more diffusive HLL Riemann solver is
+used to compute the fluxes associated with that cell, and the flux
+update is repeated.

### 2013-06-25 numerical-timestep.tex
Reformatting tex files so they're all standardized in terms of paragraph wrapping, etc. (no content changed, just formatting.)

@@
 
-In \enzo, the integration of the equations being solved is generally adaptive in time 
-as well as in space.  The timestep $\Delta t$ is set on a level-by-level
-basis by finding the largest timestep such that all of the criteria
-listed below (that are relevant for the simulation in question) are
-satisfied.  The timestep criteria are given by the following
-expressions, showing the one-dimensional case for clarity:
+In \enzo, the integration of the equations being solved is generally
+adaptive in time as well as in space.  The timestep $\Delta t$ is set
+on a level-by-level basis by finding the largest timestep such that
+all of the criteria listed below (that are relevant for the simulation
+in question) are satisfied.  The timestep criteria are given by the
+following expressions, showing the one-dimensional case for clarity:
 
@@
 
-In equations~(\ref{eqn:dthydro})-(\ref{eqn:dtcond}), the $\min ( \ldots)_L$ 
-formalism means that this value is calculated for all cells or
-particles on a given level L and the minimum overall value is taken as the timestep.
+In equations~(\ref{eqn:dthydro})-(\ref{eqn:dtcond}), the $\min (
+\ldots)_L$ formalism means that this value is calculated for all cells
+or particles on a given level L and the minimum overall value is taken
+as the timestep.
 
@@
 an explicit finite difference discretization of the Euler equations.
-In this equation, $\kappa_{\rm hydro}$ is a dimensionless
-numerical constant with a value of $0 < \kappa_{\rm hydro} \leq 1$
-(with a typical value of $\kappa_{\rm hydro} \sim 0.3-0.5$) that
-ensures that the CFL condition is always met, and $c_s$ and $v_x$ are
-the sound speed and peculiar baryon velocity in a given cell.
+In this equation, $\kappa_{\rm hydro}$ is a dimensionless numerical
+constant with a value of $0 < \kappa_{\rm hydro} \leq 1$ (with a
+typical value of $\kappa_{\rm hydro} \sim 0.3-0.5$) that ensures that
+the CFL condition is always met, and $c_s$ and $v_x$ are the sound
+speed and peculiar baryon velocity in a given cell.
 
@@
 when the equations of hydrodynamics are being solved.  For 2 or 3
-dimensions, it was shown by \cite{Godunov1959}  that using the
-harmonic average of the timestep found along each of the coordinate
-axes yields a maximum $\kappa_{\rm hydro} = 0.8$.  So letting $\Delta
-t_x$, $\Delta t_y$, and $\Delta t_z$ be the analogues of
+dimensions, it was shown by \cite{Godunov1959} that using the harmonic
+average of the timestep found along each of the coordinate axes yields
+a maximum $\kappa_{\rm hydro} = 0.8$.  So letting $\Delta t_x$,
+$\Delta t_y$, and $\Delta t_z$ be the analogues of
 equation~(\ref{eqn:dthydro}) along the $x,y$ and $z$ axes,
+
 \begin{equation}
@@
 
-For all other criteria except for equation~(\ref{eqn:dtMHD}), multiple dimensions are accounted for by
-repeating the one dimensional criterion along each axis, and taking the minimum.
+For all other criteria except for equation~(\ref{eqn:dtMHD}), multiple
+dimensions are accounted for by repeating the one dimensional
+criterion along each axis, and taking the minimum.
 
@@
 is being enforced at all times.  In this equation, $\kappa_{\rm MHD}$
-is a dimensionless
-numerical constant with a value of $0 < \kappa_{\rm MHD} \leq 1$ (with a
-typical value of $\kappa_{\rm MHD} \sim 0.5$) that ensures that the CFL
-condition is always met, and $v_f$ and $v_x$ are the ``fast wave
-speed'' and peculiar baryon velocity in a given cell.  The fast wave
-speed comes from a stability analysis of the MHD equations, and is
-given by:
+is a dimensionless numerical constant with a value of $0 < \kappa_{\rm
+MHD} \leq 1$ (with a typical value of $\kappa_{\rm MHD} \sim 0.5$)
+that ensures that the CFL condition is always met, and $v_f$ and $v_x$
+are the ``fast wave speed'' and peculiar baryon velocity in a given
+cell.  The fast wave speed comes from a stability analysis of the MHD
+equations, and is given by:
 %
@@
 %
-where $c_s$ is the sound speed and $v_A$ is the Alfven speed, calculated
-as $v_A = \sqrt{B^2/\rho_B}$ in units where $\mu_0 = 1$.
+where $c_s$ is the sound speed and $v_A$ is the Alfven speed,
+calculated as $v_A = \sqrt{B^2/\rho_B}$ in units where $\mu_0 = 1$.
 
 Equation~(\ref{eqn:dtdarkmatter}) is analogous to
-equation~(\ref{eqn:dthydro}) and helps to ensure accuracy in the N-body solver
-by requiring that no particle travels more than one cell
-width.  The parameter $\kappa_{\rm dm}$ is like $\kappa_{\rm
-hydro}$, with a similar range of values.  This criterion is used when
-massive particles are included in a simulation.
+equation~(\ref{eqn:dthydro}) and helps to ensure accuracy in the
+N-body solver by requiring that no particle travels more than one cell
+width.  The parameter $\kappa_{\rm dm}$ is like $\kappa_{\rm hydro}$,
+with a similar range of values.  This criterion is used when massive
+particles are included in a simulation.
 
-Equations~(\ref{eqn:dtaccel}) and~(\ref{eqn:dtrad}) are supplementary to equation~(\ref{eqn:dthydro}) in that they
-take into account the possibility of large accelerations due to either
-gravity (equation~\ref{eqn:dtaccel}) or radiation pressure
-(equation~\ref{eqn:dtrad}).  In equation~(\ref{eqn:dtaccel}), $\myvec{g}$ is the
-gravitational acceleration in each cell on level l.  In
-equation~(\ref{eqn:dtrad}), $\myvec{a}_{\rm rad}$ is the estimated
+Equations~(\ref{eqn:dtaccel}) and~(\ref{eqn:dtrad}) are supplementary
+to equation~(\ref{eqn:dthydro}) in that they take into account the
+possibility of large accelerations due to either gravity
+(equation~\ref{eqn:dtaccel}) or radiation pressure
+(equation~\ref{eqn:dtrad}).  In equation~(\ref{eqn:dtaccel}),
+$\myvec{g}$ is the gravitational acceleration in each cell on level l.
+In equation~(\ref{eqn:dtrad}), $\myvec{a}_{\rm rad}$ is the estimated
 acceleration due to radiation pressure in each cell on level l,
 defined as
+
 \begin{equation}
@@
 \end{equation}
+
 where the sum calculates the energy deposited in a cell during the
@@
 solution to the equation of heat conduction.  In this expression,
-$n_b$ is the baryon number density, 
-$\kappa_{\rm sp}(T)$ is the Spitzer thermal conductivity, and f$_{\rm sp}$ is a
-user-defined, dimensionless
-conduction suppression factor whose value must be f$_{\rm sp} \leq 1$.  
-%Both are defined in Section~\ref{sec.num.conductions}.  
-$k_{\rm cond}$ is a dimensionless prefactor whose
-value must be $0 < k_{\rm cond}  \leq 0.5$, and is exactly 0.5 for the
-implemention in \enzo.
-From a practical
-perspective, it is useful to note that, unlike other timestep criteria
-discussed above (which effectively scale as $\Delta x / \sqrt{T}$ with
-the grid cell size $\Delta x$ and temperature $T$),
-the timestep criterion due to thermal conduction scales as $\Delta x^2
-/ T^{2.5}$, which can result in a rapid decrease in timestep in
-regions of high resolution and/or temperature.
+$n_b$ is the baryon number density, $\kappa_{\rm sp}(T)$ is the
+Spitzer thermal conductivity, and f$_{\rm sp}$ is a user-defined,
+dimensionless conduction suppression factor whose value must be
+f$_{\rm sp} \leq 1$.  $k_{\rm cond}$ is a dimensionless prefactor
+whose value must be $0 < k_{\rm cond} \leq 0.5$, and is exactly 0.5
+for the implemention in \enzo.  From a practical perspective, it is
+useful to note that, unlike other timestep criteria discussed above
+(which effectively scale as $\Delta x / \sqrt{T}$ with the grid cell
+size $\Delta x$ and temperature $T$), the timestep criterion due to
+thermal conduction scales as $\Delta x^2 / T^{2.5}$, which can result
+in a rapid decrease in timestep in regions of high resolution and/or
+temperature.
 
-Finally, equation~(\ref{eqn:dtexpand}) is a cosmological constraint that
-limits the timestep so that the simulated universe only expands by
-some fractional amount, f$_{\rm exp}$, during a single step.  In this
-equation, $a$ and $\dot{a}$ refer to the scale factor of the universe
-and its rate of change, respectively.  This criteria is necessary
-because the expansion of the universe and its first derivative with
-respect to time both appear in the equations of cosmological
-(magneto)hydrodynamics and particle motion, and some limit is required
-for the stability of the PPM algorithm in comoving coordinates.  This
-criterion typically limits the simulation's timestep
-only during the earliest phases of a cosmological simulation, before
-substantial structure has formed.
+Finally, equation~(\ref{eqn:dtexpand}) is a cosmological constraint
+that limits the timestep so that the simulated universe only expands
+by some fractional amount, f$_{\rm exp}$, during a single step.  In
+this equation, $a$ and $\dot{a}$ refer to the scale factor of the
+universe and its rate of change, respectively.  This criteria is
+necessary because the expansion of the universe and its first
+derivative with respect to time both appear in the equations of
+cosmological (magneto)hydrodynamics and particle motion, and some
+limit is required for the stability of the PPM algorithm in comoving
+coordinates.  This criterion typically limits the simulation's
+timestep only during the earliest phases of a cosmological simulation,
+before substantial structure has formed.

### 2013-06-25 overview.tex
Reformatting tex files so they're all standardized in terms of paragraph wrapping, etc. (no content changed, just formatting.)

@@
 describe the numerical algorithms that we use to solve these
-equations.  This section is intended to be an overview of
-\enzo's capabilities: thus, we gather all of the equations
-solved into a single place and provide a brief and high-level
-introduction to the numerics of the code.  Detailed descriptions of
-the individual components are then provided in
+equations.  This section is intended to be an overview of \enzo's
+capabilities: thus, we gather all of the equations solved into a
+single place and provide a brief and high-level introduction to the
+numerics of the code.  Detailed descriptions of the individual
+components are then provided in
 Sections~\ref{sec.amr}-\ref{sec.num.analysis}. In
@@
 (AMR) capability, which allows it to reach extremely large spatial and
-temporal dynamical
-ranges with limited computational resources, opening doors to applications otherwise
-closed by finite memory and computational time. Unlike moving mesh
-methods \citep{1995ApJS..100..269P,1995ApJS...97..231G} or methods
-that subdivide individual cells \citep{Adjerid}, Berger \& Collela's
-AMR (also referred to as \emph{structured} AMR, or SAMR) utilizes an adaptive
-hierarchy of grid patches at varying levels of resolution.  Each
-rectangular grid patch (referred to as a ``grid'') covers some region
-of space requiring higher resolution than its \emph{parent grid},
-and can itself become the parent grid to an even more highly resolved
-\emph{child grid}.
+temporal dynamical ranges with limited computational resources,
+opening doors to applications otherwise closed by finite memory and
+computational time. Unlike moving mesh methods
+\citep{1995ApJS..100..269P,1995ApJS...97..231G} or methods that
+subdivide individual cells \citep{Adjerid}, Berger \& Collela's AMR
+(also referred to as \emph{structured} AMR, or SAMR) utilizes an
+adaptive hierarchy of grid patches at varying levels of resolution.
+Each rectangular grid patch (referred to as a ``grid'') covers some
+region of space requiring higher resolution than its \emph{parent
+grid}, and can itself become the parent grid to an even more highly
+resolved \emph{child grid}.
 
 The grid hierarchy begins with the root grid, which covers the entire
-domain of interest with a coarse uniform Cartesian grid. Then, as
-the solution evolves and interesting regions start to develop, finer grids are
-placed within these coarse regions.
-We restrict the ratio between cell sizes of parent and child grids to be an integer, typically 2
-or 4, and refer to a level as all grids with the same cell size.  For simplicity, the edges of subgrids must
-coincide with the cell edge of its immediate parent (coarser)
-grid. Additionally, the hierarchy can be initialized with one or more
-static grids if a higher initial resolution is required.
+domain of interest with a coarse uniform Cartesian grid. Then, as the
+solution evolves and interesting regions start to develop, finer grids
+are placed within these coarse regions.  We restrict the ratio between
+cell sizes of parent and child grids to be an integer, typically 2 or
+4, and refer to a level as all grids with the same cell size.  For
+simplicity, the edges of subgrids must coincide with the cell edge of
+its immediate parent (coarser) grid. Additionally, the hierarchy can
+be initialized with one or more static grids if a higher initial
+resolution is required.
 
@@
 accuracy and stability criteria and advance the grid by that time
-interval.  We then move down to the next level and
-advance all the grids on that level.  If there are more levels, we repeat this procedure until
-the bottom level of the hierarchy has been reached.  Once there, we
-continue advancing the grids on the lowest level until they have
-``caught up'' to the coarser parent grids.  This procedure repeats itself until all grids have
-been advanced by the desired time.
+interval.  We then move down to the next level and advance all the
+grids on that level.  If there are more levels, we repeat this
+procedure until the bottom level of the hierarchy has been reached.
+Once there, we continue advancing the grids on the lowest level until
+they have ``caught up'' to the coarser parent grids.  This procedure
+repeats itself until all grids have been advanced by the desired time.
 
@@
 machine-vision based algorithm \citep{Berger91} to find edges and
-determine a good placement of subgrids.
-Once these new subgrids have been identified, the solution from the
-next coarser grid is interpolated in order to initialize the values on
-the new grids.  Finally, any overlap between new subgrids and old ones
-is identified, and the prior solution within the regions of overlap is
-copied to the new subgrids. This procedure is then repeated on the new
-grids and in this way, the entire hierarchy (from the original level
-examined and continuing on to all finer levels) is rebuilt.
+determine a good placement of subgrids.  Once these new subgrids have
+been identified, the solution from the next coarser grid is
+interpolated in order to initialize the values on the new grids.
+Finally, any overlap between new subgrids and old ones is identified,
+and the prior solution within the regions of overlap is copied to the
+new subgrids. This procedure is then repeated on the new grids and in
+this way, the entire hierarchy (from the original level examined and
+continuing on to all finer levels) is rebuilt.
 
@@
 by~\citet{1995CoPhC..89..149B}; (ii) the MUSCL-like Godunov scheme
-\citep{1977JCoPh..23..276V} with or without magnetic fields and Dedner-based
-divergence cleaning, described in more detail in
+\citep{1977JCoPh..23..276V} with or without magnetic fields and
+Dedner-based divergence cleaning, described in more detail in
 \citet{WangAbelZhang08} and \citet{WangAbel09}; (iii) a constrained
@@
 explicit, higher-order accurate version of Godunov's method for ideal
-gas dynamics with the spatially third-order accurate piecewise parabolic monotonic
-interpolation and a nonlinear Riemann solver for shock capturing.  It
-advances the hydrodynamic equations in the following steps:
-(i) Construct monotonic parabolic interpolation of cell average data,
-for each fluid quantity;
-(ii) Compute interface states by averaging the parabola over the
-domain of dependence for each interface;
-(iii) Use interface data to solve the Riemann problem;
+gas dynamics with the spatially third-order accurate piecewise
+parabolic monotonic interpolation and a nonlinear Riemann solver for
+shock capturing.  It advances the hydrodynamic equations in the
+following steps: (i) Construct monotonic parabolic interpolation of
+cell average data, for each fluid quantity; (ii) Compute interface
+states by averaging the parabola over the domain of dependence for
+each interface; (iii) Use interface data to solve the Riemann problem;
 (iv) Difference the interface fluxes to update the cell average
@@
 directional splitting and produce a method that is formally
-second-order accurate in space and time and which explicitly conserves mass,
-linear momentum, and energy \citep{Hawley84, Norman86}.  A variety of
-Riemann solvers have been implemented.
+second-order accurate in space and time and which explicitly conserves
+mass, linear momentum, and energy \citep{Hawley84, Norman86}.  A
+variety of Riemann solvers have been implemented.
 
-As described in \citet{1995CoPhC..89..149B}, we modify the method for use in
-hypersonic flows when the thermal energy $e$ is extremely small
+As described in \citet{1995CoPhC..89..149B}, we modify the method for
+use in hypersonic flows when the thermal energy $e$ is extremely small
 compared to the total energy $E$.  This situation presents a problem
@@
 states. Consequently, this suite of solvers can be used for hydro and
-magneto-hydrodynamic simulations. This class of solvers, as
-well as a version of the PPM hydro solver, has been ported to
-nVidia's CUDA framework, allowing \enzo\ to take advantage of modern
-graphics hardware \citep{Wang:2010}.
+magneto-hydrodynamic simulations. This class of solvers, as well as a
+version of the PPM hydro solver, has been ported to nVidia's CUDA
+framework, allowing \enzo\ to take advantage of modern graphics
+hardware \citep{Wang:2010}.
 
-\paragraph{Godunov MHD with Constrained Transport (MHD)}
-This MHD method is second-order in time and space, and preserves the
-divergence constraint, $\div \vecB = 0$, to machine precision
-through the Constrained Transport (CT) method \citep{Collins10}.  CT, originally
+\paragraph{Godunov MHD with Constrained Transport (MHD)} This MHD
+method is second-order in time and space, and preserves the divergence
+constraint, $\div \vecB = 0$, to machine precision through the
+Constrained Transport (CT) method \citep{Collins10}.  CT, originally
 described by \citet{Evans88}, updates the magnetic field with the curl
-of an electric field, suitably formulated to preserve $\div
-\vecB$.  We employ the CT methods described by \citet{Balsara99} and
-\citet{Gardiner05} with the second-order hyperbolic
-solver of \citet{Li08a} and the divergence-free AMR scheme of
-\citet{Balsara01}.
+of an electric field, suitably formulated to preserve $\div \vecB$.
+We employ the CT methods described by \citet{Balsara99} and
+\citet{Gardiner05} with the second-order hyperbolic solver of
+\citet{Li08a} and the divergence-free AMR scheme of \citet{Balsara01}.
 
@@
 particles that interact with the baryons only via gravity.  These
-particles are advanced through a single timestep using a drift-kick-drift
-algorithm \citep{Hockney88} to provide second-order accuracy even in
-the presence of varying timesteps.  Since the particles follow the
-collapse of structure, they are not adaptively refined.  Nor are there
-duplicate sets of particles for each level; instead, each particle is
-associated with the highest refined level available at its position in
-the domain and particles are moved between grids as the hierarchy is
-rebuilt. Thus a particle has the same timestep and feels the same
-gravitational force as a grid cell at that refinement level.
+particles are advanced through a single timestep using a
+drift-kick-drift algorithm \citep{Hockney88} to provide second-order
+accuracy even in the presence of varying timesteps.  Since the
+particles follow the collapse of structure, they are not adaptively
+refined.  Nor are there duplicate sets of particles for each level;
+instead, each particle is associated with the highest refined level
+available at its position in the domain and particles are moved
+between grids as the hierarchy is rebuilt. Thus a particle has the
+same timestep and feels the same gravitational force as a grid cell at
+that refinement level.
 
@@
 formation (H$^-$, H$_2$ and H$_2^+$), and then HD formation (HD, D,
-D$^+$).
-The cooling and heating due to these species is included (see the next
-section). The solution of the rate equations is carried out using one
-Jacobi iteration of an implicit Euler time discretization to ensure
-stability. To ensure accuracy the rate equations are sub-cycled
+D$^+$).  The cooling and heating due to these species is included (see
+the next section). The solution of the rate equations is carried out
+using one Jacobi iteration of an implicit Euler time discretization to
+ensure stability. To ensure accuracy the rate equations are sub-cycled
 within one hydrodynamic timestep, subject to the constraint that the
@@
 multi-species flag is turned off and no individual chemical species
-are tracked, the cooling rate is computed from a
-simple temperature-dependent cooling function, taken from
-\citet{SW87}.  If chemistry is turned on, then the code can include
-cooling from all species of hydrogen and helium (including H$_2$ and
-Deuterium-related species such as HD) --
-and the primordial cooling rates are computed in the same Jacobi iteration as the
-chemistry.  It is also possible to include metal cooling based on a
-set of multi-dimensional lookup tables computed with the
-\textsc{Cloudy} code \citep{1998PASP..110..761F} as described in
+are tracked, the cooling rate is computed from a simple
+temperature-dependent cooling function, taken from \citet{SW87}.  If
+chemistry is turned on, then the code can include cooling from all
+species of hydrogen and helium (including H$_2$ and Deuterium-related
+species such as HD) -- and the primordial cooling rates are computed
+in the same Jacobi iteration as the chemistry.  It is also possible to
+include metal cooling based on a set of multi-dimensional lookup
+tables computed with the \textsc{Cloudy} code
+\citep{1998PASP..110..761F} as described in
 \citet{2008MNRAS.385.1443S} and \citet{2011ApJ...731....6S}. Note that
@@
 limits and is coupled to an ionization network of either purely
-hydrogen, or both hydrogen and helium.  The
-resulting set of linear equations is solved using the parallel HYPRE
-framework.  Full details on the \enzo\ implementation of this method
-can be found in \citet{ReynoldsHayesPaschosNorman2009}.
+hydrogen, or both hydrogen and helium.  The resulting set of linear
+equations is solved using the parallel HYPRE framework.  Full details
+on the \enzo\ implementation of this method can be found in
+\citet{ReynoldsHayesPaschosNorman2009}.
 

### 2013-06-25 tests-orbit.tex
Reformatting tex files so they're all standardized in terms of paragraph wrapping, etc. (no content changed, just formatting.)

@@
 duration of the simulation, with a zoomed box showing a small portion
-of the orbit to highlight deviations from perfect circularity.  Right panel: total specific energy
-(kinetic plus potential) for the test particle for the duration of the
-simulation.  The particle's mean energy (integrated over the length of
-the simulation) is represented by a horizontal dashed line, and +/-
-one standard deviation by horizontal dotted lines.  All quantities are
-in arbitrary units.}
+of the orbit to highlight deviations from perfect circularity.  Right
+panel: total specific energy (kinetic plus potential) for the test
+particle for the duration of the simulation.  The particle's mean
+energy (integrated over the length of the simulation) is represented
+by a horizontal dashed line, and +/- one standard deviation by
+horizontal dotted lines.  All quantities are in arbitrary units.}
 \label{fig.orbittest}
@@
 velocities such that they will orbit their common center of mass in
-the x-y plane.  The simulation is run for 200 orbits, writing
-out a dataset 
-approximately once per orbit (with the particle positions being
-written out several hundred times per orbit).
+the x-y plane.  The simulation is run for 200 orbits, writing out a
+dataset approximately once per orbit (with the particle positions
+being written out several hundred times per orbit).
 
@@
 test particle in the x-y plane (at z=$0.5$), plotted over the duration
-of the simulation.  The test particle
-maintains a path that has effectively unnoticeable deviation from
-circularity, and with no discernable drift in either net radius or
-orbit center over time.  The right panel displays the total specific
-energy of the test particle (defined as $\frac{1}{2} {\bf v}^2 +
-\phi$, where $\phi$ is the gravitational potential at the position of
-the test particle) as a function of time, plotted once per orbit.  The
-mean total specific energy is $-1.76797$ in arbitrary units, with a
-standard deviation of $0.004885$ ($0.276\%$ of the mean value), and
-with a maximum deviation from the mean of $0.546\%$ of the mean value.
-Given that the potential is calculated on $32^3$ cells, the two
-particles are separated by between 10 and 14 cells depending on their
-relative positioning in the grid, and the particles are integrated
-using a second-order leapfrog method, this level of accuracy is
-expected and acceptable.
+of the simulation.  The test particle maintains a path that has
+effectively unnoticeable deviation from circularity, and with no
+discernable drift in either net radius or orbit center over time.  The
+right panel displays the total specific energy of the test particle
+(defined as $\frac{1}{2} {\bf v}^2 + \phi$, where $\phi$ is the
+gravitational potential at the position of the test particle) as a
+function of time, plotted once per orbit.  The mean total specific
+energy is $-1.76797$ in arbitrary units, with a standard deviation of
+$0.004885$ ($0.276\%$ of the mean value), and with a maximum deviation
+from the mean of $0.546\%$ of the mean value.  Given that the
+potential is calculated on $32^3$ cells, the two particles are
+separated by between 10 and 14 cells depending on their relative
+positioning in the grid, and the particles are integrated using a
+second-order leapfrog method, this level of accuracy is expected and
+acceptable.

### 2013-06-25 tests-orszagtang.tex
Reformatting tex files so they're all standardized in terms of paragraph wrapping, etc. (no content changed, just formatting.)

@@
 \label{sec.tests.mhd}
-Figure~\ref{fig.orszag} shows the Orszag-Tang vortex problem \citep{Orszag79}.  The left panel shows
-the result using \enzo's constrained transport MHD method, while the right panel shows the result using
-\enzo's implementation of Dedner MHD.
-The Orszag-Tang vortex test is a classig MHD test problem, and shows that significant small scale structure can be generated in MHD
-from large scale initial perturbations.  It is often used to compare the
-effective resolution of different MHD schemes.  The test begins with uniform
-density, $\rho_0=25/(36\pi)$, and pressure, $P_0=5/(12\pi)$ (as with
-other tests in this section, in the absence of gravity or chemistry/cooling we use
-dimensionless units).  There is a
-single rotational mode in the velocity, and two in the magnetic field:
-${\bf v}_0 = (-\sin(2\pi y) \hat{x},\, \sin(2\pi x) \hat{y})$,
-${\bf B}_0 = (-\sin(2\pi y) \hat{x},\, \sin(4\pi x) \hat{y})$.
-The simulation is evolved to $t=0.48$.  One can see that the structures
-are accurately represented as compared to, for example,
-\citet{Toth00}, and that the resolution of shocks is comparable in
-both methods.
+
+Figure~\ref{fig.orszag} shows the Orszag-Tang vortex problem
+\citep{Orszag79}.  The left panel shows the result using \enzo's
+constrained transport MHD method, while the right panel shows the
+result using \enzo's implementation of Dedner MHD.  The Orszag-Tang
+vortex test is a classig MHD test problem, and shows that significant
+small scale structure can be generated in MHD from large scale initial
+perturbations.  It is often used to compare the effective resolution
+of different MHD schemes.  The test begins with uniform density,
+$\rho_0=25/(36\pi)$, and pressure, $P_0=5/(12\pi)$ (as with other
+tests in this section, in the absence of gravity or chemistry/cooling
+we use dimensionless units).  There is a single rotational mode in the
+velocity, and two in the magnetic field: ${\bf v}_0 = (-\sin(2\pi y)
+\hat{x},\, \sin(2\pi x) \hat{y})$, ${\bf B}_0 = (-\sin(2\pi y)
+\hat{x},\, \sin(4\pi x) \hat{y})$.  The simulation is evolved to
+$t=0.48$.  One can see that the structures are accurately represented
+as compared to, for example, \citet{Toth00}, and that the resolution
+of shocks is comparable in both methods.
 

### 2013-06-25 tests-selfsiminfall.tex
self similar infall test section edited.  also a fix for a dumb \cite error added in the homogeneous radiation background section.
Reformatting tex files so they're all standardized in terms of paragraph wrapping, etc. (no content changed, just formatting.)

@@
 This test problem is based on the self-similar solutions to the
-cosmological spherical infall problem found
-by \citet{Bertschinger1985}.  It features a small density perturbation
-in a homogeneous $\Omega = 1$ universe and is a strong test of the
+cosmological spherical infall problem found by
+\citet{Bertschinger1985}.  It features a small density perturbation in
+a homogeneous $\Omega = 1$ universe and is a strong test of the
 cosmological evolution, gravity, and hydrodynamic portions of the
@@
 For the initial conditions, we adopt a $32^3$ top grid with a single,
-initial sub grid (covering $1/8^3$ of the domain) with a refinement
+initial subgrid (covering $1/8^3$ of the domain) with a refinement
 factor of 2.  An overdensity $\delta = 40$ is placed in a single cell

### 2013-06-25 tests-sodshocktube.tex
Reformatting tex files so they're all standardized in terms of paragraph wrapping, etc. (no content changed, just formatting.)

@@
 
-We begin the paper test suite with the classic one-dimensional Sod shock tube problem
-\citep{Sod78}, which provides a good test of a hydrodynamical solver's
-ability to resolve a clean Riemann problem with clear separation
-between the three resultant waves.  These waves consist of a rarefaction fan, a
-contact discontinuity, and a moderate shock.  The initial state is 
-$\rho_{\rm L}, P_{\rm L} = 1.0, 1.0$ on the left of the boundary at 
-$x=0.5$ and $\rho_{\rm R}, P_{\rm R} = 0.125, 0.1$ on the right.
-All velocities are initially zero.  In
+We begin the paper test suite with the classic one-dimensional Sod
+shock tube problem \citep{Sod78}, which provides a good test of a
+hydrodynamical solver's ability to resolve a clean Riemann problem
+with clear separation between the three resultant waves.  These waves
+consist of a rarefaction fan, a contact discontinuity, and a moderate
+shock.  The initial state is $\rho_{\rm L}, P_{\rm L} = 1.0, 1.0$ on
+the left of the boundary at $x=0.5$ and $\rho_{\rm R}, P_{\rm R} =
+0.125, 0.1$ on the right.  All velocities are initially zero.  In
 Figure~\ref{fig.sodshocktube}, we show the density solution at the
-final time ($t=0.25$) for three of our hydro solvers -- the spatially third-order
-PPM, as well as the two second-order \zeus\ and MUSCL schemes.  We use
-100 cells across the domain, which is a relatively standard choice in
-code method papers, and show solutions both with and without adaptive
-mesh refinement.  Without AMR (top row in Figure \ref{fig.sodshocktube}), it is
-clear that the PPM scheme produces by far the cleanest solution with
-all wave families crisply reproduced (in particular, the contact
-discontinuity and shock).  \zeus\ and MUSCL produce similar results,
-with MUSCL doing a slightly better job on the rarefaction fan.  In Figure
-\ref{fig.sodshocktube}, we also show the integrated absolute deviation from 
-the exact solution, $||E_1|| = \sum_i \Delta x_i |F(x_i)-{\rm Exact}(x_i)|$.
-%, as well as the maximum deviation,
-%$||E_\infty|| = \mathrm{max_i}\left\{ \Delta x_i|F(x_i)-{\rm Exact}(x_i)| \right\}$.  
-These numbers confirm the qualitative differences noted
-previously.
+final time ($t=0.25$) for three of our hydro solvers -- the spatially
+third-order PPM, as well as the two second-order \zeus\ and MUSCL
+schemes.  We use 100 cells across the domain, which is a relatively
+standard choice in code method papers, and show solutions both with
+and without adaptive mesh refinement.  Without AMR (top row in Figure
+\ref{fig.sodshocktube}), it is clear that the PPM scheme produces by
+far the cleanest solution with all wave families crisply reproduced
+(in particular, the contact discontinuity and shock).  \zeus\ and
+MUSCL produce similar results, with MUSCL doing a slightly better job
+on the rarefaction fan.  In Figure \ref{fig.sodshocktube}, we also
+show the integrated absolute deviation from the exact solution,
+$||E_1|| = \sum_i \Delta x_i |F(x_i)-{\rm Exact}(x_i)|$.  These
+numbers confirm the qualitative differences noted previously.
 
@@
 than 10\% in the density.  This refinement criterion results in only
-the refinement of strong gradients, and
-does not include the rarefaction fan at late times.  The results are
-shown in the bottom row of Figure~\ref{fig.sodshocktube}.  Using AMR,
-the results are much better for all three methods, with much sharper
-shocks and contact discontinuities and even a better representation of
-the rarefaction wave, which is only refined beyond the root grid at
-early times.  Although the results are improved for all methods, PPM
-still produces the best result, as is shown clearly by the computed
-error norms (displayed again in each of the individual panels).
+the refinement of strong gradients, and does not include the
+rarefaction fan at late times.  The results are shown in the bottom
+row of Figure~\ref{fig.sodshocktube}.  Using AMR, the results are much
+better for all three methods, with much sharper shocks and contact
+discontinuities and even a better representation of the rarefaction
+wave, which is only refined beyond the root grid at early times.
+Although the results are improved for all methods, PPM still produces
+the best result, as is shown clearly by the computed error norms
+(displayed again in each of the individual panels).
 
@@
 the root level and the results are shown at $t=0.25$.  All cells are
-plotted and color-coded by level with purple 
-indicating level 0, green level 1 and red level 2 (at the time shown, only small region
+plotted and color-coded by level with purple indicating level 0, green
+level 1 and red level 2 (at the time shown, only small region
 surrounding the contact discontinuity and the shock are refined).  In
-each panel, we show the analytic solution as a solid line and the $E_1$
-error norm in the upper right.}
+each panel, we show the analytic solution as a solid line and the
+$E_1$ error norm in the upper right.}
 \label{fig.sodshocktube}

### 2013-06-25 tests-vandv.tex
Reformatting tex files so they're all standardized in terms of paragraph wrapping, etc. (no content changed, just formatting.)

@@
 \enzo\ is capable of simulating a large variety of problems types,
-with all but a few of these types requiring only a parameter file as an
-input.  The most notable exception is the cosmology simulation, which takes
-as input initial conditions created by other codes.  The suite of test
-problems spans a wide range in complexity.  At one end of this spectrum
-are simple problems that utilize only a single component of \enzo\ and
-for which analytic solutions exist for comparison with the simulation
-results.  At the opposite end are problems that exercise a large
-portion of \enzo's machinery in concert.  Together, the available
-problem types fully cover all of \enzo's functionality.  This enables
-them to serve as a vehicle for verifying that the code's behavior
-remains stable over time, on new computing platforms, and after modification of the codebase.
+with all but a few of these types requiring only a parameter file as
+an input.  The most notable exception is the cosmology simulation,
+which takes as input initial conditions created by other codes.  The
+suite of test problems spans a wide range in complexity.  At one end
+of this spectrum are simple problems that utilize only a single
+component of \enzo\ and for which analytic solutions exist for
+comparison with the simulation results.  At the opposite end are
+problems that exercise a large portion of \enzo's machinery in
+concert.  Together, the available problem types fully cover all of
+\enzo's functionality.  This enables them to serve as a vehicle for
+verifying that the code's behavior remains stable over time, on new
+computing platforms, and after modification of the codebase.
 
@@
 against results produced by any other version of the code.  Within the
-\enzo\ source distribution, the test problem parameter files are stored
-in a nested directory structure organized according to the primary
-functionality tested (e.g., hydrodynamics, gravity, cooling, etc.).
-Each parameter file is accompanied by a text file containing various
-descriptive keywords, such as the machinery tested, the
+\enzo\ source distribution, the test problem parameter files are
+stored in a nested directory structure organized according to the
+primary functionality tested (e.g., hydrodynamics, gravity, cooling,
+etc.).  Each parameter file is accompanied by a text file containing
+various descriptive keywords, such as the machinery tested, the
 dimensionality, and the approximate run time.  A test runner script is
@@
 are run in succession using either the \enzo\ executable contained
-within that distribution or an external executable built from another \enzo\
-version and specified by the user.  This allows for any version of the
-code to be tested with an identical set of test problems.  After
-running the test problem simulations, the test runner then performs a
-series of basic analysis tasks using the \texttt{yt} analysis toolkit
-\citep{2011ApJS..192....9T, 2011arXiv1112.4482T}.  The default
-analysis performed on all test problems includes calculation of
-various statistics (such as extrema, mean, and variance) on the fields
-present in the output data.  Custom analysis provided by scripts that
-accompany the test problem parameter files is run for special cases,
-such as when an analytical solution exists that can be compared
-against the simulation data.  After the analysis is performed, the
-results are compared against a set of gold standard results that are
-maintained on a website and downloaded on the fly by the test runner
-script.  Alternately, results from any version of \enzo\ can be stored
-locally and compared to any other version of the code.  In
-Section~\ref{sec.tests.problems}, we describe some of the key test
-problems that are used to verify proper behavior.  All of these test
-problems, as well as the scripts to generate the figures from them,
-are available at \url{http://bitbucket.org/enzo/enzo-method-paper},
-the Bitbucket repository for the \enzo\ method paper.  The
-\texttt{test\_problems} subdirectory in this repository contains all of
-the files necessary to regenerate all of the figures in
-Section~\ref{sec.tests.problems}.  We note for completeness that 
-the figures in Section~\ref{sec.tests.problems} were generated using \texttt{yt} version 2.5.3 and the
-\url{https://bitbucket.org/enzo/enzo-dev} repository of \enzo\ with
-changeset \texttt{5d8c412}, which corresponds to \enzo\ 2.3.
+within that distribution or an external executable built from another
+\enzo\ version and specified by the user.  This allows for any version
+of the code to be tested with an identical set of test problems.
+After running the test problem simulations, the test runner then
+performs a series of basic analysis tasks using the \texttt{yt}
+analysis toolkit \citep{2011ApJS..192....9T, 2011arXiv1112.4482T}.
+The default analysis performed on all test problems includes
+calculation of various statistics (such as extrema, mean, and
+variance) on the fields present in the output data.  Custom analysis
+provided by scripts that accompany the test problem parameter files is
+run for special cases, such as when an analytical solution exists that
+can be compared against the simulation data.  After the analysis is
+performed, the results are compared against a set of gold standard
+results that are maintained on a website and downloaded on the fly by
+the test runner script.  Alternately, results from any version of
+\enzo\ can be stored locally and compared to any other version of the
+code.  In Section~\ref{sec.tests.problems}, we describe some of the
+key test problems that are used to verify proper behavior.  All of
+these test problems, as well as the scripts to generate the figures
+from them, are available at
+\url{http://bitbucket.org/enzo/enzo-method-paper}, the Bitbucket
+repository for the \enzo\ method paper.  The \texttt{test\_problems}
+subdirectory in this repository contains all of the files necessary to
+regenerate all of the figures in Section~\ref{sec.tests.problems}.  We
+note for completeness that the figures in
+Section~\ref{sec.tests.problems} were generated using \texttt{yt}
+version 2.5.3 and the \url{https://bitbucket.org/enzo/enzo-dev}
+repository of \enzo\ with changeset \texttt{5d8c412}, which
+corresponds to \enzo\ 2.3.
 
@@
 comparisons with other astrophysical codes used for self-gravitating
-fluid dynamics.  In general, \enzo\ behaves in a manner similar to other
-grid-based (and particularly AMR-based) codes, as we will summarize below.
+fluid dynamics.  In general, \enzo\ behaves in a manner similar to
+other grid-based (and particularly AMR-based) codes, as we will
+summarize below.
 
@@
 \citep{2008CS&D....1a5003H}, as well as several direct comparisons
-between \enzo\ and the GADGET SPH code in a variety of
-set-ups, including N-body and adiabatic hydrodynamics
+between \enzo\ and the GADGET SPH code in a variety of set-ups,
+including N-body and adiabatic hydrodynamics
 \citep{2005ApJS..160....1O,2005MNRAS.364..909V, 2011MNRAS.418..960V}
 and simulations of the Lyman-alpha forest \citep{2007MNRAS.374..196R}.
-Compared to the other codes involved in these projects, the \enzo\ code typically has a more difficult time resolving
-small-scale self-gravitating structures (for an equivalent dark matter
-particle count and nominally equivalent force resolution), but does
-comparably well as a tree-based code for larger structure, and is
-typically superior in terms of resolving fluid features due to its
-higher-order (and artificial viscosity-free) PPM hydrodynamics 
-solver.  When examining classical test cases such
-as the Santa Barbara project \citep{SantaBarbara}, \enzo\ forms galaxy
-clusters with very similar density, temperature, and entropy profiles
-to other grid-based codes that use Godunov-type hydro methods, which
-systematically differed from particle-based codes using SPH in this
-code comparison.  Similarly, in galaxy cluster simulations that look
-specifically at the properties of cosmological shocks
-\citep[e.g.][]{2011MNRAS.418..960V}, \enzo\ produces results that are
-similar to other high-order grid-based hydrodynamics codes, and a far
-superior performance of \enzo\ is observed (in terms of resolution of
-fluid features and shock detection) in low-density regions when
-compared to a particle-based code.  In tests of the Lyman-alpha forest
-that include radiative cooling and a uniform metagalactic ultraviolet
-background, \enzo\ and GADGET provide results on metrics such as the
-matter power spectrum that are comparable to within 5\%
-\citep{2007MNRAS.374..196R}.
+Compared to the other codes involved in these projects, the \enzo\
+code typically has a more difficult time resolving small-scale
+self-gravitating structures (for an equivalent dark matter particle
+count and nominally equivalent force resolution), but does comparably
+well as a tree-based code for larger structure, and is typically
+superior in terms of resolving fluid features due to its higher-order
+(and artificial viscosity-free) PPM hydrodynamics solver.  When
+examining classical test cases such as the Santa Barbara project
+\citep{SantaBarbara}, \enzo\ forms galaxy clusters with very similar
+density, temperature, and entropy profiles to other grid-based codes
+that use Godunov-type hydro methods, which systematically differed
+from particle-based codes using SPH in this code comparison.
+Similarly, in galaxy cluster simulations that look specifically at the
+properties of cosmological shocks \citep[e.g.][]{2011MNRAS.418..960V},
+\enzo\ produces results that are similar to other high-order
+grid-based hydrodynamics codes, and a far superior performance of
+\enzo\ is observed (in terms of resolution of fluid features and shock
+detection) in low-density regions when compared to a particle-based
+code.  In tests of the Lyman-alpha forest that include radiative
+cooling and a uniform metagalactic ultraviolet background, \enzo\ and
+GADGET provide results on metrics such as the matter power spectrum
+that are comparable to within 5\% \citep{2007MNRAS.374..196R}.
 
@@
 hydrodynamics solvers and fluid behavior.  In particular, the work of
-\citet{2007MNRAS.380..963A} and \citet{Tasker2008} perform
-direct comparisons between several grid- and particle-based
-codes for a variety of fluid-centric test problems (including shocked gas clouds,
+\citet{2007MNRAS.380..963A} and \citet{Tasker2008} perform direct
+comparisons between several grid- and particle-based codes for a
+variety of fluid-centric test problems (including shocked gas clouds,
 self-gravitating, translating clouds, and Sedov-Taylor explosions),
@@
 provide useful information on the sort of practical challenges that a
-user of an
-AMR code such as \enzo\ may experience.  More specific comparisons, including one
-testing the linear and nonlinear growth of the
-Kelvin-Helmholz instability \citep{2012ApJS..201...18M}, as well as a
-comparison that more
-broadly examines Galilean invariance in grid-based codes
-\citep{2010MNRAS.401.2463R}, show that \enzo, and in particular its
-implementation of the PPM hydrodynamics solver,
+user of an AMR code such as \enzo\ may experience.  More specific
+comparisons, including one testing the linear and nonlinear growth of
+the Kelvin-Helmholz instability \citep{2012ApJS..201...18M}, as well
+as a comparison that more broadly examines Galilean invariance in
+grid-based codes \citep{2010MNRAS.401.2463R}, show that \enzo, and in
+particular its implementation of the PPM hydrodynamics solver,
 converge to the correct solution as expected, and generally provide
 less diffusive solutions than lower-order codes, including those that
-use artificial viscosity.  Finally, there have been two code comparison
-projects that focus on turbulence simulations.  The first studied the
-behavior of decaying isothermal supersonic turbulence
+use artificial viscosity.  Finally, there have been two code
+comparison projects that focus on turbulence simulations.  The first
+studied the behavior of decaying isothermal supersonic turbulence
 \citep{2009A&A...508..541K}, and the second examined supersonic
@@
 hydrodynamics and the latter both the constrained transport MHD
-implementation of \citet{Collins10} and the Dedner method of \citet{WangAbelZhang08}.  In both cases, \enzo\ performed similarly to other
-grid-based codes that use Godunov-based fluid solvers, and typically
-had better effective resolution than particle-based codes when using the
-same number of particles as the number of grid cells in the \enzo\ simulation.
+implementation of \citet{Collins10} and the Dedner method of
+\citet{WangAbelZhang08}.  In both cases, \enzo\ performed similarly to
+other grid-based codes that use Godunov-based fluid solvers, and
+typically had better effective resolution than particle-based codes
+when using the same number of particles as the number of grid cells in
+the \enzo\ simulation.
 
@@
 fluid flow.  The flux-limited diffusion radiation transport scheme was
-measured against several test problems by \citet{IlievEtAl2009},
-which involved tests with and without analytical solutions.  \enzo\ produced
-results similar to both the analytical solutions and results obtained by 
-other codes.  We note, however, that there were minor differences throughout
-the comparison between various codes, and the majority of the codes differed
-in at least a subset of the tests.
+measured against several test problems by \citet{IlievEtAl2009}, which
+involved tests with and without analytical solutions.  \enzo\ produced
+results similar to both the analytical solutions and results obtained
+by other codes.  We note, however, that there were minor differences
+throughout the comparison between various codes, and the majority of
+the codes differed in at least a subset of the tests.
 \citet{2011ApJ...726...55T} show the result of varying reaction rate
 coefficients for the formation of molecular hydrogen via the 3-body
-process in both the \enzo\ and GADGET-2 codes, observing similar trends between
-the two codes. However, at nominally equivalent resolution
-where the particle and cell gas masses are comparable, 
-\enzo\ simulations typically displayed a
-substantially higher level of gas structure.  This is unsurprising due
-to \enzo's higher-order hydro solver.  Finally,
-\citet{2012ApJ...744...52P} show the results of comparing \enzo\ in
-its non-AMR mode to a smoothed-particle hydrodynamics code, SNSPH, in
-the context of common-envelope binary stellar interactions.  The
-authors show that the codes display reasonable convergence properties
-as a function of simulation resolution, and also agree quite well with
-each other. However, the observed mass-loss rates do not agree
-particularly well with observations.
+process in both the \enzo\ and GADGET-2 codes, observing similar
+trends between the two codes. However, at nominally equivalent
+resolution where the particle and cell gas masses are comparable,
+\enzo\ simulations typically displayed a substantially higher level of
+gas structure.  This is unsurprising due to \enzo's higher-order hydro
+solver.  Finally, \citet{2012ApJ...744...52P} show the results of
+comparing \enzo\ in its non-AMR mode to a smoothed-particle
+hydrodynamics code, SNSPH, in the context of common-envelope binary
+stellar interactions.  The authors show that the codes display
+reasonable convergence properties as a function of simulation
+resolution, and also agree quite well with each other. However, the
+observed mass-loss rates do not agree particularly well with
+observations.

### 2013-06-27 gpu_plot/gpu_plot.py
Small tweaks for the gpu scaling plot.
This makes it so the fonts match the rest of the figures.

@@ -1,3 +1,6 @@
+import matplotlib as mpl
+mpl.rcParams['font.family'] = 'STIXGeneral'
+mpl.rcParams['figure.figsize'] = 4, 3
 import matplotlib.pyplot as plt
 
 X = [1.0,2.0,4.0,8.0]
@@ -18,9 +21,10 @@ plt.plot(X,cpunode,color='blue',linewidth=2.0,marker='o')
 plt.plot(X,gpunode,color='red',linewidth=2.0,marker='o')
 plt.loglog()
 plt.axis([0,10,1e6,5e7])
+plt.xlim(0.8,10.2)
 plt.xlabel('N$_{node}$')
 plt.ylabel('Cells/s/node')
-plt.text(7.0, 2.0e7, 'CPU', size=15, color='blue')
-plt.text(7.0, 3.0e7, 'GPU', size=15, color='red')
+plt.text(7.0, 2.0e7, 'CPU', size=12, color='blue')
+plt.text(7.0, 3.0e7, 'GPU', size=12, color='red')
 plt.savefig('gpu_scaling.eps')
 

### 2013-07-18 numerical-radiation-homogeneous.tex
self similar infall test section edited.  also a fix for a dumb \cite error added in the homogeneous radiation background section.
Many small changes to all files -- GB final edits
Switched rates for He and He+ photoionization in Table 4 which had been put in the wrong rows.  Current ordering is odd, but matches codes.

@@
 \end{equation}
-
 where the constant coefficients $\alpha$, $\beta$, $\gamma$, $z_0$ and
@@
 H ionization & $6.7 \times 10^{-13}$ & 0.43 & 1/1.95 & 0 & 2.3 & 0 \\
-He ionization & $6.3 \times 10^{-15}$ & 0.51 & 1/2.35 & 0 & 2.3 & 0 \\
-He$^+$ ionization & $3.2 \times 10^{-13}$ & 0.50 & 1/2.00 & 0 & 2.3 & 0 \\
+He$^+$ ionization & $6.3 \times 10^{-15}$ & 0.51 & 1/2.35 & 0 & 2.3 & 0 \\
+He ionization & $3.2 \times 10^{-13}$ & 0.50 & 1/2.00 & 0 & 2.3 & 0 \\
 H heating & $4.7 \times 10^{-24}$ & 0.43 & 1/1.95 & 0 & 2.3 & 0 \\
@@
 H ionization & $5.6 \times 10^{-13}$ & 0.43 & 1/1.95 & 0 & 2.3 & 0 \\
-He ionization & $3.2 \times 10^{-15}$ & 0.30 & 1/2.60 & 0 & 2.3 & 0 \\
-He$^+$ ionization & $4.8 \times 10^{-13}$ & 0.43 & 1/1.95 & 0 & 2.3 & 0 \\
+He$^+$ ionization & $3.2 \times 10^{-15}$ & 0.30 & 1/2.60 & 0 & 2.3 & 0 \\
+He ionization & $4.8 \times 10^{-13}$ & 0.43 & 1/1.95 & 0 & 2.3 & 0 \\
 H heating & $3.9 \times 10^{-24}$ & 0.43 & 1/1.95 & 0 & 2.3 & 0 \\
@@
 \tableline
-\multicolumn{7}{c}{Radiation Type 3 modified \citept2012ApJ...746..125H}} \\
+\multicolumn{7}{c}{Radiation Type 3 modified \citet{2012ApJ...746..125H}} \\
 \tableline
 H ionization &        $1.04 \times 10^{-12}$ & 0.231 & -0.6818 & 0.1646 & 1.855 & 0.3097 \\
-He ionization &       $1.84 \times 10^{-14}$ & -1.038 & -1.1640 & 0.1940 & 1.973 & -0.6561 \\
-He$^+$ ionization & $5.79 \times 10^{-13}$ & 0.278 & -0.8260 & 0.1730 & 1.973 & 0.2880 \\
+He$^+$ ionization &       $1.84 \times 10^{-14}$ & -1.038 & -1.1640 & 0.1940 & 1.973 & -0.6561 \\
+He ionization & $5.79 \times 10^{-13}$ & 0.278 & -0.8260 & 0.1730 & 1.973 & 0.2880 \\
 H heating &            $8.86 \times 10^{-25}$ & -0.0290 & -0.7055 & 0.1884 & 2.003 & 0.2888 \\

### 2013-08-07 table_of_variables_and_constants.tex
Correcting another error associated with the Jeans refinement criterion.  The equation doesn't include a factor of mu or gamma which are in the code.

@@
   $\tau_*$ & characteristic star formation timescale (for Schmidt-Law method) \\[3pt]
+  $\mu$ & mean molecular weight
 

### 2013-10-22 refreport/response.tex
added first draft of referee's report, which does NOT address the test problems but DOES address the length/split issues.

\documentclass[11pt]{article}

\usepackage{fullpage}
\usepackage{natbib}
\usepackage{hyperref, breakurl, color}
\newcommand{\red}[1]{\textcolor{red}{#1}}
\newcommand{\blue}[1]{\textcolor{blue}{#1}}

\newcommand{\code}[1]{\textsf{#1}}
\newcommand{\enzo}{\code{Enzo}}

\def\apj{ApJ}
\def\apjl{ApJL}
\def\apjs{ApJS}
\def\mnras{MNRAS}
\def\aj{Ast.J.}
\def\nat{Nature}
\def\na{New.A.}
\newcommand{\aap}{A\&A}

\begin{document}

\begin{center} 
\bfseries{
\begin{large}
Response to referee report for manuscript  ApJ/ApJS92163
\end{large}
}
\end{center}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

This is the response to the referee's report for ApJ/ApJS manuscript
\#92163, ``Enzo: An Adaptive Mesh Refinement Code for Astrophysics,''
by the Enzo collaboration (Bryan et al.).  We thank the referee for
their feedback, and address each comment separately below.

\section{Test problems}

\red{(Add response regarding test problems here)}

\section{Paper length and breakdown}

Finally, we would like to address the referee's feedback regarding the
paper's length and current organization.  The referee's comment is:

\begin{verbatim} 
Given the paper's length as it currently stands, it seems appropriate
to split the paper into two. One paper should document the basic
algorithms (e.g. AMR, hydro, MHD and gravity implementations) along
with an appropriate range of tests based on standard tests from the
literature. The second paper should document the microphysics options
(chemistry, radiation, thermal conductivity, etc.).
\end{verbatim}

We respectfully disagree with the referee regarding this point --
splitting the paper in two does not make sense, will not shorten the
overall page count, and will not improve the clarity of presentation. One of the strengths
of the \enzo\ code is the breadth of physics packages that are
available in the publicly-available version -- unlike, for example,
Gadget and Gadget-2 \citep{2001NewA....6...79S, 2005MNRAS.364.1105S},
where the only physics packages that are publicly available are dark
matter and adiabatic hydrodynamics.  Given this, it is logical that
the totality of the code's capabilities be documented in a single
place, for reference by future generations of computational scientists.

Regarding the paper's length, we argue that it is not excessively long
given the topics it covers, and thus does not need to be split into
multiple parts.  At present, the \enzo\ method paper is \red{60 pages
in ApJ format (BWO: we will need to update this before submission)}.
For comparison:

\begin{itemize}
\item The paper for the open-source FLASH code
\citep{2000ApJS..131..273F} is 62 pages long -- almost identical in
length to the \enzo\ method paper.  At the time of the FLASH method paper's
publication, it covered the full feature set of the FLASH code, which
is comparable to the breadth of topcs covered in the \enzo\ method
paper.

\item The AREPO method paper \citep{2010MNRAS.401..791S} is 60 pages
long, covers a narrower range of topics, and the code is
closed-source.

\item The paper for the open-source Athena code
\citep{2008ApJS..178..137S} is 41 pages long, but only documents the
hydrodynamics and MHD algorithms.  At the time of the code's release,
this paper documented the entire feature set of the publicly-available
code.

\item The paper for the open-source ZEUS-MP 2 code
\citep{2006ApJS..165..188H} is 41 pages long, but again covers a
substantially narrower range of physics modules.  The entirety of the
publicly-available code is documented in this paper.
\end{itemize}

The only \textit{comparable} open-source code that we are aware of
whose method paper is broken up into multiple segments is the series
of three ZEUS-2D papers \citep{1992ApJS...80..753S,
1992ApJS...80..791S, 1992ApJS...80..819S}.  We argue, however, that
this is not a useful comparison, as the three papers were published in
the same issue of the Astrophysical Journal in 1992, and were
published back-to-pack as consecutive articles in the same volume.  In
a time when the primary mode of consumption of articles was via paper
journals, this is effectively a single paper (with a length of
approximately 85 pages, no less).

We further note that splitting the paper up into two separate pieces
provides structural challenges, and will serve to increase the total
page count.  Structurally, Section 2 (the ``overview'' section) would
need to be split into two pieces, and the component that discusses the
microphysics will be somewhat confusing on its own, as these modules
act as sink and source terms for the fluid equations that would in
principle be described in a separate paper.  It's not clear where
Section 10 (``Analysis''), Section 12 (``Parallel strategy and
performance''), and Section 13 (``Code development methodology'')
would logically go -- some of these sections apply to all physics
packages in the code (e.g., Section 12), and others to none in
particular (e.g., Section 13).  Similarly, Section 11.1 (``Verifying
and validating the Enzo code'') describes the \enzo\ test suite, which
is used for all aspects of code testing (including both the
fluid/gravity/AMR algorithms \textit{and} the microphysics), and also
for code comparisons that test many different aspects of the code,
often in combination.  In order for these sections to make sense, and
to address test problems that are split between two papers, several of
these sections would need to be either duplicated or require an
extensive amount of cross-referencing between papers Furthermore,
separate introductions, discussions, and conclusions will by necessity
be mutually referential and, in some sense, duplicates of each other.
Together, this will not serve to either decrease page count or
increase clarity for the reader.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\newpage

\bibliographystyle{apj}
\bibliography{response}

\end{document}

### 2013-10-22 tests-ionbalance.tex
Adding ionization balance test.

\subsubsection{Ion balance test}
\label{sec.tests.ionbalance}

The one-zone collapse problem, described in \S\ref{sec.tests.1-zone},
is a good test of the chemistry solver at low temperatures where it is
difficult to calculate an analytical solution for the equilibrium
fraction of each species.  However, for a gas of primordial composition
at high temperatures, a simple analytical solution exists for the
ionization balance of H and He species.  The equations balancing
ionization and recombination for all ions of H and He result in six
equations with six unknowns \citep[e.g.,][]{1996ApJS..105...19K}, are
given by
\begin{equation} \label{eq:eqnHI}
n_{\rm HI} = \frac{n_{\rm H} \alpha_{\rm HII}}{\alpha_{\rm HII}
+ \Gamma_{\rm eHI} + \Gamma_{\gamma \rm HI} / n_{\rm e}},
\end{equation}
\begin{equation}
n_{\rm HII} = n_{\rm H} - n_{\rm HI},
\end{equation}
\begin{equation}
n_{\rm HeII} = \frac{n_{\rm He}}{1 + \frac{\alpha_{\rm HeII}
+ \alpha_{\rm d}}{\Gamma_{\rm eHeI} + \Gamma_{\gamma \rm HeI} / n_{\rm
e}} + \frac{\Gamma_{\rm eHeII} + \Gamma_{\gamma \rm HeII} / n_{\rm
e}}{\alpha_{\rm HeIII}}},
\end{equation}
\begin{equation}
n_{\rm HeI} = n_{\rm HeII} \frac{\alpha_{\rm HeII} + \alpha_{\rm
d}}{\Gamma_{\rm eHeI} + \Gamma_{\gamma \rm HeI} / n_{\rm e}},
\end{equation}
\begin{equation}
n_{\rm HeIII} = n_{\rm HeII} \frac{\Gamma_{\rm eHeII}
+ \Gamma_{\rm \gamma \rm HeII} / n_{\rm e}}{\alpha_{\rm HeIII}},
\end{equation}
\begin{equation} \label{eq:eqne}
n_{\rm e} = n_{\rm HII} + n_{\rm HeII} + 2 n_{\rm HeIII},
\end{equation}
where $n_{\rm H}$ and $n_{\rm He}$ are the total number densities of
all H and He ions, $\alpha_{\rm i}$ is the recombination rate of
species i, and $\Gamma_{ij}$ is the ionization rate of species j by
i, where i is either an electron for collisional ionization or a
photon for photo-ionization.

In this test, we initialize a one-dimension grid at a constant number
density of 1 cm$^{-3}$ with smoothly varying temperature from 10$^{4}$
K to 10$^{9}$ K.  The gas is initially fully ionized and we then
iterate the chemistry equations without allowing the gas to cool until
convergence is reached.  In Figure \ref{fig.eq_test}, we show the
equilibrium ionization fractions for H and He ions and the resulting
cooling rate.  We assume no photo-ionization for this test.  
The results from this test are shown by the solid lines 
with the analytic solution decribed by 
Equations \ref{eq:eqnHI}--\ref{eq:eqne} using the rate equations 
in \enzo, referenced in \s\ref{sec.num.chemistry}, overplotted with
the dashed lines and an analytical solution using the rates 
of \citet{1992ApJS...78..341C}, shown by the dotted lines.  As can be
seen, the \enzo\ chemistry solver and corresponding analytical
solution are in perfect agreement.  The results largely agree with
those of \citet{1992ApJS...78..341C}, save for HI, HeI, and HeII at
high temperatures, where their ion fractions are quite low.  The
slightly higher abundance of HeI in \citet{1992ApJS...78..341C}
results in a higher cooling rate by about a factor of 2 near $T \sim
10^{5}$ K due to the enhanced collisional excitation cooling of HeI.

\begin{figure}
  \begin{center}
    \includegraphics[width=0.49\textwidth]{figures/eq_ionization.eps}
    \includegraphics[width=0.49\textwidth]{figures/eq_cooling.eps}
  \end{center}
  \caption{Results of the ionization balance test
    (\S\ref{sec.tests.ionbalance}) calculating the equilibrium ionization
    balance (left) and resulting cooling rate (right) of a primordial
    gas with no incident photo-ionizing radiation.  Solid lines show
    the results of iterating the \enzo\ chemistry solver to
    convergence.  Dashed lines (directly underneath the solid lines)
    show the analytical solution using the same rates.  Dotted 
    lines show the analytical solution using the rates
    of \citet{1992ApJS...78..341C}.}
  \label{fig.eq_test}
\end{figure}

### 2013-10-22 acknowledgments.tex

@@
 AST-0507717 (MLN), 
-AST-0507768 (AK),
+AST-0507768 (AGK),
 AST-0529734 (TA),
-AST-0607675 (AK),
+AST-0607675 (AGK),
 AST-0702923 (EJH),
@@
 AST-0807215 (JB),
-AST-0808184 (MLN, AK),
+AST-0808184 (MLN, AGK),
 AST-0808398 (TA),
-AST-0908740 (AK, DCC),
+AST-0908740 (AGK, DCC),
 AST-0908819 (BWO), 
@@
 AST-1008134 (GB),
+AST-1109570 (AGK), 
 AST-1009802 (JSO), 

### 2013-10-22 numerical-gravity.tex
Removed commented-out text from lots of files - basically vestigial junk.

@@
-% EJT: small changes only.
-% dark matter particles -> massive (dark matter and star) particles
-
-% Changed:  ``Particles from subgrids are also added in this way'' to ``Particles on subgrids within the grid's volume are also added to its gravitating field using the same method.'' I expanded because I missed the point and had to clarify with Brian when I read the text.
-But if it's too long, please just cut it!
-
-% removed blank reference [REF]
-
 \subsection{Gravity}
@@
 
-Solving for the accelerations of the cells and particles on the grid due to self-gravity involves three steps: (i) computing the total gravitating mass, (ii) solving for the gravitational potential field with the appropriate boundary conditions, and (iii) differencing the potential to get the acceleration, and, if necessary, interpolating the acceleration back to the particles.
-These steps are described in detail below.
+Solving for the accelerations of the cells and particles on the grid
+due to self-gravity involves three steps: (i) computing the total
+gravitating mass, (ii) solving for the gravitational potential field
+with the appropriate boundary conditions, and (iii) differencing the
+potential to get the acceleration, and, if necessary, interpolating
+the acceleration back to the particles. These steps are described in
+detail below.
 
-First, the massive (dark matter and star) particles are distributed onto the grids using the second-order cloud-in-cell (CIC) interpolation technique \citep{Hockney88} to form a spatially-discretized density field $\rho_{\rm DM}$.
-During the CIC interpolation, particle positions are (temporarily) advanced by $0.5 v^n \Delta t$ so that we generate an estimate of the time-centered density field.
-Particles on subgrids within the grid's volume are also added to its gravitating field using the same method.
-In addition, since the gravitating field for a grid is defined beyond the grid edges (see below), massive particles from sibling grids and sub grids that lie within the entire gravitating field are used.
-This step can involve communication.
+First, the massive (dark matter and star) particles are distributed
+onto the grids using the second-order cloud-in-cell (CIC)
+interpolation technique \citep{Hockney88} to form a
+spatially-discretized density field $\rho_{\rm DM}$.  During the CIC
+interpolation, particle positions are (temporarily) advanced by $0.5
+v^n \Delta t$ so that we generate an estimate of the time-centered
+density field.  Particles on subgrids within the grid's volume are
+also added to its gravitating field using the same method. In
+addition, since the gravitating field for a grid is defined beyond the
+grid edges (see below), massive particles from sibling grids and sub
+grids that lie within the entire gravitating field are used.  This
+step can involve communication.
 
-Next, we add the baryonic grid densities in a similar fashion.
-In particular, we treat baryonic cells as virtual CIC particles that are are placed at the grid center but are advanced by $0.5 v^n \Delta t$ in order to approximately time-center the gravitating mass field.
-Cells that are covered by further-refined grids are treated in a similar way (i.e., we also use the subgrid cells as virtual CIC particles).
-This procedure results in a total gravitating mass field $\rho_{\rm total}^{n+1/2}$.
+Next, we add the baryonic grid densities in a similar fashion.  In
+particular, we treat baryonic cells as virtual CIC particles that are
+are placed at the grid center but are advanced by $0.5 v^n \Delta t$
+in order to approximately time-center the gravitating mass field.
+Cells that are covered by further-refined grids are treated in a
+similar way (i.e., we also use the subgrid cells as virtual CIC
+particles).  This procedure results in a total gravitating mass field
+$\rho_{\rm total}^{n+1/2}$.
 
-To compute the potential field from this gravitating mass field on the root grid, we use a a fast Fourier transform.
-For periodic boundary conditions, we can use either a simple Greens function kernel of $-k^{-2}$, or the finite difference equivalent \citep{Hockney88}:
+To compute the potential field from this gravitating mass field on the
+root grid, we use a a fast Fourier transform. For periodic boundary
+conditions, we can use either a simple Greens function kernel of
+$-k^{-2}$, or the finite difference equivalent \citep{Hockney88}:
+
 \begin{equation}
@@
 \end{equation}
-where $k^2 = k_x^2 + k_y^2 + k_z^2$ is the wavenumber in Fourier space and the potential is calculated in k-space as usual with $\tilde{\phi}(k) = G(k) \tilde{\rho}(k)$.
 
-For isolated boundary conditions, we use the James method \citep{James77}.
-In this case, the Greens function is generated in real-space so as to have the correct zero-padding properties and then transformed into the Fourier domain.
-In both cases, the potential is then transformed back into the real domain to get potential values at the cell centers.
-These are differenced with a two-point centered difference scheme to obtain accelerations at the cell centers (except if we are using the staggered \zeus-like solver, in which case the accelerations are computed at the cell faces to match the velocities).
-Particle accelerations are obtained using a (linear) CIC interpolation from the grid.
+where $k^2 = k_x^2 + k_y^2 + k_z^2$ is the wavenumber in Fourier space
+and the potential is calculated in k-space as usual with
+$\tilde{\phi}(k) = G(k) \tilde{\rho}(k)$.
 
-In order to calculate more accurate potentials on the subgrids, \enzo\ uses a similar but slightly different technique from the root grid.
-The generation of the total gravitating mass field is essentially identical, using CIC interpolation for both the particles and baryons, including subgrids as before.
-To compute the potential on subgrids, however, we use the standard seven-point (in three dimensions) second-order finite difference approximation to Poisson's equation.
-Boundary conditions are then interpolated from the potential values on the parent grid.
-We use either tri-linear interpolation or a natural second-order spline for this: both methods give similar results, but the default is the tri-linear interpolation, which empirically provides a resonable compromise between speed and accuracy.
-The potential equation on each subgrid is then solved with the given Dirichlet boundary conditions with a multigrid relaxation technique.
+For isolated boundary conditions, we use the James method
+\citep{James77}.  In this case, the Greens function is generated in
+real-space so as to have the correct zero-padding properties and then
+transformed into the Fourier domain.  In both cases, the potential is
+then transformed back into the real domain to get potential values at
+the cell centers.  These are differenced with a two-point centered
+difference scheme to obtain accelerations at the cell centers (except
+if we are using the staggered \zeus-like solver, in which case the
+accelerations are computed at the cell faces to match the velocities).
+Particle accelerations are obtained using a (linear) CIC interpolation
+from the grid.
+
+In order to calculate more accurate potentials on the subgrids, \enzo\
+uses a similar but slightly different technique from the root grid.
+The generation of the total gravitating mass field is essentially
+identical, using CIC interpolation for both the particles and baryons,
+including subgrids as before.  To compute the potential on subgrids,
+however, we use the standard seven-point (in three dimensions)
+second-order finite difference approximation to Poisson's equation.
+Boundary conditions are then interpolated from the potential values on
+the parent grid.  We use either tri-linear interpolation or a natural
+second-order spline for this: both methods give similar results, but
+the default is the tri-linear interpolation, which empirically
+provides a resonable compromise between speed and accuracy. The
+potential equation on each subgrid is then solved with the given
+Dirichlet boundary conditions with a multigrid relaxation technique.
 This is applied to each subgrid separately.
 
-The region immediately next to the boundary can contain unwanted oscillations \citep[e.g.,][]{Anninos94}, and so we use an expanded buffer zone around the grid, of size three parent grid boundary zones (so typically six refined zones for a refinement factor of 2).
-The density is computed in this region and the potential solved, but only the region that overlaps with the active region of the grid itself is used to calculate accelerations.
+The region immediately next to the boundary can contain unwanted
+oscillations \citep[e.g.,][]{Anninos94}, and so we use an expanded
+buffer zone around the grid, of size three parent grid boundary zones
+(so typically six refined zones for a refinement factor of 2).  The
+density is computed in this region and the potential solved, but only
+the region that overlaps with the active region of the grid itself is
+used to calculate accelerations.
 
-Simply interpolating the potential without feeding it back to higher levels leads to errors in the potential at more refined levels, due to the build-up of errors during the interpolation of coarse boundary values.
-In addition, neighboring subgrids are not guaranteed to generate the same potential values because of the lack of a coherent potential solve including the whole grid hierarchy.
-In an attempt to partially alleviate this problem, we allow for an iterative procedure across sibling grids, in which the potential values on the boundary of grids can be updated with the potential in `active' regions of neighboring subgrids.
-To prevent overshoot, we average the potential on the boundary and allow for (by default) 4 iterations, with the number of iterations determined by a parameter specified at runtime.
-This procedure can help in many cases, but does not necessarily produce a coherent solution across all grids and so does not completely solve the problem; we are working on a slower but more accurate method that does a multigrid solve across the whole grid (Reynolds \etal, in preparation).
+Simply interpolating the potential without feeding it back to higher
+levels leads to errors in the potential at more refined levels, due to
+the build-up of errors during the interpolation of coarse boundary
+values.  In addition, neighboring subgrids are not guaranteed to
+generate the same potential values because of the lack of a coherent
+potential solve including the whole grid hierarchy.  In an attempt to
+partially alleviate this problem, we allow for an iterative procedure
+across sibling grids, in which the potential values on the boundary of
+grids can be updated with the potential in `active' regions of
+neighboring subgrids.  To prevent overshoot, we average the potential
+on the boundary and allow for (by default) 4 iterations, with the
+number of iterations determined by a parameter specified at runtime.
+This procedure can help in many cases, but does not necessarily
+produce a coherent solution across all grids and so does not
+completely solve the problem; we are working on a slower but more
+accurate method that does a multigrid solve across the whole grid
+(Reynolds \etal, in preparation).
 
-At this point it is useful to emphasize that the effective force resolution of an adaptive particle-mesh calculation is approximately twice as coarse as the grid spacing at a given level of resolution.
+At this point it is useful to emphasize that the effective force
+resolution of an adaptive particle-mesh calculation is approximately
+twice as coarse as the grid spacing at a given level of resolution.

### 2013-10-22 numerical-mhd-ct.tex
Removed commented-out text from lots of files - basically vestigial junk.

@@
 The third solver we describe is an MHD method developed by
-\citet{Collins10}.  Since a full description and suite of test problems can be
-found in that reference, we only describe the method briefly here.
+\citet{Collins10}.  Since a full description and suite of test
+problems can be found in that reference, we only describe the method
+briefly here.
 
-The divergence of the magnetic field, \Divb, is
-identically zero in reality due to the fact that the evolution of the magnetic field is the
-curl of a vector, and the divergence of the curl of a vector is identically
-zero.
-%\begin{eqnarray}
-%\partial \Bvec/\partial t & =& \frac{1}{a} \curl \Evec\\
-%\Divb & =& 0.
-%\end{eqnarray}
-The Constrained Transport (CT) method \citep{Evans88, Balsara99} for
-magnetohydrodynamic (MHD) evolution employs this same vector property
-to evolve the magnetic field in a manner that preserves $\Divb=0$.  
-The electric field is computed using
+The divergence of the magnetic field, \Divb, is identically zero in
+reality due to the fact that the evolution of the magnetic field is
+the curl of a vector, and the divergence of the curl of a vector is
+identically zero.  The Constrained Transport (CT) method
+\citep{Evans88, Balsara99} for magnetohydrodynamic (MHD) evolution
+employs this same vector property to evolve the magnetic field in a
+manner that preserves $\Divb=0$.  The electric field is computed using
 the fluxes from the Riemann solver.  The curl of that electric field
@@
 method is that it preserves \Divb\ to machine precision.  The primary
-drawback is increased algorithm complexity.  Note also that since 
-only the update of the magnetic field is divergence free, any monopoles 
-created by other numerical sources (such as ill-chosen initial conditions) persist.
+drawback is increased algorithm complexity.  Note also that since only
+the update of the magnetic field is divergence free, any monopoles
+created by other numerical sources (such as ill-chosen initial
+conditions) persist.
 
-The base Godunov method, described in \citet{Li08a}, uses spatially and
-temporally second-order reconstruction (both MUSCL-Hancock and
+The base Godunov method, described in \citet{Li08a}, uses spatially
+and temporally second-order reconstruction (both MUSCL-Hancock and
 Piecewise Linear Method), and a selection of Riemann solvers including
-HLLC and HLLD \citep{Mignone07}, as described earlier.  
-The constrained transport methods
-are the first-order method described by \citet{Balsara99} and the
-second-order methods described in \citet{Gardiner05}.  The AMR
-machinery is described by \citet{Balsara01} and \citet{Collins10}.
+HLLC and HLLD \citep{Mignone07}, as described earlier.  The
+constrained transport methods are the first-order method described by
+\citet{Balsara99} and the second-order methods described in
+\citet{Gardiner05}.  The AMR machinery is described by
+\citet{Balsara01} and \citet{Collins10}.
 
@@
 Divergence-free AMR is somewhat more complex than the AMR employed
-elsewhere in \enzo.  First, the interpolation must be constrained to be
-divergence free.  Thus, all three face-centered field components are
-interpolated in concert.  Second, any magnetic field information from
-the previous timestep must be included in the interpolation, making the
-interpolation more complex than the simple parent-child relation used
-for other fields.  Third, the flux correction involves more possible
-grid relations than traditional AMR.  In order to circumvent this last
-complexity, the electric field is projected from fine grids to parent
-grids (rather than the magnetic field), and is then used to re-update
-the parent magnetic field.  This is described in detail in
-\citet{Balsara99} and \citet{Collins10}.
-
-%Cosmological MHD can be treated in a number of ways.  As described in 
-%the previous section, and equation~\ref{eq:induction} the MUSCL with Dedner
-%method treats the cosmological dilution of
-%$\vecB$ with a source term,  However, one can eliminate the right hand side of equation 
-%\ref{eq:induction} by defining
-%$\vecB^\prime = \vecB / \sqrt{a}.$
-%Then equation~\ref{eq:induction} becomes
-%$$ 
-%\frac{\partial \vecB^\prime}{\partial t} - \frac{1}{a}  \curl (\vecv \times
-%\vecB^\prime)  = 0, \label{eq:induction_cosmo_prime_nosource}
-%$$
-%and we can reuse each part of the non-cosmological algorithm (hyperbolic solver,
-%electric field creation, interpolation, projection, and flux correction) without
-%modification.  
-%GB: I removed the above because the rest of the paper now uses the definition
-% of B in MHD-CT.
+elsewhere in \enzo.  First, the interpolation must be constrained to
+be divergence free.  Thus, all three face-centered field components
+are interpolated in concert.  Second, any magnetic field information
+from the previous timestep must be included in the interpolation,
+making the interpolation more complex than the simple parent-child
+relation used for other fields.  Third, the flux correction involves
+more possible grid relations than traditional AMR.  In order to
+circumvent this last complexity, the electric field is projected from
+fine grids to parent grids (rather than the magnetic field), and is
+then used to re-update the parent magnetic field.  This is described
+in detail in \citet{Balsara99} and \citet{Collins10}.
 
 The dual energy formalism has also been incorporated in two possible
-ways -- one that uses internal
-energy, as described in Section \ref{sec.hydro.ppm}, and one that
-uses entropy \cite{TVD93, Collins10}.
-
-%The additional equations are
-%\begin{eqnarray}
-%\end{eqnarray}
-%\begin{eqnarray}
-%  \frac{\partial S}{\partial t} +\frac{1}{a} \div (S \vecv) & = &  -
-%  \frac{ 3 (\gamma - 1) \dot{a}}{a} S \\
-%  \frac{\partial \rho e}{\partial t} + \frac{1}{a} \div (\rho e \vecv) &
-%  = & -\frac{ 3(\gamma -1 ) \dot{a}}{a} + \frac{p}{a} \div \vecv,
-%\end{eqnarray}
-%where $e$ is the internal energy and $S=p/\rho^{\gamma-1}$ is the entropy.  If
-%the ratio of thermal to kinetic energy, $\eta$, is below a certain value, the
-%alternative energy is used.  As in \citet{Li08a}, we use $\eta=0.008$.
-
-%
-%\enzo's MHD-CT has been used in a variety of contexts, from galaxy
-%cluster evolution \citep{Xu11,Xu12}, synthetic observations of radio
-%halos \citep{Skillman13}, primordial star formation \citep{Xu08}, and
-%present-day star formation \citep{Collins11, Collins12a}.
-% GB NOTE: incorporate these elsewhere
+ways -- one that uses internal energy, as described in Section
+\ref{sec.hydro.ppm}, and one that uses entropy \cite{TVD93,
+Collins10}.

### 2013-10-22 numerical-radiation-fld.tex
Removed commented-out text from lots of files - basically vestigial junk.

@@
 
-%% With regards to the numerical method section, we're really looking for
-%% something that's no more than a page long: a few relevant equations,
-%% maybe a finite difference equation or two if you feel that it's
-%% relevant, but primarily a description of the algorithms used for the
-%% FLD solve and a brief explanation of how it's coupled to the
-%% chemistry, as well as references to the relevant method paper(s).
-%% Working in a reference or two to science papers is also great!
-
 In addition to the ray-tracing approach for radiation transport
-described in Section~\ref{sec.num.raytracing}, \enzo\ currently includes
-a field-based radiation transport solver for problems posed on uniform
-(i.e.~non-AMR, non-static mesh refinement) grids, which has been tuned for large-scale simulations
-involving many ionizing sources.  Detailed explanations of the model and
-solution approach may be found in
-\citet{NBHBROW2007}, \citet{ReynoldsHayesPaschosNorman2009},
-\citet{NRS2009}, and \citet{2013arXiv1306.0645N}, the salient features of which are reproduced here. 
-In addition, comparisons of this solver with other astrophysical
-radiation transport solvers may be found in \cite{IlievEtAl2009}.  
-\enzo's field-based radiation solver focuses on
+described in Section~\ref{sec.num.raytracing}, \enzo\ currently
+includes a field-based radiation transport solver for problems posed
+on uniform (i.e.~non-AMR, non-static mesh refinement) grids, which has
+been tuned for large-scale simulations involving many ionizing
+sources.  Detailed explanations of the model and solution approach may
+be found in \citet{NBHBROW2007},
+\citet{ReynoldsHayesPaschosNorman2009}, \citet{NRS2009}, and
+\citet{2013arXiv1306.0645N}, the salient features of which are
+reproduced here.  In addition, comparisons of this solver with other
+astrophysical radiation transport solvers may be found in
+\cite{IlievEtAl2009}.  \enzo's field-based radiation solver focuses on
 a flux-limited diffusion approximation for cosmological radiative
@@
 densities.
-%\begin{eqnarray}
-%  \label{eq:fld_radiation}
-%  \partial_t E_r + \frac1a \div \left(E_r \myvec{v}_b\right) &=& 
-%    \div\left(D\grad E_r\right) -
-%    \frac{\dot{a}}{a} E_r - c\kappa E_r + \eta, \\
-%  \label{eq:fld_heating}
-%  \partial_t e_c &=& -\frac{2\dot{a}}{a} e_c + G - \Lambda, \\
-%  \label{eq:fld_chemistry}
-%  \partial_t {\mathrm n}_i + 
-%    \frac1a \div\left({\mathrm n}_i \myvec{v}_b\right) &=& 
-%    \alpha_{i,j} {\mathrm n}_e {\mathrm n}_j - {\mathrm n}_i
-%    \Gamma_i^{ph}, \quad i=1,\ldots,N_s,
-%\end{eqnarray}
-%where $E_r$ is a grey radiation energy density, $e_c$ is the
-%internal energy correction due to photo-heating and chemical cooling,
-%and ${\mathrm n}_i$ correspond to the chemical number densities of
-%HI, HII, HeI, HeII and HeIII, and ${\mathrm n}_e$ is the electron
-%number density.  Here, we define the grey radiation energy density
-%through first assuming a fixed frequency spectrum, i.e.
-%$E_{\nu}(\nu,\myvec{x},t) = \tilde{E}_r(\myvec{x},t) \chi(\nu)$, and then
-%defining the integrated quantity
-%\begin{equation}
-%\label{eq:grey_radiation_energy}
-%   E_r(\myvec{x},t) \equiv \int_{\nu_0}^{\infty}
-%   E_{\nu}(\nu,\myvec{x},t)\,\mathrm d\nu \  = \ 
-%   \tilde{E}_r(\myvec{x},t) \int_{\nu_0}^{\infty} \chi(\nu)\,\mathrm d\nu.
-%\end{equation}
-%In addition, $D$ in (\ref{eq:fld_radiation}) is the {\em Larsen
-%square-root flux-limiter} (see \cite{Morel2000}), $\kappa$ is the total
-%opacity, $\eta$ is the field of radiation sources, $G$ provides the
-%radiation induced photo-heating, $\Lambda$ provides the chemical
-%cooling, $\alpha_{i,j}$ are the reaction rate coefficients defining
-%the interactions between species ${\mathrm n}_i$ and ${\mathrm n}_j$,
-%and $\Gamma_i^{ph}$ are the radiation induced photo-ionization rates.
 
-The system of equations
-(\ref{eq:fld_radiation}-\ref{eq:fld_heating}) along with the chemical network 
-(Equation~\ref{eq:species_evolution}) is solved
-independently of \enzo's hydrodynamics, gravity and dark-matter solvers
-(Sections \ref{sec.hydro.ppm}-\ref{sec.ov.nbody}), thereby allowing
-the advective portions of Equations (\ref{eq:fld_radiation}) and
-(\ref{eq:species_evolution}) to be taken care of by the fluid solvers.
-Due to the disparate time scales between radiation transport and chemical
-ionization and heating, the remainder of these equations is solved
-using an operator-split algorithm.  Within a given timestep to evolve
-$(E_r^n, e_c^n, {\mathrm n}_i^n) \to (E_r^{n+1}, e_c^{n+1}, {\mathrm
-n}_i^{n+1})$, we first evolve equation (\ref{eq:fld_radiation}):
-$(E_r^n, e_c^n, {\mathrm n}_i^n) \to (E_r^{n+1}, e_c^{n}, {\mathrm
-  n}_i^{n})$.  This uses an implicit Euler time discretization, and a
-second-order centered finite difference spatial discretization,
-resulting in a large linear system of equations.  These are solved
-using a preconditioned conjugate gradient iteration, where the
-preconditioner consists of a geometric multigrid solver.  Both of
-these linear solvers are provided by the HYPRE linear solver library
-\citep[see][]{FalgoutYang2002, hypre-manual}.
+The system of equations (\ref{eq:fld_radiation}-\ref{eq:fld_heating})
+along with the chemical network (Equation~\ref{eq:species_evolution})
+is solved independently of \enzo's hydrodynamics, gravity and
+dark-matter solvers (Sections \ref{sec.hydro.ppm}-\ref{sec.ov.nbody}),
+thereby allowing the advective portions of Equations
+(\ref{eq:fld_radiation}) and (\ref{eq:species_evolution}) to be taken
+care of by the fluid solvers.  Due to the disparate time scales
+between radiation transport and chemical ionization and heating, the
+remainder of these equations is solved using an operator-split
+algorithm.  Within a given timestep to evolve $(E_r^n, e_c^n, {\mathrm
+n}_i^n) \to (E_r^{n+1}, e_c^{n+1}, {\mathrm n}_i^{n+1})$, we first
+evolve equation (\ref{eq:fld_radiation}): $(E_r^n, e_c^n, {\mathrm
+n}_i^n) \to (E_r^{n+1}, e_c^{n}, {\mathrm n}_i^{n})$.  This uses an
+implicit Euler time discretization, and a second-order centered finite
+difference spatial discretization, resulting in a large linear system
+of equations.  These are solved using a preconditioned conjugate
+gradient iteration, where the preconditioner consists of a geometric
+multigrid solver.  Both of these linear solvers are provided by the
+HYPRE linear solver library \citep[see][]{FalgoutYang2002,
+hypre-manual}.
 
-We then evolve the heating and chemistry system, Equations~(\ref{eq:fld_heating})
-and (\ref{eq:species_evolution}):  
+We then evolve the heating and chemistry system,
+Equations~(\ref{eq:fld_heating}) and (\ref{eq:species_evolution}):
 $(E_r^{n+1}, e_c^n, {\mathrm n}_i^n) \to (E_r^{n+1}, e_c^{n+1},
@@
 iteration, for the first 50 sweeps we use $\lambda=1$. For more
-challenging problems where this does not converge, we switch to a damping
-parameter of $\lambda=0.1$.
+challenging problems where this does not converge, we switch to a
+damping parameter of $\lambda=0.1$.

### 2013-10-22 numerical-radiation-raytracing.tex
Removed commented-out text from lots of files - basically vestigial junk.

@@
 
-Stars and black holes strongly affect their surroundings through radiation.
-Radiation transport is a well-studied problem; however, its treatment in
-multidimensional calculations is difficult because of the dependence on seven
-variables -- three spatial, two angular, frequency, and time.  The non-local
-nature of the thermal and hydrodynamical response to radiation sources further
-adds to the difficulty.  Here we briefly describe \enzo's ray tracing
+Stars and black holes strongly affect their surroundings through
+radiation.  Radiation transport is a well-studied problem; however,
+its treatment in multidimensional calculations is difficult because of
+the dependence on seven variables -- three spatial, two angular,
+frequency, and time.  The non-local nature of the thermal and
+hydrodynamical response to radiation sources further adds to the
+difficulty.  Here we briefly describe \enzo's ray tracing
 implementation \moray, which is presented in full detail in
@@
 
-%The radiative transfer equation in comoving coordinates
-%\citep{Gnedin97} reads
-%%
-%\begin{equation}
-%  \label{eq:rteqn}
-%  \frac{1}{c} \; \frac{\partial I_\nu}{\partial t} + 
-%  \frac{\hat{n} \cdot \grad I_\nu}{\bar{a}} -
-%  \frac{H}{c} \; \left( \nu \frac{\partial I_\nu}{\partial \nu} -
-%  3 I_\nu \right) = -\kappa_\nu I_\nu + j_\nu .
-%\end{equation}
-%%
-%Here $I_\nu \equiv I(\nu, \mathbf{x}, \Omega, t)$ is the radiation
-%specific intensity in units of energy per time $t$ per solid angle per
-%unit area per frequency $\nu$.  $H = \dot{a}/a$ is the Hubble
-%constant, where $a$ is the scale factor.  $\bar{a} = a/a_{em}$ is the
-%ratio of scale factors at the current time and time of emission.  The
-%second term represents the propagation of radiation, where the factor
-%$1/\bar{a}$ accounts for cosmic expansion.  The third term describes
-%both the cosmological redshift and dilution of radiation.  On the
-%right hand side, the first term considers the absorption coefficient
-%$\kappa_\nu \equiv \kappa_\nu(\mathbf{x},\nu,t)$, and the second term
-%$j_\nu \equiv j_\nu(\mathbf{x},\nu,t)$ is the emission coefficient
-%that includes any point sources of radiation or diffuse radiation.
-%GB: this was moved to the equations section.
-
-We solve the radiative transfer equation in comoving coordinates (given by Equation~\ref{eq:rteqn}).
-We can make some appropriate approximations to reduce the complexity
-of this equation in order to include radiation transport
-in numerical calculations.  Typically timesteps in dynamic
-calculations are small enough so that $\Delta a/a \ll 1$, therefore
-$a_{em}/a \approx 1$ in any given timestep, reducing the second term to
-$\hat{n} \partial I_\nu/\partial \mathbf{x}$.  To determine the
-importance of the third term, we evaluate the ratio of the third term
-to the second term.  This is $HL/c$, where $L$ is the simulation box
-length.  If this ratio is $\ll 1$, we can ignore the third term.  For
-example at $z=5$, this ratio is 0.1 when $L = c/H(z=5)$ = 53 proper
-Mpc.  In large boxes where the light crossing time is comparable to
-the Hubble time, then it becomes important to consider cosmological
-redshifting and dilution of the radiation.  Thus equation
-(\ref{eq:rteqn}) reduces to the non-cosmological form in this local
-approximation,
+We solve the radiative transfer equation in comoving coordinates
+(given by Equation~\ref{eq:rteqn}).  We can make some appropriate
+approximations to reduce the complexity of this equation in order to
+include radiation transport in numerical calculations.  Typically
+timesteps in dynamic calculations are small enough so that $\Delta a/a
+\ll 1$, therefore $a_{em}/a \approx 1$ in any given timestep, reducing
+the second term to $\hat{n} \partial I_\nu/\partial \mathbf{x}$.  To
+determine the importance of the third term, we evaluate the ratio of
+the third term to the second term.  This is $HL/c$, where $L$ is the
+simulation box length.  If this ratio is $\ll 1$, we can ignore the
+third term.  For example at $z=5$, this ratio is 0.1 when $L =
+c/H(z=5)$ = 53 proper Mpc.  In large boxes where the light crossing
+time is comparable to the Hubble time, then it becomes important to
+consider cosmological redshifting and dilution of the radiation.  Thus
+equation (\ref{eq:rteqn}) reduces to the non-cosmological form in this
+local approximation,
 %
@@
 $(r/R)^2$ rays, which is computationally wasteful because only a few
-rays per cell  are required to provide an accurate calculation
-of the radiation field (as we will show later).
+rays per cell are required to provide an accurate calculation of the
+radiation field (as we will show later).
 
 We avoid this inefficiency by utilizing adaptive ray tracing
-\citep{Abel02_RT}, which is based on Hierarchical Equal Area isoLatitude
-Pixelation \citep[HEALPix;][]{HEALPix} and progressively splits rays when the sampling
-becomes too coarse.  In this approach, the rays are
-traced along normal directions of the centers of the HEALPix pixels that 
-evenly divide a sphere into equal areas.  The rays are initialized at
-each point source with the photon luminosity (photon s$^{-1}$) equally
-spread across $N_{\rm pix} = 12 \times 4^l$ rays, where $l$ is the
-initial HEALPix level.  We usually find $l$ = 0 or 1 is sufficient
-because these coarse rays will usually be split before traversing
-the first cell.
+\citep{Abel02_RT}, which is based on Hierarchical Equal Area
+isoLatitude Pixelation \citep[HEALPix;][]{HEALPix} and progressively
+splits rays when the sampling becomes too coarse.  In this approach,
+the rays are traced along normal directions of the centers of the
+HEALPix pixels that evenly divide a sphere into equal areas.  The rays
+are initialized at each point source with the photon luminosity
+(photon s$^{-1}$) equally spread across $N_{\rm pix} = 12 \times 4^l$
+rays, where $l$ is the initial HEALPix level.  We usually find $l$ = 0
+or 1 is sufficient because these coarse rays will usually be split
+before traversing the first cell.
 
@@
 
-A ray propagates and splits until at least one of the following conditions is met: (1) the photon has traveled $c
-\times dt_P$, where $dt_P$ is the radiative transfer timestep; (2) its
-photon flux is almost fully absorbed ($>99.9\%$) in a single cell,
-which significantly reduces the computational time if the radiation
-volume filling fraction is small; (3) the photon leaves the
-computational domain with isolated boundary conditions; or (4) the
-photon travels $\sqrt{3}$ of the simulation box length with periodic
-boundary conditions.
-In the first case, the photon is halted at that position and saved,
-where it will be considered in the solution of $I_\nu$ at the next
-timestep.  In the next timestep, the photon will encounter a different
-hydrodynamical and ionization state, hence $\kappa$, in its path.
-Furthermore any time variations of the luminosities will be retained
-in the radiation field.  This is how this method retains the time
-derivative of the radiative transfer equation.  The last restriction
-prevents our method from considering sources external to the
-computational domain. However, a uniform radiation background can be used
-in conjunction with ray tracing that adds the background intensity to the
-local radiation field.
+A ray propagates and splits until at least one of the following
+conditions is met: (1) the photon has traveled $c \times dt_P$, where
+$dt_P$ is the radiative transfer timestep; (2) its photon flux is
+almost fully absorbed ($>99.9\%$) in a single cell, which
+significantly reduces the computational time if the radiation volume
+filling fraction is small; (3) the photon leaves the computational
+domain with isolated boundary conditions; or (4) the photon travels
+$\sqrt{3}$ of the simulation box length with periodic boundary
+conditions.  In the first case, the photon is halted at that position
+and saved, where it will be considered in the solution of $I_\nu$ at
+the next timestep.  In the next timestep, the photon will encounter a
+different hydrodynamical and ionization state, hence $\kappa$, in its
+path.  Furthermore any time variations of the luminosities will be
+retained in the radiation field.  This is how this method retains the
+time derivative of the radiative transfer equation.  The last
+restriction prevents our method from considering sources external to
+the computational domain. However, a uniform radiation background can
+be used in conjunction with ray tracing that adds the background
+intensity to the local radiation field.
 
@@
 absorption rate, resulting in photon conservation \citep{Abel99_RT,
-  Mellema06}.  Thus the photo-ionization and
-photo-heating rates associated with a single ray ($k_{\rm ph}$ and
-$\Gamma_{\rm ph}$, respectively)  are
+Mellema06}.  Thus the photo-ionization and photo-heating rates
+associated with a single ray ($k_{\rm ph}$ and $\Gamma_{\rm ph}$,
+respectively) are
 %
@@
 In each cell, the photo-ionization and photo-heating rates from each
-ray in the calculation are summed. After the ray tracing is
-complete, these rates are used as inputs to the solver described in
-Section~\ref{sec.ov.chem} to update the ionization, chemical, and energy states
-of the gas in each cell.
-
-%%% Local Variables: 
-%%% mode: latex
-%%% TeX-master: "ms"
-%%% End:
+ray in the calculation are summed. After the ray tracing is complete,
+these rates are used as inputs to the solver described in
+Section~\ref{sec.ov.chem} to update the ionization, chemical, and
+energy states of the gas in each cell.

### 2013-10-22 numerical-zeus.tex
Removed commented-out text from lots of files - basically vestigial junk.

@@
 
-As an alternative to the previous Godunov methods, \enzo\ also includes an implementation of the finite difference hydrodynamic algorithm employed in the compressible magnetohydrodynamics code \zeus\ \citep{Stone92a, Stone92b}.
-Fluid transport is solved on a Cartesian grid using the upwind, monotonic advection scheme of \citet{1977JCoPh..23..276V} within a multistep (operator-split) solution procedure that is fully explicit in time.
-This method is formally second-order accurate in space but first-order accurate in time.
+As an alternative to the previous Godunov methods, \enzo\ also
+includes an implementation of the finite difference hydrodynamic
+algorithm employed in the compressible magnetohydrodynamics code
+\zeus\ \citep{Stone92a, Stone92b}.  Fluid transport is solved on a
+Cartesian grid using the upwind, monotonic advection scheme of
+\citet{1977JCoPh..23..276V} within a multistep (operator-split)
+solution procedure that is fully explicit in time.  This method is
+formally second-order accurate in space but first-order accurate in
+time.
 
-As discussed in the section describing the Piecewise Parabolic Method (Section~\ref{sec.hydro.ppm}), operator-split methods break the solution of the hydrodynamic equations into parts, with each part representing a single term in the equations.
-Each part is evaluated successively using the results preceding it.
-In this method, in addition to operator-splitting the expansion terms (i.e., those terms in Eqs.~(\ref{eq:mass}) -- (\ref{eq:total_energy}) that depend on $\dot{a}$), we divide the remaining terms into \emph{source} and \emph{transport} steps.
-The terms to be solved in the source step are those on the right-hand side of eqs.~(\ref{eq:mass}) -- (\ref{eq:total_energy}), while the transport terms are on the left-hand side of these equations and are responsible for the advection of mass, momentum and energy across the grid.
+As discussed in the section describing the Piecewise Parabolic Method
+(Section~\ref{sec.hydro.ppm}), operator-split methods break the
+solution of the hydrodynamic equations into parts, with each part
+representing a single term in the equations.  Each part is evaluated
+successively using the results preceding it.  In this method, in
+addition to operator-splitting the expansion terms (i.e., those terms
+in Eqs.~(\ref{eq:mass}) -- (\ref{eq:total_energy}) that depend on
+$\dot{a}$), we divide the remaining terms into \emph{source} and
+\emph{transport} steps.  The terms to be solved in the source step are
+those on the right-hand side of eqs.~(\ref{eq:mass}) --
+(\ref{eq:total_energy}), while the transport terms are on the
+left-hand side of these equations and are responsible for the
+advection of mass, momentum and energy across the grid.
 
-The \zeus\ method uses a von Neumann-Richtmyer artificial viscosity to smooth shock discontinuities that may appear in fluid flows and can cause a breakdown of the finite difference equations.
-The artificial viscosity term is added in the source terms as:
+The \zeus\ method uses a von Neumann-Richtmyer artificial viscosity to
+smooth shock discontinuities that may appear in fluid flows and can
+cause a breakdown of the finite difference equations.  The artificial
+viscosity term is added in the source terms as:
 \begin{eqnarray}
@@
 \end{eqnarray}
-Here \textbf{Q} is the artificial viscosity stress tensor, which we take to be diagonal with on-axis terms given by $l^2 \rho (\partial v / \partial x)^2$ as proposed by von Neumann \& Richtmyer.
-The length scale $l$ determines the width of shocks and is typically a few times the cell spacing.
+Here \textbf{Q} is the artificial viscosity stress tensor, which we
+take to be diagonal with on-axis terms given by $l^2 \rho (\partial v
+/ \partial x)^2$ as proposed by von Neumann \& Richtmyer.  The length
+scale $l$ determines the width of shocks and is typically a few times
+the cell spacing.
 
-In our Cartesian coordinate system, the finite difference version of these equations is particularly simple, although there is one important complication.
-In the \zeus\ formalism, the velocity is a face-centered quantity -- that is, the velocity is recorded on a grid that is staggered as compared to the density, pressure and energy, which are at the cell center.
-Therefore we must remember that $v_j$ is at position $x_{j-1/2}$ (we use this notation, rather than writing $v_{j+1/2}$ both to match the original \zeus\ paper and also to make it easier to compare these equations to what is actually in the code).
+In our Cartesian coordinate system, the finite difference version of
+these equations is particularly simple, although there is one
+important complication.  In the \zeus\ formalism, the velocity is a
+face-centered quantity -- that is, the velocity is recorded on a grid
+that is staggered as compared to the density, pressure and energy,
+which are at the cell center.  Therefore we must remember that $v_j$
+is at position $x_{j-1/2}$ (we use this notation, rather than writing
+$v_{j+1/2}$ both to match the original \zeus\ paper and also to make
+it easier to compare these equations to what is actually in the code).
 
-As in the original \zeus\ paper, the source terms are added in three steps. First we add the pressure and gravity forces:
+As in the original \zeus\ paper, the source terms are added in three
+steps. First we add the pressure and gravity forces:
 \begin{equation}
@@
 \end{equation}
-Partial updates are denoted by the $n+a$, $n+b$ notation.
-We show updates in one dimension as the extension to the multi-dimensional case is straightforward (note that all dimensions are carried out for each substep before progressing to the next substep).
-We then add the artificial viscosity:
+Partial updates are denoted by the $n+a$, $n+b$ notation.  We show
+updates in one dimension as the extension to the multi-dimensional
+case is straightforward (note that all dimensions are carried out for
+each substep before progressing to the next substep). We then add the
+artificial viscosity:
 \begin{eqnarray}
@@
 \end{equation}
-where $Q_{\rm AV}$ is a constant with a typical value of 2.
-We refer the interested reader to \citet{Stone92a} and \citet{1994ApJ...429..434A} for more details.
-We also include the option (turned off by default) of adding a linear artificial viscosity as suggested in the \zeus\ paper for stagnant flow regions.
-This is given by
+where $Q_{\rm AV}$ is a constant with a typical value of 2. We refer
+the interested reader to \citet{Stone92a} and
+\citet{1994ApJ...429..434A} for more details.  We also include the
+option (turned off by default) of adding a linear artificial viscosity
+as suggested in the \zeus\ paper for stagnant flow regions.  This is
+given by
 \begin{equation}
@@
 \end{equation}
-We have used the notation $(\div \vecv)_j$ to indicate the (potentially) multi-dimensional velocity divergence evaluated at the cell center position $x_j$.
-This equation differs from the previous ones in that in the multi-dimensional case, still only one finite difference equation is evaluated, but the divergence becomes multi-dimensional.
+We have used the notation $(\div \vecv)_j$ to indicate the
+(potentially) multi-dimensional velocity divergence evaluated at the
+cell center position $x_j$.  This equation differs from the previous
+ones in that in the multi-dimensional case, still only one finite
+difference equation is evaluated, but the divergence becomes
+multi-dimensional.
 
-We next examine the transport step, which is conservative.
-Once again, we dimensionally split the equations and present only the one-dimensional version.
-The finite difference equations actually solved are:
+We next examine the transport step, which is conservative.  Once
+again, we dimensionally split the equations and present only the
+one-dimensional version.  The finite difference equations actually
+solved are:
+
 \begin{equation}
@@
 \end{equation}
-Here $\rho^*_j$ is the correctly upwinded value of $\rho$ evaluated at the cell-face corresponding to $v_j$, making $\rho^*_j v_j$ the mass flux at the cell boundary and guaranteeing mass conservation.
-This requires interpolating each cell-centered quantity to the cell edge.
-As recommended in \citet{Stone92a}, we use the second-order van Leer scheme, which uses piecewise linear functions.
-These are given by Equations (48) and (49) of \citet{Stone92a}.
-The transport steps for the other variables are similar.
-Note that we advect the specific energy and specific momenta using the mass flux, as dictated by the principle of consistent transport.
-This requires appropriate averaging for the momenta in the perpendicular directions as outlined in equations (57)-(72) of \citet{Stone92a}.
 
-%GB: the following is not correct (it is based on the Anninos & Norman paper which uses a
-%GB: different set of definitions for the comoving coordinates).
-%\begin{eqnarray}
-%Q_{ii} & = & \left\{ 
-%   \begin{array}{ll}
-%      Q_{\rm AV} \rho_b  (a \Delta v_{i} + \dot{a} \Delta x_i)^2 \quad
-%       & \textrm{for $a \Delta v_{i} + \dot{a} \Delta x_i < 0$}  \\
-%      0 & \textrm{otherwise}\\
-%\end{array} \right. \\
-%%\end{eqnarray}
-%%and
-%%\begin{equation}
-%Q_{ij} & = & 0  \;\;\;{\rm for}\;\; i \ne j . 
-%%\end{equation}
-%\end{eqnarray}
-%$\Delta x_i$ and $\Delta v_{i}$ refer to the comoving width of the grid cell
-%along the $i$-th axis and the corresponding difference in gas
-%peculiar velocities across the grid cell, respectively, and $a$ is the
-%cosmological scale factor.
+Here $\rho^*_j$ is the correctly upwinded value of $\rho$ evaluated at
+the cell-face corresponding to $v_j$, making $\rho^*_j v_j$ the mass
+flux at the cell boundary and guaranteeing mass conservation.   This
+requires interpolating each cell-centered quantity to the cell edge.
+As recommended in \citet{Stone92a}, we use the second-order van Leer
+scheme, which uses piecewise linear functions.  These are given by
+Equations (48) and (49) of \citet{Stone92a}.  The transport steps for
+the other variables are similar.  Note that we advect the specific
+energy and specific momenta using the mass flux, as dictated by the
+principle of consistent transport.  This requires appropriate
+averaging for the momenta in the perpendicular directions as outlined
+in equations (57)-(72) of \citet{Stone92a}.
 
-A limitation of a technique that uses an artificial viscosity is that, while the correct Rankine-Hugoniot jump conditions are achieved, shocks are broadened over 3-4 mesh cells.
-This may cause unphysical pre-heating of gas upstream of the shock wave, as discussed in \citet{1994ApJ...429..434A}.
-On the other hand, it is much more robust than PPM and is easy to add additional physics.
-We also note that this method solves only the internal energy equation rather than total energy, so the dual energy formulation discussed in Section~\ref{sec.hydro.ppm} is unnecessary.
+A limitation of a technique that uses an artificial viscosity is that,
+while the correct Rankine-Hugoniot jump conditions are achieved,
+shocks are broadened over 3-4 mesh cells. This may cause unphysical
+pre-heating of gas upstream of the shock wave, as discussed in
+\citet{1994ApJ...429..434A}.  On the other hand, it is much more
+robust than PPM and is easy to add additional physics.  We also note
+that this method solves only the internal energy equation rather than
+total energy, so the dual energy formulation discussed in
+Section~\ref{sec.hydro.ppm} is unnecessary.

### 2013-10-22 parallel.tex

@@
 that the current version of \enzo\ is not threaded, although a hybrid
-MPI + OpenMP version is under development.}.  
-Parallelization is
+MPI + OpenMP version is under development.}.  Parallelization is
 accomplished by distributing grids amongst processors.  This is done
@@
 communication; however, this generally does not result in a
-well-balanced computational load.  Therefore, the code has a number of options for
-load balancing the grids on a given level. Each grid is assigned an
-estimated computational load, generally equal to the number of cells in the grid (which,
-empirically, is a good estimate of computational cost).  The first
-load-balancing option is to move a grid from the processor with the
-highest computational load to the processor with the lowest load, with the proviso
-that only grids with load factors less than half the difference
-between the highest and lowest loaded processors will be moved.  This
-continues until the load ratio between the most-to-least loaded
-processor is below 1.05 or until no suitable grid can be found to
-transfer.  A second load balancing option uses a space-filling Hilbert
-curve to order the grids by their approximate spatial position.  Then,
-once the grids have a specific one-dimensional ordering, we can divide
-up the grids into $N_p$ groups (with the division taking place as
-equally as possible).  Load balancing is done separately for each
-level.  Clearly load balancing is most successful if there are
-significantly more grids than processors; however, small grids are
-less efficient because of the large number of ghost zones compared to
-active zones, and so the code uses a
-simple heuristic in order to split up grids until there are of order
-10 grids per processor.  This generally results in good load-balancing
-while not producing grids that are wastefully small.
+well-balanced computational load.  Therefore, the code has a number of
+options for load balancing the grids on a given level. Each grid is
+assigned an estimated computational load, generally equal to the
+number of cells in the grid (which, empirically, is a good estimate of
+computational cost).  The first load-balancing option is to move a
+grid from the processor with the highest computational load to the
+processor with the lowest load, with the proviso that only grids with
+load factors less than half the difference between the highest and
+lowest loaded processors will be moved.  This continues until the load
+ratio between the most-to-least loaded processor is below 1.05 or
+until no suitable grid can be found to transfer.  A second load
+balancing option uses a space-filling Hilbert curve to order the grids
+by their approximate spatial position.  Then, once the grids have a
+specific one-dimensional ordering, we can divide up the grids into
+$N_p$ groups (with the division taking place as equally as possible).
+Load balancing is done separately for each level.  Clearly load
+balancing is most successful if there are significantly more grids
+than processors; however, small grids are less efficient because of
+the large number of ghost zones compared to active zones, and so the
+code uses a simple heuristic in order to split up grids until there
+are of order 10 grids per processor.  This generally results in good
+load-balancing while not producing grids that are wastefully small.
 
@@
 that do not `live' on a given processor only contain meta data
-(essentially just the location and size of the grid; such grids are denoted as
-`ghost' grids).  Replicating the hierarchy means that all
+(essentially just the location and size of the grid; such grids are
+denoted as `ghost' grids).  Replicating the hierarchy means that all
 communication of data from one grid to another can be identified by
@@
 
-Because of the wide variety of simulations, methods, and uses of \enzo,
-it is difficult to state in general terms which routines within the
-code will be most costly during a given simulation.  As such, we have
-designed a lightweight registering system that has been implemented
-for the most commonly used routines (such as the hydrodynamic and
-gravity solvers) as well as refinement level timers that measure the
-time spent on each level.  Beyond this minimal set of routines, we
-have designed a simple way for the user to modify the source by adding
-\texttt{TIMER\_START("Your\_Routine\_Name")} and
+Because of the wide variety of simulations, methods, and uses of
+\enzo, it is difficult to state in general terms which routines within
+the code will be most costly during a given simulation.  As such, we
+have designed a lightweight registering system that has been
+implemented for the most commonly used routines (such as the
+hydrodynamic and gravity solvers) as well as refinement level timers
+that measure the time spent on each level.  Beyond this minimal set of
+routines, we have designed a simple way for the user to modify the
+source by adding \texttt{TIMER\_START("Your\_Routine\_Name")} and
 \texttt{TIMER\_END("Your\_Routine\_Name")}.  These timers are created
@@
 
-At each complete root grid timestep (or less often if specified),
-each timer is then communicated to the root processor where it
-calculates the mean, standard deviation, minimum, and maximum for each
-of the timers of that name across all processors.  For level timers, there are additional
-attributes such as the number of cell updates, the current number of
-grids, and the average cells updates per second per MPI process.  This
-information is then 
-output to a logfile.  This provides a simplified interface to the user
-that can be used to diagnose performance issues as well as estimate a
-given problem type's scalability.  In addition to the logfile, we have
-developed a plotting interface for quickly producing figures that
-process the data from the logfile.  These capabilities are described
-in the online documentation, along with further discussion of the
-performance measurement implementation.
+At each complete root grid timestep (or less often if specified), each
+timer is then communicated to the root processor where it calculates
+the mean, standard deviation, minimum, and maximum for each of the
+timers of that name across all processors.  For level timers, there
+are additional attributes such as the number of cell updates, the
+current number of grids, and the average cells updates per second per
+MPI process.  This information is then output to a logfile.  This
+provides a simplified interface to the user that can be used to
+diagnose performance issues as well as estimate a given problem type's
+scalability.  In addition to the logfile, we have developed a plotting
+interface for quickly producing figures that process the data from the
+logfile.  These capabilities are described in the online
+documentation, along with further discussion of the performance
+measurement implementation.
 
@@
 forest cosmology simulations with constant comoving spatial resolution
-per grid cell, showing cell updates per second per processor 
-plotted as a function of the number of root grid tiles of dimension
-$128^3$ (R) in each dimension.  The number of MPI tasks is N$ = R^3$,
-so R$ = 16$ on this plot corresponds to a $2048^3$ computational mesh
-running on 4,096 MPI tasks.  This plot goes from R$ = 2$ (8 MPI tasks)
-to R$ = 24$ (13,824 MPI tasks) on two supercomputers -- NICS Kraken
-and ORNL Jaguar when they were Cray XT4 systems -- and using 1, 2 or 4
-MPI processes per node, where each compute node contained a single
+per grid cell, showing cell updates per second per processor plotted
+as a function of the number of root grid tiles of dimension $128^3$
+(R) in each dimension.  The number of MPI tasks is N$ = R^3$, so R$ =
+16$ on this plot corresponds to a $2048^3$ computational mesh running
+on 4,096 MPI tasks.  This plot goes from R$ = 2$ (8 MPI tasks) to R$ =
+24$ (13,824 MPI tasks) on two supercomputers -- NICS Kraken and ORNL
+Jaguar when they were Cray XT4 systems -- and using 1, 2 or 4 MPI
+processes per node, where each compute node contained a single
 quad-core AMD Opteron CPU with a speed of 2.1 GHz on Jaguar and 2.3
@@
 upon initialization, unigrid simulations are decomposed such that each
-MPI process has a roughly equal subvolume (and thus an equal number of grid
-cells), meaning that work is evenly distributed among computational elements.  Communication
-patterns for both the gravity solve (which uses a fast Fourier
-transform) and the fluid solves (which transfer boundary information
-between subvolumes) are predictable and straightforward, and
-rebuilding of the grid hierarchy does not take place, removing a
-substantial global operation and a great deal of communication.
+MPI process has a roughly equal subvolume (and thus an equal number of
+grid cells), meaning that work is evenly distributed among
+computational elements.  Communication patterns for both the gravity
+solve (which uses a fast Fourier transform) and the fluid solves
+(which transfer boundary information between subvolumes) are
+predictable and straightforward, and rebuilding of the grid hierarchy
+does not take place, removing a substantial global operation and a
+great deal of communication.
 
@@
 sequence of scaled unigrid Lyman-$\alpha$ forest calculations. These
-calculations include dark matter dynamics, hydrodynamics using the
-PPM solver, six-species non-equilibrium chemistry and
-radiative cooling, and a uniform metagalactic ultraviolet background.
-In this sequence of test calculations, we perform a weak scaling test
-on up to 13,824 MPI tasks on the NICS Kraken XT4 and ORNL Jaguar XT4
+calculations include dark matter dynamics, hydrodynamics using the PPM
+solver, six-species non-equilibrium chemistry and radiative cooling,
+and a uniform metagalactic ultraviolet background.  In this sequence
+of test calculations, we perform a weak scaling test on up to 13,824
+MPI tasks on the NICS Kraken XT4 and ORNL Jaguar XT4
 supercomputers\footnote{These simulations were performed prior to
@@
 approximately $128^3$ dark matter particles.  The number of grid cells
-on each processor 
-was constant throughout the calculation; the number of dark matter
-particles on each processor varies as they are moved from subvolume to subvolume as
-structure evolves.  The grid resolution was kept at a constant
-comoving size of $\simeq 40$~kpc/h, and as the core count was
-increased, so was the simulation volume.  On each machine, a compute
-node contained a single AMD Opteron quad-core chip (2.1 Ghz on Jaguar;
-2.3 Ghz on Kraken) with 2 GB/memory per core (8 GB/total per node).
-Both machines used the SeaStar2 interconnect.  In the scaling study,
-calculations were run with 1, 2, or 4 MPI tasks per node.  The figure
-shows cell updates per second per MPI process; perfect scaling would
-be a horizontal line.
+on each processor was constant throughout the calculation; the number
+of dark matter particles on each processor varies as they are moved
+from subvolume to subvolume as structure evolves.  The grid resolution
+was kept at a constant comoving size of $\simeq 40$~kpc/h, and as the
+core count was increased, so was the simulation volume.  On each
+machine, a compute node contained a single AMD Opteron quad-core chip
+(2.1 Ghz on Jaguar; 2.3 Ghz on Kraken) with 2 GB/memory per core (8
+GB/total per node).  Both machines used the SeaStar2 interconnect.  In
+the scaling study, calculations were run with 1, 2, or 4 MPI tasks per
+node.  The figure shows cell updates per second per MPI process;
+perfect scaling would be a horizontal line.
 
@@
 performance of the code is extremely good for this problem, with only
-a 20\% decrease in cell updates per second per MPI task as the code is scaled from
-8 to 4,096 MPI tasks, and a 40\% decrease in performance overall going
-from 8 to 13,824 (or $24^3$) MPI tasks.  We speculate that this
-decrease is likely to be partially due to global MPI communications
-used to, e.g., calculate the overall timestep of the simulation, and
-also likely due to load imbalances due to increasing cosmological
-power (and thus an increasingly uneven distribution of dark matter
-particles between MPI tasks at late times) as the simulation volume
-grows.  We also observe that a systematic difference in speed can be
-seen between the two machines, which can be attributed primarily to
-the slightly faster CPUs on Kraken at the time (2.3 Ghz, vs. 2.1 Ghz
-on Jaguar).  The difference in speed when using different numbers of
-MPI tasks per node can be attributed primarily due to differences in
-competing usage of shared cache on the quad-core chips used on this
-machine.
+a 20\% decrease in cell updates per second per MPI task as the code is
+scaled from 8 to 4,096 MPI tasks, and a 40\% decrease in performance
+overall going from 8 to 13,824 (or $24^3$) MPI tasks.  We speculate
+that this decrease is likely to be partially due to global MPI
+communications used to, e.g., calculate the overall timestep of the
+simulation, and also likely due to load imbalances due to increasing
+cosmological power (and thus an increasingly uneven distribution of
+dark matter particles between MPI tasks at late times) as the
+simulation volume grows.  We also observe that a systematic difference
+in speed can be seen between the two machines, which can be attributed
+primarily to the slightly faster CPUs on Kraken at the time (2.3 Ghz,
+vs. 2.1 Ghz on Jaguar).  The difference in speed when using different
+numbers of MPI tasks per node can be attributed primarily due to
+differences in competing usage of shared cache on the quad-core chips
+used on this machine.
 
@@
 been empirically determined to be roughly $128^3$ cells per core.  If
-fewer cells per core are used, the CPU is essentially data-starved 
-and poor scaling is observed due to computing units being idle while
+fewer cells per core are used, the CPU is essentially data-starved and
+poor scaling is observed due to computing units being idle while
 waiting for information to be communicated from other processes (for,
@@
 cosmological calculation.  The root grid scaling is not representative
-of the true strong scaling of \enzo\ because the root grid tiles are not
-repartitioned when a simulation is restarted with additional cores.
-The weak scaling test in Figure \ref{fig.uniscale} is more
+of the true strong scaling of \enzo\ because the root grid tiles are
+not repartitioned when a simulation is restarted with additional
+cores.  The weak scaling test in Figure \ref{fig.uniscale} is more
 representative of the scaling on the root level.  The performance in
@@
 \citep{2012ApJ...749..140H}, high-resolution disk galaxy simulations
-\citep{2011ApJ...738...54K}, high-redshift \citep{2009Sci...325..601T}, and present-day
-star formation \citep{Collins12a}, involve multi-scale physics that
-span several orders of magnitude in both space and time.  In these
-situations, using \enzo\ in its adaptive mesh refinement mode is
-beneficial.  Because the adaptive grid hierarchy is dynamic, grid
-boundaries and, thus, communication patterns can be unpredictable,
-hindering strong scaling to high core counts.
+\citep{2011ApJ...738...54K}, high-redshift
+\citep{2009Sci...325..601T}, and present-day star formation
+\citep{Collins12a}, involve multi-scale physics that span several
+orders of magnitude in both space and time.  In these situations,
+using \enzo\ in its adaptive mesh refinement mode is beneficial.
+Because the adaptive grid hierarchy is dynamic, grid boundaries and,
+thus, communication patterns can be unpredictable, hindering strong
+scaling to high core counts.
 
@@
 simulation behavior in general.  The results shown in Figure
-\ref{fig:strong_scaling} come from a single root-level timestep
-of $\Delta t = 2.1\, \textrm{Myr}$.  At this time, there are $3.04 \times
+\ref{fig:strong_scaling} come from a single root-level timestep of
+$\Delta t = 2.1\, \textrm{Myr}$.  At this time, there are $3.04 \times
 10^5$ AMR grids, $9.03 \times 10^8$ ($\sim 967^3$) computational
@@
 be easily implemented and is planned in the next major release of
-\enzo, where scaling results would be similar to the weak scaling shown
-in \S\ref{sec:weak_scaling}.  The right panel in Figure
+\enzo, where scaling results would be similar to the weak scaling
+shown in \S\ref{sec:weak_scaling}.  The right panel in Figure
 \ref{fig:strong_scaling} shows the time spent in some representative
@@
 in each routine, it is clear that the communication in the
-\texttt{SetBoundaryConditions}, \texttt{SolveForPotential} (the multi-grid
-solver in AMR subgrids), and \texttt{RebuildHierarchy} are responsible
-for the lack of strong scaling at $N_{\rm core} \ga$~4,096 in this
-particular simulation.  The transpose of the root grid tiles are
-responsible for the performance decrease because it is not optimized
-for situations where the number of tiles is greater than the number of
-MPI processes.  These results are directly applicable to simulations
-with similar computational demands.  In simulations with fewer
-computational cells, strong scaling will cease at a smaller core count
-because the CPUs will become data-starved more quickly, and the
+\texttt{SetBoundaryConditions}, \texttt{SolveForPotential} (the
+multi-grid solver in AMR subgrids), and \texttt{RebuildHierarchy} are
+responsible for the lack of strong scaling at $N_{\rm core} \ga$~4,096
+in this particular simulation.  The transpose of the root grid tiles
+are responsible for the performance decrease because it is not
+optimized for situations where the number of tiles is greater than the
+number of MPI processes.  These results are directly applicable to
+simulations with similar computational demands.  In simulations with
+fewer computational cells, strong scaling will cease at a smaller core
+count because the CPUs will become data-starved more quickly, and the
 opposite occurs with larger simulations.
@@
 drastically throughout the calculation.  For example, at the beginning
-of a cosmological simulation, when the densities are nearly uniform, only the
-static grid is required and the calculation progresses rapidly.
-However, as structure forms and dense clumps are generated, the number
-of grid points swells by orders of magnitude (an increase of $10^3$ is
-not uncommon) and most of the CPU time is consumed at late physical
-time in the simulation.
-Therefore, simply performing a few steps at the beginning of the
-calculation does not produce a good estimate of the required CPU time.
+of a cosmological simulation, when the densities are nearly uniform,
+only the static grid is required and the calculation progresses
+rapidly.  However, as structure forms and dense clumps are generated,
+the number of grid points swells by orders of magnitude (an increase
+of $10^3$ is not uncommon) and most of the CPU time is consumed at
+late physical time in the simulation.  Therefore, simply performing a
+few steps at the beginning of the calculation does not produce a good
+estimate of the required CPU time.
 
@@
 to $N_{\rm root}^4$ since the number of cells scales as $N_{\rm
-root}^3$ and, assuming the Courant condition is the 
-factor that controls the timestep, the number of steps to advance the
-calculation over a given time interval is
-proportional to $N_{\rm root}.$ Therefore, to advance a hierarchy a
-given time interval, we find, accounting for all levels and using a refinement
-factor of 2,
-\begin{equation} t_{\rm SU} = C_1 \sum_l f_l N_{\rm root}^4 2^{4l}
-\end{equation} where $C_1$ is a constant, which can be thought of as
-the time taken to advance a single cell.  The factor $f_l$ is the
-fraction of the volume on a given level that is actually refined.  By
-definition $f_0 = 1$, and $f_l \le 1$.  In writing down this equation,
-we neglect a number of costs that are not directly proportional to
-cell count, including communication between processors, the $\log{N}$
-factor for the root grid FFT, cache misses, optimizations, and other
-costs associated with processing the hierarchy.  The first item on
-this list, in particular, is clearly important for large processor
-counts; however, we neglect parallel considerations in this section.
+root}^3$ and, assuming the Courant condition is the factor that
+controls the timestep, the number of steps to advance the calculation
+over a given time interval is proportional to $N_{\rm root}.$
+Therefore, to advance a hierarchy a given time interval, we find,
+accounting for all levels and using a refinement factor of 2,
+
+\begin{equation} 
+t_{\rm SU} = C_1 \sum_l f_l N_{\rm root}^4 2^{4l}
+\end{equation}
+
+where $C_1$ is a constant, which can be thought of as the time taken
+to advance a single cell.  The factor $f_l$ is the fraction of the
+volume on a given level that is actually refined.  By definition $f_0
+= 1$, and $f_l \le 1$.  In writing down this equation, we neglect a
+number of costs that are not directly proportional to cell count,
+including communication between processors, the $\log{N}$ factor for
+the root grid FFT, cache misses, optimizations, and other costs
+associated with processing the hierarchy.  The first item on this
+list, in particular, is clearly important for large processor counts;
+however, we neglect parallel considerations in this section.
 
@@
 cosmological runs, with only slight variations depending on the exact
-problem that is being simulated.  The top-right panel
-shows the ratio of $f_{l}/f_{l+1}$ which, remarkably, hovers around
-the critical value of 0.0625 determined earlier.  This implies that
-the relative CPU usage of each level, shown in the lower-left panel,
-is nearly flat, and that going to additional levels of refinement only
-increases the amount of calculation by a factor of roughly $1/l_{\rm
-max}$.  Even in the most extreme case, the L80 run, the last level
-adds only 25\% to the computation time.
+problem that is being simulated.  The top-right panel shows the ratio
+of $f_{l}/f_{l+1}$ which, remarkably, hovers around the critical value
+of 0.0625 determined earlier.  This implies that the relative CPU
+usage of each level, shown in the lower-left panel, is nearly flat,
+and that going to additional levels of refinement only increases the
+amount of calculation by a factor of roughly $1/l_{\rm max}$.  Even in
+the most extreme case, the L80 run, the last level adds only 25\% to
+the computation time.
 
 The memory used by each level, on the other hand, scales as
+
 \begin{equation}
@@
 \end{equation}
+
 and the terms in this sum are shown in the bottom-right panel of
@@
 clouds themselves.  In particular, we note that for a $\rho \propto
-r^{-2}$ density profile, the Lagrangian refinement criteria typically used in
-such simulations produces an $f_{l+1}/f_l$ ratio of 1/8 for a single,
-resolved cloud.  Of course, at some point we would resolve the flat
-density of individual clouds and the ratio would climb, making further
-refinement more costly; however, we are not yet in this regime in this
-example.
+r^{-2}$ density profile, the Lagrangian refinement criteria typically
+used in such simulations produces an $f_{l+1}/f_l$ ratio of 1/8 for a
+single, resolved cloud.  Of course, at some point we would resolve the
+flat density of individual clouds and the ratio would climb, making
+further refinement more costly; however, we are not yet in this regime
+in this example.
 
@@
 of a galaxy formation run, as shown by the dot-dashed curves in
-Figure~\ref{fig:scaling}.  Note that grid levels 1 and 2 in this simulation are statically
-refined to ensure that the Lagrangian volume of the galaxy halo is resolved
-to at least level 2 at all times. More importantly, the
-$f_{l+1}/f_{l}$ ratio of the highest levels are systematically
-slightly larger than the critical 0.0625 value, indicating that the
-CPU time is dominated by the most refined levels, as shown in the
-bottom-left panel.  The memory usage is dominated by levels 2 and 3,
-the static levels, as the bottom-right panel demonstrates.  In this
-case, adding more levels of refinement would substantially increase the
-cost of the simulation.
+Figure~\ref{fig:scaling}.  Note that grid levels 1 and 2 in this
+simulation are statically refined to ensure that the Lagrangian volume
+of the galaxy halo is resolved to at least level 2 at all times. More
+importantly, the $f_{l+1}/f_{l}$ ratio of the highest levels are
+systematically slightly larger than the critical 0.0625 value,
+indicating that the CPU time is dominated by the most refined levels,
+as shown in the bottom-left panel.  The memory usage is dominated by
+levels 2 and 3, the static levels, as the bottom-right panel
+demonstrates.  In this case, adding more levels of refinement would
+substantially increase the cost of the simulation.
 
@@
 on the first and subsequent levels by this factor.  Again, we have
-empirically tested that this scaling is approximately valid,
-provided that we keep the maximum level constant.
+empirically tested that this scaling is approximately valid, provided
+that we keep the maximum level constant.
 
@@
 computational time required for cosmological simulations:
+
 \begin{equation}
@@
 \end{equation}
+
 This is approximately accurate, and gives users some way to estimate

### 2013-10-22 tests-aniso-conduct.tex
Finished up all of the test problems!

--- 2013-06-08 tests-aniso-conduct.tex
@@
 Figure~\ref{fig.conduct} shows a test that demonstrates the correct
-behavior of anisotropic thermal conduction (see
-Section~\ref{sec.num.conductions}) in \enzo. We initialize a
+behavior of anisotropic thermal conduction in \enzo. We initialize a
 two-dimensional, $256 \times 256$ cell simulation having a physical
@@
 \caption{Two-dimensional anisotropic conduction test in a uniform,
-constant temperature background with circular magnetic fields centered
-on (0.5, 0.5) in a $256 \times 256$ grid. The background medium has a
-density of 1 proton/cc and a temperature of $10^6$~K.  At t$ = 0$
-(left panel), a Gaussian heat pulse is injected at (0.75, 0.5) with a
-FWHM of $1/64$ (with all numbers given in units of the box size) and a
-peak temperature of $10^8$~K, and allowed to evolve without
-hydrodynamical motion (i.e., static gas) and no radiative cooling for
-300 Myr.  At t$ = 300$~Myr (right panel), heat has been transported
-along magnetic field lines with no significant diffusion perpendicular
-to field lines. Furthermore, there are no detectable oscillations in
-the temperature in regions where the magnetic field is not parallel
-with the grids.}
+constant temperature background with circular magnetic fields
+(indicated by white streamlines) centered on (0.5, 0.5) in a $256
+\times 256$ grid. The background medium has a density of 1 proton/cc
+and a temperature of $10^6$~K.  At t$ = 0$ (left panel), a Gaussian
+heat pulse is injected at (0.75, 0.5) with a FWHM of $1/64$ (with all
+numbers given in units of the box size) and a peak temperature of
+$10^8$~K, and allowed to evolve without hydrodynamical motion (i.e.,
+static gas) and no radiative cooling for 300 Myr.  At t$ = 300$~Myr
+(right panel), heat has been transported along magnetic field lines
+with no significant diffusion perpendicular to field
+lines. Furthermore, there are no detectable oscillations in the
+temperature in regions where the magnetic field is not parallel with
+the grids.}
 \label{fig.conduct}

### 2013-10-22 tests-fld-ifront.tex

@@
 \cite{ReynoldsHayesPaschosNorman2009}, run using the cosmological
-deceleration parameter $q_0 = 0.05$ and initial redshift $z_i=10$.
-In this test problem, the physics of interest is the expansion of an \ion{H}{2} region
-in a uniform medium around a single monochromatic ionizing source
-(with frequency $h\nu = 13.6$ eV).  The I-front propagates rapidly at
-first, approaching 90\% of the Str{\" o}mgren radius by a scaled
-redshift of $-\log\left[(1+z)/(1+z_i)\right] \approx 0.15$, before
-cosmological expansion overtakes ionization, pushing the Str{\"
-  o}mgren radius outward faster than the I-front can propagate.  The Str{\" o}mgren radius is analytically given by
+deceleration parameter $q_0 = 0.05$ and initial redshift $z_i=10$.  In
+this test problem, the physics of interest is the expansion of an
+\ion{H}{2} region in a uniform medium around a single monochromatic
+ionizing source (with frequency $h\nu = 13.6$ eV).  The I-front
+propagates rapidly at first, approaching 90\% of the Str{\" o}mgren
+radius by a scaled redshift of $-\log\left[(1+z)/(1+z_i)\right]
+\approx 0.15$, before cosmological expansion overtakes ionization,
+pushing the Str{\" o}mgren radius outward faster than the I-front can
+propagate.  The Str{\" o}mgren radius is analytically given by
 \[
@@
 The simulation parameters are: box size $L_i\approx 27$ kpc, Hubble
-constant $H_0 = 100$ km s$^{-1}$ Mpc$^{-1}$, and cosmological parameters $\Omega_m = 0.1$, $\Omega_\Lambda=0$, and $\Omega_b = 0.1$.  The
-initial values for the simulation are a radiation energy density
+constant $H_0 = 100$ km s$^{-1}$ Mpc$^{-1}$, and cosmological
+parameters $\Omega_m = 0.1$, $\Omega_\Lambda=0$, and $\Omega_b = 0.1$.
+The initial values for the simulation are a radiation energy density
 $E_{r,i} = 10^{-35}$ erg cm$^{-3}$, temperature $T_i = 10^4$ K,
@@
 \end{figure}
-
-%%% Local Variables: 
-%%% mode: latex
-%%% TeX-master: "ms"
-%%% End:

### 2013-10-22 tests-onezonecollapse.tex

--- 2013-06-25 tests-onezonecollapse.tex
@@
 running a true collapse simulation.  The density evolution is based on
-the self-similar Larson-Penston solution for isothermal collapse \citep{1969MNRAS.145..271L,
-1969MNRAS.144..425P} with a
-modification to account for the efficiency with which the heat
-introduced by compression can be radiated away
-\citep{1983ApJ...265.1047Y}.  Our implementation, described briefly
-here, follows the work of \citet{2005ApJ...626..627O}, and we direct
-the interested reader to this paper for further details.  The
-evolution of the gas density, $\rho$, is given by
+the self-similar Larson-Penston solution for isothermal collapse
+\citep{1969MNRAS.145..271L, 1969MNRAS.144..425P} with a modification
+to account for the efficiency with which the heat introduced by
+compression can be radiated away \citep{1983ApJ...265.1047Y}.  Our
+implementation, described briefly here, follows the work of
+\citet{2005ApJ...626..627O}, and we direct the interested reader to
+this paper for further details.  The evolution of the gas density,
+$\rho$, is given by
 
@@
  The collapse timescale is altered from the dynamical time by a factor
-$1/\sqrt{1-f}$ in Equation
-\ref{eqn.tcol}, which is an approximation of the ratio of the gas
-pressure to the force of gravity.  The value of $f$ depends on the
-effective adiabatic index, $\gamma_{\rm ef} \equiv (\partial \ln p
-/ \partial \ln \rho)$, which we linearly extrapolate
-from derivative values at the two previous timesteps.  For the value
-of $f$ in this test problem, we use the piecewise function of
-\citet{2005ApJ...626..627O}, given by
+$1/\sqrt{1-f}$ in Equation \ref{eqn.tcol}, which is an approximation
+of the ratio of the gas pressure to the force of gravity.  The value
+of $f$ depends on the effective adiabatic index, $\gamma_{\rm ef}
+\equiv (\partial \ln p / \partial \ln \rho)$, which we linearly
+extrapolate from derivative values at the two previous timesteps.  For
+the value of $f$ in this test problem, we use the piecewise function
+of \citet{2005ApJ...626..627O}, given by
 
@@
 and energy, temperature, density, and pressure are related by the
-ideal gas law, including effects from molecular hydrogen as appropriate.
-Figure \ref{fig.onezone} shows an example of the one-zone collapse test
-performed with an initial number density of 1 hydrogen atom per cm$^{-3}$ and
-an initial temperature of 100 K using the 12 species chemistry network with H,
-D, and He species and metal cooling rates calculated with the \texttt{Cloudy}
-code.  The effects of metal cooling can be clearly seen; as the metallicity
-increases from zero to $10^{-2}$~Z$_\odot$, the gas rapidly and significantly
-deviates from the primordial result (black line).  Our primordial results
-compare very well to those shown in \citet{2005ApJ...626..627O}; however, we
-use a different metal cooling method, so the lines describing the evolution of
-the metal-enriched gas are not directly comparable.
+ideal gas law, including effects from molecular hydrogen as
+appropriate.  Figure \ref{fig.onezone} shows an example of the
+one-zone collapse test performed with an initial number density of 1
+hydrogen atom per cm$^{-3}$ and an initial temperature of 100 K using
+the 12 species chemistry network with H, D, and He species and metal
+cooling rates calculated with the \texttt{Cloudy} code.  The effects
+of metal cooling can be clearly seen; as the metallicity increases
+from zero to $10^{-2}$~Z$_\odot$, the gas rapidly and significantly
+deviates from the primordial result (black line).  Our primordial
+results compare very well to those shown in
+\citet{2005ApJ...626..627O}; however, we use a different metal cooling
+method, so the lines describing the evolution of the metal-enriched
+gas are not directly comparable.
 

### 2013-10-22 tests-photoevaporation.tex

@@
 \citet{Wise11_Moray}.  The simulation domain is 6.6~kpc on a side with
-an ambient medium of pure neutral hydrogen of density $n_{\rm H} =
-2 \times 10^{-4}\; \cubecm$ and temperature $T = 8000 \unit{K}$.  We
+an ambient medium of pure neutral hydrogen of density $n_{\rm H} = 2
+\times 10^{-4}\; \cubecm$ and temperature $T = 8000 \unit{K}$.  We
 place a spherical overdensity in hydrostatic equilibrium with the
 ambient medium.  It has a radius $r = 0.8 \unit{kpc}$, hydrogen
-density $n_{\rm H} = 0.04\ \cubecm$ (i.e., overdensity of 200), and temperature $T = 40$ K, and
-is centered at $(x,y,z) = (5, 3.3, 3.3) \unit{kpc}$.
-In \citet{IlievEtAl2009} all of the codes used a fixed $128^3$ grid to
-ease the comparison, but in this test to demonstrate a higher
-resolution AMR solution, we employ a $128^3$ grid with two additional
-levels of refinement by factors of 2 for cells with a baryon mass greater than 1.5
-(method 2 in Section~\ref{sec:refinement_criteria}). This test is run
-for 15 Myr.
+density $n_{\rm H} = 0.04\ \cubecm$ (i.e., overdensity of 200), and
+temperature $T = 40$ K, and is centered at $(x,y,z) = (5, 3.3, 3.3)
+\unit{kpc}$.  In \citet{IlievEtAl2009} all of the codes used a fixed
+$128^3$ grid to ease the comparison, but in this test to demonstrate a
+higher resolution AMR solution, we employ a $128^3$ grid with two
+additional levels of refinement by factors of 2 for cells with a
+baryon mass greater than 1.5 (method 2 in
+Section~\ref{sec:refinement_criteria}). This test is run for 15 Myr.
 
-The cloud is subject to radiation from a point source at the center
-of the $x=0$ boundary with an ionizing photon luminosity
-$\dot{N}_\gamma = 3 \times 10^{51}$ photons s$^{-1}$, corresponding to
-a flux $F_0 = 10^6 \unit{photons s}^{-1} \unit{cm}^{-2}$ at the clump
-surface closest to the radiation source.  The radiation source has a
-spectrum of a $T = 10^5 \unit{K}$ blackbody, and we use four energy
-groups with the following mean energies and relative luminosities:
-$E_i = (17.98, 31.15, 49.09, 76.98) \unit{eV}, L_i/L = (0.23, 0.36,
-0.24, 0.06)$ that are optimized to reduce errors in the solutions with
-a full spectrum and energy discretization \citep{Mirocha12}.
-(Note that this choice of energy groups is different from those used
-in Wise \& Abel 2011.) \nocite{Wise11_Moray}  We use a minimum angular resolution of 10 rays
+The cloud is subject to radiation from a point source at the center of
+the $x=0$ boundary with an ionizing photon luminosity $\dot{N}_\gamma
+= 3 \times 10^{51}$ photons s$^{-1}$, corresponding to a flux $F_0 =
+10^6 \unit{photons s}^{-1} \unit{cm}^{-2}$ at the clump surface
+closest to the radiation source.  The radiation source has a spectrum
+of a $T = 10^5 \unit{K}$ blackbody, and we use four energy groups with
+the following mean energies and relative luminosities: $E_i = (17.98,
+31.15, 49.09, 76.98) \unit{eV}, L_i/L = (0.23, 0.36, 0.24, 0.06)$ that
+are optimized to reduce errors in the solutions with a full spectrum
+and energy discretization \citep{Mirocha12}.  (Note that this choice
+of energy groups is different from those used in Wise \& Abel 2011.)
+\nocite{Wise11_Moray} We use a minimum angular resolution of 10 rays
 per cell and a constant radiative transfer timestep of 25 kyr.  Figure
-\ref{fig:shadowing} depicts the clump at $t = 15$ Myr with the
-outer layers expanding after being photo-heated.  It also shows the
-sharp shadowing effects of the dense clump in the neutral fraction
-plot that is representative of ray tracing techniques.
+\ref{fig:shadowing} depicts the clump at $t = 15$ Myr with the outer
+layers expanding after being photo-heated.  It also shows the sharp
+shadowing effects of the dense clump in the neutral fraction plot that
+is representative of ray tracing techniques.
 

### 2013-10-22 refreport/response.tex
updated response to fix typo and talk about test problems.

@@
 
-\red{(Add response regarding test problems here)}
+Regarding test problems, the referee's comment is:
+
+\begin{verbatim}
+However, the paper is let down due to the lack of tests presented for
+many of the algorithms presented in the paper. In particular, the only
+test presented for the MHD algorithm is the Orszag-Tang vortex test,
+which 'does not seem to be extremely discriminating for MHD
+algorithms' (Stone et al., 2008). In addition, results from some
+tests, such as the photo-evaporation test, do not seem to be properly
+compared to prior results in the literature (e.g. Jiang et
+al. 2012). Finally, no tests appear to be presented for some tests
+entirely, e.g. the Chemistry algorithms presented in Section 6.1.
+\end{verbatim}
+
+We have modified the test problem section as follows:
+
+\begin{itemize}
+\item \red{cooling Britton fill in here}
+\item \red{radiation John fill in here}
+\item \red{MHD Dave fill in here}
+\end{itemize}
+
+We also note that many of the fundamental \enzo\ algorithms are
+extensively tested in separate method papers that focus on a single
+algorithm/physics package, and a single test problem or small number
+of test problems are included in the full \enzo\ method paper as a
+demonstration that the algorithms behave sensibly.  In the subsection
+describing these test problems, we cite the full method paper for the
+appropriate algorithm and the direct the reader to it for more
+information.  These additional algorithm method papers include ones
+for both flavors of MHD contained in \enzo\ \citep{Wang:2009a,
+Collins10}, the ray-tracing code \citep{Wise11_Moray}, and the
+flux-limited diffusion \citep{ReynoldsHayesPaschosNorman2009}.
 
@@
 the same issue of the Astrophysical Journal in 1992, and were
-published back-to-pack as consecutive articles in the same volume.  In
+published back-to-back as consecutive articles in the same volume.  In
 a time when the primary mode of consumption of articles was via paper

### 2013-10-22 tests-photoevaporation.tex
Updating the photo-shadowing test to compare with the Iliev et al. (2009) paper.  Adding the PDF of temperature and Mach number and line cuts of density, temperature, pressure, and neutral fraction.

@@
 \label{sec.tests.raytracing}
+
+\begin{figure}
+  \centering
+  \includegraphics[width=0.48\textwidth]{figures/shadowing-rays.eps}
+  \hfill
+  \includegraphics[width=0.48\textwidth]{figures/shadowing-pdf.eps}
+  \caption{Photo-evaporation of a dense clump using the Moray
+    ray-tracing module.  Left panel: Line cuts in the axis from the
+    point source to the clump center at $t = 1, 5, 10, 15$~Myr
+    (clockwise from the upper-left side) of the density, temperature,
+    pressure, and neutral fraction.  Right panel: Volume-weighted
+    probability distribution functions for the temperature (left-hand
+    panel) and flow Mach number (right-hand panel) at $t = 10
+    \unit{Myr}$.}
+  \label{fig:shadowing1}
+\end{figure}
+
+\begin{figure}
+  \centering
+  \includegraphics[width=1.0\textwidth]{figures/shadowing-slices.eps}
+  \caption{Photo-evaporation of a dense clump using the Moray
+ray-tracing module.  Clockwise from the upper left: slices
+through the clump center of density, temperature, neutral hydrogen
+fraction, and pressure at $t=10$~Myr after the initialization of the
+simulation.  A point source of radiation is at the center of the -x
+boundary and illuminates the clump with a constant luminosity, casting
+a clear shadow behind the clump, as seen in the neutral hydrogen
+fraction.}
+  \label{fig:shadowing2}
+\end{figure}
 
@@
 \nocite{Wise11_Moray} We use a minimum angular resolution of 10 rays
-per cell and a constant radiative transfer timestep of 25 kyr.  Figure
-\ref{fig:shadowing} depicts the clump at $t = 15$ Myr with the outer
-layers expanding after being photo-heated.  It also shows the sharp
-shadowing effects of the dense clump in the neutral fraction plot that
-is representative of ray tracing techniques.
+per cell and a constant radiative transfer timestep of 25 kyr.
 
-\begin{figure}
-  \centering
-  \includegraphics[width=1.0\textwidth]{figures/code-test-shadowing.eps}
-  \caption{Photo-evaporation of a dense clump using the Moray
-ray-tracing module.  Clockwise from the upper left: slices
-through the clump center of density, temperature, neutral hydrogen
-fraction, and pressure at $t=15$~Myr after the initialization of the
-simulation.  A point source of radiation is at the center of the -x
-boundary and illuminates the clump with a constant luminosity, casting
-a clear shadow behind the clump, as seen in the neutral hydrogen
-fraction.}
-  \label{fig:shadowing}
-\end{figure}
+The left panel in Figure \ref{fig:shadowing1} shows cuts of the
+density, temperature, neutral fraction, and pressure in a line
+connecting the source and the clump center at $t = 1, 5, 10, 15$~Myr.
+At 1~Myr, the ionization front has propagated through the leftmost
+500~pc of the clump.  This heated gas is now overpressurized, as seen
+in the pressure plot of Figure \ref{fig:shadowing1} that expands into
+the ambient medium.  This photo-evaporative flow becomes apparent at
+5~Myr with the increased density and pressure on the left side of the
+clump.  The right panel in Figure \ref{fig:shadowing1} shows the
+volume-weighted distributions of the temperature and flow Mach number
+in the entire domain at $t = 10 \unit{Myr}$.  Overall, our results
+agree well with the codes presented in \citet{IlievEtAl2009} and the
+uniform grid test of \moray~\citep{Wise11_Moray}.
+
+However, there are some small differences in the line cuts and
+distributions that arise from our choice of energy discretization that
+better samples the high-energy tail of a $T = 10^5 \unit{K}$ blackbody
+spectrum and does not signify possible shortcomings of \moray.  First,
+in the temperature cut at $t = 10 \unit{Myr}$, the high-energy
+radiation photo-heat the gas to $T = 10^4 \unit{K}$, but it remains
+primarily neutral.  The temperature dip at $x \simeq 0.88$ is the gas
+has not come into thermal equilibrium with the incident radiation.  A
+stronger temperature dip exists in the \citeauthor{IlievEtAl2009}
+results and varies from 1000 to 8000~K between codes.  Another minor
+difference is the larger breadth in the temperature distribution
+feature centered at $T = 4.5 \times 10^4 \unit{K}$ that corresponds to
+the ambient medium.  This difference also originates from the
+different sampling in the spectral energy distribution of the source.
+Figure \ref{fig:shadowing2} depicts the same quantities in a
+two-dimensional slice through the clump center at $t = 10$ Myr.  At
+this time, it is apparent that the outer layers are expanding after
+being photo-heated.  These slices also show the sharp shadowing
+effects of the dense clump in the neutral fraction plot that is
+representative of ray tracing techniques.
+
+%%% Local Variables: 
+%%% mode: latex
+%%% TeX-master: "ms"
+%%% End:

### 2013-10-23 test_problems/IonizationBalance/enzo_paper_equilibrium_figure.py
Adding script used to make ionization balance test figures.

@@ -1,0 +1,183 @@
+# This figure requires the grackle chemistry and cooling library.
+# Visit grackle.readthedocs.org for more information.
+
+from matplotlib import pyplot
+
+from utilities.testing import *
+
+from pygrackle.grackle_wrapper import *
+from pygrackle.fluid_container import FluidContainer
+
+from utilities.api import \
+     setup_fluid_container, \
+     calculate_mean_molecular_weight, \
+     calculate_hydrogen_number_density, \
+     set_cosmology_units, \
+     get_cooling_units
+
+from utilities.primordial_equilibrium import \
+     total_cooling, \
+     nHI, nHII, nHeI, nHeII, nHeIII, ne
+     
+from utilities.physical_constants import \
+     mass_hydrogen_cgs, \
+     sec_per_Myr, \
+     sec_per_Gyr, \
+     cm_per_mpc
+
+my_chem = chemistry_data()
+my_chem.use_chemistry = 1
+my_chem.with_radiative_cooling = 0
+my_chem.primordial_chemistry = 1
+my_chem.metal_cooling = 0
+my_chem.UVbackground = 0
+my_chem.include_metal_heating = 0
+my_chem.comoving_coordinates = 0
+my_chem.a_units = 1.0
+my_chem.density_units = mass_hydrogen_cgs
+my_chem.length_units = 1.0
+my_chem.time_units = 1.0
+my_chem.velocity_units = my_chem.length_units / my_chem.time_units
+current_redshift = 0.0
+fc = setup_fluid_container(my_chem, current_redshift=current_redshift,
+                           converge=True, tolerance=1e-6, max_iterations=np.inf,
+                           dt=(0.0001 * sec_per_Myr / my_chem.time_units))
+calculate_temperature(fc)
+a = 1.0 / (1.0 + current_redshift) / my_chem.a_units
+calculate_cooling_time(fc, a)      
+t_sort = np.argsort(fc["temperature"])
+t_cool = fc["cooling_time"][t_sort] * my_chem.time_units
+my_T = fc["temperature"][t_sort]
+my_nH = calculate_hydrogen_number_density(my_chem, fc).mean()
+cooling_rate = fc["energy"][t_sort] / t_cool * fc["density"] * \
+  my_chem.density_units / my_nH**2
[…]

### 2013-10-23 table_of_variables_and_constants.tex
Fixing format bug.

--- 2013-06-25 table_of_variables_and_constants.tex
@@
   $\rho_{\rm SF}$ & constant proper density threshold for star formation \\
-  $\tau_*$ & characteristic star formation timescale (for Schmidt-Law method) \\[3pt]
+  $\tau_*$ & characteristic star formation timescale (for Schmidt-Law method) \\
+  $\mu$ & mean molecular weight \\[3pt]
 

### 2013-10-24 tests-bw.tex
adding brio & wu mhd test

\subsubsection{MHD: Brio \& Wu Shock Tube}
\label{sec.tests.bw}

A standard test of MHD solvers is the MHD analogue of the Sod shock tube,
described by \citet{Brio88}.   This test, which uses $\gamma=2.0$ and $B_x=0.75$
throughout; $\rho=1.0$, $B_y=1.0$, and $P=1.0$ in the left state; and $\rho=0.125$,
$B_y=-1.0$ and $P=0.1$ in the right state.  A total of 800 grid points were
used, and run to a time of $t=0.08$.  Both tests were run using the HLLD Riemann
solver of \citet{Miyoshi05} and a Piecewise Linear Method reconstruction.  The
results of the test are shown in Figure \ref{fig.brio}.  In that figure, the
black lines show the results of the CT solver while the grey lines show the
reference solution, run at 10,000 points.  A light grey line shows the dedner
results, but is completely indistinguishable from the points of the CT method.
To highlight the differences between the CT and Dedner sovlers, 
the lower pane of each plot shows the difference between the two solutions,
which is at most a few percent.   Most shocks are resolved by a few grid points,
though the contact discontinuity (the large jump seen in density but not
the other quantities) is slightly more smeared.

\begin{figure}
\begin{center}
\includegraphics[width=1\textwidth]{figures/bw.eps}
\caption{The shock tube of \citet{Brio88}.  Top panels show the result at
t=0.08
for the CT scheme (black line and points) and Dedner scheme (light grey line,
entirely hidden behind black points) and reference solution, run at 10,000
zones.  Bottom panels
shows the difference between CT and Dedner.}
\label{fig.brio}
\end{center}
\end{figure}

### 2013-10-24 tests-bw.tex
adding reference, fixing typos

@@
 used, and run to a time of $t=0.08$.  Both tests were run using the HLLD Riemann
-solver of \citet{Miyoshi05} and a Piecewise Linear Method reconstruction.  The
+solver of \citet{Miyoshi05} and a Piecewise Linear Method \citep{Colella85} reconstruction.  The
 results of the test are shown in Figure \ref{fig.brio}.  In that figure, the
 black lines show the results of the CT solver while the grey lines show the
-reference solution, run at 10,000 points.  A light grey line shows the dedner
+reference solution, run at 10,000 points.  A light grey line shows the Dedner
 results, but is completely indistinguishable from the points of the CT method.
-To highlight the differences between the CT and Dedner sovlers, 
+To highlight the differences between the CT and Dedner solvers, 
 the lower pane of each plot shows the difference between the two solutions,

### 2013-10-26 tests-ionbalance.tex

@@
 Equations \ref{eq:eqnHI}--\ref{eq:eqne} using the rate equations 
-in \enzo, referenced in \s\ref{sec.num.chemistry}, overplotted with
+in \enzo, referenced in \S\ref{sec.num.chemistry}, overplotted with
 the dashed lines and an analytical solution using the rates 

### 2013-10-26 acknowledgments.tex
added missing acknowledgments and fixed a typo.

@@
 Research, 
-the National Center for Supercomputing Applications, 
+the National Center for Supercomputing Applications at the University
+of Illinois in Urbana-Champaign, 
 the Pennsylvania State University,
- the San Diego Supercomputer Center (through the Strategic Applications
+the San Diego Supercomputer Center at US San Diego (through the Strategic Applications
 Partner program and the Director’s office),
@@
 Stanford University,
+the Texas Advanced Computing Center at the University of Texas,
 the University of Arizona,
-the University of Califoria at San Diego, 
-the University of Colorado at Boulder,
+the University of California at San Diego, 
+the University of California at Santa Cruz,
+the University of Colorado at Boulder and the Janus supercomputing collaboration,
 the University of Florida,
@@
 Computational resources for \enzo\ development have come from the NSF
-XSEDE program, the NASA High Performance Computing program, the DOE INCITE
-program, and the DOE Advanced Simulation and Computing (ASC) program.
+XSEDE program (at NCSA, SDSC, and TACC), the NASA High-End Computing program (through the NASA Advanced Supercomputing (NAS) Division at Ames Research Center), the DOE INCITE
+program, the DOE Advanced Simulation and Computing (ASC) program, and
+the Oak Ridge Leadership Computing Facility.
 

### 2013-10-26 refreport/response.tex
updating referee response

@@
 \begin{itemize}
-\item \red{cooling Britton fill in here}
-\item \red{radiation John fill in here}
-\item \red{MHD Dave fill in here}
+\item We have added a section (11.2.10) showing how both \enzo\ MHD
+  algorithms perform on the standard Brio \& Wu shock tube test
+  problem, which is a standard test of MHD solvers.  Both \enzo\ MHD
+  algorithms match the reference solution to high precision.  We retain the
+  Orszag-Tang Vortex test as we fell that it is interesting to the
+  intended audience.
+
+\item We have added a test of the non-equilibrium primordial chemistry code that
+  tests the ionization balance in ionized gas as compared to an
+  analytical solution (Section 11.2.13).  As can be seen, \enzo's
+  non-equilibrium chemistry solver  reproduces the analytical solution perfectly.
+
+\item The ``Photo-evaporation of a dense clump'' test problem (Section
+  11.2.14) has been expanded to include a new figure (Fig. 18) that
+  contains much more detail about the evolution of the test problem,
+  and comparison to prior results, including the MORAY method paper
+  \citep{Wise11_Moray} and the Cosmological Radiative Transfer
+  Comparison Project paper \citep{IlievEtAl2009}.
+
 \end{itemize}
@@
 of test problems are included in the full \enzo\ method paper as a
-demonstration that the algorithms behave sensibly.  In the subsection
-describing these test problems, we cite the full method paper for the
+demonstration that the algorithms behave sensibly.  In the paper's
+section or subsection
+describing these algorithm, we cite the full method paper for the
 appropriate algorithm and the direct the reader to it for more
@@
 given the topics it covers, and thus does not need to be split into
-multiple parts.  At present, the \enzo\ method paper is \red{60 pages
-in ApJ format (BWO: we will need to update this before submission)}.
+multiple parts.  At present, the \enzo\ method paper is 62 pages long
+in ApJ format.
 For comparison:
@@
 
+\section{Additional changes}
+
+In addition to directly addressing concerns expressed by the referee,
+we have fixed additional typos in the paper.  These include:
+
+\begin{enumerate}
+\item  We have corrected a typo in the equation describing the Jeans
+  length refinement criterion (Equation 29).
+
+\item We have corrected a typo in Table 4, where two photoionization
+  rates were accidentally switched in the submitted version.
+
+\item We have added missing acknowledgments.
+
+\end{enumerate}
+
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

### 2013-10-28 tests-vandv.tex
updated the paper w/enzo 2.3 changeset hash

@@
 version 2.5.3 and the \url{https://bitbucket.org/enzo/enzo-dev}
-repository of \enzo\ with changeset \texttt{5d8c412}, which
-corresponds to \enzo\ 2.3.
+repository of \enzo\ with changeset \texttt{bad70b86fcd6}, which
+corresponds to the \enzo\ 2.3 code release.
 

### 2013-11-01 refreport/response.tex
fixed typo.

@@
   algorithms match the reference solution to high precision.  We retain the
-  Orszag-Tang Vortex test as we fell that it is interesting to the
+  Orszag-Tang Vortex test as we feel that it is interesting to the
   intended audience.
@@
 
+\item We have updated the mercurial changeset hash to correspond to
+  the \enzo\ 2.3 release (which was done slightly after the submission
+  of the paper to ApJ).
+
 \end{enumerate}

### 2013-11-06 introduction.tex

@@
 flash_method, MLAPM01, Yahagi01, RAMSES, Quilis04, Ziegler05, Zhang06,
-Astrobear09, Pluto-amr, GAMER, Nyx}. Of these, perhaps the most
+CHARM, Astrobear09, Pluto-amr, GAMER, Nyx}. Of these, perhaps the most
 comparable and widely used are FLASH \citep{flash_method}, which uses

### 2013-11-10 test_problems/PhotonShadowing/make_plots.py
Adding to the photon shadowing plotting script to include the new plots made after the referee report.

@@ -1,5 +1,40 @@
+import os, cPickle, math
+import h5py as h5
+
 from yt.mods import *
 import yt.visualization.eps_writer as EPS
+from matplotlib import rc
+import matplotlib.pyplot as plt
+font= FontProperties(size='small')
+rc('font', family='serif', serif='cmr10', size=18)
+from matplotlib.font_manager import fontManager, FontProperties
+
+from matplotlib.ticker import NullFormatter
+nullfmt = NullFormatter()
+
+########################################################################
+# Derived fields
+########################################################################
+def _MyRadius(field, data):
+    center = data.get_field_parameter("center")
+    dx = data["x"] - center[0]
+    dy = data["y"] - center[1]
+    dz = data["z"] - center[2]
+    return na.sqrt(dx*dx + dy*dy + dz*dz)
+add_field("Radius", function=_MyRadius, take_log=False, units='')
+def _MyIonizedFrac(field, data):
+    return data['HII_Fraction'] / 0.75908798
+add_field("Ionized_Fraction", function=_MyIonizedFrac, take_log=True,
+          units='')
+def _MyNeutralFrac(field, data):
+    return data['HI_Fraction'] / 0.75908798
+add_field("Neutral_Fraction", function=_MyNeutralFrac, take_log=True,
+          units='')
+########################################################################
+
+########################################################################
+# Slice multi-panel plot
+########################################################################
 
 output_num = 150
 #fields = ["Density", "HI_Fraction", "Pressure", "Temperature"]
@@ -20,3 +55,144 @@ ep = EPS.multiplot_yt(2,2,pc, margins=(0.2,0.2), bare_axes=True)
 ep.scale_line(1.0/6.6, '1 kpc')
 ep.title_box("15 Myr")
 ep.save_fig("code-test-shadowing.eps")
+
+########################################################################
+# Line cuts through the clump center at t = 1, 5, 10, 15 Myr
+########################################################################
+
+output = [10,50,100,150]
+fields = ["Density", "Temperature", "Neutral_Fraction", "Pressure"]
+
[…]

### 2013-11-11 test_problems/SedovAMR/make_plots.py
added new plots for internal energy, pressure, velocity.  also updated the density one so that everything actually looks like the plots in the paper.

@@ -13,6 +13,9 @@ execfile('sedov_outputdata.py')
 # Makes slices of density for both simulations (PPM, Zeus)
 execfile('sedov_slice.py')
 
-# plots radially-averaged density profile from both simulations, along
+# plots annulus-averaged profiles from both simulations, along
 # with the analytical result from sedov.in
-execfile('sedov_plot.py')
+execfile('sedov_density_plot.py')
+execfile('sedov_internal_energy_plot.py')
+execfile('sedov_pressure_plot.py')
+execfile('sedov_velocity_plot.py')

### 2013-11-11 test_problems/SedovAMR/sedov_internal_energy_plot.py

@@ -1,0 +1,60 @@
+from yt.mods import *
+# plots radially-averaged pressure profile from both simulations, along
+# with the analytical result from sedov.in
+
+import matplotlib.pyplot as plt
+
+################ PPM ################
+# Load the dataset.
+pf_ppm = load("PPM/DD0007/DD0007")
+
+pc_ppm = PlotCollection(pf_ppm, center=[0.5, 0.5, 0.5])
+
+pc_ppm.add_profile_sphere(0.5, "cm", ["Radius", "ThermalEnergy"], x_bins=100, x_log=False)
+thermal_energy_data_ppm = pc_ppm.plots[-1].data["ThermalEnergy"]
+radius_data_ppm =  pc_ppm.plots[-1].data["Radius"]
+
+
+################ Zeus ################
+# Load the dataset.
+pf_zeus = load("Zeus/DD0007/DD0007")
+
+pc_zeus = PlotCollection(pf_zeus, center=[0.5, 0.5, 0.5])
+
+pc_zeus.add_profile_sphere(0.5, "cm", ["Radius", "ThermalEnergy"], x_bins=100, x_log=False)
+thermal_energy_data_zeus = pc_zeus.plots[-1].data["ThermalEnergy"]
+radius_data_zeus =  pc_zeus.plots[-1].data["Radius"]
+
+################################ Make plot ################
+
+# read in anlytical solution
+sedovsol_radius = []
+sedovsol_thermal_energy = []
+file = open("sedov.in", "r")
+for line in file:
+    linesplit = line.split('\t')
+    if linesplit[0] != '#':
+        sedovsol_radius.append(float(linesplit[0]))
+        sedovsol_thermal_energy.append(float(linesplit[4]))
+        
+
+# python plot                                                                                                                     
+plotfig = plt.figure()
+
+# normalised mass by total mass                                                                                                   
+plt.plot(sedovsol_radius, sedovsol_thermal_energy,'k-', linewidth=3)  # analytic solution
+plt.plot(radius_data_ppm, thermal_energy_data_ppm, 'bo-', radius_data_zeus, thermal_energy_data_zeus, 'ro-', linewidth=1)
+plt.yscale('log')
+
+plt.axis([0.2, 0.5, 1.0, 500.0])
+
+plt.figtext(0.75,0.76,"Zeus",color='r', fontsize=14)
+plt.figtext(0.75,0.80,"PPM",color='b', fontsize=14)
+plt.figtext(0.75,0.84,"Analytic", color='k', fontsize=14)
+
[…]

### 2013-11-11 test_problems/SedovAMR/sedov_pressure_plot.py

@@ -1,5 +1,5 @@
 from yt.mods import *
-# plots radially-averaged density profile from both simulations, along
+# plots radially-averaged pressure profile from both simulations, along
 # with the analytical result from sedov.in
 
 import matplotlib.pyplot as plt
@@ -10,14 +10,9 @@ pf_ppm = load("PPM/DD0007/DD0007")
 
 pc_ppm = PlotCollection(pf_ppm, center=[0.5, 0.5, 0.5])
 
-pc_ppm.add_profile_sphere(0.5, "cm", ["Radius", "CellMass"], weight=None, x_bins=100, x_log=False)
-mass_data_ppm = pc_ppm.plots[-1].data["CellMass"]
+pc_ppm.add_profile_sphere(0.5, "cm", ["Radius", "Pressure"], x_bins=100, x_log=False)
+pressure_data_ppm = pc_ppm.plots[-1].data["Pressure"]
 radius_data_ppm =  pc_ppm.plots[-1].data["Radius"]
-
-pc_ppm.add_profile_sphere(0.5, "cm", ["Radius", "CellVolume"], weight=None, x_bins=100, x_log=False)
-volume_data_ppm = pc_ppm.plots[-1].data["CellVolume"]
-
-density_data_ppm = mass_data_ppm/volume_data_ppm
 
 
 ################ Zeus ################
@@ -26,44 +21,39 @@ pf_zeus = load("Zeus/DD0007/DD0007")
 
 pc_zeus = PlotCollection(pf_zeus, center=[0.5, 0.5, 0.5])
 
-pc_zeus.add_profile_sphere(0.5, "cm", ["Radius", "CellMass"], weight=None, x_bins=100, x_log=False)
-mass_data_zeus = pc_zeus.plots[-1].data["CellMass"]
+pc_zeus.add_profile_sphere(0.5, "cm", ["Radius", "Pressure"], x_bins=100, x_log=False)
+pressure_data_zeus = pc_zeus.plots[-1].data["Pressure"]
 radius_data_zeus =  pc_zeus.plots[-1].data["Radius"]
 
-pc_zeus.add_profile_sphere(0.5, "cm", ["Radius", "CellVolume"], weight=None, x_bins=100, x_log=False)
-volume_data_zeus = pc_zeus.plots[-1].data["CellVolume"]
-
-density_data_zeus = mass_data_zeus/volume_data_zeus
-
-
-################################ Make density plot ################
+################################ Make plot ################
 
[…]

### 2013-11-11 test_problems/SedovAMR/sedov_velocity_plot.py

@@ -1,5 +1,5 @@
 from yt.mods import *
-# plots radially-averaged density profile from both simulations, along
+# plots shell-averaged radial velocity profile from both simulations, along
 # with the analytical result from sedov.in
 
 import matplotlib.pyplot as plt
@@ -10,14 +10,9 @@ pf_ppm = load("PPM/DD0007/DD0007")
 
 pc_ppm = PlotCollection(pf_ppm, center=[0.5, 0.5, 0.5])
 
-pc_ppm.add_profile_sphere(0.5, "cm", ["Radius", "CellMass"], weight=None, x_bins=100, x_log=False)
-mass_data_ppm = pc_ppm.plots[-1].data["CellMass"]
+pc_ppm.add_profile_sphere(0.5, "cm", ["Radius", "RadialVelocity"], x_bins=100, x_log=False)
+radial_velocity_data_ppm = pc_ppm.plots[-1].data["RadialVelocity"]
 radius_data_ppm =  pc_ppm.plots[-1].data["Radius"]
-
-pc_ppm.add_profile_sphere(0.5, "cm", ["Radius", "CellVolume"], weight=None, x_bins=100, x_log=False)
-volume_data_ppm = pc_ppm.plots[-1].data["CellVolume"]
-
-density_data_ppm = mass_data_ppm/volume_data_ppm
 
 
 ################ Zeus ################
@@ -26,44 +21,38 @@ pf_zeus = load("Zeus/DD0007/DD0007")
 
 pc_zeus = PlotCollection(pf_zeus, center=[0.5, 0.5, 0.5])
 
-pc_zeus.add_profile_sphere(0.5, "cm", ["Radius", "CellMass"], weight=None, x_bins=100, x_log=False)
-mass_data_zeus = pc_zeus.plots[-1].data["CellMass"]
+pc_zeus.add_profile_sphere(0.5, "cm", ["Radius", "RadialVelocity"], x_bins=100, x_log=False)
+radial_velocity_data_zeus = pc_zeus.plots[-1].data["RadialVelocity"]
 radius_data_zeus =  pc_zeus.plots[-1].data["Radius"]
 
-pc_zeus.add_profile_sphere(0.5, "cm", ["Radius", "CellVolume"], weight=None, x_bins=100, x_log=False)
-volume_data_zeus = pc_zeus.plots[-1].data["CellVolume"]
-
-density_data_zeus = mass_data_zeus/volume_data_zeus
-
-
-################################ Make density plot ################
[…]

### 2013-11-11 test_problems/PhotonShadowing/make_plots.py
Fixing some bugs in the photon shadowing merged plotting script after trying to re-run the combined version.

@@ -1,13 +1,10 @@
-import os, cPickle, math
-import h5py as h5
-
 from yt.mods import *
 import yt.visualization.eps_writer as EPS
 from matplotlib import rc
 import matplotlib.pyplot as plt
+from matplotlib.font_manager import fontManager, FontProperties
 font= FontProperties(size='small')
 rc('font', family='serif', serif='cmr10', size=18)
-from matplotlib.font_manager import fontManager, FontProperties
 
 from matplotlib.ticker import NullFormatter
 nullfmt = NullFormatter()
@@ -54,7 +51,7 @@ for i,f in enumerate(fields):
 ep = EPS.multiplot_yt(2,2,pc, margins=(0.2,0.2), bare_axes=True)
 ep.scale_line(1.0/6.6, '1 kpc')
 ep.title_box("15 Myr")
-ep.save_fig("code-test-shadowing.eps")
+ep.save_fig("shadowing-slices.eps")
 
 ########################################################################
 # Line cuts through the clump center at t = 1, 5, 10, 15 Myr
@@ -163,7 +160,7 @@ ni = len(outputs)
 nj = len(fields)
 
 plt.clf()
-plt.subplots_adjust(right=0.97, top=0.95, hspace=1e-3)
+plt.subplots_adjust(right=0.97, top=0.95, hspace=1e-3, wspace=0.2)
 
 index = 1
 for i,out in enumerate(outputs):

### 2013-11-11 numerical-amr.tex
Correcting a typo in the Jeans refinement criterion.
Correcting another error associated with the Jeans refinement criterion.  The equation doesn't include a factor of mu or gamma which are in the code.
Modified shear refinement criteria to reflect what is currently in code.

@@
 \begin{equation}
-\Delta x < \left( \frac{\pi k_B T}{N_J^2 G \rho \mh} \right)^{1/2},
+\Delta x < \left( \frac{\gamma \pi k_B T}{N_J^2 G \rho \mu \mh} \right)^{1/2},
 \end{equation}
-where $J_J$ is the required number of cells per Jeans length (4 by default).
+where $N_J$ is the required number of cells per Jeans length (4 by default).
 
@@
 \end{equation}
-where $\epsilon_s$ is a dimensionless parameter.
+where $\epsilon_s$ is a dimensionless parameter. The current
+implementation assumes that $c_s$ is a constant and that $\Delta x$ in
+the above expression is evaluated with the cell size on the level
+being tested.
 

### 2013-11-12 numerical-amr.tex
Updating the shear criterion to match the current pending PR.  The old method written for Kritsuk et al 2006 refined on vorticity + shear, not shear alone.

@@
 
-\item{\em Shear.} As described in \citet{Kritsuk06}, turbulence
-  calculations can be successfully modeled with AMR if the shear is
-  used as a refinement parameter; in particular, refinement occurs
-  when the finite difference version of the following inequality
+\item{\em Shear.} Originally used in \citet{Kritsuk06} and later 
+  modified into a more general form, turbulence calculations and
+  Kelvin-Helmholtz instabilities can be successfully modeled with AMR if 
+  the shear is used as a refinement parameter; in particular, refinement 
+  occurs when the finite difference version of the following inequality
   holds:
@@
 \begin{equation}
-\sum_i \sum_j \left( \frac{\partial v_i}{\partial x_j} \right)^2 -  \sum_i \left( \frac{\partial v_i}{\partial x_i} \right)^2
-> \epsilon_s \frac{c_s^2}{\Delta x^2},
+\left| \sum_{i \ne j} \left( \frac{\partial v_i}{\partial x_j} + \frac{\partial v_j}{\partial x_i} \right) \right|
+> \epsilon_s \frac{c_s}{\Delta x},
 \end{equation}
-where $\epsilon_s$ is a dimensionless parameter. The current
-implementation assumes that $c_s$ is a constant and that $\Delta x$ in
-the above expression is evaluated with the cell size on the level
-being tested.
+where $\epsilon_s$ is a dimensionless parameter.
 

### 2013-11-18 refreport/response.tex
updated response to referee's report to include shear refinement fix.

@@
 \item  We have corrected a typo in the equation describing the Jeans
-  length refinement criterion (Equation 29).
+  length refinement criterion (Section 3.6, Equation 29).
+
+\item We have updated our description of the shear refinement
+  criterion (Section 3.6, Equation 30).
 

### 2013-11-18 code-tests.tex
Removed commented-out text from lots of files - basically vestigial junk.
Adding ionization balance test.
adding brio & wu mhd test
Point out that we also test the code in the other enzo method papers (MHD, radiation).

@@
 to a ``gold standard'' solution from a stable version of \enzo.
+We also note that there are many other \enzo\ test problems in the
+associated papers which describe individual methods, including:
+MHD-CT \citep{Collins10}, MHD-Dedner \citep{WangAbelZhang08,
+WangAbel09}, Radiation-FLD \citep{NBHBROW2007,
+ReynoldsHayesPaschosNorman2009, NRS2009}, Radiaton-ray-tracing
+\citep{Wise11_Moray}.
 
@@
 figures themselves.
-
-% General structure for each test problem: Outline how the test problem
-% is constructed (initial and boundary conditions), the analytical
-% solution, why we have in the paper (what does it break, or what flaw
-% does it expose (not in enzo of course)), a plot showing how well enzo
-% solves said test problem, and a brief description of the plot and how
-% awesome enzo is.
 
@@
 
+\input{tests-bw}
+
 \input{tests-orszagtang}
@@
 \input{tests-onezonecollapse}
+
+\input{tests-ionbalance}
 
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-%%% Local Variables: 
-%%% mode: latex
-%%% TeX-master: "ms"
-%%% End:

### 2013-11-18 numerical-amr.tex
Very minor clean-up.

@@
 interior and negative $x$ exterior to the domain):
-
 \begin{enumerate}
@@
   in the cell ($M_g = \rho\, (\Delta x)^d$) is larger than a specified value:
-
 \begin{equation}
@@
 \end{equation}
-
 where $\rho_{\rm flag}$ is the equivalent density on the root grid
@@
   contain shocks using the following criteria:
-
 \begin{eqnarray}
@@
 \end{eqnarray}
-
 We note that the parameters $0.33$ and $0.1$ in the equations above
@@
   more specifically, whenever the following criterion is not met:
-
 \begin{equation}
@@
   holds:
-
 \begin{equation}
-\left| \sum_{i \ne j} \left( \frac{\partial v_i}{\partial x_j} + \frac{\partial v_j}{\partial x_i} \right) \right|
-> \epsilon_s \frac{c_s}{\Delta x},
+\left( \frac{\partial v_x}{\partial y} + \frac{\partial v_y}{\partial x} \right)^2 +
+\left( \frac{\partial v_z}{\partial y} + \frac{\partial v_y}{\partial z} \right)^2 +
+\left( \frac{\partial v_x}{\partial z} + \frac{\partial v_z}{\partial x} \right)^2
+ >  \left(\epsilon_s \frac{c_s}{\Delta x}\right)^2
 \end{equation}
-where $\epsilon_s$ is a dimensionless parameter.
+where $\epsilon_s$ is a dimensionless user-defined parameter and $\Delta x$ 
+is the local cell size at the level where refinement is occurring.
 
@@
 optical depth of a cell is estimated to be:
-
 \begin{equation}
@@
   This is done by ensuring that the resistive length scale, defined as
-
 \begin{equation}
@@
 \end{equation}
-
 is always refined by at least N$_R$ cells locally (i.e.~we refine a
@@
   multi-dimensional version of the criteria takes in to account all
-  cross-derivatives, and is expressed by
-
+  cross-derivatives, and is expressed by,
  \begin{equation}
@@
  \end{equation}
-
   where $f$ is the user-defined field, $\abs{\bar{f_{ij}}}$ is an average of

### 2013-11-18 refreport/response.tex
Very minor changes to referee response.

@@
 by the Enzo collaboration (Bryan et al.).  We thank the referee for
-their feedback, and address each comment separately below.
+his or her feedback, and address each comment separately below.
 
@@
 
-We have modified the test problem section as follows:
+We have addressed this issue by modifying the test problem section as follows:
 
@@
   non-equilibrium chemistry solver  reproduces the analytical solution perfectly.
+  We also note that the chemistry solver is tested in Sections 11.2.12 and 11.2.14.
 
@@
 flux-limited diffusion \citep{ReynoldsHayesPaschosNorman2009}.
+We now explicitly point this out and list these papers in Section 
+11.2, the introduction to testing section.
 
@@
 these sections would need to be either duplicated or require an
-extensive amount of cross-referencing between papers Furthermore,
+extensive amount of cross-referencing between papers. Furthermore,
 separate introductions, discussions, and conclusions will by necessity
